### 2018-07-03 paper/Version10/draft_v10.tex
finally on git

\documentclass[12pt]{article}
\usepackage[margin=1in]{geometry}
%\usepackage{hyperref}
\usepackage{amsmath}
\usepackage{graphicx}% http://ctan.org/pkg/graphicx
\usepackage{systeme}
%% tables
\usepackage{booktabs}
\usepackage[table,xcdraw]{xcolor}
\usepackage{float}

%% Make red text
\newcommand{\com}[1]{\textcolor{red}{ #1}}

%\linespread{2}

\usepackage[authoryear]{natbib}

\usepackage{tikz}
\tikzset{
  int/.style={circle, draw, fill=blue!20, minimum size=3em},
  init/.style={pin distance=1.2cm,pin edge={loop,thin,black}}
}
\usetikzlibrary{arrows,automata}

\pagenumbering{arabic}

\newcommand{\XX}{\ensuremath{25}} % number of methods

\newcommand{\xxsir}{\ensuremath{11}} % number of SIR methods

\newcommand{\rr}{\ensuremath{\mathcal{R}_0}}

\setlength{\itemsep}{0pt}

\begin{document}

%%% draw a tree
\tikzset{
  treenode/.style = {shape=rectangle, rounded corners,
                     draw, align=center,
                     top color=white, bottom color=blue!20},
  root/.style     = {treenode, font=\Large, bottom color=red!30},
  env/.style      = {treenode, font=\ttfamily\normalsize},
  dummy/.style    = {circle,draw}
}

\title{$\XX$ dubious ways of estimating $\rr$}
\author{ Department of Statistics, Carnegie Mellon University}
\date{\today}
\maketitle

% \tableofcontents

\section{Introduction}\label{sec:intro}
What has been called ``arguably the most important quantity in the study of epidemics'', $\mathcal{R}_0$, the reproduction number (by convention pronounced ``R-naught''), for nearly a century, has been an elusive quantity for epidemic modelers to estimate \citep{Heesterbeek2002}.  Defined by \citet{anderson1992}, $\rr$ is the ``the average number of secondary infections produced when one infected individual is introduced into a host population where everyone is susceptible.''  $\rr$, in some ways, summarizes an entire outbreak of a disease; it  is used to assess whether a disease will break out and what percentage of the population needs to be vaccinated to avoid such an outbreak.  Despite a clear definition of $\rr$, epidemiologists have struggled to create a standard  estimator for $\rr$  \citep{hethcote2000}.  The primary issue in estimating $\rr$ is that the quantity is a \textit{property of the model}, meaning that $\rr$ is dependent not only on the usual noise that comes with statistical modeling but also on a variety of assumptions on how researchers assume a disease is transmitted through a population.

Often, disease models are a part of the ``SI-framework'' pioneered by Kermack and McKendrick in the 1920s \citep{getz2006}.  Here, S and I are known as compartments where S stands for ``susceptible'' and I for ``infected.''  The issue of estimating $\rr$ stems from how individuals are assumed, either implicitly or explicitly, to move from the S compartment to the I compartment and any other compartments that have been introduced such as immune, recovered, dead, or exposed, for example.  One of the simplest models in the SI framework is the SIR model where R stands for a ``recovered'' state where individuals are no longer susceptible to the disease \citep{Kermack700}.  Even in this simplified framework, there is much dispute on how $\rr$ is to be estimated arising from assumptions on how populations intermix, if any intervention has been put in place, or whether disease parameters are constant over time.

Other issues about estimating $\rr$ include mathematical versus statistical approaches (i.e. solving for $\rr$ versus estimating $\rr$), how to estimate the quantity in the presence of types of available data, and when to estimate $\rr$ in a disease's lifetime.  This depends on the ``time constant'' of the disease.  Some diseases like measles progress rapdily, HIV more slowly, prion diseases glacially so.  For instance, if only weekly incidence counts of a disease are available, then we have different methods of estimating $\rr$ available than when we have more data available such as contact networks or the serial interval of a disease.  Another major issue is that since $\rr$ is difficult to estimate in the first place, the estimated variance of the estimated $\rr$ has largely been glossed over or  ignored.  This is especially troubling because we are concerned whether $\rr > 1$, which indicates an outbreak of a disease.  A point estimate is simply not enough to estimate the truth of this inequality.

In short, $\rr$ is an important parameter in epidemiology that is very difficult to estimate.  This, in turn, makes it difficult to compare $\rr$ across different diseases or even different instances of the same disease.  Many of these challenges with estimating $\rr$ and more are summarized and expanded upon by \cite{li2011}.

We review $\XX$ ways to estimate $\rr$ across a number of model assumptions and in the presence of different sources of data.  The title of this manuscript is a reference to a seminal paper by Moler and Van Loan (1978) (and its subsequent update in 2003) in which they discuss ``Nineteen dubious ways to compute the exponential of a matrix'' \citep{moler2003}.  We not only review the methods of estimating  $\rr$ but also include recommendations on when to use the methods as well as introducing techniques which allow estimation the variance.  We compare these estimates, when applicable, in the presence of different data, both simulated and real.

The rest of this manuscript is organized as follows.  In Section \ref{sec:overview}, we introduce the methods in four broad groups: the survival function, compartment models, exponential growth models, and agent-based models.    In Section \ref{sec:details}, we give an overview of each of the methods, what assumptions are used, what data is required, and some of the places the method has been used in the past along with ways in which to estimate the variance of the various methods, when possible.  In Section \ref{sec:results}, for a number of different data sets, we give estimates of $\rr$ for the different methods and their variances over a variety of data sets.  Finally, in Section \ref{sec:dis} we discuss our conclusions and recommendations.

\section{Overview of $\XX$ Methods to estimate $\rr$}
\label{sec:overview}

In this manuscript, $\XX$ methods were chosen to give a survey of methods for estimating $\rr$.  This is not exhaustive, but these methods were chosen due to historical importance and impact in the field of epidemiology.

In order to organize the methods for estimating $\rr$, we have chosen four broad classes of estimation methods.  These partitions are meant to be guidelines for distinguishing methods from one another rather than immutable groups.  Many of these methods  fit in many of the groups.  We have chosen the four groups as application of the survival function and its close relatives, classic compartment models, exponential growth assumptions, and estimation from networks.  We list these point estimate methods below in Table \ref{tab:methods} and explain each in detail in Section \ref{sec:details}.  In Table \ref{tab:terms}, we summarize many of the common terms used throughout the document.

% Please add the following required packages to your document preamble:
% \usepackage{booktabs}
% \usepackage[table,xcdraw]{xcolor}
% If you use beamer only pass "xcolor=table" option, i.e. \documentclass[xcolor=table]{beamer}
\begin{table}[h]
  \centering
  \resizebox{\textwidth}{!}{
\begin{tabular}{@{}llll@{}}
\toprule
\textbf{Survival Function}        & \textbf{Compartment Models}            & \textbf{Exponential Growth}                       & \textbf{Networks}        \\ \midrule
Survival Function (SF)            & \textbf{SIR}                           & Exponential Growth (EG)                           & Contact Tracing (CT)     \\
Direct Parameter Estimation (DPE) & - Least Squares (SIR LS)               & Maximum Likelihood of Secondary Infections (MLSI) & Branching Process (BP)   \\
Von Foerster Equations (VFE)      & - Reparameterized LS (SIR ReLS)        & Time Dependent Reproduction Number (TDR)          & Agent-Based Models (ABM) \\
                                  & - Linear Model Approximation (SIR LMA) & Initial Growth and  Final Size (IGFS)             &                          \\
                                  & - LMA, All Time Points (SIR LMAT)      &                                                   &                          \\
                                  & - Max of Data (SIR Max)                &                                                   &                          \\
                                  & - Smooth Max (SIR SMax)                &                                                   &                          \\
                                  & - Incidence-Prevalence Ratio (SIR IPR) &                                                   &                          \\
                                  & - Smooth IPR (SIR SIPR)                &                                                   &                          \\
                                  & - Log Linear (SIR LL)                  &                                                   &                          \\
                                  & - Markov Chain Estimation (SIR MC)     &                                                   &                          \\
                                  & - Sequential Bayes (SIR SB)            &                                                   &                          \\
                                  & SEIR Model (SEIR)                      &                                                   &                          \\
                                  & MSEIR Model (MSEIR)                    &                                                   &                          \\
                                  & Amplifier Model (AMP)                  &                                                   &                          \\
                                  & Next Generation Model (NGM)            &                                                   &                          \\ \bottomrule
\end{tabular}
}
\caption{Table of methods used to estimate $\rr$, divided into four categories.}
\label{tab:methods}
\end{table}

Additionally, we describe  interval estimates for the above methods for estimating $\rr$ in order to better answer the question of whether $\rr > 1$, which implies an outbreak of the disease.  These methods are the delta method and posterior distribution.

% Please add the following required packages to your document preamble:
% \usepackage{booktabs}
\begin{table}[]
\centering
\begin{tabular}{@{}ll@{}}
\toprule
Term        & Meaning                                                                                    \\ \midrule
$\rr$     & reproduction number                                                                        \\
$\mathcal{R}_t$     & effective reproduction number at time $t$                                                  \\
$\beta$     & average rate of infection                                                                  \\
$\gamma$    & average rate of recovery                                                                   \\
$\alpha$    & rate of transfer from exposed to infectious                                                \\
$\delta$    & rate of transfer from immune infant to susceptible                                         \\
$D$         & average duration of an infection                                                           \\
$N$         & population size                                                                            \\
$X(t)$      & number of susceptible individuals at time $t$                                              \\
$Y(t)$      & number of infected individuals at time $t$                                                 \\
$Z(t)$      & number of recovered individuals at time $t$                                                \\
$M(t)$      & number of immune infants at time $t$                                                       \\
$E(t)$      & number of exposed individuals at time $t$                                                  \\
$J(t)$      & incidence at time $t$                                                                      \\
$\omega(t)$ & the time from onset of symptoms in an index case to onset of symptoms in a subsequent case \\
$\mu_\infty$       & percentage of population eventually infected                                               \\
IPR         & incidence to prevalence ratio                                                             \\ \bottomrule
\end{tabular}
\caption{Table of common terms used throughout the document and their meanings}\label{tab:terms}
\end{table}

\section{Details for the $\XX$ methods}
\label{sec:details}
\subsection{Survival function}
\label{sec:direct}

The first set of methods are those of the survival function, which is closely  tied to the history of $\rr$.  The survival function, which originates from the field of demography in the late 1800s, describes how many female offspring a woman would produce \citep{dietz1993estimation}.  In demography, we have
\begin{align*}
\rr = \int_0^\infty p(a) \beta(a) da
\end{align*}
where $p(a)$ denotes the probability of a woman surviving to age $a$ and $\beta(a)$ the rate of an individual of age $a$ giving birth to a  girl.   This interpretation also explains the origins of the name of $\rr$, which was imported to the field of epidemiology by MacDonald and Smith \citep{dietz1993estimation}.  Analogously in epidemiology, $p(a)$ is the age of a disease, and $\beta$ is the transmission rate.

This is the chronologically first, and in many ways, the most direct method to estimate $\rr$.  However, this model also seems to be the most difficult to estimate as both $p(a)$ and $\beta (a)$ are random variables.   In general, it is also difficult to estimate these parameters before or even during an ongoing disease outbreak.  As such, $\rr$ can only be estimated after an outbreak, which means it cannot be used to determine ongoing prevention strategies.  However, having a post-disease estimation of $\rr$ may be useful for preventing future outbreaks of the disease and aids in comparing the severity of many diseases.  Another advantage of this method is that it is intuitive and independent of assumptions of how the disease is transmitted from person to person.    We describe four methods which estimate $\rr$ from the survival function:  the survival function, direct estimation, and direct parameter estimation.

\subsubsection{Survival Function (SF)}
\label{sec:survival_fxn}
The first method of estimating $\rr$, is not surprisingly, the survival function itself as presented by \cite{Heffernan2005}.  We use standard notation from infectious disease epidemiology.

Let $F(a)$ be the probability that a newly infected individual remains infectious for at least time $a$, and let $b(a)$ be the average number of newly infected individuals that an infectious individual will produce per unit time when infected for total time $a$.  Then $\rr$ is given in Equation (\ref{eq:r0_survivalfxn}),

\begin{align}\label{eq:r0_survivalfxn}
  \rr = \int_0^\infty b(a)F(a)da.
\end{align}

This method presents one of the unique difficulties of estimating $\rr$: that of reconciling generations and time.  $\rr$ tells us about the generations of the disease, meaning an infector and its infected, but not necessarily when the infections occur.  This reconciliation between generations and time is made difficult as the data we have about a disease is often of a temporal nature.  On the other hand, this method is flexible in that this derivation of $\rr$  does not place assumptions on how the disease transmits through a population.  A disadvantage is that data about $F(a)$ and especially $b(a)$ is difficult to collect directly.

Estimating this quantity in Equation \ref{eq:r0_survivalfxn} is hard and, in fact, rarely attempted. It is due to this difficulty that many of the other methods for estimating $\rr$ even exist.

\subsubsection{Direct Parameter Estimation (DPE)}
\label{sec:dpe}

Direct parameter estimation, an approach for parameters that are readily available to be estimated, is a `plug-in' estimator approach as described by \cite{lipsitch2003} and \cite{dietz1993estimation}.  This method, like the survival function, is highly interpretable.  We assume that we can estimate $\rr$ if we know how many others an individual comes into contact $(k)$, the probability of transmission given contact $(b)$, and how long the disease lasts within an individual $(D)$.  Then $\rr$ is the product of these parameters, 
\begin{align}\label{eq:dpe}
\rr = kbD.
\end{align}
\cite{lipsitch2003} note that directly estimating $\rr$ in such a way may be difficult due to the available data, as in the previous  method.  They instead estimate $\rr$ using the mean serial interval, $\omega$, the time from onset of symptoms in an index case to the onset of symptoms in a subsequent case infected by the index patient in addition to the assumption of exponential growth of infected cases, as a proxy for the duration $D$.  An advantage of this method is that it reduces the problem to three parameters that we can conceivably collect data on.  A disadvantage is that to estimate $\rr$ alone,  data is only used from the initial outbreak of the disease, so often we have a small sample size, resulting in large confidence sets or credible intervals.

%%%%%%%

\subsubsection{Von Foerster Equations (VFE)}
\label{sec:direct-estim-surv}

A derivation of $\rr$ similar to the survival function comes from the idea of the von Foerster equations, which were used to calculate the number of cells in a ``deme,'' a constrained area.  The function $Y(t, \tau)$ calculates the number of cells of age $\tau$ at time $t$ \citep{trucco1965}.  In epidemiology, $Y$ represents the number of infections instead of cells. Letting $Y(t, \tau)$ be the number of infections at time $t$ of age $\tau$, the von Foerster equation is
\begin{align*}
  \frac{\partial Y(t, \tau)}{\partial t} +   \frac{\partial Y(t, \tau)}{\partial \tau} = -  \lambda \cdot n(t, a) ,
\end{align*}
where $Y(t,\tau)$ is the number of people at time $t$ who were infected time $\tau$ ago and $\lambda$ is the loss function, which we assume to be 0, that is no deaths.  The right hand side represents the net loss of cells.  This derivation is based on the ``reasoning that the number of cells which are in the age category $(a, a + \Delta a)$ at time $t$, is equal to $n(t,a) \Delta a$'' \citep{trucco1965}.     \cite{fraser2004factors} derived an expression for $\rr$ from the von Foerster equation and the additional boundary equation
\begin{align*}
  Y(t,0) &= \int_0^t \beta( \tau) Y(t, \tau) d \tau,
\end{align*}
with initial conditions $Y(0,0) = Y_0$ and $Y(t, \tau) = 0$ when $t < \tau$.  Then the reproduction number is given by Equation \ref{eq:vfr0},
\begin{align}\label{eq:vfr0}
\rr &= \int_0^\infty \beta(\tau) d\tau,
\end{align}
 where $\beta (\tau)$ represents the infectiousness at time $\tau$ since the infection.    Again, this model is more mathematical in nature than practical as $\beta (\tau)$ is difficult to estimate.  To estimate $\rr$, \cite{fraser2004factors} make an approximation that $Y(t, \tau) = K(\tau) \exp (rt)$ and further adapt the method to include prevention strategies.  The main function to be estimated is $\beta (\tau)$.  Estimating $\beta (\tau)$ has many of the same issues as trying to estimate $b(a)$ in the survival function in that it is difficult to estimate while the disease is ongoing but is also advantageous in that it does not place assumptions on how the disease is transmitted throughout the population.

%%%%%%%%%%%%%%%

\subsection{Compartment Models}
\label{sec:cms}

First introduced by \cite{Kermack700}, a large and common set of methods used estimate $\rr$ is that of fitting compartment models (CMs) to disease data.  As described in \cite{daley2001epidemic},  CMs model how objects move from one discrete state to another over time.  These models provide an aggregate model of how diseases may be spread throughout a population.  Within the CM framework, we make four essential assumptions,
\begin{enumerate}
\item The compartments are discrete and have no overlap.
\item The transition of objects into and out of compartments is described by a set of known equations, possibly dependent on unknown parameters.
\item The populations mix homogeneously
  \item The number of objects in each compartment at time $t=0$ is known
  \end{enumerate}

Another common assumption for compartment models, is the law of mass action, a property borrowed from chemistry which says that the mass of the product is proportional to the mass of the reactants.  In epidemiology, this means that the proportion of new infections is proportional to the current  number of susceptible and infected \citep{anderson1992}.  Therefore, the flow of objects from one compartment to another may be dependent on the percentage of objects within one or many of the compartments.

For disease modelling, all CMs fall into the SI-framework, meaning that there is a susceptible population that may become infected.  Thus, we require there be susceptible and infected compartments but also allow additional compartments such as the recovered state, an exposed latent state, and more.  CMs may either work with continuous or discrete time and \cite{getz2006} remark that there is no preference between the two.  Of course, data is always recorded in discrete time!

In this section, we discuss $15$ different CM-based methods used to estimate $\rr$.  Different model specifications of CMs yield different mathematical derivations of $\rr$.  That is, for example, an SIR model with non-constant population will yield a different mathematical formula for $\rr$ than the SEIR model (here ``E'' stands for exposed) with non-constant population.  The difference in these two models is that the SEIR model has an extra compartment which in turn effects the estimates of transitions from one compartment to another.  That said, even within the same compartment model framework, we can estimate $\rr$ in a variety of ways using statistical techniques.  In fact, we describe $\xxsir$ methods to estimate $\rr$, within the SIR framework alone!  Most of these methods are recipes to estimate parameters within the SIR model and may be extended to other CM frameworks such as the SEIR model.  The final method described in this section, the next generation operator (NGM) is a general method to derive $\rr$ from any valid CM.  A particular advantage to CMs compared to survival function methods is that they are easier to estimate as the data required is typically collected about diseases: namely, incidence counts.

\subsubsection{SIR Model}
\label{sec:sir-model}

We begin with the most well known CM, the SIR model.   First introduced to epidemiology by  \cite{Kermack700}, the SIR model governs how individuals transition from susceptible, infected, and recovered compartments.  The model is represented graphically in Figure \ref{fig::sir}.

\begin{figure}[h]
\centering
\begin{tikzpicture}[node distance=3cm,auto,>=latex',every node/.append style={align=center}]
    \node [int,  fill = white!70!blue] (a)              {$S$};
    \node [int,  fill = white!70!blue]           (c) [right of=a] {$I$};
    \node [int,  fill = white!70!blue] (e) [right of=c] {$R$};
    \path[->, auto=false] (a) edge node {} (c)
                          (c) edge node {} (e) ;

\end{tikzpicture}
\caption{Depiction of a SIR model.  One can only get the disease once in this model.}\label{fig::sir}
\end{figure}
In this particular  SIR model, $N$, the number of individuals is constant, as we see no arrows pointing away from the three nodes in Figure \ref{fig::sir}. Simple adaptations of the SIR model can include birth and death rates, which may correspondingly change the derivation of $\rr$ (for further discussion, see Section \ref{sec:ngm}). The parameters  $\beta$ and $\gamma$ have epidemiological meaning.  Here, $\beta$ is the average infection rate and $\gamma$ is the average recovery rate.  The transition of individuals from one compartment to another is represented through the ODEs below.  Throughout, we use $X$, $Y$, and $Z$ for stand-ins for the ODEs for $S$, $I$, and $R$, respectively, to avoid any confusion with $\rr$.
\begin{align}
\systeme{\frac{dX}{dt} = -\frac{\beta XY}{N}, \frac{dY}{dt} = \frac{\beta XY}{N} - \gamma Y, \frac{dZ}{dt} = \gamma Z}. \label{eq::sir}
\end{align}

In words, susceptible individuals become infected at a rate that is proportional to the percentage of infected individuals multiplied by $\beta$, the infection rate, and the number of susceptible individuals.  Infected individuals recover at a rate of $\gamma$ multiplied by the number of infected individuals.

An outbreak occurs if the rate of change of infected individuals is positive,
\begin{align*}
  \frac{dY}{dt} &> 0 \\
  \frac{\beta X Y}{N}  - \gamma Y &> 0 ,\\
  Y \left ( \beta \frac{X}{N} - \gamma \right ) & > 0
\end{align*}
that is,  the rate of new infections is greater than the rate of recovery.  So as long as the number of susceptibles is large, $\frac{X}{N} \approx 1$, then an outbreak will occur if $\rr >1$,
\begin{align}\label{eq:deriv_sir}
  \rr = \frac{\beta}{\gamma} > 1.
  \end{align}
Ideally, we have access data of the number of susceptibles, infecteds, and recovered individuals at different points over time.  From this we would like to estimate the $\beta$ and $\gamma$ parameters and ultimately $\rr$.

There are numerous ways to estimate the parameters in this model.  We detail $\xxsir$ of them here.  In every one of these methods, it is assumed that the SIR model is the correct specification.

\subsubsection{Least Squares ($\beta$, $\gamma$) (SIR LS)}\label{least-squares-beta-gamma}
The first approach to estimate $\rr$ in the SIR model is simply to minimize the joint mean square error for the data collected at each time point and use the plug-in estimator found in Equation \ref{eq:sirls}.  In particular, we find

\begin{align*}
(\hat{\beta}, \hat{\gamma} )&=\text{argmin}_{\beta, \gamma} \sum_{t} \left [ \left (Y_{obs}(t) - Y(t)\right )^2 + \left ( X_{obs}(t) - X(t) \right )^2 \right ]
\end{align*}
Then  the estimate for $\rr$ is given by Equation \ref{eq:sirls},
\begin{align}\label{eq:sirls}
  \hat{R}_0= \frac{\hat{\beta}}{\hat{\gamma}}.
  \end{align}
We can use gradient descent or another optimizer to find estimates for these parameters.  If we assume $Y= f(t) + \epsilon_t$ with $\epsilon_t \overset{iid}{\sim} N(0, \sigma^2)$ where $f(t) = \int_0^t \frac{dY}{ds} ds$, then this this method is equivalent to finding the maximum likelihood estimator (MLE) of $\beta$ and $\gamma$.  However, we note that the the $\epsilon_t$ are certainly not independent from one another and so valid inference is still an issue in this framework.  Even if it were correct, ratio estimatiors typically result in large, uninformative confidence intervals.   The advantage is that this is a very straightforward way to estimate $\rr$ from the SIR model.

\subsubsection{Reparametrized Least Squares ($\rr$, $\gamma$) ( SIR ReLS)}\label{reparametrized-least-squares-rux5f0-gamma}

In the previous method, we estimated $\beta$ and $\gamma$ and then estimated $\rr$.  However, it is possible to directly estimate $\rr$ if we reparametrize the ODEs in Equation \eqref{eq::sir} directly with \(\rr\) and \(\gamma\), using the relation $\rr = \frac{\beta}{\gamma}$.

\begin{align*}
  \left \{
  \begin{array}{cl}
    \frac{dX}{dt} &= - \rr \gamma Y \frac{X}{N}\\
    \frac{dY}{dt} &=  \rr \gamma Y \frac{X}{N}  - \gamma Y \\
    \frac{dZ}{dt} &=  - \gamma Y 
  \end{array}
  \right .
  \end{align*}
We find
\begin{align*}
(\hat{R}_0, \hat{\gamma} ) &= \text{argmin}_{\beta, \gamma} \sum_{t} \left [ \left (Y_{obs}(t) - Y(t)\right )^2 + \left ( X_{obs}(t) - X(t) \right )^2 \right ]
\end{align*}
We use the \(\rr\) directly from the above estimation process, which again can be done with gradient descent or another optimization process.  Again, we cannot assume $\rr$ has a normal distribution as the errors are correlated.  This method is better than SIR LS for estimating $\rr$ as there is no ratio.

\subsubsection{Linear Model Approximation (SIR LMA)}\label{linear-model-approximation-degree-10}

The SIR ODEs have no known closed form solution, and so we are already using approximations using numerical integration, albeit small approximations.  In addition to this, data collected from real diseases are typically very noisy in the first place.  \cite{chang2017} discovered that an SIR model may be well approximated by a linear model.  We use this approach here to estimate $\rr$.

Specifically, we fit two linear polynomials in \(t\) with degree \(K= 10\) to \(X_{obs}\)
and \(Y_{obs}\) using least squares to find the coefficients $\{(\hat{x}_k,
\hat{y}_k)\}_{k=1, \dots, K}$,
\begin{align*}
\hat{X}(t) &= \sum_{k=0}^K \hat{x}_k t^k\\
{\hat{Y}}(t) &= \sum_{k=0}^K \hat{y}_k t^k
\end{align*}
Then, we estimate the derivatives as
\begin{align*}
\hat{X}^\prime(t) &= \sum_{k=1}^K k \hat{x}_k t^{k-1}\\
\hat{Y}^\prime(t) &= \sum_{k=0}^K k \hat{y}_k t^{k-1}
\end{align*}
Following,  \(\rr\) is derived from the ODEs in Equation \eqref{eq::sir},
\begin{align}
  - \frac{X^\prime}{Z^\prime}&= \rr \frac{X}{N} \nonumber\\
\rr &=       \frac{X}{ X + Y} \cdot \frac{N}{X} \nonumber\\
  \hat{\rr} &= \frac{\hat{X}^\prime(0)}{ \hat{X}^\prime(0) + \hat{Y}^\prime(0)} \cdot \frac{N}{\hat{X}(0)}. \nonumber
  \end{align}
Like in the previous cases, we are assuming the true models are
\begin{align*}
X(t) &= \sum_{k=0}^K x_k t^k + \epsilon_t\\
  {Y}(t) &= \sum_{k=0}^K y_k t^k + \epsilon_t,
\end{align*}
with $\epsilon_t \overset{iid}{\sim} N(0, \sigma^2)$.  Here, $K=10$ is rather arbitrary and should be selected using some criterion such as AIC in addition with the knowledge of how many data points are available to the user.  Besides optionally deciding on the degree of polynomials to fit, this model is simple to implement and gives very comparable results to using least squares with the SIR model.  The time $t=0$ to best capture the initial outbreak.  Again, we cannot use this model for inference, as our error assumption is violated.  Another disadvantage is that we are using the first time point $t=0$ to estimate $\rr$.

\subsubsection{Linear Model Approximation, All Time Points (SIR LMAT)}\label{linear-model-approximation-all-time-points-degree-10}

The above formulation of a linear model approximation only uses the estimate at time $t=0$ to estimate $\rr$.  We can instead, use all time points available to estimate $\rr$.  We fit a linear polynomial of \(t\) with degree \(K= 10\) model to \(X\)
and \(Y\) as above, with a slight modification in how we calculate
\(\rr\),
\[\rr = \frac{1}{\# \text{ Obs }t}\sum_t \frac{\hat{X}^\prime(t)}{ \hat{X}^\prime(t) + \hat{Y}^\prime(t)} \cdot \frac{N}{X(0)} \]
The intuition is that $\frac{X^\prime}{X^\prime + Y^\prime}$ is constant in $t$, but due to our approximations with the linear model, this is no longer the case.  Here, we average over the different possible values of $\rr$, estimated at different times.  An advantage to this approach is that we have a more robust estimate of $\rr$ than just using one time point.  The disadvantages are the same as before.

\subsubsection{Max of Data (SIR Max)}\label{max-of-data}
The I ($Y$) compartment of the SIR model has one peak.    We note that \(Y^\prime(t) = \frac{\beta XY}{N} - \gamma Y = 0\) when \(\rr = \frac{\beta}{\gamma} = \frac{N(t)}{X(t)}\). The max
of \(Y\) occurs when \(Y^\prime = 0\) and hence
\[\rr = \frac{N}{X_{obs}(t^*)},\] where
\(t^* = \text{arg max}_{t} Y_{obs}(t)\).   A major disadvantage to this method is that it only uses one sample, making it a highly variable statistic.  An advantage of this method is that it places no distributional assumptions on any parameter.

\subsubsection{Smooth Maximum of Data (SIR SMax)}\label{smooth-maximum-of-data}

We use a spline with 4 degrees of freedom to fit both \(X\), and \(Y\).
We then apply the principle as above, except now,
\[\rr = \frac{N}{\hat{X}(t^*)},\] where
\(t^* = \text{arg max}_{t} \hat{Y}(t)\).   This approach is a statistical approximation of the previous method of using the smoothed max to estimate $\rr$ instead of a single point.  In this way, we can utilize all our data.  Another advantage is the flexibility in model choice in that we can use a spline, linear regression, or any other model which gives us predictions for the number infected and number of susceptibles.  However, despite this flexibility in model choice, one must remember that we assumed an underlying SIR model for the derivation of $\rr$.  Also, it may be unclear how to estimate the variance from this method.

\subsubsection{Incidence to Prevalence
  Ratio (SIR IPR)}\label{incidence-to-prevalence-ratio}
The incidence to prevalence ratio (IPR), described by \cite{Nishiura2009} is another intuitive method to calculate $\rr$ as it incorporates some of the most basic epidemiological quantities, incidence and prevalence.

In terms of data from the SIR model, incidence $J(t) \approx -(X(t+1) - X(t))$, and the IPR$(t) = \frac{J(t)}{Y(t)}$. The \textit{actual} reproduction number, $R_a(t) = \text{IPR}(t)\cdot D$ where $D = 1 /\gamma$ the duration of the infection.  This method assumes that we have some prior knowledge about $\gamma$, the transmission rate.  Thus we use as our estimate,
\begin{align*}
\rr &= \textnormal{IPR}(t) \cdot D
\end{align*}

Here we assume that the time step is small enough to approximate the incidence.  The advantage of this method is that incidence data is generally readily available as is prevalence data for certain diseases such as HIV.  However, as one is required to have prior knowledge about $\gamma$, it may be easier to directly estimate $\rr$ with one of the many other methods described that does not require a prior knowledge about $\gamma$.  Again, we are using only one time point to estimate $\rr$.

\subsubsection{Smoothed Incidence to Prevalence Ratio (SIR SIPR)}
We use the same method as above, IPR, but first fit splines with 4 degrees of freedom, $\hat{X}(t)$ and $\hat{Y}(t)$, , to fit to $X_{obs}$ and $Y_{obs}$, respectively.  Then  $J(t) \approx -(\hat{X}(t+1) - \hat{X}(t))$, and the $\hat{\textnormal{IPR}}(t) = \frac{J(t)}{\hat{Y}(t)}$.  Then
\begin{align*}
\rr &= \hat{\textnormal{IPR}}(t) \cdot D
\end{align*}
The advantage of this method is that it creates a less noisy estimate than estimating IPR using only one point.  It has the same disadvantages as the regular IPR ratio in that it requires knowledge about $\gamma$.

%%%%%%%%%%%%%%%%%%%%%
\subsubsection{Log-Linear (SIR LL)}
Recently, \cite{harko2014exact} were able to reduce the SIR model to one ODE.  From this, we can derive the following,
\begin{align}
  X(t) &=  X(0) e^{\frac{\beta}{N \gamma}Z(t)} \nonumber\\
  \log \frac{X(t)}{X(0)} &=  \frac{\beta }{\gamma N }Z(t) \nonumber\\
  N\log \frac{X(t)}{X(0)} &=  \rr Z(t) \label{eq:harko_lin}
\end{align}
If we add error into Equation (\ref{eq:harko_lin}), then we have
\begin{align}
N  \log \frac{X(t)}{X(0)} &=  \rr Z  + \epsilon_t\label{eq:r0_harko}
\end{align}
If we assume $\epsilon_t \overset{iid}{\sim}N(0, \sigma^2)$, then we can directly estimate $\rr$ via least squares linear regression.  This method greatly simplifies the computational effort needed to estimate $\rr$ as it does not require any numerical integration.  Another advantage is that unlike using the linear approximation method, the coefficient from this model is directly interpretable.  In words, $\rr$ means that for one more recovery, we expect $N \log \frac{X(t)}{X(0)}$ to increase by $\rr$.  A disadvantage is that again, $\epsilon_t$ are not likely iid.  Another disadvantage is that this does not generalize to more complicated CMs.

\subsubsection{Markov chain estimation (SIR MC)}
A natural approach to epidemic modeling is that of Markov chains (MC), since it is assumed an individual's next state is only dependent on its current state.  Much work has been done over the years in this specific field including asymptotic behavior, continuous time MC, confidence intervals, and more \citep{jacquez1991,gani1995,daley2001epidemic}.  We present one simple instantiation of the model, the discrete time case, which traces its origin back to the Reed-Frost model \citep{abbey1952}.

In this framework, the number of susceptibles at the next step, $X_{t}$, has a Binomial distribution based on the contacts with the current number of infected, $Y_{t}$ and the current number of susceptibles.  That is $X_{t+1} \sim \text{Binom}(X_t, \alpha^{Y_t}$), where $\alpha$ is the probability of avoiding infection from an infective.  \cite{barbour2004} report that the reproduction number is thus,
\begin{align}\label{eq:r0-mc}
\rr &= \log \left ( \frac{1}{1-\alpha}\right ).
\end{align}

Thus, using regression on susceptible/infection counts will lead to an estimate of $\alpha$ and hence $\rr$.  However, this framework typically allows more than just the reproduction number to be estimated.  Through recursion, one can calculate the probability of having a given number of susceptibles and infected at each time step, and hence the entire distribution may be known.

\subsubsection{Sequential Bayes (SIR SB)}\label{sec:seqbayes}

Described by \cite{bettencourt2008} and summarized in \cite{obadia2012r0}, the sequential Bayes method is a Bayesian approach to an approximation of the classic SIR model.  The approximated SIR model assumes that the incidence at $t+1$, $J(t+1)$ has a Poisson distribution, with $\gamma$ as the  average inverse of the infectious period and $\mathcal{R}$ as the effective reproduction number, which we take here to be $\rr$. In order to estimate $\rr$, we must have some idea about $\gamma$,
\begin{align*}
J(t+1)  \sim Poisson( J(t) \exp \left \{  \gamma (\rr-1)\right \})
\end{align*}
Then, the posterior distribution of $\rr$ given the previous days' incidences is
\begin{align*}
  P(\rr | J_0, \dots, J_{t+1}) = \frac{P(J_{t+1} | \rr, J_0, \dots, J_t)P(\rr| J_0, \dots, J_t)}{P(J_0, \dots, J_{t+1})}.
\end{align*}
This method is sequential in that the prior distribution for $\rr$ comes from the previous day.  The initial prior for $\rr$ is assumed to be flat.  This method results in a posterior distribution from which credible intervals may be obtained.  This method assumes, initial growth in incidence to be exponential, and homogeneous mixing of populations as with any compartment model.  The advantages of this method are that of the ability to obtain an entire posterior distribution, whereas many other methods are difficult to even find an estimate of the variance.  Disadvantages include strict assumptions about the distributions and computational time required to estimate the relevant parameters.

\subsubsection{SEIR Model (SEIR)}
\label{sec:seir-model}

The SEIR model is a common adaptation, especially in recent years, of the SIR compartment model used to model infectious diseases such as influenza and Ebola \citep{mills2004,althaus2014}.  The ``E'' compartment stands for ``exposed'' and represents the stage where individuals are infected but not yet infectious.  The model is described by the following set of equations, as shown by \cite{cintronarias2009}.  This particular model  does not include changse in population size but can be adapted to do so,
\begin{align*}
  \frac{dX}{dt} &= - \frac{\beta XY}{N} \\
  \frac{dE}{dt} &= \frac{\beta XY}{N}  - \alpha E\\
  \frac{dY}{dt} &= \alpha E - \gamma Y \\
  \frac{dZ}{dt} &= \gamma Y.
\end{align*}

Then $\rr = \beta / \gamma$, just as in the SIR model.  As in the SIR model, $\beta$ and $\gamma$ have the same interpretation and additionally, $\alpha$ is the rate at which exposed individuals become infectious.  The methods used to estimate the parameters $\beta$ and $\gamma$ are of the same types used to estimate those parameters in the SIR model such as least squares and maximum likelihood estimation.  Here, we see a very strong example of how $\rr$ is a property of the model and why it is important for this model to be correct.  An advantage of this model is that the addition of an exposed class is thought to better mimic reality of the transmission of many diseases.  A major disadvantage is that exposure counts are typically back-estimated from incidence counts, and noise is rarely considered in this estimation of the exposure counts.   Typically, some form of maximum likelihood estimation is used to estimate parameters in this model, which places strict assumptions on each of the compartments \citep{althaus2014}.

\subsubsection{MSEIR model (MSEIR)}
\label{sec:mseir-model-mseir}

\begin{figure}[h]
\centering
\begin{tikzpicture}[node distance=3cm,auto,>=latex',every node/.append style={align=center}]
  \node [int,  fill = white!70!blue] (s)              {$S$};
  \node[int, fill = white!70!blue] (m) [left of=s] {$M$};
  \node [int,  fill = white!70!blue] (e) [right of=s]             {$E$};
  \node [int,  fill = white!70!blue]           (i) [right of=e] {$I$};
  \node [int,  fill = white!70!blue] (r) [right of=i] {$R$};
  \path[->, auto=false] (s) edge node {} (e)
  (e) edge node {} (i)
  (i) edge node {} (r)
  (m) edge node {} (s)
  ++ (0,2) edge node {} (m)
  (s) ++ (0,2) edge node {} (s);
  \path[->]   (m) edge node {}  ++ (0,-2);
  \path[->]   (s) edge node {}  ++ (0,-2);
  \path[->]   (e) edge node {}  ++ (0,-2);
  \path[->]   (i) edge node {}  ++ (0,-2);
  \path[->]   (r) edge node {}  ++ (0,-2);
\end{tikzpicture}
\caption{Depiction of a MSEIR model with birth and death rates.  $M$ - passive immunity from mother; $S$ - susceptible; $E$- exposed, latent; $I$- infected; $R$ - recovered.  An arrow out of the system indicates death and an arrow into the system indicates birth.}\label{fig::mseir}
\end{figure}

This compartment model approach includes not only the exposed, latent compartment $E$ but additionally the compartment $M$ which stands for passive immunity inherited from the mother.  \cite{hethcote2000} describes the model with the following set of ODES that include birth rate $b$ and death rate $d$, respectively (reparametrized to match our notation),
\begin{align*}
  \frac{dM}{dt} &= b(N-X) - (\delta + d) M \\
  \frac{dX}{dt} &= bX + \delta M - \frac{\beta XY}{N} - dX\\
  \frac{dE}{dt} &= \frac{\beta XY}{N} - (\alpha + d) E\\ 
  \frac{dY}{dt} &= \alpha E - (\gamma + d) Y \\
  \frac{dZ}{dt} &= \gamma Y - dZ \\
  \frac{dN}{dt} &= (b-d)N,
\end{align*}
where $\delta$ is the rate of passive, immune infants transitioning to susceptible indnividuals, $\beta$ is the rate of infection, $\alpha$ is the rate of of infectious individuals becoming infected, and $\gamma$ is the rate of recovery.  The reproduction number is then,
\begin{align}\label{eq:r0-mseir}
\rr &= \frac{ \beta \alpha}{ (\gamma + b)(\alpha + b)}.
\end{align}
\cite{hethcote2000} note that this calculation of $\rr$ may be strictly derived through interpretation of the definition, as ``$\rr$ is the product of the contact rate $\beta$ per unit time, the average infectious period adjusted for population growth of $\frac{1}{\gamma + b}$ and the fraction $\frac{\alpha}{\alpha + b}$ of exposed people surviving the latent class $E$.''  They note that this model is well suited for diseases such as measles, rubella, or mumps where the mother may pass some antibodies to her newborn.  This expression may also be derived using the next generation operator described in Section \ref{sec:ngm}.

Again, this is simply a model for transmission of a disease and does not specify how $\rr$ is to be estimated once the model is written down.  Method such as least squares and MLE are common ways to estimate the parameters in this model and hence $\rr$.  When the birth rate $b=0$, note that $\rr$ reduces to the derived expression for both the SIR and SEIR models without birth and death rates, $\frac{\beta}{\gamma}$.

\subsubsection{Amplifier Model (AMP)}
As mentioned previously, compartment models may be arbitrarily complex.  We demonstrate this idea with a set of compartments for each strain of a disease like influenza or HIV.  \cite{blower2004} discuss the case where a disease may have different strains and consequently a different reproduction number for each strain.  They introduce a multistrain mathematical model called the amplifier model to deal with this case.

This model is based on the the SEIR compartment model, accounting for a number of strains simultaneously.  There is a set of ODEs for each strain of the disease with different parameters for recovery, transmission rate, and drug resistance of the disease.  Additionally, there is mixing of the different strains as ``the amplifier model also allows for immigrating and emigration of individuals of all types, as well as reinfection of latently infected individuals'' \citep{blower2004}.  Ultimately, their formula for $\rr^{(i)}$ where $i$ stands for the $i$th strain is
\begin{align*}
\rr^{(i)} = X^* \frac{ ( \beta_i^Y + \beta_i^E)\alpha_i + \beta_i^Y d_i^E}{(\alpha_i + d_i^E)(c_i + k_{i,i+1} + d_i^Y)}.
\end{align*}
Here, $X^*$ is the number of susceptibles individuals in the disease-free equilibrium in the absence of immigration or emigration; $\beta_i^E$ and $\beta_i^I$ are strain-specific transmission rates for the latent, and infected classes, respectively; $\alpha_i$ denotes the progression rate from latently infected to disease for strain $i$; $d_i^E$ and $d_i^{Y}$ are death rates for the latent and infected classes, respectively; $c_i$ is the cure rate; and $k_{i, i+1}$ is a parameter associated with the degree of amplification of drug resistance.  They note that ``each reproduction number is the product of average number of secondary infections caused per unit time, the average time a case remains infectious and the probability that an infected individual develops disease.''

In the amplifier model, $\rr$ was derived from using the stability point (disease free equilibrium state in the absence of immigration or emigration) of the system by solving the characteristic polynomial for the eigenvalues.  This derivation is explained in more depth in Section \ref{sec:ngm}.

We describe this model as opposed to other available CMs because it shows not only how complex CMs can become, but also to demonstrate how researchers approach the often controversial assumption of homogeneity.  In this case, we have strain-specific transmission rates, so we are essentially stratifying compartments into different groups, allowing for some heterogeneity.  A disadvantage is the amount of strain-specific data required to estimate all parameters, which may be estimated through the usual methods such as least squares and maximum likelihood estimation.

\subsubsection{Next Generation Model (NGM)}
\label{sec:ngm}

Very complicated models have arisen from stratification of compartments such as one with 26 compartments \citep{pandey2014}.  Thus, there is a need to be able to derive an expression for $\rr$ regardless of the structure of the compartment model.   The Next Generation Model (NGM) is a generalization of any compartment model at the infection-free steady state. This model solves the problem of having an expression for $\rr$ in terms of the epidemiological parameters for a general compartment model.  Originally, introduced by \citep{diekmann1990}, \cite{diekmann2009} posit a recipe to find $\rr$ for a wide class of compartment models, including models we discussed, SIR, SEIR, and MSEIR. We summarize that recipe here. First, define the infected subsystem as ``the equations of the ODE system that describe the production of new infections and changes in state among infected individuals.''  Let $x = (C_1, C_2, \dots, C_m)$ where $C_i$ are the different compartments of the infected subsystem.  The steps to find $\rr$ are as follows:

\begin{enumerate}
\item ``Linearize the infected subsystem of nonlinear ODEs about the infection-free steady state''
\item Decompose the linearized infected subsystem into $(T + \Sigma )x$ where $T$ is a $m\times m$ matrix of transmissions and $\Sigma$ is a $m \times m$ matrix of transitions.
\item $\rr$ is the spectral radius (i.e. the dominant eigenvalue) of $K_L=-T \Sigma^{-1}$.  
\end{enumerate}

Here, $T_{ij}$ is the rate of transmission of \textit{newly} infected individuals in state $i$ created by individuals in state $j$.  $\Sigma_{ij}$ is the transition rate of individuals into compartment $i$ from compartment $j$.

This method has the assumptions of a basic compartment model: homogeneous mixing among populations and the law of mass action, i.e., that the number of infected individuals are proportional to the number of infected individuals at the previous step.   Additionally, this method is advantageous in that the matrix $-\Sigma^{-1}$ has intuitive meaning.  The $ij$th entry of $- \Sigma^{-1}$ is the ``expected time that an individual who presently has state $j$ will spend in infected state $i$.''

Although, this method is useful for deriving an expression for $\rr$, it is unclear how to actually estimate it from these expressions.  One could use maximum likelihood if one specifies distributions for how compartments change over time or least squares may be applied. In general, adding numerous compartments is often tedious and difficult and needlessly complex.  Instead, researchers have started to lean towards ABMs when introducing heterogeneity in their models, which we describe in Section \ref{sec:network}.

%%%%%%%%%%%%%%%%%%%%%%%%

\subsection{Exponential Growth}\label{sec:exp-growth}
Typically, CMs expressed through ordinary non-linear differential equations have no known analytical solutions, which makes deriving a closed-form expression for $\rr$ difficult or impossible.  One simplification that is often introduced to the SI-framework is that of exponential growth, or that the number of infections in the beginning of an epidemic mimics that of an exponential curve.  This simplified expression has allowed for many derivations of $\rr$.

\subsubsection{Exponential Growth (EG)}
\label{sec:expgrowth}
\cite{wallinga2007generation} report that the effective reproduction number $\mathcal{R}_t$ and hence the initial reproduction number $\rr$ may derived using the fact that infection ``counts increase exponentially in the initial phase of an epidemic.''  We then have to estimate $r$, the \textit{per capita} change in the number of new cases per unit of time and $\omega$ the serial interval, the distribution of time between a primary and secondary infection. Then, we have
\begin{align}\label{eq:lotka}
\rr = \exp{(r \omega)}
\end{align}
or its first order approximation
\begin{align}\label{eq:anderson}
\rr = 1 + r \omega.
\end{align}
Equation \eqref{eq:lotka} is derived from a demographic view using the Lotka-Euler survival equations which come from the fields of demography, ecology, and evolutionary biology, whereas Equation \eqref{eq:anderson} is derived through an epidemiologist's perspective.

\cite{wallinga2007generation}, instead, use a moment generating function expression for $\rr$.  With $\omega(t)$ as the serial interval, then
\begin{align*}
\rr^{-1} &= \int_{a=0}^\infty e^{-rt}\omega(t)dt.
\end{align*}
For all these above estimates of $\rr$, we observe the duration of serial intervals in a period of exponential growth.  Deciding when exponential growth occurs and which data points to use to estimate $\rr$ is difficult.  \citeauthor{wallinga2007generation} specify $\omega$ in multiple ways.  Parametrically, they use exponential, normal, and point-wise distributions.  Additionally, they discuss using the the empirical distribution of the serial interval.

\citeauthor{wallinga2007generation} note that specifying the epidemic model ``implicitly specifies a generation interval distribution.''  The above equations are under the SIR framework, but this method may be adapted to acommodate other frameworks.

\subsubsection{Maximum Likelihood Estimator of Secondary Infections (MLSI)}\label{sec:mle-si}
A common method used to estimate parameters in the SI-framework is maximum likelihood estimation.  This method requires assuming the likelihood of a parameter given the data observed, and hence many distributions may be specified.  The MLE of secondary infections, described by \cite{forsberg2008}, finds the estimate of $\rr$ by maximizing the likelihood function under a certain assumptions.  We assume that the ``the number of secondary cases produced by an infected individual follows a Poisson distribution, and that the serial interval is described by a multinomial distribution.''  Recall, the serial interval is the distribution of time between a primary case developing symptoms and a case infected by the primary case developing symptoms.  An approximate version of this can be simplified to a thinned Poisson  where
\begin{align}\label{eq:mlesi}
  L(\rr, \mathbf{p}) = \prod_{t=1}^T \frac{e^{- \rr \sum_{j=1}^{\min(k,t)}J_{t-j}p_j}\left (\rr \sum_{j=1}^{\min(k,t)}J_{t-j}p_j \right )^{J_t}}{J_t!}.
\end{align}
Here, $k$ is the maximal amount of the serial number (the maximum amount of time between infections), $T$ is the total time and $p_j$ is the probability of displaying symptoms on day $j$ after being infected.  The number of cases on day $t$, $J_t$ are observed incidence counts.

This method assumes that the number of new cases caused by an infected individual follows a Poisson distribution with mean $\rr$ and that hte serial interval follows a multinomial distribution. Additionally, this method also assumes a period of exponential growth.  An advantage of this specification is that we can write down a nice likelihood for $\rr$. Two disadvantages are that we only have an approximation for $\rr$ and have to decide which data points to use in a certain time frame.  Additionally, we have to validate our assumptions about the specified distributions.  This method was used to estimate $\rr$ for Ebola, Avian Influenza, and Swine Flu.

%%%%%%%%%%%%% 5

\subsubsection{Time Dependent Reproduction Number (TDR)}\label{sec:timedep}
Another likelihood based approach to estimate $\rr$ is that of the time dependent reproduction number, as shown by \cite{wallinga2007generation} and summarized by \cite{obadia2012r0}.  They maximize the likelihood that case $i$ has been infected by case $j$ at certain time using the serial interval, $w(t)$ where $t_i$ is when case $i$ was infected and $N$ is the total number of cases,
\begin{align*}
  p_{ij} &= w(t_i- t_j) / \sum_{i \neq k} w(t_i - t_k),\\
  \rr &= \frac{1}{Y(0)}\sum_{i=1}^N p_{i0}
  \end{align*}
  A disadvantage of this method is that it assumed we have knowledge about the serial interval distribution, $w(t)$, but an advantage is that the method works with any specification of $\omega$.  This method also requires that the initial number of susceptibles $X(0)$ is roughly the size of the total population.  Another advantage is the explicit use of distributions to indicate that $\rr$ is indeed a random variable.  This method was used to estimate $\rr$ for Influenza A.

\subsubsection{Initial Growth Rate and Final Size (IGRFS)}
\label{sec:igr-fs}

Initial growth rate, as described by \cite{dietz1993estimation} is derived by \cite{anderson1986}, initially to study the spread of HIV.  Initial growth rate estimates $\rr$ through
\begin{align*}
\rr = \frac{D \ln 2} {t_d} + 1,
  \end{align*}
  with $D$ as the average incubation period and $t_d$ as the doubling time during its early stages.  This method assumes one knows when the ``early stage'' ends along with knowledge of the average incubation period.

  Final size, on the other hand, only looks at $\rr$ once a disease has run its course.  Here $\mu_0$ is the proportion of initially infected individuals and $\mu_\infty$ is the final, cumulative proportion of infected cases at the end of an epidemic,
\begin{align}\label{r0_attackrate}
\rr =  \frac{\log \left ( \frac{X(0)}{\mu_{\infty} N}\right )}{\mu_\infty - \frac{X(0)}{N}}
\end{align}
This method can only be used to assess $\rr$ after a disease has passed through a population.  Additionally, if either $\beta$ or $\gamma$ were affected during the disease's lifetime, such as through interventions and prevention strategies, then $\rr$ cannot be properly assessed.

  These two methods allow one to look at $\rr$ at both the beginning and the end of an epidemic.   For the initial growth rate, while one can see $\rr$ at early stages, estimating the doubling time $t_d$ is difficult.  For final size, one has to wait until the end of an epidemic to calculate $\rr$, and prevention strategies may skew the final result.  These methods were used to estimate $\rr$ for AIDS and smallpox.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

  \subsection{Estimation from Networks}
As estimating $\rr$ is concerned with the number of new infections based on the contacts of old infections, a network is a seemingly intuitive way to model such contacts and hence infections.  Networks that have been used to estimate $\rr$ include contact tracing of actual individuals, branching processes to model the spread of disease, and agent-based models (also known as individual level models), which all leverage networks along with different sources of data to generate estimates of $\rr$.  Additionally, networks provide an intuitive framework in which individuals can interact with varying levels of heterogeneity.  However, when using a network approach, we must be careful of our interpretation of $\rr$ as we often are conditioning on certain individuals.    \cite{breban2007} note that calculating $\rr$ from a network approach may give misleading results due to conditioning on the network structure.

\label{sec:network}

%%%%%%%% 5
\subsubsection{Contact Tracing (CT)}
\label{sec:contact_tracing}
A reasonable idea to estimate $\rr$ is to directly observe the number of secondary infections produced by an initial infector.  However, this approach is generally impractical as it is difficult to estimate every single contact an individual may have.   With seriological information, we could conceivably recreate the transmission of the disease throughout a population.  Even in this case it is unlikely, for one, to have a completely susceptible population, which is essential to an unbiased estimate of $\rr$.  Moreover, this would implicitly be conditioning on the initial infector's contacts which may not reflect those of the rest of the population.  We write the calculation for $\rr$ under contact tracing in Equation \ref{eq:r0_contacttracing}, where $X_0$ is a random variable denoting the number of individuals in the population individual 0 infects.  Here $X(0)=N$ indicates that the initial number of susceptibles is $N$, the size of the population,
\begin{align}\label{eq:r0_contacttracing}
\rr = E[ X_0 | X(0) = N].
\end{align}
\cite{eames2003} describe contact tracing as an ``extreme form of targeted control, where the potential next-generation cases are the primary focus.'' That is, one focuses on treating the contacts of an infected individual rather than applying uniform treatment across a population.

They describe an estimate for $\rr$ for a SIR model on a network, with $r$ as the rate of transmission across a contact multiplied by the infectious period and $n$ as the number of contacts as
\begin{align*}
\rr = r(n-2)
\end{align*}
Again, the difficulties of this method lies in one, estimating the contacts of an initial infected individual and two, extrapolating the contacts of the initial individual to the population as a whole.  As such, although important in treating an outbreak of a disease in its early stages, is not the best choice to estimate $\rr$ due to the high level of uncertainty inherit in the method and due to conditional assumptions of the initially infected individual.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\subsubsection{Branching Process (BP)}
\label{sec:branching-process}

Branching processes are models for population growth which track generations of new offspring.  We describe branching processes as adapted for epidemic modelling from \cite{grimmett1992}.  An infector $i$ may produce new infections, called offspring, over the duration of the  infector $i$'s disease.  The offspring of the original infector directly produces are called the first generation.  The offspring of the first generation are called the second generation and so on.  This process is illustrated in Figure \ref{fig:bp}.  An important assumption in branching processes is that the offspring produced from different infectors are indepedent and identically distributed, and so for a branching process to be a valid model in epidemic modeling, we must assume a large population with few infectors.

\begin{figure}
  \centering
\begin{tikzpicture}
  [
    grow                    = right,
    sibling distance        = 4em,
    level distance          = 10em,
    edge from parent/.style = {draw, -latex},
    every node/.style       = {font=\footnotesize},
    sloped
  ]
  \node {Case 0}
  child { node {Gen. 1}
    child {node{ $\dots$}}
  }
  child {node {Gen. 1}
    child{ node  {$\dots$}}
  }
  child{ node  {Gen. 1 }
    child{ node  {Gen. 2}
    }
    child{ node {Gen 2}
      child{ node  {$\vdots$}}
      child{ node  {$\vdots$}}
      child{ node  {$\vdots$}}}
  };
\end{tikzpicture}
\caption{Illustration of a branching process.}\label{fig:bp}
\end{figure}

\cite{getz2006} give a way to calculate $\rr$ from a branching process.  ``We define the offspring distribution $\{q_i \}_{i=0}^\infty$, where $q_i$  is the probability that an infectious individual infects $i$ other individuals.  Thus we require $\sum_{i=0}^\infty q_i =1$ and note that $\rr$, the mean number of cases contracting disease from each infective'', is simply given by'
  \begin{align*}
    \rr = \sum_{i=0}^\infty iq_i.
  \end{align*}
  It is assumed we know how to estimate the $q_i$, and so in this case, it is not too different of an approach than contact tracing.  The disadvantages of this method are similar to that of contact tracing as data about which individuals  infected another is very difficult to obtain.  An advantage is that this method directly gets to the ``heart'' of the definition of $\rr$ as it is concerned with generations of infected individuals rather than incidence at a certain time.

\subsubsection{Agent-Based Models (ABM)}
\label{sec:agent-based-models}
Agent-based models (ABMs) or individual level models (ILMs) are ``bottom-up'' models, meaning micro patterns are simulated and macro patterns are then inferred from these micro patterns.  ABMs consist of agents which represent individuals (e.g., people, mosquitoes, poultry) and a series of activities in which the agents can affect one another and evolve over time.

In epidemiology, ABMs may be used to simulate the spread of disease.  ABMs allow for every detail of an infection to be known, in the context of estimating $\rr$, meaning we know exactly who is infected by whom and when.  Let $L$ be the number of simulations run and $X_{n \ell}$ be the number of infections caused by agent $n = 1, 2, \dots, N$ who was infected at time $t=0$ in simulation $\ell$.   Thus $\rr$ can be estimated simply through
\begin{align*}
  \rr =& \frac{1}{LN}\sum_{\ell =1}^L\sum_{n=1}^NX_{n \ell}
\end{align*}
Examples of estimations of $\rr$ using ABMs/ILMs are in \cite{breban2007,ahmed2013variance}.

There are many advantages to using ABMs, a main one being able to run the model multiple times under the same initial conditions.  In this way, one can easily obtain variance estimates of $\rr$ in addition to empirical distributions of it.  Moreover, this estimation can accommodate any model specification of how the disease is transmitted through the population.  However, the major disadvantage of the ABM lies in calibrating the evolving of agents over time to accurately reflect reality, a problem that cannot be understated.  Thus, it is typically necessary to have knowledge of parameters such as the serial interval, transmission rate, recovery rate, etc.  Additionally, we need to have accurate agents in characteristics and their implied network structure.  ABMs are flexible in that they may be used with any model in the SI framework.  The only requirement is that we track who infects whom and can adapt for any additional states such as exposure or prevention strategies such as closing schools \citep{fred}.

\subsection{Variance Methods}
\label{sec:methods}

Estimating $\rr$ is difficult and estimating $V[\rr]$, the variance, is even more so.  We describe general methods which may be applicable to estimate the variance.  We describe the assumptions that are used along with each method and note that they may not be applicable to all of our methods.  Still, reporting confidence intervals is very important, and we hope to encourage this practice by describing the below methods used to estimate models in the SIR framework in a simple fashion.  Regardless of the method used to estimate variance, confidence intervals, or credible intervals, we encourage researchers to carefully examine all underlying assumptions of 1) specifying a model for disease transmission, 2) estimating $\rr$ from the model, and 3) the method used to estimate the variance or CIs.

\subsubsection{Delta Method}\label{delta-method}

When the method estimates \(\beta\) and \(\gamma\) instead of \(\rr\) directly, we use the delta method approximation to calculate the
variance of \(\rr\). Here, we know \(\rr = h(\beta, \gamma) = \frac{\beta}{\gamma}\). Then \(\bigtriangleup h = (\frac{1}{\gamma},  -\frac{\beta}{\gamma^2})^T\) and \(V[\rr] = \bigtriangleup h^T \Sigma_{\beta, \gamma} \bigtriangleup h\), where \(\Sigma_{\beta, \gamma}\) is covariance matrix of \(\beta\) and \(\gamma\).

This estimate assumes that the distribution of $\beta$ and $\gamma$ are asymptotically normal.  Here, we use the relationship of $\rr = \frac{\beta}{\gamma}$ in order to estimate the variance, which is specific to the SIR framework.  This method may be extended to other frameworks.  The advantages of the delta method are that as the NGM posits a recipe to derive $\rr$, then it is theoretically possible to use the delta method to derive an expression for the variance of $\rr$ for any compartment model.

\subsubsection{Posterior Distribution}
When one uses a Bayesian approach to estimate $\rr$, the result should be a distribution rather than an estimate.  One can simply look at the quantities of this distribution to form a credible interval for this estimate.  Advantages of this method include having an entire distribution compared to a point estimate.  One disadvantage is that the Bayesian approach will not guarantee frequentist properties that epidemiologists may be looking for such as being able to compare diseases to one another.

%%%%%%%%%%%%%%%

\section{Analysis and Results}
\label{sec:results}

\subsection{Comparing Models}

\textit{We compare the different ways to estimate $\rr$ to see which ones give more accurate and reliable estimates.}  As there is no way to know the ``true'' $\rr$ of an actual disease we compare our methods by fitting them to simulated data sets and look at the resulting $\rr$ estimates and standard errors. Our main comparisons are done with data generated from the SIR model with random error added so that we can compare model estimates to a known true $\rr$ value. We vary the conditions under which the SIR data is generated to see how the results from the estimators change. We also look at data from different functional models to see how the estimators perform under different circumstances.

%% What do you mean by describing the standard standard deviations?
%% SG: delta_t is normally distributed with sd sigma and similarly for other methods.  How and why did we pick the numbers that we did?

\subsubsection{SIR Data}

\textit{The datasets from the SIR model are simulated from the ODEs: \begin{align}
	\systeme{\frac{dX}{dt} = -\frac{\beta XY}{N}, \frac{dY}{dt} = \frac{\beta XY}{N} - \gamma Y, \frac{dZ}{dt} = \gamma Z}.
	\end{align}
}

We use arbitrarily selected the parameters to generate the time series data set from the SIR model. We first generate "baseline" data from which we can make comparisons by changing the  conditions under which the data were generated. The baseline SIR data are simulated with the following conditions:

\begin{center}

	$\beta$ = 0.006, $\gamma$ = 0.003; $\rr$ = 2

	Number of time steps: 365

	Starting X size ($X_0$): 99950

	Starting Y size ($Y_0$): 50

	Starting Z size: 0

	Total population: 100000.

\end{center}

%% Is this enough detail on describing the initial S, I, and R, or are we looking for something more specific?

\textbf{Our baseline initial size for $Y_0$ is 0.05\% of the total population in our simulated environment, which is fairly large compared to the starting point for the data in most epidemics. However, I/we felt that this was necessary in order to see the differences between the models.}

We then generate random errors that are added to the X and Y compartments at each time step in order to simulate some form of randomness. \textbf{The sum of these was then subtracted from the Z compartment in order to keep the total population in the system constant.} We generated data sets with two different types of errors: Gaussian and autoregressive. \textbf{The Gaussian errors are independently distributed with mean 0 and constant variance for the entire length of the time series. The autoregressive errors are generated from a first-order process of the form: $$ \epsilon_{t} = \rho * \epsilon_{t-1} + \delta_{t}$$, where $\rho$ is fixed throughout and $\delta_{t} \sim N(0, \sigma)$ with constant $\sigma$. The values of the variances are chosen somewhat arbitrarily, although they are intended to be fairly small at first.}

We also add a monotonicity adjustment to the X and Z compartments with both of the error models. \textbf{This is done using the Pool Adjacent Violators Algorithm (PAVA); the algorithm checks if the observation at each time step is greater than or equal to the previous one, and changes the value of both observations to be equal to the mean of the two if the condition is not fulfilled.} This gives us a total of four error models: autoregressive (Auto), autoregressive monotonic (AutoM), Gaussian (Norm), and Gaussian monotonic (NormM). For our baseline data, we have ($\sigma_X, \sigma_Y$) = (10, 1). We simulate 5 different baseline sets to see how consistent each of the models are.

\textit{To see the effects of different conditions for the data, we change one of the starting conditions while holding all of the other ones constant. We try changing two conditions: the starting number of susceptible and infected people in the population (holding the total population constant) and the variance of the errors added on to the The starting population sizes for the $X$ and $Y$ are set to ($X_0$, $Y_0$) =  (99999, 1), (99990, 10), (99900, 100), and (99000, 1000), holding all other conditions constant. The variance of the added errors by scaling them to: ($\sigma_X$, $\sigma_Y$) = (10, 1), (100, 10), (500, 50), (1000, 100), (2000, 200), (5000, 500), (10000, 1000), (15000, 1500); these are done with ($X_0$, $Y_0$) = (99000, 1000). The errors start very small but become very large relative to the actual compartment sizes; this allows us to more clearly see the effects of the changes.}

%% Describe the initial S, I, R values
%% Describe the different error methods
%% Describe the standard standard deviations involved
%% High level fitting comments

%% Throw fitting details into the appendix

%% Talk about top 3-5 methods
%%%%%% Honestly, all the methods seem to have flaws in one way or another - maybe point out why?
%% Talk about methods that really miss
%% Talk about the standard error reports
%% Talk about 4 categories.  Does anyone do worse/better?
%% What about the one where we cannot estimate std
%% What happens when SD is large?

%% Throw some graphs in

\subsubsection{Other Models}
We simulated data from 6 other models besides the classic SIR model. 5 of these were polynomial linear regression models, from order 0 to 4 with respect to time/day. The coefficients used to generate data from the linear regression models were originally chosen by taking the approximate coefficient estimates from fitting the linear regression models to classic SIR data. This does not necessarily guarantee that the true $\rr$ will be 1.2, especially for the 0th and 1st order linear regression models. What we are interested in for these cases is to see if the models still give reasonable estimates for $\rr$. The other model is the linear SIR model, which is a three compartment model with the following ODEs:

	\begin{eqnarray*}
		\frac{dS}{dt} &=& -\beta I \\
		\frac{dI}{dt} &=& \beta I - \gamma I \\
		\frac{dR}{dt} &=& \gamma I.
	\end{eqnarray*}

 The data from the linear SIR model were simulated with parameters $\beta$ = 0.06 and $\gamma$ = 0.05, such that $\rr$ = 1.2. This model will allow us to compare how the SIR models perform under misspecified differential equations.

 \textbf{Random error is added to the data generated here as well using the same 4 error models as above. The other specifications for these datasets are the same as for the baseline classic SIR data, namely the variance of the added errors and starting compartment sizes.}

\subsubsection{Variance Estimation}

\textbf{The method for calculating the standard errors for $\hat{\rr}$ depend on the model used. The following models take the standard error directly from the regression methods used to estimate $\rr$:}

	\begin{itemize}
		\item SIR LMA
		\item SIR LMAT
		\item SIR Log Linear
		\item SIR MC
		\item Exponential Growth
		\item Initial Growth
	\end{itemize}

\textbf{The ones that are calculated using the delta method:}

	\begin{itemize}
		\item-SIR LS
		\item SIR ReLS
		\item SEIR
		\item MSEIR
		%% \item AMP
		\item NGM
		\item Final Size
	\end{itemize}

Standard errors for the maximum likelihood and branching process models are calculated using the negative reciprocal of the Fisher information. The sequential Bayes model uses the standard deviation from the posterior distribution. We use the jackknife method to estimate the standard error of the SIR smoothed max and the SIR smoothed IPR methods. The jackknife calculates the standard error using the standard deviation of estimates of $\rr$ after leaving out a single data point.

Some of the models use only one single point in time to estimate $\rr$ and therefore do not seem to have an innate source of variance. We currently do not have standard errors for the SIR Max and SIR IPR models.

\subsection{Results}

\subsubsection{Survival Function}

\textbf{STILL NEED TO ACTUALLY DO DPE}

We do not have a direct way to estimate $/rr$ using the survival function method from the data we have. The method requires a density estimate of the survival function at each point in time. Either we would need to have some prior knowledge or assumptions about what the survival function for a particular disease was, or we would need to search over a space of possible function and possible parameter values to estimate $/rr$. For similar reasons, we can not estimate $/rr$ using the Von Forester equations using the data we have, as we would need to know the infectivity of a diseased individual over time. This may be possible for a known, well studied disease, but may not be  For the direct parameter estimation method, assuming the population is homogeneous and that we know the average duration of infection, we can estimate the probability of transmission parameter and its standard error by averaging over the ratio between percent change in the susceptible compartment and the prevalence of the disease at each time step.

\subsubsection{Compartment Models}

\paragraph{SIR Models}
%In general, the models derived from the SIR ordinary differential equations are relatively accurate. Most estimates are within the range of $\pm$ 0.2 of the true $\rr$ for the baseline SIR datasets. The magnitude of the standard errors on the baseline datasets for the different SIR models generally depend on the method used. The models that estimate $\rr$ using optimization or gradient descent, namely SIR LS and SIR ReLS, have standard errors on the order of $10^{-4}$ to $10^{-7}$. The standard errors are somewhat larger for the methods that use regression, on the order of $10^{-1}$ to $10^{-3}$.

%The SIR models as a whole also do fairly well with changing the starting number of infected people. However, they tend not to do as well when the variance of the added errors is very large; in these situations, many of the models tend to have very large standard errors, sometimes on the order of $10^{1}$ or $10^{2}$.

%The exceptions to the above are the linear, incidence-prevalence ratio, and Markov chain models. The estimates tend to be far from the true $\rr$, and they fluctuate greatly between the baseline data sets. The standard errors for these models (if they exist) also tend to be very large under some of the changing conditions.

%Many of the SIR models do not work as well with the data that come from the polynomial linear models. Specifically, for the 0th and 1st order model datasets, the estimates can be unreasonably large or negative. The Log linear, Sequential Bayes, Smoothed max, and Max models do not have this problem.

\paragraph{Other Compartment Models}

%All of these models use optimization to estimate the parameters that are used to calculate $\rr$. They tend to be fairly accurate and consistent on the baseline data, often as good as or better than those from the various SIR models. The SEIR models (least squares and next generation) do not handle the large added errors very well, as it tends to give ridiculously large results. The standard errors tend to be on the order of $10^{-2}$ to $10^{-3}$. However, as with the estimates above, the standard errors from the SEIR models can become unreasonably large when the variance of the added errors becomes large. Also, as with the SIR models, the $\rr$ estimates and standard errors from the 0th and 1st order polynomial linear model datasets are not very good, as the former is sometimes negative and the latter can be very large. The MSEIR model tends to do fairly well in all situations.

\subsubsection{Exponential Growth}

%We have assumed for these models that the number of infections follows an exponential curve throughout the entire time frame of the data sets. This assumption certainly does not hold for the data from the 0th and 1st order linear models, and may not hold for some of the other data sets either.

%These models are generally not as accurate as the compartment models above. Most of them, except for the initial size model, tend to underestimate $\rr$, with estimates between 0.8 and 1.0. The standard errors are all fairly small, on the order of magnitude from $10^{-1}$ to $10^{-3}$. This is larger than many of the compartment model standard errors. The main advantage that all of these models, except for the maximum likelihood model, seem to be more robust to changing conditions and functional forms of the data compared to the compartment models. The estimates and standard errors for $\rr$ appear to be reasonable in all cases, including for very large added errors and for the 0th and 1st order linear model data.

\subsubsection{Network Models}

%The branching process model (the only one in this category so far) is not very accurate. In particular, for the SIR datasets, the $\rr$ estimates are less than 1, while the true $\rr$ is greater than 1. The standard errors, though, are relatively small, around the same order of magnitude as the SIR least squares models. Model estimates and standard errors are fairly consistent across the baseline datasets and changing starting compartment sizes. However, increasing the standard deviation of the added errors tends to have a somewhat noticeable impact on model results. The model also does not handle the 0th order polynomial linear data well.

\section{Discussion}
\label{sec:dis}

Below, we compare the estimates and standard errors of $\rr$ given by each of the models for a few selected data sets using forest plots; they show the point estimates and 95\% confidence intervals of $\rr$ from each model for a particular data set. The true $\rr$ value is also shown by the vertical red line in the graphs. All of the figures below show results from data sets that were simulated with added Gaussian errors. The plot on the top left shows the estimates from one of the baseline data sets. For the most part, the $\rr$ values range from approximately 0.8 to 2. We note that many of the models estimate an $/rr$ value of approximately 1, and some estimate that $/rr$ is less than 1. This would indicate that the infection disease is growing very slowly or is actually dying out, when in reality the disease is spreading at a fairly rapid rate. The confidence intervals are very small relative to the estimates for the most part; exceptions to this include the SIR IPR and LMA models as well as the Markov chain model.

The plot on the top right shows the estimates for the data set where $Y_0$ is increased to 1000. The $\rr$ estimates are almost all further away from the true $\rr$ value compared to the results from the baseline data sets. The confidence intervals are also more often than not larger than those in the baseline. In particular, the standard errors for some of the SIR models increase by several orders of magnitude.

The bottom left plot shows the estimates for the data set where the error added to the generated data is increased. For most of the models, the estimates of $\rr$ decrease slightly, and overall they do not change very much compared to the baseline data set. The confidence intervals for this data set increase substantially for some models, such as the SIR LMA, while they remain approximately the same as the baseline data sets for several of the others.

On the bottom right, we see the $\rr$ estimates for the case when disease prevalence grows linearly, as opposed to exponentially as seen in the data sets generated under SIR conditions, where the disease prevalence grows exponentially. Model performance here varies depending on how well the models the aforementioned linear growth. Some of the models, such as the SIR Markov chain and the IPR models, do not handle this case very well, giving much larger estimates and standard errors than are reasonable. On the other hand, models such as the SIR maximum and final size seem to be more flexible to different trajectories.

\begin{minipage}{0.5\linewidth}
	\includegraphics[width=3in]{BaselineFP.jpeg}
\end{minipage}
\begin{minipage}{0.5\linewidth}
	\includegraphics[width=3in]{StartPopFP.jpeg}
\end{minipage}

\begin{minipage}{0.5\linewidth}
	\includegraphics[width=3in]{VarianceFP.jpeg}
\end{minipage}
\begin{minipage}{0.5\linewidth}
	\includegraphics[width=3in]{OtherFP.jpeg}
\end{minipage}

\bibliographystyle{apa}%Choose a bibliograhpic style
\bibliography{Master}

\section{Appendix}

Included here are the technical details for fitting the models in R, as well as tables of the results.

\subsection{Compartment Models}

\subsubsection{SIR Least Squares}

\paragraph{Estimation}

For the least-squares SIR model, we estimate the parameters $\beta$ and $\gamma$ of the ODEs using optim() in R to minimize the sum of squared errors using Nelder-Mead. The standard errors for $\beta$ and $\gamma$ were estimated using the Hessian matrix given to us by optim(), and then the standard error for $\rr$ was calculated using the delta method.

\paragraph{Comments}

The $\rr$ estimates for the baseline data are both close to the actual value, and the standard errors were relatively small. We would probably expect this given that this is the same model from which the data were generated.

For the changing compartment sizes, we see that the standard errors are much larger for smaller values of $Y_0$. This might be because the variance of the I compartment is much larger relative to the size of the compartment, since we hold $\sigma_Y$ at 1 no matter if $Y_0$ is 1 or 1000.

For the changing error variances, we see that the estimates for $\rr$ become less accurate for increasing $\sigma_X$ and $\sigma_Y$. However, it is interesting to note that the standard errors are increasing for the monotonic Gaussian error data, but decreasing for the autoregressive and monotonic autoregressive error data and relatively stable for the Gaussian non-monotonic data. It is not clear why this would be the case.

\begin{table}[H]
	\caption{$\rr$ Estimates and Std. Errs, SIR Model, 
		Baseline Data, $X_0 = 99950, Y_0 = 50$, 
		$\sigma_X = 10, \sigma_Y = 1$}
	\begin{footnotesize}
		\hskip -1cm
		\begin{tabular}{l|r|r|r|r|r|r|r|r}
			\hline
			Data Set & Auto Est & Auto SE & AutoM Est & AutoM SE & Norm Est & Norm SE & NormM Est & NormM SE\\
			\hline
			Baseline 1 & 1.965322 & 0.000869 & 1.983308 & 0.000882 & 2.005867 & 0.000893 & 1.994736 & 0.000884\\
			\hline
			Baseline 2 & 2.039608 & 0.000913 & 1.999245 & 0.000895 & 2.001217 & 0.000889 & 1.995960 & 0.000887\\
			\hline
			Baseline 3 & 2.005609 & 0.000901 & 1.966631 & 0.000873 & 2.016216 & 0.000901 & 1.997175 & 0.000887\\
			\hline
			Baseline 4 & 1.958037 & 0.000859 & 1.941109 & 0.000852 & 1.995544 & 0.000886 & 2.004107 & 0.000892\\
			\hline
			Baseline 5 & 2.050904 & 0.000926 & 1.964328 & 0.000866 & 2.009912 & 0.000896 & 1.990153 & 0.000881\\
			\hline
		\end{tabular}
	\end{footnotesize}
\end{table}

\begin{table}[H]

	\caption{$\rr$ Estimates and Std. Errs, SIR Model,
		Different Starting Populations, 
		$\sigma_X = 10, \sigma_Y = 1$}
	\begin{footnotesize}
		\hskip -1.7cm
		\begin{tabular}{l|r|r|r|r|r|r|r|r}
			\hline
			Data Set & Auto Est & Auto SE & AutoM Est & AutoM SE & Norm Est & Norm SE & NormM Est & NormM SE\\
			\hline
			$X_0 = 99999, Y_0 = 1$ & 1.689292 & 0.020794 & -4.167216 & 0.168853 & 1.976254 & 0.045081 & 1.688593 & 0.029680\\
			\hline
			$X_0 = 99990, Y_0 = 10$ & 2.038297 & 0.004597 & 2.207003 & 0.005165 & 2.037726 & 0.004596 & 1.883084 & 0.004049\\
			\hline
			$X_0 = 99900, Y_0 = 100$ & 2.009581 & 0.000449 & 2.006557 & 0.000448 & 1.999236 & 0.000445 & 1.995467 & 0.000444\\
			\hline
			$X_0 = 99000, Y_0 = 1000$ & 2.002521 & 0.000046 & 2.001829 & 0.000046 & 1.999519 & 0.000046 & 2.000098 & 0.000046\\
			\hline
		\end{tabular}
	\end{footnotesize}
\end{table}

\begin{table}[H]

	\caption{$\rr$ Estimates and Std. Errs, SIR Model, 
		Different Variances, $X_0 = 99000$, $Y_0 = 1000$}
	\begin{footnotesize}
		\hskip -1.7cm
		\begin{tabular}{l|r|r|r|r|r|r|r|r}
			\hline
			Data Set & Auto Est & Auto SE & AutoM Est & AutoM SE & Norm Est & Norm SE & NormM Est & NormM SE\\
			\hline
			$\sigma_X = 10, \sigma_Y = 1$ & 1.927444 & 0.000059 & 1.927090 & 5.9e-05 & 1.925768 & 5.9e-05 & 1.926576 & 5.9e-05\\
			\hline
			$\sigma_X = 100, \sigma_Y = 10$ & 1.926786 & 0.000058 & 1.916571 & 5.9e-05 & 1.919778 & 5.9e-05 & 1.929422 & 5.9e-05\\
			\hline
			$\sigma_X = 300, \sigma_Y = 30$ & 1.994211 & 0.000062 & 1.976893 & 5.8e-05 & 1.930660 & 5.8e-05 & 1.920869 & 5.9e-05\\
			\hline
			$\sigma_X = 1000, \sigma_Y = 100$ & 1.749057 & 0.000050 & 1.736588 & 5.9e-05 & 2.003654 & 6.2e-05 & 1.897049 & 5.9e-05\\
			\hline
			$\sigma_X = 3000, \sigma_Y = 300$ & 1.992674 & 0.000063 & 1.681585 & 4.9e-05 & 1.959815 & 6.7e-05 & 1.770816 & 5.3e-05\\
			\hline
			$\sigma_X = 10000, \sigma_Y = 1000$ & -42.264180 & 0.002694 & 1.773388 & 3.1e-05 & 1.889188 & 6.6e-05 & 2.649010 & 6.3e-05\\
			\hline
		\end{tabular}
	\end{footnotesize}
\end{table}

\begin{table}[H]

	\caption{$\rr$ Estimates and Std. Errs, SIR Model,
		Other Models, $X_0 = 99950, Y_0 = 50$, 
		$\sigma_X = 10, \sigma_Y = 1$}
	\begin{footnotesize}
		\hskip -1cm
		\begin{tabular}{l|r|r|r|r|r|r|r|r}
			\hline
			Data Set & Auto Est & Auto SE & AutoM Est & AutoM SE & Norm Est & Norm SE & NormM Est & NormM SE\\
			\hline
			0th & -2.032207 & 0.244058 & 1.229963 & 0.069020 & 1.021369 & 0.056558 & 0.795460 & 0.331468\\
			\hline
			1st & 1.038117 & 0.000029 & 1.037889 & 0.000029 & 1.037902 & 0.000029 & 1.037887 & 0.000029\\
			\hline
			2nd & 1.198856 & 0.000045 & 1.198589 & 0.000045 & 1.198772 & 0.000045 & 1.198501 & 0.000045\\
			\hline
			3rd & 1.202405 & 0.000043 & 1.202098 & 0.000043 & 1.202350 & 0.000043 & 1.202240 & 0.000043\\
			\hline
			4th & 1.210597 & 0.000047 & 1.210597 & 0.000047 & 1.210194 & 0.000047 & 1.210304 & 0.000047\\
			\hline
			Linear SIR & 1.274346 & 0.000034 & 1.273739 & 0.000034 & 1.274229 & 0.000034 & 1.273950 & 0.000034\\
			\hline
		\end{tabular}
	\end{footnotesize}
\end{table}

\subsubsection{SIR Reparameterized Least Squares}

\paragraph{Estimation}

For the reparameterized SIR model, we estimate the parameters $\rr$ (as well as $\gamma$) of the ODEs using optim() in R to minimize the sum of squared errors using Nelder-Mead. The standard errors were taken directly from the term in the Hessian matrix corresponding to $\rr$. (I also have the sensitivity analysis code here.)

\paragraph{Comments}

The $\rr$ estimates for the baseline data are both close to the actual value, and the estimated standard errors are extremely small. We would probably expect this given that this is the same model from which the data were generated. In general, the estimates for the standard errors are also much smaller than normal least squares. (Some of the errors are 0 because I rounded to 6 digits - all of them are around the magnitude of $10 ^ {-7}$)

The $\rr$ estimates for the data with non-monotonic errors become more accurate as ($\sigma_X$, $\sigma_Y$) increase, while they become less accurate for the monotonic values. The standard errors are also positively correlated with the estimate of $\rr$ as long as all other parts, i.e., $X_0$ and $Y_0$ are held constant.

\begin{table}[H]

	\caption{$\rr$ Estimates and Std. Errs, SIRREPAR Model,
		Baseline Data, $X_0 = 99950, Y_0 = 50$, 
		$\sigma_X = 10, \sigma_Y = 1$}
	\begin{footnotesize}
		\hskip -1cm
		\begin{tabular}{l|r|r|r|r|r|r|r|r}
			\hline
			Data Set & Auto Est & Auto SE & AutoM Est & AutoM SE & Norm Est & Norm SE & NormM Est & NormM SE\\
			\hline
			Baseline 1 & 3.226272 & 0.000000 & 1.984447 & 0.000294 & 2.006310 & 0.000298 & 1.993558 & 0.000294\\
			\hline
			Baseline 2 & 3.203012 & 0.000000 & 1.999589 & 0.000299 & 2.000868 & 0.000297 & 1.995630 & 0.000296\\
			\hline
			Baseline 3 & 4.180804 & 0.000733 & 4.180810 & 0.000724 & 5.043355 & 0.000908 & 1.997399 & 0.000296\\
			\hline
			Baseline 4 & 4.180819 & 0.000709 & 4.180816 & 0.000715 & 1.995237 & 0.000295 & 2.003821 & 0.000298\\
			\hline
			Baseline 5 & 4.180799 & 0.000741 & 4.180815 & 0.000717 & 2.008713 & 0.000300 & 4.180812 & 0.000720\\
			\hline
		\end{tabular}
	\end{footnotesize}
\end{table}

\begin{table}[H]

	\caption{$\rr$ Estimates and Std. Errs, SIRREPAR Model,
		Different Starting Populations, 
		$\sigma_X = 10, \sigma_Y = 1$}
	\begin{footnotesize}
		\hskip -1.7cm
		\begin{tabular}{l|r|r|r|r|r|r|r|r}
			\hline
			Data Set & Auto Est & Auto SE & AutoM Est & AutoM SE & Norm Est & Norm SE & NormM Est & NormM SE\\
			\hline
			$X_0 = 99999, Y_0 = 1$ & 5.899238 & 0.025009 & 5.898924 & 0.051433 & 3.394166 & 0.000021 & 3.217232 & 0.000016\\
			\hline
			$X_0 = 99990, Y_0 = 10$ & 3.162364 & 0.000001 & 3.170152 & 0.000002 & 3.131576 & 0.000001 & 3.108194 & 0.000001\\
			\hline
			$X_0 = 99900, Y_0 = 100$ & 4.564960 & 0.000404 & 2.005939 & 0.000150 & 1.998880 & 0.000149 & 1.994592 & 0.000148\\
			\hline
			$X_0 = 99000, Y_0 = 1000$ & 47.241714 & 0.002493 & 47.241714 & 0.002511 & 47.241714 & 0.002502 & 47.241714 & 0.002502\\
			\hline
		\end{tabular}
	\end{footnotesize}
\end{table}

\begin{table} [H]

	\caption{$\rr$ Estimates and Std. Errs, SIRREPAR Model,
		Different Variances, 
		$X_0 = 99000, Y_0 = 1000$}
	\begin{footnotesize}
		\hskip -1.7cm
		\begin{tabular}{l|r|r|r|r|r|r|r|r}
			\hline
			Data Set & Auto Est & Auto SE & AutoM Est & AutoM SE & Norm Est & Norm SE & NormM Est & NormM SE\\
			\hline
			$\sigma_X = 10, \sigma_Y = 1$ & 1.926733 & 4e-06 & 1.927468 & 4e-06 & 1.925555 & 4e-06 & 1.926689 & 4e-06\\
			\hline
			$\sigma_X = 100, \sigma_Y = 10$ & 1.925550 & 4e-06 & 1.916380 & 4e-06 & 1.920075 & 4e-06 & 1.928987 & 4e-06\\
			\hline
			$\sigma_X = 300, \sigma_Y = 30$ & 1.994719 & 4e-06 & 1.976273 & 4e-06 & 1.930978 & 4e-06 & 1.920290 & 4e-06\\
			\hline
			$\sigma_X = 1000, \sigma_Y = 100$ & 1.749031 & 3e-06 & 1.736589 & 3e-06 & 2.003044 & 4e-06 & 1.897328 & 4e-06\\
			\hline
			$\sigma_X = 3000, \sigma_Y = 300$ & 1.990948 & 4e-06 & 1.681748 & 3e-06 & 1.959229 & 4e-06 & 1.770892 & 3e-06\\
			\hline
			$\sigma_X = 10000, \sigma_Y = 1000$ & 5.329640 & 3e-05 & 1.773820 & 2e-06 & 1.887014 & 4e-06 & 2.648300 & 6e-06\\
			\hline
		\end{tabular}
	\end{footnotesize}
\end{table}
\begin{table}[H]

	\caption{$\rr$ Estimates and Std. Errs, SIRREPAR Model,
		Other Models, $X_0 = 99950, Y_0 = 50$, 
		$\sigma_X = 10, \sigma_Y = 1$}
	\begin{footnotesize}
		\hskip -1cm
		\begin{tabular}{l|r|r|r|r|r|r|r|r}
			\hline
			Data Set & Auto Est & Auto SE & AutoM Est & AutoM SE & Norm Est & Norm SE & NormM Est & NormM SE\\
			\hline
			0th & 3.272461 & 0e+00 & 3.275535 & 0e+00 & 3.135010 & 0e+00 & 3.235197 & 0e+00\\
			\hline
			1st & 1.038125 & 0e+00 & 1.037880 & 0e+00 & 1.037903 & 0e+00 & 1.037881 & 0e+00\\
			\hline
			2nd & 1.198760 & 1e-06 & 1.198526 & 1e-06 & 1.198864 & 1e-06 & 1.198616 & 1e-06\\
			\hline
			3rd & 1.202489 & 1e-06 & 1.202071 & 1e-06 & 1.202246 & 1e-06 & 1.202355 & 1e-06\\
			\hline
			4th & 1.210605 & 1e-06 & 1.210605 & 1e-06 & 1.210220 & 1e-06 & 1.210199 & 1e-06\\
			\hline
			Linear SIR & 1.274175 & 1e-06 & 1.273799 & 1e-06 & 1.274314 & 1e-06 & 1.274075 & 1e-06\\
			\hline
		\end{tabular}
	\end{footnotesize}
\end{table}

\subsubsection{SIR Linear Model}

\paragraph{Estimation}

To calculate the estimates for $\rr$, we use lm() in R to fit the linear polynomial regression functions for X(t) and Y(t) with respect to t. We then calculated the derivatives for the linear regression functions at time 0 using our own function. For the standard error, the jackknife was used on the data set. This calculates the standard error of the $\rr$ by using the standard deviation of the $R-0$ estimates after leaving out a single observation.

\paragraph{Comments}

The estimates for $\rr$ are very inaccurate compared to the other models. The inaccuracy is most glaring for the data with large error variances and the simulated from models besides the SIR one. The estimates seem to become less accurate for the data generated from the classic SIR model as $\sigma_X$ and $\sigma_Y$ increase and as $Y_0$ decreases relative to the total population. There are also many estimates of $\rr$ that are negative, which implies that there is a negative derivative estimate for either the susceptible or infected compartment (is this supposed to be possible?).

The standard errors are also extremely large in some cases. This seems to hold consistently when $\sigma_X$ and $\sigma_Y$ are large for the classic SIR data as well as for the 4th order linear data. However, there are also other places where large standard error estimates appear for no apparent reason. The large standard errors and the above issue with the $\rr$ estimates being negative could be because we are only using a point estimate. In general, the only time we see fairly consistent and reasonable estimates from the linear model is for the SIR data with relatively large $Y_0$ as well as small $\sigma_X$ and $\sigma_Y$ (i.e., for the data sets where ($X_0$, $Y_0$) = (99000, 1000)).

\begin{table}[H]

	\caption{$\rr$ Estimates and Std. Errs, SIRLMA Model,
		Baseline Data, $X_0 = 99950, Y_0 = 50$, 
		$\sigma_X = 10, \sigma_Y = 1$}
	\begin{footnotesize}
		\hskip -1cm
		\begin{tabular}{l|r|r|r|r|r|r|r|r}
			\hline
			Data Set & Auto Est & Auto SE & AutoM Est & AutoM SE & Norm Est & Norm SE & NormM Est & NormM SE\\
			\hline
			Baseline 1 & 0.864913 & 0.203402 & 1.739858 & 0.122632 & 1.359771 & 0.919629 & 1.173245 & 0.096355\\
			\hline
			Baseline 2 & 1.485664 & 1.654031 & 1.294412 & 0.618155 & 1.461580 & 0.734907 & 1.186621 & 0.091633\\
			\hline
			Baseline 3 & 1.082572 & 0.088142 & 1.389817 & 0.117032 & 0.615994 & 1.496674 & 1.088364 & 0.089282\\
			\hline
			Baseline 4 & 1.088362 & 0.367110 & 1.370251 & 0.159221 & 2.683109 & 27.981877 & 1.090398 & 0.059515\\
			\hline
			Baseline 5 & 0.871650 & 0.100744 & 1.438190 & 0.382650 & 0.829591 & 0.120911 & 0.924238 & 0.188305\\
			\hline
		\end{tabular}
	\end{footnotesize}
\end{table}

\begin{table}[H]

	\caption{$\rr$ Estimates and Std. Errs, SIRLMA Model,
		Different Starting Populations, 
		$\sigma_X = 10, \sigma_Y = 1$}
	\begin{footnotesize}
		\hskip -1.7cm
		\begin{tabular}{l|r|r|r|r|r|r|r|r}
			\hline
			Data Set & Auto Est & Auto SE & AutoM Est & AutoM SE & Norm Est & Norm SE & NormM Est & NormM SE\\
			\hline
			$X_0 = 99999, Y_0 = 1$ & 1.713333 & 1.757278 & 1.187769 & 0.515629 & 1.451833 & 59.746089 & 1.068633 & 1.809596\\
			\hline
			$X_0 = 99990, Y_0 = 10$ & 1.197457 & 0.192831 & 0.674259 & 0.090168 & 1.095143 & 0.143146 & 1.199437 & 0.106393\\
			\hline
			$X_0 = 99900, Y_0 = 100$ & 0.796760 & 0.488540 & 1.022290 & 0.070029 & 0.024273 & 4.564245 & 1.265712 & 0.080303\\
			\hline
			$X_0 = 99000, Y_0 = 1000$ & 1.431229 & 0.105605 & 2.091993 & 0.340798 & 2.183329 & 0.428090 & 1.812900 & 0.122106\\
			\hline
		\end{tabular}
	\end{footnotesize}
\end{table}

\begin{table}[H]

	\caption{$\rr$ Estimates and Std. Errs, SIRLMA Model,
		Different Variances, 
		$X_0 = 99000, Y_0 = 1000$}
	\begin{footnotesize}
		\hskip -2cm
		\begin{tabular}{l|r|r|r|r|r|r|r|r}
			\hline
			Data Set & Auto Est & Auto SE & AutoM Est & AutoM SE & Norm Est & Norm SE & NormM Est & NormM SE\\
			\hline
			$\sigma_X = 10, \sigma_Y = 1$ & -121.825007 & 904.236167 & 121.668347 & 308.632857 & 21.981032 & 21.204959 & 18.741699 & 17.132467\\
			\hline
			$\sigma_X = 100, \sigma_Y = 10$ & 2.918979 & 2.010263 & -1.223521 & 2.355038 & -5.683542 & 10.750736 & 4.071392 & 2.877332\\
			\hline
			$\sigma_X = 300, \sigma_Y = 30$ & 2.769860 & 1.759591 & 2.002654 & 0.949586 & 3.227956 & 3.934407 & 87.874782 & 456.259511\\
			\hline
			$\sigma_X = 1000, \sigma_Y = 100$ & 0.915434 & 0.114881 & 1.887797 & 0.672640 & 1.759919 & 1.537275 & -0.235868 & 1.519149\\
			\hline
			$\sigma_X = 3000, \sigma_Y = 300$ & -4.547791 & 970.222266 & 1.037930 & 0.112483 & 0.748209 & 0.437849 & 0.235119 & 1.489429\\
			\hline
			$\sigma_X = 10000, \sigma_Y = 1000$ & 1.033786 & 0.068722 & 0.958868 & 0.066758 & 0.862043 & 0.168471 & 0.671640 & 0.206497\\
			\hline
		\end{tabular}
	\end{footnotesize}
\end{table}

\begin{table}[H]

	\caption{$\rr$ Estimates and Std. Errs, SIRLMA Model,
		Other Models, $X_0 = 99950, Y_0 = 50$, 
		$\sigma_X = 10, \sigma_Y = 1$}
	\begin{footnotesize}
		\hskip -1cm
		\begin{tabular}{l|r|r|r|r|r|r|r|r}
			\hline
			Data Set & Auto Est & Auto SE & AutoM Est & AutoM SE & Norm Est & Norm SE & NormM Est & NormM SE\\
			\hline
			0th & 1.049994 & 0.056806 & 1.286339 & 0.479624 & 0.980772 & 0.060085 & 0.474270 & 2.099030\\
			\hline
			1st & 1.000767 & 0.125396 & 1.009129 & 0.142918 & 1.018881 & 0.127098 & 1.019178 & 0.159078\\
			\hline
			2nd & 1.467835 & 0.558738 & 0.085260 & 1.223959 & 1.215482 & 0.162666 & -6.656798 & 265.908999\\
			\hline
			3rd & 4.877978 & 57.939148 & 1.111925 & 0.120143 & 1.238831 & 0.321316 & 1.227691 & 0.221355\\
			\hline
			4th & -0.892745 & 7.391321 & -0.892745 & 7.391321 & 0.445461 & 2.016867 & 1.709272 & 2.038759\\
			\hline
			Linear SIR & 1.214648 & 0.062718 & 0.950315 & 0.106644 & 1.219144 & 0.109327 & 1.206761 & 0.048099\\
			\hline
		\end{tabular}
	\end{footnotesize}
\end{table}

\subsubsection{SIR Linear Model Average}

\paragraph{Estimation}

To calculate the estimates for $\rr$, we use lm() in R to fit the linear polynomial regression functions for X(t) and Y(t) with respect to t. We then calculated the derivatives of the linear regression functions at each time using our own function in order to find the average linear model $\rr$ estimate across all time steps $t$. The standard error was calculated by taking the standard deviation of the $\rr$ estimates across all time steps $t$ in the data set.

\paragraph{Comments}

The estimates for $\rr$ here are somewhat more accurate compared to the linear model estimates that only looked at the derivative first time point. In particular, it does a lot better for the baseline data and the data from other models. The estimates are still really bad for large values of $\sigma_X$ and $\sigma_Y$ (including some negative estimates). Interestingly, the linear model average estimates here are less accurate for large values of $Y_0$, whereas they were more accurate for the linear model.

The standard errors are not as ridiculously large compared to the linear model with only one time point, but are still extremely big compared to what we might expect to see from other models. This is especially apparent when $\sigma_X$ and $\sigma_Y$ are large.

\begin{table}[H]

	\caption{$\rr$ Estimates and Std. Errs, SIRLMAT Model,
		Baseline Data, $X_0 = 99950, Y_0 = 50$, 
		$\sigma_X = 10, \sigma_Y = 1$}
	\begin{footnotesize}
		\hskip -1cm
		\begin{tabular}{l|r|r|r|r|r|r|r|r}
			\hline
			Data Set & Auto Est & Auto SE & AutoM Est & AutoM SE & Norm Est & Norm SE & NormM Est & NormM SE\\
			\hline
			Baseline 1 & 2.015772 & 27.518508 & 2.150988 & 4.474655 & 1.988515 & 0.232390 & 2.103188 & 0.636825\\
			\hline
			Baseline 2 & 2.571202 & 21.048981 & -0.131021 & 51.293391 & 2.192838 & 1.684886 & 2.189259 & 1.514415\\
			\hline
			Baseline 3 & 1.201407 & 21.526024 & 3.023289 & 26.322188 & 2.079618 & 1.027536 & 2.007048 & 0.269511\\
			\hline
			Baseline 4 & 1.991370 & 1.164132 & 3.604925 & 35.619793 & 2.028025 & 0.205708 & 1.875085 & 2.080861\\
			\hline
			Baseline 5 & 1.873964 & 7.146565 & 1.032642 & 11.871804 & 1.372848 & 10.164198 & 1.926002 & 0.166963\\
			\hline
		\end{tabular}
	\end{footnotesize}
\end{table}

\begin{table}[H]

	\caption{$\rr$ Estimates and Std. Errs, SIRLMAT Model,
		Different Starting Populations, 
		$\sigma_X = 10, \sigma_Y = 1$}
	\begin{footnotesize}
		\hskip -1.7cm
		\begin{tabular}{l|r|r|r|r|r|r|r|r}
			Data Set & Auto Est & Auto SE & AutoM Est & AutoM SE & Norm Est & Norm SE & NormM Est & NormM SE\\
			\hline
			$X_0 = 99999, Y_0 = 1$ & 2.558602 & 32.007328 & 1.296284 & 9.652181 & 1.178839 & 3.901942 & 1.287102 & 0.524929\\
			\hline
			$X_0 = 99990, Y_0 = 10$ & 0.745127 & 3.487290 & 1.541772 & 1.793288 & 3.227689 & 26.102558 & 2.489264 & 26.755104\\
			\hline
			$X_0 = 99900, Y_0 = 100$ & 2.082023 & 2.181500 & -1.745769 & 81.529981 & 1.946617 & 1.232307 & 2.008181 & 0.191259\\
			\hline
			$X_0 = 99000, Y_0 = 1000$ & 1.969496 & 0.088997 & 1.979588 & 0.093847 & 1.972052 & 0.036176 & 1.964677 & 0.028405\\
			\hline
		\end{tabular}
	\end{footnotesize}
\end{table}

\begin{table}[H]

	\caption{$\rr$ Estimates and Std. Errs, SIRLMAT Model,
		Different Variances, 
		$X_0 = 99000, Y_0 = 1000$}
	\begin{footnotesize}
		\hskip -1.7cm
		\begin{tabular}{l|r|r|r|r|r|r|r|r}
			\hline
			Data Set & Auto Est & Auto SE & AutoM Est & AutoM SE & Norm Est & Norm SE & NormM Est & NormM SE\\
			\hline
			$\sigma_X = 10, \sigma_Y = 1$ & 5.556294 & 63.868004 & 2.843995 & 7.555433 & 2.277135 & 2.022040 & 2.234907 & 1.759312\\
			\hline
			$\sigma_X = 100, \sigma_Y = 10$ & 2.020666 & 0.483996 & 1.802373 & 5.248770 & 1.891754 & 5.445302 & 2.034306 & 0.527914\\
			\hline
			$\sigma_X = 300, \sigma_Y = 30$ & -1.064869 & 50.543183 & 4.685491 & 53.111296 & 2.076773 & 0.646978 & 2.521401 & 5.276258\\
			\hline
			$\sigma_X = 1000, \sigma_Y = 100$ & 0.999874 & 3.992411 & 2.390182 & 3.320493 & 1.456545 & 6.796468 & 1.703567 & 2.395295\\
			\hline
			$\sigma_X = 3000, \sigma_Y = 300$ & -2.740783 & 70.416107 & 1.371659 & 1.347889 & 0.030795 & 21.459099 & 2.023080 & 6.341932\\
			\hline
			$\sigma_X = 10000, \sigma_Y = 1000$ & 0.877306 & 0.815504 & 1.106179 & 2.530188 & 1.183491 & 3.063114 & 0.949453 & 6.173692\\
			\hline
		\end{tabular}
	\end{footnotesize}
\end{table}
\begin{table}[H]

	\caption{$\rr$ Estimates and Std. Errs, SIRLMAT Model,
		Other Models, $X_0 = 99950, Y_0 = 50$, 
		$\sigma_X = 10, \sigma_Y = 1$}
	\begin{footnotesize}
		\hskip -1cm
		\begin{tabular}{l|r|r|r|r|r|r|r|r}
			\hline
			Data Set & Auto Est & Auto SE & AutoM Est & AutoM SE & Norm Est & Norm SE & NormM Est & NormM SE\\
			\hline
			0th & 0.848941 & 3.448691 & 0.925502 & 18.031398 & -0.885541 & 17.053815 & 0.898319 & 0.676024\\
			\hline
			1st & 1.152006 & 0.018441 & 1.152364 & 0.017047 & 1.151996 & 0.016274 & 1.152495 & 0.015856\\
			\hline
			2nd & 1.164984 & 0.747602 & 1.181705 & 1.419549 & 1.233825 & 1.421868 & 3.337101 & 40.581184\\
			\hline
			3rd & 1.184468 & 0.202637 & 1.168692 & 0.036327 & 1.170062 & 0.035413 & 1.170215 & 0.034933\\
			\hline
			4th & 1.188048 & 0.446227 & 1.188048 & 0.446227 & 1.172531 & 0.353343 & 1.189808 & 0.052711\\
			\hline
			Linear SIR & 1.203159 & 0.014809 & 1.195875 & 0.024693 & 1.200980 & 0.003110 & 1.200130 & 0.004076\\
			\hline
		\end{tabular}
	\end{footnotesize}
\end{table}

\subsubsection{SIR Maximum}

\paragraph{Estimation}

We use the which.max() function in R to find the point in time where the number people in the infected compartment is at a maximum. We need to come up with a way to calculate the variance (sensitivity analysis?).

\paragraph{Comments}

The SIR max model gives decent results. Most of the estimates are fairly close to the true $\rr$ value. It appears to give more accurate $\rr$ estimates better for higher values of $Y_0$ relative to the total population as well as smaller values of $\sigma_X$ and $\sigma_Y$. This model also consistently gives reasonable results, unlike for other models, which sometimes fail miserably when dealing with larger errors or data generated from other models (especially for 0th and 1st order data).

\begin{table}[H]

	\caption{$\rr$ Estimates and Std. Errs, SIRMAX Model,
		Baseline Data, $X_0 = 99950, Y_0 = 50$, 
		$\sigma_X = 10, \sigma_Y = 1$}
	\begin{footnotesize}
		\hskip -1cm
		\begin{tabular}{l|r|r|r|r|r|r|r|r}
			\hline
			Data Set & Auto Est & Auto SE & AutoM Est & AutoM SE & Norm Est & Norm SE & NormM Est & NormM SE\\
			\hline
			Baseline 1 & 1.002232 & NA & 1.002359 & NA & 1.002498 & NA & 1.002531 & NA\\
			\hline
			Baseline 2 & 1.002405 & NA & 1.002802 & NA & 1.002364 & NA & 1.002561 & NA\\
			\hline
			Baseline 3 & 1.002261 & NA & 1.002393 & NA & 1.002516 & NA & 1.002501 & NA\\
			\hline
			Baseline 4 & 1.002469 & NA & 1.002494 & NA & 1.002511 & NA & 1.002481 & NA\\
			\hline
			Baseline 5 & 1.002392 & NA & 1.002496 & NA & 1.002509 & NA & 1.002529 & NA\\
			\hline
		\end{tabular}
	\end{footnotesize}
\end{table}

\begin{table}[H]

	\caption{$\rr$ Estimates and Std. Errs, SIRMAX Model,
		Different Starting Populations, 
		$\sigma_X = 10, \sigma_Y = 1$}
	\begin{footnotesize}
		\hskip -1.7cm
		\begin{tabular}{l|r|r|r|r|r|r|r|r}
			\hline
			Data Set & Auto Est & Auto SE & AutoM Est & AutoM SE & Norm Est & Norm SE & NormM Est & NormM SE\\
			\hline
			$X_0 = 99999, Y_0 = 1$ & 0.999831 & NA & 1.000028 & NA & 0.999962 & NA & 1.000079 & NA\\
			\hline
			$X_0 = 99990, Y_0 = 10$ & 1.000442 & NA & 1.000473 & NA & 1.000514 & NA & 1.000527 & NA\\
			\hline
			$X_0 = 99900, Y_0 = 100$ & 1.005071 & NA & 1.004938 & NA & 1.005222 & NA & 1.004948 & NA\\
			\hline
			$X_0 = 99000, Y_0 = 1000$ & 1.049518 & NA & 1.050139 & NA & 1.050143 & NA & 1.050157 & NA\\
			\hline
		\end{tabular}
	\end{footnotesize}
\end{table}

\begin{table}[H]

	\caption{$\rr$ Estimates and Std. Errs, SIRMAX Model,
		Different Variances, 
		$X_0 = 99000, Y_0 = 1000$}
	\begin{footnotesize}
		\hskip -1.7cm
		\begin{tabular}{l|r|r|r|r|r|r|r|r}
			\hline
			Data Set & Auto Est & Auto SE & AutoM Est & AutoM SE & Norm Est & Norm SE & NormM Est & NormM SE\\
			\hline
			$\sigma_X = 10, \sigma_Y = 1$ & 1.050097 & NA & 1.049897 & NA & 1.050034 & NA & 1.050098 & NA\\
			\hline
			$\sigma_X = 100, \sigma_Y = 10$ & 1.048816 & NA & 1.051191 & NA & 1.051560 & NA & 1.051580 & NA\\
			\hline
			$\sigma_X = 300, \sigma_Y = 30$ & 1.045130 & NA & 1.054815 & NA & 1.042187 & NA & 1.049682 & NA\\
			\hline
			$\sigma_X = 1000, \sigma_Y = 100$ & 1.038396 & NA & 1.051101 & NA & 1.057586 & NA & 1.049022 & NA\\
			\hline
			$\sigma_X = 3000, \sigma_Y = 300$ & 1.005515 & NA & 1.081383 & NA & 1.002521 & NA & 1.064796 & NA\\
			\hline
			$\sigma_X = 10000, \sigma_Y = 1000$ & 1.076870 & NA & 1.114340 & NA & 1.077148 & NA & 1.037306 & NA\\
			\hline
		\end{tabular}
	\end{footnotesize}
\end{table}

\begin{table}[H]

	\caption{$\rr$ Estimates and Std. Errs, SIRMAX Model,
		Other Models, $X_0 = 99950, Y_0 = 50$, 
		$\sigma_X = 10, \sigma_Y = 1$}
	\begin{footnotesize}
		\hskip -1cm
		\begin{tabular}{l|r|r|r|r|r|r|r|r}
			\hline
			Data Set & Auto Est & Auto SE & AutoM Est & AutoM SE & Norm Est & Norm SE & NormM Est & NormM SE\\
			\hline
			0th & 1.000604 & NA & 1.000563 & NA & 1.000366 & NA & 1.000543 & NA\\
			\hline
			1st & 1.083647 & NA & 1.083855 & NA & 1.083627 & NA & 1.083315 & NA\\
			\hline
			2nd & 1.086218 & NA & 1.085875 & NA & 1.085994 & NA & 1.085891 & NA\\
			\hline
			3rd & 1.096363 & NA & 1.095844 & NA & 1.096062 & NA & 1.095990 & NA\\
			\hline
			4th & 1.085509 & NA & 1.085509 & NA & 1.085289 & NA & 1.085296 & NA\\
			\hline
			Linear SIR & 1.127272 & NA & 1.127403 & NA & 1.127313 & NA & 1.127194 & NA\\
			\hline
		\end{tabular}
	\end{footnotesize}
\end{table}

\subsubsection{SIR Smoothed Max}

\paragraph{Estimation}

We use the smooth.spline() function in R to get estimates for X(t) and Y(t), then used the which.max() function to find the point in time where the number of people in the infected compartment is at a maximum. The standard error for the $\rr$ estimates was done using the jackknife method; we estimate a new spline function after removing one observation from the data set and calculate a new $\rr$ estimate, then use the standard deviation of those estimates as the estimate of the standard error.

\paragraph{Comments}

The results are fairly similar to the non-smoothed SIR maximum model.  While the model does not give perfectly accurate estimates for $\rr$, the results are decent. It does not give us any crazy results. The model is more accurate for larger values of $Y_0$ relative to the total size of the population and for smaller $\sigma_X$ and $\sigma_Y$.

The standard errors are reasonable values, and they seem to be fairly consistent under repeated estimation with the same conditions for the generated data.  They increase with increases in $Y_0$ as well as $\sigma_X$ and $\sigma_Y$.

\begin{table}[H]

	\caption{$\rr$ Estimates and Std. Errs, SIRSMAX Model,
		Baseline Data, $X_0 = 99950, Y_0 = 50$, 
		$\sigma_X = 10, \sigma_Y = 1$}
	\begin{footnotesize}
		\hskip -1cm
		\begin{tabular}{l|r|r|r|r|r|r|r|r}
			\hline
			Data Set & Auto Est & Auto SE & AutoM Est & AutoM SE & Norm Est & Norm SE & NormM Est & NormM SE\\
			\hline
			Baseline 1 & 1.002424 & 2.3e-05 & 1.002460 & 1.1e-05 & 1.002435 & 1.5e-05 & 1.002459 & 8e-06\\
			\hline
			Baseline 2 & 1.002507 & 2.2e-05 & 1.002444 & 1.8e-05 & 1.002462 & 1.9e-05 & 1.002436 & 7e-06\\
			\hline
			Baseline 3 & 1.002374 & 1.9e-05 & 1.002418 & 5.0e-06 & 1.002439 & 1.7e-05 & 1.002445 & 5e-06\\
			\hline
			Baseline 4 & 1.002434 & 1.5e-05 & 1.002479 & 8.0e-06 & 1.002461 & 1.5e-05 & 1.002449 & 6e-06\\
			\hline
			Baseline 5 & 1.002403 & 2.0e-05 & 1.002469 & 9.0e-06 & 1.002439 & 1.6e-05 & 1.002468 & 6e-06\\
			\hline
		\end{tabular}
	\end{footnotesize}
\end{table}

\begin{table}[H]

	\caption{$\rr$ Estimates and Std. Errs, SIRSMAX Model,
		Different Starting Populations, 
		$\sigma_X = 10, \sigma_Y = 1$}
	\begin{footnotesize}
		\hskip -1.7cm
		\begin{tabular}{l|r|r|r|r|r|r|r|r}
			\hline
			Data Set & Auto Est & Auto SE & AutoM Est & AutoM SE & Norm Est & Norm SE & NormM Est & NormM SE\\
			\hline
			$X_0 = 99999, Y_0 = 1$ & 1.000116 & 1.7e-05 & 1.000063 & 2.0e-06 & 1.000040 & 1.5e-05 & 1.000064 & 1.0e-06\\
			\hline
			$X_0 = 99990, Y_0 = 10$ & 1.000474 & 2.1e-05 & 1.000477 & 8.0e-06 & 1.000473 & 1.6e-05 & 1.000507 & 4.0e-06\\
			\hline
			$X_0 = 99900, Y_0 = 100$ & 1.004902 & 1.8e-05 & 1.004880 & 1.4e-05 & 1.004899 & 1.9e-05 & 1.004895 & 8.0e-06\\
			\hline
			$X_0 = 99000, Y_0 = 1000$ & 1.049155 & 7.4e-05 & 1.049349 & 6.6e-05 & 1.049271 & 6.9e-05 & 1.049277 & 6.6e-05\\
			\hline
		\end{tabular}
	\end{footnotesize}
\end{table}

\begin{table}[H]

	\caption{$\rr$ Estimates and Std. Errs, SIRSMAX Model,
		Different Variances, 
		$X_0 = 99000, Y_0 = 1000$}
	\begin{footnotesize}
		\hskip -1.7cm
		\begin{tabular}{l|r|r|r|r|r|r|r|r}
			\hline
			Data Set & Auto Est & Auto SE & AutoM Est & AutoM SE & Norm Est & Norm SE & NormM Est & NormM SE\\
			\hline
			$\sigma_X = 10, \sigma_Y = 1$ & 1.049351 & 0.000076 & 1.049219 & 0.000065 & 1.049294 & 0.000069 & 1.049283 & 0.000069\\
			\hline
			$\sigma_X = 100, \sigma_Y = 10$ & 1.049675 & 0.000154 & 1.049504 & 0.000132 & 1.049443 & 0.000182 & 1.049389 & 0.000131\\
			\hline
			$\sigma_X = 300, \sigma_Y = 30$ & 1.046190 & 0.000566 & 1.051219 & 0.000676 & 1.050137 & 0.000623 & 1.049671 & 0.000141\\
			\hline
			$\sigma_X = 1000, \sigma_Y = 100$ & 1.061907 & 0.002419 & 1.048793 & 0.000363 & 1.046010 & 0.001880 & 1.048815 & 0.000172\\
			\hline
			$\sigma_X = 3000, \sigma_Y = 300$ & 1.045117 & 0.005651 & 1.068145 & 0.004304 & 1.039168 & 0.005444 & 1.057089 & 0.000372\\
			\hline
			$\sigma_X = 10000, \sigma_Y = 1000$ & 0.979174 & 0.014889 & 1.111583 & 0.002687 & 1.039000 & 0.015012 & 1.070210 & 0.004720\\
			\hline
		\end{tabular}
	\end{footnotesize}
\end{table}

\begin{table}[H]

	\caption{$\rr$ Estimates and Std. Errs, SIRSMAX Model,
		Other Models, $X_0 = 99950, Y_0 = 50$, 
		$\sigma_X = 10, \sigma_Y = 1$}
	\begin{footnotesize}
		\hskip -1cm
		\begin{tabular}{l|r|r|r|r|r|r|r|r}
			\hline
			Data Set & Auto Est & Auto SE & AutoM Est & AutoM SE & Norm Est & Norm SE & NormM Est & NormM SE\\
			\hline
			0th & 1.000515 & 0.000022 & 1.000486 & 0.000000 & 1.000466 & 0.000015 & 1.000462 & 0.000003\\
			\hline
			1st & 1.083710 & 0.000021 & 1.083642 & 0.000032 & 1.083606 & 0.000022 & 1.083580 & 0.000018\\
			\hline
			2nd & 1.082579 & 0.000320 & 1.082543 & 0.000312 & 1.082600 & 0.000314 & 1.082560 & 0.000310\\
			\hline
			3rd & 1.089734 & 0.000559 & 1.089601 & 0.000550 & 1.089649 & 0.000554 & 1.089634 & 0.000553\\
			\hline
			4th & 1.079535 & 0.000524 & 1.079535 & 0.000524 & 1.079467 & 0.000516 & 1.079486 & 0.000511\\
			\hline
			Linear SIR & 1.112544 & 0.001242 & 1.112644 & 0.001239 & 1.112546 & 0.001238 & 1.112544 & 0.001236\\
			\hline
		\end{tabular}
	\end{footnotesize}
\end{table}

\subsubsection{SIR Incidence-Prevalence Ratio}

\paragraph{Estimation}

We use basic arithmetic functions in R to calculate the incidence to prevalence ratio for each observation in time and to derive the $\rr$ estimate.

I don't currently have a standard error calculation. One thing we could do is use the standard deviation of the $\frac{IPR(t)}{D}$ values for across all $t$, since we take the mean of them as the estimate for $\rr$.

\paragraph{Comments}

The estimates for $\rr$ seem to be very imprecise for this model. Looking at the baseline data, the $\rr$ estimates vary fairly widely given that the 5 data sets were generated under the same conditions. We also see that the estimates change massively as we change $Y_0$ or $\sigma_X$ and $\sigma_Y$. As we see below, the estimate for $\rr$ drops from around 45 for $Y_0 = 1$ to 0.9 for $Y_0 = 1000$. Similarly, the estimate for $\rr$ grows from around 0.9 for ($\sigma_X$, $\sigma_Y$) = (10, 1) to 200 for ($\sigma_X$, $\sigma_Y$) = (15000, 1500). In general, we see that the estimates for $\rr$ decrease as $Y_0$ increases relative to the total population and that they increase as we increase 
$\sigma_X$ and $\sigma_Y$.

\begin{table}[H]

	\caption{$\rr$ Estimates and Std. Errs, SIRIPR Model,
		Baseline Data, $X_0 = 99950, Y_0 = 50$, 
		$\sigma_X = 10, \sigma_Y = 1$}
	\begin{footnotesize}
		\hskip -1cm
		\begin{tabular}{l|r|r|r|r|r|r|r|r}
			\hline
			Data Set & Auto Est & Auto SE & AutoM Est & AutoM SE & Norm Est & Norm SE & NormM Est & NormM SE\\
			\hline
			Data Set & Auto Est & Auto SE & AutoM Est & AutoM SE & Norm Est & Norm SE & NormM Est & NormM SE\\
			\hline
			Baseline 1 & 1.170709 & 1.726236 & 0.118299 & 0.501111 & 1.566640 & 2.461139 & 0.129870 & 0.410478\\
			\hline
			Baseline 2 & 1.169642 & 1.736295 & 0.134667 & 0.531973 & 1.405544 & 1.947442 & 0.143524 & 0.470087\\
			\hline
			Baseline 3 & 1.024440 & 1.566376 & 0.147574 & 0.616955 & 1.327697 & 2.039847 & 0.135604 & 0.415505\\
			\hline
			Baseline 4 & 1.078543 & 1.635074 & 0.126040 & 0.518177 & 1.538235 & 2.350329 & 0.136590 & 0.461818\\
			\hline
			Baseline 5 & 1.082233 & 1.600320 & 0.130253 & 0.464518 & 1.287763 & 2.121584 & 0.124392 & 0.379142\\
			\hline
		\end{tabular}
	\end{footnotesize}
\end{table}

\begin{table}[H]

	\caption{$\rr$ Estimates and Std. Errs, SIRIPR Model,
		Different Starting Populations, 
		$\sigma_X = 10, \sigma_Y = 1$}
	\begin{footnotesize}
		\hskip -1.7cm
		\begin{tabular}{l|r|r|r|r|r|r|r|r}
			\hline
			Data Set & Auto Est & Auto SE & AutoM Est & AutoM SE & Norm Est & Norm SE & NormM Est & NormM SE\\
			\hline
			$X_0 = 99999, Y_0 = 1$ & 132.003557 & 979.057271 & 0.637136 & 7.591563 & 571.022230 & 8311.989460 & 0.917525 & 10.386777\\
			\hline
			$X_0 = 99990, Y_0 = 10$ & 5.566424 & 8.249684 & 0.169114 & 1.056522 & 6.817065 & 11.495275 & 0.152642 & 0.759693\\
			\hline
			$X_0 = 99900, Y_0 = 100$ & 0.595913 & 0.859870 & 0.129120 & 0.359819 & 0.764240 & 1.162178 & 0.121305 & 0.305982\\
			\hline
			$X_0 = 99000, Y_0 = 1000$ & 0.132387 & 0.115126 & 0.117280 & 0.105368 & 0.146490 & 0.145274 & 0.117695 & 0.116146\\
			\hline
		\end{tabular}
	\end{footnotesize}
\end{table}

\begin{table}[H]

	\caption{$\rr$ Estimates and Std. Errs, SIRIPR Model,
		Different Variances, 
		$X_0 = 99000, Y_0 = 1000$}
	\begin{footnotesize}
		\hskip -1.7cm
		\begin{tabular}{l|r|r|r|r|r|r|r|r}
			\hline
			Data Set & Auto Est & Auto SE & AutoM Est & AutoM SE & Norm Est & Norm SE & NormM Est & NormM SE\\
			\hline
			$\sigma_X = 10, \sigma_Y = 1$ & 1.178938 & 20.02418 & 1.185635 & 20.439407 & 1.219909 & 20.39467 & 1.167064 & 20.057929\\
			\hline
			$\sigma_X = 100, \sigma_Y = 10$ & 1.356012 & 15.38756 & 1.089050 & 18.415936 & 1.859016 & 20.12416 & 1.150239 & 19.649800\\
			\hline
			$\sigma_X = 300, \sigma_Y = 30$ & 3.142258 & 28.52150 & 0.969976 & 15.934410 & 3.201584 & 21.35673 & 1.085100 & 18.379016\\
			\hline
			$\sigma_X = 1000, \sigma_Y = 100$ & 7.305927 & 39.63369 & 1.088018 & 18.236155 & 7.978626 & 16.64830 & 0.787073 & 12.417881\\
			\hline
			$\sigma_X = 3000, \sigma_Y = 300$ & 16.009523 & 27.38954 & 0.424419 & 3.054379 & 30.351952 & 130.75442 & 1.357313 & 23.125797\\
			\hline
			$\sigma_X = 10000, \sigma_Y = 1000$ & 142.030295 & 894.32779 & 0.718062 & 8.362043 & 174.629326 & 706.02904 & 0.390140 & 3.444363\\
			\hline
		\end{tabular}
	\end{footnotesize}
\end{table}

\begin{table}[H]

	\caption{$\rr$ Estimates and Std. Errs, SIRIPR Model,
		Other Models, $X_0 = 99950, Y_0 = 50$, 
		$\sigma_X = 10, \sigma_Y = 1$}
	\begin{footnotesize}
		\hskip -1cm
		\begin{tabular}{l|r|r|r|r|r|r|r|r}
			\hline
			Data Set & Auto Est & Auto SE & AutoM Est & AutoM SE & Norm Est & Norm SE & NormM Est & NormM SE\\
			\hline
			0th & 1.779121 & 2.482162 & 0.009267 & 0.089052 & 2.352857 & 3.236806 & 0.021922 & 0.200162\\
			\hline
			1st & 3.392178 & 21.127161 & 5.199764 & 44.528822 & 4.615877 & 45.803986 & 4.032633 & 33.360591\\
			\hline
			2nd & 1.453143 & 5.808155 & 1.073020 & 2.586030 & 1.608484 & 5.985549 & 1.076302 & 2.657303\\
			\hline
			3rd & 1.360154 & 1.299049 & 1.192788 & 1.033645 & 1.453000 & 1.464758 & 1.195854 & 1.134779\\
			\hline
			4th & 1.300569 & 1.215989 & 1.300569 & 1.215989 & 1.426296 & 1.590604 & 1.153380 & 1.398890\\
			\hline
			Linear SIR & 1.379871 & 1.311406 & 1.208521 & 1.003065 & 1.516803 & 1.554241 & 1.208578 & 1.125440\\
			\hline
		\end{tabular}
	\end{footnotesize}
\end{table}

\subsubsection{SIR Smoothed Incidence-Prevalence Ratio}

\paragraph{Estimation}

This estimation method is similar to the incidence-prevalence ratio estimator above, except that we first smooth the data for the number of susceptible and infected using a smoothing spline. We use basic arithmetic functions in R to calculate the incidence to prevalence ratio for each observation in time and to derive the $\rr$ estimate.

We estimate the standard error using the jackknife method. For every observation, we fit a smoothed spline to X(t) and Y(t) after taking out that specific observation, then take the standard deviation of $t$ the $\rr$ estimates.

\paragraph{Comments}

Smoothing the data actually makes the $\rr$ estimates much more precise. All of them are now within $\pm$ 0.5 of the true $\rr$ for the data generated from the classic SIR model. We still see the same pattern as in the non-smoothed incidence-prevalence ratio estimates, that the estimates for $\rr$ decrease as $Y_0$ increases relative to the total population and that they increase as we increase $\sigma_X$ and $\sigma_Y$. For the data generated from other models, we see that there are some negative $\rr$ estimates for the 0th and 1st order linear model data. (This does not happen to be a problem with the non-smoothed IPR estimates.) This might be the results of having some negative incidences in the data affecting the results of the smoothing spline.

The standard errors for this model are large compared to the SIR models that are soled using optim() (such as least squares). We also see that the standard error increases dramatically as $\sigma_X$ and $\sigma_Y$ increase.

\begin{table}[H]

	\caption{$\rr$ Estimates and Std. Errs, SIRSIPR Model,
		Baseline Data, $X_0 = 99950, Y_0 = 50$, 
		$\sigma_X = 10, \sigma_Y = 1$}
	\begin{footnotesize}
		\hskip -1cm
		\begin{tabular}{l|r|r|r|r|r|r|r|r}
			\hline
			Data Set & Auto Est & Auto SE & AutoM Est & AutoM SE & Norm Est & Norm SE & NormM Est & NormM SE\\
			\hline
			Baseline 1 & 0.111589 & 0.013332 & 0.116714 & 0.004233 & 0.121731 & 0.008296 & 0.122992 & 0.002911\\
			\hline
			Baseline 2 & 0.113638 & 0.011308 & 0.115175 & 0.003540 & 0.119221 & 0.007058 & 0.120945 & 0.005199\\
			\hline
			Baseline 3 & 0.118958 & 0.009720 & 0.126134 & 0.011717 & 0.121108 & 0.007978 & 0.122298 & 0.004135\\
			\hline
			Baseline 4 & 0.129073 & 0.010889 & 0.125410 & 0.004537 & 0.120610 & 0.008410 & 0.120497 & 0.003618\\
			\hline
			Baseline 5 & 0.122424 & 0.011326 & 0.129906 & 0.004269 & 0.118352 & 0.007518 & 0.124123 & 0.002414\\
			\hline
		\end{tabular}
	\end{footnotesize}
\end{table}

\begin{table}[H]

	\caption{$\rr$ Estimates and Std. Errs, SIRSIPR Model,
		Different Starting Populations, 
		$\sigma_X = 10, \sigma_Y = 1$}
	\begin{footnotesize}
		\hskip -1.7cm
		\begin{tabular}{l|r|r|r|r|r|r|r|r}
			\hline
			Data Set & Auto Est & Auto SE & AutoM Est & AutoM SE & Norm Est & Norm SE & NormM Est & NormM SE\\
			\hline
			$X_0 = 99999, Y_0 = 1$ & 0.496556 & 0.404930 & 0.249343 & 0.080732 & 0.244347 & 0.534660 & 0.279914 & 0.121538\\
			\hline
			$X_0 = 99990, Y_0 = 10$ & 0.082031 & 0.045989 & 0.130633 & 0.007387 & 0.116519 & 0.052715 & 0.119705 & 0.006003\\
			\hline
			$X_0 = 99900, Y_0 = 100$ & 0.120003 & 0.006022 & 0.119230 & 0.003320 & 0.121229 & 0.004296 & 0.120600 & 0.001531\\
			\hline
			$X_0 = 99000, Y_0 = 1000$ & 0.117786 & 0.001351 & 0.117888 & 0.001247 & 0.117755 & 0.001314 & 0.117804 & 0.001219\\
			\hline
		\end{tabular}
	\end{footnotesize}
\end{table}

\begin{table}[H]

	\caption{$\rr$ Estimates and Std. Errs, SIRSIPR Model,
		Different Variances, 
		$X_0 = 99000, Y_0 = 1000$}
	\begin{footnotesize}
		\hskip -1.7cm
		\begin{tabular}{l|r|r|r|r|r|r|r|r}
			\hline
			Data Set & Auto Est & Auto SE & AutoM Est & AutoM SE & Norm Est & Norm SE & NormM Est & NormM SE\\
			\hline
			$\sigma_X = 10, \sigma_Y = 1$ & 0.119798 & 0.015080 & 0.119388 & 0.015098 & 0.119629 & 0.015084 & 0.119720 & 0.015063\\
			\hline
			$\sigma_X = 100, \sigma_Y = 10$ & 0.123180 & 0.015455 & 0.120830 & 0.015114 & 0.119951 & 0.015663 & 0.118726 & 0.015203\\
			\hline
			$\sigma_X = 300, \sigma_Y = 30$ & 0.108430 & 0.022619 & 0.119020 & 0.014395 & 0.118138 & 0.018715 & 0.120683 & 0.015369\\
			\hline
			$\sigma_X = 1000, \sigma_Y = 100$ & 0.151905 & 0.067340 & 0.122619 & 0.016990 & 0.120618 & 0.053277 & 0.138444 & 0.013843\\
			\hline
			$\sigma_X = 3000, \sigma_Y = 300$ & 0.148289 & 0.141669 & 0.237888 & 0.054042 & 0.121580 & 0.147039 & 0.135574 & 0.015871\\
			\hline
			$\sigma_X = 10000, \sigma_Y = 1000$ & -0.253735 & 0.457021 & 0.636736 & 0.092124 & 0.113880 & 0.555507 & 0.208459 & 0.018114\\
			\hline
		\end{tabular}
	\end{footnotesize}
\end{table}

\begin{table}[H]

	\caption{$\rr$ Estimates and Std. Errs, SIRSIPR Model,
		Other Models, $X_0 = 99950, Y_0 = 50$, 
		$\sigma_X = 10, \sigma_Y = 1$}
	\begin{footnotesize}
		\hskip -1cm
		\begin{tabular}{l|r|r|r|r|r|r|r|r}
			\hline
			Data Set & Auto Est & Auto SE & AutoM Est & AutoM SE & Norm Est & Norm SE & NormM Est & NormM SE\\
			\hline
			0th & 0.014270 & 0.013561 & 0.002865 & 0.000486 & -0.006574 & 0.010632 & 0.005140 & 0.001792\\
			\hline
			1st & -9.099024 & 94554.286222 & 2.655271 & 153.459186 & 2.339653 & 190.604213 & 2.285211 & 200.382597\\
			\hline
			2nd & 1.045203 & 0.201211 & 1.076390 & 0.156022 & 1.042439 & 0.205421 & 1.074918 & 0.152346\\
			\hline
			3rd & 1.279959 & 0.299658 & 1.290916 & 0.315063 & 1.279248 & 0.295345 & 1.280916 & 0.294307\\
			\hline
			4th & 1.192312 & 0.187455 & 1.192312 & 0.187455 & 1.203649 & 0.200467 & 1.202999 & 0.198628\\
			\hline
			Linear SIR & 1.296731 & 0.221560 & 1.304489 & 0.229988 & 1.294079 & 0.215113 & 1.294364 & 0.215810\\
			\hline
		\end{tabular}
	\end{footnotesize}
\end{table}

\subsubsection{SIR Log Linear}

\paragraph{Estimation}

We use lm() in R to fit a linear regression model to $log(\frac{X(t)}{X(0)})$ on $Z(t)$. We use the coefficient estimate and standard error from lm() as our estimate of and standard error of $\rr$.

\paragraph{Comments}

The $\rr$ estimates here are very accurate and precise. The estimates from the baseline and changing $Y_0$ data sets are all very close to the true $\rr$ value. Increasing $\sigma_X$ and $\sigma_Y$ does decrease the model accuracy. It is not quite as good for data generated from other models, but it gives reasonable results for the 0th and 1st order linear models unlike some of the other models.

The standard errors for the model are relatively small as well, and there are no unusually large values. They tend to decrease as $Y_0$ increases and they tend to increase as $\sigma_X$ and $\sigma_Y$ increase.

\begin{table}[H]

	\caption{$\rr$ Estimates and Std. Errs, HARKO Model, 
		Baseline Data, $X_0 = 99950, Y_0 = 50$, 
		$\sigma_X = 10, \sigma_Y = 1$}
	\begin{footnotesize}
		\hskip -1cm
		\begin{tabular}{l|r|r|r|r|r|r|r|r}
			\hline
			Data Set & Auto Est & Auto SE & AutoM Est & AutoM SE & Norm Est & Norm SE & NormM Est & NormM SE\\
			\hline
			Baseline 1 & 1.780655 & 0.023436 & 1.983971 & 0.006583 & 1.887483 & 0.016523 & 1.977806 & 0.004671\\
			\hline
			Baseline 2 & 1.831444 & 0.021164 & 2.002378 & 0.010261 & 1.902733 & 0.017365 & 2.003881 & 0.003998\\
			\hline
			Baseline 3 & 1.839113 & 0.022491 & 1.914380 & 0.010098 & 1.906137 & 0.016586 & 1.987737 & 0.003900\\
			\hline
			Baseline 4 & 1.794680 & 0.017916 & 1.840147 & 0.007965 & 1.894689 & 0.016695 & 1.990857 & 0.003619\\
			\hline
			Baseline 5 & 1.823206 & 0.020518 & 1.892921 & 0.011397 & 1.914908 & 0.016957 & 1.965487 & 0.003611\\
			\hline
		\end{tabular}
	\end{footnotesize}
\end{table}

\begin{table}[H]

	\caption{$\rr$ Estimates and Std. Errs, HARKO Model,
		Different Starting Populations, 
		$\sigma_X = 10, \sigma_Y = 1$}
	\begin{footnotesize}
		\hskip -1.7cm
		\begin{tabular}{l|r|r|r|r|r|r|r|r}
			\hline
			Data Set & Auto Est & Auto SE & AutoM Est & AutoM SE & Norm Est & Norm SE & NormM Est & NormM SE\\
			\hline
			$X_0 = 99999, Y_0 = 1$ & 0.990097 & 0.006292 & 1.257974 & 0.013850 & 1.007582 & 0.005853 & 1.191026 & 0.003584\\
			\hline
			$X_0 = 99990, Y_0 = 10$ & 1.123018 & 0.021783 & 1.735611 & 0.013290 & 1.245655 & 0.024246 & 1.923291 & 0.010741\\
			\hline
			$X_0 = 99900, Y_0 = 100$ & 1.951514 & 0.012785 & 2.016594 & 0.005770 & 1.966805 & 0.008724 & 1.983918 & 0.002413\\
			\hline
			$X_0 = 99000, Y_0 = 1000$ & 2.006289 & 0.001318 & 1.995221 & 0.001014 & 2.000268 & 0.000951 & 2.000158 & 0.000544\\
			\hline
		\end{tabular}
	\end{footnotesize}
\end{table}

\begin{table}[H]

	\caption{$\rr$ Estimates and Std. Errs, HARKO Model, 
		Different Variances, $X_0 = 99000$, $Y_0 = 1000$}
	\begin{footnotesize}
		\hskip -1.7cm
		\begin{tabular}{l|r|r|r|r|r|r|r|r}
			\hline
			Data Set & Auto Est & Auto SE & AutoM Est & AutoM SE & Norm Est & Norm SE & NormM Est & NormM SE\\
			\hline
			$\sigma_X = 10, \sigma_Y = 1$ & 1.997497 & 0.001357 & 2.002070 & 0.001141 & 2.000316 & 0.001015 & 2.000543 & 0.000521\\
			\hline
			$\sigma_X = 100, \sigma_Y = 10$ & 1.918862 & 0.009993 & 1.990296 & 0.004519 & 1.968331 & 0.009515 & 2.002762 & 0.002228\\
			\hline
			$\sigma_X = 300, \sigma_Y = 30$ & 1.688924 & 0.027688 & 1.822846 & 0.012972 & 1.786604 & 0.021708 & 1.999747 & 0.003704\\
			\hline
			$\sigma_X = 1000, \sigma_Y = 100$ & 1.202346 & 0.017004 & 1.860911 & 0.008679 & 1.244529 & 0.023070 & 1.863444 & 0.009948\\
			\hline
			$\sigma_X = 3000, \sigma_Y = 300$ & 1.043406 & 0.009591 & 1.414679 & 0.009365 & 1.027725 & 0.010276 & 1.696223 & 0.003871\\
			\hline
			$\sigma_X = 10000, \sigma_Y = 1000$ & 1.044556 & 0.008612 & 1.093328 & 0.004175 & 1.021139 & 0.007466 & 1.203591 & 0.007232\\
			\hline
		\end{tabular}
	\end{footnotesize}
\end{table}

\begin{table}[H]

	\caption{$\rr$ Estimates and Std. Errs, HARKO Model,
		Other Models, $X_0 = 99950, Y_0 = 50$, 
		$\sigma_X = 10, \sigma_Y = 1$}
	\begin{footnotesize}
		\hskip -1cm
		\begin{tabular}{l|r|r|r|r|r|r|r|r}
			\hline
			Data Set & Auto Est & Auto SE & AutoM Est & AutoM SE & Norm Est & Norm SE & NormM Est & NormM SE\\
			\hline
			0th & 0.992417 & 0.005898 & 0.801563 & 0.009989 & 0.984522 & 0.004772 & 0.960479 & 0.003184\\
			\hline
			1st & 1.200537 & 0.000641 & 1.200834 & 0.000664 & 1.201039 & 0.000648 & 1.200963 & 0.000653\\
			\hline
			2nd & 1.198169 & 0.000451 & 1.198486 & 0.000407 & 1.198510 & 0.000451 & 1.198652 & 0.000403\\
			\hline
			3rd & 1.207309 & 0.000196 & 1.207754 & 0.000188 & 1.207402 & 0.000184 & 1.207447 & 0.000187\\
			\hline
			4th & 1.210499 & 0.000136 & 1.210499 & 0.000136 & 1.210671 & 0.000121 & 1.210571 & 0.000114\\
			\hline
			Linear SIR & 1.258920 & 0.000994 & 1.258462 & 0.001020 & 1.258587 & 0.001008 & 1.258670 & 0.001007\\
			\hline
		\end{tabular}
	\end{footnotesize}
\end{table}

\subsubsection{SIR Markov Chain}

\paragraph{Estimation}

I'm not sure what I did was correct in this case. Based on what I understood from the paper, the method I came up with for estimating $\rr$ here was using lm() in R. My logic in algebra:

$$X_{t+1} \sim Binom(X_{t}, \alpha^Y)$$ $$ X_{t+1} = X_{t} * \alpha^Y$$ $$X _{t+1} / X_{t} = \alpha^Y$$ $$log(X _{t+1}) - log(X_{t}) = Y * log(\alpha)$$

We would regress $log(X _{t+1}) - log(X_{t})$ on $Y$. However, that regression function returned unreasonable estimates for $\rr$. I then attempted scale $Y$ by the number of contacts per day of each member of the susceptible compartment such that $Y^*$ would be the expected number of infectious people contacted for the day (i.e., $log(X _{t+1}) - log(X_{t}) = Y^* log(\alpha)$ where $Y^* = Y/N * contacts$) The results below were on the assumption that each susceptible person would contact one other person per time step (i.e., $Y^* = Y/N$).

\begin{table}[H]

	\caption{$\rr$ Estimates and Std. Errs, SIRMC Model,
		Baseline Data, $X_0 = 99950, Y_0 = 50$, 
		$\sigma_X = 10, \sigma_Y = 1$}
	\begin{footnotesize}
		\hskip -1cm
		\begin{tabular}{l|r|r|r|r|r|r|r|r}
			\hline
			Data Set & Auto Est & Auto SE & AutoM Est & AutoM SE & Norm Est & Norm SE & NormM Est & NormM SE\\
			\hline
			Baseline 1 & 5.127170 & 179.2844 & 5.234313 & 40.61000 & 5.086225 & 216.3083 & 5.100115 & 26.56028\\
			\hline
			Baseline 2 & 5.158918 & 189.7917 & 4.921625 & 26.13471 & 5.126360 & 217.2991 & 5.074875 & 29.08225\\
			\hline
			Baseline 3 & 5.160983 & 169.1971 & 5.207777 & 47.15165 & 5.083438 & 195.1501 & 5.119470 & 28.63475\\
			\hline
			Baseline 4 & 5.161219 & 170.6186 & 5.127197 & 34.24464 & 5.207647 & 269.4176 & 5.135429 & 31.76258\\
			\hline
			Baseline 5 & 5.187114 & 187.2816 & 5.131911 & 31.43095 & 5.077370 & 186.0883 & 5.098254 & 23.90021\\
			\hline
		\end{tabular}
	\end{footnotesize}
\end{table}

\begin{table}[H]

	\caption{$\rr$ Estimates and Std. Errs, SIRMC Model,
		Different Starting Populations, 
		$\sigma_X = 10, \sigma_Y = 1$}
	\begin{footnotesize}
		\hskip -1.7cm
		\begin{tabular}{l|r|r|r|r|r|r|r|r}
			\hline
			Data Set & Auto Est & Auto SE & AutoM Est & AutoM SE & Norm Est & Norm SE & NormM Est & NormM SE\\
			\hline
			$X_0 = 99999, Y_0 = 1$ & 2.502983 & 31.703174 & 4.587050 & 141.618120 & NaN & 1.53910 & 4.584606 & 211.645976\\
			\hline
			$X_0 = 99990, Y_0 = 10$ & 5.370030 & 1402.395267 & 4.745881 & 37.893736 & 5.079299 & 1000.80267 & 4.808880 & 30.644371\\
			\hline
			$X_0 = 99900, Y_0 = 100$ & 5.083530 & 78.047548 & 5.141349 & 24.913580 & 5.036936 & 94.46507 & 5.137375 & 21.960267\\
			\hline
			$X_0 = 99000, Y_0 = 1000$ & 5.134886 & 9.346665 & 5.112739 & 7.046171 & 5.112847 & 11.30402 & 5.113115 & 7.906187\\
			\hline
		\end{tabular}
	\end{footnotesize}
\end{table}

\begin{table}[H]

	\caption{$\rr$ Estimates and Std. Errs, SIRMC Model,
		Different Variances, 
		$X_0 = 99000, Y_0 = 1000$}
	\begin{footnotesize}
		\hskip -1.7cm
		\begin{tabular}{l|r|r|r|r|r|r|r|r}
			\hline
			Data Set & Auto Est & Auto SE & AutoM Est & AutoM SE & Norm Est & Norm SE & NormM Est & NormM SE\\
			\hline
			$\sigma_X = 10, \sigma_Y = 1$ & 5.107658 & 40.07322 & 5.118340 & 41.49349 & 5.112780 & 42.18333 & 5.109027 & 40.04422\\
			\hline
			$\sigma_X = 100, \sigma_Y = 10$ & 5.047351 & 74.96928 & 5.074418 & 40.22261 & 5.055666 & 112.53820 & 5.060957 & 39.62932\\
			\hline
			$\sigma_X = 300, \sigma_Y = 30$ & 5.183856 & 304.55113 & 4.929632 & 34.41019 & 5.459944 & 664.73372 & 5.129073 & 49.24831\\
			\hline
			$\sigma_X = 1000, \sigma_Y = 100$ & 4.267588 & 158.79219 & 5.122059 & 60.44409 & 5.741570 & 4073.99286 & 5.133013 & 56.70565\\
			\hline
			$\sigma_X = 3000, \sigma_Y = 300$ & NaN & 74731.85279 & 3.918439 & 18.49635 & 3.200282 & 80.21230 & 4.771752 & 45.08632\\
			\hline
			$\sigma_X = 10000, \sigma_Y = 1000$ & 2.120699 & 17.46286 & 3.767783 & 35.76247 & NaN & 83.73480 & 3.884264 & 23.04767\\
			\hline
		\end{tabular}
	\end{footnotesize}
\end{table}

\begin{table}[H]

	\caption{$\rr$ Estimates and Std. Errs, SIRMC Model,
		Other Models, $X_0 = 99950, Y_0 = 50$, 
		$\sigma_X = 10, \sigma_Y = 1$}
	\begin{footnotesize}
		\hskip -1cm
		\begin{tabular}{l|r|r|r|r|r|r|r|r}
			\hline
			Data Set & Auto Est & Auto SE & AutoM Est & AutoM SE & Norm Est & Norm SE & NormM Est & NormM SE\\
			\hline
			0th & NaN & 6.511016e+05 & 7.976281 & 2207.762640 & NaN & 3.490885e+05 & 7.207505 & 1149.944668\\
			\hline
			1st & 3.436204 & 1.350819e+00 & 3.433055 & 1.341851 & 3.437569 & 1.599346e+00 & 3.444295 & 1.441849\\
			\hline
			2nd & 2.934223 & 6.682100e-01 & 2.939971 & 0.443031 & 2.938814 & 7.756630e-01 & 2.939860 & 0.504643\\
			\hline
			3rd & 2.859786 & 3.352930e-01 & 2.867682 & 0.292399 & 2.864071 & 4.244820e-01 & 2.865659 & 0.388652\\
			\hline
			4th & 2.907566 & 3.910230e-01 & 2.907566 & 0.391023 & 2.912020 & 5.038200e-01 & 2.912098 & 0.438172\\
			\hline
			Linear SIR & 2.762911 & 2.063760e-01 & 2.762471 & 0.193776 & 2.762182 & 2.720380e-01 & 2.764241 & 0.229030\\
			\hline
		\end{tabular}
	\end{footnotesize}
\end{table}

\subsubsection{SIR Sequential Bayes}

\paragraph{Estimation}

For the results below, I assume that the prior distribution is distributed as Gamma(1,1). Thus, multiplying by the Poisson likelihood will give us a posterior that is also gamma distributed. The distribution is updated for every day in the data. The final estimate and standard error for $\rr$ are the mean and standard deviation of the final posterior distribution, respectively. In this case, these are the expected value and standard deviation of the final gamma distribution.

\paragraph{Comments}

The $\rr$ estimates and standard errors are dependent on the choice of prior distribution and parameters. The Gamma(1,1) was chosen only for mathematical simplicity, and other choices could give us vastly different results.

The $\rr$ estimates here are all fairly reasonable, though they are not as close to the true $\rr$ as some of the other SIR models. Also, in general, the model seems to underestimate $\rr$. There is a spread in the estimates for the baseline data of around 0.2. The estimates appear to increase as $\sigma_X$ and $\sigma_Y$ increase. On the other hand, there does not seem to be a consistent effect with changing $Y_0$

The standard errors are larger than some of the other SIR models, especially those that use lm() or optim() for estimation. There does not seem to be any noticeable patterns with changing the variance of the errors or the starting compartment populations.

\begin{table}[H]

	\caption{$\rr$ Estimates and Std. Errs, SEQBAYES Model, 
		Baseline Data, $X_0 = 99950, Y_0 = 50$, 
		$\sigma_X = 10, \sigma_Y = 1$}
	\begin{footnotesize}
		\hskip -1cm
		\begin{tabular}{l|r|r|r|r|r|r|r|r}
			Data Set & Auto Est & Auto SE & AutoM Est & AutoM SE & Norm Est & Norm SE & NormM Est & NormM SE\\
			\hline
			Baseline 1 & 1.120551 & 0.055408 & 0.834692 & 0.047821 & 1.085113 & 0.054524 & 1.000000 & 0.052342\\
			\hline
			Baseline 2 & 1.163677 & 0.056464 & 0.963486 & 0.051378 & 1.086715 & 0.054565 & 1.266142 & 0.058897\\
			\hline
			Baseline 3 & 1.006208 & 0.052505 & 1.000000 & 0.052342 & 1.000000 & 0.052342 & 1.200584 & 0.057352\\
			\hline
			Baseline 4 & 0.874006 & 0.048934 & 1.200129 & 0.057341 & 0.840416 & 0.047984 & 1.000000 & 0.052342\\
			\hline
			Baseline 5 & 1.000000 & 0.052342 & 1.000000 & 0.052342 & 1.200097 & 0.057341 & 1.169565 & 0.056606\\
			\hline
		\end{tabular}
	\end{footnotesize}
\end{table}

\begin{table}[H]

	\caption{$\rr$ Estimates and Std. Errs, SEQBAYES Model,
		Different Starting Populations, 
		$\sigma_X = 10, \sigma_Y = 1$}
	\begin{footnotesize}
		\hskip -1.7cm
		\begin{tabular}{l|r|r|r|r|r|r|r|r}
			\hline
			Data Set & Auto Est & Auto SE & AutoM Est & AutoM SE & Norm Est & Norm SE & NormM Est & NormM SE\\
			\hline
			$X_0 = 99999, Y_0 = 1$ & 1.284355 & 0.059319 & 1.000000 & 0.052342 & 1.255944 & 0.058660 & 1.129035 & 0.055617\\
			\hline
			$X_0 = 99990, Y_0 = 10$ & 0.745920 & 0.045206 & 1.000000 & 0.052342 & 1.000000 & 0.052342 & 1.241239 & 0.058315\\
			\hline
			$X_0 = 99900, Y_0 = 100$ & 0.716836 & 0.044316 & 1.236492 & 0.058204 & 1.064823 & 0.054012 & 1.000000 & 0.052342\\
			\hline
			$X_0 = 99000, Y_0 = 1000$ & 0.753867 & 0.045447 & 1.158019 & 0.056326 & 1.048705 & 0.053602 & 1.322988 & 0.060205\\
			\hline
		\end{tabular}
	\end{footnotesize}
\end{table}

\begin{table}[H]

	\caption{$\rr$ Estimates and Std. Errs, SEQBAYES Model, 
		Different Variances, $X_0 = 99000$, $Y_0 = 1000$}
	\begin{footnotesize}
		\hskip -1.7cm
		\begin{tabular}{l|r|r|r|r|r|r|r|r}
			\hline
			Data Set & Auto Est & Auto SE & AutoM Est & AutoM SE & Norm Est & Norm SE & NormM Est & NormM SE\\
			\hline
			$\sigma_X = 10, \sigma_Y = 1$ & 0.793499 & 0.046626 & 0.787881 & 0.046461 & 0.762945 & 0.045719 & 0.796512 & 0.046714\\
			\hline
			$\sigma_X = 100, \sigma_Y = 10$ & 0.925325 & 0.050350 & 0.891440 & 0.049420 & 0.888104 & 0.049327 & 0.885803 & 0.049263\\
			\hline
			$\sigma_X = 300, \sigma_Y = 30$ & 0.478432 & 0.036205 & 0.510298 & 0.037391 & 0.494531 & 0.036809 & 0.502460 & 0.037103\\
			\hline
			$\sigma_X = 1000, \sigma_Y = 100$ & 0.897133 & 0.049577 & 0.973510 & 0.051644 & 0.524686 & 0.037914 & 0.524018 & 0.037890\\
			\hline
			$\sigma_X = 3000, \sigma_Y = 300$ & 1.000000 & 0.052342 & 1.000000 & 0.052342 & 1.024154 & 0.052971 & 0.984151 & 0.051926\\
			\hline
			$\sigma_X = 10000, \sigma_Y = 1000$ & 1.686398 & 0.067973 & 1.000000 & 0.052342 & 0.379716 & 0.032254 & 1.000000 & 0.052342\\
			\hline
		\end{tabular}
	\end{footnotesize}
\end{table}

\begin{table}[H]

	\caption{$\rr$ Estimates and Std. Errs, SEQBAYES Model,
		Other Models, $X_0 = 99950, Y_0 = 50$, 
		$\sigma_X = 10, \sigma_Y = 1$}
	\begin{footnotesize}
		\hskip -1cm
		\begin{tabular}{l|r|r|r|r|r|r|r|r}
			\hline
			Data Set & Auto Est & Auto SE & AutoM Est & AutoM SE & Norm Est & Norm SE & NormM Est & NormM SE\\
			\hline
			0th & 0.751296 & 0.045369 & 1.177502 & 0.056798 & 1.000000 & 0.052342 & 1.234447 & 0.058155\\
			\hline
			1st & 0.997717 & 0.052283 & 0.999939 & 0.052341 & 1.251535 & 0.058556 & 0.895793 & 0.049540\\
			\hline
			2nd & 0.919672 & 0.050196 & 0.946849 & 0.050932 & 0.899505 & 0.049643 & 0.935652 & 0.050630\\
			\hline
			3rd & 1.368874 & 0.061240 & 1.343871 & 0.060678 & 1.359729 & 0.061035 & 1.365769 & 0.061171\\
			\hline
			4th & 1.357390 & 0.060983 & 1.357390 & 0.060983 & 1.346107 & 0.060729 & 1.345567 & 0.060716\\
			\hline
			Linear SIR & 1.123600 & 0.055483 & 1.156827 & 0.056297 & 1.121743 & 0.055437 & 1.194507 & 0.057207\\
			\hline
		\end{tabular}
	\end{footnotesize}
\end{table}

\subsubsection{SEIR Least Squares}

\paragraph{Estimation}

The parameters of the SEIR model ($\beta$, $\gamma$, and $\alpha$) were estimated using optim() in R to minimize the sum of squared errors using Nelder-Mead. The number of iterations was limited to 50 so that the estimate for $\alpha$ did not reach infinity and break the function. The covariance matrix of the estimates of $\alpha$, $\beta$, an $\gamma$ were calculated using sensitivity analysis. The standard error for $\rr$ was then calculated by resampling $\beta$ and $\gamma$ from a multivariate normal distribution with the aforementioned covariance matrix from sensitivity analysis, then taking the standard deviation of the new $\rr$ estimates derived from the set of resampled parameters.

\paragraph{Comments}

In general, we might have issues with fitting the SEIR model to data generated from the classic SIR model, since $\alpha$ will go off to infinity with too many iterations of unconstrained optimization.

The estimates from the SEIR model are relatively consistent for the baseline data, but they are all slightly larger than the true $\rr$ value. The estimates tend to become more accurate as $Y_0$ increases relative to the total population. However, we can see that increasing the variance of the added errors to the simulated data dramatically decreases the accuracy and precision of the $\rr$ estimates. This includes estimates that are negative, estimates that are orders of magnitude larger than they should be, and estimates that are NA.

The standard errors are an order of magnitude larger than for the SIR model. They are fairly consistent between the multiple baseline set. The standard errors decrease dramatically as $Y_0$ increases. Also, as we see from the estimates above, the standard deviations fluctuate by several orders of magnitude as we change the variance of the added errors.

\begin{table}[H]

	\caption{$\rr$ Estimates and Std. Errs, SEIR Model, 
		Baseline Data, $X_0 = 99950, Y_0 = 50$, 
		$\sigma_X = 10, \sigma_Y = 1$}
	\begin{footnotesize}
		\hskip -1cm
		\begin{tabular}{l|r|r|r|r|r|r|r|r}
			\hline
			Data Set & Auto Est & Auto SE & AutoM Est & AutoM SE & Norm Est & Norm SE & NormM Est & NormM SE\\
			\hline
			Baseline 1 & 1.546608 & 0.171494 & 1.326568 & 0.013100 & 3.206236 & 2.451373 & 2.455754 & 0.066867\\
			\hline
			Baseline 2 & 1.772431 & 0.354807 & -0.080128 & 0.031367 & -12.937346 & 150.847519 & 3.675647 & 0.123981\\
			\hline
			Baseline 3 & 1.781222 & 0.395661 & 1.375838 & 0.024409 & 2.482198 & 1.391814 & 2.149920 & 0.029209\\
			\hline
			Baseline 4 & 1.636096 & 0.594579 & 1.737715 & 0.020414 & 0.616769 & 9.237073 & 1.982802 & 0.025452\\
			\hline
			Baseline 5 & -10.920596 & 123.661450 & 1.846941 & 0.027824 & 4.892252 & 12.254994 & 2.457392 & 0.039093\\
			\hline
		\end{tabular}
	\end{footnotesize}
\end{table}

\begin{table}[H]

	\caption{$\rr$ Estimates and Std. Errs, SEIR Model,
		Different Starting Populations, 
		$\sigma_X = 10, \sigma_Y = 1$}
	\begin{footnotesize}
		\hskip -1.7cm
		\begin{tabular}{l|r|r|r|r|r|r|r|r}
			\hline
			Data Set & Auto Est & Auto SE & AutoM Est & AutoM SE & Norm Est & Norm SE & NormM Est & NormM SE\\
			\hline
			$X_0 = 99999, Y_0 = 1$ & NA & NA & -0.182830 & 0.225877 & NA & NA & -0.484025 & 0.614292\\
			\hline
			$X_0 = 99990, Y_0 = 10$ & 0.973745 & 0.546827 & NA & NA & 1.052351 & 444.645186 & NA & NA\\
			\hline
			$X_0 = 99900, Y_0 = 100$ & -6.416001 & 25.905717 & 2.451224 & 0.118072 & -3.124388 & 6.766299 & -1.475048 & 1.343083\\
			\hline
			$X_0 = 99000, Y_0 = 1000$ & -1.665131 & 1.139871 & 2.132087 & 0.009851 & 2.146884 & 0.018722 & 2.114676 & 0.010444\\
			\hline
		\end{tabular}
	\end{footnotesize}
\end{table}

\begin{table}[H]

	\caption{$\rr$ Estimates and Std. Errs, SEIR Model, 
		Different Variances, $X_0 = 99000$, $Y_0 = 1000$}
	\begin{footnotesize}
		\hskip -1.7cm
		\begin{tabular}{l|r|r|r|r|r|r|r|r}
			\hline
			Data Set & Auto Est & Auto SE & AutoM Est & AutoM SE & Norm Est & Norm SE & NormM Est & NormM SE\\
			\hline
			$\sigma_X = 10, \sigma_Y = 1$ & 19.226063 & 0.289480 & 21.685544 & 0.346537 & 18.468777 & 0.256194 & 23.787444 & 0.195589\\
			\hline
			$\sigma_X = 100, \sigma_Y = 10$ & 16.976120 & 0.530197 & 19.896647 & 0.226174 & 13.179504 & 0.570207 & 15.331206 & 0.206823\\
			\hline
			$\sigma_X = 300, \sigma_Y = 30$ & 9.925354 & 0.770734 & 16.239531 & 0.234976 & -4.473057 & 0.190822 & 22.862570 & 0.525209\\
			\hline
			$\sigma_X = 1000, \sigma_Y = 100$ & -0.007144 & 0.000245 & 25.560238 & 0.933150 & -4.082639 & 0.436790 & 11.261214 & 0.216712\\
			\hline
			$\sigma_X = 3000, \sigma_Y = 300$ & -0.262553 & 0.003032 & -0.012345 & 0.000093 & -0.000062 & 0.000004 & 12.886121 & 0.277206\\
			\hline
			$\sigma_X = 10000, \sigma_Y = 1000$ & -0.000217 & 0.000031 & -5.959291 & 0.225965 & NA & NA & -0.000064 & 0.000000\\
			\hline
		\end{tabular}
	\end{footnotesize}
\end{table}

\begin{table}[H]

	\caption{$\rr$ Estimates and Std. Errs, SEIR Model,
		Other Models, $X_0 = 99950, Y_0 = 50$, 
		$\sigma_X = 10, \sigma_Y = 1$}
	\begin{footnotesize}
		\hskip -1cm
		\begin{tabular}{l|r|r|r|r|r|r|r|r}
			\hline
			Data Set & Auto Est & Auto SE & AutoM Est & AutoM SE & Norm Est & Norm SE & NormM Est & NormM SE\\
			\hline
			0th & -1.034177 & 0.143642 & -0.055180 & 0.000244 & 0.055060 & 0.029086 & -0.165649 & 0.000683\\
			\hline
			1st & 1.040105 & 0.004954 & 1.190938 & 1.173853 & 1.043084 & 0.004767 & 1.040387 & 0.008804\\
			\hline
			2nd & 1.134694 & 0.000507 & 1.150572 & 0.000496 & 1.132432 & 0.000509 & 1.143565 & 0.000470\\
			\hline
			3rd & 1.303457 & 0.004479 & 1.288333 & 0.004306 & 1.289472 & 0.005720 & 1.287972 & 0.004548\\
			\hline
			4th & 1.264807 & 0.006362 & 1.253277 & 0.004502 & 1.271573 & 0.006568 & 1.247322 & 0.005298\\
			\hline
			Linear SIR & 1.495457 & 0.006932 & 1.496117 & 0.006626 & 1.493676 & 0.009262 & 1.488063 & 0.008507\\
			\hline
		\end{tabular}
	\end{footnotesize}
\end{table}

\subsubsection{MSEIR Least Squares}

\paragraph{Estimation}

The parameters of the MSEIR model were estimated by finding the values that minimizes the sum of squared errors using Nelder-Mead through the constrOptim() in R. This function is the same as optim(), except that it allows us to constrain the search space for the parameters by specifying a constraint matrix. In this case, all of the parameters were constrained to be positive, as they should be by definition. The maximum number of iterations for the optimization was chosen to be 50 in order to limit runtime.

The standard error was calculated using the delta method. We find the standard deviation of all of the parameters using the Hessian matrix, then use that as the covariance matrix in our delta method calculations.

\paragraph{Comments}

The estimates of $\rr$ are very precise under the same conditions for the classic SIR model. The change in $\sigma_X$ and $\sigma_Y$ do not change the estimates much. They do differ somewhat largely as $Y_0$ changes. The estimates from the data generated from other models are actually consistent and reasonable as well.

The standard errors are large relative to the SEIR or SIR models solved by least squares. There are also NAs for some of the standard error estimates because of singular Hessian matrices.

One thing to note is that the estimates and standard errors may also be highly dependent on the starting parameter estimates given to the optimization function, as there are 5 parameters for the MSEIR model to optimize over.

\begin{table}[H]

	\caption{$\rr$ Estimates and Std. Errs, COMP Model, 
		Baseline Data, $X_0 = 99950, Y_0 = 50$, 
		$\sigma_X = 10, \sigma_Y = 1$}
	\begin{footnotesize}
		\hskip -1cm
		\begin{tabular}{l|r|r|r|r|r|r|r|r}
			\hline
			Data Set & Auto Est & Auto SE & AutoM Est & AutoM SE & Norm Est & Norm SE & NormM Est & NormM SE\\
			\hline
			Baseline 1 & 1.151396 & 0.007840 & 1.156166 & 0.075838 & 1.168638 & 0.061978 & 1.168638 & 0.069916\\
			\hline
			Baseline 2 & 1.151396 & 0.015427 & 1.151396 & 0.015825 & 1.168638 & 0.068289 & 1.168638 & 0.056477\\
			\hline
			Baseline 3 & 1.155559 & 0.127438 & 1.151396 & 0.012891 & 1.168638 & 0.068609 & 1.168638 & 0.052676\\
			\hline
			Baseline 4 & 1.156588 & 0.049150 & 1.154959 & 0.026360 & 1.168638 & 0.052003 & 1.126363 & 0.038690\\
			\hline
			Baseline 5 & 1.149516 & 0.024329 & 1.151396 & 0.014140 & 1.168638 & 0.060267 & 1.149516 & 0.029827\\
			\hline
		\end{tabular}
	\end{footnotesize}
\end{table}

\begin{table}[H]

	\caption{$\rr$ Estimates and Std. Errs, COMP Model,
		Different Starting Populations, 
		$\sigma_X = 10, \sigma_Y = 1$}
	\begin{footnotesize}
		\hskip -1.7cm
		\begin{tabular}{l|r|r|r|r|r|r|r|r}
			\hline
			Data Set & Auto Est & Auto SE & AutoM Est & AutoM SE & Norm Est & Norm SE & NormM Est & NormM SE\\
			\hline
			$X_0 = 99999, Y_0 = 1$ & 0.205516 & NA & 0.220109 & NA & 0.203940 & NA & 0.220109 & NA\\
			\hline
			$X_0 = 99990, Y_0 = 10$ & 1.049647 & 0.040841 & 1.043576 & 0.109889 & 1.049647 & 0.042400 & 1.049647 & 0.070514\\
			\hline
			$X_0 = 99900, Y_0 = 100$ & 1.215624 & 0.008808 & 1.163593 & 0.005915 & 1.163593 & 0.004472 & 1.163593 & 0.007400\\
			\hline
			$X_0 = 99000, Y_0 = 1000$ & 1.153396 & 0.000940 & 1.143722 & 0.000599 & 1.103664 & 0.000472 & 1.153396 & 0.000970\\
			\hline
		\end{tabular}
	\end{footnotesize}
\end{table}

\begin{table}[H]

	\caption{$\rr$ Estimates and Std. Errs, COMP Model, 
		Different Variances, $X_0 = 99000$, $Y_0 = 1000$}
	\begin{footnotesize}
		\hskip -1.7cm
		\begin{tabular}{l|r|r|r|r|r|r|r|r}
			\hline
			Data Set & Auto Est & Auto SE & AutoM Est & AutoM SE & Norm Est & Norm SE & NormM Est & NormM SE\\
			\hline
			$\sigma_X = 10, \sigma_Y = 1$ & 1.143722 & 0.000584 & 1.147266 & 0.002023 & 1.132778 & 0.000684 & 1.132778 & 0.000696\\
			\hline
			$\sigma_X = 100, \sigma_Y = 10$ & 1.140700 & 0.000566 & 1.171033 & 0.000785 & 1.155093 & 0.000925 & 1.183014 & 0.001369\\
			\hline
			$\sigma_X = 300, \sigma_Y = 30$ & 1.174448 & 0.001166 & 1.137020 & 0.002796 & 1.236429 & 0.003236 & 1.122842 & 0.001041\\
			\hline
			$\sigma_X = 1000, \sigma_Y = 100$ & 1.142865 & 0.000803 & 1.154025 & 0.001744 & 1.205823 & 0.003666 & 1.399630 & 0.002187\\
			\hline
			$\sigma_X = 3000, \sigma_Y = 300$ & 1.142642 & 0.001985 & 1.140222 & 0.000500 & 1.150901 & 0.000913 & 1.140190 & 0.000416\\
			\hline
			$\sigma_X = 10000, \sigma_Y = 1000$ & 1.198172 & NA & 1.147908 & 0.000934 & 1.162161 & 0.001444 & 1.157407 & 0.000959\\
			\hline
		\end{tabular}
	\end{footnotesize}
\end{table}

\begin{table}[H]

	\caption{$\rr$ Estimates and Std. Errs, COMP Model,
		Other Models, $X_0 = 99950, Y_0 = 50$, 
		$\sigma_X = 10, \sigma_Y = 1$}
	\begin{footnotesize}
		\hskip -1cm
		\begin{tabular}{l|r|r|r|r|r|r|r|r}
			\hline
			Data Set & Auto Est & Auto SE & AutoM Est & AutoM SE & Norm Est & Norm SE & NormM Est & NormM SE\\
			\hline
			0th & 1.040710 & 0.042623 & 1.038403 & 0.142913 & 1.036463 & 0.053749 & 1.036463 & 0.049341\\
			\hline
			1st & 1.287607 & 0.010336 & 1.287607 & 0.009681 & 1.287607 & 0.009128 & 1.287607 & 0.010393\\
			\hline
			2nd & 1.397864 & 0.000200 & 1.397864 & 0.002296 & 1.397864 & 0.000618 & 1.397864 & 0.001171\\
			\hline
			3rd & 1.364520 & 0.004846 & 1.364520 & 0.004999 & 1.364520 & 0.004918 & 1.364520 & 0.004876\\
			\hline
			4th & 1.397864 & 0.002116 & 1.397864 & 0.002116 & 1.397864 & 0.002426 & 1.397864 & 0.001674\\
			\hline
			Linear SIR & 1.463399 & 0.011059 & 1.463399 & 0.011044 & 1.463399 & 0.010852 & 1.463399 & 0.010955\\
			\hline
		\end{tabular}
	\end{footnotesize}
\end{table}

\subsubsection{Next Generation (SEIR)}

\paragraph{Estimation}

The next generation model below is implemented for the SEIR model. The parameters of the SEIR model ($\beta$, $\gamma$, and $\alpha$) were estimated using constrOptim() in R to minimize the sum of squared errors using Nelder-Mead. The number of iterations was limited to 50 to limit the run time. The constrOptim() function is the same as optim(), except that it allows us to constrain the search space for the parameters by specifying a constraint matrix. In this case, all of the parameters were constrained to be positive, as they should be by definition. We then plug in the parameter estimates to the $F$ and $V$ matrices used to calculate the estimates. The standard errors are calculated by resampling the parameters of the SEIR model and taking the standard deviation of the $\rr$ estimates from the resampled parameters. The standard deviation of the parameters are calculated using the Hessian matrix given by constrOptim() for Nelder-Mead, and resampling is done under the assumption that the parameters are distributed as a multivariate Gaussian.

\paragraph{Comments}

Since this is the SEIR model fit to SIR data, we might have issues with fitting. A lot of the $\rr$ estimates are exactly 1.4. It is not clear why this is or why this is not the case for some of the other data sets. Like we see from the SEIR least squares model, there are some NA estimates, though there are no negative estimates. The estimates do not fluctuate as much as those from the SEIR least squares model either. The $\rr$ estimates also become less accurate for large values of $\sigma_X$ and $\sigma_Y$. Looking at the data generated from other models, the next generation does not seem to work well for the 0th and 1st order data, where the number of people in the S and aI compartment change linearly rather than exponentially.

The standard errors are approximately of the same order of magnitude as those from the SEIR least squares model. Also, like for the SEIR least squares model, they are fairly consistent between the multiple baseline set, the standard errors decrease dramatically as $Y_0$ increases, and the standard deviations fluctuate by several orders of magnitude as we change the variance of the added errors. The most extreme values of standard errors that we see here are not as unreasonably gigantic as for the SEIR least squares model either.

\begin{table}[H]

	\caption{$\rr$ Estimates and Std. Errs, NEXTGEN Model, 
		Baseline Data, $X_0 = 99950, Y_0 = 50$, 
		$\sigma_X = 10, \sigma_Y = 1$}
	\begin{footnotesize}
		\hskip -1cm
		\begin{tabular}{l|r|r|r|r|r|r|r|r}
			\hline
			Data Set & Auto Est & Auto SE & AutoM Est & AutoM SE & Norm Est & Norm SE & NormM Est & NormM SE\\
			\hline
			Baseline 1 & 1.421896 & 0.110167 & 1.246864 & 0.010846 & 1.715318 & 0.798366 & 1.781744 & 0.024938\\
			\hline
			Baseline 2 & 1.758858 & 0.366701 & 5.512315 & 0.320605 & 1.963196 & 0.846297 & 1.625233 & 0.018011\\
			\hline
			Baseline 3 & 1.758858 & 0.569505 & 1.421896 & 0.016966 & 2.297508 & 2.211150 & 1.858192 & 0.036113\\
			\hline
			Baseline 4 & 1.447204 & 0.091904 & 1.712243 & 0.020892 & 1.421896 & 0.341627 & 2.172764 & 0.038152\\
			\hline
			Baseline 5 & 1.421896 & 0.109030 & 2.162264 & 0.034685 & 1.625233 & 0.626916 & 2.011729 & 0.027506\\
			\hline
		\end{tabular}
	\end{footnotesize}
\end{table}

\begin{table}[H]

	\caption{$\rr$ Estimates and Std. Errs, NEXTGEN Model,
		Different Starting Populations, 
		$\sigma_X = 10, \sigma_Y = 1$}
	\begin{footnotesize}
		\hskip -1.7cm
		\begin{tabular}{l|r|r|r|r|r|r|r|r}
			\hline
			Data Set & Auto Est & Auto SE & AutoM Est & AutoM SE & Norm Est & Norm SE & NormM Est & NormM SE\\
			\hline
			$X_0 = 99999, Y_0 = 1$ & 575.476314 & 5450.536541 & 205.872462 & 2972.609438 & 16.358566 & 9841.298666 & 34.387930 & 5805.567900\\
			\hline
			$X_0 = 99990, Y_0 = 10$ & 1.383532 & 25.019898 & 206.028474 & 18394.966469 & 3.214261 & 12206.722105 & 741.368136 & 15921.351448\\
			\hline
			$X_0 = 99900, Y_0 = 100$ & 4.757366 & 2.846210 & 3.819417 & 0.635151 & 15.683199 & 137.249782 & 3.874259 & 0.588683\\
			\hline
			$X_0 = 99000, Y_0 = 1000$ & 4.072926 & 0.317373 & 4.130517 & 0.226424 & 4.145316 & 0.359158 & 4.012320 & 0.239506\\
			\hline
		\end{tabular}
	\end{footnotesize}
\end{table}

\begin{table}[H]

	\caption{$\rr$ Estimates and Std. Errs, NEXTGEN Model, 
		Different Variances, $X_0 = 99000$, $Y_0 = 1000$}
	\begin{footnotesize}
		\hskip -1.7cm
		\begin{tabular}{l|r|r|r|r|r|r|r|r}
			\hline
			Data Set & Auto Est & Auto SE & AutoM Est & AutoM SE & Norm Est & Norm SE & NormM Est & NormM SE\\
			\hline
			$\sigma_X = 10, \sigma_Y = 1$ & 12.998364 & 8.755960e-01 & 8.061987 & 0.486050 & 9.400660 & 6.204380e-01 & 5.752931 & 0.261811\\
			\hline
			$\sigma_X = 100, \sigma_Y = 10$ & 5.404625 & 5.198610e-01 & 7.387983 & 0.436634 & 8.531874 & 1.705313e+00 & 5.978214 & 1.108848\\
			\hline
			$\sigma_X = 300, \sigma_Y = 30$ & 23.186330 & 6.441964e+02 & 5.963903 & 1.266200 & 322.144190 & 2.088921e+05 & 7.276865 & 0.663915\\
			\hline
			$\sigma_X = 1000, \sigma_Y = 100$ & 23.983516 & 1.278927e+04 & 67.517641 & 52.327269 & 11.529412 & 4.001174e+02 & 7.010464 & 0.567235\\
			\hline
			$\sigma_X = 3000, \sigma_Y = 300$ & 33.846154 & 1.163559e+07 & 21.915757 & 1527.656692 & 162.022472 & 4.124435e+03 & 12.140594 & 4.474694\\
			\hline
			$\sigma_X = 10000, \sigma_Y = 1000$ & 174.921348 & 3.898656e+03 & 11.524763 & 165.981987 & 1.025641 & 2.676852e+02 & 101.413031 & 1760.215196\\
			\hline
		\end{tabular}
	\end{footnotesize}
\end{table}

\begin{table}[H]

	\caption{$\rr$ Estimates and Std. Errs, NEXTGEN Model,
		Other Models, $X_0 = 99950, Y_0 = 50$, 
		$\sigma_X = 10, \sigma_Y = 1$}
	\begin{footnotesize}
		\hskip -1cm
		\begin{tabular}{l|r|r|r|r|r|r|r|r}
			\hline
			Data Set & Auto Est & Auto SE & AutoM Est & AutoM SE & Norm Est & Norm SE & NormM Est & NormM SE\\
			\hline
			0th & 0.001360 & 0.497415 & 0.000040 & 0.004153 & 0.037223 & 146.510831 & 0.000040 & 0.010014\\
			\hline
			1st & 6.217726 & 78.702007 & 13.734754 & 240.363417 & 111.232168 & 4202.221696 & 8.742980 & 66.867285\\
			\hline
			2nd & 2.451548 & 0.016035 & 5.220397 & 0.437529 & 8.020012 & 2.063975 & 5.220397 & 0.475516\\
			\hline
			3rd & 5.406024 & 0.554855 & 4.833617 & 0.256668 & 7.333991 & 1.363208 & 5.536633 & 0.660877\\
			\hline
			4th & 9.415812 & 2.964507 & 9.415812 & 2.964507 & 5.964387 & 0.856505 & 5.964387 & 0.748410\\
			\hline
			Linear SIR & 14.488463 & 6.089165 & 14.488463 & 5.828112 & 14.488463 & 6.513183 & 5.383303 & 0.308536\\
			\hline
		\end{tabular}
	\end{footnotesize}
\end{table}

\subsection{Exponential Growth}

\subsubsection{Exponential Growth}

\paragraph{Estimation}

We use Poisson regression with glm() in R to fit the incidence on the time step, without an intercept. We then use the coefficient estimate from the regression to find the estimate for $\rr$ by plugging in to the moment generating function. The standard error is calculated using the standard error of the coefficient from glm() and the delta method.

\paragraph{Comments}

It is assumed here that the exponential growth in incidence is seen throughout the entire time span of the data set. This may not be true, but there doesn't seem to be a mathematical way to determine when "exponential growth" starts or stops. We also assume that the generation time distribution is Uniform(10,30). This is somewhat arbitrarily chosen. While we know that the mean time between a primary and secondary case is 20 days for the classic SIR data sets that we generated, we have to choose the distribution around the mean by ourselves.

The estimates here for $\rr$ are consistently too small compared to the true $\rr$. In particular, the estimates are almost all under 1 when the true $\rr$ is above 1, meaning we would predict that there is no epidemic when there actually is. It seems that $\rr$ increases as $Y_0$ increases as well as when $\sigma_X$ and $\sigma_Y$ increase. It is unclear whether this is because the model becomes more accurate or if it just gets larger in general. The model does not break when handling linearly shaped incidence data (although one could argue that this isn't exponential growth, so this model does not apply here in these cases.)

The standard errors are all fairly small, among the order of $10^{-3}$. They seem to decrease as $Y_0$ increases as well as when $\sigma_X$ and $\sigma_Y$ increases.

\begin{table}[H]

	\caption{$\rr$ Estimates and Std. Errs, EXPGROWTH Model, 
		Baseline Data, $X_0 = 99950, Y_0 = 50$, 
		$\sigma_X = 10, \sigma_Y = 1$}
	\begin{footnotesize}
		\hskip -1cm
		\begin{tabular}{l|r|r|r|r|r|r|r|r}
			\hline
			Data Set & Auto Est & Auto SE & AutoM Est & AutoM SE & Norm Est & Norm SE & NormM Est & NormM SE\\
			\hline
			Baseline 1 & 1.012007 & 0.010027 & 0.929414 & 0.030175 & 1.018400 & 0.009172 & 0.935858 & 0.027818\\
			\hline
			Baseline 2 & 1.010507 & 0.010239 & 0.943961 & 0.025078 & 1.018896 & 0.009109 & 0.936850 & 0.027469\\
			\hline
			Baseline 3 & 1.006199 & 0.010870 & 0.930872 & 0.029627 & 1.016426 & 0.009428 & 0.934886 & 0.028163\\
			\hline
			Baseline 4 & 1.006496 & 0.010825 & 0.934524 & 0.028292 & 1.018095 & 0.009211 & 0.934295 & 0.028375\\
			\hline
			Baseline 5 & 1.009068 & 0.010445 & 0.934444 & 0.028321 & 1.015362 & 0.009569 & 0.935845 & 0.027819\\
			\hline
		\end{tabular}
	\end{footnotesize}
\end{table}

\begin{table}[H]

	\caption{$\rr$ Estimates and Std. Errs, EXPGROWTH Model,
		Different Starting Populations, 
		$\sigma_X = 10, \sigma_Y = 1$}
	\begin{footnotesize}
		\hskip -1.7cm
		\begin{tabular}{l|r|r|r|r|r|r|r|r}
			Data Set & Auto Est & Auto SE & AutoM Est & AutoM SE & Norm Est & Norm SE & NormM Est & NormM SE\\
			\hline
			$X_0 = 99999, Y_0 = 1$ & 1.004407 & 0.011143 & 0.767095 & 0.146339 & 1.016822 & 0.009376 & 0.757875 & 0.155300\\
			\hline
			$X_0 = 99990, Y_0 = 10$ & 1.008824 & 0.010481 & 0.885239 & 0.051071 & 1.015486 & 0.009553 & 0.877557 & 0.055606\\
			\hline
			$X_0 = 99900, Y_0 = 100$ & 1.012554 & 0.009951 & 0.961057 & 0.020060 & 1.019685 & 0.009009 & 0.961228 & 0.020014\\
			\hline
			$X_0 = 99000, Y_0 = 1000$ & 1.044120 & 0.006387 & 1.042065 & 0.006576 & 1.045714 & 0.006245 & 1.042041 & 0.006578\\
			\hline
		\end{tabular}
	\end{footnotesize}
\end{table}

\begin{table}[H]

	\caption{$\rr$ Estimates and Std. Errs, EXPGROWTH Model, 
		Different Variances, $X_0 = 99000$, $Y_0 = 1000$}
	\begin{footnotesize}
		\hskip -1.7cm
		\begin{tabular}{l|r|r|r|r|r|r|r|r}
			\hline
			Data Set & Auto Est & Auto SE & AutoM Est & AutoM SE & Norm Est & Norm SE & NormM Est & NormM SE\\
			\hline
			$\sigma_X = 10, \sigma_Y = 1$ & 1.044244 & 0.006376 & 1.041733 & 0.006606 & 1.047118 & 0.006122 & 1.042010 & 0.006580\\
			\hline
			$\sigma_X = 100, \sigma_Y = 10$ & 1.088170 & 0.003410 & 1.043282 & 0.006463 & 1.099191 & 0.002912 & 1.044029 & 0.006395\\
			\hline
			$\sigma_X = 300, \sigma_Y = 30$ & 1.123132 & 0.002065 & 1.049199 & 0.005944 & 1.130609 & 0.001855 & 1.041324 & 0.006645\\
			\hline
			$\sigma_X = 1000, \sigma_Y = 100$ & 1.161414 & 0.001192 & 1.041609 & 0.006618 & 1.171423 & 0.001033 & 1.040663 & 0.006707\\
			\hline
			$\sigma_X = 3000, \sigma_Y = 300$ & 1.194876 & 0.000738 & 1.077426 & 0.003976 & 1.210073 & 0.000594 & 1.056925 & 0.005326\\
			\hline
			$\sigma_X = 10000, \sigma_Y = 1000$ & 1.237467 & 0.000402 & 1.092848 & 0.003189 & 1.246201 & 0.000355 & 1.093502 & 0.003159\\
			\hline
		\end{tabular}
	\end{footnotesize}
\end{table}

\begin{table}[H]

	\caption{$\rr$ Estimates and Std. Errs, EXPGROWTH Model,
		Other Models, $X_0 = 99950, Y_0 = 50$, 
		$\sigma_X = 10, \sigma_Y = 1$}
	\begin{footnotesize}
		\hskip -1cm
		\begin{tabular}{l|r|r|r|r|r|r|r|r}
			\hline
			Data Set & Auto Est & Auto SE & AutoM Est & AutoM SE & Norm Est & Norm SE & NormM Est & NormM SE\\
			\hline
			0th & 1.006482 & 0.010827 & 0.782299 & 0.131800 & 1.016043 & 0.009479 & 0.829597 & 0.089876\\
			\hline
			1st & 1.060996 & 0.005026 & 1.061004 & 0.005025 & 1.061431 & 0.004995 & 1.060633 & 0.005052\\
			\hline
			2nd & 1.073885 & 0.004183 & 1.073323 & 0.004216 & 1.074063 & 0.004172 & 1.073328 & 0.004216\\
			\hline
			3rd & 1.077530 & 0.003970 & 1.077039 & 0.003998 & 1.077724 & 0.003959 & 1.077108 & 0.003994\\
			\hline
			4th & 1.073888 & 0.004182 & 1.073888 & 0.004182 & 1.074161 & 0.004166 & 1.073508 & 0.004205\\
			\hline
			Linear SIR & 1.087538 & 0.003441 & 1.087399 & 0.003448 & 1.087742 & 0.003431 & 1.087338 & 0.003451\\
			\hline04 & 0.001114\\
			\hline
		\end{tabular}
	\end{footnotesize}
\end{table}

\subsubsection{Maximum Likelihood}

\paragraph{Estimation}

We assume that the generation time distribution is Uniform(10,30). This is somewhat arbitrarily chosen. While we know that the mean time between a primary and secondary case is 20 days for the classic SIR data sets that we generated, we have to choose the distribution around the mean by ourselves. We also round to the nearest integer because our data contains non-integer counts of cases.

Note that the mathematical calculations are the same here as for the branching process.

We use basic algebra in R to find the estimates for $\rr$, and we use the Fisher information to calculate the standard error for the MLE.

\paragraph{Comments}

All of the $\rr$ estimates are less than 1, while the true $\rr$ is greater than 1. Changing $Y_0$ does not appear to change much, but increasing $\sigma_X$ and $\sigma_Y$ seems to increase the estimate of $\rr$ towards the true value. It is not clear whether the estimate becomes more accurate or simply increases.

The standard errors are really small, around the same order of magnitude as the optim() models. They decrease as $Y_0$ increases and increase as $\sigma_X$ and $\sigma_Y$ increase. We have some negative standard error estimates for the 0th order model data because the sum of the incidences is actually negative due to our added random error. This would probably not happen with actual data.

\begin{table}[H]

	\caption{$\rr$ Estimates and Std. Errs, MLE Model, 
		Baseline Data, $X_0 = 99950, Y_0 = 50$, 
		$\sigma_X = 10, \sigma_Y = 1$}
	\begin{footnotesize}
		\hskip -1cm
		\begin{tabular}{l|r|r|r|r|r|r|r|r}
			\hline
			Data Set & Auto Est & Auto SE & AutoM Est & AutoM SE & Norm Est & Norm SE & NormM Est & NormM SE\\
			\hline
			Baseline 1 & 1.045203 & 0.005596 & 0.954158 & 0.004911 & 1.009321 & 0.005045 & 1.010310 & 0.005043\\
			\hline
			Baseline 2 & 0.927409 & 0.004482 & 1.132450 & 0.005591 & 0.985546 & 0.004868 & 1.038093 & 0.005246\\
			\hline
			Baseline 3 & 1.035850 & 0.005555 & 0.969459 & 0.004980 & 1.001299 & 0.005062 & 1.006797 & 0.005083\\
			\hline
			Baseline 4 & 0.974099 & 0.004834 & 0.971615 & 0.004748 & 0.944385 & 0.004705 & 0.994621 & 0.005009\\
			\hline
			Baseline 5 & 0.988996 & 0.005185 & 1.020904 & 0.005239 & 1.053116 & 0.005433 & 1.002879 & 0.004973\\
			\hline
		\end{tabular}
	\end{footnotesize}
\end{table}

\begin{table}[H]

	\caption{$\rr$ Estimates and Std. Errs, MLE Model,
		Different Starting Populations, 
		$\sigma_X = 10, \sigma_Y = 1$}
	\begin{footnotesize}
		\hskip -1.7cm
		\begin{tabular}{l|r|r|r|r|r|r|r|r}
			\hline
			Data Set & Auto Est & Auto SE & AutoM Est & AutoM SE & Norm Est & Norm SE & NormM Est & NormM SE\\
			\hline
			$X_0 = 99999, Y_0 = 1$ & 1.232756 & 0.069833 & 0.976411 & 0.163888 & 1.615820 & 0.467853 & 1.192249 & 0.205440\\
			\hline
			$X_0 = 99990, Y_0 = 10$ & 0.949905 & 0.027784 & 1.301035 & 0.031985 & 0.884468 & 0.022576 & 1.208762 & 0.028717\\
			\hline
			$X_0 = 99900, Y_0 = 100$ & 1.016057 & 0.002553 & 0.984204 & 0.002475 & 1.055094 & 0.002654 & 0.990380 & 0.002500\\
			\hline
			$X_0 = 99000, Y_0 = 1000$ & 0.983594 & 0.000260 & 0.996016 & 0.000263 & 0.997396 & 0.000264 & 0.998214 & 0.000264\\
			\hline
		\end{tabular}
	\end{footnotesize}
\end{table}

\begin{table}[H]

	\caption{$\rr$ Estimates and Std. Errs, MLE Model, 
		Different Variances, $X_0 = 99000$, $Y_0 = 1000$}
	\begin{footnotesize}
		\hskip -1.7cm
		\begin{tabular}{l|r|r|r|r|r|r|r|r}
			\hline
			Data Set & Auto Est & Auto SE & AutoM Est & AutoM SE & Norm Est & Norm SE & NormM Est & NormM SE\\
			\hline
			$\sigma_X = 10, \sigma_Y = 1$ & 0.985737 & 0.000206 & 0.985711 & 0.000207 & 0.985510 & 0.000206 & 0.986968 & 0.000206\\
			\hline
			$\sigma_X = 100, \sigma_Y = 10$ & 0.999005 & 0.000208 & 1.001050 & 0.000208 & 1.006205 & 0.000209 & 1.013000 & 0.000211\\
			\hline
			$\sigma_X = 300, \sigma_Y = 30$ & 1.051749 & 0.000236 & 0.966436 & 0.000181 & 0.816122 & 0.000167 & 0.966859 & 0.000200\\
			\hline
			$\sigma_X = 1000, \sigma_Y = 100$ & 1.060553 & 0.000161 & 1.065634 & 0.000236 & 0.777445 & 0.000177 & 0.975480 & 0.000206\\
			\hline
			$\sigma_X = 3000, \sigma_Y = 300$ & 0.511344 & 0.000110 & 1.003435 & 0.000113 & 2.385157 & 0.000672 & 1.095206 & 0.000199\\
			\hline
			$\sigma_X = 10000, \sigma_Y = 1000$ & -3.235191 & 0.000840 & 0.952381 & 0.000089 & -2.850294 & -0.001262 & 1.532468 & 0.000209\\
			\hline
		\end{tabular}
	\end{footnotesize}
\end{table}

\begin{table}[H]

	\caption{$\rr$ Estimates and Std. Errs, MLE Model,
		Other Models, $X_0 = 99950, Y_0 = 50$, 
		$\sigma_X = 10, \sigma_Y = 1$}
	\begin{footnotesize}
		\hskip -1cm
		\begin{tabular}{l|r|r|r|r|r|r|r|r}
			\hline
			Data Set & Auto Est & Auto SE & AutoM Est & AutoM SE & Norm Est & Norm SE & NormM Est & NormM SE\\
			\hline
			0th & -12.609343 & -14.168066 & 16.325348 & 42.366553 & 3.152735 & -0.768473 & 10.431474 & 8.031043\\
			\hline
			1st & 0.978683 & 0.000125 & 0.982726 & 0.000126 & 0.979523 & 0.000125 & 0.976522 & 0.000125\\
			\hline
			2nd & 1.017042 & 0.000131 & 1.013913 & 0.000131 & 1.014664 & 0.000131 & 1.014455 & 0.000131\\
			\hline
			3rd & 1.029675 & 0.000121 & 1.026645 & 0.000121 & 1.027745 & 0.000121 & 1.027322 & 0.000121\\
			\hline
			4th & 1.028890 & 0.000135 & 1.028890 & 0.000135 & 1.027876 & 0.000135 & 1.027989 & 0.000135\\
			\hline
			Linear SIR & 1.052763 & 0.000099 & 1.053736 & 0.000099 & 1.053703 & 0.000099 & 1.052769 & 0.000099\\
			\hline
		\end{tabular}
	\end{footnotesize}
\end{table}

\subsubsection{Initial Growth}

\paragraph{Estimation}

We use Poisson regression with glm() in R to fit the incidence on the time step, without an intercept. We then use the coefficient estimate from the regression to find the estimate for $\rr$ by plugging in to the given equation. The duration time is assumed to be $\frac{1}{\gamma}$ from the true data. We use the delta method to solve for the standard errors, using the standard error given from the regression from glm().

\paragraph{Comments}

This model assumes that the incidence is in a period of exponential growth throughout the entire time period. I have fitted the model on the entire data sets.

The estimates here are actually all pretty close to the true $\rr$, and they are extremely consistent for the baseline data sets. They also seem to increase as $Y_0$ increases and as $\sigma_X$ and $\sigma_Y$ increase. This seems to occur regardless of accuracy. The model also handles the 0th and 1st order data sets well, giving us somewhat reasonable estimates (although one could argue that this isn't exponential growth, so this model does not apply here in these cases.)

The standard error estimates are all reasonable. They seem to decrease slightly as $Y_0$ increases and as $\sigma_X$ and $\sigma_Y$ increase.

\begin{table}[H]

	\caption{$\rr$ Estimates and Std. Errs, INITGROWTH Model, 
		Baseline Data, $X_0 = 99950, Y_0 = 50$, 
		$\sigma_X = 10, \sigma_Y = 1$}
	\begin{footnotesize}
		\hskip -1cm
		\begin{tabular}{l|r|r|r|r|r|r|r|r}
			\hline
			Data Set & Auto Est & Auto SE & AutoM Est & AutoM SE & Norm Est & Norm SE & NormM Est & NormM SE\\
			\hline
			Baseline 1 & 1.122458 & 0.044665 & 0.951335 & 0.152044 & 1.135276 & 0.040490 & 0.965079 & 0.138753\\
			\hline
			Baseline 2 & 1.119443 & 0.045703 & 0.982261 & 0.123521 & 1.136269 & 0.040183 & 0.967189 & 0.136801\\
			\hline
			Baseline 3 & 1.110763 & 0.048817 & 0.954452 & 0.148942 & 1.131325 & 0.041737 & 0.963011 & 0.140689\\
			\hline
			Baseline 4 & 1.111362 & 0.048596 & 0.962238 & 0.141417 & 1.134666 & 0.040680 & 0.961751 & 0.141879\\
			\hline
			Baseline 5 & 1.116547 & 0.046721 & 0.962069 & 0.141578 & 1.129191 & 0.042424 & 0.965051 & 0.138765\\
			\hline
		\end{tabular}
	\end{footnotesize}
\end{table}

\begin{table}[H]

	\caption{$\rr$ Estimates and Std. Errs, INITGROWTH Model,
		Different Starting Populations, 
		$\sigma_X = 10, \sigma_Y = 1$}
	\begin{footnotesize}
		\hskip -1.7cm
		\begin{tabular}{l|r|r|r|r|r|r|r|r}
			\hline
			Data Set & Auto Est & Auto SE & AutoM Est & AutoM SE & Norm Est & Norm SE & NormM Est & NormM SE\\
			\hline
			$X_0 = 99999, Y_0 = 1$ & 1.107146 & 0.050171 & 0.578787 & 0.985754 & 1.132118 & 0.041484 & 0.555865 & 1.065990\\
			\hline
			$X_0 = 99990, Y_0 = 10$ & 1.116055 & 0.046896 & 0.855171 & 0.276575 & 1.129441 & 0.042344 & 0.838084 & 0.305078\\
			\hline
			$X_0 = 99900, Y_0 = 100$ & 1.123557 & 0.044293 & 1.018159 & 0.096252 & 1.137845 & 0.039698 & 1.018516 & 0.096009\\
			\hline
			$X_0 = 99000, Y_0 = 1000$ & 1.186268 & 0.027217 & 1.182227 & 0.028098 & 1.189399 & 0.026553 & 1.182180 & 0.028108\\
			\hline
		\end{tabular}
	\end{footnotesize}
\end{table}

\begin{table}[H]

	\caption{$\rr$ Estimates and Std. Errs, INITGROWTH Model, 
		Different Variances, $X_0 = 99000$, $Y_0 = 1000$}
	\begin{footnotesize}
		\hskip -1.7cm
		\begin{tabular}{l|r|r|r|r|r|r|r|r}
			\hline
			Data Set & Auto Est & Auto SE & AutoM Est & AutoM SE & Norm Est & Norm SE & NormM Est & NormM SE\\
			\hline
			$\sigma_X = 10, \sigma_Y = 1$ & 1.186512 & 0.027164 & 1.181573 & 0.028243 & 1.192153 & 0.025981 & 1.182118 & 0.028121\\
			\hline
			$\sigma_X = 100, \sigma_Y = 10$ & 1.271572 & 0.013714 & 1.184621 & 0.027573 & 1.292533 & 0.011548 & 1.186089 & 0.027256\\
			\hline
			$\sigma_X = 300, \sigma_Y = 30$ & 1.337570 & 0.007951 & 1.196232 & 0.025155 & 1.351500 & 0.007077 & 1.180770 & 0.028422\\
			\hline
			$\sigma_X = 1000, \sigma_Y = 100$ & 1.408230 & 0.004385 & 1.181330 & 0.028297 & 1.426440 & 0.003755 & 1.179468 & 0.028713\\
			\hline
			$\sigma_X = 3000, \sigma_Y = 300$ & 1.468704 & 0.002614 & 1.250992 & 0.016213 & 1.495791 & 0.002069 & 1.211323 & 0.022310\\
			\hline
			$\sigma_X = 10000, \sigma_Y = 1000$ & 1.544051 & 0.001360 & 1.280487 & 0.012748 & 1.559289 & 0.001190 & 1.281732 & 0.012619\\
			\hline

		\end{tabular}
	\end{footnotesize}
\end{table}

\begin{table}[H]

	\caption{$\rr$ Estimates and Std. Errs, INITGROWTH Model,
		Other Models, $X_0 = 99950, Y_0 = 50$, 
		$\sigma_X = 10, \sigma_Y = 1$}
	\begin{footnotesize}
		\hskip -1cm
		\begin{tabular}{l|r|r|r|r|r|r|r|r}
			\hline
			Data Set & Auto Est & Auto SE & AutoM Est & AutoM SE & Norm Est & Norm SE & NormM Est & NormM SE\\
			\hline
			0th & 1.111335 & 0.048605 & 0.616117 & 0.861217 & 1.130557 & 0.041983 & 0.728769 & 0.536725\\
			\hline
			1st & 1.219243 & 0.020940 & 1.219259 & 0.020937 & 1.220088 & 0.020800 & 1.218537 & 0.021060\\
			\hline
			2nd & 1.244179 & 0.017134 & 1.243097 & 0.017284 & 1.244522 & 0.017086 & 1.243106 & 0.017283\\
			\hline
			3rd & 1.251193 & 0.016187 & 1.250249 & 0.016312 & 1.251565 & 0.016139 & 1.250381 & 0.016295\\
			\hline
			4th & 1.244185 & 0.017133 & 1.244185 & 0.017133 & 1.244711 & 0.017060 & 1.243452 & 0.017235\\
			\hline
			Linear SIR & 1.270365 & 0.013849 & 1.270099 & 0.013879 & 1.270753 & 0.013805 & 1.269983 & 0.013893\\
			\hline
		\end{tabular}
	\end{footnotesize}
\end{table}

\subsubsection{Final Size}

\paragraph{Estimation}

We use basic arithmetic functions in R to calculate the $\rr$ estimates. We use the delta method to calculate the standard error; treating the final size a proportion of the total population, we use the variance of proportions equation($ \sqrt{\frac{p(1-p)}{N}}$) to calculate the variance of the final size.

\paragraph{Comments}

The model depends on our estimate for the mean infectious period. In this case, we use the true infectious period from the classic SIR data ($\frac{1}{\gamma}$). The estimates and standard errors for the baseline data sets are almost exactly the same between the different ones. The estimates tend to become less accurate as $\sigma_X$ and $\sigma_Y$ increase. They increase as $Y_0$ increases as well. However, this may be due to the fact that the model does not take starting compartment sizes in to account; thus, it does not differentiate between the change in final size caused by changing $Y_0$ when the data are generated from the classic SIR model. Also, the model does not give crazy results for data that are not generated from the classic SIR model, as we might expect since all it uses for estimation is the final number of people who were infected.

The standard errors are all relatively small in magnitude given the way I calculated them, and they increase as $Y_0$ and as $\sigma_X$ and $\sigma_Y$ increase.

\begin{table}[H]

	\caption{$\rr$ Estimates and Std. Errs, FINALSIZE Model,
		Baseline Data$X_0 = 99950, Y_0 = 50$, 
		$\sigma_X = 10, \sigma_Y = 1$}
	\centering
	\begin{footnotesize}
		\hskip -1cm
	\begin{tabular}[t]{l|r|r|r|r|r|r|r|r}
		\hline
		Data Set & Auto Est & Auto SE & AutoM Est & AutoM SE & Norm Est & Norm SE & NormM Est & NormM SE\\
		\hline
		Baseline 1 & 1.001228 & 7.8e-05 & 1.001179 & 7.7e-05 & 1.001262 & 8.0e-05 & 1.001264 & 8.0e-05\\
		\hline
		Baseline 2 & 1.001211 & 7.8e-05 & 1.001400 & 8.4e-05 & 1.001250 & 7.9e-05 & 1.001279 & 8.0e-05\\
		\hline
		Baseline 3 & 1.001218 & 7.8e-05 & 1.001196 & 7.7e-05 & 1.001242 & 7.9e-05 & 1.001249 & 7.9e-05\\
		\hline
		Baseline 4 & 1.001233 & 7.9e-05 & 1.001246 & 7.9e-05 & 1.001200 & 7.8e-05 & 1.001239 & 7.9e-05\\
		\hline
		Baseline 5 & 1.001195 & 7.7e-05 & 1.001247 & 7.9e-05 & 1.001273 & 8.0e-05 & 1.001263 & 8.0e-05\\
		\hline
	\end{tabular}
	\end{footnotesize}
\end{table}

\begin{table}[H]

	\caption{$\rr$ Estimates and Std. Errs, FINALSIZE Model,
		Different Starting Populations, 
		$\sigma_X = 10, \sigma_Y = 1$}
	\begin{footnotesize}
		\hskip -1.7cm
		\begin{tabular}{l|r|r|r|r|r|r|r|r}
			\hline
			Data Set & Auto Est & Auto SE & AutoM Est & AutoM SE & Norm Est & Norm SE & NormM Est & NormM SE\\
			\hline
			$X_0 = 99999, Y_0 = 1$ & 1.000114 & 0.000024 & 1.000034 & 0.000013 & 1.000033 & 0.000013 & 1.000040 & 0.000014\\
			\hline
			$X_0 = 99990, Y_0 = 10$ & 1.000212 & 0.000033 & 1.000315 & 0.000040 & 1.000223 & 0.000033 & 1.000305 & 0.000039\\
			\hline
			$X_0 = 99900, Y_0 = 100$ & 1.002530 & 0.000113 & 1.002465 & 0.000111 & 1.002606 & 0.000114 & 1.002470 & 0.000111\\
			\hline
			$X_0 = 99000, Y_0 = 1000$ & 1.024360 & 0.000358 & 1.024661 & 0.000360 & 1.024662 & 0.000360 & 1.024670 & 0.000360\\
			\hline
		\end{tabular}
	\end{footnotesize}
\end{table}

\begin{table}[H]

	\caption{$\rr$ Estimates and Std. Errs, FINALSIZE Model, 
		Different Variances, $X_0 = 99000$, $Y_0 = 1000$}
	\begin{footnotesize}
		\hskip -1.7cm
		\begin{tabular}{l|r|r|r|r|r|r|r|r}
			\hline
			Data Set & Auto Est & Auto SE & AutoM Est & AutoM SE & Norm Est & Norm SE & NormM Est & NormM SE\\
			\hline
			$\sigma_X = 10, \sigma_Y = 1$ & 1.024640 & 0.000360 & 1.024544 & 0.000359 & 1.024610 & 0.000359 & 1.024641 & 0.000360\\
			\hline
			$\sigma_X = 100, \sigma_Y = 10$ & 1.025024 & 0.000363 & 1.025169 & 0.000364 & 1.025348 & 0.000365 & 1.025358 & 0.000365\\
			\hline
			$\sigma_X = 300, \sigma_Y = 30$ & 1.024494 & 0.000359 & 1.026920 & 0.000377 & 1.020803 & 0.000329 & 1.024440 & 0.000358\\
			\hline
			$\sigma_X = 1000, \sigma_Y = 100$ & 1.036821 & 0.000445 & 1.025126 & 0.000363 & 1.017766 & 0.000303 & 1.024120 & 0.000356\\
			\hline
			$\sigma_X = 3000, \sigma_Y = 300$ & 1.012305 & 0.000251 & 1.047894 & 0.000513 & 1.045165 & 0.000497 & 1.031720 & 0.000411\\
			\hline
			$\sigma_X = 10000, \sigma_Y = 1000$ & 1.068277 & 0.000624 & 1.055107 & 0.000554 & 0.969364 & NaN & 1.061044 & 0.000586\\
			\hline
		\end{tabular}
	\end{footnotesize}
\end{table}

\begin{table}[H]

	\caption{$\rr$ Estimates and Std. Errs, FINALSIZE Model,
		Other Models, $X_0 = 99950, Y_0 = 50$, 
		$\sigma_X = 10, \sigma_Y = 1$}
	\begin{footnotesize}
		\hskip -1cm
		\begin{tabular}{l|r|r|r|r|r|r|r|r}
			\hline
			Data Set & Auto Est & Auto SE & AutoM Est & AutoM SE & Norm Est & Norm SE & NormM Est & NormM SE\\
			\hline
			0th & 1.000194 & 0.000031 & 1.000282 & 0.000038 & 1.000185 & 0.000030 & 1.000318 & 0.000040\\
			\hline
			1st & 1.040704 & 0.000470 & 1.040802 & 0.000470 & 1.040694 & 0.000469 & 1.040546 & 0.000469\\
			\hline
			2nd & 1.041921 & 0.000477 & 1.041759 & 0.000476 & 1.041815 & 0.000476 & 1.041766 & 0.000476\\
			\hline
			3rd & 1.046705 & 0.000506 & 1.046460 & 0.000504 & 1.046563 & 0.000505 & 1.046529 & 0.000505\\
			\hline
			4th & 1.041585 & 0.000475 & 1.041585 & 0.000475 & 1.041481 & 0.000474 & 1.041485 & 0.000474\\
			\hline
			Linear SIR & 1.061096 & 0.000587 & 1.061156 & 0.000587 & 1.061115 & 0.000587 & 1.061060 & 0.000586\\
			\hline
		\end{tabular}
	\end{footnotesize}
\end{table}

\subsection{Networks}

\subsubsection{Branching Process}

\paragraph{Estimation}

We assume that the generation time distribution is Uniform(10,30). This is somewhat arbitrarily chosen. While we know that the mean time between a primary and secondary case is 20 days for the classic SIR data sets that we generated, we have to choose the distribution around the mean by ourselves. We also round to the nearest integer because our data contains non-integer counts of cases.

Note that the mathematical calculations are the same here as for the MLE.

We use basic algebra in R to find the estimates for $\rr$, and we use the second derivative to calculate the standard error for the branching process.

\paragraph{Comments}

All of the $\rr$ estimates are less than 1, while the true $\rr$ is greater than 1. Changing $Y_0$ does not appear to change much, but increasing $\sigma_X$ and $\sigma_Y$ seems to increase the estimate of $\rr$ towards the true value. It is not clear whether the estimate becomes more accurate or simply increases.

The standard errors are really small, around the same order of magnitude as the optim() models. They decrease as $Y_0$ increases and increase as $\sigma_X$ and $\sigma_Y$ increase. We have some negative standard error estimates for the 0th order model data because the sum of the incidences is actually negative due to our added random error. This would probably not happen with actual data.

\begin{table}[H]

	\caption{$\rr$ Estimates and Std. Errs, BRANCH Model,
		Baseline Data$X_0 = 99950, Y_0 = 50$, 
		$\sigma_X = 10, \sigma_Y = 1$}
	\centering
	\begin{footnotesize}
		\hskip -1cm
	\begin{tabular}[t]{l|r|r|r|r|r|r|r|r}
		\hline
		Data Set & Auto Est & Auto SE & AutoM Est & AutoM SE & Norm Est & Norm SE & NormM Est & NormM SE\\
		\hline
		Baseline 1 & 1.045203 & 0.005596 & 0.954158 & 0.004911 & 1.009321 & 0.005045 & 1.010310 & 0.005043\\
		\hline
		Baseline 2 & 0.927409 & 0.004482 & 1.132450 & 0.005591 & 0.985546 & 0.004868 & 1.038093 & 0.005246\\
		\hline
		Baseline 3 & 1.035850 & 0.005555 & 0.969459 & 0.004980 & 1.001299 & 0.005062 & 1.006797 & 0.005083\\
		\hline
		Baseline 4 & 0.974099 & 0.004834 & 0.971615 & 0.004748 & 0.944385 & 0.004705 & 0.994621 & 0.005009\\
		\hline
		Baseline 5 & 0.988996 & 0.005185 & 1.020904 & 0.005239 & 1.053116 & 0.005433 & 1.002879 & 0.004973\\
		\hline
	\end{tabular}
\end{footnotesize}
\end{table}

\begin{table}[H]

	\caption{$\rr$ Estimates and Std. Errs, BRANCH Model,
		Different Starting Populations, 
		$\sigma_X = 10, \sigma_Y = 1$}
	\begin{footnotesize}
		\hskip -1.7cm
		\begin{tabular}{l|r|r|r|r|r|r|r|r}
			\hline
			Data Set & Auto Est & Auto SE & AutoM Est & AutoM SE & Norm Est & Norm SE & NormM Est & NormM SE\\
			\hline
			$X_0 = 99999, Y_0 = 1$ & 1.232756 & 0.069833 & 0.976411 & 0.163888 & 1.615820 & 0.467853 & 1.192249 & 0.205440\\
			\hline
			$X_0 = 99990, Y_0 = 10$ & 0.949905 & 0.027784 & 1.301035 & 0.031985 & 0.884468 & 0.022576 & 1.208762 & 0.028717\\
			\hline
			$X_0 = 99900, Y_0 = 100$ & 1.016057 & 0.002553 & 0.984204 & 0.002475 & 1.055094 & 0.002654 & 0.990380 & 0.002500\\
			\hline
			$X_0 = 99000, Y_0 = 1000$ & 0.983594 & 0.000260 & 0.996016 & 0.000263 & 0.997396 & 0.000264 & 0.998214 & 0.000264\\
			\hline
		\end{tabular}
	\end{footnotesize}
\end{table}

\begin{table}[H]

	\caption{$\rr$ Estimates and Std. Errs, BRANCH Model, 
		Different Variances, $X_0 = 99000$, $Y_0 = 1000$}
	\begin{footnotesize}
		\hskip -1.7cm
		\begin{tabular}{l|r|r|r|r|r|r|r|r}
			\hline
			Data Set & Auto Est & Auto SE & AutoM Est & AutoM SE & Norm Est & Norm SE & NormM Est & NormM SE\\
			\hline
			$\sigma_X = 10, \sigma_Y = 1$ & 0.985737 & 0.000206 & 0.985711 & 0.000207 & 0.985510 & 0.000206 & 0.986968 & 0.000206\\
			\hline
			$\sigma_X = 100, \sigma_Y = 10$ & 0.999005 & 0.000208 & 1.001050 & 0.000208 & 1.006205 & 0.000209 & 1.013000 & 0.000211\\
			\hline
			$\sigma_X = 300, \sigma_Y = 30$ & 1.051749 & 0.000236 & 0.966436 & 0.000181 & 0.816122 & 0.000167 & 0.966859 & 0.000200\\
			\hline
			$\sigma_X = 1000, \sigma_Y = 100$ & 1.060553 & 0.000161 & 1.065634 & 0.000236 & 0.777445 & 0.000177 & 0.975480 & 0.000206\\
			\hline
			$\sigma_X = 3000, \sigma_Y = 300$ & 0.511344 & 0.000110 & 1.003435 & 0.000113 & 2.385157 & 0.000672 & 1.095206 & 0.000199\\
			\hline
			$\sigma_X = 10000, \sigma_Y = 1000$& -3.235191 & 0.000840 & 0.952381 & 0.000089 & -2.850294 & 0.001262 & 1.532468 & 0.000209\\
			\hline
		\end{tabular}
	\end{footnotesize}
\end{table}

\begin{table}[H]

	\caption{$\rr$ Estimates and Std. Errs, BRANCH Model,
		Other Models, $X_0 = 99950, Y_0 = 50$, 
		$\sigma_X = 10, \sigma_Y = 1$}
	\begin{footnotesize}
		\hskip -1cm
		\begin{tabular}{l|r|r|r|r|r|r|r|r}
			\hline
			Data Set & Auto Est & Auto SE & AutoM Est & AutoM SE & Norm Est & Norm SE & NormM Est & NormM SE\\
			\hline
			0th & -12.609343 & -14.168066 & 16.325348 & 42.366553 & 3.152735 & -0.768473 & 10.431474 & 8.031043\\
			\hline
			1st & 0.978683 & 0.000125 & 0.982726 & 0.000126 & 0.979523 & 0.000125 & 0.976522 & 0.000125\\
			\hline
			2nd & 1.017042 & 0.000131 & 1.013913 & 0.000131 & 1.014664 & 0.000131 & 1.014455 & 0.000131\\
			\hline
			3rd & 1.029675 & 0.000121 & 1.026645 & 0.000121 & 1.027745 & 0.000121 & 1.027322 & 0.000121\\
			\hline
			4th & 1.028890 & 0.000135 & 1.028890 & 0.000135 & 1.027876 & 0.000135 & 1.027989 & 0.000135\\
			\hline
			Linear SIR & 1.052763 & 0.000099 & 1.053736 & 0.000099 & 1.053703 & 0.000099 & 1.052769 & 0.000099\\
			\hline
		\end{tabular}
	\end{footnotesize}
\end{table}

\end{document}

%%%%%%%%%%%%%%%%%%%%%%
\Textbf{Title:}

\textbf{Author:}

\textbf{Citation:}

\textbf{Major themes:}

\textbf{Notes:}
\\
%%%%%%%%%%%%%%%%%%%%%%

\begin{figure}[h]
\begin{center}
\includegraphics[width=4in]{mvhw7_3c.pdf}
\end{center}
\caption{
Biplots of the different continuous variables and their correlations.}\label{fig3c}
\end{figure}

%% two pictures whoa

\begin{figure}[h]
\centering

\begin{figure}[h]
\centering

\begin{subfigure}{.5\textwidth}
  \centering
  \includegraphics[width=1\linewidth]{mt_eda_cont_hists.pdf}
  \caption{Histograms of Arrival Delay and continuous covariates.  Arrival delay seems to have a right skewed distribution.  This may indicate that we will be transforming this variable later on.  After transforming Air Time and Distance by a log transformation, we don't really seem to have many outliers in our covariates.  We seem to have outliers in the CRS Dep. Time and Arrival Time; however, time is cyclical and so these are not, in fact outliers.}
  \label{hists}
\end{subfigure}%
\begin{subfigure}{.5\textwidth}
  \centering
  \includegraphics[width=1\linewidth]{mt_eda_cont_hists.pdf}
  \caption{\textcolor{red}{Placeholder}}
  \label{tabs1}
\end{subfigure}
\caption{}
\end{figure}

\begin{table}
\begin{tabular}{l | a | b | a | b}
\hline
\rowcolor{LightCyan}
\mc{1}{}  & \mc{1}{x} & \mc{1}{y} & \mc{1}{w} & \mc{1}{z} \    \hline
variable 1 & a & b & c & d \    variable 2 & a & b & c & d \\ \hline
\end{tabular}
\end{table}

%%
\begin{figure}
\centering
\begin{subfigure}{.5\textwidth}
  \centering
  \includegraphics[width=1\linewidth]{resids_full.pdf}
  \caption{}
  \label{residsf}
\end{subfigure}%
\begin{subfigure}{.5\textwidth}
  \centering
  \includegraphics[width=1\linewidth]{diags_full.pdf}
  \caption{ }
  \label{diagsf}
\end{subfigure}
\caption{}
\end{figure}

### 2018-07-03 paper/Version11/draft_v11.tex
finally on git

\documentclass[12pt]{article}
\usepackage[margin=1in]{geometry}
%\usepackage{hyperref}
\usepackage{amsmath}
\usepackage{graphicx}% http://ctan.org/pkg/graphicx
\usepackage{systeme}
%% tables
\usepackage{booktabs}
\usepackage[table,xcdraw]{xcolor}
\usepackage{float}

%% Make red text
\newcommand{\com}[1]{\textcolor{red}{ #1}}

%\linespread{2}

\usepackage[authoryear]{natbib}

\usepackage{tikz}
\tikzset{
  int/.style={circle, draw, fill=blue!20, minimum size=3em},
  init/.style={pin distance=1.2cm,pin edge={loop,thin,black}}
}
\usetikzlibrary{arrows,automata}

\pagenumbering{arabic}

\newcommand{\XX}{\ensuremath{25}} % number of methods

\newcommand{\xxsir}{\ensuremath{9}} % number of SIR methods

\newcommand{\rr}{\ensuremath{\mathcal{R}_0}}

\setlength{\itemsep}{0pt}

\begin{document}

%%% draw a tree
\tikzset{
  treenode/.style = {shape=rectangle, rounded corners,
                     draw, align=center,
                     top color=white, bottom color=blue!20},
  root/.style     = {treenode, font=\Large, bottom color=red!30},
  env/.style      = {treenode, font=\ttfamily\normalsize},
  dummy/.style    = {circle,draw}
}

\title{An overview of $\xxsir$ methods to estimate $\rr$ from the SIR model and comments on more complex models}
\author{ Department of Statistics, Carnegie Mellon University}
\date{\today}
\maketitle

% \tableofcontents

\section{Introduction}\label{sec:intro}
What has been called ``arguably the most important quantity in the study of epidemics'' \citep{Heesterbeek2002},  $\mathcal{R}_0$, the reproduction number (by convention pronounced ``R-naught'') is a persistently elusive quantity for epidemic modelers to estimate.  As defined by \citet{anderson1992}, $\rr$ is the ``the average number of secondary infections produced when one infected individual is introduced into a host population where everyone is susceptible.''  $\rr$, in some ways, summarizes an entire outbreak of a disease; it  is used to assess whether a disease epidemic will occur.  Additionally, it describes what percentage of the population needs to be vaccinated to avoid such an epidemic.  Despite a clear definition of $\rr$, epidemiologists have struggled to create a standard  estimator for $\rr$  \citep{hethcote2000}.  A major issue in estimating $\rr$ is that the quantity is a \textit{property of the model}, meaning that $\rr$ is dependent not only on the usual noise that comes with statistical modeling but also on a variety of assumptions on how researchers assume a disease is transmitted through a population.

To elaborate on $\rr$ being the property of the model, we first need to introduce the concept of the ``SI-framework,'' which was pioneered by Kermack and McKendrick in the 1920s \citep{getz2006}.   Here, S and I are known as compartments where S stands for ``susceptible'' and I for ``infectious.'' Infectious disease models then specify how individuals move from the S compartment to the I compartment.  Often, ancillary compartments are added such as R, which stands for recovered or dead, or E, which stands for exposed but not yet infectious.  The estimate of $\rr$ from a SIR model will not be directly comparable to the estimate of $\rr$ from a SEIR model, for example.  The difficulties of estimating $\rr$ are summarized and expanded upon by \cite{li2011}.   As a result of $\rr$ being a property of the model, we limit our review to methods for estimating $\rr$ for the SIR model introduced by \cite{Kermack700}.

Even when limiting $\rr$ estimates to those from SIR models, estimation is still difficult.  Problems include mathematical versus statistical methods (i.e. solving for $\rr$ versus estimating $\rr$), the number of observations used to estimate $\rr$ (i.e. time), boundary cases of infection and recovery rates, population size, initial SI ratio, and restrictive assumptions placed upon the noise.  These are all issues for estimating single points, to say nothing of confidence or credible intervals (CI).

We review $\xxsir$ methods for estimating $\rr$.  We note that these methods are not exhaustive but are chosen due to their impact and use in epidemiology.  We find that these methods also help to highlight the problems discussed above.  In addition to making a point estimate of $\rr$, we are particularly interested in the hypothesis  $H_0: \rr > 1$, as a value greater than 1 indicates an outbreak of a disease.  As such, we discuss techniques for estimating CIs for each of the different methods.  These techniques include, the delta method, block bootstrap, and the posterior distribution.

Following, we present a series of simulations in which we examine each of our $\rr$ estimation methods.  Within these simulations, we vary the number of observations used to estimate $\rr$, infection and recovery rates, the total population size, initial SI ratio, and several assumptions about the noise.

We find that...

Finally, we discuss our recommendations in estimating $\rr$ for the SIR model.  Afterwards, we comment on how these recommendations extend to more complex models and how many of the problems with estimating $\rr$  become exacerbated with the increase in complexity.

The rest of this manuscript is organized as follows.  In Section \ref{sec:r0}, we briefly discuss the origin of $\rr$.  In Section \ref{sec:sir-intro}, we introduce the SIR model as described by \cite{Kermack700}.  In Section \ref{sec:methods}, we overview the $\xxsir$ methods for estimating $\rr$. In Section \ref{sec:ci}, we discuss how to estimate CIs for each of the $\xxsir$ methods.  Following in Section \ref{sec:sim-res}, we present the results of our simulations.  Finally, in Section \ref{sec:discussion}, we provide final analysis of estimating $\rr$ for the SIR model and offer comments on how these conclusions extend to more complex models.

\section{Origins of $\rr$}
\label{sec:r0}

The origins of $\rr$ are tied to the survival function.  This function, which originates from the field of demography in the late 1800s, describes how many female offspring a woman is expected to produce in her lifetime \citep{dietz1993estimation}.  In demography, we have
\begin{align}\label{eq:surv}
\rr = \int_0^\infty p(a) \beta(a) da
\end{align}
where $p(a)$ denotes the probability of a woman surviving to age $a$ and $\beta(a)$ the rate of an individual of age $a$ giving birth to a  girl.   This interpretation also explains the origins of the name of $\rr$.  The concept of the reproduction number was later imported to the field of epidemiology by MacDonald and Smith \citep{dietz1993estimation}.  Analogously in epidemiology, $p(a)$ is the age of a disease, and $\beta(a)$ is the infection rate.

This is the chronologically first, and in many ways, the most direct method to estimate $\rr$.  As we see here, the SIR model is not even mentioned in this derivation.  The premise of this whole paper is based upon the fact that estimating $\rr$ from the survival function in Equation \eqref{eq:surv} is a difficult task.  It is due to this difficulty that other estimates even exist and why we must discuss $\rr$ as a property of the SIR model.

\section{Kermack and McKendrick's SIR Model}
\label{sec:sir-intro}

The SIR model introduced by \cite{Kermack700} is a compartment model, where individuals move from susceptible, to infectious, and finally recovered states.  We make four essential assumptions,
\begin{enumerate}
\item The compartments are discrete and have no overlap.
\item The transition of objects into and out of compartments is described by a set of known equations, possibly dependent on unknown parameters.
\item The populations mix homogeneously
  \item The number of objects in each compartment at time $t=0$ is known
  \end{enumerate}

Another common assumption for compartment models, is the law of mass action, a property borrowed from chemistry which says that the mass of the product is proportional to the mass of the reactants.  In epidemiology, this means that the proportion of new infections is proportional to the current  number of susceptible \citep{anderson1992}.  Therefore, the flow of objects from one compartment to another may be dependent on the percentage of objects within one or many of the compartments.

In this particular  SIR model, $N$, the number of individuals is constant, as we see no arrows pointing away from the three nodes in Figure \ref{fig::sir}. Simple adaptations of the SIR model can include birth and death rates, which may correspondingly change the derivation of $\rr$ (for further discussion, see Section \ref{sec:discussion}). The parameters  $\beta$ and $\gamma$ have epidemiological meaning.  Here, $\beta$ is the average infection rate and $\gamma$ is the average recovery rate.  The movement of individuals from one compartment to another is represented through the ordinary differential equations below.  Both $\beta$ and $\gamma$ are constrained to be between $0$ and $1$.  Throughout, we use $X$, $Y$, and $Z$ for stand-ins for the ODEs for $S$, $I$, and $R$, respectively, to avoid any confusion with $\rr$.
\begin{align}
\systeme{\frac{dX}{dt} = -\frac{\beta XY}{N}, \frac{dY}{dt} = \frac{\beta XY}{N} - \gamma Y, \frac{dZ}{dt} = \gamma Z}. \label{eq:sir}
\end{align}
In words, susceptible individuals become infected at a rate that is proportional to the percentage of infected individuals multiplied by $\beta$, the infection rate, and the number of susceptible individuals.  Infectious individuals recover at a rate of $\gamma$ multiplied by the number of infected individuals.

The SIR model is represented graphically in Figure \ref{fig::sir}. 
\begin{figure}[h]
\centering
\begin{tikzpicture}[node distance=3cm,auto,>=latex',every node/.append style={align=center}]
    \node [int,  fill = white!70!blue] (a)              {$X$};
    \node [int,  fill = white!70!blue]           (c) [right of=a] {$Y$};
    \node [int,  fill = white!70!blue] (e) [right of=c] {$Z$};
    \path[->, auto=false] (a) edge node {} (c)
                          (c) edge node {} (e) ;

\end{tikzpicture}
\caption{Depiction of a SIR model where $X=S, Y=I,$ and $Z=R$.  One can only get the disease once in this model.}\label{fig::sir}
\end{figure}
An outbreak occurs if the rate of change of infectious individuals is positive,
\begin{align*}
  \frac{dY}{dt} &> 0 \\
  \frac{\beta X Y}{N}  - \gamma Y &> 0 ,\\
  Y \left ( \beta \frac{X}{N} - \gamma \right ) & > 0\\
   \frac{\beta}{\gamma} &> \frac{N}{X}.
\end{align*}
That is,  the rate of new infections is greater than the rate of recovery.  So as long as the number of susceptibles is large, $\frac{X}{N} \approx 1$, then an outbreak will occur if $\rr >1$,
\begin{align*}
  \rr \overset{def}{=} \frac{\beta}{\gamma} > 1.
  \end{align*}
  In order to incorporate randomness, we add noise, namely,
  \begin{align}\label{eq:sir-noise}
    X_{obs}(t) &= X(t) + \epsilon_{X,t}\\
    Y_{obs}(t) &=  N - X_{obs}(t) -Y_{obs}(t)  \nonumber\\
    Z_{obs}(t) &= Z(t) + \epsilon_{Z,t}. \nonumber
  \end{align}
We are assuming the observations are generated based on the deterministic ODEs presented in Equation \ref{eq:sir} with the addition of time ($t=0, 1, \dots, T$0) and compartment dependent noise $\epsilon_{X,t}$ and $\epsilon_{Z,t}$.  Since, $N$, the total population, is constant, then $Y_{obs}$ is adjusted accordingly.

In summary, when we discuss estimators for $\rr$ for the SIR model, we mean to say we are forming an estimator of $\rr$ from the given set of data in Eq. \ref{eq:sir-noise},
\begin{align*}
\textnormal{data} = \left \{\left (X_{obs}(t), Y_{obs}(t), Z_{obs}(t) \right ) : t=0, 1, \dots, T\right \}.
  \end{align*}.

\section{Overview of $\xxsir$ methods to estimate $\rr$ for the SIR model}
\label{sec:methods}

In this section, we describe the $\xxsir$ methods in detail.

\subsection{Least Squares ($\beta$, $\gamma$) (LS)}\label{least-squares-beta-gamma}
The first approach to estimate $\rr$ in the SIR model is to minimize the joint mean square error for the data collected at each time point and use the plug-in estimator found in Equation \ref{eq:sirls}.  In particular, we find

\begin{align*}
(\hat{\beta}, \hat{\gamma} )&=\arg \min_{\beta, \gamma} \sum_{t} \left [ \left (X_{obs}(t) - X(t)\right )^2 + \left ( Z_{obs}(t) - Z(t) \right )^2 \right ]
\end{align*}
Then  the estimate for $\rr$ is given by Equation \ref{eq:sirls},
\begin{align}\label{eq:sirls}
  \hat{\rr}= \frac{\hat{\beta}}{\hat{\gamma}}.
  \end{align}

\subsection{Reparametrized Least Squares ($\rr$, $\gamma$) (ReLS)}\label{reparametrized-least-squares-rux5f0-gamma}

In the previous method (Section \ref{least-squares-beta-gamma}), we estimated $\beta$ and $\gamma$ and then estimated $\rr$.  However, it is possible to directly estimate $\rr$ if we reparametrize the ODEs in Equation \eqref{eq:sir} directly with \(\rr\) and \(\gamma\), using the relation $\rr = \frac{\beta}{\gamma}$,

\begin{align*}
  \left \{
  \begin{array}{cl}
    \frac{dX}{dt} &= - \rr \gamma Y \frac{X}{N}\vspace{.5em}\\
    \frac{dY}{dt} &=  \rr \gamma Y \frac{X}{N}  - \gamma Y\vspace{.5em} \\
    \frac{dZ}{dt} &=  - \gamma Y 
  \end{array}
  \right . .
  \end{align*}
We find
\begin{align*}
(\hat{\rr}, \hat{\gamma} ) &= \text{argmin}_{\beta, \gamma} \sum_{t} \left [ \left (X_{obs}(t) - X(t)\right )^2 + \left ( Z_{obs}(t) - Z(t) \right )^2 \right ]
\end{align*}
We use the $\hat{\rr}$ directly from the above estimation process, which again can be done with gradient descent or another optimization process.

\subsection{Linear Model Approximation (LMA)}\label{linear-model-approximation-degree-10}

The SIR ODEs in Eq. \ref{eq:sir} have no known closed form solution, and so we are already using approximations using numerical integration, albeit small ones.  In addition to this, data collected from real diseases are typically very noisy to begin with.  \cite{chang2017} discovered that an SIR model may be well approximated by a linear model.  We use this approach here to estimate $\rr$.

Specifically, we estimate two linear polynomials in \(t\) with degree \(K= 10\) to \(X_{obs}\)
and \(Y_{obs}\) using least squares to find the coefficients $\{(\hat{x}_k,
\hat{y}_k)\}_{k=1, \dots, K}$,
\begin{align*}
\hat{X}(t) &= \sum_{k=0}^K \hat{x}_k t^k\\
{\hat{Y}}(t) &= \sum_{k=0}^K \hat{y}_k t^k
\end{align*}
Then, we estimate the derivatives as
\begin{align*}
\hat{X}^\prime(t) &= \sum_{k=1}^K k \hat{x}_k t^{k-1}\\
\hat{Y}^\prime(t) &= \sum_{k=0}^K k \hat{y}_k t^{k-1}
\end{align*}
Following,  an estimator for \(\rr\) is derived from the ODEs in Equation \eqref{eq:sir},
\begin{align}
  - \frac{X^\prime}{Z^\prime}&= \rr \frac{X}{N} \nonumber\\
  \rr &=       -\frac{X^\prime}{
        Z^\prime} \cdot \frac{N}{X} \nonumber\\
  \hat{\rr} &= -\frac{\hat{X}^\prime(0)}{ \hat{Z}^\prime(t)} \cdot \frac{N}{\hat{X}(0)}. \nonumber
  \end{align}
Here, $K=10$ is rather arbitrary and should be selected using some criterion such as AIC in addition with the knowledge of how many data points are available to the user.  Besides optionally deciding on the degree of polynomials to fit, this model is simple to implement and gives very comparable results to using least squares with the SIR model.  The time $t=0$ is used to best capture the initial outbreak.

\subsection{Linear Model Approximation, All Time Points (LMAT)}\label{linear-model-approximation-all-time-points-degree-10}

The above formulation (Section \ref{linear-model-approximation-degree-10} of a linear model approximation only uses the estimate at time $t=0$ to estimate $\rr$.  We can instead, use all time points available to estimate $\rr$.  We fit a linear polynomial of \(t\) with degree \(K= 10\) model to \(X\)
and \(Z\) as above, with a slight modification in how we estimate
\(\rr\),
\begin{align*}
  \hat{\rr} &= \frac{1}{\# \text{ Obs }t} \sum_t \frac{-\hat{X}^\prime(t)}{\hat{Z}^\prime(t)} \cdot \frac{N}{X(0)} 
\end{align*}
The intuition is that $\frac{-X^\prime}{Z^\prime(t)}$ is constant in $t$, but due to our approximations with the linear model, this is no longer the case.  Here, we average over the different possible values of $\rr$, estimated at different times.  An advantage to this approach is that we have a more robust estimate of $\rr$ than just using one time point.

\subsection{Incidence to Prevalence Ratio (IPR)}\label{incidence-to-prevalence-ratio}
The incidence to prevalence ratio (IPR), described by \cite{Nishiura2009} is another intuitive method to calculate $\rr$ as it incorporates some of the most basic epidemiological quantities, incidence and prevalence.

In terms of data from the SIR model, incidence $J(t) \approx -(X(t+1) - X(t))$, and the IPR$(t) = \frac{J(t)}{Y(t)}$.  This method assumes that we have some prior knowledge about $\gamma$, the transmission rate.  Thus we use as our estimate,
\begin{align*}
\hat{\rr} &= \textnormal{IPR}(t) \cdot \frac{1}{\gamma}
\end{align*}

Here we assume that the time step is small enough to approximate the incidence.  The advantage of this method is that incidence data is generally readily available as is prevalence data for certain diseases such as HIV.  However, as one is required to have prior knowledge about $\gamma$, it may be easier to directly estimate $\rr$ with one of the many other methods described that does not require a prior knowledge about $\gamma$.  Again, we are using only one time point to estimate $\rr$.

\subsection{Smoothed Incidence to Prevalence Ratio (SIPR)}
We use the same method as above, IPR, but first estimate splines with 4 degrees of freedom, $\hat{X}(t)$ and $\hat{Y}(t)$, , to fit to $X_{obs}$ and $Y_{obs}$, respectively.  Then  $J(t) \approx -(\hat{X}(t+1) - \hat{X}(t))$, and the $\hat{\textnormal{IPR}}(t) = \frac{J(t)}{\hat{Y}(t)}$.  Then
\begin{align*}
\rr &= \hat{\textnormal{IPR}}(t) \cdot D
\end{align*}
The advantage of this method is that it creates a less noisy estimate than estimating IPR using only one point.  It has the same disadvantages as the regular IPR ratio in that it requires knowledge about $\gamma$.

%%%%%%%%%%%%%%%%%%%%%
\subsection{Log-Linear (LL)}
\cite{harko2014exact} were able to reduce the SIR model to one ODE.  From this, we can derive the following,
\begin{align}
  X(t) &=  X(0) e^{\frac{\beta}{N \gamma}Z(t)} \nonumber\\
  \log \frac{X(t)}{X(0)} &=  \frac{\beta }{\gamma N }Z(t) \nonumber\\
  N\log \frac{X(t)}{X(0)} &=  \rr Z(t). \label{eq:harko_lin}
\end{align}

Thus, we can regress the left hand side in Eq. \ref{eq:harko_lin} on $Z(t)$, the number of recovered individuals, with $\rr$ as the coefficient of $Z(t)$ and no intercept term to obtain an estimate of $\hat{\rr}$.

\subsection{Markov chain estimation (MC)}
A natural approach to epidemic modeling is that of Markov chains (MC), since it is assumed an individual's next state is only dependent on its current state.  Much work has been done over the years in this specific field including asymptotic behavior, continuous time MC, confidence intervals, and more \citep{jacquez1991,gani1995,daley2001epidemic}.  We present one simple instantiation of the model, the discrete time case, which traces its origin back to the Reed-Frost model \citep{abbey1952}.

In this framework, the number of susceptibles at the next step, $X_{t}$, has a Binomial distribution based on the contacts with the current number of infectious, $Y_{t}$ and the current number of susceptibles.  That is $X_{t+1} \sim \text{Binom}(X_t, \alpha^{Y_t}$), where $\alpha$ is the probability of avoiding infection from an infective.  \cite{barbour2004} report that the reproduction number is thus,
\begin{align}\label{eq:r0-mc}
\rr &= \log \left ( \frac{1}{1-\alpha}\right ).
\end{align}

Thus, using regression on susceptible/infection counts will lead to an estimate of $\hat{\alpha}$ and hence $\hat{\rr}$.  However, this framework typically allows more than just the reproduction number to be estimated.  Through recursion, one can calculate the probability of having a given number of susceptibles and infected at each time step, and hence the entire probability distribution may be known.

\subsection{Sequential Bayes (SB)}\label{sec:seqbayes}

Described by \cite{bettencourt2008} and summarized in \cite{obadia2012r0}, the sequential Bayes method is a Bayesian approach to an approximation of the classic SIR model.  The approximated SIR model assumes that the incidence at $t+1$, $J(t+1)$ has a Poisson distribution, with $\gamma$ as the  average inverse of the infectious period. In order to estimate $\rr$, we must have some idea about $\gamma$,
\begin{align*}
J(t+1)  \sim Poisson( J(t) \exp \left \{  \gamma (\rr-1)\right \})
\end{align*}
Then, the posterior distribution of $\rr$ given the previous days' incidences is
\begin{align*}
  P(\rr | J_0, \dots, J_{t+1}) = \frac{P(J_{t+1} | \rr, J_0, \dots, J_t)P(\rr| J_0, \dots, J_t)}{P(J_0, \dots, J_{t+1})}.
\end{align*}
This method is sequential in that the prior distribution for $\rr$ comes from the previous day.  The initial prior for $\rr$ is assumed to be flat.  This method results in a posterior distribution from which credible intervals may be obtained.  This method assumes, initial growth in incidence to be exponential, and homogeneous mixing of populations as with any compartment model.  The advantages of this method are that of the ability to obtain an entire posterior distribution, whereas many other methods are difficult to even find an estimate of the variance.  Disadvantages include strict assumptions about the distributions and computational time required to estimate the relevant parameters.

\subsection{MLE}

\section{Confidence and Credible Intervals}
\label{sec:ci}

Estimating $\rr$ is difficult and estimating $V[\rr]$, the variance and subsequently CIs, are even more so.  We describe general methods which may be applicable to estimate the variance.  In particular, we describe three methods 1) the delta method, 2) the block strap, and 3) posterior distributions.

\subsection{Delta Method}\label{delta-method}

When the method estimates \(\beta\) and \(\gamma\) instead of \(\rr\) directly, we use the delta method approximation to calculate the
variance of \(\rr\). Here, we know \(\rr = h(\beta, \gamma) = \frac{\beta}{\gamma}\). Then \(\bigtriangleup h = (\frac{1}{\gamma},  -\frac{\beta}{\gamma^2})^T\) and \(V[\rr] = \bigtriangleup h^T \Sigma_{\beta, \gamma} \bigtriangleup h\), where \(\Sigma_{\beta, \gamma}\) is covariance matrix of \(\beta\) and \(\gamma\).

This estimate assumes that the distribution of $\beta$ and $\gamma$ are asymptotically normal.  Here, we use the relationship of $\rr = \frac{\beta}{\gamma}$ in order to estimate the variance, which is specific to the SIR framework.  This method may be extended to other frameworks.  The advantages of the delta method are that as the NGM posits a recipe to derive $\rr$, then it is theoretically possible to use the delta method to derive an expression for the variance of $\rr$ for any compartment model.

\subsection{Block Bootstrap}

The block bootstrap is a variant of the bootstrap (cite) in which the $n$ observations are dependent on one another.  In contrast to the original bootstrap, the block bootstrap partitions consectutive observations into $k$ blocks with block-length $b$ ($n=kb$).  These are non-overlapping partitions or blocks, although there are variants for which blocks can overlap (cite).  Following, for iteration $\ell$, one samples uniformly replacement $k$ blocks and estimates $\hat{\rr}_b$ from those $k$ blocks, repeating the sampling and estimation for a total of $L$ estimates of $\hat{\rr}$.  Then
\begin{align*}
  V\left [ \hat{\rr} \right ] &\approx V\left [\hat{\rr}_\ell \right ].
\end{align*}
More about the asympotic distributions of the blockwise-bootstrap can be found in \cite{cao1999}, along with descriptions of variants.

The assympotic properties hold when the stochastic process is $m$-dependent, that is when observations $t, t+1, \dots$ and $s, s+1, \dots$ are independent whenver $s+m < t$ and under some smoothness conditions of the statistic used to estimate $\rr$.

A choice one must make in the block bootstrap is the block size $b$.  One simple recommendation is to take $b= n^{1/3}$, which is derived from the asymptotic mean square error of the method.

\subsection{Posterior Distribution}
When one uses a Bayesian approach to estimate $\rr$, the result should be a distribution rather than an estimate.  One can simply look at the quantities of this distribution to form a credible interval for this estimate.  Advantages of this method include having an entire distribution compared to a point estimate.  One disadvantage is that the Bayesian approach will not guarantee frequentist properties that epidemiologists may be looking for such as being able to compare diseases to one another.

%%%%%%%%%%%%%%%

\section{Simulation Results}\label{sec:sim-res}

\section{Discussion}\label{discussion}

\bibliographystyle{apa}%Choose a bibliograhpic style
\bibliography{Master}

\section{Appendix}

\subsection{Temporary Hold: Baseline Data}

\begin{table}[H]

	\caption{\label{tab:}$\rr$ Estimates and Std. Errs, SIRLS Model,
		$X_0 = 99950, Y_0 = 50$, $\sigma_X = 100, \sigma_Y = 5$,$\beta = 0.06, \gamma = 0.03$}
	\centering
	\begin{footnotesize}
	\begin{tabular}[t]{l|r|r|r|r|r|r|r|r}
		\hline
		Data Set & Norm Est & Norm SE & Norm-M Est & Norm-M SE & AR Est & AR SE & AR-M Est & AR-M SE\\
		\hline
		Baseline1 & 1.999919 & 0.0058322 & 2.000036 & 0.0044639 & 2.000020 & 0.0078308 & 2.000859 & 0.0060766\\
		\hline
		Baseline2 & 1.999971 & 0.0055923 & 2.000246 & 0.0044456 & 1.999739 & 0.0071509 & 1.999992 & 0.0073664\\
		\hline
		Baseline3 & 2.000327 & 0.0054549 & 1.999861 & 0.0048111 & 1.999165 & 0.0078453 & 2.000160 & 0.0063235\\
		\hline
		Baseline4 & 2.000520 & 0.0057238 & 1.999501 & 0.0046540 & 1.999282 & 0.0077363 & 1.999289 & 0.0065290\\
		\hline
		Baseline5 & 1.999919 & 0.0056334 & 2.000053 & 0.0045020 & 2.001443 & 0.0081239 & 2.002397 & 0.0075056\\
		\hline
	\end{tabular}
\end{footnotesize}
\end{table}
\begin{table}[H]

	\caption{\label{tab:}$\rr$ Estimates and Std. Errs, SIRREPAR Model,
		$X_0 = 99950, Y_0 = 50$, $\sigma_X = 100, \sigma_Y = 5$,$\beta = 0.06, \gamma = 0.03$}
	\centering
	\begin{footnotesize}
	\begin{tabular}[t]{l|r|r|r|r|r|r|r|r}
		\hline
		Data Set & Norm Est & Norm SE & Norm-M Est & Norm-M SE & AR Est & AR SE & AR-M Est & AR-M SE\\
		\hline
		Baseline1 & 1.999708 & 0.0052113 & 2.000030 & 0.0039910 & 1.999885 & 0.0069901 & 2.001023 & 0.0054369\\
		\hline
		Baseline2 & 1.999734 & 0.0049939 & 2.000314 & 0.0039684 & 2.000046 & 0.0063859 & 1.999684 & 0.0065745\\
		\hline
		Baseline3 & 2.000604 & 0.0048757 & 2.000025 & 0.0042948 & 1.999356 & 0.0069903 & 2.000438 & 0.0056503\\
		\hline
		Baseline4 & 2.000410 & 0.0051162 & 1.999830 & 0.0041504 & 1.999885 & 0.0069068 & 1.999013 & 0.0058261\\
		\hline
		Baseline5 & 1.999830 & 0.0050288 & 1.999830 & 0.0040212 & 2.001592 & 0.0072630 & 2.002090 & 0.0067284\\
		\hline
	\end{tabular}
\end{footnotesize}
\end{table}
\begin{table}[H]

	\caption{\label{tab:}$\rr$ Estimates and Std. Errs, SIRLMA Model,
		$X_0 = 99950, Y_0 = 50$, $\sigma_X = 100, \sigma_Y = 5$,$\beta = 0.06, \gamma = 0.03$}
	\centering
	\begin{footnotesize}
	\begin{tabular}[t]{l|r|r|r|r|r|r|r|r}
		\hline
		Data Set & Norm Est & Norm SE & Norm-M Est & Norm-M SE & AR Est & AR SE & AR-M Est & AR-M SE\\
		\hline
		Baseline1 & 1.999502 & 0.3896154 & 2.694876 & 0.5083823 & 35.700310 & 4820.9883994 & 1.669865 & 0.1717277\\
		\hline
		Baseline2 & 1.688318 & 0.1882819 & 1.681966 & 0.1102752 & 1.477492 & 0.1124589 & 1.537388 & 0.1495794\\
		\hline
		Baseline3 & 1.574728 & 0.1339888 & 1.746144 & 0.1668073 & 5.340898 & 33.0770287 & 2.314664 & 0.6038287\\
		\hline
		Baseline4 & 1.642471 & 0.1949104 & 2.047336 & 0.3952098 & 1.328799 & 0.0667234 & -3.007100 & 7.5904183\\
		\hline
		Baseline5 & 1.593838 & 0.1684652 & 1.843829 & 0.2199302 & 1.490924 & 0.1280881 & 1.871116 & 0.2218524\\
		\hline
	\end{tabular}
\end{footnotesize}
\end{table}
\begin{table}[H]

	\caption{\label{tab:}$\rr$ Estimates and Std. Errs, SIRLMAT Model,
		$X_0 = 99950, Y_0 = 50$, $\sigma_X = 100, \sigma_Y = 5$,$\beta = 0.06, \gamma = 0.03$}
	\centering
	\begin{footnotesize}
	\begin{tabular}[t]{l|r|r|r|r|r|r|r|r}
		\hline
		Data Set & Norm Est & Norm SE & Norm-M Est & Norm-M SE & AR Est & AR SE & AR-M Est & AR-M SE\\
		\hline
		Baseline1 & 1.399476 & 0.6034464 & 1.422638 & 0.7411130 & 1.050638 & 4.8889559 & 1.403266 & 0.8205110\\
		\hline
		Baseline2 & 1.407952 & 0.6316481 & 1.444502 & 0.9255988 & 1.372240 & 0.8699573 & 1.451384 & 0.7475245\\
		\hline
		Baseline3 & 1.410617 & 0.6223034 & 1.360441 & 0.8507911 & 1.374595 & 1.2564522 & 1.437625 & 0.6731037\\
		\hline
		Baseline4 & 1.389717 & 0.6186280 & 1.391034 & 0.5732198 & 1.598858 & 1.3621554 & 1.324797 & 0.6542186\\
		\hline
		Baseline5 & 1.390334 & 0.6670626 & 1.387943 & 0.5756532 & 1.362033 & 1.0034553 & 1.541135 & 0.9082263\\
		\hline
	\end{tabular}
\end{footnotesize}
\end{table}
\begin{table}[H]

	\caption{\label{tab:}$\rr$ Estimates and Std. Errs, SIRIPR Model,
		$X_0 = 99950, Y_0 = 50$, $\sigma_X = 100, \sigma_Y = 5$,$\beta = 0.06, \gamma = 0.03$}
	\centering
	\begin{footnotesize}
	\begin{tabular}[t]{l|r|r|r|r|r|r|r|r}
		\hline
		Data Set & Norm Est & Norm SE & Norm-M Est & Norm-M SE & AR Est & AR SE & AR-M Est & AR-M SE\\
		\hline
		Baseline1 & 3.111569 & 10.849009 & 0.8738855 & 1.543847 & 2.620045 & 8.877271 & 0.9175130 & 2.234642\\
		\hline
		Baseline2 & 2.652218 & 7.415605 & 0.8982691 & 1.805619 & 2.055229 & 6.085662 & 0.9016198 & 1.872737\\
		\hline
		Baseline3 & 2.767507 & 8.021874 & 0.8648371 & 1.477029 & 2.062531 & 5.512523 & 1.0049512 & 3.050472\\
		\hline
		Baseline4 & 3.057198 & 11.812401 & 1.0174550 & 2.864584 & 2.180584 & 6.925265 & 1.0325816 & 2.768722\\
		\hline
		Baseline5 & 2.668934 & 7.472743 & 0.9696804 & 2.452165 & 2.231211 & 5.995580 & 0.9224206 & 2.770667\\
		\hline
	\end{tabular}
\end{footnotesize}
\end{table}
\begin{table}[H]

	\caption{\label{tab:}$\rr$ Estimates and Std. Errs, SIRSIPR Model,
		$X_0 = 99950, Y_0 = 50$, $\sigma_X = 100, \sigma_Y = 5$,$\beta = 0.06, \gamma = 0.03$}
	\centering
	\begin{footnotesize}
	\begin{tabular}[t]{l|r|r|r|r|r|r|r|r}
		\hline
		Data Set & Norm Est & Norm SE & Norm-M Est & Norm-M SE & AR Est & AR SE & AR-M Est & AR-M SE\\
		\hline
		Baseline1 & 1.110211 & 0 & 1.030592 & 0 & 1.006362 & 0 & 0.9295163 & 0\\
		\hline
		Baseline2 & 1.132249 & 0 & 1.089786 & 0 & 1.055082 & 0 & 0.8371411 & 0\\
		\hline
		Baseline3 & 1.163383 & 0 & 1.102025 & 0 & 1.207890 & 0 & 0.8555288 & 0\\
		\hline
		Baseline4 & 1.177337 & 0 & 1.049182 & 0 & 1.039892 & 0 & 1.1049484 & 0\\
		\hline
		Baseline5 & 1.113234 & 0 & 1.092939 & 0 & 1.381775 & 0 & 0.9232862 & 0\\
		\hline
	\end{tabular}
\end{footnotesize}
\end{table}
\begin{table}[H]

	\caption{\label{tab:}$\rr$ Estimates and Std. Errs, LOGLINEAR Model,
		$X_0 = 99950, Y_0 = 50$, $\sigma_X = 100, \sigma_Y = 5$,$\beta = 0.06, \gamma = 0.03$}
	\centering
	\begin{footnotesize}
	\begin{tabular}[t]{l|r|r|r|r|r|r|r|r}
		\hline
		Data Set & Norm Est & Norm SE & Norm-M Est & Norm-M SE & AR Est & AR SE & AR-M Est & AR-M SE\\
		\hline
		Baseline1 & 1.999840 & 0.0002227 & 1.999761 & 0.0001427 & 1.999820 & 0.0003227 & 2.000443 & 0.0001910\\
		\hline
		Baseline2 & 1.999807 & 0.0002198 & 2.000303 & 0.0001342 & 2.000000 & 0.0002885 & 1.999556 & 0.0002342\\
		\hline
		Baseline3 & 2.000386 & 0.0002191 & 2.000042 & 0.0001598 & 1.998138 & 0.0003215 & 1.999213 & 0.0002316\\
		\hline
		Baseline4 & 2.000039 & 0.0002252 & 1.999710 & 0.0001506 & 1.999217 & 0.0002895 & 1.999525 & 0.0002144\\
		\hline
		Baseline5 & 2.000148 & 0.0002194 & 1.999993 & 0.0001479 & 2.001889 & 0.0003495 & 2.001873 & 0.0002438\\
		\hline
	\end{tabular}
\end{footnotesize}
\end{table}
\begin{table}[H]

	\caption{\label{tab:}$\rr$ Estimates and Std. Errs, SIRMC Model,
		$X_0 = 99950, Y_0 = 50$, $\sigma_X = 100, \sigma_Y = 5$,$\beta = 0.06, \gamma = 0.03$}
	\centering
	\begin{footnotesize}
	\begin{tabular}[t]{l|r|r|r|r|r|r|r|r}
		\hline
		Data Set & Norm Est & Norm SE & Norm-M Est & Norm-M SE & AR Est & AR SE & AR-M Est & AR-M SE\\
		\hline
		Baseline1 & 0.9485555 & 0 & 0.9485555 & 0 & 0.9485555 & 0 & 0.9485555 & 0\\
		\hline
		Baseline2 & 0.9485555 & 0 & 0.9485555 & 0 & 0.9485555 & 0 & 0.9485555 & 0\\
		\hline
		Baseline3 & 0.9485555 & 0 & 0.9485555 & 0 & 0.9485555 & 0 & 0.9485555 & 0\\
		\hline
		Baseline4 & 0.9485555 & 0 & 0.9485555 & 0 & 0.9485555 & 0 & 0.9485555 & 0\\
		\hline
		Baseline5 & 0.9485555 & 0 & 0.9485555 & 0 & 0.9485555 & 0 & 0.9485555 & 0\\
		\hline
	\end{tabular}
\end{footnotesize}
\end{table}
\begin{table}[H]

	\caption{\label{tab:}$\rr$ Estimates and Std. Errs, SEQBAYES Model,
		$X_0 = 99950, Y_0 = 50$, $\sigma_X = 100, \sigma_Y = 5$, $\beta = 0.06, \gamma = 0.03$}
	\centering
	\begin{footnotesize}
	\begin{tabular}[t]{l|r|r|r|r|r|r|r|r}
		\hline
		Data Set & Norm Est & Norm SE & Norm-M Est & Norm-M SE & AR Est & AR SE & AR-M Est & AR-M SE\\
		\hline
		Baseline1 & 1.0000000 & 0.0523424 & 0.8099669 & 0.0471072 & 1.0000000 & 0.0523424 & 1.000000 & 0.0523424\\
		\hline
		Baseline2 & 1.3936378 & 0.0617915 & 1.0000000 & 0.0523424 & 1.0000000 & 0.0523424 & 1.432940 & 0.0626567\\
		\hline
		Baseline3 & 1.4504175 & 0.0630377 & 1.1185890 & 0.0553591 & 1.4331404 & 0.0626611 & 1.000000 & 0.0523424\\
		\hline
		Baseline4 & 1.0000000 & 0.0523424 & 1.3993848 & 0.0619187 & 0.8221337 & 0.0474597 & 1.386606 & 0.0616354\\
		\hline
		Baseline5 & 0.9935711 & 0.0521739 & 1.4389154 & 0.0627872 & 1.3236479 & 0.0602199 & 1.014455 & 0.0527193\\
		\hline
	\end{tabular}
\end{footnotesize}
\end{table}

\subsection{Temporary Hold: Parameter Change}

\begin{table}[H]

	\caption{\label{tab:}$\rr$ Estimates and Std. Errs, SIRLS Model,
		$X_0 = 99950, Y_0 = 50$, $\sigma_X = 100, \sigma_Y = 5$, $\beta = 0.06$}
	\centering
	\begin{footnotesize}
	\begin{tabular}[t]{l|r|r|r|r|r|r|r|r}
		\hline
		Data Set & Norm Est & Norm SE & Norm-M Est & Norm-M SE & AR Est & AR SE & AR-M Est & AR-M SE\\
		\hline
		$\gamma$ = 0.001 & 59.9764682 & 0.0434705 & 59.9764682 & 0.0386733 & 60.0151171 & 0.0697888 & 59.9589126 & 0.0631448\\
		\hline
		$\gamma$ = 0.04 & 1.5000763 & 0.0126817 & 1.5001978 & 0.0090880 & 1.4989700 & 0.0182852 & 1.4993603 & 0.0165168\\
		\hline
		$\gamma$ = 0.06 & 1.0043054 & 0.4140424 & 0.9955554 & 0.1086448 & 0.9915431 & 0.6272644 & 0.9941392 & 0.1891812\\
		\hline
		$\gamma$ = 0.24 & 0.2363274 & 4.1042813 & 0.2327691 & 0.6701185 & -0.0695410 & 2.8291362 & 0.3021083 & 0.5657139\\
		\hline
	\end{tabular}
\end{footnotesize}
\end{table}
\begin{table}[H]

	\caption{\label{tab:}$\rr$ Estimates and Std. Errs, SIRREPAR Model,
		$X_0 = 99950, Y_0 = 50$, $\sigma_X = 100, \sigma_Y = 5$, $\beta = 0.06$}
	\centering
	\begin{footnotesize}
	\begin{tabular}[t]{l|r|r|r|r|r|r|r|r}
		\hline
		Data Set & Norm Est & Norm SE & Norm-M Est & Norm-M SE & AR Est & AR SE & AR-M Est & AR-M SE\\
		\hline
		1st Order & 59.995188 & 0.4343984 & 59.9951876 & 0.3874879 & 60.025452 & 0.6911933 & 59.9960910 & 0.6214841\\
		\hline
		4th Order & 1.499883 & 0.0062529 & 1.5001705 & 0.0044936 & 1.499068 & 0.0090313 & 1.4993177 & 0.0081651\\
		\hline
		Linear SIR & 1.004306 & 0.0501444 & 0.9955446 & 0.0117215 & 0.991558 & 0.0645788 & 0.9941511 & 0.0190525\\
		\hline
		& 4.179525 & 2.7455569 & 3.5709590 & 0.3171284 & 5.192951 & 3.5249847 & 9.8415034 & 2.5646198\\
		\hline
	\end{tabular}
\end{footnotesize}
\end{table}
\begin{table}[H]

	\caption{\label{tab:}$\rr$ Estimates and Std. Errs, SIRLMA Model,
		$X_0 = 99950, Y_0 = 50$, $\sigma_X = 100, \sigma_Y = 5$, $\beta = 0.06$}
	\centering
	\begin{footnotesize}
	\begin{tabular}[t]{l|r|r|r|r|r|r|r|r}
		\hline
		Data Set & Norm Est & Norm SE & Norm-M Est & Norm-M SE & AR Est & AR SE & AR-M Est & AR-M SE\\
		\hline
		$\gamma$ = 0.001 & -122.0810136 & 104.7144676 & -57.2648513 & 11.6741827 & -19.8476503 & 6.3159285 & -190.2995380 & 290.3182374\\
		\hline
		$\gamma$ = 0.04 & 1.1085550 & 0.0868180 & -3.9714925 & 1115.8099120 & 0.9776237 & 0.0216463 & 0.7343152 & 0.1793076\\
		\hline
		$\gamma$ = 0.06 & 1.0835046 & 0.1530378 & 0.9903843 & 0.1184626 & 0.9273120 & 0.0602626 & 1.0387809 & 0.0361673\\
		\hline
		$\gamma$ = 0.24 & 0.7069533 & 0.2023701 & 0.5215129 & 0.1714423 & 0.8747402 & 0.0586876 & -49.7565356 & 5300.3981000\\
		\hline
	\end{tabular}
\end{footnotesize}
\end{table}
\begin{table}[H]

	\caption{\label{tab:}$\rr$ Estimates and Std. Errs, SIRLMAT Model,
		$X_0 = 99950, Y_0 = 50$, $\sigma_X = 100, \sigma_Y = 5$, $\beta = 0.06$}
	\centering
	\begin{footnotesize}
	\begin{tabular}[t]{l|r|r|r|r|r|r|r|r}
		\hline
		Data Set & Norm Est & Norm SE & Norm-M Est & Norm-M SE & AR Est & AR SE & AR-M Est & AR-M SE\\
		\hline
		$\gamma$ = 0.001 & -2.9530478 & 200.4018824 & 8.0337261 & 106.7435722 & 78.891674 & 488.2036231 & -72.3228279 & 1659.2071496\\
		\hline
		$\gamma$ = 0.04 & 1.4367115 & 0.7496644 & 1.4714250 & 3.0708882 & 1.295474 & 0.3349151 & 1.2785848 & 0.2451543\\
		\hline
		$\gamma$ = 0.06 & 0.9853118 & 0.1423857 & 0.9943726 & 0.0160834 & 1.007444 & 0.5983482 & 1.0258649 & 0.5010848\\
		\hline
		$\gamma$ = 0.24 & 0.6742069 & 6.2426393 & 0.3870527 & 4.8804441 & 1.785998 & 15.9375643 & 0.2220084 & 5.4085242\\
		\hline
	\end{tabular}
\end{footnotesize}
\end{table}
\begin{table}[H]

	\caption{\label{tab:}$\rr$ Estimates and Std. Errs, SIRIPR Model,
		$X_0 = 99950, Y_0 = 50$, $\sigma_X = 100, \sigma_Y = 5$, $\beta = 0.06$}
	\centering
	\begin{footnotesize}
	\begin{tabular}[t]{l|r|r|r|r|r|r|r|r}
		\hline
		Data Set & Norm Est & Norm SE & Norm-M Est & Norm-M SE & AR Est & AR SE & AR-M Est & AR-M SE\\
		\hline
		$\gamma$ = 0.001 & 1.082930 & 4.005191 & 0.4881735 & 1.323335 & 1.174475 & 4.399781 & 0.5704615 & 2.418576\\
		\hline
		$\gamma$ = 0.04 & 3.817206 & 10.282636 & 1.1102025 & 2.104624 & 2.891003 & 6.829224 & 1.1172480 & 2.272797\\
		\hline
		$\gamma$ = 0.06 & 23.793692 & 34.952641 & 1.3460001 & 5.280598 & 18.781262 & 28.247185 & 1.3502216 & 5.880026\\
		\hline
		$\gamma$ = 0.24 & 868.875851 & 3562.122383 & 32.0292478 & 431.693917 & 711.882081 & 3552.087545 & 3.9097302 & 56.980358\\
		\hline
	\end{tabular}
\end{footnotesize}
\end{table}
\begin{table}[H]

	\caption{\label{tab:}$\rr$ Estimates and Std. Errs, SIRSIPR Model,
		$X_0 = 99950, Y_0 = 50$, $\sigma_X = 100, \sigma_Y = 5$, $\beta = 0.06$}
	\centering
	\begin{footnotesize}
	\begin{tabular}[t]{l|r|r|r|r|r|r|r|r}
		\hline
		Data Set & Norm Est & Norm SE & Norm-M Est & Norm-M SE & AR Est & AR SE & AR-M Est & AR-M SE\\
		\hline
		$\gamma$ = 0.001 & 0.0792904 & 0.0000000 & 0.0765708 & 0.0000000 & 0.0781112 & 0.0000000 & 0.0742256 & 0.0000000\\
		\hline
		$\gamma$ = 0.04 & 1.3960820 & 0.0000000 & 2.3028451 & 0.0000000 & 54.9912000 & 0.0000000 & 0.9972600 & 0.0000000\\
		\hline
		$\gamma$ = 0.06 & 1.1903748 & 0.0670862 & 1.1910744 & 0.0357340 & 1.0461934 & 0.1047191 & 1.2352948 & 0.0612702\\
		\hline
		$\gamma$ = 0.24 & -1.1924562 & 0.5566890 & -0.1523352 & 0.1610466 & -13.7485924 & 0.0000000 & 3.5354100 & 0.3241589\\
		\hline
	\end{tabular}
\end{footnotesize}
\end{table}
\begin{table}[H]

	\caption{\label{tab:}$\rr$ Estimates and Std. Errs, LOGLINEAR Model,
		$X_0 = 99950, Y_0 = 50$, $\sigma_X = 100, \sigma_Y = 5$, $\beta = 0.06$}
	\centering
	\begin{footnotesize}
	\begin{tabular}[t]{l|r|r|r|r|r|r|r|r}
		\hline
		Data Set & Norm Est & Norm SE & Norm-M Est & Norm-M SE & AR Est & AR SE & AR-M Est & AR-M SE\\
		\hline
		$\gamma$ = 0.001 & 46.3209595 & 0.7797216 & 59.8300371 & 0.1079267 & 41.897714 & 0.9767785 & 70.7075929 & 0.9517239\\
		\hline
		$\gamma$ = 0.04 & 1.5002728 & 0.0001905 & 1.5000436 & 0.0000928 & 1.499516 & 0.0002653 & 1.5000445 & 0.0001481\\
		\hline
		$\gamma$ = 0.06 & 0.9989449 & 0.0007428 & 0.9994831 & 0.0003320 & 1.000275 & 0.0010605 & 1.0008545 & 0.0004910\\
		\hline
		$\gamma$ = 0.24 & 0.9912397 & 0.0030505 & 0.8623034 & 0.0087543 & 1.001367 & 0.0025488 & 0.9198973 & 0.0052616\\
		\hline
	\end{tabular}
\end{footnotesize}
\end{table}
\begin{table}[H]

	\caption{\label{tab:}$\rr$ Estimates and Std. Errs, SIRMC Model,
		$X_0 = 99950, Y_0 = 50$, $\sigma_X = 100, \sigma_Y = 5$, $\beta = 0.06$}
	\centering
	\begin{footnotesize}
	\begin{tabular}[t]{l|r|r|r|r|r|r|r|r}
		\hline
		Data Set & Norm Est & Norm SE & Norm-M Est & Norm-M SE & AR Est & AR SE & AR-M Est & AR-M SE\\
		\hline
		$\gamma$ = 0.001 & 0.9485554 & 1.0e-07 & 0.9485555 & 0.0e+00 & 0.9485555 & 0.00e+00 & 0.9485555 & 0e+00\\
		\hline
		$\gamma$ = 0.04 & 0.9485555 & 0.0e+00 & 0.9485555 & 0.0e+00 & 0.9485555 & 0.00e+00 & 0.9485555 & 0e+00\\
		\hline
		$\gamma$ = 0.06 & 0.9485555 & 2.0e-07 & 0.9485555 & 0.0e+00 & 0.9485555 & 2.00e-07 & 0.9485555 & 0e+00\\
		\hline
		$\gamma$ = 0.24 & 0.9485812 & 1.5e-05 & 0.9485605 & 1.4e-06 & 0.9486128 & 1.26e-05 & 0.9485558 & 2e-07\\
		\hline
	\end{tabular}
\end{footnotesize}
\end{table}
\begin{table}[H]

	\caption{\label{tab:}$\rr$ Estimates and Std. Errs, SEQBAYES Model,
		$X_0 = 99950, Y_0 = 50$, $\sigma_X = 100, \sigma_Y = 5$, $\beta = 0.06$}
	\centering
	\begin{footnotesize}
	\begin{tabular}[t]{l|r|r|r|r|r|r|r|r}
		\hline
		Data Set & Norm Est & Norm SE & Norm-M Est & Norm-M SE & AR Est & AR SE & AR-M Est & AR-M SE\\
		\hline
		$\gamma$ = 0.001 & 0.9769731 & 0.0517362 & 1.000000 & 0.0523424 & 1.000000 & 0.0523424 & 1.000000 & 0.0523424\\
		\hline
		$\gamma$ = 0.04 & 1.2060946 & 0.0574836 & 1.000000 & 0.0523424 & 1.073894 & 0.0542418 & 1.115316 & 0.0552780\\
		\hline
		$\gamma$ = 0.06 & 1.0000000 & 0.0523424 & 1.000000 & 0.0523424 & 0.640393 & 0.0418868 & 1.000000 & 0.0523424\\
		\hline
		$\gamma$ = 0.24 & 1.2354952 & 0.0581800 & 1.338591 & 0.0605588 & 1.009694 & 0.0525955 & 1.000000 & 0.0523424\\
		\hline
	\end{tabular}
\end{footnotesize}
\end{table}

\subsection{Temporary Hold: Error Change}

\begin{table}[H]
	\caption{\label{tab:}$\rr$ Estimates and Std. Errs, SIRLS Model,
		$X_0 = 99950, Y_0 = 50$, $\beta = 0.06, \gamma = 0.03$}
	\centering
	\begin{footnotesize}
	\begin{tabular}[t]{l|r|r|r|r|r|r|r|r}
		\hline
		Data Set & Norm Est & Norm SE & Norm-M Est & Norm-M SE & AR Est & AR SE & AR-M Est & AR-M SE\\
		\hline
		$\sigma_X = 50, \sigma_Y = 2$ & 1.993455 & 0.5784530 & 1.979966 & 0.1204126 & 2.033934 & 0.7559710 & 2.066706 & 0.3382834\\
		\hline
		$\sigma_X = 500, \sigma_Y = 20$ & 1.998088 & 0.1353803 & 1.996950 & 0.0541332 & 1.999686 & 0.1893888 & 2.002135 & 0.0927952\\
		\hline
		$\sigma_X = 2500, \sigma_Y = 100$ & 1.999755 & 0.0026502 & 1.999834 & 0.0026397 & 1.999555 & 0.0037664 & 1.999038 & 0.0036473\\
		\hline
		$\sigma_X = 10000, \sigma_Y = 500$ & 2.001870 & 0.0259428 & 1.997748 & 0.0162536 & 2.002792 & 0.0416203 & 1.997918 & 0.0257997\\
		\hline
	\end{tabular}
\end{footnotesize}
\end{table}
\begin{table}[H]

	\caption{\label{tab:}$\rr$ Estimates and Std. Errs, SIRREPAR Model,
		$X_0 = 99950, Y_0 = 50$, $\beta = 0.06, \gamma = 0.03$}
	\centering
	\begin{footnotesize}
	\begin{tabular}[t]{l|r|r|r|r|r|r|r|r}
		\hline
		Data Set & Norm Est & Norm SE & Norm-M Est & Norm-M SE & AR Est & AR SE & AR-M Est & AR-M SE\\
		\hline
		$\sigma_X = 50, \sigma_Y = 2$ & 1.993660 & 0.5115266 & 1.979973 & 0.1053955 & 2.033636 & 0.7059701 & 2.066476 & 0.3172823\\
		\hline
		$\sigma_X = 500, \sigma_Y = 20$ & 1.998145 & 0.1207935 & 1.997147 & 0.0481545 & 2.000194 & 0.1680126 & 2.002244 & 0.0826643\\
		\hline
		$\sigma_X = 2500, \sigma_Y = 100$ & 1.999829 & 0.0023646 & 1.999946 & 0.0023591 & 1.999885 & 0.0033514 & 1.999483 & 0.0032200\\
		\hline
		$\sigma_X = 10000, \sigma_Y = 500$ & 2.002305 & 0.0232636 & 1.997996 & 0.0144744 & 2.003065 & 0.0372826 & 1.997732 & 0.0230069\\
		\hline
	\end{tabular}
\end{footnotesize}
\end{table}
\begin{table}[H]

	\caption{\label{tab:}$\rr$ Estimates and Std. Errs, SIRLMA Model,
		$X_0 = 99950, Y_0 = 50$, $\beta = 0.06, \gamma = 0.03$}
	\centering
	\begin{footnotesize}
	\begin{tabular}[t]{l|r|r|r|r|r|r|r|r}
		\hline
		Data Set & Norm Est & Norm SE & Norm-M Est & Norm-M SE & AR Est & AR SE & AR-M Est & AR-M SE\\
		\hline
		$\sigma_X = 50, \sigma_Y = 2$ & 1.1308571 & 0.1012907 & 3.899869 & 4.3543097 & 0.6625391 & 0.2163070 & 1.2910663 & 1.0083954\\
		\hline
		$\sigma_X = 500, \sigma_Y = 20$ & 0.4627433 & 7.5197341 & 4.178743 & 65.7755318 & 1.0790398 & 0.0165707 & -0.1152434 & 1.8870865\\
		\hline
		$\sigma_X = 2500, \sigma_Y = 100$ & 2.0417370 & 0.2771862 & 1.629287 & 0.0865344 & 1.4661799 & 0.0749495 & 2.8537066 & 0.5542235\\
		\hline
		$\sigma_X = 10000, \sigma_Y = 500$ & 1.6650841 & 1.0679885 & 1.686458 & 0.2567413 & 1.1579492 & 0.0983696 & -4.9577781 & 32.2547059\\
		\hline
	\end{tabular}
\end{footnotesize}
\end{table}
\begin{table}[H]

	\caption{\label{tab:}$\rr$ Estimates and Std. Errs, SIRLMAT Model,
		$X_0 = 99950, Y_0 = 50$, $\beta = 0.06, \gamma = 0.03$}
	\centering
	\begin{footnotesize}
	\begin{tabular}[t]{l|r|r|r|r|r|r|r|r}
		\hline
		Data Set & Norm Est & Norm SE & Norm-M Est & Norm-M SE & AR Est & AR SE & AR-M Est & AR-M SE\\
		\hline
		$\sigma_X = 50, \sigma_Y = 2$ & 4.991704 & 73.3084024 & 1.471488 & 0.9852195 & 0.8769744 & 1.2146547 & 1.0159519 & 2.9569665\\
		\hline
		$\sigma_X = 500, \sigma_Y = 20$ & 1.281010 & 4.3716385 & 1.356436 & 3.2462324 & 1.2444787 & 1.6725599 & 0.9326791 & 4.3459096\\
		\hline
		$\sigma_X = 2500, \sigma_Y = 100$ & 1.416200 & 0.5785114 & 1.512228 & 2.1276407 & 1.4016766 & 0.7202478 & 1.4274038 & 0.6787529\\
		\hline
		$\sigma_X = 10000, \sigma_Y = 500$ & 1.357813 & 1.5667226 & 1.447490 & 0.7230416 & 1.0226014 & 3.5724900 & 1.9986845 & 15.7532083\\
		\hline
	\end{tabular}
\end{footnotesize}
\end{table}
\begin{table}[H]

	\caption{\label{tab:}$\rr$ Estimates and Std. Errs, SIRIPR Model,
		$X_0 = 99950, Y_0 = 50$, $\beta = 0.06, \gamma = 0.03$}
	\centering
	\begin{footnotesize}
	\begin{tabular}[t]{l|r|r|r|r|r|r|r|r}
		\hline
		Data Set & Norm Est & Norm SE & Norm-M Est & Norm-M SE & AR Est & AR SE & AR-M Est & AR-M SE\\
		\hline
		$\sigma_X = 50, \sigma_Y = 2$ & 143.885460 & 778.881724 & 2.7003338 & 35.642540 & 55.265406 & 163.492483 & 2.1210664 & 13.550223\\
		\hline
		$\sigma_X = 500, \sigma_Y = 20$ & 48.894100 & 209.241051 & 1.3602975 & 10.234817 & 117.222483 & 870.624089 & 1.7544551 & 11.060645\\
		\hline
		$\sigma_X = 2500, \sigma_Y = 100$ & 1.656451 & 4.879121 & 0.8705396 & 1.231344 & 1.318824 & 3.274114 & 0.8679029 & 1.309243\\
		\hline
		$\sigma_X = 10000, \sigma_Y = 500$ & 10.426754 & 42.824909 & 1.1338742 & 4.906623 & 7.750396 & 24.637815 & 1.4241606 & 12.589987\\
		\hline
	\end{tabular}
\end{footnotesize}
\end{table}
\begin{table}[H]

	\caption{\label{tab:}$\rr$ Estimates and Std. Errs, SIRSIPR Model,
		$X_0 = 99950, Y_0 = 50$, $\beta = 0.06, \gamma = 0.03$}
	\centering
	\begin{footnotesize}
	\begin{tabular}[t]{l|r|r|r|r|r|r|r|r}
		\hline
		Data Set & Norm Est & Norm SE & Norm-M Est & Norm-M SE & AR Est & AR SE & AR-M Est & AR-M SE\\
		\hline
		$\sigma_X = 50, \sigma_Y = 2$ & -1.5215678 & 0 & 0.4256483 & 0 & 0.5346892 & 0 & -0.6623638 & 0\\
		\hline
		$\sigma\_X = 500, \sigma_Y = 20$ & 0.6363657 & 0 & 0.7407530 & 0 & 1.0216527 & 0 & 1.5576959 & 0\\
		\hline
		$\sigma\_X = 2500, \sigma_Y = 100$ & 1.1020944 & 0 & 1.0666890 & 0 & 1.0562382 & 0 & 1.1053737 & 0\\
		\hline
		$\sigma\_X = 10000, \sigma_Y = 500$ & 1.9158393 & 0 & 0.7926880 & 0 & 0.8105835 & 0 & 0.8006750 & 0\\
		\hline
	\end{tabular}
\end{footnotesize}
\end{table}
\begin{table}[H]

	\caption{\label{tab:}$\rr$ Estimates and Std. Errs, LOGLINEAR Model,
		$X_0 = 99950, Y_0 = 50$, $\beta = 0.06, \gamma = 0.03$}
	\centering
	\begin{footnotesize}
	\begin{tabular}[t]{l|r|r|r|r|r|r|r|r}
		\hline
		Data Set & Norm Est & Norm SE & Norm-M Est & Norm-M SE & AR Est & AR SE & AR-M Est & AR-M SE\\
		\hline
		$\sigma_X = 50, \sigma_Y = 2$ & 2.059934 & 0.0246402 & 1.995133 & 0.0038894 & 2.209575 & 0.0439100 & 2.545819 & 0.0568537\\
		\hline
		$\sigma\_X = 500, \sigma_Y = 20$ & 2.001882 & 0.0054085 & 1.995742 & 0.0015419 & 2.038646 & 0.0081562 & 2.015675 & 0.0031322\\
		\hline
		$\sigma\_X = 2500, \sigma_Y = 100$ & 1.999891 & 0.0001037 & 2.000036 & 0.0000930 & 1.999143 & 0.0001394 & 1.999749 & 0.0001328\\
		\hline
		$\sigma\_X = 10000, \sigma_Y = 500$ & 2.003402 & 0.0010320 & 1.998765 & 0.0003682 & 2.003763 & 0.0015505 & 1.999829 & 0.0008545\\
		\hline
	\end{tabular}
\end{footnotesize}
\end{table}
\begin{table}[H]

	\caption{\label{tab:}$\rr$ Estimates and Std. Errs, SIRMC Model,
		$X_0 = 99950, Y_0 = 50$, $\beta = 0.06, \gamma = 0.03$}
	\centering
	\begin{footnotesize}
	\begin{tabular}[t]{l|r|r|r|r|r|r|r|r}
		\hline
		Data Set & Norm Est & Norm SE & Norm-M Est & Norm-M SE & AR Est & AR SE & AR-M Est & AR-M SE\\
		\hline
		$\sigma_X = 50, \sigma_Y = 2$ & 0.9485506 & 3.8e-06 & 0.9485556 & 2e-07 & 0.9485428 & 2.6e-06 & 0.9485538 & 8e-07\\
		\hline
		$\sigma\_X = 500, \sigma_Y = 20$ & 0.9485562 & 1.0e-06 & 0.9485554 & 1e-07 & 0.9485640 & 3.6e-06 & 0.9485554 & 1e-07\\
		\hline
		$\sigma\_X = 2500, \sigma_Y = 100$ & 0.9485555 & 0.0e+00 & 0.9485555 & 0e+00 & 0.9485555 & 0.0e+00 & 0.9485555 & 0e+00\\
		\hline
		$\sigma\_X = 10000, \sigma_Y = 500$ & 0.9485554 & 2.0e-07 & 0.9485555 & 0e+00 & 0.9485554 & 1.0e-07 & 0.9485555 & 1e-07\\
		\hline
	\end{tabular}
\end{footnotesize}
\end{table}
\begin{table}[H]

	\caption{\label{tab:}$\rr$ Estimates and Std. Errs, SEQBAYES Model,
		$X_0 = 99950, Y_0 = 50$, $\beta = 0.06, \gamma = 0.03$}
	\centering
	\begin{footnotesize}
	\begin{tabular}[t]{l|r|r|r|r|r|r|r|r}
		\hline
		Data Set & Norm Est & Norm SE & Norm-M Est & Norm-M SE & AR Est & AR SE & AR-M Est & AR-M SE\\
		\hline
		$\sigma_X = 50, \sigma_Y = 2$ & 1.127679 & 0.0555836 & 1.000000 & 0.0523424 & 1.540929 & 0.0649748 & 1.000000 & 0.0523424\\
		\hline
		$\sigma\_X = 500, \sigma_Y = 20$ & 1.519223 & 0.0645155 & 1.000000 & 0.0523424 & 1.549862 & 0.0651629 & 1.000000 & 0.0523424\\
		\hline
		$\sigma\_X = 2500, \sigma_Y = 100$ & 1.409094 & 0.0621332 & 1.141402 & 0.0559207 & 1.386553 & 0.0616342 & 1.336668 & 0.0605153\\
		\hline
		$\sigma\_X = 10000, \sigma_Y = 500$ & 1.000000 & 0.0523424 & 1.415487 & 0.0622740 & 1.000000 & 0.0523424 & 1.000000 & 0.0523424\\
		\hline
	\end{tabular}
\end{footnotesize}
\end{table}

\subsection{Temporary Hold: Start Change}

\begin{table}[H]

	\caption{\label{tab:}$\rr$ Estimates and Std. Errs, SIRLS Model,
		$\sigma_X = 100, \sigma_Y = 5$, $\beta = 0.06, \gamma = 0.03$}
	\centering
	\begin{footnotesize}
	\begin{tabular}[t]{l|r|r|r|r|r|r|r|r}
		\hline
		Data Set & Norm Est & Norm SE & Norm-M Est & Norm-M SE & AR Est & AR SE & AR-M Est & AR-M SE\\
		\hline
		$X_0 = 99000, Y_0 = 1000$ & 1.999588 & 0.0049609 & 2.000284 & 0.0046294 & 2.000461 & 0.0078398 & 2.000122 & 0.0068626\\
		\hline
		$X_0 = 99900, Y_0 = 100$ & 2.022501 & 0.3064263 & 2.021865 & 0.3064148 & 2.022296 & 0.3063477 & 2.022665 & 0.3068512\\
		\hline
		$X_0 = 99990, Y_0 = 10$ & 2.116865 & 0.6626380 & 2.116551 & 0.6621184 & 2.116959 & 0.6623142 & 2.116803 & 0.6620448\\
		\hline
		$X_0 = 99999, Y_0 = 1$ & 2.468304 & 1.5370731 & 2.466776 & 1.5348981 & 2.470607 & 1.5371890 & 2.466700 & 1.5348542\\
		\hline
	\end{tabular}
\end{footnotesize}
\end{table}
\begin{table}[H]

	\caption{\label{tab:}$\rr$ Estimates and Std. Errs, SIRREPAR Model,
		$\sigma_X = 100, \sigma_Y = 5$, $\beta = 0.06, \gamma = 0.03$}
	\centering
	\begin{footnotesize}
	\begin{tabular}[t]{l|r|r|r|r|r|r|r|r}
		\hline
		Data Set & Norm Est & Norm SE & Norm-M Est & Norm-M SE & AR Est & AR SE & AR-M Est & AR-M SE\\
		\hline
		$X_0 = 99000, Y_0 = 1000$ & 1.999712 & 0.0034303 & 2.000274 & 0.0032086 & 2.000082 & 0.0053846 & 2.000014 & 0.0047509\\
		\hline
		$X_0 = 99900, Y_0 = 100$ & 2.022359 & 0.2698034 & 2.021866 & 0.2694871 & 2.022178 & 0.2696002 & 2.022530 & 0.2701860\\
		\hline
		$X_0 = 99990, Y_0 = 10$ & 2.117059 & 0.7291460 & 2.116858 & 0.7290133 & 2.116829 & 0.7288841 & 2.116534 & 0.7286543\\
		\hline
		$X_0 = 99999, Y_0 = 1$  & 2.469189 & 2.7293667 & 2.466967 & 2.7219457 & 2.470707 & 2.7245397 & 2.466483 & 2.7349248\\
		\hline
	\end{tabular}
\end{footnotesize}
\end{table}
\begin{table}[H]

	\caption{\label{tab:}$\rr$ Estimates and Std. Errs, SIRLMA Model,
		$\sigma_X = 100, \sigma_Y = 5$, $\beta = 0.06, \gamma = 0.03$}
	\centering
	\begin{footnotesize}
	\begin{tabular}[t]{l|r|r|r|r|r|r|r|r}
		\hline
		Data Set & Norm Est & Norm SE & Norm-M Est & Norm-M SE & AR Est & AR SE & AR-M Est & AR-M SE\\
		\hline
		$X_0 = 99000, Y_0 = 1000$ & -87.5945650 & 66413.128046 & -0.9999761 & 1.422640 & 4.1817846 & 3.8812899 & -10.8458320 & 288.1028935\\
		\hline
		$X_0 = 99900, Y_0 = 100$ & 0.5237886 & 1.632971 & 0.2436375 & 3.754091 & 0.7149415 & 0.9374275 & -0.3051038 & 8.8837193\\
		\hline
		$X_0 = 99990, Y_0 = 10$ & -33.0051124 & 433.594227 & -1.3382741 & 3.047220 & -0.2463189 & 1.5875310 & 0.4368556 & 0.7686276\\
		\hline
		$X_0 = 99999, Y_0 = 1$ & 10.9282204 & 19.297359 & -98.6422337 & 3113.332617 & -0.8194023 & 1.7121668 & 16.6102769 & 18.1045393\\
		\hline
	\end{tabular}
\end{footnotesize}
\end{table}
\begin{table}[H]

	\caption{\label{tab:}$\rr$ Estimates and Std. Errs, SIRLMAT Model,
		$\sigma_X = 100, \sigma_Y = 5$, $\beta = 0.06, \gamma = 0.03$}
	\centering
	\begin{footnotesize}
	\begin{tabular}[t]{l|r|r|r|r|r|r|r|r}
		\hline
		Data Set & Norm Est & Norm SE & Norm-M Est & Norm-M SE & AR Est & AR SE & AR-M Est & AR-M SE\\
		\hline
		$X_0 = 99000, Y_0 = 1000$ & 0.730610 & 4.6810416 & 0.9720592 & 0.6586069 & 3.163816 & 45.776770 & 0.9367865 & 0.8857030\\
		\hline
		$X_0 = 99900, Y_0 = 100$ & 1.236415 & 0.7388933 & 1.1744675 & 1.3957615 & 1.249219 & 0.813166 & 1.1960662 & 0.6589163\\
		\hline
		$X_0 = 99990, Y_0 = 10$ & 1.459703 & 10.9219088 & 0.8531016 & 12.5573016 & -3.846910 & 71.297410 & -0.5884883 & 39.4915091\\
		\hline
		$X_0 = 99999, Y_0 = 1$ & 1.624645 & 5.0359845 & 0.4174208 & 12.8380228 & 1.301346 & 25.850056 & 3.1746775 & 12.8142151\\
		\hline
	\end{tabular}
\end{footnotesize}
\end{table}
\begin{table}[H]

	\caption{\label{tab:}$\rr$ Estimates and Std. Errs, SIRIPR Model,
		$\sigma_X = 100, \sigma_Y = 5$, $\beta = 0.06, \gamma = 0.03$}
	\centering
	\begin{footnotesize}
	\begin{tabular}[t]{l|r|r|r|r|r|r|r|r}
		\hline
		Data Set & Norm Est & Norm SE & Norm-M Est & Norm-M SE & AR Est & AR SE & AR-M Est & AR-M SE\\
		\hline
		$X_0 = 99000, Y_0 = 1000$ & 0.7710423 & 0.8645882 & 0.5919617 & 0.6342256 & 0.7161564 & 0.7174921 & 0.5976212 & 0.5992238\\
		\hline
		$X_0 = 99900, Y_0 = 100$ & 1.7659585 & 4.6419825 & 0.8824341 & 2.0203472 & 1.3381463 & 2.9156688 & 0.8541963 & 1.6873018\\
		\hline
		$X_0 = 99990, Y_0 = 10$ & 11.1978083 & 40.2941067 & 1.1073477 & 3.2407038 & 10.0438581 & 43.1896138 & 0.9850750 & 2.6487228\\
		\hline
		$X_0 = 99999, Y_0 = 1$ & 464.5647956 & 4494.6030482 & 2.1636551 & 15.9245640 & 137.5104852 & 719.4933920 & 1.5049398 & 7.7183465\\
		\hline
	\end{tabular}
\end{footnotesize}
\end{table}
\begin{table}[H]

	\caption{\label{tab:}$\rr$ Estimates and Std. Errs, SIRSIPR Model,
		$\sigma_X = 100, \sigma_Y = 5$, $\beta = 0.06, \gamma = 0.03$}
	\centering
	\begin{footnotesize}
	\begin{tabular}[t]{l|r|r|r|r|r|r|r|r}
		\hline
		Data Set & Norm Est & Norm SE & Norm-M Est & Norm-M SE & AR Est & AR SE & AR-M Est & AR-M SE\\
		\hline
		$X_0 = 99000, Y_0 = 1000$ & 1.2522106 & 34.7537893 & 1.2099648 & 34.6892205 & 1.3712547 & 34.4629789 & 1.1533154 & 34.2274520\\
		\hline
		$X_0 = 99900, Y_0 = 100$ & -0.5284892 & 0.0000000 & -0.6902358 & 0.0000000 & -3.9381962 & 0.0000000 & -1.0663894 & 0.0000000\\
		\hline
		$X_0 = 99990, Y_0 = 10$ & 0.8071191 & 0.1942094 & 0.8129547 & 0.1895219 & 0.8287186 & 0.1855934 & 0.8070414 & 0.1954849\\
		\hline
		$X_0 = 99999, Y_0 = 1$ & 1.4634454 & 0.0000000 & 3.2469490 & 0.0000000 & 0.7228135 & 0.0000000 & 0.6213745 & 0.0000000\\
		\hline
	\end{tabular}
\end{footnotesize}
\end{table}
\begin{table}[H]

	\caption{\label{tab:}$\rr$ Estimates and Std. Errs, LOGLINEAR Model,
		$\sigma_X = 100, \sigma_Y = 5$, $\beta = 0.06, \gamma = 0.03$}
	\centering
	\begin{footnotesize}
	\begin{tabular}[t]{l|r|r|r|r|r|r|r|r}
		\hline
		Data Set & Norm Est & Norm SE & Norm-M Est & Norm-M SE & AR Est & AR SE & AR-M Est & AR-M SE\\
		\hline
		$X_0 = 99000, Y_0 = 1000$ & 1.999515 & 0.0002828 & 2.000144 & 0.0001594 & 1.999753 & 0.0004197 & 2.000586 & 0.0003024\\
		\hline
		$X_0 = 99900, Y_0 = 100$ & 2.000011 & 0.0002315 & 1.999732 & 0.0001526 & 2.000473 & 0.0002776 & 2.000640 & 0.0002656\\
		\hline
		$X_0 = 99990, Y_0 = 10$ & 1.999886 & 0.0002517 & 1.999599 & 0.0001361 & 2.000608 & 0.0003289 & 1.999049 & 0.0002664\\
		\hline
		$X_0 = 99999, Y_0 = 1$ & 1.999690 & 0.0007731 & 2.000503 & 0.0003700 & 2.002716 & 0.0010572 & 1.999919 & 0.0005663\\
		\hline
	\end{tabular}
\end{footnotesize}
\end{table}
\begin{table}[H]

	\caption{\label{tab:}$\rr$ Estimates and Std. Errs, SIRMC Model,
		$\sigma_X = 100, \sigma_Y = 5$, $\beta = 0.06, \gamma = 0.03$}
	\centering
	\begin{footnotesize}
	\begin{tabular}[t]{l|r|r|r|r|r|r|r|r}
		\hline
		Data Set & Norm Est & Norm SE & Norm-M Est & Norm-M SE & AR Est & AR SE & AR-M Est & AR-M SE\\
		\hline
		$X_0 = 99000, Y_0 = 1000$ & 0.9485555 & 0.00e+00 & 0.9485555 & 0 & 0.9485555 & 0.0e+00 & 0.9485555 & 0\\
		\hline
		$X_0 = 99900, Y_0 = 100$ & 0.9485555 & 0.00e+00 & 0.9485555 & 0 & 0.9485555 & 0.0e+00 & 0.9485555 & 0\\
		\hline
		$X_0 = 99990, Y_0 = 10$ & 0.9485552 & 2.00e-07 & 0.9485555 & 0 & 0.9485558 & 2.0e-07 & 0.9485555 & 0\\
		\hline
		$X_0 = 99999, Y_0 = 1$ & 0.9486030 & 1.56e-05 & 0.9485557 & 0 & 0.9485701 & 2.6e-06 & 0.9485555 & 0\\
		\hline
	\end{tabular}
\end{footnotesize}
\end{table}
\begin{table}[H]

	\caption{\label{tab:}$\rr$ Estimates and Std. Errs, SEQBAYES Model,
		$\sigma_X = 100, \sigma_Y = 5$, $\beta = 0.06, \gamma = 0.03$}
	\centering
	\begin{footnotesize}
	\begin{tabular}[t]{l|r|r|r|r|r|r|r|r}
		\hline
		Data Set & Norm Est & Norm SE & Norm-M Est & Norm-M SE & AR Est & AR SE & AR-M Est & AR-M SE\\
		\hline
		$X_0 = 99000, Y_0 = 1000$ & 0.9728534 & 0.0516270 & 1.052579 & 0.0537008 & 0.5850802 & 0.0400370 & 0.6236582 & 0.0413358\\
		\hline
		$X_0 = 99900, Y_0 = 100$ & 1.0000000 & 0.0523424 & 1.000000 & 0.0523424 & 1.4175193 & 0.0623186 & 1.4017333 & 0.0619707\\
		\hline
		$X_0 = 99990, Y_0 = 10$ & 1.3877095 & 0.0616599 & 1.430888 & 0.0626118 & 1.3948200 & 0.0618177 & 1.4507203 & 0.0630442\\
		\hline
		$X_0 = 99999, Y_0 = 1$ & 1.4714813 & 0.0634938 & 1.465728 & 0.0633695 & 1.4665442 & 0.0633871 & 1.4723265 & 0.0635120\\
		\hline
	\end{tabular}
\end{footnotesize}
\end{table}

\subsection{Temporary Hold: Other Models}

\begin{table}[H]

	\caption{\label{tab:}$\rr$ Estimates and Std. Errs, SIRLS Model,
		$X_0 = 99950, Y_0 = 50$, $\sigma_X = 100, \sigma_Y = 5$}
	\centering
	\begin{footnotesize}
	\begin{tabular}[t]{l|r|r|r|r|r|r|r|r}
		\hline
		Data Set & Norm Est & Norm SE & Norm-M Est & Norm-M SE & AR Est & AR SE & AR-M Est & AR-M SE\\
		\hline
		1st Order & 1.051438 & 0.3782910 & 1.051420 & 0.3784628 & 1.051864 & 0.3775363 & 1.051293 & 0.3782570\\
		\hline
		4th Order & 1.769347 & 0.4226615 & 1.769136 & 0.4215768 & 1.769002 & 0.4227615 & 1.769136 & 0.4215910\\
		\hline
		Linear SIR & 1.274258 & 0.0979741 & 1.273917 & 0.0976920 & 1.273654 & 0.1000387 & 1.274007 & 0.0970809\\
		\hline
	\end{tabular}
\end{footnotesize}
\end{table}
\begin{table}[H]

	\caption{\label{tab:}$\rr$ Estimates and Std. Errs, SIRREPAR Model,
		$X_0 = 99950, Y_0 = 50$, $\sigma_X = 100, \sigma_Y = 5$}
	\centering
	\begin{footnotesize}
	\begin{tabular}[t]{l|r|r|r|r|r|r|r|r}
		\hline
		Data Set & Norm Est & Norm SE & Norm-M Est & Norm-M SE & AR Est & AR SE & AR-M Est & AR-M SE\\
		\hline
		1st Order & 1.051435 & 0.0259416 & 1.051397 & 0.0258690 & 1.051845 & 0.0260146 & 1.051255 & 0.0257166\\
		\hline
		4th Order & 1.769331 & 0.2796903 & 1.769079 & 0.2791191 & 1.769095 & 0.2799270 & 1.769111 & 0.2788789\\
		\hline
		Linear SIR & 1.274314 & 0.0339775 & 1.274075 & 0.0337499 & 1.274174 & 0.0345629 & 1.273799 & 0.0334192\\
		\hline
	\end{tabular}
\end{footnotesize}
\end{table}
\begin{table}[H]

	\caption{\label{tab:}$\rr$ Estimates and Std. Errs, SIRLMA Model,
		$X_0 = 99950, Y_0 = 50$, $\sigma_X = 100, \sigma_Y = 5$}
	\centering
	\begin{footnotesize}
	\begin{tabular}[t]{l|r|r|r|r|r|r|r|r}
		\hline
		Data Set & Norm Est & Norm SE & Norm-M Est & Norm-M SE & AR Est & AR SE & AR-M Est & AR-M SE\\
		\hline
		1st Order & 1.2559522 & 0.1296269 & 1.2209276 & 0.0597736 & 1.2670255 & 0.1588846 & 1.2476445 & 0.1150267\\
		\hline
		4th Order & 0.3632202 & 0.1230895 & -0.0201429 & 0.1133200 & 0.3580085 & 0.1258904 & -0.0161666 & 0.1103940\\
		\hline
		Linear SIR & 1.2191443 & 0.1093267 & 1.2067608 & 0.0480986 & 1.2146478 & 0.0627181 & 0.9503153 & 0.1066437\\
		\hline
	\end{tabular}
\end{footnotesize}
\end{table}
\begin{table}[H]

	\caption{\label{tab:}$\rr$ Estimates and Std. Errs, SIRLMAT Model,
		$X_0 = 99950, Y_0 = 50$, $\sigma_X = 100, \sigma_Y = 5$}
	\centering
	\begin{footnotesize}
	\begin{tabular}[t]{l|r|r|r|r|r|r|r|r}
		\hline
		Data Set & Norm Est & Norm SE & Norm-M Est & Norm-M SE & AR Est & AR SE & AR-M Est & AR-M SE\\
		\hline
		1st Order & 1.1509447 & 0.1012249 & 1.138621 & 0.1738250 & 1.147184 & 0.0681524 & 1.145189 & 0.0811484\\
		\hline
		4th Order & 0.8246203 & 6.0710436 & 1.000597 & 2.5841000 & 3.255953 & 26.3713818 & 1.385674 & 8.5430506\\
		\hline
		Linear SIR & 1.2009801 & 0.0031100 & 1.200130 & 0.0040763 & 1.203159 & 0.0148085 & 1.195875 & 0.0246927\\
		\hline
	\end{tabular}
\end{footnotesize}
\end{table}
\begin{table}[H]

	\caption{\label{tab:}$\rr$ Estimates and Std. Errs, SIRIPR Model,
		$X_0 = 99950, Y_0 = 50$, $\sigma_X = 100, \sigma_Y = 5$}
	\centering
	\begin{footnotesize}
	\begin{tabular}[t]{l|r|r|r|r|r|r|r|r}
		\hline
		Data Set & Norm Est & Norm SE & Norm-M Est & Norm-M SE & AR Est & AR SE & AR-M Est & AR-M SE\\
		\hline
		1st Order & 2.699973 & 9.131131 & 2.297208 & 7.14196 & 4.334767 & 41.325146 & 4.194341 & 36.201142\\
		\hline
		4th Order & 6.260963 & 20.337163 & 5.924029 & 17.81136 & 6.268500 & 20.536192 & 5.892582 & 17.341763\\
		\hline
		Linear SIR & 1.516803 & 1.554241 & 1.208578 & 1.12544 & 1.379871 & 1.311406 & 1.208521 & 1.003065\\
		\hline
	\end{tabular}
\end{footnotesize}
\end{table}
\begin{table}[H]

	\caption{\label{tab:}$\rr$ Estimates and Std. Errs, SIRSIPR Model,
		$X_0 = 99950, Y_0 = 50$, $\sigma_X = 100, \sigma_Y = 5$}
	\centering
	\begin{footnotesize}
	\begin{tabular}[t]{l|r|r|r|r|r|r|r|r}
		\hline
		Data Set & Norm Est & Norm SE & Norm-M Est & Norm-M SE & AR Est & AR SE & AR-M Est & AR-M SE\\
		\hline
		1st Order & 58.836635 & 0.0000000 & 27.429685 & 0.0000000 & 1.498805 & 0.0000000 & -14.135196 & 0.0000000\\
		\hline
		4th Order & 2.848624 & 0.0000000 & 15.083621 & 0.0000000 & 2.766152 & 0.0000000 & 14.972580 & 0.0000000\\
		\hline
		Linear SIR & 1.294079 & 0.2151131 & 1.294364 & 0.2158095 & 1.296731 & 0.2215603 & 1.304489 & 0.2299876\\
		\hline
	\end{tabular}
\end{footnotesize}
\end{table}
\begin{table}[H]

	\caption{\label{tab:}$\rr$ Estimates and Std. Errs, LOGLINEAR Model,
		$X_0 = 99950, Y_0 = 50$, $\sigma_X = 100, \sigma_Y = 5$}
	\centering
	\begin{footnotesize}
	\begin{tabular}[t]{l|r|r|r|r|r|r|r|r}
		\hline
		Data Set & Norm Est & Norm SE & Norm-M Est & Norm-M SE & AR Est & AR SE & AR-M Est & AR-M SE\\
		\hline
		1st Order & 1.188050 & 0.0006341 & 1.187976 & 0.0006389 & 1.187553 & 0.0006273 & 1.187848 & 0.0006497\\
		\hline
		4th Order & 1.955180 & 0.0164608 & 1.955313 & 0.0164212 & 1.955130 & 0.0164623 & 1.955302 & 0.0164165\\
		\hline
		Linear SIR & 1.258587 & 0.0010079 & 1.258670 & 0.0010069 & 1.258920 & 0.0009938 & 1.258462 & 0.0010199\\
		\hline
	\end{tabular}
\end{footnotesize}
\end{table}
\begin{table}[H]

	\caption{\label{tab:}$\rr$ Estimates and Std. Errs, SIRMC Model,
		$X_0 = 99950, Y_0 = 50$, $\sigma_X = 100, \sigma_Y = 5$}
	\centering
	\begin{footnotesize}
	\begin{tabular}[t]{l|r|r|r|r|r|r|r|r}
		\hline
		Data Set & Norm Est & Norm SE & Norm-M Est & Norm-M SE & AR Est & AR SE & AR-M Est & AR-M SE\\
		\hline
		1st Order & 0.9485551 & 1e-07 & 0.9485540 & 4e-07 & 0.9485555 & 2e-07 & 0.9485553 & 2.0e-07\\
		\hline
		4th Order & 0.9485545 & 6e-07 & 0.9485533 & 1e-06 & 0.9485535 & 9e-07 & 0.9485527 & 1.2e-06\\
		\hline
		Linear SIR & 0.9485555 & 0e+00 & 0.9485555 & 0e+00 & 0.9485555 & 0e+00 & 0.9485555 & 0.0e+00\\
		\hline
	\end{tabular}
\end{footnotesize}
\end{table}
\begin{table}[H]

	\caption{\label{tab:}$\rr$ Estimates and Std. Errs, SEQBAYES Model,
		$X_0 = 99950, Y_0 = 50$, $\sigma_X = 100, \sigma_Y = 5$}
	\centering
	\begin{footnotesize}
	\begin{tabular}[t]{l|r|r|r|r|r|r|r|r}
		\hline
		Data Set & Norm Est & Norm SE & Norm-M Est & Norm-M SE & AR Est & AR SE & AR-M Est & AR-M SE\\
		\hline
		1st Order & 1.2515351 & 0.0585565 & 1.221050 & 0.0578389 & 1.301757 & 0.0597198 & 1.2848494 & 0.0593307\\
		\hline
		4th Order & 0.5223345 & 0.0378293 & 0.550362 & 0.0388309 & 0.523122 & 0.0378578 & 0.5506554 & 0.0388413\\
		\hline
		Linear SIR & 1.1217433 & 0.0554371 & 1.194507 & 0.0572068 & 1.123600 & 0.0554829 & 1.1568274 & 0.0562973\\
		\hline
	\end{tabular}
	\end{footnotesize}
\end{table}

\end{document}

%%%%%%%%%%%%%%%%%%%%%%
\Textbf{Title:}

\textbf{Author:}

\textbf{Citation:}

\textbf{Major themes:}

\textbf{Notes:}
\\
%%%%%%%%%%%%%%%%%%%%%%

\begin{figure}[h]
\begin{center}
\includegraphics[width=4in]{mvhw7_3c.pdf}
\end{center}
\caption{
Biplots of the different continuous variables and their correlations.}\label{fig3c}
\end{figure}

%% two pictures whoa

\begin{figure}[h]
\centering

\begin{figure}[h]
\centering

\begin{subfigure}{.5\textwidth}
  \centering
  \includegraphics[width=1\linewidth]{mt_eda_cont_hists.pdf}
  \caption{Histograms of Arrival Delay and continuous covariates.  Arrival delay seems to have a right skewed distribution.  This may indicate that we will be transforming this variable later on.  After transforming Air Time and Distance by a log transformation, we don't really seem to have many outliers in our covariates.  We seem to have outliers in the CRS Dep. Time and Arrival Time; however, time is cyclical and so these are not, in fact outliers.}
  \label{hists}
\end{subfigure}%
\begin{subfigure}{.5\textwidth}
  \centering
  \includegraphics[width=1\linewidth]{mt_eda_cont_hists.pdf}
  \caption{\textcolor{red}{Placeholder}}
  \label{tabs1}
\end{subfigure}
\caption{}
\end{figure}

\begin{table}
\begin{tabular}{l | a | b | a | b}
\hline
\rowcolor{LightCyan}
\mc{1}{}  & \mc{1}{x} & \mc{1}{y} & \mc{1}{w} & \mc{1}{z} \    \hline
variable 1 & a & b & c & d \    variable 2 & a & b & c & d \\ \hline
\end{tabular}
\end{table}

%%
\begin{figure}
\centering
\begin{subfigure}{.5\textwidth}
  \centering
  \includegraphics[width=1\linewidth]{resids_full.pdf}
  \caption{}
  \label{residsf}
\end{subfigure}%
\begin{subfigure}{.5\textwidth}
  \centering
  \includegraphics[width=1\linewidth]{diags_full.pdf}
  \caption{ }
  \label{diagsf}
\end{subfigure}
\caption{}
\end{figure}

### 2018-07-03 paper/Version12/draft_v12 (Shannon Gallagher's conflicted copy 2018-05-01).tex
finally on git

--- 2018-07-03 paper/Version11/draft_v11.tex
@@
 \usepackage{amsmath}
+\usepackage{amsfonts}
 \usepackage{graphicx}% http://ctan.org/pkg/graphicx
@@
 \newcommand{\xxsir}{\ensuremath{9}} % number of SIR methods
+\newcommand{\wxxsir}{nine }
+\newcommand{\Wxxsir}{Nine }
 
@@
 
-\title{An overview of $\xxsir$ methods to estimate $\rr$ from the SIR model and comments on more complex models}
-\author{ Department of Statistics, Carnegie Mellon University}
+\title{\Wxxsir methods to estimate $\rr$ in the SIR model}
+\author{ Shannon Gallagher$^{\dag}$, Andersen Chang$^{\ddag}$, and William F.
+Eddy$^{\dag}$ \\$\dag$ Department of Statistics and Data Science, Carnegie Mellon University\\ $\ddag$ Department of Statistics, Rice University}
 \date{\today}
@@
 As defined by \citet{anderson1992}, $\rr$ is the ``the average number of secondary infections produced when one infected individual is introduced into a host population where everyone is susceptible.''  $\rr$, in some ways, summarizes an entire outbreak of a disease; it  is used to assess whether a disease epidemic will occur.
-Additionally, it describes what percentage of the population needs to be vaccinated to avoid such an epidemic.
+Additionally, it describes what percentage of the population needs to be vaccinated to avoid such an epidemic, roughly $1-\rr^{-1}$ \citep{anderson1992}).
 Despite a clear definition of $\rr$, epidemiologists have struggled to create a standard  estimator for $\rr$  \citep{hethcote2000}.
@@
 time), boundary cases of infection and recovery rates, population size, initial SI ratio, and restrictive assumptions placed upon the noise.
-These are all issues for estimating single points, to say nothing of confidence or credible intervals (CI).
-
-We review $\xxsir$ methods for estimating $\rr$.
+These are all issues for producing point estimates, to say nothing of confidence or credible intervals (CI).
+
+We review \wxxsir methods for estimating $\rr$.
 We note that these methods are not exhaustive but are chosen due to their impact and use in epidemiology.
@@
 As such, we discuss techniques for estimating CIs for each of the different methods.
-These techniques include, the delta method, block bootstrap, and the posterior distribution.
-
-Following, we present a series of simulations in which we examine each of our $\rr$ estimation methods.
+These CI estimates are the delta method, block bootstrap, and the posterior distribution.
+
+In the following sections, we present a series of simulations in which we examine each of our $\rr$ estimation methods.
 Within these simulations, we vary the number of observations used to estimate $\rr$, infection and recovery rates, the total population size, initial SI ratio, and several assumptions about the noise.
@@
 
-Finally, we discuss our recommendations in estimating $\rr$ for the SIR model.
-Afterwards, we comment on how these recommendations extend to more complex models and how many of the problems with estimating $\rr$  become exacerbated with the increase in complexity.
+Finally, we discuss our recommendations in estimating $\rr$ for the SIR model, and we comment on how these recommendations extend to more complex models and how many of the problems with estimating $\rr$  become exacerbated with the increase in complexity.
 
@@
 In Section \ref{sec:sir-intro}, we introduce the SIR model as described by \cite{Kermack700}.
-In Section \ref{sec:methods}, we overview the $\xxsir$ methods for estimating $\rr$.
-In Section \ref{sec:ci}, we discuss how to estimate CIs for each of the $\xxsir$ methods.
+In Section \ref{sec:methods}, we overview the \wxxsir methods for estimating $\rr$.
+In Section \ref{sec:ci}, we discuss how to estimate CIs for each of the \wxxsir methods.
 Following in Section \ref{sec:sim-res}, we present the results of our simulations.
@@
 
-\section{Origins of $\rr$}
+\section{Origins and Difficulties of Estimating $\rr$}
 \label{sec:r0}
 
-The origins of $\rr$ are tied to the survival function.
-This function, which originates from the field of demography in the late 1800s, describes how many female offspring a woman is expected to produce in her lifetime \citep{dietz1993estimation}.
+The origins of $\rr$ are tied to the survival function, which originates from the field of demography in the late 1800s.
+The survival function describes how many female offspring a woman is expected to produce in her lifetime \citep{dietz1993estimation}.
 In demography, we have
@@
 This is the chronologically first, and in many ways, the most direct method to estimate $\rr$.
-As we see here, the SIR model is not even mentioned in this derivation.
 The premise of this whole paper is based upon the fact that estimating $\rr$ from the survival function in Equation \eqref{eq:surv} is a difficult task.
-It is due to this difficulty that other estimates even exist and why we must discuss $\rr$ as a property of the SIR model.
+It is due to this difficulty that other estimates even exist and why we must treat $\rr$ as a property of the SIR model.
+
+Other difficulties in estimating $\rr$ arise from model assumptions on how randomness enters the model and sensitivity of the estimate.
+For example, the SIR model is unique in that two of its compartments, S and R are monotonic.
+The number of susceptibles is non-increasing and the number of susceptibles is non-decreasing.
+Models should enforce this monotonicity of the two compartments.
+As a consequence, this restricts assumptions on how randomness enters the model.
+For example, Gaussian error, a common assumption in statistics and machine learning, becomes highly suspect due to the non-symmetrical nature of the noise imposed by the monotonic $S$ and $R$ compartments.
+Without assuming Gaussian noise, it becomes much more difficult to guarantee properties such as unbiased and consist estimates of $\rr$.
+
+These difficulties in estimating $\rr$ have led a number of researchers to analyze the sensitivity of $\rr$ \citep{lash2003,epstein2007agent,capaldi2012}.
+Even this is a difficult task because we are looking at the rate of change of the compartments with respect to $\beta$ and $\gamma$.
+Since there is no closed-form solution of the SIR model, estimating the set of derivatives with respect to $\beta$ and $\gamma$ must be done numerically, which can substantially increase computation time.
+We should note that this sensitivity analysis is done before even adding any distributional assumptions on how randomness enters the model.
+
+% A final difficulty in estimating $\rr$ arises from the data.
+Ideally, we would have counts of the number of susceptibles, infectious, and recovered at each time step.
+However, sometimes only the incidence of the disease is available, where we actually need the prevalence.
+Then, the number of susceptibles is back-estimated from the number of infectious which requires an estimate of the number of susceptibles, and even the US Census is known to be at least 10\% off in both directions of undercounting and overcounting (CITE).
+Finally, the distinction between susceptible and infectious and infectious and recovered may not be clear.
+As asymptomatic individuals may infect others and be mistaken to be in either the susceptible or infectious compartment.
+
+As a result of these difficulties, accurately estimating $\rr$ is a monumentous task.
+We overview \wxxsir methods introduced by a number of researchers used to solve some of the problems described above.
 
@@
 
-The SIR model introduced by \cite{Kermack700} is a compartment model, where individuals move from susceptible, to infectious, and finally recovered states.  We make four essential assumptions,
+The SIR model introduced by \cite{Kermack700} is a compartment model, where individuals move from susceptible, to infectious, and finally recovered states.  We make five essential assumptions,
 \begin{enumerate}
@@
 \item The transition of objects into and out of compartments is described by a set of known equations, possibly dependent on unknown parameters.
-\item The populations mix homogeneously
-  \item The number of objects in each compartment at time $t=0$ is known
+\item The populations mix homogeneously.
+\item The number of objects in each compartment at time $t=0$ is known.
+  \item The law of mass action is followed.
   \end{enumerate}
 
-Another common assumption for compartment models, is the law of mass action, a property borrowed from chemistry which says that the mass of the product is proportional to the mass of the reactants.
-In epidemiology, this means that the proportion of new infections is proportional to the current  number of susceptible \citep{anderson1992}.
+The law of mass action is a property borrowed from chemistry which says that the mass of the product per unit time is proportional to the mass of the reactants \citep{lotka1920}.
+In epidemiology, this means that the proportion of new infections per unit time is proportional to the current  number of susceptible \citep{anderson1992}.
 Therefore, the flow of objects from one compartment to another may be dependent on the percentage of objects within one or many of the compartments.
@@
 Both $\beta$ and $\gamma$ are constrained to be between $0$ and $1$.
-Throughout, we use $X$, $Y$, and $Z$ for stand-ins for the ODEs for $S$, $I$, and $R$, respectively, to avoid any confusion with $\rr$.
+For the remainder of this paper, we use $X$, $Y$, and $Z$ for stand-ins for the ODEs for $S$, $I$, and $R$, respectively, to avoid any confusion with $\rr$.
 \begin{align}
-\systeme{\frac{dX}{dt} = -\frac{\beta XY}{N}, \frac{dY}{dt} = \frac{\beta XY}{N} - \gamma Y, \frac{dZ}{dt} = \gamma Z}. \label{eq:sir}
+\systeme{\frac{dX}{dt} = -\frac{\beta XY}{N}, \frac{dY}{dt} = \frac{\beta XY}{N} - \gamma Y, \frac{dZ}{dt} = \gamma Y}. \label{eq:sir}
 \end{align}
@@
 \end{figure}
-An outbreak occurs if the rate of change of infectious individuals is positive,
+An outbreak occurs if the rate of change of infectious individuals is positive, $\frac{dY}{dt} > 0$, or equivalently,
 \begin{align*}
   \frac{dY}{dt} &> 0 \\
-  \frac{\beta X Y}{N}  - \gamma Y &> 0 ,\\
-  Y \left ( \beta \frac{X}{N} - \gamma \right ) & > 0\\
-   \frac{\beta}{\gamma} &> \frac{N}{X}.
-\end{align*}
-That is,  the rate of new infections is greater than the rate of recovery.  So as long as the number of susceptibles is large, $\frac{X}{N} \approx 1$, then an outbreak will occur if $\rr >1$,
+  \implies \frac{\beta X Y}{N}  - \gamma Y &> 0 ,\\
+\implies  Y \left ( \beta \frac{X}{N} - \gamma \right ) & > 0\\
+\implies   \frac{\beta}{\gamma} &> \frac{N}{X}.
+\end{align*}
+That is,  the rate of new infections is greater than the rate of recovery.
+So as long as the number of susceptibles is large compared to the total population, $\frac{X}{N} \approx 1$, then an outbreak will occur if $\rr >1$,
 \begin{align*}
@@
   \end{align*}
-  In order to incorporate randomness, we add noise, namely,
+  In order to incorporate randomness into the model, we add noise, namely,
   \begin{align}\label{eq:sir-noise}
@@
     Y_{obs}(t) &=  N - X_{obs}(t) -Y_{obs}(t)  \nonumber\\
-    Z_{obs}(t) &= Z(t) + \epsilon_{Z,t}. \nonumber
+    Z_{obs}(t) &= Z(t) - \epsilon_{Z,t}. \nonumber
   \end{align}
-We are assuming the observations are generated based on the deterministic ODEs presented in Equation \ref{eq:sir} with the addition of time ($t=0, 1, \dots, T$0) and compartment dependent noise $\epsilon_{X,t}$ and $\epsilon_{Z,t}$.
+We are assuming the observations are generated based on the deterministic ODEs presented in Equation \ref{eq:sir} with the addition of time ($t=0, 1, \dots, T$) and compartment dependent noise $\epsilon_{X,t}$ and $\epsilon_{Z,t}$.
 Since, $N$, the total population, is constant, then $Y_{obs}$ is adjusted accordingly.
@@
 \begin{align*}
-\textnormal{data} = \left \{\left (X_{obs}(t), Y_{obs}(t), Z_{obs}(t) \right ) : t=0, 1, \dots, T\right \}.
-  \end{align*}.
-
-\section{Overview of $\xxsir$ methods to estimate $\rr$ for the SIR model}
+  \textnormal{Data} &= \left \{\left (X_{obs}(t), Y_{obs}(t), Z_{obs}(t) \right ) : t=0, 1, \dots, T\right \}, \\
+  \hat{\rr} &= m(Data),
+\end{align*}
+where $m$ is a function of the data.
+
+Good properties of an estimator may include being (1) being a consistent estimator
+\begin{align*}
+  \hat{\rr} \overset{P}{\to} 0 \textnormal{ as } T\to \infty
+\end{align*}
+or (2) converging to a known probability distribution,
+\begin{align*}
+\hat{\rr} \leadsto F.  
+\end{align*}
+As we are interested in the hypothesis
+\begin{align*}
+  H_0:\;& \rr > 1 \\
+  H_A:\;& \rr \le 1
+\end{align*}
+Then for some $\alpha \in [0,1]$, we would like to know for some $a, b >0$
+\begin{align*}
+P(a \le \hat{\rr} \le b) = 1 - \alpha,  
+\end{align*}
+the second property is particularly desirable.  At any rate, we would like to be able to estimate $a$ and $b$ for a given $\alpha$-level.
+
+\section{Overview of \wxxsir methods to estimate $\rr$ for the SIR model}
 \label{sec:methods}
 
-In this section, we describe the $\xxsir$ methods in detail.
+In this section, we describe the \wxxsir methods in detail, including advantages and disadvantages of each.
 
 \subsection{Least Squares ($\beta$, $\gamma$) (LS)}\label{least-squares-beta-gamma}
-The first approach to estimate $\rr$ in the SIR model is to minimize the joint mean square error for the data collected at each time point and use the plug-in estimator found in Equation \ref{eq:sirls}.
+The first approach to estimate $\rr$ in the SIR model is to minimize the joint sum of squares as seen in approaches similar to \cite{malik2008} and \cite{canto2017}.
+In this method, the least squares estimate of $\rr$ is formed from the data collected at each time point to estimate $\hat{\beta}$ and $\hat{\gamma}$ and uses the  subsequent plug-in estimator found in Equation \ref{eq:sirls} to estimate $\rr$.
 In particular, we find
@@
 \end{align}
+
+This method is sometimes used to estimate $\hat{\beta}$ and $\hat{\gamma}$ by using a grid search but can be done with more sophisticated minimization algorithms.
+Due to the ODE format of the equations, we can only use minimization algorithms which do not rely on an explicit gradient such as Nelder-Mead simplex minimization \citep{nelder-mead1965}.
+
+A reasonable question one may ask is why use the $L_2$-norm as opposed to $L_1$ or some other similarity score.
+A possible answer is that if we were to assume Gaussian noise, then the $L_2$-norm would be equivalent to the maximum likelihood estimation.
+Another answer is that $L_1$ is a continuously differentiable function and hence is easier to compute, which allows for sensitivity analysis to more easily be conducted.
+Finally, LS can be done without writing down explicit assumptions on the noise within the model.
+However, we cannot guarantee properties of consistency or converging to a known distribution without more explicit assumptions.
+
+On the other hand, least squares is easy to implement in software such as base \texttt{R}'s \texttt{optim()} function, and as we see in Section \ref{sec:results}, produces comparable methods to other methods.
 
@@
 \begin{align*}
-(\hat{\rr}, \hat{\gamma} ) &= \text{argmin}_{\beta, \gamma} \sum_{t} \left [ \left (X_{obs}(t) - X(t)\right )^2 + \left ( Z_{obs}(t) - Z(t) \right )^2 \right ]
-\end{align*}
-We use the $\hat{\rr}$ directly from the above estimation process, which again can be done with gradient descent or another optimization process.
+(\hat{\rr}, \hat{\gamma} ) &= \text{argmin}_{\rr, \gamma} \sum_{t} \left [ \left (X_{obs}(t) - X(t)\right )^2 + \left ( Z_{obs}(t) - Z(t) \right )^2 \right ]
+\end{align*}
+We use the $\hat{\rr}$ directly from the above estimation process, which again can be done with a grid search or another optimization process.
+It may be surprising that this method leads to different results than in simply using least squares.
+One reason we obtain different results may be attributed to better numerical precision as there is one fewer step to obtain an estimate for $\rr$.
+However, we also see that this method results in smaller standard errors than in LS, which we discuss more in Section \ref{sec:results}.
+
+Like LS, while we have no explicit parametric assumptions on the noise, we also are uncertain of any theoretical guarantees of our resulting estimator.
+Likewise, the same difficulties in sensitivity analysis arise as in LS.
 
@@
 The SIR ODEs in Eq.
-\ref{eq:sir} have no known closed form solution, and so we are already using approximations using numerical integration, albeit small ones.
+\ref{eq:sir} have no known closed form solution, and so we are already using approximations using numerical integration.
 In addition to this, data collected from real diseases are typically very noisy to begin with.
@@
         Z^\prime} \cdot \frac{N}{X} \nonumber\\
-  \hat{\rr} &= -\frac{\hat{X}^\prime(0)}{ \hat{Z}^\prime(t)} \cdot \frac{N}{\hat{X}(0)}. \nonumber
+  \hat{\rr} &= -\frac{\hat{X}^\prime(0)}{ \hat{Z}^\prime(0)} \cdot \frac{N}{\hat{X}(0)}. \nonumber
   \end{align}
-Here, $K=10$ is rather arbitrary and should be selected using some criterion such as AIC in addition with the knowledge of how many data points are available to the user.
+  Here, $K=10$ is arbitrary and should be selected using some criterion such as AIC in addition with the knowledge of how many data points are available to the user.
 Besides optionally deciding on the degree of polynomials to fit, this model is simple to implement and gives very comparable results to using least squares with the SIR model.
@@
 
+  An advantage of using this estimation method is that it is simple to implement using any linear modelling software.
+On the other hand, we are assuming $X(t) = \hat{X}(t)$ and so if the data truly follows a SIR model, then we will have biased estimators for $\rr$.
+Using this method, we are able to check how sensitive $\rr$ is to the degree of the fitted polynomial.
+
 \subsection{Linear Model Approximation, All Time Points (LMAT)}\label{linear-model-approximation-all-time-points-degree-10}
 
-The above formulation (Section \ref{linear-model-approximation-degree-10} of a linear model approximation only uses the estimate at time $t=0$ to estimate $\rr$.
+The above formulation (Section \ref{linear-model-approximation-degree-10}) of a linear model approximation only uses the estimate at time $t=0$ to estimate $\rr$.
 We can instead, use all time points available to estimate $\rr$.
@@
 \end{align*}
-The intuition is that $\frac{-X^\prime}{Z^\prime(t)}$ is constant in $t$, but due to our approximations with the linear model, this is no longer the case.
+The intuition is that $\frac{-X^\prime(t)}{Z^\prime(t)}$ is constant in $t$, but due to our approximations with the linear model, this is no longer the case.
 Here, we average over the different possible values of $\rr$, estimated at different times.
 An advantage to this approach is that we have a more robust estimate of $\rr$ than just using one time point.
+However, the estimate will still be biased.
+Like in LMAT, we can examine how sensitive $\rr$ is to the order of the polynomial $K$.
 
 \subsection{Incidence to Prevalence Ratio (IPR)}\label{incidence-to-prevalence-ratio}
-The incidence to prevalence ratio (IPR), described by \cite{Nishiura2009} is another intuitive method to calculate $\rr$ as it incorporates some of the most basic epidemiological quantities, incidence and prevalence.
+The incidence to prevalence ratio (IPR), described by \cite{Nishiura2009}, is another intuitive method to calculate $\rr$ as it incorporates some of the most basic epidemiological quantities, incidence and prevalence.
 
@@
 Here we assume that the time step is small enough to approximate the incidence.
-The advantage of this method is that incidence data is generally readily available as is prevalence data for certain diseases such as HIV.
+The advantage of this method is that incidence data is generally available as is prevalence data for certain diseases such as HIV \citep{hall2008}.
 However, as one is required to have prior knowledge about $\gamma$, it may be easier to directly estimate $\rr$ with one of the many other methods described that does not require a prior knowledge about $\gamma$.
 Again, we are using only one time point to estimate $\rr$.
+This model is agnostic to assumptions on the noise, but again, it is unclear if this is an unbiased or consistent estimator of $\rr$.
 
@@
 It has the same disadvantages as the regular IPR ratio in that it requires knowledge about $\gamma$.
+Also, ratio estimators usually behave badly when the denominator is small.
 
@@
 \begin{align}
-  X(t) &=  X(0) e^{\frac{\beta}{N \gamma}Z(t)} \nonumber\\
-  \log \frac{X(t)}{X(0)} &=  \frac{\beta }{\gamma N }Z(t) \nonumber\\
-  N\log \frac{X(t)}{X(0)} &=  \rr Z(t). \label{eq:harko_lin}
+  X(t) &=  X(0) e^{\frac{-\beta}{\gamma}\frac{Z(t)}{N}} \nonumber\\
+  \log \frac{X(t)}{X(0)} &=  \frac{-\beta }{\gamma}\frac{Z(t)}{N} \nonumber\\
+  \log \frac{X(t)}{X(0)} &=  -\rr \frac{Z(t)}{N}. \label{eq:harko_lin}
 \end{align}
@@
 Thus, we can regress the left hand side in Eq.
-\ref{eq:harko_lin} on $Z(t)$, the number of recovered individuals, with $\rr$ as the coefficient of $Z(t)$ and no intercept term to obtain an estimate of $\hat{\rr}$.
+\ref{eq:harko_lin} on $Z(t)$, the number of recovered individuals, with $\rr$ as the coefficient of $Z(t)/N$ and no intercept term to obtain an estimate of $\hat{\rr}$,
+\begin{align*}
+  \hat{\rr} = -\frac{\sum_{t=0}^T \log \frac{ X(t)}{X(0)}}{\sum_{t=0}^T\frac{Z(t)}{N}}.
+\end{align*}
+This method says that for a one percent increase in number of recovered individuals, we expect the ratio of number of old susceptibles to number of new susceptibles to increase by $e^{\rr}$,
+\begin{align*}
+  \log \left ( X(t+1)/ X(0) \right ) - \log \left ( X(t)/X(0) \right ) &= - .01\rr\\
+  \log \left ( X(t+1) \right ) - \log \left ( X(t) \right )  &=- .01\rr\\
+  \log \left ( X(t) / X(t+1) \right ) &= .01\rr\\
+  \frac{X(t)}{X(t+1)}  &\approx e^{\rr}.
+\end{align*}
+
+The advantages of this method are many.
+One, the estimate of $\rr$ is highly interpretable.
+Two, an assumption of Gaussian noise in Equation \ref{eq:harko_lin} is not as egregious as in the other methods to estimate $\rr$ because we are looking at a log transformation of the the percent of suceptibles instead of untransformed variables.
+As a consequence, we may be able to make a case where we the resulting estimate of $\rr$ is unbiased and a consistent estimator.
+Finally,  this method is easily implemented using any linear modelling software.
 
@@
 
-In this framework, the number of susceptibles at the next step, $X_{t}$, has a Binomial distribution based on the contacts with the current number of infectious, $Y_{t}$ and the current number of susceptibles.
-That is $X_{t+1} \sim \text{Binom}(X_t, \alpha^{Y_t}$), where $\alpha$ is the probability of avoiding infection from an infective.
-\cite{barbour2004} report that the reproduction number is thus,
+In this method, the number of susceptibles at the next step, $X(t+1)$, has a Binomial distribution based on the contacts with the current number of infectious, $Y(t)$ and the current number of susceptibles.
+That is $X(t+1) \sim \text{Binomial}\left(X(t), \alpha^{Y(t)}\right)$, where $\alpha$ is the probability of avoiding infection from an infective.
+One can write out the likelihood for $\alpha$ in this model, namely,
+\begin{align*}
+\mathcal{L}\left ( \alpha ; \textnormal{data}\right ) \propto \prod_{t=0}^{T-1}\left ( \alpha^{Y(t)} \right )^{X(t+1)} \left (1- \alpha^{Y(t)} \right )^{X(t) - X(t+1)}.
+\end{align*}
+Maximizing the likelihood yields an estimate for $\alpha$,
+\begin{align*}
+\hat{\alpha} = \arg \max_{\alpha} \mathcal{L}\left(\alpha; \textnormal{data} \right ).
+  \end{align*}
+ \cite{barbour2004} report that the estimated reproduction number is thus,
 \begin{align}\label{eq:r0-mc}
-\rr &= \log \left ( \frac{1}{1-\alpha}\right ).
+\hat{\rr} &= \log \left ( \frac{1}{1-\hat{\alpha}}\right ).
 \end{align}
 
-Thus, using regression on susceptible/infection counts will lead to an estimate of $\hat{\alpha}$ and hence $\hat{\rr}$.
-However, this framework typically allows more than just the reproduction number to be estimated.
+This method typically allows for more than just the reproduction number to be estimated.
 Through recursion, one can calculate the probability of having a given number of susceptibles and infected at each time step, and hence the entire probability distribution may be known.
+
+The advantages of this model are its simplicity of interpretation and ability to generate a whole probability distrubtion for $(X(t), Y(t))$.
+We have explicitly stated how randomness enters the model so it is up to the researcher to check whether this assumption is a good fit.
+However, it is possible to simulate how sensitive $\hat{\rr}$ is to $\hat{\alpha}$.
 
@@
 \begin{align*}
-J(t+1)  \sim Poisson( J(t) \exp \left \{  \gamma (\rr-1)\right \})
+J(t+1)  \sim \textnormal{Poisson}( J(t) \exp \left \{  \gamma (\rr-1)\right \})
 \end{align*}
@@
 \begin{align*}
-  P(\rr | J_0, \dots, J_{t+1}) = \frac{P(J_{t+1} | \rr, J_0, \dots, J_t)P(\rr| J_0, \dots, J_t)}{P(J_0, \dots, J_{t+1})}.
+  P(\rr | J(0), \dots, J(t+1)) = \frac{P(J(t+1) | \rr, J(0), \dots, J(t))P(\rr| J(0), \dots, J(t))}{P(J(0), \dots, J(t+1))}.
 \end{align*}
@@
 
-\subsection{MLE}
-
 \section{Confidence and Credible Intervals}
@@
 
-When the method estimates \(\beta\) and \(\gamma\) instead of \(\rr\) directly, we use the delta method approximation to calculate the
+When a method estimates \(\beta\) and \(\gamma\) instead of \(\rr\) directly, we use the delta method approximation to calculate the
 variance of \(\rr\).
@@
 Here, we use the relationship of $\rr = \frac{\beta}{\gamma}$ in order to estimate the variance, which is specific to the SIR framework.
-This method may be extended to other frameworks.
-The advantages of the delta method are that as the NGM posits a recipe to derive $\rr$, then it is theoretically possible to use the delta method to derive an expression for the variance of $\rr$ for any compartment model.
+A disadvantage of this method is that it provides no insight on how to estimate $\Sigma_{\beta,\gamma}$.
 
@@
 
-The block bootstrap is a variant of the bootstrap (cite) in which the $n$ observations are dependent on one another.
+The block bootstrap is a variant of the bootstrap in which the $n$ observations are dependent on one another.
 In contrast to the original bootstrap, the block bootstrap partitions consectutive observations into $k$ blocks with block-length $b$ ($n=kb$).
-These are non-overlapping partitions or blocks, although there are variants for which blocks can overlap (cite).
+These are non-overlapping partitions or blocks, although there are variants for which blocks can overlap.
 Following, for iteration $\ell$, one samples uniformly replacement $k$ blocks and estimates $\hat{\rr}_b$ from those $k$ blocks, repeating the sampling and estimation for a total of $L$ estimates of $\hat{\rr}$.
@@
 
-The assympotic properties hold when the stochastic process is $m$-dependent, that is when observations $t, t+1, \dots$ and $s, s+1, \dots$ are independent whenver $s+m < t$ and under some smoothness conditions of the statistic used to estimate $\rr$.
+The assympotic properties hold when the stochastic process is $m$-dependent, that is when observations $t, t+1, \dots$ and $s, s+1, \dots$ are independent whenever $s+m < t$ and under some smoothness conditions of the statistic used to estimate $\rr$.
 
@@
 When one uses a Bayesian approach to estimate $\rr$, the result should be a distribution rather than an estimate.
-One can simply look at the quantities of this distribution to form a credible interval for this estimate.
+One can simply look at the quantiles of this distribution to form a credible interval for this estimate.
 Advantages of this method include having an entire distribution compared to a point estimate.
@@
 
-\section{Simulation Results}\label{sec:sim-res}
-
-\section{Discussion}\label{discussion}
+\section{Methods}\label{sec:sim-res}
+
+We compare the performance of the SIR models on simulated data from both the SIR as well as other models.
+In all of these simulations, we first generate data from the models under known conditions, then we add on errors with known variances to each time point.
+We simulate with two different types of error distributions, Gaussian and autoregressive.
+In the Gaussian case, we add on independently and identically distributed errors to the $X$ and $Y$ compartments with variances $\sigma_X$ and $\sigma_Y$, respectively; the sum of these errors is then subtracted from the $Z$ compartment to maintain constant population.
+The same process was followed for the autoregressive case, except that we generate errors from an AR(1) model with a correlation coefficient $\rho = 0.7$ added to Gaussian white noise.
+For each error distribution, we simulate an additional data set which adds a monotonicity adjustment to the X and Z compartments after adding on the  using the Pool Adjacent Violators Algorithm (PAVA).
+The sum of the populations of these compartments is then subtracted from the $Y$ compartment, again to maintain a constant population.
+This gives us a total of four error structures for each set of conditions: Gaussian (Norm), Gaussian monotone (Norm-M), autoregressive (AR), and autoregressive monotone (AR-M).
+
+\subsection{SIR Data}
+
+We first simulate 5 ``baseline'' data sets from the SIR model under the following conditions:
+
+\begin{center}
+
+	$\beta$ = 0.06, $\gamma$ = 0.03; $\rr$ = 2
+
+	Number of time steps: 365
+
+	Starting X size ($X_0$): 99950
+
+	Starting Y size ($Y_0$): 50
+
+	Starting Z size: 0
+
+	$\sigma_X$: 100
+
+	$\sigma_Y$: 5
+
+	Total population: 100000.
+
+\end{center}
+
+We then generate different SIR data sets, changing the one of the conditions for simulation each time and using the baseline SIR data as our basis for comparison.
+The different data sets generate are shown in the following table:
+
+\begin{table}[h]
+	\centering
+	\resizebox{\textwidth}{!}{
+		\begin{tabular}{@{}lll@{}}
+			\toprule
+			\textbf{Parameter Values ($\beta, \gamma$)}        & \textbf{Starting Compartment Sizes ($X_0, Y_0$)}            & \textbf{Variance of Added Errors ($\sigma_X, \sigma_Y$ )}       \\ \midrule
+			(0.06, 0.001)            & (99000, 1000)                           & (50, 2)                     \\
+			(0.06, 0.04)             & (99900, 100)               & (500, 20)     \\
+			(0.06, 0.06)             & (99990, 10)             & (2500, 100)           \\
+			(0.06, 0.24) & (99999, 1) & (10000, 500)                                      \\ \bottomrule
+		\end{tabular}
+	}
+	\caption{Table of data sets.}
+	\label{tab:datasets}
+\end{table}
+
+\subsection{Other Models}
+
+To test how well the SIR models perform under misspecification, we simulate data from other models.
+While we will not have a ``known'' $\rr$ value to compare our results to, this will allow us to see if the models give us reasonable results when the data do not come from the SIR model.
+We simulate data from linear and quartic models with respect to the time step, using a monotonically increasing function for the $X$ compartment and a monotonically decreasing function for the $Z$ compartment.
+The sum of these is subtracted from the constant total population to get the $Y$ compartment.
+We also generated data from a linearized version of the three compartment SIR model with the following ODEs:
+
+\begin{eqnarray*}
+	\frac{dX}{dt} &=& -\beta Y \\
+	\frac{dY}{dt} &=& \beta Y - \gamma Y \\
+	\frac{dZ}{dt} &=& \gamma Y.
+\end{eqnarray*}
+
+The data from the linear SIR model were simulated under the following conditions:
+
+\begin{center}
+
+	$\beta$ = 0.06, $\gamma$ = 0.05; $\rr$ = 1.2
+
+	Number of time steps: 365
+
+	Starting X size ($X_0$): 99950
+
+	Starting Y size ($Y_0$): 50
+
+	Starting Z size: 0
+
+	$\sigma_X$: 10
+
+	$\sigma_Y$: 1
+
+	Total population: 100000.
+
+\end{center}
+
+\section{Results}\label{sec:results}
+There are numerous aspects to consider when using a method to estimate $\rr$.
+Among these include values of both the infection and recovery rate $(\beta, \gamma)$, the number of data points available (time $T$), the total size of the population ($N$), the initial percent of susceptible individuals $\left (\frac{X(0)}{N}\right)$, the initial percent of infectious individuals $\left (\frac{Y(0)}{N}\right )$, and  the magnitude of the variance of both the number of susceptibles and the number of infectious ($\sigma_X, \sigma_Y$).
+We examine each of these aspects individually.
+
+\subsection{Baseline}\label{sec:res-base}
+
+\subsection{Infection and Recovery Rate ($\beta, \gamma$)}\label{sec:res-beta-gamma}
+
+\subsection{Number of Time Steps ($T$)}\label{sec:res-time}
+
+\subsection{Initial Percentage of Infectious $\left (\frac{Y(0)}{N}\right)$}\label{sec:res-inf}
+
+\subsection{Magnitude of Variance $(\sigma^2_X, \sigma^2_Y)$}\label{sec:res-var}
+
+\subsection{Other Models}\label{sec:res-oth}
+
+\section{Discussion}\label{sec:discussion}
 
@@
 
-\section{Appendix}
-
-\subsection{Temporary Hold: Baseline Data}
+\appendix
+
+\section{Beyond the SIR Model}
+
+\section{Fitting the SIR Model}
+
+\subsection{SIR Least Squares}
 
@@
 \end{table}
+
+\begin{table}[H]
+
+	\caption{\label{tab:}$\rr$ Estimates and Std. Errs, SIRLS Model,
+		$X_0 = 99950, Y_0 = 50$, $\sigma_X = 100, \sigma_Y = 5$, $\beta = 0.06$}
+	\centering
+	\begin{footnotesize}
+		\begin{tabular}[t]{l|r|r|r|r|r|r|r|r}
+			\hline
+			Data Set & Norm Est & Norm SE & Norm-M Est & Norm-M SE & AR Est & AR SE & AR-M Est & AR-M SE\\
+			\hline
+			$\gamma$ = 0.001 & 59.9764682 & 0.0434705 & 59.9764682 & 0.0386733 & 60.0151171 & 0.0697888 & 59.9589126 & 0.0631448\\
+			\hline
+			$\gamma$ = 0.04 & 1.5000763 & 0.0126817 & 1.5001978 & 0.0090880 & 1.4989700 & 0.0182852 & 1.4993603 & 0.0165168\\
+			\hline
+			$\gamma$ = 0.06 & 1.0043054 & 0.4140424 & 0.9955554 & 0.1086448 & 0.9915431 & 0.6272644 & 0.9941392 & 0.1891812\\
+			\hline
+			$\gamma$ = 0.24 & 0.2363274 & 4.1042813 & 0.2327691 & 0.6701185 & -0.0695410 & 2.8291362 & 0.3021083 & 0.5657139\\
+			\hline
+		\end{tabular}
+	\end{footnotesize}
+\end{table}
+
+\begin{table}[H]
+
+	\caption{\label{tab:}$\rr$ Estimates and Std. Errs, SIRLS Model,
+		$\sigma_X = 100, \sigma_Y = 5$, $\beta = 0.06, \gamma = 0.03$}
+	\centering
+	\begin{footnotesize}
+		\begin{tabular}[t]{l|r|r|r|r|r|r|r|r}
+			\hline
+			Data Set & Norm Est & Norm SE & Norm-M Est & Norm-M SE & AR Est & AR SE & AR-M Est & AR-M SE\\
+			\hline
+			$X_0 = 99000, Y_0 = 1000$ & 1.999588 & 0.0049609 & 2.000284 & 0.0046294 & 2.000461 & 0.0078398 & 2.000122 & 0.0068626\\
+			\hline
+			$X_0 = 99900, Y_0 = 100$ & 2.022501 & 0.3064263 & 2.021865 & 0.3064148 & 2.022296 & 0.3063477 & 2.022665 & 0.3068512\\
+			\hline
+			$X_0 = 99990, Y_0 = 10$ & 2.116865 & 0.6626380 & 2.116551 & 0.6621184 & 2.116959 & 0.6623142 & 2.116803 & 0.6620448\\
+			\hline
+			$X_0 = 99999, Y_0 = 1$ & 2.468304 & 1.5370731 & 2.466776 & 1.5348981 & 2.470607 & 1.5371890 & 2.466700 & 1.5348542\\
+			\hline
+		\end{tabular}
+	\end{footnotesize}
+\end{table}
+
+\begin{table}[H]
+	\caption{\label{tab:}$\rr$ Estimates and Std. Errs, SIRLS Model,
+		$X_0 = 99950, Y_0 = 50$, $\beta = 0.06, \gamma = 0.03$}
+	\centering
+	\begin{footnotesize}
+		\begin{tabular}[t]{l|r|r|r|r|r|r|r|r}
+			\hline
+			Data Set & Norm Est & Norm SE & Norm-M Est & Norm-M SE & AR Est & AR SE & AR-M Est & AR-M SE\\
+			\hline
+			$\sigma_X = 50, \sigma_Y = 2$ & 1.993455 & 0.5784530 & 1.979966 & 0.1204126 & 2.033934 & 0.7559710 & 2.066706 & 0.3382834\\
+			\hline
+			$\sigma_X = 500, \sigma_Y = 20$ & 1.998088 & 0.1353803 & 1.996950 & 0.0541332 & 1.999686 & 0.1893888 & 2.002135 & 0.0927952\\
+			\hline
+			$\sigma_X = 2500, \sigma_Y = 100$ & 1.999755 & 0.0026502 & 1.999834 & 0.0026397 & 1.999555 & 0.0037664 & 1.999038 & 0.0036473\\
+			\hline
+			$\sigma_X = 10000, \sigma_Y = 500$ & 2.001870 & 0.0259428 & 1.997748 & 0.0162536 & 2.002792 & 0.0416203 & 1.997918 & 0.0257997\\
+			\hline
+		\end{tabular}
+	\end{footnotesize}
+\end{table}
+
+\begin{table}[H]
+
+	\caption{\label{tab:}$\rr$ Estimates and Std. Errs, SIRLS Model,
+		$X_0 = 99950, Y_0 = 50$, $\sigma_X = 100, \sigma_Y = 5$}
+	\centering
+	\begin{footnotesize}
+		\begin{tabular}[t]{l|r|r|r|r|r|r|r|r}
+			\hline
+			Data Set & Norm Est & Norm SE & Norm-M Est & Norm-M SE & AR Est & AR SE & AR-M Est & AR-M SE\\
+			\hline
+			1st Order & 1.051438 & 0.3782910 & 1.051420 & 0.3784628 & 1.051864 & 0.3775363 & 1.051293 & 0.3782570\\
+			\hline
+			4th Order & 1.769347 & 0.4226615 & 1.769136 & 0.4215768 & 1.769002 & 0.4227615 & 1.769136 & 0.4215910\\
+			\hline
+			Linear SIR & 1.274258 & 0.0979741 & 1.273917 & 0.0976920 & 1.273654 & 0.1000387 & 1.274007 & 0.0970809\\
+			\hline
+		\end{tabular}
+	\end{footnotesize}
+\end{table}
+
+\begin{table}[H]
+
+	\caption{\label{tab:}$\rr$ Estimates and Std. Errs, SIRLS Model,
+		$\beta = 0.06, \gamma = 0.03$, $X_0 = 99950, Y_0 = 50$, $\sigma_X = 100, \sigma_Y = 5$}
+	\centering
+	\begin{tabular}[t]{l|r|r|r|r|r|r|r|r}
+		\hline
+		Data Set & Norm Est & Norm SE & Norm-M Est & Norm-M SE & AR Est & AR SE & AR-M Est & AR-M SE\\
+		\hline
+		T = 200 & 1.999339 & 0.0286060 & 2.000583 & 0.0204020 & 2.0006557 & 0.0357152 & 2.000722 & 0.0364659\\
+		\hline
+		T = 100 & 1.990754 & 0.2922125 & 1.993375 & 0.1463550 & 1.9098406 & 0.3811467 & 1.967388 & 0.2324481\\
+		\hline
+		T = 50 & 1.978167 & 1.1596532 & 1.856227 & 0.3825275 & 2.2720601 & 1.2900031 & 1.699976 & 0.6064311\\
+		\hline
+		T = 20 & 38.689552 & 106.2705193 & 5.036600 & 1.3672212 & 0.0098756 & 2.1687945 & 1.681588 & 0.4747083\\
+		\hline
+	\end{tabular}
+\end{table}
+
+\subsection{SIR Reparameterization}
+
 \begin{table}[H]
@@
 \end{table}
+
+\begin{table}[H]
+
+	\caption{\label{tab:}$\rr$ Estimates and Std. Errs, SIRREPAR Model,
+		$X_0 = 99950, Y_0 = 50$, $\sigma_X = 100, \sigma_Y = 5$, $\beta = 0.06$}
+	\centering
+	\begin{footnotesize}
+		\begin{tabular}[t]{l|r|r|r|r|r|r|r|r}
+			\hline
+			Data Set & Norm Est & Norm SE & Norm-M Est & Norm-M SE & AR Est & AR SE & AR-M Est & AR-M SE\\
+			\hline
+			$\gamma$ = 0.001 & 59.995188 & 0.4343984 & 59.9951876 & 0.3874879 & 60.025452 & 0.6911933 & 59.9960910 & 0.6214841\\
+			\hline
+			$\gamma$ = 0.04 & 1.499883 & 0.0062529 & 1.5001705 & 0.0044936 & 1.499068 & 0.0090313 & 1.4993177 & 0.0081651\\
+			\hline
+			$\gamma$ = 0.06 & 1.004306 & 0.0501444 & 0.9955446 & 0.0117215 & 0.991558 & 0.0645788 & 0.9941511 & 0.0190525\\
+			\hline
+			$\gamma$ = 0.24 & 4.179525 & 2.7455569 & 3.5709590 & 0.3171284 & 5.192951 & 3.5249847 & 9.8415034 & 2.5646198\\
+			\hline
+		\end{tabular}
+	\end{footnotesize}
+\end{table}
+
+\begin{table}[H]
+
+	\caption{\label{tab:}$\rr$ Estimates and Std. Errs, SIRREPAR Model,
+		$\sigma_X = 100, \sigma_Y = 5$, $\beta = 0.06, \gamma = 0.03$}
+	\centering
+	\begin{footnotesize}
+		\begin{tabular}[t]{l|r|r|r|r|r|r|r|r}
+			\hline
+			Data Set & Norm Est & Norm SE & Norm-M Est & Norm-M SE & AR Est & AR SE & AR-M Est & AR-M SE\\
+			\hline
+			$X_0 = 99000, Y_0 = 1000$ & 1.999712 & 0.0034303 & 2.000274 & 0.0032086 & 2.000082 & 0.0053846 & 2.000014 & 0.0047509\\
+			\hline
+			$X_0 = 99900, Y_0 = 100$ & 2.022359 & 0.2698034 & 2.021866 & 0.2694871 & 2.022178 & 0.2696002 & 2.022530 & 0.2701860\\
+			\hline
+			$X_0 = 99990, Y_0 = 10$ & 2.117059 & 0.7291460 & 2.116858 & 0.7290133 & 2.116829 & 0.7288841 & 2.116534 & 0.7286543\\
+			\hline
+			$X_0 = 99999, Y_0 = 1$  & 2.469189 & 2.7293667 & 2.466967 & 2.7219457 & 2.470707 & 2.7245397 & 2.466483 & 2.7349248\\
+			\hline
+		\end{tabular}
+	\end{footnotesize}
+\end{table}
+
+\begin{table}[H]
+
+	\caption{\label{tab:}$\rr$ Estimates and Std. Errs, SIRREPAR Model,
+		$X_0 = 99950, Y_0 = 50$, $\beta = 0.06, \gamma = 0.03$}
+	\centering
+	\begin{footnotesize}
+		\begin{tabular}[t]{l|r|r|r|r|r|r|r|r}
+			\hline
+			Data Set & Norm Est & Norm SE & Norm-M Est & Norm-M SE & AR Est & AR SE & AR-M Est & AR-M SE\\
+			\hline
+			$\sigma_X = 50, \sigma_Y = 2$ & 1.993660 & 0.5115266 & 1.979973 & 0.1053955 & 2.033636 & 0.7059701 & 2.066476 & 0.3172823\\
+			\hline
+			$\sigma_X = 500, \sigma_Y = 20$ & 1.998145 & 0.1207935 & 1.997147 & 0.0481545 & 2.000194 & 0.1680126 & 2.002244 & 0.0826643\\
+			\hline
+			$\sigma_X = 2500, \sigma_Y = 100$ & 1.999829 & 0.0023646 & 1.999946 & 0.0023591 & 1.999885 & 0.0033514 & 1.999483 & 0.0032200\\
+			\hline
+			$\sigma_X = 10000, \sigma_Y = 500$ & 2.002305 & 0.0232636 & 1.997996 & 0.0144744 & 2.003065 & 0.0372826 & 1.997732 & 0.0230069\\
+			\hline
+		\end{tabular}
+	\end{footnotesize}
+\end{table}
+
+\begin{table}[H]
+
+	\caption{\label{tab:}$\rr$ Estimates and Std. Errs, SIRREPAR Model,
+		$X_0 = 99950, Y_0 = 50$, $\sigma_X = 100, \sigma_Y = 5$}
+	\centering
+	\begin{footnotesize}
+		\begin{tabular}[t]{l|r|r|r|r|r|r|r|r}
+			\hline
+			Data Set & Norm Est & Norm SE & Norm-M Est & Norm-M SE & AR Est & AR SE & AR-M Est & AR-M SE\\
+			\hline
+			1st Order & 1.051435 & 0.0259416 & 1.051397 & 0.0258690 & 1.051845 & 0.0260146 & 1.051255 & 0.0257166\\
+			\hline
+			4th Order & 1.769331 & 0.2796903 & 1.769079 & 0.2791191 & 1.769095 & 0.2799270 & 1.769111 & 0.2788789\\
+			\hline
+			Linear SIR & 1.274314 & 0.0339775 & 1.274075 & 0.0337499 & 1.274174 & 0.0345629 & 1.273799 & 0.0334192\\
+			\hline
+		\end{tabular}
+	\end{footnotesize}
+\end{table}
+
+\begin{table}[H]
+
+	\caption{\label{tab:}$\rr$ Estimates and Std. Errs, SIRREPAR Model,
+		$\beta = 0.06, \gamma = 0.03$, $X_0 = 99950, Y_0 = 50$, $\sigma_X = 100, \sigma_Y = 5$}
+	\centering
+	\begin{tabular}[t]{l|r|r|r|r|r|r|r|r}
+		\hline
+		Data Set & Norm Est & Norm SE & Norm-M Est & Norm-M SE & AR Est & AR SE & AR-M Est & AR-M SE\\
+		\hline
+		T = 200 & 1.999428 & 0.0312814 & 2.000386 & 0.0223765 & 2.000419 & 0.0390441 & 2.000199 & 0.0402489\\
+		\hline
+		T = 100 & 1.991014 & 0.3940097 & 1.993501 & 0.1974773 & 1.910435 & 0.4773240 & 1.967858 & 0.3065049\\
+		\hline
+		T = 50 & 1.977508 & 1.8877725 & 1.856480 & 0.5810168 & 2.273215 & 2.3912219 & 1.700025 & 0.7995792\\
+		\hline
+		T = 20 & 22.964587 & NaN & 5.044869 & 2.1398099 & NA & NA & 1.681269 & 0.6543603\\
+		\hline
+	\end{tabular}
+\end{table}
+
+\subsection{SIR Linear Model Approximation}
+
 \begin{table}[H]
@@
 			\hline
-		Baseline1 & 1.999502 & 0.3896154 & 2.694876 & 0.5083823 & 35.700310 & 4820.9883994 & 1.669865 & 0.1717277\\
-		\hline
-		Baseline2 & 1.688318 & 0.1882819 & 1.681966 & 0.1102752 & 1.477492 & 0.1124589 & 1.537388 & 0.1495794\\
-		\hline
-		Baseline3 & 1.574728 & 0.1339888 & 1.746144 & 0.1668073 & 5.340898 & 33.0770287 & 2.314664 & 0.6038287\\
-		\hline
-		Baseline4 & 1.642471 & 0.1949104 & 2.047336 & 0.3952098 & 1.328799 & 0.0667234 & -3.007100 & 7.5904183\\
-		\hline
-		Baseline5 & 1.593838 & 0.1684652 & 1.843829 & 0.2199302 & 1.490924 & 0.1280881 & 1.871116 & 0.2218524\\
-		\hline
-	\end{tabular}
-\end{footnotesize}
-\end{table}
+			Baseline 1 & 1.997416 & 0.3900006 & 2.692152 & 0.5081138 & 35.656866 & 4815.2163162 & 1.667689 & 0.1718689\\
+			\hline
+			Baseline 2 & 1.686364 & 0.1886132 & 1.679359 & 0.1102722 & 1.474460 & 0.1130752 & 1.534979 & 0.1496990\\
+			\hline
+			Baseline 3 & 1.572708 & 0.1340966 & 1.743175 & 0.1667602 & 5.345243 & 33.1125597 & 2.313202 & 0.6045124\\
+			\hline
+			Baseline 4 & 1.640228 & 0.1956211 & 2.046183 & 0.3959176 & 1.327043 & 0.0671306 & -3.008847 & 7.5919771\\
+			\hline
+			Baseline 5 & 1.590702 & 0.1685914 & 1.842150 & 0.2204031 & 1.488547 & 0.1285866 & 1.868254 & 0.2217948\\
+			\hline
+		\end{tabular}
+	\end{footnotesize}
+\end{table}
+
+\begin{table}[H]
+
+	\caption{\label{tab:}$\rr$ Estimates and Std. Errs, SIRLMA Model,
+		$X_0 = 99950, Y_0 = 50$, $\sigma_X = 100, \sigma_Y = 5$, $\beta = 0.06$}
+	\centering
+	\begin{footnotesize}
+		\begin{tabular}[t]{l|r|r|r|r|r|r|r|r}
+			\hline
+			Data Set & Norm Est & Norm SE & Norm-M Est & Norm-M SE & AR Est & AR SE & AR-M Est & AR-M SE\\
+			\hline
+			$\gamma$ = 0.001 & -117.9694856 & 101.1885824 & -55.3748542 & 10.9872264 & -19.2057935 & 5.9751126 & -183.9423408 & 277.8291528\\
+			\hline
+			$\gamma$ = 0.04 & 1.1087585 & 0.0863592 & -3.9728056 & 1116.1887158 & 0.9757401 & 0.0213324 & 0.7333872 & 0.1787735\\
+			\hline
+			$\gamma$ = 0.06 & 1.0829034 & 0.1533121 & 0.9905684 & 0.1182606 & 0.9279273 & 0.0605799 & 1.0397917 & 0.0359875\\
+			\hline
+			$\gamma$ = 0.24 & 0.7074613 & 0.2028765 & 0.5216530 & 0.1715628 & 0.8749860 & 0.0590753 & -49.7813412 & 5303.0302298\\
+			\hline
+		\end{tabular}
+	\end{footnotesize}
+\end{table}
+
+\begin{table}[H]
+
+	\caption{\label{tab:}$\rr$ Estimates and Std. Errs, SIRLMA Model,
+		$\sigma_X = 100, \sigma_Y = 5$, $\beta = 0.06, \gamma = 0.03$}
+	\centering
+	\begin{footnotesize}
+		\begin{tabular}[t]{l|r|r|r|r|r|r|r|r}
+			\hline
+			Data Set & Norm Est & Norm SE & Norm-M Est & Norm-M SE & AR Est & AR SE & AR-M Est & AR-M SE\\
+			\hline
+			$X_0 = 99000, Y_0 = 1000$ & -87.5945650 & 66413.128046 & -0.9999761 & 1.422640 & 4.1817846 & 3.8812899 & -10.8458320 & 288.1028935\\
+			\hline
+			$X_0 = 99900, Y_0 = 100$ & 0.5237886 & 1.632971 & 0.2436375 & 3.754091 & 0.7149415 & 0.9374275 & -0.3051038 & 8.8837193\\
+			\hline
+			$X_0 = 99990, Y_0 = 10$ & -33.0051124 & 433.594227 & -1.3382741 & 3.047220 & -0.2463189 & 1.5875310 & 0.4368556 & 0.7686276\\
+			\hline
+			$X_0 = 99999, Y_0 = 1$ & 10.9282204 & 19.297359 & -98.6422337 & 3113.332617 & -0.8194023 & 1.7121668 & 16.6102769 & 18.1045393\\
+			\hline
+		\end{tabular}
+	\end{footnotesize}
+\end{table}
+
+\begin{table}[H]
+
+	\caption{\label{tab:}$\rr$ Estimates and Std. Errs, SIRLMA Model,
+		$X_0 = 99950, Y_0 = 50$, $\beta = 0.06, \gamma = 0.03$}
+	\centering
+	\begin{footnotesize}
+		\begin{tabular}[t]{l|r|r|r|r|r|r|r|r}
+			\hline
+			Data Set & Norm Est & Norm SE & Norm-M Est & Norm-M SE & AR Est & AR SE & AR-M Est & AR-M SE\\
+			\hline
+			$\sigma_X = 50, \sigma_Y = 2$ & 2.0398087 & 0.2773968 & 1.626788 & 0.0864885 & 1.4643659 & 0.0750183 & 2.8515449 & 0.5540996\\
+			\hline
+			$\sigma_X = 500, \sigma_Y = 20$ & 1.6600538 & 1.0693161 & 1.684919 & 0.2571470 & 1.1531754 & 0.0997937 & -4.9533560 & 32.2198830\\
+			\hline
+			$\sigma_X = 2500, \sigma_Y = 100$ & 0.4651914 & 7.5508447 & 4.206167 & 66.2206340 & 1.0391560 & 0.0166472 & -0.1164499 & 1.9126019\\
+			\hline
+			$\sigma_X = 10000, \sigma_Y = 500$ & 1.0899949 & 0.1202859 & 3.976724 & 4.4279629 & 0.6981674 & 0.2429844 & 1.3944715 & 1.0825033\\
+			\hline
+		\end{tabular}
+	\end{footnotesize}
+\end{table}
+
+\begin{table}[H]
+
+	\caption{\label{tab:}$\rr$ Estimates and Std. Errs, SIRLMA Model,
+		$X_0 = 99950, Y_0 = 50$, $\sigma_X = 100, \sigma_Y = 5$}
+	\centering
+	\begin{footnotesize}
+		\begin{tabular}[t]{l|r|r|r|r|r|r|r|r}
+			\hline
+			Data Set & Norm Est & Norm SE & Norm-M Est & Norm-M SE & AR Est & AR SE & AR-M Est & AR-M SE\\
+			\hline
+			1st Order & 1.2559522 & 0.1296269 & 1.2209276 & 0.0597736 & 1.2670255 & 0.1588846 & 1.2476445 & 0.1150267\\
+			\hline
+			4th Order & 0.3632202 & 0.1230895 & -0.0201429 & 0.1133200 & 0.3580085 & 0.1258904 & -0.0161666 & 0.1103940\\
+			\hline
+			Linear SIR & 1.2191443 & 0.1093267 & 1.2067608 & 0.0480986 & 1.2146478 & 0.0627181 & 0.9503153 & 0.1066437\\
+			\hline
+		\end{tabular}
+	\end{footnotesize}
+\end{table}
+
+\begin{table}[H]
+
+	\caption{\label{tab:}$\rr$ Estimates and Std. Errs, SIRLMA Model,
+		$\beta = 0.06, \gamma = 0.03$, $X_0 = 99950, Y_0 = 50$, $\sigma_X = 100, \sigma_Y = 5$}
+	\centering
+	\begin{tabular}[t]{l|r|r|r|r|r|r|r|r}
+		\hline
+		Data Set & Norm Est & Norm SE & Norm-M Est & Norm-M SE & AR Est & AR SE & AR-M Est & AR-M SE\\
+		\hline
+		T = 200 & 0.9496498 & 0.0825419 & 1.052892 & 0.2237478 & 1.0611334 & 0.0328849 & 0.7221436 & 0.1920598\\
+		\hline
+		T = 100 & 1.2374397 & 0.4953607 & 1.202445 & 0.5438859 & 0.9140701 & 0.2563191 & 1.1584230 & 0.3660429\\
+		\hline
+		T = 50 & 0.9720834 & 0.0507265 & 1.924122 & 1.4342312 & 3.7359315 & 59.4952531 & 0.5305768 & 1.0293919\\
+		\hline
+		T = 20 & 1.4418031 & 7.3163878 & 1.121779 & 1.7267097 & -5.2984368 & 65.6502768 & 0.8899288 & 0.7605429\\
+		\hline
+	\end{tabular}
+\end{table}
+
+\subsection{SIR Linear Model Approximation, All Time Points}
+
 \begin{table}[H]
@@
 			\hline
-		Baseline1 & 1.399476 & 0.6034464 & 1.422638 & 0.7411130 & 1.050638 & 4.8889559 & 1.403266 & 0.8205110\\
-		\hline
-		Baseline2 & 1.407952 & 0.6316481 & 1.444502 & 0.9255988 & 1.372240 & 0.8699573 & 1.451384 & 0.7475245\\
-		\hline
-		Baseline3 & 1.410617 & 0.6223034 & 1.360441 & 0.8507911 & 1.374595 & 1.2564522 & 1.437625 & 0.6731037\\
-		\hline
-		Baseline4 & 1.389717 & 0.6186280 & 1.391034 & 0.5732198 & 1.598858 & 1.3621554 & 1.324797 & 0.6542186\\
-		\hline
-		Baseline5 & 1.390334 & 0.6670626 & 1.387943 & 0.5756532 & 1.362033 & 1.0034553 & 1.541135 & 0.9082263\\
-		\hline
-	\end{tabular}
-\end{footnotesize}
-\end{table}
+			Baseline 1 & 1.398016 & 0.6028170 & 1.421201 & 0.7403641 & 1.049360 & 4.8830065 & 1.401437 & 0.8194417\\
+			\hline
+			Baseline 2 & 1.406323 & 0.6309170 & 1.442263 & 0.9241642 & 1.369425 & 0.8681724 & 1.449109 & 0.7463529\\
+			\hline
+			Baseline 3 & 1.408807 & 0.6215050 & 1.358127 & 0.8493445 & 1.375713 & 1.2574743 & 1.436717 & 0.6726785\\
+			\hline
+			Baseline 4 & 1.387819 & 0.6177830 & 1.390251 & 0.5728970 & 1.596744 & 1.3603545 & 1.325567 & 0.6545989\\
+			\hline
+			Baseline 5 & 1.387599 & 0.6657501 & 1.386679 & 0.5751290 & 1.359862 & 1.0018556 & 1.538778 & 0.9068372\\
+			\hline
+		\end{tabular}
+	\end{footnotesize}
+\end{table}
+
+\begin{table}[H]
+
+	\caption{\label{tab:}$\rr$ Estimates and Std. Errs, SIRLMAT Model,
+		$X_0 = 99950, Y_0 = 50$, $\sigma_X = 100, \sigma_Y = 5$, $\beta = 0.06$}
+	\centering
+	\begin{footnotesize}
+		\begin{tabular}[t]{l|r|r|r|r|r|r|r|r}
+			\hline
+			Data Set & Norm Est & Norm SE & Norm-M Est & Norm-M SE & AR Est & AR SE & AR-M Est & AR-M SE\\
+			\hline
+			$\gamma$ = 0.001 & -2.8535930 & 193.6526107 & 7.7685771 & 103.2205552 & 76.340382 & 472.4155164 & -69.9067922 & 1603.7792325\\
+			\hline
+			$\gamma$ = 0.04 & 1.4369752 & 0.7498020 & 1.4719115 & 3.0719036 & 1.292978 & 0.3342698 & 1.2769691 & 0.2448445\\
+			\hline
+			$\gamma$ = 0.06 & 0.9847650 & 0.1423066 & 0.9945574 & 0.0160864 & 1.008112 & 0.5987452 & 1.0268632 & 0.5015724\\
+			\hline
+			$\gamma$ = 0.24 & 0.6746914 & 6.2471252 & 0.3871567 & 4.8817556 & 1.786500 & 15.9420420 & 0.2221191 & 5.4112205\\
+			\hline
+		\end{tabular}
+	\end{footnotesize}
+\end{table}
+
+\begin{table}[H]
+
+	\caption{\label{tab:}$\rr$ Estimates and Std. Errs, SIRLMAT Model,
+		$\sigma_X = 100, \sigma_Y = 5$, $\beta = 0.06, \gamma = 0.03$}
+	\centering
+	\begin{footnotesize}
+		\begin{tabular}[t]{l|r|r|r|r|r|r|r|r}
+			\hline
+			Data Set & Norm Est & Norm SE & Norm-M Est & Norm-M SE & AR Est & AR SE & AR-M Est & AR-M SE\\
+			\hline
+			$X_0 = 99000, Y_0 = 1000$ & 0.730610 & 4.6810416 & 0.9720592 & 0.6586069 & 3.163816 & 45.776770 & 0.9367865 & 0.8857030\\
+			\hline
+			$X_0 = 99900, Y_0 = 100$ & 1.236415 & 0.7388933 & 1.1744675 & 1.3957615 & 1.249219 & 0.813166 & 1.1960662 & 0.6589163\\
+			\hline
+			$X_0 = 99990, Y_0 = 10$ & 1.459703 & 10.9219088 & 0.8531016 & 12.5573016 & -3.846910 & 71.297410 & -0.5884883 & 39.4915091\\
+			\hline
+			$X_0 = 99999, Y_0 = 1$ & 1.624645 & 5.0359845 & 0.4174208 & 12.8380228 & 1.301346 & 25.850056 & 3.1746775 & 12.8142151\\
+			\hline
+		\end{tabular}
+	\end{footnotesize}
+\end{table}
+
+\begin{table}[H]
+
+	\caption{\label{tab:}$\rr$ Estimates and Std. Errs, SIRLMAT Model,
+		$X_0 = 99950, Y_0 = 50$, $\beta = 0.06, \gamma = 0.03$}
+	\centering
+	\begin{footnotesize}
+		\begin{tabular}[t]{l|r|r|r|r|r|r|r|r}
+			\hline
+			Data Set & Norm Est & Norm SE & Norm-M Est & Norm-M SE & AR Est & AR SE & AR-M Est & AR-M SE\\
+			\hline
+			$\sigma_X = 50, \sigma_Y = 2$ & 1.414863 & 0.5779651 & 1.509908 & 2.1243768 & 1.3999424 & 0.7193567 & 1.426323 & 0.6782388\\
+			\hline
+			$\sigma_X = 500, \sigma_Y = 20$ & 1.353711 & 1.5619894 & 1.446169 & 0.7223815 & 1.0183856 & 3.5577619 & 1.996902 & 15.7391571\\
+			\hline
+			$\sigma_X = 2500, \sigma_Y = 100$ & 1.287786 & 4.3947658 & 1.365338 & 3.2675366 & 1.1984800 & 1.6107382 & 0.942443 & 4.3914050\\
+			\hline
+			$\sigma_X = 10000, \sigma_Y = 500$ & 4.811335 & 70.6594883 & 1.500486 & 1.0046353 & 0.9241342 & 1.2799734 & 1.097322 & 3.1937985\\
+			\hline
+		\end{tabular}
+	\end{footnotesize}
+\end{table}
+
+\begin{table}[H]
+
+	\caption{\label{tab:}$\rr$ Estimates and Std. Errs, SIRLMAT Model,
+		$X_0 = 99950, Y_0 = 50$, $\sigma_X = 100, \sigma_Y = 5$}
+	\centering
+	\begin{footnotesize}
+		\begin{tabular}[t]{l|r|r|r|r|r|r|r|r}
+			\hline
+			Data Set & Norm Est & Norm SE & Norm-M Est & Norm-M SE & AR Est & AR SE & AR-M Est & AR-M SE\\
+			\hline
+			1st Order & 1.1509447 & 0.1012249 & 1.138621 & 0.1738250 & 1.147184 & 0.0681524 & 1.145189 & 0.0811484\\
+			\hline
+			4th Order & 0.8246203 & 6.0710436 & 1.000597 & 2.5841000 & 3.255953 & 26.3713818 & 1.385674 & 8.5430506\\
+			\hline
+			Linear SIR & 1.2009801 & 0.0031100 & 1.200130 & 0.0040763 & 1.203159 & 0.0148085 & 1.195875 & 0.0246927\\
+			\hline
+		\end{tabular}
+	\end{footnotesize}
+\end{table}
+
+\begin{table}[H]
+
+	\caption{\label{tab:}$\rr$ Estimates and Std. Errs, SIRLMAT Model,
+		$\beta = 0.06, \gamma = 0.03$, $X_0 = 99950, Y_0 = 50$, $\sigma_X = 100, \sigma_Y = 5$}
+	\centering
+	\begin{tabular}[t]{l|r|r|r|r|r|r|r|r}
+		\hline
+		Data Set & Norm Est & Norm SE & Norm-M Est & Norm-M SE & AR Est & AR SE & AR-M Est & AR-M SE\\
+		\hline
+		T = 200 & 3.3442696 & 21.4895422 & 1.8327212 & 0.2855541 & 1.7170357 & 0.4970127 & 2.294206 & 6.234463\\
+		\hline
+		T = 100 & 1.2354159 & 3.0461522 & 2.7980661 & 3.8629146 & 0.7524086 & 2.5622573 & 3.434636 & 14.604145\\
+		\hline
+		T = 50 & 0.7050038 & 2.8719153 & 0.3376738 & 7.9042199 & 0.6897879 & 2.7333396 & 1.274668 & 3.840504\\
+		\hline
+		T = 20 & 1.0078807 & 0.1424793 & 2.7045722 & 4.8031306 & 0.5716857 & 1.3981528 & 1.449643 & 1.542269\\
+		\hline
+	\end{tabular}
+\end{table}
+
+\subsection{Incidence to Prevalence Ratio}
+
 \begin{table}[H]
@@
 			\hline
-		Baseline1 & 3.111569 & 10.849009 & 0.8738855 & 1.543847 & 2.620045 & 8.877271 & 0.9175130 & 2.234642\\
-		\hline
-		Baseline2 & 2.652218 & 7.415605 & 0.8982691 & 1.805619 & 2.055229 & 6.085662 & 0.9016198 & 1.872737\\
-		\hline
-		Baseline3 & 2.767507 & 8.021874 & 0.8648371 & 1.477029 & 2.062531 & 5.512523 & 1.0049512 & 3.050472\\
-		\hline
-		Baseline4 & 3.057198 & 11.812401 & 1.0174550 & 2.864584 & 2.180584 & 6.925265 & 1.0325816 & 2.768722\\
-		\hline
-		Baseline5 & 2.668934 & 7.472743 & 0.9696804 & 2.452165 & 2.231211 & 5.995580 & 0.9224206 & 2.770667\\
-		\hline
-	\end{tabular}
-\end{footnotesize}
-\end{table}
+			Baseline 1 & 2.592974 & 9.040841 & 0.7282379 & 1.286539 & 2.183371 & 7.397726 & 0.7645942 & 1.862201\\
+			\hline
+			Baseline 2 & 2.210182 & 6.179671 & 0.7485576 & 1.504683 & 1.712691 & 5.071385 & 0.7513498 & 1.560614\\
+			\hline
+			Baseline 3 & 2.306256 & 6.684895 & 0.7206976 & 1.230857 & 1.718776 & 4.593769 & 0.8374593 & 2.542060\\
+			\hline
+			Baseline 4 & 2.547665 & 9.843667 & 0.8478792 & 2.387153 & 1.817153 & 5.771054 & 0.8604847 & 2.307269\\
+			\hline
+			Baseline 5 & 2.224111 & 6.227286 & 0.8080670 & 2.043471 & 1.859343 & 4.996316 & 0.7686838 & 2.308889\\
+			\hline
+		\end{tabular}
+	\end{footnotesize}
+\end{table}
+
+\begin{table}[H]
+
+	\caption{\label{tab:}$\rr$ Estimates and Std. Errs, SIRIPR Model,
+		$X_0 = 99950, Y_0 = 50$, $\sigma_X = 100, \sigma_Y = 5$, $\beta = 0.06$}
+	\centering
+	\begin{footnotesize}
+		\begin{tabular}[t]{l|r|r|r|r|r|r|r|r}
+			\hline
+			Data Set & Norm Est & Norm SE & Norm-M Est & Norm-M SE & AR Est & AR SE & AR-M Est & AR-M SE\\
+			\hline
+			$\gamma$ = 0.001 & 54.146506 & 200.2595 & 24.408675 & 66.166769 & 58.723744 & 219.989055 & 28.5230747 & 120.928788\\
+			\hline
+			$\gamma$ = 0.04 & 4.771508 & 12.8533 & 1.387753 & 2.630779 & 3.613754 & 8.536531 & 1.3965600 & 2.840997\\
+			\hline
+			$\gamma$ = 0.06 & 19.828076 & 29.1272 & 1.121667 & 4.400499 & 15.651052 & 23.539321 & 1.1251847 & 4.900022\\
+			\hline
+			$\gamma$ = 0.24 & 181.015802 & 742.1088 & 6.672760 & 89.936233 & 148.308767 & 740.018239 & 0.8145271 & 11.870908\\
+			\hline
+		\end{tabular}
+	\end{footnotesize}
+\end{table}
+
+\begin{table}[H]
+
+	\caption{\label{tab:}$\rr$ Estimates and Std. Errs, SIRIPR Model,
+		$\sigma_X = 100, \sigma_Y = 5$, $\beta = 0.06, \gamma = 0.03$}
+	\centering
+	\begin{footnotesize}
+		\begin{tabular}[t]{l|r|r|r|r|r|r|r|r}
+			\hline
+			Data Set & Norm Est & Norm SE & Norm-M Est & Norm-M SE & AR Est & AR SE & AR-M Est & AR-M SE\\
+			\hline
+			$X_0 = 99000, Y_0 = 1000$ & 1.285071 & 1.440980 & 0.9866029 & 1.057043 & 1.193594 & 1.195820 & 0.9960353 & 0.9987062\\
+			\hline
+			$X_0 = 99900, Y_0 = 100$ & 2.943264 & 7.736638 & 1.4707234 & 3.367245 & 2.230244 & 4.859448 & 1.4236604 & 2.8121697\\
+			\hline
+			$X_0 = 99990, Y_0 = 10$ & 18.663014 & 67.156845 & 1.8455795 & 5.401173 & 16.739763 & 71.982690 & 1.6417916 & 4.4145381\\
+			\hline
+			$X_0 = 99999, Y_0 = 1$ & 774.274659 & 7491.005080 & 3.6060918 & 26.540940 & 229.184142 & 1199.155653 & 2.5082330 & 12.8639109\\
+			\hline
+		\end{tabular}
+	\end{footnotesize}
+\end{table}
+
+\begin{table}[H]
+
+	\caption{\label{tab:}$\rr$ Estimates and Std. Errs, SIRIPR Model,
+		$X_0 = 99950, Y_0 = 50$, $\beta = 0.06, \gamma = 0.03$}
+	\centering
+	\begin{footnotesize}
+		\begin{tabular}[t]{l|r|r|r|r|r|r|r|r}
+			\hline
+			Data Set & Norm Est & Norm SE & Norm-M Est & Norm-M SE & AR Est & AR SE & AR-M Est & AR-M SE\\
+			\hline
+			$\sigma_X = 50, \sigma_Y = 2$ & 2.760752 & 8.131869 & 1.450899 & 2.052241 & 2.19804 & 5.456857 & 1.446505 & 2.182072\\
+			\hline
+			$\sigma_X = 50, \sigma_Y = 25$ & 17.377923 & 71.374848 & 1.889790 & 8.177704 & 12.91733 & 41.063025 & 2.373601 & 20.983311\\
+			\hline
+			$\sigma_X = 2500, \sigma_Y = 100$ & 81.490167 & 348.735085 & 2.267162 & 17.058029 & 195.37080 & 1451.040149 & 2.924092 & 18.434409\\
+			\hline
+			$\sigma_X = 10000, \sigma_Y = 500$ & 239.809101 & 1298.136206 & 4.500556 & 59.404233 & 92.10901 & 272.487471 & 3.535111 & 22.583706\\
+			\hline
+		\end{tabular}
+	\end{footnotesize}
+\end{table}
+
+\begin{table}[H]
+
+	\caption{\label{tab:}$\rr$ Estimates and Std. Errs, SIRIPR Model,
+		$X_0 = 99950, Y_0 = 50$, $\sigma_X = 100, \sigma_Y = 5$}
+	\centering
+	\begin{footnotesize}
+		\begin{tabular}[t]{l|r|r|r|r|r|r|r|r}
+			\hline
+			Data Set & Norm Est & Norm SE & Norm-M Est & Norm-M SE & AR Est & AR SE & AR-M Est & AR-M SE\\
+			\hline
+			1st Order & 2.699973 & 9.131131 & 2.297208 & 7.14196 & 4.334767 & 41.325146 & 4.194341 & 36.201142\\
+			\hline
+			4th Order & 6.260963 & 20.337163 & 5.924029 & 17.81136 & 6.268500 & 20.536192 & 5.892582 & 17.341763\\
+			\hline
+			Linear SIR & 1.516803 & 1.554241 & 1.208578 & 1.12544 & 1.379871 & 1.311406 & 1.208521 & 1.003065\\
+			\hline
+		\end{tabular}
+	\end{footnotesize}
+\end{table}
+
+\begin{table}[H]
+
+	\caption{\label{tab:}$\rr$ Estimates and Std. Errs, SIRIPR Model,
+		$\beta = 0.06, \gamma = 0.03$, $X_0 = 99950, Y_0 = 50$, $\sigma_X = 100, \sigma_Y = 5$}
+	\centering
+	\begin{tabular}[t]{l|r|r|r|r|r|r|r|r}
+		\hline
+		Data Set & Norm Est & Norm SE & Norm-M Est & Norm-M SE & AR Est & AR SE & AR-M Est & AR-M SE\\
+		\hline
+		T = 200 & 7.345813 & 16.12889 & 2.048281 & 3.959862 & 5.553952 & 13.33904 & 2.052063 & 4.119090\\
+		\hline
+		T = 100 & 12.683604 & 21.50385 & 2.278429 & 5.329775 & 9.221913 & 18.11852 & 2.245441 & 5.708995\\
+		\hline
+		T = 50 & 19.220693 & 27.75807 & 2.470106 & 6.581896 & 14.757521 & 24.04262 & 2.616444 & 7.417433\\
+		\hline
+		T = 20 & 25.731306 & 35.29261 & 1.415537 & 4.565460 & 20.489247 & 33.72596 & 3.885644 & 8.289876\\
+		\hline
+	\end{tabular}
+\end{table}
+
+\subsection{Smoothed Incidence to Prevalence Ratio}
+
 \begin{table}[H]
@@
 			\hline
-		Baseline1 & 1.110211 & 0 & 1.030592 & 0 & 1.006362 & 0 & 0.9295163 & 0\\
-		\hline
-		Baseline2 & 1.132249 & 0 & 1.089786 & 0 & 1.055082 & 0 & 0.8371411 & 0\\
-		\hline
-		Baseline3 & 1.163383 & 0 & 1.102025 & 0 & 1.207890 & 0 & 0.8555288 & 0\\
-		\hline
-		Baseline4 & 1.177337 & 0 & 1.049182 & 0 & 1.039892 & 0 & 1.1049484 & 0\\
-		\hline
-		Baseline5 & 1.113234 & 0 & 1.092939 & 0 & 1.381775 & 0 & 0.9232862 & 0\\
-		\hline
-	\end{tabular}
-\end{footnotesize}
-\end{table}
+			Baseline 1 & 0.9251762 & $<$ 1e-07 & 0.8588266 & $<$ 1e-07 & 0.8386350 & $<$ 1e-07 & 0.7745969 & $<$ 1e-07\\
+			\hline
+			Baseline 2 & 0.9435412 & $<$ 1e-07 & 0.9081552 & $<$ 1e-07 & 0.8792348 & $<$ 1e-07 & 0.6976176 & $<$ 1e-07\\
+			\hline
+			Baseline 3 & 0.9694855 & $<$ 1e-07 & 0.9183543 & $<$ 1e-07 & 1.0065755 & $<$ 1e-07 & 0.7129407 & $<$ 1e-07\\
+			\hline
+			Baseline 4 & 0.9811143 & $<$ 1e-07 & 0.8743186 & $<$ 1e-07 & 0.8665770 & $<$ 1e-07 & 0.9207903 & $<$ 1e-07\\
+			\hline
+			Baseline 5 & 0.9276948 & $<$ 1e-07 & 0.9107829 & $<$ 1e-07 & 1.1514790 & $<$ 1e-07 & 0.7694052 & $<$ 1e-07\\
+			\hline
+		\end{tabular}
+	\end{footnotesize}
+\end{table}
+
+\begin{table}[H]
+
+	\caption{\label{tab:}$\rr$ Estimates and Std. Errs, SIRSIPR Model,
+		$X_0 = 99950, Y_0 = 50$, $\sigma_X = 100, \sigma_Y = 5$, $\beta = 0.06$}
+	\centering
+	\begin{footnotesize}
+		\begin{tabular}[t]{l|r|r|r|r|r|r|r|r}
+			\hline
+			Data Set & Norm Est & Norm SE & Norm-M Est & Norm-M SE & AR Est & AR SE & AR-M Est & AR-M SE\\
+			\hline
+			$\gamma$ = 0.001 & 3.9645187 & $<$ 1e-07 & 3.8285414 & $<$ 1e-07 & 3.9055582 & 0.000000 & 3.7112781 & $<$ 1e-07 \\
+			\hline
+			$\gamma$ = 0.04 & 1.7451025 & $<$ 1e-07 & 2.8785564 & $<$ 1e-07 & 68.7390000 & 0.000000 & 1.2465750 & $<$ 1e-07 \\
+			\hline
+			$\gamma$ = 0.06 & 0.9919790 & 0.0559052 & 0.9925620 & 0.0297783 & 0.8718278 & 0.087266 & 1.0294124 & 0.0510585\\
+			\hline
+			$\gamma$ = 0.24 & -0.2484284 & 0.1159769 & -0.0317365 & 0.0335514 & -2.8642901 & 0.000000 & 0.7365437 & 0.0675331\\
+			\hline
+		\end{tabular}
+	\end{footnotesize}
+\end{table}
+
+\begin{table}[H]
+
+	\caption{\label{tab:}$\rr$ Estimates and Std. Errs, SIRSIPR Model,
+		$\sigma_X = 100, \sigma_Y = 5$, $\beta = 0.06, \gamma = 0.03$}
+	\centering
+	\begin{footnotesize}
+		\begin{tabular}[t]{l|r|r|r|r|r|r|r|r}
+			\hline
+			Data Set & Norm Est & Norm SE & Norm-M Est & Norm-M SE & AR Est & AR SE & AR-M Est & AR-M SE\\
+			\hline
+			$X_0 = 99000, Y_0 = 1000$ & 2.0870176 & 57.9229822 & 2.016608 & 57.8153675 & 2.285424 & 57.4382982 & 1.922192 & 57.0457534\\
+			\hline
+			$X_0 = 99900, Y_0 = 100$ & -0.8808153 & $<$ 1e-07 & -1.150393 & $<$ 1e-07 & -6.563660 & $<$ 1e-07 & -1.777316 & $<$ 1e-07 \\
+			\hline
+			$X_0 = 99990, Y_0 = 10$ & 1.3451986 & 0.3236824 & 1.354925 & 0.3158698 & 1.381198 & 0.3093223 & 1.345069 & 0.3258082\\
+			\hline
+			$X_0 = 99999, Y_0 = 1$ & 2.4390757 & $<$ 1e-07 & 5.411582 & $<$ 1e-07 & 1.204689 & $<$ 1e-07 & 1.035624 & $<$ 1e-07 \\
+			\hline
+		\end{tabular}
+	\end{footnotesize}
+\end{table}
+
+\begin{table}[H]
+
+	\caption{\label{tab:}$\rr$ Estimates and Std. Errs, SIRSIPR Model,
+		$X_0 = 99950, Y_0 = 50$, $\beta = 0.06, \gamma = 0.03$}
+	\centering
+	\begin{footnotesize}
+		\begin{tabular}[t]{l|r|r|r|r|r|r|r|r}
+			\hline
+			Data Set & Norm Est & Norm SE & Norm-M Est & Norm-M SE & AR Est & AR SE & AR-M Est & AR-M SE\\
+			\hline
+			$\sigma_X = 50, \sigma_Y = 2$ & 1.836824 & $<$ 1e-07 & 1.7778150 & $<$ 1e-07 & 1.7603970 & $<$ 1e-07 & 1.842289 & $<$ 1e-07\\
+			\hline
+			$\sigma_X = 500, \sigma_Y = 25$ & 3.193066 & $<$ 1e-07 & 1.3211468 & $<$ 1e-07 & 1.3509726 & $<$ 1e-07 & 1.334458 & $<$ 1e-07\\
+			\hline
+			$\sigma_X = 2500, \sigma_Y = 100$ & 1.060609 & $<$ 1e-07 & 1.2345884 & $<$ 1e-07 & 1.7027546 & $<$ 1e-07 & 2.596160 & $<$ 1e-07\\
+			\hline
+			$\sigma_X = 10000, \sigma_Y = 500$ & -2.535946 & $<$ 1e-07 & 0.7094139 & $<$ 1e-07 & 0.8911487 & $<$ 1e-07 & -1.103940 & $<$ 1e-07\\
+			\hline
+		\end{tabular}
+	\end{footnotesize}
+\end{table}
+
+\begin{table}[H]
+
+	\caption{\label{tab:}$\rr$ Estimates and Std. Errs, SIRSIPR Model,
+		$X_0 = 99950, Y_0 = 50$, $\sigma_X = 100, \sigma_Y = 5$}
+	\centering
+	\begin{footnotesize}
+		\begin{tabular}[t]{l|r|r|r|r|r|r|r|r}
+			\hline
+			Data Set & Norm Est & Norm SE & Norm-M Est & Norm-M SE & AR Est & AR SE & AR-M Est & AR-M SE\\
+			\hline
+			1st Order & 58.836635 & $<$ 1e-07 & 27.429685 & $<$ 1e-07 & 1.498805 & $<$ 1e-07 & -14.135196 & $<$ 1e-07 \\
+			\hline
+			4th Order & 2.848624 & $<$ 1e-07 & 15.083621 & $<$ 1e-07 & 2.766152 & $<$ 1e-07 & 14.972580 & $<$ 1e-07 \\
+			\hline
+			Linear SIR & 1.294079 & 0.2151131 & 1.294364 & 0.2158095 & 1.296731 & 0.2215603 & 1.304489 & 0.2299876\\
+			\hline
+		\end{tabular}
+	\end{footnotesize}
+\end{table}
+
+\begin{table}[H]
+
+	\caption{\label{tab:}$\rr$ Estimates and Std. Errs, SIRSIPR Model,
+		$\beta = 0.06, \gamma = 0.03$, $X_0 = 99950, Y_0 = 50$, $\sigma_X = 100, \sigma_Y = 5$}
+	\centering
+	\begin{tabular}[t]{l|r|r|r|r|r|r|r|r}
+		\hline
+		Data Set & Norm Est & Norm SE & Norm-M Est & Norm-M SE & AR Est & AR SE & AR-M Est & AR-M SE\\
+		\hline
+		T = 200 & -2.119890 & 0.000000 & 2.6624046 & 0.0000000 & 3.3187716 & 0.0000000 & 2.363125 & 32.2727752\\
+		\hline
+		T = 100 & 2.341119 & 1.168462 & 2.2020092 & 0.5643447 & 2.0919450 & 1.0699219 & 1.963112 & 0.3489431\\
+		\hline
+		T = 50 & 2.190565 & 2.238582 & 1.9477092 & 0.4198946 & 1.8701913 & 0.7519526 & 2.279451 & 0.6584392\\
+		\hline
+		T = 20 & 4.177245 & 7.320297 & 0.9702882 & 0.8673670 & 0.3570221 & 7.6611297 & 3.984916 & 2.1350792\\
+		\hline
+	\end{tabular}
+\end{table}
+
+\subsection{SIR Log Linear}
+
 \begin{table}[H]
@@
 \end{table}
+
+\begin{table}[H]
+
+	\caption{\label{tab:}$\rr$ Estimates and Std. Errs, LOGLINEAR Model,
+		$X_0 = 99950, Y_0 = 50$, $\sigma_X = 100, \sigma_Y = 5$, $\beta = 0.06$}
+	\centering
+	\begin{footnotesize}
+		\begin{tabular}[t]{l|r|r|r|r|r|r|r|r}
+			\hline
+			Data Set & Norm Est & Norm SE & Norm-M Est & Norm-M SE & AR Est & AR SE & AR-M Est & AR-M SE\\
+			\hline
+			$\gamma$ = 0.001 & 46.3209595 & 0.7797216 & 59.8300371 & 0.1079267 & 41.897714 & 0.9767785 & 70.7075929 & 0.9517239\\
+			\hline
+			$\gamma$ = 0.04 & 1.5002728 & 0.0001905 & 1.5000436 & 0.0000928 & 1.499516 & 0.0002653 & 1.5000445 & 0.0001481\\
+			\hline
+			$\gamma$ = 0.06 & 0.9989449 & 0.0007428 & 0.9994831 & 0.0003320 & 1.000275 & 0.0010605 & 1.0008545 & 0.0004910\\
+			\hline
+			$\gamma$ = 0.24 & 0.9912397 & 0.0030505 & 0.8623034 & 0.0087543 & 1.001367 & 0.0025488 & 0.9198973 & 0.0052616\\
+			\hline
+		\end{tabular}
+	\end{footnotesize}
+\end{table}
+
+\begin{table}[H]
+
+	\caption{\label{tab:}$\rr$ Estimates and Std. Errs, LOGLINEAR Model,
+		$\sigma_X = 100, \sigma_Y = 5$, $\beta = 0.06, \gamma = 0.03$}
+	\centering
+	\begin{footnotesize}
+		\begin{tabular}[t]{l|r|r|r|r|r|r|r|r}
+			\hline
+			Data Set & Norm Est & Norm SE & Norm-M Est & Norm-M SE & AR Est & AR SE & AR-M Est & AR-M SE\\
+			\hline
+			$X_0 = 99000, Y_0 = 1000$ & 1.999515 & 0.0002828 & 2.000144 & 0.0001594 & 1.999753 & 0.0004197 & 2.000586 & 0.0003024\\
+			\hline
+			$X_0 = 99900, Y_0 = 100$ & 2.000011 & 0.0002315 & 1.999732 & 0.0001526 & 2.000473 & 0.0002776 & 2.000640 & 0.0002656\\
+			\hline
+			$X_0 = 99990, Y_0 = 10$ & 1.999886 & 0.0002517 & 1.999599 & 0.0001361 & 2.000608 & 0.0003289 & 1.999049 & 0.0002664\\
+			\hline
+			$X_0 = 99999, Y_0 = 1$ & 1.999690 & 0.0007731 & 2.000503 & 0.0003700 & 2.002716 & 0.0010572 & 1.999919 & 0.0005663\\
+			\hline
+		\end{tabular}
+	\end{footnotesize}
+\end{table}
+
+\begin{table}[H]
+
+	\caption{\label{tab:}$\rr$ Estimates and Std. Errs, LOGLINEAR Model,
+		$X_0 = 99950, Y_0 = 50$, $\beta = 0.06, \gamma = 0.03$}
+	\centering
+	\begin{footnotesize}
+		\begin{tabular}[t]{l|r|r|r|r|r|r|r|r}
+			\hline
+			Data Set & Norm Est & Norm SE & Norm-M Est & Norm-M SE & AR Est & AR SE & AR-M Est & AR-M SE\\
+			\hline
+			$\sigma_X = 50, \sigma_Y = 2$ & 2.059934 & 0.0246402 & 1.995133 & 0.0038894 & 2.209575 & 0.0439100 & 2.545819 & 0.0568537\\
+			\hline
+			$\sigma\_X = 500, \sigma_Y = 20$ & 2.001882 & 0.0054085 & 1.995742 & 0.0015419 & 2.038646 & 0.0081562 & 2.015675 & 0.0031322\\
+			\hline
+			$\sigma\_X = 2500, \sigma_Y = 100$ & 1.999891 & 0.0001037 & 2.000036 & 0.0000930 & 1.999143 & 0.0001394 & 1.999749 & 0.0001328\\
+			\hline
+			$\sigma\_X = 10000, \sigma_Y = 500$ & 2.003402 & 0.0010320 & 1.998765 & 0.0003682 & 2.003763 & 0.0015505 & 1.999829 & 0.0008545\\
+			\hline
+		\end{tabular}
+	\end{footnotesize}
+\end{table}
+
+\begin{table}[H]
+
+	\caption{\label{tab:}$\rr$ Estimates and Std. Errs, LOGLINEAR Model,
+		$X_0 = 99950, Y_0 = 50$, $\sigma_X = 100, \sigma_Y = 5$}
+	\centering
+	\begin{footnotesize}
+		\begin{tabular}[t]{l|r|r|r|r|r|r|r|r}
+			\hline
+			Data Set & Norm Est & Norm SE & Norm-M Est & Norm-M SE & AR Est & AR SE & AR-M Est & AR-M SE\\
+			\hline
+			1st Order & 1.188050 & 0.0006341 & 1.187976 & 0.0006389 & 1.187553 & 0.0006273 & 1.187848 & 0.0006497\\
+			\hline
+			4th Order & 1.955180 & 0.0164608 & 1.955313 & 0.0164212 & 1.955130 & 0.0164623 & 1.955302 & 0.0164165\\
+			\hline
+			Linear SIR & 1.258587 & 0.0010079 & 1.258670 & 0.0010069 & 1.258920 & 0.0009938 & 1.258462 & 0.0010199\\
+			\hline
+		\end{tabular}
+	\end{footnotesize}
+\end{table}
+
+\begin{table}[H]
+
+	\caption{\label{tab:}$\rr$ Estimates and Std. Errs, LOGLINEAR Model,
+		$\beta = 0.06, \gamma = 0.03$, $X_0 = 99950, Y_0 = 50$, $\sigma_X = 100, \sigma_Y = 5$}
+	\centering
+	\begin{tabular}[t]{l|r|r|r|r|r|r|r|r}
+		\hline
+		Data Set & Norm Est & Norm SE & Norm-M Est & Norm-M SE & AR Est & AR SE & AR-M Est & AR-M SE\\
+		\hline
+		T = 200 & 2.001415 & 0.0017620 & 2.000649 & 0.0009166 & 2.0053322 & 0.0023026 & 1.992983 & 0.0019266\\
+		\hline
+		T = 100 & 1.869996 & 0.0322729 & 1.992759 & 0.0104984 & 1.7390224 & 0.0357848 & 2.052393 & 0.0208558\\
+		\hline
+		T = 50 & 1.193299 & 0.0560189 & 1.741553 & 0.0454525 & 1.2098418 & 0.0420061 & 1.690908 & 0.0406272\\
+		\hline
+		T = 20 & 1.042976 & 0.0276261 & 1.747193 & 0.3635417 & 0.9170623 & 0.0337569 & 1.432791 & 0.0397426\\
+		\hline
+	\end{tabular}
+\end{table}
+
+\subsection{SIR Markov Chain}
+
 \begin{table}[H]
@@
 			\hline
-		Baseline1 & 0.9485555 & 0 & 0.9485555 & 0 & 0.9485555 & 0 & 0.9485555 & 0\\
-		\hline
-		Baseline2 & 0.9485555 & 0 & 0.9485555 & 0 & 0.9485555 & 0 & 0.9485555 & 0\\
-		\hline
-		Baseline3 & 0.9485555 & 0 & 0.9485555 & 0 & 0.9485555 & 0 & 0.9485555 & 0\\
-		\hline
-		Baseline4 & 0.9485555 & 0 & 0.9485555 & 0 & 0.9485555 & 0 & 0.9485555 & 0\\
-		\hline
-		Baseline5 & 0.9485555 & 0 & 0.9485555 & 0 & 0.9485555 & 0 & 0.9485555 & 0\\
-		\hline
-	\end{tabular}
-\end{footnotesize}
-\end{table}
+			Baseline1 & 0.9485555 & $<$ 1e-07 & 0.9485555 & $<$ 1e-07 & 0.9485555 & $<$ 1e-07 & 0.9485555 & $<$ 1e-07\\
+			\hline
+			Baseline2 & 0.9485555 & $<$ 1e-07 & 0.9485555 & $<$ 1e-07 & 0.9485555 & $<$ 1e-07 & 0.9485555 & $<$ 1e-07\\
+			\hline
+			Baseline3 & 0.9485555 & $<$ 1e-07 & 0.9485555 & $<$ 1e-07 & 0.9485555 & $<$ 1e-07 & 0.9485555 & $<$ 1e-07\\
+			\hline
+			Baseline4 & 0.9485555 & $<$ 1e-07 & 0.9485555 & $<$ 1e-07 & 0.9485555 & $<$ 1e-07 & 0.9485555 & $<$ 1e-07\\
+			\hline
+			Baseline5 & 0.9485555 & $<$ 1e-07 & 0.9485555 & $<$ 1e-07 & 0.9485555 & $<$ 1e-07 & 0.9485555 & $<$ 1e-07\\
+			\hline
+		\end{tabular}
+	\end{footnotesize}
+\end{table}
+
+\begin{table}[H]
+
+	\caption{\label{tab:}$\rr$ Estimates and Std. Errs, SIRMC Model,
+		$X_0 = 99950, Y_0 = 50$, $\sigma_X = 100, \sigma_Y = 5$, $\beta = 0.06$}
+	\centering
+	\begin{footnotesize}
+		\begin{tabular}[t]{l|r|r|r|r|r|r|r|r}
+			\hline
+			Data Set & Norm Est & Norm SE & Norm-M Est & Norm-M SE & AR Est & AR SE & AR-M Est & AR-M SE\\
+			\hline
+			$\gamma$ = 0.001 & 0.9485554 & 1.0e-07 & 0.9485555 & $<$ 1e-07 & 0.9485555 & $<$ 1e-07 & 0.9485555 & 0e+00\\
+			\hline
+			$\gamma$ = 0.04 & 0.9485555 & $<$ 1e-07 & 0.9485555 & $<$ 1e-07 & 0.9485555 & $<$ 1e-07 & 0.9485555 & 0e+00\\
+			\hline
+			$\gamma$ = 0.06 & 0.9485555 & 2.0e-07 & 0.9485555 & $<$ 1e-07 & 0.9485555 & 2.00e-07 & 0.9485555 & 0e+00\\
+			\hline
+			$\gamma$ = 0.24 & 0.9485812 & 1.5e-05 & 0.9485605 & 1.4e-06 & 0.9486128 & 1.26e-05 & 0.9485558 & 2e-07\\
+			\hline
+		\end{tabular}
+	\end{footnotesize}
+\end{table}
+
+\begin{table}[H]
+
+	\caption{\label{tab:}$\rr$ Estimates and Std. Errs, SIRMC Model,
+		$\sigma_X = 100, \sigma_Y = 5$, $\beta = 0.06, \gamma = 0.03$}
+	\centering
+	\begin{footnotesize}
+		\begin{tabular}[t]{l|r|r|r|r|r|r|r|r}
+			\hline
+			Data Set & Norm Est & Norm SE & Norm-M Est & Norm-M SE & AR Est & AR SE & AR-M Est & AR-M SE\\
+			\hline
+			$X_0 = 99000, Y_0 = 1000$ & 0.9485555 & $<$ 1e-07 & 0.9485555 & $<$ 1e-07 & 0.9485555 & $<$ 1e-07 & 0.9485555 & $<$ 1e-07\\
+			\hline
+			$X_0 = 99900, Y_0 = 100$ & 0.9485555 & $<$ 1e-07 & 0.9485555 & $<$ 1e-07 & 0.9485555 & $<$ 1e-07 & 0.9485555 & $<$ 1e-07\\
+			\hline
+			$X_0 = 99990, Y_0 = 10$ & 0.9485552 & 2.00e-07 & 0.9485555 & $<$ 1e-07 & 0.9485558 & 2.0e-07 & 0.9485555 & $<$ 1e-07\\
+			\hline
+			$X_0 = 99999, Y_0 = 1$ & 0.9486030 & 1.56e-05 & 0.9485557 & $<$ 1e-07 & 0.9485701 & 2.6e-06 & 0.9485555 & $<$ 1e-07\\
+			\hline
+		\end{tabular}
+	\end{footnotesize}
+\end{table}
+
+\begin{table}[H]
+
+	\caption{\label{tab:}$\rr$ Estimates and Std. Errs, SIRMC Model,
+		$X_0 = 99950, Y_0 = 50$, $\beta = 0.06, \gamma = 0.03$}
+	\centering
+	\begin{footnotesize}
+		\begin{tabular}[t]{l|r|r|r|r|r|r|r|r}
+			\hline
+			Data Set & Norm Est & Norm SE & Norm-M Est & Norm-M SE & AR Est & AR SE & AR-M Est & AR-M SE\\
+			\hline
+			$\sigma_X = 50, \sigma_Y = 2$ & 0.9485506 & 3.8e-06 & 0.9485556 & 2e-07 & 0.9485428 & 2.6e-06 & 0.9485538 & 8e-07\\
+			\hline
+			$\sigma\_X = 500, \sigma_Y = 20$ & 0.9485562 & 1.0e-06 & 0.9485554 & 1e-07 & 0.9485640 & 3.6e-06 & 0.9485554 & 1e-07\\
+			\hline
+			$\sigma\_X = 2500, \sigma_Y = 100$ & 0.9485555 & $<$ 1e-07 & 0.9485555 & $<$ 1e-07 & 0.9485555 & $<$ 1e-07 & 0.9485555 & 0e+00\\
+			\hline
+			$\sigma\_X = 10000, \sigma_Y = 500$ & 0.9485554 & 2.0e-07 & 0.9485555 & $<$ 1e-07 & 0.9485554 & 1.0e-07 & 0.9485555 & 1e-07\\
+			\hline
+		\end{tabular}
+	\end{footnotesize}
+\end{table}
+
+\begin{table}[H]
+
+	\caption{\label{tab:}$\rr$ Estimates and Std. Errs, SIRMC Model,
+		$X_0 = 99950, Y_0 = 50$, $\sigma_X = 100, \sigma_Y = 5$}
+	\centering
+	\begin{footnotesize}
+		\begin{tabular}[t]{l|r|r|r|r|r|r|r|r}
+			\hline
+			Data Set & Norm Est & Norm SE & Norm-M Est & Norm-M SE & AR Est & AR SE & AR-M Est & AR-M SE\\
+			\hline
+			1st Order & 0.9485551 & 1e-07 & 0.9485540 & 4e-07 & 0.9485555 & 2e-07 & 0.9485553 & 2.0e-07\\
+			\hline
+			4th Order & 0.9485545 & 6e-07 & 0.9485533 & 1e-06 & 0.9485535 & 9e-07 & 0.9485527 & 1.2e-06\\
+			\hline
+			Linear SIR & 0.9485555 & $<$ 1e-07 & 0.9485555 & $<$ 1e-07 & 0.9485555 & $<$ 1e-07 & 0.9485555 & 0.0e+00\\
+			\hline
+		\end{tabular}
+	\end{footnotesize}
+\end{table}
+
+\begin{table}[H]
+
+	\caption{\label{tab:}$\rr$ Estimates and Std. Errs, SIRMC Model,
+		$\beta = 0.06, \gamma = 0.03$, $X_0 = 99950, Y_0 = 50$, $\sigma_X = 100, \sigma_Y = 5$}
+	\centering
+	\begin{tabular}[t]{l|r|r|r|r|r|r|r|r}
+		\hline
+		Data Set & Norm Est & Norm SE & Norm-M Est & Norm-M SE & AR Est & AR SE & AR-M Est & AR-M SE\\
+		\hline
+		T = 200 & 0.9485555 & 1e-07 & 0.9485555 & 0 & 0.9485555 & 0e+00 & 0.9485555 & 0e+00\\
+		\hline
+		T = 100 & 0.9485555 & 1e-07 & 0.9485555 & 0 & 0.9485555 & 1e-07 & 0.9485555 & 0e+00\\
+		\hline
+		T = 50 & 0.9485554 & 2e-07 & 0.9485555 & 0 & 0.9485555 & 2e-07 & 0.9485555 & 0e+00\\
+		\hline
+		T = 20 & 0.9485556 & 5e-07 & 0.9485554 & 0 & 0.9485552 & 5e-07 & 0.9485556 & 1e-07\\
+		\hline
+	\end{tabular}
+\end{table}
+
+\subsection{SIR Sequential Bayes}
+
 \begin{table}[H]
@@
 
-\subsection{Temporary Hold: Parameter Change}
-
-\begin{table}[H]
-
-	\caption{\label{tab:}$\rr$ Estimates and Std. Errs, SIRLS Model,
-		$X_0 = 99950, Y_0 = 50$, $\sigma_X = 100, \sigma_Y = 5$, $\beta = 0.06$}
-	\centering
-	\begin{footnotesize}
-	\begin{tabular}[t]{l|r|r|r|r|r|r|r|r}
-		\hline
-		Data Set & Norm Est & Norm SE & Norm-M Est & Norm-M SE & AR Est & AR SE & AR-M Est & AR-M SE\\
-		\hline
-		$\gamma$ = 0.001 & 59.9764682 & 0.0434705 & 59.9764682 & 0.0386733 & 60.0151171 & 0.0697888 & 59.9589126 & 0.0631448\\
-		\hline
-		$\gamma$ = 0.04 & 1.5000763 & 0.0126817 & 1.5001978 & 0.0090880 & 1.4989700 & 0.0182852 & 1.4993603 & 0.0165168\\
-		\hline
-		$\gamma$ = 0.06 & 1.0043054 & 0.4140424 & 0.9955554 & 0.1086448 & 0.9915431 & 0.6272644 & 0.9941392 & 0.1891812\\
-		\hline
-		$\gamma$ = 0.24 & 0.2363274 & 4.1042813 & 0.2327691 & 0.6701185 & -0.0695410 & 2.8291362 & 0.3021083 & 0.5657139\\
-		\hline
-	\end{tabular}
-\end{footnotesize}
-\end{table}
-\begin{table}[H]
-
-	\caption{\label{tab:}$\rr$ Estimates and Std. Errs, SIRREPAR Model,
-		$X_0 = 99950, Y_0 = 50$, $\sigma_X = 100, \sigma_Y = 5$, $\beta = 0.06$}
-	\centering
-	\begin{footnotesize}
-	\begin{tabular}[t]{l|r|r|r|r|r|r|r|r}
-		\hline
-		Data Set & Norm Est & Norm SE & Norm-M Est & Norm-M SE & AR Est & AR SE & AR-M Est & AR-M SE\\
-		\hline
-		1st Order & 59.995188 & 0.4343984 & 59.9951876 & 0.3874879 & 60.025452 & 0.6911933 & 59.9960910 & 0.6214841\\
-		\hline
-		4th Order & 1.499883 & 0.0062529 & 1.5001705 & 0.0044936 & 1.499068 & 0.0090313 & 1.4993177 & 0.0081651\\
-		\hline
-		Linear SIR & 1.004306 & 0.0501444 & 0.9955446 & 0.0117215 & 0.991558 & 0.0645788 & 0.9941511 & 0.0190525\\
-		\hline
-		& 4.179525 & 2.7455569 & 3.5709590 & 0.3171284 & 5.192951 & 3.5249847 & 9.8415034 & 2.5646198\\
-		\hline
-	\end{tabular}
-\end{footnotesize}
-\end{table}
-\begin{table}[H]
-
-	\caption{\label{tab:}$\rr$ Estimates and Std. Errs, SIRLMA Model,
-		$X_0 = 99950, Y_0 = 50$, $\sigma_X = 100, \sigma_Y = 5$, $\beta = 0.06$}
-	\centering
-	\begin{footnotesize}
-	\begin{tabular}[t]{l|r|r|r|r|r|r|r|r}
-		\hline
-		Data Set & Norm Est & Norm SE & Norm-M Est & Norm-M SE & AR Est & AR SE & AR-M Est & AR-M SE\\
-		\hline
-		$\gamma$ = 0.001 & -122.0810136 & 104.7144676 & -57.2648513 & 11.6741827 & -19.8476503 & 6.3159285 & -190.2995380 & 290.3182374\\
-		\hline
-		$\gamma$ = 0.04 & 1.1085550 & 0.0868180 & -3.9714925 & 1115.8099120 & 0.9776237 & 0.0216463 & 0.7343152 & 0.1793076\\
-		\hline
-		$\gamma$ = 0.06 & 1.0835046 & 0.1530378 & 0.9903843 & 0.1184626 & 0.9273120 & 0.0602626 & 1.0387809 & 0.0361673\\
-		\hline
-		$\gamma$ = 0.24 & 0.7069533 & 0.2023701 & 0.5215129 & 0.1714423 & 0.8747402 & 0.0586876 & -49.7565356 & 5300.3981000\\
-		\hline
-	\end{tabular}
-\end{footnotesize}
-\end{table}
-\begin{table}[H]
-
-	\caption{\label{tab:}$\rr$ Estimates and Std. Errs, SIRLMAT Model,
-		$X_0 = 99950, Y_0 = 50$, $\sigma_X = 100, \sigma_Y = 5$, $\beta = 0.06$}
-	\centering
-	\begin{footnotesize}
-	\begin{tabular}[t]{l|r|r|r|r|r|r|r|r}
-		\hline
-		Data Set & Norm Est & Norm SE & Norm-M Est & Norm-M SE & AR Est & AR SE & AR-M Est & AR-M SE\\
-		\hline
-		$\gamma$ = 0.001 & -2.9530478 & 200.4018824 & 8.0337261 & 106.7435722 & 78.891674 & 488.2036231 & -72.3228279 & 1659.2071496\\
-		\hline
-		$\gamma$ = 0.04 & 1.4367115 & 0.7496644 & 1.4714250 & 3.0708882 & 1.295474 & 0.3349151 & 1.2785848 & 0.2451543\\
-		\hline
-		$\gamma$ = 0.06 & 0.9853118 & 0.1423857 & 0.9943726 & 0.0160834 & 1.007444 & 0.5983482 & 1.0258649 & 0.5010848\\
-		\hline
-		$\gamma$ = 0.24 & 0.6742069 & 6.2426393 & 0.3870527 & 4.8804441 & 1.785998 & 15.9375643 & 0.2220084 & 5.4085242\\
-		\hline
-	\end{tabular}
-\end{footnotesize}
-\end{table}
-\begin{table}[H]
-
-	\caption{\label{tab:}$\rr$ Estimates and Std. Errs, SIRIPR Model,
-		$X_0 = 99950, Y_0 = 50$, $\sigma_X = 100, \sigma_Y = 5$, $\beta = 0.06$}
-	\centering
-	\begin{footnotesize}
-	\begin{tabular}[t]{l|r|r|r|r|r|r|r|r}
-		\hline
-		Data Set & Norm Est & Norm SE & Norm-M Est & Norm-M SE & AR Est & AR SE & AR-M Est & AR-M SE\\
-		\hline
-		$\gamma$ = 0.001 & 1.082930 & 4.005191 & 0.4881735 & 1.323335 & 1.174475 & 4.399781 & 0.5704615 & 2.418576\\
-		\hline
-		$\gamma$ = 0.04 & 3.817206 & 10.282636 & 1.1102025 & 2.104624 & 2.891003 & 6.829224 & 1.1172480 & 2.272797\\
-		\hline
-		$\gamma$ = 0.06 & 23.793692 & 34.952641 & 1.3460001 & 5.280598 & 18.781262 & 28.247185 & 1.3502216 & 5.880026\\
-		\hline
-		$\gamma$ = 0.24 & 868.875851 & 3562.122383 & 32.0292478 & 431.693917 & 711.882081 & 3552.087545 & 3.9097302 & 56.980358\\
-		\hline
-	\end{tabular}
-\end{footnotesize}
-\end{table}
-\begin{table}[H]
-
-	\caption{\label{tab:}$\rr$ Estimates and Std. Errs, SIRSIPR Model,
-		$X_0 = 99950, Y_0 = 50$, $\sigma_X = 100, \sigma_Y = 5$, $\beta = 0.06$}
-	\centering
-	\begin{footnotesize}
-	\begin{tabular}[t]{l|r|r|r|r|r|r|r|r}
-		\hline
-		Data Set & Norm Est & Norm SE & Norm-M Est & Norm-M SE & AR Est & AR SE & AR-M Est & AR-M SE\\
-		\hline
-		$\gamma$ = 0.001 & 0.0792904 & 0.0000000 & 0.0765708 & 0.0000000 & 0.0781112 & 0.0000000 & 0.0742256 & 0.0000000\\
-		\hline
-		$\gamma$ = 0.04 & 1.3960820 & 0.0000000 & 2.3028451 & 0.0000000 & 54.9912000 & 0.0000000 & 0.9972600 & 0.0000000\\
-		\hline
-		$\gamma$ = 0.06 & 1.1903748 & 0.0670862 & 1.1910744 & 0.0357340 & 1.0461934 & 0.1047191 & 1.2352948 & 0.0612702\\
-		\hline
-		$\gamma$ = 0.24 & -1.1924562 & 0.5566890 & -0.1523352 & 0.1610466 & -13.7485924 & 0.0000000 & 3.5354100 & 0.3241589\\
-		\hline
-	\end{tabular}
-\end{footnotesize}
-\end{table}
-\begin{table}[H]
-
-	\caption{\label{tab:}$\rr$ Estimates and Std. Errs, LOGLINEAR Model,
-		$X_0 = 99950, Y_0 = 50$, $\sigma_X = 100, \sigma_Y = 5$, $\beta = 0.06$}
-	\centering
-	\begin{footnotesize}
-	\begin{tabular}[t]{l|r|r|r|r|r|r|r|r}
-		\hline
-		Data Set & Norm Est & Norm SE & Norm-M Est & Norm-M SE & AR Est & AR SE & AR-M Est & AR-M SE\\
-		\hline
-		$\gamma$ = 0.001 & 46.3209595 & 0.7797216 & 59.8300371 & 0.1079267 & 41.897714 & 0.9767785 & 70.7075929 & 0.9517239\\
-		\hline
-		$\gamma$ = 0.04 & 1.5002728 & 0.0001905 & 1.5000436 & 0.0000928 & 1.499516 & 0.0002653 & 1.5000445 & 0.0001481\\
-		\hline
-		$\gamma$ = 0.06 & 0.9989449 & 0.0007428 & 0.9994831 & 0.0003320 & 1.000275 & 0.0010605 & 1.0008545 & 0.0004910\\
-		\hline
-		$\gamma$ = 0.24 & 0.9912397 & 0.0030505 & 0.8623034 & 0.0087543 & 1.001367 & 0.0025488 & 0.9198973 & 0.0052616\\
-		\hline
-	\end{tabular}
-\end{footnotesize}
-\end{table}
-\begin{table}[H]
-
-	\caption{\label{tab:}$\rr$ Estimates and Std. Errs, SIRMC Model,
-		$X_0 = 99950, Y_0 = 50$, $\sigma_X = 100, \sigma_Y = 5$, $\beta = 0.06$}
-	\centering
-	\begin{footnotesize}
-	\begin{tabular}[t]{l|r|r|r|r|r|r|r|r}
-		\hline
-		Data Set & Norm Est & Norm SE & Norm-M Est & Norm-M SE & AR Est & AR SE & AR-M Est & AR-M SE\\
-		\hline
-		$\gamma$ = 0.001 & 0.9485554 & 1.0e-07 & 0.9485555 & 0.0e+00 & 0.9485555 & 0.00e+00 & 0.9485555 & 0e+00\\
-		\hline
-		$\gamma$ = 0.04 & 0.9485555 & 0.0e+00 & 0.9485555 & 0.0e+00 & 0.9485555 & 0.00e+00 & 0.9485555 & 0e+00\\
-		\hline
-		$\gamma$ = 0.06 & 0.9485555 & 2.0e-07 & 0.9485555 & 0.0e+00 & 0.9485555 & 2.00e-07 & 0.9485555 & 0e+00\\
-		\hline
-		$\gamma$ = 0.24 & 0.9485812 & 1.5e-05 & 0.9485605 & 1.4e-06 & 0.9486128 & 1.26e-05 & 0.9485558 & 2e-07\\
-		\hline
-	\end{tabular}
-\end{footnotesize}
-\end{table}
 \begin{table}[H]
@@
 
-\subsection{Temporary Hold: Error Change}
-
-\begin{table}[H]
-	\caption{\label{tab:}$\rr$ Estimates and Std. Errs, SIRLS Model,
-		$X_0 = 99950, Y_0 = 50$, $\beta = 0.06, \gamma = 0.03$}
-	\centering
-	\begin{footnotesize}
-	\begin{tabular}[t]{l|r|r|r|r|r|r|r|r}
-		\hline
-		Data Set & Norm Est & Norm SE & Norm-M Est & Norm-M SE & AR Est & AR SE & AR-M Est & AR-M SE\\
-		\hline
-		$\sigma_X = 50, \sigma_Y = 2$ & 1.993455 & 0.5784530 & 1.979966 & 0.1204126 & 2.033934 & 0.7559710 & 2.066706 & 0.3382834\\
-		\hline
-		$\sigma_X = 500, \sigma_Y = 20$ & 1.998088 & 0.1353803 & 1.996950 & 0.0541332 & 1.999686 & 0.1893888 & 2.002135 & 0.0927952\\
-		\hline
-		$\sigma_X = 2500, \sigma_Y = 100$ & 1.999755 & 0.0026502 & 1.999834 & 0.0026397 & 1.999555 & 0.0037664 & 1.999038 & 0.0036473\\
-		\hline
-		$\sigma_X = 10000, \sigma_Y = 500$ & 2.001870 & 0.0259428 & 1.997748 & 0.0162536 & 2.002792 & 0.0416203 & 1.997918 & 0.0257997\\
-		\hline
-	\end{tabular}
-\end{footnotesize}
-\end{table}
-\begin{table}[H]
-
-	\caption{\label{tab:}$\rr$ Estimates and Std. Errs, SIRREPAR Model,
-		$X_0 = 99950, Y_0 = 50$, $\beta = 0.06, \gamma = 0.03$}
-	\centering
-	\begin{footnotesize}
-	\begin{tabular}[t]{l|r|r|r|r|r|r|r|r}
-		\hline
-		Data Set & Norm Est & Norm SE & Norm-M Est & Norm-M SE & AR Est & AR SE & AR-M Est & AR-M SE\\
-		\hline
-		$\sigma_X = 50, \sigma_Y = 2$ & 1.993660 & 0.5115266 & 1.979973 & 0.1053955 & 2.033636 & 0.7059701 & 2.066476 & 0.3172823\\
-		\hline
-		$\sigma_X = 500, \sigma_Y = 20$ & 1.998145 & 0.1207935 & 1.997147 & 0.0481545 & 2.000194 & 0.1680126 & 2.002244 & 0.0826643\\
-		\hline
-		$\sigma_X = 2500, \sigma_Y = 100$ & 1.999829 & 0.0023646 & 1.999946 & 0.0023591 & 1.999885 & 0.0033514 & 1.999483 & 0.0032200\\
-		\hline
-		$\sigma_X = 10000, \sigma_Y = 500$ & 2.002305 & 0.0232636 & 1.997996 & 0.0144744 & 2.003065 & 0.0372826 & 1.997732 & 0.0230069\\
-		\hline
-	\end{tabular}
-\end{footnotesize}
-\end{table}
-\begin{table}[H]
-
-	\caption{\label{tab:}$\rr$ Estimates and Std. Errs, SIRLMA Model,
-		$X_0 = 99950, Y_0 = 50$, $\beta = 0.06, \gamma = 0.03$}
-	\centering
-	\begin{footnotesize}
-	\begin{tabular}[t]{l|r|r|r|r|r|r|r|r}
-		\hline
-		Data Set & Norm Est & Norm SE & Norm-M Est & Norm-M SE & AR Est & AR SE & AR-M Est & AR-M SE\\
-		\hline
-		$\sigma_X = 50, \sigma_Y = 2$ & 1.1308571 & 0.1012907 & 3.899869 & 4.3543097 & 0.6625391 & 0.2163070 & 1.2910663 & 1.0083954\\
-		\hline
-		$\sigma_X = 500, \sigma_Y = 20$ & 0.4627433 & 7.5197341 & 4.178743 & 65.7755318 & 1.0790398 & 0.0165707 & -0.1152434 & 1.8870865\\
-		\hline
-		$\sigma_X = 2500, \sigma_Y = 100$ & 2.0417370 & 0.2771862 & 1.629287 & 0.0865344 & 1.4661799 & 0.0749495 & 2.8537066 & 0.5542235\\
-		\hline
-		$\sigma_X = 10000, \sigma_Y = 500$ & 1.6650841 & 1.0679885 & 1.686458 & 0.2567413 & 1.1579492 & 0.0983696 & -4.9577781 & 32.2547059\\
-		\hline
-	\end{tabular}
-\end{footnotesize}
-\end{table}
-\begin{table}[H]
-
-	\caption{\label{tab:}$\rr$ Estimates and Std. Errs, SIRLMAT Model,
-		$X_0 = 99950, Y_0 = 50$, $\beta = 0.06, \gamma = 0.03$}
-	\centering
-	\begin{footnotesize}
-	\begin{tabular}[t]{l|r|r|r|r|r|r|r|r}
-		\hline
-		Data Set & Norm Est & Norm SE & Norm-M Est & Norm-M SE & AR Est & AR SE & AR-M Est & AR-M SE\\
-		\hline
-		$\sigma_X = 50, \sigma_Y = 2$ & 4.991704 & 73.3084024 & 1.471488 & 0.9852195 & 0.8769744 & 1.2146547 & 1.0159519 & 2.9569665\\
-		\hline
-		$\sigma_X = 500, \sigma_Y = 20$ & 1.281010 & 4.3716385 & 1.356436 & 3.2462324 & 1.2444787 & 1.6725599 & 0.9326791 & 4.3459096\\
-		\hline
-		$\sigma_X = 2500, \sigma_Y = 100$ & 1.416200 & 0.5785114 & 1.512228 & 2.1276407 & 1.4016766 & 0.7202478 & 1.4274038 & 0.6787529\\
-		\hline
-		$\sigma_X = 10000, \sigma_Y = 500$ & 1.357813 & 1.5667226 & 1.447490 & 0.7230416 & 1.0226014 & 3.5724900 & 1.9986845 & 15.7532083\\
-		\hline
-	\end{tabular}
-\end{footnotesize}
-\end{table}
-\begin{table}[H]
-
-	\caption{\label{tab:}$\rr$ Estimates and Std. Errs, SIRIPR Model,
-		$X_0 = 99950, Y_0 = 50$, $\beta = 0.06, \gamma = 0.03$}
-	\centering
-	\begin{footnotesize}
-	\begin{tabular}[t]{l|r|r|r|r|r|r|r|r}
-		\hline
-		Data Set & Norm Est & Norm SE & Norm-M Est & Norm-M SE & AR Est & AR SE & AR-M Est & AR-M SE\\
-		\hline
-		$\sigma_X = 50, \sigma_Y = 2$ & 143.885460 & 778.881724 & 2.7003338 & 35.642540 & 55.265406 & 163.492483 & 2.1210664 & 13.550223\\
-		\hline
-		$\sigma_X = 500, \sigma_Y = 20$ & 48.894100 & 209.241051 & 1.3602975 & 10.234817 & 117.222483 & 870.624089 & 1.7544551 & 11.060645\\
-		\hline
-		$\sigma_X = 2500, \sigma_Y = 100$ & 1.656451 & 4.879121 & 0.8705396 & 1.231344 & 1.318824 & 3.274114 & 0.8679029 & 1.309243\\
-		\hline
-		$\sigma_X = 10000, \sigma_Y = 500$ & 10.426754 & 42.824909 & 1.1338742 & 4.906623 & 7.750396 & 24.637815 & 1.4241606 & 12.589987\\
-		\hline
-	\end{tabular}
-\end{footnotesize}
-\end{table}
-\begin{table}[H]
-
-	\caption{\label{tab:}$\rr$ Estimates and Std. Errs, SIRSIPR Model,
-		$X_0 = 99950, Y_0 = 50$, $\beta = 0.06, \gamma = 0.03$}
-	\centering
-	\begin{footnotesize}
-	\begin{tabular}[t]{l|r|r|r|r|r|r|r|r}
-		\hline
-		Data Set & Norm Est & Norm SE & Norm-M Est & Norm-M SE & AR Est & AR SE & AR-M Est & AR-M SE\\
-		\hline
-		$\sigma_X = 50, \sigma_Y = 2$ & -1.5215678 & 0 & 0.4256483 & 0 & 0.5346892 & 0 & -0.6623638 & 0\\
-		\hline
-		$\sigma\_X = 500, \sigma_Y = 20$ & 0.6363657 & 0 & 0.7407530 & 0 & 1.0216527 & 0 & 1.5576959 & 0\\
-		\hline
-		$\sigma\_X = 2500, \sigma_Y = 100$ & 1.1020944 & 0 & 1.0666890 & 0 & 1.0562382 & 0 & 1.1053737 & 0\\
-		\hline
-		$\sigma\_X = 10000, \sigma_Y = 500$ & 1.9158393 & 0 & 0.7926880 & 0 & 0.8105835 & 0 & 0.8006750 & 0\\
-		\hline
-	\end{tabular}
-\end{footnotesize}
-\end{table}
-\begin{table}[H]
-
-	\caption{\label{tab:}$\rr$ Estimates and Std. Errs, LOGLINEAR Model,
-		$X_0 = 99950, Y_0 = 50$, $\beta = 0.06, \gamma = 0.03$}
-	\centering
-	\begin{footnotesize}
-	\begin{tabular}[t]{l|r|r|r|r|r|r|r|r}
-		\hline
-		Data Set & Norm Est & Norm SE & Norm-M Est & Norm-M SE & AR Est & AR SE & AR-M Est & AR-M SE\\
-		\hline
-		$\sigma_X = 50, \sigma_Y = 2$ & 2.059934 & 0.0246402 & 1.995133 & 0.0038894 & 2.209575 & 0.0439100 & 2.545819 & 0.0568537\\
-		\hline
-		$\sigma\_X = 500, \sigma_Y = 20$ & 2.001882 & 0.0054085 & 1.995742 & 0.0015419 & 2.038646 & 0.0081562 & 2.015675 & 0.0031322\\
-		\hline
-		$\sigma\_X = 2500, \sigma_Y = 100$ & 1.999891 & 0.0001037 & 2.000036 & 0.0000930 & 1.999143 & 0.0001394 & 1.999749 & 0.0001328\\
-		\hline
-		$\sigma\_X = 10000, \sigma_Y = 500$ & 2.003402 & 0.0010320 & 1.998765 & 0.0003682 & 2.003763 & 0.0015505 & 1.999829 & 0.0008545\\
-		\hline
-	\end{tabular}
-\end{footnotesize}
-\end{table}
-\begin{table}[H]
-
-	\caption{\label{tab:}$\rr$ Estimates and Std. Errs, SIRMC Model,
-		$X_0 = 99950, Y_0 = 50$, $\beta = 0.06, \gamma = 0.03$}
-	\centering
-	\begin{footnotesize}
-	\begin{tabular}[t]{l|r|r|r|r|r|r|r|r}
-		\hline
-		Data Set & Norm Est & Norm SE & Norm-M Est & Norm-M SE & AR Est & AR SE & AR-M Est & AR-M SE\\
-		\hline
-		$\sigma_X = 50, \sigma_Y = 2$ & 0.9485506 & 3.8e-06 & 0.9485556 & 2e-07 & 0.9485428 & 2.6e-06 & 0.9485538 & 8e-07\\
-		\hline
-		$\sigma\_X = 500, \sigma_Y = 20$ & 0.9485562 & 1.0e-06 & 0.9485554 & 1e-07 & 0.9485640 & 3.6e-06 & 0.9485554 & 1e-07\\
-		\hline
-		$\sigma\_X = 2500, \sigma_Y = 100$ & 0.9485555 & 0.0e+00 & 0.9485555 & 0e+00 & 0.9485555 & 0.0e+00 & 0.9485555 & 0e+00\\
-		\hline
-		$\sigma\_X = 10000, \sigma_Y = 500$ & 0.9485554 & 2.0e-07 & 0.9485555 & 0e+00 & 0.9485554 & 1.0e-07 & 0.9485555 & 1e-07\\
-		\hline
-	\end{tabular}
-\end{footnotesize}
-\end{table}
+\begin{table}[H]
+
+	\caption{\label{tab:}$\rr$ Estimates and Std. Errs, SEQBAYES Model,
+		$\sigma_X = 100, \sigma_Y = 5$, $\beta = 0.06, \gamma = 0.03$}
+	\centering
+	\begin{footnotesize}
+		\begin{tabular}[t]{l|r|r|r|r|r|r|r|r}
+			\hline
+			Data Set & Norm Est & Norm SE & Norm-M Est & Norm-M SE & AR Est & AR SE & AR-M Est & AR-M SE\\
+			\hline
+			$X_0 = 99000, Y_0 = 1000$ & 0.9728534 & 0.0516270 & 1.052579 & 0.0537008 & 0.5850802 & 0.0400370 & 0.6236582 & 0.0413358\\
+			\hline
+			$X_0 = 99900, Y_0 = 100$ & 1.0000000 & 0.0523424 & 1.000000 & 0.0523424 & 1.4175193 & 0.0623186 & 1.4017333 & 0.0619707\\
+			\hline
+			$X_0 = 99990, Y_0 = 10$ & 1.3877095 & 0.0616599 & 1.430888 & 0.0626118 & 1.3948200 & 0.0618177 & 1.4507203 & 0.0630442\\
+			\hline
+			$X_0 = 99999, Y_0 = 1$ & 1.4714813 & 0.0634938 & 1.465728 & 0.0633695 & 1.4665442 & 0.0633871 & 1.4723265 & 0.0635120\\
+			\hline
+		\end{tabular}
+	\end{footnotesize}
+\end{table}
+
 \begin{table}[H]
@@
 
-\subsection{Temporary Hold: Start Change}
-
-\begin{table}[H]
-
-	\caption{\label{tab:}$\rr$ Estimates and Std. Errs, SIRLS Model,
-		$\sigma_X = 100, \sigma_Y = 5$, $\beta = 0.06, \gamma = 0.03$}
-	\centering
-	\begin{footnotesize}
-	\begin{tabular}[t]{l|r|r|r|r|r|r|r|r}
-		\hline
-		Data Set & Norm Est & Norm SE & Norm-M Est & Norm-M SE & AR Est & AR SE & AR-M Est & AR-M SE\\
-		\hline
-		$X_0 = 99000, Y_0 = 1000$ & 1.999588 & 0.0049609 & 2.000284 & 0.0046294 & 2.000461 & 0.0078398 & 2.000122 & 0.0068626\\
-		\hline
-		$X_0 = 99900, Y_0 = 100$ & 2.022501 & 0.3064263 & 2.021865 & 0.3064148 & 2.022296 & 0.3063477 & 2.022665 & 0.3068512\\
-		\hline
-		$X_0 = 99990, Y_0 = 10$ & 2.116865 & 0.6626380 & 2.116551 & 0.6621184 & 2.116959 & 0.6623142 & 2.116803 & 0.6620448\\
-		\hline
-		$X_0 = 99999, Y_0 = 1$ & 2.468304 & 1.5370731 & 2.466776 & 1.5348981 & 2.470607 & 1.5371890 & 2.466700 & 1.5348542\\
-		\hline
-	\end{tabular}
-\end{footnotesize}
-\end{table}
-\begin{table}[H]
-
-	\caption{\label{tab:}$\rr$ Estimates and Std. Errs, SIRREPAR Model,
-		$\sigma_X = 100, \sigma_Y = 5$, $\beta = 0.06, \gamma = 0.03$}
-	\centering
-	\begin{footnotesize}
-	\begin{tabular}[t]{l|r|r|r|r|r|r|r|r}
-		\hline
-		Data Set & Norm Est & Norm SE & Norm-M Est & Norm-M SE & AR Est & AR SE & AR-M Est & AR-M SE\\
-		\hline
-		$X_0 = 99000, Y_0 = 1000$ & 1.999712 & 0.0034303 & 2.000274 & 0.0032086 & 2.000082 & 0.0053846 & 2.000014 & 0.0047509\\
-		\hline
-		$X_0 = 99900, Y_0 = 100$ & 2.022359 & 0.2698034 & 2.021866 & 0.2694871 & 2.022178 & 0.2696002 & 2.022530 & 0.2701860\\
-		\hline
-		$X_0 = 99990, Y_0 = 10$ & 2.117059 & 0.7291460 & 2.116858 & 0.7290133 & 2.116829 & 0.7288841 & 2.116534 & 0.7286543\\
-		\hline
-		$X_0 = 99999, Y_0 = 1$  & 2.469189 & 2.7293667 & 2.466967 & 2.7219457 & 2.470707 & 2.7245397 & 2.466483 & 2.7349248\\
-		\hline
-	\end{tabular}
-\end{footnotesize}
-\end{table}
-\begin{table}[H]
-
-	\caption{\label{tab:}$\rr$ Estimates and Std. Errs, SIRLMA Model,
-		$\sigma_X = 100, \sigma_Y = 5$, $\beta = 0.06, \gamma = 0.03$}
-	\centering
-	\begin{footnotesize}
-	\begin{tabular}[t]{l|r|r|r|r|r|r|r|r}
-		\hline
-		Data Set & Norm Est & Norm SE & Norm-M Est & Norm-M SE & AR Est & AR SE & AR-M Est & AR-M SE\\
-		\hline
-		$X_0 = 99000, Y_0 = 1000$ & -87.5945650 & 66413.128046 & -0.9999761 & 1.422640 & 4.1817846 & 3.8812899 & -10.8458320 & 288.1028935\\
-		\hline
-		$X_0 = 99900, Y_0 = 100$ & 0.5237886 & 1.632971 & 0.2436375 & 3.754091 & 0.7149415 & 0.9374275 & -0.3051038 & 8.8837193\\
-		\hline
-		$X_0 = 99990, Y_0 = 10$ & -33.0051124 & 433.594227 & -1.3382741 & 3.047220 & -0.2463189 & 1.5875310 & 0.4368556 & 0.7686276\\
-		\hline
-		$X_0 = 99999, Y_0 = 1$ & 10.9282204 & 19.297359 & -98.6422337 & 3113.332617 & -0.8194023 & 1.7121668 & 16.6102769 & 18.1045393\\
-		\hline
-	\end{tabular}
-\end{footnotesize}
-\end{table}
-\begin{table}[H]
-
-	\caption{\label{tab:}$\rr$ Estimates and Std. Errs, SIRLMAT Model,
-		$\sigma_X = 100, \sigma_Y = 5$, $\beta = 0.06, \gamma = 0.03$}
-	\centering
-	\begin{footnotesize}
-	\begin{tabular}[t]{l|r|r|r|r|r|r|r|r}
-		\hline
-		Data Set & Norm Est & Norm SE & Norm-M Est & Norm-M SE & AR Est & AR SE & AR-M Est & AR-M SE\\
-		\hline
-		$X_0 = 99000, Y_0 = 1000$ & 0.730610 & 4.6810416 & 0.9720592 & 0.6586069 & 3.163816 & 45.776770 & 0.9367865 & 0.8857030\\
-		\hline
-		$X_0 = 99900, Y_0 = 100$ & 1.236415 & 0.7388933 & 1.1744675 & 1.3957615 & 1.249219 & 0.813166 & 1.1960662 & 0.6589163\\
-		\hline
-		$X_0 = 99990, Y_0 = 10$ & 1.459703 & 10.9219088 & 0.8531016 & 12.5573016 & -3.846910 & 71.297410 & -0.5884883 & 39.4915091\\
-		\hline
-		$X_0 = 99999, Y_0 = 1$ & 1.624645 & 5.0359845 & 0.4174208 & 12.8380228 & 1.301346 & 25.850056 & 3.1746775 & 12.8142151\\
-		\hline
-	\end{tabular}
-\end{footnotesize}
-\end{table}
-\begin{table}[H]
-
-	\caption{\label{tab:}$\rr$ Estimates and Std. Errs, SIRIPR Model,
-		$\sigma_X = 100, \sigma_Y = 5$, $\beta = 0.06, \gamma = 0.03$}
-	\centering
-	\begin{footnotesize}
-	\begin{tabular}[t]{l|r|r|r|r|r|r|r|r}
-		\hline
-		Data Set & Norm Est & Norm SE & Norm-M Est & Norm-M SE & AR Est & AR SE & AR-M Est & AR-M SE\\
-		\hline
-		$X_0 = 99000, Y_0 = 1000$ & 0.7710423 & 0.8645882 & 0.5919617 & 0.6342256 & 0.7161564 & 0.7174921 & 0.5976212 & 0.5992238\\
-		\hline
-		$X_0 = 99900, Y_0 = 100$ & 1.7659585 & 4.6419825 & 0.8824341 & 2.0203472 & 1.3381463 & 2.9156688 & 0.8541963 & 1.6873018\\
-		\hline
-		$X_0 = 99990, Y_0 = 10$ & 11.1978083 & 40.2941067 & 1.1073477 & 3.2407038 & 10.0438581 & 43.1896138 & 0.9850750 & 2.6487228\\
-		\hline
-		$X_0 = 99999, Y_0 = 1$ & 464.5647956 & 4494.6030482 & 2.1636551 & 15.9245640 & 137.5104852 & 719.4933920 & 1.5049398 & 7.7183465\\
-		\hline
-	\end{tabular}
-\end{footnotesize}
-\end{table}
-\begin{table}[H]
-
-	\caption{\label{tab:}$\rr$ Estimates and Std. Errs, SIRSIPR Model,
-		$\sigma_X = 100, \sigma_Y = 5$, $\beta = 0.06, \gamma = 0.03$}
-	\centering
-	\begin{footnotesize}
-	\begin{tabular}[t]{l|r|r|r|r|r|r|r|r}
-		\hline
-		Data Set & Norm Est & Norm SE & Norm-M Est & Norm-M SE & AR Est & AR SE & AR-M Est & AR-M SE\\
-		\hline
-		$X_0 = 99000, Y_0 = 1000$ & 1.2522106 & 34.7537893 & 1.2099648 & 34.6892205 & 1.3712547 & 34.4629789 & 1.1533154 & 34.2274520\\
-		\hline
-		$X_0 = 99900, Y_0 = 100$ & -0.5284892 & 0.0000000 & -0.6902358 & 0.0000000 & -3.9381962 & 0.0000000 & -1.0663894 & 0.0000000\\
-		\hline
-		$X_0 = 99990, Y_0 = 10$ & 0.8071191 & 0.1942094 & 0.8129547 & 0.1895219 & 0.8287186 & 0.1855934 & 0.8070414 & 0.1954849\\
-		\hline
-		$X_0 = 99999, Y_0 = 1$ & 1.4634454 & 0.0000000 & 3.2469490 & 0.0000000 & 0.7228135 & 0.0000000 & 0.6213745 & 0.0000000\\
-		\hline
-	\end{tabular}
-\end{footnotesize}
-\end{table}
-\begin{table}[H]
-
-	\caption{\label{tab:}$\rr$ Estimates and Std. Errs, LOGLINEAR Model,
-		$\sigma_X = 100, \sigma_Y = 5$, $\beta = 0.06, \gamma = 0.03$}
-	\centering
-	\begin{footnotesize}
-	\begin{tabular}[t]{l|r|r|r|r|r|r|r|r}
-		\hline
-		Data Set & Norm Est & Norm SE & Norm-M Est & Norm-M SE & AR Est & AR SE & AR-M Est & AR-M SE\\
-		\hline
-		$X_0 = 99000, Y_0 = 1000$ & 1.999515 & 0.0002828 & 2.000144 & 0.0001594 & 1.999753 & 0.0004197 & 2.000586 & 0.0003024\\
-		\hline
-		$X_0 = 99900, Y_0 = 100$ & 2.000011 & 0.0002315 & 1.999732 & 0.0001526 & 2.000473 & 0.0002776 & 2.000640 & 0.0002656\\
-		\hline
-		$X_0 = 99990, Y_0 = 10$ & 1.999886 & 0.0002517 & 1.999599 & 0.0001361 & 2.000608 & 0.0003289 & 1.999049 & 0.0002664\\
-		\hline
-		$X_0 = 99999, Y_0 = 1$ & 1.999690 & 0.0007731 & 2.000503 & 0.0003700 & 2.002716 & 0.0010572 & 1.999919 & 0.0005663\\
-		\hline
-	\end{tabular}
-\end{footnotesize}
-\end{table}
-\begin{table}[H]
-
-	\caption{\label{tab:}$\rr$ Estimates and Std. Errs, SIRMC Model,
-		$\sigma_X = 100, \sigma_Y = 5$, $\beta = 0.06, \gamma = 0.03$}
-	\centering
-	\begin{footnotesize}
-	\begin{tabular}[t]{l|r|r|r|r|r|r|r|r}
-		\hline
-		Data Set & Norm Est & Norm SE & Norm-M Est & Norm-M SE & AR Est & AR SE & AR-M Est & AR-M SE\\
-		\hline
-		$X_0 = 99000, Y_0 = 1000$ & 0.9485555 & 0.00e+00 & 0.9485555 & 0 & 0.9485555 & 0.0e+00 & 0.9485555 & 0\\
-		\hline
-		$X_0 = 99900, Y_0 = 100$ & 0.9485555 & 0.00e+00 & 0.9485555 & 0 & 0.9485555 & 0.0e+00 & 0.9485555 & 0\\
-		\hline
-		$X_0 = 99990, Y_0 = 10$ & 0.9485552 & 2.00e-07 & 0.9485555 & 0 & 0.9485558 & 2.0e-07 & 0.9485555 & 0\\
-		\hline
-		$X_0 = 99999, Y_0 = 1$ & 0.9486030 & 1.56e-05 & 0.9485557 & 0 & 0.9485701 & 2.6e-06 & 0.9485555 & 0\\
-		\hline
-	\end{tabular}
-\end{footnotesize}
-\end{table}
-\begin{table}[H]
-
-	\caption{\label{tab:}$\rr$ Estimates and Std. Errs, SEQBAYES Model,
-		$\sigma_X = 100, \sigma_Y = 5$, $\beta = 0.06, \gamma = 0.03$}
-	\centering
-	\begin{footnotesize}
-	\begin{tabular}[t]{l|r|r|r|r|r|r|r|r}
-		\hline
-		Data Set & Norm Est & Norm SE & Norm-M Est & Norm-M SE & AR Est & AR SE & AR-M Est & AR-M SE\\
-		\hline
-		$X_0 = 99000, Y_0 = 1000$ & 0.9728534 & 0.0516270 & 1.052579 & 0.0537008 & 0.5850802 & 0.0400370 & 0.6236582 & 0.0413358\\
-		\hline
-		$X_0 = 99900, Y_0 = 100$ & 1.0000000 & 0.0523424 & 1.000000 & 0.0523424 & 1.4175193 & 0.0623186 & 1.4017333 & 0.0619707\\
-		\hline
-		$X_0 = 99990, Y_0 = 10$ & 1.3877095 & 0.0616599 & 1.430888 & 0.0626118 & 1.3948200 & 0.0618177 & 1.4507203 & 0.0630442\\
-		\hline
-		$X_0 = 99999, Y_0 = 1$ & 1.4714813 & 0.0634938 & 1.465728 & 0.0633695 & 1.4665442 & 0.0633871 & 1.4723265 & 0.0635120\\
-		\hline
-	\end{tabular}
-\end{footnotesize}
-\end{table}
-
-\subsection{Temporary Hold: Other Models}
-
-\begin{table}[H]
-
-	\caption{\label{tab:}$\rr$ Estimates and Std. Errs, SIRLS Model,
-		$X_0 = 99950, Y_0 = 50$, $\sigma_X = 100, \sigma_Y = 5$}
-	\centering
-	\begin{footnotesize}
-	\begin{tabular}[t]{l|r|r|r|r|r|r|r|r}
-		\hline
-		Data Set & Norm Est & Norm SE & Norm-M Est & Norm-M SE & AR Est & AR SE & AR-M Est & AR-M SE\\
-		\hline
-		1st Order & 1.051438 & 0.3782910 & 1.051420 & 0.3784628 & 1.051864 & 0.3775363 & 1.051293 & 0.3782570\\
-		\hline
-		4th Order & 1.769347 & 0.4226615 & 1.769136 & 0.4215768 & 1.769002 & 0.4227615 & 1.769136 & 0.4215910\\
-		\hline
-		Linear SIR & 1.274258 & 0.0979741 & 1.273917 & 0.0976920 & 1.273654 & 0.1000387 & 1.274007 & 0.0970809\\
-		\hline
-	\end{tabular}
-\end{footnotesize}
-\end{table}
-\begin{table}[H]
-
-	\caption{\label{tab:}$\rr$ Estimates and Std. Errs, SIRREPAR Model,
-		$X_0 = 99950, Y_0 = 50$, $\sigma_X = 100, \sigma_Y = 5$}
-	\centering
-	\begin{footnotesize}
-	\begin{tabular}[t]{l|r|r|r|r|r|r|r|r}
-		\hline
-		Data Set & Norm Est & Norm SE & Norm-M Est & Norm-M SE & AR Est & AR SE & AR-M Est & AR-M SE\\
-		\hline
-		1st Order & 1.051435 & 0.0259416 & 1.051397 & 0.0258690 & 1.051845 & 0.0260146 & 1.051255 & 0.0257166\\
-		\hline
-		4th Order & 1.769331 & 0.2796903 & 1.769079 & 0.2791191 & 1.769095 & 0.2799270 & 1.769111 & 0.2788789\\
-		\hline
-		Linear SIR & 1.274314 & 0.0339775 & 1.274075 & 0.0337499 & 1.274174 & 0.0345629 & 1.273799 & 0.0334192\\
-		\hline
-	\end{tabular}
-\end{footnotesize}
-\end{table}
-\begin{table}[H]
-
-	\caption{\label{tab:}$\rr$ Estimates and Std. Errs, SIRLMA Model,
-		$X_0 = 99950, Y_0 = 50$, $\sigma_X = 100, \sigma_Y = 5$}
-	\centering
-	\begin{footnotesize}
-	\begin{tabular}[t]{l|r|r|r|r|r|r|r|r}
-		\hline
-		Data Set & Norm Est & Norm SE & Norm-M Est & Norm-M SE & AR Est & AR SE & AR-M Est & AR-M SE\\
-		\hline
-		1st Order & 1.2559522 & 0.1296269 & 1.2209276 & 0.0597736 & 1.2670255 & 0.1588846 & 1.2476445 & 0.1150267\\
-		\hline
-		4th Order & 0.3632202 & 0.1230895 & -0.0201429 & 0.1133200 & 0.3580085 & 0.1258904 & -0.0161666 & 0.1103940\\
-		\hline
-		Linear SIR & 1.2191443 & 0.1093267 & 1.2067608 & 0.0480986 & 1.2146478 & 0.0627181 & 0.9503153 & 0.1066437\\
-		\hline
-	\end{tabular}
-\end{footnotesize}
-\end{table}
-\begin{table}[H]
-
-	\caption{\label{tab:}$\rr$ Estimates and Std. Errs, SIRLMAT Model,
-		$X_0 = 99950, Y_0 = 50$, $\sigma_X = 100, \sigma_Y = 5$}
-	\centering
-	\begin{footnotesize}
-	\begin{tabular}[t]{l|r|r|r|r|r|r|r|r}
-		\hline
-		Data Set & Norm Est & Norm SE & Norm-M Est & Norm-M SE & AR Est & AR SE & AR-M Est & AR-M SE\\
-		\hline
-		1st Order & 1.1509447 & 0.1012249 & 1.138621 & 0.1738250 & 1.147184 & 0.0681524 & 1.145189 & 0.0811484\\
-		\hline
-		4th Order & 0.8246203 & 6.0710436 & 1.000597 & 2.5841000 & 3.255953 & 26.3713818 & 1.385674 & 8.5430506\\
-		\hline
-		Linear SIR & 1.2009801 & 0.0031100 & 1.200130 & 0.0040763 & 1.203159 & 0.0148085 & 1.195875 & 0.0246927\\
-		\hline
-	\end{tabular}
-\end{footnotesize}
-\end{table}
-\begin{table}[H]
-
-	\caption{\label{tab:}$\rr$ Estimates and Std. Errs, SIRIPR Model,
-		$X_0 = 99950, Y_0 = 50$, $\sigma_X = 100, \sigma_Y = 5$}
-	\centering
-	\begin{footnotesize}
-	\begin{tabular}[t]{l|r|r|r|r|r|r|r|r}
-		\hline
-		Data Set & Norm Est & Norm SE & Norm-M Est & Norm-M SE & AR Est & AR SE & AR-M Est & AR-M SE\\
-		\hline
-		1st Order & 2.699973 & 9.131131 & 2.297208 & 7.14196 & 4.334767 & 41.325146 & 4.194341 & 36.201142\\
-		\hline
-		4th Order & 6.260963 & 20.337163 & 5.924029 & 17.81136 & 6.268500 & 20.536192 & 5.892582 & 17.341763\\
-		\hline
-		Linear SIR & 1.516803 & 1.554241 & 1.208578 & 1.12544 & 1.379871 & 1.311406 & 1.208521 & 1.003065\\
-		\hline
-	\end{tabular}
-\end{footnotesize}
-\end{table}
-\begin{table}[H]
-
-	\caption{\label{tab:}$\rr$ Estimates and Std. Errs, SIRSIPR Model,
-		$X_0 = 99950, Y_0 = 50$, $\sigma_X = 100, \sigma_Y = 5$}
-	\centering
-	\begin{footnotesize}
-	\begin{tabular}[t]{l|r|r|r|r|r|r|r|r}
-		\hline
-		Data Set & Norm Est & Norm SE & Norm-M Est & Norm-M SE & AR Est & AR SE & AR-M Est & AR-M SE\\
-		\hline
-		1st Order & 58.836635 & 0.0000000 & 27.429685 & 0.0000000 & 1.498805 & 0.0000000 & -14.135196 & 0.0000000\\
-		\hline
-		4th Order & 2.848624 & 0.0000000 & 15.083621 & 0.0000000 & 2.766152 & 0.0000000 & 14.972580 & 0.0000000\\
-		\hline
-		Linear SIR & 1.294079 & 0.2151131 & 1.294364 & 0.2158095 & 1.296731 & 0.2215603 & 1.304489 & 0.2299876\\
-		\hline
-	\end{tabular}
-\end{footnotesize}
-\end{table}
-\begin{table}[H]
-
-	\caption{\label{tab:}$\rr$ Estimates and Std. Errs, LOGLINEAR Model,
-		$X_0 = 99950, Y_0 = 50$, $\sigma_X = 100, \sigma_Y = 5$}
-	\centering
-	\begin{footnotesize}
-	\begin{tabular}[t]{l|r|r|r|r|r|r|r|r}
-		\hline
-		Data Set & Norm Est & Norm SE & Norm-M Est & Norm-M SE & AR Est & AR SE & AR-M Est & AR-M SE\\
-		\hline
-		1st Order & 1.188050 & 0.0006341 & 1.187976 & 0.0006389 & 1.187553 & 0.0006273 & 1.187848 & 0.0006497\\
-		\hline
-		4th Order & 1.955180 & 0.0164608 & 1.955313 & 0.0164212 & 1.955130 & 0.0164623 & 1.955302 & 0.0164165\\
-		\hline
-		Linear SIR & 1.258587 & 0.0010079 & 1.258670 & 0.0010069 & 1.258920 & 0.0009938 & 1.258462 & 0.0010199\\
-		\hline
-	\end{tabular}
-\end{footnotesize}
-\end{table}
-\begin{table}[H]
-
-	\caption{\label{tab:}$\rr$ Estimates and Std. Errs, SIRMC Model,
-		$X_0 = 99950, Y_0 = 50$, $\sigma_X = 100, \sigma_Y = 5$}
-	\centering
-	\begin{footnotesize}
-	\begin{tabular}[t]{l|r|r|r|r|r|r|r|r}
-		\hline
-		Data Set & Norm Est & Norm SE & Norm-M Est & Norm-M SE & AR Est & AR SE & AR-M Est & AR-M SE\\
-		\hline
-		1st Order & 0.9485551 & 1e-07 & 0.9485540 & 4e-07 & 0.9485555 & 2e-07 & 0.9485553 & 2.0e-07\\
-		\hline
-		4th Order & 0.9485545 & 6e-07 & 0.9485533 & 1e-06 & 0.9485535 & 9e-07 & 0.9485527 & 1.2e-06\\
-		\hline
-		Linear SIR & 0.9485555 & 0e+00 & 0.9485555 & 0e+00 & 0.9485555 & 0e+00 & 0.9485555 & 0.0e+00\\
-		\hline
-	\end{tabular}
-\end{footnotesize}
-\end{table}
 \begin{table}[H]
@@
 	\end{footnotesize}
+\end{table}
+
+\begin{table}[H]
+
+	\caption{\label{tab:}$\rr$ Estimates and Std. Errs, SEQBAYES Model,
+		$\beta = 0.06, \gamma = 0.03$, $X_0 = 99950, Y_0 = 50$, $\sigma_X = 100, \sigma_Y = 5$}
+	\centering
+	\begin{tabular}[t]{l|r|r|r|r|r|r|r|r}
+		\hline
+		Data Set & Norm Est & Norm SE & Norm-M Est & Norm-M SE & AR Est & AR SE & AR-M Est & AR-M SE\\
+		\hline
+		T = 200 & 2.428938 & 0.1102030 & 2.444793 & 0.1105620 & 2.420992 & 0.1100225 & 2.353856 & 0.1084863\\
+		\hline
+		T = 100 & 1.000000 & 0.1000000 & 1.000000 & 0.1000000 & 3.363462 & 0.1833974 & 1.000000 & 0.1000000\\
+		\hline
+		T = 50 & 1.000000 & 0.1414214 & 5.598703 & 0.3346252 & 1.000000 & 0.1414214 & 5.390413 & 0.3283417\\
+		\hline
+		T = 20 & 1.000000 & 0.2236068 & 1.000000 & 0.2236068 & 11.379288 & 0.7542973 & 1.000000 & 0.2236068\\
+		\hline
+	\end{tabular}
 \end{table}

### 2018-07-03 paper/Version12/draft_v12.tex
finally on git

--- 2018-07-03 paper/Version12/draft_v12 (Shannon Gallagher's conflicted copy 2018-05-01).tex
@@
 \usepackage{booktabs}
+\usepackage{lscape}
 \usepackage[table,xcdraw]{xcolor}
@@
 Despite a clear definition of $\rr$, epidemiologists have struggled to create a standard  estimator for $\rr$  \citep{hethcote2000}.
-A major issue in estimating $\rr$ is that the quantity is a \textit{property of the model}, meaning that $\rr$ is dependent not only on the usual noise that comes with statistical modeling but also on a variety of assumptions on how researchers assume a disease is transmitted through a population.
-
-To elaborate on $\rr$ being the property of the model, we first need to introduce the concept of the ``SI-framework,'' which was pioneered by Kermack and McKendrick in the 1920s \citep{getz2006}.
+A major issue in estimating $\rr$ is that the quantity is a \textit{property of the model}, meaning that $\rr$ is dependent not only on the usual noise that comes with statistical modeling but also on a variety of assumptions on how researchers assume a disease is transmitted through a population \citep{diekmann2009}.
+
+To elaborate on $\rr$ being a property of the model, we first need to introduce the concept of the ``SI-framework,'' which was pioneered by Kermack and McKendrick in the 1920s \citep{getz2006}.
 Here, S and I are known as compartments where S stands for ``susceptible'' and I for ``infectious.'' Infectious disease models then specify how individuals move from the S compartment to the I compartment.
@@
 As such, we discuss techniques for estimating CIs for each of the different methods.
-These CI estimates are the delta method, block bootstrap, and the posterior distribution.
-
-In the following sections, we present a series of simulations in which we examine each of our $\rr$ estimation methods.
+These CI estimates are created using the delta method, the block bootstrap, and the posterior distribution.
+
+In the following later, we present a series of simulations in which we examine each of our $\rr$ estimation methods.
 Within these simulations, we vary the number of observations used to estimate $\rr$, infection and recovery rates, the total population size, initial SI ratio, and several assumptions about the noise.
 
-We find that...
-
-Finally, we discuss our recommendations in estimating $\rr$ for the SIR model, and we comment on how these recommendations extend to more complex models and how many of the problems with estimating $\rr$  become exacerbated with the increase in complexity.
+Finally, we discuss our recommendations for estimating $\rr$ for the SIR model, and we comment on how these recommendations extend to more complex models and how many of the problems with estimating $\rr$  become exacerbated with the increase in complexity.
 
@@
 In Section \ref{sec:ci}, we discuss how to estimate CIs for each of the \wxxsir methods.
-Following in Section \ref{sec:sim-res}, we present the results of our simulations.
-Finally, in Section \ref{sec:discussion}, we provide final analysis of estimating $\rr$ for the SIR model and offer comments on how these conclusions extend to more complex models.
+In Section \ref{sec:sim-res}, we present the results of our simulations.
+Finally, in Section \ref{sec:discussion}, we provide recommendations of estimating $\rr$ for the SIR model and offer comments on how these conclusions extend to more complex models.
 
@@
 For example, Gaussian error, a common assumption in statistics and machine learning, becomes highly suspect due to the non-symmetrical nature of the noise imposed by the monotonic $S$ and $R$ compartments.
-Without assuming Gaussian noise, it becomes much more difficult to guarantee properties such as unbiased and consist estimates of $\rr$.
+Without assuming Gaussian noise, it becomes much more difficult to guarantee properties such as unbiased and consistent estimates of $\rr$.
 
@@
 
-% A final difficulty in estimating $\rr$ arises from the data.
-Ideally, we would have counts of the number of susceptibles, infectious, and recovered at each time step.
-However, sometimes only the incidence of the disease is available, where we actually need the prevalence.
-Then, the number of susceptibles is back-estimated from the number of infectious which requires an estimate of the number of susceptibles, and even the US Census is known to be at least 10\% off in both directions of undercounting and overcounting (CITE).
-Finally, the distinction between susceptible and infectious and infectious and recovered may not be clear.
-As asymptomatic individuals may infect others and be mistaken to be in either the susceptible or infectious compartment.
-
-As a result of these difficulties, accurately estimating $\rr$ is a monumentous task.
-We overview \wxxsir methods introduced by a number of researchers used to solve some of the problems described above.
+Below in Table \ref{tab:r0-real-ex}, we provide examples of estimates $\rr$ from publications over the past three decades, spanning various locations over the world.
+We display estimates for HIV, Zika (ZIK-V), Ebola (EVD), seasonal influenza, and H1N1 influenza.
+These estimates of $\rr$ are not limited to those of SIR model, although most results come from this framework.
+Of these, Zika seems to have the largest estimated values of $\rr$, followed by Ebola, H1N1 influenza, seasonal influenza, and HIV.
+We see that African locations seem to have larger values of $\rr$ compared to the rest of the world, regardless of the disease.
+Some of the accompanying lower and upper bound intervals (not necessarily CIs) are extremely large such as the case of HIV in Uganda (Interval: [.12, 14.17]) where as some are extremely small such as for Ebola in Guinea [1.5, 1.52].
+We see that estimates for H1N1 influenza are generally higher than that of seasonal influenza, which is evidence that H1N1 was a bigger concern than the typical seasonal influenza cycle.
+The lowest possible value for $\rr$ reported is .12 and the largest is 14.17, both for HIV in Uganda.
+Within diseases, we see that estimates for $\rr$ seem to be similar, although typically do not lie in another's lower and upper bound, suggesting that there is great disparity in estimates of $\rr$.
+
+% Please add the following required packages to your document preamble:
+% \usepackage{booktabs}
+% \usepackage{graphicx}
+% \usepackage{lscape}
+\begin{landscape}
+\begin{table}[H]
+\centering
+\begin{tabular}{@{}lllrrrl@{}}
+\toprule
+\textbf{Disease}    & \textbf{Location}                & \textbf{Year}                 & \textbf{$\rr$} & \textbf{Lower} & \textbf{Upper} & \textbf{Source}                                                                                                                  \\ \midrule
+HIV                 & Belgium                          & 2004-2010                     & 1.10          & 1.09           & 1.16           & \cite{coelho2011}   \\
+HIV                 & Netherlands                      & 2004-2010                     & 1.11        & 1.10            & 1.18           & \cite{coelho2011}    \\
+HIV                 & Portugal                         & 2004-2010                     & 1.08        & 1.06           & 1.15           & \cite{coelho2011}   \\
+HIV                 & San Francisco, USA               & 1993                          & 5.00           & 2.00              & 8.00              & \cite{blower1994}    \\
+HIV                 & Uganda                           & 2002-2008                     & 3.31        & 0.12           & 14.17          & \cite{nsubuga2014}    \\ \hline
+ZIK-V                & French Polynesia                 & 2013                          & 2.75        & 2.53           & 2.98           & \cite{zhang2017}        \\
+ZIK-V                & French Polynesia                 & 2013-2014                     & 3.70         & 2.60            & 4.80            & \cite{kucharski2016}  \\
+ZIK-V                & Florida                          & 2016                          & 0.16        & 0.13           & 0.19           & \cite{dinh2016}       \\
+ZIK-V                & Colombia                         & 2015-2016                     & 2.50         & 3.00              & 3.60            & \cite{nishiura2016}    \\
+ZIK-V                & Barranquilla, Colombia           & 2015                          & 2.40         & 3.80            & 5.60            & \cite{towers2016}    \\
+ZIK-V                & Colombia                         & 2015                          & 1.42        & 2.56           & 3.83           & \cite{majumder2016}  \\ \hline
+EVD               & Western Africa                   & 2014                          & 1.78        & 1.60            & 2.00              & \cite{fisman2014}   \\
+EVD & Western Africa                   & 2014                          & 1.60         & 1.40            & 1.80            & \cite{towers2014}  \\
+EVD                 & Democratic Republic of the Congo & 1995                          & 2.70         & 1.90            & 2.80            & \cite{legrand2007}  \\
+EVD                 & Uganda                           & 2000                          & 2.70         & 2.50            & 4.10            & \cite{legrand2007}  \\
+EVD                 & Guinea                           & 2014                          & 1.51        & 1.50            & 1.52           & \cite{althaus2014}   \\
+EVD                 & Sierra Leone                     & 2014                          & 2.53        & 2.41           & 2.67           & \cite{althaus2014}      \\
+EVD                 & Liberia                          & 2014                          & 1.59        & 1.57           & 1.60            & \cite{althaus2014}    \\ \hline
+Seasonal Influenza  & USA, France, Australia           & 1972-2002 & 1.30         & 1.20            & 1.40            & \cite{chowell2008}  \\
+Seasonal Influenza  & French Territory                 & 1987-1995                     & 1.50         & 1.09           & 1.73           & \cite{bonabeau1998}  \\ \hline
+H1N1 Influenza      & Thailand                         & 2009                          & 2.07        & 1.92           & 2.22           & \cite{desilva2009}     \\
+H1N1 Influenza     & Peru                             & 2009                          & 1.37        & 1.20            & 1.70            & \cite{desilva2009}        \\
+H1N1 Influenza      & La Gloria, Mexico                  & 2009                          & 1.58        & 1.34           & 1.60            & \cite{fraser2009}     \\ \bottomrule
+\end{tabular}
+\caption{Estimates of $\rr$ from various researchers for the infectious diseases of HIV, ZIK-V, Ebola (EVD), seasonal influenza, and H1N1 pandemic influenza ranging from the years 1972-2016 in various locations.
+We also report the lower and upper bounds reported by the researchers.}
+\label{tab:r0-real-ex}
+\end{table}
+\end{landscape}
 
@@
 Both $\beta$ and $\gamma$ are constrained to be between $0$ and $1$.
-For the remainder of this paper, we use $X$, $Y$, and $Z$ for stand-ins for the ODEs for $S$, $I$, and $R$, respectively, to avoid any confusion with $\rr$.
+For the remainder of this paper, we use $X$, $Y$, and $Z$ to denote the number of individuals in the $S$, $I$, and $R$ compartments, respectively, to avoid any confusion with $\rr$.
 \begin{align}
@@
   \textnormal{Data} &= \left \{\left (X_{obs}(t), Y_{obs}(t), Z_{obs}(t) \right ) : t=0, 1, \dots, T\right \}, \\
-  \hat{\rr} &= m(Data),
+  \hat{\rr} &= m(\textnormal{Data}),
 \end{align*}
@@
 
-Good properties of an estimator may include being (1) being a consistent estimator
-\begin{align*}
-  \hat{\rr} \overset{P}{\to} 0 \textnormal{ as } T\to \infty
+Good properties of an estimator may include (1) being a consistent estimator
+\begin{align*}
+  \hat{\rr} \overset{P}{\to} \rr \textnormal{ as } T\to \infty
 \end{align*}
@@
 \begin{align*}
-\hat{\rr} \leadsto F.  
+\hat{\rr} \overset{d}{\to} F.  
 \end{align*}
@@
 \end{align*}
-the second property is particularly desirable.  At any rate, we would like to be able to estimate $a$ and $b$ for a given $\alpha$-level.
+the second property is particularly desirable, as we could find $a$ and $b$ from the distribution $F$.
+At any rate, we would like to be able to estimate $a$ and $b$ for a given $\alpha$-level.
+We could also make similar hypothesis tests such as that the $\rr$ for H1N1 influenza is greater than the reported $\rr$ for seasonal influenza.
 
@@
 \subsection{Least Squares ($\beta$, $\gamma$) (LS)}\label{least-squares-beta-gamma}
-The first approach to estimate $\rr$ in the SIR model is to minimize the joint sum of squares as seen in approaches similar to \cite{malik2008} and \cite{canto2017}.
-In this method, the least squares estimate of $\rr$ is formed from the data collected at each time point to estimate $\hat{\beta}$ and $\hat{\gamma}$ and uses the  subsequent plug-in estimator found in Equation \ref{eq:sirls} to estimate $\rr$.
+Our first approach to estimate $\rr$ in the SIR model is to minimize the joint mean square error for the data collected at each time point and use the plug-in estimator found in Equation \ref{eq:sirls}.
 In particular, we find
@@
 \end{align*}
-Then  the estimate for $\rr$ is given by Equation \ref{eq:sirls},
+Then the least squares (LS) estimate for $\rr$ is given by Equation \ref{eq:sirls},
 \begin{align}\label{eq:sirls}
@@
 
-This method is sometimes used to estimate $\hat{\beta}$ and $\hat{\gamma}$ by using a grid search but can be done with more sophisticated minimization algorithms.
-Due to the ODE format of the equations, we can only use minimization algorithms which do not rely on an explicit gradient such as Nelder-Mead simplex minimization \citep{nelder-mead1965}.
+This method is sometimes used to estimate $\hat{\beta}$ and $\hat{\gamma}$ via a grid search of $\beta$ and $\gamma$ but can also be done with more sophisticated optimiization algorithms.
+Since we cannot explicitly write down the partial derivatives of $S$, $I$, and $R$ with respect to $\beta$ and $\gamma$, we can only use minimization algorithms which do not rely on an explicit gradient such as Nelder-Mead simplex minimization \citep{nelder-mead1965}.
 
@@
 A possible answer is that if we were to assume Gaussian noise, then the $L_2$-norm would be equivalent to the maximum likelihood estimation.
-Another answer is that $L_1$ is a continuously differentiable function and hence is easier to compute, which allows for sensitivity analysis to more easily be conducted.
+Another answer is that $L_2$ is a continuously differentiable function and hence is easier to compute, which allows for sensitivity analysis to  be conducted more easily.
 Finally, LS can be done without writing down explicit assumptions on the noise within the model.
-However, we cannot guarantee properties of consistency or converging to a known distribution without more explicit assumptions.
-
-On the other hand, least squares is easy to implement in software such as base \texttt{R}'s \texttt{optim()} function, and as we see in Section \ref{sec:results}, produces comparable methods to other methods.
+However, we cannot guarantee properties of consistency or convergence to a known distribution without more explicit assumptions.
+
+On the other hand, least squares is easy to implement in software such as \texttt{R}'s \texttt{optim()} function, and as we see in Section \ref{sec:results} produces comparable methods to other methods.
+Least squares and similar variants have been used to estimate $\rr$ in \cite{majumder2016}.
 
@@
 \end{align*}
-We use the $\hat{\rr}$ directly from the above estimation process, which again can be done with a grid search or another optimization process.
-It may be surprising that this method leads to different results than in simply using least squares.
+We use the $\hat{\rr}$ directly from the above estimation problem, which again can be solved with a grid search or another optimization process.
+It may be surprising that this method leads to different results than simply using least squares.
 One reason we obtain different results may be attributed to better numerical precision as there is one fewer step to obtain an estimate for $\rr$.
-However, we also see that this method results in smaller standard errors than in LS, which we discuss more in Section \ref{sec:results}.
-
-Like LS, while we have no explicit parametric assumptions on the noise, we also are uncertain of any theoretical guarantees of our resulting estimator.
+However, we also see that this method further results in smaller standard errors than in LS, which we discuss more in Section \ref{sec:results}.
+
+Like LS, while we have no explicit parametric assumptions on the noise, we also are uncertain of any theoretical properties of our resulting estimator.
 Likewise, the same difficulties in sensitivity analysis arise as in LS.
@@
 The SIR ODEs in Eq.
-\ref{eq:sir} have no known closed form solution, and so we are already using approximations using numerical integration.
+\ref{eq:sir} have no known closed form solution, and so we use numerical integration to obtain approximate solutions.
 In addition to this, data collected from real diseases are typically very noisy to begin with.
-\cite{chang2017} discovered that an SIR model may be well approximated by a linear model.
+\cite{chang2017} showed that an SIR model may be well approximated by a linear model.
 We use this approach here to estimate $\rr$.
 
-Specifically, we estimate two linear polynomials in \(t\) with degree \(K= 10\) to \(X_{obs}\)
-and \(Y_{obs}\) using least squares to find the coefficients $\{(\hat{x}_k,
-\hat{y}_k)\}_{k=1, \dots, K}$,
+Specifically, we estimate two linear polynomials in \(t\) with degree $K$  to \(X_{obs}\)
+and \(Z_{obs}\) using least squares to find the coefficients $\{(\hat{x}_k,
+\hat{z}_k)\}_{k=1, \dots, K}$,
 \begin{align*}
 \hat{X}(t) &= \sum_{k=0}^K \hat{x}_k t^k\\
-{\hat{Y}}(t) &= \sum_{k=0}^K \hat{y}_k t^k
+{\hat{Z}}(t) &= \sum_{k=0}^K \hat{z}_k t^k
 \end{align*}
@@
 \hat{X}^\prime(t) &= \sum_{k=1}^K k \hat{x}_k t^{k-1}\\
-\hat{Y}^\prime(t) &= \sum_{k=0}^K k \hat{y}_k t^{k-1}
+\hat{Z}^\prime(t) &= \sum_{k=0}^K k \hat{z}_k t^{k-1}
 \end{align*}
@@
   \end{align}
-  Here, $K=10$ is arbitrary and should be selected using some criterion such as AIC in addition with the knowledge of how many data points are available to the user.
-Besides optionally deciding on the degree of polynomials to fit, this model is simple to implement and gives very comparable results to using least squares with the SIR model.
+  Here, $K$ is arbitrary and should be selected using some criterion such as AIC.
+Besides optionally deciding on the degree of polynomials to fit, this model is simple to implement and gives comparable results to using least squares with the SIR model.
 The time $t=0$ is used to best capture the initial outbreak.
@@
   An advantage of using this estimation method is that it is simple to implement using any linear modelling software.
-On the other hand, we are assuming $X(t) = \hat{X}(t)$ and so if the data truly follows a SIR model, then we will have biased estimators for $\rr$.
+On the other hand, we are assuming $X(t) = \hat{X}(t)$ and so if the data truly follows a SIR model, then we will likely have a biased result for $\rr$.
 Using this method, we are able to check how sensitive $\rr$ is to the degree of the fitted polynomial.
@@
 The above formulation (Section \ref{linear-model-approximation-degree-10}) of a linear model approximation only uses the estimate at time $t=0$ to estimate $\rr$.
-We can instead, use all time points available to estimate $\rr$.
-We fit a linear polynomial of \(t\) with degree \(K= 10\) model to \(X\)
-and \(Z\) as above, with a slight modification in how we estimate
+We can instead, use all time points $O$ available to estimate $\rr$.
+We fit a polynomial in \(t\) with degree \(K\) to \(X_{obs}\)
+and \(Z_{obs}\) as above, with a slight modification in how we estimate
 \(\rr\),
 \begin{align*}
-  \hat{\rr} &= \frac{1}{\# \text{ Obs }t} \sum_t \frac{-\hat{X}^\prime(t)}{\hat{Z}^\prime(t)} \cdot \frac{N}{X(0)} 
+  \hat{\rr} &= \frac{1}{O} \sum_t \frac{-\hat{X}^\prime(t)}{\hat{Z}^\prime(t)} \cdot \frac{N}{X(0)} 
 \end{align*}
@@
 An advantage to this approach is that we have a more robust estimate of $\rr$ than just using one time point.
-However, the estimate will still be biased.
-Like in LMAT, we can examine how sensitive $\rr$ is to the order of the polynomial $K$.
+Like in LMA, we can examine how sensitive $\rr$ is to the order of the polynomial $K$.
 
@@
 Here we assume that the time step is small enough to approximate the incidence.
-The advantage of this method is that incidence data is generally available as is prevalence data for certain diseases such as HIV \citep{hall2008}.
+The advantage of this method is that incidence data is generally readily available as is prevalence data for certain diseases such as HIV.
 However, as one is required to have prior knowledge about $\gamma$, it may be easier to directly estimate $\rr$ with one of the many other methods described that does not require a prior knowledge about $\gamma$.
@@
 \subsection{Smoothed Incidence to Prevalence Ratio (SIPR)}
-We use the same method as above, IPR, but first estimate splines with 4 degrees of freedom, $\hat{X}(t)$ and $\hat{Y}(t)$, , to fit to $X_{obs}$ and $Y_{obs}$, respectively.
+We use the same method as above, IPR, but first estimate splines with $K$ degrees of freedom, where $K$ is small, $\hat{X}(t)$ and $\hat{Y}(t)$, to fit to $X_{obs}$ and $Y_{obs}$, respectively.
 Then  $J(t) \approx -(\hat{X}(t+1) - \hat{X}(t))$, and the $\hat{\textnormal{IPR}}(t) = \frac{J(t)}{\hat{Y}(t)}$.
@@
 \end{align*}
-The advantage of this method is that it creates a less noisy estimate than estimating IPR using only one point.
+The advantage of this method is that it creates a less variable estimate than estimating IPR using only one point.
 It has the same disadvantages as the regular IPR ratio in that it requires knowledge about $\gamma$.
-Also, ratio estimators usually behave badly when the denominator is small.
 
@@
 Thus, we can regress the left hand side in Eq.
-\ref{eq:harko_lin} on $Z(t)$, the number of recovered individuals, with $\rr$ as the coefficient of $Z(t)/N$ and no intercept term to obtain an estimate of $\hat{\rr}$,
+\ref{eq:harko_lin} on $Z(t)$, the number of recovered individuals, with $\rr$ as the coefficient of $Z(t)/N$ and with an intercept term of $\beta_0=0$ to obtain an estimate of $\rr$,
 \begin{align*}
@@
 \end{align*}
-This method says that for a one percent increase in number of recovered individuals, we expect the ratio of number of old susceptibles to number of new susceptibles to increase by $e^{\rr}$,
+This method says that for a one percent increase in number of recovered individuals, we expect the ratio of number of previous susceptibles to number of new susceptibles to increase by $e^{\rr}$,
 \begin{align*}
@@
 One, the estimate of $\rr$ is highly interpretable.
-Two, an assumption of Gaussian noise in Equation \ref{eq:harko_lin} is not as egregious as in the other methods to estimate $\rr$ because we are looking at a log transformation of the the percent of suceptibles instead of untransformed variables.
+Two, an assumption of Gaussian noise in Equation \ref{eq:harko_lin} is not as egregious as in the other methods to estimate $\rr$ because we are looking at a log transformation of the the percent of suceptibles instead of raw variables.
 As a consequence, we may be able to make a case where we the resulting estimate of $\rr$ is unbiased and a consistent estimator.
@@
 \subsection{Markov chain estimation (MC)}
-A natural approach to epidemic modeling is that of Markov chains (MC), since it is assumed an individual's next state is only dependent on its current state.
+A natural approach to epidemic modelling is that of Markov chains (MC), since it is assumed an individual's next state is only dependent on its current state and the current states of other individuals.
 Much work has been done over the years in this specific field including asymptotic behavior, continuous time MC, confidence intervals, and more \citep{jacquez1991,gani1995,daley2001epidemic}.
@@
 We have explicitly stated how randomness enters the model so it is up to the researcher to check whether this assumption is a good fit.
-However, it is possible to simulate how sensitive $\hat{\rr}$ is to $\hat{\alpha}$.
+Also, it is possible to simulate how sensitive $\hat{\rr}$ is to $\hat{\alpha}$.
 
@@
 \end{align*}
-Then, the posterior distribution of $\rr$ given the previous days' incidences is
-\begin{align*}
-  P(\rr | J(0), \dots, J(t+1)) = \frac{P(J(t+1) | \rr, J(0), \dots, J(t))P(\rr| J(0), \dots, J(t))}{P(J(0), \dots, J(t+1))}.
+Then, the posterior distribution of $\rr$ given the previous days' incidence is
+\begin{align*}
+  P(\rr | J_0, \dots, J_{t+1}) = \frac{P(J_{t+1} | \rr, J_0, \dots, J_t)P(\rr| J_0, \dots, J_t)}{P(J_0, \dots, J_{t+1})}.
 \end{align*}
@@
 
+\subsection{Exponential Growth}\label{sec:expgrowth}
+%% nishiura, majumder, IDEA model -- sum of squares (santillana2016)
+%% An IDEA for short term outbreak projection: nearcasting using the basic reproduction number.  (fisman2013, fisman2014)
+%% chowell 2008 - parametric bootstrap for a model they made up, but sort of empirical.
+
 \section{Confidence and Credible Intervals}
@@
 
-Estimating $\rr$ is difficult and estimating $V[\rr]$, the variance and subsequently CIs, are even more so.
+Estimating $\rr$ is difficult and estimating $V[\rr]$, the variance and confidence or credible intervals (CI) is even more so.
 We describe general methods which may be applicable to estimate the variance.
-In particular, we describe three methods 1) the delta method, 2) the block strap, and 3) posterior distributions.
+In particular, we describe three methods 1) the delta method, 2) the block bootstrap, and 3) posterior distributions.
 
@@
 
-When a method estimates \(\beta\) and \(\gamma\) instead of \(\rr\) directly, we use the delta method approximation to calculate the
+When the method estimates \(\beta\) and \(\gamma\) instead of \(\rr\) directly, we use the delta method approximation to calculate the
 variance of \(\rr\).
@@
 This estimate assumes that the distribution of $\beta$ and $\gamma$ are asymptotically normal.
-Here, we use the relationship of $\rr = \frac{\beta}{\gamma}$ in order to estimate the variance, which is specific to the SIR framework.
-A disadvantage of this method is that it provides no insight on how to estimate $\Sigma_{\beta,\gamma}$.
+Here, we use the relationship of $\rr = \frac{\beta}{\gamma}$ in order to estimate the variance, which is specific to the SIR model.
+This method may be extended to other frameworks such as the SEIR model.
+The advantages of the delta method are that since the Next Generation Matrix \citep{diekmann2009} posits a recipe to derive $\rr$, then it is theoretically possible to use the delta method to derive an expression for the variance of $\rr$ for any compartment model.
 
@@
 These are non-overlapping partitions or blocks, although there are variants for which blocks can overlap.
-Following, for iteration $\ell$, one samples uniformly replacement $k$ blocks and estimates $\hat{\rr}_b$ from those $k$ blocks, repeating the sampling and estimation for a total of $L$ estimates of $\hat{\rr}$.
+For iteration $\ell$, one samples $k$ blocks uniformly with replacement and estimates $\hat{\rr}_b$, repeating the sampling and estimation for a total of $L$ estimates of $\hat{\rr}$.
 Then
@@
 \end{align*}
-More about the asympotic distributions of the blockwise-bootstrap can be found in \cite{cao1999}, along with descriptions of variants.
-
-The assympotic properties hold when the stochastic process is $m$-dependent, that is when observations $t, t+1, \dots$ and $s, s+1, \dots$ are independent whenever $s+m < t$ and under some smoothness conditions of the statistic used to estimate $\rr$.
+More about the asympotic distributions of the blockwise-bootstrap can be found in \cite{cao1999}, along with descriptions of variants of the block bootstrap.
+
+The asympotic properties hold when the stochastic process is $m$-dependent, that is when observations $t, t+1, \dots$ and $s, s+1, \dots$ are independent whenever $s+m < t$ and under some smoothness conditions of the statistic used to estimate $\rr$.
 
@@
 \subsection{Posterior Distribution}
-When one uses a Bayesian approach to estimate $\rr$, the result should be a distribution rather than an estimate.
-One can simply look at the quantiles of this distribution to form a credible interval for this estimate.
+Here, we assume $\rr$ is a random variable and the data is fixed.
+Following the notation of \cite{wasserman2004}, then the posterior distribution $f(\rr | \textnormal{data})$ is is proportional to the likelihood of observing the data given $\rr$, $\mathcal{L}(\textnormal{data}|\rr)$, multiplied by our prior on $\rr$, $f(\rr)$.
+\begin{align*}
+f(\rr | \textnormal{data} ) \propto \mathcal{L}(\textnormal{data} | \rr) f(\rr)
+\end{align*}
+ Typically, the posterior is either estimated using conjugate distributions for the likelihood and prior or through Markov Chain Monte Carlo (MCMC) to simulate a posterior distribution $\hat{F}$ such that
+\begin{align*}
+\hat{F} \overset{d}{\to} f(\rr| \textnormal{data}).
+\end{align*}
+Then commonly, $\hat{\rr} = \hat{F}_{.5}$, the median of $\hat{F}$.
+Correspondingly, the 95\% credible interval is $\left[\hat{F}_{.05}, \hat{F}_{.95} \right ]$.
 Advantages of this method include having an entire distribution compared to a point estimate.
-One disadvantage is that the Bayesian approach will not guarantee frequentist properties that epidemiologists may be looking for such as being able to compare diseases to one another.
+A disadvantage is that using conjugate priors may not be justified for the given data and MCMC may be computationally intractable.
+One should note that these 95\% credible intervalss are not the same as confidence intervals as the ``true'' value of $\rr$ may not be contained in the credible intervals 95\% of the time.
 
@@
 We compare the performance of the SIR models on simulated data from both the SIR as well as other models.
-In all of these simulations, we first generate data from the models under known conditions, then we add on errors with known variances to each time point.
-We simulate with two different types of error distributions, Gaussian and autoregressive.
-In the Gaussian case, we add on independently and identically distributed errors to the $X$ and $Y$ compartments with variances $\sigma_X$ and $\sigma_Y$, respectively; the sum of these errors is then subtracted from the $Z$ compartment to maintain constant population.
-The same process was followed for the autoregressive case, except that we generate errors from an AR(1) model with a correlation coefficient $\rho = 0.7$ added to Gaussian white noise.
-For each error distribution, we simulate an additional data set which adds a monotonicity adjustment to the X and Z compartments after adding on the  using the Pool Adjacent Violators Algorithm (PAVA).
-The sum of the populations of these compartments is then subtracted from the $Y$ compartment, again to maintain a constant population.
-This gives us a total of four error structures for each set of conditions: Gaussian (Norm), Gaussian monotone (Norm-M), autoregressive (AR), and autoregressive monotone (AR-M).
+In all of these simulations, we first generate data from the models under known conditions, then we add on errors with known variances to each time point,
+\begin{align}\label{eq:sim-models}
+  X_{model}(t) &= f(t) + \epsilon_{X,t} \\
+  Y_{model}(t) &= g(t) + \epsilon_{Y,t} \nonumber\\
+  Z_{model}(t) &= N - X(t) - Y(t)\nonumber 
+\end{align}
+We adjust the recovered compartment $Z$ so the total population is constant. 
+We simulate with two different types of error distributions, Gaussian and autoregressive,
+
+\noindent \textbf{Gaussian (Norm) ($X_{G}$, $Y_{G}$, $Z_{G}$)}
+\begin{align*}
+  f(t) &= X(t) \\
+  g(t) &= Y(t) \\
+  \epsilon_{X,t} &\sim N(0, \sigma_X^2) \\
+  \epsilon_{Y,t} &\sim N(0, \sigma_Y^2)
+\end{align*}
+\textbf{Autoregressive (AR) ($X_{AR}$, $Y_{AR}$, $Z_{AR}$)}
+\begin{align*}
+  f(t) &= \rho X_{obs}(t-1) \\
+  g(t) &= \rho Y_{obs}(t-1) \\
+  \epsilon_{X,t} &\sim N(0, \sigma_X^2) \\
+  \epsilon_{Y,t} &\sim N(0, \sigma_Y^2).\\
+\end{align*}
+For both the Gaussian and AR dataset, we form an additional dataset which enforces the monotonicity of X and Z via the Pool Adjacent Violators Algorithm (PAVA), described by \cite{friedman1984}).
+Here, we adjust $Y$ so the constant population is maintained.
+
+\noindent \textbf{Gaussian-PAVA (Norm-M) ($X_{GP}$, $Y_{GP}$, $Z_{GP}$)}
+\begin{align*}
+ X_{GP}(t) &= \textnormal{PAVA}(X_G(t)) \\
+  Y_{GP}(t) &= N - X_{GP}(t) - Z_{GP}(t) \\
+  Z_{GP}(t) &= \textnormal{PAVA}(Z_G(t)) \\
+\end{align*}
+\textbf{Autoregressive-PAVA (AR-M) ($X_{ARP}$, $Y_{ARP}$, $Z_{ARP}$)}
+\begin{align*}
+  X_{ARP}(t) &= \textnormal{PAVA}(X_{AR}(t)) \\
+  Y_{ARP}(t) &= N - X_{ARP}(t) - Z_{ARP}(t) \\
+  Z_{ARP}(t) &= \textnormal{PAVA}(Z_{AR}(t)) 
+\end{align*}
+This gives us a total of four error structures for each set of conditions: Gaussian (Norm), Gaussian monotone (Norm-M), autoregressive (AR), and autoregressive monotone (AR-M) for a given set of parameters $(\beta, \gamma, O, X(0), Y(0), Z(0), \sigma_X, \sigma_Y, N)$.
 
@@
 
-	$\beta$ = 0.06, $\gamma$ = 0.03; $\rr$ = 2
-
-	Number of time steps: 365
-
-	Starting X size ($X_0$): 99950
-
-	Starting Y size ($Y_0$): 50
-
-	Starting Z size: 0
+	$\beta$ = 0.06, $\gamma$ = 0.03 ($\rr$ = 2)
+
+	Number of time steps $O$: 365
+
+	Starting X size ($X(0)$): 99950
+
+	Starting Y size ($Y(0)$): 50
+
+	Starting Z size ($Z(0)$): 0
 
@@
 
-	Total population: 100000.
+	Total population $N$: 100000.
 
@@
                   (0.06, 0.04)             & (99900, 100)               & (500, 20)     \\
+                  (0.06,.03) & (99950, 50) & (100, 5)\\
                   (0.06, 0.06)             & (99990, 10)             & (2500, 100)           \\
@@
 While we will not have a ``known'' $\rr$ value to compare our results to, this will allow us to see if the models give us reasonable results when the data do not come from the SIR model.
-We simulate data from linear and quartic models with respect to the time step, using a monotonically increasing function for the $X$ compartment and a monotonically decreasing function for the $Z$ compartment.
+We simulate data from polynomials of degree 1 and 4 with respect to the time step, using a monotonically increasing function for the $X$ compartment and a monotonically decreasing function for the $Z$ compartment.
 The sum of these is subtracted from the constant total population to get the $Y$ compartment.
@@
 
+\begin{table}[H]
+	\includegraphics[scale=0.5]{BaseBase.jpeg}
+
+	\caption{\label{tab:}Baseline Data, $\rr$ Estimates and Standard Errors}
+	\centering
+	\begin{tabular}[t]{l|r|r}
+		\hline
+		Model & Estimate & Std. Err\\
+		\hline
+		SIRLS & 1.9999711 & 0.0055923\\
+		\hline
+		SIRREPAR & 1.9997340 & 0.0049939\\
+		\hline
+		SIRLMA & 1.6863640 & 0.1886132\\
+		\hline
+		SIRLMAT & 1.4063227 & 0.6309170\\
+		\hline
+		SIRIPR & 4.4203632 & 12.3593415\\
+		\hline
+		SIRSIPR & 1.8870824 & $<$ 1e-07 \\
+		\hline
+		LOGLINEAR & 1.9998072 & 0.0002198\\
+		\hline
+		SIRMC & 0.9485555 &  $<$ 1e-07 \\
+		\hline
+		SEQBAYES & 1.6495023 & 0.0672249\\
+		\hline
+	\end{tabular}
+\end{table}
+
+Overall, we see a wide variation in the estimates of and the standard errors for $/rr$ depending on the method used.
+The methods that rely upon least squares optimization and the log-linear model are the closest to the true $\rr$ value, and they also have fairly small standard error estimates.
+On the other hand, a lot of the methods appear to underestimate the true $/rr$, and many of them have larger standard errors as well.
+Some notable exceptions include the IPR model, which tends to overestimate $/rr$, and the Markov chain and SIPR models, which appear to have unreasonably small standard errors.
+
+\subsection{Autoregressive Errors}\label{sec:res-AR}
+
+\includegraphics[scale=0.5]{AR.jpeg}
+\begin{table}[H]
+
+	\caption{\label{tab:}Autoregressive Errors, $\rr$ Estimates and Standard Errors}
+	\centering
+	\begin{tabular}[t]{l|r|r}
+		\hline
+		Model & Estimate & Std. Err\\
+		\hline
+		SIRLS & 2.0000198 & 0.0078308\\
+		\hline
+		SIRREPAR & 1.9998852 & 0.0069901 \\
+		\hline
+		SIRLMA & 35.656866 & 4815.2163162\\
+		\hline
+		SIRLMAT & 1.049360 & 4.8830065 \\
+		\hline
+		SIRIPR & 2.183371 & 7.397726 \\
+		\hline
+		SIRSIPR & 0.8386350 & $<$ 1e-07 \\
+		\hline
+		LOGLINEAR & 1.9998203 & 0.0003227\\
+		\hline
+		SIRMC & 0.9485555 & $<$ 1e-07 \\
+		\hline
+		SEQBAYES & 1.0000000 & 0.0523424\\
+		\hline
+	\end{tabular}
+\end{table}
+
+With an autoregressive error structure in the simulated data, the log-linear and least squares estimation methods appear to have a similar accuracy.
+On the other hand, the other models mostly appear to be somewhat less accurate.
+In particular, the linear regression based models do much worse than in the Gaussian error case.
+For all of these models, the standard error is larger, which would expect given that the data have a large positive autocorrelation coefficient.
+
 \subsection{Infection and Recovery Rate ($\beta, \gamma$)}\label{sec:res-beta-gamma}
 
+\textbf{Need to figure out incorporating different $/rr$ scales in to graph. Facets won't work.}
+
+\begin{table}[H]
+
+	\caption{\label{tab:}$\gamma = 0.001$, $\rr$ Estimates and Standard Errors}
+	\centering
+	\begin{tabular}[t]{l|r|r}
+		\hline
+		Model & Estimate & Std. Err\\
+		\hline
+		SIRLS & 59.9764682 & 0.0434705\\
+		\hline
+		SIRREPAR & 59.9951876 & 0.4343984\\
+		\hline
+		SIRLMA &  -117.9694856 & 101.1885824\\
+		\hline
+		SIRLMAT & -2.8535930 & 193.6526107 \\
+		\hline
+		SIRIPR & 54.146506 & 200.2595\\
+		\hline
+		SIRSIPR & 3.9645187 & $<$ 1e-07\\
+		\hline
+		LOGLINEAR & 46.3209596 & 0.7797216\\
+		\hline
+		SIRMC & 0.9485554 & 0.0000001\\
+		\hline
+		SEQBAYES & 0.9769731 & 0.0517362\\
+		\hline
+	\end{tabular}
+\end{table}
+
+Looking at the case where $\gamma = 0.001$, it appears that most of the models do not handle a very large $/rr$ well.
+Only the least squares and reparameterized least squares models give very close answers; the IPR and log-linear models give fairly close answers as well.
+However, the other models do very poorly; they ted to greatly underestimate $/rr$.
+The linear model approximation does particularly poorly, likely because the linear model does not fit the data well.
+Several of the models also return extremely large standard errors, meaning they are not very reliable in these cases.
+
+\begin{table}[H]
+
+	\caption{\label{tab:}$\gamma = 0.24$, $\rr$ Estimates and Standard Errors}
+	\centering
+	\begin{tabular}[t]{l|r|r}
+		\hline
+		Model & Estimate & Std. Err\\
+		\hline
+		SIRLS & 0.2363274 & 4.1042813\\
+		\hline
+		SIRREPAR & 4.1795252 & 2.7455569\\
+		\hline
+		SIRLMA &  0.7074613 & 0.2028765 \\
+		\hline
+		SIRLMAT & 0.6746914 & 6.2471252 \\
+		\hline
+		SIRIPR & 181.015802 & 742.1088 \\
+		\hline
+		SIRSIPR & -0.2484284 & 0.1159769 \\
+		\hline
+		LOGLINEAR & 0.9912397 & 0.0030505\\
+		\hline
+		SIRMC & 0.9485811 & 0.0000150\\
+		\hline
+		SEQBAYES & 1.2354952 & 0.0581801\\
+		\hline
+	\end{tabular}
+\end{table}
+
+The above table corresponds to the case where $\gamma = 0.24$.
+It appears that a lot of estimation methods do not handle very small $/rr$ well either.
+None of them are extremely close to truth, except for regular least squares, as most of the tend to overestimate the truth.
+The IPR is particularly bad in this case.
+Also, in general, the standard errors are several orders of magnitude larger than in the baseline data.
+
 \subsection{Number of Time Steps ($T$)}\label{sec:res-time}
 
+\includegraphics[scale=0.5]{time.jpeg}
+
+\begin{table}[H]
+
+	\caption{\label{tab:} $T$ = 100, $\rr$ Estimates and Standard Errors}
+	\centering
+	\begin{tabular}[t]{l|r|r}
+		\hline
+		Model & Estimate & Std. Err\\
+		\hline
+		SIRLS & 1.9993392 & 0.0286060\\
+		\hline
+		SIRREPAR & 1.9994284 & 0.0312814\\
+		\hline
+		SIRLMA & 0.9496498 & 0.0825419\\
+		\hline
+		SIRLMAT & 3.3442696 & 21.4895422\\
+		\hline
+		SIRIPR & 7.3458131 & 16.1288906\\
+		\hline
+		SIRSIPR & -2.1198897 & $<$ 1e-07\\
+		\hline
+		LOGLINEAR & 2.0014146 & 0.0017620\\
+		\hline
+		SIRMC & 0.9485555 & 0.0000001\\
+		\hline
+		SEQBAYES & 2.4289382 & 0.1102030\\
+		\hline
+	\end{tabular}
+\end{table}
+
+The table above shows the methods applied to the first 100 time steps, out of 365, of a baseline data set.
+The estimates and standard errors are worse, which we would expect given less data.
+The results are still relatively close to the full baseline data.
+There also appears to be less tendency to underestimate $/rr$ compared to a longer time scale.
+
+\begin{table}[H]
+
+	\caption{\label{tab:} $T$ = 20, $\rr$ Estimates and Standard Errors}
+	\centering
+	\begin{tabular}[t]{l|r|r}
+		\hline
+		Model & Estimate & Std. Err\\
+		\hline
+		SIRLS & 38.6895520 & 106.2705193\\
+		\hline
+		SIRREPAR & 22.9645867 & 6.9816713\\
+		\hline
+		SIRLMA & 1.4418031 & 7.3163878\\
+		\hline
+		SIRLMAT & 1.0078807 & 0.1424793\\
+		\hline
+		SIRIPR & 25.7313065 & 35.2926077\\
+		\hline
+		SIRSIPR & 4.1772451 & 7.3202970\\
+		\hline
+		LOGLINEAR & 1.0429764 & 0.0276261\\
+		\hline
+		SIRMC & 0.9485556 & 0.0000005\\
+		\hline
+		SEQBAYES & 1.0000000 & 0.2236068\\
+		\hline
+	\end{tabular}
+\end{table}
+
+The table above shows the methods applied to the first 20 time steps of a baseline data set.
+As might be expected, the $/rr$ estimates are wildly inaccurate, and none are very close.
+The least squares methods are particularly worse in this situation compared to the other data sets.
+The standard errors are much larger than the baseline, which is to be expected as well due to the small number of data points.
+This shows that it is extremely difficult to create accurate online estimates of $/rr$ or to estimate $/rr$ without adequate data.
+
 \subsection{Initial Percentage of Infectious $\left (\frac{Y(0)}{N}\right)$}\label{sec:res-inf}
 
+\includegraphics[scale=0.5]{start.jpeg}
+
+\begin{table}[H]
+
+	\caption{\label{tab:}$Y(0) = 1000$, $\rr$ Estimates and Standard Errors}
+	\centering
+	\begin{tabular}[t]{l|r|r}
+		\hline
+		Model & Estimate & Std. Err\\
+		\hline
+		SIRLS & 1.9995884 & 0.0049609\\
+		\hline
+		SIRREPAR & 1.9997123 & 0.0034303\\
+		\hline
+		SIRLMA & -87.5945650 & 66413.128046\\
+		\hline
+		SIRLMAT & 0.7306100 & 4.681042\\
+		\hline
+		SIRIPR & 1.285071 & 1.440980\\
+		\hline
+		SIRSIPR & 1.2522106 & 57.9229822 \\
+		\hline
+		LOGLINEAR & 1.9995148 & 0.0002828\\
+		\hline
+		SIRMC & 0.9485555 & $<$ 1e-07\\
+		\hline
+		SEQBAYES & 0.9728534 & 0.0516270\\
+		\hline
+	\end{tabular}
+\end{table}
+
+The table above shows the methods applied to data where $Y(0) = 1000$, essentially starting at the middle of an epidemic.
+This data has very large underestimation of $/rr$.
+The LMA method once again gives a very far estimate, potentially indicating that a linear approximation does not apply in this case.
+The sample standard errors in this case are actually relatively close to the baseline, and even smaller for some of the methods.
+
+\begin{table}[H]
+
+	\caption{\label{tab:}$Y(0) = 1$, $\rr$ Estimates and Standard Errors}
+	\centering
+	\begin{tabular}[t]{l|r|r}
+		\hline
+		Model & Estimate & Std. Err\\
+		\hline
+		SIRLS & 2.468304 & 1.5370731\\
+		\hline
+		SIRREPAR & 2.469189 & 2.7293667\\
+		\hline
+		SIRLMA & 10.9282204 & 19.297359 \\
+		\hline
+		SIRLMAT & 1.624645 & 5.0359845\\
+		\hline
+		SIRIPR & 774.274659 & 7491.005080 \\
+		\hline
+		SIRSIPR & 2.4390757 & $<$ 1e-07 \\
+		\hline
+		LOGLINEAR & 1.999690 & 0.0007731\\
+		\hline
+		SIRMC & 0.948603 & 0.0000156\\
+		\hline
+		SEQBAYES & 1.471481 & 0.0634937\\
+		\hline
+	\end{tabular}
+\end{table}
+
+The table above shows the methods applied to data where $Y(0) = 1$, starting at the very beginning of an epidemic.
+The methods give many larger estimates, tending to overestimate the true $/rr$.
+In particular, the IPR and LMA methods give very large results.
+We also see a large standard error than the baseline.
+
 \subsection{Magnitude of Variance $(\sigma^2_X, \sigma^2_Y)$}\label{sec:res-var}
 
+\includegraphics[scale=0.5]{var.jpeg}
+
+\begin{table}[H]
+
+	\caption{\label{tab:} $(\sigma^2_X, \sigma^2_Y) = (50, 2)$, $\rr$ Estimates and Standard Errors}
+	\centering
+	\begin{tabular}[t]{l|r|r}
+		\hline
+		Model & Estimate & Std. Err\\
+		\hline
+		SIRLS & 1.9997552 & 0.0026502\\
+		\hline
+		SIRREPAR & 1.9998295 & 0.0023646\\
+		\hline
+		SIRLMA & 2.0398087 & 0.2773968\\
+		\hline
+		SIRLMAT & 1.4148628 & 0.5779651\\
+		\hline
+		SIRIPR & 2.7607518 & 8.1318687\\
+		\hline
+		SIRSIPR & 1.8368240 & $<$ 1e-07\\
+		\hline
+		LOGLINEAR & 1.9998907 & 0.0001037\\
+		\hline
+		SIRMC & 0.9485555 & $<$ 1e-07\\
+		\hline
+		SEQBAYES & 1.6750053 & 0.0677426\\
+		\hline
+	\end{tabular}
+\end{table}
+
+In the table above, the magnitude of the variance of the noise is smaller than in the baseline.
+The estimates are closer to the true $/rr$ and the standard errors are smaller across the board.
+This indicates that the methods give more accurate and precise estimates when the data are closer to following underlying true SIR model.
+
+\begin{table}[H]
+
+	\caption{\label{tab:} $(\sigma^2_X, \sigma^2_Y) = (2500, 100)$, $\rr$ Estimates and Standard Errors}
+	\centering
+	\begin{tabular}[t]{l|r|r}
+		\hline
+		Model & Estimate & Std. Err\\
+		\hline
+		SIRLS & 1.9980880 & 0.1353803\\
+		\hline
+		SIRREPAR & 1.9981446 & 0.1207935\\
+		\hline
+		SIRLMA & 0.4651913 & 7.5508447\\
+		\hline
+		SIRLMAT & 1.2877865 & 4.3947658\\
+		\hline
+		SIRIPR & 81.4901666 & 348.7350849\\
+		\hline
+		SIRSIPR & 1.0606095 & $<$ 1e-07\\
+		\hline
+		LOGLINEAR & 2.0018816 & 0.0054085\\
+		\hline
+		SIRMC & 0.9485562 & 0.0000010\\
+		\hline
+		SEQBAYES & 1.8567176 & 0.0713225\\
+		\hline
+	\end{tabular}
+\end{table}
+
+In the table above, the magnitude of the variance of the noise is larger than in the baseline.
+The estimates are worse and the standard errors are larger compared to the baseline data set, which we might expect.
+These estimates are not as terrible as we might imagine given the magnitude of the change - it seems that these methods are somewhat more robust to large noise compared to other changes.
+One exception in the IPR model, which may do worse due to larger fluctuations in the incidence prevalence ratio.
+
 \subsection{Other Models}\label{sec:res-oth}
 
+\begin{table}[H]
+
+  \begin{center}
+\includegraphics[scale=0.5]{other.jpeg}
+
+\end{center}
+
+\caption{\label{tab:} 4th Order Model, $\rr$ Estimates and Standard Errors}
+\centering
+\begin{tabular}[t]{l|r|r}
+	\hline
+	Model & Estimate & Std. Dev\\
+	\hline
+	SIRLS & 1.7693473 & 0.4226615\\
+	\hline
+	SIRREPAR & 1.7693312 & 0.2796903\\
+	\hline
+	SIRLMA & 0.3617606 & 0.1231325\\
+	\hline
+	SIRLMAT & 0.8213066 & 6.0466473\\
+	\hline
+	SIRIPR & 10.4349388 & 33.8952710\\
+	\hline
+	SIRSIPR & 4.7477067 &  $<$ 1e-07\\
+	\hline
+	LOGLINEAR & 1.9551795 & 0.0164608\\
+	\hline
+	SIRMC & 0.9485545 & 0.0000006\\
+	\hline
+	SEQBAYES & 0.2118519 & 0.0240918\\
+	\hline
+\end{tabular}
+\end{table}
+
+The table above shows the methods applied to data from a 4th order linear regression model with respect to time.
+Obviously, there is no true $\rr$ to compare to.
+What interests us in this case is whether the methods give similar results, since all the estimation methods are derived from the same model.
+However, the results vary much more compared to baseline data estimates, where underlying data came from SIR.
+We also get massive standard errors, even though actual noise variance magnitudes are the same as baseline.
+
+\subsection{Repeated Sampling}
+
+\begin{table}[H]
+
+	\caption{\label{tab:}Mean $\rr$ Estimates and Std. Devs, Repeated Sampling}
+	\centering
+	\begin{tabular}[t]{l|r|r|r|r}
+		\hline
+		Model & Mean Estimate & Std. Dev. & Baseline Estimate & Baseline SE\\
+		\hline
+		SIRLS & 2.0000175 & 0.0002734 & 1.9999190 & 0.0058322\\
+		\hline
+		SIRREPAR & 2.0000289 & 0.0003092 & 1.9997083 & 0.0052113\\
+		\hline
+		SIRLMA & 1.8497054 & 0.3338774 & 1.997416 & 0.3900006\\
+		\hline
+		SIRLMAT & 1.4562107 & 0.3576736 & 1.398016 & 0.6028170 \\
+		\hline
+		SIRIPR & 4.5214294 & 0.3705241 & 2.592974 & 9.040841\\
+		\hline
+		SIRSIPR & 1.7972938 & 0.0951042 & 0.9251762 & $<$ 1e-07 \\
+		\hline
+		LOGLINEAR & 2.0000079 & 0.0003191 & 1.9998403 & 0.0002227\\
+		\hline
+		SIRMC & 0.9485555 & $<$ 1e-07 & 0.9485555 & $<$ 1e-07\\
+		\hline
+		SEQBAYES & 1.1330264 & 0.4163031 & 1.0000000 & 0.0523424\\
+		\hline
+	\end{tabular}
+\end{table}
+
+The table above shows the mean and standard deviation of $/rr$ estimates produced by the models under 100 repetitions of the baseline conditions, with new noise generated each time.
+We want to compare the standard deviations of the estimates to the standard errors returned by each of the models for a single data set to asses how reliable those standard errors are.
+Overall, it appears than most of the models have standard error estimates that are smaller than the actual standard deviation of the estimates under repeated sampling.
+In some cases, such as the log linear method and the linear model approximations, are fairly close.
+The least squares methods, on the other hand, tend to overestimate the error by an order of magnitude.
+The SIPR and sequential Bayes model, on the other hand, have individual standard error estimates that are smaller than the standard deviation of repeated estimates, meaning that they understate how much error is in the estimates.
+
 \section{Discussion}\label{sec:discussion}
+
+The SIR estimation methods generally give reliable estimates when the data are from an underlying SIR model with little or no noise.
+However, the estimates from the various methods are generally much worse when the magnitude of the noise is larger, when the parameters of the model are unusual, and when the model is misspecified.
+We also see a large difference between the $\rr$ estimates and standard errors on the exact same data sets as well, even though all derived from the same model.
+Thus, when estimating $/rr$ based on an SIR or any other kind of compartment model, we need to take careful consideration.
 
@@
 			\hline
-			$\sigma_X = 50, \sigma_Y = 2$ & 1.993455 & 0.5784530 & 1.979966 & 0.1204126 & 2.033934 & 0.7559710 & 2.066706 & 0.3382834\\
-			\hline
-			$\sigma_X = 500, \sigma_Y = 20$ & 1.998088 & 0.1353803 & 1.996950 & 0.0541332 & 1.999686 & 0.1893888 & 2.002135 & 0.0927952\\
-			\hline
-			$\sigma_X = 2500, \sigma_Y = 100$ & 1.999755 & 0.0026502 & 1.999834 & 0.0026397 & 1.999555 & 0.0037664 & 1.999038 & 0.0036473\\
-			\hline
-			$\sigma_X = 10000, \sigma_Y = 500$ & 2.001870 & 0.0259428 & 1.997748 & 0.0162536 & 2.002792 & 0.0416203 & 1.997918 & 0.0257997\\
+			$\sigma_X = 50, \sigma_Y = 2$ & 1.999755 & 0.0026502 & 1.999834 & 0.0026397 & 1.999555 & 0.0037664 & 1.999038 & 0.0036473\\
+			\hline
+			$\sigma_X = 500, \sigma_Y = 20$ & 2.001870 & 0.0259428 & 1.997748 & 0.0162536 & 2.002792 & 0.0416203 & 1.997918 & 0.0257997\\
+			\hline
+			$\sigma_X = 2500, \sigma_Y = 100$ & 1.998088 & 0.1353803 & 1.996950 & 0.0541332 & 1.999686 & 0.1893888 & 2.002135 & 0.0927952\\
+			\hline
+			$\sigma_X = 10000, \sigma_Y = 500$ & 1.993455 & 0.5784530 & 1.979966 & 0.1204126 & 2.033934 & 0.7559710 & 2.066706 & 0.3382834\\
 			\hline
@@
 			\hline
-			$\sigma_X = 50, \sigma_Y = 2$ & 1.993660 & 0.5115266 & 1.979973 & 0.1053955 & 2.033636 & 0.7059701 & 2.066476 & 0.3172823\\
-			\hline
-			$\sigma_X = 500, \sigma_Y = 20$ & 1.998145 & 0.1207935 & 1.997147 & 0.0481545 & 2.000194 & 0.1680126 & 2.002244 & 0.0826643\\
-			\hline
-			$\sigma_X = 2500, \sigma_Y = 100$ & 1.999829 & 0.0023646 & 1.999946 & 0.0023591 & 1.999885 & 0.0033514 & 1.999483 & 0.0032200\\
-			\hline
-			$\sigma_X = 10000, \sigma_Y = 500$ & 2.002305 & 0.0232636 & 1.997996 & 0.0144744 & 2.003065 & 0.0372826 & 1.997732 & 0.0230069\\
+			$\sigma_X = 50, \sigma_Y = 2$ & 1.999829 & 0.0023646 & 1.999946 & 0.0023591 & 1.999885 & 0.0033514 & 1.999483 & 0.0032200\\
+			\hline
+			$\sigma_X = 500, \sigma_Y = 20$ & 2.002305 & 0.0232636 & 1.997996 & 0.0144744 & 2.003065 & 0.0372826 & 1.997732 & 0.0230069\\
+			\hline
+			$\sigma_X = 2500, \sigma_Y = 100$ & 1.998145 & 0.1207935 & 1.997147 & 0.0481545 & 2.000194 & 0.1680126 & 2.002244 & 0.0826643\\
+			\hline
+			$\sigma_X = 10000, \sigma_Y = 500$ & 1.993660 & 0.5115266 & 1.979973 & 0.1053955 & 2.033636 & 0.7059701 & 2.066476 & 0.3172823\\
 			\hline
@@
 		\hline
-		T = 20 & 22.964587 & NaN & 5.044869 & 2.1398099 & NA & NA & 1.681269 & 0.6543603\\
+		T = 20 & 22.964587 & 6.9816713 & 5.044869 & 2.1398099 & NA & NA & 1.681269 & 0.6543603\\
 		\hline
@@
 			\hline
-			$\sigma_X = 50, \sigma_Y = 2$ & 2.760752 & 8.131869 & 1.450899 & 2.052241 & 2.19804 & 5.456857 & 1.446505 & 2.182072\\
-			\hline
-			$\sigma_X = 50, \sigma_Y = 25$ & 17.377923 & 71.374848 & 1.889790 & 8.177704 & 12.91733 & 41.063025 & 2.373601 & 20.983311\\
-			\hline
-			$\sigma_X = 2500, \sigma_Y = 100$ & 81.490167 & 348.735085 & 2.267162 & 17.058029 & 195.37080 & 1451.040149 & 2.924092 & 18.434409\\
-			\hline
-			$\sigma_X = 10000, \sigma_Y = 500$ & 239.809101 & 1298.136206 & 4.500556 & 59.404233 & 92.10901 & 272.487471 & 3.535111 & 22.583706\\
+			$\sigma_X = 50, \sigma_Y = 2$ & 1.656451 & 4.879121 & 0.8705396 & 1.231344 & 1.318824 & 3.274114 & 0.8679029 & 1.309243\\
+			\hline
+			$\sigma_X = 500, \sigma_Y = 20$ & 10.426754 & 42.824909 & 1.1338742 & 4.906623 & 7.750396 & 24.637815 & 1.4241606 & 12.589987\\
+			\hline
+			$\sigma_X = 2500, \sigma_Y = 100$ & 48.894100 & 209.241051 & 1.3602975 & 10.234817 & 117.222483 & 870.624089 & 1.7544551 & 11.060645\\
+			\hline
+			$\sigma_X = 10000, \sigma_Y = 500$ & 143.885460 & 778.881724 & 2.7003338 & 35.642540 & 55.265406 & 163.492483 & 2.1210664 & 13.550223\\
 			\hline
@@
 			\hline
-			$X_0 = 99900, Y_0 = 100$ & -0.8808153 & $<$ 1e-07 & -1.150393 & $<$ 1e-07 & -6.563660 & $<$ 1e-07 & -1.777316 & $<$ 1e-07 \\
+			$X_0 = 99900, Y_0 = 100$ & -0.8808153 & 0.0000000 & -1.150393 & 0.0000000 & -6.563660 & 0.0000000 & -1.777316 & 0.0000000\\
 			\hline
@@
 			\hline
-			$X_0 = 99999, Y_0 = 1$ & 2.4390757 & $<$ 1e-07 & 5.411582 & $<$ 1e-07 & 1.204689 & $<$ 1e-07 & 1.035624 & $<$ 1e-07 \\
+			$X_0 = 99999, Y_0 = 1$ & 2.4390757 & 0.0000000 & 5.411582 & 0.0000000 & 1.204689 & 0.0000000 & 1.035624 & 0.0000000\\
 			\hline
@@
 			\hline
-			$\sigma_X = 50, \sigma_Y = 2$ & 1.836824 & $<$ 1e-07 & 1.7778150 & $<$ 1e-07 & 1.7603970 & $<$ 1e-07 & 1.842289 & $<$ 1e-07\\
-			\hline
-			$\sigma_X = 500, \sigma_Y = 25$ & 3.193066 & $<$ 1e-07 & 1.3211468 & $<$ 1e-07 & 1.3509726 & $<$ 1e-07 & 1.334458 & $<$ 1e-07\\
-			\hline
-			$\sigma_X = 2500, \sigma_Y = 100$ & 1.060609 & $<$ 1e-07 & 1.2345884 & $<$ 1e-07 & 1.7027546 & $<$ 1e-07 & 2.596160 & $<$ 1e-07\\
-			\hline
-			$\sigma_X = 10000, \sigma_Y = 500$ & -2.535946 & $<$ 1e-07 & 0.7094139 & $<$ 1e-07 & 0.8911487 & $<$ 1e-07 & -1.103940 & $<$ 1e-07\\
+			$\sigma_X = 50, \sigma_Y = 2$ & 1.836824 & $<$ 1e-07 & 1.7778150 & $<$ 1e-07 & 1.7603970 & $<$ 1e-07 & 1.842289 & $<$ 1e-07 \\
+			\hline
+			$\sigma_X = 500, \sigma_Y = 20$ & 3.193066 & $<$ 1e-07 & 1.3211468 & $<$ 1e-07 & 1.3509726 & $<$ 1e-07 & 1.334458 & $<$ 1e-07 \\
+			\hline
+			$\sigma_X = 2500, \sigma_Y = 100$ & 1.060609 & $<$ 1e-07 & 1.2345884 & $<$ 1e-07 & 1.7027546 & $<$ 1e-07 & 2.596160 & $<$ 1e-07 \\
+			\hline
+			$\sigma_X = 10000, \sigma_Y = 500$ & -2.535946 & $<$ 1e-07 & 0.7094139 & $<$ 1e-07 & 0.8911487 & $<$ 1e-07 & -1.103940 & $<$ 1e-07 \\
 			\hline
@@
 			\hline
-			$\sigma_X = 50, \sigma_Y = 2$ & 2.059934 & 0.0246402 & 1.995133 & 0.0038894 & 2.209575 & 0.0439100 & 2.545819 & 0.0568537\\
-			\hline
-			$\sigma\_X = 500, \sigma_Y = 20$ & 2.001882 & 0.0054085 & 1.995742 & 0.0015419 & 2.038646 & 0.0081562 & 2.015675 & 0.0031322\\
-			\hline
-			$\sigma\_X = 2500, \sigma_Y = 100$ & 1.999891 & 0.0001037 & 2.000036 & 0.0000930 & 1.999143 & 0.0001394 & 1.999749 & 0.0001328\\
-			\hline
-			$\sigma\_X = 10000, \sigma_Y = 500$ & 2.003402 & 0.0010320 & 1.998765 & 0.0003682 & 2.003763 & 0.0015505 & 1.999829 & 0.0008545\\
+			$\sigma_X = 50, \sigma_Y = 2$ & 1.999891 & 0.0001037 & 2.000036 & 0.0000930 & 1.999143 & 0.0001394 & 1.999749 & 0.0001328\\
+			\hline
+			$\sigma_X = 500, \sigma_Y = 20$ & 2.003402 & 0.0010320 & 1.998765 & 0.0003682 & 2.003763 & 0.0015505 & 1.999829 & 0.0008545\\
+			\hline
+			$\sigma_X = 2500, \sigma_Y = 100$ & 2.001882 & 0.0054085 & 1.995742 & 0.0015419 & 2.038646 & 0.0081562 & 2.015675 & 0.0031322\\
+			\hline
+			$\sigma_X = 10000, \sigma_Y = 500$ & 2.059934 & 0.0246402 & 1.995133 & 0.0038894 & 2.209575 & 0.0439100 & 2.545819 & 0.0568537\\
 			\hline
@@
 			\hline
-			$\gamma$ = 0.001 & 0.9485554 & 1.0e-07 & 0.9485555 & $<$ 1e-07 & 0.9485555 & $<$ 1e-07 & 0.9485555 & 0e+00\\
-			\hline
-			$\gamma$ = 0.04 & 0.9485555 & $<$ 1e-07 & 0.9485555 & $<$ 1e-07 & 0.9485555 & $<$ 1e-07 & 0.9485555 & 0e+00\\
-			\hline
-			$\gamma$ = 0.06 & 0.9485555 & 2.0e-07 & 0.9485555 & $<$ 1e-07 & 0.9485555 & 2.00e-07 & 0.9485555 & 0e+00\\
+			$\gamma$ = 0.001 & 0.9485554 & 1.0e-07 & 0.9485555 & $<$ 1e-07 & 0.9485555 & $<$ 1e-07 & 0.9485555 & $<$ 1e-07\\
+			\hline
+			$\gamma$ = 0.04 & 0.9485555 & $<$ 1e-07 & 0.9485555 & $<$ 1e-07 & 0.9485555 & $<$ 1e-07 & 0.9485555 & $<$ 1e-07\\
+			\hline
+			$\gamma$ = 0.06 & 0.9485555 & 2.0e-07 & 0.9485555 & $<$ 1e-07 & 0.9485555 & 2.00e-07 & 0.9485555 & $<$ 1e-07\\
 			\hline
@@
 			\hline
-			$\sigma_X = 50, \sigma_Y = 2$ & 0.9485506 & 3.8e-06 & 0.9485556 & 2e-07 & 0.9485428 & 2.6e-06 & 0.9485538 & 8e-07\\
-			\hline
-			$\sigma\_X = 500, \sigma_Y = 20$ & 0.9485562 & 1.0e-06 & 0.9485554 & 1e-07 & 0.9485640 & 3.6e-06 & 0.9485554 & 1e-07\\
-			\hline
-			$\sigma\_X = 2500, \sigma_Y = 100$ & 0.9485555 & $<$ 1e-07 & 0.9485555 & $<$ 1e-07 & 0.9485555 & $<$ 1e-07 & 0.9485555 & 0e+00\\
-			\hline
-			$\sigma\_X = 10000, \sigma_Y = 500$ & 0.9485554 & 2.0e-07 & 0.9485555 & $<$ 1e-07 & 0.9485554 & 1.0e-07 & 0.9485555 & 1e-07\\
+			$\sigma_X = 50, \sigma_Y = 2$ & 0.9485555 & 0.$<$ 1e-07 & 0.9485555 & $<$ 1e-07 & 0.9485555 & 0.$<$ 1e-07 & 0.9485555 & $<$ 1e-07\\
+			\hline
+			$\sigma_X = 500, \sigma_Y = 20$ & 0.9485554 & 2.0e-07 & 0.9485555 & $<$ 1e-07 & 0.9485554 & 1.0e-07 & 0.9485555 & 1e-07\\
+			\hline
+			$\sigma_X = 2500, \sigma_Y = 100$ & 0.9485562 & 1.0e-06 & 0.9485554 & 1e-07 & 0.9485640 & 3.6e-06 & 0.9485554 & 1e-07\\
+			\hline
+			$\sigma_X = 10000, \sigma_Y = 500$ & 0.9485506 & 3.8e-06 & 0.9485556 & 2e-07 & 0.9485428 & 2.6e-06 & 0.9485538 & 8e-07\\
 			\hline
@@
 			\hline
-			Linear SIR & 0.9485555 & $<$ 1e-07 & 0.9485555 & $<$ 1e-07 & 0.9485555 & $<$ 1e-07 & 0.9485555 & 0.0e+00\\
+			Linear SIR & 0.9485555 & $<$ 1e-07 & 0.9485555 & $<$ 1e-07 & 0.9485555 & $<$ 1e-07 & 0.9485555 & 0.$<$ 1e-07\\
 			\hline
@@
 		\hline
-		T = 200 & 0.9485555 & 1e-07 & 0.9485555 & 0 & 0.9485555 & 0e+00 & 0.9485555 & 0e+00\\
-		\hline
-		T = 100 & 0.9485555 & 1e-07 & 0.9485555 & 0 & 0.9485555 & 1e-07 & 0.9485555 & 0e+00\\
-		\hline
-		T = 50 & 0.9485554 & 2e-07 & 0.9485555 & 0 & 0.9485555 & 2e-07 & 0.9485555 & 0e+00\\
-		\hline
-		T = 20 & 0.9485556 & 5e-07 & 0.9485554 & 0 & 0.9485552 & 5e-07 & 0.9485556 & 1e-07\\
+		T = 200 & 0.9485555 & 1e-07 & 0.9485555 & $<$ 1e-07 & 0.9485555 & $<$ 1e-07 & 0.9485555 & $<$ 1e-07\\
+		\hline
+		T = 100 & 0.9485555 & 1e-07 & 0.9485555 & $<$ 1e-07 & 0.9485555 & 1e-07 & 0.9485555 & $<$ 1e-07\\
+		\hline
+		T = 50 & 0.9485554 & 2e-07 & 0.9485555 & $<$ 1e-07 & 0.9485555 & 2e-07 & 0.9485555 & $<$ 1e-07\\
+		\hline
+		T = 20 & 0.9485556 & 5e-07 & 0.9485554 & $<$ 1e-07 & 0.9485552 & 5e-07 & 0.9485556 & 1e-07\\
 		\hline
@@
 		\hline
-		$\sigma_X = 50, \sigma_Y = 2$ & 1.127679 & 0.0555836 & 1.000000 & 0.0523424 & 1.540929 & 0.0649748 & 1.000000 & 0.0523424\\
-		\hline
-		$\sigma\_X = 500, \sigma_Y = 20$ & 1.519223 & 0.0645155 & 1.000000 & 0.0523424 & 1.549862 & 0.0651629 & 1.000000 & 0.0523424\\
-		\hline
-		$\sigma\_X = 2500, \sigma_Y = 100$ & 1.409094 & 0.0621332 & 1.141402 & 0.0559207 & 1.386553 & 0.0616342 & 1.336668 & 0.0605153\\
-		\hline
-		$\sigma\_X = 10000, \sigma_Y = 500$ & 1.000000 & 0.0523424 & 1.415487 & 0.0622740 & 1.000000 & 0.0523424 & 1.000000 & 0.0523424\\
+		$\sigma_X = 50, \sigma_Y = 2$ & 1.409094 & 0.0621332 & 1.141402 & 0.0559207 & 1.386553 & 0.0616342 & 1.336668 & 0.0605153\\
+		\hline
+		$\sigma_X = 500, \sigma_Y = 20$ & 1.000000 & 0.0523424 & 1.415487 & 0.0622740 & 1.000000 & 0.0523424 & 1.000000 & 0.0523424\\
+		\hline
+		$\sigma_X = 2500, \sigma_Y = 100$ & 1.519223 & 0.0645155 & 1.000000 & 0.0523424 & 1.549862 & 0.0651629 & 1.000000 & 0.0523424\\
+		\hline
+		$\sigma_X = 10000, \sigma_Y = 500$ & 1.127679 & 0.0555836 & 1.000000 & 0.0523424 & 1.540929 & 0.0649748 & 1.000000 & 0.0523424\\
 		\hline
@@
 		\hline
-		T = 200 & 2.428938 & 0.1102030 & 2.444793 & 0.1105620 & 2.420992 & 0.1100225 & 2.353856 & 0.1084863\\
-		\hline
-		T = 100 & 1.000000 & 0.1000000 & 1.000000 & 0.1000000 & 3.363462 & 0.1833974 & 1.000000 & 0.1000000\\
-		\hline
-		T = 50 & 1.000000 & 0.1414214 & 5.598703 & 0.3346252 & 1.000000 & 0.1414214 & 5.390413 & 0.3283417\\
-		\hline
-		T = 20 & 1.000000 & 0.2236068 & 1.000000 & 0.2236068 & 11.379288 & 0.7542973 & 1.000000 & 0.2236068\\
-		\hline
-	\end{tabular}
-\end{table}
-
+		T = 200 & 1.675005 & 0.0677426 & 1.233313 & 0.0581286 & 1.637813 & 0.0669862 & 1.555502 & 0.0652813\\
+		\hline
+		T = 100 & 1.000000 & 0.0523424 & 1.685553 & 0.0679555 & 1.000000 & 0.0523424 & 1.000000 & 0.0523424\\
+		\hline
+		T = 50 & 1.856718 & 0.0713225 & 1.000000 & 0.0523424 & 1.907273 & 0.0722870 & 1.000000 & 0.0523424\\
+		\hline
+		T = 20 & 1.210671 & 0.0575926 & 1.000000 & 0.0523424 & 1.892533 & 0.0720071 & 1.000000 & 0.0523424\\
+		\hline
+	\end{tabular}
+\end{table}
+
+\subsection{Exponential Growth}
+
+\begin{table}
+
+	\caption{\label{tab:}$\rr$ Estimates and Std. Errs, EXPGROWTH Model,
+		$\beta = 0.06, \gamma = 0.03$, $X_0 = 99950, Y_0 = 50$, $\sigma_X = 100, \sigma_Y = 5$}
+	\centering
+	\begin{tabular}[t]{l|r|r|r|r|r|r|r|r}
+		\hline
+		Data Set & Norm Est & Norm SE & Norm-M Est & Norm-M SE & AR Est & AR SE & AR-M Est & AR-M SE\\
+		\hline
+		Baseline1 & 4.421794 & 0.5328029 & 2.280974 & 1.6428406 & 4.442458 & 0.5255489 & 2.609343 & 1.4541148\\
+		\hline
+		Baseline2 & 1.145541 & 0.0014971 & 1.144060 & 0.0015292 & 1.144700 & 0.0015153 & 1.144115 & 0.0015280\\
+		\hline
+		Baseline3 & 1.145483 & 0.0014983 & 1.144182 & 0.0015266 & 1.144805 & 0.0015130 & 1.143958 & 0.0015315\\
+		\hline
+		Baseline4 & 1.145608 & 0.0014956 & 1.144130 & 0.0015277 & 1.144702 & 0.0015152 & 1.143986 & 0.0015309\\
+		\hline
+		Baseline5 & 1.145550 & 0.0014969 & 1.144154 & 0.0015272 & 1.144711 & 0.0015150 & 1.143915 & 0.0015324\\
+		\hline
+	\end{tabular}
+\end{table}
+
+\begin{table}
+
+	\caption{\label{tab:}$\rr$ Estimates and Std. Errs, EXPGROWTH Model,
+		$X_0 = 99950, Y_0 = 50$, $\sigma_X = 100, \sigma_Y = 5$}
+	\centering
+	\begin{tabular}[t]{l|r|r|r|r|r|r|r|r}
+		\hline
+		Data Set & Norm Est & Norm SE & Norm-M Est & Norm-M SE & AR Est & AR SE & AR-M Est & AR-M SE\\
+		\hline
+		$\gamma$ = 0.001 & 1.141015 & 0.0015976 & 1.1349235 & 0.0017432 & 1.139834 & 0.0016246 & 1.1348807 & 0.0017443\\
+		\hline
+		$\gamma$ = 0.04 & 1.133308 & 0.0017845 & 1.1310087 & 0.0018443 & 1.132539 & 0.0018042 & 1.1313483 & 0.0018354\\
+		\hline
+		$\gamma$ = 0.06 & 1.093381 & 0.0031645 & 0.9905347 & 0.0134865 & 1.085491 & 0.0035430 & 0.9899559 & 0.0135941\\
+		\hline
+		$\gamma$ = 0.24 & 1.092493 & 0.0032048 & 0.9206087 & 0.0336647 & 1.085583 & 0.0035384 & 0.8860333 & 0.0506171\\
+		\hline
+	\end{tabular}
+\end{table}
+
+\begin{table}
+
+	\caption{\label{tab:}$\rr$ Estimates and Std. Errs, EXPGROWTH Model,
+		$\beta = 0.06, \gamma = 0.03$, $\sigma_X = 100, \sigma_Y = 5$}
+	\centering
+	\begin{tabular}[t]{l|r|r|r|r|r|r|r|r}
+		\hline
+		Data Set & Norm Est & Norm SE & Norm-M Est & Norm-M SE & AR Est & AR SE & AR-M Est & AR-M SE\\
+		\hline
+		$X_0 = 99000, Y_0 = 1000$ & 1.132250 & 0.0018118 & 1.127900 & 0.0019286 & 1.131379 & 0.0018346 & 1.127954 & 0.0019271\\
+		\hline
+		$X_0 = 99900, Y_0 = 100$ & 1.143191 & 0.0015483 & 1.141440 & 0.0015878 & 1.142493 & 0.0015640 & 1.141621 & 0.0015836\\
+		\hline
+		$X_0 = 99990, Y_0 = 10$ & 1.148203 & 0.0014410 & 1.147130 & 0.0014633 & 1.148093 & 0.0014432 & 1.147197 & 0.0014618\\
+		\hline
+		$X_0 = 99999, Y_0 = 1$ & 1.140076 & 0.0016192 & 1.136283 & 0.0017099 & 1.138602 & 0.0016539 & 1.136402 & 0.0017070\\
+		\hline
+	\end{tabular}
+\end{table}
+
+\begin{table}
+
+	\caption{\label{tab:}$\rr$ Estimates and Std. Errs, EXPGROWTH Model,
+		$\beta = 0.06, \gamma = 0.03$, $X_0 = 99950, Y_0 = 50$}
+	\centering
+	\begin{tabular}[t]{l|r|r|r|r|r|r|r|r}
+		\hline
+		Data Set & Norm Est & Norm SE & Norm-M Est & Norm-M SE & AR Est & AR SE & AR-M Est & AR-M SE\\
+		\hline
+		$\sigma_X = 50, \sigma_Y = 2$ & 1.144401 & 0.0015218 & 1.144116 & 0.0015280 & 1.144199 & 0.0015262 & 1.144083 & 0.0015287\\
+		\hline
+		$\sigma_X = 500, \sigma_Y = 20$ & 1.161010 & 0.0011989 & 1.144155 & 0.0015272 & 1.157193 & 0.0012665 & 1.143996 & 0.0015306\\
+		\hline
+		$\sigma_X = 2500, \sigma_Y = 100$ & 1.203070 & 0.0006567 & 1.144496 & 0.0015197 & 1.195816 & 0.0007284 & 1.147499 & 0.0014556\\
+		\hline
+		$\sigma_X = 10000, \sigma_Y = 500$ & 1.248181 & 0.0003455 & 1.146211 & 0.0014828 & 1.237044 & 0.0004047 & 1.156277 & 0.0012833\\
+		\hline
+	\end{tabular}
+\end{table}
+
+\begin{table}
+
+	\caption{\label{tab:}$\rr$ Estimates and Std. Errs, EXPGROWTH Model,
+		$X_0 = 99950, Y_0 = 50$, $\sigma_X = 100, \sigma_Y = 5$}
+	\centering
+	\begin{tabular}[t]{l|r|r|r|r|r|r|r|r}
+		\hline
+		Data Set & Norm Est & Norm SE & Norm-M Est & Norm-M SE & AR Est & AR SE & AR-M Est & AR-M SE\\
+		\hline
+		1st Order & 1.061431 & 0.0049954 & 1.060632 & 0.0050525 & 1.060995 & 0.0050262 & 1.061004 & 0.0050255\\
+		\hline
+		4th Order & 1.148524 & 0.0014343 & 1.148405 & 0.0014368 & 1.148513 & 0.0014346 & 1.148399 & 0.0014369\\
+		\hline
+		Linear SIR & 1.087741 & 0.0034308 & 1.087338 & 0.0034506 & 1.087538 & 0.0034408 & 1.087399 & 0.0034476\\
+		\hline
+	\end{tabular}
+\end{table}
+
+\begin{table}
+
+	\caption{\label{tab:}$\rr$ Estimates and Std. Errs, EXPGROWTH Model,
+		$\beta = 0.06, \gamma = 0.03$, $X_0 = 99950, Y_0 = 50$, $\sigma_X = 100, \sigma_Y = 5$}
+	\centering
+	\begin{tabular}[t]{l|r|r|r|r|r|r|r|r}
+		\hline
+		Data Set & Norm Est & Norm SE & Norm-M Est & Norm-M SE & AR Est & AR SE & AR-M Est & AR-M SE\\
+		\hline
+		T = 200 & 1.306607 & 0.0047484 & 1.301989 & 0.0049042 & 1.304421 & 0.0048215 & 1.301390 & 0.0049248\\
+		\hline
+		T = 100 & 1.560029 & 0.0278141 & 1.418887 & 0.0419730 & 1.517223 & 0.0315180 & 1.390702 & 0.0455494\\
+		\hline
+		T = 50 & 2.238028 & 0.1132619 & 1.623615 & 0.2357995 & 2.186315 & 0.1209265 & 1.634726 & 0.2329303\\
+		\hline
+		T = 20 & 4.472654 & 0.5150738 & 1.108447 & 1.8062680 & 4.159546 & 0.6317608 & 2.711197 & 1.3929029\\
+		\hline
+	\end{tabular}
+\end{table}
 \end{document}

### 2018-07-03 paper/Version13/_region_.tex
finally on git

--- 2018-07-03 paper/Version12/draft_v12.tex
@@
-\documentclass[12pt]{article}
+\message{ !name(draft_v13.tex)}\documentclass[12pt]{article}
 \usepackage[margin=1in]{geometry}
@@
 
-\newcommand{\xxsir}{\ensuremath{9}} % number of SIR methods
-\newcommand{\wxxsir}{nine }
-\newcommand{\Wxxsir}{Nine }
+\newcommand{\xxsir}{\ensuremath{11} } % number of SIR methods
+\newcommand{\wxxsir}{11 }
+\newcommand{\Wxxsir}{11 }
 
@@
 \begin{document}
+
+\message{ !name(draft_v13.tex) !offset(-3) }
 
@@
 In terms of data from the SIR model, incidence $J(t) \approx -(X(t+1) - X(t))$, and the IPR$(t) = \frac{J(t)}{Y(t)}$.
-This method assumes that we have some prior knowledge about $\gamma$, the transmission rate.
+This method assumes that we have some prior knowledge about $\gamma$, the recovery rate.
 Thus we use as our estimate,
@@
 
-\subsection{Exponential Growth}\label{sec:expgrowth}
+\subsection{Exponential Growth (EG)}\label{sec:expgrowth}
 %% nishiura, majumder, IDEA model -- sum of squares (santillana2016)
@@
 %% chowell 2008 - parametric bootstrap for a model they made up, but sort of empirical.
+Many researchers employ an assumption of exponential growth of the number infectious through times $t=0, 1, \dots, T^* <T$ \citep{wallinga2007generation,fisman2014,nishiura2016,majumder2016,towers2016}.
+\cite{wallinga2007generation} report that the effective reproduction number $\mathcal{R}_t$ and hence the initial reproduction number $\rr$ may derived using the fact that infection ``counts increase exponentially in the initial phase of an epidemic.''  We then have to estimate $r$, the \textit{per capita} change in the number of new cases per unit of time and $\omega$ the serial interval, the distribution of time between a primary and secondary infection.
+Then, we have
+\begin{align}\label{eq:lotka}
+\rr = \exp{(r \omega)}.
+\end{align}
+Equation \eqref{eq:lotka} is derived from a demographic view using the Lotka-Euler survival equations which come from the fields of demography, ecology, and evolutionary biology.
+
+\cite{wallinga2007generation}, instead, use a moment generating function expression for $\rr$.  With $\omega(t)$ as the serial interval, then
+\begin{align*}
+\rr^{-1} &= \frac{1}{M(-r)} = \int_{a=0}^\infty e^{-rt}\omega(t)dt.
+\end{align*}
+\cite{wallinga2007generation} specifically derive an estimate for $\rr$ for the SIR model.
+They assume that the rate of leaving the infectious stsate, $\gamma$ is constant as is the rate of making contacts during the infectious stage.
+Then, the duration of a serial interval is an exponential distribution with mean $\gamma^{-1}$, the moment generating function is then,
+\begin{align*}
+  M(-r) = \frac{\gamma}{\gamma + r},
+\end{align*}
+and the resulting estimate of $\rr$ is
+\begin{align}\label{eq:r0-eg}
+  \hat{\rr} = 1 + \frac{\hat{r}}{\hat{\gamma}},
+\end{align}
+where $\hat{r}$ and $\hat{\gamma}$ are estimates of the $r$ and $\gamma$, respectively.
+The parameter $\hat{\gamma}$ is often assumed to be fixed or tried with a sequence of serial intervals \citep{majumder2016}.
+We may estimate $\hat{r}$ using ordinary linear regression, namely,
+\begin{align*}
+  \log \left  (\frac{Y(t)}{Y(0)} \right )&= rt + \epsilon_t\\
+  \epsilon_t &\sim N(0, \sigma^2),
+\end{align*}
+which results in
+\begin{align}\label{eq:exp-growth}
+  \hat{r} &= \frac{\sum_{t=0}^{T^*}\left  (\frac{Y(t)}{Y(0)} \right )}{\sum_{t=0}^{T^*}t}.
+\end{align}
+
+For the above estimate of $\rr$, we observe the duration of serial intervals in a period of exponential growth of the number of infectious individuals.
+Deciding when exponential growth occurs and which data points to use to estimate $\rr$ may be difficult.
+We can use sensitivity analysis to see how the estimate $\rr$ changes as we increase or decrease $T^{*}$, the number of time points used.
+The reader may notice some similarities to the estimate for $\rr$ in Equation \eqref{eq:r0-eg} and from the log-linear model in Equation \eqref{eq:harko_lin}.
+This method is a simplification of the SIR model as the $R$ compartment is not used to estimate $r$, only the recovery rate $\gamma$.
+
+An advantage to this method are its (and similar adaptations) prevalence in the literature, which allows one to to compare estimates $\rr$ directly to one another.
+Another advantage is its flexibility.
+Here, we assumed the serial interval has an exponential distribution, but this is not a requirement and other distributions may be used (see \cite{wallinga2007generation}).
+However, a disadvantage to this method is that is a simplification of the SIR model and we have to decide how many time points to use to estimate the exponential growth rate of infectious.
+Another disadvantage to this method is that one also has to decide how to treat the additional paramter $\gamma$, whether to treat it as fixed, a range of fixed values, or even having a probability distribution.
+
+\subsection{Incidence Decay and Exponential Adjustment (IDEA)}\label{sec:idea}
+The Incidence Decay and Exponential Adjustment (IDEA) model is a way to account for a declining effective reproduction number of time due to both depletion of susceptible individuals and spontaneous and planned control activities and behaviors'' \citep{fisman2013}.
+They model the incidence counts during the $t$th interval, $J(t),$ as
+\begin{align*}
+  J(t) &= \left [ \frac{\rr}{(1 + d)^t}\right ]^t,
+\end{align*}
+where $d$ is a dampening or discount factor.  As $d$ increases, the incidence at time $t$  decreases. The estimate for $\rr$ is obtained by minimizing the sum of squares over $\rr$ and $d$,
+\begin{align*}
+  \hat{\rr} &=  \arg \min_{\rr,d} \sum_{t=0}^{T^*}\left (J_{obs}(t) - J(t) \right)^2.
+\end{align*}
+The method presented in \cite{fisman2013} is a mathematical, rather than statistical model.
+Sensitivity analysis is conducted with respect to $T^*$, $d$, and $\rr$, and lower and upper bounds are reported as a result of this analysis.
+
+Advantages of this method include its usefulness in estimating $\rr$ \citep{fisman2014,majumder2016}.
+Additionally, this model may be employed to address the effect of control measures and better estimate the final size of an epidemic.
+A disadvantage to this model is that its upper and lower bounds do not correspond to CIs.
+Also, in this method we must select the interval size of the incidence counts.
 
@@
 We compare the performance of the SIR models on simulated data from both the SIR as well as other models.
-In all of these simulations, we first generate data from the models under known conditions, then we add on errors with known variances to each time point,
+In all of these simulations, we first generate data from the models under known conditions, then we add on errors drawn from known distributions variances to each time point,
 \begin{align}\label{eq:sim-models}
@@
 \end{align*}
-This gives us a total of four error structures for each set of conditions: Gaussian (Norm), Gaussian monotone (Norm-M), autoregressive (AR), and autoregressive monotone (AR-M) for a given set of parameters $(\beta, \gamma, O, X(0), Y(0), Z(0), \sigma_X, \sigma_Y, N)$.
+This gives us a total of four error structures for each set of conditions: Gaussian (Norm), Gaussian monotone (Norm-M), autoregressive (AR), and autoregressive monotone (AR-M) for a given set of parameters $(\beta, \gamma, O, X(0), Y(0), Z(0), \sigma_X, \sigma_Y, N)$, the infection rate; recovery rate; total number of time points; initial conditions for number of susceptibles, infectious, and recovered individuals, the standard deviation for both the susceptible and infectious compartments, and the total population, respectively.
 
@@
 
-We first simulate 5 ``baseline'' data sets from the SIR model under the following conditions:
-
+We first simulate a ``baseline'' data set from the SIR model under the following conditions:
+
+\textbf{Baseline Dataset}
 \begin{center}
@@
 
-	Total population $N$: 100000.
+	Total population $N$: 100000
+
+        Error: Gaussian.
 
@@
 
-We then generate different SIR data sets, changing the one of the conditions for simulation each time and using the baseline SIR data as our basis for comparison.
-The different data sets generate are shown in the following table:
-
-\begin{table}[h]
-	\centering
-	\resizebox{\textwidth}{!}{
-		\begin{tabular}{@{}lll@{}}
-			\toprule
-                  \textbf{Parameter Values ($\beta, \gamma$)}        & \textbf{Starting Compartment Sizes ($X_0, Y_0$)}            & \textbf{Variance of Added Errors ($\sigma_X, \sigma_Y$ )}       \\ \midrule
-                  (0.06, 0.001)            & (99000, 1000)                           & (50, 2)                     \\
-                  (0.06, 0.04)             & (99900, 100)               & (500, 20)     \\
-                  (0.06,.03) & (99950, 50) & (100, 5)\\
-                  (0.06, 0.06)             & (99990, 10)             & (2500, 100)           \\
-                  (0.06, 0.24) & (99999, 1) & (10000, 500)                                      \\ \bottomrule
-		\end{tabular}
-	}
-	\caption{Table of data sets.}
-	\label{tab:datasets}
-\end{table}
+This baseline data set
+
+% We then generate different SIR data sets, changing the one of the conditions for simulation each time and using the baseline SIR data as our basis for comparison.
+The different data sets generated are shown in the following table:
+
+% \begin{table}[h]
+% 	\centering
+% 	\resizebox{\textwidth}{!}{
+% 		\begin{tabular}{@{}lll@{}}
+% 			\toprule
+%                   \textbf{Parameter Values ($\beta, \gamma$)}        & \textbf{Starting Compartment Sizes ($X_0, Y_0$)}            & \textbf{Variance of Added Errors ($\sigma_X, \sigma_Y$ )}       \\ \midrule
+%                   (0.06, 0.001)            & (99000, 1000)                           & (50, 2)                     \\
+%                   (0.06, 0.04)             & (99900, 100)               & (500, 20)     \\
+%                   (0.06,.03) & (99950, 50) & (100, 5)\\
+%                   (0.06, 0.06)             & (99990, 10)             & (2500, 100)           \\
+%                   (0.06, 0.24) & (99999, 1) & (10000, 500)                                      \\ \bottomrule
+% 		\end{tabular}
+% 	}
+% 	\caption{Table of data sets.}
+% 	\label{tab:datasets}
+% \end{table}
 
@@
 Among these include values of both the infection and recovery rate $(\beta, \gamma)$, the number of data points available (time $T$), the total size of the population ($N$), the initial percent of susceptible individuals $\left (\frac{X(0)}{N}\right)$, the initial percent of infectious individuals $\left (\frac{Y(0)}{N}\right )$, and  the magnitude of the variance of both the number of susceptibles and the number of infectious ($\sigma_X, \sigma_Y$).
-We examine each of these aspects individually.
+We first examine our baseline data, and then we examine each of these aspects individually.
 
@@
 
-\begin{table}[H]
-	\includegraphics[scale=0.5]{BaseBase.jpeg}
-
-	\caption{\label{tab:}Baseline Data, $\rr$ Estimates and Standard Errors}
+\begin{figure}
+  \centering
+  \includegraphics[scale=0.5]{images/BaseBase.jpeg}
+  \caption{Forest plots of estimates of $\rr$ for the \xxsir methods for the baseline data set $(\beta=.06, \gamma=.03, O=365, X(0)=99950, Y(0)=50, Z(0)=0, \sigma_X=100, \sigma_Y=5, N=10^5)$.
+The points are the estimates of $\rr$ and the lines denote $\pm 2\cdot $Std.
+Err.
+The red line is the true $\rr$ value.}
+  \end{figure}
+
+\begin{table}[H]	
 	\centering
@@
 	\end{tabular}
+        \caption{Baseline Data $(\beta=.06, \gamma=.03, O=365, X(0)=99950, Y(0)=50, Z(0)=0, \sigma_X=100, \sigma_Y=5, N=10^5)$, $\rr$ Estimates and Standard Errors}\label{tab:baseline}
 \end{table}
@@
 
-\includegraphics[scale=0.5]{AR.jpeg}
+\includegraphics[scale=0.5]{images/AR.jpeg}
 \begin{table}[H]
@@
 
-\textbf{Need to figure out incorporating different $/rr$ scales in to graph. Facets won't work.}
+\includegraphics[scale=0.5]{images/parchange.jpeg}
 
@@
 
-\includegraphics[scale=0.5]{time.jpeg}
+\includegraphics[scale=0.5]{images/time.jpeg}
 
@@
 
-\includegraphics[scale=0.5]{start.jpeg}
+\includegraphics[scale=0.5]{images/start.jpeg}
 
@@
 
-\includegraphics[scale=0.5]{var.jpeg}
+\includegraphics[scale=0.5]{images/var.jpeg}
 
@@
   \begin{center}
-\includegraphics[scale=0.5]{other.jpeg}
+\includegraphics[scale=0.5]{images/other.jpeg}
 
@@
 	\caption{\label{tab:}$\rr$ Estimates and Std. Errs, SIRLS Model,
-		$X_0 = 99950, Y_0 = 50$, $\sigma_X = 100, \sigma_Y = 5$,$\beta = 0.06, \gamma = 0.03$}
-	\centering
-	\begin{footnotesize}
-		\begin{tabular}[t]{l|r|r|r|r|r|r|r|r}
-			\hline
-			Data Set & Norm Est & Norm SE & Norm-M Est & Norm-M SE & AR Est & AR SE & AR-M Est & AR-M SE\\
-			\hline
-			Baseline1 & 1.999919 & 0.0058322 & 2.000036 & 0.0044639 & 2.000020 & 0.0078308 & 2.000859 & 0.0060766\\
-			\hline
-			Baseline2 & 1.999971 & 0.0055923 & 2.000246 & 0.0044456 & 1.999739 & 0.0071509 & 1.999992 & 0.0073664\\
-			\hline
-			Baseline3 & 2.000327 & 0.0054549 & 1.999861 & 0.0048111 & 1.999165 & 0.0078453 & 2.000160 & 0.0063235\\
-			\hline
-			Baseline4 & 2.000520 & 0.0057238 & 1.999501 & 0.0046540 & 1.999282 & 0.0077363 & 1.999289 & 0.0065290\\
-			\hline
-			Baseline5 & 1.999919 & 0.0056334 & 2.000053 & 0.0045020 & 2.001443 & 0.0081239 & 2.002397 & 0.0075056\\
-			\hline
-		\end{tabular}
-	\end{footnotesize}
-\end{table}
-
-\begin{table}[H]
-
-	\caption{\label{tab:}$\rr$ Estimates and Std. Errs, SIRLS Model,
 		$X_0 = 99950, Y_0 = 50$, $\sigma_X = 100, \sigma_Y = 5$, $\beta = 0.06$}
@@
 	\caption{\label{tab:}$\rr$ Estimates and Std. Errs, SIRREPAR Model,
-		$X_0 = 99950, Y_0 = 50$, $\sigma_X = 100, \sigma_Y = 5$,$\beta = 0.06, \gamma = 0.03$}
-	\centering
-	\begin{footnotesize}
-		\begin{tabular}[t]{l|r|r|r|r|r|r|r|r}
-			\hline
-			Data Set & Norm Est & Norm SE & Norm-M Est & Norm-M SE & AR Est & AR SE & AR-M Est & AR-M SE\\
-			\hline
-			Baseline1 & 1.999708 & 0.0052113 & 2.000030 & 0.0039910 & 1.999885 & 0.0069901 & 2.001023 & 0.0054369\\
-			\hline
-			Baseline2 & 1.999734 & 0.0049939 & 2.000314 & 0.0039684 & 2.000046 & 0.0063859 & 1.999684 & 0.0065745\\
-			\hline
-			Baseline3 & 2.000604 & 0.0048757 & 2.000025 & 0.0042948 & 1.999356 & 0.0069903 & 2.000438 & 0.0056503\\
-			\hline
-			Baseline4 & 2.000410 & 0.0051162 & 1.999830 & 0.0041504 & 1.999885 & 0.0069068 & 1.999013 & 0.0058261\\
-			\hline
-			Baseline5 & 1.999830 & 0.0050288 & 1.999830 & 0.0040212 & 2.001592 & 0.0072630 & 2.002090 & 0.0067284\\
-			\hline
-		\end{tabular}
-	\end{footnotesize}
-\end{table}
-
-\begin{table}[H]
-
-	\caption{\label{tab:}$\rr$ Estimates and Std. Errs, SIRREPAR Model,
 		$X_0 = 99950, Y_0 = 50$, $\sigma_X = 100, \sigma_Y = 5$, $\beta = 0.06$}
@@
 	\caption{\label{tab:}$\rr$ Estimates and Std. Errs, SIRLMA Model,
-		$X_0 = 99950, Y_0 = 50$, $\sigma_X = 100, \sigma_Y = 5$,$\beta = 0.06, \gamma = 0.03$}
-	\centering
-	\begin{footnotesize}
-		\begin{tabular}[t]{l|r|r|r|r|r|r|r|r}
-			\hline
-			Data Set & Norm Est & Norm SE & Norm-M Est & Norm-M SE & AR Est & AR SE & AR-M Est & AR-M SE\\
-			\hline
-			Baseline 1 & 1.997416 & 0.3900006 & 2.692152 & 0.5081138 & 35.656866 & 4815.2163162 & 1.667689 & 0.1718689\\
-			\hline
-			Baseline 2 & 1.686364 & 0.1886132 & 1.679359 & 0.1102722 & 1.474460 & 0.1130752 & 1.534979 & 0.1496990\\
-			\hline
-			Baseline 3 & 1.572708 & 0.1340966 & 1.743175 & 0.1667602 & 5.345243 & 33.1125597 & 2.313202 & 0.6045124\\
-			\hline
-			Baseline 4 & 1.640228 & 0.1956211 & 2.046183 & 0.3959176 & 1.327043 & 0.0671306 & -3.008847 & 7.5919771\\
-			\hline
-			Baseline 5 & 1.590702 & 0.1685914 & 1.842150 & 0.2204031 & 1.488547 & 0.1285866 & 1.868254 & 0.2217948\\
-			\hline
-		\end{tabular}
-	\end{footnotesize}
-\end{table}
-
-\begin{table}[H]
-
-	\caption{\label{tab:}$\rr$ Estimates and Std. Errs, SIRLMA Model,
 		$X_0 = 99950, Y_0 = 50$, $\sigma_X = 100, \sigma_Y = 5$, $\beta = 0.06$}
@@
 	\caption{\label{tab:}$\rr$ Estimates and Std. Errs, SIRLMAT Model,
-		$X_0 = 99950, Y_0 = 50$, $\sigma_X = 100, \sigma_Y = 5$,$\beta = 0.06, \gamma = 0.03$}
-	\centering
-	\begin{footnotesize}
-		\begin{tabular}[t]{l|r|r|r|r|r|r|r|r}
-			\hline
-			Data Set & Norm Est & Norm SE & Norm-M Est & Norm-M SE & AR Est & AR SE & AR-M Est & AR-M SE\\
-			\hline
-			Baseline 1 & 1.398016 & 0.6028170 & 1.421201 & 0.7403641 & 1.049360 & 4.8830065 & 1.401437 & 0.8194417\\
-			\hline
-			Baseline 2 & 1.406323 & 0.6309170 & 1.442263 & 0.9241642 & 1.369425 & 0.8681724 & 1.449109 & 0.7463529\\
-			\hline
-			Baseline 3 & 1.408807 & 0.6215050 & 1.358127 & 0.8493445 & 1.375713 & 1.2574743 & 1.436717 & 0.6726785\\
-			\hline
-			Baseline 4 & 1.387819 & 0.6177830 & 1.390251 & 0.5728970 & 1.596744 & 1.3603545 & 1.325567 & 0.6545989\\
-			\hline
-			Baseline 5 & 1.387599 & 0.6657501 & 1.386679 & 0.5751290 & 1.359862 & 1.0018556 & 1.538778 & 0.9068372\\
-			\hline
-		\end{tabular}
-	\end{footnotesize}
-\end{table}
-
-\begin{table}[H]
-
-	\caption{\label{tab:}$\rr$ Estimates and Std. Errs, SIRLMAT Model,
 		$X_0 = 99950, Y_0 = 50$, $\sigma_X = 100, \sigma_Y = 5$, $\beta = 0.06$}
@@
 	\caption{\label{tab:}$\rr$ Estimates and Std. Errs, SIRIPR Model,
-		$X_0 = 99950, Y_0 = 50$, $\sigma_X = 100, \sigma_Y = 5$,$\beta = 0.06, \gamma = 0.03$}
-	\centering
-	\begin{footnotesize}
-		\begin{tabular}[t]{l|r|r|r|r|r|r|r|r}
-			\hline
-			Data Set & Norm Est & Norm SE & Norm-M Est & Norm-M SE & AR Est & AR SE & AR-M Est & AR-M SE\\
-			\hline
-			Baseline 1 & 2.592974 & 9.040841 & 0.7282379 & 1.286539 & 2.183371 & 7.397726 & 0.7645942 & 1.862201\\
-			\hline
-			Baseline 2 & 2.210182 & 6.179671 & 0.7485576 & 1.504683 & 1.712691 & 5.071385 & 0.7513498 & 1.560614\\
-			\hline
-			Baseline 3 & 2.306256 & 6.684895 & 0.7206976 & 1.230857 & 1.718776 & 4.593769 & 0.8374593 & 2.542060\\
-			\hline
-			Baseline 4 & 2.547665 & 9.843667 & 0.8478792 & 2.387153 & 1.817153 & 5.771054 & 0.8604847 & 2.307269\\
-			\hline
-			Baseline 5 & 2.224111 & 6.227286 & 0.8080670 & 2.043471 & 1.859343 & 4.996316 & 0.7686838 & 2.308889\\
-			\hline
-		\end{tabular}
-	\end{footnotesize}
-\end{table}
-
-\begin{table}[H]
-
-	\caption{\label{tab:}$\rr$ Estimates and Std. Errs, SIRIPR Model,
 		$X_0 = 99950, Y_0 = 50$, $\sigma_X = 100, \sigma_Y = 5$, $\beta = 0.06$}
@@
 	\caption{\label{tab:}$\rr$ Estimates and Std. Errs, SIRSIPR Model,
-		$X_0 = 99950, Y_0 = 50$, $\sigma_X = 100, \sigma_Y = 5$,$\beta = 0.06, \gamma = 0.03$}
-	\centering
-	\begin{footnotesize}
-		\begin{tabular}[t]{l|r|r|r|r|r|r|r|r}
-			\hline
-			Data Set & Norm Est & Norm SE & Norm-M Est & Norm-M SE & AR Est & AR SE & AR-M Est & AR-M SE\\
-			\hline
-			Baseline 1 & 0.9251762 & $<$ 1e-07 & 0.8588266 & $<$ 1e-07 & 0.8386350 & $<$ 1e-07 & 0.7745969 & $<$ 1e-07\\
-			\hline
-			Baseline 2 & 0.9435412 & $<$ 1e-07 & 0.9081552 & $<$ 1e-07 & 0.8792348 & $<$ 1e-07 & 0.6976176 & $<$ 1e-07\\
-			\hline
-			Baseline 3 & 0.9694855 & $<$ 1e-07 & 0.9183543 & $<$ 1e-07 & 1.0065755 & $<$ 1e-07 & 0.7129407 & $<$ 1e-07\\
-			\hline
-			Baseline 4 & 0.9811143 & $<$ 1e-07 & 0.8743186 & $<$ 1e-07 & 0.8665770 & $<$ 1e-07 & 0.9207903 & $<$ 1e-07\\
-			\hline
-			Baseline 5 & 0.9276948 & $<$ 1e-07 & 0.9107829 & $<$ 1e-07 & 1.1514790 & $<$ 1e-07 & 0.7694052 & $<$ 1e-07\\
-			\hline
-		\end{tabular}
-	\end{footnotesize}
-\end{table}
-
-\begin{table}[H]
-
-	\caption{\label{tab:}$\rr$ Estimates and Std. Errs, SIRSIPR Model,
 		$X_0 = 99950, Y_0 = 50$, $\sigma_X = 100, \sigma_Y = 5$, $\beta = 0.06$}
@@
 	\caption{\label{tab:}$\rr$ Estimates and Std. Errs, LOGLINEAR Model,
-		$X_0 = 99950, Y_0 = 50$, $\sigma_X = 100, \sigma_Y = 5$,$\beta = 0.06, \gamma = 0.03$}
-	\centering
-	\begin{footnotesize}
-		\begin{tabular}[t]{l|r|r|r|r|r|r|r|r}
-			\hline
-			Data Set & Norm Est & Norm SE & Norm-M Est & Norm-M SE & AR Est & AR SE & AR-M Est & AR-M SE\\
-			\hline
-			Baseline1 & 1.999840 & 0.0002227 & 1.999761 & 0.0001427 & 1.999820 & 0.0003227 & 2.000443 & 0.0001910\\
-			\hline
-			Baseline2 & 1.999807 & 0.0002198 & 2.000303 & 0.0001342 & 2.000000 & 0.0002885 & 1.999556 & 0.0002342\\
-			\hline
-			Baseline3 & 2.000386 & 0.0002191 & 2.000042 & 0.0001598 & 1.998138 & 0.0003215 & 1.999213 & 0.0002316\\
-			\hline
-			Baseline4 & 2.000039 & 0.0002252 & 1.999710 & 0.0001506 & 1.999217 & 0.0002895 & 1.999525 & 0.0002144\\
-			\hline
-			Baseline5 & 2.000148 & 0.0002194 & 1.999993 & 0.0001479 & 2.001889 & 0.0003495 & 2.001873 & 0.0002438\\
-			\hline
-		\end{tabular}
-	\end{footnotesize}
-\end{table}
-
-\begin{table}[H]
-
-	\caption{\label{tab:}$\rr$ Estimates and Std. Errs, LOGLINEAR Model,
 		$X_0 = 99950, Y_0 = 50$, $\sigma_X = 100, \sigma_Y = 5$, $\beta = 0.06$}
@@
 	\caption{\label{tab:}$\rr$ Estimates and Std. Errs, SIRMC Model,
-		$X_0 = 99950, Y_0 = 50$, $\sigma_X = 100, \sigma_Y = 5$,$\beta = 0.06, \gamma = 0.03$}
-	\centering
-	\begin{footnotesize}
-		\begin{tabular}[t]{l|r|r|r|r|r|r|r|r}
-			\hline
-			Data Set & Norm Est & Norm SE & Norm-M Est & Norm-M SE & AR Est & AR SE & AR-M Est & AR-M SE\\
-			\hline
-			Baseline1 & 0.9485555 & $<$ 1e-07 & 0.9485555 & $<$ 1e-07 & 0.9485555 & $<$ 1e-07 & 0.9485555 & $<$ 1e-07\\
-			\hline
-			Baseline2 & 0.9485555 & $<$ 1e-07 & 0.9485555 & $<$ 1e-07 & 0.9485555 & $<$ 1e-07 & 0.9485555 & $<$ 1e-07\\
-			\hline
-			Baseline3 & 0.9485555 & $<$ 1e-07 & 0.9485555 & $<$ 1e-07 & 0.9485555 & $<$ 1e-07 & 0.9485555 & $<$ 1e-07\\
-			\hline
-			Baseline4 & 0.9485555 & $<$ 1e-07 & 0.9485555 & $<$ 1e-07 & 0.9485555 & $<$ 1e-07 & 0.9485555 & $<$ 1e-07\\
-			\hline
-			Baseline5 & 0.9485555 & $<$ 1e-07 & 0.9485555 & $<$ 1e-07 & 0.9485555 & $<$ 1e-07 & 0.9485555 & $<$ 1e-07\\
-			\hline
-		\end{tabular}
-	\end{footnotesize}
-\end{table}
-
-\begin{table}[H]
-
-	\caption{\label{tab:}$\rr$ Estimates and Std. Errs, SIRMC Model,
 		$X_0 = 99950, Y_0 = 50$, $\sigma_X = 100, \sigma_Y = 5$, $\beta = 0.06$}
@@
 	\caption{\label{tab:}$\rr$ Estimates and Std. Errs, SEQBAYES Model,
-		$X_0 = 99950, Y_0 = 50$, $\sigma_X = 100, \sigma_Y = 5$, $\beta = 0.06, \gamma = 0.03$}
-	\centering
-	\begin{footnotesize}
-	\begin{tabular}[t]{l|r|r|r|r|r|r|r|r}
-		\hline
-		Data Set & Norm Est & Norm SE & Norm-M Est & Norm-M SE & AR Est & AR SE & AR-M Est & AR-M SE\\
-		\hline
-		Baseline1 & 1.0000000 & 0.0523424 & 0.8099669 & 0.0471072 & 1.0000000 & 0.0523424 & 1.000000 & 0.0523424\\
-		\hline
-		Baseline2 & 1.3936378 & 0.0617915 & 1.0000000 & 0.0523424 & 1.0000000 & 0.0523424 & 1.432940 & 0.0626567\\
-		\hline
-		Baseline3 & 1.4504175 & 0.0630377 & 1.1185890 & 0.0553591 & 1.4331404 & 0.0626611 & 1.000000 & 0.0523424\\
-		\hline
-		Baseline4 & 1.0000000 & 0.0523424 & 1.3993848 & 0.0619187 & 0.8221337 & 0.0474597 & 1.386606 & 0.0616354\\
-		\hline
-		Baseline5 & 0.9935711 & 0.0521739 & 1.4389154 & 0.0627872 & 1.3236479 & 0.0602199 & 1.014455 & 0.0527193\\
-		\hline
-	\end{tabular}
-\end{footnotesize}
-\end{table}
-
-\begin{table}[H]
-
-	\caption{\label{tab:}$\rr$ Estimates and Std. Errs, SEQBAYES Model,
 		$X_0 = 99950, Y_0 = 50$, $\sigma_X = 100, \sigma_Y = 5$, $\beta = 0.06$}
@@
 
-\begin{table}
+\begin{table}[H]
 
@@
 		\hline
-		Baseline1 & 4.421794 & 0.5328029 & 2.280974 & 1.6428406 & 4.442458 & 0.5255489 & 2.609343 & 1.4541148\\
-		\hline
-		Baseline2 & 1.145541 & 0.0014971 & 1.144060 & 0.0015292 & 1.144700 & 0.0015153 & 1.144115 & 0.0015280\\
-		\hline
-		Baseline3 & 1.145483 & 0.0014983 & 1.144182 & 0.0015266 & 1.144805 & 0.0015130 & 1.143958 & 0.0015315\\
-		\hline
-		Baseline4 & 1.145608 & 0.0014956 & 1.144130 & 0.0015277 & 1.144702 & 0.0015152 & 1.143986 & 0.0015309\\
-		\hline
-		Baseline5 & 1.145550 & 0.0014969 & 1.144154 & 0.0015272 & 1.144711 & 0.0015150 & 1.143915 & 0.0015324\\
-		\hline
-	\end{tabular}
-\end{table}
-
-\begin{table}
+		$\gamma$ = 0.001 & 1.141015 & 0.0001503 & 1.1349235 & 0.0001645 & 1.139834 & 0.0001529 & 1.1348807 & 0.0001646\\
+		\hline
+		$\gamma$ = 0.04 & 1.133308 & 0.0001686 & 1.1310087 & 0.0001744 & 1.132539 & 0.0001705 & 1.1313483 & 0.0001735\\
+		\hline
+		$\gamma$ = 0.06 & 1.093381 & 0.0003053 & 0.9905347 & 0.0013816 & 1.085491 & 0.0003433 & 0.9899559 & 0.0013931\\
+		\hline
+		$\gamma$ = 0.24 & 1.092493 & 0.0003093 & 0.9206087 & 0.0036121 & 1.085583 & 0.0003428 & 0.8860333 & 0.0055673\\
+		\hline
+	\end{tabular}
+\end{table}
+
+\begin{table}[H]
+
+	\caption{\label{tab:}$\rr$ Estimates and Std. Errs, EXPGROWTH Model,
+		$\beta = 0.06, \gamma = 0.03$, $X_0 = 99950, Y_0 = 50$, $\sigma_X = 100, \sigma_Y = 5$}
+	\centering
+	\begin{tabular}[t]{l|r|r|r|r|r|r|r|r}
+		\hline
+		Data Set & Norm Est & Norm SE & Norm-M Est & Norm-M SE & AR Est & AR SE & AR-M Est & AR-M SE\\
+		\hline
+		$X_0 = 99000, Y_0 = 1000$ & 1.132250 & 0.0001712 & 1.127900 & 0.0001827 & 1.131379 & 0.0001734 & 1.127954 & 0.0001825\\
+		\hline
+		$X_0 = 99900, Y_0 = 100$ & 1.143191 & 0.0001455 & 1.141440 & 0.0001494 & 1.142493 & 0.0001470 & 1.141621 & 0.0001489\\
+		\hline
+		$X_0 = 99990, Y_0 = 10$ & 1.148203 & 0.0001351 & 1.147130 & 0.0001372 & 1.148093 & 0.0001353 & 1.147197 & 0.0001371\\
+		\hline
+		$X_0 = 99999, Y_0 = 1$& 1.140076 & 0.0001524 & 1.136283 & 0.0001612 & 1.138602 & 0.0001558 & 1.136402 & 0.0001610\\
+		\hline
+	\end{tabular}
+\end{table}
+
+\begin{table}[H]
+
+	\caption{\label{tab:}$\rr$ Estimates and Std. Errs, EXPGROWTH Model,
+		$\beta = 0.06, \gamma = 0.03$, $X_0 = 99950, Y_0 = 50$}
+	\centering
+	\begin{tabular}[t]{l|r|r|r|r|r|r|r|r}
+		\hline
+		Data Set & Norm Est & Norm SE & Norm-M Est & Norm-M SE & AR Est & AR SE & AR-M Est & AR-M SE\\
+		\hline
+		$\sigma_X = 50, \sigma_Y = 2$ & 1.144401 & 0.0001429 & 1.144116 & 0.0001435 & 1.144199 & 0.0001433 & 1.144083 & 0.0001436\\
+		\hline
+		$\sigma_X = 500, \sigma_Y = 20$ & 1.161010 & 0.0001117 & 1.144155 & 0.0001434 & 1.157193 & 0.0001182 & 1.143996 & 0.0001438\\
+		\hline
+		$\sigma_X = 2500, \sigma_Y = 100$ & 1.203070 & 0.0000599 & 1.144496 & 0.0001427 & 1.195816 & 0.0000667 & 1.147499 & 0.0001365\\
+		\hline
+		$\sigma_X = 10000, \sigma_Y = 500$ & 1.248181 & 0.0000309 & 1.146211 & 0.0001391 & 1.237044 & 0.0000364 & 1.156277 & 0.0001198\\
+		\hline
+	\end{tabular}
+\end{table}
+
+\begin{table}[H]
 
@@
 		\hline
-		$\gamma$ = 0.001 & 1.141015 & 0.0015976 & 1.1349235 & 0.0017432 & 1.139834 & 0.0016246 & 1.1348807 & 0.0017443\\
-		\hline
-		$\gamma$ = 0.04 & 1.133308 & 0.0017845 & 1.1310087 & 0.0018443 & 1.132539 & 0.0018042 & 1.1313483 & 0.0018354\\
-		\hline
-		$\gamma$ = 0.06 & 1.093381 & 0.0031645 & 0.9905347 & 0.0134865 & 1.085491 & 0.0035430 & 0.9899559 & 0.0135941\\
-		\hline
-		$\gamma$ = 0.24 & 1.092493 & 0.0032048 & 0.9206087 & 0.0336647 & 1.085583 & 0.0035384 & 0.8860333 & 0.0506171\\
-		\hline
-	\end{tabular}
-\end{table}
-
-\begin{table}
-
-	\caption{\label{tab:}$\rr$ Estimates and Std. Errs, EXPGROWTH Model,
-		$\beta = 0.06, \gamma = 0.03$, $\sigma_X = 100, \sigma_Y = 5$}
-	\centering
-	\begin{tabular}[t]{l|r|r|r|r|r|r|r|r}
-		\hline
-		Data Set & Norm Est & Norm SE & Norm-M Est & Norm-M SE & AR Est & AR SE & AR-M Est & AR-M SE\\
-		\hline
-		$X_0 = 99000, Y_0 = 1000$ & 1.132250 & 0.0018118 & 1.127900 & 0.0019286 & 1.131379 & 0.0018346 & 1.127954 & 0.0019271\\
-		\hline
-		$X_0 = 99900, Y_0 = 100$ & 1.143191 & 0.0015483 & 1.141440 & 0.0015878 & 1.142493 & 0.0015640 & 1.141621 & 0.0015836\\
-		\hline
-		$X_0 = 99990, Y_0 = 10$ & 1.148203 & 0.0014410 & 1.147130 & 0.0014633 & 1.148093 & 0.0014432 & 1.147197 & 0.0014618\\
-		\hline
-		$X_0 = 99999, Y_0 = 1$ & 1.140076 & 0.0016192 & 1.136283 & 0.0017099 & 1.138602 & 0.0016539 & 1.136402 & 0.0017070\\
-		\hline
-	\end{tabular}
-\end{table}
-
-\begin{table}
-
-	\caption{\label{tab:}$\rr$ Estimates and Std. Errs, EXPGROWTH Model,
-		$\beta = 0.06, \gamma = 0.03$, $X_0 = 99950, Y_0 = 50$}
-	\centering
-	\begin{tabular}[t]{l|r|r|r|r|r|r|r|r}
-		\hline
-		Data Set & Norm Est & Norm SE & Norm-M Est & Norm-M SE & AR Est & AR SE & AR-M Est & AR-M SE\\
-		\hline
-		$\sigma_X = 50, \sigma_Y = 2$ & 1.144401 & 0.0015218 & 1.144116 & 0.0015280 & 1.144199 & 0.0015262 & 1.144083 & 0.0015287\\
-		\hline
-		$\sigma_X = 500, \sigma_Y = 20$ & 1.161010 & 0.0011989 & 1.144155 & 0.0015272 & 1.157193 & 0.0012665 & 1.143996 & 0.0015306\\
-		\hline
-		$\sigma_X = 2500, \sigma_Y = 100$ & 1.203070 & 0.0006567 & 1.144496 & 0.0015197 & 1.195816 & 0.0007284 & 1.147499 & 0.0014556\\
-		\hline
-		$\sigma_X = 10000, \sigma_Y = 500$ & 1.248181 & 0.0003455 & 1.146211 & 0.0014828 & 1.237044 & 0.0004047 & 1.156277 & 0.0012833\\
-		\hline
-	\end{tabular}
-\end{table}
-
-\begin{table}
-
-	\caption{\label{tab:}$\rr$ Estimates and Std. Errs, EXPGROWTH Model,
-		$X_0 = 99950, Y_0 = 50$, $\sigma_X = 100, \sigma_Y = 5$}
-	\centering
-	\begin{tabular}[t]{l|r|r|r|r|r|r|r|r}
-		\hline
-		Data Set & Norm Est & Norm SE & Norm-M Est & Norm-M SE & AR Est & AR SE & AR-M Est & AR-M SE\\
-		\hline
-		1st Order & 1.061431 & 0.0049954 & 1.060632 & 0.0050525 & 1.060995 & 0.0050262 & 1.061004 & 0.0050255\\
-		\hline
-		4th Order & 1.148524 & 0.0014343 & 1.148405 & 0.0014368 & 1.148513 & 0.0014346 & 1.148399 & 0.0014369\\
-		\hline
-		Linear SIR & 1.087741 & 0.0034308 & 1.087338 & 0.0034506 & 1.087538 & 0.0034408 & 1.087399 & 0.0034476\\
-		\hline
-	\end{tabular}
-\end{table}
-
-\begin{table}
+		1st Order & 1.061431 & 0.0004905 & 1.060632 & 0.0004964 & 1.060995 & 0.0004937 & 1.061004 & 0.0004936\\
+		\hline
+		4th Order & 1.148524 & 0.0001344 & 1.148405 & 0.0001347 & 1.148513 & 0.0001344 & 1.148399 & 0.0001347\\
+		\hline
+		Linear SIR & 1.087741 & 0.0003320 & 1.087338 & 0.0003340 & 1.087538 & 0.0003330 & 1.087399 & 0.0003337\\
+		\hline
+	\end{tabular}
+\end{table}
+
+\begin{table}[H]
 
@@
 		\hline
-		T = 200 & 1.306607 & 0.0047484 & 1.301989 & 0.0049042 & 1.304421 & 0.0048215 & 1.301390 & 0.0049248\\
-		\hline
-		T = 100 & 1.560029 & 0.0278141 & 1.418887 & 0.0419730 & 1.517223 & 0.0315180 & 1.390702 & 0.0455494\\
-		\hline
-		T = 50 & 2.238028 & 0.1132619 & 1.623615 & 0.2357995 & 2.186315 & 0.1209265 & 1.634726 & 0.2329303\\
-		\hline
-		T = 20 & 4.472654 & 0.5150738 & 1.108447 & 1.8062680 & 4.159546 & 0.6317608 & 2.711197 & 1.3929029\\
-		\hline
-	\end{tabular}
-\end{table}
+		T = 200 & 1.306607 & 0.0004142 & 1.301989 & 0.0004286 & 1.304421 & 0.0004209 & 1.301390 & 0.0004305\\
+		\hline
+		T = 100 & 1.560029 & 0.0022192 & 1.418887 & 0.0035072 & 1.517223 & 0.0025481 & 1.390702 & 0.0038451\\
+		\hline
+		T = 50 & 2.238028 & 0.0078639 & 1.623615 & 0.0184718 & 2.186315 & 0.0084553 & 1.634726 & 0.0181912\\
+		\hline
+		T = 20 & 4.472654 & 0.0339963 & 1.108447 & 0.1728373 & 4.159546 & 0.0412621 & 2.711197 & 0.0923793\\
+		\hline
+	\end{tabular}
+\end{table}
+
+\subsection{Wallinga}
+
+\begin{table}[H]
+
+	\caption{\label{tab:}$\rr$ Estimates and Std. Errs, WALLINGA Model,
+		$\beta = 0.06, \gamma = 0.03$, $X_0 = 99950, Y_0 = 50$, $\sigma_X = 100, \sigma_Y = 5$}
+	\centering
+	\begin{tabular}[t]{l|r|r|r|r|r|r|r|r}
+		\hline
+		Data Set & Norm Est & Norm SE & Norm-M Est & Norm-M SE & AR Est & AR SE & AR-M Est & AR-M SE\\
+		\hline
+		$\gamma$ = 0.001 & 1.114829 & 0.0459625 & 0.9591728 & 0.0488597 & 1.140804 & 0.0459639 & 0.8574921 & 0.0500578\\
+		\hline
+		$\gamma$ = 0.04 & 2.008614 & 0.0318940 & 2.0086026 & 0.0318958 & 2.008640 & 0.0318947 & 2.0086160 & 0.0318934\\
+		\hline
+		$\gamma$ = 0.06 & 2.038747 & 0.0313772 & 2.0387516 & 0.0313766 & 2.038764 & 0.0313758 & 2.0387101 & 0.0313764\\
+		\hline
+		$\gamma$ = 0.24 & 2.039700 & 0.0313775 & 2.0397001 & 0.0313775 & 2.039712 & 0.0313791 & 2.0396914 & 0.0313784\\
+		\hline
+	\end{tabular}
+\end{table}
+
+\begin{table}[H]
+
+	\caption{\label{tab:}$\rr$ Estimates and Std. Errs, WALLINGA Model,
+		$\beta = 0.06, \gamma = 0.03$, $X_0 = 99950, Y_0 = 50$, $\sigma_X = 100, \sigma_Y = 5$}
+	\centering
+	\begin{tabular}[t]{l|r|r|r|r|r|r|r|r}
+		\hline
+		Data Set & Norm Est & Norm SE & Norm-M Est & Norm-M SE & AR Est & AR SE & AR-M Est & AR-M SE\\
+		\hline
+		$X_0 = 99000, Y_0 = 1000$ & 1.873533 & 0.0318531 & 1.873446 & 0.0318541 & 1.873475 & 0.0318524 & 1.873496 & 0.0318549\\
+		\hline
+		$X_0 = 99900, Y_0 = 100$ & 1.923607 & 0.0328651 & 1.923653 & 0.0328648 & 1.923560 & 0.0328664 & 1.923576 & 0.0328694\\
+		\hline
+		$X_0 = 99990, Y_0 = 10$ & 1.979338 & 0.0325801 & 1.979345 & 0.0325796 & 1.979353 & 0.0325789 & 1.979404 & 0.0325774\\
+		\hline
+		$X_0 = 99999, Y_0 = 1$& 2.022337 & 0.0317890 & 2.022325 & 0.0317883 & 2.022367 & 0.0317860 & 2.022322 & 0.0317898\\
+		\hline
+	\end{tabular}
+\end{table}
+
+\begin{table}[H]
+
+	\caption{\label{tab:}$\rr$ Estimates and Std. Errs, WALLINGA Model,
+		$\beta = 0.06, \gamma = 0.03$, $X_0 = 99950, Y_0 = 50$}
+	\centering
+	\begin{tabular}[t]{l|r|r|r|r|r|r|r|r}
+		\hline
+		Data Set & Norm Est & Norm SE & Norm-M Est & Norm-M SE & AR Est & AR SE & AR-M Est & AR-M SE\\
+		\hline
+		$\sigma_X = 50, \sigma_Y = 2$ & 1.940295 & 0.0328946 & 1.940280 & 0.0328946 & 1.940313 & 0.0328958 & 1.940303 & 0.0328948\\
+		\hline
+		$\sigma_X = 500, \sigma_Y = 20$ & 1.940066 & 0.0328982 & 1.940463 & 0.0328877 & 1.939755 & 0.0329040 & 1.940506 & 0.0328904\\
+		\hline
+		$\sigma_X = 2500, \sigma_Y = 100$ & 1.940497 & 0.0328856 & 1.940101 & 0.0329012 & 1.937796 & 0.0329248 & 1.938458 & 0.0329298\\
+		\hline
+		$\sigma_X = 10000, \sigma_Y = 500$ & 1.934924 & 0.0329764 & 1.942861 & 0.0328652 & 1.910502 & 0.0346809 & 1.913664 & 0.0339201\\
+		\hline
+	\end{tabular}
+\end{table}
+
+\begin{table}[H]
+
+	\caption{\label{tab:}$\rr$ Estimates and Std. Errs, WALLINGA Model,
+		$X_0 = 99950, Y_0 = 50$, $\sigma_X = 100, \sigma_Y = 5$}
+	\centering
+	\begin{tabular}[t]{l|r|r|r|r|r|r|r|r}
+		\hline
+		Data Set & Norm Est & Norm SE & Norm-M Est & Norm-M SE & AR Est & AR SE & AR-M Est & AR-M SE\\
+		\hline
+		1st Order & 2.034020 & 0.0314227 & 2.034020 & 0.0314227 & 2.034018 & 0.0314231 & 2.034013 & 0.0314227\\
+		\hline
+		4th Order & 1.938669 & 0.0331254 & 1.938646 & 0.0331271 & 1.938678 & 0.0331252 & 1.938644 & 0.0331271\\
+		\hline
+		Linear SIR & 2.033555 & 0.0314612 & 2.033554 & 0.0314611 & 2.033555 & 0.0314609 & 2.033547 & 0.0314615\\
+		\hline
+	\end{tabular}
+\end{table}
+
+\begin{table}[H]
+
+	\caption{\label{tab:}$\rr$ Estimates and Std. Errs, WALLINGA Model,
+		$\beta = 0.06, \gamma = 0.03$, $X_0 = 99950, Y_0 = 50$, $\sigma_X = 100, \sigma_Y = 5$}
+	\centering
+	\begin{tabular}[t]{l|r|r|r|r|r|r|r|r}
+		\hline
+		Data Set & Norm Est & Norm SE & Norm-M Est & Norm-M SE & AR Est & AR SE & AR-M Est & AR-M SE\\
+		\hline
+		T = 200 & 2.872029 & 0.0777506 & 2.872039 & 0.0777500 & 2.872024 & 0.0777475 & 2.871995 & 0.0777522\\
+		\hline
+		T = 100 & 4.777225 & 0.2173201 & 4.777183 & 0.2173166 & 4.777046 & 0.2173274 & 4.777196 & 0.2173002\\
+		\hline
+		T = 50 & 8.523307 & 0.6084195 & 8.523224 & 0.6084023 & 8.523328 & 0.6084437 & 8.523118 & 0.6083556\\
+		\hline
+		T = 20 & 19.536946 & 2.3356802 & 19.536830 & 2.3354926 & 19.537515 & 2.3353411 & 19.534388 & 2.3356299\\
+		\hline
+	\end{tabular}
+\end{table}
+
 \end{document}
@@
 %%%%%%%%%%%%%%%%%%%%%%
-\Textbf{Title:}
+\textbf{Title:}
 
@@
 \begin{center}
-\includegraphics[width=4in]{mvhw7_3c.pdf}
+\includegraphics[width=4in]{images/mvhw7_3c.pdf}
 \end{center}
@@
   \centering
-  \includegraphics[width=1\linewidth]{mt_eda_cont_hists.pdf}
+  \includegraphics[width=1\linewidth]{images/mt_eda_cont_hists.pdf}
   \caption{Histograms of Arrival Delay and continuous covariates.
@@
   \centering
-  \includegraphics[width=1\linewidth]{mt_eda_cont_hists.pdf}
+  \includegraphics[width=1\linewidth]{images/mt_eda_cont_hists.pdf}
   \caption{\textcolor{red}{Placeholder}}
@@
   \centering
-  \includegraphics[width=1\linewidth]{resids_full.pdf}
+  \includegraphics[width=1\linewidth]{images/resids_full.pdf}
   \caption{}
@@
   \centering
-  \includegraphics[width=1\linewidth]{diags_full.pdf}
+  \includegraphics[width=1\linewidth]{images/diags_full.pdf}
   \caption{ }
@@
 \end{figure}
+
+\message{ !name(draft_v13.tex) !offset(-2294) }

### 2018-07-03 paper/Version13/draft_v13.tex
finally on git

--- 2018-07-03 paper/Version13/_region_.tex
@@
-\message{ !name(draft_v13.tex)}\documentclass[12pt]{article}
+\documentclass[12pt]{article}
 \usepackage[margin=1in]{geometry}
@@
 
-\newcommand{\xxsir}{\ensuremath{11} } % number of SIR methods
-\newcommand{\wxxsir}{11 }
-\newcommand{\Wxxsir}{11 }
+\newcommand{\xxsir}{\ensuremath{9} } % number of SIR methods
+\newcommand{\wxxsir}{nine } % lower case number as word
+\newcommand{\Wxxsir}{Nine } % capitalized word
 
@@
 \begin{document}
-
-\message{ !name(draft_v13.tex) !offset(-3) }
 
@@
 \section{Introduction}\label{sec:intro}
-What has been called ``arguably the most important quantity in the study of epidemics'' \citep{Heesterbeek2002},  $\mathcal{R}_0$, the reproduction number (by convention pronounced ``R-naught'') is a persistently elusive quantity for epidemic modelers to estimate.
-As defined by \citet{anderson1992}, $\rr$ is the ``the average number of secondary infections produced when one infected individual is introduced into a host population where everyone is susceptible.''  $\rr$, in some ways, summarizes an entire outbreak of a disease; it  is used to assess whether a disease epidemic will occur.
+What has been called ``arguably the most important quantity in the study of epidemics'' \citep{Heesterbeek2002},  $\mathcal{R}_0$, the reproduction number (by convention pronounced ``R-naught'') is an elusive quantity for epidemic modelers to estimate.
+As defined by \citet{anderson1992}, $\rr$ is the ``the average number of secondary infections produced when one infected individual is introduced into a host population where everyone is susceptible.''  In some ways,$\rr$ summarizes an entire outbreak of a disease; it  is used to assess whether a disease epidemic will occur.
 Additionally, it describes what percentage of the population needs to be vaccinated to avoid such an epidemic, roughly $1-\rr^{-1}$ \citep{anderson1992}).
@@
 The difficulties of estimating $\rr$ are summarized and expanded upon by \cite{li2011}.
+\cite{driessche2017} overviews a number of methods used to estimate $\rr$ but looks at systems beyond the SIR model and so these methods may not be directly compared.
 As a result of $\rr$ being a property of the model, we limit our review to methods for estimating $\rr$ for the SIR model introduced by \cite{Kermack700}.
@@
 We find that these methods also help to highlight the problems discussed above.
-In addition to making a point estimate of $\rr$, we are particularly interested in the hypothesis  $H_0: \rr > 1$, as a value greater than 1 indicates an outbreak of a disease.
-As such, we discuss techniques for estimating CIs for each of the different methods.
-These CI estimates are created using the delta method, the block bootstrap, and the posterior distribution.
-
-In the following later, we present a series of simulations in which we examine each of our $\rr$ estimation methods.
+
+In addition to making a point estimate of $\rr$, we are particularly interested in the hypothesis  $H_0: \rr < 1$, as a value less than 1 indicates that a disease will not outbreak.
+We also discuss techniques for estimating CIs for each of the different methods.
+These CI estimates are created using the delta method, the block bootstrap, and the posterior distribution \citep{cao1999,wasserman2004}.
+
+In the following sections, we present a series of simulations in which we examine each of our $\rr$ estimation methods.
 Within these simulations, we vary the number of observations used to estimate $\rr$, infection and recovery rates, the total population size, initial SI ratio, and several assumptions about the noise.
 
-Finally, we discuss our recommendations for estimating $\rr$ for the SIR model, and we comment on how these recommendations extend to more complex models and how many of the problems with estimating $\rr$  become exacerbated with the increase in complexity.
+Finally, we discuss our recommendations for estimating $\rr$ in the SIR model, and we comment on how these recommendations extend to more complex epidemiological models beyond the SIR model and on how many of the problems with estimating $\rr$  become exacerbated in such models.
 
@@
 The origins of $\rr$ are tied to the survival function, which originates from the field of demography in the late 1800s.
-The survival function describes how many female offspring a woman is expected to produce in her lifetime \citep{dietz1993estimation}.
+The survival function describes how many female offspring a woman is expected to produce in her lifetime, literally the reproduction number \citep{dietz1993estimation}.
 In demography, we have
@@
 where $p(a)$ denotes the probability of a woman surviving to age $a$ and $\beta(a)$ the rate of an individual of age $a$ giving birth to a  girl.
-This interpretation also explains the origins of the name of $\rr$.
 The concept of the reproduction number was later imported to the field of epidemiology by MacDonald and Smith \citep{dietz1993estimation}.
-Analogously in epidemiology, $p(a)$ is the age of a disease, and $\beta(a)$ is the infection rate.
-
-This is the chronologically first, and in many ways, the most direct method to estimate $\rr$.
+Analogously in epidemiology, $p(a)$ is the age of a disease, and $\beta(a)$ is the infection rate at time $a$.
+
+Equation \ref{eq:surv} is the chronologically first, and in many ways, the most direct method to estimate $\rr$.
 The premise of this whole paper is based upon the fact that estimating $\rr$ from the survival function in Equation \eqref{eq:surv} is a difficult task.
@@
 
-Other difficulties in estimating $\rr$ arise from model assumptions on how randomness enters the model and sensitivity of the estimate.
-For example, the SIR model is unique in that two of its compartments, S and R are monotonic.
+Other difficulties in estimating $\rr$ arise from model assumptions on how randomness enters the model and sensitivity of the estimate to changes in the model parameters.
+For example, the SIR model is special in that two of its compartments, S and R are monotonic.
 The number of susceptibles is non-increasing and the number of susceptibles is non-decreasing.
-Models should enforce this monotonicity of the two compartments.
+Logically, models should enforce this monotonicity of the two compartments.
 As a consequence, this restricts assumptions on how randomness enters the model.
@@
 
-These difficulties in estimating $\rr$ have led a number of researchers to analyze the sensitivity of $\rr$ \citep{lash2003,epstein2007agent,capaldi2012}.
-Even this is a difficult task because we are looking at the rate of change of the compartments with respect to $\beta$ and $\gamma$.
-Since there is no closed-form solution of the SIR model, estimating the set of derivatives with respect to $\beta$ and $\gamma$ must be done numerically, which can substantially increase computation time.
+These difficulties in estimating $\rr$ have led a number of researchers to analyze the sensitivity of $\rr$ with respect to small perturbations in parameters like $\beta$ and $\gamma$ \citep{lash2003,epstein2007agent,capaldi2012}.
+Even this is a difficult task.
+Since there is no analytic solution to the SIR model, estimating the set of derivatives with respect to $\beta$ and $\gamma$ must be done numerically, which can substantially increase computation time.
 We should note that this sensitivity analysis is done before even adding any distributional assumptions on how randomness enters the model.
 
-Below in Table \ref{tab:r0-real-ex}, we provide examples of estimates $\rr$ from publications over the past three decades, spanning various locations over the world.
-We display estimates for HIV, Zika (ZIK-V), Ebola (EVD), seasonal influenza, and H1N1 influenza.
+Below in Table \ref{tab:r0-real-ex}, we provide examples of estimates of $\rr$ for specific diseases from publications over the past three decades, spanning various locations over the world.
+We display estimates for HIV, Zika (ZIK-V), Ebola (EVD), seasonal influenza, H1N1 influenza, and measles.
 These estimates of $\rr$ are not limited to those of SIR model, although most results come from this framework.
 Of these, Zika seems to have the largest estimated values of $\rr$, followed by Ebola, H1N1 influenza, seasonal influenza, and HIV.
-We see that African locations seem to have larger values of $\rr$ compared to the rest of the world, regardless of the disease.
-Some of the accompanying lower and upper bound intervals (not necessarily CIs) are extremely large such as the case of HIV in Uganda (Interval: [.12, 14.17]) where as some are extremely small such as for Ebola in Guinea [1.5, 1.52].
-We see that estimates for H1N1 influenza are generally higher than that of seasonal influenza, which is evidence that H1N1 was a bigger concern than the typical seasonal influenza cycle.
-The lowest possible value for $\rr$ reported is .12 and the largest is 14.17, both for HIV in Uganda.
-Within diseases, we see that estimates for $\rr$ seem to be similar, although typically do not lie in another's lower and upper bound, suggesting that there is great disparity in estimates of $\rr$.
+We see that African locations seem to have larger estimates of $\rr$ compared to the rest of the world, regardless of the disease.
+Some of the accompanying intervals (not necessarily CIs) are extremely large such as the case of HIV in Uganda (Interval: [0.12, 14.17]) whereas some are extremely small such as for Ebola in Guinea (Interval: [1.50, 1.52]).
+We see that estimates for H1N1 influenza are generally higher than that of seasonal influenza, which is evidence that H1N1 should have been treated as a more critical concern than the typical seasonal influenza cycle.
+The lowest possible estimate for $\rr$ reported is 0.12 for HIV in Uganda, and the largest is 17.00 for measles in England and Wales.
+Within diseases, we see that estimates for $\rr$ seem to be similar, although typically do not lie in one another's intervals, suggesting that there is disparity in estimates of $\rr$.
 
@@
   H1N1 Influenza     & Peru                             & 2009                          & 1.37        & 1.20            & 1.70            & \cite{desilva2009}        \\ 
-H1N1 Influenza      & La Gloria, Mexico                  & 2009                          & 1.58        & 1.34           & 1.60            & \cite{fraser2009}     \\ \bottomrule
-\end{tabular}
-\caption{Estimates of $\rr$ from various researchers for the infectious diseases of HIV, ZIK-V, Ebola (EVD), seasonal influenza, and H1N1 pandemic influenza ranging from the years 1972-2016 in various locations.
+  H1N1 Influenza & USA & 2009 & 1.50 & 1.30 & 1.70 &\cite{towers2009} \\ \hline
+Measles      & Cirencester, England                 & 1947-1950                          & 13.50   & 13.00           & 14.00            & \cite{fraser2009}     \\ 
+  Measles     & England and Wales                  & 1950-1968                          & 17.00      & 16.00          & 18.00         & \cite{anderson1992}     \\
+  Measles     & Kansas, USA                  & 1918-1921                          & 5.50    & 5.00          & 6.00          & \cite{anderson1992}     \\
+  Measles     & Ontario, Canada                  & 1912-1913                         & 11.50        & 11.00           & 12.00            & \cite{anderson1992}     \\
+  Measles     & Willesden, England                  & 1912-1913                          & 11.00        & 12.00       & 1.60            & \cite{anderson1992}     \\
+  Measles     & Ghana                 & 1960-1968                          & 14.50       & 14.00          & 15.00            & \cite{anderson1992}     \\
+  Measles     & Eastern Nigeria                  & 1960-1968                          & 16.50        & 16.00          & 17.00           & \cite{anderson1992}     \\
+\bottomrule
+\end{tabular}
+\caption{Estimates of $\rr$ from various researchers for the infectious diseases of HIV, (Zika) ZIK-V, Ebola (EVD), seasonal influenza, and H1N1 pandemic influenza ranging from the years 1972-2016 in various locations.
 We also report the lower and upper bounds reported by the researchers.}
@@
 In epidemiology, this means that the proportion of new infections per unit time is proportional to the current  number of susceptible \citep{anderson1992}.
-Therefore, the flow of objects from one compartment to another may be dependent on the percentage of objects within one or many of the compartments.
-
-In this particular  SIR model, $N$, the number of individuals is constant, as we see no arrows pointing away from the three nodes in Figure \ref{fig::sir}.
-Simple adaptations of the SIR model can include birth and death rates, which may correspondingly change the derivation of $\rr$ (for further discussion, see Section \ref{sec:discussion}).
+
+In this particular  SIR model (displayed in Figure \ref{fig::sir}), the number of individuals ($N$) is constant.
+Adaptations of the SIR model can include birth and death rates, which may correspondingly change the derivation of $\rr$ (for further discussion, see Section \ref{sec:discussion}).
 The parameters  $\beta$ and $\gamma$ have epidemiological meaning.
@@
   \frac{dY}{dt} &> 0 \\
+  \implies -\frac{dX}{dt} - \frac{dZ}{dt} &> 0 \\
   \implies \frac{\beta X Y}{N}  - \gamma Y &> 0 ,\\
@@
 That is,  the rate of new infections is greater than the rate of recovery.
-So as long as the number of susceptibles is large compared to the total population, $\frac{X}{N} \approx 1$, then an outbreak will occur if $\rr >1$,
-\begin{align*}
-  \rr \overset{def}{=} \frac{\beta}{\gamma} > 1.
-  \end{align*}
-  In order to incorporate randomness into the model, we add noise, namely,
+So as long as the number of initially susceptible individuals is large compared to the total population, $\frac{X}{N} \approx 1$, then an outbreak will occur if $\rr >1$,
+\begin{align*}
+  \rr \overset{def}{=} \frac{\beta}{\gamma}.
+  \end{align*}
+  To incorporate randomness into the model, we add noise, namely,
   \begin{align}\label{eq:sir-noise}
@@
   \end{align}
-We are assuming the observations are generated based on the deterministic ODEs presented in Equation \ref{eq:sir} with the addition of time ($t=0, 1, \dots, T$) and compartment dependent noise $\epsilon_{X,t}$ and $\epsilon_{Z,t}$.
+We are assuming the observations are generated based on the discretizied time, deterministic ODEs presented in Equation \ref{eq:sir} with the addition of time ($t=0, 1, \dots, T$) and compartment dependent noise $\epsilon_{X,t}$ and $\epsilon_{Z,t}$.
 Since, $N$, the total population, is constant, then $Y_{obs}$ is adjusted accordingly.
@@
 \begin{align*}
-  \hat{\rr} \overset{P}{\to} \rr \textnormal{ as } T\to \infty
-\end{align*}
-or (2) converging to a known probability distribution,
+  \hat{\rr} \overset{P}{\to} \rr \textnormal{ as } T\to \infty,
+\end{align*}
+where `$\overset{P}{\to}$' stands for convergence in probability \citep{wasserman2004} or (2) converging to a known probability distribution (convergence in distribution `$\overset{d}{\to}$' \citep{wasserman2004}),
 \begin{align*}
@@
 \begin{align*}
-  H_0:\;& \rr > 1 \\
-  H_A:\;& \rr \le 1
-\end{align*}
-Then for some $\alpha \in [0,1]$, we would like to know for some $a, b >0$
-\begin{align*}
-P(a \le \hat{\rr} \le b) = 1 - \alpha,  
-\end{align*}
-the second property is particularly desirable, as we could find $a$ and $b$ from the distribution $F$.
-At any rate, we would like to be able to estimate $a$ and $b$ for a given $\alpha$-level.
+  H_0:\;& \rr < 1 \\
+  H_A:\;& \rr \ge 1
+\end{align*}
+Then for some $\alpha \in [0,1]$, we would like to know for some $\hat{a}, \hat{b} >0$
+\begin{align*}
+P(\hat{a} \le \rr \le \hat{b}) \le 1 - \alpha,  
+\end{align*}
+the second property is particularly desirable, as we can estimate $\hat{a}$ and $\hat{b}$ from the distribution $F$.
+At any rate, we would like to be able to estimate $\hat{a}$ and $\hat{b}$ for a given $\alpha$-level.
 We could also make similar hypothesis tests such as that the $\rr$ for H1N1 influenza is greater than the reported $\rr$ for seasonal influenza.
@@
 
-This method is sometimes used to estimate $\hat{\beta}$ and $\hat{\gamma}$ via a grid search of $\beta$ and $\gamma$ but can also be done with more sophisticated optimiization algorithms.
-Since we cannot explicitly write down the partial derivatives of $S$, $I$, and $R$ with respect to $\beta$ and $\gamma$, we can only use minimization algorithms which do not rely on an explicit gradient such as Nelder-Mead simplex minimization \citep{nelder-mead1965}.
+The estimates from this method are often found via a grid search of $\beta$ and $\gamma$ values to minimize the squared error.
+The estimates can also be found using more sophisticated optimization algorithms.
+Since we typically cannot explicitly write down the partial derivatives of $S$, $I$, and $R$ with respect to $\beta$ and $\gamma$, we can only use minimization algorithms which do not rely on an explicit gradient such as Nelder-Mead simplex minimization \citep{nelder-mead1965}.
 
@@
 Another answer is that $L_2$ is a continuously differentiable function and hence is easier to compute, which allows for sensitivity analysis to  be conducted more easily.
-Finally, LS can be done without writing down explicit assumptions on the noise within the model.
+Finally, LS can be done without writing down explicit assumptions about the noise within the model.
 However, we cannot guarantee properties of consistency or convergence to a known distribution without more explicit assumptions.
 
-On the other hand, least squares is easy to implement in software such as \texttt{R}'s \texttt{optim()} function, and as we see in Section \ref{sec:results} produces comparable methods to other methods.
+On the other hand, least squares is easy to implement using software such as \texttt{R}'s \texttt{optim()} function, and as we see in Section \ref{sec:results} produces results comparable methods to other methods.
 Least squares and similar variants have been used to estimate $\rr$ in \cite{majumder2016}.
@@
 In the previous method (Section \ref{least-squares-beta-gamma}), we estimated $\beta$ and $\gamma$ and then estimated $\rr$.
-However, it is possible to directly estimate $\rr$ if we reparametrize the ODEs in Equation \eqref{eq:sir} directly with \(\rr\) and \(\gamma\), using the relation $\rr = \frac{\beta}{\gamma}$,
+However, it is possible to directly estimate $\rr$ if we reparametrize the ODEs in Equation \eqref{eq:sir} directly with \(\rr\) and \(\gamma\), using the relation $\rr = \frac{\beta}{\gamma}$; this yields
 
@@
 \end{align*}
-We use the $\hat{\rr}$ directly from the above estimation problem, which again can be solved with a grid search or another optimization process.
+We use the $\hat{\rr}$ directly from the above estimation problem, which again can be found with a grid search or another optimization process.
 It may be surprising that this method leads to different results than simply using least squares.
 One reason we obtain different results may be attributed to better numerical precision as there is one fewer step to obtain an estimate for $\rr$.
-However, we also see that this method further results in smaller standard errors than in LS, which we discuss more in Section \ref{sec:results}.
+However, we also see that this method further results in different standard errors than in LS, which we discuss further in Section \ref{sec:results}.
 
@@
 In addition to this, data collected from real diseases are typically very noisy to begin with.
-\cite{chang2017} showed that an SIR model may be well approximated by a linear model.
-We use this approach here to estimate $\rr$.
-
-Specifically, we estimate two linear polynomials in \(t\) with degree $K$  to \(X_{obs}\)
+\cite{chang2017} showed that the SIR model may be well approximated by a linear model.
+We use this method to estimate $\rr$.
+
+Specifically, we estimate two polynomials in \(t\) with degree $K$  to \(X_{obs}\)
 and \(Z_{obs}\) using least squares to find the coefficients $\{(\hat{x}_k,
@@
   An advantage of using this estimation method is that it is simple to implement using any linear modelling software.
-On the other hand, we are assuming $X(t) = \hat{X}(t)$ and so if the data truly follows a SIR model, then we will likely have a biased result for $\rr$.
+On the other hand, we assume $X(t) = \hat{X}(t)$ and so if the data truly follows the SIR model, then we will likely have a biased result for $\rr$.
 Using this method, we are able to check how sensitive $\rr$ is to the degree of the fitted polynomial.
@@
 \subsection{Incidence to Prevalence Ratio (IPR)}\label{incidence-to-prevalence-ratio}
-The incidence to prevalence ratio (IPR), described by \cite{Nishiura2009}, is another intuitive method to calculate $\rr$ as it incorporates some of the most basic epidemiological quantities, incidence and prevalence.
-
-In terms of data from the SIR model, incidence $J(t) \approx -(X(t+1) - X(t))$, and the IPR$(t) = \frac{J(t)}{Y(t)}$.
+The incidence to prevalence ratio (IPR), described by \cite{Nishiura2009}, is another intuitive method to estimate $\rr$.
+It incorporates some of the most basic epidemiological quantities: incidence and prevalence.
+
+In terms of data from the SIR model, the incidence $J(t) \approx -(X(t+1) - X(t))$, and the IPR$(t) = \frac{J(t)}{Y(t)}$.
 This method assumes that we have some prior knowledge about $\gamma$, the recovery rate.
@@
 The advantage of this method is that incidence data is generally readily available as is prevalence data for certain diseases such as HIV.
-However, as one is required to have prior knowledge about $\gamma$, it may be easier to directly estimate $\rr$ with one of the many other methods described that does not require a prior knowledge about $\gamma$.
+However, as one is required to have prior knowledge about $\gamma$, it may be easier to directly estimate $\rr$ with one of the many other methods described that does not require  prior knowledge.
 Again, we are using only one time point to estimate $\rr$.
@@
 \end{align*}
-The advantage of this method is that it creates a less variable estimate than estimating IPR using only one point.
-It has the same disadvantages as the regular IPR ratio in that it requires knowledge about $\gamma$.
+Splines are used because they typically can be used to make good approximations of epidemic curves (CITE).
+However, the researcher could use their method of choice to estimate $\hat{X}(t)$ and $\hat{Y}(t)$.
+An advantage of this method is that it creates a less variable estimate than estimating IPR using only one point.
+It has the same disadvantages as the usual IPR ratio in that it requires knowledge about $\gamma$.
 
@@
 
-Thus, we can regress the left hand side in Eq.
-\ref{eq:harko_lin} on $Z(t)$, the number of recovered individuals, with $\rr$ as the coefficient of $Z(t)/N$ and with an intercept term of $\beta_0=0$ to obtain an estimate of $\rr$,
+We can regress the left hand side in Eq.
+\ref{eq:harko_lin} on $Z(t)$ (using least squares), the number of recovered individuals, with $\rr$ as the coefficient of $Z(t)/N$ and with an intercept term of $\beta_0=0$ to obtain an estimate of $\rr$,
 \begin{align*}
@@
 \end{align*}
-This method says that for a one percent increase in number of recovered individuals, we expect the ratio of number of previous susceptibles to number of new susceptibles to increase by $e^{\rr}$,
+This method says that for a one percent increase in number of recovered individuals, we expect the ratio of the number of previous susceptibles to number of new susceptibles to increase by $e^{\rr}$,
 \begin{align*}
@@
 One, the estimate of $\rr$ is highly interpretable.
-Two, an assumption of Gaussian noise in Equation \ref{eq:harko_lin} is not as egregious as in the other methods to estimate $\rr$ because we are looking at a log transformation of the the percent of suceptibles instead of raw variables.
-As a consequence, we may be able to make a case where we the resulting estimate of $\rr$ is unbiased and a consistent estimator.
-Finally,  this method is easily implemented using any linear modelling software.
+Two, an assumption of Gaussian noise in Equation \ref{eq:harko_lin} is not as strong as an assumption in the other methods to estimate $\rr$ because we are looking at a log transformation of the the percent of suceptibles instead of raw variables.
+If the assumption holds, we can show that the resulting estimate of $\rr$ is an unbiased and a consistent estimator.
+Finally,  this method is easily implemented using any linear modeling software.
 
@@
 
-In this method, the number of susceptibles at the next step, $X(t+1)$, has a Binomial distribution based on the contacts with the current number of infectious, $Y(t)$ and the current number of susceptibles.
-That is $X(t+1) \sim \text{Binomial}\left(X(t), \alpha^{Y(t)}\right)$, where $\alpha$ is the probability of avoiding infection from an infective.
+In this method, the number of susceptibles at the next step, $X_{obs}(t+1)$, has a Binomial distribution based on the contacts with the current number of infectious, $Y(t)$ and the current number of susceptibles.
+That is $X_{obs}(t+1) \sim \text{Binomial}\left(X_{obs}(t), \alpha^{Y_{obs}(t)}\right)$, where $\alpha$ is the probability of avoiding infection from an infective.
 One can write out the likelihood for $\alpha$ in this model, namely,
 \begin{align*}
-\mathcal{L}\left ( \alpha ; \textnormal{data}\right ) \propto \prod_{t=0}^{T-1}\left ( \alpha^{Y(t)} \right )^{X(t+1)} \left (1- \alpha^{Y(t)} \right )^{X(t) - X(t+1)}.
+\mathcal{L}\left ( \alpha ; \textnormal{data}\right ) \propto \prod_{t=0}^{T-1}\left ( \alpha^{Y_{obs}(t)} \right )^{X_{obs}(t+1)} \left (1- \alpha^{Y_{obs}(t)} \right )^{X_{obs}(t) - X_{obs}(t+1)}.
 \end{align*}
@@
 This method typically allows for more than just the reproduction number to be estimated.
-Through recursion, one can calculate the probability of having a given number of susceptibles and infected at each time step, and hence the entire probability distribution may be known.
-
-The advantages of this model are its simplicity of interpretation and ability to generate a whole probability distrubtion for $(X(t), Y(t))$.
+Through recursion, one can estimate the probability of having a given number of susceptibles and infected at each time step, and hence the entire probability distribution may be known.
+
+The advantages of this model are its simplicity of interpretation and ability to generate an estimate for the probability distrubtion for $(X(t), Y(t))$.
 We have explicitly stated how randomness enters the model so it is up to the researcher to check whether this assumption is a good fit.
-Also, it is possible to simulate how sensitive $\hat{\rr}$ is to $\hat{\alpha}$.
+Using this method, it is possible to simulate how sensitive $\hat{\rr}$ is to changes in $\hat{\alpha}$.
 
@@
 
-\subsection{Exponential Growth (EG)}\label{sec:expgrowth}
-%% nishiura, majumder, IDEA model -- sum of squares (santillana2016)
-%% An IDEA for short term outbreak projection: nearcasting using the basic reproduction number.  (fisman2013, fisman2014)
-%% chowell 2008 - parametric bootstrap for a model they made up, but sort of empirical.
-Many researchers employ an assumption of exponential growth of the number infectious through times $t=0, 1, \dots, T^* <T$ \citep{wallinga2007generation,fisman2014,nishiura2016,majumder2016,towers2016}.
-\cite{wallinga2007generation} report that the effective reproduction number $\mathcal{R}_t$ and hence the initial reproduction number $\rr$ may derived using the fact that infection ``counts increase exponentially in the initial phase of an epidemic.''  We then have to estimate $r$, the \textit{per capita} change in the number of new cases per unit of time and $\omega$ the serial interval, the distribution of time between a primary and secondary infection.
-Then, we have
-\begin{align}\label{eq:lotka}
-\rr = \exp{(r \omega)}.
-\end{align}
-Equation \eqref{eq:lotka} is derived from a demographic view using the Lotka-Euler survival equations which come from the fields of demography, ecology, and evolutionary biology.
-
-\cite{wallinga2007generation}, instead, use a moment generating function expression for $\rr$.  With $\omega(t)$ as the serial interval, then
-\begin{align*}
-\rr^{-1} &= \frac{1}{M(-r)} = \int_{a=0}^\infty e^{-rt}\omega(t)dt.
-\end{align*}
-\cite{wallinga2007generation} specifically derive an estimate for $\rr$ for the SIR model.
-They assume that the rate of leaving the infectious stsate, $\gamma$ is constant as is the rate of making contacts during the infectious stage.
-Then, the duration of a serial interval is an exponential distribution with mean $\gamma^{-1}$, the moment generating function is then,
-\begin{align*}
-  M(-r) = \frac{\gamma}{\gamma + r},
-\end{align*}
-and the resulting estimate of $\rr$ is
-\begin{align}\label{eq:r0-eg}
-  \hat{\rr} = 1 + \frac{\hat{r}}{\hat{\gamma}},
-\end{align}
-where $\hat{r}$ and $\hat{\gamma}$ are estimates of the $r$ and $\gamma$, respectively.
-The parameter $\hat{\gamma}$ is often assumed to be fixed or tried with a sequence of serial intervals \citep{majumder2016}.
-We may estimate $\hat{r}$ using ordinary linear regression, namely,
-\begin{align*}
-  \log \left  (\frac{Y(t)}{Y(0)} \right )&= rt + \epsilon_t\\
-  \epsilon_t &\sim N(0, \sigma^2),
-\end{align*}
-which results in
-\begin{align}\label{eq:exp-growth}
-  \hat{r} &= \frac{\sum_{t=0}^{T^*}\left  (\frac{Y(t)}{Y(0)} \right )}{\sum_{t=0}^{T^*}t}.
-\end{align}
-
-For the above estimate of $\rr$, we observe the duration of serial intervals in a period of exponential growth of the number of infectious individuals.
-Deciding when exponential growth occurs and which data points to use to estimate $\rr$ may be difficult.
-We can use sensitivity analysis to see how the estimate $\rr$ changes as we increase or decrease $T^{*}$, the number of time points used.
-The reader may notice some similarities to the estimate for $\rr$ in Equation \eqref{eq:r0-eg} and from the log-linear model in Equation \eqref{eq:harko_lin}.
-This method is a simplification of the SIR model as the $R$ compartment is not used to estimate $r$, only the recovery rate $\gamma$.
-
-An advantage to this method are its (and similar adaptations) prevalence in the literature, which allows one to to compare estimates $\rr$ directly to one another.
-Another advantage is its flexibility.
-Here, we assumed the serial interval has an exponential distribution, but this is not a requirement and other distributions may be used (see \cite{wallinga2007generation}).
-However, a disadvantage to this method is that is a simplification of the SIR model and we have to decide how many time points to use to estimate the exponential growth rate of infectious.
-Another disadvantage to this method is that one also has to decide how to treat the additional paramter $\gamma$, whether to treat it as fixed, a range of fixed values, or even having a probability distribution.
-
-\subsection{Incidence Decay and Exponential Adjustment (IDEA)}\label{sec:idea}
-The Incidence Decay and Exponential Adjustment (IDEA) model is a way to account for a declining effective reproduction number of time due to both depletion of susceptible individuals and spontaneous and planned control activities and behaviors'' \citep{fisman2013}.
-They model the incidence counts during the $t$th interval, $J(t),$ as
-\begin{align*}
-  J(t) &= \left [ \frac{\rr}{(1 + d)^t}\right ]^t,
-\end{align*}
-where $d$ is a dampening or discount factor.  As $d$ increases, the incidence at time $t$  decreases. The estimate for $\rr$ is obtained by minimizing the sum of squares over $\rr$ and $d$,
-\begin{align*}
-  \hat{\rr} &=  \arg \min_{\rr,d} \sum_{t=0}^{T^*}\left (J_{obs}(t) - J(t) \right)^2.
-\end{align*}
-The method presented in \cite{fisman2013} is a mathematical, rather than statistical model.
-Sensitivity analysis is conducted with respect to $T^*$, $d$, and $\rr$, and lower and upper bounds are reported as a result of this analysis.
-
-Advantages of this method include its usefulness in estimating $\rr$ \citep{fisman2014,majumder2016}.
-Additionally, this model may be employed to address the effect of control measures and better estimate the final size of an epidemic.
-A disadvantage to this model is that its upper and lower bounds do not correspond to CIs.
-Also, in this method we must select the interval size of the incidence counts.
-
 \section{Confidence and Credible Intervals}
@@
 
-When the method estimates \(\beta\) and \(\gamma\) instead of \(\rr\) directly, we use the delta method approximation to calculate the
+When the method estimates \(\beta\) and \(\gamma\) instead of \(\rr\) directly, we use the delta method approximation to estimate the
 variance of \(\rr\).
@@
 This method may be extended to other frameworks such as the SEIR model.
-The advantages of the delta method are that since the Next Generation Matrix \citep{diekmann2009} posits a recipe to derive $\rr$, then it is theoretically possible to use the delta method to derive an expression for the variance of $\rr$ for any compartment model.
+The advantages of the delta method are that since the Next Generation Matrix \citep{diekmann2009} gives recipe for deriving $\rr$ from the deterministic ODEs, then it is theoretically possible to use the delta method to derive an expression for the variance of $\rr$ for any compartment model.
 
@@
 
-The asympotic properties hold when the stochastic process is $m$-dependent, that is when observations $t, t+1, \dots$ and $s, s+1, \dots$ are independent whenever $s+m < t$ and under some smoothness conditions of the statistic used to estimate $\rr$.
+The asympotic properties hold when the stochastic process is $m$-dependent, that is when observations $t, t+1, \dots$ and $s, s+1, \dots$ are independent whenever $s+m < t$ and under some smoothness conditions on the statistic used to estimate $\rr$.
 
@@
 \subsection{Posterior Distribution}
-Here, we assume $\rr$ is a random variable and the data is fixed.
-Following the notation of \cite{wasserman2004}, then the posterior distribution $f(\rr | \textnormal{data})$ is is proportional to the likelihood of observing the data given $\rr$, $\mathcal{L}(\textnormal{data}|\rr)$, multiplied by our prior on $\rr$, $f(\rr)$.
+In this setting, we assume $\rr$ is a random variable and the data is fixed.
+Following the notation of \cite{wasserman2004}, then the posterior distribution $f(\rr | \textnormal{data})$ is proportional to the likelihood of observing the data given $\rr$, $\mathcal{L}(\textnormal{data}|\rr)$, multiplied by our prior on $\rr$, $f(\rr)$.
 \begin{align*}
@@
 Correspondingly, the 95\% credible interval is $\left[\hat{F}_{.05}, \hat{F}_{.95} \right ]$.
-Advantages of this method include having an entire distribution compared to a point estimate.
+Advantages of this method include having an entire distribution instead of a point estimate.
 A disadvantage is that using conjugate priors may not be justified for the given data and MCMC may be computationally intractable.
-One should note that these 95\% credible intervalss are not the same as confidence intervals as the ``true'' value of $\rr$ may not be contained in the credible intervals 95\% of the time.
+One should note that these 95\% credible intervals are not the same as confidence intervals since the ``true'' value of $\rr$ may not be contained in the credible intervals 95\% of the time.
 
@@
 We compare the performance of the SIR models on simulated data from both the SIR as well as other models.
-In all of these simulations, we first generate data from the models under known conditions, then we add on errors drawn from known distributions variances to each time point,
+In all of these simulations, we first generate data from the models under known conditions, then we add noise drawn from known distributions to each time point,
 \begin{align}\label{eq:sim-models}
@@
   Y_{model}(t) &= g(t) + \epsilon_{Y,t} \nonumber\\
-  Z_{model}(t) &= N - X(t) - Y(t)\nonumber 
+  Z_{model}(t) &= N - X_{model}(t) - Y_{model}(t)\nonumber 
 \end{align}
 We adjust the recovered compartment $Z$ so the total population is constant. 
-We simulate with two different types of error distributions, Gaussian and autoregressive,
+We simulate with two different types of error distributions, Gaussian (Norm) and autoregressive (AR),
 
@@
 \end{align*}
-For both the Gaussian and AR dataset, we form an additional dataset which enforces the monotonicity of X and Z via the Pool Adjacent Violators Algorithm (PAVA), described by \cite{friedman1984}).
+For each of the Gaussian and AR datasets, we form an additional dataset which enforces the monotonicity of X and Z via the Pool Adjacent Violators Algorithm (PAVA), described by \cite{friedman1984}).
 Here, we adjust $Y$ so the constant population is maintained.
@@
 \end{align*}
-This gives us a total of four error structures for each set of conditions: Gaussian (Norm), Gaussian monotone (Norm-M), autoregressive (AR), and autoregressive monotone (AR-M) for a given set of parameters $(\beta, \gamma, O, X(0), Y(0), Z(0), \sigma_X, \sigma_Y, N)$, the infection rate; recovery rate; total number of time points; initial conditions for number of susceptibles, infectious, and recovered individuals, the standard deviation for both the susceptible and infectious compartments, and the total population, respectively.
+This gives us a total of four error structures for each set of conditions: (1) Gaussian (Norm), (2) Gaussian monotone (Norm-M),  (3) autoregressive (AR), and (4) autoregressive monotone (AR-M) for a given set of parameters $\left (\beta, \gamma, O, X(0), Y(0), Z(0), \sigma_X, \sigma_Y, N \right )$, the infection rate; recovery rate; total number of time points; initial conditions for number of susceptibles, infectious, and recovered individuals, the standard deviation for both the susceptible and infectious compartments, and the total population, respectively.
 
@@
 
-\textbf{Baseline Dataset}
+\textbf{Baseline Simulated Data}
 \begin{center}
@@
 
-	Total population $N$: 100000
+	Total population $N$: 100,000
 
@@
 
-This baseline data set
-
-% We then generate different SIR data sets, changing the one of the conditions for simulation each time and using the baseline SIR data as our basis for comparison.
-The different data sets generated are shown in the following table:
+      This baseline data set is chosen due to its moderate $\rr$ size of 2, which indicates an outbreak of an epidemic similar to HIV, ZIK-V, EVD, and H1N1 pandemic influenza (see Table \ref{tab:r0-real-ex}).
+The infection rate and recovery rate ($\beta, \gamma$) are relatively small.
+The duration of the disease, under the true model, spans nearly a whole year (see Figure \ref{fig:baseline}) and so we use a year's worth of daily observations of (X,Y, Z) values.
+Initially, 50 (0.05\%) diseased individuals are introduced to the population of 100,000 individuals (approximately the average population size of a US county in 2015).
+The standard deviation of the X values is 0.1\% of the magnitude of the initial number of susceptible individuals, and the standard deviation of the Y values is 10\% of the magnitude of the initial number of infectious individuals, although in this setting the number of infectious individuals increases rapidly.
+We use Gaussian error as our baseline error because of its known statistical properties so we can more easily compare methods to one another.
+Later on, we explore the autoregressive and PAVA simulated data, which may better simulate data from  a real epidemic.
+
+      \subsection{Other Simulated Data}
+
+We then generate different SIR data sets, changing one of the conditions for simulation each time and using the baseline SIR data as our basis for comparison.
+The different data sets generated are shown in Table \ref{tab:simulated-data}, for a total of 27 simulated data sets.
 
@@
 
-\subsection{Other Models}
-
-To test how well the SIR models perform under misspecification, we simulate data from other models.
-While we will not have a ``known'' $\rr$ value to compare our results to, this will allow us to see if the models give us reasonable results when the data do not come from the SIR model.
+% % Please add the following required packages to your document preamble:
+% % \usepackage{booktabs}
+% \begin{table}[]
+% \centering
+
+% Please add the following required packages to your document preamble:
+% \usepackage{booktabs}
+\begin{table}[]
+\centering
+\begin{tabular}{@{}llllllll@{}}
+\toprule
+\textbf{Name} & \textbf{Model} & \textbf{Error} & \textbf{($\beta, \gamma$)} & \textbf{O} & \textbf{(X(0),Y(0))} & \textbf{N} & \textbf{($\sigma_X, \sigma_Y$)} \\ \midrule
+  Baseline        & SIR      &   Norm             & (.06, .03)                          &     365       &  (99950, 50)                    &  $10^5$          &(100, 5)                                \\ \midrule
+  AR       & SIR      &   AR ($\rho=.7)$             & (.06, .03)                          &     365       &  (99950, 50)                    &  $10^5$          &(100, 5)                            \\ \midrule
+  \textcolor{red}{Norm-M}      & SIR      &   Norm-M ($\rho=.7)$             & (.06, .03)                          &     365       &  (99950, 50)                    &  $10^5$          &(100, 5)                            \\ \midrule
+   \textcolor{red}{AR-M}       & SIR      &   AR-M ($\rho=.7)$             & (.06, .03)                          &     365       &  (99950, 50)                    &  $10^5$          &(100, 5)                            \\ \midrule
+  InfRec1       & SIR      &   Norm             & (60000, .001)                          &     365       &  (99950, 50)                    &  $10^5$          &(100, 5)\\
+  InfRec2      & SIR      &   Norm             & (1500, .001)                          &     365       &  (99950, 50)                    &  $10^5$          &(100, 5)                                \\
+  InfRec3      & SIR      &   Norm             & (1000, .001)                          &     365       &  (99950, 50)                    &  $10^5$          &(100, 5)                                \\
+InfRec4      & SIR      &   Norm             & (06, .24)                          &     365       &  (99950, 50)                    &  $10^5$          &(100, 5)                                \\ \midrule
+  Time1 & SIR &  Norm   & (.06, .03)                          &     200       &  (99950, 50)                    &  $10^5$          &(100, 5)                                \\
+  Time2 & SIR &  Norm   & (.06, .03)                          &     100       &  (99950, 50)                    &  $10^5$          &(100, 5)                                \\
+  Time3 & SIR &  Norm   & (.06, .03)                          &     50       &  (99950, 50)                    &  $10^5$          &(100, 5)                                \\
+  Time4 & SIR &  Norm   & (.06, .03)                          &     20       &  (99950, 50)                    &  $10^5$          &(100, 5)                                \\ \midrule
+  Initial1    & SIR      &   Norm             & (.06, .03)                          &     365       &  (99999, 1)                    &  $10^5$          &(100, 5)                                \\
+  Initial2    & SIR      &   Norm             & (.06, .03)                          &     365       &  (99990, 10)                    &  $10^5$          &(100, 5)                                \\
+  Initial3    & SIR      &   Norm             & (.06, .03)                          &     365       &  (99900, 100)                    &  $10^5$          &(100, 5)                                \\
+  Initial4    & SIR      &   Norm             & (.06, .03)                          &     365       &  (99000, 1000)                    &  $10^5$          &(100, 5)                                \\ \midrule
+  \textcolor{red}{PopSize1}        & SIR      &   Norm             & (.06, .03)                          &     365       &  (999500, 500)                    &  $10^6$          &(100, 5)                                \\
+  \textcolor{red}{PopSize2}       & SIR      &   Norm             & (.06, .03)                          &     365       &  (9995, 5)                    &  $10^4$          &(100, 5)                                \\
+  \textcolor{red}{PopSize3}        & SIR      &   Norm             & (.06, .03)                          &     365       &  (990, 10)                    &  $10^3$          &(100, 5)                                \\
+  \textcolor{red}{PopSize4}        & SIR      &   Norm             & (.06, .03)                          &     365       &  (99, 1)                    &  $10^2$          &(100, 5)                                \\
+ \midrule
+  Var1      & SIR      &   Norm             & (.06, .03)                          &     365       &  (99950, 50)                    &  $10^5$          &(50, 2)                                \\
+  Var2      & SIR      &   Norm             & (.06, .03)                          &     365       &  (99950, 50)                    &  $10^5$          &(500, 20)                                \\
+  Var3      & SIR      &   Norm             & (.06, .03)                          &     365       &  (99950, 50)                    &  $10^5$          &(2500, 1000)                                \\
+  Var4      & SIR      &   Norm             & (.06, .03)                          &     365       &  (99950, 50)                    &  $10^5$          &(10000, 500)                                \\ \midrule
+  \textcolor{red}{Poly1}     & Linear      &   Norm             & (.06, .03)                          &     365       &  (99950, 50)                    &  $10^5$          &(100, 5)                            \\
+  \textcolor{red}{Poly2}    & Quartic      &   Norm             & (.06, .03)                          &     365       &  (99950, 50)                    &  $10^5$          &(100, 5)                            \\
+   LSIR     & LSIR      &   Norm             & (.06, .03)                          &     365       &  (99950, 50)                    &  $10^5$          &(100, 5)                            \\ 
+  \bottomrule
+\end{tabular}
+\caption{Table of simulated data sets.  We simulate 27 total data sets with the given parameters.}
+\label{tab:simulated-data}
+\end{table}
+
+To test how well the SIR models perform under misspecification, we simulate data from three other models.
+While we will not have a ``known'' $\rr$ value to compare our results to, this will allow us to see if the models give us reasonable (and comparable) results when the data do not come from the SIR model.
 We simulate data from polynomials of degree 1 and 4 with respect to the time step, using a monotonically increasing function for the $X$ compartment and a monotonically decreasing function for the $Z$ compartment.
-The sum of these is subtracted from the constant total population to get the $Y$ compartment.
-We also generated data from a linearized version of the three compartment SIR model with the following ODEs:
+The sum of these is subtracted from the constant total population to obtain the $Y$ compartment.
+We call these models polynomial of degree 1 (Poly1) and polynomial of degree 4 (Poly4), respectively.
+We also generate data from a linearized version (in terms of the ODEs and not to be confused with LMA) of the three compartment SIR model (LSIR) with the following linearized ODEs:
 
@@
 
-The data from the linear SIR model were simulated under the following conditions:
-
-\begin{center}
-
-	$\beta$ = 0.06, $\gamma$ = 0.05; $\rr$ = 1.2
-
-	Number of time steps: 365
-
-	Starting X size ($X_0$): 99950
-
-	Starting Y size ($Y_0$): 50
-
-	Starting Z size: 0
-
-	$\sigma_X$: 10
-
-	$\sigma_Y$: 1
-
-	Total population: 100000.
-
-\end{center}
-
-\section{Results}\label{sec:results}
+% The data from the linear SIR model were simulated under the following conditions:
+
+% \begin{center}
+
+% 	$\beta$ = 0.06, $\gamma$ = 0.05; $\rr$ = 1.2
+
+% 	Number of time steps: 365
+
+% 	Starting X size ($X_0$): 99950
+
+% 	Starting Y size ($Y_0$): 50
+
+% 	Starting Z size: 0
+
+% 	$\sigma_X$: 10
+
+% 	$\sigma_Y$: 1
+
+% 	Total population: 100000.
+
+%       \end{center}
+
+\subsection{Model Fitting Details}\label{sec:model-fit}
+
+Table \ref{tab:baseline-res} shows the ways that standard errors are acquired for the different methods.
+The IPR and SIPR models require knowledge of the length of the average infectious period.
+For the results below, we assume the best case scenario that the true average infectious period is known for the data set.
+
+\begin{table}[H]
+	\centering
+	\begin{tabular}{@{}ll@{}}
+		\toprule
+		\textbf{Model} & \textbf{Default Standard Error Method} \\ \midrule
+		Least Squares (LS) & Delta method with Hessian matrix\\
+		Reparameterized Least Squares (ReLS) & Hessian matrix \\
+		Linear Model Approximation (LMA) & Delta method with linear regression standard error \\
+		LMA, All Time Points  (LMAT)& Delta method with linear regression standard error \\
+		Incidence-Prevalence Ratio (IPR) & Jackknife resampling \\
+		Smoothed IPR (SIPR) & Jackknife resampling \\
+		Log-Linear (LL) & Linear regression standard error \\
+		Markov Chain (MC) & Delta method with GLM regression standard error \\
+		Sequential Bayes (SB) & Standard deviation of posterior distribution\\
+		% Exponential Growth (EG) & Delta method with linear regression standard error \\
+		% IDEA & Delta method with linear regression standard error \\
+		\bottomrule
+	\end{tabular}
+	\caption{Table of methods for standard errors.}
+	\label{tab:se-methods}
+\end{table}
+
+Table \ref{tab:baseline-res} shows the ways that standard errors are acquired for the different methods.
+The IPR, SIPR, and IDEA models require knowledge of the length of the average infectious period.
+For the results below, we assume the best case scenario that the true average infectious period is known for the data set.
+The exponential growth model requires knowledge of the probability distribution of the infectious period.
+In this case, we assume a uniform probability distribution with a mean equivalent to the true average infectious period.
+
+\section{Results of Simulations}\label{sec:results}
 There are numerous aspects to consider when using a method to estimate $\rr$.
-Among these include values of both the infection and recovery rate $(\beta, \gamma)$, the number of data points available (time $T$), the total size of the population ($N$), the initial percent of susceptible individuals $\left (\frac{X(0)}{N}\right)$, the initial percent of infectious individuals $\left (\frac{Y(0)}{N}\right )$, and  the magnitude of the variance of both the number of susceptibles and the number of infectious ($\sigma_X, \sigma_Y$).
+Among these include values of both the infection and recovery rate $(\beta, \gamma)$, the number of data points available ($O$), the total size of the population ($N$), the initial percent of susceptible individuals $\left (\frac{X(0)}{N}\right)$, the initial percent of infectious individuals $\left (\frac{Y(0)}{N}\right )$, the magnitude of the variance of both the number of susceptibles and the number of infectious ($\sigma_X^2, \sigma_Y^2$), the error modeling assumptions, data generated from a non-SIR model, and finally methods of estimating the variance.
 We first examine our baseline data, and then we examine each of these aspects individually.
 
-\subsection{Baseline}\label{sec:res-base}
-
-\begin{figure}
+\subsection{Baseline Simulated Data}\label{sec:res-base}
+We implement the \wxxsir methods to estimate $\rr$ using the baseline data set.  The results are displayed in Figure \ref{fig:baseline-res} and Table \ref{tab:baseline-res}.
+\begin{figure}[H]
   \centering
@@
 Err.
-The red line is the true $\rr$ value.}
+The vertical line is the true $\rr$ value.}\label{fig:baseline-res}
   \end{figure}
@@
 		\hline
-		SIRLS & 1.9999711 & 0.0055923\\
-		\hline
-		SIRREPAR & 1.9997340 & 0.0049939\\
+		LS & 1.9999711 & 0.0055923\\
+		\hline
+		ReLS & 1.9997340 & 0.0049939\\
 		\hline
@@
 		\hline
-		SIRLMAT & 1.4063227 & 0.6309170\\
-		\hline
-		SIRIPR & 4.4203632 & 12.3593415\\
-		\hline
-		SIRSIPR & 1.8870824 & $<$ 1e-07 \\
-		\hline
-		LOGLINEAR & 1.9998072 & 0.0002198\\
-		\hline
-		SIRMC & 0.9485555 &  $<$ 1e-07 \\
-		\hline
-		SEQBAYES & 1.6495023 & 0.0672249\\
-		\hline
-	\end{tabular}
-        \caption{Baseline Data $(\beta=.06, \gamma=.03, O=365, X(0)=99950, Y(0)=50, Z(0)=0, \sigma_X=100, \sigma_Y=5, N=10^5)$, $\rr$ Estimates and Standard Errors}\label{tab:baseline}
-\end{table}
-
-Overall, we see a wide variation in the estimates of and the standard errors for $/rr$ depending on the method used.
-The methods that rely upon least squares optimization and the log-linear model are the closest to the true $\rr$ value, and they also have fairly small standard error estimates.
-On the other hand, a lot of the methods appear to underestimate the true $/rr$, and many of them have larger standard errors as well.
-Some notable exceptions include the IPR model, which tends to overestimate $/rr$, and the Markov chain and SIPR models, which appear to have unreasonably small standard errors.
+		LMAT & 1.4063227 & 0.6309170\\
+		\hline
+		IPR & 4.4203632 & 12.3593415\\
+		\hline
+		SIPR & 1.8870824 & $<$ 1e-07 \\
+		\hline
+		LL & 1.9998072 & 0.0002198\\
+		\hline
+		MC & 0.9485555 &  $<$ 1e-07 \\
+		\hline
+		SB & 1.6495023 & 0.0672249\\
+		\hline
+	\end{tabular}
+        \caption{Baseline Data $(\beta=.06, \gamma=.03, O=365, X(0)=99950, Y(0)=50, Z(0)=0, \sigma_X=100, \sigma_Y=5, N=10^5)$, $\rr$ Estimates and Standard Errors}\label{tab:baseline-res}
+\end{table}
+
+Overall, in Figure \ref{fig:baseline-res}, we see a wide variation in the estimates of and the standard errors for $\rr$ depending on the method used.
+The methods of LS, ReLS, LMA, LMAT, IPR, and LL all result in a CI that covers the true value of $\rr=2$.
+Of these, LMAT and IPR result in a CI which contains 1, so we would not be able to say whether or not we would have an outbreak of a disease or not.
+On the other hand, many methods appear to underestimate the true $\rr$, and many of them have larger standard errors as well.
+We would say that IPR's CI is uninformative.
+Additionally, MC and SIPR models seem to have unreasonably small standard errors.
 
 \subsection{Autoregressive Errors}\label{sec:res-AR}
-
+We next examine the estimates of $\rr$ when fit to the autoregressive simulated data with $\rho=0.7$.
+In this data, we may see errors accumulate over time, which may more accurately reflect the course of infection in an epidemic.
+For example, we may expect to see more variation around the peak of an epidemic than the beginning.
+Besides the AR error, the other parameters are the same as in the baseline model.
+The results of the estimates are displayed in Figure \ref{fig:ar-res} and Table \ref{tab:ar-res}.
+
+In this setting, the CIs from LS, ReLS, LMAT, SIPR, and LL cover the true $\rr=2$.
+In contrast to the baseline data, the CI for LMA no longer covers the true estimate.
+For all of these models, the standard error is larger, which we would expect given that the data have a large positive autocorrelation coefficient.
+
+\begin{figure}[H]
+  \centering
   \includegraphics[scale=0.5]{images/AR.jpeg}
-\begin{table}[H]
-
-	\caption{\label{tab:}Autoregressive Errors, $\rr$ Estimates and Standard Errors}
+  \caption{Forest plots of estimates of $\rr$ for the \xxsir methods for the autoregressive (AR) ($\rho=.7)$ simulated data set $(\beta=.06, \gamma=.03, O=365, X(0)=99950, Y(0)=50, Z(0)=0, \sigma_X=100, \sigma_Y=5, N=10^5)$.
+The points are the estimates of $\rr$ and the lines denote $\pm 2\cdot $Std.
+Err.
+The vertical line is the true $\rr$ value.}
+  \end{figure}\label{fig:ar-res}
+\begin{table}[H]
+
 	\centering
@@
 		\hline
-		SIRLS & 2.0000198 & 0.0078308\\
-		\hline
-		SIRREPAR & 1.9998852 & 0.0069901 \\
-		\hline
-		SIRLMA & 35.656866 & 4815.2163162\\
-		\hline
-		SIRLMAT & 1.049360 & 4.8830065 \\
-		\hline
-		SIRIPR & 2.183371 & 7.397726 \\
-		\hline
-		SIRSIPR & 0.8386350 & $<$ 1e-07 \\
-		\hline
-		LOGLINEAR & 1.9998203 & 0.0003227\\
-		\hline
-		SIRMC & 0.9485555 & $<$ 1e-07 \\
-		\hline
-		SEQBAYES & 1.0000000 & 0.0523424\\
-		\hline
-	\end{tabular}
-\end{table}
-
-With an autoregressive error structure in the simulated data, the log-linear and least squares estimation methods appear to have a similar accuracy.
-On the other hand, the other models mostly appear to be somewhat less accurate.
-In particular, the linear regression based models do much worse than in the Gaussian error case.
-For all of these models, the standard error is larger, which would expect given that the data have a large positive autocorrelation coefficient.
+		LS & 1.9997387 & 0.0071509\\
+		\hline
+		ReLS & 2.0000460 & 0.0063859\\
+		\hline
+		SIRLMA & 1.4744605 & 0.1130753\\
+		\hline
+		LMAT & 1.3694249 & 0.8681724\\
+		\hline
+		IPR & 1.7126909 & 5.0713852\\
+		\hline
+		SIPR & 0.8792348 & $<$ 1e-07 \\
+		\hline
+		LL & 2.0000004 & 0.0002885\\
+		\hline
+		MC & 0.9485555 & $<$ 1e-07 \\
+		\hline
+		SB & 1.0000000 & 0.0523424\\
+		\hline
+	\end{tabular}
+        \caption{Autoregressive Errors, $\rr$ Estimates and Standard Errors}\label{tab:ar-res}
+\end{table}
+
+\subsection{PAVA Monotonicity}\label{sec:res-PAVA}
+We examine the case of the PAVA error assumptions which enforces monotonic assumptions on the suceptible and recovered compartments.
+This in turn may more accurately model the course of an epidemic, but as a consequence, we may not be able to guarantee that a 95\% CI covers the true value of $\rr$ 95\% of the time.
+
+In Figure \ref{fig:pava-res} and Table \ref{tab:pava-res}, we display the results of fitting the methods on two data sets, those generated from the Norm-G and Norm-AR error assumptions.
+We first see that there is very little difference among the estimates of $\rr$ in both Norm-G and Norm-AR error assumptions.
+We see that LMA and IPR have smaller estimates of $\rr$ and SB has a larger estimate of $\rr$.
+The standard errors for Norm-AR error assumption are greater than or equal to than the standard error in the Norm-G error assumption with the exceptions of LMAT and SB.
+
+The methods that contain the true value of $\rr$ are LS, ReLS, LMAT, IPR for both error assumptions and LL with Norm-AR error assumption.
+The methods LS, ReLS, LMA, and LL would be able to reject the hypothesis that $\rr < 1$.
+
+\begin{figure}[H]
+	\centering
+	\includegraphics[scale=0.5]{images/mono.jpeg}
+	\caption{Forest plots of estimates of $\rr$ for the \xxsir methods for the Norm-M and AR-M data sets $(\beta=.06, \gamma=.03, O=365, X(0)=99950, Y(0)=50, Z(0)=0, \sigma_X=100, \sigma_Y=5, N=10^5)$.
+The points are the estimates of $\rr$ and the lines denote $\pm 2\cdot $Std.
+Err.
+The vertical lines are the true $\rr$ value.}
+	\label{fig:pava-res}
+\end{figure}
+
+\begin{table}[H]
+
+	\centering
+	\begin{tabular}[t]{l|r|r|r|r}
+		\hline
+		Model & Norm-M Est. & Norm-M SE & AR-M Est. & AR-E SE\\
+		\hline
+		LS & 2.0002458 & 0.0044456 & 1.9999920 & 0.0073664\\
+		\hline
+		ReLS & 2.0003143 & 0.0039684 & 1.9996837 & 0.0065745\\
+		\hline
+		SIRLMA & 1.6793591 & 0.1102722 & 1.5349785 & 0.1496989\\
+		\hline
+		LMAT & 1.4422630 & 0.9241642 & 1.4491091 & 0.7463529\\
+		\hline
+		IPR & 0.7485576 & 1.5046827 & 0.7513498 & 1.5606142\\
+		\hline
+		SIPR & 0.9081552 & 0.0000000 & 0.6976176 & 0.0000000\\
+		\hline
+		LL & 2.0003027 & 0.0001342 & 1.9995563 & 0.0002342\\
+		\hline
+		MC & 0.9485555 & 0.0000000 & 0.9485555 & 0.0000000\\
+		\hline
+		SB & 1.0000000 & 0.0523424 & 1.7143505 & 0.0685336\\
+		\hline
+	\end{tabular}
+\caption{Gaussian and Autoregressive Data with PAVA Applied, $\rr$ Estimates and Standard Errors}\label{tab:pava-res}
+\end{table}
+
+In comparison to the baseline data set, the estimates of $\rr$ from the methods fit on the Norm-G error assumption data set have smaller CIs.
+The estimates of $\rr$ for the Norm-G error assumption are very similar to those from the baseline assumption.
+LMA and LMAT have estimates closer to the true $\rr$ in the Norm-G error assumption data setting compared to the baseline data and SB has an estimate further away.
+
+Overall, there is not much difference between the baseline and monotone data sets, at least in the case with $\sigma_X=100$ and $\sigma_Y=5$.
 
@@
 
+We next examine the simulated data sets InfRec1-4, three of which have different levels of $\beta$ for a given level of $\gamma=.001$ and the fourth is simulated with $\rr=0.25 < 1$, a case where an outbreak would not occur.
+The idea behind this is that since $\rr=\frac{\beta}{\gamma}$ and ratio estimators typically have large variance, we want to examine how a small value of $\gamma$, that is a small recovery rate, would effect the resulting estimates of $\rr$.
+Additionally, we would like to know when a disease will die out on its own accord, that we can be certain the estimate of $\rr$ is less than 1.
+
+\begin{figure}[H]
+  \centering
   \includegraphics[scale=0.5]{images/parchange.jpeg}
-
-\begin{table}[H]
-
-	\caption{\label{tab:}$\gamma = 0.001$, $\rr$ Estimates and Standard Errors}
+  \caption{Forest plots of estimates of $\rr$ for the \xxsir methods for the infrec1-4 data sets $(\beta=.06, \gamma=.03, O=365, X(0)=99950, Y(0)=50, Z(0)=0, \sigma_X=100, \sigma_Y=5, N=10^5)$.
+The points are the estimates of $\rr$ and the lines denote $\pm 2\cdot $Std.
+Err.
+The vertical line is the true $\rr$ value.}
+  \label{fig:infrec1-res}
+\end{figure}
+\begin{table}[H]
+
 	\centering
@@
 		\hline
-		SIRREPAR & 59.9951876 & 0.4343984\\
+		ReLS & 59.9951876 & 0.4343984\\
 		\hline
@@
 		\hline
-		SIRLMAT & -2.8535930 & 193.6526107 \\
-		\hline
-		SIRIPR & 54.146506 & 200.2595\\
-		\hline
-		SIRSIPR & 3.9645187 & $<$ 1e-07\\
-		\hline
-		LOGLINEAR & 46.3209596 & 0.7797216\\
-		\hline
-		SIRMC & 0.9485554 & 0.0000001\\
-		\hline
-		SEQBAYES & 0.9769731 & 0.0517362\\
-		\hline
-	\end{tabular}
-\end{table}
-
-Looking at the case where $\gamma = 0.001$, it appears that most of the models do not handle a very large $/rr$ well.
-Only the least squares and reparameterized least squares models give very close answers; the IPR and log-linear models give fairly close answers as well.
-However, the other models do very poorly; they ted to greatly underestimate $/rr$.
-The linear model approximation does particularly poorly, likely because the linear model does not fit the data well.
+		LMAT & -2.8535930 & 193.6526107 \\
+		\hline
+		IPR & 54.146506 & 200.2595\\
+		\hline
+		SIPR & 3.9645187 & $<$ 1e-07\\
+		\hline
+		LL & 46.3209596 & 0.7797216\\
+		\hline
+		MC & 0.9485554 & 0.0000001\\
+		\hline
+		SB & 0.9769731 & 0.0517362\\
+		\hline
+	\end{tabular}
+        \caption{\label{tab:infrec1-res}$\gamma = 0.001$, $\rr$ Estimates and Standard Errors}
+\end{table}
+
+Looking in Figure \ref{fig:infrec1-res} and Table \ref{tab:infrec1-res} at the cases where $\gamma = 0.001$ with $\beta=60000, 1500$, and 1000, it appears that most of the models do not handle a small $\gamma$ well.
+We examine an exaggerated extreme case of when $\rr=60$ (the largest reported $\rr$ value we found is $\hat{\rr}=14$ for measles as reported in \cite{anderson1992}).
+Only LS and ReLS result in CIs that cover the true $\rr$.
+IPR and LL give fairly close point estimates to the true $\rr=60$ as well.
+However, the other models do very poorly; they tend to greatly underestimate $\rr$.
+The linear model approximation does particularly poorly, likely because the linear model does not fit the data well, and is not even constrained to positive values of $X$, $Y$, and $Z$, which is why we can result in a negative estimate of $\rr$.
 Several of the models also return extremely large standard errors, meaning they are not very reliable in these cases.
-
-\begin{table}[H]
-
-	\caption{\label{tab:}$\gamma = 0.24$, $\rr$ Estimates and Standard Errors}
+Only LS, ReLS, and LL would reject the hypothesis that $\rr <1$, when this is clearly not the case at all.
+
+When $\rr=1.5$ we see that the estimates from the methods give results much closer to the true value, and the resulting CIs are also smaller.
+The methods LS, ReLS, SIPR, LL, and SB would all reject the hypothesis that $\rr > 1$.
+Although if we are testing the null hypothesis for  all methods simultaneously, we should use a multiple-testing correction such as Bonferroni.
+Under the Bonferroni correction in this case, the only method which may then fail to reject would be SB as the other SEs are very small).
+
+When $\rr=1$, we are unsure whether there will be an epidemic or not.
+In this case, all the CIs of the methods cover the true value.
+We also see that the SE of LS is much greater than ReLS, which may be surprising.
+((WHY?))
+
+\begin{table}[H]
+
 	\centering
@@
 		\hline
-		SIRLS & 0.2363274 & 4.1042813\\
-		\hline
-		SIRREPAR & 4.1795252 & 2.7455569\\
+		LS & 0.2363274 & 4.1042813\\
+		\hline
+		ReLS & 4.1795252 & 2.7455569\\
 		\hline
@@
 		\hline
-		SIRLMAT & 0.6746914 & 6.2471252 \\
-		\hline
-		SIRIPR & 181.015802 & 742.1088 \\
-		\hline
-		SIRSIPR & -0.2484284 & 0.1159769 \\
-		\hline
-		LOGLINEAR & 0.9912397 & 0.0030505\\
-		\hline
-		SIRMC & 0.9485811 & 0.0000150\\
-		\hline
-		SEQBAYES & 1.2354952 & 0.0581801\\
-		\hline
-	\end{tabular}
-\end{table}
-
-The above table corresponds to the case where $\gamma = 0.24$.
-It appears that a lot of estimation methods do not handle very small $/rr$ well either.
-None of them are extremely close to truth, except for regular least squares, as most of the tend to overestimate the truth.
+		LMAT & 0.6746914 & 6.2471252 \\
+		\hline
+		IPR & 181.015802 & 742.1088 \\
+		\hline
+		SIPR & -0.2484284 & 0.1159769 \\
+		\hline
+		LL & 0.9912397 & 0.0030505\\
+		\hline
+		MC & 0.9485811 & 0.0000150\\
+		\hline
+		SB & 1.2354952 & 0.0581801\\
+		\hline
+	\end{tabular}
+        \caption{$\gamma = 0.24$, $\rr$ Estimates and Standard Errors}\label{tab:infrec4-res}
+\end{table}
+
+Table \ref{tab:infrec4-res} corresponds to the case where $\gamma = 0.24$ and $\beta=0.06$.
+It appears that a lot of estimation methods do not handle small $\rr$ well either.
+None of them are extremely close to truth, except for ReLS, as most of the tend to overestimate the truth.
 The IPR is particularly bad in this case.
@@
 
-\subsection{Number of Time Steps ($T$)}\label{sec:res-time}
-
+\subsection{Number of Time Points ($O$)}\label{sec:res-time}
+We examine how the number of time steps the estimates of $\rr$ in the different methods.
+Ideally, the more data we have, the more likely we are to have lower-variance estimates for $\rr$.
+Since we would like to  infer how a disease will spread through a population, the sooner we can be sure whether $\rr> 1$, the sooner we can implement interventions and control measures for a disease.
+Another reason to analyze the number of time steps is that some of the methods assume linear or exponential growth of an epidemic in a certain time period.
+If the growth is no longer exponential, we will obtain a poor estimate of $\rr$.
+In this case, we can be using too many time points above a certain time.
+
+In Figure \ref{fig:time-res} and Tables \ref{tab:time-res1} and \ref{tab:time-res2} we examine the methods fit to different amounts of time periods where we observe the progress of a disease, including $T=200$, 100, 50, and 20.
+
+Again, we see that IPR does very poorly in estimating $\rr$, regardless of the amount of data observed.
+\begin{figure}[H]
+  \centering
   \includegraphics[scale=0.5]{images/time.jpeg}
-
-\begin{table}[H]
-
-	\caption{\label{tab:} $T$ = 100, $\rr$ Estimates and Standard Errors}
+  \caption{Forest plots of estimates of $\rr$ for the \xxsir methods for the Time1-4 data sets $(\beta=.06, \gamma=.03, O=365, X(0)=99950, Y(0)=50, Z(0)=0, \sigma_X=100, \sigma_Y=5, N=10^5)$.
+The points are the estimates of $\rr$ and the lines denote $\pm 2\cdot $Std.
+Err.
+The vertical line is the true $\rr$ value.}\label{fig:time-res}
+  \end{figure}
+
+\begin{table}[H]
+
 	\centering
@@
 		\hline
-		SIRLS & 1.9993392 & 0.0286060\\
-		\hline
-		SIRREPAR & 1.9994284 & 0.0312814\\
+		LS & 1.9993392 & 0.0286060\\
+		\hline
+		ReLS & 1.9994284 & 0.0312814\\
 		\hline
@@
 		\hline
-		SIRLMAT & 3.3442696 & 21.4895422\\
-		\hline
-		SIRIPR & 7.3458131 & 16.1288906\\
-		\hline
-		SIRSIPR & -2.1198897 & $<$ 1e-07\\
-		\hline
-		LOGLINEAR & 2.0014146 & 0.0017620\\
-		\hline
-		SIRMC & 0.9485555 & 0.0000001\\
-		\hline
-		SEQBAYES & 2.4289382 & 0.1102030\\
-		\hline
-	\end{tabular}
-\end{table}
-
-The table above shows the methods applied to the first 100 time steps, out of 365, of a baseline data set.
-The estimates and standard errors are worse, which we would expect given less data.
+		LMAT & 3.3442696 & 21.4895422\\
+		\hline
+		IPR & 7.3458131 & 16.1288906\\
+		\hline
+		SIPR & -2.1198897 & $<$ 1e-07\\
+		\hline
+		LL & 2.0014146 & 0.0017620\\
+		\hline
+		MC & 0.9485555 & 0.0000001\\
+		\hline
+		SB & 2.4289382 & 0.1102030\\
+		\hline
+	\end{tabular}
+        \caption{ $T$ = 100, $\rr$ Estimates and Standard Errors}\label{tab:time-res1}
+\end{table}
+
+Table \ref{tab:time-res1} shows the methods applied to the first 100 time steps, out of 365, of the baseline data set .
+The estimates and standard errors are worse than baseline, which we would expect given less data.
 The results are still relatively close to the full baseline data.
-There also appears to be less tendency to underestimate $/rr$ compared to a longer time scale.
-
-\begin{table}[H]
-
-	\caption{\label{tab:} $T$ = 20, $\rr$ Estimates and Standard Errors}
+There also appears to be less tendency to underestimate $\rr$ compared to a longer time scale.
+In general, we see that that the length of CIs decreases as we use more and more time points.
+Only LL would reject the hypothesis that $\rr < 1$ for 50 or fewer time points, but its CI does not cover the true $\rr$ value.
+With 50 and greater time points, we see that LS and ReLS provide close point estimates to the true value and by 100 time points, would reject the hypothesis that $\rr < 1$.
+The methods and MC and SB consistently have small CIs, but never cover the true value.
+
+\begin{table}[H]
   \centering
@@
 		\hline
-		SIRLS & 38.6895520 & 106.2705193\\
-		\hline
-		SIRREPAR & 22.9645867 & 6.9816713\\
+		LS & 38.6895520 & 106.2705193\\
+		\hline
+		ReLS & 22.9645867 & 6.9816713\\
 		\hline
@@
 		\hline
-		SIRLMAT & 1.0078807 & 0.1424793\\
-		\hline
-		SIRIPR & 25.7313065 & 35.2926077\\
-		\hline
-		SIRSIPR & 4.1772451 & 7.3202970\\
-		\hline
-		LOGLINEAR & 1.0429764 & 0.0276261\\
-		\hline
-		SIRMC & 0.9485556 & 0.0000005\\
-		\hline
-		SEQBAYES & 1.0000000 & 0.2236068\\
-		\hline
-	\end{tabular}
-\end{table}
-
-The table above shows the methods applied to the first 20 time steps of a baseline data set.
-As might be expected, the $/rr$ estimates are wildly inaccurate, and none are very close.
-The least squares methods are particularly worse in this situation compared to the other data sets.
+		LMAT & 1.0078807 & 0.1424793\\
+		\hline
+		IPR & 25.7313065 & 35.2926077\\
+		\hline
+		SIPR & 4.1772451 & 7.3202970\\
+		\hline
+		LL & 1.0429764 & 0.0276261\\
+		\hline
+		MC & 0.9485556 & 0.0000005\\
+		\hline
+		SB & 1.0000000 & 0.2236068\\
+		\hline
+	\end{tabular}
+        \caption{ $T$ = 20, $\rr$ Estimates and Standard Errors}\label{tab:time-res2}
+\end{table}
+
+Table \ref{tab:time-res2} shows the methods applied to the first 20 time steps of a baseline data set.
+As might be expected, the $\rr$ estimates are wildly inaccurate, and none are very close to the true value.
+The methods of LS and ReLS are particularly worse in this situation compared to the other data sets.
 The standard errors are much larger than the baseline, which is to be expected as well due to the small number of data points.
-This shows that it is extremely difficult to create accurate online estimates of $/rr$ or to estimate $/rr$ without adequate data.
+This shows that it is extremely difficult to create accurate online estimates of $\rr$ or to estimate $\rr$ without adequate data.
 
 \subsection{Initial Percentage of Infectious $\left (\frac{Y(0)}{N}\right)$}\label{sec:res-inf}
-
+We next examine the initial conditions of the number of susceptible, infectious, and recovered.
+Deterministically, different initial conditions in the number of infectious can cause very different epidemic courses (see Figures in appendix).
+
+In Figure \ref{fig:inits-res} and Tables \ref{tab:inits-res1} and \ref{tab:inits-res2}, we display the results of different number of initially infectious individuals of 1, 10, 100, and 1000 with a constant population of 100,000.
+We see that for a smaller percentage of initially infectious individuals, we have larger CIs in our estimates of $\rr$.
+In fact, LL is the only method that has a consistently small CI around the true $\rr=2$, regardless of the starting number of infectious.
+
+\begin{figure}[H]
+  \centering
   \includegraphics[scale=0.5]{images/start.jpeg}
-
-\begin{table}[H]
-
-	\caption{\label{tab:}$Y(0) = 1000$, $\rr$ Estimates and Standard Errors}
+    \caption{Forest plots of estimates of $\rr$ for the \xxsir methods for the Inits1-4 data sets $(\beta=.06, \gamma=.03, O=365, X(0)=99950, Y(0)=50, Z(0)=0, \sigma_X=100, \sigma_Y=5, N=10^5)$.
+The points are the estimates of $\rr$ and the lines denote $\pm 2\cdot $Std.
+Err.
+The vertical line is the true $\rr$ value.}\label{fig:inits-res}
+  \end{figure}
+
+  Table \ref{tab:inits-res1} shows the methods applied to data where $Y(0) = 1000$, essentially starting at the middle of an epidemic.
+This data has very large underestimation of $\rr$.
+The LMA method once again gives an estimate very far from the true $\rr$, potentially indicating that a linear approximation does not apply in this case.
+The sample standard errors in this case are actually relatively close to the baseline, and even smaller for some of the methods.
+
+Table \ref{tab:inits-res2} shows the methods applied to data where $Y(0) = 1$, starting at the very beginning of an epidemic, and the true technical definition of $\rr$ where we introduce exactly one infectious individual into a population.
+The methods give many larger estimates, tending to overestimate the true $\rr$.
+In particular, IPR and LMA methods give very large results.
+We also see a larger standard error than the baseline.
+
+\begin{table}[H]
+
 	\centering
@@
 		\hline
-		SIRLS & 1.9995884 & 0.0049609\\
-		\hline
-		SIRREPAR & 1.9997123 & 0.0034303\\
+		LS & 1.9995884 & 0.0049609\\
+		\hline
+		ReLS & 1.9997123 & 0.0034303\\
 		\hline
@@
 		\hline
-		SIRLMAT & 0.7306100 & 4.681042\\
-		\hline
-		SIRIPR & 1.285071 & 1.440980\\
-		\hline
-		SIRSIPR & 1.2522106 & 57.9229822 \\
-		\hline
-		LOGLINEAR & 1.9995148 & 0.0002828\\
-		\hline
-		SIRMC & 0.9485555 & $<$ 1e-07\\
-		\hline
-		SEQBAYES & 0.9728534 & 0.0516270\\
-		\hline
-	\end{tabular}
-\end{table}
-
-The table above shows the methods applied to data where $Y(0) = 1000$, essentially starting at the middle of an epidemic.
-This data has very large underestimation of $/rr$.
-The LMA method once again gives a very far estimate, potentially indicating that a linear approximation does not apply in this case.
-The sample standard errors in this case are actually relatively close to the baseline, and even smaller for some of the methods.
-
-\begin{table}[H]
-
-	\caption{\label{tab:}$Y(0) = 1$, $\rr$ Estimates and Standard Errors}
+		LMAT & 0.7306100 & 4.681042\\
+		\hline
+		IPR & 1.285071 & 1.440980\\
+		\hline
+		SIPR & 1.2522106 & 57.9229822 \\
+		\hline
+		LL & 1.9995148 & 0.0002828\\
+		\hline
+		MC & 0.9485555 & $<$ 1e-07\\
+		\hline
+		SB & 0.9728534 & 0.0516270\\
+		\hline
+	\end{tabular}
+        \caption{$Y(0) = 1000$, $\rr$ Estimates and Standard Errors}\label{tab:inits-res1}
+\end{table}
+
+\begin{table}[H]
+
 	\centering
@@
 		\hline
-		SIRLS & 2.468304 & 1.5370731\\
-		\hline
-		SIRREPAR & 2.469189 & 2.7293667\\
+		LS & 2.468304 & 1.5370731\\
+		\hline
+		ReLS & 2.469189 & 2.7293667\\
 		\hline
@@
 		\hline
-		SIRLMAT & 1.624645 & 5.0359845\\
-		\hline
-		SIRIPR & 774.274659 & 7491.005080 \\
-		\hline
-		SIRSIPR & 2.4390757 & $<$ 1e-07 \\
-		\hline
-		LOGLINEAR & 1.999690 & 0.0007731\\
-		\hline
-		SIRMC & 0.948603 & 0.0000156\\
-		\hline
-		SEQBAYES & 1.471481 & 0.0634937\\
-		\hline
-	\end{tabular}
-\end{table}
-
-The table above shows the methods applied to data where $Y(0) = 1$, starting at the very beginning of an epidemic.
-The methods give many larger estimates, tending to overestimate the true $/rr$.
-In particular, the IPR and LMA methods give very large results.
-We also see a large standard error than the baseline.
+		LMAT & 1.624645 & 5.0359845\\
+		\hline
+		IPR & 774.274659 & 7491.005080 \\
+		\hline
+		SIPR & 2.4390757 & $<$ 1e-07 \\
+		\hline
+		LL & 1.999690 & 0.0007731\\
+		\hline
+		MC & 0.948603 & 0.0000156\\
+		\hline
+		SB & 1.471481 & 0.0634937\\
+		\hline
+	\end{tabular}
+        \caption{$Y(0) = 1$, $\rr$ Estimates and Standard Errors}\label{tab:inits-res2}
+\end{table}
+
+\subsection{Population Size $N$}\label{sec:res-n}
+We next examine the effects on the estimates of $\rr$ when changing the total population size, $N$.
+The percent of initially infectious individuals is 1\% when $N=10,000$ and $N=100,000$ and .5\% when $N=1,000$ and $N=100$.
+Intuitively, we would expect to see the variance decrease and to have more accurate estimates with a larger $N$ compared to a smaller one, due to the central limit theorem.
+This is exactly what we see for some of the methods, but not all of them.
+
+We display the results of estimating $\rr$ from the methods fit on the baseline data set with varying $N$ in Figure \ref{fig:inits-res} and Tables \ref{tab:n1-res2} and \ref{tab:n2-res2}.
+When $N=100$, LS, ReLS, LMA, LMAT, IPR, and SIPR's 95\% CIs all cover the true $\rr$.
+However, we would say that all of these intervals are uninformative.
+We see that CIs of LS are wider than those of ReLS, although the difference becomes less substantial as $N$ grows larger.
+
+When $N=100$, we would be able to say the estimate for $\rr$ from LL is greater than 1 with 95\% confidence, but the estimate is much smaller than the true value.
+MC and SB would actually allow us to say $\rr < 1$ with 95\% confidence.
+It is not until $N=10,000$ that we are able to say with certainty that LS and ReLS, in addition to LL, give informative estimates of $\rr$.
+
+Thus, it seems that if one is estimating for $\rr$ for a small population such as a small village or crusie ship, LL may be the most appropriate method, but when the population is larger (~10,000), we can be confident in using LS or ReLS.
+
+\begin{figure}[H]
+	\centering
+	\includegraphics[scale=0.5]{images/popsize.jpeg}
+	\caption{Forest plots of estimates of $\rr$ for the \xxsir methods for the PopSize1-4 data sets $(\beta=.06, \gamma=.03, O=365, \sigma_X=100, \sigma_Y=5)$.
+The points are the estimates of $\rr$ and the lines denote $\pm 2\cdot $Std.
+Err.
+The vertical line is the true $\rr$ value.}\label{fig:inits-res2}
+\end{figure}
+
+\begin{table}[H]
+
+	\centering
+	\begin{tabular}[t]{l|r|r}
+		\hline
+		Model & Estimate & Std. Err\\
+		\hline
+		LS & 2.0007741 & 0.0590423\\
+		\hline
+		ReLS & 2.0005189 & 0.0528419\\
+		\hline
+		SIRLMA & 1.9976309 & 2606.0798123\\
+		\hline
+		LMAT & 1.2944136 & 1.4304706\\
+		\hline
+		IPR & 40.8072373 & 133.2798314\\
+		\hline
+		SIPR & 1.0628910 &  $<$ 1e-07\\
+		\hline
+		LL & 1.9988393 & 0.0023193\\
+		\hline
+		MC & 0.9485626 & 0.0000040\\
+		\hline
+		SB & 0.4418182 & 0.0347917\\
+		\hline
+	\end{tabular}
+\caption{$N = 10000$, $\rr$ Estimates and Standard Errors}\label{tab:n1-res2}
+\end{table}
+
+\begin{table}[H]
+
+	\centering
+	\begin{tabular}[t]{l|r|r}
+		\hline
+		Model & Estimate & Std. Err\\
+		\hline
+		LS & 1.9862533 & 4.861988\\
+		\hline
+		ReLS & 1.9871049 & 3.456891\\
+		\hline
+		SIRLMA & 0.8023469 & 18.94166\\
+		\hline
+		LMAT & 1.2208744 & 4.769929\\
+		\hline
+		IPR & 3620.8676795 & 57608.17\\
+		\hline
+		SIPR & 3.3678247 & 22.41867\\
+		\hline
+		LL & 1.1413175 & 0.0422219\\
+		\hline
+		MC & 0.8455334 & 0.0124837\\
+		\hline
+		SB & 0.3887115 & 0.0326338\\
+		\hline
+	\end{tabular}
+\caption{$N = 100$, $\rr$ Estimates and Standard Errors}\label{tab:n2-res2}
+\end{table}
 
 \subsection{Magnitude of Variance $(\sigma^2_X, \sigma^2_Y)$}\label{sec:res-var}
-
+We examine the estimates of $\rr$ when changing the values of the variance of the number of susceptibles and number of infectious.
+We would expect to see wider CIs when we have larger variances in the $X$ and $Y$ compartments, but we would like to determine how sensitive estimates of $\rr$ are to these variance values.
+
+\begin{figure}[H]
+\begin{center}
   \includegraphics[scale=0.5]{images/var.jpeg}
-
-\begin{table}[H]
-
-	\caption{\label{tab:} $(\sigma^2_X, \sigma^2_Y) = (50, 2)$, $\rr$ Estimates and Standard Errors}
+  \caption{Forest plots of estimates of $\rr$ for the \xxsir methods for the Var1-4 data sets $(\beta=.06, \gamma=.03, O=365, T=365, X(0)=99950, Y(0)=50, Z(0)=0, N=10^5)$.
+The points are the estimates of $\rr$ and the lines denote $\pm 2\cdot $Std.
+Err.
+The vertical line is the true $\rr$ value.}
+  \label{fig:mag-var-res}
+\end{center}
+\end{figure}
+
+We display the results in Figure \ref{fig:mag-var-res} and Tables \ref{tab:mag-var-res1}-\ref{tab:mag-var-res2}.
+Generally, we see wider CIs and more biased estimates of $\rr$ when the variance levels are large.
+In particular, LMA and LMAT seem to be particularly effected as the CIs grow substantially as the variance grows.
+Again we see that LS, ReLS, and LL provide the most accurate estimates with relatively small CIs.
+
+\begin{table}[H]
+
 	\centering
@@
 		\hline
-		SIRLS & 1.9997552 & 0.0026502\\
-		\hline
-		SIRREPAR & 1.9998295 & 0.0023646\\
+		LS & 1.9997552 & 0.0026502\\
+		\hline
+		ReLS & 1.9998295 & 0.0023646\\
 		\hline
@@
 		\hline
-		SIRLMAT & 1.4148628 & 0.5779651\\
-		\hline
-		SIRIPR & 2.7607518 & 8.1318687\\
-		\hline
-		SIRSIPR & 1.8368240 & $<$ 1e-07\\
-		\hline
-		LOGLINEAR & 1.9998907 & 0.0001037\\
-		\hline
-		SIRMC & 0.9485555 & $<$ 1e-07\\
-		\hline
-		SEQBAYES & 1.6750053 & 0.0677426\\
-		\hline
-	\end{tabular}
-\end{table}
-
-In the table above, the magnitude of the variance of the noise is smaller than in the baseline.
-The estimates are closer to the true $/rr$ and the standard errors are smaller across the board.
+		LMAT & 1.4148628 & 0.5779651\\
+		\hline
+		IPR & 2.7607518 & 8.1318687\\
+		\hline
+		SIPR & 1.8368240 & $<$ 1e-07\\
+		\hline
+		LL & 1.9998907 & 0.0001037\\
+		\hline
+		MC & 0.9485555 & $<$ 1e-07\\
+		\hline
+		SB & 1.6750053 & 0.0677426\\
+		\hline
+	\end{tabular}
+        	\caption{ $(\sigma^2_X, \sigma^2_Y) = (50, 2)$, $\rr$ Estimates and Standard Errors}\label{tab:mag-var-res1}
+\end{table}
+
+In Table \ref{tab:mag-var-res1}, the magnitude of the variance of the noise is smaller than in the baseline.
+The estimates are closer to the true $\rr$ and the standard errors are smaller across the board.
 This indicates that the methods give more accurate and precise estimates when the data are closer to following underlying true SIR model.
@@
 
-	\caption{\label{tab:} $(\sigma^2_X, \sigma^2_Y) = (2500, 100)$, $\rr$ Estimates and Standard Errors}
 	\centering
@@
 		\hline
-		SIRLS & 1.9980880 & 0.1353803\\
-		\hline
-		SIRREPAR & 1.9981446 & 0.1207935\\
+		LS & 1.9980880 & 0.1353803\\
+		\hline
+		ReLS & 1.9981446 & 0.1207935\\
 		\hline
@@
 		\hline
-		SIRLMAT & 1.2877865 & 4.3947658\\
-		\hline
-		SIRIPR & 81.4901666 & 348.7350849\\
-		\hline
-		SIRSIPR & 1.0606095 & $<$ 1e-07\\
-		\hline
-		LOGLINEAR & 2.0018816 & 0.0054085\\
-		\hline
-		SIRMC & 0.9485562 & 0.0000010\\
-		\hline
-		SEQBAYES & 1.8567176 & 0.0713225\\
-		\hline
-	\end{tabular}
-\end{table}
-
-In the table above, the magnitude of the variance of the noise is larger than in the baseline.
+		LMAT & 1.2877865 & 4.3947658\\
+		\hline
+		IPR & 81.4901666 & 348.7350849\\
+		\hline
+		SIPR & 1.0606095 & $<$ 1e-07\\
+		\hline
+		LL & 2.0018816 & 0.0054085\\
+		\hline
+		MC & 0.9485562 & 0.0000010\\
+		\hline
+		SB & 1.8567176 & 0.0713225\\
+		\hline
+	\end{tabular}
+        	\caption{$(\sigma^2_X, \sigma^2_Y) = (2500, 100)$, $\rr$ Estimates and Standard Errors}\label{tab:mag-var-res2}
+\end{table}
+
+In Table \ref{tab:mag-var-res2}, the magnitude of the variance of the noise is larger than in the baseline.
 The estimates are worse and the standard errors are larger compared to the baseline data set, which we might expect.
@@
 \subsection{Other Models}\label{sec:res-oth}
-
-\begin{table}[H]
-
+Until this point, we have only examined results while assuming that the epidemic truly follows a SIR model.
+However, this may not be the case.
+We examine the estimates of $\rr$ under different underlying models including a polynomial of degree 1, a polynomial of degree 4, and a linearized SIR model.
+We present the results of the estimates of $\rr$ on these models in Figure \ref{fig:other-res} and Tables \ref{tab:other-res1} - \ref{tab:other-res3}.
+
+\begin{figure}[H]
   \begin{center}
     \includegraphics[scale=0.5]{images/other.jpeg}
-
+  \caption{Forest plots of estimates of $\rr$ for the \xxsir methods for the Poly1, Poly4, and LSIR data sets $(\beta=.06, \gamma=.03, O=365, T=365, X(0)=99950, Y(0)=50, Z(0)=0, N=10^5, \sigma_X=100, \sigma_Y=5)$.
+The points are the estimates of $\rr$ and the lines denote $\pm 2\cdot $Std.
+Err.
+The vertical line is the true $\rr$ value.}
+  \label{fig:other-res}	
   \end{center}
-
-\caption{\label{tab:} 4th Order Model, $\rr$ Estimates and Standard Errors}
+\end{figure}
+
+For both the Poly1 and LSIR data sets, all the methods besides IPR and SIPR underestimate $\rr$.
+Moreover, only the CI of IPR covers that of the true value of $\rr=2$ and the CIs are uninformative.
+For the quartic model, only ReLS has a CI that both covers the true $\rr$ while rejecting the hypothesis that $\rr < 1$.
+LS covers the true value with its CI but also includes $\rr=1$.
+LL rejects the hypothesis that $\rr <$ but its CI does not contain $\rr=2$.
+
+\begin{table}[H]
+
 	\centering
@@
 		\hline
+		Model & Estimate & Std. Err\\
+		\hline
+		LS & 1.0514375 & 0.3782909\\
+		\hline
+		ReLS & 1.0514351 & 0.0259416\\
+		\hline
+		SIRLMA & 1.2655140 & 0.1278435\\
+		\hline
+		LMAT & 1.1597070 & 0.1019956\\
+		\hline
+		IPR & 4.4999548 & 15.2185520\\
+		\hline
+		SIPR & 98.0610578 & $<$ 1e-07\\
+		\hline
+		LL & 1.1880504 & 0.0006341\\
+		\hline
+		MC & 0.9485551 & 0.0000001\\
+		\hline
+		SB & 1.4150329 & 0.0622640\\
+		\hline
+	\end{tabular}
+	\caption{Linear Data, $\rr$ Estimates and Standard Errors}\label{tab:other-res1}
+\end{table}
+
+\begin{table}[H]
+
+\centering
+\begin{tabular}[t]{l|r|r}
+	\hline
 	Model & Estimate & Std. Dev\\
 	\hline
-	SIRLS & 1.7693473 & 0.4226615\\
-	\hline
-	SIRREPAR & 1.7693312 & 0.2796903\\
+	LS & 1.7693473 & 0.4226615\\
+	\hline
+	ReLS & 1.7693312 & 0.2796903\\
 	\hline
@@
 	\hline
-	SIRLMAT & 0.8213066 & 6.0466473\\
-	\hline
-	SIRIPR & 10.4349388 & 33.8952710\\
-	\hline
-	SIRSIPR & 4.7477067 &  $<$ 1e-07\\
-	\hline
-	LOGLINEAR & 1.9551795 & 0.0164608\\
-	\hline
-	SIRMC & 0.9485545 & 0.0000006\\
-	\hline
-	SEQBAYES & 0.2118519 & 0.0240918\\
-	\hline
-\end{tabular}
-\end{table}
-
-The table above shows the methods applied to data from a 4th order linear regression model with respect to time.
+	LMAT & 0.8213066 & 6.0466473\\
+	\hline
+	IPR & 10.4349388 & 33.8952710\\
+	\hline
+	SIPR & 4.7477067 &  $<$ 1e-07\\
+	\hline
+	LL & 1.9551795 & 0.0164608\\
+	\hline
+	MC & 0.9485545 & 0.0000006\\
+	\hline
+	SB & 0.2118519 & 0.0240918\\
+	\hline
+\end{tabular}
+\caption{4th Order Model, $\rr$ Estimates and Standard Errors}\label{tab:other-res2} 
+\end{table}
+
+Table \ref{tab:other-res2} shows the methods applied to data from a 4th order linear regression model with respect to time.
 Obviously, there is no true $\rr$ to compare to.
@@
 
+\begin{table}[H]
+
+	\centering
+	\begin{tabular}[t]{l|r|r}
+		\hline
+		Model & Estimate & Std. Err\\
+		\hline
+		LS & 1.2742577 & 0.0979741\\
+		\hline
+		ReLS & 1.2743142 & 0.0339775\\
+		\hline
+		SIRLMA & 1.2191008 & 0.1092962\\
+		\hline
+		LMAT & 1.2009373 & 0.0031099\\
+		\hline
+		IPR & 2.5280045 & 2.5904017\\
+		\hline
+		SIPR & 2.1567978 & 0.3585219\\
+		\hline
+		LL & 1.2585872 & 0.0010079\\
+		\hline
+		MC & 0.9485555 & $<$ 1e-07 \\
+		\hline
+		SB & 1.2008765 & 0.0573592\\
+		\hline
+	\end{tabular}
+	\caption{Linear SIR Data, $\rr$ Estimates and Standard Errors}\label{tab:other-res3}
+\end{table}
+
+\section{Variance Estimation Methods}
+We examine the variance methods used to estimate CIs for our estimates of $\rr$.
+From the above results, we saw that methods such as LS, ReLS, LL, and MC result in very small CIs.
+On the other hand, methods such as IPR have consistently large CIs, to the point of being uninformative.
+The other methods CI width seems to depend heavily on the initial parameters used.
+Unsurprisingly, an $\sigma_X$, $\sigma_Y$, $T$, or $N$ seems to be positively associated with a larger CI.
+Having a larger percentage of initially infected individuals also seems to be associated with a larger CI.
+We also see that having a very large $\rr$ value seems to have less effect on creating a large CI than does having a small value of $\rr$.
+Perhaps surprisingly, the width of the CI seems to be independent of the error structure.
+
+We estimate the variance through the delta method, block bootstrap, and posterior and compare it to the empirical variance of $\rr$ for each of the methods, using 1000 data sets generated from the baseline data structure.
+
 \subsection{Repeated Sampling}
 
-\begin{table}[H]
-
-	\caption{\label{tab:}Mean $\rr$ Estimates and Std. Devs, Repeated Sampling}
+Table \ref{tab:rep-samp} shows the mean and standard deviation of $\rr$ estimates produced by the models under 100 repetitions of the baseline conditions, with new noise generated each time.
+We want to compare the standard deviations of the estimates to the standard errors returned by each of the models for a single data set to asses how reliable those standard errors are.
+Overall, it appears than most of the models have standard error estimates that are larger than the actual standard deviation of the estimates under repeated sampling.
+In some cases, such as LL, LMA, and LMAT, the standard deviation and standard error estimates are fairly close.
+The least squares methods, LS and ReLS, on the other hand, tend to overestimate the error by an order of magnitude.
+Finally, SIPR and SB, have individual standard error estimates that are smaller than the standard deviation of repeated estimates, meaning that they understate how much error is in the estimates.
+
+\begin{table}[H]
+
 	\centering
@@
 		\hline
-		SIRLS & 2.0000175 & 0.0002734 & 1.9999190 & 0.0058322\\
-		\hline
-		SIRREPAR & 2.0000289 & 0.0003092 & 1.9997083 & 0.0052113\\
+		LS & 2.0000175 & 0.0002734 & 1.9999190 & 0.0058322\\
+		\hline
+		ReLS & 2.0000289 & 0.0003092 & 1.9997083 & 0.0052113\\
 		\hline
@@
 		\hline
-		SIRLMAT & 1.4562107 & 0.3576736 & 1.398016 & 0.6028170 \\
-		\hline
-		SIRIPR & 4.5214294 & 0.3705241 & 2.592974 & 9.040841\\
-		\hline
-		SIRSIPR & 1.7972938 & 0.0951042 & 0.9251762 & $<$ 1e-07 \\
-		\hline
-		LOGLINEAR & 2.0000079 & 0.0003191 & 1.9998403 & 0.0002227\\
-		\hline
-		SIRMC & 0.9485555 & $<$ 1e-07 & 0.9485555 & $<$ 1e-07\\
-		\hline
-		SEQBAYES & 1.1330264 & 0.4163031 & 1.0000000 & 0.0523424\\
-		\hline
-	\end{tabular}
-\end{table}
-
-The table above shows the mean and standard deviation of $/rr$ estimates produced by the models under 100 repetitions of the baseline conditions, with new noise generated each time.
-We want to compare the standard deviations of the estimates to the standard errors returned by each of the models for a single data set to asses how reliable those standard errors are.
-Overall, it appears than most of the models have standard error estimates that are smaller than the actual standard deviation of the estimates under repeated sampling.
-In some cases, such as the log linear method and the linear model approximations, are fairly close.
-The least squares methods, on the other hand, tend to overestimate the error by an order of magnitude.
-The SIPR and sequential Bayes model, on the other hand, have individual standard error estimates that are smaller than the standard deviation of repeated estimates, meaning that they understate how much error is in the estimates.
+		LMAT & 1.4562107 & 0.3576736 & 1.398016 & 0.6028170 \\
+		\hline
+		IPR & 4.5214294 & 0.3705241 & 2.592974 & 9.040841\\
+		\hline
+		SIPR & 1.7972938 & 0.0951042 & 0.9251762 & $<$ 1e-07 \\
+		\hline
+		LL & 2.0000079 & 0.0003191 & 1.9998403 & 0.0002227\\
+		\hline
+		MC & 0.9485555 & $<$ 1e-07 & 0.9485555 & $<$ 1e-07\\
+		\hline
+		SB & 1.1330264 & 0.4163031 & 1.0000000 & 0.0523424\\
+		\hline
+	\end{tabular}
+        \caption{Mean $\rr$ Estimates and Std. Devs, Repeated Sampling}
+        \label{tab:rep-samp}
+\end{table}
+
+\subsection{Block Bootstrap}
+
+Table \ref{tab:bb-samp} shows the mean and standard deviation of $\rr$ estimates produced by the models under 100 block bootstrap repetitions of the baseline data.
+The data were first detrended using the predictions of the compartment sizes at each time point from each of the models on the original data set.
+The block bootstrap algorithm was then applied on to the residuals from the predictions using a fixed block length.
+Only 5 of the models provide predictions of the size of the compartments and thus allow for detrending of the data.
+The least squares and reparameterized model have a fairly accurate estimates of $\rr$ as well as very small standard errors.
+On the other hand, the linear approximation and smoothed IPR models both have very inaccurate estimates as well as very large standard errors of $\rr$.
+
+\begin{table}[H]
+
+	\centering
+	\begin{tabular}[t]{l|r|r|r|r}
+		\hline
+		Model & Mean Estimate & Std. Dev. & Baseline Estimate & Baseline SE\\
+		\hline
+		LS & 1.999903 & 0.0003271 & 1.9999190 & 0.0058322\\
+		\hline
+		ReLS & 1.999802 & 0.0002793 & 1.9997083 & 0.0052113\\
+		\hline
+		SIRLMA & -0.9975459 & 21.71612 & 1.997416 & 0.3900006\\
+		\hline
+		LMAT & 1.513824 & 1.048486 & 1.398016 & 0.6028170 \\
+		\hline
+		SIPR & 0.7761622 & 2.662394 & 0.9251762 & $<$ 1e-07 \\
+		\hline
+	\end{tabular}
+	\caption{Mean $\rr$ Estimates and Std. Devs, Block Bootstrap}
+	\label{tab:bb-samp}
+\end{table}
+
+Figure \ref{fig:coverage} shows the proportion of repetitions for both repeated sampling and block bootstrapping in which the 95\% confidence intervals of the estimates produced by the models covers the true $\rr$.
+The least squares, reparameterized, and linear model approximation with all time points contain the true $\rr$ value in all of the confidence intervals.
+On the other hand, the smoothed IPR, Markov chain, and sequential Bayes models never cover the true $\rr$ in any of the confidence intervals.
+
+\begin{figure}[H]
+	\begin{center}
+		\includegraphics[scale=0.5]{images/coverage.jpeg}
+		\caption{Proportion of coverage of true $\rr$ of confidence intervals produced by repeated sampling and block bootstrapping.}
+		\label{fig:coverage}	
+	\end{center}
+\end{figure}
 
 \section{Discussion}\label{sec:discussion}
+
+%% say what we said
+%% say it
+%% say it again
+
+%% what is r0 and why it is important
+In this paper, we examine $\rr$, the average number of secondary infections a primary individual generates when introduced into a completely susceptible population, in the context of the SIR model.
+The quantity $\rr$ summarizes an epidemic into a single estimate in the hopes that infectious disease outbreaks may be compared to one another.
+A value of $\rr > 1$ indicates an outbreak of a disease whereas a value of $\rr < 1$ indicates that no outbreak will occur.
+Thus, it is critical to not only estimate the value of $\rr$ but also a corresponding CI.
+Overestimating $\rr > 1$ during an ongoing epidemic may result in a waste of resources in prevention and control measures; underestimating $\rr < 1$ may result in a worldwide crisis.
+
+In Table \ref{tab:r0-real-ex}, we provide estimates for $\rr$ that have appeared over the past three decades for HIV, Zika, Ebola, influenza, H1N1 influenza, and measles in different locations throughout the world.
+Each of these estimates is greater than $1$, which makes sense as we tend to care more when a disease has the potential outbreak.
+The largest $\rr$ point estimate we have seen is for measles at $\rr=17$ in England in the 20th century, and the smallest is $\rr=1.08$ for HIV in Portugal in the 21st century.
+%% r0 as a property of the model
+The quantity $\rr$ may summarize an epidemic, accounting for temperature, seasonality, location, and genetics of a disease, but it is still a property of the model.
+Namely, how individuals progress through an epidemic (e.g.
+susceptible to infectious to recovered compartments) influences how we view $\rr$.
+We thus should compare estimates for $\rr$ for a fixed epidemic model.
+Here, we limit our analysis to estimates of $\rr$ derived from the SIR model.
+Then $\rr = \frac{\beta}{\gamma}$ where $\beta$ is the rate of infection and $\gamma$ is the rate of recovery.
+As a consequence, we need to estimate $\rr$ from not only the number of infections but also the number of individuals recovering at each time step.
+
+%% the methods we chose and why we chose them
+Additionally, the SIR model assumes monotoncity of S and R compartments which can make estimating $\rr$ more difficult, especially in terms of the CIs since the usual assumption of Gaussian noise may be violated due to the non-symmetric nature of the noise.
+We examine \wxxsir methods to estimate $\rr$ which include least squares optimization, polynomial approximations, estimates from incidence and variance, a log-linear model, a Markov chain method, and a simple Bayesian method.
+Although, this is not an exhaustive list of methods used to estimate $\rr$ from the SIR model, we believe it covers common and popular methods used throughout the decades.
+We overview the details of methods in Section \ref{sec:methods}.
+
+%% the importance of variance and testing
+Each of the \wxxsir methods results in an estimate of $\rr$, $\hat{\rr}$.
+We also estimate a 95\% CI for the  $\rr$ estimate.
+Although, many researchers report intervals via sensitivity analysis (how the estimate changes with small variations in the underlying parameters), the ranges of these parameters often do not incorporate any distributional assumptions about these estimated ranges and hence do not correspond to 95\% CIs.
+We provide three statistical methods to estimate the standard error of $\hat{\rr}$ and use the interval $\left [\hat{\rr} \pm 2\hat{SE}\left (\hat{\rr} \right ) \right ]$ as our 95\% CI.
+We hope that these intervals will provide guidance on how to create more meaningful intervals when estimating $\rr$ and discerning which diseases are more severe than others, and whether the true $\rr >1$.
+
+We compare the the estimates of $\rr$ for the \wxxsir methods using a series of simulations.
+We generate different simulated data sets under different noise assumptions and estimate $\rr$ using each of these data sets.
+We then analyze estimates of $\rr$ while varying individual parameters of the model including the infection and recovery parameters, the total amount of data, the initial percent of infectious and susceptible individuals, the standard deviation of the noise used to generate the simulated data, and the total number of individuals in a population.
+We also analyze the model under misspecification, that is when a model other than the SIR model is used to generate the simulated data sets.
+Finally, we compare our variance methods to one another and analyze the coverage of our methods.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%RESULTS
+%%%%%%%%%%%%%%%%%
+
+Overall, we recommend based on our analysis:
+\begin{itemize}
+\item processes (e.g. Gaussian error, AR error, and PAVA)
+  \item LS and ReLS provide the most accurate estimates of $\rr$
+  \item LL results in close point estimates but occasionally has too small standard errors
+  \item The specific MC and SB methods presented here do not seem to fit our generated data well
+  \item There is not much difference in estimations of $\rr$ due to different error generating processes
+  \item Very small values of $\gamma$ can result in very poor estimates of $\rr$
+  \item 100 data points seems to be a reasonable number of data points to have enough power for testing that $\rr > 1$ and 50 seems to be too few
+  \item The larger the number of initially infectious individuals is, the smaller the CIs for $\rr$ are
+  \item The noisier the epidemic data is, the larger our CIs for $\rr$ are
+  \item Estimations of $\rr$ are fairly robust to the misspecified data sets we presented, at least for the methods of LS, ReLS, and LL
+  \item Block bootstrap matches the empirical variance through repeated sampling.
+  \end{itemize}
+
+$\rr$ is a difficult quantity to estimate, but we hope that our methods of estimation and simulations provide some guidance for researchers on how to generate statistically reliable estimates for $\rr$ and its CIs.
+Our analysis is limited in scope to the SIR model, and in the future we would like to see how methods of estimating $\rr$ compare for models such as the SEIR and beyond.
+
+%% results of simulations:
+%% - data
+%% - what parameters make a difference
+%% - what methods perform best when
+%% - misspecification
+%% variance
+%% beyond the sir model
 
@@
 We also see a large difference between the $\rr$ estimates and standard errors on the exact same data sets as well, even though all derived from the same model.
-Thus, when estimating $/rr$ based on an SIR or any other kind of compartment model, we need to take careful consideration.
+Thus, when estimating $\rr$ based on an SIR or any other kind of compartment model, we need to take careful consideration.
 
@@
 
+\subsection{Exponential Growth (EG)}\label{sec:expgrowth}
+%% nishiura, majumder, IDEA model -- sum of squares (santillana2016)
+%% An IDEA for short term outbreak projection: nearcasting using the basic reproduction number.  (fisman2013, fisman2014)
+%% chowell 2008 - parametric bootstrap for a model they made up, but sort of empirical.
+Many researchers employ an assumption of exponential growth of the number infectious through times $t=0, 1, \dots, T^* <T$ \citep{wallinga2007generation,fisman2014,nishiura2016,majumder2016,towers2016}.
+\cite{wallinga2007generation} report that the effective reproduction number $\mathcal{R}_t$ and hence the initial reproduction number $\rr$ may derived using the fact that infection ``counts increase exponentially in the initial phase of an epidemic.''  We then have to estimate $r$, the \textit{per capita} change in the number of new cases per unit of time and $\omega$ the serial interval, the distribution of time between a primary and secondary infection.
+Then, we have
+\begin{align}\label{eq:lotka}
+\rr = \exp{(r \omega)}.
+\end{align}
+Equation \eqref{eq:lotka} is derived from a demographic view using the Lotka-Euler survival equations which come from the fields of demography, ecology, and evolutionary biology.
+
+\cite{wallinga2007generation}, instead, use a moment generating function expression for $\rr$.  With $\omega(t)$ as the serial interval, then
+\begin{align*}
+\rr^{-1} &= \frac{1}{M(-r)} = \int_{a=0}^\infty e^{-rt}\omega(t)dt.
+\end{align*}
+\cite{wallinga2007generation} specifically derive an estimate for $\rr$ for the SIR model.
+They assume that the rate of leaving the infectious stsate, $\gamma$ is constant as is the rate of making contacts during the infectious stage.
+Then, the duration of a serial interval is an exponential distribution with mean $\gamma^{-1}$, the moment generating function is then,
+\begin{align*}
+  M(-r) = \frac{\gamma}{\gamma + r},
+\end{align*}
+and the resulting estimate of $\rr$ is
+\begin{align}\label{eq:r0-eg}
+  \hat{\rr} = 1 + \frac{\hat{r}}{\hat{\gamma}},
+\end{align}
+where $\hat{r}$ and $\hat{\gamma}$ are estimates of the $r$ and $\gamma$, respectively.
+The parameter $\hat{\gamma}$ is often assumed to be fixed or tried with a sequence of serial intervals \citep{majumder2016}.
+We may estimate $\hat{r}$ using ordinary linear regression, namely,
+\begin{align*}
+  \log \left  (\frac{Y(t)}{Y(0)} \right )&= rt + \epsilon_t\\
+  \epsilon_t &\sim N(0, \sigma^2),
+\end{align*}
+which results in
+\begin{align}\label{eq:exp-growth}
+  \hat{r} &= \frac{\sum_{t=0}^{T^*}\left  (\frac{Y(t)}{Y(0)} \right )}{\sum_{t=0}^{T^*}t}.
+\end{align}
+
+For the above estimate of $\rr$, we observe the duration of serial intervals in a period of exponential growth of the number of infectious individuals.
+Deciding when exponential growth occurs and which data points to use to estimate $\rr$ may be difficult.
+We can use sensitivity analysis to see how the estimate $\rr$ changes as we increase or decrease $T^{*}$, the number of time points used.
+The reader may notice some similarities to the estimate for $\rr$ in Equation \eqref{eq:r0-eg} and from the log-linear model in Equation \eqref{eq:harko_lin}.
+This method is a simplification of the SIR model as the $R$ compartment is not used to estimate $r$, only the recovery rate $\gamma$.
+
+An advantage to this method are its (and similar adaptations) prevalence in the literature, which allows one to to compare estimates $\rr$ directly to one another.
+Another advantage is its flexibility.
+Here, we assumed the serial interval has an exponential distribution, but this is not a requirement and other distributions may be used (see \cite{wallinga2007generation}).
+However, a disadvantage to this method is that is a simplification of the SIR model and we have to decide how many time points to use to estimate the exponential growth rate of infectious.
+Another disadvantage to this method is that one also has to decide how to treat the additional paramter $\gamma$, whether to treat it as fixed, a range of fixed values, or even having a probability distribution.
+
+\subsection{Incidence Decay and Exponential Adjustment (IDEA)}\label{sec:idea}
+The Incidence Decay and Exponential Adjustment (IDEA) model is a way to account for a declining effective reproduction number of time due to both depletion of susceptible individuals and spontaneous and planned control activities and behaviors'' \citep{fisman2013}.
+They model the incidence counts during the $t$th interval, $J(t),$ as
+\begin{align*}
+  J(t) &= \left [ \frac{\rr}{(1 + d)^t}\right ]^t,
+\end{align*}
+where $d$ is a dampening or discount factor.  As $d$ increases, the incidence at time $t$  decreases. The estimate for $\rr$ is obtained by minimizing the sum of squares over $\rr$ and $d$,
+\begin{align*}
+  \hat{\rr} &=  \arg \min_{\rr,d} \sum_{t=0}^{T^*}\left (J_{obs}(t) - J(t) \right)^2.
+\end{align*}
+The method presented in \cite{fisman2013} is a mathematical, rather than statistical model.
+Sensitivity analysis is conducted with respect to $T^*$, $d$, and $\rr$, and lower and upper bounds are reported as a result of this analysis.
+
+Advantages of this method include its usefulness in estimating $\rr$ \citep{fisman2014,majumder2016}.
+Additionally, this model may be employed to address the effect of control measures and better estimate the final size of an epidemic.
+A disadvantage to this model is that its upper and lower bounds do not correspond to CIs.
+Also, in this method we must select the interval size of the incidence counts.
+
 \section{Fitting the SIR Model}
@@
 
-	\caption{\label{tab:}$\rr$ Estimates and Std. Errs, SIRLS Model,
+	\caption{\label{tab:}$\rr$ Estimates and Std. Errs, LS Model,
 		$X_0 = 99950, Y_0 = 50$, $\sigma_X = 100, \sigma_Y = 5$, $\beta = 0.06$}
@@
 
-	\caption{\label{tab:}$\rr$ Estimates and Std. Errs, SIRLS Model,
+	\caption{\label{tab:}$\rr$ Estimates and Std. Errs, LS Model,
 		$\sigma_X = 100, \sigma_Y = 5$, $\beta = 0.06, \gamma = 0.03$}
@@
 \begin{table}[H]
-	\caption{\label{tab:}$\rr$ Estimates and Std. Errs, SIRLS Model,
+	\caption{\label{tab:}$\rr$ Estimates and Std. Errs, LS Model,
 		$X_0 = 99950, Y_0 = 50$, $\beta = 0.06, \gamma = 0.03$}
@@
 
-	\caption{\label{tab:}$\rr$ Estimates and Std. Errs, SIRLS Model,
+	\caption{\label{tab:}$\rr$ Estimates and Std. Errs, LS Model,
 		$X_0 = 99950, Y_0 = 50$, $\sigma_X = 100, \sigma_Y = 5$}
@@
 
-	\caption{\label{tab:}$\rr$ Estimates and Std. Errs, SIRLS Model,
+	\caption{\label{tab:}$\rr$ Estimates and Std. Errs, LS Model,
 		$\beta = 0.06, \gamma = 0.03$, $X_0 = 99950, Y_0 = 50$, $\sigma_X = 100, \sigma_Y = 5$}
@@
 
+\begin{table}[H]
+
+	\caption{\label{tab:}$\rr$ Estimates and Std. Errs, LS Model,
+		$\beta = 0.06, \gamma = 0.03$, $X_0 = 99950, Y_0 = 50$, $\sigma_X = 100, \sigma_Y = 5$}
+	\centering
+	\begin{tabular}[t]{l|r|r|r|r|r|r|r|r}
+		\hline
+		Data Set & Norm Est & Norm SE & Norm-M Est & Norm-M SE & AR Est & AR SE & AR-M Est & AR-M SE\\
+		\hline
+		$X_0 = 99, Y_0 = 1$ & 1.986253 & 4.8619876 & 2.038011 & 0.4537907 & 2.106225 & 7.1198939 & 2.105517 & 1.5447687\\
+		\hline
+		$X_0 = 990, Y_0 = 10$ & 1.997222 & 0.5406162 & 2.017454 & 0.1429851 & 2.122109 & 0.7362701 & 2.097491 & 0.2921481\\
+		\hline
+		$X_0 = 9995, Y_0 = 5$ & 2.000774 & 0.0590423 & 1.995300 & 0.0239310 & 2.000782 & 0.0744317 & 1.996394 & 0.0489651\\
+		\hline
+		$X_0 = 999500, Y_0 = 500$ & 2.000050 & 0.0005467 & 1.999834 & 0.0005748 & 1.999833 & 0.0008762 & 1.999959 & 0.0007396\\
+		\hline
+	\end{tabular}
+\end{table}
+
 \subsection{SIR Reparameterization}
@@
 
-	\caption{\label{tab:}$\rr$ Estimates and Std. Errs, SIRREPAR Model,
+	\caption{\label{tab:}$\rr$ Estimates and Std. Errs, ReLS Model,
 		$X_0 = 99950, Y_0 = 50$, $\sigma_X = 100, \sigma_Y = 5$, $\beta = 0.06$}
@@
 
-	\caption{\label{tab:}$\rr$ Estimates and Std. Errs, SIRREPAR Model,
+	\caption{\label{tab:}$\rr$ Estimates and Std. Errs, ReLS Model,
 		$\sigma_X = 100, \sigma_Y = 5$, $\beta = 0.06, \gamma = 0.03$}
@@
 
-	\caption{\label{tab:}$\rr$ Estimates and Std. Errs, SIRREPAR Model,
+	\caption{\label{tab:}$\rr$ Estimates and Std. Errs, ReLS Model,
 		$X_0 = 99950, Y_0 = 50$, $\beta = 0.06, \gamma = 0.03$}
@@
 
-	\caption{\label{tab:}$\rr$ Estimates and Std. Errs, SIRREPAR Model,
+	\caption{\label{tab:}$\rr$ Estimates and Std. Errs, ReLS Model,
 		$X_0 = 99950, Y_0 = 50$, $\sigma_X = 100, \sigma_Y = 5$}
@@
 
-	\caption{\label{tab:}$\rr$ Estimates and Std. Errs, SIRREPAR Model,
+	\caption{\label{tab:}$\rr$ Estimates and Std. Errs, ReLS Model,
 		$\beta = 0.06, \gamma = 0.03$, $X_0 = 99950, Y_0 = 50$, $\sigma_X = 100, \sigma_Y = 5$}
@@
 		T = 20 & 22.964587 & 6.9816713 & 5.044869 & 2.1398099 & NA & NA & 1.681269 & 0.6543603\\
+		\hline
+	\end{tabular}
+\end{table}
+
+\begin{table}[H]
+
+	\caption{\label{tab:}$\rr$ Estimates and Std. Errs, ReLS Model,
+		$\beta = 0.06, \gamma = 0.03$, $X_0 = 99950, Y_0 = 50$, $\sigma_X = 100, \sigma_Y = 5$}
+	\centering
+	\begin{tabular}[t]{l|r|r|r|r|r|r|r|r}
+		\hline
+		Data Set & Norm Est & Norm SE & Norm-M Est & Norm-M SE & AR Est & AR SE & AR-M Est & AR-M SE\\
+		\hline
+		$X_0 = 99, Y_0 = 1$ & 1.987105 & 3.4568905 & 2.037779 & 0.3348238 & 1.658070 & 2.0918740 & 2.105172 & 1.1781993\\
+		\hline
+		$X_0 = 990, Y_0 = 10$ & 1.997065 & 0.3756188 & 2.017454 & 0.1020655 & 2.121902 & 0.5659528 & 2.097579 & 0.2210569\\
+		\hline
+		$X_0 = 9995, Y_0 = 5$ & 2.000519 & 0.0528419 & 1.995631 & 0.0212989 & 2.000955 & 0.0665574 & 1.996249 & 0.0434920\\
+		\hline
+		$X_0 = 999500, Y_0 = 500$ & 1.999946 & 0.0004956 & 1.999946 & 0.0005270 & 2.000104 & 0.0007140 & 1.999734 & 0.0006529\\
 		\hline
@@
 
+\begin{table}[H]
+
+	\caption{\label{tab:}$\rr$ Estimates and Std. Errs, SIRLMA Model,
+		$\beta = 0.06, \gamma = 0.03$, $\sigma_X = 100, \sigma_Y = 5$}
+	\centering
+	\begin{tabular}[t]{l|r|r|r|r|r|r|r|r}
+		\hline
+		Data Set & Norm Est & Norm SE & Norm-M Est & Norm-M SE & AR Est & AR SE & AR-M Est & AR-M SE\\
+		\hline
+		$X_0 = 99, Y_0 = 1$ & 0.8023469 & 18.9416598 & 0.8332037 & 0.6714657 & 1.0679323 & 0.0510976 & 1.038611 & 1.5641167\\
+		\hline
+		$X_0 = 990, Y_0 = 10$ & 1.0682053 & 0.0327648 & 0.8256301 & 1.4444021 & 1.2044940 & 0.3878046 & 1.172003 & 0.0511127\\
+		\hline
+		$X_0 = 9995, Y_0 = 5$ & 1.9976309 & 2606.0798123 & 1.3632453 & 0.3709415 & 0.7923714 & 2.5234696 & 1.398184 & 0.2967151\\
+		\hline
+		$X_0 = 999500, Y_0 = 500$ & 1.7316488 & 0.0919930 & 1.7673457 & 0.0976337 & 1.6873477 & 0.0810172 & 1.833159 & 0.1199462\\
+		\hline
+	\end{tabular}
+\end{table}
+
 \subsection{SIR Linear Model Approximation, All Time Points}
@@
 
-	\caption{\label{tab:}$\rr$ Estimates and Std. Errs, SIRLMAT Model,
+	\caption{\label{tab:}$\rr$ Estimates and Std. Errs, LMAT Model,
 		$X_0 = 99950, Y_0 = 50$, $\sigma_X = 100, \sigma_Y = 5$, $\beta = 0.06$}
@@
 
-	\caption{\label{tab:}$\rr$ Estimates and Std. Errs, SIRLMAT Model,
+	\caption{\label{tab:}$\rr$ Estimates and Std. Errs, LMAT Model,
 		$\sigma_X = 100, \sigma_Y = 5$, $\beta = 0.06, \gamma = 0.03$}
@@
 
-	\caption{\label{tab:}$\rr$ Estimates and Std. Errs, SIRLMAT Model,
+	\caption{\label{tab:}$\rr$ Estimates and Std. Errs, LMAT Model,
 		$X_0 = 99950, Y_0 = 50$, $\beta = 0.06, \gamma = 0.03$}
@@
 
-	\caption{\label{tab:}$\rr$ Estimates and Std. Errs, SIRLMAT Model,
+	\caption{\label{tab:}$\rr$ Estimates and Std. Errs, LMAT Model,
 		$X_0 = 99950, Y_0 = 50$, $\sigma_X = 100, \sigma_Y = 5$}
@@
 
-	\caption{\label{tab:}$\rr$ Estimates and Std. Errs, SIRLMAT Model,
+	\caption{\label{tab:}$\rr$ Estimates and Std. Errs, LMAT Model,
 		$\beta = 0.06, \gamma = 0.03$, $X_0 = 99950, Y_0 = 50$, $\sigma_X = 100, \sigma_Y = 5$}
@@
 
+\begin{table}[H]
+
+	\caption{\label{tab:}$\rr$ Estimates and Std. Errs, LMAT Model,
+		$\beta = 0.06, \gamma = 0.03$, $\sigma_X = 100, \sigma_Y = 5$}
+	\centering
+	\begin{tabular}[t]{l|r|r|r|r|r|r|r|r}
+		\hline
+		Data Set & Norm Est & Norm SE & Norm-M Est & Norm-M SE & AR Est & AR SE & AR-M Est & AR-M SE\\
+		\hline
+		$X_0 = 99, Y_0 = 1$ & 1.220874 & 4.7699293 & -0.0614019 & 15.5900050 & -0.0217592 & 18.7796434 & 0.9203832 & 1.5845866\\
+		\hline
+		$X_0 = 990, Y_0 = 10$ & 2.248607 & 29.7561744 & 1.0457445 & 0.6365320 & 1.0880985 & 3.9238850 & -13.6905770 & 281.9114385\\
+		\hline
+		$X_0 = 9995, Y_0 = 5$ & 1.294414 & 1.4304706 & 1.3710979 & 0.6597925 & 1.5307307 & 1.0343598 & 1.3865816 & 1.1453533\\
+		\hline
+		$X_0 = 999500, Y_0 = 500$ & 1.370406 & 0.8821226 & 1.9731686 & 10.8871603 & 1.3885429 & 0.6854633 & 1.4164038 & 0.6347757\\
+		\hline
+	\end{tabular}
+\end{table}
+
 \subsection{Incidence to Prevalence Ratio}
@@
 
-	\caption{\label{tab:}$\rr$ Estimates and Std. Errs, SIRIPR Model,
+	\caption{\label{tab:}$\rr$ Estimates and Std. Errs, IPR Model,
 		$X_0 = 99950, Y_0 = 50$, $\sigma_X = 100, \sigma_Y = 5$, $\beta = 0.06$}
@@
 
-	\caption{\label{tab:}$\rr$ Estimates and Std. Errs, SIRIPR Model,
+	\caption{\label{tab:}$\rr$ Estimates and Std. Errs, IPR Model,
 		$\sigma_X = 100, \sigma_Y = 5$, $\beta = 0.06, \gamma = 0.03$}
@@
 
-	\caption{\label{tab:}$\rr$ Estimates and Std. Errs, SIRIPR Model,
+	\caption{\label{tab:}$\rr$ Estimates and Std. Errs, IPR Model,
 		$X_0 = 99950, Y_0 = 50$, $\beta = 0.06, \gamma = 0.03$}
@@
 
-	\caption{\label{tab:}$\rr$ Estimates and Std. Errs, SIRIPR Model,
+	\caption{\label{tab:}$\rr$ Estimates and Std. Errs, IPR Model,
 		$X_0 = 99950, Y_0 = 50$, $\sigma_X = 100, \sigma_Y = 5$}
@@
 
-	\caption{\label{tab:}$\rr$ Estimates and Std. Errs, SIRIPR Model,
+	\caption{\label{tab:}$\rr$ Estimates and Std. Errs, IPR Model,
 		$\beta = 0.06, \gamma = 0.03$, $X_0 = 99950, Y_0 = 50$, $\sigma_X = 100, \sigma_Y = 5$}
@@
 
+\begin{table}[H]
+
+	\caption{\label{tab:}$\rr$ Estimates and Std. Errs, IPR Model,
+		$\beta = 0.06, \gamma = 0.03$, $\sigma_X = 100, \sigma_Y = 5$}
+	\centering
+	\begin{tabular}[t]{l|r|r|r|r|r|r|r|r}
+		\hline
+		Data Set & Norm Est & Norm SE & Norm-M Est & Norm-M SE & AR Est & AR SE & AR-M Est & AR-M SE\\
+		\hline
+		$X_0 = 99, Y_0 = 1$ & 3620.867680 & 57608.167596 & 5.239294 & 63.654679 & 729.654107 & 2978.983072 & 15.800132 & 284.613763\\
+		\hline
+		$X_0 = 990, Y_0 = 10$ & 67.597632 & 157.628039 & 1.734334 & 12.726031 & 164.611097 & 1887.125297 & 1.707676 & 11.777508\\
+		\hline
+		$X_0 = 9995, Y_0 = 5$ & 40.807237 & 133.279831 & 1.797818 & 7.233937 & 976.308278 & 17227.979404 & 1.700549 & 9.887431\\
+		\hline
+		$X_0 = 999500, Y_0 = 500$ & 1.577592 & 1.422029 & 1.434415 & 1.131136 & 1.567228 & 1.594393 & 1.449019 & 1.180341\\
+		\hline
+	\end{tabular}
+\end{table}
+
 \subsection{Smoothed Incidence to Prevalence Ratio}
@@
 
-	\caption{\label{tab:}$\rr$ Estimates and Std. Errs, SIRSIPR Model,
+	\caption{\label{tab:}$\rr$ Estimates and Std. Errs, SIPR Model,
 		$X_0 = 99950, Y_0 = 50$, $\sigma_X = 100, \sigma_Y = 5$, $\beta = 0.06$}
@@
 
-	\caption{\label{tab:}$\rr$ Estimates and Std. Errs, SIRSIPR Model,
+	\caption{\label{tab:}$\rr$ Estimates and Std. Errs, SIPR Model,
 		$\sigma_X = 100, \sigma_Y = 5$, $\beta = 0.06, \gamma = 0.03$}
@@
 
-	\caption{\label{tab:}$\rr$ Estimates and Std. Errs, SIRSIPR Model,
+	\caption{\label{tab:}$\rr$ Estimates and Std. Errs, SIPR Model,
 		$X_0 = 99950, Y_0 = 50$, $\beta = 0.06, \gamma = 0.03$}
@@
 
-	\caption{\label{tab:}$\rr$ Estimates and Std. Errs, SIRSIPR Model,
+	\caption{\label{tab:}$\rr$ Estimates and Std. Errs, SIPR Model,
 		$X_0 = 99950, Y_0 = 50$, $\sigma_X = 100, \sigma_Y = 5$}
@@
 
-	\caption{\label{tab:}$\rr$ Estimates and Std. Errs, SIRSIPR Model,
+	\caption{\label{tab:}$\rr$ Estimates and Std. Errs, SIPR Model,
 		$\beta = 0.06, \gamma = 0.03$, $X_0 = 99950, Y_0 = 50$, $\sigma_X = 100, \sigma_Y = 5$}
@@
 
+\begin{table}[H]
+
+	\caption{\label{tab:}$\rr$ Estimates and Std. Errs, SIPR Model,
+		$\beta = 0.06, \gamma = 0.03$, $\sigma_X = 100, \sigma_Y = 5$}
+	\centering
+	\begin{tabular}[t]{l|r|r|r|r|r|r|r|r}
+		\hline
+		Data Set & Norm Est & Norm SE & Norm-M Est & Norm-M SE & AR Est & AR SE & AR-M Est & AR-M SE\\
+		\hline
+		$X_0 = 99, Y_0 = 1$ & 3.367825 & 22.41867 & 0.9850654 & 0.5665987 & 1.5591972 & 3959.42264 & 4.207729 & 13.26187\\
+		\hline
+		$X_0 = 990, Y_0 = 10$ & 1.912639 & 390.58481 & 1.6931741 & 12.0122249 & 0.5334608 & 13.65189 & 2.898150 & 284.49327\\
+		\hline
+		$X_0 = 9995, Y_0 = 5$ & 1.062891 & 0.00000 & 1.8635858 & 0.0000000 & 1.0938066 & 0.00000 & 1.572230 & 0.00000\\
+		\hline
+		$X_0 = 999500, Y_0 = 500$ & 1.792115 & 0.00000 & 1.7907398 & 0.0000000 & 1.7902743 & 0.00000 & 1.750503 & 0.00000\\
+		\hline
+	\end{tabular}
+\end{table}
+
 \subsection{SIR Log Linear}
@@
 
-	\caption{\label{tab:}$\rr$ Estimates and Std. Errs, LOGLINEAR Model,
+	\caption{\label{tab:}$\rr$ Estimates and Std. Errs, LL Model,
 		$X_0 = 99950, Y_0 = 50$, $\sigma_X = 100, \sigma_Y = 5$, $\beta = 0.06$}
@@
 
-	\caption{\label{tab:}$\rr$ Estimates and Std. Errs, LOGLINEAR Model,
+	\caption{\label{tab:}$\rr$ Estimates and Std. Errs, LL Model,
 		$\sigma_X = 100, \sigma_Y = 5$, $\beta = 0.06, \gamma = 0.03$}
@@
 
-	\caption{\label{tab:}$\rr$ Estimates and Std. Errs, LOGLINEAR Model,
+	\caption{\label{tab:}$\rr$ Estimates and Std. Errs, LL Model,
 		$X_0 = 99950, Y_0 = 50$, $\beta = 0.06, \gamma = 0.03$}
@@
 
-	\caption{\label{tab:}$\rr$ Estimates and Std. Errs, LOGLINEAR Model,
+	\caption{\label{tab:}$\rr$ Estimates and Std. Errs, LL Model,
 		$X_0 = 99950, Y_0 = 50$, $\sigma_X = 100, \sigma_Y = 5$}
@@
 
-	\caption{\label{tab:}$\rr$ Estimates and Std. Errs, LOGLINEAR Model,
+	\caption{\label{tab:}$\rr$ Estimates and Std. Errs, LL Model,
 		$\beta = 0.06, \gamma = 0.03$, $X_0 = 99950, Y_0 = 50$, $\sigma_X = 100, \sigma_Y = 5$}
@@
 
+\begin{table}[H]
+
+	\caption{\label{tab:}$\rr$ Estimates and Std. Errs, LL Model,
+		$\beta = 0.06, \gamma = 0.03$, $\sigma_X = 100, \sigma_Y = 5$}
+	\centering
+	\begin{tabular}[t]{l|r|r|r|r|r|r|r|r}
+		\hline
+		Data Set & Norm Est & Norm SE & Norm-M Est & Norm-M SE & AR Est & AR SE & AR-M Est & AR-M SE\\
+		\hline
+		$X_0 = 99, Y_0 = 1$ & 1.141317 & 0.0422218 & 2.031666 & 0.0113166 & 1.056918 & 0.0447463 & 1.771771 & 0.0561071\\
+		\hline
+		$X_0 = 990, Y_0 = 10$ & 2.255041 & 0.0379147 & 2.010183 & 0.0055668 & 2.470092 & 0.0545897 & 2.280617 & 0.0229552\\
+		\hline
+		$X_0 = 9995, Y_0 = 5$ & 1.998839 & 0.0023193 & 1.997778 & 0.0007420 & 2.004827 & 0.0027816 & 2.001058 & 0.0015548\\
+		\hline
+		$X_0 = 999500, Y_0 = 500$ & 1.999983 & 0.0000215 & 1.999989 & 0.0000203 & 1.999924 & 0.0000314 & 1.999829 & 0.0000294\\
+		\hline
+	\end{tabular}
+\end{table}
+
 \subsection{SIR Markov Chain}
@@
 
-	\caption{\label{tab:}$\rr$ Estimates and Std. Errs, SIRMC Model,
+	\caption{\label{tab:}$\rr$ Estimates and Std. Errs, MC Model,
 		$X_0 = 99950, Y_0 = 50$, $\sigma_X = 100, \sigma_Y = 5$, $\beta = 0.06$}
@@
 
-	\caption{\label{tab:}$\rr$ Estimates and Std. Errs, SIRMC Model,
+	\caption{\label{tab:}$\rr$ Estimates and Std. Errs, MC Model,
 		$\sigma_X = 100, \sigma_Y = 5$, $\beta = 0.06, \gamma = 0.03$}
@@
 
-	\caption{\label{tab:}$\rr$ Estimates and Std. Errs, SIRMC Model,
+	\caption{\label{tab:}$\rr$ Estimates and Std. Errs, MC Model,
 		$X_0 = 99950, Y_0 = 50$, $\beta = 0.06, \gamma = 0.03$}
@@
 
-	\caption{\label{tab:}$\rr$ Estimates and Std. Errs, SIRMC Model,
+	\caption{\label{tab:}$\rr$ Estimates and Std. Errs, MC Model,
 		$X_0 = 99950, Y_0 = 50$, $\sigma_X = 100, \sigma_Y = 5$}
@@
 
-	\caption{\label{tab:}$\rr$ Estimates and Std. Errs, SIRMC Model,
+	\caption{\label{tab:}$\rr$ Estimates and Std. Errs, MC Model,
 		$\beta = 0.06, \gamma = 0.03$, $X_0 = 99950, Y_0 = 50$, $\sigma_X = 100, \sigma_Y = 5$}
@@
 
+\begin{table}[H]
+
+	\caption{\label{tab:}$\rr$ Estimates and Std. Errs, MC Model,
+		$\beta = 0.06, \gamma = 0.03$, $\sigma_X = 100, \sigma_Y = 5$}
+	\centering
+	\begin{tabular}[t]{l|r|r|r|r|r|r|r|r}
+		\hline
+		Data Set & Norm Est & Norm SE & Norm-M Est & Norm-M SE & AR Est & AR SE & AR-M Est & AR-M SE\\
+		\hline
+		$X_0 = 99, Y_0 = 1$ & 0.8455334 & 0.0124837 & 0.9481149 & 0.0002788 & NaN & NaN & 0.9480033 & 0.0005127\\
+		\hline
+		$X_0 = 990, Y_0 = 10$ & 0.9432094 & 0.0014184 & 0.9485736 & 0.0000060 & 0.9500365 & 0.0003484 & 0.9486116 & 0.0000141\\
+		\hline
+		$X_0 = 9995, Y_0 = 5$ & 0.9485626 & 0.0000040 & 0.9485575 & 0.0000002 & 0.9492072 & 0.0002225 & 0.9485575 & 0.0000002\\
+		\hline
+		$X_0 = 999500, Y_0 = 500$ & 0.9485553 & 0.0000000 & 0.9485553 & 0.0000000 & 0.9485553 & 0.0000000 & 0.9485553 & 0.0000000\\
+		\hline
+	\end{tabular}
+\end{table}
+
 \subsection{SIR Sequential Bayes}
@@
 
-	\caption{\label{tab:}$\rr$ Estimates and Std. Errs, SEQBAYES Model,
+	\caption{\label{tab:}$\rr$ Estimates and Std. Errs, SB Model,
 		$X_0 = 99950, Y_0 = 50$, $\sigma_X = 100, \sigma_Y = 5$, $\beta = 0.06$}
@@
 
-	\caption{\label{tab:}$\rr$ Estimates and Std. Errs, SEQBAYES Model,
+	\caption{\label{tab:}$\rr$ Estimates and Std. Errs, SB Model,
 		$\sigma_X = 100, \sigma_Y = 5$, $\beta = 0.06, \gamma = 0.03$}
@@
 
-	\caption{\label{tab:}$\rr$ Estimates and Std. Errs, SEQBAYES Model,
+	\caption{\label{tab:}$\rr$ Estimates and Std. Errs, SB Model,
 		$X_0 = 99950, Y_0 = 50$, $\beta = 0.06, \gamma = 0.03$}
@@
 
-	\caption{\label{tab:}$\rr$ Estimates and Std. Errs, SEQBAYES Model,
+	\caption{\label{tab:}$\rr$ Estimates and Std. Errs, SB Model,
 		$X_0 = 99950, Y_0 = 50$, $\sigma_X = 100, \sigma_Y = 5$}
@@
 
-	\caption{\label{tab:}$\rr$ Estimates and Std. Errs, SEQBAYES Model,
+	\caption{\label{tab:}$\rr$ Estimates and Std. Errs, SB Model,
 		$\beta = 0.06, \gamma = 0.03$, $X_0 = 99950, Y_0 = 50$, $\sigma_X = 100, \sigma_Y = 5$}
@@
 \end{table}
+
+\begin{table}[H]
+
+	\caption{\label{tab:}$\rr$ Estimates and Std. Errs, SB Model,
+		$\beta = 0.06, \gamma = 0.03$, $\sigma_X = 100, \sigma_Y = 5$}
+	\centering
+	\begin{tabular}[t]{l|r|r|r|r|r|r|r|r}
+		\hline
+		Data Set & Norm Est & Norm SE & Norm-M Est & Norm-M SE & AR Est & AR SE & AR-M Est & AR-M SE\\
+		\hline
+		$X_0 = 99, Y_0 = 1$ & 0.3887115 & 0.0326338 & 1.574343 & 0.0656755 & 0.4241419 & 0.0340886 & 1.000000 & 0.0523424\\
+		\hline
+		$X_0 = 990, Y_0 = 10$ & 0.9028353 & 0.0497345 & 1.000000 & 0.0523424 & 1.5457192 & 0.0650757 & 1.000000 & 0.0523424\\
+		\hline
+		$X_0 = 9995, Y_0 = 5$ & 0.4418182 & 0.0347917 & 1.000000 & 0.0523424 & 1.0000000 & 0.0523424 & 1.452822 & 0.0630899\\
+		\hline
+		$X_0 = 999500, Y_0 = 500$ & 1.1774169 & 0.0567961 & 1.270930 & 0.0590085 & 1.1876846 & 0.0570432 & 1.807844 & 0.0703775\\
+		\hline
+	\end{tabular}
+\end{table}
+
+\section{Fitting the Exponential Model}
 
@@
 
-\subsection{Wallinga}
-
-\begin{table}[H]
-
-	\caption{\label{tab:}$\rr$ Estimates and Std. Errs, WALLINGA Model,
+\begin{table}[H]
+
+	\caption{\label{tab:}$\rr$ Estimates and Std. Errs, EXPGROWTH Model,
+		$\beta = 0.06, \gamma = 0.03$, $\sigma_X = 100, \sigma_Y = 5$}
+	\centering
+	\begin{tabular}[t]{l|r|r|r|r|r|r|r|r}
+		\hline
+		Data Set & Norm Est & Norm SE & Norm-M Est & Norm-M SE & AR Est & AR SE & AR-M Est & AR-M SE\\
+		\hline
+		$X_0 = 99, Y_0 = 1$ & 1.091297 & 0.0003149 & 0.9202789 & 0.0036275 & 1.083304 & 0.0003546 & 0.8941523 & 0.0050462\\
+		\hline
+		$X_0 = 990, Y_0 = 10$ & 1.094724 & 0.0002992 & 0.9728240 & 0.0017765 & 1.086413 & 0.0003386 & 0.9799270 & 0.0016069\\
+		\hline
+		$X_0 = 9995, Y_0 = 5$ & 1.102157 & 0.0002679 & 1.0678388 & 0.0004461 & 1.094024 & 0.0003024 & 1.0676423 & 0.0004474\\
+		\hline
+		$X_0 = 999500, Y_0 = 500$ & 1.220466 & 0.0000464 & 1.2204454 & 0.0000464 & 1.220457 & 0.0000464 & 1.2204480 & 0.0000464\\
+		\hline
+	\end{tabular}
+\end{table}
+
+\subsection{IDEA}
+
+\begin{table}[H]
+
+	\caption{\label{tab:}$\rr$ Estimates and Std. Errs, IDEA Model,
 		$\beta = 0.06, \gamma = 0.03$, $X_0 = 99950, Y_0 = 50$, $\sigma_X = 100, \sigma_Y = 5$}
@@
 
-	\caption{\label{tab:}$\rr$ Estimates and Std. Errs, WALLINGA Model,
+	\caption{\label{tab:}$\rr$ Estimates and Std. Errs, IDEA Model,
 		$\beta = 0.06, \gamma = 0.03$, $X_0 = 99950, Y_0 = 50$, $\sigma_X = 100, \sigma_Y = 5$}
@@
 
-	\caption{\label{tab:}$\rr$ Estimates and Std. Errs, WALLINGA Model,
+	\caption{\label{tab:}$\rr$ Estimates and Std. Errs, IDEA Model,
 		$\beta = 0.06, \gamma = 0.03$, $X_0 = 99950, Y_0 = 50$}
@@
 
-	\caption{\label{tab:}$\rr$ Estimates and Std. Errs, WALLINGA Model,
+	\caption{\label{tab:}$\rr$ Estimates and Std. Errs, IDEA Model,
 		$X_0 = 99950, Y_0 = 50$, $\sigma_X = 100, \sigma_Y = 5$}
@@
 		T = 20 & 19.536946 & 2.3356802 & 19.536830 & 2.3354926 & 19.537515 & 2.3353411 & 19.534388 & 2.3356299\\
+		\hline
+	\end{tabular}
+\end{table}
+
+\begin{table}[h]
+
+	\caption{\label{tab:}$\rr$ Estimates and Std. Errs, IDEA Model,
+		$\beta = 0.06, \gamma = 0.03$, $\sigma_X = 100, \sigma_Y = 5$}
+	\centering
+	\begin{tabular}[t]{l|r|r|r|r|r|r|r|r}
+		\hline
+		Data Set & Norm Est & Norm SE & Norm-M Est & Norm-M SE & AR Est & AR SE & AR-M Est & AR-M SE\\
+		\hline
+		$X_0 = 99, Y_0 = 1$ & 1.247573 & 0.0297242 & 1.454198 & 0.0194568 & 1.212651 & 0.0297805 & 1.411583 & 0.0210353\\
+		\hline
+		$X_0 = 990, Y_0 = 10$ & 1.440702 & 0.0207630 & 1.461154 & 0.0196739 & 1.369004 & 0.0237707 & 1.436235 & 0.0202001\\
+		\hline
+		$X_0 = 9995, Y_0 = 5$ & 1.940396 & 0.0329020 & 1.940602 & 0.0328856 & 1.939845 & 0.0328895 & 1.939680 & 0.0328886\\
+		\hline
+		$X_0 = 999500, Y_0 = 500$ & 1.940281 & 0.0328949 & 1.940280 & 0.0328949 & 1.940286 & 0.0328948 & 1.940299 & 0.0328944\\
 		\hline
@@
 \end{figure}
-
-\message{ !name(draft_v13.tex) !offset(-2294) }

### 2018-07-03 paper/Version13/watch.tex
finally on git

--- 2018-07-03 paper/Version13/_region_.tex
@@
 The difficulties of estimating $\rr$ are summarized and expanded upon by \cite{li2011}.
+\cite{driessche2017} overview a number of methods used to estimate $\rr$ but looks at systems beyond the SIR model and so these methods may not be directly compared.
 As a result of $\rr$ being a property of the model, we limit our review to methods for estimating $\rr$ for the SIR model introduced by \cite{Kermack700}.
@@
 We find that these methods also help to highlight the problems discussed above.
-In addition to making a point estimate of $\rr$, we are particularly interested in the hypothesis  $H_0: \rr > 1$, as a value greater than 1 indicates an outbreak of a disease.
+In addition to making a point estimate of $\rr$, we are particularly interested in the hypothesis  $H_0: \rr < 1$, as a value less than 1 indicates that a disease will not outbreak.
 As such, we discuss techniques for estimating CIs for each of the different methods.
@@
 
-These difficulties in estimating $\rr$ have led a number of researchers to analyze the sensitivity of $\rr$ \citep{lash2003,epstein2007agent,capaldi2012}.
+These difficulties in estimating $\rr$ have led a number of researchers to analyze the sensitivity of $\rr$ with respect to small perturbations in parameters like $\beta$ and $\gamma$ \citep{lash2003,epstein2007agent,capaldi2012}.
 Even this is a difficult task because we are looking at the rate of change of the compartments with respect to $\beta$ and $\gamma$.
@@
 \begin{align*}
-  H_0:\;& \rr > 1 \\
-  H_A:\;& \rr \le 1
-\end{align*}
-Then for some $\alpha \in [0,1]$, we would like to know for some $a, b >0$
-\begin{align*}
-P(a \le \hat{\rr} \le b) = 1 - \alpha,  
-\end{align*}
-the second property is particularly desirable, as we could find $a$ and $b$ from the distribution $F$.
-At any rate, we would like to be able to estimate $a$ and $b$ for a given $\alpha$-level.
+  H_0:\;& \rr < 1 \\
+  H_A:\;& \rr \ge 1
+\end{align*}
+Then for some $\alpha \in [0,1]$, we would like to know for some $\hat{a}, \hat{b} >0$
+\begin{align*}
+P(\hat{a} \le \rr \le \hat{b}) \le 1 - \alpha,  
+\end{align*}
+the second property is particularly desirable, as we could estimate $\hat{a}$ and $\hat{b}$ from the distribution $F$.
+At any rate, we would like to be able to estimate $\hat{a}$ and $\hat{b}$ for a given $\alpha$-level.
 We could also make similar hypothesis tests such as that the $\rr$ for H1N1 influenza is greater than the reported $\rr$ for seasonal influenza.
@@
 As a consequence, we may be able to make a case where we the resulting estimate of $\rr$ is unbiased and a consistent estimator.
-Finally,  this method is easily implemented using any linear modelling software.
+Finally,  this method is easily implemented using any linear modeling software.
 
@@
 Disadvantages include strict assumptions about the distributions and computational time required to estimate the relevant parameters.
-
-\subsection{Exponential Growth (EG)}\label{sec:expgrowth}
-%% nishiura, majumder, IDEA model -- sum of squares (santillana2016)
-%% An IDEA for short term outbreak projection: nearcasting using the basic reproduction number.  (fisman2013, fisman2014)
-%% chowell 2008 - parametric bootstrap for a model they made up, but sort of empirical.
-Many researchers employ an assumption of exponential growth of the number infectious through times $t=0, 1, \dots, T^* <T$ \citep{wallinga2007generation,fisman2014,nishiura2016,majumder2016,towers2016}.
-\cite{wallinga2007generation} report that the effective reproduction number $\mathcal{R}_t$ and hence the initial reproduction number $\rr$ may derived using the fact that infection ``counts increase exponentially in the initial phase of an epidemic.''  We then have to estimate $r$, the \textit{per capita} change in the number of new cases per unit of time and $\omega$ the serial interval, the distribution of time between a primary and secondary infection.
-Then, we have
-\begin{align}\label{eq:lotka}
-\rr = \exp{(r \omega)}.
-\end{align}
-Equation \eqref{eq:lotka} is derived from a demographic view using the Lotka-Euler survival equations which come from the fields of demography, ecology, and evolutionary biology.
-
-\cite{wallinga2007generation}, instead, use a moment generating function expression for $\rr$.  With $\omega(t)$ as the serial interval, then
-\begin{align*}
-\rr^{-1} &= \frac{1}{M(-r)} = \int_{a=0}^\infty e^{-rt}\omega(t)dt.
-\end{align*}
-\cite{wallinga2007generation} specifically derive an estimate for $\rr$ for the SIR model.
-They assume that the rate of leaving the infectious stsate, $\gamma$ is constant as is the rate of making contacts during the infectious stage.
-Then, the duration of a serial interval is an exponential distribution with mean $\gamma^{-1}$, the moment generating function is then,
-\begin{align*}
-  M(-r) = \frac{\gamma}{\gamma + r},
-\end{align*}
-and the resulting estimate of $\rr$ is
-\begin{align}\label{eq:r0-eg}
-  \hat{\rr} = 1 + \frac{\hat{r}}{\hat{\gamma}},
-\end{align}
-where $\hat{r}$ and $\hat{\gamma}$ are estimates of the $r$ and $\gamma$, respectively.
-The parameter $\hat{\gamma}$ is often assumed to be fixed or tried with a sequence of serial intervals \citep{majumder2016}.
-We may estimate $\hat{r}$ using ordinary linear regression, namely,
-\begin{align*}
-  \log \left  (\frac{Y(t)}{Y(0)} \right )&= rt + \epsilon_t\\
-  \epsilon_t &\sim N(0, \sigma^2),
-\end{align*}
-which results in
-\begin{align}\label{eq:exp-growth}
-  \hat{r} &= \frac{\sum_{t=0}^{T^*}\left  (\frac{Y(t)}{Y(0)} \right )}{\sum_{t=0}^{T^*}t}.
-\end{align}
-
-For the above estimate of $\rr$, we observe the duration of serial intervals in a period of exponential growth of the number of infectious individuals.
-Deciding when exponential growth occurs and which data points to use to estimate $\rr$ may be difficult.
-We can use sensitivity analysis to see how the estimate $\rr$ changes as we increase or decrease $T^{*}$, the number of time points used.
-The reader may notice some similarities to the estimate for $\rr$ in Equation \eqref{eq:r0-eg} and from the log-linear model in Equation \eqref{eq:harko_lin}.
-This method is a simplification of the SIR model as the $R$ compartment is not used to estimate $r$, only the recovery rate $\gamma$.
-
-An advantage to this method are its (and similar adaptations) prevalence in the literature, which allows one to to compare estimates $\rr$ directly to one another.
-Another advantage is its flexibility.
-Here, we assumed the serial interval has an exponential distribution, but this is not a requirement and other distributions may be used (see \cite{wallinga2007generation}).
-However, a disadvantage to this method is that is a simplification of the SIR model and we have to decide how many time points to use to estimate the exponential growth rate of infectious.
-Another disadvantage to this method is that one also has to decide how to treat the additional paramter $\gamma$, whether to treat it as fixed, a range of fixed values, or even having a probability distribution.
-
-\subsection{Incidence Decay and Exponential Adjustment (IDEA)}\label{sec:idea}
-The Incidence Decay and Exponential Adjustment (IDEA) model is a way to account for a declining effective reproduction number of time due to both depletion of susceptible individuals and spontaneous and planned control activities and behaviors'' \citep{fisman2013}.
-They model the incidence counts during the $t$th interval, $J(t),$ as
-\begin{align*}
-  J(t) &= \left [ \frac{\rr}{(1 + d)^t}\right ]^t,
-\end{align*}
-where $d$ is a dampening or discount factor.  As $d$ increases, the incidence at time $t$  decreases. The estimate for $\rr$ is obtained by minimizing the sum of squares over $\rr$ and $d$,
-\begin{align*}
-  \hat{\rr} &=  \arg \min_{\rr,d} \sum_{t=0}^{T^*}\left (J_{obs}(t) - J(t) \right)^2.
-\end{align*}
-The method presented in \cite{fisman2013} is a mathematical, rather than statistical model.
-Sensitivity analysis is conducted with respect to $T^*$, $d$, and $\rr$, and lower and upper bounds are reported as a result of this analysis.
-
-Advantages of this method include its usefulness in estimating $\rr$ \citep{fisman2014,majumder2016}.
-Additionally, this model may be employed to address the effect of control measures and better estimate the final size of an epidemic.
-A disadvantage to this model is that its upper and lower bounds do not correspond to CIs.
-Also, in this method we must select the interval size of the incidence counts.
 
@@
 
-This baseline data set
-
-% We then generate different SIR data sets, changing the one of the conditions for simulation each time and using the baseline SIR data as our basis for comparison.
-The different data sets generated are shown in the following table:
+      This baseline data set is chosen due to its moderate $\rr$ size of 2, which indicates an outbreak of an epidemic on par with that of HIV, ZIK-V, EVD, and H1N1 Influenza (see Table \ref{tab:r0-real-ex}).
+The infection rate and recovery rate ($\beta, \gamma$) are relatively small.
+The duration of the disease, under the true model, spans nearly a whole year (see Figure \ref{fig:baseline}) and so we use a year's worth of daily observations of (X,Y, Z) values.
+Initially, 50 (0.05\%) diseased individuals are introduced to the population of 100,000 individuals (approximately the average population size of a US county in 2015).
+The standard deviation of the X values is 0.1\% of the magnitude of the initial number of susceptible individuals, and the standard deviation of the Y values is 10\% of the magnitude of the initial number of infectious individuals, although in this setting the number of infectious individuals increases rapidly.
+We use Gaussian error as our baseline error because of its known statistical properties so we can more easily compare methods to one another.
+Later on, we explore the autoregressive and PAVA simulated data, which may better simulate data from  a real epidemic.
+
+      \subsection{Other Simulated Data}
+
+We then generate different SIR data sets, changing the one of the conditions for simulation each time and using the baseline SIR data as our basis for comparison.
+The different data sets generated are shown in Table \ref{tab:simulated-data}, for a total of 27 simulated data sets.
 
@@
 
-\subsection{Other Models}
-
-To test how well the SIR models perform under misspecification, we simulate data from other models.
-While we will not have a ``known'' $\rr$ value to compare our results to, this will allow us to see if the models give us reasonable results when the data do not come from the SIR model.
+% % Please add the following required packages to your document preamble:
+% % \usepackage{booktabs}
+% \begin{table}[]
+% \centering
+
+% Please add the following required packages to your document preamble:
+% \usepackage{booktabs}
+\begin{table}[]
+\centering
+\begin{tabular}{@{}llllllll@{}}
+\toprule
+\textbf{Name} & \textbf{Model} & \textbf{Error} & \textbf{($\beta, \gamma$)} & \textbf{O} & \textbf{(X(0),Y(0))} & \textbf{N} & \textbf{($\sigma_X, \sigma_Y$)} \\ \midrule
+  Baseline        & SIR      &   Norm             & (.06, .03)                          &     365       &  (99950, 50)                    &  $10^5$          &(100, 5)                                \\ \midrule
+  AR       & SIR      &   AR ($\rho=.7)$             & (.06, .03)                          &     365       &  (99950, 50)                    &  $10^5$          &(100, 5)                            \\ \midrule
+  \textcolor{red}{Norm-M}      & SIR      &   Norm-M ($\rho=.7)$             & (.06, .03)                          &     365       &  (99950, 50)                    &  $10^5$          &(100, 5)                            \\ \midrule
+   \textcolor{red}{AR-M}       & SIR      &   AR-M ($\rho=.7)$             & (.06, .03)                          &     365       &  (99950, 50)                    &  $10^5$          &(100, 5)                            \\ \midrule
+  InfRec1       & SIR      &   Norm             & (60000, .001)                          &     365       &  (99950, 50)                    &  $10^5$          &(100, 5)\\
+  InfRec2      & SIR      &   Norm             & (1500, .001)                          &     365       &  (99950, 50)                    &  $10^5$          &(100, 5)                                \\
+  InfRec3      & SIR      &   Norm             & (1000, .001)                          &     365       &  (99950, 50)                    &  $10^5$          &(100, 5)                                \\
+InfRec4      & SIR      &   Norm             & (06, .24)                          &     365       &  (99950, 50)                    &  $10^5$          &(100, 5)                                \\ \midrule
+  Time1 & SIR &  Norm   & (.06, .03)                          &     200       &  (99950, 50)                    &  $10^5$          &(100, 5)                                \\
+  Time2 & SIR &  Norm   & (.06, .03)                          &     100       &  (99950, 50)                    &  $10^5$          &(100, 5)                                \\
+  Time3 & SIR &  Norm   & (.06, .03)                          &     50       &  (99950, 50)                    &  $10^5$          &(100, 5)                                \\
+  Time4 & SIR &  Norm   & (.06, .03)                          &     20       &  (99950, 50)                    &  $10^5$          &(100, 5)                                \\ \midrule
+  Initial1    & SIR      &   Norm             & (.06, .03)                          &     365       &  (99999, 1)                    &  $10^5$          &(100, 5)                                \\
+  Initial2    & SIR      &   Norm             & (.06, .03)                          &     365       &  (99990, 10)                    &  $10^5$          &(100, 5)                                \\
+  Initial3    & SIR      &   Norm             & (.06, .03)                          &     365       &  (99900, 100)                    &  $10^5$          &(100, 5)                                \\
+  Initial4    & SIR      &   Norm             & (.06, .03)                          &     365       &  (99000, 1000)                    &  $10^5$          &(100, 5)                                \\ \midrule
+  \textcolor{red}{PopSize1}        & SIR      &   Norm             & (.06, .03)                          &     365       &  (999500, 500)                    &  $10^6$          &(100, 5)                                \\
+  \textcolor{red}{PopSize2}       & SIR      &   Norm             & (.06, .03)                          &     365       &  (9995, 5)                    &  $10^4$          &(100, 5)                                \\
+  \textcolor{red}{PopSize3}        & SIR      &   Norm             & (.06, .03)                          &     365       &  (990, 10)                    &  $10^3$          &(100, 5)                                \\
+  \textcolor{red}{PopSize4}        & SIR      &   Norm             & (.06, .03)                          &     365       &  (99, 1)                    &  $10^2$          &(100, 5)                                \\
+ \midrule
+  Var1      & SIR      &   Norm             & (.06, .03)                          &     365       &  (99950, 50)                    &  $10^5$          &(50, 2)                                \\
+  Var2      & SIR      &   Norm             & (.06, .03)                          &     365       &  (99950, 50)                    &  $10^5$          &(500, 20)                                \\
+  Var3      & SIR      &   Norm             & (.06, .03)                          &     365       &  (99950, 50)                    &  $10^5$          &(2500, 1000)                                \\
+  Var4      & SIR      &   Norm             & (.06, .03)                          &     365       &  (99950, 50)                    &  $10^5$          &(10000, 500)                                \\ \midrule
+  \textcolor{red}{Poly1}     & Linear      &   Norm             & (.06, .03)                          &     365       &  (99950, 50)                    &  $10^5$          &(100, 5)                            \\
+  \textcolor{red}{Poly2}    & Quartic      &   Norm             & (.06, .03)                          &     365       &  (99950, 50)                    &  $10^5$          &(100, 5)                            \\
+   LSIR     & LSIR      &   Norm             & (.06, .03)                          &     365       &  (99950, 50)                    &  $10^5$          &(100, 5)                            \\ 
+  \bottomrule
+\end{tabular}
+\caption{Table of simulated data sets.  We simulate 25 total data sets with the given parameters.}
+\label{tab:simulated-data}
+\end{table}
+
+To test how well the SIR models perform under misspecification, we simulate data from three other models.
+While we will not have a ``known'' $\rr$ value to compare our results to, this will allow us to see if the models give us reasonable (and comparable) results when the data do not come from the SIR model.
 We simulate data from polynomials of degree 1 and 4 with respect to the time step, using a monotonically increasing function for the $X$ compartment and a monotonically decreasing function for the $Z$ compartment.
-The sum of these is subtracted from the constant total population to get the $Y$ compartment.
-We also generated data from a linearized version of the three compartment SIR model with the following ODEs:
+The sum of these is subtracted from the constant total population to obtain the $Y$ compartment.
+We call these models polynomial of degree 1 (Poly1) and polynomial of degree 4 (Poly4), respectively.
+We also generated data from a linearized version of the three compartment SIR model (LSIR) with the following linearized ODEs:
 
@@
 
-The data from the linear SIR model were simulated under the following conditions:
-
-\begin{center}
-
-	$\beta$ = 0.06, $\gamma$ = 0.05; $\rr$ = 1.2
-
-	Number of time steps: 365
-
-	Starting X size ($X_0$): 99950
-
-	Starting Y size ($Y_0$): 50
-
-	Starting Z size: 0
-
-	$\sigma_X$: 10
-
-	$\sigma_Y$: 1
-
-	Total population: 100000.
-
-\end{center}
-
-\section{Results}\label{sec:results}
+% The data from the linear SIR model were simulated under the following conditions:
+
+% \begin{center}
+
+% 	$\beta$ = 0.06, $\gamma$ = 0.05; $\rr$ = 1.2
+
+% 	Number of time steps: 365
+
+% 	Starting X size ($X_0$): 99950
+
+% 	Starting Y size ($Y_0$): 50
+
+% 	Starting Z size: 0
+
+% 	$\sigma_X$: 10
+
+% 	$\sigma_Y$: 1
+
+% 	Total population: 100000.
+
+%       \end{center}
+
+\subsection{Model Fitting Details}\label{sec:model-fit}
+
+\begin{table}[H]
+	\centering
+	\begin{tabular}{@{}ll@{}}
+		\toprule
+		\textbf{Model} & \textbf{Default Standard Error Method} \\ \midrule
+		Least Squares & Delta method with Hessian matrix\\
+		Reparameterized Least Squares & Hessian matrix \\
+		Linear Model Approximation & Delta method with linear regression standard error \\
+		LMA, All Time Points & Delta method with linear regression standard error \\
+		Incidence-Prevalence Ratio & Jackknife resampling \\
+		Smoothed IPR & Jackknife resampling \\
+		Log-Linear & Linear regression standard error \\
+		Markov Chain & Delta method with GLM regression standard error \\
+		Sequential Bayes & Standard deviation of posterior distribution\\
+		Exponential Growth & Delta method with linear regression standard error \\
+		IDEA & Delta method with linear regression standard error \\
+		\bottomrule
+	\end{tabular}
+	\caption{Table of methods for standard errors.}
+	\label{tab:se-methods}
+\end{table}
+
+Table \ref{tab:baseline-res} shows the ways that standard errors are acquired for the different methods.
+The IPR, SIPR, and IDEA models require knowledge of the length of the average infectious period.
+For the results below, we assume the best case scenario that the true average infectious period is known for the data set.
+The exponential growth model requires knowledge of the probability distribution of the infectious period.
+In this case, we assume a uniform probability distribution with a mean equivalent to the true average infectious period.
+
+\section{Results of Simulations}\label{sec:results}
 There are numerous aspects to consider when using a method to estimate $\rr$.
-Among these include values of both the infection and recovery rate $(\beta, \gamma)$, the number of data points available (time $T$), the total size of the population ($N$), the initial percent of susceptible individuals $\left (\frac{X(0)}{N}\right)$, the initial percent of infectious individuals $\left (\frac{Y(0)}{N}\right )$, and  the magnitude of the variance of both the number of susceptibles and the number of infectious ($\sigma_X, \sigma_Y$).
+Among these include values of both the infection and recovery rate $(\beta, \gamma)$, the number of data points available (time $T$), the total size of the population ($N$), the initial percent of susceptible individuals $\left (\frac{X(0)}{N}\right)$, the initial percent of infectious individuals $\left (\frac{Y(0)}{N}\right )$, the magnitude of the variance of both the number of susceptibles and the number of infectious ($\sigma_X^2, \sigma_Y^2$), the error modeling assumptions, data generated from a non-SIR model, and finally methods of estimating the variance.
 We first examine our baseline data, and then we examine each of these aspects individually.
 
-\subsection{Baseline}\label{sec:res-base}
-
-\begin{figure}
+\subsection{Baseline Simulated Data}\label{sec:res-base}
+We implement our \wxxsir methods to estimate $\rr$ using the baseline data set.  The results are displayed in Figure \ref{fig:baseline-res} and Table \ref{tab:baseline-res}
+\begin{figure}[H]
   \centering
@@
 Err.
+The red line is the true $\rr$ value.}\label{fig:baseline-res}
+  \end{figure}
+
+\begin{table}[H]	
+	\centering
+	\begin{tabular}[t]{l|r|r}
+		\hline
+		Model & Estimate & Std. Err\\
+		\hline
+		SIRLS & 1.9999711 & 0.0055923\\
+		\hline
+		SIRREPAR & 1.9997340 & 0.0049939\\
+		\hline
+		SIRLMA & 1.6863640 & 0.1886132\\
+		\hline
+		SIRLMAT & 1.4063227 & 0.6309170\\
+		\hline
+		SIRIPR & 4.4203632 & 12.3593415\\
+		\hline
+		SIRSIPR & 1.8870824 & $<$ 1e-07 \\
+		\hline
+		LOGLINEAR & 1.9998072 & 0.0002198\\
+		\hline
+		SIRMC & 0.9485555 &  $<$ 1e-07 \\
+		\hline
+		SEQBAYES & 1.6495023 & 0.0672249\\
+		\hline
+		EXPGROWTH & 1.145541 & 0.0001405 \\
+		\hline
+		IDEA & 1.940302 & 0.0328945 \\
+		\hline
+	\end{tabular}
+        \caption{Baseline Data $(\beta=.06, \gamma=.03, O=365, X(0)=99950, Y(0)=50, Z(0)=0, \sigma_X=100, \sigma_Y=5, N=10^5)$, $\rr$ Estimates and Standard Errors}\label{tab:baseline-res}
+\end{table}
+
+Overall, in Figure \ref{fig:baseline-res}, we see a wide variation in the estimates of and the standard errors for $\rr$ depending on the method used.
+The methods of LS, ReLS, LMA, LMAT, and IPR all result in a CI that covers the true value of $\rr=2$.
+However, LMAT and IPR result in a CI which contains 1, so we would not be able to say whether or not we would have an outbreak of a disease or not.
+On the other hand, many methods appear to underestimate the true $\rr$, and many of them have larger standard errors as well.
+We would say that IPR's CI is uninformative.
+Some notable exceptions include the IPR model, which tends to overestimate $\rr$, and the MC and IPR models, which appear to have unreasonably small standard errors.
+
+\subsection{Autoregressive Errors}\label{sec:res-AR}
+We next examine the estimates of $\rr$ when fit to the autoregressive simulated data with $\rho=0.7$.
+In this data, we may see errors accumulate over time, which may more accurately reflect the course of infection in an epidemic.
+For example, we may expect to see more variation around the peak of an epidemic than the beginning.
+Besides the AR error, the other parameters are the same as in the baseline model.
+The results of the estimates are displayed in Figure \ref{fig:ar-res} and Table \ref{tab:ar-res}.
+
+In this setting, the CIs from LS, ReLS, LMAT, and SIPR cover the true $\rr=2$.
+In contrast to the baseline data, the CI for LMA no longer covers the true estimate.
+For all of these models, the standard error is larger, which would expect given that the data have a large positive autocorrelation coefficient.
+
+\begin{figure}[H]
+  \centering
+  \includegraphics[scale=0.5]{images/AR.jpeg}
+  \caption{Forest plots of estimates of $\rr$ for the \xxsir methods for the autoregressive (AR) ($\rho=.7)$ simulated data set $(\beta=.06, \gamma=.03, O=365, X(0)=99950, Y(0)=50, Z(0)=0, \sigma_X=100, \sigma_Y=5, N=10^5)$.
+The points are the estimates of $\rr$ and the lines denote $\pm 2\cdot $Std.
+Err.
 The red line is the true $\rr$ value.}
+  \end{figure}\label{fig:ar-res}
+\begin{table}[H]
+
+	\centering
+	\begin{tabular}[t]{l|r|r}
+		\hline
+		Model & Estimate & Std. Err\\
+		\hline
+		SIRLS & 1.9997387 & 0.0071509\\
+		\hline
+		SIRREPAR & 2.0000460 & 0.0063859\\
+		\hline
+		SIRLMA & 1.4744605 & 0.1130753\\
+		\hline
+		SIRLMAT & 1.3694249 & 0.8681724\\
+		\hline
+		SIRIPR & 1.7126909 & 5.0713852\\
+		\hline
+		SIRSIPR & 0.8792348 & $<$ 1e-07 \\
+		\hline
+		LOGLINEAR & 2.0000004 & 0.0002885\\
+		\hline
+		SIRMC & 0.9485555 & $<$ 1e-07 \\
+		\hline
+		SEQBAYES & 1.0000000 & 0.0523424\\
+		\hline
+		EXPGROWTH & 1.1446998 & 0.0001423\\
+		\hline
+		IDEA & 1.4701361 & 0.0164468\\
+		\hline
+	\end{tabular}
+        \caption{Autoregressive Errors, $\rr$ Estimates and Standard Errors}\label{tab:ar-res}
+\end{table}
+
+\subsection{PAVA Monotonicity}\label{sec:res-PAVA}
+We examine the case of the PAVA error assumptions which enforces monotonic assumptions on the suceptible and recovered compartments.
+This is turn may more accurately model the course of an epidemic.
+As a consequence, we may not be able to guarantee that a 95\% CI covers the true value of $\rr$ 95\% of the time, for example.
+
+In Figure \ref{fig:pava-res} and Table \ref{tab:pava-res}, we display the results of fitting the methods on two data sets, those generated from the Norm-G and Norm-AR error assumptions.
+We first see that there is very little difference among the estimates of $\rr$ in both Norm-G and Norm-AR error assumptions.
+We see that LMA and IPR have smaller estimates of $\rr$ and SB has a larger estimate of $\rr$.
+The standard errors for Norm-AR error assumption are greater than or equal to than the standard error in the Norm-G error assumption with the exceptions of LMAT and SB.
+
+The methods that contain the true value of $\rr$ are LS, ReLS, LMAT, IPR for both error assumptions and LL with Norm-AR error assumption.
+The methods LS, ReLS, LMA, and LL would be able to reject the hypothesis that $\rr < 1$.
+
+In comparison to the baseline data set,
+
+\begin{figure}[H]
+	\centering
+	\includegraphics[scale=0.5]{images/mono.jpeg}
+	\caption{Forest plots of estimates of $\rr$ for the \xxsir methods for the Norm-M and AR-M data sets $(\beta=.06, \gamma=.03, O=365, X(0)=99950, Y(0)=50, Z(0)=0, \sigma_X=100, \sigma_Y=5, N=10^5)$.
+The points are the estimates of $\rr$ and the lines denote $\pm 2\cdot $Std.
+Err.
+The red line is the true $\rr$ value.}
+	\label{fig:pava-res}
 \end{figure}
@@
 \begin{table}[H]
+
+	\centering
+	\begin{tabular}[t]{l|r|r|r|r}
+		\hline
+		Model & Norm-M Est. & Norm-M SE & AR-M Est. & AR-E SE\\
+		\hline
+		SIRLS & 2.0002458 & 0.0044456 & 1.9999920 & 0.0073664\\
+		\hline
+		SIRREPAR & 2.0003143 & 0.0039684 & 1.9996837 & 0.0065745\\
+		\hline
+		SIRLMA & 1.6793591 & 0.1102722 & 1.5349785 & 0.1496989\\
+		\hline
+		SIRLMAT & 1.4422630 & 0.9241642 & 1.4491091 & 0.7463529\\
+		\hline
+		SIRIPR & 0.7485576 & 1.5046827 & 0.7513498 & 1.5606142\\
+		\hline
+		SIRSIPR & 0.9081552 & 0.0000000 & 0.6976176 & 0.0000000\\
+		\hline
+		LOGLINEAR & 2.0003027 & 0.0001342 & 1.9995563 & 0.0002342\\
+		\hline
+		SIRMC & 0.9485555 & 0.0000000 & 0.9485555 & 0.0000000\\
+		\hline
+		SEQBAYES & 1.0000000 & 0.0523424 & 1.7143505 & 0.0685336\\
+		\hline
+		EXPGROWTH & 1.1440598 & 0.0001436 & 1.1441154 & 0.0001435\\
+		\hline
+		IDEA & 1.4701289 & 0.0164478 & 1.4701281 & 0.0164473\\
+		\hline
+	\end{tabular}
+\caption{Gaussian and Autoregressive Data with PAVA Applied, $\rr$ Estimates and Standard Errors}\label{tab:pava-res}
+\end{table}
+
+\subsection{Infection and Recovery Rate ($\beta, \gamma$)}\label{sec:res-beta-gamma}
+
+We next examine the simulated data sets InfRec1-4, three of which have different levels of $\beta$ for a given level of $\gamma=.001$ and the fourth is simulated with $\rr=0.25 < 1$, a case where an outbreak would not occur.
+The idea behind this is that since $\rr=\frac{\beta}{\gamma}$ and ratio estimators typically have large variance, we want to examine how a small value of $\gamma$, that is a small recovery rate, would effect the resulting estimates of $\rr$.
+Additionally, we would like to know when a disease will die out on its own accord, that we can be certain the estimate of $\rr$ is less than 1.
+
+\begin{figure}[H]
+  \centering
+  \includegraphics[scale=0.5]{images/parchange.jpeg}
+  \caption{Forest plots of estimates of $\rr$ for the \xxsir methods for the infrec1-4 data sets $(\beta=.06, \gamma=.03, O=365, X(0)=99950, Y(0)=50, Z(0)=0, \sigma_X=100, \sigma_Y=5, N=10^5)$.
+The points are the estimates of $\rr$ and the lines denote $\pm 2\cdot $Std.
+Err.
+The red line is the true $\rr$ value.}
+  \label{fig:infrec1-res}
+\end{figure}
+\begin{table}[H]
+
 	\centering
@@
 		\hline
-		SIRLS & 1.9999711 & 0.0055923\\
-		\hline
-		SIRREPAR & 1.9997340 & 0.0049939\\
-		\hline
-		SIRLMA & 1.6863640 & 0.1886132\\
-		\hline
-		SIRLMAT & 1.4063227 & 0.6309170\\
-		\hline
-		SIRIPR & 4.4203632 & 12.3593415\\
-		\hline
-		SIRSIPR & 1.8870824 & $<$ 1e-07 \\
-		\hline
-		LOGLINEAR & 1.9998072 & 0.0002198\\
-		\hline
-		SIRMC & 0.9485555 &  $<$ 1e-07 \\
-		\hline
-		SEQBAYES & 1.6495023 & 0.0672249\\
-		\hline
-	\end{tabular}
-        \caption{Baseline Data $(\beta=.06, \gamma=.03, O=365, X(0)=99950, Y(0)=50, Z(0)=0, \sigma_X=100, \sigma_Y=5, N=10^5)$, $\rr$ Estimates and Standard Errors}\label{tab:baseline}
-\end{table}
-
-Overall, we see a wide variation in the estimates of and the standard errors for $/rr$ depending on the method used.
-The methods that rely upon least squares optimization and the log-linear model are the closest to the true $\rr$ value, and they also have fairly small standard error estimates.
-On the other hand, a lot of the methods appear to underestimate the true $/rr$, and many of them have larger standard errors as well.
-Some notable exceptions include the IPR model, which tends to overestimate $/rr$, and the Markov chain and SIPR models, which appear to have unreasonably small standard errors.
-
-\subsection{Autoregressive Errors}\label{sec:res-AR}
-
-\includegraphics[scale=0.5]{images/AR.jpeg}
-\begin{table}[H]
-
-	\caption{\label{tab:}Autoregressive Errors, $\rr$ Estimates and Standard Errors}
+		SIRLS & 59.9764682 & 0.0434705\\
+		\hline
+		SIRREPAR & 59.9951876 & 0.4343984\\
+		\hline
+		SIRLMA &  -117.9694856 & 101.1885824\\
+		\hline
+		SIRLMAT & -2.8535930 & 193.6526107 \\
+		\hline
+		SIRIPR & 54.146506 & 200.2595\\
+		\hline
+		SIRSIPR & 3.9645187 & $<$ 1e-07\\
+		\hline
+		LOGLINEAR & 46.3209596 & 0.7797216\\
+		\hline
+		SIRMC & 0.9485554 & 0.0000001\\
+		\hline
+		SEQBAYES & 0.9769731 & 0.0517362\\
+		\hline
+		EXPGROWTH & 1.1410149 & 0.0001503\\
+		\hline
+		IDEA & 4.4448600 & 1.3788751\\
+		\hline
+	\end{tabular}
+        \caption{\label{tab:infrec1-res}$\gamma = 0.001$, $\rr$ Estimates and Standard Errors}
+\end{table}
+
+Looking in Figure \ref{fig:infrec1-res} and Table \ref{tab:infrec1-res} at the cases where $\gamma = 0.001$ with $\beta=60000, 1500$, and 1000, it appears that most of the models do not handle a small $\gamma$ well.
+We examine an exaggerated extreme case of when $\rr=60$ (the largest reported $\rr$ value we found is in \cite{anderson1992} for measles with $\hat{\rr}$=14) Only LS and ReLS result in CIs that cover the true $\rr$.
+IPR and LL give fairly close point estimates to the true $\rr=60$ as well.
+However, the other models do very poorly; they tend to greatly underestimate $\rr$.
+The linear model approximation does particularly poorly, likely because the linear model does not fit the data well, and is not even constrained to positive values of $X$, $Y$, and $Z$, which is why we can result in a negative estimate of $\rr$.
+Several of the models also return extremely large standard errors, meaning they are not very reliable in these cases.
+Only LS, ReLS, and LL would reject the hypothesis that $\rr <1$, when this is clearly not the case at all.
+
+When $\rr=1.5$ we see that the estimates from the methods give results much closer to the true value, and the resulting CIs are also smaller.
+The methods LS, ReLS, SIPR, LL, and SB would all reject the hypothesis that $\rr > 1$ (although if testing all methods at once, we should use a multiple-testing correction such as Bonferroni, but the only method which may then fail to reject would be SB as the other SEs are very small).
+
+When $\rr=1$, we are unsure whether there will be an epidemic or not.
+In this case, all the CIs of the methods cover the true value.
+We also see that the SE of LS is much greater than ReLS, which may be surprising.
+((WHY?))
+
+\begin{table}[H]
+
 	\centering
@@
 		\hline
-		SIRLS & 2.0000198 & 0.0078308\\
-		\hline
-		SIRREPAR & 1.9998852 & 0.0069901 \\
-		\hline
-		SIRLMA & 35.656866 & 4815.2163162\\
-		\hline
-		SIRLMAT & 1.049360 & 4.8830065 \\
-		\hline
-		SIRIPR & 2.183371 & 7.397726 \\
-		\hline
-		SIRSIPR & 0.8386350 & $<$ 1e-07 \\
-		\hline
-		LOGLINEAR & 1.9998203 & 0.0003227\\
-		\hline
-		SIRMC & 0.9485555 & $<$ 1e-07 \\
-		\hline
-		SEQBAYES & 1.0000000 & 0.0523424\\
-		\hline
-	\end{tabular}
-\end{table}
-
-With an autoregressive error structure in the simulated data, the log-linear and least squares estimation methods appear to have a similar accuracy.
-On the other hand, the other models mostly appear to be somewhat less accurate.
-In particular, the linear regression based models do much worse than in the Gaussian error case.
-For all of these models, the standard error is larger, which would expect given that the data have a large positive autocorrelation coefficient.
-
-\subsection{Infection and Recovery Rate ($\beta, \gamma$)}\label{sec:res-beta-gamma}
-
-\includegraphics[scale=0.5]{images/parchange.jpeg}
-
-\begin{table}[H]
-
-	\caption{\label{tab:}$\gamma = 0.001$, $\rr$ Estimates and Standard Errors}
-	\centering
-	\begin{tabular}[t]{l|r|r}
-		\hline
-		Model & Estimate & Std. Err\\
-		\hline
-		SIRLS & 59.9764682 & 0.0434705\\
-		\hline
-		SIRREPAR & 59.9951876 & 0.4343984\\
-		\hline
-		SIRLMA &  -117.9694856 & 101.1885824\\
-		\hline
-		SIRLMAT & -2.8535930 & 193.6526107 \\
-		\hline
-		SIRIPR & 54.146506 & 200.2595\\
-		\hline
-		SIRSIPR & 3.9645187 & $<$ 1e-07\\
-		\hline
-		LOGLINEAR & 46.3209596 & 0.7797216\\
-		\hline
-		SIRMC & 0.9485554 & 0.0000001\\
-		\hline
-		SEQBAYES & 0.9769731 & 0.0517362\\
-		\hline
-	\end{tabular}
-\end{table}
-
-Looking at the case where $\gamma = 0.001$, it appears that most of the models do not handle a very large $/rr$ well.
-Only the least squares and reparameterized least squares models give very close answers; the IPR and log-linear models give fairly close answers as well.
-However, the other models do very poorly; they ted to greatly underestimate $/rr$.
-The linear model approximation does particularly poorly, likely because the linear model does not fit the data well.
-Several of the models also return extremely large standard errors, meaning they are not very reliable in these cases.
-
-\begin{table}[H]
-
-	\caption{\label{tab:}$\gamma = 0.24$, $\rr$ Estimates and Standard Errors}
-	\centering
-	\begin{tabular}[t]{l|r|r}
-		\hline
-		Model & Estimate & Std. Err\\
-		\hline
 		SIRLS & 0.2363274 & 4.1042813\\
@@
 		\hline
-	\end{tabular}
-\end{table}
-
-The above table corresponds to the case where $\gamma = 0.24$.
-It appears that a lot of estimation methods do not handle very small $/rr$ well either.
-None of them are extremely close to truth, except for regular least squares, as most of the tend to overestimate the truth.
+		EXPGROWTH & 1.0924931 & 0.0003093\\
+		\hline
+		IDEA & 1.1299625 & 0.0039222\\
+		\hline
+	\end{tabular}
+        \caption{$\gamma = 0.24$, $\rr$ Estimates and Standard Errors}\label{tab:infrec4-res}
+\end{table}
+
+Table \ref{tab:infrec4-res} corresponds to the case where $\gamma = 0.24$ and $\beta=0.06$.
+It appears that a lot of estimation methods do not handle small $\rr$ well either.
+None of them are extremely close to truth, except for ReLS, as most of the tend to overestimate the truth.
 The IPR is particularly bad in this case.
@@
 \subsection{Number of Time Steps ($T$)}\label{sec:res-time}
-
+We examine how the number of time steps the estimates of $\rr$ in the different methods.
+Ideally, the more data we have, the more likely we are to have unbiased and low-variance estimates for $\rr$.
+Since we would like to  infer how a disease will spread through a population, the sooner we can be sure whether $\rr> 1$, the sooner we can provide interventions and control measures for a disease.
+Another reason to analyze the number of time steps is that some of the methods assume linear or exponential growth of an epidemic in a certain time period.
+If the growth is no longer exponential, we will obtain a poor estimate of $\rr$.
+In this case, we can be using too many time points.
+
+In Figure \ref{fig:time-res} and Tables \ref{tab:time-res1} and \ref{tab:time-res2} we examine the methods fit to different amounts of time periods where we observe the progress of a disease, including $T=200$, 100, 50, and 20.
+
+Again, we see that IPR does very poorly in estimating $\rr$, regardless of the amount of data observed.
+\begin{figure}[H]
+  \centering
   \includegraphics[scale=0.5]{images/time.jpeg}
-
-\begin{table}[H]
-
-	\caption{\label{tab:} $T$ = 100, $\rr$ Estimates and Standard Errors}
+  \caption{Forest plots of estimates of $\rr$ for the \xxsir methods for the Time1-4 data sets $(\beta=.06, \gamma=.03, O=365, X(0)=99950, Y(0)=50, Z(0)=0, \sigma_X=100, \sigma_Y=5, N=10^5)$.
+The points are the estimates of $\rr$ and the lines denote $\pm 2\cdot $Std.
+Err.
+The red line is the true $\rr$ value.}\label{fig:time-res}
+  \end{figure}
+
+\begin{table}[H]
+
 	\centering
@@
 		\hline
-	\end{tabular}
-\end{table}
-
-The table above shows the methods applied to the first 100 time steps, out of 365, of a baseline data set.
-The estimates and standard errors are worse, which we would expect given less data.
+		EXPGROWTH & 1.5600292 & 0.0022192\\
+		\hline
+		IDEA & 4.7772246 & 0.2173201\\
+		\hline
+	\end{tabular}
+        \caption{ $T$ = 100, $\rr$ Estimates and Standard Errors}\label{tab:time-res1}
+\end{table}
+
+Table \ref{tab:time-res1} shows the methods applied to the first 100 time steps, out of 365, of the baseline data set .
+The estimates and standard errors are worse than baseline, which we would expect given less data.
 The results are still relatively close to the full baseline data.
-There also appears to be less tendency to underestimate $/rr$ compared to a longer time scale.
-
-\begin{table}[H]
-
-	\caption{\label{tab:} $T$ = 20, $\rr$ Estimates and Standard Errors}
+There also appears to be less tendency to underestimate $\rr$ compared to a longer time scale.
+In general, we see that that the length of CIs decreases as we use more and more time points.
+Only LL would reject the hypothesis that $\rr < 1$ for 50 or fewer time points, but its CI does not cover the true $\rr$ value.
+With 50 and greater time points, we see that LS and ReLS provide close point estimates to the true value and by 100 time points, would reject the hypothesis that $\rr < 1$.
+The methods and MC and SB consistently have small CIs, but never cover the true value.
+
+\begin{table}[H]
   \centering
@@
 		\hline
-	\end{tabular}
-\end{table}
-
-The table above shows the methods applied to the first 20 time steps of a baseline data set.
-As might be expected, the $/rr$ estimates are wildly inaccurate, and none are very close.
-The least squares methods are particularly worse in this situation compared to the other data sets.
+		EXPGROWTH & 4.4726539 & 0.0339963\\
+		\hline
+		IDEA & 19.5369460 & 2.3356802\\
+		\hline
+	\end{tabular}
+        \caption{ $T$ = 20, $\rr$ Estimates and Standard Errors}\label{tab:time-res2}
+\end{table}
+
+Table \ref{tab:time-res2} shows the methods applied to the first 20 time steps of a baseline data set.
+As might be expected, the $\rr$ estimates are wildly inaccurate, and none are very close to the true value.
+The methods of LS and ReLS are particularly worse in this situation compared to the other data sets.
 The standard errors are much larger than the baseline, which is to be expected as well due to the small number of data points.
-This shows that it is extremely difficult to create accurate online estimates of $/rr$ or to estimate $/rr$ without adequate data.
+This shows that it is extremely difficult to create accurate online estimates of $\rr$ or to estimate $\rr$ without adequate data.
 
 \subsection{Initial Percentage of Infectious $\left (\frac{Y(0)}{N}\right)$}\label{sec:res-inf}
-
+We next examine the initial conditions of the number of susceptible, infectious, and recovered.
+Deterministically, different initial conditions in the number of infectious can cause very different epidemic courses (see Figures in appendix).
+
+In Figure \ref{fig:inits-res} and Tables \ref{tab:inits-res1} and \ref{tab:inits-res2}, we display the results of different number of initially infectious individuals of 1, 10, 100, and 1000 with a constant population of 100000.
+We see that for a smaller percentage of initially infectious individuals, we have larger CIs in our estimates of $\rr$.
+In fact, LL is the only method that has a consistently small CI around the true $\rr=2$, regardless of the starting number of infectious.
+
+\begin{figure}[H]
+  \centering
   \includegraphics[scale=0.5]{images/start.jpeg}
-
-\begin{table}[H]
-
-	\caption{\label{tab:}$Y(0) = 1000$, $\rr$ Estimates and Standard Errors}
+    \caption{Forest plots of estimates of $\rr$ for the \xxsir methods for the Inits1-4 data sets $(\beta=.06, \gamma=.03, O=365, X(0)=99950, Y(0)=50, Z(0)=0, \sigma_X=100, \sigma_Y=5, N=10^5)$.
+The points are the estimates of $\rr$ and the lines denote $\pm 2\cdot $Std.
+Err.
+The red line is the true $\rr$ value.}\label{fig:inits-res}
+  \end{figure}
+
+  Table \ref{tab:inits-res1} shows the methods applied to data where $Y(0) = 1000$, essentially starting at the middle of an epidemic.
+This data has very large underestimation of $\rr$.
+The LMA method once again gives an estimate very far from the true $\rr$, potentially indicating that a linear approximation does not apply in this case.
+The sample standard errors in this case are actually relatively close to the baseline, and even smaller for some of the methods.
+
+Table \ref{tab:inits-res2} shows the methods applied to data where $Y(0) = 1$, starting at the very beginning of an epidemic, and the true technical definition of $\rr$ where we introduce exactly one infectious individual into a population.
+The methods give many larger estimates, tending to overestimate the true $\rr$.
+In particular, IPR and LMA methods give very large results.
+We also see a large standard error than the baseline.
+
+\begin{table}[H]
+
 	\centering
@@
 		\hline
-	\end{tabular}
-\end{table}
-
-The table above shows the methods applied to data where $Y(0) = 1000$, essentially starting at the middle of an epidemic.
-This data has very large underestimation of $/rr$.
-The LMA method once again gives a very far estimate, potentially indicating that a linear approximation does not apply in this case.
-The sample standard errors in this case are actually relatively close to the baseline, and even smaller for some of the methods.
-
-\begin{table}[H]
-
-	\caption{\label{tab:}$Y(0) = 1$, $\rr$ Estimates and Standard Errors}
+		EXPGROWTH & 1.5600292 & 0.0022192\\
+		\hline
+		IDEA & 4.7772246 & 0.2173201\\
+		\hline
+	\end{tabular}
+        \caption{$Y(0) = 1000$, $\rr$ Estimates and Standard Errors}\label{tab:inits-res1}
+\end{table}
+
+\begin{table}[H]
+
 	\centering
@@
 		\hline
-	\end{tabular}
-\end{table}
-
-The table above shows the methods applied to data where $Y(0) = 1$, starting at the very beginning of an epidemic.
-The methods give many larger estimates, tending to overestimate the true $/rr$.
-In particular, the IPR and LMA methods give very large results.
-We also see a large standard error than the baseline.
+		EXPGROWTH & 2.2380282 & 0.0078639\\
+		\hline
+		IDEA & 8.5233073 & 0.6084195\\
+		\hline
+	\end{tabular}
+        \caption{$Y(0) = 1$, $\rr$ Estimates and Standard Errors}\label{tab:inits-res2}
+\end{table}
+
+\subsection{Population Size $N$}\label{sec:res-n}
+
+\begin{figure}[H]
+	\centering
+	\includegraphics[scale=0.5]{images/popsize.jpeg}
+	\caption{Forest plots of estimates of $\rr$ for the \xxsir methods for the PopSize1-4 data sets $(\beta=.06, \gamma=.03, O=365, \sigma_X=100, \sigma_Y=5)$.
+The points are the estimates of $\rr$ and the lines denote $\pm 2\cdot $Std.
+Err.
+The red line is the true $\rr$ value.}\label{fig:inits-res}
+\end{figure}
+
+\begin{table}[H]
+
+	\centering
+	\begin{tabular}[t]{l|r|r}
+		\hline
+		Model & Estimate & Std. Err\\
+		\hline
+		SIRLS & 2.0007741 & 0.0590423\\
+		\hline
+		SIRREPAR & 2.0005189 & 0.0528419\\
+		\hline
+		SIRLMA & 1.9976309 & 2606.0798123\\
+		\hline
+		SIRLMAT & 1.2944136 & 1.4304706\\
+		\hline
+		SIRIPR & 40.8072373 & 133.2798314\\
+		\hline
+		SIRSIPR & 1.0628910 &  $<$ 1e-07\\
+		\hline
+		LOGLINEAR & 1.9988393 & 0.0023193\\
+		\hline
+		SIRMC & 0.9485626 & 0.0000040\\
+		\hline
+		SEQBAYES & 0.4418182 & 0.0347917\\
+		\hline
+		EXPGROWTH & 1.1021571 & 0.0002679\\
+		\hline
+		IDEA & 1.9403960 & 0.0329020\\
+		\hline
+	\end{tabular}
+\caption{$N = 10000$, $\rr$ Estimates and Standard Errors}\label{tab:n1-res2}
+\end{table}
+
+\begin{table}[H]
+
+	\centering
+	\begin{tabular}[t]{l|r|r}
+		\hline
+		Model & Estimate & Std. Err\\
+		\hline
+		SIRLS & 1.9862533 & 4.861988\\
+		\hline
+		SIRREPAR & 1.9871049 & 3.456891\\
+		\hline
+		SIRLMA & 0.8023469 & 18.94166\\
+		\hline
+		SIRLMAT & 1.2208744 & 4.769929\\
+		\hline
+		SIRIPR & 3620.8676795 & 57608.17\\
+		\hline
+		SIRSIPR & 3.3678247 & 22.41867\\
+		\hline
+		LOGLINEAR & 1.1413175 & 0.0422219\\
+		\hline
+		SIRMC & 0.8455334 & 0.0124837\\
+		\hline
+		SEQBAYES & 0.3887115 & 0.0326338\\
+		\hline
+		EXPGROWTH & 1.0912970 & 0.0003149\\
+		\hline
+		IDEA & 1.2475728 & 0.0297242\\
+		\hline
+	\end{tabular}
+\caption{$N = 100$, $\rr$ Estimates and Standard Errors}\label{tab:n2-res2}
+\end{table}
 
@@
 
+\begin{center}
 	\includegraphics[scale=0.5]{images/var.jpeg}
+\end{center}
 
@@
 		SEQBAYES & 1.6750053 & 0.0677426\\
+		\hline
+		EXPGROWTH & 1.1444008 & 0.0001429\\
+		\hline
+		IDEA & 1.9402950 & 0.0328946\\
 		\hline
@@
 		SEQBAYES & 1.8567176 & 0.0713225\\
+		\hline
+		EXPGROWTH & 1.2030700 & 0.0000599\\
+		\hline
+		IDEA & 1.9404973 & 0.0328855\\
 		\hline
@@
 
-\begin{table}[H]
-
 \begin{center}
@@
 
-\caption{\label{tab:} 4th Order Model, $\rr$ Estimates and Standard Errors}
+\begin{table}[H]
+
 	\centering
@@
 		\hline
+		Model & Estimate & Std. Err\\
+		\hline
+		SIRLS & 1.0514375 & 0.3782909\\
+		\hline
+		SIRREPAR & 1.0514351 & 0.0259416\\
+		\hline
+		SIRLMA & 1.2655140 & 0.1278435\\
+		\hline
+		SIRLMAT & 1.1597070 & 0.1019956\\
+		\hline
+		SIRIPR & 4.4999548 & 15.2185520\\
+		\hline
+		SIRSIPR & 98.0610578 & $<$ 1e-07\\
+		\hline
+		LOGLINEAR & 1.1880504 & 0.0006341\\
+		\hline
+		SIRMC & 0.9485551 & 0.0000001\\
+		\hline
+		SEQBAYES & 1.4150329 & 0.0622640\\
+		\hline
+		EXPGROWTH & 1.0614311 & 0.0004905\\
+		\hline
+		IDEA & 2.0340199 & 0.0314227\\
+		\hline
+	\end{tabular}
+	\caption{Linear Data, $\rr$ Estimates and Standard Errors}\label{tab:lin-res}
+\end{table}
+
+\begin{table}[H]
+
+\centering
+\begin{tabular}[t]{l|r|r}
+	\hline
 	Model & Estimate & Std. Dev\\
@@
 	\hline
-\end{tabular}
+	EXPGROWTH & 1.1485237 & 0.0001344\\
+	\hline
+	IDEA & 1.5632013 & 0.0198752\\
+	\hline
+\end{tabular}
+\caption{\label{tab:f-ord} 4th Order Model, $\rr$ Estimates and Standard Errors}
 \end{table}
@@
 
+\begin{table}[H]
+
+	\centering
+	\begin{tabular}[t]{l|r|r}
+		\hline
+		Model & Estimate & Std. Err\\
+		\hline
+		SIRLS & 1.2742577 & 0.0979741\\
+		\hline
+		SIRREPAR & 1.2743142 & 0.0339775\\
+		\hline
+		SIRLMA & 1.2191008 & 0.1092962\\
+		\hline
+		SIRLMAT & 1.2009373 & 0.0031099\\
+		\hline
+		SIRIPR & 2.5280045 & 2.5904017\\
+		\hline
+		SIRSIPR & 2.1567978 & 0.3585219\\
+		\hline
+		LOGLINEAR & 1.2585872 & 0.0010079\\
+		\hline
+		SIRMC & 0.9485555 & $<$ 1e-07 \\
+		\hline
+		SEQBAYES & 1.2008765 & 0.0573592\\
+		\hline
+		EXPGROWTH & 1.0877415 & 0.0003320\\
+		\hline
+		IDEA & 2.0335555 & 0.0314612\\
+		\hline
+	\end{tabular}
+	\caption{Linear SIR Data, $\rr$ Estimates and Standard Errors}\label{tab:lsir-res}
+\end{table}
+
+\subsection{Variance Estimation Methods}
+
 \subsection{Repeated Sampling}
@@
 
-	\caption{\label{tab:}Mean $\rr$ Estimates and Std. Devs, Repeated Sampling}
 	\centering
@@
 		\hline
-	\end{tabular}
+		EXPGROWTH & 1.1453460 & 0.0001981 & 1.1455408 & 0.0001405 \\
+		\hline
+		IDEA & 1.9402790 & 0.0000256 & 1.4701510 & 0.0164472 \\
+		\hline
+	\end{tabular}
+\caption{\label{tab:rep-samp}Mean $\rr$ Estimates and Std. Devs, Repeated Sampling}
 \end{table}
@@
 
+\begin{table}[H]
+
+	\caption{\label{tab:}$\rr$ Estimates and Std. Errs, SIRLS Model,
+		$\beta = 0.06, \gamma = 0.03$, $X_0 = 99950, Y_0 = 50$, $\sigma_X = 100, \sigma_Y = 5$}
+	\centering
+	\begin{tabular}[t]{l|r|r|r|r|r|r|r|r}
+		\hline
+		Data Set & Norm Est & Norm SE & Norm-M Est & Norm-M SE & AR Est & AR SE & AR-M Est & AR-M SE\\
+		\hline
+		$X_0 = 99, Y_0 = 1$ & 1.986253 & 4.8619876 & 2.038011 & 0.4537907 & 2.106225 & 7.1198939 & 2.105517 & 1.5447687\\
+		\hline
+		$X_0 = 990, Y_0 = 10$ & 1.997222 & 0.5406162 & 2.017454 & 0.1429851 & 2.122109 & 0.7362701 & 2.097491 & 0.2921481\\
+		\hline
+		$X_0 = 9995, Y_0 = 5$ & 2.000774 & 0.0590423 & 1.995300 & 0.0239310 & 2.000782 & 0.0744317 & 1.996394 & 0.0489651\\
+		\hline
+		$X_0 = 999500, Y_0 = 500$ & 2.000050 & 0.0005467 & 1.999834 & 0.0005748 & 1.999833 & 0.0008762 & 1.999959 & 0.0007396\\
+		\hline
+	\end{tabular}
+\end{table}
+
 \subsection{SIR Reparameterization}
@@
 
+\begin{table}[H]
+
+	\caption{\label{tab:}$\rr$ Estimates and Std. Errs, SIRREPAR Model,
+		$\beta = 0.06, \gamma = 0.03$, $X_0 = 99950, Y_0 = 50$, $\sigma_X = 100, \sigma_Y = 5$}
+	\centering
+	\begin{tabular}[t]{l|r|r|r|r|r|r|r|r}
+		\hline
+		Data Set & Norm Est & Norm SE & Norm-M Est & Norm-M SE & AR Est & AR SE & AR-M Est & AR-M SE\\
+		\hline
+		$X_0 = 99, Y_0 = 1$ & 1.987105 & 3.4568905 & 2.037779 & 0.3348238 & 1.658070 & 2.0918740 & 2.105172 & 1.1781993\\
+		\hline
+		$X_0 = 990, Y_0 = 10$ & 1.997065 & 0.3756188 & 2.017454 & 0.1020655 & 2.121902 & 0.5659528 & 2.097579 & 0.2210569\\
+		\hline
+		$X_0 = 9995, Y_0 = 5$ & 2.000519 & 0.0528419 & 1.995631 & 0.0212989 & 2.000955 & 0.0665574 & 1.996249 & 0.0434920\\
+		\hline
+		$X_0 = 999500, Y_0 = 500$ & 1.999946 & 0.0004956 & 1.999946 & 0.0005270 & 2.000104 & 0.0007140 & 1.999734 & 0.0006529\\
+		\hline
+	\end{tabular}
+\end{table}
+
 \subsection{SIR Linear Model Approximation}
@@
 
+\begin{table}[H]
+
+	\caption{\label{tab:}$\rr$ Estimates and Std. Errs, SIRLMAT Model,
+		$\beta = 0.06, \gamma = 0.03$, $\sigma_X = 100, \sigma_Y = 5$}
+	\centering
+	\begin{tabular}[t]{l|r|r|r|r|r|r|r|r}
+		\hline
+		Data Set & Norm Est & Norm SE & Norm-M Est & Norm-M SE & AR Est & AR SE & AR-M Est & AR-M SE\\
+		\hline
+		$X_0 = 99, Y_0 = 1$ & 1.220874 & 4.7699293 & -0.0614019 & 15.5900050 & -0.0217592 & 18.7796434 & 0.9203832 & 1.5845866\\
+		\hline
+		$X_0 = 990, Y_0 = 10$ & 2.248607 & 29.7561744 & 1.0457445 & 0.6365320 & 1.0880985 & 3.9238850 & -13.6905770 & 281.9114385\\
+		\hline
+		$X_0 = 9995, Y_0 = 5$ & 1.294414 & 1.4304706 & 1.3710979 & 0.6597925 & 1.5307307 & 1.0343598 & 1.3865816 & 1.1453533\\
+		\hline
+		$X_0 = 999500, Y_0 = 500$ & 1.370406 & 0.8821226 & 1.9731686 & 10.8871603 & 1.3885429 & 0.6854633 & 1.4164038 & 0.6347757\\
+		\hline
+	\end{tabular}
+\end{table}
+
 \subsection{Incidence to Prevalence Ratio}
@@
 
+\begin{table}[H]
+
+	\caption{\label{tab:}$\rr$ Estimates and Std. Errs, SIRIPR Model,
+		$\beta = 0.06, \gamma = 0.03$, $\sigma_X = 100, \sigma_Y = 5$}
+	\centering
+	\begin{tabular}[t]{l|r|r|r|r|r|r|r|r}
+		\hline
+		Data Set & Norm Est & Norm SE & Norm-M Est & Norm-M SE & AR Est & AR SE & AR-M Est & AR-M SE\\
+		\hline
+		$X_0 = 99, Y_0 = 1$ & 3620.867680 & 57608.167596 & 5.239294 & 63.654679 & 729.654107 & 2978.983072 & 15.800132 & 284.613763\\
+		\hline
+		$X_0 = 990, Y_0 = 10$ & 67.597632 & 157.628039 & 1.734334 & 12.726031 & 164.611097 & 1887.125297 & 1.707676 & 11.777508\\
+		\hline
+		$X_0 = 9995, Y_0 = 5$ & 40.807237 & 133.279831 & 1.797818 & 7.233937 & 976.308278 & 17227.979404 & 1.700549 & 9.887431\\
+		\hline
+		$X_0 = 999500, Y_0 = 500$ & 1.577592 & 1.422029 & 1.434415 & 1.131136 & 1.567228 & 1.594393 & 1.449019 & 1.180341\\
+		\hline
+	\end{tabular}
+\end{table}
+
 \subsection{Smoothed Incidence to Prevalence Ratio}
@@
 
+\begin{table}[H]
+
+	\caption{\label{tab:}$\rr$ Estimates and Std. Errs, SIRSIPR Model,
+		$\beta = 0.06, \gamma = 0.03$, $\sigma_X = 100, \sigma_Y = 5$}
+	\centering
+	\begin{tabular}[t]{l|r|r|r|r|r|r|r|r}
+		\hline
+		Data Set & Norm Est & Norm SE & Norm-M Est & Norm-M SE & AR Est & AR SE & AR-M Est & AR-M SE\\
+		\hline
+		$X_0 = 99, Y_0 = 1$ & 3.367825 & 22.41867 & 0.9850654 & 0.5665987 & 1.5591972 & 3959.42264 & 4.207729 & 13.26187\\
+		\hline
+		$X_0 = 990, Y_0 = 10$ & 1.912639 & 390.58481 & 1.6931741 & 12.0122249 & 0.5334608 & 13.65189 & 2.898150 & 284.49327\\
+		\hline
+		$X_0 = 9995, Y_0 = 5$ & 1.062891 & 0.00000 & 1.8635858 & 0.0000000 & 1.0938066 & 0.00000 & 1.572230 & 0.00000\\
+		\hline
+		$X_0 = 999500, Y_0 = 500$ & 1.792115 & 0.00000 & 1.7907398 & 0.0000000 & 1.7902743 & 0.00000 & 1.750503 & 0.00000\\
+		\hline
+	\end{tabular}
+\end{table}
+
 \subsection{SIR Log Linear}
@@
 
+\begin{table}[H]
+
+	\caption{\label{tab:}$\rr$ Estimates and Std. Errs, LOGLINEAR Model,
+		$\beta = 0.06, \gamma = 0.03$, $\sigma_X = 100, \sigma_Y = 5$}
+	\centering
+	\begin{tabular}[t]{l|r|r|r|r|r|r|r|r}
+		\hline
+		Data Set & Norm Est & Norm SE & Norm-M Est & Norm-M SE & AR Est & AR SE & AR-M Est & AR-M SE\\
+		\hline
+		$X_0 = 99, Y_0 = 1$ & 1.141317 & 0.0422218 & 2.031666 & 0.0113166 & 1.056918 & 0.0447463 & 1.771771 & 0.0561071\\
+		\hline
+		$X_0 = 990, Y_0 = 10$ & 2.255041 & 0.0379147 & 2.010183 & 0.0055668 & 2.470092 & 0.0545897 & 2.280617 & 0.0229552\\
+		\hline
+		$X_0 = 9995, Y_0 = 5$ & 1.998839 & 0.0023193 & 1.997778 & 0.0007420 & 2.004827 & 0.0027816 & 2.001058 & 0.0015548\\
+		\hline
+		$X_0 = 999500, Y_0 = 500$ & 1.999983 & 0.0000215 & 1.999989 & 0.0000203 & 1.999924 & 0.0000314 & 1.999829 & 0.0000294\\
+		\hline
+	\end{tabular}
+\end{table}
+
 \subsection{SIR Markov Chain}
@@
 
+\begin{table}[H]
+
+	\caption{\label{tab:}$\rr$ Estimates and Std. Errs, SIRMC Model,
+		$\beta = 0.06, \gamma = 0.03$, $\sigma_X = 100, \sigma_Y = 5$}
+	\centering
+	\begin{tabular}[t]{l|r|r|r|r|r|r|r|r}
+		\hline
+		Data Set & Norm Est & Norm SE & Norm-M Est & Norm-M SE & AR Est & AR SE & AR-M Est & AR-M SE\\
+		\hline
+		$X_0 = 99, Y_0 = 1$ & 0.8455334 & 0.0124837 & 0.9481149 & 0.0002788 & NaN & NaN & 0.9480033 & 0.0005127\\
+		\hline
+		$X_0 = 990, Y_0 = 10$ & 0.9432094 & 0.0014184 & 0.9485736 & 0.0000060 & 0.9500365 & 0.0003484 & 0.9486116 & 0.0000141\\
+		\hline
+		$X_0 = 9995, Y_0 = 5$ & 0.9485626 & 0.0000040 & 0.9485575 & 0.0000002 & 0.9492072 & 0.0002225 & 0.9485575 & 0.0000002\\
+		\hline
+		$X_0 = 999500, Y_0 = 500$ & 0.9485553 & 0.0000000 & 0.9485553 & 0.0000000 & 0.9485553 & 0.0000000 & 0.9485553 & 0.0000000\\
+		\hline
+	\end{tabular}
+\end{table}
+
 \subsection{SIR Sequential Bayes}
@@
 
+\begin{table}[H]
+
+	\caption{\label{tab:}$\rr$ Estimates and Std. Errs, SEQBAYES Model,
+		$\beta = 0.06, \gamma = 0.03$, $\sigma_X = 100, \sigma_Y = 5$}
+	\centering
+	\begin{tabular}[t]{l|r|r|r|r|r|r|r|r}
+		\hline
+		Data Set & Norm Est & Norm SE & Norm-M Est & Norm-M SE & AR Est & AR SE & AR-M Est & AR-M SE\\
+		\hline
+		$X_0 = 99, Y_0 = 1$ & 0.3887115 & 0.0326338 & 1.574343 & 0.0656755 & 0.4241419 & 0.0340886 & 1.000000 & 0.0523424\\
+		\hline
+		$X_0 = 990, Y_0 = 10$ & 0.9028353 & 0.0497345 & 1.000000 & 0.0523424 & 1.5457192 & 0.0650757 & 1.000000 & 0.0523424\\
+		\hline
+		$X_0 = 9995, Y_0 = 5$ & 0.4418182 & 0.0347917 & 1.000000 & 0.0523424 & 1.0000000 & 0.0523424 & 1.452822 & 0.0630899\\
+		\hline
+		$X_0 = 999500, Y_0 = 500$ & 1.1774169 & 0.0567961 & 1.270930 & 0.0590085 & 1.1876846 & 0.0570432 & 1.807844 & 0.0703775\\
+		\hline
+	\end{tabular}
+\end{table}
+
 \subsection{Exponential Growth}

### 2018-07-03 paper/Version14/draft_v14 (Shannon Gallagher's conflicted copy 2018-06-26).tex
finally on git

--- 2018-07-03 paper/Version13/draft_v13.tex
@@
 \usepackage{float}
+\usepackage{url}
 
@@
 Despite a clear definition of $\rr$, epidemiologists have struggled to create a standard  estimator for $\rr$  \citep{hethcote2000}.
-A major issue in estimating $\rr$ is that the quantity is a \textit{property of the model}, meaning that $\rr$ is dependent not only on the usual noise that comes with statistical modeling but also on a variety of assumptions on how researchers assume a disease is transmitted through a population \citep{diekmann2009}.
+A major issue in estimating $\rr$ is that the quantity is a \textit{property of the model}, meaning that the estimate $\hat{\rr}$ is dependent not only on the usual noise that comes with statistical modeling but also on a variety of assumptions on how researchers assume a disease is transmitted through a population \citep{diekmann2009}.
 
@@
 Here, S and I are known as compartments where S stands for ``susceptible'' and I for ``infectious.'' Infectious disease models then specify how individuals move from the S compartment to the I compartment.
-Often, ancillary compartments are added such as R, which stands for recovered or dead, or E, which stands for exposed but not yet infectious.
+Often, ancillary compartments are added such as R, which stands for recovered or removed, or E, which stands for exposed but not yet infectious.
 The estimate of $\rr$ from a SIR model will not be directly comparable to the estimate of $\rr$ from a SEIR model, for example.
@@
 We note that these methods are not exhaustive but are chosen due to their impact and use in epidemiology.
-We find that these methods also help to highlight the problems discussed above.
+We find that these methods also help to highlight the problems mentioned above.
 
@@
 Within these simulations, we vary the number of observations used to estimate $\rr$, infection and recovery rates, the total population size, initial SI ratio, and several assumptions about the noise.
+After analyzing our simulations, we apply our methods to data from the USA H1N1 influenza pandemic.
 
@@
 In Section \ref{sec:ci}, we discuss how to estimate CIs for each of the \wxxsir methods.
-In Section \ref{sec:sim-res}, we present the results of our simulations.
+In Sections \ref{sec:sim-res} - \ref{sec:sim-var-res}, we present the results of our simulations.
+In Section \ref{sec:real-data}, we evaluate the methods on real data--from the USA H1N1 influenza pandemic of 2009.
 Finally, in Section \ref{sec:discussion}, we provide recommendations of estimating $\rr$ for the SIR model and offer comments on how these conclusions extend to more complex models.
@@
 For example, the SIR model is special in that two of its compartments, S and R are monotonic.
-The number of susceptibles is non-increasing and the number of susceptibles is non-decreasing.
+The number of susceptibles is non-increasing and the number of recovereds is non-decreasing.
 Logically, models should enforce this monotonicity of the two compartments.
 As a consequence, this restricts assumptions on how randomness enters the model.
-For example, Gaussian error, a common assumption in statistics and machine learning, becomes highly suspect due to the non-symmetrical nature of the noise imposed by the monotonic $S$ and $R$ compartments.
+For example, Gaussian error, a common assumption in statistics and machine learning, becomes highly suspect due to the assymmetrical nature of the noise imposed by the monotonic $S$ and $R$ compartments.
 Without assuming Gaussian noise, it becomes much more difficult to guarantee properties such as unbiased and consistent estimates of $\rr$.
@@
 The lowest possible estimate for $\rr$ reported is 0.12 for HIV in Uganda, and the largest is 17.00 for measles in England and Wales.
-Within diseases, we see that estimates for $\rr$ seem to be similar, although typically do not lie in one another's intervals, suggesting that there is disparity in estimates of $\rr$.
+Within diseases, we see that estimates for $\rr$ seem to be similar, although the respective CIs typically do not intersect, suggesting that there is disparity in estimates of $\rr$.
 
@@
 \begin{landscape}
-\begin{table}[H]
+\begin{table}
 \centering
@@
 \toprule
-\textbf{Disease}    & \textbf{Location}                & \textbf{Year}                 & \textbf{$\rr$} & \textbf{Lower} & \textbf{Upper} & \textbf{Source}                                                                                                                  \\ \midrule
+\textbf{Disease}    & \textbf{Location}                & \textbf{Year(s)}                 & \textbf{$\rr$} & \textbf{Lower} & \textbf{Upper} & \textbf{Source}                                                                                                                  \\ \midrule
 HIV                 & Belgium                          & 2004-2010                     & 1.10          & 1.09           & 1.16           & \cite{coelho2011}   \\
@@
 The parameters  $\beta$ and $\gamma$ have epidemiological meaning.
-Here, $\beta$ is the average infection rate and $\gamma$ is the average recovery rate.
+Here, $\beta$ is the inverse of expected time to infection rate and $\gamma$ is the inverse of the expected time to recovery.
 The movement of individuals from one compartment to another is represented through the ordinary differential equations below.
@@
 \end{align}
-In words, susceptible individuals become infected at a rate that is proportional to the percentage of infected individuals multiplied by $\beta$, the infection rate, and the number of susceptible individuals.
+In words, susceptible individuals become infected at a rate that is proportional to the percentage of infected individuals multiplied by $\beta$, the inverse of expected time to infection, and the number of susceptible individuals.
 Infectious individuals recover at a rate of $\gamma$ multiplied by the number of infected individuals.
@@
 That is,  the rate of new infections is greater than the rate of recovery.
-So as long as the number of initially susceptible individuals is large compared to the total population, $\frac{X}{N} \approx 1$, then an outbreak will occur if $\rr >1$,
+So as long as the number of initially susceptible individuals is large compared to the total population, $\frac{N}{X} \approx 1$, then an outbreak will occur if $\rr >1$,
 \begin{align*}
@@
   \end{align}
-We are assuming the observations are generated based on the discretizied time, deterministic ODEs presented in Equation \ref{eq:sir} with the addition of time ($t=0, 1, \dots, T$) and compartment dependent noise $\epsilon_{X,t}$ and $\epsilon_{Z,t}$.
+We are assuming the observations are generated based on the discretized time (replacing $d$ with $\Delta$), deterministic ODEs presented in Equation \ref{eq:sir} with the addition of time ($t=0, 1, \dots, T$) and compartment dependent noise $\epsilon_{X,t}$ and $\epsilon_{Z,t}$.
 Since, $N$, the total population, is constant, then $Y_{obs}$ is adjusted accordingly.
@@
 \begin{align*}
-  H_0:\;& \rr < 1 \\
-  H_A:\;& \rr \ge 1
+  H_0:\;& \rr > 1 \\
+  H_A:\;& \rr \le 1
 \end{align*}
@@
 \begin{align*}
-P(\hat{a} \le \rr \le \hat{b}) \le 1 - \alpha,  
-\end{align*}
-the second property is particularly desirable, as we can estimate $\hat{a}$ and $\hat{b}$ from the distribution $F$.
+P(\hat{a} \le \rr \le \hat{b}) \le 1 - \alpha.
+\end{align*}
+This property is particularly desirable, as we can estimate $\hat{a}$ and $\hat{b}$ from the distribution $F$.
 At any rate, we would like to be able to estimate $\hat{a}$ and $\hat{b}$ for a given $\alpha$-level.
@@
 
+\subsection{$\rr$ is a property of the model}\label{sec:prop}
+
+We provide an example of how $\rr$ is a property of the model.
+Consider the situation of data generated from an SEIR (XEYZ) model (described by \cite{cintronarias2009}), generated from the following ODEs,
+
+\begin{align*}
+  \frac{dX}{dt} &= - \frac{\beta XY}{N} \\
+  \frac{dE}{dt} &= \frac{\beta XY}{N}  - \alpha E\\
+  \frac{dY}{dt} &= \alpha E - \gamma Y \\
+  \frac{dZ}{dt} &= \gamma Y,
+\end{align*}
+again, with a constant population of $N$ and with known initival values $(X(0), E(0), Y(0), Z(t))$.
+The parameter $\alpha$ is the rate at which latent exposed individuals become infectious.
+As $\alpha \to \infty$, the XEYZ model converges to the XYZ model.
+The reproduction number for the XEYZ model, is the same as in the XYZ model, that is $\rr = \frac{\beta}{\gamma}$.
+
+Consider the case when $N=10000, (\beta, \gamma) = (.06, .03)$, $(X(0), E(0), Y(0), Z(t)) = (9500, 0, 500, 0)$ for times $t=1, \dots, 365$.
+Let the parameter $\alpha$ range between [0.005, 0.25] with step size $0.005$.
+This generates a first data set
+$$D_{\alpha, \textnormal{XEYZ}} = \left \{(X_{XEYZ}(t), E_{XEYZ}(t), Y_{XEYZ}(t), Z_{XEYZ}(t): t=1, \dots, 365\right \},$$
+which is observed by researcher 1.
+Now assume that researcher 2 misspecifies the model for this data as a XYZ model rather than a XEYZ model.
+That is, she observes all exposed individuals as susceptible and generates a second data set
+$$D_{\alpha, XYZ} = \left \{X_{XYZ}(t), Y_{XYZ}(t), Z_{XEYZ}(t) : t = 1, \dots, 365 \right \},$$
+where $X_{XYZ}(t) = X_{XEYZ}(t) + E_{XEYZ}(t)$, $Y_{XYZ}(t) = Y_{XEYZ}(t)$, and  $Z_{XEYZ}(t) = Z_{XYZ}(t)$. This situation is illustrated in Figure \ref{fig:sir-vs-seir} with $\alpha = 0.01$.
+
+\begin{figure}
+  \centering
+  \includegraphics[width=\textwidth]{images/seir-sir-data.pdf}
+  \caption{Top: XEYZ generated data over the course of a year.  Bottom: data tranformed from the XEYZ model to the XYZ model by combining all X and E compartments.}\label{fig:sir-vs-seir}
+  \end{figure}
+
+Both researchers want to estimate $\rr$ and so fit models to their data sets.
+The details of how the model are not so important at the moment (but will be in the later sections), but to fairly compare assume both researchers use least squares to fit the models.
+That is researcher 1 estimates $\rr$ with
+\begin{align*}
+  (\hat{\beta}_{XEYZ}, \hat{\gamma}_{XEYZ}, \hat{\alpha}_{XEYZ} )&=\arg \min_{\beta, \gamma, \alpha} \sum_{t} \left [ \left (X_{XEYZ}(t) - X(t)\right )^2  \right . \\
+                                                                 &+\left ( E_{XEYZ}(t) - E(t) \right )^2  \\
+  &\left . + \left ( Z_{XEYZ}(t) - Z(t) \right )^2\right ]
+\end{align*}
+Then the least squares (LS) estimate for $\rr^{XYZ}$ is given by Equation \ref{eq:sirls},
+\begin{align*}
+  \hat{\rr}^{XEYZ}= \frac{\hat{\beta}_{XEYZ}}{\hat{\gamma}_{XEYZ}}.
+\end{align*}
+Researcher 2 estimates $\rr$ with 
+\begin{align*}
+(\hat{\beta}_{XYZ}, \hat{\gamma}_{XYZ} )&=\arg \min_{\beta, \gamma} \sum_{t} \left [ \left (X_{XYZ}(t) - X(t)\right )^2 + \left ( Z_{XYZ}(t) - Z(t) \right )^2 \right ].
+\end{align*}
+Then the least squares (LS) estimate for $\rr^{XYZ}$ is given by Equation \ref{eq:sirls},
+\begin{align*}
+  \hat{\rr}^{XYZ}= \frac{\hat{\beta}_{XYZ}}{\hat{\gamma}_{XYZ}}.
+\end{align*}
+
+We plot the two estimates of $\rr$ as a function of $\alpha$ in Figure \ref{fig:r0-alpha}.
+We should note we also estimated the standard errors of the estimates of $\rr$, but they do not show up on the figure because they are of an order of magnitude of $10^{-4}$.
+Unsurprisingly, under the SEIR model speficiation, $\hat{\rr}=2$ because there is no noise in the data, and the data is generated from an SEIR model.
+However, the SIR model for small values of $\alpha$ can result in double the size of $\rr$!
+
+\begin{figure}
+  \centering
+  \includegraphics[width=\textwidth]{images/r0-est.pdf}
+  \caption{Estimates of $\rr$ from both the XEYZ and XYZ model specification for different values of $\alpha$ using Least Squares with $\beta=0.06$, $\gamma=0.03$, and $(X(0), E(0), Y(0)) = (9500, 500, 0)$.}\label{fig:r0-alpha}
+  \end{figure}
+
+For those readers not convinced by our small example, we offer this mathematical intuition of the next generation model (NGM) described by \cite{diekmann1990} and \cite{diekmann2009}.
+
+ The NGM is a generalization of any compartment model at the infection-free steady state.
+This model solves the problem of having an expression for $\rr$ in terms of the epidemiological parameters for a general compartment model.
+Originally, introduced by \citep{diekmann1990}, \cite{diekmann2009} posit a recipe to find $\rr$ for a wide class of compartment models, including SIR and SEIR.
+First, define the infected subsystem as ``the equations of the ODE system that describe the production of new infections and changes in state among infected individuals.''  Let $x = (C_1, C_2, \dots, C_m)$ where $C_i$ are the different compartments of the infected subsystem.
+The steps to find $\rr$ are as follows:
+
+\begin{enumerate}
+\item ``Linearize the infected subsystem of nonlinear ODEs about the infection-free steady state''
+\item Decompose the linearized infected subsystem into $(T + \Sigma )x$ where $T$ is a $m\times m$ matrix of transmissions and $\Sigma$ is a $m \times m$ matrix of transitions.
+\item $\rr$ is the spectral radius (i.e. the dominant eigenvalue) of $-T \Sigma^{-1}$.  
+\end{enumerate}
+
+Here, $T_{ij}$ is the rate of transmission of \textit{newly} infected individuals in state $i$ created by individuals in state $j$.
+$\Sigma_{ij}$ is the transition rate of individuals into compartment $i$ from compartment $j$.
+
+This method has the assumptions of a basic compartment model: homogeneous mixing among populations and the law of mass action, i.e., that the number of infected individuals are proportional to the number of infected individuals at the previous step.
+Additionally, this method is advantageous in that the matrix $-\Sigma^{-1}$ has intuitive meaning.
+The $ij$th entry of $- \Sigma^{-1}$ is the ``expected time that an individual who presently has state $j$ will spend in infected state $i$.''
+
+The NGM shows that $\rr$ is the spectral radius of a matrix composed from the transmission of disease from diseased individuals to susceptible individuals and transition of diseased individuals from among the compartments and is indeed a property of the model.
+
 \section{Overview of \wxxsir methods to estimate $\rr$ for the SIR model}
@@
 
-In this section, we describe the \wxxsir methods in detail, including advantages and disadvantages of each.
+In Section \ref{sec:prop} we described how $\rr$ is a property of the model.
+To fairly compare estimation methods to one another, we restrict our methods to estimate $\rr$ to the SIR model.
+Also in Section \ref{sec:prop}, we examined simulated data that was generated from a SEIR model without noise.
+However, we know that data is very noisy.
+In this section, we describe the \wxxsir methods to estimate $\rr$ for the SIR model in the presence of noise in detail, including advantages and disadvantages of each.
 
@@
 \end{align*}
-Then the least squares (LS) estimate for $\rr$ is given by Equation \ref{eq:sirls},
+Then the least squares (LS) estimate for $\hat{\beta}$ and $\hat{\gamma}$ is then combined to give an estimate of $\hat{\rr}$ in Equation \ref{eq:sirls},
 \begin{align}\label{eq:sirls}
@@
 We find
-\begin{align*}
+\begin{align}
 (\hat{\rr}, \hat{\gamma} ) &= \text{argmin}_{\rr, \gamma} \sum_{t} \left [ \left (X_{obs}(t) - X(t)\right )^2 + \left ( Z_{obs}(t) - Z(t) \right )^2 \right ]
-\end{align*}
+\end{align}
 We use the $\hat{\rr}$ directly from the above estimation problem, which again can be found with a grid search or another optimization process.
+
+We can also write $\hat{\rr}$ as a function of $\gamma^*$,
+\begin{align*}
+\hat{\rr}(\gamma^*)&= \text{argmin}_{\rr} \sum_{t} \left [ \left (X_{obs}(t) - X(t; \gamma^*)\right )^2 + \left ( Z_{obs}(t) - Z(t; \gamma^*) \right )^2 \right ].
+  \end{align*}
+
 It may be surprising that this method leads to different results than simply using least squares.
-One reason we obtain different results may be attributed to better numerical precision as there is one fewer step to obtain an estimate for $\rr$.
-However, we also see that this method further results in different standard errors than in LS, which we discuss further in Section \ref{sec:results}.
+We also see that this method further results in different standard errors than in LS, which we discuss further in Section \ref{sec:results}.
 
@@
 \end{align*}
-Following,  an estimator for \(\rr\) is derived from the ODEs in Equation \eqref{eq:sir},
+An estimator for $\rr$ is derived from the ODEs in Equation \eqref{eq:sir},
 \begin{align}
@@
   \end{align}
-  Here, $K$ is arbitrary and should be selected using some criterion such as AIC.
+  Here, $K$ is arbitrary and should be selected using some criterion such as AIC (REF).
 Besides optionally deciding on the degree of polynomials to fit, this model is simple to implement and gives comparable results to using least squares with the SIR model.
@@
 The above formulation (Section \ref{linear-model-approximation-degree-10}) of a linear model approximation only uses the estimate at time $t=0$ to estimate $\rr$.
-We can instead, use all time points $O$ available to estimate $\rr$.
+We can instead, use all time points $T$ available to estimate $\rr$.
 We fit a polynomial in \(t\) with degree \(K\) to \(X_{obs}\)
@@
 \begin{align*}
-  \hat{\rr} &= \frac{1}{O} \sum_t \frac{-\hat{X}^\prime(t)}{\hat{Z}^\prime(t)} \cdot \frac{N}{X(0)} 
+  \hat{\rr} &= \frac{1}{T} \sum_t \frac{-\hat{X}^\prime(t)}{\hat{Z}^\prime(t)} \cdot \frac{N}{X(0)} 
 \end{align*}
@@
 An advantage to this approach is that we have a more robust estimate of $\rr$ than just using one time point.
-Like in LMA, we can examine how sensitive $\rr$ is to the order of the polynomial $K$.
+As in LMA, we can examine how sensitive $\rr$ is to the order of the polynomial $K$.
 
@@
 \begin{align*}
-\rr &= \hat{\textnormal{IPR}}(t) \cdot D
+\rr &= \hat{\textnormal{IPR}}(t) \cdot \frac{1}{\gamma}
 \end{align*}
@@
 One, the estimate of $\rr$ is highly interpretable.
-Two, an assumption of Gaussian noise in Equation \ref{eq:harko_lin} is not as strong as an assumption in the other methods to estimate $\rr$ because we are looking at a log transformation of the the percent of suceptibles instead of raw variables.
-If the assumption holds, we can show that the resulting estimate of $\rr$ is an unbiased and a consistent estimator.
+Two, an assumption of Gaussian noise in Equation \ref{eq:harko_lin} is not as strong as an assumption in the other methods to estimate $\rr$ because we are looking at a log transformation of the percent of suceptibles instead of raw variables.
+If the assumption holds, we can show that the resulting estimate of $\rr$ is unbiased and consistent.
 Finally,  this method is easily implemented using any linear modeling software.
@@
 We have explicitly stated how randomness enters the model so it is up to the researcher to check whether this assumption is a good fit.
-Using this method, it is possible to simulate how sensitive $\hat{\rr}$ is to changes in $\hat{\alpha}$.
+Using this method, it is easy to see how sensitive $\hat{\rr}$ is to changes in $\hat{\alpha}$ as $\hat{\rr}$ is a function of $\alpha$.
 
@@
 \begin{align*}
-J(t+1)  \sim \textnormal{Poisson}( J(t) \exp \left \{  \gamma (\rr-1)\right \})
-\end{align*}
-Then, the posterior distribution of $\rr$ given the previous days' incidence is
+J(t+1) | \rr  \sim \textnormal{Poisson}( J(t) \exp \left \{  \gamma (\rr-1)\right \})
+\end{align*}
+Then, the posterior distribution of $\rr$ given the previous time period's incidence is
 \begin{align*}
@@
 \end{align*}
-This method is sequential in that the prior distribution for $\rr$ comes from the previous day.
+This method is sequential in that the prior distribution for $\rr$ comes from the previous time period.
 The initial prior for $\rr$ is assumed to be flat.
@@
 
+Although, the Poisson distribution shown above is often used to estimate $\rr$, we have found that this model does not perform well in practice (see Section \ref{sec:results}).
+Thus, we would recommend using a more sophisitcated Bayesian model such as the one in \cite{bauer2017}.
+
 \section{Confidence and Credible Intervals}
@@
 When the method estimates \(\beta\) and \(\gamma\) instead of \(\rr\) directly, we use the delta method approximation to estimate the
-variance of \(\rr\).
-Here, we know \(\rr = h(\beta, \gamma) = \frac{\beta}{\gamma}\).
-Then \(\bigtriangleup h = (\frac{1}{\gamma},  -\frac{\beta}{\gamma^2})^T\) and \(V[\rr] = \bigtriangleup h^T \Sigma_{\beta, \gamma} \bigtriangleup h\), where \(\Sigma_{\beta, \gamma}\) is covariance matrix of \(\beta\) and \(\gamma\).
+variance of \(\rr\).   From \cite{wasserman2004}, if
+\begin{align*}
+  \sqrt{T} \left ( \hat{\mathbf{\theta}}_o - \mathbf{\theta} \right ) \overset{d}{\to} N\left ( \mathbf{0}, \Sigma \right  )
+\end{align*}
+and $h$ is a continuously differentiable function then
+\begin{align*}
+  \sqrt{T} \left (h(\hat{\theta}_n) - h(\theta) \right ) \overset{d}{\to} N \left (0,  \nabla^T h(\theta)\Sigma \nabla h(\theta)\right ),
+\end{align*}
+where $\nabla h$ is the gradient of $h$.
+
+For the method LS, we know $\rr = h(\beta, \gamma) = \frac{\beta}{\gamma}$.
+Then, if $E\left [\frac{\hat{\beta}}{\hat{\gamma}}\right] \to \frac{\beta}{\gamma}$, then  $(\nabla h = (\frac{1}{\gamma},  -\frac{\beta}{\gamma^2})^T)$ and
+\begin{align*}
+  V[\rr] &\to \nabla h^T(\beta, \gamma) \Sigma_{\beta, \gamma} \nabla h(\beta, \gamma) \\
+  &= \frac{\Sigma_{\beta, \beta}}{\gamma^2} - \frac{\Sigma_{\beta, \gamma}(1 + \beta)}{\gamma^3} + \frac{\beta \Sigma_{\gamma, \gamma}}{\gamma^4}
+\end{align*}
+where $\Sigma_{\beta, \gamma}$ is covariance matrix of \(\beta\) and \(\gamma\).
+.
+We can use plug-in estimators for $\hat{\beta}$, $\hat{\gamma}$, and $\hat{\Sigma}$ where $\hat{\Sigma}$ is taken to the inverse of the Hessian from the least squares optimzation of $\beta$ and $\gamma$.
+
+For ReLS, we use the estimate of $\Sigma_{\rr, \gamma}$ taken from the optmization of $\rr$ and $\gamma$ as to find our estimate for $\hat{\Sigma}_{\rr}$.
 
@@
 This method may be extended to other frameworks such as the SEIR model.
-The advantages of the delta method are that since the Next Generation Matrix \citep{diekmann2009} gives recipe for deriving $\rr$ from the deterministic ODEs, then it is theoretically possible to use the delta method to derive an expression for the variance of $\rr$ for any compartment model.
+The advantages of the delta method are that since the Next Generation Matrix \citep{diekmann2009} gives a recipe for deriving $\rr$ from the deterministic ODEs, then it is theoretically possible to use the delta method to derive an expression for the variance of $\rr$ for any compartment model.
+
+We can also use the Delta method to derive variance estimates of LMA, LMAT, LL, and MC as for all these estimates first estimate parameters $\theta$ since $\rr= h(\theta)$
 
@@
 \begin{align*}
-  V\left [ \hat{\rr} \right ] &\approx V\left [\hat{\rr}_\ell \right ].
+  V\left [ \hat{\rr} \right ] &\approx V\left [\hat{\rr}_\ell \right ], \; \ell = 1, \dots, L
 \end{align*}
@@
 
-A choice one must make in the block bootstrap is the block size $b$.  One simple recommendation is to take $b= n^{1/3}$, which is derived from the asymptotic mean square error of the method.
+A choice one must make in the block bootstrap is the block size $b$.
+One simple recommendation is to take $b= Cn^{1/3}$, which is derived from the asymptotic mean square error of the method.
+\cite{hall1995} describe how to calculate $C$, which is context dependent.
 
@@
 Then commonly, $\hat{\rr} = \hat{F}_{.5}$, the median of $\hat{F}$.
-Correspondingly, the 95\% credible interval is $\left[\hat{F}_{.05}, \hat{F}_{.95} \right ]$.
+Correspondingly, the 95\% credible interval is $\left[\hat{F}_{.025}, \hat{F}_{.975} \right ]$.
 Advantages of this method include having an entire distribution instead of a point estimate.
@@
 One should note that these 95\% credible intervals are not the same as confidence intervals since the ``true'' value of $\rr$ may not be contained in the credible intervals 95\% of the time.
+We only use the posterior distribution to form a CI for the SB method, as it is the only Bayesian method presented here.
 
@@
 
-	Number of time steps $O$: 365
+	Number of time steps $T$: 365
 
@@
       \end{center}
-
+      The baseline data set is generated from the deterministic SIR curves in Figure \ref{fig:baseline-data} and then noise is added to two of the compartments, and the third is adjusted to maintain a constant population value.
+In this scenario, we see that the infection begins slowly but over the course of a year infects over 75\% of the population.
+      \begin{figure}
+        \centering
+        \includegraphics[width=.8\textwidth]{images/baseline_sir.pdf}
+        \caption{SIR (XYZ) curves for the deterministic model with $N=100,000$, $(\beta, \gamma) = (0.06, 0.03)$, $(X(0), Y(0), Z(0))= (99,950, 50,0)$  The baseline data is generated by using this model and adding noise to the $X$ and $Y$ compartments and adjusting $Z$ at each time step to maintain a constant population.}\label{fig:baseline-data}
+        \end{figure}
       This baseline data set is chosen due to its moderate $\rr$ size of 2, which indicates an outbreak of an epidemic similar to HIV, ZIK-V, EVD, and H1N1 pandemic influenza (see Table \ref{tab:r0-real-ex}).
 The infection rate and recovery rate ($\beta, \gamma$) are relatively small.
-The duration of the disease, under the true model, spans nearly a whole year (see Figure \ref{fig:baseline}) and so we use a year's worth of daily observations of (X,Y, Z) values.
+The duration of the disease, under the true model, spans nearly a whole year (see Figure \ref{fig:baseline-data}) and so we use a year's worth of daily observations of (X,Y, Z) values.
 Initially, 50 (0.05\%) diseased individuals are introduced to the population of 100,000 individuals (approximately the average population size of a US county in 2015).
@@
 \toprule
-\textbf{Name} & \textbf{Model} & \textbf{Error} & \textbf{($\beta, \gamma$)} & \textbf{O} & \textbf{(X(0),Y(0))} & \textbf{N} & \textbf{($\sigma_X, \sigma_Y$)} \\ \midrule
+\textbf{Name} & \textbf{Model} & \textbf{Error} & \textbf{($\beta, \gamma$)} & \textbf{T} & \textbf{(X(0),Y(0))} & \textbf{N} & \textbf{($\sigma_X, \sigma_Y$)} \\ \midrule
   Baseline        & SIR      &   Norm             & (.06, .03)                          &     365       &  (99950, 50)                    &  $10^5$          &(100, 5)                                \\ \midrule
@@
 There are numerous aspects to consider when using a method to estimate $\rr$.
-Among these include values of both the infection and recovery rate $(\beta, \gamma)$, the number of data points available ($O$), the total size of the population ($N$), the initial percent of susceptible individuals $\left (\frac{X(0)}{N}\right)$, the initial percent of infectious individuals $\left (\frac{Y(0)}{N}\right )$, the magnitude of the variance of both the number of susceptibles and the number of infectious ($\sigma_X^2, \sigma_Y^2$), the error modeling assumptions, data generated from a non-SIR model, and finally methods of estimating the variance.
+Among these include values of both the infection and recovery rate $(\beta, \gamma)$, the number of data points available ($T$), the total size of the population ($N$), the initial percent of susceptible individuals $\left (\frac{X(0)}{N}\right)$, the initial percent of infectious individuals $\left (\frac{Y(0)}{N}\right )$, the magnitude of the variance of both the number of susceptibles and the number of infectious ($\sigma_X^2, \sigma_Y^2$), the error modeling assumptions, data generated from a non-SIR model, and finally methods of estimating the variance.
 We first examine our baseline data, and then we examine each of these aspects individually.
@@
   \includegraphics[scale=0.5]{images/BaseBase.jpeg}
-  \caption{Forest plots of estimates of $\rr$ for the \xxsir methods for the baseline data set $(\beta=.06, \gamma=.03, O=365, X(0)=99950, Y(0)=50, Z(0)=0, \sigma_X=100, \sigma_Y=5, N=10^5)$.
+  \caption{Forest plots of estimates of $\rr$ for the \xxsir methods for the baseline data set $(\beta=.06, \gamma=.03, T=365, X(0)=99950, Y(0)=50, Z(0)=0, \sigma_X=100, \sigma_Y=5, N=10^5)$.
 The points are the estimates of $\rr$ and the lines denote $\pm 2\cdot $Std.
@@
 		\hline
-		LS & 1.9999711 & 0.0055923\\
-		\hline
-		ReLS & 1.9997340 & 0.0049939\\
-		\hline
-		SIRLMA & 1.6863640 & 0.1886132\\
-		\hline
-		LMAT & 1.4063227 & 0.6309170\\
-		\hline
-		IPR & 4.4203632 & 12.3593415\\
-		\hline
-		SIPR & 1.8870824 & $<$ 1e-07 \\
-		\hline
-		LL & 1.9998072 & 0.0002198\\
-		\hline
-		MC & 0.9485555 &  $<$ 1e-07 \\
-		\hline
-		SB & 1.6495023 & 0.0672249\\
-		\hline
-	\end{tabular}
-        \caption{Baseline Data $(\beta=.06, \gamma=.03, O=365, X(0)=99950, Y(0)=50, Z(0)=0, \sigma_X=100, \sigma_Y=5, N=10^5)$, $\rr$ Estimates and Standard Errors}\label{tab:baseline-res}
+		LS & 1.9999711 & 0.0056\\
+		\hline
+		ReLS & 1.9997340 & 0.0050\\
+		\hline
+		LMA & 1.6863640 & 0.1886\\
+		\hline
+		LMAT & 1.4063227 & 0.6309\\
+		\hline
+		IPR & 4.4203632 & 12.3593\\
+		\hline
+		SIPR & 1.8870824 & $<$ 1e-04 \\
+		\hline
+		LL & 1.9998072 & 0.0002\\
+		\hline
+		MC & 0.9485555 &  $<$ 1e-04 \\
+		\hline
+		SB & 1.6495023 & 0.0672\\
+		\hline
+	\end{tabular}
+        \caption{Baseline Data $(\beta=.06, \gamma=.03, T=365, X(0)=99950, Y(0)=50, Z(0)=0, \sigma_X=100, \sigma_Y=5, N=10^5)$, $\rr$ Estimates and Standard Errors}\label{tab:baseline-res}
 \end{table}
@@
   \includegraphics[scale=0.5]{images/AR.jpeg}
-  \caption{Forest plots of estimates of $\rr$ for the \xxsir methods for the autoregressive (AR) ($\rho=.7)$ simulated data set $(\beta=.06, \gamma=.03, O=365, X(0)=99950, Y(0)=50, Z(0)=0, \sigma_X=100, \sigma_Y=5, N=10^5)$.
+  \caption{Forest plots of estimates of $\rr$ for the \xxsir methods for the autoregressive (AR) ($\rho=.7)$ simulated data set $(\beta=.06, \gamma=.03, T=365, X(0)=99950, Y(0)=50, Z(0)=0, \sigma_X=100, \sigma_Y=5, N=10^5)$.
 The points are the estimates of $\rr$ and the lines denote $\pm 2\cdot $Std.
@@
 		\hline
-		LS & 1.9997387 & 0.0071509\\
-		\hline
-		ReLS & 2.0000460 & 0.0063859\\
-		\hline
-		SIRLMA & 1.4744605 & 0.1130753\\
-		\hline
-		LMAT & 1.3694249 & 0.8681724\\
-		\hline
-		IPR & 1.7126909 & 5.0713852\\
-		\hline
-		SIPR & 0.8792348 & $<$ 1e-07 \\
-		\hline
-		LL & 2.0000004 & 0.0002885\\
-		\hline
-		MC & 0.9485555 & $<$ 1e-07 \\
-		\hline
-		SB & 1.0000000 & 0.0523424\\
+		LS & 1.9997 & 0.0072\\
+		\hline
+		ReLS & 2.0001 & 0.0063\\
+		\hline
+		LMA & 1.4745 & 0.1130\\
+		\hline
+		LMAT & 1.3694 & 0.8682\\
+		\hline
+		IPR & 1.7127 & 5.0713\\
+		\hline
+		SIPR & 0.8792 & $<$ 1e-04 \\
+		\hline
+		LL & 2.0000 & 0.0003\\
+		\hline
+		MC & 0.9486 & $<$ 1e-04 \\
+		\hline
+		SB & 1.0000 & 0.0523\\
 		\hline
@@
 	\includegraphics[scale=0.5]{images/mono.jpeg}
-	\caption{Forest plots of estimates of $\rr$ for the \xxsir methods for the Norm-M and AR-M data sets $(\beta=.06, \gamma=.03, O=365, X(0)=99950, Y(0)=50, Z(0)=0, \sigma_X=100, \sigma_Y=5, N=10^5)$.
+	\caption{Forest plots of estimates of $\rr$ for the \xxsir methods for the Norm-M and AR-M data sets $(\beta=.06, \gamma=.03, T=365, X(0)=99950, Y(0)=50, Z(0)=0, \sigma_X=100, \sigma_Y=5, N=10^5)$.
 The points are the estimates of $\rr$ and the lines denote $\pm 2\cdot $Std.
@@
 		\hline
-		LS & 2.0002458 & 0.0044456 & 1.9999920 & 0.0073664\\
-		\hline
-		ReLS & 2.0003143 & 0.0039684 & 1.9996837 & 0.0065745\\
-		\hline
-		SIRLMA & 1.6793591 & 0.1102722 & 1.5349785 & 0.1496989\\
-		\hline
-		LMAT & 1.4422630 & 0.9241642 & 1.4491091 & 0.7463529\\
-		\hline
-		IPR & 0.7485576 & 1.5046827 & 0.7513498 & 1.5606142\\
-		\hline
-		SIPR & 0.9081552 & 0.0000000 & 0.6976176 & 0.0000000\\
-		\hline
-		LL & 2.0003027 & 0.0001342 & 1.9995563 & 0.0002342\\
-		\hline
-		MC & 0.9485555 & 0.0000000 & 0.9485555 & 0.0000000\\
-		\hline
-		SB & 1.0000000 & 0.0523424 & 1.7143505 & 0.0685336\\
+		LS & 2.0002 & 0.0044 & 1.9999 & 0.0074\\
+		\hline
+		ReLS & 2.0003 & 0.0040 & 1.9997 & 0.0065\\
+		\hline
+		LMA & 1.6794 & 0.1103 & 1.5350 & 0.1497\\
+		\hline
+		LMAT & 1.4422 & 0.9242 & 1.4491 & 0.7464\\
+		\hline
+		IPR & 0.7486 & 1.5047 & 0.7513 & 1.5606\\
+		\hline
+		SIPR & 0.9082 & $<$ 1e-04 & 0.6976 & $<$ 1e-04\\
+		\hline
+		LL & 2.0003 & 0.0001 & 1.9996 & 0.0002\\
+		\hline
+		MC & 0.9486 & $<$ 1e-04 & 0.9486 & $<$ 1e-04\\
+		\hline
+		SB & 1.0000 & 0.0523 & 1.7143505 & 0.0685\\
 		\hline
@@
 
-We next examine the simulated data sets InfRec1-4, three of which have different levels of $\beta$ for a given level of $\gamma=.001$ and the fourth is simulated with $\rr=0.25 < 1$, a case where an outbreak would not occur.
+We next examine the simulated data sets InfRec1-4, three of which have different levels of $\gamma$ for a given level of $\beta = 0.06$ and the fourth is simulated with $\rr=0.25 < 1$, a case where an outbreak would not occur.
 The idea behind this is that since $\rr=\frac{\beta}{\gamma}$ and ratio estimators typically have large variance, we want to examine how a small value of $\gamma$, that is a small recovery rate, would effect the resulting estimates of $\rr$.
@@
 
+\subsubsection{Gaussian Errors}
+
 \begin{figure}[H]
   \centering
-  \includegraphics[scale=0.5]{images/parchange.jpeg}
-  \caption{Forest plots of estimates of $\rr$ for the \xxsir methods for the infrec1-4 data sets $(\beta=.06, \gamma=.03, O=365, X(0)=99950, Y(0)=50, Z(0)=0, \sigma_X=100, \sigma_Y=5, N=10^5)$.
+  \includegraphics[scale=0.5]{images/parchange_n.jpeg}
+  \caption{Forest plots of estimates of $\rr$ for the \xxsir methods for the infrec1-4 data sets $(\beta=.06, \gamma=.03, T=365, X(0)=99950, Y(0)=50, Z(0)=0, \sigma_X=100, \sigma_Y=5, N=10^5)$.
 The points are the estimates of $\rr$ and the lines denote $\pm 2\cdot $Std.
@@
 		\hline
-		SIRLS & 59.9764682 & 0.0434705\\
-		\hline
-		ReLS & 59.9951876 & 0.4343984\\
-		\hline
-		SIRLMA &  -117.9694856 & 101.1885824\\
-		\hline
-		LMAT & -2.8535930 & 193.6526107 \\
-		\hline
-		IPR & 54.146506 & 200.2595\\
-		\hline
-		SIPR & 3.9645187 & $<$ 1e-07\\
-		\hline
-		LL & 46.3209596 & 0.7797216\\
-		\hline
-		MC & 0.9485554 & 0.0000001\\
-		\hline
-		SB & 0.9769731 & 0.0517362\\
+		SIRLS & 59.9765 & 0.0435\\
+		\hline
+		ReLS & 59.9952 & 0.4344\\
+		\hline
+		LMA &  -117.9695 & 101.1886 \\
+		\hline
+		LMAT & -2.8536 & 193.6526 \\
+		\hline
+		IPR & 54.1465 & 200.2595\\
+		\hline
+		SIPR & 3.9645 & $<$ 1e-04\\
+		\hline
+		LL & 46.3210 & 0.7797\\
+		\hline
+		MC & 0.9486 & $<$ 1e-04 \\
+		\hline
+		SB & 0.9767 & 0.0517\\
 		\hline
@@
 
-Looking in Figure \ref{fig:infrec1-res} and Table \ref{tab:infrec1-res} at the cases where $\gamma = 0.001$ with $\beta=60000, 1500$, and 1000, it appears that most of the models do not handle a small $\gamma$ well.
+Looking in Figure \ref{fig:infrec1-res} and Table \ref{tab:infrec1-res} at the cases where $\beta = 0.06$ with $\gamma = 0.001, 0.04, 0.01, 0.24$, it appears that most of the models do not handle a small $\gamma$ well.
 We examine an exaggerated extreme case of when $\rr=60$ (the largest reported $\rr$ value we found is $\hat{\rr}=14$ for measles as reported in \cite{anderson1992}).
@@
 
-When $\rr=1$, we are unsure whether there will be an epidemic or not.
-In this case, all the CIs of the methods cover the true value.
-We also see that the SE of LS is much greater than ReLS, which may be surprising.
-((WHY?))
+When $\rr=1$, we are unsure whether there will be an epidemic or not.  In this case, all the CIs of the methods cover the true value.  We also see that the SE of LS is much greater than ReLS.
 
@@
 		\hline
-		LS & 0.2363274 & 4.1042813\\
-		\hline
-		ReLS & 4.1795252 & 2.7455569\\
-		\hline
-		SIRLMA &  0.7074613 & 0.2028765 \\
-		\hline
-		LMAT & 0.6746914 & 6.2471252 \\
-		\hline
-		IPR & 181.015802 & 742.1088 \\
-		\hline
-		SIPR & -0.2484284 & 0.1159769 \\
-		\hline
-		LL & 0.9912397 & 0.0030505\\
-		\hline
-		MC & 0.9485811 & 0.0000150\\
-		\hline
-		SB & 1.2354952 & 0.0581801\\
+		LS & 0.2363 & 4.1043 \\
+		\hline
+		ReLS & 4.1795 & 2.7456\\
+		\hline
+		LMA &  0.7075 & 0.2029 \\
+		\hline
+		LMAT & 0.6747 & 6.2471 \\
+		\hline
+		IPR & 181.0158 & 742.1088 \\
+		\hline
+		SIPR & -0.2484 & 0.1160 \\
+		\hline
+		LL & 0.9912 & 0.0031\\
+		\hline
+		MC & 0.9486 & 0.0001\\
+		\hline
+		SB & 1.2355 & 0.0582\\
 		\hline
@@
 
-\subsection{Number of Time Points ($O$)}\label{sec:res-time}
+\subsubsection{Gaussian Monotonic Errors}
+
+\begin{figure}[H]
+	\centering
+	\includegraphics[scale=0.5]{images/parchange_nm.jpeg}
+	\caption{Forest plots of estimates of $\rr$ for the \xxsir methods for the infrec1-4 data sets $(\beta=.06, \gamma=.03, T=365, X(0)=99950, Y(0)=50, Z(0)=0, \sigma_X=100, \sigma_Y=5, N=10^5)$.
+The points are the estimates of $\rr$ and the lines denote $\pm 2\cdot $Std.
+Err.
+The vertical line is the true $\rr$ value.}
+\end{figure}
+\begin{table}[H]
+
+	\centering
+	\begin{tabular}[t]{l|r|r}
+		\hline
+		Model & Estimate & Std. Dev\\
+		\hline
+		LS & 59.9765 & 0.0387\\
+		\hline
+		ReLS & 59.9952 & 0.3875\\
+		\hline
+		LMA & -55.3749 & 10.9872\\
+		\hline
+		LMAT & 7.7686 & 103.2206\\
+		\hline
+		IPR & 24.4087 & 66.1668\\
+		\hline
+		SIPR & 3.8285 & $<$ 1e-04\\
+		\hline
+		LL & 59.8300 & 0.1079\\
+		\hline
+		MC & 0.9486 & $<$ 1e-04\\
+		\hline
+		SB & 1.0000 & 0.0523\\
+		\hline
+	\end{tabular}
+	\caption{$\gamma = 0.001$, $\rr$ Estimates and Standard Errors}
+\end{table}
+
+\begin{table}[H]
+
+	\centering
+	\begin{tabular}[t]{l|r|r}
+		\hline
+		Model & Estimate & Std. Dev\\
+		\hline
+		LS & 0.2328 & 0.6701\\
+		\hline
+		ReLS & 3.5710 & 0.3171\\
+		\hline
+		LMA & 0.5217 & 0.1716\\
+		\hline
+		LMAT & 0.3872 & 4.8818\\
+		\hline
+		IPR & 6.6728 & 89.9362\\
+		\hline
+		SIPR & -0.0317 & 0.0336\\
+		\hline
+		LL & 0.8623 & 0.0088\\
+		\hline
+		MC & 0.9486 & $<$ 1e-04\\
+		\hline
+		SB & 1.5587 & 0.0653\\
+		\hline
+	\end{tabular}
+	\caption{$\gamma = 0.24$, $\rr$ Estimates and Standard Errors}
+\end{table}
+
+\subsubsection{Autoregressive Errors}
+
+\begin{figure}[H]
+	\centering
+	\includegraphics[scale=0.5]{images/parchange_ar.jpeg}
+	\caption{Forest plots of estimates of $\rr$ for the \xxsir methods for the infrec1-4 data sets $(\beta=.06, \gamma=.03, T=365, X(0)=99950, Y(0)=50, Z(0)=0, \sigma_X=100, \sigma_Y=5, N=10^5)$.
+The points are the estimates of $\rr$ and the lines denote $\pm 2\cdot $Std.
+Err.
+The vertical line is the true $\rr$ value.}
+\end{figure}
+\begin{table}[H]
+
+	\centering
+	\begin{tabular}[t]{l|r|r}
+		\hline
+		Model & Estimate & Std. Dev\\
+		\hline
+		LS & 60.0151 & 0.0698\\
+		\hline
+		ReLS & 60.0255 & 0.6912\\
+		\hline
+		LMA & -19.2058 & 5.9751\\
+		\hline
+		LMAT & 76.3404 & 472.4155\\
+		\hline
+		IPR & 58.7237 & 219.9891\\
+		\hline
+		SIPR & 3.9056 & $<$ 1e-04\\
+		\hline
+		LL & 41.8977 & 0.9768\\
+		\hline
+		MC & 0.9486 & $<$ 1e-04\\
+		\hline
+		SB & 1.0000 & 0.0523\\
+		\hline
+	\end{tabular}
+	\caption{$\gamma = 0.001$, $\rr$ Estimates and Standard Errors}
+\end{table}
+
+\begin{table}[H]
+
+	\centering
+	\begin{tabular}[t]{l|r|r}
+		\hline
+		Model & Estimate & Std. Dev\\
+		\hline
+		LS & -0.0695 & 2.8291\\
+		\hline
+		ReLS & 5.1930 & 3.5250\\
+		\hline
+		LMA & 0.8750 & 0.0591\\
+		\hline
+		LMAT & 1.7865 & 15.9420\\
+		\hline
+		IPR & 52.5000 & 740.0182\\
+		\hline
+		SIPR & -2.8643 & $<$ 1e-04\\
+		\hline
+		LL & 1.0014 & 0.0025\\
+		\hline
+		MC & 0.9486 & $<$ 1e-04\\
+		\hline
+		SB & 1.0160 & 0.0528\\
+		\hline
+	\end{tabular}
+	\caption{$\gamma = 0.24$, $\rr$ Estimates and Standard Errors}
+\end{table}
+
+\subsubsection{Autoregressive Monotonic Errors}
+
+\begin{figure}[H]
+	\centering
+	\includegraphics[scale=0.5]{images/parchange_arm.jpeg}
+	\caption{Forest plots of estimates of $\rr$ for the \xxsir methods for the infrec1-4 data sets $(\beta=.06, \gamma=.03, T=365, X(0)=99950, Y(0)=50, Z(0)=0, \sigma_X=100, \sigma_Y=5, N=10^5)$.
+The points are the estimates of $\rr$ and the lines denote $\pm 2\cdot $Std.
+Err.
+The vertical line is the true $\rr$ value.}
+\end{figure}
+\begin{table}[H]
+
+	\centering
+	\begin{tabular}[t]{l|r|r}
+		\hline
+		Model & Estimate & Std. Dev\\
+		\hline
+		LS & 59.9589 & 0.0631\\
+		\hline
+		ReLS & 59.9961 & 0.6215\\
+		\hline
+		LMA & -183.9423 & 277.8292\\
+		\hline
+		LMAT & -69.9068 & 1603.7792\\
+		\hline
+		IPR & 28.5231 & 120.9288\\
+		\hline
+		SIPR & 3.7113 & $<$ 1e-04\\
+		\hline
+		LL & 70.7076 & 0.9517\\
+		\hline
+		MC & 0.9486 & $<$ 1e-04\\
+		\hline
+		SB & 1.0000 & 0.0523\\
+		\hline
+	\end{tabular}
+	\caption{$\gamma = 0.001$, $\rr$ Estimates and Standard Errors}
+\end{table}
+
+\begin{table}[H]
+
+	\centering
+	\begin{tabular}[t]{l|r|r}
+		\hline
+		Model & Estimate & Std. Dev\\
+		\hline
+		LS & 0.3021 & 0.5657\\
+		\hline
+		ReLS & 9.8415 & 1.7074\\
+		\hline
+		LMA & -49.7813 & 5303.0302\\
+		\hline
+		LMAT & 0.2221 & 5.4112\\
+		\hline
+		IPR & 0.8145 & 11.8709\\
+		\hline
+		SIPR & 0.7365 & 0.0675\\
+		\hline
+		LL & 0.9199 & 0.0053\\
+		\hline
+		MC & 0.9486 & $<$ 1e-04\\
+		\hline
+		SB & 1.0000 & 0.0523\\
+		\hline
+	\end{tabular}
+	\caption{$\gamma = 0.24$, $\rr$ Estimates and Standard Errors}
+\end{table}
+
+\subsection{Number of Time Points ($T$)}\label{sec:res-time}
 We examine how the number of time steps the estimates of $\rr$ in the different methods.
@@
 Again, we see that IPR does very poorly in estimating $\rr$, regardless of the amount of data observed.
+
+\subsubsection{Gaussian Errors}
+
 \begin{figure}[H]
   \centering
-  \includegraphics[scale=0.5]{images/time.jpeg}
-  \caption{Forest plots of estimates of $\rr$ for the \xxsir methods for the Time1-4 data sets $(\beta=.06, \gamma=.03, O=365, X(0)=99950, Y(0)=50, Z(0)=0, \sigma_X=100, \sigma_Y=5, N=10^5)$.
+  \includegraphics[scale=0.5]{images/time_n.jpeg}
+  \caption{Forest plots of estimates of $\rr$ for the \xxsir methods for the Time1-4 data sets with Gaussian errors $(\beta=.06, \gamma=.03, T=365, X(0)=99950, Y(0)=50, Z(0)=0, \sigma_X=100, \sigma_Y=5, N=10^5)$.
 The points are the estimates of $\rr$ and the lines denote $\pm 2\cdot $Std.
@@
 		\hline
-		LS & 1.9993392 & 0.0286060\\
-		\hline
-		ReLS & 1.9994284 & 0.0312814\\
-		\hline
-		SIRLMA & 0.9496498 & 0.0825419\\
-		\hline
-		LMAT & 3.3442696 & 21.4895422\\
-		\hline
-		IPR & 7.3458131 & 16.1288906\\
-		\hline
-		SIPR & -2.1198897 & $<$ 1e-07\\
-		\hline
-		LL & 2.0014146 & 0.0017620\\
-		\hline
-		MC & 0.9485555 & 0.0000001\\
-		\hline
-		SB & 2.4289382 & 0.1102030\\
+		LS & 1.9993 & 0.0286\\
+		\hline
+		ReLS & 1.9994 & 0.0313\\
+		\hline
+		LMA & 0.9496 & 0.0825\\
+		\hline
+		LMAT & 3.3443 & 21.4895\\
+		\hline
+		IPR & 7.3458 & 16.1289\\
+		\hline
+		SIPR & -2.1199 & $<$ 1e-04\\
+		\hline
+		LL & 2.0014 & 0.0018\\
+		\hline
+		MC & 0.9486 & $<$ 1e-04\\
+		\hline
+		SB & 2.4289 & 0.1102\\
 		\hline
@@
 		\hline
-		LS & 38.6895520 & 106.2705193\\
-		\hline
-		ReLS & 22.9645867 & 6.9816713\\
-		\hline
-		SIRLMA & 1.4418031 & 7.3163878\\
-		\hline
-		LMAT & 1.0078807 & 0.1424793\\
-		\hline
-		IPR & 25.7313065 & 35.2926077\\
-		\hline
-		SIPR & 4.1772451 & 7.3202970\\
-		\hline
-		LL & 1.0429764 & 0.0276261\\
-		\hline
-		MC & 0.9485556 & 0.0000005\\
-		\hline
-		SB & 1.0000000 & 0.2236068\\
-		\hline
-	\end{tabular}
-        \caption{ $T$ = 20, $\rr$ Estimates and Standard Errors}\label{tab:time-res2}
+		LS & 38.6896 & 106.2705\\
+		\hline
+		ReLS & 22.9646 & 6.9817\\
+		\hline
+		LMA & 1.4418 & 7.3164\\
+		\hline
+		LMAT & 1.0079 & 0.1425\\
+		\hline
+		IPR & 25.7313 & 35.2926\\
+		\hline
+		SIPR & 4.1772 & 7.3203\\
+		\hline
+		LL & 1.0430 & 0.0276\\
+		\hline
+		MC & 0.9486 & 5.0e-07\\
+		\hline
+		SB & 1.0000 & 0.2236\\
+		\hline
+	\end{tabular}
+        \caption{$T$ = 20, $\rr$ Estimates and Standard Errors}\label{tab:time-res2}
 \end{table}
@@
 
+\subsubsection{Gaussian Monotonic Errors}
+\begin{figure}[H]
+	\centering
+	\includegraphics[scale=0.5]{images/time_nm.jpeg}
+	\caption{Forest plots of estimates of $\rr$ for the \xxsir methods for the Time1-4 data sets with Gaussian errors $(\beta=.06, \gamma=.03, T=365, X(0)=99950, Y(0)=50, Z(0)=0, \sigma_X=100, \sigma_Y=5, N=10^5)$.
+The points are the estimates of $\rr$ and the lines denote $\pm 2\cdot $Std.
+Err.
+The vertical line is the true $\rr$ value.}
+\end{figure}
+
+\begin{table}[H]
+
+	\centering
+	\begin{tabular}[t]{l|r|r}
+		\hline
+		Model & Estimate & Std. Dev\\
+		\hline
+		LS & 1.9934 & 0.1464\\
+		\hline
+		ReLS & 1.9935 & 0.1975\\
+		\hline
+		LMA & 1.2024 & 0.5439\\
+		\hline
+		LMAT & 2.7981 & 3.8629\\
+		\hline
+		IPR & 2.2784 & 5.3298\\
+		\hline
+		SIPR & 2.2020 & 0.5643\\
+		\hline
+		LL & 1.9928 & 0.0105\\
+		\hline
+		MC & 0.9486 & $<$ 1e-04\\
+		\hline
+		SB & 1.0000 & 0.1000\\
+		\hline
+	\end{tabular}
+	\caption{ $T$ = 100, $\rr$ Estimates and Standard Errors}
+\end{table}
+
+\begin{table}[H]
+	\centering
+	\begin{tabular}[t]{l|r|r}
+		\hline
+		Model & Estimate & Std. Dev\\
+		\hline
+		LS & 5.0366 & 1.3672\\
+		\hline
+		ReLS & 5.0449 & 2.1398\\
+		\hline
+		LMA & 1.1218 & 1.7267\\
+		\hline
+		LMAT & 2.7046 & 4.8031\\
+		\hline
+		IPR & 1.4155 & 4.5655\\
+		\hline
+		SIPR & 0.9703 & 0.8674\\
+		\hline
+		LL & 1.7472 & 0.3635\\
+		\hline
+		MC & 0.9486 & $<$ 1e-04\\
+		\hline
+		SB & 1.0000 & 0.2236\\
+		\hline
+	\end{tabular}
+	\caption{$T$ = 20, $\rr$ Estimates and Standard Errors}
+\end{table}
+
+\subsubsection{Autoregressive Errors}
+\begin{figure}[H]
+	\centering
+	\includegraphics[scale=0.5]{images/time_ar.jpeg}
+	\caption{Forest plots of estimates of $\rr$ for the \xxsir methods for the Time1-4 data sets with Gaussian errors $(\beta=.06, \gamma=.03, T=365, X(0)=99950, Y(0)=50, Z(0)=0, \sigma_X=100, \sigma_Y=5, N=10^5)$.
+The points are the estimates of $\rr$ and the lines denote $\pm 2\cdot $Std.
+Err.
+The vertical line is the true $\rr$ value.}
+\end{figure}
+
+\begin{table}[H]
+
+	\centering
+	\begin{tabular}[t]{l|r|r}
+		\hline
+		Model & Estimate & Std. Dev\\
+		\hline
+		LS & 1.9098 & 0.3811\\
+		\hline
+		ReLS & 1.9104 & 0.4773\\
+		\hline
+		LMA & 0.9141 & 0.2563\\
+		\hline
+		LMAT & 0.7524 & 2.5623\\
+		\hline
+		IPR & 9.2219 & 18.1185\\
+		\hline
+		SIPR & 2.0919 & 1.0699\\
+		\hline
+		LL & 1.7390 & 0.0358\\
+		\hline
+		MC & 0.9486 & $<$ 1e-04\\
+		\hline
+		SB & 3.3635 & 0.1834\\
+		\hline
+	\end{tabular}
+	\caption{ $T$ = 100, $\rr$ Estimates and Standard Errors}
+\end{table}
+
+\begin{table}[H]
+	\centering
+	\begin{tabular}[t]{l|r|r}
+		\hline
+		Model & Estimate & Std. Dev\\
+		\hline
+		LS & 0.0099 & 2.1688\\
+		\hline
+		ReLS & -0.1539 & 2.7692\\
+		\hline
+		LMA & -5.2984 & 65.6503\\
+		\hline
+		LMAT & 0.5717 & 1.3982\\
+		\hline
+		IPR & 20.4892 & 33.7260\\
+		\hline
+		SIPR & 0.3570 & 7.6611\\
+		\hline
+		LL & 0.9171 & 0.0338\\
+		\hline
+		MC & 0.9486 & $<$ 1e-04\\
+		\hline
+		SB & 11.3793 & 0.7543\\
+		\hline
+	\end{tabular}
+	\caption{$T$ = 20, $\rr$ Estimates and Standard Errors}
+\end{table}
+
+\subsubsection{Autoregressive Monotonic Errors}
+\begin{figure}[H]
+	\centering
+	\includegraphics[scale=0.5]{images/time_arm.jpeg}
+	\caption{Forest plots of estimates of $\rr$ for the \xxsir methods for the Time1-4 data sets with Gaussian errors $(\beta=.06, \gamma=.03, T=365, X(0)=99950, Y(0)=50, Z(0)=0, \sigma_X=100, \sigma_Y=5, N=10^5)$.
+The points are the estimates of $\rr$ and the lines denote $\pm 2\cdot $Std.
+Err.
+The vertical line is the true $\rr$ value.}
+\end{figure}
+
+\begin{table}[H]
+
+	\centering
+	\begin{tabular}[t]{l|r|r}
+		\hline
+		Model & Estimate & Std. Dev\\
+		\hline
+		LS & 1.9674 & 0.2324\\
+		\hline
+		ReLS & 1.9679 & 0.3065\\
+		\hline
+		LMA & 1.1584 & 0.3660\\
+		\hline
+		LMAT & 3.4346 & 14.6041\\
+		\hline
+		IPR & 2.2454 & 5.7090\\
+		\hline
+		SIPR & 1.9631 & 0.3489\\
+		\hline
+		LL & 2.0524 & 0.0209\\
+		\hline
+		MC & 0.9486 & $<$ 1e-04\\
+		\hline
+		SB & 1.0000 & 0.1000\\
+		\hline
+	\end{tabular}
+	\caption{ $T$ = 100, $\rr$ Estimates and Standard Errors}
+\end{table}
+
+\begin{table}[H]
+	\centering
+	\begin{tabular}[t]{l|r|r}
+		\hline
+		Model & Estimate & Std. Dev\\
+		\hline
+		LS & 1.6816 & 0.4747\\
+		\hline
+		ReLS & 1.6813 & 0.6544\\
+		\hline
+		LMA & 0.8899 & 0.7605\\
+		\hline
+		LMAT & 1.4496 & 1.5423\\
+		\hline
+		IPR & 3.8856 & 8.2899\\
+		\hline
+		SIPR & 3.9849 & 2.1351\\
+		\hline
+		LL & 1.4328 & 0.0397\\
+		\hline
+		MC & 0.9486 & $<$ 1e-04\\
+		\hline
+		SB & 1.0000 & 0.2236\\
+		\hline
+	\end{tabular}
+	\caption{$T$ = 20, $\rr$ Estimates and Standard Errors}
+\end{table}
+
 \subsection{Initial Percentage of Infectious $\left (\frac{Y(0)}{N}\right)$}\label{sec:res-inf}
@@
 
+\subsubsection{Gaussian Errors}
+
 \begin{figure}[H]
   \centering
-  \includegraphics[scale=0.5]{images/start.jpeg}
-    \caption{Forest plots of estimates of $\rr$ for the \xxsir methods for the Inits1-4 data sets $(\beta=.06, \gamma=.03, O=365, X(0)=99950, Y(0)=50, Z(0)=0, \sigma_X=100, \sigma_Y=5, N=10^5)$.
+  \includegraphics[scale=0.5]{images/start_n.jpeg}
+    \caption{Forest plots of estimates of $\rr$ for the \xxsir methods for the Inits1-4 data sets $(\beta=.06, \gamma=.03, T=365, X(0)=99950, Y(0)=50, Z(0)=0, \sigma_X=100, \sigma_Y=5, N=10^5)$.
 The points are the estimates of $\rr$ and the lines denote $\pm 2\cdot $Std.
@@
 		\hline
-		LS & 1.9995884 & 0.0049609\\
-		\hline
-		ReLS & 1.9997123 & 0.0034303\\
-		\hline
-		SIRLMA & -87.5945650 & 66413.128046\\
-		\hline
-		LMAT & 0.7306100 & 4.681042\\
-		\hline
-		IPR & 1.285071 & 1.440980\\
-		\hline
-		SIPR & 1.2522106 & 57.9229822 \\
-		\hline
-		LL & 1.9995148 & 0.0002828\\
-		\hline
-		MC & 0.9485555 & $<$ 1e-07\\
-		\hline
-		SB & 0.9728534 & 0.0516270\\
+		LS & 1.9996 & 0.0050\\
+		\hline
+		ReLS & 1.9998 & 0.0034\\
+		\hline
+		LMA & -87.5946 & 66413.1280\\
+		\hline
+		LMAT & 0.7306 & 4.6810\\
+		\hline
+		IPR & 1.2851 & 1.4410\\
+		\hline
+		SIPR & 1.2522 & 57.9230\\
+		\hline
+		LL & 1.9996 & 0.0003 \\
+		\hline
+		MC & 0.9486 & $<$ 1e-04\\
+		\hline
+		SB & 0.9729 & 0.0516\\
 		\hline
@@
 		\hline
-		LS & 2.468304 & 1.5370731\\
-		\hline
-		ReLS & 2.469189 & 2.7293667\\
-		\hline
-		SIRLMA & 10.9282204 & 19.297359 \\
-		\hline
-		LMAT & 1.624645 & 5.0359845\\
-		\hline
-		IPR & 774.274659 & 7491.005080 \\
-		\hline
-		SIPR & 2.4390757 & $<$ 1e-07 \\
-		\hline
-		LL & 1.999690 & 0.0007731\\
-		\hline
-		MC & 0.948603 & 0.0000156\\
-		\hline
-		SB & 1.471481 & 0.0634937\\
+		LS & 2.4683 & 1.5371\\
+		\hline
+		ReLS & 2.4692 & 2.7294 \\
+		\hline
+		LMA & 10.9282 & 19.2974 \\
+		\hline
+		LMAT & 1.6246 & 5.0360\\
+		\hline
+		IPR & 774.2747 & 7491.0051 \\
+		\hline
+		SIPR & 2.4391 & $<$ 1e-04 \\
+		\hline
+		LL & 1.9997 & 0.0008\\
+		\hline
+		MC & 0.9486 & 1.56e-05\\
+		\hline
+		SB & 1.4715 & 0.0635\\
 		\hline
@@
         \caption{$Y(0) = 1$, $\rr$ Estimates and Standard Errors}\label{tab:inits-res2}
+\end{table}
+
+\subsubsection{Gaussian Montonic Errors}
+
+\begin{figure}[H]
+	\centering
+	\includegraphics[scale=0.5]{images/start_nm.jpeg}
+	\caption{Forest plots of estimates of $\rr$ for the \xxsir methods for the Inits1-4 data sets $(\beta=.06, \gamma=.03, T=365, X(0)=99950, Y(0)=50, Z(0)=0, \sigma_X=100, \sigma_Y=5, N=10^5)$.
+The points are the estimates of $\rr$ and the lines denote $\pm 2\cdot $Std.
+Err.
+The vertical line is the true $\rr$ value.}
+\end{figure}
+
+\begin{table}[H]
+
+	\centering
+	\begin{tabular}[t]{l|r|r}
+		\hline
+		Model & Estimate & Std. Dev\\
+		\hline
+		LS & 2.0003 & 0.0046\\
+		\hline
+		ReLS & 2.0003 & 0.0032\\
+		\hline
+		LMA & -0.9982 & 1.4189\\
+		\hline
+		LMAT & 0.9703 & 0.6574\\
+		\hline
+		IPR & 0.9866 & 1.0570\\
+		\hline
+		SIPR & 2.0166 & 57.8154\\
+		\hline
+		LL & 2.0001 & 0.0002\\
+		\hline
+		MC & 0.9486 & $<$ 1e-04\\
+		\hline
+		SB & 1.0868 & 0.0546\\
+		\hline
+	\end{tabular}
+	\caption{$Y(0) = 1000$, $\rr$ Estimates and Standard Errors}
+\end{table}
+
+\begin{table}[H]
+
+	\centering
+	\begin{tabular}[t]{l|r|r}
+		\hline
+		Model & Estimate & Std. Dev\\
+		\hline
+		LS & 2.4668 & 1.5349\\
+		\hline
+		ReLS & 2.4670 & 2.7219\\
+		\hline
+		LMA & -99.3726 & 3136.3141\\
+		\hline
+		LMAT & 0.4205 & 12.9331\\
+		\hline
+		IPR & 3.6061 & 26.5409\\
+		\hline
+		SIPR & 5.4116 & $<$ 1e-04\\
+		\hline
+		LL & 2.0005 & 0.0004\\
+		\hline
+		MC & 0.9486 & $<$ 1e-04\\
+		\hline
+		SB & 1.7685 & 0.0696\\
+		\hline
+	\end{tabular}
+	\caption{$Y(0) = 1$, $\rr$ Estimates and Standard Errors}
+\end{table}
+
+\subsubsection{Autoregressive Errors}
+
+\begin{figure}[H]
+	\centering
+	\includegraphics[scale=0.5]{images/start_ar.jpeg}
+	\caption{Forest plots of estimates of $\rr$ for the \xxsir methods for the Inits1-4 data sets $(\beta=.06, \gamma=.03, T=365, X(0)=99950, Y(0)=50, Z(0)=0, \sigma_X=100, \sigma_Y=5, N=10^5)$.
+The points are the estimates of $\rr$ and the lines denote $\pm 2\cdot $Std.
+Err.
+The vertical line is the true $\rr$ value.}
+\end{figure}
+
+\begin{table}[H]
+
+	\centering
+	\begin{tabular}[t]{l|r|r}
+		\hline
+		Model & Estimate & Std. Dev\\
+		\hline
+		LS & 2.0005 & 0.0078\\
+		\hline
+		ReLS & 2.0001 & 0.0054\\
+		\hline
+		LMA & 4.1646 & 3.8688\\
+		\hline
+		LMAT & 3.1508 & 45.5885\\
+		\hline
+		IPR & 1.1936 & 1.1958\\
+		\hline
+		SIPR & 2.2854 & 57.4383\\
+		\hline
+		LL & 1.9998 & 0.0004\\
+		\hline
+		MC & 0.9486 & $<$ 1e-04\\
+		\hline
+		SB & 0.3154 & 0.0294\\
+		\hline
+	\end{tabular}
+	\caption{$Y(0) = 1000$, $\rr$ Estimates and Standard Errors}
+\end{table}
+
+\begin{table}[H]
+
+	\centering
+	\begin{tabular}[t]{l|r|r}
+		\hline
+		Model & Estimate & Std. Dev\\
+		\hline
+		LS & 2.4706 & 1.5372\\
+		\hline
+		ReLS & 2.4707 & 2.7245\\
+		\hline
+		LMA & -0.8259 & 1.7277\\
+		\hline
+		LMAT & 1.3117 & 26.0563\\
+		\hline
+		IPR & 5.0000 & 1199.1557\\
+		\hline
+		SIPR & 1.2047 & $<$ 1e-04\\
+		\hline
+		LL & 2.0027 & 0.0011\\
+		\hline
+		MC & 0.9486 & $<$ 1e-04\\
+		\hline
+		SB & 1.7698 & 0.0696\\
+		\hline
+	\end{tabular}
+	\caption{$Y(0) = 1$, $\rr$ Estimates and Standard Errors}
+\end{table}
+
+\subsubsection{Autoregressive Montonic Errors}
+
+\begin{figure}[H]
+	\centering
+	\includegraphics[scale=0.5]{images/start_arm.jpeg}
+	\caption{Forest plots of estimates of $\rr$ for the \xxsir methods for the Inits1-4 data sets $(\beta=.06, \gamma=.03, T=365, X(0)=99950, Y(0)=50, Z(0)=0, \sigma_X=100, \sigma_Y=5, N=10^5)$.
+The points are the estimates of $\rr$ and the lines denote $\pm 2\cdot $Std.
+Err.
+The vertical line is the true $\rr$ value.}
+\end{figure}
+
+\begin{table}[H]
+
+	\centering
+	\begin{tabular}[t]{l|r|r}
+		\hline
+		Model & Estimate & Std. Dev\\
+		\hline
+		LS & 2.0001 & 0.0069\\
+		\hline
+		ReLS & 2.0000 & 0.0048\\
+		\hline
+		LMA & -10.8054 & 286.9861\\
+		\hline
+		LMAT & 0.9333 & 0.8824\\
+		\hline
+		IPR & 0.9960 & 0.9987\\
+		\hline
+		SIPR & 1.9222 & 57.0458\\
+		\hline
+		LL & 2.0006 & 0.0003\\
+		\hline
+		MC & 0.9486 & $<$ 1e-04\\
+		\hline
+		SB & 0.3790 & 0.0322\\
+		\hline
+	\end{tabular}
+	\caption{$Y(0) = 1000$, $\rr$ Estimates and Standard Errors}
+\end{table}
+
+\begin{table}[H]
+
+	\centering
+	\begin{tabular}[t]{l|r|r}
+		\hline
+		Model & Estimate & Std. Dev\\
+		\hline
+		LS & 2.4667 & 1.5349\\
+		\hline
+		ReLS & 2.4665 & 2.7349\\
+		\hline
+		LMA & 16.7344 & 18.2294\\
+		\hline
+		LMAT & 3.1984 & 12.9100\\
+		\hline
+		IPR & 2.5082 & 12.8639\\
+		\hline
+		SIPR & 1.0356 & $<$ 1e-04\\
+		\hline
+		LL & 1.9999 & 0.0006\\
+		\hline
+		MC & 0.9486 & $<$ 1e-04\\
+		\hline
+		SB & 1.7793 & 0.0698\\
+		\hline
+	\end{tabular}
+	\caption{$Y(0) = 1$, $\rr$ Estimates and Standard Errors}
 \end{table}
@@
 
+\subsubsection{Gaussian Errors}
+
 \begin{figure}[H]
 	\centering
-	\includegraphics[scale=0.5]{images/popsize.jpeg}
-	\caption{Forest plots of estimates of $\rr$ for the \xxsir methods for the PopSize1-4 data sets $(\beta=.06, \gamma=.03, O=365, \sigma_X=100, \sigma_Y=5)$.
+	\includegraphics[scale=0.5]{images/popsize_n.jpeg}
+	\caption{Forest plots of estimates of $\rr$ for the \xxsir methods for the PopSize1-4 data sets $(\beta=.06, \gamma=.03, T=365, \sigma_X=100, \sigma_Y=5)$.
 The points are the estimates of $\rr$ and the lines denote $\pm 2\cdot $Std.
@@
 		\hline
-		LS & 2.0007741 & 0.0590423\\
-		\hline
-		ReLS & 2.0005189 & 0.0528419\\
-		\hline
-		SIRLMA & 1.9976309 & 2606.0798123\\
-		\hline
-		LMAT & 1.2944136 & 1.4304706\\
-		\hline
-		IPR & 40.8072373 & 133.2798314\\
-		\hline
-		SIPR & 1.0628910 &  $<$ 1e-07\\
-		\hline
-		LL & 1.9988393 & 0.0023193\\
-		\hline
-		MC & 0.9485626 & 0.0000040\\
-		\hline
-		SB & 0.4418182 & 0.0347917\\
+		LS & 2.0008 & 0.0590\\
+		\hline
+		ReLS & 2.0005 & 0.0528\\
+		\hline
+		LMA & 1.9976 & 2606.0798\\
+		\hline
+		LMAT & 1.2944 & 1.4305\\
+		\hline
+		IPR & 40.8072 & 133.2798\\
+		\hline
+		SIPR & 1.0629 &  $<$ 1e-04\\
+		\hline
+		LL & 1.9988 & 0.0023\\
+		\hline
+		MC & 0.9486 & 4.0e-06\\
+		\hline
+		SB & 0.4418 & 0.0348\\
 		\hline
@@
 		\hline
-		LS & 1.9862533 & 4.861988\\
-		\hline
-		ReLS & 1.9871049 & 3.456891\\
-		\hline
-		SIRLMA & 0.8023469 & 18.94166\\
-		\hline
-		LMAT & 1.2208744 & 4.769929\\
-		\hline
-		IPR & 3620.8676795 & 57608.17\\
-		\hline
-		SIPR & 3.3678247 & 22.41867\\
-		\hline
-		LL & 1.1413175 & 0.0422219\\
-		\hline
-		MC & 0.8455334 & 0.0124837\\
-		\hline
-		SB & 0.3887115 & 0.0326338\\
+		LS & 1.9863 & 4.8620\\
+		\hline
+		ReLS & 1.9871 & 3.4569\\
+		\hline
+		LMA & 0.8023 & 18.9417\\
+		\hline
+		LMAT & 1.2209 & 4.7699\\
+		\hline
+		IPR & 3620.8677 & 57608.1700\\
+		\hline
+		SIPR & 3.3678 & 22.4187\\
+		\hline
+		LL & 1.1413 & 0.0422\\
+		\hline
+		MC & 0.8456 & 0.0125\\
+		\hline
+		SB & 0.3887 & 0.0326\\
 		\hline
@@
 \caption{$N = 100$, $\rr$ Estimates and Standard Errors}\label{tab:n2-res2}
+\end{table}
+
+\subsubsection{Gaussian Monotonic Errors}
+
+\begin{figure}[H]
+	\centering
+	\includegraphics[scale=0.5]{images/popsize_nm.jpeg}
+	\caption{Forest plots of estimates of $\rr$ for the \xxsir methods for the PopSize1-4 data sets $(\beta=.06, \gamma=.03, T=365, \sigma_X=100, \sigma_Y=5)$.
+The points are the estimates of $\rr$ and the lines denote $\pm 2\cdot $Std.
+Err.
+The vertical line is the true $\rr$ value.}
+\end{figure}
+
+\begin{table}[H]
+
+	\centering
+	\begin{tabular}[t]{l|r|r}
+		\hline
+		Model & Estimate & Std. Dev\\
+		\hline
+		LS & 2.0175 & 0.1430\\
+		\hline
+		ReLS & 2.0175 & 0.1021\\
+		\hline
+		LMA & 0.8256 & 1.4444\\
+		\hline
+		LMAT & 1.0457 & 0.6365\\
+		\hline
+		IPR & 1.7343 & 12.7260\\
+		\hline
+		SIPR & 1.6932 & 12.0122\\
+		\hline
+		LL & 2.0102 & 0.0056\\
+		\hline
+		MC & 0.9486 & $<$ 1e-04\\
+		\hline
+		SB & 1.0000 & 0.0523\\
+		\hline
+	\end{tabular}
+	\caption{$N = 10000$, $\rr$ Estimates and Standard Errors}
+\end{table}
+
+\begin{table}[H]
+
+	\centering
+	\begin{tabular}[t]{l|r|r}
+		\hline
+		Model & Estimate & Std. Dev\\
+		\hline
+		LS & 1.9998 & 0.0006\\
+		\hline
+		ReLS & 1.9999 & 0.0005\\
+		\hline
+		LMA & 1.7673 & 0.0976\\
+		\hline
+		LMAT & 1.9732 & 10.8872\\
+		\hline
+		IPR & 1.4344 & 1.1311\\
+		\hline
+		SIPR & 1.7907 & $<$ 1e-04\\
+		\hline
+		LL & 2.0000 & $<$ 1e-04\\
+		\hline
+		MC & 0.9486 & $<$ 1e-04\\
+		\hline
+		SB & 1.2709 & 0.0590\\
+		\hline
+	\end{tabular}
+	\caption{$N = 100$, $\rr$ Estimates and Standard Errors}
+\end{table}
+
+\subsubsection{Autoregressive Errors}
+
+\begin{figure}[H]
+	\centering
+	\includegraphics[scale=0.5]{images/popsize_ar.jpeg}
+	\caption{Forest plots of estimates of $\rr$ for the \xxsir methods for the PopSize1-4 data sets $(\beta=.06, \gamma=.03, T=365, \sigma_X=100, \sigma_Y=5)$.
+The points are the estimates of $\rr$ and the lines denote $\pm 2\cdot $Std.
+Err.
+The vertical line is the true $\rr$ value.}
+\end{figure}
+
+\begin{table}[H]
+
+	\centering
+	\begin{tabular}[t]{l|r|r}
+		\hline
+		Model & Estimate & Std. Dev\\
+		\hline
+		LS & 2.1221 & 0.7363\\
+		\hline
+		ReLS & 2.1219 & 0.5660\\
+		\hline
+		LMA & 1.2045 & 0.3878\\
+		\hline
+		LMAT & 1.0881 & 3.9239\\
+		\hline
+		IPR & 164.6111 & 1887.1253\\
+		\hline
+		SIPR & 0.5335 & 13.6519\\
+		\hline
+		LL & 2.4701 & 0.0546\\
+		\hline
+		MC & 0.9500 & 0.0003\\
+		\hline
+		SB & 1.5457 & 0.0651\\
+		\hline
+	\end{tabular}
+	\caption{$N = 10000$, $\rr$ Estimates and Standard Errors}
+\end{table}
+
+\begin{table}[H]
+
+	\centering
+	\begin{tabular}[t]{l|r|r}
+		\hline
+		Model & Estimate & Std. Dev\\
+		\hline
+		LS & 1.9998 & 0.0009\\
+		\hline
+		ReLS & 2.0001 & 0.0007\\
+		\hline
+		LMA & 1.6873 & 0.0810\\
+		\hline
+		LMAT & 1.3885 & 0.6855\\
+		\hline
+		IPR & 1.5672 & 1.5944\\
+		\hline
+		SIPR & 1.7903 & $<$ 1e-04\\
+		\hline
+		LL & 1.9999 & $<$ 1e-04\\
+		\hline
+		MC & 0.9486 & $<$ 1e-04\\
+		\hline
+		SB & 1.1877 & 0.0570\\
+		\hline
+	\end{tabular}
+	\caption{$N = 100$, $\rr$ Estimates and Standard Errors}
+\end{table}
+
+\subsubsection{Autoregressive Monotonic Errors}
+
+\begin{figure}[H]
+	\centering
+	\includegraphics[scale=0.5]{images/popsize_arm.jpeg}
+	\caption{Forest plots of estimates of $\rr$ for the \xxsir methods for the PopSize1-4 data sets $(\beta=.06, \gamma=.03, T=365, \sigma_X=100, \sigma_Y=5)$.
+The points are the estimates of $\rr$ and the lines denote $\pm 2\cdot $Std.
+Err.
+The vertical line is the true $\rr$ value.}
+\end{figure}
+
+\begin{table}[H]
+
+	\centering
+	\begin{tabular}[t]{l|r|r}
+		\hline
+		Model & Estimate & Std. Dev\\
+		\hline
+		LS & 2.0975 & 0.2921\\
+		\hline
+		ReLS & 2.0976 & 0.2211\\
+		\hline
+		LMA & 1.1720 & 0.0511\\
+		\hline
+		LMAT & -13.6906 & 281.9114\\
+		\hline
+		IPR & 1.7077 & 11.7775\\
+		\hline
+		SIPR & 2.8981 & 284.4933\\
+		\hline
+		LL & 2.2806 & 0.0230\\
+		\hline
+		MC & 0.9486 & $<$ 1e-04\\
+		\hline
+		SB & 1.0000 & 0.0523\\
+		\hline
+	\end{tabular}
+	\caption{$N = 10000$, $\rr$ Estimates and Standard Errors}
+\end{table}
+
+\begin{table}[H]
+
+	\centering
+	\begin{tabular}[t]{l|r|r}
+		\hline
+		Model & Estimate & Std. Dev\\
+		\hline
+		LS & 2.0000 & 0.0007\\
+		\hline
+		ReLS & 1.9997 & 0.0007\\
+		\hline
+		LMA & 1.8332 & 0.1199\\
+		\hline
+		LMAT & 1.4164 & 0.6348\\
+		\hline
+		IPR & 1.4490 & 1.1803\\
+		\hline
+		SIPR & 1.7505 & $<$ 1e-04\\
+		\hline
+		LL & 1.9998 & $<$ 1e-04\\
+		\hline
+		MC & 0.9486 & $<$ 1e-04\\
+		\hline
+		SB & 1.8078 & 0.0704\\
+		\hline
+	\end{tabular}
+	\caption{$N = 100$, $\rr$ Estimates and Standard Errors}
 \end{table}
@@
 
+\subsubsection{Gaussian Errors}
+
 \begin{figure}[H]
 \begin{center}
-  \includegraphics[scale=0.5]{images/var.jpeg}
-  \caption{Forest plots of estimates of $\rr$ for the \xxsir methods for the Var1-4 data sets $(\beta=.06, \gamma=.03, O=365, T=365, X(0)=99950, Y(0)=50, Z(0)=0, N=10^5)$.
+  \includegraphics[scale=0.5]{images/var_n.jpeg}
+  \caption{Forest plots of estimates of $\rr$ for the \xxsir methods for the Var1-4 data sets $(\beta=.06, \gamma=.03, T=365, X(0)=99950, Y(0)=50, Z(0)=0, N=10^5)$.
 The points are the estimates of $\rr$ and the lines denote $\pm 2\cdot $Std.
@@
 		\hline
-		LS & 1.9997552 & 0.0026502\\
-		\hline
-		ReLS & 1.9998295 & 0.0023646\\
-		\hline
-		SIRLMA & 2.0398087 & 0.2773968\\
-		\hline
-		LMAT & 1.4148628 & 0.5779651\\
-		\hline
-		IPR & 2.7607518 & 8.1318687\\
-		\hline
-		SIPR & 1.8368240 & $<$ 1e-07\\
-		\hline
-		LL & 1.9998907 & 0.0001037\\
-		\hline
-		MC & 0.9485555 & $<$ 1e-07\\
-		\hline
-		SB & 1.6750053 & 0.0677426\\
+		LS & 1.9998 & 0.0027\\
+		\hline
+		ReLS & 1.9998 & 0.0024\\
+		\hline
+		LMA & 2.0398 & 0.2774\\
+		\hline
+		LMAT & 1.4149 & 0.5780\\
+		\hline
+		IPR & 2.7608 & 8.1319\\
+		\hline
+		SIPR & 1.8368 & $<$ 1e-04\\
+		\hline
+		LL & 1.9999 & 0.0001\\
+		\hline
+		MC & 0.9486 & $<$ 1e-04\\
+		\hline
+		SB & 1.6750 & 0.0677\\
 		\hline
@@
 		\hline
-		LS & 1.9980880 & 0.1353803\\
-		\hline
-		ReLS & 1.9981446 & 0.1207935\\
-		\hline
-		SIRLMA & 0.4651913 & 7.5508447\\
-		\hline
-		LMAT & 1.2877865 & 4.3947658\\
-		\hline
-		IPR & 81.4901666 & 348.7350849\\
-		\hline
-		SIPR & 1.0606095 & $<$ 1e-07\\
-		\hline
-		LL & 2.0018816 & 0.0054085\\
-		\hline
-		MC & 0.9485562 & 0.0000010\\
-		\hline
-		SB & 1.8567176 & 0.0713225\\
+		LS & 1.9981 & 0.1354\\
+		\hline
+		ReLS & 1.9981 & 0.1208\\
+		\hline
+		LMA & 0.4652 & 7.5508\\
+		\hline
+		LMAT & 1.2878 & 4.3948\\
+		\hline
+		IPR & 81.4902 & 348.7351\\
+		\hline
+		SIPR & 1.0606 & $<$ 1e-04\\
+		\hline
+		LL & 2.0019 & 0.0054\\
+		\hline
+		MC & 0.9486 & 1.0e-06\\
+		\hline
+		SB & 1.8567 & 0.0713\\
 		\hline
@@
 One exception in the IPR model, which may do worse due to larger fluctuations in the incidence prevalence ratio.
+
+\subsubsection{Gaussian Monotonic Errors}
+
+\begin{figure}[H]
+	\begin{center}
+		\includegraphics[scale=0.5]{images/var_nm.jpeg}
+		\caption{Forest plots of estimates of $\rr$ for the \xxsir methods for the Var1-4 data sets $(\beta=.06, \gamma=.03, T=365, X(0)=99950, Y(0)=50, Z(0)=0, N=10^5)$.
+The points are the estimates of $\rr$ and the lines denote $\pm 2\cdot $Std.
+Err.
+The vertical line is the true $\rr$ value.}
+	\end{center}
+\end{figure}
+
+\begin{table}[H]
+
+	\centering
+	\begin{tabular}[t]{l|r|r}
+		\hline
+		Model & Estimate & Std. Dev\\
+		\hline
+		LS & 1.9998 & 0.0026\\
+		\hline
+		ReLS & 1.9999 & 0.0024\\
+		\hline
+		LMA & 1.6268 & 0.0865\\
+		\hline
+		LMAT & 1.5099 & 2.1244\\
+		\hline
+		IPR & 1.4509 & 2.0522\\
+		\hline
+		SIPR & 1.7778 & $<$ 1e-04\\
+		\hline
+		LL & 2.0000 & 0.0001\\
+		\hline
+		MC & 0.9486 & $<$ 1e-04\\
+		\hline
+		SB & 1.2333 & 0.0581\\
+		\hline
+	\end{tabular}
+	\caption{ $(\sigma^2_X, \sigma^2_Y) = (50, 2)$, $\rr$ Estimates and Standard Errors}
+\end{table}
+
+\begin{table}[H]
+
+	\centering
+	\begin{tabular}[t]{l|r|r}
+		\hline
+		Model & Estimate & Std. Dev\\
+		\hline
+		LS & 1.9970 & 0.0541\\
+		\hline
+		ReLS & 1.9971 & 0.0482\\
+		\hline
+		LMA & 4.2062 & 66.2206\\
+		\hline
+		LMAT & 1.3653 & 3.2675\\
+		\hline
+		IPR & 2.2672 & 17.0580\\
+		\hline
+		SIPR & 1.2346 & $<$ 1e-04\\
+		\hline
+		LL & 1.9957 & 0.0015\\
+		\hline
+		MC & 0.9486 & $<$ 1e-04\\
+		\hline
+		SB & 1.0000 & 0.0523\\
+		\hline
+	\end{tabular}
+	\caption{$(\sigma^2_X, \sigma^2_Y) = (2500, 100)$, $\rr$ Estimates and Standard Errors}
+\end{table}
+
+\subsubsection{Autoregressive Errors}
+
+\begin{figure}[H]
+	\begin{center}
+		\includegraphics[scale=0.5]{images/var_ar.jpeg}
+		\caption{Forest plots of estimates of $\rr$ for the \xxsir methods for the Var1-4 data sets $(\beta=.06, \gamma=.03, T=365, X(0)=99950, Y(0)=50, Z(0)=0, N=10^5)$.
+The points are the estimates of $\rr$ and the lines denote $\pm 2\cdot $Std.
+Err.
+The vertical line is the true $\rr$ value.}
+	\end{center}
+\end{figure}
+
+\begin{table}[H]
+
+	\centering
+	\begin{tabular}[t]{l|r|r}
+		\hline
+		Model & Estimate & Std. Dev\\
+		\hline
+		LS & 1.9996 & 0.0038\\
+		\hline
+		ReLS & 1.9999 & 0.0034\\
+		\hline
+		LMA & 1.4644 & 0.0750\\
+		\hline
+		LMAT & 1.3999 & 0.7194\\
+		\hline
+		IPR & 2.1980 & 5.4569\\
+		\hline
+		SIPR & 1.7604 & $<$ 1e-04\\
+		\hline
+		LL & 1.9991 & 0.0001\\
+		\hline
+		MC & 0.9486 & $<$ 1e-04\\
+		\hline
+		SB & 1.6378 & 0.0670\\
+		\hline
+	\end{tabular}
+	\caption{ $(\sigma^2_X, \sigma^2_Y) = (50, 2)$, $\rr$ Estimates and Standard Errors}
+\end{table}
+
+\begin{table}[H]
+
+	\centering
+	\begin{tabular}[t]{l|r|r}
+		\hline
+		Model & Estimate & Std. Dev\\
+		\hline
+		LS & 1.9997 & 0.1894\\
+		\hline
+		ReLS & 2.0002 & 0.1680\\
+		\hline
+		LMA & 1.0392 & 0.0166\\
+		\hline
+		LMAT & 1.1985 & 1.6107\\
+		\hline
+		IPR & 195.3708 & 1451.0401\\
+		\hline
+		SIPR & 1.7028 & $<$ 1e-04\\
+		\hline
+		LL & 2.0386 & 0.0082\\
+		\hline
+		MC & 0.9486 & $<$ 1e-04\\
+		\hline
+		SB & 1.9073 & 0.0723\\
+		\hline
+	\end{tabular}
+	\caption{$(\sigma^2_X, \sigma^2_Y) = (2500, 100)$, $\rr$ Estimates and Standard Errors}
+\end{table}
+
+\subsubsection{Autoregressive Monotonic Errors}
+
+\begin{figure}[H]
+	\begin{center}
+		\includegraphics[scale=0.5]{images/var_arm.jpeg}
+		\caption{Forest plots of estimates of $\rr$ for the \xxsir methods for the Var1-4 data sets $(\beta=.06, \gamma=.03, T=365, X(0)=99950, Y(0)=50, Z(0)=0, N=10^5)$.
+The points are the estimates of $\rr$ and the lines denote $\pm 2\cdot $Std.
+Err.
+The vertical line is the true $\rr$ value.}
+	\end{center}
+\end{figure}
+
+\begin{table}[H]
+
+	\centering
+	\begin{tabular}[t]{l|r|r}
+		\hline
+		Model & Estimate & Std. Dev\\
+		\hline
+		LS & 1.9990 & 0.0036\\
+		\hline
+		ReLS & 1.9995 & 0.0032\\
+		\hline
+		LMA & 2.8515 & 0.5541\\
+		\hline
+		LMAT & 1.4263 & 0.6782\\
+		\hline
+		IPR & 1.4465 & 2.1821\\
+		\hline
+		SIPR & 1.8423 & $<$ 1e-04\\
+		\hline
+		LL & 1.9997 & 0.0001\\
+		\hline
+		MC & 0.9486 & $<$ 1e-04\\
+		\hline
+		SB & 1.5555 & 0.0653\\
+		\hline
+	\end{tabular}
+	\caption{ $(\sigma^2_X, \sigma^2_Y) = (50, 2)$, $\rr$ Estimates and Standard Errors}
+\end{table}
+
+\begin{table}[H]
+
+	\centering
+	\begin{tabular}[t]{l|r|r}
+		\hline
+		Model & Estimate & Std. Dev\\
+		\hline
+		LS & 2.0021 & 0.0928\\
+		\hline
+		ReLS & 2.0022 & 0.0827\\
+		\hline
+		LMA & -0.1164 & 1.9126\\
+		\hline
+		LMAT & 0.9424 & 4.3914\\
+		\hline
+		IPR & 2.9241 & 18.4344\\
+		\hline
+		SIPR & 2.5962 & $<$ 1e-04\\
+		\hline
+		LL & 2.0157 & 0.0031\\
+		\hline
+		MC & 0.9486 & $<$ 1e-04\\
+		\hline
+		SB & 1.0000 & 0.0523\\
+		\hline
+	\end{tabular}
+	\caption{$(\sigma^2_X, \sigma^2_Y) = (2500, 100)$, $\rr$ Estimates and Standard Errors}
+\end{table}
 
@@
 
+\subsubsection{Gaussian Errors}
+
 \begin{figure}[H]
   \begin{center}
-    \includegraphics[scale=0.5]{images/other.jpeg}
-  \caption{Forest plots of estimates of $\rr$ for the \xxsir methods for the Poly1, Poly4, and LSIR data sets $(\beta=.06, \gamma=.03, O=365, T=365, X(0)=99950, Y(0)=50, Z(0)=0, N=10^5, \sigma_X=100, \sigma_Y=5)$.
+    \includegraphics[scale=0.5]{images/other_n.jpeg}
+  \caption{Forest plots of estimates of $\rr$ for the \xxsir methods for the Poly1, Poly4, and LSIR data sets $(\beta=.06, \gamma=.03, T=365, X(0)=99950, Y(0)=50, Z(0)=0, N=10^5, \sigma_X=100, \sigma_Y=5)$.
 The points are the estimates of $\rr$ and the lines denote $\pm 2\cdot $Std.
@@
 		\hline
-		LS & 1.0514375 & 0.3782909\\
-		\hline
-		ReLS & 1.0514351 & 0.0259416\\
-		\hline
-		SIRLMA & 1.2655140 & 0.1278435\\
-		\hline
-		LMAT & 1.1597070 & 0.1019956\\
-		\hline
-		IPR & 4.4999548 & 15.2185520\\
-		\hline
-		SIPR & 98.0610578 & $<$ 1e-07\\
-		\hline
-		LL & 1.1880504 & 0.0006341\\
-		\hline
-		MC & 0.9485551 & 0.0000001\\
-		\hline
-		SB & 1.4150329 & 0.0622640\\
+		LS & 1.0514 & 0.3783\\
+		\hline
+		ReLS & 1.0514 & 0.0259\\
+		\hline
+		LMA & 1.2656 & 0.1278\\
+		\hline
+		LMAT & 1.1597 & 0.1020\\
+		\hline
+		IPR & 4.5000 & 15.2186\\
+		\hline
+		SIPR & 98.0611 & $<$ 1e-04\\
+		\hline
+		LL & 1.1881 & 0.0006\\
+		\hline
+		MC & 0.9486 & 1e-07 \\
+		\hline
+		SB & 1.4150 & 0.0622\\
 		\hline
@@
 	\hline
-	LS & 1.7693473 & 0.4226615\\
-	\hline
-	ReLS & 1.7693312 & 0.2796903\\
-	\hline
-	SIRLMA & 0.3617606 & 0.1231325\\
-	\hline
-	LMAT & 0.8213066 & 6.0466473\\
-	\hline
-	IPR & 10.4349388 & 33.8952710\\
-	\hline
-	SIPR & 4.7477067 &  $<$ 1e-07\\
-	\hline
-	LL & 1.9551795 & 0.0164608\\
-	\hline
-	MC & 0.9485545 & 0.0000006\\
-	\hline
-	SB & 0.2118519 & 0.0240918\\
+	LS & 1.7693 & 0.4227\\
+	\hline
+	ReLS & 1.7693 & 0.2797\\
+	\hline
+	LMA & 0.3618 & 0.1231\\
+	\hline
+	LMAT & 0.8213 & 6.0466\\
+	\hline
+	IPR & 10.4349 & 33.8953 \\
+	\hline
+	SIPR & 4.7477 &  $<$ 1e-04\\
+	\hline
+	LL & 1.9552 & 0.0165\\
+	\hline
+	MC & 0.9486 & 6e-07\\
+	\hline
+	SB & 0.2119 & 0.02409\\
 	\hline
@@
 		\hline
-		LS & 1.2742577 & 0.0979741\\
-		\hline
-		ReLS & 1.2743142 & 0.0339775\\
-		\hline
-		SIRLMA & 1.2191008 & 0.1092962\\
-		\hline
-		LMAT & 1.2009373 & 0.0031099\\
-		\hline
-		IPR & 2.5280045 & 2.5904017\\
-		\hline
-		SIPR & 2.1567978 & 0.3585219\\
-		\hline
-		LL & 1.2585872 & 0.0010079\\
-		\hline
-		MC & 0.9485555 & $<$ 1e-07 \\
-		\hline
-		SB & 1.2008765 & 0.0573592\\
+		LS & 1.2743 & 0.0980\\
+		\hline
+		ReLS & 1.2743 & 0.0340\\
+		\hline
+		LMA & 1.2191 & 0.1093\\
+		\hline
+		LMAT & 1.2009 & 0.0031\\
+		\hline
+		IPR & 2.5280 & 2.5904\\
+		\hline
+		SIPR & 2.1568 & 0.3585\\
+		\hline
+		LL & 1.2586 & 0.0010\\
+		\hline
+		MC & 0.9486 & $<$ 1e-04 \\
+		\hline
+		SB & 1.2009 & 0.0574\\
 		\hline
@@
 
-\section{Variance Estimation Methods}
+\subsubsection{Gaussian Monotonic Errors}
+
+\begin{figure}[H]
+	\begin{center}
+		\includegraphics[scale=0.5]{images/other_nm.jpeg}
+		\caption{Forest plots of estimates of $\rr$ for the \xxsir methods for the Poly1, Poly4, and LSIR data sets $(\beta=.06, \gamma=.03, T=365, X(0)=99950, Y(0)=50, Z(0)=0, N=10^5, \sigma_X=100, \sigma_Y=5)$.
+The points are the estimates of $\rr$ and the lines denote $\pm 2\cdot $Std.
+Err.
+The vertical line is the true $\rr$ value.}
+	\end{center}
+\end{figure}
+
+\begin{table}[H]
+
+	\centering
+	\begin{tabular}[t]{l|r|r}
+		\hline
+		Model & Estimate & Std. Dev\\
+		\hline
+		LS & 1.0514 & 0.3785\\
+		\hline
+		ReLS & 1.0514 & 0.0259\\
+		\hline
+		LMA & 1.2300 & 0.0572\\
+		\hline
+		LMAT & 1.1471 & 0.1751\\
+		\hline
+		IPR & 2.2972 & 7.1420\\
+		\hline
+		SIPR & 27.4297 & $<$ 1e-04\\
+		\hline
+		LL & 1.1880 & 0.0006\\
+		\hline
+		MC & 0.9486 & $<$ 1e-04\\
+		\hline
+		SB & 1.3647 & 0.0611\\
+		\hline
+	\end{tabular}
+	\caption{Linear Data, $\rr$ Estimates and Standard Errors}
+\end{table}
+
+\begin{table}[H]
+
+	\centering
+	\begin{tabular}[t]{l|r|r}
+		\hline
+		Model & Estimate & Std. Dev\\
+		\hline
+		LS & 1.7691 & 0.4216\\
+		\hline
+		ReLS & 1.7691 & 0.2791\\
+		\hline
+		LMA & -0.0201 & 0.1128\\
+		\hline
+		LMAT & 0.9972 & 2.5752\\
+		\hline
+		IPR & 5.9240 & 17.8114\\
+		\hline
+		SIPR & 15.0836 & $<$ 1e-04\\
+		\hline
+		LL & 1.9553 & 0.0164\\
+		\hline
+		MC & 0.9486 & $<$ 1e-04\\
+		\hline
+		SB & 0.2581 & 0.0266\\
+		\hline
+	\end{tabular}
+	\caption{4th Order Model, $\rr$ Estimates and Standard Errors}
+\end{table}
+
+\begin{table}[H]
+
+	\centering
+	\begin{tabular}[t]{l|r|r}
+		\hline
+		Model & Estimate & Std. Dev\\
+		\hline
+		LS & 1.2739 & 0.0977\\
+		\hline
+		ReLS & 1.2741 & 0.0337\\
+		\hline
+		LMA & 1.2068 & 0.0481\\
+		\hline
+		LMAT & 1.2002 & 0.0041\\
+		\hline
+		IPR & 1.2086 & 1.1254\\
+		\hline
+		SIPR & 1.2944 & 0.2158\\
+		\hline
+		LL & 1.2587 & 0.0010\\
+		\hline
+		MC & 0.9486 & $<$ 1e-04\\
+		\hline
+		SB & 1.3209 & 0.0602\\
+		\hline
+	\end{tabular}
+	\caption{Linear SIR Data, $\rr$ Estimates and Standard Errors}
+\end{table}
+
+\subsubsection{Autoregressive Errors}
+
+\begin{figure}[H]
+	\begin{center}
+		\includegraphics[scale=0.5]{images/other_ar.jpeg}
+		\caption{Forest plots of estimates of $\rr$ for the \xxsir methods for the Poly1, Poly4, and LSIR data sets $(\beta=.06, \gamma=.03, T=365, X(0)=99950, Y(0)=50, Z(0)=0, N=10^5, \sigma_X=100, \sigma_Y=5)$.
+The points are the estimates of $\rr$ and the lines denote $\pm 2\cdot $Std.
+Err.
+The vertical line is the true $\rr$ value.}
+	\end{center}
+\end{figure}
+
+\begin{table}[H]
+
+	\centering
+	\begin{tabular}[t]{l|r|r}
+		\hline
+		Model & Estimate & Std. Dev\\
+		\hline
+		LS & 1.0519 & 0.3775\\
+		\hline
+		ReLS & 1.0518 & 0.0260\\
+		\hline
+		LMA & 1.2766 & 0.1573\\
+		\hline
+		LMAT & 1.1558 & 0.0687\\
+		\hline
+		IPR & 4.3348 & 41.3251\\
+		\hline
+		SIPR & 1.4988 & $<$ 1e-04\\
+		\hline
+		LL & 1.1876 & 0.0006\\
+		\hline
+		MC & 0.9486 & $<$ 1e-04\\
+		\hline
+		SB & 1.4979 & 0.0641\\
+		\hline
+	\end{tabular}
+	\caption{Linear Data, $\rr$ Estimates and Standard Errors}
+\end{table}
+
+\begin{table}[H]
+
+	\centering
+	\begin{tabular}[t]{l|r|r}
+		\hline
+		Model & Estimate & Std. Dev\\
+		\hline
+		LS & 1.7690 & 0.4228\\
+		\hline
+		ReLS & 1.7691 & 0.2799\\
+		\hline
+		LMA & 0.3565 & 0.1259\\
+		\hline
+		LMAT & 3.2425 & 26.2628\\
+		\hline
+		IPR & 6.2685 & 20.5362\\
+		\hline
+		SIPR & 2.7662 & $<$ 1e-04\\
+		\hline
+		LL & 1.9551 & 0.0165\\
+		\hline
+		MC & 0.9486 & $<$ 1e-04\\
+		\hline
+		SB & 0.2132 & 0.0242\\
+		\hline
+	\end{tabular}
+	\caption{4th Order Model, $\rr$ Estimates and Standard Errors}
+\end{table}
+
+\begin{table}[H]
+
+	\centering
+	\begin{tabular}[t]{l|r|r}
+		\hline
+		Model & Estimate & Std. Dev\\
+		\hline
+		LS & 1.2737 & 0.1000\\
+		\hline
+		ReLS & 1.2742 & 0.0346\\
+		\hline
+		LMA & 1.2146 & 0.0627\\
+		\hline
+		LMAT & 1.2032 & 0.0148\\
+		\hline
+		IPR & 1.3799 & 1.3114\\
+		\hline
+		SIPR & 1.2967 & 0.2216\\
+		\hline
+		LL & 1.2589 & 0.0010\\
+		\hline
+		MC & 0.9486 & $<$ 1e-04\\
+		\hline
+		SB & 1.2039 & 0.0574\\
+		\hline
+	\end{tabular}
+	\caption{Linear SIR Data, $\rr$ Estimates and Standard Errors}
+\end{table}
+
+\subsubsection{Autoregressive Monotonic Errors}
+
+\begin{figure}[H]
+	\begin{center}
+		\includegraphics[scale=0.5]{images/other_arm.jpeg}
+		\caption{Forest plots of estimates of $\rr$ for the \xxsir methods for the Poly1, Poly4, and LSIR data sets $(\beta=.06, \gamma=.03, T=365, X(0)=99950, Y(0)=50, Z(0)=0, N=10^5, \sigma_X=100, \sigma_Y=5)$.
+The points are the estimates of $\rr$ and the lines denote $\pm 2\cdot $Std.
+Err.
+The vertical line is the true $\rr$ value.}
+	\end{center}
+\end{figure}
+
+\begin{table}[H]
+
+	\centering
+	\begin{tabular}[t]{l|r|r}
+		\hline
+		Model & Estimate & Std. Dev\\
+		\hline
+		LS & 1.0513 & 0.3783\\
+		\hline
+		ReLS & 1.0513 & 0.0257\\
+		\hline
+		LMA & 1.2569 & 0.1132\\
+		\hline
+		LMAT & 1.1537 & 0.0818\\
+		\hline
+		IPR & 4.1943 & 36.2011\\
+		\hline
+		SIPR & -14.1352 & $<$ 1e-04\\
+		\hline
+		LL & 1.1878 & 0.0006\\
+		\hline
+		MC & 0.9486 & $<$ 1e-04\\
+		\hline
+		SB & 1.4700 & 0.0635\\
+		\hline
+	\end{tabular}
+	\caption{Linear Data, $\rr$ Estimates and Standard Errors}
+\end{table}
+
+\begin{table}[H]
+
+	\centering
+	\begin{tabular}[t]{l|r|r}
+		\hline
+		Model & Estimate & Std. Dev\\
+		\hline
+		LS & 1.7691 & 0.4216\\
+		\hline
+		ReLS & 1.7691 & 0.2789\\
+		\hline
+		LMA & -0.0161 & 0.1099\\
+		\hline
+		LMAT & 1.3810 & 8.5140\\
+		\hline
+		IPR & 5.8926 & 17.3418\\
+		\hline
+		SIPR & 14.9726 & $<$ 1e-04\\
+		\hline
+		LL & 1.9553 & 0.0164\\
+		\hline
+		MC & 0.9486 & $<$ 1e-04\\
+		\hline
+		SB & 0.2586 & 0.0266\\
+		\hline
+	\end{tabular}
+	\caption{4th Order Model, $\rr$ Estimates and Standard Errors}
+\end{table}
+
+\begin{table}[H]
+
+	\centering
+	\begin{tabular}[t]{l|r|r}
+		\hline
+		Model & Estimate & Std. Dev\\
+		\hline
+		LS & 1.2740 & 0.0971\\
+		\hline
+		ReLS & 1.2738 & 0.0334\\
+		\hline
+		LMA & 0.9503 & 0.1067\\
+		\hline
+		LMAT & 1.1959 & 0.0247\\
+		\hline
+		IPR & 1.2085 & 1.0031\\
+		\hline
+		SIPR & 1.3045 & 0.2300\\
+		\hline
+		LL & 1.2585 & 0.0010\\
+		\hline
+		MC & 0.9486 & $<$ 1e-04\\
+		\hline
+		SB & 1.2588 & 0.0587\\
+		\hline
+	\end{tabular}
+	\caption{Linear SIR Data, $\rr$ Estimates and Standard Errors}
+\end{table}
+
+\section{Variance Estimation Methods}\label{sec:sim-var-res}
 We examine the variance methods used to estimate CIs for our estimates of $\rr$.
@@
 		\hline
-		LS & 2.0000175 & 0.0002734 & 1.9999190 & 0.0058322\\
-		\hline
-		ReLS & 2.0000289 & 0.0003092 & 1.9997083 & 0.0052113\\
-		\hline
-		SIRLMA & 1.8497054 & 0.3338774 & 1.997416 & 0.3900006\\
-		\hline
-		LMAT & 1.4562107 & 0.3576736 & 1.398016 & 0.6028170 \\
-		\hline
-		IPR & 4.5214294 & 0.3705241 & 2.592974 & 9.040841\\
-		\hline
-		SIPR & 1.7972938 & 0.0951042 & 0.9251762 & $<$ 1e-07 \\
-		\hline
-		LL & 2.0000079 & 0.0003191 & 1.9998403 & 0.0002227\\
-		\hline
-		MC & 0.9485555 & $<$ 1e-07 & 0.9485555 & $<$ 1e-07\\
-		\hline
-		SB & 1.1330264 & 0.4163031 & 1.0000000 & 0.0523424\\
+		LS & 2.0000 & 0.0003 & 1.9999 & 0.0058\\
+		\hline
+		ReLS & 2.0000 & 0.0003 & 1.9997 & 0.0052\\
+		\hline
+		LMA & 1.8497 & 0.3339 & 1.9974 & 0.3900\\
+		\hline
+		LMAT & 1.4562 & 0.3577 & 1.3980 & 0.6028 \\
+		\hline
+		IPR & 4.5214 & 0.3705 & 2.5930 & 9.0408\\
+		\hline
+		SIPR & 1.7973 & 0.0951 & 0.9251 & $<$ 1e-04 \\
+		\hline
+		LL & 2.0000 & 0.0003 & 1.9998 & 0.0002\\
+		\hline
+		MC & 0.9486 & $<$ 1e-04 & 0.9486 & $<$ 1e-04\\
+		\hline
+		SB & 1.1330 & 0.4163 & 1.0000 & 0.0523\\
 		\hline
@@
 \end{table}
+
+\subsection{Delta Method}
 
@@
 		\hline
-		LS & 1.999903 & 0.0003271 & 1.9999190 & 0.0058322\\
-		\hline
-		ReLS & 1.999802 & 0.0002793 & 1.9997083 & 0.0052113\\
-		\hline
-		SIRLMA & -0.9975459 & 21.71612 & 1.997416 & 0.3900006\\
-		\hline
-		LMAT & 1.513824 & 1.048486 & 1.398016 & 0.6028170 \\
-		\hline
-		SIPR & 0.7761622 & 2.662394 & 0.9251762 & $<$ 1e-07 \\
+		LS & 1.9999 & 0.0003 & 1.9999 & 0.0058\\
+		\hline
+		ReLS & 1.9998 & 0.0003 & 1.9997 & 0.0052\\
+		\hline
+		LMA & -0.9975 & 21.7161 & 1.9974 & 0.3900\\
+		\hline
+		LMAT & 1.5138 & 1.0485 & 1.3980 & 0.6028 \\
+		\hline
+		SIPR & 0.7762 & 2.6624 & 0.9252 & $<$ 1e-04 \\
 		\hline
@@
       \end{figure}
+
+      \subsection{Posterior}
+
+      \section{Application to USA H1N1 data}\label{sec:real-data}
+      We apply our methods using data from the 2009 H1N1 pandemic in the United States.
+
+      \subsection{USA H1N1 Influenza 2009 Data}
+      The data is from the Center of Disease Control and Prevention's (CDC) FluView \citep{cdc-fluview}.
+The data source is ILINet and is the national level data.
+We use the same date range as in \cite{towers2009}, namely Epiweeks 21-33 (May 23, 2009-August 22, 2009).
+From the United States Census Bureau, we use the population estimate of the USA as of April 1, 2010 of $N=308,740,000$ individuals \citep{census-2010}.
+The main features in this data set are weighted influenza like illness (wILI) for a given week (Epiweek).
+The wILI values are estimated based on the number of patients diagnosed with influenza like illness and weighted based on the reporting locations.
+To obtain weekly incidence from wILI, we use $J(t) = \textnormal{wILI}(t) / 100 \cdot N$.
+Then to obtain prevalence $I(t)$, we use the relation of $I(t) = \gamma^{-1}J(t)$, where $\gamma$ is the inverse of the expected time to recovery, which we here use as $\gamma^{-1} = \frac{3}{7}=.429$ weeks \citep{vespignani2007}.
+Then $S(t) = N - \sum_{s=0}^{t}J(s)$, the total population size minus the cumulative sum of the incidence, and $R(t) = N - S(t) - I(t)$.
+As prior knowledge of $\gamma$ is used to estimate our data for $I$, we expect it to influence our estimates of $\rr$.
+The data is summarized in Table \ref{tab:h1n1-data}.
+
+% Please add the following required packages to your document preamble:
+% \usepackage{booktabs}
+\begin{table}[H]
+\centering
+\begin{tabular}{@{}ll@{}}
+\toprule
+Feature       & Notes                                                                   \\ \midrule
+Disease       & H1N1 Pandemic Influenza                                                 \\
+Source        & CDC FluView (\url{https://www.cdc.gov/flu/weekly/})      \\
+Year          & 2009                                                                    \\
+Date(s)          & May 23-August 22 (Epiweeks 21-33)                                       \\
+Location      & USA                                                                     \\
+Population    & $N$=308,740,000 people                                                  \\
+  Avg. Duration $\left ( \gamma^{-1}\right )$ & $\frac{3}{7}$ weeks\\
+  Raw data & Weekly wILI reports\\
+  Final data & S$(t)$, I$(t)$, R$(t)$ \\\bottomrule
+\end{tabular}
+\caption{Summary of the data used to estimate $\rr$ during the 2009 H1N1 influenza pandemic in the USA.  We follow the data collection procedure described in \cite{towers2009}.}
+\label{tab:h1n1-data}
+\end{table}
+
+\begin{figure}[H]
+  \centering
+  \includegraphics[width=\textwidth]{images/h1n1-sir.pdf}
+  \caption{Data from H1N1 influenza 2009 in the USA.  Weekly wILI is reported, which we transform into S, I, and R counts for epiweeks 21-33 using $\gamma^{-1} = 3/7$ weeks.}
+  \end{figure}
+
+      \subsection{Results}
+      The results are shown below.  \cite{towers2009} report an estimate of $\hat{\rr} \in [1.50, 1.70]$.
 
@@
 \begin{itemize}
-\item processes (e.g. Gaussian error, AR error, and PAVA)
   \item LS and ReLS provide the most accurate estimates of $\rr$
@@
 
-\section{Fitting the SIR Model}
-
-\subsection{SIR Least Squares}
-
-\begin{table}[H]
-
-	\caption{\label{tab:}$\rr$ Estimates and Std. Errs, LS Model,
-		$X_0 = 99950, Y_0 = 50$, $\sigma_X = 100, \sigma_Y = 5$, $\beta = 0.06$}
-	\centering
-	\begin{footnotesize}
-		\begin{tabular}[t]{l|r|r|r|r|r|r|r|r}
-			\hline
-			Data Set & Norm Est & Norm SE & Norm-M Est & Norm-M SE & AR Est & AR SE & AR-M Est & AR-M SE\\
-			\hline
-			$\gamma$ = 0.001 & 59.9764682 & 0.0434705 & 59.9764682 & 0.0386733 & 60.0151171 & 0.0697888 & 59.9589126 & 0.0631448\\
-			\hline
-			$\gamma$ = 0.04 & 1.5000763 & 0.0126817 & 1.5001978 & 0.0090880 & 1.4989700 & 0.0182852 & 1.4993603 & 0.0165168\\
-			\hline
-			$\gamma$ = 0.06 & 1.0043054 & 0.4140424 & 0.9955554 & 0.1086448 & 0.9915431 & 0.6272644 & 0.9941392 & 0.1891812\\
-			\hline
-			$\gamma$ = 0.24 & 0.2363274 & 4.1042813 & 0.2327691 & 0.6701185 & -0.0695410 & 2.8291362 & 0.3021083 & 0.5657139\\
-			\hline
-		\end{tabular}
-	\end{footnotesize}
-\end{table}
-
-\begin{table}[H]
-
-	\caption{\label{tab:}$\rr$ Estimates and Std. Errs, LS Model,
-		$\sigma_X = 100, \sigma_Y = 5$, $\beta = 0.06, \gamma = 0.03$}
-	\centering
-	\begin{footnotesize}
-		\begin{tabular}[t]{l|r|r|r|r|r|r|r|r}
-			\hline
-			Data Set & Norm Est & Norm SE & Norm-M Est & Norm-M SE & AR Est & AR SE & AR-M Est & AR-M SE\\
-			\hline
-			$X_0 = 99000, Y_0 = 1000$ & 1.999588 & 0.0049609 & 2.000284 & 0.0046294 & 2.000461 & 0.0078398 & 2.000122 & 0.0068626\\
-			\hline
-			$X_0 = 99900, Y_0 = 100$ & 2.022501 & 0.3064263 & 2.021865 & 0.3064148 & 2.022296 & 0.3063477 & 2.022665 & 0.3068512\\
-			\hline
-			$X_0 = 99990, Y_0 = 10$ & 2.116865 & 0.6626380 & 2.116551 & 0.6621184 & 2.116959 & 0.6623142 & 2.116803 & 0.6620448\\
-			\hline
-			$X_0 = 99999, Y_0 = 1$ & 2.468304 & 1.5370731 & 2.466776 & 1.5348981 & 2.470607 & 1.5371890 & 2.466700 & 1.5348542\\
-			\hline
-		\end{tabular}
-	\end{footnotesize}
-\end{table}
-
-\begin{table}[H]
-	\caption{\label{tab:}$\rr$ Estimates and Std. Errs, LS Model,
-		$X_0 = 99950, Y_0 = 50$, $\beta = 0.06, \gamma = 0.03$}
-	\centering
-	\begin{footnotesize}
-		\begin{tabular}[t]{l|r|r|r|r|r|r|r|r}
-			\hline
-			Data Set & Norm Est & Norm SE & Norm-M Est & Norm-M SE & AR Est & AR SE & AR-M Est & AR-M SE\\
-			\hline
-			$\sigma_X = 50, \sigma_Y = 2$ & 1.999755 & 0.0026502 & 1.999834 & 0.0026397 & 1.999555 & 0.0037664 & 1.999038 & 0.0036473\\
-			\hline
-			$\sigma_X = 500, \sigma_Y = 20$ & 2.001870 & 0.0259428 & 1.997748 & 0.0162536 & 2.002792 & 0.0416203 & 1.997918 & 0.0257997\\
-			\hline
-			$\sigma_X = 2500, \sigma_Y = 100$ & 1.998088 & 0.1353803 & 1.996950 & 0.0541332 & 1.999686 & 0.1893888 & 2.002135 & 0.0927952\\
-			\hline
-			$\sigma_X = 10000, \sigma_Y = 500$ & 1.993455 & 0.5784530 & 1.979966 & 0.1204126 & 2.033934 & 0.7559710 & 2.066706 & 0.3382834\\
-			\hline
-		\end{tabular}
-	\end{footnotesize}
-\end{table}
-
-\begin{table}[H]
-
-	\caption{\label{tab:}$\rr$ Estimates and Std. Errs, LS Model,
-		$X_0 = 99950, Y_0 = 50$, $\sigma_X = 100, \sigma_Y = 5$}
-	\centering
-	\begin{footnotesize}
-		\begin{tabular}[t]{l|r|r|r|r|r|r|r|r}
-			\hline
-			Data Set & Norm Est & Norm SE & Norm-M Est & Norm-M SE & AR Est & AR SE & AR-M Est & AR-M SE\\
-			\hline
-			1st Order & 1.051438 & 0.3782910 & 1.051420 & 0.3784628 & 1.051864 & 0.3775363 & 1.051293 & 0.3782570\\
-			\hline
-			4th Order & 1.769347 & 0.4226615 & 1.769136 & 0.4215768 & 1.769002 & 0.4227615 & 1.769136 & 0.4215910\\
-			\hline
-			Linear SIR & 1.274258 & 0.0979741 & 1.273917 & 0.0976920 & 1.273654 & 0.1000387 & 1.274007 & 0.0970809\\
-			\hline
-		\end{tabular}
-	\end{footnotesize}
-\end{table}
-
-\begin{table}[H]
-
-	\caption{\label{tab:}$\rr$ Estimates and Std. Errs, LS Model,
-		$\beta = 0.06, \gamma = 0.03$, $X_0 = 99950, Y_0 = 50$, $\sigma_X = 100, \sigma_Y = 5$}
-	\centering
-	\begin{tabular}[t]{l|r|r|r|r|r|r|r|r}
-		\hline
-		Data Set & Norm Est & Norm SE & Norm-M Est & Norm-M SE & AR Est & AR SE & AR-M Est & AR-M SE\\
-		\hline
-		T = 200 & 1.999339 & 0.0286060 & 2.000583 & 0.0204020 & 2.0006557 & 0.0357152 & 2.000722 & 0.0364659\\
-		\hline
-		T = 100 & 1.990754 & 0.2922125 & 1.993375 & 0.1463550 & 1.9098406 & 0.3811467 & 1.967388 & 0.2324481\\
-		\hline
-		T = 50 & 1.978167 & 1.1596532 & 1.856227 & 0.3825275 & 2.2720601 & 1.2900031 & 1.699976 & 0.6064311\\
-		\hline
-		T = 20 & 38.689552 & 106.2705193 & 5.036600 & 1.3672212 & 0.0098756 & 2.1687945 & 1.681588 & 0.4747083\\
-		\hline
-	\end{tabular}
-\end{table}
-
-\begin{table}[H]
-
-	\caption{\label{tab:}$\rr$ Estimates and Std. Errs, LS Model,
-		$\beta = 0.06, \gamma = 0.03$, $X_0 = 99950, Y_0 = 50$, $\sigma_X = 100, \sigma_Y = 5$}
-	\centering
-	\begin{tabular}[t]{l|r|r|r|r|r|r|r|r}
-		\hline
-		Data Set & Norm Est & Norm SE & Norm-M Est & Norm-M SE & AR Est & AR SE & AR-M Est & AR-M SE\\
-		\hline
-		$X_0 = 99, Y_0 = 1$ & 1.986253 & 4.8619876 & 2.038011 & 0.4537907 & 2.106225 & 7.1198939 & 2.105517 & 1.5447687\\
-		\hline
-		$X_0 = 990, Y_0 = 10$ & 1.997222 & 0.5406162 & 2.017454 & 0.1429851 & 2.122109 & 0.7362701 & 2.097491 & 0.2921481\\
-		\hline
-		$X_0 = 9995, Y_0 = 5$ & 2.000774 & 0.0590423 & 1.995300 & 0.0239310 & 2.000782 & 0.0744317 & 1.996394 & 0.0489651\\
-		\hline
-		$X_0 = 999500, Y_0 = 500$ & 2.000050 & 0.0005467 & 1.999834 & 0.0005748 & 1.999833 & 0.0008762 & 1.999959 & 0.0007396\\
-		\hline
-	\end{tabular}
-\end{table}
-
-\subsection{SIR Reparameterization}
-
-\begin{table}[H]
-
-	\caption{\label{tab:}$\rr$ Estimates and Std. Errs, ReLS Model,
-		$X_0 = 99950, Y_0 = 50$, $\sigma_X = 100, \sigma_Y = 5$, $\beta = 0.06$}
-	\centering
-	\begin{footnotesize}
-		\begin{tabular}[t]{l|r|r|r|r|r|r|r|r}
-			\hline
-			Data Set & Norm Est & Norm SE & Norm-M Est & Norm-M SE & AR Est & AR SE & AR-M Est & AR-M SE\\
-			\hline
-			$\gamma$ = 0.001 & 59.995188 & 0.4343984 & 59.9951876 & 0.3874879 & 60.025452 & 0.6911933 & 59.9960910 & 0.6214841\\
-			\hline
-			$\gamma$ = 0.04 & 1.499883 & 0.0062529 & 1.5001705 & 0.0044936 & 1.499068 & 0.0090313 & 1.4993177 & 0.0081651\\
-			\hline
-			$\gamma$ = 0.06 & 1.004306 & 0.0501444 & 0.9955446 & 0.0117215 & 0.991558 & 0.0645788 & 0.9941511 & 0.0190525\\
-			\hline
-			$\gamma$ = 0.24 & 4.179525 & 2.7455569 & 3.5709590 & 0.3171284 & 5.192951 & 3.5249847 & 9.8415034 & 2.5646198\\
-			\hline
-		\end{tabular}
-	\end{footnotesize}
-\end{table}
-
-\begin{table}[H]
-
-	\caption{\label{tab:}$\rr$ Estimates and Std. Errs, ReLS Model,
-		$\sigma_X = 100, \sigma_Y = 5$, $\beta = 0.06, \gamma = 0.03$}
-	\centering
-	\begin{footnotesize}
-		\begin{tabular}[t]{l|r|r|r|r|r|r|r|r}
-			\hline
-			Data Set & Norm Est & Norm SE & Norm-M Est & Norm-M SE & AR Est & AR SE & AR-M Est & AR-M SE\\
-			\hline
-			$X_0 = 99000, Y_0 = 1000$ & 1.999712 & 0.0034303 & 2.000274 & 0.0032086 & 2.000082 & 0.0053846 & 2.000014 & 0.0047509\\
-			\hline
-			$X_0 = 99900, Y_0 = 100$ & 2.022359 & 0.2698034 & 2.021866 & 0.2694871 & 2.022178 & 0.2696002 & 2.022530 & 0.2701860\\
-			\hline
-			$X_0 = 99990, Y_0 = 10$ & 2.117059 & 0.7291460 & 2.116858 & 0.7290133 & 2.116829 & 0.7288841 & 2.116534 & 0.7286543\\
-			\hline
-			$X_0 = 99999, Y_0 = 1$  & 2.469189 & 2.7293667 & 2.466967 & 2.7219457 & 2.470707 & 2.7245397 & 2.466483 & 2.7349248\\
-			\hline
-		\end{tabular}
-	\end{footnotesize}
-\end{table}
-
-\begin{table}[H]
-
-	\caption{\label{tab:}$\rr$ Estimates and Std. Errs, ReLS Model,
-		$X_0 = 99950, Y_0 = 50$, $\beta = 0.06, \gamma = 0.03$}
-	\centering
-	\begin{footnotesize}
-		\begin{tabular}[t]{l|r|r|r|r|r|r|r|r}
-			\hline
-			Data Set & Norm Est & Norm SE & Norm-M Est & Norm-M SE & AR Est & AR SE & AR-M Est & AR-M SE\\
-			\hline
-			$\sigma_X = 50, \sigma_Y = 2$ & 1.999829 & 0.0023646 & 1.999946 & 0.0023591 & 1.999885 & 0.0033514 & 1.999483 & 0.0032200\\
-			\hline
-			$\sigma_X = 500, \sigma_Y = 20$ & 2.002305 & 0.0232636 & 1.997996 & 0.0144744 & 2.003065 & 0.0372826 & 1.997732 & 0.0230069\\
-			\hline
-			$\sigma_X = 2500, \sigma_Y = 100$ & 1.998145 & 0.1207935 & 1.997147 & 0.0481545 & 2.000194 & 0.1680126 & 2.002244 & 0.0826643\\
-			\hline
-			$\sigma_X = 10000, \sigma_Y = 500$ & 1.993660 & 0.5115266 & 1.979973 & 0.1053955 & 2.033636 & 0.7059701 & 2.066476 & 0.3172823\\
-			\hline
-		\end{tabular}
-	\end{footnotesize}
-\end{table}
-
-\begin{table}[H]
-
-	\caption{\label{tab:}$\rr$ Estimates and Std. Errs, ReLS Model,
-		$X_0 = 99950, Y_0 = 50$, $\sigma_X = 100, \sigma_Y = 5$}
-	\centering
-	\begin{footnotesize}
-		\begin{tabular}[t]{l|r|r|r|r|r|r|r|r}
-			\hline
-			Data Set & Norm Est & Norm SE & Norm-M Est & Norm-M SE & AR Est & AR SE & AR-M Est & AR-M SE\\
-			\hline
-			1st Order & 1.051435 & 0.0259416 & 1.051397 & 0.0258690 & 1.051845 & 0.0260146 & 1.051255 & 0.0257166\\
-			\hline
-			4th Order & 1.769331 & 0.2796903 & 1.769079 & 0.2791191 & 1.769095 & 0.2799270 & 1.769111 & 0.2788789\\
-			\hline
-			Linear SIR & 1.274314 & 0.0339775 & 1.274075 & 0.0337499 & 1.274174 & 0.0345629 & 1.273799 & 0.0334192\\
-			\hline
-		\end{tabular}
-	\end{footnotesize}
-\end{table}
-
-\begin{table}[H]
-
-	\caption{\label{tab:}$\rr$ Estimates and Std. Errs, ReLS Model,
-		$\beta = 0.06, \gamma = 0.03$, $X_0 = 99950, Y_0 = 50$, $\sigma_X = 100, \sigma_Y = 5$}
-	\centering
-	\begin{tabular}[t]{l|r|r|r|r|r|r|r|r}
-		\hline
-		Data Set & Norm Est & Norm SE & Norm-M Est & Norm-M SE & AR Est & AR SE & AR-M Est & AR-M SE\\
-		\hline
-		T = 200 & 1.999428 & 0.0312814 & 2.000386 & 0.0223765 & 2.000419 & 0.0390441 & 2.000199 & 0.0402489\\
-		\hline
-		T = 100 & 1.991014 & 0.3940097 & 1.993501 & 0.1974773 & 1.910435 & 0.4773240 & 1.967858 & 0.3065049\\
-		\hline
-		T = 50 & 1.977508 & 1.8877725 & 1.856480 & 0.5810168 & 2.273215 & 2.3912219 & 1.700025 & 0.7995792\\
-		\hline
-		T = 20 & 22.964587 & 6.9816713 & 5.044869 & 2.1398099 & NA & NA & 1.681269 & 0.6543603\\
-		\hline
-	\end{tabular}
-\end{table}
-
-\begin{table}[H]
-
-	\caption{\label{tab:}$\rr$ Estimates and Std. Errs, ReLS Model,
-		$\beta = 0.06, \gamma = 0.03$, $X_0 = 99950, Y_0 = 50$, $\sigma_X = 100, \sigma_Y = 5$}
-	\centering
-	\begin{tabular}[t]{l|r|r|r|r|r|r|r|r}
-		\hline
-		Data Set & Norm Est & Norm SE & Norm-M Est & Norm-M SE & AR Est & AR SE & AR-M Est & AR-M SE\\
-		\hline
-		$X_0 = 99, Y_0 = 1$ & 1.987105 & 3.4568905 & 2.037779 & 0.3348238 & 1.658070 & 2.0918740 & 2.105172 & 1.1781993\\
-		\hline
-		$X_0 = 990, Y_0 = 10$ & 1.997065 & 0.3756188 & 2.017454 & 0.1020655 & 2.121902 & 0.5659528 & 2.097579 & 0.2210569\\
-		\hline
-		$X_0 = 9995, Y_0 = 5$ & 2.000519 & 0.0528419 & 1.995631 & 0.0212989 & 2.000955 & 0.0665574 & 1.996249 & 0.0434920\\
-		\hline
-		$X_0 = 999500, Y_0 = 500$ & 1.999946 & 0.0004956 & 1.999946 & 0.0005270 & 2.000104 & 0.0007140 & 1.999734 & 0.0006529\\
-		\hline
-	\end{tabular}
-\end{table}
-
-\subsection{SIR Linear Model Approximation}
-
-\begin{table}[H]
-
-	\caption{\label{tab:}$\rr$ Estimates and Std. Errs, SIRLMA Model,
-		$X_0 = 99950, Y_0 = 50$, $\sigma_X = 100, \sigma_Y = 5$, $\beta = 0.06$}
-	\centering
-	\begin{footnotesize}
-		\begin{tabular}[t]{l|r|r|r|r|r|r|r|r}
-			\hline
-			Data Set & Norm Est & Norm SE & Norm-M Est & Norm-M SE & AR Est & AR SE & AR-M Est & AR-M SE\\
-			\hline
-			$\gamma$ = 0.001 & -117.9694856 & 101.1885824 & -55.3748542 & 10.9872264 & -19.2057935 & 5.9751126 & -183.9423408 & 277.8291528\\
-			\hline
-			$\gamma$ = 0.04 & 1.1087585 & 0.0863592 & -3.9728056 & 1116.1887158 & 0.9757401 & 0.0213324 & 0.7333872 & 0.1787735\\
-			\hline
-			$\gamma$ = 0.06 & 1.0829034 & 0.1533121 & 0.9905684 & 0.1182606 & 0.9279273 & 0.0605799 & 1.0397917 & 0.0359875\\
-			\hline
-			$\gamma$ = 0.24 & 0.7074613 & 0.2028765 & 0.5216530 & 0.1715628 & 0.8749860 & 0.0590753 & -49.7813412 & 5303.0302298\\
-			\hline
-		\end{tabular}
-	\end{footnotesize}
-\end{table}
-
-\begin{table}[H]
-
-	\caption{\label{tab:}$\rr$ Estimates and Std. Errs, SIRLMA Model,
-		$\sigma_X = 100, \sigma_Y = 5$, $\beta = 0.06, \gamma = 0.03$}
-	\centering
-	\begin{footnotesize}
-		\begin{tabular}[t]{l|r|r|r|r|r|r|r|r}
-			\hline
-			Data Set & Norm Est & Norm SE & Norm-M Est & Norm-M SE & AR Est & AR SE & AR-M Est & AR-M SE\\
-			\hline
-			$X_0 = 99000, Y_0 = 1000$ & -87.5945650 & 66413.128046 & -0.9999761 & 1.422640 & 4.1817846 & 3.8812899 & -10.8458320 & 288.1028935\\
-			\hline
-			$X_0 = 99900, Y_0 = 100$ & 0.5237886 & 1.632971 & 0.2436375 & 3.754091 & 0.7149415 & 0.9374275 & -0.3051038 & 8.8837193\\
-			\hline
-			$X_0 = 99990, Y_0 = 10$ & -33.0051124 & 433.594227 & -1.3382741 & 3.047220 & -0.2463189 & 1.5875310 & 0.4368556 & 0.7686276\\
-			\hline
-			$X_0 = 99999, Y_0 = 1$ & 10.9282204 & 19.297359 & -98.6422337 & 3113.332617 & -0.8194023 & 1.7121668 & 16.6102769 & 18.1045393\\
-			\hline
-		\end{tabular}
-	\end{footnotesize}
-\end{table}
-
-\begin{table}[H]
-
-	\caption{\label{tab:}$\rr$ Estimates and Std. Errs, SIRLMA Model,
-		$X_0 = 99950, Y_0 = 50$, $\beta = 0.06, \gamma = 0.03$}
-	\centering
-	\begin{footnotesize}
-		\begin{tabular}[t]{l|r|r|r|r|r|r|r|r}
-			\hline
-			Data Set & Norm Est & Norm SE & Norm-M Est & Norm-M SE & AR Est & AR SE & AR-M Est & AR-M SE\\
-			\hline
-			$\sigma_X = 50, \sigma_Y = 2$ & 2.0398087 & 0.2773968 & 1.626788 & 0.0864885 & 1.4643659 & 0.0750183 & 2.8515449 & 0.5540996\\
-			\hline
-			$\sigma_X = 500, \sigma_Y = 20$ & 1.6600538 & 1.0693161 & 1.684919 & 0.2571470 & 1.1531754 & 0.0997937 & -4.9533560 & 32.2198830\\
-			\hline
-			$\sigma_X = 2500, \sigma_Y = 100$ & 0.4651914 & 7.5508447 & 4.206167 & 66.2206340 & 1.0391560 & 0.0166472 & -0.1164499 & 1.9126019\\
-			\hline
-			$\sigma_X = 10000, \sigma_Y = 500$ & 1.0899949 & 0.1202859 & 3.976724 & 4.4279629 & 0.6981674 & 0.2429844 & 1.3944715 & 1.0825033\\
-			\hline
-		\end{tabular}
-	\end{footnotesize}
-\end{table}
-
-\begin{table}[H]
-
-	\caption{\label{tab:}$\rr$ Estimates and Std. Errs, SIRLMA Model,
-		$X_0 = 99950, Y_0 = 50$, $\sigma_X = 100, \sigma_Y = 5$}
-	\centering
-	\begin{footnotesize}
-		\begin{tabular}[t]{l|r|r|r|r|r|r|r|r}
-			\hline
-			Data Set & Norm Est & Norm SE & Norm-M Est & Norm-M SE & AR Est & AR SE & AR-M Est & AR-M SE\\
-			\hline
-			1st Order & 1.2559522 & 0.1296269 & 1.2209276 & 0.0597736 & 1.2670255 & 0.1588846 & 1.2476445 & 0.1150267\\
-			\hline
-			4th Order & 0.3632202 & 0.1230895 & -0.0201429 & 0.1133200 & 0.3580085 & 0.1258904 & -0.0161666 & 0.1103940\\
-			\hline
-			Linear SIR & 1.2191443 & 0.1093267 & 1.2067608 & 0.0480986 & 1.2146478 & 0.0627181 & 0.9503153 & 0.1066437\\
-			\hline
-		\end{tabular}
-	\end{footnotesize}
-\end{table}
-
-\begin{table}[H]
-
-	\caption{\label{tab:}$\rr$ Estimates and Std. Errs, SIRLMA Model,
-		$\beta = 0.06, \gamma = 0.03$, $X_0 = 99950, Y_0 = 50$, $\sigma_X = 100, \sigma_Y = 5$}
-	\centering
-	\begin{tabular}[t]{l|r|r|r|r|r|r|r|r}
-		\hline
-		Data Set & Norm Est & Norm SE & Norm-M Est & Norm-M SE & AR Est & AR SE & AR-M Est & AR-M SE\\
-		\hline
-		T = 200 & 0.9496498 & 0.0825419 & 1.052892 & 0.2237478 & 1.0611334 & 0.0328849 & 0.7221436 & 0.1920598\\
-		\hline
-		T = 100 & 1.2374397 & 0.4953607 & 1.202445 & 0.5438859 & 0.9140701 & 0.2563191 & 1.1584230 & 0.3660429\\
-		\hline
-		T = 50 & 0.9720834 & 0.0507265 & 1.924122 & 1.4342312 & 3.7359315 & 59.4952531 & 0.5305768 & 1.0293919\\
-		\hline
-		T = 20 & 1.4418031 & 7.3163878 & 1.121779 & 1.7267097 & -5.2984368 & 65.6502768 & 0.8899288 & 0.7605429\\
-		\hline
-	\end{tabular}
-\end{table}
-
-\begin{table}[H]
-
-	\caption{\label{tab:}$\rr$ Estimates and Std. Errs, SIRLMA Model,
-		$\beta = 0.06, \gamma = 0.03$, $\sigma_X = 100, \sigma_Y = 5$}
-	\centering
-	\begin{tabular}[t]{l|r|r|r|r|r|r|r|r}
-		\hline
-		Data Set & Norm Est & Norm SE & Norm-M Est & Norm-M SE & AR Est & AR SE & AR-M Est & AR-M SE\\
-		\hline
-		$X_0 = 99, Y_0 = 1$ & 0.8023469 & 18.9416598 & 0.8332037 & 0.6714657 & 1.0679323 & 0.0510976 & 1.038611 & 1.5641167\\
-		\hline
-		$X_0 = 990, Y_0 = 10$ & 1.0682053 & 0.0327648 & 0.8256301 & 1.4444021 & 1.2044940 & 0.3878046 & 1.172003 & 0.0511127\\
-		\hline
-		$X_0 = 9995, Y_0 = 5$ & 1.9976309 & 2606.0798123 & 1.3632453 & 0.3709415 & 0.7923714 & 2.5234696 & 1.398184 & 0.2967151\\
-		\hline
-		$X_0 = 999500, Y_0 = 500$ & 1.7316488 & 0.0919930 & 1.7673457 & 0.0976337 & 1.6873477 & 0.0810172 & 1.833159 & 0.1199462\\
-		\hline
-	\end{tabular}
-\end{table}
-
-\subsection{SIR Linear Model Approximation, All Time Points}
-
-\begin{table}[H]
-
-	\caption{\label{tab:}$\rr$ Estimates and Std. Errs, LMAT Model,
-		$X_0 = 99950, Y_0 = 50$, $\sigma_X = 100, \sigma_Y = 5$, $\beta = 0.06$}
-	\centering
-	\begin{footnotesize}
-		\begin{tabular}[t]{l|r|r|r|r|r|r|r|r}
-			\hline
-			Data Set & Norm Est & Norm SE & Norm-M Est & Norm-M SE & AR Est & AR SE & AR-M Est & AR-M SE\\
-			\hline
-			$\gamma$ = 0.001 & -2.8535930 & 193.6526107 & 7.7685771 & 103.2205552 & 76.340382 & 472.4155164 & -69.9067922 & 1603.7792325\\
-			\hline
-			$\gamma$ = 0.04 & 1.4369752 & 0.7498020 & 1.4719115 & 3.0719036 & 1.292978 & 0.3342698 & 1.2769691 & 0.2448445\\
-			\hline
-			$\gamma$ = 0.06 & 0.9847650 & 0.1423066 & 0.9945574 & 0.0160864 & 1.008112 & 0.5987452 & 1.0268632 & 0.5015724\\
-			\hline
-			$\gamma$ = 0.24 & 0.6746914 & 6.2471252 & 0.3871567 & 4.8817556 & 1.786500 & 15.9420420 & 0.2221191 & 5.4112205\\
-			\hline
-		\end{tabular}
-	\end{footnotesize}
-\end{table}
-
-\begin{table}[H]
-
-	\caption{\label{tab:}$\rr$ Estimates and Std. Errs, LMAT Model,
-		$\sigma_X = 100, \sigma_Y = 5$, $\beta = 0.06, \gamma = 0.03$}
-	\centering
-	\begin{footnotesize}
-		\begin{tabular}[t]{l|r|r|r|r|r|r|r|r}
-			\hline
-			Data Set & Norm Est & Norm SE & Norm-M Est & Norm-M SE & AR Est & AR SE & AR-M Est & AR-M SE\\
-			\hline
-			$X_0 = 99000, Y_0 = 1000$ & 0.730610 & 4.6810416 & 0.9720592 & 0.6586069 & 3.163816 & 45.776770 & 0.9367865 & 0.8857030\\
-			\hline
-			$X_0 = 99900, Y_0 = 100$ & 1.236415 & 0.7388933 & 1.1744675 & 1.3957615 & 1.249219 & 0.813166 & 1.1960662 & 0.6589163\\
-			\hline
-			$X_0 = 99990, Y_0 = 10$ & 1.459703 & 10.9219088 & 0.8531016 & 12.5573016 & -3.846910 & 71.297410 & -0.5884883 & 39.4915091\\
-			\hline
-			$X_0 = 99999, Y_0 = 1$ & 1.624645 & 5.0359845 & 0.4174208 & 12.8380228 & 1.301346 & 25.850056 & 3.1746775 & 12.8142151\\
-			\hline
-		\end{tabular}
-	\end{footnotesize}
-\end{table}
-
-\begin{table}[H]
-
-	\caption{\label{tab:}$\rr$ Estimates and Std. Errs, LMAT Model,
-		$X_0 = 99950, Y_0 = 50$, $\beta = 0.06, \gamma = 0.03$}
-	\centering
-	\begin{footnotesize}
-		\begin{tabular}[t]{l|r|r|r|r|r|r|r|r}
-			\hline
-			Data Set & Norm Est & Norm SE & Norm-M Est & Norm-M SE & AR Est & AR SE & AR-M Est & AR-M SE\\
-			\hline
-			$\sigma_X = 50, \sigma_Y = 2$ & 1.414863 & 0.5779651 & 1.509908 & 2.1243768 & 1.3999424 & 0.7193567 & 1.426323 & 0.6782388\\
-			\hline
-			$\sigma_X = 500, \sigma_Y = 20$ & 1.353711 & 1.5619894 & 1.446169 & 0.7223815 & 1.0183856 & 3.5577619 & 1.996902 & 15.7391571\\
-			\hline
-			$\sigma_X = 2500, \sigma_Y = 100$ & 1.287786 & 4.3947658 & 1.365338 & 3.2675366 & 1.1984800 & 1.6107382 & 0.942443 & 4.3914050\\
-			\hline
-			$\sigma_X = 10000, \sigma_Y = 500$ & 4.811335 & 70.6594883 & 1.500486 & 1.0046353 & 0.9241342 & 1.2799734 & 1.097322 & 3.1937985\\
-			\hline
-		\end{tabular}
-	\end{footnotesize}
-\end{table}
-
-\begin{table}[H]
-
-	\caption{\label{tab:}$\rr$ Estimates and Std. Errs, LMAT Model,
-		$X_0 = 99950, Y_0 = 50$, $\sigma_X = 100, \sigma_Y = 5$}
-	\centering
-	\begin{footnotesize}
-		\begin{tabular}[t]{l|r|r|r|r|r|r|r|r}
-			\hline
-			Data Set & Norm Est & Norm SE & Norm-M Est & Norm-M SE & AR Est & AR SE & AR-M Est & AR-M SE\\
-			\hline
-			1st Order & 1.1509447 & 0.1012249 & 1.138621 & 0.1738250 & 1.147184 & 0.0681524 & 1.145189 & 0.0811484\\
-			\hline
-			4th Order & 0.8246203 & 6.0710436 & 1.000597 & 2.5841000 & 3.255953 & 26.3713818 & 1.385674 & 8.5430506\\
-			\hline
-			Linear SIR & 1.2009801 & 0.0031100 & 1.200130 & 0.0040763 & 1.203159 & 0.0148085 & 1.195875 & 0.0246927\\
-			\hline
-		\end{tabular}
-	\end{footnotesize}
-\end{table}
-
-\begin{table}[H]
-
-	\caption{\label{tab:}$\rr$ Estimates and Std. Errs, LMAT Model,
-		$\beta = 0.06, \gamma = 0.03$, $X_0 = 99950, Y_0 = 50$, $\sigma_X = 100, \sigma_Y = 5$}
-	\centering
-	\begin{tabular}[t]{l|r|r|r|r|r|r|r|r}
-		\hline
-		Data Set & Norm Est & Norm SE & Norm-M Est & Norm-M SE & AR Est & AR SE & AR-M Est & AR-M SE\\
-		\hline
-		T = 200 & 3.3442696 & 21.4895422 & 1.8327212 & 0.2855541 & 1.7170357 & 0.4970127 & 2.294206 & 6.234463\\
-		\hline
-		T = 100 & 1.2354159 & 3.0461522 & 2.7980661 & 3.8629146 & 0.7524086 & 2.5622573 & 3.434636 & 14.604145\\
-		\hline
-		T = 50 & 0.7050038 & 2.8719153 & 0.3376738 & 7.9042199 & 0.6897879 & 2.7333396 & 1.274668 & 3.840504\\
-		\hline
-		T = 20 & 1.0078807 & 0.1424793 & 2.7045722 & 4.8031306 & 0.5716857 & 1.3981528 & 1.449643 & 1.542269\\
-		\hline
-	\end{tabular}
-\end{table}
-
-\begin{table}[H]
-
-	\caption{\label{tab:}$\rr$ Estimates and Std. Errs, LMAT Model,
-		$\beta = 0.06, \gamma = 0.03$, $\sigma_X = 100, \sigma_Y = 5$}
-	\centering
-	\begin{tabular}[t]{l|r|r|r|r|r|r|r|r}
-		\hline
-		Data Set & Norm Est & Norm SE & Norm-M Est & Norm-M SE & AR Est & AR SE & AR-M Est & AR-M SE\\
-		\hline
-		$X_0 = 99, Y_0 = 1$ & 1.220874 & 4.7699293 & -0.0614019 & 15.5900050 & -0.0217592 & 18.7796434 & 0.9203832 & 1.5845866\\
-		\hline
-		$X_0 = 990, Y_0 = 10$ & 2.248607 & 29.7561744 & 1.0457445 & 0.6365320 & 1.0880985 & 3.9238850 & -13.6905770 & 281.9114385\\
-		\hline
-		$X_0 = 9995, Y_0 = 5$ & 1.294414 & 1.4304706 & 1.3710979 & 0.6597925 & 1.5307307 & 1.0343598 & 1.3865816 & 1.1453533\\
-		\hline
-		$X_0 = 999500, Y_0 = 500$ & 1.370406 & 0.8821226 & 1.9731686 & 10.8871603 & 1.3885429 & 0.6854633 & 1.4164038 & 0.6347757\\
-		\hline
-	\end{tabular}
-\end{table}
-
-\subsection{Incidence to Prevalence Ratio}
-
-\begin{table}[H]
-
-	\caption{\label{tab:}$\rr$ Estimates and Std. Errs, IPR Model,
-		$X_0 = 99950, Y_0 = 50$, $\sigma_X = 100, \sigma_Y = 5$, $\beta = 0.06$}
-	\centering
-	\begin{footnotesize}
-		\begin{tabular}[t]{l|r|r|r|r|r|r|r|r}
-			\hline
-			Data Set & Norm Est & Norm SE & Norm-M Est & Norm-M SE & AR Est & AR SE & AR-M Est & AR-M SE\\
-			\hline
-			$\gamma$ = 0.001 & 54.146506 & 200.2595 & 24.408675 & 66.166769 & 58.723744 & 219.989055 & 28.5230747 & 120.928788\\
-			\hline
-			$\gamma$ = 0.04 & 4.771508 & 12.8533 & 1.387753 & 2.630779 & 3.613754 & 8.536531 & 1.3965600 & 2.840997\\
-			\hline
-			$\gamma$ = 0.06 & 19.828076 & 29.1272 & 1.121667 & 4.400499 & 15.651052 & 23.539321 & 1.1251847 & 4.900022\\
-			\hline
-			$\gamma$ = 0.24 & 181.015802 & 742.1088 & 6.672760 & 89.936233 & 148.308767 & 740.018239 & 0.8145271 & 11.870908\\
-			\hline
-		\end{tabular}
-	\end{footnotesize}
-\end{table}
-
-\begin{table}[H]
-
-	\caption{\label{tab:}$\rr$ Estimates and Std. Errs, IPR Model,
-		$\sigma_X = 100, \sigma_Y = 5$, $\beta = 0.06, \gamma = 0.03$}
-	\centering
-	\begin{footnotesize}
-		\begin{tabular}[t]{l|r|r|r|r|r|r|r|r}
-			\hline
-			Data Set & Norm Est & Norm SE & Norm-M Est & Norm-M SE & AR Est & AR SE & AR-M Est & AR-M SE\\
-			\hline
-			$X_0 = 99000, Y_0 = 1000$ & 1.285071 & 1.440980 & 0.9866029 & 1.057043 & 1.193594 & 1.195820 & 0.9960353 & 0.9987062\\
-			\hline
-			$X_0 = 99900, Y_0 = 100$ & 2.943264 & 7.736638 & 1.4707234 & 3.367245 & 2.230244 & 4.859448 & 1.4236604 & 2.8121697\\
-			\hline
-			$X_0 = 99990, Y_0 = 10$ & 18.663014 & 67.156845 & 1.8455795 & 5.401173 & 16.739763 & 71.982690 & 1.6417916 & 4.4145381\\
-			\hline
-			$X_0 = 99999, Y_0 = 1$ & 774.274659 & 7491.005080 & 3.6060918 & 26.540940 & 229.184142 & 1199.155653 & 2.5082330 & 12.8639109\\
-			\hline
-		\end{tabular}
-	\end{footnotesize}
-\end{table}
-
-\begin{table}[H]
-
-	\caption{\label{tab:}$\rr$ Estimates and Std. Errs, IPR Model,
-		$X_0 = 99950, Y_0 = 50$, $\beta = 0.06, \gamma = 0.03$}
-	\centering
-	\begin{footnotesize}
-		\begin{tabular}[t]{l|r|r|r|r|r|r|r|r}
-			\hline
-			Data Set & Norm Est & Norm SE & Norm-M Est & Norm-M SE & AR Est & AR SE & AR-M Est & AR-M SE\\
-			\hline
-			$\sigma_X = 50, \sigma_Y = 2$ & 1.656451 & 4.879121 & 0.8705396 & 1.231344 & 1.318824 & 3.274114 & 0.8679029 & 1.309243\\
-			\hline
-			$\sigma_X = 500, \sigma_Y = 20$ & 10.426754 & 42.824909 & 1.1338742 & 4.906623 & 7.750396 & 24.637815 & 1.4241606 & 12.589987\\
-			\hline
-			$\sigma_X = 2500, \sigma_Y = 100$ & 48.894100 & 209.241051 & 1.3602975 & 10.234817 & 117.222483 & 870.624089 & 1.7544551 & 11.060645\\
-			\hline
-			$\sigma_X = 10000, \sigma_Y = 500$ & 143.885460 & 778.881724 & 2.7003338 & 35.642540 & 55.265406 & 163.492483 & 2.1210664 & 13.550223\\
-			\hline
-		\end{tabular}
-	\end{footnotesize}
-\end{table}
-
-\begin{table}[H]
-
-	\caption{\label{tab:}$\rr$ Estimates and Std. Errs, IPR Model,
-		$X_0 = 99950, Y_0 = 50$, $\sigma_X = 100, \sigma_Y = 5$}
-	\centering
-	\begin{footnotesize}
-		\begin{tabular}[t]{l|r|r|r|r|r|r|r|r}
-			\hline
-			Data Set & Norm Est & Norm SE & Norm-M Est & Norm-M SE & AR Est & AR SE & AR-M Est & AR-M SE\\
-			\hline
-			1st Order & 2.699973 & 9.131131 & 2.297208 & 7.14196 & 4.334767 & 41.325146 & 4.194341 & 36.201142\\
-			\hline
-			4th Order & 6.260963 & 20.337163 & 5.924029 & 17.81136 & 6.268500 & 20.536192 & 5.892582 & 17.341763\\
-			\hline
-			Linear SIR & 1.516803 & 1.554241 & 1.208578 & 1.12544 & 1.379871 & 1.311406 & 1.208521 & 1.003065\\
-			\hline
-		\end{tabular}
-	\end{footnotesize}
-\end{table}
-
-\begin{table}[H]
-
-	\caption{\label{tab:}$\rr$ Estimates and Std. Errs, IPR Model,
-		$\beta = 0.06, \gamma = 0.03$, $X_0 = 99950, Y_0 = 50$, $\sigma_X = 100, \sigma_Y = 5$}
-	\centering
-	\begin{tabular}[t]{l|r|r|r|r|r|r|r|r}
-		\hline
-		Data Set & Norm Est & Norm SE & Norm-M Est & Norm-M SE & AR Est & AR SE & AR-M Est & AR-M SE\\
-		\hline
-		T = 200 & 7.345813 & 16.12889 & 2.048281 & 3.959862 & 5.553952 & 13.33904 & 2.052063 & 4.119090\\
-		\hline
-		T = 100 & 12.683604 & 21.50385 & 2.278429 & 5.329775 & 9.221913 & 18.11852 & 2.245441 & 5.708995\\
-		\hline
-		T = 50 & 19.220693 & 27.75807 & 2.470106 & 6.581896 & 14.757521 & 24.04262 & 2.616444 & 7.417433\\
-		\hline
-		T = 20 & 25.731306 & 35.29261 & 1.415537 & 4.565460 & 20.489247 & 33.72596 & 3.885644 & 8.289876\\
-		\hline
-	\end{tabular}
-\end{table}
-
-\begin{table}[H]
-
-	\caption{\label{tab:}$\rr$ Estimates and Std. Errs, IPR Model,
-		$\beta = 0.06, \gamma = 0.03$, $\sigma_X = 100, \sigma_Y = 5$}
-	\centering
-	\begin{tabular}[t]{l|r|r|r|r|r|r|r|r}
-		\hline
-		Data Set & Norm Est & Norm SE & Norm-M Est & Norm-M SE & AR Est & AR SE & AR-M Est & AR-M SE\\
-		\hline
-		$X_0 = 99, Y_0 = 1$ & 3620.867680 & 57608.167596 & 5.239294 & 63.654679 & 729.654107 & 2978.983072 & 15.800132 & 284.613763\\
-		\hline
-		$X_0 = 990, Y_0 = 10$ & 67.597632 & 157.628039 & 1.734334 & 12.726031 & 164.611097 & 1887.125297 & 1.707676 & 11.777508\\
-		\hline
-		$X_0 = 9995, Y_0 = 5$ & 40.807237 & 133.279831 & 1.797818 & 7.233937 & 976.308278 & 17227.979404 & 1.700549 & 9.887431\\
-		\hline
-		$X_0 = 999500, Y_0 = 500$ & 1.577592 & 1.422029 & 1.434415 & 1.131136 & 1.567228 & 1.594393 & 1.449019 & 1.180341\\
-		\hline
-	\end{tabular}
-\end{table}
-
-\subsection{Smoothed Incidence to Prevalence Ratio}
-
-\begin{table}[H]
-
-	\caption{\label{tab:}$\rr$ Estimates and Std. Errs, SIPR Model,
-		$X_0 = 99950, Y_0 = 50$, $\sigma_X = 100, \sigma_Y = 5$, $\beta = 0.06$}
-	\centering
-	\begin{footnotesize}
-		\begin{tabular}[t]{l|r|r|r|r|r|r|r|r}
-			\hline
-			Data Set & Norm Est & Norm SE & Norm-M Est & Norm-M SE & AR Est & AR SE & AR-M Est & AR-M SE\\
-			\hline
-			$\gamma$ = 0.001 & 3.9645187 & $<$ 1e-07 & 3.8285414 & $<$ 1e-07 & 3.9055582 & 0.000000 & 3.7112781 & $<$ 1e-07 \\
-			\hline
-			$\gamma$ = 0.04 & 1.7451025 & $<$ 1e-07 & 2.8785564 & $<$ 1e-07 & 68.7390000 & 0.000000 & 1.2465750 & $<$ 1e-07 \\
-			\hline
-			$\gamma$ = 0.06 & 0.9919790 & 0.0559052 & 0.9925620 & 0.0297783 & 0.8718278 & 0.087266 & 1.0294124 & 0.0510585\\
-			\hline
-			$\gamma$ = 0.24 & -0.2484284 & 0.1159769 & -0.0317365 & 0.0335514 & -2.8642901 & 0.000000 & 0.7365437 & 0.0675331\\
-			\hline
-		\end{tabular}
-	\end{footnotesize}
-\end{table}
-
-\begin{table}[H]
-
-	\caption{\label{tab:}$\rr$ Estimates and Std. Errs, SIPR Model,
-		$\sigma_X = 100, \sigma_Y = 5$, $\beta = 0.06, \gamma = 0.03$}
-	\centering
-	\begin{footnotesize}
-		\begin{tabular}[t]{l|r|r|r|r|r|r|r|r}
-			\hline
-			Data Set & Norm Est & Norm SE & Norm-M Est & Norm-M SE & AR Est & AR SE & AR-M Est & AR-M SE\\
-			\hline
-			$X_0 = 99000, Y_0 = 1000$ & 2.0870176 & 57.9229822 & 2.016608 & 57.8153675 & 2.285424 & 57.4382982 & 1.922192 & 57.0457534\\
-			\hline
-			$X_0 = 99900, Y_0 = 100$ & -0.8808153 & 0.0000000 & -1.150393 & 0.0000000 & -6.563660 & 0.0000000 & -1.777316 & 0.0000000\\
-			\hline
-			$X_0 = 99990, Y_0 = 10$ & 1.3451986 & 0.3236824 & 1.354925 & 0.3158698 & 1.381198 & 0.3093223 & 1.345069 & 0.3258082\\
-			\hline
-			$X_0 = 99999, Y_0 = 1$ & 2.4390757 & 0.0000000 & 5.411582 & 0.0000000 & 1.204689 & 0.0000000 & 1.035624 & 0.0000000\\
-			\hline
-		\end{tabular}
-	\end{footnotesize}
-\end{table}
-
-\begin{table}[H]
-
-	\caption{\label{tab:}$\rr$ Estimates and Std. Errs, SIPR Model,
-		$X_0 = 99950, Y_0 = 50$, $\beta = 0.06, \gamma = 0.03$}
-	\centering
-	\begin{footnotesize}
-		\begin{tabular}[t]{l|r|r|r|r|r|r|r|r}
-			\hline
-			Data Set & Norm Est & Norm SE & Norm-M Est & Norm-M SE & AR Est & AR SE & AR-M Est & AR-M SE\\
-			\hline
-			$\sigma_X = 50, \sigma_Y = 2$ & 1.836824 & $<$ 1e-07 & 1.7778150 & $<$ 1e-07 & 1.7603970 & $<$ 1e-07 & 1.842289 & $<$ 1e-07 \\
-			\hline
-			$\sigma_X = 500, \sigma_Y = 20$ & 3.193066 & $<$ 1e-07 & 1.3211468 & $<$ 1e-07 & 1.3509726 & $<$ 1e-07 & 1.334458 & $<$ 1e-07 \\
-			\hline
-			$\sigma_X = 2500, \sigma_Y = 100$ & 1.060609 & $<$ 1e-07 & 1.2345884 & $<$ 1e-07 & 1.7027546 & $<$ 1e-07 & 2.596160 & $<$ 1e-07 \\
-			\hline
-			$\sigma_X = 10000, \sigma_Y = 500$ & -2.535946 & $<$ 1e-07 & 0.7094139 & $<$ 1e-07 & 0.8911487 & $<$ 1e-07 & -1.103940 & $<$ 1e-07 \\
-			\hline
-		\end{tabular}
-	\end{footnotesize}
-\end{table}
-
-\begin{table}[H]
-
-	\caption{\label{tab:}$\rr$ Estimates and Std. Errs, SIPR Model,
-		$X_0 = 99950, Y_0 = 50$, $\sigma_X = 100, \sigma_Y = 5$}
-	\centering
-	\begin{footnotesize}
-		\begin{tabular}[t]{l|r|r|r|r|r|r|r|r}
-			\hline
-			Data Set & Norm Est & Norm SE & Norm-M Est & Norm-M SE & AR Est & AR SE & AR-M Est & AR-M SE\\
-			\hline
-			1st Order & 58.836635 & $<$ 1e-07 & 27.429685 & $<$ 1e-07 & 1.498805 & $<$ 1e-07 & -14.135196 & $<$ 1e-07 \\
-			\hline
-			4th Order & 2.848624 & $<$ 1e-07 & 15.083621 & $<$ 1e-07 & 2.766152 & $<$ 1e-07 & 14.972580 & $<$ 1e-07 \\
-			\hline
-			Linear SIR & 1.294079 & 0.2151131 & 1.294364 & 0.2158095 & 1.296731 & 0.2215603 & 1.304489 & 0.2299876\\
-			\hline
-		\end{tabular}
-	\end{footnotesize}
-\end{table}
-
-\begin{table}[H]
-
-	\caption{\label{tab:}$\rr$ Estimates and Std. Errs, SIPR Model,
-		$\beta = 0.06, \gamma = 0.03$, $X_0 = 99950, Y_0 = 50$, $\sigma_X = 100, \sigma_Y = 5$}
-	\centering
-	\begin{tabular}[t]{l|r|r|r|r|r|r|r|r}
-		\hline
-		Data Set & Norm Est & Norm SE & Norm-M Est & Norm-M SE & AR Est & AR SE & AR-M Est & AR-M SE\\
-		\hline
-		T = 200 & -2.119890 & 0.000000 & 2.6624046 & 0.0000000 & 3.3187716 & 0.0000000 & 2.363125 & 32.2727752\\
-		\hline
-		T = 100 & 2.341119 & 1.168462 & 2.2020092 & 0.5643447 & 2.0919450 & 1.0699219 & 1.963112 & 0.3489431\\
-		\hline
-		T = 50 & 2.190565 & 2.238582 & 1.9477092 & 0.4198946 & 1.8701913 & 0.7519526 & 2.279451 & 0.6584392\\
-		\hline
-		T = 20 & 4.177245 & 7.320297 & 0.9702882 & 0.8673670 & 0.3570221 & 7.6611297 & 3.984916 & 2.1350792\\
-		\hline
-	\end{tabular}
-\end{table}
-
-\begin{table}[H]
-
-	\caption{\label{tab:}$\rr$ Estimates and Std. Errs, SIPR Model,
-		$\beta = 0.06, \gamma = 0.03$, $\sigma_X = 100, \sigma_Y = 5$}
-	\centering
-	\begin{tabular}[t]{l|r|r|r|r|r|r|r|r}
-		\hline
-		Data Set & Norm Est & Norm SE & Norm-M Est & Norm-M SE & AR Est & AR SE & AR-M Est & AR-M SE\\
-		\hline
-		$X_0 = 99, Y_0 = 1$ & 3.367825 & 22.41867 & 0.9850654 & 0.5665987 & 1.5591972 & 3959.42264 & 4.207729 & 13.26187\\
-		\hline
-		$X_0 = 990, Y_0 = 10$ & 1.912639 & 390.58481 & 1.6931741 & 12.0122249 & 0.5334608 & 13.65189 & 2.898150 & 284.49327\\
-		\hline
-		$X_0 = 9995, Y_0 = 5$ & 1.062891 & 0.00000 & 1.8635858 & 0.0000000 & 1.0938066 & 0.00000 & 1.572230 & 0.00000\\
-		\hline
-		$X_0 = 999500, Y_0 = 500$ & 1.792115 & 0.00000 & 1.7907398 & 0.0000000 & 1.7902743 & 0.00000 & 1.750503 & 0.00000\\
-		\hline
-	\end{tabular}
-\end{table}
-
-\subsection{SIR Log Linear}
-
-\begin{table}[H]
-
-	\caption{\label{tab:}$\rr$ Estimates and Std. Errs, LL Model,
-		$X_0 = 99950, Y_0 = 50$, $\sigma_X = 100, \sigma_Y = 5$, $\beta = 0.06$}
-	\centering
-	\begin{footnotesize}
-		\begin{tabular}[t]{l|r|r|r|r|r|r|r|r}
-			\hline
-			Data Set & Norm Est & Norm SE & Norm-M Est & Norm-M SE & AR Est & AR SE & AR-M Est & AR-M SE\\
-			\hline
-			$\gamma$ = 0.001 & 46.3209595 & 0.7797216 & 59.8300371 & 0.1079267 & 41.897714 & 0.9767785 & 70.7075929 & 0.9517239\\
-			\hline
-			$\gamma$ = 0.04 & 1.5002728 & 0.0001905 & 1.5000436 & 0.0000928 & 1.499516 & 0.0002653 & 1.5000445 & 0.0001481\\
-			\hline
-			$\gamma$ = 0.06 & 0.9989449 & 0.0007428 & 0.9994831 & 0.0003320 & 1.000275 & 0.0010605 & 1.0008545 & 0.0004910\\
-			\hline
-			$\gamma$ = 0.24 & 0.9912397 & 0.0030505 & 0.8623034 & 0.0087543 & 1.001367 & 0.0025488 & 0.9198973 & 0.0052616\\
-			\hline
-		\end{tabular}
-	\end{footnotesize}
-\end{table}
-
-\begin{table}[H]
-
-	\caption{\label{tab:}$\rr$ Estimates and Std. Errs, LL Model,
-		$\sigma_X = 100, \sigma_Y = 5$, $\beta = 0.06, \gamma = 0.03$}
-	\centering
-	\begin{footnotesize}
-		\begin{tabular}[t]{l|r|r|r|r|r|r|r|r}
-			\hline
-			Data Set & Norm Est & Norm SE & Norm-M Est & Norm-M SE & AR Est & AR SE & AR-M Est & AR-M SE\\
-			\hline
-			$X_0 = 99000, Y_0 = 1000$ & 1.999515 & 0.0002828 & 2.000144 & 0.0001594 & 1.999753 & 0.0004197 & 2.000586 & 0.0003024\\
-			\hline
-			$X_0 = 99900, Y_0 = 100$ & 2.000011 & 0.0002315 & 1.999732 & 0.0001526 & 2.000473 & 0.0002776 & 2.000640 & 0.0002656\\
-			\hline
-			$X_0 = 99990, Y_0 = 10$ & 1.999886 & 0.0002517 & 1.999599 & 0.0001361 & 2.000608 & 0.0003289 & 1.999049 & 0.0002664\\
-			\hline
-			$X_0 = 99999, Y_0 = 1$ & 1.999690 & 0.0007731 & 2.000503 & 0.0003700 & 2.002716 & 0.0010572 & 1.999919 & 0.0005663\\
-			\hline
-		\end{tabular}
-	\end{footnotesize}
-\end{table}
-
-\begin{table}[H]
-
-	\caption{\label{tab:}$\rr$ Estimates and Std. Errs, LL Model,
-		$X_0 = 99950, Y_0 = 50$, $\beta = 0.06, \gamma = 0.03$}
-	\centering
-	\begin{footnotesize}
-		\begin{tabular}[t]{l|r|r|r|r|r|r|r|r}
-			\hline
-			Data Set & Norm Est & Norm SE & Norm-M Est & Norm-M SE & AR Est & AR SE & AR-M Est & AR-M SE\\
-			\hline
-			$\sigma_X = 50, \sigma_Y = 2$ & 1.999891 & 0.0001037 & 2.000036 & 0.0000930 & 1.999143 & 0.0001394 & 1.999749 & 0.0001328\\
-			\hline
-			$\sigma_X = 500, \sigma_Y = 20$ & 2.003402 & 0.0010320 & 1.998765 & 0.0003682 & 2.003763 & 0.0015505 & 1.999829 & 0.0008545\\
-			\hline
-			$\sigma_X = 2500, \sigma_Y = 100$ & 2.001882 & 0.0054085 & 1.995742 & 0.0015419 & 2.038646 & 0.0081562 & 2.015675 & 0.0031322\\
-			\hline
-			$\sigma_X = 10000, \sigma_Y = 500$ & 2.059934 & 0.0246402 & 1.995133 & 0.0038894 & 2.209575 & 0.0439100 & 2.545819 & 0.0568537\\
-			\hline
-		\end{tabular}
-	\end{footnotesize}
-\end{table}
-
-\begin{table}[H]
-
-	\caption{\label{tab:}$\rr$ Estimates and Std. Errs, LL Model,
-		$X_0 = 99950, Y_0 = 50$, $\sigma_X = 100, \sigma_Y = 5$}
-	\centering
-	\begin{footnotesize}
-		\begin{tabular}[t]{l|r|r|r|r|r|r|r|r}
-			\hline
-			Data Set & Norm Est & Norm SE & Norm-M Est & Norm-M SE & AR Est & AR SE & AR-M Est & AR-M SE\\
-			\hline
-			1st Order & 1.188050 & 0.0006341 & 1.187976 & 0.0006389 & 1.187553 & 0.0006273 & 1.187848 & 0.0006497\\
-			\hline
-			4th Order & 1.955180 & 0.0164608 & 1.955313 & 0.0164212 & 1.955130 & 0.0164623 & 1.955302 & 0.0164165\\
-			\hline
-			Linear SIR & 1.258587 & 0.0010079 & 1.258670 & 0.0010069 & 1.258920 & 0.0009938 & 1.258462 & 0.0010199\\
-			\hline
-		\end{tabular}
-	\end{footnotesize}
-\end{table}
-
-\begin{table}[H]
-
-	\caption{\label{tab:}$\rr$ Estimates and Std. Errs, LL Model,
-		$\beta = 0.06, \gamma = 0.03$, $X_0 = 99950, Y_0 = 50$, $\sigma_X = 100, \sigma_Y = 5$}
-	\centering
-	\begin{tabular}[t]{l|r|r|r|r|r|r|r|r}
-		\hline
-		Data Set & Norm Est & Norm SE & Norm-M Est & Norm-M SE & AR Est & AR SE & AR-M Est & AR-M SE\\
-		\hline
-		T = 200 & 2.001415 & 0.0017620 & 2.000649 & 0.0009166 & 2.0053322 & 0.0023026 & 1.992983 & 0.0019266\\
-		\hline
-		T = 100 & 1.869996 & 0.0322729 & 1.992759 & 0.0104984 & 1.7390224 & 0.0357848 & 2.052393 & 0.0208558\\
-		\hline
-		T = 50 & 1.193299 & 0.0560189 & 1.741553 & 0.0454525 & 1.2098418 & 0.0420061 & 1.690908 & 0.0406272\\
-		\hline
-		T = 20 & 1.042976 & 0.0276261 & 1.747193 & 0.3635417 & 0.9170623 & 0.0337569 & 1.432791 & 0.0397426\\
-		\hline
-	\end{tabular}
-\end{table}
-
-\begin{table}[H]
-
-	\caption{\label{tab:}$\rr$ Estimates and Std. Errs, LL Model,
-		$\beta = 0.06, \gamma = 0.03$, $\sigma_X = 100, \sigma_Y = 5$}
-	\centering
-	\begin{tabular}[t]{l|r|r|r|r|r|r|r|r}
-		\hline
-		Data Set & Norm Est & Norm SE & Norm-M Est & Norm-M SE & AR Est & AR SE & AR-M Est & AR-M SE\\
-		\hline
-		$X_0 = 99, Y_0 = 1$ & 1.141317 & 0.0422218 & 2.031666 & 0.0113166 & 1.056918 & 0.0447463 & 1.771771 & 0.0561071\\
-		\hline
-		$X_0 = 990, Y_0 = 10$ & 2.255041 & 0.0379147 & 2.010183 & 0.0055668 & 2.470092 & 0.0545897 & 2.280617 & 0.0229552\\
-		\hline
-		$X_0 = 9995, Y_0 = 5$ & 1.998839 & 0.0023193 & 1.997778 & 0.0007420 & 2.004827 & 0.0027816 & 2.001058 & 0.0015548\\
-		\hline
-		$X_0 = 999500, Y_0 = 500$ & 1.999983 & 0.0000215 & 1.999989 & 0.0000203 & 1.999924 & 0.0000314 & 1.999829 & 0.0000294\\
-		\hline
-	\end{tabular}
-\end{table}
-
-\subsection{SIR Markov Chain}
-
-\begin{table}[H]
-
-	\caption{\label{tab:}$\rr$ Estimates and Std. Errs, MC Model,
-		$X_0 = 99950, Y_0 = 50$, $\sigma_X = 100, \sigma_Y = 5$, $\beta = 0.06$}
-	\centering
-	\begin{footnotesize}
-		\begin{tabular}[t]{l|r|r|r|r|r|r|r|r}
-			\hline
-			Data Set & Norm Est & Norm SE & Norm-M Est & Norm-M SE & AR Est & AR SE & AR-M Est & AR-M SE\\
-			\hline
-			$\gamma$ = 0.001 & 0.9485554 & 1.0e-07 & 0.9485555 & $<$ 1e-07 & 0.9485555 & $<$ 1e-07 & 0.9485555 & $<$ 1e-07\\
-			\hline
-			$\gamma$ = 0.04 & 0.9485555 & $<$ 1e-07 & 0.9485555 & $<$ 1e-07 & 0.9485555 & $<$ 1e-07 & 0.9485555 & $<$ 1e-07\\
-			\hline
-			$\gamma$ = 0.06 & 0.9485555 & 2.0e-07 & 0.9485555 & $<$ 1e-07 & 0.9485555 & 2.00e-07 & 0.9485555 & $<$ 1e-07\\
-			\hline
-			$\gamma$ = 0.24 & 0.9485812 & 1.5e-05 & 0.9485605 & 1.4e-06 & 0.9486128 & 1.26e-05 & 0.9485558 & 2e-07\\
-			\hline
-		\end{tabular}
-	\end{footnotesize}
-\end{table}
-
-\begin{table}[H]
-
-	\caption{\label{tab:}$\rr$ Estimates and Std. Errs, MC Model,
-		$\sigma_X = 100, \sigma_Y = 5$, $\beta = 0.06, \gamma = 0.03$}
-	\centering
-	\begin{footnotesize}
-		\begin{tabular}[t]{l|r|r|r|r|r|r|r|r}
-			\hline
-			Data Set & Norm Est & Norm SE & Norm-M Est & Norm-M SE & AR Est & AR SE & AR-M Est & AR-M SE\\
-			\hline
-			$X_0 = 99000, Y_0 = 1000$ & 0.9485555 & $<$ 1e-07 & 0.9485555 & $<$ 1e-07 & 0.9485555 & $<$ 1e-07 & 0.9485555 & $<$ 1e-07\\
-			\hline
-			$X_0 = 99900, Y_0 = 100$ & 0.9485555 & $<$ 1e-07 & 0.9485555 & $<$ 1e-07 & 0.9485555 & $<$ 1e-07 & 0.9485555 & $<$ 1e-07\\
-			\hline
-			$X_0 = 99990, Y_0 = 10$ & 0.9485552 & 2.00e-07 & 0.9485555 & $<$ 1e-07 & 0.9485558 & 2.0e-07 & 0.9485555 & $<$ 1e-07\\
-			\hline
-			$X_0 = 99999, Y_0 = 1$ & 0.9486030 & 1.56e-05 & 0.9485557 & $<$ 1e-07 & 0.9485701 & 2.6e-06 & 0.9485555 & $<$ 1e-07\\
-			\hline
-		\end{tabular}
-	\end{footnotesize}
-\end{table}
-
-\begin{table}[H]
-
-	\caption{\label{tab:}$\rr$ Estimates and Std. Errs, MC Model,
-		$X_0 = 99950, Y_0 = 50$, $\beta = 0.06, \gamma = 0.03$}
-	\centering
-	\begin{footnotesize}
-		\begin{tabular}[t]{l|r|r|r|r|r|r|r|r}
-			\hline
-			Data Set & Norm Est & Norm SE & Norm-M Est & Norm-M SE & AR Est & AR SE & AR-M Est & AR-M SE\\
-			\hline
-			$\sigma_X = 50, \sigma_Y = 2$ & 0.9485555 & 0.$<$ 1e-07 & 0.9485555 & $<$ 1e-07 & 0.9485555 & 0.$<$ 1e-07 & 0.9485555 & $<$ 1e-07\\
-			\hline
-			$\sigma_X = 500, \sigma_Y = 20$ & 0.9485554 & 2.0e-07 & 0.9485555 & $<$ 1e-07 & 0.9485554 & 1.0e-07 & 0.9485555 & 1e-07\\
-			\hline
-			$\sigma_X = 2500, \sigma_Y = 100$ & 0.9485562 & 1.0e-06 & 0.9485554 & 1e-07 & 0.9485640 & 3.6e-06 & 0.9485554 & 1e-07\\
-			\hline
-			$\sigma_X = 10000, \sigma_Y = 500$ & 0.9485506 & 3.8e-06 & 0.9485556 & 2e-07 & 0.9485428 & 2.6e-06 & 0.9485538 & 8e-07\\
-			\hline
-		\end{tabular}
-	\end{footnotesize}
-\end{table}
-
-\begin{table}[H]
-
-	\caption{\label{tab:}$\rr$ Estimates and Std. Errs, MC Model,
-		$X_0 = 99950, Y_0 = 50$, $\sigma_X = 100, \sigma_Y = 5$}
-	\centering
-	\begin{footnotesize}
-		\begin{tabular}[t]{l|r|r|r|r|r|r|r|r}
-			\hline
-			Data Set & Norm Est & Norm SE & Norm-M Est & Norm-M SE & AR Est & AR SE & AR-M Est & AR-M SE\\
-			\hline
-			1st Order & 0.9485551 & 1e-07 & 0.9485540 & 4e-07 & 0.9485555 & 2e-07 & 0.9485553 & 2.0e-07\\
-			\hline
-			4th Order & 0.9485545 & 6e-07 & 0.9485533 & 1e-06 & 0.9485535 & 9e-07 & 0.9485527 & 1.2e-06\\
-			\hline
-			Linear SIR & 0.9485555 & $<$ 1e-07 & 0.9485555 & $<$ 1e-07 & 0.9485555 & $<$ 1e-07 & 0.9485555 & 0.$<$ 1e-07\\
-			\hline
-		\end{tabular}
-	\end{footnotesize}
-\end{table}
-
-\begin{table}[H]
-
-	\caption{\label{tab:}$\rr$ Estimates and Std. Errs, MC Model,
-		$\beta = 0.06, \gamma = 0.03$, $X_0 = 99950, Y_0 = 50$, $\sigma_X = 100, \sigma_Y = 5$}
-	\centering
-	\begin{tabular}[t]{l|r|r|r|r|r|r|r|r}
-		\hline
-		Data Set & Norm Est & Norm SE & Norm-M Est & Norm-M SE & AR Est & AR SE & AR-M Est & AR-M SE\\
-		\hline
-		T = 200 & 0.9485555 & 1e-07 & 0.9485555 & $<$ 1e-07 & 0.9485555 & $<$ 1e-07 & 0.9485555 & $<$ 1e-07\\
-		\hline
-		T = 100 & 0.9485555 & 1e-07 & 0.9485555 & $<$ 1e-07 & 0.9485555 & 1e-07 & 0.9485555 & $<$ 1e-07\\
-		\hline
-		T = 50 & 0.9485554 & 2e-07 & 0.9485555 & $<$ 1e-07 & 0.9485555 & 2e-07 & 0.9485555 & $<$ 1e-07\\
-		\hline
-		T = 20 & 0.9485556 & 5e-07 & 0.9485554 & $<$ 1e-07 & 0.9485552 & 5e-07 & 0.9485556 & 1e-07\\
-		\hline
-	\end{tabular}
-\end{table}
-
-\begin{table}[H]
-
-	\caption{\label{tab:}$\rr$ Estimates and Std. Errs, MC Model,
-		$\beta = 0.06, \gamma = 0.03$, $\sigma_X = 100, \sigma_Y = 5$}
-	\centering
-	\begin{tabular}[t]{l|r|r|r|r|r|r|r|r}
-		\hline
-		Data Set & Norm Est & Norm SE & Norm-M Est & Norm-M SE & AR Est & AR SE & AR-M Est & AR-M SE\\
-		\hline
-		$X_0 = 99, Y_0 = 1$ & 0.8455334 & 0.0124837 & 0.9481149 & 0.0002788 & NaN & NaN & 0.9480033 & 0.0005127\\
-		\hline
-		$X_0 = 990, Y_0 = 10$ & 0.9432094 & 0.0014184 & 0.9485736 & 0.0000060 & 0.9500365 & 0.0003484 & 0.9486116 & 0.0000141\\
-		\hline
-		$X_0 = 9995, Y_0 = 5$ & 0.9485626 & 0.0000040 & 0.9485575 & 0.0000002 & 0.9492072 & 0.0002225 & 0.9485575 & 0.0000002\\
-		\hline
-		$X_0 = 999500, Y_0 = 500$ & 0.9485553 & 0.0000000 & 0.9485553 & 0.0000000 & 0.9485553 & 0.0000000 & 0.9485553 & 0.0000000\\
-		\hline
-	\end{tabular}
-\end{table}
-
-\subsection{SIR Sequential Bayes}
-
-\begin{table}[H]
-
-	\caption{\label{tab:}$\rr$ Estimates and Std. Errs, SB Model,
-		$X_0 = 99950, Y_0 = 50$, $\sigma_X = 100, \sigma_Y = 5$, $\beta = 0.06$}
-	\centering
-	\begin{footnotesize}
-	\begin{tabular}[t]{l|r|r|r|r|r|r|r|r}
-		\hline
-		Data Set & Norm Est & Norm SE & Norm-M Est & Norm-M SE & AR Est & AR SE & AR-M Est & AR-M SE\\
-		\hline
-		$\gamma$ = 0.001 & 0.9769731 & 0.0517362 & 1.000000 & 0.0523424 & 1.000000 & 0.0523424 & 1.000000 & 0.0523424\\
-		\hline
-		$\gamma$ = 0.04 & 1.2060946 & 0.0574836 & 1.000000 & 0.0523424 & 1.073894 & 0.0542418 & 1.115316 & 0.0552780\\
-		\hline
-		$\gamma$ = 0.06 & 1.0000000 & 0.0523424 & 1.000000 & 0.0523424 & 0.640393 & 0.0418868 & 1.000000 & 0.0523424\\
-		\hline
-		$\gamma$ = 0.24 & 1.2354952 & 0.0581800 & 1.338591 & 0.0605588 & 1.009694 & 0.0525955 & 1.000000 & 0.0523424\\
-		\hline
-	\end{tabular}
-\end{footnotesize}
-\end{table}
-
-\begin{table}[H]
-
-	\caption{\label{tab:}$\rr$ Estimates and Std. Errs, SB Model,
-		$\sigma_X = 100, \sigma_Y = 5$, $\beta = 0.06, \gamma = 0.03$}
-	\centering
-	\begin{footnotesize}
-		\begin{tabular}[t]{l|r|r|r|r|r|r|r|r}
-			\hline
-			Data Set & Norm Est & Norm SE & Norm-M Est & Norm-M SE & AR Est & AR SE & AR-M Est & AR-M SE\\
-			\hline
-			$X_0 = 99000, Y_0 = 1000$ & 0.9728534 & 0.0516270 & 1.052579 & 0.0537008 & 0.5850802 & 0.0400370 & 0.6236582 & 0.0413358\\
-			\hline
-			$X_0 = 99900, Y_0 = 100$ & 1.0000000 & 0.0523424 & 1.000000 & 0.0523424 & 1.4175193 & 0.0623186 & 1.4017333 & 0.0619707\\
-			\hline
-			$X_0 = 99990, Y_0 = 10$ & 1.3877095 & 0.0616599 & 1.430888 & 0.0626118 & 1.3948200 & 0.0618177 & 1.4507203 & 0.0630442\\
-			\hline
-			$X_0 = 99999, Y_0 = 1$ & 1.4714813 & 0.0634938 & 1.465728 & 0.0633695 & 1.4665442 & 0.0633871 & 1.4723265 & 0.0635120\\
-			\hline
-		\end{tabular}
-	\end{footnotesize}
-\end{table}
-
-\begin{table}[H]
-
-	\caption{\label{tab:}$\rr$ Estimates and Std. Errs, SB Model,
-		$X_0 = 99950, Y_0 = 50$, $\beta = 0.06, \gamma = 0.03$}
-	\centering
-	\begin{footnotesize}
-	\begin{tabular}[t]{l|r|r|r|r|r|r|r|r}
-		\hline
-		Data Set & Norm Est & Norm SE & Norm-M Est & Norm-M SE & AR Est & AR SE & AR-M Est & AR-M SE\\
-		\hline
-		$\sigma_X = 50, \sigma_Y = 2$ & 1.409094 & 0.0621332 & 1.141402 & 0.0559207 & 1.386553 & 0.0616342 & 1.336668 & 0.0605153\\
-		\hline
-		$\sigma_X = 500, \sigma_Y = 20$ & 1.000000 & 0.0523424 & 1.415487 & 0.0622740 & 1.000000 & 0.0523424 & 1.000000 & 0.0523424\\
-		\hline
-		$\sigma_X = 2500, \sigma_Y = 100$ & 1.519223 & 0.0645155 & 1.000000 & 0.0523424 & 1.549862 & 0.0651629 & 1.000000 & 0.0523424\\
-		\hline
-		$\sigma_X = 10000, \sigma_Y = 500$ & 1.127679 & 0.0555836 & 1.000000 & 0.0523424 & 1.540929 & 0.0649748 & 1.000000 & 0.0523424\\
-		\hline
-	\end{tabular}
-\end{footnotesize}
-\end{table}
-
-\begin{table}[H]
-
-	\caption{\label{tab:}$\rr$ Estimates and Std. Errs, SB Model,
-		$X_0 = 99950, Y_0 = 50$, $\sigma_X = 100, \sigma_Y = 5$}
-	\centering
-	\begin{footnotesize}
-	\begin{tabular}[t]{l|r|r|r|r|r|r|r|r}
-		\hline
-		Data Set & Norm Est & Norm SE & Norm-M Est & Norm-M SE & AR Est & AR SE & AR-M Est & AR-M SE\\
-		\hline
-		1st Order & 1.2515351 & 0.0585565 & 1.221050 & 0.0578389 & 1.301757 & 0.0597198 & 1.2848494 & 0.0593307\\
-		\hline
-		4th Order & 0.5223345 & 0.0378293 & 0.550362 & 0.0388309 & 0.523122 & 0.0378578 & 0.5506554 & 0.0388413\\
-		\hline
-		Linear SIR & 1.1217433 & 0.0554371 & 1.194507 & 0.0572068 & 1.123600 & 0.0554829 & 1.1568274 & 0.0562973\\
-		\hline
-	\end{tabular}
-	\end{footnotesize}
-\end{table}
-
-\begin{table}[H]
-
-	\caption{\label{tab:}$\rr$ Estimates and Std. Errs, SB Model,
-		$\beta = 0.06, \gamma = 0.03$, $X_0 = 99950, Y_0 = 50$, $\sigma_X = 100, \sigma_Y = 5$}
-	\centering
-	\begin{tabular}[t]{l|r|r|r|r|r|r|r|r}
-		\hline
-		Data Set & Norm Est & Norm SE & Norm-M Est & Norm-M SE & AR Est & AR SE & AR-M Est & AR-M SE\\
-		\hline
-		T = 200 & 1.675005 & 0.0677426 & 1.233313 & 0.0581286 & 1.637813 & 0.0669862 & 1.555502 & 0.0652813\\
-		\hline
-		T = 100 & 1.000000 & 0.0523424 & 1.685553 & 0.0679555 & 1.000000 & 0.0523424 & 1.000000 & 0.0523424\\
-		\hline
-		T = 50 & 1.856718 & 0.0713225 & 1.000000 & 0.0523424 & 1.907273 & 0.0722870 & 1.000000 & 0.0523424\\
-		\hline
-		T = 20 & 1.210671 & 0.0575926 & 1.000000 & 0.0523424 & 1.892533 & 0.0720071 & 1.000000 & 0.0523424\\
-		\hline
-	\end{tabular}
-\end{table}
-
-\begin{table}[H]
-
-	\caption{\label{tab:}$\rr$ Estimates and Std. Errs, SB Model,
-		$\beta = 0.06, \gamma = 0.03$, $\sigma_X = 100, \sigma_Y = 5$}
-	\centering
-	\begin{tabular}[t]{l|r|r|r|r|r|r|r|r}
-		\hline
-		Data Set & Norm Est & Norm SE & Norm-M Est & Norm-M SE & AR Est & AR SE & AR-M Est & AR-M SE\\
-		\hline
-		$X_0 = 99, Y_0 = 1$ & 0.3887115 & 0.0326338 & 1.574343 & 0.0656755 & 0.4241419 & 0.0340886 & 1.000000 & 0.0523424\\
-		\hline
-		$X_0 = 990, Y_0 = 10$ & 0.9028353 & 0.0497345 & 1.000000 & 0.0523424 & 1.5457192 & 0.0650757 & 1.000000 & 0.0523424\\
-		\hline
-		$X_0 = 9995, Y_0 = 5$ & 0.4418182 & 0.0347917 & 1.000000 & 0.0523424 & 1.0000000 & 0.0523424 & 1.452822 & 0.0630899\\
-		\hline
-		$X_0 = 999500, Y_0 = 500$ & 1.1774169 & 0.0567961 & 1.270930 & 0.0590085 & 1.1876846 & 0.0570432 & 1.807844 & 0.0703775\\
-		\hline
-	\end{tabular}
-\end{table}
-
-\section{Fitting the Exponential Model}
-
-\subsection{Exponential Growth}
-
-\begin{table}[H]
-
-	\caption{\label{tab:}$\rr$ Estimates and Std. Errs, EXPGROWTH Model,
-		$\beta = 0.06, \gamma = 0.03$, $X_0 = 99950, Y_0 = 50$, $\sigma_X = 100, \sigma_Y = 5$}
-	\centering
-	\begin{tabular}[t]{l|r|r|r|r|r|r|r|r}
-		\hline
-		Data Set & Norm Est & Norm SE & Norm-M Est & Norm-M SE & AR Est & AR SE & AR-M Est & AR-M SE\\
-		\hline
-		$\gamma$ = 0.001 & 1.141015 & 0.0001503 & 1.1349235 & 0.0001645 & 1.139834 & 0.0001529 & 1.1348807 & 0.0001646\\
-		\hline
-		$\gamma$ = 0.04 & 1.133308 & 0.0001686 & 1.1310087 & 0.0001744 & 1.132539 & 0.0001705 & 1.1313483 & 0.0001735\\
-		\hline
-		$\gamma$ = 0.06 & 1.093381 & 0.0003053 & 0.9905347 & 0.0013816 & 1.085491 & 0.0003433 & 0.9899559 & 0.0013931\\
-		\hline
-		$\gamma$ = 0.24 & 1.092493 & 0.0003093 & 0.9206087 & 0.0036121 & 1.085583 & 0.0003428 & 0.8860333 & 0.0055673\\
-		\hline
-	\end{tabular}
-\end{table}
-
-\begin{table}[H]
-
-	\caption{\label{tab:}$\rr$ Estimates and Std. Errs, EXPGROWTH Model,
-		$\beta = 0.06, \gamma = 0.03$, $X_0 = 99950, Y_0 = 50$, $\sigma_X = 100, \sigma_Y = 5$}
-	\centering
-	\begin{tabular}[t]{l|r|r|r|r|r|r|r|r}
-		\hline
-		Data Set & Norm Est & Norm SE & Norm-M Est & Norm-M SE & AR Est & AR SE & AR-M Est & AR-M SE\\
-		\hline
-		$X_0 = 99000, Y_0 = 1000$ & 1.132250 & 0.0001712 & 1.127900 & 0.0001827 & 1.131379 & 0.0001734 & 1.127954 & 0.0001825\\
-		\hline
-		$X_0 = 99900, Y_0 = 100$ & 1.143191 & 0.0001455 & 1.141440 & 0.0001494 & 1.142493 & 0.0001470 & 1.141621 & 0.0001489\\
-		\hline
-		$X_0 = 99990, Y_0 = 10$ & 1.148203 & 0.0001351 & 1.147130 & 0.0001372 & 1.148093 & 0.0001353 & 1.147197 & 0.0001371\\
-		\hline
-		$X_0 = 99999, Y_0 = 1$& 1.140076 & 0.0001524 & 1.136283 & 0.0001612 & 1.138602 & 0.0001558 & 1.136402 & 0.0001610\\
-		\hline
-	\end{tabular}
-\end{table}
-
-\begin{table}[H]
-
-	\caption{\label{tab:}$\rr$ Estimates and Std. Errs, EXPGROWTH Model,
-		$\beta = 0.06, \gamma = 0.03$, $X_0 = 99950, Y_0 = 50$}
-	\centering
-	\begin{tabular}[t]{l|r|r|r|r|r|r|r|r}
-		\hline
-		Data Set & Norm Est & Norm SE & Norm-M Est & Norm-M SE & AR Est & AR SE & AR-M Est & AR-M SE\\
-		\hline
-		$\sigma_X = 50, \sigma_Y = 2$ & 1.144401 & 0.0001429 & 1.144116 & 0.0001435 & 1.144199 & 0.0001433 & 1.144083 & 0.0001436\\
-		\hline
-		$\sigma_X = 500, \sigma_Y = 20$ & 1.161010 & 0.0001117 & 1.144155 & 0.0001434 & 1.157193 & 0.0001182 & 1.143996 & 0.0001438\\
-		\hline
-		$\sigma_X = 2500, \sigma_Y = 100$ & 1.203070 & 0.0000599 & 1.144496 & 0.0001427 & 1.195816 & 0.0000667 & 1.147499 & 0.0001365\\
-		\hline
-		$\sigma_X = 10000, \sigma_Y = 500$ & 1.248181 & 0.0000309 & 1.146211 & 0.0001391 & 1.237044 & 0.0000364 & 1.156277 & 0.0001198\\
-		\hline
-	\end{tabular}
-\end{table}
-
-\begin{table}[H]
-
-	\caption{\label{tab:}$\rr$ Estimates and Std. Errs, EXPGROWTH Model,
-		$X_0 = 99950, Y_0 = 50$, $\sigma_X = 100, \sigma_Y = 5$}
-	\centering
-	\begin{tabular}[t]{l|r|r|r|r|r|r|r|r}
-		\hline
-		Data Set & Norm Est & Norm SE & Norm-M Est & Norm-M SE & AR Est & AR SE & AR-M Est & AR-M SE\\
-		\hline
-		1st Order & 1.061431 & 0.0004905 & 1.060632 & 0.0004964 & 1.060995 & 0.0004937 & 1.061004 & 0.0004936\\
-		\hline
-		4th Order & 1.148524 & 0.0001344 & 1.148405 & 0.0001347 & 1.148513 & 0.0001344 & 1.148399 & 0.0001347\\
-		\hline
-		Linear SIR & 1.087741 & 0.0003320 & 1.087338 & 0.0003340 & 1.087538 & 0.0003330 & 1.087399 & 0.0003337\\
-		\hline
-	\end{tabular}
-\end{table}
-
-\begin{table}[H]
-
-	\caption{\label{tab:}$\rr$ Estimates and Std. Errs, EXPGROWTH Model,
-		$\beta = 0.06, \gamma = 0.03$, $X_0 = 99950, Y_0 = 50$, $\sigma_X = 100, \sigma_Y = 5$}
-	\centering
-	\begin{tabular}[t]{l|r|r|r|r|r|r|r|r}
-		\hline
-		Data Set & Norm Est & Norm SE & Norm-M Est & Norm-M SE & AR Est & AR SE & AR-M Est & AR-M SE\\
-		\hline
-		T = 200 & 1.306607 & 0.0004142 & 1.301989 & 0.0004286 & 1.304421 & 0.0004209 & 1.301390 & 0.0004305\\
-		\hline
-		T = 100 & 1.560029 & 0.0022192 & 1.418887 & 0.0035072 & 1.517223 & 0.0025481 & 1.390702 & 0.0038451\\
-		\hline
-		T = 50 & 2.238028 & 0.0078639 & 1.623615 & 0.0184718 & 2.186315 & 0.0084553 & 1.634726 & 0.0181912\\
-		\hline
-		T = 20 & 4.472654 & 0.0339963 & 1.108447 & 0.1728373 & 4.159546 & 0.0412621 & 2.711197 & 0.0923793\\
-		\hline
-	\end{tabular}
-\end{table}
-
-\begin{table}[H]
-
-	\caption{\label{tab:}$\rr$ Estimates and Std. Errs, EXPGROWTH Model,
-		$\beta = 0.06, \gamma = 0.03$, $\sigma_X = 100, \sigma_Y = 5$}
-	\centering
-	\begin{tabular}[t]{l|r|r|r|r|r|r|r|r}
-		\hline
-		Data Set & Norm Est & Norm SE & Norm-M Est & Norm-M SE & AR Est & AR SE & AR-M Est & AR-M SE\\
-		\hline
-		$X_0 = 99, Y_0 = 1$ & 1.091297 & 0.0003149 & 0.9202789 & 0.0036275 & 1.083304 & 0.0003546 & 0.8941523 & 0.0050462\\
-		\hline
-		$X_0 = 990, Y_0 = 10$ & 1.094724 & 0.0002992 & 0.9728240 & 0.0017765 & 1.086413 & 0.0003386 & 0.9799270 & 0.0016069\\
-		\hline
-		$X_0 = 9995, Y_0 = 5$ & 1.102157 & 0.0002679 & 1.0678388 & 0.0004461 & 1.094024 & 0.0003024 & 1.0676423 & 0.0004474\\
-		\hline
-		$X_0 = 999500, Y_0 = 500$ & 1.220466 & 0.0000464 & 1.2204454 & 0.0000464 & 1.220457 & 0.0000464 & 1.2204480 & 0.0000464\\
-		\hline
-	\end{tabular}
-\end{table}
-
-\subsection{IDEA}
-
-\begin{table}[H]
-
-	\caption{\label{tab:}$\rr$ Estimates and Std. Errs, IDEA Model,
-		$\beta = 0.06, \gamma = 0.03$, $X_0 = 99950, Y_0 = 50$, $\sigma_X = 100, \sigma_Y = 5$}
-	\centering
-	\begin{tabular}[t]{l|r|r|r|r|r|r|r|r}
-		\hline
-		Data Set & Norm Est & Norm SE & Norm-M Est & Norm-M SE & AR Est & AR SE & AR-M Est & AR-M SE\\
-		\hline
-		$\gamma$ = 0.001 & 1.114829 & 0.0459625 & 0.9591728 & 0.0488597 & 1.140804 & 0.0459639 & 0.8574921 & 0.0500578\\
-		\hline
-		$\gamma$ = 0.04 & 2.008614 & 0.0318940 & 2.0086026 & 0.0318958 & 2.008640 & 0.0318947 & 2.0086160 & 0.0318934\\
-		\hline
-		$\gamma$ = 0.06 & 2.038747 & 0.0313772 & 2.0387516 & 0.0313766 & 2.038764 & 0.0313758 & 2.0387101 & 0.0313764\\
-		\hline
-		$\gamma$ = 0.24 & 2.039700 & 0.0313775 & 2.0397001 & 0.0313775 & 2.039712 & 0.0313791 & 2.0396914 & 0.0313784\\
-		\hline
-	\end{tabular}
-\end{table}
-
-\begin{table}[H]
-
-	\caption{\label{tab:}$\rr$ Estimates and Std. Errs, IDEA Model,
-		$\beta = 0.06, \gamma = 0.03$, $X_0 = 99950, Y_0 = 50$, $\sigma_X = 100, \sigma_Y = 5$}
-	\centering
-	\begin{tabular}[t]{l|r|r|r|r|r|r|r|r}
-		\hline
-		Data Set & Norm Est & Norm SE & Norm-M Est & Norm-M SE & AR Est & AR SE & AR-M Est & AR-M SE\\
-		\hline
-		$X_0 = 99000, Y_0 = 1000$ & 1.873533 & 0.0318531 & 1.873446 & 0.0318541 & 1.873475 & 0.0318524 & 1.873496 & 0.0318549\\
-		\hline
-		$X_0 = 99900, Y_0 = 100$ & 1.923607 & 0.0328651 & 1.923653 & 0.0328648 & 1.923560 & 0.0328664 & 1.923576 & 0.0328694\\
-		\hline
-		$X_0 = 99990, Y_0 = 10$ & 1.979338 & 0.0325801 & 1.979345 & 0.0325796 & 1.979353 & 0.0325789 & 1.979404 & 0.0325774\\
-		\hline
-		$X_0 = 99999, Y_0 = 1$& 2.022337 & 0.0317890 & 2.022325 & 0.0317883 & 2.022367 & 0.0317860 & 2.022322 & 0.0317898\\
-		\hline
-	\end{tabular}
-\end{table}
-
-\begin{table}[H]
-
-	\caption{\label{tab:}$\rr$ Estimates and Std. Errs, IDEA Model,
-		$\beta = 0.06, \gamma = 0.03$, $X_0 = 99950, Y_0 = 50$}
-	\centering
-	\begin{tabular}[t]{l|r|r|r|r|r|r|r|r}
-		\hline
-		Data Set & Norm Est & Norm SE & Norm-M Est & Norm-M SE & AR Est & AR SE & AR-M Est & AR-M SE\\
-		\hline
-		$\sigma_X = 50, \sigma_Y = 2$ & 1.940295 & 0.0328946 & 1.940280 & 0.0328946 & 1.940313 & 0.0328958 & 1.940303 & 0.0328948\\
-		\hline
-		$\sigma_X = 500, \sigma_Y = 20$ & 1.940066 & 0.0328982 & 1.940463 & 0.0328877 & 1.939755 & 0.0329040 & 1.940506 & 0.0328904\\
-		\hline
-		$\sigma_X = 2500, \sigma_Y = 100$ & 1.940497 & 0.0328856 & 1.940101 & 0.0329012 & 1.937796 & 0.0329248 & 1.938458 & 0.0329298\\
-		\hline
-		$\sigma_X = 10000, \sigma_Y = 500$ & 1.934924 & 0.0329764 & 1.942861 & 0.0328652 & 1.910502 & 0.0346809 & 1.913664 & 0.0339201\\
-		\hline
-	\end{tabular}
-\end{table}
-
-\begin{table}[H]
-
-	\caption{\label{tab:}$\rr$ Estimates and Std. Errs, IDEA Model,
-		$X_0 = 99950, Y_0 = 50$, $\sigma_X = 100, \sigma_Y = 5$}
-	\centering
-	\begin{tabular}[t]{l|r|r|r|r|r|r|r|r}
-		\hline
-		Data Set & Norm Est & Norm SE & Norm-M Est & Norm-M SE & AR Est & AR SE & AR-M Est & AR-M SE\\
-		\hline
-		1st Order & 2.034020 & 0.0314227 & 2.034020 & 0.0314227 & 2.034018 & 0.0314231 & 2.034013 & 0.0314227\\
-		\hline
-		4th Order & 1.938669 & 0.0331254 & 1.938646 & 0.0331271 & 1.938678 & 0.0331252 & 1.938644 & 0.0331271\\
-		\hline
-		Linear SIR & 2.033555 & 0.0314612 & 2.033554 & 0.0314611 & 2.033555 & 0.0314609 & 2.033547 & 0.0314615\\
-		\hline
-	\end{tabular}
-\end{table}
-
-\begin{table}[H]
-
-	\caption{\label{tab:}$\rr$ Estimates and Std. Errs, IDEA Model,
-		$\beta = 0.06, \gamma = 0.03$, $X_0 = 99950, Y_0 = 50$, $\sigma_X = 100, \sigma_Y = 5$}
-	\centering
-	\begin{tabular}[t]{l|r|r|r|r|r|r|r|r}
-		\hline
-		Data Set & Norm Est & Norm SE & Norm-M Est & Norm-M SE & AR Est & AR SE & AR-M Est & AR-M SE\\
-		\hline
-		T = 200 & 2.872029 & 0.0777506 & 2.872039 & 0.0777500 & 2.872024 & 0.0777475 & 2.871995 & 0.0777522\\
-		\hline
-		T = 100 & 4.777225 & 0.2173201 & 4.777183 & 0.2173166 & 4.777046 & 0.2173274 & 4.777196 & 0.2173002\\
-		\hline
-		T = 50 & 8.523307 & 0.6084195 & 8.523224 & 0.6084023 & 8.523328 & 0.6084437 & 8.523118 & 0.6083556\\
-		\hline
-		T = 20 & 19.536946 & 2.3356802 & 19.536830 & 2.3354926 & 19.537515 & 2.3353411 & 19.534388 & 2.3356299\\
-		\hline
-	\end{tabular}
-\end{table}
-
-\begin{table}[h]
-
-	\caption{\label{tab:}$\rr$ Estimates and Std. Errs, IDEA Model,
-		$\beta = 0.06, \gamma = 0.03$, $\sigma_X = 100, \sigma_Y = 5$}
-	\centering
-	\begin{tabular}[t]{l|r|r|r|r|r|r|r|r}
-		\hline
-		Data Set & Norm Est & Norm SE & Norm-M Est & Norm-M SE & AR Est & AR SE & AR-M Est & AR-M SE\\
-		\hline
-		$X_0 = 99, Y_0 = 1$ & 1.247573 & 0.0297242 & 1.454198 & 0.0194568 & 1.212651 & 0.0297805 & 1.411583 & 0.0210353\\
-		\hline
-		$X_0 = 990, Y_0 = 10$ & 1.440702 & 0.0207630 & 1.461154 & 0.0196739 & 1.369004 & 0.0237707 & 1.436235 & 0.0202001\\
-		\hline
-		$X_0 = 9995, Y_0 = 5$ & 1.940396 & 0.0329020 & 1.940602 & 0.0328856 & 1.939845 & 0.0328895 & 1.939680 & 0.0328886\\
-		\hline
-		$X_0 = 999500, Y_0 = 500$ & 1.940281 & 0.0328949 & 1.940280 & 0.0328949 & 1.940286 & 0.0328948 & 1.940299 & 0.0328944\\
-		\hline
-	\end{tabular}
-\end{table}
-
 \end{document}

### 2018-07-03 paper/Version14/draft_v14.tex
finally on git

--- 2018-07-03 paper/Version14/draft_v14 (Shannon Gallagher's conflicted copy 2018-06-26).tex
@@
 Despite a clear definition of $\rr$, epidemiologists have struggled to create a standard  estimator for $\rr$  \citep{hethcote2000}.
-A major issue in estimating $\rr$ is that the quantity is a \textit{property of the model}, meaning that the estimate $\hat{\rr}$ is dependent not only on the usual noise that comes with statistical modeling but also on a variety of assumptions on how researchers assume a disease is transmitted through a population \citep{diekmann2009}.
+A major issue in estimating $\rr$ is that the quantity is a \textit{property of the model}, meaning that $\rr$ is dependent not only on the usual noise that comes with statistical modeling but also on a variety of assumptions on how researchers assume a disease is transmitted through a population \citep{diekmann2009}.
 
@@
 Here, S and I are known as compartments where S stands for ``susceptible'' and I for ``infectious.'' Infectious disease models then specify how individuals move from the S compartment to the I compartment.
-Often, ancillary compartments are added such as R, which stands for recovered or removed, or E, which stands for exposed but not yet infectious.
+Often, ancillary compartments are added such as R, which stands for recovered or dead, or E, which stands for exposed but not yet infectious.
 The estimate of $\rr$ from a SIR model will not be directly comparable to the estimate of $\rr$ from a SEIR model, for example.
@@
 We note that these methods are not exhaustive but are chosen due to their impact and use in epidemiology.
-We find that these methods also help to highlight the problems mentioned above.
+We find that these methods also help to highlight the problems discussed above.
 
@@
 For example, the SIR model is special in that two of its compartments, S and R are monotonic.
-The number of susceptibles is non-increasing and the number of recovereds is non-decreasing.
+The number of susceptibles is non-increasing and the number of susceptibles is non-decreasing.
 Logically, models should enforce this monotonicity of the two compartments.
 As a consequence, this restricts assumptions on how randomness enters the model.
-For example, Gaussian error, a common assumption in statistics and machine learning, becomes highly suspect due to the assymmetrical nature of the noise imposed by the monotonic $S$ and $R$ compartments.
+For example, Gaussian error, a common assumption in statistics and machine learning, becomes highly suspect due to the non-symmetrical nature of the noise imposed by the monotonic $S$ and $R$ compartments.
 Without assuming Gaussian noise, it becomes much more difficult to guarantee properties such as unbiased and consistent estimates of $\rr$.
@@
 The lowest possible estimate for $\rr$ reported is 0.12 for HIV in Uganda, and the largest is 17.00 for measles in England and Wales.
-Within diseases, we see that estimates for $\rr$ seem to be similar, although the respective CIs typically do not intersect, suggesting that there is disparity in estimates of $\rr$.
+Within diseases, we see that estimates for $\rr$ seem to be similar, although typically do not lie in one another's intervals, suggesting that there is disparity in estimates of $\rr$.
 
@@
 The parameters  $\beta$ and $\gamma$ have epidemiological meaning.
-Here, $\beta$ is the inverse of expected time to infection rate and $\gamma$ is the inverse of the expected time to recovery.
+Here, $\beta$ is the average infection rate and $\gamma$ is the average recovery rate.
 The movement of individuals from one compartment to another is represented through the ordinary differential equations below.
@@
 \end{align}
-In words, susceptible individuals become infected at a rate that is proportional to the percentage of infected individuals multiplied by $\beta$, the inverse of expected time to infection, and the number of susceptible individuals.
+In words, susceptible individuals become infected at a rate that is proportional to the percentage of infected individuals multiplied by $\beta$, the infection rate, and the number of susceptible individuals.
 Infectious individuals recover at a rate of $\gamma$ multiplied by the number of infected individuals.
@@
 That is,  the rate of new infections is greater than the rate of recovery.
-So as long as the number of initially susceptible individuals is large compared to the total population, $\frac{N}{X} \approx 1$, then an outbreak will occur if $\rr >1$,
+So as long as the number of initially susceptible individuals is large compared to the total population, $\frac{X}{N} \approx 1$, then an outbreak will occur if $\rr >1$,
 \begin{align*}
@@
   \end{align}
-We are assuming the observations are generated based on the discretized time (replacing $d$ with $\Delta$), deterministic ODEs presented in Equation \ref{eq:sir} with the addition of time ($t=0, 1, \dots, T$) and compartment dependent noise $\epsilon_{X,t}$ and $\epsilon_{Z,t}$.
+We are assuming the observations are generated based on the discretizied time, deterministic ODEs presented in Equation \ref{eq:sir} with the addition of time ($t=0, 1, \dots, T$) and compartment dependent noise $\epsilon_{X,t}$ and $\epsilon_{Z,t}$.
 Since, $N$, the total population, is constant, then $Y_{obs}$ is adjusted accordingly.
@@
 \begin{align*}
-  H_0:\;& \rr > 1 \\
-  H_A:\;& \rr \le 1
+  H_0:\;& \rr < 1 \\
+  H_A:\;& \rr \ge 1
 \end{align*}
@@
 \begin{align*}
-P(\hat{a} \le \rr \le \hat{b}) \le 1 - \alpha.
-\end{align*}
-This property is particularly desirable, as we can estimate $\hat{a}$ and $\hat{b}$ from the distribution $F$.
+P(\hat{a} \le \rr \le \hat{b}) \le 1 - \alpha,  
+\end{align*}
+the second property is particularly desirable, as we can estimate $\hat{a}$ and $\hat{b}$ from the distribution $F$.
 At any rate, we would like to be able to estimate $\hat{a}$ and $\hat{b}$ for a given $\alpha$-level.
@@
 
-\subsection{$\rr$ is a property of the model}\label{sec:prop}
+\subsection{$\rr$ is a property of the model}
 
@@
 Consider the case when $N=10000, (\beta, \gamma) = (.06, .03)$, $(X(0), E(0), Y(0), Z(t)) = (9500, 0, 500, 0)$ for times $t=1, \dots, 365$.
-Let the parameter $\alpha$ range between [0.005, 0.25] with step size $0.005$.
-This generates a first data set
+Let the parameter $\alpha$ range between [.005, 1].
+This generates data set
 $$D_{\alpha, \textnormal{XEYZ}} = \left \{(X_{XEYZ}(t), E_{XEYZ}(t), Y_{XEYZ}(t), Z_{XEYZ}(t): t=1, \dots, 365\right \},$$
@@
 Now assume that researcher 2 misspecifies the model for this data as a XYZ model rather than a XEYZ model.
-That is, she observes all exposed individuals as susceptible and generates a second data set
-$$D_{\alpha, XYZ} = \left \{X_{XYZ}(t), Y_{XYZ}(t), Z_{XEYZ}(t) : t = 1, \dots, 365 \right \},$$
-where $X_{XYZ}(t) = X_{XEYZ}(t) + E_{XEYZ}(t)$, $Y_{XYZ}(t) = Y_{XEYZ}(t)$, and  $Z_{XEYZ}(t) = Z_{XYZ}(t)$. This situation is illustrated in Figure \ref{fig:sir-vs-seir} with $\alpha = 0.01$.
+That is, she places all exposed individuals in the X compartment to have data set
+$$D_{\alpha, XYZ} = \left \{X_{XYZ}(t) = X_{XEYZ} + E_{XEYZ}, Y_{XYZ} = Y_{XEYZ}, Z_{XEYZ} = Z_{XYZ} : t = 1, \dots, 365 \right \}.$$
+This situation is illustrated in Figure \ref{fig:sir-vs-seir} with $\alpha = 0.01$.
 
@@
 
+For those
+
 \begin{figure}
@@
   \includegraphics[width=\textwidth]{images/r0-est.pdf}
-  \caption{Estimates of $\rr$ from both the XEYZ and XYZ model specification for different values of $\alpha$ using Least Squares with $\beta=0.06$, $\gamma=0.03$, and $(X(0), E(0), Y(0)) = (9500, 500, 0)$.}\label{fig:r0-alpha}
-  \end{figure}
-
-For those readers not convinced by our small example, we offer this mathematical intuition of the next generation model (NGM) described by \cite{diekmann1990} and \cite{diekmann2009}.
-
- The NGM is a generalization of any compartment model at the infection-free steady state.
-This model solves the problem of having an expression for $\rr$ in terms of the epidemiological parameters for a general compartment model.
-Originally, introduced by \citep{diekmann1990}, \cite{diekmann2009} posit a recipe to find $\rr$ for a wide class of compartment models, including SIR and SEIR.
-First, define the infected subsystem as ``the equations of the ODE system that describe the production of new infections and changes in state among infected individuals.''  Let $x = (C_1, C_2, \dots, C_m)$ where $C_i$ are the different compartments of the infected subsystem.
-The steps to find $\rr$ are as follows:
-
-\begin{enumerate}
-\item ``Linearize the infected subsystem of nonlinear ODEs about the infection-free steady state''
-\item Decompose the linearized infected subsystem into $(T + \Sigma )x$ where $T$ is a $m\times m$ matrix of transmissions and $\Sigma$ is a $m \times m$ matrix of transitions.
-\item $\rr$ is the spectral radius (i.e. the dominant eigenvalue) of $-T \Sigma^{-1}$.  
-\end{enumerate}
-
-Here, $T_{ij}$ is the rate of transmission of \textit{newly} infected individuals in state $i$ created by individuals in state $j$.
-$\Sigma_{ij}$ is the transition rate of individuals into compartment $i$ from compartment $j$.
-
-This method has the assumptions of a basic compartment model: homogeneous mixing among populations and the law of mass action, i.e., that the number of infected individuals are proportional to the number of infected individuals at the previous step.
-Additionally, this method is advantageous in that the matrix $-\Sigma^{-1}$ has intuitive meaning.
-The $ij$th entry of $- \Sigma^{-1}$ is the ``expected time that an individual who presently has state $j$ will spend in infected state $i$.''
-
-The NGM shows that $\rr$ is the spectral radius of a matrix composed from the transmission of disease from diseased individuals to susceptible individuals and transition of diseased individuals from among the compartments and is indeed a property of the model.
+  \caption{Estimates of $\rr$ from both the XEYZ and XYZ model specification for different values of $\alpha$.}\label{fig:r0-alpha}
+  \end{figure}
 
@@
 
-In Section \ref{sec:prop} we described how $\rr$ is a property of the model.
-To fairly compare estimation methods to one another, we restrict our methods to estimate $\rr$ to the SIR model.
-Also in Section \ref{sec:prop}, we examined simulated data that was generated from a SEIR model without noise.
-However, we know that data is very noisy.
-In this section, we describe the \wxxsir methods to estimate $\rr$ for the SIR model in the presence of noise in detail, including advantages and disadvantages of each.
+In this section, we describe the \wxxsir methods in detail, including advantages and disadvantages of each.
 
@@
 \end{align*}
-Then the least squares (LS) estimate for $\hat{\beta}$ and $\hat{\gamma}$ is then combined to give an estimate of $\hat{\rr}$ in Equation \ref{eq:sirls},
+Then the least squares (LS) estimate for $\rr$ is given by Equation \ref{eq:sirls},
 \begin{align}\label{eq:sirls}
@@
 We find
-\begin{align}
+\begin{align*}
 (\hat{\rr}, \hat{\gamma} ) &= \text{argmin}_{\rr, \gamma} \sum_{t} \left [ \left (X_{obs}(t) - X(t)\right )^2 + \left ( Z_{obs}(t) - Z(t) \right )^2 \right ]
-\end{align}
+\end{align*}
 We use the $\hat{\rr}$ directly from the above estimation problem, which again can be found with a grid search or another optimization process.
-
-We can also write $\hat{\rr}$ as a function of $\gamma^*$,
-\begin{align*}
-\hat{\rr}(\gamma^*)&= \text{argmin}_{\rr} \sum_{t} \left [ \left (X_{obs}(t) - X(t; \gamma^*)\right )^2 + \left ( Z_{obs}(t) - Z(t; \gamma^*) \right )^2 \right ].
-  \end{align*}
-
 It may be surprising that this method leads to different results than simply using least squares.
-We also see that this method further results in different standard errors than in LS, which we discuss further in Section \ref{sec:results}.
+One reason we obtain different results may be attributed to better numerical precision as there is one fewer step to obtain an estimate for $\rr$.
+However, we also see that this method further results in different standard errors than in LS, which we discuss further in Section \ref{sec:results}.
 
@@
 \end{align*}
-An estimator for $\rr$ is derived from the ODEs in Equation \eqref{eq:sir},
+Following,  an estimator for \(\rr\) is derived from the ODEs in Equation \eqref{eq:sir},
 \begin{align}
@@
   \end{align}
-  Here, $K$ is arbitrary and should be selected using some criterion such as AIC (REF).
+  Here, $K$ is arbitrary and should be selected using some criterion such as AIC.
 Besides optionally deciding on the degree of polynomials to fit, this model is simple to implement and gives comparable results to using least squares with the SIR model.
@@
 The above formulation (Section \ref{linear-model-approximation-degree-10}) of a linear model approximation only uses the estimate at time $t=0$ to estimate $\rr$.
-We can instead, use all time points $T$ available to estimate $\rr$.
+We can instead, use all time points $O$ available to estimate $\rr$.
 We fit a polynomial in \(t\) with degree \(K\) to \(X_{obs}\)
@@
 \begin{align*}
-  \hat{\rr} &= \frac{1}{T} \sum_t \frac{-\hat{X}^\prime(t)}{\hat{Z}^\prime(t)} \cdot \frac{N}{X(0)} 
+  \hat{\rr} &= \frac{1}{O} \sum_t \frac{-\hat{X}^\prime(t)}{\hat{Z}^\prime(t)} \cdot \frac{N}{X(0)} 
 \end{align*}
@@
 An advantage to this approach is that we have a more robust estimate of $\rr$ than just using one time point.
-As in LMA, we can examine how sensitive $\rr$ is to the order of the polynomial $K$.
+Like in LMA, we can examine how sensitive $\rr$ is to the order of the polynomial $K$.
 
@@
 One, the estimate of $\rr$ is highly interpretable.
-Two, an assumption of Gaussian noise in Equation \ref{eq:harko_lin} is not as strong as an assumption in the other methods to estimate $\rr$ because we are looking at a log transformation of the percent of suceptibles instead of raw variables.
-If the assumption holds, we can show that the resulting estimate of $\rr$ is unbiased and consistent.
+Two, an assumption of Gaussian noise in Equation \ref{eq:harko_lin} is not as strong as an assumption in the other methods to estimate $\rr$ because we are looking at a log transformation of the the percent of suceptibles instead of raw variables.
+If the assumption holds, we can show that the resulting estimate of $\rr$ is an unbiased and a consistent estimator.
 Finally,  this method is easily implemented using any linear modeling software.
@@
 We have explicitly stated how randomness enters the model so it is up to the researcher to check whether this assumption is a good fit.
-Using this method, it is easy to see how sensitive $\hat{\rr}$ is to changes in $\hat{\alpha}$ as $\hat{\rr}$ is a function of $\alpha$.
+Using this method, it is possible to simulate how sensitive $\hat{\rr}$ is to changes in $\hat{\alpha}$.
 
@@
 \end{align*}
-Then, the posterior distribution of $\rr$ given the previous time period's incidence is
+Then, the posterior distribution of $\rr$ given the previous days' incidence is
 \begin{align*}
@@
 \end{align*}
-This method is sequential in that the prior distribution for $\rr$ comes from the previous time period.
+This method is sequential in that the prior distribution for $\rr$ comes from the previous day.
 The initial prior for $\rr$ is assumed to be flat.
@@
 \begin{align*}
-  \sqrt{T} \left ( \hat{\mathbf{\theta}}_o - \mathbf{\theta} \right ) \overset{d}{\to} N\left ( \mathbf{0}, \Sigma \right  )
+  \sqrt{O} \left ( \hat{\mathbf{\theta}}_o - \mathbf{\theta} \right ) \overset{d}{\to} N\left ( \mathbf{0}, \Sigma \right  )
 \end{align*}
@@
 \begin{align*}
-  \sqrt{T} \left (h(\hat{\theta}_n) - h(\theta) \right ) \overset{d}{\to} N \left (0,  \nabla^T h(\theta)\Sigma \nabla h(\theta)\right ),
+  \sqrt{O} \left (h(\hat{\theta}_n) - h(\theta) \right ) \overset{d}{\to} N \left (0,  \nabla^T h(\theta)\Sigma \nabla h(\theta)\right ),
 \end{align*}
@@
 This method may be extended to other frameworks such as the SEIR model.
-The advantages of the delta method are that since the Next Generation Matrix \citep{diekmann2009} gives a recipe for deriving $\rr$ from the deterministic ODEs, then it is theoretically possible to use the delta method to derive an expression for the variance of $\rr$ for any compartment model.
+The advantages of the delta method are that since the Next Generation Matrix \citep{diekmann2009} gives recipe for deriving $\rr$ from the deterministic ODEs, then it is theoretically possible to use the delta method to derive an expression for the variance of $\rr$ for any compartment model.
 
@@
 \begin{align*}
-  V\left [ \hat{\rr} \right ] &\approx V\left [\hat{\rr}_\ell \right ], \; \ell = 1, \dots, L
+  V\left [ \hat{\rr} \right ] &\approx V\left [\hat{\rr}_\ell \right ].
 \end{align*}
@@
 
-A choice one must make in the block bootstrap is the block size $b$.
-One simple recommendation is to take $b= Cn^{1/3}$, which is derived from the asymptotic mean square error of the method.
-\cite{hall1995} describe how to calculate $C$, which is context dependent.
+A choice one must make in the block bootstrap is the block size $b$.  One simple recommendation is to take $b= n^{1/3}$, which is derived from the asymptotic mean square error of the method.
 
@@
 Then commonly, $\hat{\rr} = \hat{F}_{.5}$, the median of $\hat{F}$.
-Correspondingly, the 95\% credible interval is $\left[\hat{F}_{.025}, \hat{F}_{.975} \right ]$.
+Correspondingly, the 95\% credible interval is $\left[\hat{F}_{.05}, \hat{F}_{.95} \right ]$.
 Advantages of this method include having an entire distribution instead of a point estimate.
@@
 
-	Number of time steps $T$: 365
+	Number of time steps $O$: 365
 
@@
 \toprule
-\textbf{Name} & \textbf{Model} & \textbf{Error} & \textbf{($\beta, \gamma$)} & \textbf{T} & \textbf{(X(0),Y(0))} & \textbf{N} & \textbf{($\sigma_X, \sigma_Y$)} \\ \midrule
+\textbf{Name} & \textbf{Model} & \textbf{Error} & \textbf{($\beta, \gamma$)} & \textbf{O} & \textbf{(X(0),Y(0))} & \textbf{N} & \textbf{($\sigma_X, \sigma_Y$)} \\ \midrule
   Baseline        & SIR      &   Norm             & (.06, .03)                          &     365       &  (99950, 50)                    &  $10^5$          &(100, 5)                                \\ \midrule
@@
 There are numerous aspects to consider when using a method to estimate $\rr$.
-Among these include values of both the infection and recovery rate $(\beta, \gamma)$, the number of data points available ($T$), the total size of the population ($N$), the initial percent of susceptible individuals $\left (\frac{X(0)}{N}\right)$, the initial percent of infectious individuals $\left (\frac{Y(0)}{N}\right )$, the magnitude of the variance of both the number of susceptibles and the number of infectious ($\sigma_X^2, \sigma_Y^2$), the error modeling assumptions, data generated from a non-SIR model, and finally methods of estimating the variance.
+Among these include values of both the infection and recovery rate $(\beta, \gamma)$, the number of data points available ($O$), the total size of the population ($N$), the initial percent of susceptible individuals $\left (\frac{X(0)}{N}\right)$, the initial percent of infectious individuals $\left (\frac{Y(0)}{N}\right )$, the magnitude of the variance of both the number of susceptibles and the number of infectious ($\sigma_X^2, \sigma_Y^2$), the error modeling assumptions, data generated from a non-SIR model, and finally methods of estimating the variance.
 We first examine our baseline data, and then we examine each of these aspects individually.
@@
   \includegraphics[scale=0.5]{images/BaseBase.jpeg}
-  \caption{Forest plots of estimates of $\rr$ for the \xxsir methods for the baseline data set $(\beta=.06, \gamma=.03, T=365, X(0)=99950, Y(0)=50, Z(0)=0, \sigma_X=100, \sigma_Y=5, N=10^5)$.
+  \caption{Forest plots of estimates of $\rr$ for the \xxsir methods for the baseline data set $(\beta=.06, \gamma=.03, O=365, X(0)=99950, Y(0)=50, Z(0)=0, \sigma_X=100, \sigma_Y=5, N=10^5)$.
 The points are the estimates of $\rr$ and the lines denote $\pm 2\cdot $Std.
@@
 	\end{tabular}
-        \caption{Baseline Data $(\beta=.06, \gamma=.03, T=365, X(0)=99950, Y(0)=50, Z(0)=0, \sigma_X=100, \sigma_Y=5, N=10^5)$, $\rr$ Estimates and Standard Errors}\label{tab:baseline-res}
+        \caption{Baseline Data $(\beta=.06, \gamma=.03, O=365, X(0)=99950, Y(0)=50, Z(0)=0, \sigma_X=100, \sigma_Y=5, N=10^5)$, $\rr$ Estimates and Standard Errors}\label{tab:baseline-res}
 \end{table}
@@
   \includegraphics[scale=0.5]{images/AR.jpeg}
-  \caption{Forest plots of estimates of $\rr$ for the \xxsir methods for the autoregressive (AR) ($\rho=.7)$ simulated data set $(\beta=.06, \gamma=.03, T=365, X(0)=99950, Y(0)=50, Z(0)=0, \sigma_X=100, \sigma_Y=5, N=10^5)$.
+  \caption{Forest plots of estimates of $\rr$ for the \xxsir methods for the autoregressive (AR) ($\rho=.7)$ simulated data set $(\beta=.06, \gamma=.03, O=365, X(0)=99950, Y(0)=50, Z(0)=0, \sigma_X=100, \sigma_Y=5, N=10^5)$.
 The points are the estimates of $\rr$ and the lines denote $\pm 2\cdot $Std.
@@
 	\includegraphics[scale=0.5]{images/mono.jpeg}
-	\caption{Forest plots of estimates of $\rr$ for the \xxsir methods for the Norm-M and AR-M data sets $(\beta=.06, \gamma=.03, T=365, X(0)=99950, Y(0)=50, Z(0)=0, \sigma_X=100, \sigma_Y=5, N=10^5)$.
+	\caption{Forest plots of estimates of $\rr$ for the \xxsir methods for the Norm-M and AR-M data sets $(\beta=.06, \gamma=.03, O=365, X(0)=99950, Y(0)=50, Z(0)=0, \sigma_X=100, \sigma_Y=5, N=10^5)$.
 The points are the estimates of $\rr$ and the lines denote $\pm 2\cdot $Std.
@@
   \includegraphics[scale=0.5]{images/parchange_n.jpeg}
-  \caption{Forest plots of estimates of $\rr$ for the \xxsir methods for the infrec1-4 data sets $(\beta=.06, \gamma=.03, T=365, X(0)=99950, Y(0)=50, Z(0)=0, \sigma_X=100, \sigma_Y=5, N=10^5)$.
+  \caption{Forest plots of estimates of $\rr$ for the \xxsir methods for the infrec1-4 data sets $(\beta=.06, \gamma=.03, O=365, X(0)=99950, Y(0)=50, Z(0)=0, \sigma_X=100, \sigma_Y=5, N=10^5)$.
 The points are the estimates of $\rr$ and the lines denote $\pm 2\cdot $Std.
@@
 	\includegraphics[scale=0.5]{images/parchange_nm.jpeg}
-	\caption{Forest plots of estimates of $\rr$ for the \xxsir methods for the infrec1-4 data sets $(\beta=.06, \gamma=.03, T=365, X(0)=99950, Y(0)=50, Z(0)=0, \sigma_X=100, \sigma_Y=5, N=10^5)$.
+	\caption{Forest plots of estimates of $\rr$ for the \xxsir methods for the infrec1-4 data sets $(\beta=.06, \gamma=.03, O=365, X(0)=99950, Y(0)=50, Z(0)=0, \sigma_X=100, \sigma_Y=5, N=10^5)$.
 The points are the estimates of $\rr$ and the lines denote $\pm 2\cdot $Std.
@@
 	\includegraphics[scale=0.5]{images/parchange_ar.jpeg}
-	\caption{Forest plots of estimates of $\rr$ for the \xxsir methods for the infrec1-4 data sets $(\beta=.06, \gamma=.03, T=365, X(0)=99950, Y(0)=50, Z(0)=0, \sigma_X=100, \sigma_Y=5, N=10^5)$.
+	\caption{Forest plots of estimates of $\rr$ for the \xxsir methods for the infrec1-4 data sets $(\beta=.06, \gamma=.03, O=365, X(0)=99950, Y(0)=50, Z(0)=0, \sigma_X=100, \sigma_Y=5, N=10^5)$.
 The points are the estimates of $\rr$ and the lines denote $\pm 2\cdot $Std.
@@
 	\includegraphics[scale=0.5]{images/parchange_arm.jpeg}
-	\caption{Forest plots of estimates of $\rr$ for the \xxsir methods for the infrec1-4 data sets $(\beta=.06, \gamma=.03, T=365, X(0)=99950, Y(0)=50, Z(0)=0, \sigma_X=100, \sigma_Y=5, N=10^5)$.
+	\caption{Forest plots of estimates of $\rr$ for the \xxsir methods for the infrec1-4 data sets $(\beta=.06, \gamma=.03, O=365, X(0)=99950, Y(0)=50, Z(0)=0, \sigma_X=100, \sigma_Y=5, N=10^5)$.
 The points are the estimates of $\rr$ and the lines denote $\pm 2\cdot $Std.
@@
 
-\subsection{Number of Time Points ($T$)}\label{sec:res-time}
+\subsection{Number of Time Points ($O$)}\label{sec:res-time}
 We examine how the number of time steps the estimates of $\rr$ in the different methods.
@@
   \includegraphics[scale=0.5]{images/time_n.jpeg}
-  \caption{Forest plots of estimates of $\rr$ for the \xxsir methods for the Time1-4 data sets with Gaussian errors $(\beta=.06, \gamma=.03, T=365, X(0)=99950, Y(0)=50, Z(0)=0, \sigma_X=100, \sigma_Y=5, N=10^5)$.
+  \caption{Forest plots of estimates of $\rr$ for the \xxsir methods for the Time1-4 data sets with Gaussian errors $(\beta=.06, \gamma=.03, O=365, X(0)=99950, Y(0)=50, Z(0)=0, \sigma_X=100, \sigma_Y=5, N=10^5)$.
 The points are the estimates of $\rr$ and the lines denote $\pm 2\cdot $Std.
@@
 	\includegraphics[scale=0.5]{images/time_nm.jpeg}
-	\caption{Forest plots of estimates of $\rr$ for the \xxsir methods for the Time1-4 data sets with Gaussian errors $(\beta=.06, \gamma=.03, T=365, X(0)=99950, Y(0)=50, Z(0)=0, \sigma_X=100, \sigma_Y=5, N=10^5)$.
+	\caption{Forest plots of estimates of $\rr$ for the \xxsir methods for the Time1-4 data sets with Gaussian errors $(\beta=.06, \gamma=.03, O=365, X(0)=99950, Y(0)=50, Z(0)=0, \sigma_X=100, \sigma_Y=5, N=10^5)$.
 The points are the estimates of $\rr$ and the lines denote $\pm 2\cdot $Std.
@@
 	\includegraphics[scale=0.5]{images/time_ar.jpeg}
-	\caption{Forest plots of estimates of $\rr$ for the \xxsir methods for the Time1-4 data sets with Gaussian errors $(\beta=.06, \gamma=.03, T=365, X(0)=99950, Y(0)=50, Z(0)=0, \sigma_X=100, \sigma_Y=5, N=10^5)$.
+	\caption{Forest plots of estimates of $\rr$ for the \xxsir methods for the Time1-4 data sets with Gaussian errors $(\beta=.06, \gamma=.03, O=365, X(0)=99950, Y(0)=50, Z(0)=0, \sigma_X=100, \sigma_Y=5, N=10^5)$.
 The points are the estimates of $\rr$ and the lines denote $\pm 2\cdot $Std.
@@
 	\includegraphics[scale=0.5]{images/time_arm.jpeg}
-	\caption{Forest plots of estimates of $\rr$ for the \xxsir methods for the Time1-4 data sets with Gaussian errors $(\beta=.06, \gamma=.03, T=365, X(0)=99950, Y(0)=50, Z(0)=0, \sigma_X=100, \sigma_Y=5, N=10^5)$.
+	\caption{Forest plots of estimates of $\rr$ for the \xxsir methods for the Time1-4 data sets with Gaussian errors $(\beta=.06, \gamma=.03, O=365, X(0)=99950, Y(0)=50, Z(0)=0, \sigma_X=100, \sigma_Y=5, N=10^5)$.
 The points are the estimates of $\rr$ and the lines denote $\pm 2\cdot $Std.
@@
   \includegraphics[scale=0.5]{images/start_n.jpeg}
-    \caption{Forest plots of estimates of $\rr$ for the \xxsir methods for the Inits1-4 data sets $(\beta=.06, \gamma=.03, T=365, X(0)=99950, Y(0)=50, Z(0)=0, \sigma_X=100, \sigma_Y=5, N=10^5)$.
+    \caption{Forest plots of estimates of $\rr$ for the \xxsir methods for the Inits1-4 data sets $(\beta=.06, \gamma=.03, O=365, X(0)=99950, Y(0)=50, Z(0)=0, \sigma_X=100, \sigma_Y=5, N=10^5)$.
 The points are the estimates of $\rr$ and the lines denote $\pm 2\cdot $Std.
@@
 	\includegraphics[scale=0.5]{images/start_nm.jpeg}
-	\caption{Forest plots of estimates of $\rr$ for the \xxsir methods for the Inits1-4 data sets $(\beta=.06, \gamma=.03, T=365, X(0)=99950, Y(0)=50, Z(0)=0, \sigma_X=100, \sigma_Y=5, N=10^5)$.
+	\caption{Forest plots of estimates of $\rr$ for the \xxsir methods for the Inits1-4 data sets $(\beta=.06, \gamma=.03, O=365, X(0)=99950, Y(0)=50, Z(0)=0, \sigma_X=100, \sigma_Y=5, N=10^5)$.
 The points are the estimates of $\rr$ and the lines denote $\pm 2\cdot $Std.
@@
 	\includegraphics[scale=0.5]{images/start_ar.jpeg}
-	\caption{Forest plots of estimates of $\rr$ for the \xxsir methods for the Inits1-4 data sets $(\beta=.06, \gamma=.03, T=365, X(0)=99950, Y(0)=50, Z(0)=0, \sigma_X=100, \sigma_Y=5, N=10^5)$.
+	\caption{Forest plots of estimates of $\rr$ for the \xxsir methods for the Inits1-4 data sets $(\beta=.06, \gamma=.03, O=365, X(0)=99950, Y(0)=50, Z(0)=0, \sigma_X=100, \sigma_Y=5, N=10^5)$.
 The points are the estimates of $\rr$ and the lines denote $\pm 2\cdot $Std.
@@
 	\includegraphics[scale=0.5]{images/start_arm.jpeg}
-	\caption{Forest plots of estimates of $\rr$ for the \xxsir methods for the Inits1-4 data sets $(\beta=.06, \gamma=.03, T=365, X(0)=99950, Y(0)=50, Z(0)=0, \sigma_X=100, \sigma_Y=5, N=10^5)$.
+	\caption{Forest plots of estimates of $\rr$ for the \xxsir methods for the Inits1-4 data sets $(\beta=.06, \gamma=.03, O=365, X(0)=99950, Y(0)=50, Z(0)=0, \sigma_X=100, \sigma_Y=5, N=10^5)$.
 The points are the estimates of $\rr$ and the lines denote $\pm 2\cdot $Std.
@@
 	\includegraphics[scale=0.5]{images/popsize_n.jpeg}
-	\caption{Forest plots of estimates of $\rr$ for the \xxsir methods for the PopSize1-4 data sets $(\beta=.06, \gamma=.03, T=365, \sigma_X=100, \sigma_Y=5)$.
+	\caption{Forest plots of estimates of $\rr$ for the \xxsir methods for the PopSize1-4 data sets $(\beta=.06, \gamma=.03, O=365, \sigma_X=100, \sigma_Y=5)$.
 The points are the estimates of $\rr$ and the lines denote $\pm 2\cdot $Std.
@@
 	\includegraphics[scale=0.5]{images/popsize_nm.jpeg}
-	\caption{Forest plots of estimates of $\rr$ for the \xxsir methods for the PopSize1-4 data sets $(\beta=.06, \gamma=.03, T=365, \sigma_X=100, \sigma_Y=5)$.
+	\caption{Forest plots of estimates of $\rr$ for the \xxsir methods for the PopSize1-4 data sets $(\beta=.06, \gamma=.03, O=365, \sigma_X=100, \sigma_Y=5)$.
 The points are the estimates of $\rr$ and the lines denote $\pm 2\cdot $Std.
@@
 	\includegraphics[scale=0.5]{images/popsize_ar.jpeg}
-	\caption{Forest plots of estimates of $\rr$ for the \xxsir methods for the PopSize1-4 data sets $(\beta=.06, \gamma=.03, T=365, \sigma_X=100, \sigma_Y=5)$.
+	\caption{Forest plots of estimates of $\rr$ for the \xxsir methods for the PopSize1-4 data sets $(\beta=.06, \gamma=.03, O=365, \sigma_X=100, \sigma_Y=5)$.
 The points are the estimates of $\rr$ and the lines denote $\pm 2\cdot $Std.
@@
 	\includegraphics[scale=0.5]{images/popsize_arm.jpeg}
-	\caption{Forest plots of estimates of $\rr$ for the \xxsir methods for the PopSize1-4 data sets $(\beta=.06, \gamma=.03, T=365, \sigma_X=100, \sigma_Y=5)$.
+	\caption{Forest plots of estimates of $\rr$ for the \xxsir methods for the PopSize1-4 data sets $(\beta=.06, \gamma=.03, O=365, \sigma_X=100, \sigma_Y=5)$.
 The points are the estimates of $\rr$ and the lines denote $\pm 2\cdot $Std.
@@
   \includegraphics[scale=0.5]{images/var_n.jpeg}
-  \caption{Forest plots of estimates of $\rr$ for the \xxsir methods for the Var1-4 data sets $(\beta=.06, \gamma=.03, T=365, X(0)=99950, Y(0)=50, Z(0)=0, N=10^5)$.
+  \caption{Forest plots of estimates of $\rr$ for the \xxsir methods for the Var1-4 data sets $(\beta=.06, \gamma=.03, O=365, T=365, X(0)=99950, Y(0)=50, Z(0)=0, N=10^5)$.
 The points are the estimates of $\rr$ and the lines denote $\pm 2\cdot $Std.
@@
 		\includegraphics[scale=0.5]{images/var_nm.jpeg}
-		\caption{Forest plots of estimates of $\rr$ for the \xxsir methods for the Var1-4 data sets $(\beta=.06, \gamma=.03, T=365, X(0)=99950, Y(0)=50, Z(0)=0, N=10^5)$.
+		\caption{Forest plots of estimates of $\rr$ for the \xxsir methods for the Var1-4 data sets $(\beta=.06, \gamma=.03, O=365, T=365, X(0)=99950, Y(0)=50, Z(0)=0, N=10^5)$.
 The points are the estimates of $\rr$ and the lines denote $\pm 2\cdot $Std.
@@
 		\includegraphics[scale=0.5]{images/var_ar.jpeg}
-		\caption{Forest plots of estimates of $\rr$ for the \xxsir methods for the Var1-4 data sets $(\beta=.06, \gamma=.03, T=365, X(0)=99950, Y(0)=50, Z(0)=0, N=10^5)$.
+		\caption{Forest plots of estimates of $\rr$ for the \xxsir methods for the Var1-4 data sets $(\beta=.06, \gamma=.03, O=365, T=365, X(0)=99950, Y(0)=50, Z(0)=0, N=10^5)$.
 The points are the estimates of $\rr$ and the lines denote $\pm 2\cdot $Std.
@@
 		\includegraphics[scale=0.5]{images/var_arm.jpeg}
-		\caption{Forest plots of estimates of $\rr$ for the \xxsir methods for the Var1-4 data sets $(\beta=.06, \gamma=.03, T=365, X(0)=99950, Y(0)=50, Z(0)=0, N=10^5)$.
+		\caption{Forest plots of estimates of $\rr$ for the \xxsir methods for the Var1-4 data sets $(\beta=.06, \gamma=.03, O=365, T=365, X(0)=99950, Y(0)=50, Z(0)=0, N=10^5)$.
 The points are the estimates of $\rr$ and the lines denote $\pm 2\cdot $Std.
@@
     \includegraphics[scale=0.5]{images/other_n.jpeg}
-  \caption{Forest plots of estimates of $\rr$ for the \xxsir methods for the Poly1, Poly4, and LSIR data sets $(\beta=.06, \gamma=.03, T=365, X(0)=99950, Y(0)=50, Z(0)=0, N=10^5, \sigma_X=100, \sigma_Y=5)$.
+  \caption{Forest plots of estimates of $\rr$ for the \xxsir methods for the Poly1, Poly4, and LSIR data sets $(\beta=.06, \gamma=.03, O=365, T=365, X(0)=99950, Y(0)=50, Z(0)=0, N=10^5, \sigma_X=100, \sigma_Y=5)$.
 The points are the estimates of $\rr$ and the lines denote $\pm 2\cdot $Std.
@@
 		\includegraphics[scale=0.5]{images/other_nm.jpeg}
-		\caption{Forest plots of estimates of $\rr$ for the \xxsir methods for the Poly1, Poly4, and LSIR data sets $(\beta=.06, \gamma=.03, T=365, X(0)=99950, Y(0)=50, Z(0)=0, N=10^5, \sigma_X=100, \sigma_Y=5)$.
+		\caption{Forest plots of estimates of $\rr$ for the \xxsir methods for the Poly1, Poly4, and LSIR data sets $(\beta=.06, \gamma=.03, O=365, T=365, X(0)=99950, Y(0)=50, Z(0)=0, N=10^5, \sigma_X=100, \sigma_Y=5)$.
 The points are the estimates of $\rr$ and the lines denote $\pm 2\cdot $Std.
@@
 		\includegraphics[scale=0.5]{images/other_ar.jpeg}
-		\caption{Forest plots of estimates of $\rr$ for the \xxsir methods for the Poly1, Poly4, and LSIR data sets $(\beta=.06, \gamma=.03, T=365, X(0)=99950, Y(0)=50, Z(0)=0, N=10^5, \sigma_X=100, \sigma_Y=5)$.
+		\caption{Forest plots of estimates of $\rr$ for the \xxsir methods for the Poly1, Poly4, and LSIR data sets $(\beta=.06, \gamma=.03, O=365, T=365, X(0)=99950, Y(0)=50, Z(0)=0, N=10^5, \sigma_X=100, \sigma_Y=5)$.
 The points are the estimates of $\rr$ and the lines denote $\pm 2\cdot $Std.
@@
 		\includegraphics[scale=0.5]{images/other_arm.jpeg}
-		\caption{Forest plots of estimates of $\rr$ for the \xxsir methods for the Poly1, Poly4, and LSIR data sets $(\beta=.06, \gamma=.03, T=365, X(0)=99950, Y(0)=50, Z(0)=0, N=10^5, \sigma_X=100, \sigma_Y=5)$.
+		\caption{Forest plots of estimates of $\rr$ for the \xxsir methods for the Poly1, Poly4, and LSIR data sets $(\beta=.06, \gamma=.03, O=365, T=365, X(0)=99950, Y(0)=50, Z(0)=0, N=10^5, \sigma_X=100, \sigma_Y=5)$.
 The points are the estimates of $\rr$ and the lines denote $\pm 2\cdot $Std.
@@
       The results are shown below.  \cite{towers2009} report an estimate of $\hat{\rr} \in [1.50, 1.70]$.
+
+\begin{figure}[H]
+	\centering
+	\includegraphics[width=\textwidth]{images/h1n1.jpeg}
+	\caption{Results from SIR models on H1N1 data.}
+\end{figure}
+
+\begin{table}[H]
+
+	\centering
+	\begin{tabular}[t]{l|r|r}
+		\hline
+		Model & Estimate & Std. Dev\\
+		\hline
+		LS & 0.0280 & 0.0549\\
+		\hline
+		ReLS & 0.3332 & 0.0001\\
+		\hline
+		LMA & 0.9900 & 3.2589\\
+		\hline
+		LMAT & 0.9928 & 0.0276\\
+		\hline
+		IPR & 5.3078 & 1.7141\\
+		\hline
+		SIPR & -17.9646 & 5.1240\\
+		\hline
+		LL & 0.0036 & 0.0000\\
+		\hline
+		MC & 0.9453 & 0.0011\\
+		\hline
+		SB & 39.9036 & 1.7520\\
+		\hline
+	\end{tabular}
+	\caption{H1N1 Data, $\rr$ Estimates and Standard Errors}
+\end{table}
 

### 2018-07-03 paper/Version15/draft_v15.tex
finally on git

--- 2018-07-03 paper/Version14/draft_v14.tex
@@
 
-% \tableofcontents
+\tableofcontents
 
@@
 
-\section{Origins and Difficulties of Estimating $\rr$}
+\subsection{Origins and Difficulties of Estimating $\rr$}
 \label{sec:r0}
@@
 
-\section{Kermack and McKendrick's SIR Model}
+\subsection{Kermack and McKendrick's SIR Model}
 \label{sec:sir-intro}
@@
   \end{align}
-We are assuming the observations are generated based on the discretizied time, deterministic ODEs presented in Equation \ref{eq:sir} with the addition of time ($t=0, 1, \dots, T$) and compartment dependent noise $\epsilon_{X,t}$ and $\epsilon_{Z,t}$.
+We are assuming the observations are generated based on the discrete-time,  deterministic ODEs presented in Equation \ref{eq:sir} with time and compartment dependent noise $\epsilon_{X,t}$ and $\epsilon_{Z,t}$.
 Since, $N$, the total population, is constant, then $Y_{obs}$ is adjusted accordingly.
@@
 \end{align*}
-where `$\overset{P}{\to}$' stands for convergence in probability \citep{wasserman2004} or (2) converging to a known probability distribution (convergence in distribution `$\overset{d}{\to}$' \citep{wasserman2004}),
+where `$\overset{P}{\to}$' stands for convergence in probability or (2) converging to a known probability distribution (convergence in distribution `$\overset{d}{\to}$' \citep{wasserman2004}),
 \begin{align*}
@@
   \frac{dX}{dt} &= - \frac{\beta XY}{N} \\
-  \frac{dE}{dt} &= \frac{\beta XY}{N}  - \alpha E\\
-  \frac{dY}{dt} &= \alpha E - \gamma Y \\
+  \frac{dE}{dt} &= \frac{\beta XY}{N}  - \mu E\\
+  \frac{dY}{dt} &= \mu E - \gamma Y \\
   \frac{dZ}{dt} &= \gamma Y,
 \end{align*}
-again, with a constant population of $N$ and with known initival values $(X(0), E(0), Y(0), Z(t))$.
-The parameter $\alpha$ is the rate at which latent exposed individuals become infectious.
-As $\alpha \to \infty$, the XEYZ model converges to the XYZ model.
+again, with a constant population of $N$ and with known initial values $(X(0), E(0), Y(0), Z(t))$.
+The parameter $\mu$ is the rate at which latently exposed individuals become infectious.
+As $\frac{1}{\mu} \to 0$, the XEYZ model converges to the XYZ model.
 The reproduction number for the XEYZ model, is the same as in the XYZ model, that is $\rr = \frac{\beta}{\gamma}$.
@@
 Consider the case when $N=10000, (\beta, \gamma) = (.06, .03)$, $(X(0), E(0), Y(0), Z(t)) = (9500, 0, 500, 0)$ for times $t=1, \dots, 365$.
-Let the parameter $\alpha$ range between [.005, 1].
+Let the parameter $\mu$ range between [0.005, 0.250].
 This generates data set
-$$D_{\alpha, \textnormal{XEYZ}} = \left \{(X_{XEYZ}(t), E_{XEYZ}(t), Y_{XEYZ}(t), Z_{XEYZ}(t): t=1, \dots, 365\right \},$$
+$$D_{\mu, \textnormal{XEYZ}} = \left \{(X_{XEYZ}(t), E_{XEYZ}(t), Y_{XEYZ}(t), Z_{XEYZ}(t): t=1, \dots, 365\right \},$$
 which is observed by researcher 1.
-Now assume that researcher 2 misspecifies the model for this data as a XYZ model rather than a XEYZ model.
-That is, she places all exposed individuals in the X compartment to have data set
-$$D_{\alpha, XYZ} = \left \{X_{XYZ}(t) = X_{XEYZ} + E_{XEYZ}, Y_{XYZ} = Y_{XEYZ}, Z_{XEYZ} = Z_{XYZ} : t = 1, \dots, 365 \right \}.$$
-This situation is illustrated in Figure \ref{fig:sir-vs-seir} with $\alpha = 0.01$.
+Now assume that researcher 2 misspecifies the model for this data as a SIR (XYZ) model rather than a SEIR (XEYZ) model.
+That is, she observes all non-infectious individuals as susceptibles and generates data set
+$$D_{\mu, XYZ} = \left \{X_{XYZ}(t) = X_{XEYZ} + E_{XEYZ}, Y_{XYZ} = Y_{XEYZ}, Z_{XEYZ} = Z_{XYZ} : t = 1, \dots, 365 \right \}.$$
+This situation is illustrated in Figure \ref{fig:sir-vs-seir} with $\mu = 0.01$.
 
@@
   \centering
-  \includegraphics[width=\textwidth]{images/seir-sir-data.pdf}
-  \caption{Top: XEYZ generated data over the course of a year.  Bottom: data tranformed from the XEYZ model to the XYZ model by combining all X and E compartments.}\label{fig:sir-vs-seir}
+  \includegraphics[width=.9\textwidth]{images/seir-sir-data.pdf}
+  \caption{Top: XEYZ generated data over the course of a year.
+Bottom: data tranformed from the XEYZ model to the XYZ model by combining all X and E compartments from the XEYZ model for the X compartment in the XYZ model.}\label{fig:sir-vs-seir}
   \end{figure}
@@
 Both researchers want to estimate $\rr$ and so fit models to their data sets.
-The details of how the model are not so important at the moment (but will be in the later sections), but to fairly compare assume both researchers use least squares to fit the models.
+The details of how the models are fit are not so important at the moment (but will be in the later sections), but to fairly compare one model to another assume both researchers use ratio estimator to fit the models.
 That is researcher 1 estimates $\rr$ with
 \begin{align*}
-  (\hat{\beta}_{XEYZ}, \hat{\gamma}_{XEYZ}, \hat{\alpha}_{XEYZ} )&=\arg \min_{\beta, \gamma, \alpha} \sum_{t} \left [ \left (X_{XEYZ}(t) - X(t)\right )^2  \right . \\
+  (\hat{\beta}_{XEYZ}, \hat{\gamma}_{XEYZ}, \hat{\mu}_{XEYZ} )&=\arg \min_{\beta, \gamma, \mu} \sum_{t} \left [ \left (X_{XEYZ}(t) - X(t)\right )^2  \right . \\
                                                                  &+\left ( E_{XEYZ}(t) - E(t) \right )^2  \\
@@
 \end{align*}
-Then the least squares (LS) estimate for $\rr^{XYZ}$ is given by Equation \ref{eq:sirls},
+Then the ratio estimator (RE) for $\rr^{XYZ}$ is given by Equation \ref{eq:sirls},
 \begin{align*}
@@
 \end{align*}
-Then the least squares (LS) estimate for $\rr^{XYZ}$ is given by Equation \ref{eq:sirls},
+Then the RE for $\rr^{XYZ}$ is given by Equation \ref{eq:sirls},
 \begin{align*}
@@
 
-We plot the two estimates of $\rr$ as a function of $\alpha$ in Figure \ref{fig:r0-alpha}.
+We plot the two estimates of $\rr$ as a function of $\mu$ in Figure \ref{fig:r0-alpha}.
 We should note we also estimated the standard errors of the estimates of $\rr$, but they do not show up on the figure because they are of an order of magnitude of $10^{-4}$.
 Unsurprisingly, under the SEIR model speficiation, $\hat{\rr}=2$ because there is no noise in the data, and the data is generated from an SEIR model.
-However, the SIR model for small values of $\alpha$ can result in double the size of $\rr$!
-
-For those
+However, as we see in Figure \ref{fig:r0-alpha} the RE of $\rr$ from the SIR model for small values of $\mu$ can result in double the size of $\rr$!
 
@@
   \includegraphics[width=\textwidth]{images/r0-est.pdf}
-  \caption{Estimates of $\rr$ from both the XEYZ and XYZ model specification for different values of $\alpha$.}\label{fig:r0-alpha}
-  \end{figure}
+  \caption{Estimates of $\rr$ from both the XEYZ and XYZ model specification for different values of $\mu$.}\label{fig:r0-alpha}
+  \end{figure}
+
+For those readers not convinced by our small example, we offer this mathematical intuition of the next generation model (NGM) described by \cite{diekmann1990} and \cite{diekmann2009}.
+
+ The NGM is a generalization of any compartment model at the infection-free steady state ($Y(t)=0$ and all derivatives with respect to time sum equal zero).
+Originally, introduced by \citep{diekmann1990}, \cite{diekmann2009} posit a recipe to find $\rr$ for a wide class of compartment models, including SIR and SEIR.
+First, define the infected subsystem as ``the equations of the ODE system that describe the production of new infections and changes in state among infected individuals.''  Let $x = (C_1, C_2, \dots, C_m)^T$ where $C_i$ are the different compartments of the infected subsystem.
+The steps to find $\rr$ are as follows:
+
+\begin{enumerate}
+\item ``Linearize the infected subsystem of nonlinear ODEs about the infection-free steady state''
+\item Decompose the linearized infected subsystem into $(T + \Sigma )x$ where $T$ is a $m\times m$ matrix of transmissions and $\Sigma$ is a $m \times m$ matrix of transitions.
+\item $\rr$ is the spectral radius (i.e. the dominant eigenvalue) of $-T \Sigma^{-1}$.  
+\end{enumerate}
+
+Here, $T_{ij}$ is the rate of transmission of \textit{newly} infected individuals in state $i$ created by individuals in state $j$.
+$\Sigma_{ij}$ is the transition rate of individuals into compartment $i$ from compartment $j$.
+
+This method has the assumptions of a basic compartment model: homogeneous mixing among populations and the law of mass action, i.e., that the number of infected individuals are proportional to the number of infected individuals at the previous step.
+Additionally, this method is advantageous in that the matrix $-\Sigma^{-1}$ has intuitive meaning.
+The $ij$th entry of $- \Sigma^{-1}$ is the ``expected time that an individual who presently has state $j$ will spend in infected state $i$.''
+
+The NGM shows that $\rr$ is the spectral radius of a matrix composed from the transmission of disease from diseased individuals to susceptible individuals and transition of diseased individuals from among the compartments and is indeed a property of the model.
+
+Following the examples presented in \cite{diekmann1990}, for the SIR (XYZ) model, the linearized infection-free sub-system occurs when $Y(t)=0$ and $X(t)=N$ and is
+\begin{align*}
+\frac{dY}{dt} &= - \beta Y + \gamma .
+  \end{align*}
+  Then $T= \beta$ (the rate of new infections) and $\Sigma = -\gamma$.  Then $K = \frac{\beta}{\gamma}$ which is also, trivially, the dominant eigenvalue, and hence, $R_0 = \frac{\beta}{\gamma}$.
+
+We can also derive the NGM for the SEIR (XEYZ) model.  We have an infection-free steady state when $E=Y=0$ and set $X=N$.  Then $x = (E, Y)^T$ is the infected subsystem.  The linearized subsystem is
+  \begin{align*}
+  \left \{
+  \begin{array}{cl}
+    \frac{dE}{dt} &= \beta Y  - \mu E\\
+    \frac{dY}{dt} &=  \mu E  - \gamma Y \\
+  \end{array}
+  \right .
+  \end{align*}
+  Then $T = \left ( \begin{array}{cc} 0 & \beta \\ 0 & 0  \end{array} \right )$ and $\Sigma = \left ( \begin{array}{cc} -\mu & 0 \\ \mu & - \gamma \end{array} \right )$, and $K =\left ( \begin{array}{cc} 0& \beta / \gamma \\ 0 & 0 \end{array} \right  )$.
+
+  Then the spectral radius of $K$ is $R_0  = \frac{\beta}{\gamma}$, which makes sense as exposed individuals are not contributing to the number of new infections.
+
+  Although the $\rr$ ends up being the same in both the SIR (XYZ) and SEIR (XEYZ) case, the derivations were different because of the addition of the exposed compartment.
+In general, more complicated models will result in different derivations of $T$ and $\Sigma$ and hence different derivations of $\rr$.
 
@@
 
-\subsection{Least Squares ($\beta$, $\gamma$) (LS)}\label{least-squares-beta-gamma}
+\subsection{Ratio Estimator ($\beta$, $\gamma$) (RE)}\label{least-squares-beta-gamma}
 Our first approach to estimate $\rr$ in the SIR model is to minimize the joint mean square error for the data collected at each time point and use the plug-in estimator found in Equation \ref{eq:sirls}.
@@
 \end{align*}
-Then the least squares (LS) estimate for $\rr$ is given by Equation \ref{eq:sirls},
+Then the ratio estimator (RE) estimate for $\rr$ is given by Equation \ref{eq:sirls},
 \begin{align}\label{eq:sirls}
@@
 Another answer is that $L_2$ is a continuously differentiable function and hence is easier to compute, which allows for sensitivity analysis to  be conducted more easily.
-Finally, LS can be done without writing down explicit assumptions about the noise within the model.
+Finally, RE can be done without writing down explicit assumptions about the noise within the model.
 However, we cannot guarantee properties of consistency or convergence to a known distribution without more explicit assumptions.
 
-On the other hand, least squares is easy to implement using software such as \texttt{R}'s \texttt{optim()} function, and as we see in Section \ref{sec:results} produces results comparable methods to other methods.
-Least squares and similar variants have been used to estimate $\rr$ in \cite{majumder2016}.
-
-\subsection{Reparametrized Least Squares ($\rr$, $\gamma$) (ReLS)}\label{reparametrized-least-squares-rux5f0-gamma}
+On the other hand, ratio estimator is easy to implement using software such as \texttt{R}'s \texttt{optim()} function, and as we see in Section \ref{sec:results} produces results comparable methods to other methods.
+ratio estimator and similar variants have been used to estimate $\rr$ in \cite{majumder2016}.
+
+\subsection{Reparametrized Ratio Estimator ($\rr$, $\gamma$) (rRE)}\label{reparametrized-least-squares-rux5f0-gamma}
 
@@
 We use the $\hat{\rr}$ directly from the above estimation problem, which again can be found with a grid search or another optimization process.
-It may be surprising that this method leads to different results than simply using least squares.
+It may be surprising that this method leads to different results than simply using the ratio estimator.
 One reason we obtain different results may be attributed to better numerical precision as there is one fewer step to obtain an estimate for $\rr$.
-However, we also see that this method further results in different standard errors than in LS, which we discuss further in Section \ref{sec:results}.
-
-Like LS, while we have no explicit parametric assumptions on the noise, we also are uncertain of any theoretical properties of our resulting estimator.
-Likewise, the same difficulties in sensitivity analysis arise as in LS.
+However, we also see that this method further results in different standard errors than in RE, which we discuss further in Section \ref{sec:results}.
+
+Like RE, while we have no explicit parametric assumptions on the noise, we also are uncertain of any theoretical properties of our resulting estimator.
+Likewise, the same difficulties in sensitivity analysis arise as in RE.
 
@@
   Here, $K$ is arbitrary and should be selected using some criterion such as AIC.
-Besides optionally deciding on the degree of polynomials to fit, this model is simple to implement and gives comparable results to using least squares with the SIR model.
+Besides optionally deciding on the degree of polynomials to fit, this model is simple to implement and gives comparable results to using the ratio estimator with the SIR model.
 The time $t=0$ is used to best capture the initial outbreak.
@@
 We can regress the left hand side in Eq.
-\ref{eq:harko_lin} on $Z(t)$ (using least squares), the number of recovered individuals, with $\rr$ as the coefficient of $Z(t)/N$ and with an intercept term of $\beta_0=0$ to obtain an estimate of $\rr$,
+\ref{eq:harko_lin} on $Z(t)$ (using least squares to estimate the coefficients), the number of recovered individuals, with $\rr$ as the coefficient of $Z(t)/N$ and with an intercept term of $\beta_0=0$ to obtain an estimate of $\rr$,
 \begin{align*}
@@
 
-For the method LS, we know $\rr = h(\beta, \gamma) = \frac{\beta}{\gamma}$.
+For the method RE, we know $\rr = h(\beta, \gamma) = \frac{\beta}{\gamma}$.
 Then, if $E\left [\frac{\hat{\beta}}{\hat{\gamma}}\right] \to \frac{\beta}{\gamma}$, then  $(\nabla h = (\frac{1}{\gamma},  -\frac{\beta}{\gamma^2})^T)$ and
@@
 
-For ReLS, we use the estimate of $\Sigma_{\rr, \gamma}$ taken from the optmization of $\rr$ and $\gamma$ as to find our estimate for $\hat{\Sigma}_{\rr}$.
+For rRE, we use the estimate of $\Sigma_{\rr, \gamma}$ taken from the optmization of $\rr$ and $\gamma$ as to find our estimate for $\hat{\Sigma}_{\rr}$.
 
@@
 
-\section{Methods}\label{sec:sim-res}
+\section{Fitting the models to data}\label{sec:sim-res}
 
@@
 We then generate different SIR data sets, changing one of the conditions for simulation each time and using the baseline SIR data as our basis for comparison.
-The different data sets generated are shown in Table \ref{tab:simulated-data}, for a total of 27 simulated data sets.
+The different data sets for a given error structure are shown in Table \ref{tab:simulated-data}.
+We repeat  Table \ref{tab:simulated-data}, replacing the ``Error'' with  one of the other three error structures for a total of 96 simulated data sets.
 
@@
   Baseline        & SIR      &   Norm             & (.06, .03)                          &     365       &  (99950, 50)                    &  $10^5$          &(100, 5)                                \\ \midrule
-  AR       & SIR      &   AR ($\rho=.7)$             & (.06, .03)                          &     365       &  (99950, 50)                    &  $10^5$          &(100, 5)                            \\ \midrule
-  \textcolor{red}{Norm-M}      & SIR      &   Norm-M ($\rho=.7)$             & (.06, .03)                          &     365       &  (99950, 50)                    &  $10^5$          &(100, 5)                            \\ \midrule
-   \textcolor{red}{AR-M}       & SIR      &   AR-M ($\rho=.7)$             & (.06, .03)                          &     365       &  (99950, 50)                    &  $10^5$          &(100, 5)                            \\ \midrule
+%  AR       & SIR      &   AR ($\rho=.7)$             & (.06, .03)                          &     365       &  (99950, 50)                    &  $10^5$          &(100, 5)                            \\ \midrule
+%  \textcolor{red}{Norm-M}      & SIR      &   Norm-M ($\rho=.7)$             & (.06, .03)                          &     365       &  (99950, 50)                    &  $10^5$          &(100, 5)                            \\ \midrule
+%   \textcolor{red}{AR-M}       & SIR      &   AR-M ($\rho=.7)$             & (.06, .03)                          &     365       &  (99950, 50)                    &  $10^5$          &(100, 5)                            \\ \midrule
   InfRec1       & SIR      &   Norm             & (60000, .001)                          &     365       &  (99950, 50)                    &  $10^5$          &(100, 5)\\
@@
   Initial4    & SIR      &   Norm             & (.06, .03)                          &     365       &  (99000, 1000)                    &  $10^5$          &(100, 5)                                \\ \midrule
-  \textcolor{red}{PopSize1}        & SIR      &   Norm             & (.06, .03)                          &     365       &  (999500, 500)                    &  $10^6$          &(100, 5)                                \\
-  \textcolor{red}{PopSize2}       & SIR      &   Norm             & (.06, .03)                          &     365       &  (9995, 5)                    &  $10^4$          &(100, 5)                                \\
-  \textcolor{red}{PopSize3}        & SIR      &   Norm             & (.06, .03)                          &     365       &  (990, 10)                    &  $10^3$          &(100, 5)                                \\
-  \textcolor{red}{PopSize4}        & SIR      &   Norm             & (.06, .03)                          &     365       &  (99, 1)                    &  $10^2$          &(100, 5)                                \\
+PopSize1      & SIR      &   Norm             & (.06, .03)                          &     365       &  (999500, 500)                    &  $10^6$          &(100, 5)                                \\
+PopSize2      & SIR      &   Norm             & (.06, .03)                          &     365       &  (9995, 5)                    &  $10^4$          &(100, 5)                                \\
+PopSize3        & SIR      &   Norm             & (.06, .03)                          &     365       &  (990, 10)                    &  $10^3$          &(100, 5)                                \\
+PopSize4       & SIR      &   Norm             & (.06, .03)                          &     365       &  (99, 1)                    &  $10^2$          &(100, 5)                                \\
  \midrule
@@
   Var4      & SIR      &   Norm             & (.06, .03)                          &     365       &  (99950, 50)                    &  $10^5$          &(10000, 500)                                \\ \midrule
-  \textcolor{red}{Poly1}     & Linear      &   Norm             & (.06, .03)                          &     365       &  (99950, 50)                    &  $10^5$          &(100, 5)                            \\
-  \textcolor{red}{Poly2}    & Quartic      &   Norm             & (.06, .03)                          &     365       &  (99950, 50)                    &  $10^5$          &(100, 5)                            \\
+ Poly1     & Linear      &   Norm             & (.06, .03)                          &     365       &  (99950, 50)                    &  $10^5$          &(100, 5)                            \\
+Poly2    & Quartic      &   Norm             & (.06, .03)                          &     365       &  (99950, 50)                    &  $10^5$          &(100, 5)                            \\ 
    LSIR     & LSIR      &   Norm             & (.06, .03)                          &     365       &  (99950, 50)                    &  $10^5$          &(100, 5)                            \\ 
@@
 \end{tabular}
-\caption{Table of simulated data sets.  We simulate 27 total data sets with the given parameters.}
+\caption{Table of simulated data sets for the ``Norm'' error.
+We also repeat the tables with the other three error structures: AR, Norm-M, and AR-M.
+We simulate $4 \times 24=96$ total data sets with the given parameters.}
 \label{tab:simulated-data}
@@
 
-Table \ref{tab:baseline-res} shows the ways that standard errors are acquired for the different methods.
-The IPR and SIPR models require knowledge of the length of the average infectious period.
+Table \ref{tab:se-methods} shows the ways that standard errors are acquired for the different methods.
+The IPR, SIPR, and IDEA models require knowledge of the length of the average infectious period.
 For the results below, we assume the best case scenario that the true average infectious period is known for the data set.
+The exponential growth model requires knowledge of the probability distribution of the infectious period.
+In this case, we assume a uniform probability distribution with a mean equivalent to the true average infectious period.
 
@@
 		\textbf{Model} & \textbf{Default Standard Error Method} \\ \midrule
-		Least Squares (LS) & Delta method with Hessian matrix\\
-		Reparameterized Least Squares (ReLS) & Hessian matrix \\
-		Linear Model Approximation (LMA) & Delta method with linear regression standard error \\
-		LMA, All Time Points  (LMAT)& Delta method with linear regression standard error \\
-		Incidence-Prevalence Ratio (IPR) & Jackknife resampling \\
-		Smoothed IPR (SIPR) & Jackknife resampling \\
-		Log-Linear (LL) & Linear regression standard error \\
-		Markov Chain (MC) & Delta method with GLM regression standard error \\
-		Sequential Bayes (SB) & Standard deviation of posterior distribution\\
+		Ratio Estimator (RE) & Delta (with inverse Hessian)\\
+		Reparameterized Ratio Estimator (rRE) & Delta (with inverse Hessian) \\
+		Linear Model Approximation (LMA) & Delta (linear regression SE) \\
+		LMA, All Time Points  (LMAT)& Delta (linear regression SE)\\
+		Incidence-Prevalence Ratio (IPR) & Block bootstrap (Jackknife resampling) \\
+		Smoothed IPR (SIPR) & Block bootstrap (Jackknife resampling) \\
+		Log-Linear (LL) & Delta (linear regression SE) \\
+		Markov Chain (MC) & Delta (GLM regression SE) \\
+		Sequential Bayes (SB) & Posterior\\
 		% Exponential Growth (EG) & Delta method with linear regression standard error \\
@@
 	\end{tabular}
-	\caption{Table of methods for standard errors.}
+	\caption{Table of methods for estimating standard errors for each of our $\rr$ estimates.}
 	\label{tab:se-methods}
@@
 
-Table \ref{tab:baseline-res} shows the ways that standard errors are acquired for the different methods.
-The IPR, SIPR, and IDEA models require knowledge of the length of the average infectious period.
-For the results below, we assume the best case scenario that the true average infectious period is known for the data set.
-The exponential growth model requires knowledge of the probability distribution of the infectious period.
-In this case, we assume a uniform probability distribution with a mean equivalent to the true average infectious period.
-
-\section{Results of Simulations}\label{sec:results}
+\section{Results}\label{sec:results-top}
+\subsection{$\rr$ estimation methods}\label{sec:results}
 There are numerous aspects to consider when using a method to estimate $\rr$.
-Among these include values of both the infection and recovery rate $(\beta, \gamma)$, the number of data points available ($O$), the total size of the population ($N$), the initial percent of susceptible individuals $\left (\frac{X(0)}{N}\right)$, the initial percent of infectious individuals $\left (\frac{Y(0)}{N}\right )$, the magnitude of the variance of both the number of susceptibles and the number of infectious ($\sigma_X^2, \sigma_Y^2$), the error modeling assumptions, data generated from a non-SIR model, and finally methods of estimating the variance.
-We first examine our baseline data, and then we examine each of these aspects individually.
-
-\subsection{Baseline Simulated Data}\label{sec:res-base}
-We implement the \wxxsir methods to estimate $\rr$ using the baseline data set.  The results are displayed in Figure \ref{fig:baseline-res} and Table \ref{tab:baseline-res}.
+Among these include values of both the infection and recovery rate $(\beta, \gamma)$, the number of data points available ($T$), the total size of the population ($N$), the initial percent of susceptible individuals $\left (\frac{X(0)}{N}\right)$, the initial percent of infectious individuals $\left (\frac{Y(0)}{N}\right )$, the magnitude of the variance of both the number of susceptibles and the number of infectious ($\sigma_X^2, \sigma_Y^2$), the error modeling assumptions, data generated from a non-SIR model, and finally methods of estimating the variance.
+For the sake of length, we do not provide the full results here, and instead examine the results at a high-level.
+However, for those interested readers, we provide results in the form of forest plots and tables of estimates of $\rr$ for each data set in the Supplementary Material.
+
+We implement the \wxxsir methods to estimate $\rr$ using the baseline data set with normal error.  The results are displayed in Figure \ref{fig:baseline-res} and Table \ref{tab:baseline-res}.
 \begin{figure}[H]
@@
 		\hline
-		LS & 1.9999711 & 0.0056\\
-		\hline
-		ReLS & 1.9997340 & 0.0050\\
+		RE & 1.9999711 & 0.0056\\
+		\hline
+		rRE & 1.9997340 & 0.0050\\
 		\hline
@@
 	\end{tabular}
-        \caption{Baseline Data $(\beta=.06, \gamma=.03, O=365, X(0)=99950, Y(0)=50, Z(0)=0, \sigma_X=100, \sigma_Y=5, N=10^5)$, $\rr$ Estimates and Standard Errors}\label{tab:baseline-res}
+        \caption{Baseline Data $(\beta=.06, \gamma=.03, T=365, X(0)=99950, Y(0)=50, Z(0)=0, \sigma_X=100, \sigma_Y=5, N=10^5)$, $\rr$ Estimates and Standard Errors}\label{tab:baseline-res}
 \end{table}
@@
 Overall, in Figure \ref{fig:baseline-res}, we see a wide variation in the estimates of and the standard errors for $\rr$ depending on the method used.
-The methods of LS, ReLS, LMA, LMAT, IPR, and LL all result in a CI that covers the true value of $\rr=2$.
+The methods of RE, rRE, LMA, LMAT, IPR, and LL all result in a CI that covers the true value of $\rr=2$.
 Of these, LMAT and IPR result in a CI which contains 1, so we would not be able to say whether or not we would have an outbreak of a disease or not.
@@
 
-\subsection{Autoregressive Errors}\label{sec:res-AR}
-We next examine the estimates of $\rr$ when fit to the autoregressive simulated data with $\rho=0.7$.
-In this data, we may see errors accumulate over time, which may more accurately reflect the course of infection in an epidemic.
-For example, we may expect to see more variation around the peak of an epidemic than the beginning.
-Besides the AR error, the other parameters are the same as in the baseline model.
-The results of the estimates are displayed in Figure \ref{fig:ar-res} and Table \ref{tab:ar-res}.
-
-In this setting, the CIs from LS, ReLS, LMAT, SIPR, and LL cover the true $\rr=2$.
-In contrast to the baseline data, the CI for LMA no longer covers the true estimate.
-For all of these models, the standard error is larger, which we would expect given that the data have a large positive autocorrelation coefficient.
+The results of the baseline data set with normal error are a good summary of the methods as a whole.
+In Figure \ref{fig:baseline-res}, we see that the clear best methods to estimate $\rr$ are RE, rRE, and LL.
+These three methods consistently outperform the other methods of accurately estimating $\rr$ with a reasonably small CI when varying a single parameter of our generated data sets and holding the others as fixed.
+
+In particular, MC, SB almost never capture the true estimate and typically underestimate the true value of $\rr$.
+This leads us to believe that these models are misspecified.
+Both models rely on assumptions of simple underlying statistical distributions for the course of infection (binomial and Poisson, respectively).
+These are good examples of why we need to consider other reasons for using a model other than the math is nice.
+That is not to say that variants of these models are unviable for modelling $\rr$, but they do not make good defaults.
+It is also an issue that these two methods consistently underestimate $\rr$, which could possibly lead to an unchecked epidemic.
+
+IPR often produces reasonable point estimates for $\rr$ but simply has too large of a CI to be considered a useful estimate.
+On the other hand, SIPR has a small CI that typically underestimates the true value of $\rr$.
+Since SIPR smoothes the SIR curves, the peak of the infectious curve in particular is shrunken.
+Since the peak value of number of infectious is crucial in estimating $\rr$ for SIPR, it makes sense that SIPR underestimates this result.
+Additionally, LMA and LMAT are generally prone to underestimating $\rr$ and have large and generally uninformative CIs.
+
+Of the three well-performing methods, RE and rRE seem to be slightly more robust to parameter variation than LL.
+This is mostly due to the fact that LL produces such small CIs that even a small discrepancy in estimation of $\rr$ can cause the CI to not capture the true value of $\rr$.
+
+However, we do see instances when LL completely outperforms RE and rRE such as in the case where the initial percent of infectious is $\frac{I(0)}{N} \cdot 100\%$ $< .01$\%.
+In Figure \ref{fig:ar-small-I}, we examine the data set generated with autoregressive monotonic error, our most stringent data, with varying levels of initial percent of infectious individuals.
+When $\frac{I(0)}{N} \cdot 100\%$ $< .01$\%, RE and rRE tend to overestimate $\rr$ and have large CIs.
+However, LL accurately estimates $\rr$ even when the initial percentage of infectious is very small.
+This is also true for the other error sets (see the Supplementary Materials for more details).
+
+One interesting feature we see in the results is that RE and rRE can result in very different estimations and CIs, which may seem unusual since rRE is a simple reparametrization of RE.
+For example, in Figure \ref{fig:ar-r0}, we see that when the true value of $\rr$ is 1, rRE has a much smaller CI than RE, which may be one reason to prefer rRE to RE.
+This figure also helps to show that RE and rRE are robust to extreme values of $\beta$ and $\gamma$ and hence extreme values of $\rr$.
 
@@
 	\centering
-  \includegraphics[scale=0.5]{images/AR.jpeg}
-  \caption{Forest plots of estimates of $\rr$ for the \xxsir methods for the autoregressive (AR) ($\rho=.7)$ simulated data set $(\beta=.06, \gamma=.03, O=365, X(0)=99950, Y(0)=50, Z(0)=0, \sigma_X=100, \sigma_Y=5, N=10^5)$.
+	\includegraphics[scale=0.5]{images/parchange_ar.jpeg}
+	\caption{Forest plots of estimates of $\rr$ for the \xxsir methods for the infrec1-4 data sets $(T=365, X(0)=99950, Y(0)=50, Z(0)=0, \sigma_X=100, \sigma_Y=5, N=10^5)$.
 The points are the estimates of $\rr$ and the lines denote $\pm 2\cdot $Std.
 Err.
-The vertical line is the true $\rr$ value.}
-  \end{figure}\label{fig:ar-res}
-\begin{table}[H]
-
-	\centering
-	\begin{tabular}[t]{l|r|r}
-		\hline
-		Model & Estimate & Std. Err\\
-		\hline
-		LS & 1.9997 & 0.0072\\
-		\hline
-		ReLS & 2.0001 & 0.0063\\
-		\hline
-		LMA & 1.4745 & 0.1130\\
-		\hline
-		LMAT & 1.3694 & 0.8682\\
-		\hline
-		IPR & 1.7127 & 5.0713\\
-		\hline
-		SIPR & 0.8792 & $<$ 1e-04 \\
-		\hline
-		LL & 2.0000 & 0.0003\\
-		\hline
-		MC & 0.9486 & $<$ 1e-04 \\
-		\hline
-		SB & 1.0000 & 0.0523\\
-		\hline
-	\end{tabular}
-        \caption{Autoregressive Errors, $\rr$ Estimates and Standard Errors}\label{tab:ar-res}
-\end{table}
-
-\subsection{PAVA Monotonicity}\label{sec:res-PAVA}
-We examine the case of the PAVA error assumptions which enforces monotonic assumptions on the suceptible and recovered compartments.
-This in turn may more accurately model the course of an epidemic, but as a consequence, we may not be able to guarantee that a 95\% CI covers the true value of $\rr$ 95\% of the time.
-
-In Figure \ref{fig:pava-res} and Table \ref{tab:pava-res}, we display the results of fitting the methods on two data sets, those generated from the Norm-G and Norm-AR error assumptions.
-We first see that there is very little difference among the estimates of $\rr$ in both Norm-G and Norm-AR error assumptions.
-We see that LMA and IPR have smaller estimates of $\rr$ and SB has a larger estimate of $\rr$.
-The standard errors for Norm-AR error assumption are greater than or equal to than the standard error in the Norm-G error assumption with the exceptions of LMAT and SB.
-
-The methods that contain the true value of $\rr$ are LS, ReLS, LMAT, IPR for both error assumptions and LL with Norm-AR error assumption.
-The methods LS, ReLS, LMA, and LL would be able to reject the hypothesis that $\rr < 1$.
+The vertical line is the true $\rr$ value.}\label{fig:ar-r0}
+\end{figure}
+
+%% robust to different error assumptions
+Reassuringly, we find that $\hat{\rr}$ seems to be generally robust to the error structure.
 
@@
 	\centering
-	\includegraphics[scale=0.5]{images/mono.jpeg}
-	\caption{Forest plots of estimates of $\rr$ for the \xxsir methods for the Norm-M and AR-M data sets $(\beta=.06, \gamma=.03, O=365, X(0)=99950, Y(0)=50, Z(0)=0, \sigma_X=100, \sigma_Y=5, N=10^5)$.
+	\includegraphics[scale=0.5]{images/err_robust.jpeg}
+	\caption{}\label{fig:err-rob}
+\end{figure}
+
+%% model misspecification is fairly robust
+We also examine circumstances where the SIR is misspecified.  The Poly1 and LSIR data sets result is good estimates of $\hat{\rr}$ but Poly4 results in
+
+\begin{figure}[H]
+	\centering
+	\includegraphics[scale=0.5]{images/model_robust.jpeg}
+	\caption{}\label{fig:mod-rob}
+\end{figure}
+
+%% small N or small  T can lead to bad results. Also large sig X, sig Y, and small gamma
+
+We find that when $N$, the number of individuals decreases, then the CIs of our estimates generally increase.
+This is likely due to the central limit theorem (CLT).
+When the noise parameters $\sigma_X$ or $\sigma_Y$ increase then the CIs of our estimates also generally increase.
+For both small $N$ and large $\sigma_X$ or $\sigma_Y$, we find that the point estimates of $\rr$ seem to be unbiased, it is just the CIs that are large.
+
+We find the parameter $T$ to be particularly important.
+For one, it can play a role similar to that of $N$, directly giving the researcher more data for a larger $T$.
+Thus, larger $T$ may result in smaller CIs due to the CLT.
+However, $T$ also plays an even greater role in estimating $\rr$.
+
+\begin{figure}[H]
+	\centering
+	\includegraphics[scale=0.5]{images/timediff.jpeg}
+	\caption{}\label{fig:timediff}
+\end{figure}
+
+\begin{figure}[H]
+	\centering
+	\includegraphics[scale=0.5]{images/obs.jpeg}
+	\caption{}\label{fig:obs}
+\end{figure}
+
+\begin{figure}[H]
+	\centering
+	\includegraphics[scale=0.5]{images/start_arm.jpeg}
+	\caption{Forest plots of estimates of $\rr$ for the \xxsir methods for the Inits1-4 data sets $(\beta=.06, \gamma=.03, T=365,  Z(0)=0, \sigma_X=100, \sigma_Y=5, N=10^5)$.
 The points are the estimates of $\rr$ and the lines denote $\pm 2\cdot $Std.
 Err.
-The vertical lines are the true $\rr$ value.}
-	\label{fig:pava-res}
-\end{figure}
-
-\begin{table}[H]
-
-	\centering
-	\begin{tabular}[t]{l|r|r|r|r}
-		\hline
-		Model & Norm-M Est. & Norm-M SE & AR-M Est. & AR-E SE\\
-		\hline
-		LS & 2.0002 & 0.0044 & 1.9999 & 0.0074\\
-		\hline
-		ReLS & 2.0003 & 0.0040 & 1.9997 & 0.0065\\
-		\hline
-		LMA & 1.6794 & 0.1103 & 1.5350 & 0.1497\\
-		\hline
-		LMAT & 1.4422 & 0.9242 & 1.4491 & 0.7464\\
-		\hline
-		IPR & 0.7486 & 1.5047 & 0.7513 & 1.5606\\
-		\hline
-		SIPR & 0.9082 & $<$ 1e-04 & 0.6976 & $<$ 1e-04\\
-		\hline
-		LL & 2.0003 & 0.0001 & 1.9996 & 0.0002\\
-		\hline
-		MC & 0.9486 & $<$ 1e-04 & 0.9486 & $<$ 1e-04\\
-		\hline
-		SB & 1.0000 & 0.0523 & 1.7143505 & 0.0685\\
-		\hline
-	\end{tabular}
-\caption{Gaussian and Autoregressive Data with PAVA Applied, $\rr$ Estimates and Standard Errors}\label{tab:pava-res}
-\end{table}
-
-In comparison to the baseline data set, the estimates of $\rr$ from the methods fit on the Norm-G error assumption data set have smaller CIs.
-The estimates of $\rr$ for the Norm-G error assumption are very similar to those from the baseline assumption.
-LMA and LMAT have estimates closer to the true $\rr$ in the Norm-G error assumption data setting compared to the baseline data and SB has an estimate further away.
-
-Overall, there is not much difference between the baseline and monotone data sets, at least in the case with $\sigma_X=100$ and $\sigma_Y=5$.
-
-\subsection{Infection and Recovery Rate ($\beta, \gamma$)}\label{sec:res-beta-gamma}
-
-We next examine the simulated data sets InfRec1-4, three of which have different levels of $\gamma$ for a given level of $\beta = 0.06$ and the fourth is simulated with $\rr=0.25 < 1$, a case where an outbreak would not occur.
-The idea behind this is that since $\rr=\frac{\beta}{\gamma}$ and ratio estimators typically have large variance, we want to examine how a small value of $\gamma$, that is a small recovery rate, would effect the resulting estimates of $\rr$.
-Additionally, we would like to know when a disease will die out on its own accord, that we can be certain the estimate of $\rr$ is less than 1.
-
-\subsubsection{Gaussian Errors}
-
-\begin{figure}[H]
-  \centering
-  \includegraphics[scale=0.5]{images/parchange_n.jpeg}
-  \caption{Forest plots of estimates of $\rr$ for the \xxsir methods for the infrec1-4 data sets $(\beta=.06, \gamma=.03, O=365, X(0)=99950, Y(0)=50, Z(0)=0, \sigma_X=100, \sigma_Y=5, N=10^5)$.
-The points are the estimates of $\rr$ and the lines denote $\pm 2\cdot $Std.
-Err.
-The vertical line is the true $\rr$ value.}
-  \label{fig:infrec1-res}
-\end{figure}
-\begin{table}[H]
-
-	\centering
-	\begin{tabular}[t]{l|r|r}
-		\hline
-		Model & Estimate & Std. Err\\
-		\hline
-		SIRLS & 59.9765 & 0.0435\\
-		\hline
-		ReLS & 59.9952 & 0.4344\\
-		\hline
-		LMA &  -117.9695 & 101.1886 \\
-		\hline
-		LMAT & -2.8536 & 193.6526 \\
-		\hline
-		IPR & 54.1465 & 200.2595\\
-		\hline
-		SIPR & 3.9645 & $<$ 1e-04\\
-		\hline
-		LL & 46.3210 & 0.7797\\
-		\hline
-		MC & 0.9486 & $<$ 1e-04 \\
-		\hline
-		SB & 0.9767 & 0.0517\\
-		\hline
-	\end{tabular}
-        \caption{\label{tab:infrec1-res}$\gamma = 0.001$, $\rr$ Estimates and Standard Errors}
-\end{table}
-
-Looking in Figure \ref{fig:infrec1-res} and Table \ref{tab:infrec1-res} at the cases where $\beta = 0.06$ with $\gamma = 0.001, 0.04, 0.01, 0.24$, it appears that most of the models do not handle a small $\gamma$ well.
-We examine an exaggerated extreme case of when $\rr=60$ (the largest reported $\rr$ value we found is $\hat{\rr}=14$ for measles as reported in \cite{anderson1992}).
-Only LS and ReLS result in CIs that cover the true $\rr$.
-IPR and LL give fairly close point estimates to the true $\rr=60$ as well.
-However, the other models do very poorly; they tend to greatly underestimate $\rr$.
-The linear model approximation does particularly poorly, likely because the linear model does not fit the data well, and is not even constrained to positive values of $X$, $Y$, and $Z$, which is why we can result in a negative estimate of $\rr$.
-Several of the models also return extremely large standard errors, meaning they are not very reliable in these cases.
-Only LS, ReLS, and LL would reject the hypothesis that $\rr <1$, when this is clearly not the case at all.
-
-When $\rr=1.5$ we see that the estimates from the methods give results much closer to the true value, and the resulting CIs are also smaller.
-The methods LS, ReLS, SIPR, LL, and SB would all reject the hypothesis that $\rr > 1$.
-Although if we are testing the null hypothesis for  all methods simultaneously, we should use a multiple-testing correction such as Bonferroni.
-Under the Bonferroni correction in this case, the only method which may then fail to reject would be SB as the other SEs are very small).
-
-When $\rr=1$, we are unsure whether there will be an epidemic or not.  In this case, all the CIs of the methods cover the true value.  We also see that the SE of LS is much greater than ReLS.
-
-\begin{table}[H]
-
-	\centering
-	\begin{tabular}[t]{l|r|r}
-		\hline
-		Model & Estimate & Std. Err\\
-		\hline
-		LS & 0.2363 & 4.1043 \\
-		\hline
-		ReLS & 4.1795 & 2.7456\\
-		\hline
-		LMA &  0.7075 & 0.2029 \\
-		\hline
-		LMAT & 0.6747 & 6.2471 \\
-		\hline
-		IPR & 181.0158 & 742.1088 \\
-		\hline
-		SIPR & -0.2484 & 0.1160 \\
-		\hline
-		LL & 0.9912 & 0.0031\\
-		\hline
-		MC & 0.9486 & 0.0001\\
-		\hline
-		SB & 1.2355 & 0.0582\\
-		\hline
-	\end{tabular}
-        \caption{$\gamma = 0.24$, $\rr$ Estimates and Standard Errors}\label{tab:infrec4-res}
-\end{table}
-
-Table \ref{tab:infrec4-res} corresponds to the case where $\gamma = 0.24$ and $\beta=0.06$.
-It appears that a lot of estimation methods do not handle small $\rr$ well either.
-None of them are extremely close to truth, except for ReLS, as most of the tend to overestimate the truth.
-The IPR is particularly bad in this case.
-Also, in general, the standard errors are several orders of magnitude larger than in the baseline data.
-
-\subsubsection{Gaussian Monotonic Errors}
-
-\begin{figure}[H]
-	\centering
-	\includegraphics[scale=0.5]{images/parchange_nm.jpeg}
-	\caption{Forest plots of estimates of $\rr$ for the \xxsir methods for the infrec1-4 data sets $(\beta=.06, \gamma=.03, O=365, X(0)=99950, Y(0)=50, Z(0)=0, \sigma_X=100, \sigma_Y=5, N=10^5)$.
-The points are the estimates of $\rr$ and the lines denote $\pm 2\cdot $Std.
-Err.
-The vertical line is the true $\rr$ value.}
-\end{figure}
-\begin{table}[H]
-
-	\centering
-	\begin{tabular}[t]{l|r|r}
-		\hline
-		Model & Estimate & Std. Dev\\
-		\hline
-		LS & 59.9765 & 0.0387\\
-		\hline
-		ReLS & 59.9952 & 0.3875\\
-		\hline
-		LMA & -55.3749 & 10.9872\\
-		\hline
-		LMAT & 7.7686 & 103.2206\\
-		\hline
-		IPR & 24.4087 & 66.1668\\
-		\hline
-		SIPR & 3.8285 & $<$ 1e-04\\
-		\hline
-		LL & 59.8300 & 0.1079\\
-		\hline
-		MC & 0.9486 & $<$ 1e-04\\
-		\hline
-		SB & 1.0000 & 0.0523\\
-		\hline
-	\end{tabular}
-	\caption{$\gamma = 0.001$, $\rr$ Estimates and Standard Errors}
-\end{table}
-
-\begin{table}[H]
-
-	\centering
-	\begin{tabular}[t]{l|r|r}
-		\hline
-		Model & Estimate & Std. Dev\\
-		\hline
-		LS & 0.2328 & 0.6701\\
-		\hline
-		ReLS & 3.5710 & 0.3171\\
-		\hline
-		LMA & 0.5217 & 0.1716\\
-		\hline
-		LMAT & 0.3872 & 4.8818\\
-		\hline
-		IPR & 6.6728 & 89.9362\\
-		\hline
-		SIPR & -0.0317 & 0.0336\\
-		\hline
-		LL & 0.8623 & 0.0088\\
-		\hline
-		MC & 0.9486 & $<$ 1e-04\\
-		\hline
-		SB & 1.5587 & 0.0653\\
-		\hline
-	\end{tabular}
-	\caption{$\gamma = 0.24$, $\rr$ Estimates and Standard Errors}
-\end{table}
-
-\subsubsection{Autoregressive Errors}
-
-\begin{figure}[H]
-	\centering
-	\includegraphics[scale=0.5]{images/parchange_ar.jpeg}
-	\caption{Forest plots of estimates of $\rr$ for the \xxsir methods for the infrec1-4 data sets $(\beta=.06, \gamma=.03, O=365, X(0)=99950, Y(0)=50, Z(0)=0, \sigma_X=100, \sigma_Y=5, N=10^5)$.
-The points are the estimates of $\rr$ and the lines denote $\pm 2\cdot $Std.
-Err.
-The vertical line is the true $\rr$ value.}
-\end{figure}
-\begin{table}[H]
-
-	\centering
-	\begin{tabular}[t]{l|r|r}
-		\hline
-		Model & Estimate & Std. Dev\\
-		\hline
-		LS & 60.0151 & 0.0698\\
-		\hline
-		ReLS & 60.0255 & 0.6912\\
-		\hline
-		LMA & -19.2058 & 5.9751\\
-		\hline
-		LMAT & 76.3404 & 472.4155\\
-		\hline
-		IPR & 58.7237 & 219.9891\\
-		\hline
-		SIPR & 3.9056 & $<$ 1e-04\\
-		\hline
-		LL & 41.8977 & 0.9768\\
-		\hline
-		MC & 0.9486 & $<$ 1e-04\\
-		\hline
-		SB & 1.0000 & 0.0523\\
-		\hline
-	\end{tabular}
-	\caption{$\gamma = 0.001$, $\rr$ Estimates and Standard Errors}
-\end{table}
-
-\begin{table}[H]
-
-	\centering
-	\begin{tabular}[t]{l|r|r}
-		\hline
-		Model & Estimate & Std. Dev\\
-		\hline
-		LS & -0.0695 & 2.8291\\
-		\hline
-		ReLS & 5.1930 & 3.5250\\
-		\hline
-		LMA & 0.8750 & 0.0591\\
-		\hline
-		LMAT & 1.7865 & 15.9420\\
-		\hline
-		IPR & 52.5000 & 740.0182\\
-		\hline
-		SIPR & -2.8643 & $<$ 1e-04\\
-		\hline
-		LL & 1.0014 & 0.0025\\
-		\hline
-		MC & 0.9486 & $<$ 1e-04\\
-		\hline
-		SB & 1.0160 & 0.0528\\
-		\hline
-	\end{tabular}
-	\caption{$\gamma = 0.24$, $\rr$ Estimates and Standard Errors}
-\end{table}
-
-\subsubsection{Autoregressive Monotonic Errors}
-
-\begin{figure}[H]
-	\centering
-	\includegraphics[scale=0.5]{images/parchange_arm.jpeg}
-	\caption{Forest plots of estimates of $\rr$ for the \xxsir methods for the infrec1-4 data sets $(\beta=.06, \gamma=.03, O=365, X(0)=99950, Y(0)=50, Z(0)=0, \sigma_X=100, \sigma_Y=5, N=10^5)$.
-The points are the estimates of $\rr$ and the lines denote $\pm 2\cdot $Std.
-Err.
-The vertical line is the true $\rr$ value.}
-\end{figure}
-\begin{table}[H]
-
-	\centering
-	\begin{tabular}[t]{l|r|r}
-		\hline
-		Model & Estimate & Std. Dev\\
-		\hline
-		LS & 59.9589 & 0.0631\\
-		\hline
-		ReLS & 59.9961 & 0.6215\\
-		\hline
-		LMA & -183.9423 & 277.8292\\
-		\hline
-		LMAT & -69.9068 & 1603.7792\\
-		\hline
-		IPR & 28.5231 & 120.9288\\
-		\hline
-		SIPR & 3.7113 & $<$ 1e-04\\
-		\hline
-		LL & 70.7076 & 0.9517\\
-		\hline
-		MC & 0.9486 & $<$ 1e-04\\
-		\hline
-		SB & 1.0000 & 0.0523\\
-		\hline
-	\end{tabular}
-	\caption{$\gamma = 0.001$, $\rr$ Estimates and Standard Errors}
-\end{table}
-
-\begin{table}[H]
-
-	\centering
-	\begin{tabular}[t]{l|r|r}
-		\hline
-		Model & Estimate & Std. Dev\\
-		\hline
-		LS & 0.3021 & 0.5657\\
-		\hline
-		ReLS & 9.8415 & 1.7074\\
-		\hline
-		LMA & -49.7813 & 5303.0302\\
-		\hline
-		LMAT & 0.2221 & 5.4112\\
-		\hline
-		IPR & 0.8145 & 11.8709\\
-		\hline
-		SIPR & 0.7365 & 0.0675\\
-		\hline
-		LL & 0.9199 & 0.0053\\
-		\hline
-		MC & 0.9486 & $<$ 1e-04\\
-		\hline
-		SB & 1.0000 & 0.0523\\
-		\hline
-	\end{tabular}
-	\caption{$\gamma = 0.24$, $\rr$ Estimates and Standard Errors}
-\end{table}
-
-\subsection{Number of Time Points ($O$)}\label{sec:res-time}
-We examine how the number of time steps the estimates of $\rr$ in the different methods.
-Ideally, the more data we have, the more likely we are to have lower-variance estimates for $\rr$.
-Since we would like to  infer how a disease will spread through a population, the sooner we can be sure whether $\rr> 1$, the sooner we can implement interventions and control measures for a disease.
-Another reason to analyze the number of time steps is that some of the methods assume linear or exponential growth of an epidemic in a certain time period.
-If the growth is no longer exponential, we will obtain a poor estimate of $\rr$.
-In this case, we can be using too many time points above a certain time.
-
-In Figure \ref{fig:time-res} and Tables \ref{tab:time-res1} and \ref{tab:time-res2} we examine the methods fit to different amounts of time periods where we observe the progress of a disease, including $T=200$, 100, 50, and 20.
-
-Again, we see that IPR does very poorly in estimating $\rr$, regardless of the amount of data observed.
-
-\subsubsection{Gaussian Errors}
-
-\begin{figure}[H]
-  \centering
-  \includegraphics[scale=0.5]{images/time_n.jpeg}
-  \caption{Forest plots of estimates of $\rr$ for the \xxsir methods for the Time1-4 data sets with Gaussian errors $(\beta=.06, \gamma=.03, O=365, X(0)=99950, Y(0)=50, Z(0)=0, \sigma_X=100, \sigma_Y=5, N=10^5)$.
-The points are the estimates of $\rr$ and the lines denote $\pm 2\cdot $Std.
-Err.
-The vertical line is the true $\rr$ value.}\label{fig:time-res}
-  \end{figure}
-
-\begin{table}[H]
-
-	\centering
-	\begin{tabular}[t]{l|r|r}
-		\hline
-		Model & Estimate & Std. Err\\
-		\hline
-		LS & 1.9993 & 0.0286\\
-		\hline
-		ReLS & 1.9994 & 0.0313\\
-		\hline
-		LMA & 0.9496 & 0.0825\\
-		\hline
-		LMAT & 3.3443 & 21.4895\\
-		\hline
-		IPR & 7.3458 & 16.1289\\
-		\hline
-		SIPR & -2.1199 & $<$ 1e-04\\
-		\hline
-		LL & 2.0014 & 0.0018\\
-		\hline
-		MC & 0.9486 & $<$ 1e-04\\
-		\hline
-		SB & 2.4289 & 0.1102\\
-		\hline
-	\end{tabular}
-        \caption{ $T$ = 100, $\rr$ Estimates and Standard Errors}\label{tab:time-res1}
-\end{table}
-
-Table \ref{tab:time-res1} shows the methods applied to the first 100 time steps, out of 365, of the baseline data set .
-The estimates and standard errors are worse than baseline, which we would expect given less data.
-The results are still relatively close to the full baseline data.
-There also appears to be less tendency to underestimate $\rr$ compared to a longer time scale.
-In general, we see that that the length of CIs decreases as we use more and more time points.
-Only LL would reject the hypothesis that $\rr < 1$ for 50 or fewer time points, but its CI does not cover the true $\rr$ value.
-With 50 and greater time points, we see that LS and ReLS provide close point estimates to the true value and by 100 time points, would reject the hypothesis that $\rr < 1$.
-The methods and MC and SB consistently have small CIs, but never cover the true value.
-
-\begin{table}[H]
-  \centering
-	\begin{tabular}[t]{l|r|r}
-		\hline
-		Model & Estimate & Std. Err\\
-		\hline
-		LS & 38.6896 & 106.2705\\
-		\hline
-		ReLS & 22.9646 & 6.9817\\
-		\hline
-		LMA & 1.4418 & 7.3164\\
-		\hline
-		LMAT & 1.0079 & 0.1425\\
-		\hline
-		IPR & 25.7313 & 35.2926\\
-		\hline
-		SIPR & 4.1772 & 7.3203\\
-		\hline
-		LL & 1.0430 & 0.0276\\
-		\hline
-		MC & 0.9486 & 5.0e-07\\
-		\hline
-		SB & 1.0000 & 0.2236\\
-		\hline
-	\end{tabular}
-        \caption{$T$ = 20, $\rr$ Estimates and Standard Errors}\label{tab:time-res2}
-\end{table}
-
-Table \ref{tab:time-res2} shows the methods applied to the first 20 time steps of a baseline data set.
-As might be expected, the $\rr$ estimates are wildly inaccurate, and none are very close to the true value.
-The methods of LS and ReLS are particularly worse in this situation compared to the other data sets.
-The standard errors are much larger than the baseline, which is to be expected as well due to the small number of data points.
-This shows that it is extremely difficult to create accurate online estimates of $\rr$ or to estimate $\rr$ without adequate data.
-
-\subsubsection{Gaussian Monotonic Errors}
-\begin{figure}[H]
-	\centering
-	\includegraphics[scale=0.5]{images/time_nm.jpeg}
-	\caption{Forest plots of estimates of $\rr$ for the \xxsir methods for the Time1-4 data sets with Gaussian errors $(\beta=.06, \gamma=.03, O=365, X(0)=99950, Y(0)=50, Z(0)=0, \sigma_X=100, \sigma_Y=5, N=10^5)$.
-The points are the estimates of $\rr$ and the lines denote $\pm 2\cdot $Std.
-Err.
-The vertical line is the true $\rr$ value.}
-\end{figure}
-
-\begin{table}[H]
-
-	\centering
-	\begin{tabular}[t]{l|r|r}
-		\hline
-		Model & Estimate & Std. Dev\\
-		\hline
-		LS & 1.9934 & 0.1464\\
-		\hline
-		ReLS & 1.9935 & 0.1975\\
-		\hline
-		LMA & 1.2024 & 0.5439\\
-		\hline
-		LMAT & 2.7981 & 3.8629\\
-		\hline
-		IPR & 2.2784 & 5.3298\\
-		\hline
-		SIPR & 2.2020 & 0.5643\\
-		\hline
-		LL & 1.9928 & 0.0105\\
-		\hline
-		MC & 0.9486 & $<$ 1e-04\\
-		\hline
-		SB & 1.0000 & 0.1000\\
-		\hline
-	\end{tabular}
-	\caption{ $T$ = 100, $\rr$ Estimates and Standard Errors}
-\end{table}
-
-\begin{table}[H]
-	\centering
-	\begin{tabular}[t]{l|r|r}
-		\hline
-		Model & Estimate & Std. Dev\\
-		\hline
-		LS & 5.0366 & 1.3672\\
-		\hline
-		ReLS & 5.0449 & 2.1398\\
-		\hline
-		LMA & 1.1218 & 1.7267\\
-		\hline
-		LMAT & 2.7046 & 4.8031\\
-		\hline
-		IPR & 1.4155 & 4.5655\\
-		\hline
-		SIPR & 0.9703 & 0.8674\\
-		\hline
-		LL & 1.7472 & 0.3635\\
-		\hline
-		MC & 0.9486 & $<$ 1e-04\\
-		\hline
-		SB & 1.0000 & 0.2236\\
-		\hline
-	\end{tabular}
-	\caption{$T$ = 20, $\rr$ Estimates and Standard Errors}
-\end{table}
-
-\subsubsection{Autoregressive Errors}
-\begin{figure}[H]
-	\centering
-	\includegraphics[scale=0.5]{images/time_ar.jpeg}
-	\caption{Forest plots of estimates of $\rr$ for the \xxsir methods for the Time1-4 data sets with Gaussian errors $(\beta=.06, \gamma=.03, O=365, X(0)=99950, Y(0)=50, Z(0)=0, \sigma_X=100, \sigma_Y=5, N=10^5)$.
-The points are the estimates of $\rr$ and the lines denote $\pm 2\cdot $Std.
-Err.
-The vertical line is the true $\rr$ value.}
-\end{figure}
-
-\begin{table}[H]
-
-	\centering
-	\begin{tabular}[t]{l|r|r}
-		\hline
-		Model & Estimate & Std. Dev\\
-		\hline
-		LS & 1.9098 & 0.3811\\
-		\hline
-		ReLS & 1.9104 & 0.4773\\
-		\hline
-		LMA & 0.9141 & 0.2563\\
-		\hline
-		LMAT & 0.7524 & 2.5623\\
-		\hline
-		IPR & 9.2219 & 18.1185\\
-		\hline
-		SIPR & 2.0919 & 1.0699\\
-		\hline
-		LL & 1.7390 & 0.0358\\
-		\hline
-		MC & 0.9486 & $<$ 1e-04\\
-		\hline
-		SB & 3.3635 & 0.1834\\
-		\hline
-	\end{tabular}
-	\caption{ $T$ = 100, $\rr$ Estimates and Standard Errors}
-\end{table}
-
-\begin{table}[H]
-	\centering
-	\begin{tabular}[t]{l|r|r}
-		\hline
-		Model & Estimate & Std. Dev\\
-		\hline
-		LS & 0.0099 & 2.1688\\
-		\hline
-		ReLS & -0.1539 & 2.7692\\
-		\hline
-		LMA & -5.2984 & 65.6503\\
-		\hline
-		LMAT & 0.5717 & 1.3982\\
-		\hline
-		IPR & 20.4892 & 33.7260\\
-		\hline
-		SIPR & 0.3570 & 7.6611\\
-		\hline
-		LL & 0.9171 & 0.0338\\
-		\hline
-		MC & 0.9486 & $<$ 1e-04\\
-		\hline
-		SB & 11.3793 & 0.7543\\
-		\hline
-	\end{tabular}
-	\caption{$T$ = 20, $\rr$ Estimates and Standard Errors}
-\end{table}
-
-\subsubsection{Autoregressive Monotonic Errors}
-\begin{figure}[H]
-	\centering
-	\includegraphics[scale=0.5]{images/time_arm.jpeg}
-	\caption{Forest plots of estimates of $\rr$ for the \xxsir methods for the Time1-4 data sets with Gaussian errors $(\beta=.06, \gamma=.03, O=365, X(0)=99950, Y(0)=50, Z(0)=0, \sigma_X=100, \sigma_Y=5, N=10^5)$.
-The points are the estimates of $\rr$ and the lines denote $\pm 2\cdot $Std.
-Err.
-The vertical line is the true $\rr$ value.}
-\end{figure}
-
-\begin{table}[H]
-
-	\centering
-	\begin{tabular}[t]{l|r|r}
-		\hline
-		Model & Estimate & Std. Dev\\
-		\hline
-		LS & 1.9674 & 0.2324\\
-		\hline
-		ReLS & 1.9679 & 0.3065\\
-		\hline
-		LMA & 1.1584 & 0.3660\\
-		\hline
-		LMAT & 3.4346 & 14.6041\\
-		\hline
-		IPR & 2.2454 & 5.7090\\
-		\hline
-		SIPR & 1.9631 & 0.3489\\
-		\hline
-		LL & 2.0524 & 0.0209\\
-		\hline
-		MC & 0.9486 & $<$ 1e-04\\
-		\hline
-		SB & 1.0000 & 0.1000\\
-		\hline
-	\end{tabular}
-	\caption{ $T$ = 100, $\rr$ Estimates and Standard Errors}
-\end{table}
-
-\begin{table}[H]
-	\centering
-	\begin{tabular}[t]{l|r|r}
-		\hline
-		Model & Estimate & Std. Dev\\
-		\hline
-		LS & 1.6816 & 0.4747\\
-		\hline
-		ReLS & 1.6813 & 0.6544\\
-		\hline
-		LMA & 0.8899 & 0.7605\\
-		\hline
-		LMAT & 1.4496 & 1.5423\\
-		\hline
-		IPR & 3.8856 & 8.2899\\
-		\hline
-		SIPR & 3.9849 & 2.1351\\
-		\hline
-		LL & 1.4328 & 0.0397\\
-		\hline
-		MC & 0.9486 & $<$ 1e-04\\
-		\hline
-		SB & 1.0000 & 0.2236\\
-		\hline
-	\end{tabular}
-	\caption{$T$ = 20, $\rr$ Estimates and Standard Errors}
-\end{table}
-
-\subsection{Initial Percentage of Infectious $\left (\frac{Y(0)}{N}\right)$}\label{sec:res-inf}
-We next examine the initial conditions of the number of susceptible, infectious, and recovered.
-Deterministically, different initial conditions in the number of infectious can cause very different epidemic courses (see Figures in appendix).
-
-In Figure \ref{fig:inits-res} and Tables \ref{tab:inits-res1} and \ref{tab:inits-res2}, we display the results of different number of initially infectious individuals of 1, 10, 100, and 1000 with a constant population of 100,000.
-We see that for a smaller percentage of initially infectious individuals, we have larger CIs in our estimates of $\rr$.
-In fact, LL is the only method that has a consistently small CI around the true $\rr=2$, regardless of the starting number of infectious.
-
-\subsubsection{Gaussian Errors}
-
-\begin{figure}[H]
-  \centering
-  \includegraphics[scale=0.5]{images/start_n.jpeg}
-    \caption{Forest plots of estimates of $\rr$ for the \xxsir methods for the Inits1-4 data sets $(\beta=.06, \gamma=.03, O=365, X(0)=99950, Y(0)=50, Z(0)=0, \sigma_X=100, \sigma_Y=5, N=10^5)$.
-The points are the estimates of $\rr$ and the lines denote $\pm 2\cdot $Std.
-Err.
-The vertical line is the true $\rr$ value.}\label{fig:inits-res}
-  \end{figure}
-
-  Table \ref{tab:inits-res1} shows the methods applied to data where $Y(0) = 1000$, essentially starting at the middle of an epidemic.
-This data has very large underestimation of $\rr$.
-The LMA method once again gives an estimate very far from the true $\rr$, potentially indicating that a linear approximation does not apply in this case.
-The sample standard errors in this case are actually relatively close to the baseline, and even smaller for some of the methods.
-
-Table \ref{tab:inits-res2} shows the methods applied to data where $Y(0) = 1$, starting at the very beginning of an epidemic, and the true technical definition of $\rr$ where we introduce exactly one infectious individual into a population.
-The methods give many larger estimates, tending to overestimate the true $\rr$.
-In particular, IPR and LMA methods give very large results.
-We also see a larger standard error than the baseline.
-
-\begin{table}[H]
-
-	\centering
-	\begin{tabular}[t]{l|r|r}
-		\hline
-		Model & Estimate & Std. Err\\
-		\hline
-		LS & 1.9996 & 0.0050\\
-		\hline
-		ReLS & 1.9998 & 0.0034\\
-		\hline
-		LMA & -87.5946 & 66413.1280\\
-		\hline
-		LMAT & 0.7306 & 4.6810\\
-		\hline
-		IPR & 1.2851 & 1.4410\\
-		\hline
-		SIPR & 1.2522 & 57.9230\\
-		\hline
-		LL & 1.9996 & 0.0003 \\
-		\hline
-		MC & 0.9486 & $<$ 1e-04\\
-		\hline
-		SB & 0.9729 & 0.0516\\
-		\hline
-	\end{tabular}
-        \caption{$Y(0) = 1000$, $\rr$ Estimates and Standard Errors}\label{tab:inits-res1}
-\end{table}
-
-\begin{table}[H]
-
-	\centering
-	\begin{tabular}[t]{l|r|r}
-		\hline
-		Model & Estimate & Std. Err\\
-		\hline
-		LS & 2.4683 & 1.5371\\
-		\hline
-		ReLS & 2.4692 & 2.7294 \\
-		\hline
-		LMA & 10.9282 & 19.2974 \\
-		\hline
-		LMAT & 1.6246 & 5.0360\\
-		\hline
-		IPR & 774.2747 & 7491.0051 \\
-		\hline
-		SIPR & 2.4391 & $<$ 1e-04 \\
-		\hline
-		LL & 1.9997 & 0.0008\\
-		\hline
-		MC & 0.9486 & 1.56e-05\\
-		\hline
-		SB & 1.4715 & 0.0635\\
-		\hline
-	\end{tabular}
-        \caption{$Y(0) = 1$, $\rr$ Estimates and Standard Errors}\label{tab:inits-res2}
-\end{table}
-
-\subsubsection{Gaussian Montonic Errors}
-
-\begin{figure}[H]
-	\centering
-	\includegraphics[scale=0.5]{images/start_nm.jpeg}
-	\caption{Forest plots of estimates of $\rr$ for the \xxsir methods for the Inits1-4 data sets $(\beta=.06, \gamma=.03, O=365, X(0)=99950, Y(0)=50, Z(0)=0, \sigma_X=100, \sigma_Y=5, N=10^5)$.
-The points are the estimates of $\rr$ and the lines denote $\pm 2\cdot $Std.
-Err.
-The vertical line is the true $\rr$ value.}
-\end{figure}
-
-\begin{table}[H]
-
-	\centering
-	\begin{tabular}[t]{l|r|r}
-		\hline
-		Model & Estimate & Std. Dev\\
-		\hline
-		LS & 2.0003 & 0.0046\\
-		\hline
-		ReLS & 2.0003 & 0.0032\\
-		\hline
-		LMA & -0.9982 & 1.4189\\
-		\hline
-		LMAT & 0.9703 & 0.6574\\
-		\hline
-		IPR & 0.9866 & 1.0570\\
-		\hline
-		SIPR & 2.0166 & 57.8154\\
-		\hline
-		LL & 2.0001 & 0.0002\\
-		\hline
-		MC & 0.9486 & $<$ 1e-04\\
-		\hline
-		SB & 1.0868 & 0.0546\\
-		\hline
-	\end{tabular}
-	\caption{$Y(0) = 1000$, $\rr$ Estimates and Standard Errors}
-\end{table}
-
-\begin{table}[H]
-
-	\centering
-	\begin{tabular}[t]{l|r|r}
-		\hline
-		Model & Estimate & Std. Dev\\
-		\hline
-		LS & 2.4668 & 1.5349\\
-		\hline
-		ReLS & 2.4670 & 2.7219\\
-		\hline
-		LMA & -99.3726 & 3136.3141\\
-		\hline
-		LMAT & 0.4205 & 12.9331\\
-		\hline
-		IPR & 3.6061 & 26.5409\\
-		\hline
-		SIPR & 5.4116 & $<$ 1e-04\\
-		\hline
-		LL & 2.0005 & 0.0004\\
-		\hline
-		MC & 0.9486 & $<$ 1e-04\\
-		\hline
-		SB & 1.7685 & 0.0696\\
-		\hline
-	\end{tabular}
-	\caption{$Y(0) = 1$, $\rr$ Estimates and Standard Errors}
-\end{table}
-
-\subsubsection{Autoregressive Errors}
-
-\begin{figure}[H]
-	\centering
-	\includegraphics[scale=0.5]{images/start_ar.jpeg}
-	\caption{Forest plots of estimates of $\rr$ for the \xxsir methods for the Inits1-4 data sets $(\beta=.06, \gamma=.03, O=365, X(0)=99950, Y(0)=50, Z(0)=0, \sigma_X=100, \sigma_Y=5, N=10^5)$.
-The points are the estimates of $\rr$ and the lines denote $\pm 2\cdot $Std.
-Err.
-The vertical line is the true $\rr$ value.}
-\end{figure}
-
-\begin{table}[H]
-
-	\centering
-	\begin{tabular}[t]{l|r|r}
-		\hline
-		Model & Estimate & Std. Dev\\
-		\hline
-		LS & 2.0005 & 0.0078\\
-		\hline
-		ReLS & 2.0001 & 0.0054\\
-		\hline
-		LMA & 4.1646 & 3.8688\\
-		\hline
-		LMAT & 3.1508 & 45.5885\\
-		\hline
-		IPR & 1.1936 & 1.1958\\
-		\hline
-		SIPR & 2.2854 & 57.4383\\
-		\hline
-		LL & 1.9998 & 0.0004\\
-		\hline
-		MC & 0.9486 & $<$ 1e-04\\
-		\hline
-		SB & 0.3154 & 0.0294\\
-		\hline
-	\end{tabular}
-	\caption{$Y(0) = 1000$, $\rr$ Estimates and Standard Errors}
-\end{table}
-
-\begin{table}[H]
-
-	\centering
-	\begin{tabular}[t]{l|r|r}
-		\hline
-		Model & Estimate & Std. Dev\\
-		\hline
-		LS & 2.4706 & 1.5372\\
-		\hline
-		ReLS & 2.4707 & 2.7245\\
-		\hline
-		LMA & -0.8259 & 1.7277\\
-		\hline
-		LMAT & 1.3117 & 26.0563\\
-		\hline
-		IPR & 5.0000 & 1199.1557\\
-		\hline
-		SIPR & 1.2047 & $<$ 1e-04\\
-		\hline
-		LL & 2.0027 & 0.0011\\
-		\hline
-		MC & 0.9486 & $<$ 1e-04\\
-		\hline
-		SB & 1.7698 & 0.0696\\
-		\hline
-	\end{tabular}
-	\caption{$Y(0) = 1$, $\rr$ Estimates and Standard Errors}
-\end{table}
-
-\subsubsection{Autoregressive Montonic Errors}
-
-\begin{figure}[H]
-	\centering
-	\includegraphics[scale=0.5]{images/start_arm.jpeg}
-	\caption{Forest plots of estimates of $\rr$ for the \xxsir methods for the Inits1-4 data sets $(\beta=.06, \gamma=.03, O=365, X(0)=99950, Y(0)=50, Z(0)=0, \sigma_X=100, \sigma_Y=5, N=10^5)$.
-The points are the estimates of $\rr$ and the lines denote $\pm 2\cdot $Std.
-Err.
-The vertical line is the true $\rr$ value.}
-\end{figure}
-
-\begin{table}[H]
-
-	\centering
-	\begin{tabular}[t]{l|r|r}
-		\hline
-		Model & Estimate & Std. Dev\\
-		\hline
-		LS & 2.0001 & 0.0069\\
-		\hline
-		ReLS & 2.0000 & 0.0048\\
-		\hline
-		LMA & -10.8054 & 286.9861\\
-		\hline
-		LMAT & 0.9333 & 0.8824\\
-		\hline
-		IPR & 0.9960 & 0.9987\\
-		\hline
-		SIPR & 1.9222 & 57.0458\\
-		\hline
-		LL & 2.0006 & 0.0003\\
-		\hline
-		MC & 0.9486 & $<$ 1e-04\\
-		\hline
-		SB & 0.3790 & 0.0322\\
-		\hline
-	\end{tabular}
-	\caption{$Y(0) = 1000$, $\rr$ Estimates and Standard Errors}
-\end{table}
-
-\begin{table}[H]
-
-	\centering
-	\begin{tabular}[t]{l|r|r}
-		\hline
-		Model & Estimate & Std. Dev\\
-		\hline
-		LS & 2.4667 & 1.5349\\
-		\hline
-		ReLS & 2.4665 & 2.7349\\
-		\hline
-		LMA & 16.7344 & 18.2294\\
-		\hline
-		LMAT & 3.1984 & 12.9100\\
-		\hline
-		IPR & 2.5082 & 12.8639\\
-		\hline
-		SIPR & 1.0356 & $<$ 1e-04\\
-		\hline
-		LL & 1.9999 & 0.0006\\
-		\hline
-		MC & 0.9486 & $<$ 1e-04\\
-		\hline
-		SB & 1.7793 & 0.0698\\
-		\hline
-	\end{tabular}
-	\caption{$Y(0) = 1$, $\rr$ Estimates and Standard Errors}
-\end{table}
-
-\subsection{Population Size $N$}\label{sec:res-n}
-We next examine the effects on the estimates of $\rr$ when changing the total population size, $N$.
-The percent of initially infectious individuals is 1\% when $N=10,000$ and $N=100,000$ and .5\% when $N=1,000$ and $N=100$.
-Intuitively, we would expect to see the variance decrease and to have more accurate estimates with a larger $N$ compared to a smaller one, due to the central limit theorem.
-This is exactly what we see for some of the methods, but not all of them.
-
-We display the results of estimating $\rr$ from the methods fit on the baseline data set with varying $N$ in Figure \ref{fig:inits-res} and Tables \ref{tab:n1-res2} and \ref{tab:n2-res2}.
-When $N=100$, LS, ReLS, LMA, LMAT, IPR, and SIPR's 95\% CIs all cover the true $\rr$.
-However, we would say that all of these intervals are uninformative.
-We see that CIs of LS are wider than those of ReLS, although the difference becomes less substantial as $N$ grows larger.
-
-When $N=100$, we would be able to say the estimate for $\rr$ from LL is greater than 1 with 95\% confidence, but the estimate is much smaller than the true value.
-MC and SB would actually allow us to say $\rr < 1$ with 95\% confidence.
-It is not until $N=10,000$ that we are able to say with certainty that LS and ReLS, in addition to LL, give informative estimates of $\rr$.
-
-Thus, it seems that if one is estimating for $\rr$ for a small population such as a small village or crusie ship, LL may be the most appropriate method, but when the population is larger (~10,000), we can be confident in using LS or ReLS.
-
-\subsubsection{Gaussian Errors}
-
-\begin{figure}[H]
-	\centering
-	\includegraphics[scale=0.5]{images/popsize_n.jpeg}
-	\caption{Forest plots of estimates of $\rr$ for the \xxsir methods for the PopSize1-4 data sets $(\beta=.06, \gamma=.03, O=365, \sigma_X=100, \sigma_Y=5)$.
-The points are the estimates of $\rr$ and the lines denote $\pm 2\cdot $Std.
-Err.
-The vertical line is the true $\rr$ value.}\label{fig:inits-res2}
-\end{figure}
-
-\begin{table}[H]
-
-	\centering
-	\begin{tabular}[t]{l|r|r}
-		\hline
-		Model & Estimate & Std. Err\\
-		\hline
-		LS & 2.0008 & 0.0590\\
-		\hline
-		ReLS & 2.0005 & 0.0528\\
-		\hline
-		LMA & 1.9976 & 2606.0798\\
-		\hline
-		LMAT & 1.2944 & 1.4305\\
-		\hline
-		IPR & 40.8072 & 133.2798\\
-		\hline
-		SIPR & 1.0629 &  $<$ 1e-04\\
-		\hline
-		LL & 1.9988 & 0.0023\\
-		\hline
-		MC & 0.9486 & 4.0e-06\\
-		\hline
-		SB & 0.4418 & 0.0348\\
-		\hline
-	\end{tabular}
-\caption{$N = 10000$, $\rr$ Estimates and Standard Errors}\label{tab:n1-res2}
-\end{table}
-
-\begin{table}[H]
-
-	\centering
-	\begin{tabular}[t]{l|r|r}
-		\hline
-		Model & Estimate & Std. Err\\
-		\hline
-		LS & 1.9863 & 4.8620\\
-		\hline
-		ReLS & 1.9871 & 3.4569\\
-		\hline
-		LMA & 0.8023 & 18.9417\\
-		\hline
-		LMAT & 1.2209 & 4.7699\\
-		\hline
-		IPR & 3620.8677 & 57608.1700\\
-		\hline
-		SIPR & 3.3678 & 22.4187\\
-		\hline
-		LL & 1.1413 & 0.0422\\
-		\hline
-		MC & 0.8456 & 0.0125\\
-		\hline
-		SB & 0.3887 & 0.0326\\
-		\hline
-	\end{tabular}
-\caption{$N = 100$, $\rr$ Estimates and Standard Errors}\label{tab:n2-res2}
-\end{table}
-
-\subsubsection{Gaussian Monotonic Errors}
-
-\begin{figure}[H]
-	\centering
-	\includegraphics[scale=0.5]{images/popsize_nm.jpeg}
-	\caption{Forest plots of estimates of $\rr$ for the \xxsir methods for the PopSize1-4 data sets $(\beta=.06, \gamma=.03, O=365, \sigma_X=100, \sigma_Y=5)$.
-The points are the estimates of $\rr$ and the lines denote $\pm 2\cdot $Std.
-Err.
-The vertical line is the true $\rr$ value.}
-\end{figure}
-
-\begin{table}[H]
-
-	\centering
-	\begin{tabular}[t]{l|r|r}
-		\hline
-		Model & Estimate & Std. Dev\\
-		\hline
-		LS & 2.0175 & 0.1430\\
-		\hline
-		ReLS & 2.0175 & 0.1021\\
-		\hline
-		LMA & 0.8256 & 1.4444\\
-		\hline
-		LMAT & 1.0457 & 0.6365\\
-		\hline
-		IPR & 1.7343 & 12.7260\\
-		\hline
-		SIPR & 1.6932 & 12.0122\\
-		\hline
-		LL & 2.0102 & 0.0056\\
-		\hline
-		MC & 0.9486 & $<$ 1e-04\\
-		\hline
-		SB & 1.0000 & 0.0523\\
-		\hline
-	\end{tabular}
-	\caption{$N = 10000$, $\rr$ Estimates and Standard Errors}
-\end{table}
-
-\begin{table}[H]
-
-	\centering
-	\begin{tabular}[t]{l|r|r}
-		\hline
-		Model & Estimate & Std. Dev\\
-		\hline
-		LS & 1.9998 & 0.0006\\
-		\hline
-		ReLS & 1.9999 & 0.0005\\
-		\hline
-		LMA & 1.7673 & 0.0976\\
-		\hline
-		LMAT & 1.9732 & 10.8872\\
-		\hline
-		IPR & 1.4344 & 1.1311\\
-		\hline
-		SIPR & 1.7907 & $<$ 1e-04\\
-		\hline
-		LL & 2.0000 & $<$ 1e-04\\
-		\hline
-		MC & 0.9486 & $<$ 1e-04\\
-		\hline
-		SB & 1.2709 & 0.0590\\
-		\hline
-	\end{tabular}
-	\caption{$N = 100$, $\rr$ Estimates and Standard Errors}
-\end{table}
-
-\subsubsection{Autoregressive Errors}
-
-\begin{figure}[H]
-	\centering
-	\includegraphics[scale=0.5]{images/popsize_ar.jpeg}
-	\caption{Forest plots of estimates of $\rr$ for the \xxsir methods for the PopSize1-4 data sets $(\beta=.06, \gamma=.03, O=365, \sigma_X=100, \sigma_Y=5)$.
-The points are the estimates of $\rr$ and the lines denote $\pm 2\cdot $Std.
-Err.
-The vertical line is the true $\rr$ value.}
-\end{figure}
-
-\begin{table}[H]
-
-	\centering
-	\begin{tabular}[t]{l|r|r}
-		\hline
-		Model & Estimate & Std. Dev\\
-		\hline
-		LS & 2.1221 & 0.7363\\
-		\hline
-		ReLS & 2.1219 & 0.5660\\
-		\hline
-		LMA & 1.2045 & 0.3878\\
-		\hline
-		LMAT & 1.0881 & 3.9239\\
-		\hline
-		IPR & 164.6111 & 1887.1253\\
-		\hline
-		SIPR & 0.5335 & 13.6519\\
-		\hline
-		LL & 2.4701 & 0.0546\\
-		\hline
-		MC & 0.9500 & 0.0003\\
-		\hline
-		SB & 1.5457 & 0.0651\\
-		\hline
-	\end{tabular}
-	\caption{$N = 10000$, $\rr$ Estimates and Standard Errors}
-\end{table}
-
-\begin{table}[H]
-
-	\centering
-	\begin{tabular}[t]{l|r|r}
-		\hline
-		Model & Estimate & Std. Dev\\
-		\hline
-		LS & 1.9998 & 0.0009\\
-		\hline
-		ReLS & 2.0001 & 0.0007\\
-		\hline
-		LMA & 1.6873 & 0.0810\\
-		\hline
-		LMAT & 1.3885 & 0.6855\\
-		\hline
-		IPR & 1.5672 & 1.5944\\
-		\hline
-		SIPR & 1.7903 & $<$ 1e-04\\
-		\hline
-		LL & 1.9999 & $<$ 1e-04\\
-		\hline
-		MC & 0.9486 & $<$ 1e-04\\
-		\hline
-		SB & 1.1877 & 0.0570\\
-		\hline
-	\end{tabular}
-	\caption{$N = 100$, $\rr$ Estimates and Standard Errors}
-\end{table}
-
-\subsubsection{Autoregressive Monotonic Errors}
-
-\begin{figure}[H]
-	\centering
-	\includegraphics[scale=0.5]{images/popsize_arm.jpeg}
-	\caption{Forest plots of estimates of $\rr$ for the \xxsir methods for the PopSize1-4 data sets $(\beta=.06, \gamma=.03, O=365, \sigma_X=100, \sigma_Y=5)$.
-The points are the estimates of $\rr$ and the lines denote $\pm 2\cdot $Std.
-Err.
-The vertical line is the true $\rr$ value.}
-\end{figure}
-
-\begin{table}[H]
-
-	\centering
-	\begin{tabular}[t]{l|r|r}
-		\hline
-		Model & Estimate & Std. Dev\\
-		\hline
-		LS & 2.0975 & 0.2921\\
-		\hline
-		ReLS & 2.0976 & 0.2211\\
-		\hline
-		LMA & 1.1720 & 0.0511\\
-		\hline
-		LMAT & -13.6906 & 281.9114\\
-		\hline
-		IPR & 1.7077 & 11.7775\\
-		\hline
-		SIPR & 2.8981 & 284.4933\\
-		\hline
-		LL & 2.2806 & 0.0230\\
-		\hline
-		MC & 0.9486 & $<$ 1e-04\\
-		\hline
-		SB & 1.0000 & 0.0523\\
-		\hline
-	\end{tabular}
-	\caption{$N = 10000$, $\rr$ Estimates and Standard Errors}
-\end{table}
-
-\begin{table}[H]
-
-	\centering
-	\begin{tabular}[t]{l|r|r}
-		\hline
-		Model & Estimate & Std. Dev\\
-		\hline
-		LS & 2.0000 & 0.0007\\
-		\hline
-		ReLS & 1.9997 & 0.0007\\
-		\hline
-		LMA & 1.8332 & 0.1199\\
-		\hline
-		LMAT & 1.4164 & 0.6348\\
-		\hline
-		IPR & 1.4490 & 1.1803\\
-		\hline
-		SIPR & 1.7505 & $<$ 1e-04\\
-		\hline
-		LL & 1.9998 & $<$ 1e-04\\
-		\hline
-		MC & 0.9486 & $<$ 1e-04\\
-		\hline
-		SB & 1.8078 & 0.0704\\
-		\hline
-	\end{tabular}
-	\caption{$N = 100$, $\rr$ Estimates and Standard Errors}
-\end{table}
-
-\subsection{Magnitude of Variance $(\sigma^2_X, \sigma^2_Y)$}\label{sec:res-var}
-We examine the estimates of $\rr$ when changing the values of the variance of the number of susceptibles and number of infectious.
-We would expect to see wider CIs when we have larger variances in the $X$ and $Y$ compartments, but we would like to determine how sensitive estimates of $\rr$ are to these variance values.
-
-\subsubsection{Gaussian Errors}
-
-\begin{figure}[H]
-\begin{center}
-  \includegraphics[scale=0.5]{images/var_n.jpeg}
-  \caption{Forest plots of estimates of $\rr$ for the \xxsir methods for the Var1-4 data sets $(\beta=.06, \gamma=.03, O=365, T=365, X(0)=99950, Y(0)=50, Z(0)=0, N=10^5)$.
-The points are the estimates of $\rr$ and the lines denote $\pm 2\cdot $Std.
-Err.
-The vertical line is the true $\rr$ value.}
-  \label{fig:mag-var-res}
-\end{center}
-\end{figure}
-
-We display the results in Figure \ref{fig:mag-var-res} and Tables \ref{tab:mag-var-res1}-\ref{tab:mag-var-res2}.
-Generally, we see wider CIs and more biased estimates of $\rr$ when the variance levels are large.
-In particular, LMA and LMAT seem to be particularly effected as the CIs grow substantially as the variance grows.
-Again we see that LS, ReLS, and LL provide the most accurate estimates with relatively small CIs.
-
-\begin{table}[H]
-
-	\centering
-	\begin{tabular}[t]{l|r|r}
-		\hline
-		Model & Estimate & Std. Err\\
-		\hline
-		LS & 1.9998 & 0.0027\\
-		\hline
-		ReLS & 1.9998 & 0.0024\\
-		\hline
-		LMA & 2.0398 & 0.2774\\
-		\hline
-		LMAT & 1.4149 & 0.5780\\
-		\hline
-		IPR & 2.7608 & 8.1319\\
-		\hline
-		SIPR & 1.8368 & $<$ 1e-04\\
-		\hline
-		LL & 1.9999 & 0.0001\\
-		\hline
-		MC & 0.9486 & $<$ 1e-04\\
-		\hline
-		SB & 1.6750 & 0.0677\\
-		\hline
-	\end{tabular}
-        	\caption{ $(\sigma^2_X, \sigma^2_Y) = (50, 2)$, $\rr$ Estimates and Standard Errors}\label{tab:mag-var-res1}
-\end{table}
-
-In Table \ref{tab:mag-var-res1}, the magnitude of the variance of the noise is smaller than in the baseline.
-The estimates are closer to the true $\rr$ and the standard errors are smaller across the board.
-This indicates that the methods give more accurate and precise estimates when the data are closer to following underlying true SIR model.
-
-\begin{table}[H]
-
-	\centering
-	\begin{tabular}[t]{l|r|r}
-		\hline
-		Model & Estimate & Std. Err\\
-		\hline
-		LS & 1.9981 & 0.1354\\
-		\hline
-		ReLS & 1.9981 & 0.1208\\
-		\hline
-		LMA & 0.4652 & 7.5508\\
-		\hline
-		LMAT & 1.2878 & 4.3948\\
-		\hline
-		IPR & 81.4902 & 348.7351\\
-		\hline
-		SIPR & 1.0606 & $<$ 1e-04\\
-		\hline
-		LL & 2.0019 & 0.0054\\
-		\hline
-		MC & 0.9486 & 1.0e-06\\
-		\hline
-		SB & 1.8567 & 0.0713\\
-		\hline
-	\end{tabular}
-        	\caption{$(\sigma^2_X, \sigma^2_Y) = (2500, 100)$, $\rr$ Estimates and Standard Errors}\label{tab:mag-var-res2}
-\end{table}
-
-In Table \ref{tab:mag-var-res2}, the magnitude of the variance of the noise is larger than in the baseline.
-The estimates are worse and the standard errors are larger compared to the baseline data set, which we might expect.
-These estimates are not as terrible as we might imagine given the magnitude of the change - it seems that these methods are somewhat more robust to large noise compared to other changes.
-One exception in the IPR model, which may do worse due to larger fluctuations in the incidence prevalence ratio.
-
-\subsubsection{Gaussian Monotonic Errors}
-
-\begin{figure}[H]
-	\begin{center}
-		\includegraphics[scale=0.5]{images/var_nm.jpeg}
-		\caption{Forest plots of estimates of $\rr$ for the \xxsir methods for the Var1-4 data sets $(\beta=.06, \gamma=.03, O=365, T=365, X(0)=99950, Y(0)=50, Z(0)=0, N=10^5)$.
-The points are the estimates of $\rr$ and the lines denote $\pm 2\cdot $Std.
-Err.
-The vertical line is the true $\rr$ value.}
-	\end{center}
-\end{figure}
-
-\begin{table}[H]
-
-	\centering
-	\begin{tabular}[t]{l|r|r}
-		\hline
-		Model & Estimate & Std. Dev\\
-		\hline
-		LS & 1.9998 & 0.0026\\
-		\hline
-		ReLS & 1.9999 & 0.0024\\
-		\hline
-		LMA & 1.6268 & 0.0865\\
-		\hline
-		LMAT & 1.5099 & 2.1244\\
-		\hline
-		IPR & 1.4509 & 2.0522\\
-		\hline
-		SIPR & 1.7778 & $<$ 1e-04\\
-		\hline
-		LL & 2.0000 & 0.0001\\
-		\hline
-		MC & 0.9486 & $<$ 1e-04\\
-		\hline
-		SB & 1.2333 & 0.0581\\
-		\hline
-	\end{tabular}
-	\caption{ $(\sigma^2_X, \sigma^2_Y) = (50, 2)$, $\rr$ Estimates and Standard Errors}
-\end{table}
-
-\begin{table}[H]
-
-	\centering
-	\begin{tabular}[t]{l|r|r}
-		\hline
-		Model & Estimate & Std. Dev\\
-		\hline
-		LS & 1.9970 & 0.0541\\
-		\hline
-		ReLS & 1.9971 & 0.0482\\
-		\hline
-		LMA & 4.2062 & 66.2206\\
-		\hline
-		LMAT & 1.3653 & 3.2675\\
-		\hline
-		IPR & 2.2672 & 17.0580\\
-		\hline
-		SIPR & 1.2346 & $<$ 1e-04\\
-		\hline
-		LL & 1.9957 & 0.0015\\
-		\hline
-		MC & 0.9486 & $<$ 1e-04\\
-		\hline
-		SB & 1.0000 & 0.0523\\
-		\hline
-	\end{tabular}
-	\caption{$(\sigma^2_X, \sigma^2_Y) = (2500, 100)$, $\rr$ Estimates and Standard Errors}
-\end{table}
-
-\subsubsection{Autoregressive Errors}
-
-\begin{figure}[H]
-	\begin{center}
-		\includegraphics[scale=0.5]{images/var_ar.jpeg}
-		\caption{Forest plots of estimates of $\rr$ for the \xxsir methods for the Var1-4 data sets $(\beta=.06, \gamma=.03, O=365, T=365, X(0)=99950, Y(0)=50, Z(0)=0, N=10^5)$.
-The points are the estimates of $\rr$ and the lines denote $\pm 2\cdot $Std.
-Err.
-The vertical line is the true $\rr$ value.}
-	\end{center}
-\end{figure}
-
-\begin{table}[H]
-
-	\centering
-	\begin{tabular}[t]{l|r|r}
-		\hline
-		Model & Estimate & Std. Dev\\
-		\hline
-		LS & 1.9996 & 0.0038\\
-		\hline
-		ReLS & 1.9999 & 0.0034\\
-		\hline
-		LMA & 1.4644 & 0.0750\\
-		\hline
-		LMAT & 1.3999 & 0.7194\\
-		\hline
-		IPR & 2.1980 & 5.4569\\
-		\hline
-		SIPR & 1.7604 & $<$ 1e-04\\
-		\hline
-		LL & 1.9991 & 0.0001\\
-		\hline
-		MC & 0.9486 & $<$ 1e-04\\
-		\hline
-		SB & 1.6378 & 0.0670\\
-		\hline
-	\end{tabular}
-	\caption{ $(\sigma^2_X, \sigma^2_Y) = (50, 2)$, $\rr$ Estimates and Standard Errors}
-\end{table}
-
-\begin{table}[H]
-
-	\centering
-	\begin{tabular}[t]{l|r|r}
-		\hline
-		Model & Estimate & Std. Dev\\
-		\hline
-		LS & 1.9997 & 0.1894\\
-		\hline
-		ReLS & 2.0002 & 0.1680\\
-		\hline
-		LMA & 1.0392 & 0.0166\\
-		\hline
-		LMAT & 1.1985 & 1.6107\\
-		\hline
-		IPR & 195.3708 & 1451.0401\\
-		\hline
-		SIPR & 1.7028 & $<$ 1e-04\\
-		\hline
-		LL & 2.0386 & 0.0082\\
-		\hline
-		MC & 0.9486 & $<$ 1e-04\\
-		\hline
-		SB & 1.9073 & 0.0723\\
-		\hline
-	\end{tabular}
-	\caption{$(\sigma^2_X, \sigma^2_Y) = (2500, 100)$, $\rr$ Estimates and Standard Errors}
-\end{table}
-
-\subsubsection{Autoregressive Monotonic Errors}
-
-\begin{figure}[H]
-	\begin{center}
-		\includegraphics[scale=0.5]{images/var_arm.jpeg}
-		\caption{Forest plots of estimates of $\rr$ for the \xxsir methods for the Var1-4 data sets $(\beta=.06, \gamma=.03, O=365, T=365, X(0)=99950, Y(0)=50, Z(0)=0, N=10^5)$.
-The points are the estimates of $\rr$ and the lines denote $\pm 2\cdot $Std.
-Err.
-The vertical line is the true $\rr$ value.}
-	\end{center}
-\end{figure}
-
-\begin{table}[H]
-
-	\centering
-	\begin{tabular}[t]{l|r|r}
-		\hline
-		Model & Estimate & Std. Dev\\
-		\hline
-		LS & 1.9990 & 0.0036\\
-		\hline
-		ReLS & 1.9995 & 0.0032\\
-		\hline
-		LMA & 2.8515 & 0.5541\\
-		\hline
-		LMAT & 1.4263 & 0.6782\\
-		\hline
-		IPR & 1.4465 & 2.1821\\
-		\hline
-		SIPR & 1.8423 & $<$ 1e-04\\
-		\hline
-		LL & 1.9997 & 0.0001\\
-		\hline
-		MC & 0.9486 & $<$ 1e-04\\
-		\hline
-		SB & 1.5555 & 0.0653\\
-		\hline
-	\end{tabular}
-	\caption{ $(\sigma^2_X, \sigma^2_Y) = (50, 2)$, $\rr$ Estimates and Standard Errors}
-\end{table}
-
-\begin{table}[H]
-
-	\centering
-	\begin{tabular}[t]{l|r|r}
-		\hline
-		Model & Estimate & Std. Dev\\
-		\hline
-		LS & 2.0021 & 0.0928\\
-		\hline
-		ReLS & 2.0022 & 0.0827\\
-		\hline
-		LMA & -0.1164 & 1.9126\\
-		\hline
-		LMAT & 0.9424 & 4.3914\\
-		\hline
-		IPR & 2.9241 & 18.4344\\
-		\hline
-		SIPR & 2.5962 & $<$ 1e-04\\
-		\hline
-		LL & 2.0157 & 0.0031\\
-		\hline
-		MC & 0.9486 & $<$ 1e-04\\
-		\hline
-		SB & 1.0000 & 0.0523\\
-		\hline
-	\end{tabular}
-	\caption{$(\sigma^2_X, \sigma^2_Y) = (2500, 100)$, $\rr$ Estimates and Standard Errors}
-\end{table}
-
-\subsection{Other Models}\label{sec:res-oth}
-Until this point, we have only examined results while assuming that the epidemic truly follows a SIR model.
-However, this may not be the case.
-We examine the estimates of $\rr$ under different underlying models including a polynomial of degree 1, a polynomial of degree 4, and a linearized SIR model.
-We present the results of the estimates of $\rr$ on these models in Figure \ref{fig:other-res} and Tables \ref{tab:other-res1} - \ref{tab:other-res3}.
-
-\subsubsection{Gaussian Errors}
-
-\begin{figure}[H]
-  \begin{center}
-    \includegraphics[scale=0.5]{images/other_n.jpeg}
-  \caption{Forest plots of estimates of $\rr$ for the \xxsir methods for the Poly1, Poly4, and LSIR data sets $(\beta=.06, \gamma=.03, O=365, T=365, X(0)=99950, Y(0)=50, Z(0)=0, N=10^5, \sigma_X=100, \sigma_Y=5)$.
-The points are the estimates of $\rr$ and the lines denote $\pm 2\cdot $Std.
-Err.
-The vertical line is the true $\rr$ value.}
-  \label{fig:other-res}	
-  \end{center}
-\end{figure}
-
-For both the Poly1 and LSIR data sets, all the methods besides IPR and SIPR underestimate $\rr$.
-Moreover, only the CI of IPR covers that of the true value of $\rr=2$ and the CIs are uninformative.
-For the quartic model, only ReLS has a CI that both covers the true $\rr$ while rejecting the hypothesis that $\rr < 1$.
-LS covers the true value with its CI but also includes $\rr=1$.
-LL rejects the hypothesis that $\rr <$ but its CI does not contain $\rr=2$.
-
-\begin{table}[H]
-
-	\centering
-	\begin{tabular}[t]{l|r|r}
-		\hline
-		Model & Estimate & Std. Err\\
-		\hline
-		LS & 1.0514 & 0.3783\\
-		\hline
-		ReLS & 1.0514 & 0.0259\\
-		\hline
-		LMA & 1.2656 & 0.1278\\
-		\hline
-		LMAT & 1.1597 & 0.1020\\
-		\hline
-		IPR & 4.5000 & 15.2186\\
-		\hline
-		SIPR & 98.0611 & $<$ 1e-04\\
-		\hline
-		LL & 1.1881 & 0.0006\\
-		\hline
-		MC & 0.9486 & 1e-07 \\
-		\hline
-		SB & 1.4150 & 0.0622\\
-		\hline
-	\end{tabular}
-	\caption{Linear Data, $\rr$ Estimates and Standard Errors}\label{tab:other-res1}
-\end{table}
-
-\begin{table}[H]
-
-\centering
-\begin{tabular}[t]{l|r|r}
-	\hline
-	Model & Estimate & Std. Dev\\
-	\hline
-	LS & 1.7693 & 0.4227\\
-	\hline
-	ReLS & 1.7693 & 0.2797\\
-	\hline
-	LMA & 0.3618 & 0.1231\\
-	\hline
-	LMAT & 0.8213 & 6.0466\\
-	\hline
-	IPR & 10.4349 & 33.8953 \\
-	\hline
-	SIPR & 4.7477 &  $<$ 1e-04\\
-	\hline
-	LL & 1.9552 & 0.0165\\
-	\hline
-	MC & 0.9486 & 6e-07\\
-	\hline
-	SB & 0.2119 & 0.02409\\
-	\hline
-\end{tabular}
-\caption{4th Order Model, $\rr$ Estimates and Standard Errors}\label{tab:other-res2} 
-\end{table}
-
-Table \ref{tab:other-res2} shows the methods applied to data from a 4th order linear regression model with respect to time.
-Obviously, there is no true $\rr$ to compare to.
-What interests us in this case is whether the methods give similar results, since all the estimation methods are derived from the same model.
-However, the results vary much more compared to baseline data estimates, where underlying data came from SIR.
-We also get massive standard errors, even though actual noise variance magnitudes are the same as baseline.
-
-\begin{table}[H]
-
-	\centering
-	\begin{tabular}[t]{l|r|r}
-		\hline
-		Model & Estimate & Std. Err\\
-		\hline
-		LS & 1.2743 & 0.0980\\
-		\hline
-		ReLS & 1.2743 & 0.0340\\
-		\hline
-		LMA & 1.2191 & 0.1093\\
-		\hline
-		LMAT & 1.2009 & 0.0031\\
-		\hline
-		IPR & 2.5280 & 2.5904\\
-		\hline
-		SIPR & 2.1568 & 0.3585\\
-		\hline
-		LL & 1.2586 & 0.0010\\
-		\hline
-		MC & 0.9486 & $<$ 1e-04 \\
-		\hline
-		SB & 1.2009 & 0.0574\\
-		\hline
-	\end{tabular}
-	\caption{Linear SIR Data, $\rr$ Estimates and Standard Errors}\label{tab:other-res3}
-\end{table}
-
-\subsubsection{Gaussian Monotonic Errors}
-
-\begin{figure}[H]
-	\begin{center}
-		\includegraphics[scale=0.5]{images/other_nm.jpeg}
-		\caption{Forest plots of estimates of $\rr$ for the \xxsir methods for the Poly1, Poly4, and LSIR data sets $(\beta=.06, \gamma=.03, O=365, T=365, X(0)=99950, Y(0)=50, Z(0)=0, N=10^5, \sigma_X=100, \sigma_Y=5)$.
-The points are the estimates of $\rr$ and the lines denote $\pm 2\cdot $Std.
-Err.
-The vertical line is the true $\rr$ value.}
-	\end{center}
-\end{figure}
-
-\begin{table}[H]
-
-	\centering
-	\begin{tabular}[t]{l|r|r}
-		\hline
-		Model & Estimate & Std. Dev\\
-		\hline
-		LS & 1.0514 & 0.3785\\
-		\hline
-		ReLS & 1.0514 & 0.0259\\
-		\hline
-		LMA & 1.2300 & 0.0572\\
-		\hline
-		LMAT & 1.1471 & 0.1751\\
-		\hline
-		IPR & 2.2972 & 7.1420\\
-		\hline
-		SIPR & 27.4297 & $<$ 1e-04\\
-		\hline
-		LL & 1.1880 & 0.0006\\
-		\hline
-		MC & 0.9486 & $<$ 1e-04\\
-		\hline
-		SB & 1.3647 & 0.0611\\
-		\hline
-	\end{tabular}
-	\caption{Linear Data, $\rr$ Estimates and Standard Errors}
-\end{table}
-
-\begin{table}[H]
-
-	\centering
-	\begin{tabular}[t]{l|r|r}
-		\hline
-		Model & Estimate & Std. Dev\\
-		\hline
-		LS & 1.7691 & 0.4216\\
-		\hline
-		ReLS & 1.7691 & 0.2791\\
-		\hline
-		LMA & -0.0201 & 0.1128\\
-		\hline
-		LMAT & 0.9972 & 2.5752\\
-		\hline
-		IPR & 5.9240 & 17.8114\\
-		\hline
-		SIPR & 15.0836 & $<$ 1e-04\\
-		\hline
-		LL & 1.9553 & 0.0164\\
-		\hline
-		MC & 0.9486 & $<$ 1e-04\\
-		\hline
-		SB & 0.2581 & 0.0266\\
-		\hline
-	\end{tabular}
-	\caption{4th Order Model, $\rr$ Estimates and Standard Errors}
-\end{table}
-
-\begin{table}[H]
-
-	\centering
-	\begin{tabular}[t]{l|r|r}
-		\hline
-		Model & Estimate & Std. Dev\\
-		\hline
-		LS & 1.2739 & 0.0977\\
-		\hline
-		ReLS & 1.2741 & 0.0337\\
-		\hline
-		LMA & 1.2068 & 0.0481\\
-		\hline
-		LMAT & 1.2002 & 0.0041\\
-		\hline
-		IPR & 1.2086 & 1.1254\\
-		\hline
-		SIPR & 1.2944 & 0.2158\\
-		\hline
-		LL & 1.2587 & 0.0010\\
-		\hline
-		MC & 0.9486 & $<$ 1e-04\\
-		\hline
-		SB & 1.3209 & 0.0602\\
-		\hline
-	\end{tabular}
-	\caption{Linear SIR Data, $\rr$ Estimates and Standard Errors}
-\end{table}
-
-\subsubsection{Autoregressive Errors}
-
-\begin{figure}[H]
-	\begin{center}
-		\includegraphics[scale=0.5]{images/other_ar.jpeg}
-		\caption{Forest plots of estimates of $\rr$ for the \xxsir methods for the Poly1, Poly4, and LSIR data sets $(\beta=.06, \gamma=.03, O=365, T=365, X(0)=99950, Y(0)=50, Z(0)=0, N=10^5, \sigma_X=100, \sigma_Y=5)$.
-The points are the estimates of $\rr$ and the lines denote $\pm 2\cdot $Std.
-Err.
-The vertical line is the true $\rr$ value.}
-	\end{center}
-\end{figure}
-
-\begin{table}[H]
-
-	\centering
-	\begin{tabular}[t]{l|r|r}
-		\hline
-		Model & Estimate & Std. Dev\\
-		\hline
-		LS & 1.0519 & 0.3775\\
-		\hline
-		ReLS & 1.0518 & 0.0260\\
-		\hline
-		LMA & 1.2766 & 0.1573\\
-		\hline
-		LMAT & 1.1558 & 0.0687\\
-		\hline
-		IPR & 4.3348 & 41.3251\\
-		\hline
-		SIPR & 1.4988 & $<$ 1e-04\\
-		\hline
-		LL & 1.1876 & 0.0006\\
-		\hline
-		MC & 0.9486 & $<$ 1e-04\\
-		\hline
-		SB & 1.4979 & 0.0641\\
-		\hline
-	\end{tabular}
-	\caption{Linear Data, $\rr$ Estimates and Standard Errors}
-\end{table}
-
-\begin{table}[H]
-
-	\centering
-	\begin{tabular}[t]{l|r|r}
-		\hline
-		Model & Estimate & Std. Dev\\
-		\hline
-		LS & 1.7690 & 0.4228\\
-		\hline
-		ReLS & 1.7691 & 0.2799\\
-		\hline
-		LMA & 0.3565 & 0.1259\\
-		\hline
-		LMAT & 3.2425 & 26.2628\\
-		\hline
-		IPR & 6.2685 & 20.5362\\
-		\hline
-		SIPR & 2.7662 & $<$ 1e-04\\
-		\hline
-		LL & 1.9551 & 0.0165\\
-		\hline
-		MC & 0.9486 & $<$ 1e-04\\
-		\hline
-		SB & 0.2132 & 0.0242\\
-		\hline
-	\end{tabular}
-	\caption{4th Order Model, $\rr$ Estimates and Standard Errors}
-\end{table}
-
-\begin{table}[H]
-
-	\centering
-	\begin{tabular}[t]{l|r|r}
-		\hline
-		Model & Estimate & Std. Dev\\
-		\hline
-		LS & 1.2737 & 0.1000\\
-		\hline
-		ReLS & 1.2742 & 0.0346\\
-		\hline
-		LMA & 1.2146 & 0.0627\\
-		\hline
-		LMAT & 1.2032 & 0.0148\\
-		\hline
-		IPR & 1.3799 & 1.3114\\
-		\hline
-		SIPR & 1.2967 & 0.2216\\
-		\hline
-		LL & 1.2589 & 0.0010\\
-		\hline
-		MC & 0.9486 & $<$ 1e-04\\
-		\hline
-		SB & 1.2039 & 0.0574\\
-		\hline
-	\end{tabular}
-	\caption{Linear SIR Data, $\rr$ Estimates and Standard Errors}
-\end{table}
-
-\subsubsection{Autoregressive Monotonic Errors}
-
-\begin{figure}[H]
-	\begin{center}
-		\includegraphics[scale=0.5]{images/other_arm.jpeg}
-		\caption{Forest plots of estimates of $\rr$ for the \xxsir methods for the Poly1, Poly4, and LSIR data sets $(\beta=.06, \gamma=.03, O=365, T=365, X(0)=99950, Y(0)=50, Z(0)=0, N=10^5, \sigma_X=100, \sigma_Y=5)$.
-The points are the estimates of $\rr$ and the lines denote $\pm 2\cdot $Std.
-Err.
-The vertical line is the true $\rr$ value.}
-	\end{center}
-\end{figure}
-
-\begin{table}[H]
-
-	\centering
-	\begin{tabular}[t]{l|r|r}
-		\hline
-		Model & Estimate & Std. Dev\\
-		\hline
-		LS & 1.0513 & 0.3783\\
-		\hline
-		ReLS & 1.0513 & 0.0257\\
-		\hline
-		LMA & 1.2569 & 0.1132\\
-		\hline
-		LMAT & 1.1537 & 0.0818\\
-		\hline
-		IPR & 4.1943 & 36.2011\\
-		\hline
-		SIPR & -14.1352 & $<$ 1e-04\\
-		\hline
-		LL & 1.1878 & 0.0006\\
-		\hline
-		MC & 0.9486 & $<$ 1e-04\\
-		\hline
-		SB & 1.4700 & 0.0635\\
-		\hline
-	\end{tabular}
-	\caption{Linear Data, $\rr$ Estimates and Standard Errors}
-\end{table}
-
-\begin{table}[H]
-
-	\centering
-	\begin{tabular}[t]{l|r|r}
-		\hline
-		Model & Estimate & Std. Dev\\
-		\hline
-		LS & 1.7691 & 0.4216\\
-		\hline
-		ReLS & 1.7691 & 0.2789\\
-		\hline
-		LMA & -0.0161 & 0.1099\\
-		\hline
-		LMAT & 1.3810 & 8.5140\\
-		\hline
-		IPR & 5.8926 & 17.3418\\
-		\hline
-		SIPR & 14.9726 & $<$ 1e-04\\
-		\hline
-		LL & 1.9553 & 0.0164\\
-		\hline
-		MC & 0.9486 & $<$ 1e-04\\
-		\hline
-		SB & 0.2586 & 0.0266\\
-		\hline
-	\end{tabular}
-	\caption{4th Order Model, $\rr$ Estimates and Standard Errors}
-\end{table}
-
-\begin{table}[H]
-
-	\centering
-	\begin{tabular}[t]{l|r|r}
-		\hline
-		Model & Estimate & Std. Dev\\
-		\hline
-		LS & 1.2740 & 0.0971\\
-		\hline
-		ReLS & 1.2738 & 0.0334\\
-		\hline
-		LMA & 0.9503 & 0.1067\\
-		\hline
-		LMAT & 1.1959 & 0.0247\\
-		\hline
-		IPR & 1.2085 & 1.0031\\
-		\hline
-		SIPR & 1.3045 & 0.2300\\
-		\hline
-		LL & 1.2585 & 0.0010\\
-		\hline
-		MC & 0.9486 & $<$ 1e-04\\
-		\hline
-		SB & 1.2588 & 0.0587\\
-		\hline
-	\end{tabular}
-	\caption{Linear SIR Data, $\rr$ Estimates and Standard Errors}
-\end{table}
-
-\section{Variance Estimation Methods}\label{sec:sim-var-res}
+The vertical line is the true $\rr$ value.}\label{fig:ar-small-I}
+\end{figure}
+
+\subsection{Variance of $\rr$ estimation methods}\label{sec:sim-var-res}
 We examine the variance methods used to estimate CIs for our estimates of $\rr$.
-From the above results, we saw that methods such as LS, ReLS, LL, and MC result in very small CIs.
+From the above results, we saw that methods such as RE, rRE, LL, and MC result in very small CIs.
 On the other hand, methods such as IPR have consistently large CIs, to the point of being uninformative.
@@
 
-\subsection{Repeated Sampling}
+\subsubsection{Delta Method and Posterior}
 
@@
 In some cases, such as LL, LMA, and LMAT, the standard deviation and standard error estimates are fairly close.
-The least squares methods, LS and ReLS, on the other hand, tend to overestimate the error by an order of magnitude.
+The least squares methods, RE and rRE, on the other hand, tend to overestimate the error by an order of magnitude.
 Finally, SIPR and SB, have individual standard error estimates that are smaller than the standard deviation of repeated estimates, meaning that they understate how much error is in the estimates.
@@
 		\hline
-		LS & 2.0000 & 0.0003 & 1.9999 & 0.0058\\
-		\hline
-		ReLS & 2.0000 & 0.0003 & 1.9997 & 0.0052\\
+		RE & 2.0000 & 0.0003 & 1.9999 & 0.0058\\
+		\hline
+		rRE & 2.0000 & 0.0003 & 1.9997 & 0.0052\\
 		\hline
@@
 	\end{tabular}
-        \caption{Mean $\rr$ Estimates and Std. Devs, Repeated Sampling}
+        \caption{Mean $\rr$ Estimates and Std. Devs, Delta Method/Posterior}
         \label{tab:rep-samp}
@@
 
-\subsection{Delta Method}
-
-\subsection{Block Bootstrap}
+\subsubsection{Block Bootstrap}
 
@@
 Only 5 of the models provide predictions of the size of the compartments and thus allow for detrending of the data.
-The least squares and reparameterized model have a fairly accurate estimates of $\rr$ as well as very small standard errors.
+The ratio estimator and reparameterized model have a fairly accurate estimates of $\rr$ as well as very small standard errors.
 On the other hand, the linear approximation and smoothed IPR models both have very inaccurate estimates as well as very large standard errors of $\rr$.
@@
 		\hline
-		LS & 1.9999 & 0.0003 & 1.9999 & 0.0058\\
-		\hline
-		ReLS & 1.9998 & 0.0003 & 1.9997 & 0.0052\\
+		RE & 1.9999 & 0.0003 & 1.9999 & 0.0058\\
+		\hline
+		rRE & 1.9998 & 0.0003 & 1.9997 & 0.0052\\
 		\hline
@@
 Figure \ref{fig:coverage} shows the proportion of repetitions for both repeated sampling and block bootstrapping in which the 95\% confidence intervals of the estimates produced by the models covers the true $\rr$.
-The least squares, reparameterized, and linear model approximation with all time points contain the true $\rr$ value in all of the confidence intervals.
+The ratio estimator, reparameterized, and linear model approximation with all time points contain the true $\rr$ value in all of the confidence intervals.
 On the other hand, the smoothed IPR, Markov chain, and sequential Bayes models never cover the true $\rr$ in any of the confidence intervals.
+
+\begin{figure}[H]
+	\begin{center}
+		\includegraphics[scale=0.5]{images/coverage2.jpeg}
+		\caption{Proportion of coverage of true $\rr$ of confidence intervals produced by repeated sampling for the applicable method (delta method or posterior).}
+		\label{fig:coverage}	
+	\end{center}
+\end{figure}
 
@@
 		\includegraphics[scale=0.5]{images/coverage.jpeg}
-		\caption{Proportion of coverage of true $\rr$ of confidence intervals produced by repeated sampling and block bootstrapping.}
+		\caption{Proportion of coverage of true $\rr$ of confidence intervals produced by repeated sampling for block bootstrapping.}
 		\label{fig:coverage}	
@@
 
-      \subsection{Posterior}
-
-      \section{Application to USA H1N1 data}\label{sec:real-data}
+      \subsubsection{Posterior}
+
+      \subsection{Application to USA H1N1 data}\label{sec:real-data}
       We apply our methods using data from the 2009 H1N1 pandemic in the United States.
 
-      \subsection{USA H1N1 Influenza 2009 Data}
+      \subsubsection{USA H1N1 Influenza 2009 data}
       The data is from the Center of Disease Control and Prevention's (CDC) FluView \citep{cdc-fluview}.
@@
   \includegraphics[width=\textwidth]{images/h1n1-sir.pdf}
-  \caption{Data from H1N1 influenza 2009 in the USA.  Weekly wILI is reported, which we transform into S, I, and R counts for epiweeks 21-33 using $\gamma^{-1} = 3/7$ weeks.}
-  \end{figure}
-
-      \subsection{Results}
+  \caption{Data from H1N1 influenza 2009 in the USA.
+Weekly wILI is reported, which we transform into S, I, and R counts shown for weeks (start of weeks: April 16-Aug.
+27)  using $\gamma^{-1} = 3/7$ weeks.
+The observations between the dotted lines denote the points used to estimate $\hat{\rr}$ in \cite{towers2009}.}
+  \end{figure}
+
+      \subsubsection{Results of H1N1 data}
       The results are shown below.  \cite{towers2009} report an estimate of $\hat{\rr} \in [1.50, 1.70]$.
@@
 		\hline
-		LS & 0.0280 & 0.0549\\
-		\hline
-		ReLS & 0.3332 & 0.0001\\
-		\hline
-		LMA & 0.9900 & 3.2589\\
-		\hline
-		LMAT & 0.9928 & 0.0276\\
-		\hline
-		IPR & 5.3078 & 1.7141\\
-		\hline
-		SIPR & -17.9646 & 5.1240\\
-		\hline
-		LL & 0.0036 & 0.0000\\
-		\hline
-		MC & 0.9453 & 0.0011\\
-		\hline
-		SB & 39.9036 & 1.7520\\
+		RE & 0.0625 & 0.0134 \\
+		\hline
+		rRE & 0.0278 & 0.0220 \\
+		\hline
+		LMA & 0.9857 & 0.3465 \\
+		\hline
+		LMAT & 0.9701 & 0.0549 \\
+		\hline
+		IPR & 5.5463 & 1.7744 \\
+		\hline
+		SIPR & -3.1438 & 4.1129 \\
+		\hline
+		LL & 0.0040 & $<$ 1e-04\\
+		\hline
+		MC & 0.9462 & 0.0008 \\
+		\hline
+		SB & 29.9850 & 1.2907 \\
 		\hline
@@
 \begin{itemize}
-  \item LS and ReLS provide the most accurate estimates of $\rr$
+  \item RE and rRE provide the most accurate estimates of $\rr$
   \item LL results in close point estimates but occasionally has too small standard errors
@@
   \item The noisier the epidemic data is, the larger our CIs for $\rr$ are
-  \item Estimations of $\rr$ are fairly robust to the misspecified data sets we presented, at least for the methods of LS, ReLS, and LL
+  \item Estimations of $\rr$ are fairly robust to the misspecified data sets we presented, at least for the methods of RE, rRE, and LL
   \item Block bootstrap matches the empirical variance through repeated sampling.

### 2018-07-03 paper/Version15/supplementary.tex
finally on git

--- 2018-07-03 paper/Version14/draft_v14.tex
@@
 
-\begin{document}
-
-%%% draw a tree
-\tikzset{
-  treenode/.style = {shape=rectangle, rounded corners,
-                     draw, align=center,
-                     top color=white, bottom color=blue!20},
-  root/.style     = {treenode, font=\Large, bottom color=red!30},
-  env/.style      = {treenode, font=\ttfamily\normalsize},
-  dummy/.style    = {circle,draw}
-}
-
-\title{\Wxxsir methods to estimate $\rr$ in the SIR model}
+\title{\Wxxsir methods to estimate $\rr$ in the SIR model: \\ Supplementary material}
 \author{ Shannon Gallagher$^{\dag}$, Andersen Chang$^{\ddag}$, and William F.
@@
 \date{\today}
+
+\begin{document}
+
 \maketitle
 
-% \tableofcontents
-
-\section{Introduction}\label{sec:intro}
-What has been called ``arguably the most important quantity in the study of epidemics'' \citep{Heesterbeek2002},  $\mathcal{R}_0$, the reproduction number (by convention pronounced ``R-naught'') is an elusive quantity for epidemic modelers to estimate.
-As defined by \citet{anderson1992}, $\rr$ is the ``the average number of secondary infections produced when one infected individual is introduced into a host population where everyone is susceptible.''  In some ways,$\rr$ summarizes an entire outbreak of a disease; it  is used to assess whether a disease epidemic will occur.
-Additionally, it describes what percentage of the population needs to be vaccinated to avoid such an epidemic, roughly $1-\rr^{-1}$ \citep{anderson1992}).
-Despite a clear definition of $\rr$, epidemiologists have struggled to create a standard  estimator for $\rr$  \citep{hethcote2000}.
-A major issue in estimating $\rr$ is that the quantity is a \textit{property of the model}, meaning that $\rr$ is dependent not only on the usual noise that comes with statistical modeling but also on a variety of assumptions on how researchers assume a disease is transmitted through a population \citep{diekmann2009}.
-
-To elaborate on $\rr$ being a property of the model, we first need to introduce the concept of the ``SI-framework,'' which was pioneered by Kermack and McKendrick in the 1920s \citep{getz2006}.
-Here, S and I are known as compartments where S stands for ``susceptible'' and I for ``infectious.'' Infectious disease models then specify how individuals move from the S compartment to the I compartment.
-Often, ancillary compartments are added such as R, which stands for recovered or dead, or E, which stands for exposed but not yet infectious.
-The estimate of $\rr$ from a SIR model will not be directly comparable to the estimate of $\rr$ from a SEIR model, for example.
-The difficulties of estimating $\rr$ are summarized and expanded upon by \cite{li2011}.
-\cite{driessche2017} overviews a number of methods used to estimate $\rr$ but looks at systems beyond the SIR model and so these methods may not be directly compared.
-As a result of $\rr$ being a property of the model, we limit our review to methods for estimating $\rr$ for the SIR model introduced by \cite{Kermack700}.
-
-Even when limiting $\rr$ estimates to those from SIR models, estimation is still difficult.
-Problems include mathematical versus statistical methods (i.e.
-solving for $\rr$ versus estimating $\rr$), the number of observations used to estimate $\rr$ (i.e.
-time), boundary cases of infection and recovery rates, population size, initial SI ratio, and restrictive assumptions placed upon the noise.
-These are all issues for producing point estimates, to say nothing of confidence or credible intervals (CI).
-
-We review \wxxsir methods for estimating $\rr$.
-We note that these methods are not exhaustive but are chosen due to their impact and use in epidemiology.
-We find that these methods also help to highlight the problems discussed above.
-
-In addition to making a point estimate of $\rr$, we are particularly interested in the hypothesis  $H_0: \rr < 1$, as a value less than 1 indicates that a disease will not outbreak.
-We also discuss techniques for estimating CIs for each of the different methods.
-These CI estimates are created using the delta method, the block bootstrap, and the posterior distribution \citep{cao1999,wasserman2004}.
-
-In the following sections, we present a series of simulations in which we examine each of our $\rr$ estimation methods.
-Within these simulations, we vary the number of observations used to estimate $\rr$, infection and recovery rates, the total population size, initial SI ratio, and several assumptions about the noise.
-After analyzing our simulations, we apply our methods to data from the USA H1N1 influenza pandemic.
-
-Finally, we discuss our recommendations for estimating $\rr$ in the SIR model, and we comment on how these recommendations extend to more complex epidemiological models beyond the SIR model and on how many of the problems with estimating $\rr$  become exacerbated in such models.
-
-The rest of this manuscript is organized as follows.
-In Section \ref{sec:r0}, we briefly discuss the origin of $\rr$.
-In Section \ref{sec:sir-intro}, we introduce the SIR model as described by \cite{Kermack700}.
-In Section \ref{sec:methods}, we overview the \wxxsir methods for estimating $\rr$.
-In Section \ref{sec:ci}, we discuss how to estimate CIs for each of the \wxxsir methods.
-In Sections \ref{sec:sim-res} - \ref{sec:sim-var-res}, we present the results of our simulations.
-In Section \ref{sec:real-data}, we evaluate the methods on real data--from the USA H1N1 influenza pandemic of 2009.
-Finally, in Section \ref{sec:discussion}, we provide recommendations of estimating $\rr$ for the SIR model and offer comments on how these conclusions extend to more complex models.
-
-\section{Origins and Difficulties of Estimating $\rr$}
-\label{sec:r0}
-
-The origins of $\rr$ are tied to the survival function, which originates from the field of demography in the late 1800s.
-The survival function describes how many female offspring a woman is expected to produce in her lifetime, literally the reproduction number \citep{dietz1993estimation}.
-In demography, we have
-\begin{align}\label{eq:surv}
-\rr = \int_0^\infty p(a) \beta(a) da
-\end{align}
-where $p(a)$ denotes the probability of a woman surviving to age $a$ and $\beta(a)$ the rate of an individual of age $a$ giving birth to a  girl.
-The concept of the reproduction number was later imported to the field of epidemiology by MacDonald and Smith \citep{dietz1993estimation}.
-Analogously in epidemiology, $p(a)$ is the age of a disease, and $\beta(a)$ is the infection rate at time $a$.
-
-Equation \ref{eq:surv} is the chronologically first, and in many ways, the most direct method to estimate $\rr$.
-The premise of this whole paper is based upon the fact that estimating $\rr$ from the survival function in Equation \eqref{eq:surv} is a difficult task.
-It is due to this difficulty that other estimates even exist and why we must treat $\rr$ as a property of the SIR model.
-
-Other difficulties in estimating $\rr$ arise from model assumptions on how randomness enters the model and sensitivity of the estimate to changes in the model parameters.
-For example, the SIR model is special in that two of its compartments, S and R are monotonic.
-The number of susceptibles is non-increasing and the number of susceptibles is non-decreasing.
-Logically, models should enforce this monotonicity of the two compartments.
-As a consequence, this restricts assumptions on how randomness enters the model.
-For example, Gaussian error, a common assumption in statistics and machine learning, becomes highly suspect due to the non-symmetrical nature of the noise imposed by the monotonic $S$ and $R$ compartments.
-Without assuming Gaussian noise, it becomes much more difficult to guarantee properties such as unbiased and consistent estimates of $\rr$.
-
-These difficulties in estimating $\rr$ have led a number of researchers to analyze the sensitivity of $\rr$ with respect to small perturbations in parameters like $\beta$ and $\gamma$ \citep{lash2003,epstein2007agent,capaldi2012}.
-Even this is a difficult task.
-Since there is no analytic solution to the SIR model, estimating the set of derivatives with respect to $\beta$ and $\gamma$ must be done numerically, which can substantially increase computation time.
-We should note that this sensitivity analysis is done before even adding any distributional assumptions on how randomness enters the model.
-
-Below in Table \ref{tab:r0-real-ex}, we provide examples of estimates of $\rr$ for specific diseases from publications over the past three decades, spanning various locations over the world.
-We display estimates for HIV, Zika (ZIK-V), Ebola (EVD), seasonal influenza, H1N1 influenza, and measles.
-These estimates of $\rr$ are not limited to those of SIR model, although most results come from this framework.
-Of these, Zika seems to have the largest estimated values of $\rr$, followed by Ebola, H1N1 influenza, seasonal influenza, and HIV.
-We see that African locations seem to have larger estimates of $\rr$ compared to the rest of the world, regardless of the disease.
-Some of the accompanying intervals (not necessarily CIs) are extremely large such as the case of HIV in Uganda (Interval: [0.12, 14.17]) whereas some are extremely small such as for Ebola in Guinea (Interval: [1.50, 1.52]).
-We see that estimates for H1N1 influenza are generally higher than that of seasonal influenza, which is evidence that H1N1 should have been treated as a more critical concern than the typical seasonal influenza cycle.
-The lowest possible estimate for $\rr$ reported is 0.12 for HIV in Uganda, and the largest is 17.00 for measles in England and Wales.
-Within diseases, we see that estimates for $\rr$ seem to be similar, although typically do not lie in one another's intervals, suggesting that there is disparity in estimates of $\rr$.
-
-% Please add the following required packages to your document preamble:
-% \usepackage{booktabs}
-% \usepackage{graphicx}
-% \usepackage{lscape}
-\begin{landscape}
-\begin{table}
-\centering
-\begin{tabular}{@{}lllrrrl@{}}
-\toprule
-\textbf{Disease}    & \textbf{Location}                & \textbf{Year(s)}                 & \textbf{$\rr$} & \textbf{Lower} & \textbf{Upper} & \textbf{Source}                                                                                                                  \\ \midrule
-HIV                 & Belgium                          & 2004-2010                     & 1.10          & 1.09           & 1.16           & \cite{coelho2011}   \\
-HIV                 & Netherlands                      & 2004-2010                     & 1.11        & 1.10            & 1.18           & \cite{coelho2011}    \\
-HIV                 & Portugal                         & 2004-2010                     & 1.08        & 1.06           & 1.15           & \cite{coelho2011}   \\
-HIV                 & San Francisco, USA               & 1993                          & 5.00           & 2.00              & 8.00              & \cite{blower1994}    \\
-HIV                 & Uganda                           & 2002-2008                     & 3.31        & 0.12           & 14.17          & \cite{nsubuga2014}    \\ \hline
-ZIK-V                & French Polynesia                 & 2013                          & 2.75        & 2.53           & 2.98           & \cite{zhang2017}        \\
-ZIK-V                & French Polynesia                 & 2013-2014                     & 3.70         & 2.60            & 4.80            & \cite{kucharski2016}  \\
-ZIK-V                & Florida                          & 2016                          & 0.16        & 0.13           & 0.19           & \cite{dinh2016}       \\
-ZIK-V                & Colombia                         & 2015-2016                     & 2.50         & 3.00              & 3.60            & \cite{nishiura2016}    \\
-ZIK-V                & Barranquilla, Colombia           & 2015                          & 2.40         & 3.80            & 5.60            & \cite{towers2016}    \\
-ZIK-V                & Colombia                         & 2015                          & 1.42        & 2.56           & 3.83           & \cite{majumder2016}  \\ \hline
-EVD               & Western Africa                   & 2014                          & 1.78        & 1.60            & 2.00              & \cite{fisman2014}   \\
-EVD & Western Africa                   & 2014                          & 1.60         & 1.40            & 1.80            & \cite{towers2014}  \\
-EVD                 & Democratic Republic of the Congo & 1995                          & 2.70         & 1.90            & 2.80            & \cite{legrand2007}  \\
-EVD                 & Uganda                           & 2000                          & 2.70         & 2.50            & 4.10            & \cite{legrand2007}  \\
-EVD                 & Guinea                           & 2014                          & 1.51        & 1.50            & 1.52           & \cite{althaus2014}   \\
-EVD                 & Sierra Leone                     & 2014                          & 2.53        & 2.41           & 2.67           & \cite{althaus2014}      \\
-EVD                 & Liberia                          & 2014                          & 1.59        & 1.57           & 1.60            & \cite{althaus2014}    \\ \hline
-Seasonal Influenza  & USA, France, Australia           & 1972-2002 & 1.30         & 1.20            & 1.40            & \cite{chowell2008}  \\
-Seasonal Influenza  & French Territory                 & 1987-1995                     & 1.50         & 1.09           & 1.73           & \cite{bonabeau1998}  \\ \hline
-H1N1 Influenza      & Thailand                         & 2009                          & 2.07        & 1.92           & 2.22           & \cite{desilva2009}     \\
-  H1N1 Influenza     & Peru                             & 2009                          & 1.37        & 1.20            & 1.70            & \cite{desilva2009}        \\ 
-  H1N1 Influenza & USA & 2009 & 1.50 & 1.30 & 1.70 &\cite{towers2009} \\ \hline
-Measles      & Cirencester, England                 & 1947-1950                          & 13.50   & 13.00           & 14.00            & \cite{fraser2009}     \\ 
-  Measles     & England and Wales                  & 1950-1968                          & 17.00      & 16.00          & 18.00         & \cite{anderson1992}     \\
-  Measles     & Kansas, USA                  & 1918-1921                          & 5.50    & 5.00          & 6.00          & \cite{anderson1992}     \\
-  Measles     & Ontario, Canada                  & 1912-1913                         & 11.50        & 11.00           & 12.00            & \cite{anderson1992}     \\
-  Measles     & Willesden, England                  & 1912-1913                          & 11.00        & 12.00       & 1.60            & \cite{anderson1992}     \\
-  Measles     & Ghana                 & 1960-1968                          & 14.50       & 14.00          & 15.00            & \cite{anderson1992}     \\
-  Measles     & Eastern Nigeria                  & 1960-1968                          & 16.50        & 16.00          & 17.00           & \cite{anderson1992}     \\
-\bottomrule
-\end{tabular}
-\caption{Estimates of $\rr$ from various researchers for the infectious diseases of HIV, (Zika) ZIK-V, Ebola (EVD), seasonal influenza, and H1N1 pandemic influenza ranging from the years 1972-2016 in various locations.
-We also report the lower and upper bounds reported by the researchers.}
-\label{tab:r0-real-ex}
-\end{table}
-\end{landscape}
-
-\section{Kermack and McKendrick's SIR Model}
-\label{sec:sir-intro}
-
-The SIR model introduced by \cite{Kermack700} is a compartment model, where individuals move from susceptible, to infectious, and finally recovered states.  We make five essential assumptions,
-\begin{enumerate}
-\item The compartments are discrete and have no overlap.
-\item The transition of objects into and out of compartments is described by a set of known equations, possibly dependent on unknown parameters.
-\item The populations mix homogeneously.
-\item The number of objects in each compartment at time $t=0$ is known.
-  \item The law of mass action is followed.
-  \end{enumerate}
-
-The law of mass action is a property borrowed from chemistry which says that the mass of the product per unit time is proportional to the mass of the reactants \citep{lotka1920}.
-In epidemiology, this means that the proportion of new infections per unit time is proportional to the current  number of susceptible \citep{anderson1992}.
-
-In this particular  SIR model (displayed in Figure \ref{fig::sir}), the number of individuals ($N$) is constant.
-Adaptations of the SIR model can include birth and death rates, which may correspondingly change the derivation of $\rr$ (for further discussion, see Section \ref{sec:discussion}).
-The parameters  $\beta$ and $\gamma$ have epidemiological meaning.
-Here, $\beta$ is the average infection rate and $\gamma$ is the average recovery rate.
-The movement of individuals from one compartment to another is represented through the ordinary differential equations below.
-Both $\beta$ and $\gamma$ are constrained to be between $0$ and $1$.
-For the remainder of this paper, we use $X$, $Y$, and $Z$ to denote the number of individuals in the $S$, $I$, and $R$ compartments, respectively, to avoid any confusion with $\rr$.
-\begin{align}
-\systeme{\frac{dX}{dt} = -\frac{\beta XY}{N}, \frac{dY}{dt} = \frac{\beta XY}{N} - \gamma Y, \frac{dZ}{dt} = \gamma Y}. \label{eq:sir}
-\end{align}
-In words, susceptible individuals become infected at a rate that is proportional to the percentage of infected individuals multiplied by $\beta$, the infection rate, and the number of susceptible individuals.
-Infectious individuals recover at a rate of $\gamma$ multiplied by the number of infected individuals.
-
-The SIR model is represented graphically in Figure \ref{fig::sir}. 
-\begin{figure}[h]
-\centering
-\begin{tikzpicture}[node distance=3cm,auto,>=latex',every node/.append style={align=center}]
-    \node [int,  fill = white!70!blue] (a)              {$X$};
-    \node [int,  fill = white!70!blue]           (c) [right of=a] {$Y$};
-    \node [int,  fill = white!70!blue] (e) [right of=c] {$Z$};
-    \path[->, auto=false] (a) edge node {} (c)
-                          (c) edge node {} (e) ;
-
-\end{tikzpicture}
-\caption{Depiction of a SIR model where $X=S, Y=I,$ and $Z=R$.  One can only get the disease once in this model.}\label{fig::sir}
-\end{figure}
-An outbreak occurs if the rate of change of infectious individuals is positive, $\frac{dY}{dt} > 0$, or equivalently,
-\begin{align*}
-  \frac{dY}{dt} &> 0 \\
-  \implies -\frac{dX}{dt} - \frac{dZ}{dt} &> 0 \\
-  \implies \frac{\beta X Y}{N}  - \gamma Y &> 0 ,\\
-\implies  Y \left ( \beta \frac{X}{N} - \gamma \right ) & > 0\\
-\implies   \frac{\beta}{\gamma} &> \frac{N}{X}.
-\end{align*}
-That is,  the rate of new infections is greater than the rate of recovery.
-So as long as the number of initially susceptible individuals is large compared to the total population, $\frac{X}{N} \approx 1$, then an outbreak will occur if $\rr >1$,
-\begin{align*}
-  \rr \overset{def}{=} \frac{\beta}{\gamma}.
-  \end{align*}
-  To incorporate randomness into the model, we add noise, namely,
-  \begin{align}\label{eq:sir-noise}
-    X_{obs}(t) &= X(t) + \epsilon_{X,t}\\
-    Y_{obs}(t) &=  N - X_{obs}(t) -Y_{obs}(t)  \nonumber\\
-    Z_{obs}(t) &= Z(t) - \epsilon_{Z,t}. \nonumber
-  \end{align}
-We are assuming the observations are generated based on the discretizied time, deterministic ODEs presented in Equation \ref{eq:sir} with the addition of time ($t=0, 1, \dots, T$) and compartment dependent noise $\epsilon_{X,t}$ and $\epsilon_{Z,t}$.
-Since, $N$, the total population, is constant, then $Y_{obs}$ is adjusted accordingly.
-
-In summary, when we discuss estimators for $\rr$ for the SIR model, we mean to say we are forming an estimator of $\rr$ from the given set of data in Eq. \ref{eq:sir-noise},
-\begin{align*}
-  \textnormal{Data} &= \left \{\left (X_{obs}(t), Y_{obs}(t), Z_{obs}(t) \right ) : t=0, 1, \dots, T\right \}, \\
-  \hat{\rr} &= m(\textnormal{Data}),
-\end{align*}
-where $m$ is a function of the data.
-
-Good properties of an estimator may include (1) being a consistent estimator
-\begin{align*}
-  \hat{\rr} \overset{P}{\to} \rr \textnormal{ as } T\to \infty,
-\end{align*}
-where `$\overset{P}{\to}$' stands for convergence in probability \citep{wasserman2004} or (2) converging to a known probability distribution (convergence in distribution `$\overset{d}{\to}$' \citep{wasserman2004}),
-\begin{align*}
-\hat{\rr} \overset{d}{\to} F.  
-\end{align*}
-As we are interested in the hypothesis
-\begin{align*}
-  H_0:\;& \rr < 1 \\
-  H_A:\;& \rr \ge 1
-\end{align*}
-Then for some $\alpha \in [0,1]$, we would like to know for some $\hat{a}, \hat{b} >0$
-\begin{align*}
-P(\hat{a} \le \rr \le \hat{b}) \le 1 - \alpha,  
-\end{align*}
-the second property is particularly desirable, as we can estimate $\hat{a}$ and $\hat{b}$ from the distribution $F$.
-At any rate, we would like to be able to estimate $\hat{a}$ and $\hat{b}$ for a given $\alpha$-level.
-We could also make similar hypothesis tests such as that the $\rr$ for H1N1 influenza is greater than the reported $\rr$ for seasonal influenza.
-
-\subsection{$\rr$ is a property of the model}
-
-We provide an example of how $\rr$ is a property of the model.
-Consider the situation of data generated from an SEIR (XEYZ) model (described by \cite{cintronarias2009}), generated from the following ODEs,
-
-\begin{align*}
-  \frac{dX}{dt} &= - \frac{\beta XY}{N} \\
-  \frac{dE}{dt} &= \frac{\beta XY}{N}  - \alpha E\\
-  \frac{dY}{dt} &= \alpha E - \gamma Y \\
-  \frac{dZ}{dt} &= \gamma Y,
-\end{align*}
-again, with a constant population of $N$ and with known initival values $(X(0), E(0), Y(0), Z(t))$.
-The parameter $\alpha$ is the rate at which latent exposed individuals become infectious.
-As $\alpha \to \infty$, the XEYZ model converges to the XYZ model.
-The reproduction number for the XEYZ model, is the same as in the XYZ model, that is $\rr = \frac{\beta}{\gamma}$.
-
-Consider the case when $N=10000, (\beta, \gamma) = (.06, .03)$, $(X(0), E(0), Y(0), Z(t)) = (9500, 0, 500, 0)$ for times $t=1, \dots, 365$.
-Let the parameter $\alpha$ range between [.005, 1].
-This generates data set
-$$D_{\alpha, \textnormal{XEYZ}} = \left \{(X_{XEYZ}(t), E_{XEYZ}(t), Y_{XEYZ}(t), Z_{XEYZ}(t): t=1, \dots, 365\right \},$$
-which is observed by researcher 1.
-Now assume that researcher 2 misspecifies the model for this data as a XYZ model rather than a XEYZ model.
-That is, she places all exposed individuals in the X compartment to have data set
-$$D_{\alpha, XYZ} = \left \{X_{XYZ}(t) = X_{XEYZ} + E_{XEYZ}, Y_{XYZ} = Y_{XEYZ}, Z_{XEYZ} = Z_{XYZ} : t = 1, \dots, 365 \right \}.$$
-This situation is illustrated in Figure \ref{fig:sir-vs-seir} with $\alpha = 0.01$.
-
-\begin{figure}
-  \centering
-  \includegraphics[width=\textwidth]{images/seir-sir-data.pdf}
-  \caption{Top: XEYZ generated data over the course of a year.  Bottom: data tranformed from the XEYZ model to the XYZ model by combining all X and E compartments.}\label{fig:sir-vs-seir}
-  \end{figure}
-
-Both researchers want to estimate $\rr$ and so fit models to their data sets.
-The details of how the model are not so important at the moment (but will be in the later sections), but to fairly compare assume both researchers use least squares to fit the models.
-That is researcher 1 estimates $\rr$ with
-\begin{align*}
-  (\hat{\beta}_{XEYZ}, \hat{\gamma}_{XEYZ}, \hat{\alpha}_{XEYZ} )&=\arg \min_{\beta, \gamma, \alpha} \sum_{t} \left [ \left (X_{XEYZ}(t) - X(t)\right )^2  \right . \\
-                                                                 &+\left ( E_{XEYZ}(t) - E(t) \right )^2  \\
-  &\left . + \left ( Z_{XEYZ}(t) - Z(t) \right )^2\right ]
-\end{align*}
-Then the least squares (LS) estimate for $\rr^{XYZ}$ is given by Equation \ref{eq:sirls},
-\begin{align*}
-  \hat{\rr}^{XEYZ}= \frac{\hat{\beta}_{XEYZ}}{\hat{\gamma}_{XEYZ}}.
-\end{align*}
-Researcher 2 estimates $\rr$ with 
-\begin{align*}
-(\hat{\beta}_{XYZ}, \hat{\gamma}_{XYZ} )&=\arg \min_{\beta, \gamma} \sum_{t} \left [ \left (X_{XYZ}(t) - X(t)\right )^2 + \left ( Z_{XYZ}(t) - Z(t) \right )^2 \right ].
-\end{align*}
-Then the least squares (LS) estimate for $\rr^{XYZ}$ is given by Equation \ref{eq:sirls},
-\begin{align*}
-  \hat{\rr}^{XYZ}= \frac{\hat{\beta}_{XYZ}}{\hat{\gamma}_{XYZ}}.
-\end{align*}
-
-We plot the two estimates of $\rr$ as a function of $\alpha$ in Figure \ref{fig:r0-alpha}.
-We should note we also estimated the standard errors of the estimates of $\rr$, but they do not show up on the figure because they are of an order of magnitude of $10^{-4}$.
-Unsurprisingly, under the SEIR model speficiation, $\hat{\rr}=2$ because there is no noise in the data, and the data is generated from an SEIR model.
-However, the SIR model for small values of $\alpha$ can result in double the size of $\rr$!
-
-For those
-
-\begin{figure}
-  \centering
-  \includegraphics[width=\textwidth]{images/r0-est.pdf}
-  \caption{Estimates of $\rr$ from both the XEYZ and XYZ model specification for different values of $\alpha$.}\label{fig:r0-alpha}
-  \end{figure}
-
-\section{Overview of \wxxsir methods to estimate $\rr$ for the SIR model}
-\label{sec:methods}
-
-In this section, we describe the \wxxsir methods in detail, including advantages and disadvantages of each.
-
-\subsection{Least Squares ($\beta$, $\gamma$) (LS)}\label{least-squares-beta-gamma}
-Our first approach to estimate $\rr$ in the SIR model is to minimize the joint mean square error for the data collected at each time point and use the plug-in estimator found in Equation \ref{eq:sirls}.
-In particular, we find
-
-\begin{align*}
-(\hat{\beta}, \hat{\gamma} )&=\arg \min_{\beta, \gamma} \sum_{t} \left [ \left (X_{obs}(t) - X(t)\right )^2 + \left ( Z_{obs}(t) - Z(t) \right )^2 \right ]
-\end{align*}
-Then the least squares (LS) estimate for $\rr$ is given by Equation \ref{eq:sirls},
-\begin{align}\label{eq:sirls}
-  \hat{\rr}= \frac{\hat{\beta}}{\hat{\gamma}}.
-\end{align}
-
-The estimates from this method are often found via a grid search of $\beta$ and $\gamma$ values to minimize the squared error.
-The estimates can also be found using more sophisticated optimization algorithms.
-Since we typically cannot explicitly write down the partial derivatives of $S$, $I$, and $R$ with respect to $\beta$ and $\gamma$, we can only use minimization algorithms which do not rely on an explicit gradient such as Nelder-Mead simplex minimization \citep{nelder-mead1965}.
-
-A reasonable question one may ask is why use the $L_2$-norm as opposed to $L_1$ or some other similarity score.
-A possible answer is that if we were to assume Gaussian noise, then the $L_2$-norm would be equivalent to the maximum likelihood estimation.
-Another answer is that $L_2$ is a continuously differentiable function and hence is easier to compute, which allows for sensitivity analysis to  be conducted more easily.
-Finally, LS can be done without writing down explicit assumptions about the noise within the model.
-However, we cannot guarantee properties of consistency or convergence to a known distribution without more explicit assumptions.
-
-On the other hand, least squares is easy to implement using software such as \texttt{R}'s \texttt{optim()} function, and as we see in Section \ref{sec:results} produces results comparable methods to other methods.
-Least squares and similar variants have been used to estimate $\rr$ in \cite{majumder2016}.
-
-\subsection{Reparametrized Least Squares ($\rr$, $\gamma$) (ReLS)}\label{reparametrized-least-squares-rux5f0-gamma}
-
-In the previous method (Section \ref{least-squares-beta-gamma}), we estimated $\beta$ and $\gamma$ and then estimated $\rr$.
-However, it is possible to directly estimate $\rr$ if we reparametrize the ODEs in Equation \eqref{eq:sir} directly with \(\rr\) and \(\gamma\), using the relation $\rr = \frac{\beta}{\gamma}$; this yields
-
-\begin{align*}
-  \left \{
-  \begin{array}{cl}
-    \frac{dX}{dt} &= - \rr \gamma Y \frac{X}{N}\vspace{.5em}\\
-    \frac{dY}{dt} &=  \rr \gamma Y \frac{X}{N}  - \gamma Y\vspace{.5em} \\
-    \frac{dZ}{dt} &=  - \gamma Y 
-  \end{array}
-  \right . .
-  \end{align*}
-We find
-\begin{align*}
-(\hat{\rr}, \hat{\gamma} ) &= \text{argmin}_{\rr, \gamma} \sum_{t} \left [ \left (X_{obs}(t) - X(t)\right )^2 + \left ( Z_{obs}(t) - Z(t) \right )^2 \right ]
-\end{align*}
-We use the $\hat{\rr}$ directly from the above estimation problem, which again can be found with a grid search or another optimization process.
-It may be surprising that this method leads to different results than simply using least squares.
-One reason we obtain different results may be attributed to better numerical precision as there is one fewer step to obtain an estimate for $\rr$.
-However, we also see that this method further results in different standard errors than in LS, which we discuss further in Section \ref{sec:results}.
-
-Like LS, while we have no explicit parametric assumptions on the noise, we also are uncertain of any theoretical properties of our resulting estimator.
-Likewise, the same difficulties in sensitivity analysis arise as in LS.
-
-\subsection{Linear Model Approximation (LMA)}\label{linear-model-approximation-degree-10}
-
-The SIR ODEs in Eq.
-\ref{eq:sir} have no known closed form solution, and so we use numerical integration to obtain approximate solutions.
-In addition to this, data collected from real diseases are typically very noisy to begin with.
-\cite{chang2017} showed that the SIR model may be well approximated by a linear model.
-We use this method to estimate $\rr$.
-
-Specifically, we estimate two polynomials in \(t\) with degree $K$  to \(X_{obs}\)
-and \(Z_{obs}\) using least squares to find the coefficients $\{(\hat{x}_k,
-\hat{z}_k)\}_{k=1, \dots, K}$,
-\begin{align*}
-\hat{X}(t) &= \sum_{k=0}^K \hat{x}_k t^k\\
-{\hat{Z}}(t) &= \sum_{k=0}^K \hat{z}_k t^k
-\end{align*}
-Then, we estimate the derivatives as
-\begin{align*}
-\hat{X}^\prime(t) &= \sum_{k=1}^K k \hat{x}_k t^{k-1}\\
-\hat{Z}^\prime(t) &= \sum_{k=0}^K k \hat{z}_k t^{k-1}
-\end{align*}
-Following,  an estimator for \(\rr\) is derived from the ODEs in Equation \eqref{eq:sir},
-\begin{align}
-  - \frac{X^\prime}{Z^\prime}&= \rr \frac{X}{N} \nonumber\\
-  \rr &=       -\frac{X^\prime}{
-        Z^\prime} \cdot \frac{N}{X} \nonumber\\
-  \hat{\rr} &= -\frac{\hat{X}^\prime(0)}{ \hat{Z}^\prime(0)} \cdot \frac{N}{\hat{X}(0)}. \nonumber
-  \end{align}
-  Here, $K$ is arbitrary and should be selected using some criterion such as AIC.
-Besides optionally deciding on the degree of polynomials to fit, this model is simple to implement and gives comparable results to using least squares with the SIR model.
-The time $t=0$ is used to best capture the initial outbreak.
-
-  An advantage of using this estimation method is that it is simple to implement using any linear modelling software.
-On the other hand, we assume $X(t) = \hat{X}(t)$ and so if the data truly follows the SIR model, then we will likely have a biased result for $\rr$.
-Using this method, we are able to check how sensitive $\rr$ is to the degree of the fitted polynomial.
-
-\subsection{Linear Model Approximation, All Time Points (LMAT)}\label{linear-model-approximation-all-time-points-degree-10}
-
-The above formulation (Section \ref{linear-model-approximation-degree-10}) of a linear model approximation only uses the estimate at time $t=0$ to estimate $\rr$.
-We can instead, use all time points $O$ available to estimate $\rr$.
-We fit a polynomial in \(t\) with degree \(K\) to \(X_{obs}\)
-and \(Z_{obs}\) as above, with a slight modification in how we estimate
-\(\rr\),
-\begin{align*}
-  \hat{\rr} &= \frac{1}{O} \sum_t \frac{-\hat{X}^\prime(t)}{\hat{Z}^\prime(t)} \cdot \frac{N}{X(0)} 
-\end{align*}
-The intuition is that $\frac{-X^\prime(t)}{Z^\prime(t)}$ is constant in $t$, but due to our approximations with the linear model, this is no longer the case.
-Here, we average over the different possible values of $\rr$, estimated at different times.
-An advantage to this approach is that we have a more robust estimate of $\rr$ than just using one time point.
-Like in LMA, we can examine how sensitive $\rr$ is to the order of the polynomial $K$.
-
-\subsection{Incidence to Prevalence Ratio (IPR)}\label{incidence-to-prevalence-ratio}
-The incidence to prevalence ratio (IPR), described by \cite{Nishiura2009}, is another intuitive method to estimate $\rr$.
-It incorporates some of the most basic epidemiological quantities: incidence and prevalence.
-
-In terms of data from the SIR model, the incidence $J(t) \approx -(X(t+1) - X(t))$, and the IPR$(t) = \frac{J(t)}{Y(t)}$.
-This method assumes that we have some prior knowledge about $\gamma$, the recovery rate.
-Thus we use as our estimate,
-\begin{align*}
-\hat{\rr} &= \textnormal{IPR}(t) \cdot \frac{1}{\gamma}
-\end{align*}
-
-Here we assume that the time step is small enough to approximate the incidence.
-The advantage of this method is that incidence data is generally readily available as is prevalence data for certain diseases such as HIV.
-However, as one is required to have prior knowledge about $\gamma$, it may be easier to directly estimate $\rr$ with one of the many other methods described that does not require  prior knowledge.
-Again, we are using only one time point to estimate $\rr$.
-This model is agnostic to assumptions on the noise, but again, it is unclear if this is an unbiased or consistent estimator of $\rr$.
-
-\subsection{Smoothed Incidence to Prevalence Ratio (SIPR)}
-We use the same method as above, IPR, but first estimate splines with $K$ degrees of freedom, where $K$ is small, $\hat{X}(t)$ and $\hat{Y}(t)$, to fit to $X_{obs}$ and $Y_{obs}$, respectively.
-Then  $J(t) \approx -(\hat{X}(t+1) - \hat{X}(t))$, and the $\hat{\textnormal{IPR}}(t) = \frac{J(t)}{\hat{Y}(t)}$.
-Then
-\begin{align*}
-\rr &= \hat{\textnormal{IPR}}(t) \cdot \frac{1}{\gamma}
-\end{align*}
-Splines are used because they typically can be used to make good approximations of epidemic curves (CITE).
-However, the researcher could use their method of choice to estimate $\hat{X}(t)$ and $\hat{Y}(t)$.
-An advantage of this method is that it creates a less variable estimate than estimating IPR using only one point.
-It has the same disadvantages as the usual IPR ratio in that it requires knowledge about $\gamma$.
-
-%%%%%%%%%%%%%%%%%%%%%
-\subsection{Log-Linear (LL)}
-\cite{harko2014exact} were able to reduce the SIR model to one ODE.  From this, we can derive the following,
-\begin{align}
-  X(t) &=  X(0) e^{\frac{-\beta}{\gamma}\frac{Z(t)}{N}} \nonumber\\
-  \log \frac{X(t)}{X(0)} &=  \frac{-\beta }{\gamma}\frac{Z(t)}{N} \nonumber\\
-  \log \frac{X(t)}{X(0)} &=  -\rr \frac{Z(t)}{N}. \label{eq:harko_lin}
-\end{align}
-
-We can regress the left hand side in Eq.
-\ref{eq:harko_lin} on $Z(t)$ (using least squares), the number of recovered individuals, with $\rr$ as the coefficient of $Z(t)/N$ and with an intercept term of $\beta_0=0$ to obtain an estimate of $\rr$,
-\begin{align*}
-  \hat{\rr} = -\frac{\sum_{t=0}^T \log \frac{ X(t)}{X(0)}}{\sum_{t=0}^T\frac{Z(t)}{N}}.
-\end{align*}
-This method says that for a one percent increase in number of recovered individuals, we expect the ratio of the number of previous susceptibles to number of new susceptibles to increase by $e^{\rr}$,
-\begin{align*}
-  \log \left ( X(t+1)/ X(0) \right ) - \log \left ( X(t)/X(0) \right ) &= - .01\rr\\
-  \log \left ( X(t+1) \right ) - \log \left ( X(t) \right )  &=- .01\rr\\
-  \log \left ( X(t) / X(t+1) \right ) &= .01\rr\\
-  \frac{X(t)}{X(t+1)}  &\approx e^{\rr}.
-\end{align*}
-
-The advantages of this method are many.
-One, the estimate of $\rr$ is highly interpretable.
-Two, an assumption of Gaussian noise in Equation \ref{eq:harko_lin} is not as strong as an assumption in the other methods to estimate $\rr$ because we are looking at a log transformation of the the percent of suceptibles instead of raw variables.
-If the assumption holds, we can show that the resulting estimate of $\rr$ is an unbiased and a consistent estimator.
-Finally,  this method is easily implemented using any linear modeling software.
-
-\subsection{Markov chain estimation (MC)}
-A natural approach to epidemic modelling is that of Markov chains (MC), since it is assumed an individual's next state is only dependent on its current state and the current states of other individuals.
-Much work has been done over the years in this specific field including asymptotic behavior, continuous time MC, confidence intervals, and more \citep{jacquez1991,gani1995,daley2001epidemic}.
-We present one simple instantiation of the model, the discrete time case, which traces its origin back to the Reed-Frost model \citep{abbey1952}.
-
-In this method, the number of susceptibles at the next step, $X_{obs}(t+1)$, has a Binomial distribution based on the contacts with the current number of infectious, $Y(t)$ and the current number of susceptibles.
-That is $X_{obs}(t+1) \sim \text{Binomial}\left(X_{obs}(t), \alpha^{Y_{obs}(t)}\right)$, where $\alpha$ is the probability of avoiding infection from an infective.
-One can write out the likelihood for $\alpha$ in this model, namely,
-\begin{align*}
-\mathcal{L}\left ( \alpha ; \textnormal{data}\right ) \propto \prod_{t=0}^{T-1}\left ( \alpha^{Y_{obs}(t)} \right )^{X_{obs}(t+1)} \left (1- \alpha^{Y_{obs}(t)} \right )^{X_{obs}(t) - X_{obs}(t+1)}.
-\end{align*}
-Maximizing the likelihood yields an estimate for $\alpha$,
-\begin{align*}
-\hat{\alpha} = \arg \max_{\alpha} \mathcal{L}\left(\alpha; \textnormal{data} \right ).
-  \end{align*}
- \cite{barbour2004} report that the estimated reproduction number is thus,
-\begin{align}\label{eq:r0-mc}
-\hat{\rr} &= \log \left ( \frac{1}{1-\hat{\alpha}}\right ).
-\end{align}
-
-This method typically allows for more than just the reproduction number to be estimated.
-Through recursion, one can estimate the probability of having a given number of susceptibles and infected at each time step, and hence the entire probability distribution may be known.
-
-The advantages of this model are its simplicity of interpretation and ability to generate an estimate for the probability distrubtion for $(X(t), Y(t))$.
-We have explicitly stated how randomness enters the model so it is up to the researcher to check whether this assumption is a good fit.
-Using this method, it is possible to simulate how sensitive $\hat{\rr}$ is to changes in $\hat{\alpha}$.
-
-\subsection{Sequential Bayes (SB)}\label{sec:seqbayes}
-
-Described by \cite{bettencourt2008} and summarized in \cite{obadia2012r0}, the sequential Bayes method is a Bayesian approach to an approximation of the classic SIR model.
-The approximated SIR model assumes that the incidence at $t+1$, $J(t+1)$ has a Poisson distribution, with $\gamma$ as the  average inverse of the infectious period.
-In order to estimate $\rr$, we must have some idea about $\gamma$,
-\begin{align*}
-J(t+1) | \rr  \sim \textnormal{Poisson}( J(t) \exp \left \{  \gamma (\rr-1)\right \})
-\end{align*}
-Then, the posterior distribution of $\rr$ given the previous days' incidence is
-\begin{align*}
-  P(\rr | J_0, \dots, J_{t+1}) = \frac{P(J_{t+1} | \rr, J_0, \dots, J_t)P(\rr| J_0, \dots, J_t)}{P(J_0, \dots, J_{t+1})}.
-\end{align*}
-This method is sequential in that the prior distribution for $\rr$ comes from the previous day.
-The initial prior for $\rr$ is assumed to be flat.
-This method results in a posterior distribution from which credible intervals may be obtained.
-This method assumes, initial growth in incidence to be exponential, and homogeneous mixing of populations as with any compartment model.
-The advantages of this method are that of the ability to obtain an entire posterior distribution, whereas many other methods are difficult to even find an estimate of the variance.
-Disadvantages include strict assumptions about the distributions and computational time required to estimate the relevant parameters.
-
-Although, the Poisson distribution shown above is often used to estimate $\rr$, we have found that this model does not perform well in practice (see Section \ref{sec:results}).
-Thus, we would recommend using a more sophisitcated Bayesian model such as the one in \cite{bauer2017}.
-
-\section{Confidence and Credible Intervals}
-\label{sec:ci}
-
-Estimating $\rr$ is difficult and estimating $V[\rr]$, the variance and confidence or credible intervals (CI) is even more so.
-We describe general methods which may be applicable to estimate the variance.
-In particular, we describe three methods 1) the delta method, 2) the block bootstrap, and 3) posterior distributions.
-
-\subsection{Delta Method}\label{delta-method}
-
-When the method estimates \(\beta\) and \(\gamma\) instead of \(\rr\) directly, we use the delta method approximation to estimate the
-variance of \(\rr\).   From \cite{wasserman2004}, if
-\begin{align*}
-  \sqrt{O} \left ( \hat{\mathbf{\theta}}_o - \mathbf{\theta} \right ) \overset{d}{\to} N\left ( \mathbf{0}, \Sigma \right  )
-\end{align*}
-and $h$ is a continuously differentiable function then
-\begin{align*}
-  \sqrt{O} \left (h(\hat{\theta}_n) - h(\theta) \right ) \overset{d}{\to} N \left (0,  \nabla^T h(\theta)\Sigma \nabla h(\theta)\right ),
-\end{align*}
-where $\nabla h$ is the gradient of $h$.
-
-For the method LS, we know $\rr = h(\beta, \gamma) = \frac{\beta}{\gamma}$.
-Then, if $E\left [\frac{\hat{\beta}}{\hat{\gamma}}\right] \to \frac{\beta}{\gamma}$, then  $(\nabla h = (\frac{1}{\gamma},  -\frac{\beta}{\gamma^2})^T)$ and
-\begin{align*}
-  V[\rr] &\to \nabla h^T(\beta, \gamma) \Sigma_{\beta, \gamma} \nabla h(\beta, \gamma) \\
-  &= \frac{\Sigma_{\beta, \beta}}{\gamma^2} - \frac{\Sigma_{\beta, \gamma}(1 + \beta)}{\gamma^3} + \frac{\beta \Sigma_{\gamma, \gamma}}{\gamma^4}
-\end{align*}
-where $\Sigma_{\beta, \gamma}$ is covariance matrix of \(\beta\) and \(\gamma\).
-.
-We can use plug-in estimators for $\hat{\beta}$, $\hat{\gamma}$, and $\hat{\Sigma}$ where $\hat{\Sigma}$ is taken to the inverse of the Hessian from the least squares optimzation of $\beta$ and $\gamma$.
-
-For ReLS, we use the estimate of $\Sigma_{\rr, \gamma}$ taken from the optmization of $\rr$ and $\gamma$ as to find our estimate for $\hat{\Sigma}_{\rr}$.
-
-This estimate assumes that the distribution of $\beta$ and $\gamma$ are asymptotically normal.
-Here, we use the relationship of $\rr = \frac{\beta}{\gamma}$ in order to estimate the variance, which is specific to the SIR model.
-This method may be extended to other frameworks such as the SEIR model.
-The advantages of the delta method are that since the Next Generation Matrix \citep{diekmann2009} gives recipe for deriving $\rr$ from the deterministic ODEs, then it is theoretically possible to use the delta method to derive an expression for the variance of $\rr$ for any compartment model.
-
-We can also use the Delta method to derive variance estimates of LMA, LMAT, LL, and MC as for all these estimates first estimate parameters $\theta$ since $\rr= h(\theta)$
-
-\subsection{Block Bootstrap}
-
-The block bootstrap is a variant of the bootstrap in which the $n$ observations are dependent on one another.
-In contrast to the original bootstrap, the block bootstrap partitions consectutive observations into $k$ blocks with block-length $b$ ($n=kb$).
-These are non-overlapping partitions or blocks, although there are variants for which blocks can overlap.
-For iteration $\ell$, one samples $k$ blocks uniformly with replacement and estimates $\hat{\rr}_b$, repeating the sampling and estimation for a total of $L$ estimates of $\hat{\rr}$.
-Then
-\begin{align*}
-  V\left [ \hat{\rr} \right ] &\approx V\left [\hat{\rr}_\ell \right ].
-\end{align*}
-More about the asympotic distributions of the blockwise-bootstrap can be found in \cite{cao1999}, along with descriptions of variants of the block bootstrap.
-
-The asympotic properties hold when the stochastic process is $m$-dependent, that is when observations $t, t+1, \dots$ and $s, s+1, \dots$ are independent whenever $s+m < t$ and under some smoothness conditions on the statistic used to estimate $\rr$.
-
-A choice one must make in the block bootstrap is the block size $b$.  One simple recommendation is to take $b= n^{1/3}$, which is derived from the asymptotic mean square error of the method.
-
-\subsection{Posterior Distribution}
-In this setting, we assume $\rr$ is a random variable and the data is fixed.
-Following the notation of \cite{wasserman2004}, then the posterior distribution $f(\rr | \textnormal{data})$ is proportional to the likelihood of observing the data given $\rr$, $\mathcal{L}(\textnormal{data}|\rr)$, multiplied by our prior on $\rr$, $f(\rr)$.
-\begin{align*}
-f(\rr | \textnormal{data} ) \propto \mathcal{L}(\textnormal{data} | \rr) f(\rr)
-\end{align*}
- Typically, the posterior is either estimated using conjugate distributions for the likelihood and prior or through Markov Chain Monte Carlo (MCMC) to simulate a posterior distribution $\hat{F}$ such that
-\begin{align*}
-\hat{F} \overset{d}{\to} f(\rr| \textnormal{data}).
-\end{align*}
-Then commonly, $\hat{\rr} = \hat{F}_{.5}$, the median of $\hat{F}$.
-Correspondingly, the 95\% credible interval is $\left[\hat{F}_{.05}, \hat{F}_{.95} \right ]$.
-Advantages of this method include having an entire distribution instead of a point estimate.
-A disadvantage is that using conjugate priors may not be justified for the given data and MCMC may be computationally intractable.
-One should note that these 95\% credible intervals are not the same as confidence intervals since the ``true'' value of $\rr$ may not be contained in the credible intervals 95\% of the time.
-We only use the posterior distribution to form a CI for the SB method, as it is the only Bayesian method presented here.
-
-%%%%%%%%%%%%%%%
-
-\section{Methods}\label{sec:sim-res}
-
-We compare the performance of the SIR models on simulated data from both the SIR as well as other models.
-In all of these simulations, we first generate data from the models under known conditions, then we add noise drawn from known distributions to each time point,
-\begin{align}\label{eq:sim-models}
-  X_{model}(t) &= f(t) + \epsilon_{X,t} \\
-  Y_{model}(t) &= g(t) + \epsilon_{Y,t} \nonumber\\
-  Z_{model}(t) &= N - X_{model}(t) - Y_{model}(t)\nonumber 
-\end{align}
-We adjust the recovered compartment $Z$ so the total population is constant. 
-We simulate with two different types of error distributions, Gaussian (Norm) and autoregressive (AR),
-
-\noindent \textbf{Gaussian (Norm) ($X_{G}$, $Y_{G}$, $Z_{G}$)}
-\begin{align*}
-  f(t) &= X(t) \\
-  g(t) &= Y(t) \\
-  \epsilon_{X,t} &\sim N(0, \sigma_X^2) \\
-  \epsilon_{Y,t} &\sim N(0, \sigma_Y^2)
-\end{align*}
-\textbf{Autoregressive (AR) ($X_{AR}$, $Y_{AR}$, $Z_{AR}$)}
-\begin{align*}
-  f(t) &= \rho X_{obs}(t-1) \\
-  g(t) &= \rho Y_{obs}(t-1) \\
-  \epsilon_{X,t} &\sim N(0, \sigma_X^2) \\
-  \epsilon_{Y,t} &\sim N(0, \sigma_Y^2).\\
-\end{align*}
-For each of the Gaussian and AR datasets, we form an additional dataset which enforces the monotonicity of X and Z via the Pool Adjacent Violators Algorithm (PAVA), described by \cite{friedman1984}).
-Here, we adjust $Y$ so the constant population is maintained.
-
-\noindent \textbf{Gaussian-PAVA (Norm-M) ($X_{GP}$, $Y_{GP}$, $Z_{GP}$)}
-\begin{align*}
- X_{GP}(t) &= \textnormal{PAVA}(X_G(t)) \\
-  Y_{GP}(t) &= N - X_{GP}(t) - Z_{GP}(t) \\
-  Z_{GP}(t) &= \textnormal{PAVA}(Z_G(t)) \\
-\end{align*}
-\textbf{Autoregressive-PAVA (AR-M) ($X_{ARP}$, $Y_{ARP}$, $Z_{ARP}$)}
-\begin{align*}
-  X_{ARP}(t) &= \textnormal{PAVA}(X_{AR}(t)) \\
-  Y_{ARP}(t) &= N - X_{ARP}(t) - Z_{ARP}(t) \\
-  Z_{ARP}(t) &= \textnormal{PAVA}(Z_{AR}(t)) 
-\end{align*}
-This gives us a total of four error structures for each set of conditions: (1) Gaussian (Norm), (2) Gaussian monotone (Norm-M),  (3) autoregressive (AR), and (4) autoregressive monotone (AR-M) for a given set of parameters $\left (\beta, \gamma, O, X(0), Y(0), Z(0), \sigma_X, \sigma_Y, N \right )$, the infection rate; recovery rate; total number of time points; initial conditions for number of susceptibles, infectious, and recovered individuals, the standard deviation for both the susceptible and infectious compartments, and the total population, respectively.
-
-\subsection{SIR Data}
-
-We first simulate a ``baseline'' data set from the SIR model under the following conditions:
-
-\textbf{Baseline Simulated Data}
-\begin{center}
-
-	$\beta$ = 0.06, $\gamma$ = 0.03 ($\rr$ = 2)
-
-	Number of time steps $O$: 365
-
-	Starting X size ($X(0)$): 99950
-
-	Starting Y size ($Y(0)$): 50
-
-	Starting Z size ($Z(0)$): 0
-
-	$\sigma_X$: 100
-
-	$\sigma_Y$: 5
-
-	Total population $N$: 100,000
-
-        Error: Gaussian.
-
-      \end{center}
-      The baseline data set is generated from the deterministic SIR curves in Figure \ref{fig:baseline-data} and then noise is added to two of the compartments, and the third is adjusted to maintain a constant population value.
-In this scenario, we see that the infection begins slowly but over the course of a year infects over 75\% of the population.
-      \begin{figure}
-        \centering
-        \includegraphics[width=.8\textwidth]{images/baseline_sir.pdf}
-        \caption{SIR (XYZ) curves for the deterministic model with $N=100,000$, $(\beta, \gamma) = (0.06, 0.03)$, $(X(0), Y(0), Z(0))= (99,950, 50,0)$  The baseline data is generated by using this model and adding noise to the $X$ and $Y$ compartments and adjusting $Z$ at each time step to maintain a constant population.}\label{fig:baseline-data}
-        \end{figure}
-      This baseline data set is chosen due to its moderate $\rr$ size of 2, which indicates an outbreak of an epidemic similar to HIV, ZIK-V, EVD, and H1N1 pandemic influenza (see Table \ref{tab:r0-real-ex}).
-The infection rate and recovery rate ($\beta, \gamma$) are relatively small.
-The duration of the disease, under the true model, spans nearly a whole year (see Figure \ref{fig:baseline-data}) and so we use a year's worth of daily observations of (X,Y, Z) values.
-Initially, 50 (0.05\%) diseased individuals are introduced to the population of 100,000 individuals (approximately the average population size of a US county in 2015).
-The standard deviation of the X values is 0.1\% of the magnitude of the initial number of susceptible individuals, and the standard deviation of the Y values is 10\% of the magnitude of the initial number of infectious individuals, although in this setting the number of infectious individuals increases rapidly.
-We use Gaussian error as our baseline error because of its known statistical properties so we can more easily compare methods to one another.
-Later on, we explore the autoregressive and PAVA simulated data, which may better simulate data from  a real epidemic.
-
-      \subsection{Other Simulated Data}
-
-We then generate different SIR data sets, changing one of the conditions for simulation each time and using the baseline SIR data as our basis for comparison.
-The different data sets generated are shown in Table \ref{tab:simulated-data}, for a total of 27 simulated data sets.
-
-% \begin{table}[h]
-% 	\centering
-% 	\resizebox{\textwidth}{!}{
-% 		\begin{tabular}{@{}lll@{}}
-% 			\toprule
-%                   \textbf{Parameter Values ($\beta, \gamma$)}        & \textbf{Starting Compartment Sizes ($X_0, Y_0$)}            & \textbf{Variance of Added Errors ($\sigma_X, \sigma_Y$ )}       \\ \midrule
-%                   (0.06, 0.001)            & (99000, 1000)                           & (50, 2)                     \\
-%                   (0.06, 0.04)             & (99900, 100)               & (500, 20)     \\
-%                   (0.06,.03) & (99950, 50) & (100, 5)\\
-%                   (0.06, 0.06)             & (99990, 10)             & (2500, 100)           \\
-%                   (0.06, 0.24) & (99999, 1) & (10000, 500)                                      \\ \bottomrule
-% 		\end{tabular}
-% 	}
-% 	\caption{Table of data sets.}
-% 	\label{tab:datasets}
-% \end{table}
-
-% % Please add the following required packages to your document preamble:
-% % \usepackage{booktabs}
-% \begin{table}[]
-% \centering
-
-% Please add the following required packages to your document preamble:
-% \usepackage{booktabs}
-\begin{table}[]
-\centering
-\begin{tabular}{@{}llllllll@{}}
-\toprule
-\textbf{Name} & \textbf{Model} & \textbf{Error} & \textbf{($\beta, \gamma$)} & \textbf{O} & \textbf{(X(0),Y(0))} & \textbf{N} & \textbf{($\sigma_X, \sigma_Y$)} \\ \midrule
-  Baseline        & SIR      &   Norm             & (.06, .03)                          &     365       &  (99950, 50)                    &  $10^5$          &(100, 5)                                \\ \midrule
-  AR       & SIR      &   AR ($\rho=.7)$             & (.06, .03)                          &     365       &  (99950, 50)                    &  $10^5$          &(100, 5)                            \\ \midrule
-  \textcolor{red}{Norm-M}      & SIR      &   Norm-M ($\rho=.7)$             & (.06, .03)                          &     365       &  (99950, 50)                    &  $10^5$          &(100, 5)                            \\ \midrule
-   \textcolor{red}{AR-M}       & SIR      &   AR-M ($\rho=.7)$             & (.06, .03)                          &     365       &  (99950, 50)                    &  $10^5$          &(100, 5)                            \\ \midrule
-  InfRec1       & SIR      &   Norm             & (60000, .001)                          &     365       &  (99950, 50)                    &  $10^5$          &(100, 5)\\
-  InfRec2      & SIR      &   Norm             & (1500, .001)                          &     365       &  (99950, 50)                    &  $10^5$          &(100, 5)                                \\
-  InfRec3      & SIR      &   Norm             & (1000, .001)                          &     365       &  (99950, 50)                    &  $10^5$          &(100, 5)                                \\
-InfRec4      & SIR      &   Norm             & (06, .24)                          &     365       &  (99950, 50)                    &  $10^5$          &(100, 5)                                \\ \midrule
-  Time1 & SIR &  Norm   & (.06, .03)                          &     200       &  (99950, 50)                    &  $10^5$          &(100, 5)                                \\
-  Time2 & SIR &  Norm   & (.06, .03)                          &     100       &  (99950, 50)                    &  $10^5$          &(100, 5)                                \\
-  Time3 & SIR &  Norm   & (.06, .03)                          &     50       &  (99950, 50)                    &  $10^5$          &(100, 5)                                \\
-  Time4 & SIR &  Norm   & (.06, .03)                          &     20       &  (99950, 50)                    &  $10^5$          &(100, 5)                                \\ \midrule
-  Initial1    & SIR      &   Norm             & (.06, .03)                          &     365       &  (99999, 1)                    &  $10^5$          &(100, 5)                                \\
-  Initial2    & SIR      &   Norm             & (.06, .03)                          &     365       &  (99990, 10)                    &  $10^5$          &(100, 5)                                \\
-  Initial3    & SIR      &   Norm             & (.06, .03)                          &     365       &  (99900, 100)                    &  $10^5$          &(100, 5)                                \\
-  Initial4    & SIR      &   Norm             & (.06, .03)                          &     365       &  (99000, 1000)                    &  $10^5$          &(100, 5)                                \\ \midrule
-  \textcolor{red}{PopSize1}        & SIR      &   Norm             & (.06, .03)                          &     365       &  (999500, 500)                    &  $10^6$          &(100, 5)                                \\
-  \textcolor{red}{PopSize2}       & SIR      &   Norm             & (.06, .03)                          &     365       &  (9995, 5)                    &  $10^4$          &(100, 5)                                \\
-  \textcolor{red}{PopSize3}        & SIR      &   Norm             & (.06, .03)                          &     365       &  (990, 10)                    &  $10^3$          &(100, 5)                                \\
-  \textcolor{red}{PopSize4}        & SIR      &   Norm             & (.06, .03)                          &     365       &  (99, 1)                    &  $10^2$          &(100, 5)                                \\
- \midrule
-  Var1      & SIR      &   Norm             & (.06, .03)                          &     365       &  (99950, 50)                    &  $10^5$          &(50, 2)                                \\
-  Var2      & SIR      &   Norm             & (.06, .03)                          &     365       &  (99950, 50)                    &  $10^5$          &(500, 20)                                \\
-  Var3      & SIR      &   Norm             & (.06, .03)                          &     365       &  (99950, 50)                    &  $10^5$          &(2500, 1000)                                \\
-  Var4      & SIR      &   Norm             & (.06, .03)                          &     365       &  (99950, 50)                    &  $10^5$          &(10000, 500)                                \\ \midrule
-  \textcolor{red}{Poly1}     & Linear      &   Norm             & (.06, .03)                          &     365       &  (99950, 50)                    &  $10^5$          &(100, 5)                            \\
-  \textcolor{red}{Poly2}    & Quartic      &   Norm             & (.06, .03)                          &     365       &  (99950, 50)                    &  $10^5$          &(100, 5)                            \\
-   LSIR     & LSIR      &   Norm             & (.06, .03)                          &     365       &  (99950, 50)                    &  $10^5$          &(100, 5)                            \\ 
-  \bottomrule
-\end{tabular}
-\caption{Table of simulated data sets.  We simulate 27 total data sets with the given parameters.}
-\label{tab:simulated-data}
-\end{table}
-
-To test how well the SIR models perform under misspecification, we simulate data from three other models.
-While we will not have a ``known'' $\rr$ value to compare our results to, this will allow us to see if the models give us reasonable (and comparable) results when the data do not come from the SIR model.
-We simulate data from polynomials of degree 1 and 4 with respect to the time step, using a monotonically increasing function for the $X$ compartment and a monotonically decreasing function for the $Z$ compartment.
-The sum of these is subtracted from the constant total population to obtain the $Y$ compartment.
-We call these models polynomial of degree 1 (Poly1) and polynomial of degree 4 (Poly4), respectively.
-We also generate data from a linearized version (in terms of the ODEs and not to be confused with LMA) of the three compartment SIR model (LSIR) with the following linearized ODEs:
-
-\begin{eqnarray*}
-	\frac{dX}{dt} &=& -\beta Y \\
-	\frac{dY}{dt} &=& \beta Y - \gamma Y \\
-	\frac{dZ}{dt} &=& \gamma Y.
-\end{eqnarray*}
-
-% The data from the linear SIR model were simulated under the following conditions:
-
-% \begin{center}
-
-% 	$\beta$ = 0.06, $\gamma$ = 0.05; $\rr$ = 1.2
-
-% 	Number of time steps: 365
-
-% 	Starting X size ($X_0$): 99950
-
-% 	Starting Y size ($Y_0$): 50
-
-% 	Starting Z size: 0
-
-% 	$\sigma_X$: 10
-
-% 	$\sigma_Y$: 1
-
-% 	Total population: 100000.
-
-%       \end{center}
-
-\subsection{Model Fitting Details}\label{sec:model-fit}
-
-Table \ref{tab:baseline-res} shows the ways that standard errors are acquired for the different methods.
-The IPR and SIPR models require knowledge of the length of the average infectious period.
-For the results below, we assume the best case scenario that the true average infectious period is known for the data set.
-
-\begin{table}[H]
-	\centering
-	\begin{tabular}{@{}ll@{}}
-		\toprule
-		\textbf{Model} & \textbf{Default Standard Error Method} \\ \midrule
-		Least Squares (LS) & Delta method with Hessian matrix\\
-		Reparameterized Least Squares (ReLS) & Hessian matrix \\
-		Linear Model Approximation (LMA) & Delta method with linear regression standard error \\
-		LMA, All Time Points  (LMAT)& Delta method with linear regression standard error \\
-		Incidence-Prevalence Ratio (IPR) & Jackknife resampling \\
-		Smoothed IPR (SIPR) & Jackknife resampling \\
-		Log-Linear (LL) & Linear regression standard error \\
-		Markov Chain (MC) & Delta method with GLM regression standard error \\
-		Sequential Bayes (SB) & Standard deviation of posterior distribution\\
-		% Exponential Growth (EG) & Delta method with linear regression standard error \\
-		% IDEA & Delta method with linear regression standard error \\
-		\bottomrule
-	\end{tabular}
-	\caption{Table of methods for standard errors.}
-	\label{tab:se-methods}
-\end{table}
-
-Table \ref{tab:baseline-res} shows the ways that standard errors are acquired for the different methods.
-The IPR, SIPR, and IDEA models require knowledge of the length of the average infectious period.
-For the results below, we assume the best case scenario that the true average infectious period is known for the data set.
-The exponential growth model requires knowledge of the probability distribution of the infectious period.
-In this case, we assume a uniform probability distribution with a mean equivalent to the true average infectious period.
-
-\section{Results of Simulations}\label{sec:results}
-There are numerous aspects to consider when using a method to estimate $\rr$.
-Among these include values of both the infection and recovery rate $(\beta, \gamma)$, the number of data points available ($O$), the total size of the population ($N$), the initial percent of susceptible individuals $\left (\frac{X(0)}{N}\right)$, the initial percent of infectious individuals $\left (\frac{Y(0)}{N}\right )$, the magnitude of the variance of both the number of susceptibles and the number of infectious ($\sigma_X^2, \sigma_Y^2$), the error modeling assumptions, data generated from a non-SIR model, and finally methods of estimating the variance.
-We first examine our baseline data, and then we examine each of these aspects individually.
+\section{Full Simulation Results}
 
@@
 
-In this setting, the CIs from LS, ReLS, LMAT, SIPR, and LL cover the true $\rr=2$.
+In this setting, the CIs from RE, rRE, LMAT, SIPR, and LL cover the true $\rr=2$.
 In contrast to the baseline data, the CI for LMA no longer covers the true estimate.
@@
 		\hline
-		LS & 1.9997 & 0.0072\\
-		\hline
-		ReLS & 2.0001 & 0.0063\\
+		RE & 1.9997 & 0.0072\\
+		\hline
+		rRE & 2.0001 & 0.0063\\
 		\hline
@@
 
-The methods that contain the true value of $\rr$ are LS, ReLS, LMAT, IPR for both error assumptions and LL with Norm-AR error assumption.
-The methods LS, ReLS, LMA, and LL would be able to reject the hypothesis that $\rr < 1$.
+The methods that contain the true value of $\rr$ are RE, rRE, LMAT, IPR for both error assumptions and LL with Norm-AR error assumption.
+The methods RE, rRE, LMA, and LL would be able to reject the hypothesis that $\rr < 1$.
 
@@
 		\hline
-		LS & 2.0002 & 0.0044 & 1.9999 & 0.0074\\
-		\hline
-		ReLS & 2.0003 & 0.0040 & 1.9997 & 0.0065\\
+		RE & 2.0002 & 0.0044 & 1.9999 & 0.0074\\
+		\hline
+		rRE & 2.0003 & 0.0040 & 1.9997 & 0.0065\\
 		\hline
@@
 		\hline
-		SIRLS & 59.9765 & 0.0435\\
-		\hline
-		ReLS & 59.9952 & 0.4344\\
+		SIRRE & 59.9765 & 0.0435\\
+		\hline
+		rRE & 59.9952 & 0.4344\\
 		\hline
@@
 We examine an exaggerated extreme case of when $\rr=60$ (the largest reported $\rr$ value we found is $\hat{\rr}=14$ for measles as reported in \cite{anderson1992}).
-Only LS and ReLS result in CIs that cover the true $\rr$.
+Only RE and rRE result in CIs that cover the true $\rr$.
 IPR and LL give fairly close point estimates to the true $\rr=60$ as well.
@@
 Several of the models also return extremely large standard errors, meaning they are not very reliable in these cases.
-Only LS, ReLS, and LL would reject the hypothesis that $\rr <1$, when this is clearly not the case at all.
+Only RE, rRE, and LL would reject the hypothesis that $\rr <1$, when this is clearly not the case at all.
 
 When $\rr=1.5$ we see that the estimates from the methods give results much closer to the true value, and the resulting CIs are also smaller.
-The methods LS, ReLS, SIPR, LL, and SB would all reject the hypothesis that $\rr > 1$.
+The methods RE, rRE, SIPR, LL, and SB would all reject the hypothesis that $\rr > 1$.
 Although if we are testing the null hypothesis for  all methods simultaneously, we should use a multiple-testing correction such as Bonferroni.
@@
 
-When $\rr=1$, we are unsure whether there will be an epidemic or not.  In this case, all the CIs of the methods cover the true value.  We also see that the SE of LS is much greater than ReLS.
+When $\rr=1$, we are unsure whether there will be an epidemic or not.  In this case, all the CIs of the methods cover the true value.  We also see that the SE of RE is much greater than rRE.
 
@@
 		\hline
-		LS & 0.2363 & 4.1043 \\
-		\hline
-		ReLS & 4.1795 & 2.7456\\
+		RE & 0.2363 & 4.1043 \\
+		\hline
+		rRE & 4.1795 & 2.7456\\
 		\hline
@@
 It appears that a lot of estimation methods do not handle small $\rr$ well either.
-None of them are extremely close to truth, except for ReLS, as most of the tend to overestimate the truth.
+None of them are extremely close to truth, except for rRE, as most of the tend to overestimate the truth.
 The IPR is particularly bad in this case.
@@
 		\hline
-		LS & 59.9765 & 0.0387\\
-		\hline
-		ReLS & 59.9952 & 0.3875\\
+		RE & 59.9765 & 0.0387\\
+		\hline
+		rRE & 59.9952 & 0.3875\\
 		\hline
@@
 		\hline
-		LS & 0.2328 & 0.6701\\
-		\hline
-		ReLS & 3.5710 & 0.3171\\
+		RE & 0.2328 & 0.6701\\
+		\hline
+		rRE & 3.5710 & 0.3171\\
 		\hline
@@
 		\hline
-		LS & 60.0151 & 0.0698\\
-		\hline
-		ReLS & 60.0255 & 0.6912\\
+		RE & 60.0151 & 0.0698\\
+		\hline
+		rRE & 60.0255 & 0.6912\\
 		\hline
@@
 		\hline
-		LS & -0.0695 & 2.8291\\
-		\hline
-		ReLS & 5.1930 & 3.5250\\
+		RE & -0.0695 & 2.8291\\
+		\hline
+		rRE & 5.1930 & 3.5250\\
 		\hline
@@
 		\hline
-		LS & 59.9589 & 0.0631\\
-		\hline
-		ReLS & 59.9961 & 0.6215\\
+		RE & 59.9589 & 0.0631\\
+		\hline
+		rRE & 59.9961 & 0.6215\\
 		\hline
@@
 		\hline
-		LS & 0.3021 & 0.5657\\
-		\hline
-		ReLS & 9.8415 & 1.7074\\
+		RE & 0.3021 & 0.5657\\
+		\hline
+		rRE & 9.8415 & 1.7074\\
 		\hline
@@
   \includegraphics[scale=0.5]{images/time_n.jpeg}
-  \caption{Forest plots of estimates of $\rr$ for the \xxsir methods for the Time1-4 data sets with Gaussian errors $(\beta=.06, \gamma=.03, O=365, X(0)=99950, Y(0)=50, Z(0)=0, \sigma_X=100, \sigma_Y=5, N=10^5)$.
+  \caption{Forest plots of estimates of $\rr$ for the \xxsir methods for the Time1-4 data sets with Gaussian errors $(\beta=.06, \gamma=.03, X(0)=99950, Y(0)=50, Z(0)=0, \sigma_X=100, \sigma_Y=5, N=10^5)$.
 The points are the estimates of $\rr$ and the lines denote $\pm 2\cdot $Std.
@@
 		\hline
-		LS & 1.9993 & 0.0286\\
-		\hline
-		ReLS & 1.9994 & 0.0313\\
+		RE & 1.9993 & 0.0286\\
+		\hline
+		rRE & 1.9994 & 0.0313\\
 		\hline
@@
 Only LL would reject the hypothesis that $\rr < 1$ for 50 or fewer time points, but its CI does not cover the true $\rr$ value.
-With 50 and greater time points, we see that LS and ReLS provide close point estimates to the true value and by 100 time points, would reject the hypothesis that $\rr < 1$.
+With 50 and greater time points, we see that RE and rRE provide close point estimates to the true value and by 100 time points, would reject the hypothesis that $\rr < 1$.
 The methods and MC and SB consistently have small CIs, but never cover the true value.
@@
 		\hline
-		LS & 38.6896 & 106.2705\\
-		\hline
-		ReLS & 22.9646 & 6.9817\\
+		RE & 38.6896 & 106.2705\\
+		\hline
+		rRE & 22.9646 & 6.9817\\
 		\hline
@@
 As might be expected, the $\rr$ estimates are wildly inaccurate, and none are very close to the true value.
-The methods of LS and ReLS are particularly worse in this situation compared to the other data sets.
+The methods of RE and rRE are particularly worse in this situation compared to the other data sets.
 The standard errors are much larger than the baseline, which is to be expected as well due to the small number of data points.
@@
 	\includegraphics[scale=0.5]{images/time_nm.jpeg}
-	\caption{Forest plots of estimates of $\rr$ for the \xxsir methods for the Time1-4 data sets with Gaussian errors $(\beta=.06, \gamma=.03, O=365, X(0)=99950, Y(0)=50, Z(0)=0, \sigma_X=100, \sigma_Y=5, N=10^5)$.
+	\caption{Forest plots of estimates of $\rr$ for the \xxsir methods for the Time1-4 data sets with Gaussian errors $(\beta=.06, \gamma=.03, X(0)=99950, Y(0)=50, Z(0)=0, \sigma_X=100, \sigma_Y=5, N=10^5)$.
 The points are the estimates of $\rr$ and the lines denote $\pm 2\cdot $Std.
@@
 		\hline
-		LS & 1.9934 & 0.1464\\
-		\hline
-		ReLS & 1.9935 & 0.1975\\
+		RE & 1.9934 & 0.1464\\
+		\hline
+		rRE & 1.9935 & 0.1975\\
 		\hline
@@
 		\hline
-		LS & 5.0366 & 1.3672\\
-		\hline
-		ReLS & 5.0449 & 2.1398\\
+		RE & 5.0366 & 1.3672\\
+		\hline
+		rRE & 5.0449 & 2.1398\\
 		\hline
@@
 	\includegraphics[scale=0.5]{images/time_ar.jpeg}
-	\caption{Forest plots of estimates of $\rr$ for the \xxsir methods for the Time1-4 data sets with Gaussian errors $(\beta=.06, \gamma=.03, O=365, X(0)=99950, Y(0)=50, Z(0)=0, \sigma_X=100, \sigma_Y=5, N=10^5)$.
+	\caption{Forest plots of estimates of $\rr$ for the \xxsir methods for the Time1-4 data sets with Gaussian errors $(\beta=.06, \gamma=.03, X(0)=99950, Y(0)=50, Z(0)=0, \sigma_X=100, \sigma_Y=5, N=10^5)$.
 The points are the estimates of $\rr$ and the lines denote $\pm 2\cdot $Std.
@@
 		\hline
-		LS & 1.9098 & 0.3811\\
-		\hline
-		ReLS & 1.9104 & 0.4773\\
+		RE & 1.9098 & 0.3811\\
+		\hline
+		rRE & 1.9104 & 0.4773\\
 		\hline
@@
 		\hline
-		LS & 0.0099 & 2.1688\\
-		\hline
-		ReLS & -0.1539 & 2.7692\\
+		RE & 0.0099 & 2.1688\\
+		\hline
+		rRE & -0.1539 & 2.7692\\
 		\hline
@@
 	\includegraphics[scale=0.5]{images/time_arm.jpeg}
-	\caption{Forest plots of estimates of $\rr$ for the \xxsir methods for the Time1-4 data sets with Gaussian errors $(\beta=.06, \gamma=.03, O=365, X(0)=99950, Y(0)=50, Z(0)=0, \sigma_X=100, \sigma_Y=5, N=10^5)$.
+	\caption{Forest plots of estimates of $\rr$ for the \xxsir methods for the Time1-4 data sets with Gaussian errors $(\beta=.06, \gamma=.03, X(0)=99950, Y(0)=50, Z(0)=0, \sigma_X=100, \sigma_Y=5, N=10^5)$.
 The points are the estimates of $\rr$ and the lines denote $\pm 2\cdot $Std.
@@
 		\hline
-		LS & 1.9674 & 0.2324\\
-		\hline
-		ReLS & 1.9679 & 0.3065\\
+		RE & 1.9674 & 0.2324\\
+		\hline
+		rRE & 1.9679 & 0.3065\\
 		\hline
@@
 		\hline
-		LS & 1.6816 & 0.4747\\
-		\hline
-		ReLS & 1.6813 & 0.6544\\
+		RE & 1.6816 & 0.4747\\
+		\hline
+		rRE & 1.6813 & 0.6544\\
 		\hline
@@
   \includegraphics[scale=0.5]{images/start_n.jpeg}
-    \caption{Forest plots of estimates of $\rr$ for the \xxsir methods for the Inits1-4 data sets $(\beta=.06, \gamma=.03, O=365, X(0)=99950, Y(0)=50, Z(0)=0, \sigma_X=100, \sigma_Y=5, N=10^5)$.
+    \caption{Forest plots of estimates of $\rr$ for the \xxsir methods for the Inits1-4 data sets $(\beta=.06, \gamma=.03, O=365, \sigma_X=100, \sigma_Y=5, N=10^5)$.
 The points are the estimates of $\rr$ and the lines denote $\pm 2\cdot $Std.
@@
 		\hline
-		LS & 1.9996 & 0.0050\\
-		\hline
-		ReLS & 1.9998 & 0.0034\\
+		RE & 1.9996 & 0.0050\\
+		\hline
+		rRE & 1.9998 & 0.0034\\
 		\hline
@@
 		\hline
-		LS & 2.4683 & 1.5371\\
-		\hline
-		ReLS & 2.4692 & 2.7294 \\
+		RE & 2.4683 & 1.5371\\
+		\hline
+		rRE & 2.4692 & 2.7294 \\
 		\hline
@@
 	\includegraphics[scale=0.5]{images/start_nm.jpeg}
-	\caption{Forest plots of estimates of $\rr$ for the \xxsir methods for the Inits1-4 data sets $(\beta=.06, \gamma=.03, O=365, X(0)=99950, Y(0)=50, Z(0)=0, \sigma_X=100, \sigma_Y=5, N=10^5)$.
+	\caption{Forest plots of estimates of $\rr$ for the \xxsir methods for the Inits1-4 data sets $(\beta=.06, \gamma=.03, O=365, \sigma_X=100, \sigma_Y=5, N=10^5)$.
 The points are the estimates of $\rr$ and the lines denote $\pm 2\cdot $Std.
@@
 		\hline
-		LS & 2.0003 & 0.0046\\
-		\hline
-		ReLS & 2.0003 & 0.0032\\
+		RE & 2.0003 & 0.0046\\
+		\hline
+		rRE & 2.0003 & 0.0032\\
 		\hline
@@
 		\hline
-		LS & 2.4668 & 1.5349\\
-		\hline
-		ReLS & 2.4670 & 2.7219\\
+		RE & 2.4668 & 1.5349\\
+		\hline
+		rRE & 2.4670 & 2.7219\\
 		\hline
@@
 	\includegraphics[scale=0.5]{images/start_ar.jpeg}
-	\caption{Forest plots of estimates of $\rr$ for the \xxsir methods for the Inits1-4 data sets $(\beta=.06, \gamma=.03, O=365, X(0)=99950, Y(0)=50, Z(0)=0, \sigma_X=100, \sigma_Y=5, N=10^5)$.
+	\caption{Forest plots of estimates of $\rr$ for the \xxsir methods for the Inits1-4 data sets $(\beta=.06, \gamma=.03, O=365, \sigma_X=100, \sigma_Y=5, N=10^5)$.
 The points are the estimates of $\rr$ and the lines denote $\pm 2\cdot $Std.
@@
 		\hline
-		LS & 2.0005 & 0.0078\\
-		\hline
-		ReLS & 2.0001 & 0.0054\\
+		RE & 2.0005 & 0.0078\\
+		\hline
+		rRE & 2.0001 & 0.0054\\
 		\hline
@@
 		\hline
-		LS & 2.4706 & 1.5372\\
-		\hline
-		ReLS & 2.4707 & 2.7245\\
+		RE & 2.4706 & 1.5372\\
+		\hline
+		rRE & 2.4707 & 2.7245\\
 		\hline
@@
 	\includegraphics[scale=0.5]{images/start_arm.jpeg}
-	\caption{Forest plots of estimates of $\rr$ for the \xxsir methods for the Inits1-4 data sets $(\beta=.06, \gamma=.03, O=365, X(0)=99950, Y(0)=50, Z(0)=0, \sigma_X=100, \sigma_Y=5, N=10^5)$.
+	\caption{Forest plots of estimates of $\rr$ for the \xxsir methods for the Inits1-4 data sets $(\beta=.06, \gamma=.03, O=365, \sigma_X=100, \sigma_Y=5, N=10^5)$.
 The points are the estimates of $\rr$ and the lines denote $\pm 2\cdot $Std.
@@
 		\hline
-		LS & 2.0001 & 0.0069\\
-		\hline
-		ReLS & 2.0000 & 0.0048\\
+		RE & 2.0001 & 0.0069\\
+		\hline
+		rRE & 2.0000 & 0.0048\\
 		\hline
@@
 		\hline
-		LS & 2.4667 & 1.5349\\
-		\hline
-		ReLS & 2.4665 & 2.7349\\
+		RE & 2.4667 & 1.5349\\
+		\hline
+		rRE & 2.4665 & 2.7349\\
 		\hline
@@
 We display the results of estimating $\rr$ from the methods fit on the baseline data set with varying $N$ in Figure \ref{fig:inits-res} and Tables \ref{tab:n1-res2} and \ref{tab:n2-res2}.
-When $N=100$, LS, ReLS, LMA, LMAT, IPR, and SIPR's 95\% CIs all cover the true $\rr$.
+When $N=100$, RE, rRE, LMA, LMAT, IPR, and SIPR's 95\% CIs all cover the true $\rr$.
 However, we would say that all of these intervals are uninformative.
-We see that CIs of LS are wider than those of ReLS, although the difference becomes less substantial as $N$ grows larger.
+We see that CIs of RE are wider than those of rRE, although the difference becomes less substantial as $N$ grows larger.
 
@@
 MC and SB would actually allow us to say $\rr < 1$ with 95\% confidence.
-It is not until $N=10,000$ that we are able to say with certainty that LS and ReLS, in addition to LL, give informative estimates of $\rr$.
-
-Thus, it seems that if one is estimating for $\rr$ for a small population such as a small village or crusie ship, LL may be the most appropriate method, but when the population is larger (~10,000), we can be confident in using LS or ReLS.
+It is not until $N=10,000$ that we are able to say with certainty that RE and rRE, in addition to LL, give informative estimates of $\rr$.
+
+Thus, it seems that if one is estimating for $\rr$ for a small population such as a small village or crusie ship, LL may be the most appropriate method, but when the population is larger (~10,000), we can be confident in using RE or rRE.
 
@@
 		\hline
-		LS & 2.0008 & 0.0590\\
-		\hline
-		ReLS & 2.0005 & 0.0528\\
+		RE & 2.0008 & 0.0590\\
+		\hline
+		rRE & 2.0005 & 0.0528\\
 		\hline
@@
 		\hline
-		LS & 1.9863 & 4.8620\\
-		\hline
-		ReLS & 1.9871 & 3.4569\\
+		RE & 1.9863 & 4.8620\\
+		\hline
+		rRE & 1.9871 & 3.4569\\
 		\hline
@@
 		\hline
-		LS & 2.0175 & 0.1430\\
-		\hline
-		ReLS & 2.0175 & 0.1021\\
+		RE & 2.0175 & 0.1430\\
+		\hline
+		rRE & 2.0175 & 0.1021\\
 		\hline
@@
 		\hline
-		LS & 1.9998 & 0.0006\\
-		\hline
-		ReLS & 1.9999 & 0.0005\\
+		RE & 1.9998 & 0.0006\\
+		\hline
+		rRE & 1.9999 & 0.0005\\
 		\hline
@@
 		\hline
-		LS & 2.1221 & 0.7363\\
-		\hline
-		ReLS & 2.1219 & 0.5660\\
+		RE & 2.1221 & 0.7363\\
+		\hline
+		rRE & 2.1219 & 0.5660\\
 		\hline
@@
 		\hline
-		LS & 1.9998 & 0.0009\\
-		\hline
-		ReLS & 2.0001 & 0.0007\\
+		RE & 1.9998 & 0.0009\\
+		\hline
+		rRE & 2.0001 & 0.0007\\
 		\hline
@@
 		\hline
-		LS & 2.0975 & 0.2921\\
-		\hline
-		ReLS & 2.0976 & 0.2211\\
+		RE & 2.0975 & 0.2921\\
+		\hline
+		rRE & 2.0976 & 0.2211\\
 		\hline
@@
 		\hline
-		LS & 2.0000 & 0.0007\\
-		\hline
-		ReLS & 1.9997 & 0.0007\\
+		RE & 2.0000 & 0.0007\\
+		\hline
+		rRE & 1.9997 & 0.0007\\
 		\hline
@@
 In particular, LMA and LMAT seem to be particularly effected as the CIs grow substantially as the variance grows.
-Again we see that LS, ReLS, and LL provide the most accurate estimates with relatively small CIs.
+Again we see that RE, rRE, and LL provide the most accurate estimates with relatively small CIs.
 
@@
 		\hline
-		LS & 1.9998 & 0.0027\\
-		\hline
-		ReLS & 1.9998 & 0.0024\\
+		RE & 1.9998 & 0.0027\\
+		\hline
+		rRE & 1.9998 & 0.0024\\
 		\hline
@@
 		\hline
-		LS & 1.9981 & 0.1354\\
-		\hline
-		ReLS & 1.9981 & 0.1208\\
+		RE & 1.9981 & 0.1354\\
+		\hline
+		rRE & 1.9981 & 0.1208\\
 		\hline
@@
 		\hline
-		LS & 1.9998 & 0.0026\\
-		\hline
-		ReLS & 1.9999 & 0.0024\\
+		RE & 1.9998 & 0.0026\\
+		\hline
+		rRE & 1.9999 & 0.0024\\
 		\hline
@@
 		\hline
-		LS & 1.9970 & 0.0541\\
-		\hline
-		ReLS & 1.9971 & 0.0482\\
+		RE & 1.9970 & 0.0541\\
+		\hline
+		rRE & 1.9971 & 0.0482\\
 		\hline
@@
 		\hline
-		LS & 1.9996 & 0.0038\\
-		\hline
-		ReLS & 1.9999 & 0.0034\\
+		RE & 1.9996 & 0.0038\\
+		\hline
+		rRE & 1.9999 & 0.0034\\
 		\hline
@@
 		\hline
-		LS & 1.9997 & 0.1894\\
-		\hline
-		ReLS & 2.0002 & 0.1680\\
+		RE & 1.9997 & 0.1894\\
+		\hline
+		rRE & 2.0002 & 0.1680\\
 		\hline
@@
 		\hline
-		LS & 1.9990 & 0.0036\\
-		\hline
-		ReLS & 1.9995 & 0.0032\\
+		RE & 1.9990 & 0.0036\\
+		\hline
+		rRE & 1.9995 & 0.0032\\
 		\hline
@@
 		\hline
-		LS & 2.0021 & 0.0928\\
-		\hline
-		ReLS & 2.0022 & 0.0827\\
+		RE & 2.0021 & 0.0928\\
+		\hline
+		rRE & 2.0022 & 0.0827\\
 		\hline
@@
 Moreover, only the CI of IPR covers that of the true value of $\rr=2$ and the CIs are uninformative.
-For the quartic model, only ReLS has a CI that both covers the true $\rr$ while rejecting the hypothesis that $\rr < 1$.
-LS covers the true value with its CI but also includes $\rr=1$.
+For the quartic model, only rRE has a CI that both covers the true $\rr$ while rejecting the hypothesis that $\rr < 1$.
+RE covers the true value with its CI but also includes $\rr=1$.
 LL rejects the hypothesis that $\rr <$ but its CI does not contain $\rr=2$.
@@
 		\hline
-		LS & 1.0514 & 0.3783\\
-		\hline
-		ReLS & 1.0514 & 0.0259\\
+		RE & 1.0514 & 0.3783\\
+		\hline
+		rRE & 1.0514 & 0.0259\\
 		\hline
@@
 	\hline
-	LS & 1.7693 & 0.4227\\
-	\hline
-	ReLS & 1.7693 & 0.2797\\
+	RE & 1.7693 & 0.4227\\
+	\hline
+	rRE & 1.7693 & 0.2797\\
 	\hline
@@
 		\hline
-		LS & 1.2743 & 0.0980\\
-		\hline
-		ReLS & 1.2743 & 0.0340\\
+		RE & 1.2743 & 0.0980\\
+		\hline
+		rRE & 1.2743 & 0.0340\\
 		\hline
@@
 		\hline
-		LS & 1.0514 & 0.3785\\
-		\hline
-		ReLS & 1.0514 & 0.0259\\
+		RE & 1.0514 & 0.3785\\
+		\hline
+		rRE & 1.0514 & 0.0259\\
 		\hline
@@
 		\hline
-		LS & 1.7691 & 0.4216\\
-		\hline
-		ReLS & 1.7691 & 0.2791\\
+		RE & 1.7691 & 0.4216\\
+		\hline
+		rRE & 1.7691 & 0.2791\\
 		\hline
@@
 		\hline
-		LS & 1.2739 & 0.0977\\
-		\hline
-		ReLS & 1.2741 & 0.0337\\
+		RE & 1.2739 & 0.0977\\
+		\hline
+		rRE & 1.2741 & 0.0337\\
 		\hline
@@
 		\hline
-		LS & 1.0519 & 0.3775\\
-		\hline
-		ReLS & 1.0518 & 0.0260\\
+		RE & 1.0519 & 0.3775\\
+		\hline
+		rRE & 1.0518 & 0.0260\\
 		\hline
@@
 		\hline
-		LS & 1.7690 & 0.4228\\
-		\hline
-		ReLS & 1.7691 & 0.2799\\
+		RE & 1.7690 & 0.4228\\
+		\hline
+		rRE & 1.7691 & 0.2799\\
 		\hline
@@
 		\hline
-		LS & 1.2737 & 0.1000\\
-		\hline
-		ReLS & 1.2742 & 0.0346\\
+		RE & 1.2737 & 0.1000\\
+		\hline
+		rRE & 1.2742 & 0.0346\\
 		\hline
@@
 		\hline
-		LS & 1.0513 & 0.3783\\
-		\hline
-		ReLS & 1.0513 & 0.0257\\
+		RE & 1.0513 & 0.3783\\
+		\hline
+		rRE & 1.0513 & 0.0257\\
 		\hline
@@
 		\hline
-		LS & 1.7691 & 0.4216\\
-		\hline
-		ReLS & 1.7691 & 0.2789\\
+		RE & 1.7691 & 0.4216\\
+		\hline
+		rRE & 1.7691 & 0.2789\\
 		\hline
@@
 		\hline
-		LS & 1.2740 & 0.0971\\
-		\hline
-		ReLS & 1.2738 & 0.0334\\
+		RE & 1.2740 & 0.0971\\
+		\hline
+		rRE & 1.2738 & 0.0334\\
 		\hline
@@
 
-\section{Variance Estimation Methods}\label{sec:sim-var-res}
-We examine the variance methods used to estimate CIs for our estimates of $\rr$.
-From the above results, we saw that methods such as LS, ReLS, LL, and MC result in very small CIs.
-On the other hand, methods such as IPR have consistently large CIs, to the point of being uninformative.
-The other methods CI width seems to depend heavily on the initial parameters used.
-Unsurprisingly, an $\sigma_X$, $\sigma_Y$, $T$, or $N$ seems to be positively associated with a larger CI.
-Having a larger percentage of initially infected individuals also seems to be associated with a larger CI.
-We also see that having a very large $\rr$ value seems to have less effect on creating a large CI than does having a small value of $\rr$.
-Perhaps surprisingly, the width of the CI seems to be independent of the error structure.
-
-We estimate the variance through the delta method, block bootstrap, and posterior and compare it to the empirical variance of $\rr$ for each of the methods, using 1000 data sets generated from the baseline data structure.
-
-\subsection{Repeated Sampling}
-
-Table \ref{tab:rep-samp} shows the mean and standard deviation of $\rr$ estimates produced by the models under 100 repetitions of the baseline conditions, with new noise generated each time.
-We want to compare the standard deviations of the estimates to the standard errors returned by each of the models for a single data set to asses how reliable those standard errors are.
-Overall, it appears than most of the models have standard error estimates that are larger than the actual standard deviation of the estimates under repeated sampling.
-In some cases, such as LL, LMA, and LMAT, the standard deviation and standard error estimates are fairly close.
-The least squares methods, LS and ReLS, on the other hand, tend to overestimate the error by an order of magnitude.
-Finally, SIPR and SB, have individual standard error estimates that are smaller than the standard deviation of repeated estimates, meaning that they understate how much error is in the estimates.
-
-\begin{table}[H]
-
-	\centering
-	\begin{tabular}[t]{l|r|r|r|r}
-		\hline
-		Model & Mean Estimate & Std. Dev. & Baseline Estimate & Baseline SE\\
-		\hline
-		LS & 2.0000 & 0.0003 & 1.9999 & 0.0058\\
-		\hline
-		ReLS & 2.0000 & 0.0003 & 1.9997 & 0.0052\\
-		\hline
-		LMA & 1.8497 & 0.3339 & 1.9974 & 0.3900\\
-		\hline
-		LMAT & 1.4562 & 0.3577 & 1.3980 & 0.6028 \\
-		\hline
-		IPR & 4.5214 & 0.3705 & 2.5930 & 9.0408\\
-		\hline
-		SIPR & 1.7973 & 0.0951 & 0.9251 & $<$ 1e-04 \\
-		\hline
-		LL & 2.0000 & 0.0003 & 1.9998 & 0.0002\\
-		\hline
-		MC & 0.9486 & $<$ 1e-04 & 0.9486 & $<$ 1e-04\\
-		\hline
-		SB & 1.1330 & 0.4163 & 1.0000 & 0.0523\\
-		\hline
-	\end{tabular}
-        \caption{Mean $\rr$ Estimates and Std. Devs, Repeated Sampling}
-        \label{tab:rep-samp}
-\end{table}
-
-\subsection{Delta Method}
-
-\subsection{Block Bootstrap}
-
-Table \ref{tab:bb-samp} shows the mean and standard deviation of $\rr$ estimates produced by the models under 100 block bootstrap repetitions of the baseline data.
-The data were first detrended using the predictions of the compartment sizes at each time point from each of the models on the original data set.
-The block bootstrap algorithm was then applied on to the residuals from the predictions using a fixed block length.
-Only 5 of the models provide predictions of the size of the compartments and thus allow for detrending of the data.
-The least squares and reparameterized model have a fairly accurate estimates of $\rr$ as well as very small standard errors.
-On the other hand, the linear approximation and smoothed IPR models both have very inaccurate estimates as well as very large standard errors of $\rr$.
-
-\begin{table}[H]
-
-	\centering
-	\begin{tabular}[t]{l|r|r|r|r}
-		\hline
-		Model & Mean Estimate & Std. Dev. & Baseline Estimate & Baseline SE\\
-		\hline
-		LS & 1.9999 & 0.0003 & 1.9999 & 0.0058\\
-		\hline
-		ReLS & 1.9998 & 0.0003 & 1.9997 & 0.0052\\
-		\hline
-		LMA & -0.9975 & 21.7161 & 1.9974 & 0.3900\\
-		\hline
-		LMAT & 1.5138 & 1.0485 & 1.3980 & 0.6028 \\
-		\hline
-		SIPR & 0.7762 & 2.6624 & 0.9252 & $<$ 1e-04 \\
-		\hline
-	\end{tabular}
-	\caption{Mean $\rr$ Estimates and Std. Devs, Block Bootstrap}
-	\label{tab:bb-samp}
-\end{table}
-
-Figure \ref{fig:coverage} shows the proportion of repetitions for both repeated sampling and block bootstrapping in which the 95\% confidence intervals of the estimates produced by the models covers the true $\rr$.
-The least squares, reparameterized, and linear model approximation with all time points contain the true $\rr$ value in all of the confidence intervals.
-On the other hand, the smoothed IPR, Markov chain, and sequential Bayes models never cover the true $\rr$ in any of the confidence intervals.
-
-\begin{figure}[H]
-	\begin{center}
-		\includegraphics[scale=0.5]{images/coverage.jpeg}
-		\caption{Proportion of coverage of true $\rr$ of confidence intervals produced by repeated sampling and block bootstrapping.}
-		\label{fig:coverage}	
-	\end{center}
-      \end{figure}
-
-      \subsection{Posterior}
-
-      \section{Application to USA H1N1 data}\label{sec:real-data}
-      We apply our methods using data from the 2009 H1N1 pandemic in the United States.
-
-      \subsection{USA H1N1 Influenza 2009 Data}
-      The data is from the Center of Disease Control and Prevention's (CDC) FluView \citep{cdc-fluview}.
-The data source is ILINet and is the national level data.
-We use the same date range as in \cite{towers2009}, namely Epiweeks 21-33 (May 23, 2009-August 22, 2009).
-From the United States Census Bureau, we use the population estimate of the USA as of April 1, 2010 of $N=308,740,000$ individuals \citep{census-2010}.
-The main features in this data set are weighted influenza like illness (wILI) for a given week (Epiweek).
-The wILI values are estimated based on the number of patients diagnosed with influenza like illness and weighted based on the reporting locations.
-To obtain weekly incidence from wILI, we use $J(t) = \textnormal{wILI}(t) / 100 \cdot N$.
-Then to obtain prevalence $I(t)$, we use the relation of $I(t) = \gamma^{-1}J(t)$, where $\gamma$ is the inverse of the expected time to recovery, which we here use as $\gamma^{-1} = \frac{3}{7}=.429$ weeks \citep{vespignani2007}.
-Then $S(t) = N - \sum_{s=0}^{t}J(s)$, the total population size minus the cumulative sum of the incidence, and $R(t) = N - S(t) - I(t)$.
-As prior knowledge of $\gamma$ is used to estimate our data for $I$, we expect it to influence our estimates of $\rr$.
-The data is summarized in Table \ref{tab:h1n1-data}.
-
-% Please add the following required packages to your document preamble:
-% \usepackage{booktabs}
-\begin{table}[H]
-\centering
-\begin{tabular}{@{}ll@{}}
-\toprule
-Feature       & Notes                                                                   \\ \midrule
-Disease       & H1N1 Pandemic Influenza                                                 \\
-Source        & CDC FluView (\url{https://www.cdc.gov/flu/weekly/})      \\
-Year          & 2009                                                                    \\
-Date(s)          & May 23-August 22 (Epiweeks 21-33)                                       \\
-Location      & USA                                                                     \\
-Population    & $N$=308,740,000 people                                                  \\
-  Avg. Duration $\left ( \gamma^{-1}\right )$ & $\frac{3}{7}$ weeks\\
-  Raw data & Weekly wILI reports\\
-  Final data & S$(t)$, I$(t)$, R$(t)$ \\\bottomrule
-\end{tabular}
-\caption{Summary of the data used to estimate $\rr$ during the 2009 H1N1 influenza pandemic in the USA.  We follow the data collection procedure described in \cite{towers2009}.}
-\label{tab:h1n1-data}
-\end{table}
-
-\begin{figure}[H]
-  \centering
-  \includegraphics[width=\textwidth]{images/h1n1-sir.pdf}
-  \caption{Data from H1N1 influenza 2009 in the USA.  Weekly wILI is reported, which we transform into S, I, and R counts for epiweeks 21-33 using $\gamma^{-1} = 3/7$ weeks.}
-  \end{figure}
-
-      \subsection{Results}
-      The results are shown below.  \cite{towers2009} report an estimate of $\hat{\rr} \in [1.50, 1.70]$.
-
-\begin{figure}[H]
-	\centering
-	\includegraphics[width=\textwidth]{images/h1n1.jpeg}
-	\caption{Results from SIR models on H1N1 data.}
-\end{figure}
-
-\begin{table}[H]
-
-	\centering
-	\begin{tabular}[t]{l|r|r}
-		\hline
-		Model & Estimate & Std. Dev\\
-		\hline
-		LS & 0.0280 & 0.0549\\
-		\hline
-		ReLS & 0.3332 & 0.0001\\
-		\hline
-		LMA & 0.9900 & 3.2589\\
-		\hline
-		LMAT & 0.9928 & 0.0276\\
-		\hline
-		IPR & 5.3078 & 1.7141\\
-		\hline
-		SIPR & -17.9646 & 5.1240\\
-		\hline
-		LL & 0.0036 & 0.0000\\
-		\hline
-		MC & 0.9453 & 0.0011\\
-		\hline
-		SB & 39.9036 & 1.7520\\
-		\hline
-	\end{tabular}
-	\caption{H1N1 Data, $\rr$ Estimates and Standard Errors}
-\end{table}
-
-\section{Discussion}\label{sec:discussion}
-
-%% say what we said
-%% say it
-%% say it again
-
-%% what is r0 and why it is important
-In this paper, we examine $\rr$, the average number of secondary infections a primary individual generates when introduced into a completely susceptible population, in the context of the SIR model.
-The quantity $\rr$ summarizes an epidemic into a single estimate in the hopes that infectious disease outbreaks may be compared to one another.
-A value of $\rr > 1$ indicates an outbreak of a disease whereas a value of $\rr < 1$ indicates that no outbreak will occur.
-Thus, it is critical to not only estimate the value of $\rr$ but also a corresponding CI.
-Overestimating $\rr > 1$ during an ongoing epidemic may result in a waste of resources in prevention and control measures; underestimating $\rr < 1$ may result in a worldwide crisis.
-
-In Table \ref{tab:r0-real-ex}, we provide estimates for $\rr$ that have appeared over the past three decades for HIV, Zika, Ebola, influenza, H1N1 influenza, and measles in different locations throughout the world.
-Each of these estimates is greater than $1$, which makes sense as we tend to care more when a disease has the potential outbreak.
-The largest $\rr$ point estimate we have seen is for measles at $\rr=17$ in England in the 20th century, and the smallest is $\rr=1.08$ for HIV in Portugal in the 21st century.
-%% r0 as a property of the model
-The quantity $\rr$ may summarize an epidemic, accounting for temperature, seasonality, location, and genetics of a disease, but it is still a property of the model.
-Namely, how individuals progress through an epidemic (e.g.
-susceptible to infectious to recovered compartments) influences how we view $\rr$.
-We thus should compare estimates for $\rr$ for a fixed epidemic model.
-Here, we limit our analysis to estimates of $\rr$ derived from the SIR model.
-Then $\rr = \frac{\beta}{\gamma}$ where $\beta$ is the rate of infection and $\gamma$ is the rate of recovery.
-As a consequence, we need to estimate $\rr$ from not only the number of infections but also the number of individuals recovering at each time step.
-
-%% the methods we chose and why we chose them
-Additionally, the SIR model assumes monotoncity of S and R compartments which can make estimating $\rr$ more difficult, especially in terms of the CIs since the usual assumption of Gaussian noise may be violated due to the non-symmetric nature of the noise.
-We examine \wxxsir methods to estimate $\rr$ which include least squares optimization, polynomial approximations, estimates from incidence and variance, a log-linear model, a Markov chain method, and a simple Bayesian method.
-Although, this is not an exhaustive list of methods used to estimate $\rr$ from the SIR model, we believe it covers common and popular methods used throughout the decades.
-We overview the details of methods in Section \ref{sec:methods}.
-
-%% the importance of variance and testing
-Each of the \wxxsir methods results in an estimate of $\rr$, $\hat{\rr}$.
-We also estimate a 95\% CI for the  $\rr$ estimate.
-Although, many researchers report intervals via sensitivity analysis (how the estimate changes with small variations in the underlying parameters), the ranges of these parameters often do not incorporate any distributional assumptions about these estimated ranges and hence do not correspond to 95\% CIs.
-We provide three statistical methods to estimate the standard error of $\hat{\rr}$ and use the interval $\left [\hat{\rr} \pm 2\hat{SE}\left (\hat{\rr} \right ) \right ]$ as our 95\% CI.
-We hope that these intervals will provide guidance on how to create more meaningful intervals when estimating $\rr$ and discerning which diseases are more severe than others, and whether the true $\rr >1$.
-
-We compare the the estimates of $\rr$ for the \wxxsir methods using a series of simulations.
-We generate different simulated data sets under different noise assumptions and estimate $\rr$ using each of these data sets.
-We then analyze estimates of $\rr$ while varying individual parameters of the model including the infection and recovery parameters, the total amount of data, the initial percent of infectious and susceptible individuals, the standard deviation of the noise used to generate the simulated data, and the total number of individuals in a population.
-We also analyze the model under misspecification, that is when a model other than the SIR model is used to generate the simulated data sets.
-Finally, we compare our variance methods to one another and analyze the coverage of our methods.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%RESULTS
-%%%%%%%%%%%%%%%%%
-
-Overall, we recommend based on our analysis:
-\begin{itemize}
-  \item LS and ReLS provide the most accurate estimates of $\rr$
-  \item LL results in close point estimates but occasionally has too small standard errors
-  \item The specific MC and SB methods presented here do not seem to fit our generated data well
-  \item There is not much difference in estimations of $\rr$ due to different error generating processes
-  \item Very small values of $\gamma$ can result in very poor estimates of $\rr$
-  \item 100 data points seems to be a reasonable number of data points to have enough power for testing that $\rr > 1$ and 50 seems to be too few
-  \item The larger the number of initially infectious individuals is, the smaller the CIs for $\rr$ are
-  \item The noisier the epidemic data is, the larger our CIs for $\rr$ are
-  \item Estimations of $\rr$ are fairly robust to the misspecified data sets we presented, at least for the methods of LS, ReLS, and LL
-  \item Block bootstrap matches the empirical variance through repeated sampling.
-  \end{itemize}
-
-$\rr$ is a difficult quantity to estimate, but we hope that our methods of estimation and simulations provide some guidance for researchers on how to generate statistically reliable estimates for $\rr$ and its CIs.
-Our analysis is limited in scope to the SIR model, and in the future we would like to see how methods of estimating $\rr$ compare for models such as the SEIR and beyond.
-
-%% results of simulations:
-%% - data
-%% - what parameters make a difference
-%% - what methods perform best when
-%% - misspecification
-%% variance
-%% beyond the sir model
-
-The SIR estimation methods generally give reliable estimates when the data are from an underlying SIR model with little or no noise.
-However, the estimates from the various methods are generally much worse when the magnitude of the noise is larger, when the parameters of the model are unusual, and when the model is misspecified.
-We also see a large difference between the $\rr$ estimates and standard errors on the exact same data sets as well, even though all derived from the same model.
-Thus, when estimating $\rr$ based on an SIR or any other kind of compartment model, we need to take careful consideration.
-
-\bibliographystyle{apa}%Choose a bibliograhpic style
-\bibliography{Master}
-
-\appendix
-
-\section{Beyond the SIR Model}
-
-\subsection{Exponential Growth (EG)}\label{sec:expgrowth}
-%% nishiura, majumder, IDEA model -- sum of squares (santillana2016)
-%% An IDEA for short term outbreak projection: nearcasting using the basic reproduction number.  (fisman2013, fisman2014)
-%% chowell 2008 - parametric bootstrap for a model they made up, but sort of empirical.
-Many researchers employ an assumption of exponential growth of the number infectious through times $t=0, 1, \dots, T^* <T$ \citep{wallinga2007generation,fisman2014,nishiura2016,majumder2016,towers2016}.
-\cite{wallinga2007generation} report that the effective reproduction number $\mathcal{R}_t$ and hence the initial reproduction number $\rr$ may derived using the fact that infection ``counts increase exponentially in the initial phase of an epidemic.''  We then have to estimate $r$, the \textit{per capita} change in the number of new cases per unit of time and $\omega$ the serial interval, the distribution of time between a primary and secondary infection.
-Then, we have
-\begin{align}\label{eq:lotka}
-\rr = \exp{(r \omega)}.
-\end{align}
-Equation \eqref{eq:lotka} is derived from a demographic view using the Lotka-Euler survival equations which come from the fields of demography, ecology, and evolutionary biology.
-
-\cite{wallinga2007generation}, instead, use a moment generating function expression for $\rr$.  With $\omega(t)$ as the serial interval, then
-\begin{align*}
-\rr^{-1} &= \frac{1}{M(-r)} = \int_{a=0}^\infty e^{-rt}\omega(t)dt.
-\end{align*}
-\cite{wallinga2007generation} specifically derive an estimate for $\rr$ for the SIR model.
-They assume that the rate of leaving the infectious stsate, $\gamma$ is constant as is the rate of making contacts during the infectious stage.
-Then, the duration of a serial interval is an exponential distribution with mean $\gamma^{-1}$, the moment generating function is then,
-\begin{align*}
-  M(-r) = \frac{\gamma}{\gamma + r},
-\end{align*}
-and the resulting estimate of $\rr$ is
-\begin{align}\label{eq:r0-eg}
-  \hat{\rr} = 1 + \frac{\hat{r}}{\hat{\gamma}},
-\end{align}
-where $\hat{r}$ and $\hat{\gamma}$ are estimates of the $r$ and $\gamma$, respectively.
-The parameter $\hat{\gamma}$ is often assumed to be fixed or tried with a sequence of serial intervals \citep{majumder2016}.
-We may estimate $\hat{r}$ using ordinary linear regression, namely,
-\begin{align*}
-  \log \left  (\frac{Y(t)}{Y(0)} \right )&= rt + \epsilon_t\\
-  \epsilon_t &\sim N(0, \sigma^2),
-\end{align*}
-which results in
-\begin{align}\label{eq:exp-growth}
-  \hat{r} &= \frac{\sum_{t=0}^{T^*}\left  (\frac{Y(t)}{Y(0)} \right )}{\sum_{t=0}^{T^*}t}.
-\end{align}
-
-For the above estimate of $\rr$, we observe the duration of serial intervals in a period of exponential growth of the number of infectious individuals.
-Deciding when exponential growth occurs and which data points to use to estimate $\rr$ may be difficult.
-We can use sensitivity analysis to see how the estimate $\rr$ changes as we increase or decrease $T^{*}$, the number of time points used.
-The reader may notice some similarities to the estimate for $\rr$ in Equation \eqref{eq:r0-eg} and from the log-linear model in Equation \eqref{eq:harko_lin}.
-This method is a simplification of the SIR model as the $R$ compartment is not used to estimate $r$, only the recovery rate $\gamma$.
-
-An advantage to this method are its (and similar adaptations) prevalence in the literature, which allows one to to compare estimates $\rr$ directly to one another.
-Another advantage is its flexibility.
-Here, we assumed the serial interval has an exponential distribution, but this is not a requirement and other distributions may be used (see \cite{wallinga2007generation}).
-However, a disadvantage to this method is that is a simplification of the SIR model and we have to decide how many time points to use to estimate the exponential growth rate of infectious.
-Another disadvantage to this method is that one also has to decide how to treat the additional paramter $\gamma$, whether to treat it as fixed, a range of fixed values, or even having a probability distribution.
-
-\subsection{Incidence Decay and Exponential Adjustment (IDEA)}\label{sec:idea}
-The Incidence Decay and Exponential Adjustment (IDEA) model is a way to account for a declining effective reproduction number of time due to both depletion of susceptible individuals and spontaneous and planned control activities and behaviors'' \citep{fisman2013}.
-They model the incidence counts during the $t$th interval, $J(t),$ as
-\begin{align*}
-  J(t) &= \left [ \frac{\rr}{(1 + d)^t}\right ]^t,
-\end{align*}
-where $d$ is a dampening or discount factor.  As $d$ increases, the incidence at time $t$  decreases. The estimate for $\rr$ is obtained by minimizing the sum of squares over $\rr$ and $d$,
-\begin{align*}
-  \hat{\rr} &=  \arg \min_{\rr,d} \sum_{t=0}^{T^*}\left (J_{obs}(t) - J(t) \right)^2.
-\end{align*}
-The method presented in \cite{fisman2013} is a mathematical, rather than statistical model.
-Sensitivity analysis is conducted with respect to $T^*$, $d$, and $\rr$, and lower and upper bounds are reported as a result of this analysis.
-
-Advantages of this method include its usefulness in estimating $\rr$ \citep{fisman2014,majumder2016}.
-Additionally, this model may be employed to address the effect of control measures and better estimate the final size of an epidemic.
-A disadvantage to this model is that its upper and lower bounds do not correspond to CIs.
-Also, in this method we must select the interval size of the incidence counts.
-
 \end{document}
-
-%%%%%%%%%%%%%%%%%%%%%%
-\textbf{Title:}
-
-\textbf{Author:}
-
-\textbf{Citation:}
-
-\textbf{Major themes:}
-
-\textbf{Notes:}
-\\
-%%%%%%%%%%%%%%%%%%%%%%
-
-\begin{figure}[h]
-\begin{center}
-\includegraphics[width=4in]{images/mvhw7_3c.pdf}
-\end{center}
-\caption{
-Biplots of the different continuous variables and their correlations.}\label{fig3c}
-\end{figure}
-
-%% two pictures whoa
-
-\begin{figure}[h]
-\centering
-
-\begin{figure}[h]
-\centering
-
-\begin{subfigure}{.5\textwidth}
-  \centering
-  \includegraphics[width=1\linewidth]{images/mt_eda_cont_hists.pdf}
-  \caption{Histograms of Arrival Delay and continuous covariates.
-Arrival delay seems to have a right skewed distribution.
-This may indicate that we will be transforming this variable later on.
-After transforming Air Time and Distance by a log transformation, we don't really seem to have many outliers in our covariates.
-We seem to have outliers in the CRS Dep.
-Time and Arrival Time; however, time is cyclical and so these are not, in fact outliers.}
-  \label{hists}
-\end{subfigure}%
-\begin{subfigure}{.5\textwidth}
-  \centering
-  \includegraphics[width=1\linewidth]{images/mt_eda_cont_hists.pdf}
-  \caption{\textcolor{red}{Placeholder}}
-  \label{tabs1}
-\end{subfigure}
-\caption{}
-\end{figure}
-
-\begin{table}
-\begin{tabular}{l | a | b | a | b}
-\hline
-\rowcolor{LightCyan}
-\mc{1}{}  & \mc{1}{x} & \mc{1}{y} & \mc{1}{w} & \mc{1}{z} \    \hline
-variable 1 & a & b & c & d \    variable 2 & a & b & c & d \\ \hline
-\end{tabular}
-\end{table}
-
-%%
-\begin{figure}
-\centering
-\begin{subfigure}{.5\textwidth}
-  \centering
-  \includegraphics[width=1\linewidth]{images/resids_full.pdf}
-  \caption{}
-  \label{residsf}
-\end{subfigure}%
-\begin{subfigure}{.5\textwidth}
-  \centering
-  \includegraphics[width=1\linewidth]{images/diags_full.pdf}
-  \caption{ }
-  \label{diagsf}
-\end{subfigure}
-\caption{}
-\end{figure}

### 2018-07-03 paper/Version9/draft_v9.tex
finally on git

--- 2018-07-03 paper/Version10/draft_v10.tex
@@
 
-\textit{We compare the different ways to estimate $\rr$ to see which ones give more accurate and reliable estimates.}  As there is no way to know the ``true'' $\rr$ of an actual disease we compare our methods by fitting them to simulated data sets and look at the resulting $\rr$ estimates and standard errors.
+\textit{With the many different ways to estimate $\rr$, we want to compare them all to see which ones give more accurate and reliable estimates.}  As there is no way to know the ``true'' $\rr$ of an actual disease we compare our methods by fitting them to simulated data sets and look at the resulting $\rr$ estimates and standard errors.
 Our main comparisons are done with data generated from the SIR model with random error added so that we can compare model estimates to a known true $\rr$ value.
@@
 %% SG: delta_t is normally distributed with sd sigma and similarly for other methods.  How and why did we pick the numbers that we did?
-
-\subsubsection{SIR Data}
-
-\textit{The datasets from the SIR model are simulated from the ODEs: \begin{align}
-	\systeme{\frac{dX}{dt} = -\frac{\beta XY}{N}, \frac{dY}{dt} = \frac{\beta XY}{N} - \gamma Y, \frac{dZ}{dt} = \gamma Z}.
-	\end{align}
-}
-
-We use arbitrarily selected the parameters to generate the time series data set from the SIR model.
-We first generate "baseline" data from which we can make comparisons by changing the  conditions under which the data were generated.
-The baseline SIR data are simulated with the following conditions:
-
-\begin{center}
-
-	$\beta$ = 0.006, $\gamma$ = 0.003; $\rr$ = 2
-
-	Number of time steps: 365
-
-	Starting X size ($X_0$): 99950
-
-	Starting Y size ($Y_0$): 50
-
-	Starting Z size: 0
-
-	Total population: 100000.
-
-\end{center}
-
-%% Is this enough detail on describing the initial S, I, and R, or are we looking for something more specific?
-
-\textbf{Our baseline initial size for $Y_0$ is 0.05\% of the total population in our simulated environment, which is fairly large compared to the starting point for the data in most epidemics.
-However, I/we felt that this was necessary in order to see the differences between the models.}
-
-We then generate random errors that are added to the X and Y compartments at each time step in order to simulate some form of randomness.
-\textbf{The sum of these was then subtracted from the Z compartment in order to keep the total population in the system constant.} We generated data sets with two different types of errors: Gaussian and autoregressive.
-\textbf{The Gaussian errors are independently distributed with mean 0 and constant variance for the entire length of the time series.
-The autoregressive errors are generated from a first-order process of the form: $$ \epsilon_{t} = \rho * \epsilon_{t-1} + \delta_{t}$$, where $\rho$ is fixed throughout and $\delta_{t} \sim N(0, \sigma)$ with constant $\sigma$.
-The values of the variances are chosen somewhat arbitrarily, although they are intended to be fairly small at first.}
-
-We also add a monotonicity adjustment to the X and Z compartments with both of the error models.
-\textbf{This is done using the Pool Adjacent Violators Algorithm (PAVA); the algorithm checks if the observation at each time step is greater than or equal to the previous one, and changes the value of both observations to be equal to the mean of the two if the condition is not fulfilled.} This gives us a total of four error models: autoregressive (Auto), autoregressive monotonic (AutoM), Gaussian (Norm), and Gaussian monotonic (NormM).
-For our baseline data, we have ($\sigma_X, \sigma_Y$) = (10, 1).
-We simulate 5 different baseline sets to see how consistent each of the models are.
 
@@
 
+\subsubsection{SIR Data}
+
+\textit{The datasets from the SIR model are simulated from the ODEs: \begin{align}
+	\systeme{\frac{dX}{dt} = -\frac{\beta XY}{N}, \frac{dY}{dt} = \frac{\beta XY}{N} - \gamma Y, \frac{dZ}{dt} = \gamma Z}.
+	\end{align}
+}
+
+We use arbitrarily selected the parameters to generate the time series data set from the SIR model.
+We first generate "baseline" data from which we can make comparisons by changing the  conditions under which the data were generated.
+The baseline SIR data are simulated with the following conditions:
+
+\begin{center}
+
+	$\beta$ = 0.006, $\gamma$ = 0.003; $\rr$ = 2
+
+	Number of time steps: 365
+
+	Starting X size ($X_0$): 99950
+
+	Starting Y size ($Y_0$): 50
+
+	Starting Z size: 0
+
+	Total population: 100000.
+
+\end{center}
+
+%% Is this enough detail on describing the initial S, I, and R, or are we looking for something more specific?
+
+\textbf{Our baseline initial size for $Y_0$ is 0.05\% of the total population in our simulated environment, which is fairly large compared to the starting point for the data in most epidemics.
+However, I/we felt that this was necessary in order to see the differences between the models.}
+
+We then generate random errors that are added to the X and Y compartments at each time step in order to simulate some form of randomness.
+\textbf{The sum of these was then subtracted from the Z compartment in order to keep the total population in the system constant.} We generated data sets with two different types of errors: Gaussian and autoregressive.
+\textbf{The Gaussian errors are independently distributed with mean 0 and constant variance for the entire length of the time series.
+The autoregressive errors are generated from a first-order process of the form: $$ \epsilon_{t} = \rho * \epsilon_{t-1} + \delta_{t}$$, where $\rho$ is fixed throughout and $\delta_{t} \sim N(0, \sigma)$ with constant $\sigma$.
+The values of the variances are chosen somewhat arbitrarily, although they are intended to be fairly small at first.}
+
+We also add a monotonicity adjustment to the X and Z compartments with both of the error models.
+\textbf{This is done using the Pool Adjacent Violators Algorithm (PAVA); the algorithm checks if the observation at each time step is greater than or equal to the previous one, and changes the value of both observations to be equal to the mean of the two if the condition is not fulfilled.} This gives us a total of four error models: autoregressive (Auto), autoregressive monotonic (AutoM), Gaussian (Norm), and Gaussian monotonic (NormM).
+For our baseline data, we have ($\sigma_X, \sigma_Y$) = (10, 1).
+We simulate 5 different baseline sets to see how consistent each of the models are.
+
 \subsubsection{Other Models}
@@
  The data from the linear SIR model were simulated with parameters $\beta$ = 0.06 and $\gamma$ = 0.05, such that $\rr$ = 1.2.
-This model will allow us to compare how the SIR models perform under misspecified differential equations.
-
- \textbf{Random error is added to the data generated here as well using the same 4 error models as above.
+
+ \textbf{Random error is added to the data generated here as well, using the same 4 error models as above.
 The other specifications for these datasets are the same as for the baseline classic SIR data, namely the variance of the added errors and starting compartment sizes.}
@@
 
-Standard errors for the maximum likelihood and branching process models are calculated using the negative reciprocal of the Fisher information.
+\textbf{Standard errors for the maximum likelihood and branching process models are calculated using the negative reciprocal of the Fisher information.
 The sequential Bayes model uses the standard deviation from the posterior distribution.
 We use the jackknife method to estimate the standard error of the SIR smoothed max and the SIR smoothed IPR methods.
-The jackknife calculates the standard error using the standard deviation of estimates of $\rr$ after leaving out a single data point.
-
-Some of the models use only one single point in time to estimate $\rr$ and therefore do not seem to have an innate source of variance.
-We currently do not have standard errors for the SIR Max and SIR IPR models.
+The jackknife calculates the standard error using the standard deviation of estimates of $\rr$ after leaving out a single data point.}
+
+\textbf{Some of the models use only one single point in time to estimate $\rr$ and therefore do not seem to have an innate source of variance.
+We currently do not have standard errors for the SIR Max and SIR IPR models.}
 
@@
 
-\textbf{STILL NEED TO ACTUALLY DO DPE}
-
-We do not have a direct way to estimate $/rr$ using the survival function method from the data we have.
-The method requires a density estimate of the survival function at each point in time.
-Either we would need to have some prior knowledge or assumptions about what the survival function for a particular disease was, or we would need to search over a space of possible function and possible parameter values to estimate $/rr$.
-For similar reasons, we can not estimate $/rr$ using the Von Forester equations using the data we have, as we would need to know the infectivity of a diseased individual over time.
-This may be possible for a known, well studied disease, but may not be  For the direct parameter estimation method, assuming the population is homogeneous and that we know the average duration of infection, we can estimate the probability of transmission parameter and its standard error by averaging over the ratio between percent change in the susceptible compartment and the prevalence of the disease at each time step.
+%% How do we want to estimate this?????????
+%% SG:
+%% 3.1.1:  So I think this one we should mention there is no direct way to estimate this from the simple data we have.
+Alternatively, you could do a parameter sweep with different beta distributions for F(a) and beta for b(a).
+%% 3.1.2:  So assuming the population is homogeneous, k = N, D = gamma^{-1}.
+So we have to estimmate $b$ from the data.
+So I would average over (change in % susceptible ) / old # infected and use that as $b$.
+%% 3.1.3:  No clue.
 
@@
 \paragraph{SIR Models}
-%In general, the models derived from the SIR ordinary differential equations are relatively accurate.
+In general, the models derived from the SIR ordinary differential equations are relatively accurate.
 Most estimates are within the range of $\pm$ 0.2 of the true $\rr$ for the baseline SIR datasets.
@@
 
-%The SIR models as a whole also do fairly well with changing the starting number of infected people.
+The SIR models as a whole also do fairly well with changing the starting number of infected people.
 However, they tend not to do as well when the variance of the added errors is very large; in these situations, many of the models tend to have very large standard errors, sometimes on the order of $10^{1}$ or $10^{2}$.
 
-%The exceptions to the above are the linear, incidence-prevalence ratio, and Markov chain models.
+The exceptions to the above are the linear, incidence-prevalence ratio, and Markov chain models.
 The estimates tend to be far from the true $\rr$, and they fluctuate greatly between the baseline data sets.
@@
 
-%Many of the SIR models do not work as well with the data that come from the polynomial linear models.
+Many of the SIR models do not work as well with the data that come from the polynomial linear models.
 Specifically, for the 0th and 1st order model datasets, the estimates can be unreasonably large or negative.
@@
 
-%All of these models use optimization to estimate the parameters that are used to calculate $\rr$.
+All of these models use optimization to estimate the parameters that are used to calculate $\rr$.
 They tend to be fairly accurate and consistent on the baseline data, often as good as or better than those from the various SIR models.
@@
 
-%We have assumed for these models that the number of infections follows an exponential curve throughout the entire time frame of the data sets.
+We have assumed for these models that the number of infections follows an exponential curve throughout the entire time frame of the data sets.
 This assumption certainly does not hold for the data from the 0th and 1st order linear models, and may not hold for some of the other data sets either.
 
-%These models are generally not as accurate as the compartment models above.
+These models are generally not as accurate as the compartment models above.
 Most of them, except for the initial size model, tend to underestimate $\rr$, with estimates between 0.8 and 1.0.
@@
 
-%The branching process model (the only one in this category so far) is not very accurate.
+The branching process model (the only one in this category so far) is not very accurate.
 In particular, for the SIR datasets, the $\rr$ estimates are less than 1, while the true $\rr$ is greater than 1.
@@
 
-Below, we compare the estimates and standard errors of $\rr$ given by each of the models for a few selected data sets using forest plots; they show the point estimates and 95\% confidence intervals of $\rr$ from each model for a particular data set.
+Below, we compare the estimates and standard errors of $\rr$ given by each of the models for a few selected data sets.
+The forest plots below show the point estimates and 95\% confidence intervals of $\rr$ from each model for a particular data set.
 The true $\rr$ value is also shown by the vertical red line in the graphs.
+
 All of the figures below show results from data sets that were simulated with added Gaussian errors.
 The plot on the top left shows the estimates from one of the baseline data sets.
-For the most part, the $\rr$ values range from approximately 0.8 to 2.
-We note that many of the models estimate an $/rr$ value of approximately 1, and some estimate that $/rr$ is less than 1.
-This would indicate that the infection disease is growing very slowly or is actually dying out, when in reality the disease is spreading at a fairly rapid rate.
-The confidence intervals are very small relative to the estimates for the most part; exceptions to this include the SIR IPR and LMA models as well as the Markov chain model.
-
-The plot on the top right shows the estimates for the data set where $Y_0$ is increased to 1000.
-The $\rr$ estimates are almost all further away from the true $\rr$ value compared to the results from the baseline data sets.
+For the most part, the $\rr$ values range from approximately 0.8 to 1.4.
+We note that the range of values includes $\rr$ values of less than 1, meaning that some models predict that a disease is not spreading when it actually is.
+However, on the whole, the variation of estimates between models is smaller than we originally expected.
+The confidence intervals are very small relative to the estimates for the most part; exceptions to this include the SIR IPR and the Markov chain model as the exception.
+
+The plot on the top right shows the estimates for the data set where $Y_0$ is increased to 100.
+The $\rr$ estimates generally seem to be further away from the true $\rr$ value compared to the results from the baseline data sets, and the overall .
 The confidence intervals are also more often than not larger than those in the baseline.
@@
 The bottom left plot shows the estimates for the data set where the error added to the generated data is increased.
-For most of the models, the estimates of $\rr$ decrease slightly, and overall they do not change very much compared to the baseline data set.
-The confidence intervals for this data set increase substantially for some models, such as the SIR LMA, while they remain approximately the same as the baseline data sets for several of the others.
+For most of the models, the estimates of $\rr$ do not change very much compared to the baseline data set.
+The confidence intervals for this data set increase for some of the models, such as the SIR LMA, while they remain approximately the same as the baseline data sets for most of the others.
 
@@
 On the other hand, models such as the SIR maximum and final size seem to be more flexible to different trajectories.
+[The error bars do not show up if they go outside of the x-axis limits - how do we fix this?]
 

### 2018-07-03 paper/draft_v6.tex
finally on git

--- 2018-07-03 paper/Version9/draft_v9.tex
@@
 \usepackage[table,xcdraw]{xcolor}
-\usepackage{float}
 
@@
 What has been called ``arguably the most important quantity in the study of epidemics'', $\mathcal{R}_0$, the reproduction number (by convention pronounced ``R-naught''), for nearly a century, has been an elusive quantity for epidemic modelers to estimate \citep{Heesterbeek2002}.
-Defined by \citet{anderson1992}, $\rr$ is the ``the average number of secondary infections produced when one infected individual is introduced into a host population where everyone is susceptible.''  $\rr$, in some ways, summarizes an entire outbreak of a disease; it  is used to assess whether a disease will break out and what percentage of the population needs to be vaccinated to avoid such an outbreak.
+Defined by \citet{anderson1992} is the ``the average number of secondary infections produced when one infected individual is introduced into a host population where everyone is susceptible.''  $\rr$, in some ways, summarizes an entire outbreak of a disease as this quantity is used to assess if a disease will break out and what percentage of the population needs to be vaccinated to avoid such an outbreak.
 Despite a clear definition of $\rr$, epidemiologists have struggled to create a standard  estimator for $\rr$  \citep{hethcote2000}.
@@
 Additionally, we describe  interval estimates for the above methods for estimating $\rr$ in order to better answer the question of whether $\rr > 1$, which implies an outbreak of the disease.
-These methods are the delta method and posterior distribution.
+These methods are (1) the delta method, (2) sensitivity analysis, and (3) posterior distribution.
 
@@
 The first set of methods are those of the survival function, which is closely  tied to the history of $\rr$.
-The survival function, which originates from the field of demography in the late 1800s, describes how many female offspring a woman would produce \citep{dietz1993estimation}.
+The survival function, which originates from the field of demography in the late 1800s, described how many female offspring a woman would produce \citep{dietz1993estimation}.
 In demography, we have
@@
 However, having a post-disease estimation of $\rr$ may be useful for preventing future outbreaks of the disease and aids in comparing the severity of many diseases.
-Another advantage of this method is that it is intuitive and independent of assumptions of how the disease is transmitted from person to person.
+Another advantage of this method is it intuitive and independent of assumptions of how the disease is transmitted from person to person.
 We describe four methods which estimate $\rr$ from the survival function:  the survival function, direct estimation, and direct parameter estimation.
@@
 
-Direct parameter estimation, an approach for parameters that are readily available to be estimated, is a `plug-in' estimator approach as described by \cite{lipsitch2003} and \cite{dietz1993estimation}.
-This method, like the survival function, is highly interpretable.
+This approach for parameters that are readily available to be estimated, direct parameter estimation, is a `plug-in' estimator approach as described by \cite{lipsitch2003} and \cite{dietz1993estimation}.
+This method has an intuitive derivation, similar to that of the survival function.
 We assume that we can estimate $\rr$ if we know how many others an individual comes into contact $(k)$, the probability of transmission given contact $(b)$, and how long the disease lasts within an individual $(D)$.
@@
 An advantage of this method is that it reduces the problem to three parameters that we can conceivably collect data on.
-A disadvantage is that to estimate $\rr$ alone,  data is only used from the initial outbreak of the disease, so often we have a small sample size, resulting in large confidence sets or credible intervals.
+A disadvantage is thatto estimate $\rr$ alone,  data is only used from the initial outbreak of the disease, so often we have a small sample size, resulting in large confidence sets or credible intervals.
 
@@
 The right hand side represents the net loss of cells.
-This derivation is based on the ``reasoning that the number of cells which are in the age category $(a, a + \Delta a)$ at time $t$, is equal to $n(t,a) \Delta a$'' \citep{trucco1965}.
+This derivation is based on the ``reasoning that the number of cells which are in the age category $(a, a + \Delta a)$ at time $t$, is equal to $n(t,a) \Delta a$.
 \cite{fraser2004factors} derived an expression for $\rr$ from the von Foerster equation and the additional boundary equation
@@
 \end{align*}
-with initial conditions $Y(0,0) = Y_0$ and $Y(t, \tau) = 0$ when $t < \tau$.  Then the reproduction number is given by Equation \ref{eq:vfr0},
+and initial conditions $Y(0,0) = Y_0$ and $Y(t, \tau) = 0$ when $t < \tau$.  Then the reproduction number is given by Equation \ref{eq:vfr0},
 \begin{align}\label{eq:vfr0}
@@
 Again, this model is more mathematical in nature than practical as $\beta (\tau)$ is difficult to estimate.
-To estimate $\rr$, \cite{fraser2004factors} make an approximation that $Y(t, \tau) = K(\tau) \exp (rt)$ and further adapt the method to include prevention strategies.
-The main function to be estimated is $\beta (\tau)$.
+To estimate $\rr$, \cite{fraser2004factors} make an approximation that $Y(t, \tau) = K(\tau) \exp (rt)$ and further adapt the method to include prevention strategies, but the main function to be estimated is $\beta (\tau)$.
 Estimating $\beta (\tau)$ has many of the same issues as trying to estimate $b(a)$ in the survival function in that it is difficult to estimate while the disease is ongoing but is also advantageous in that it does not place assumptions on how the disease is transmitted throughout the population.
@@
 
-First introduced by \cite{Kermack700}, a large and common set of methods used estimate $\rr$ is that of fitting compartment models (CMs) to disease data.
+First introduced by \cite{Kermack700}, a large and the most common set of methods used estimate $\rr$ is that of fitting compartment models (CMs) to disease data.
 As described in \cite{daley2001epidemic},  CMs model how objects move from one discrete state to another over time.
@@
 The final method described in this section, the next generation operator (NGM) is a general method to derive $\rr$ from any valid CM.
-A particular advantage to CMs compared to survival function methods is that they are easier to estimate as the data required is typically collected about diseases: namely, incidence counts.
+A particular advantage to CMs compared to survival function methods is that it easier to estimate as the data required is typically collected about diseases: namely, incidence counts.
 
@@
 
-In words, susceptible individuals become infected at a rate that is proportional to the percentage of infected individuals multiplied by $\beta$, the infection rate, and the number of susceptible individuals.
+In words, susceptible individuals become infected at a rate that is proportional to the proportion of infected individuals multiplied by $\beta$, the infection rate, and the number of susceptible individuals.
 Infected individuals recover at a rate of $\gamma$ multiplied by the number of infected individuals.
@@
   \frac{dY}{dt} &> 0 \\
-  \frac{\beta X Y}{N}  - \gamma Y &> 0 ,\\
+  \frac{\beta X Y}{N}  - \gamma I &> 0 ,\\
   Y \left ( \beta \frac{X}{N} - \gamma \right ) & > 0
@@
 We can use gradient descent or another optimizer to find estimates for these parameters.
-If we assume $Y= f(t) + \epsilon_t$ with $\epsilon_t \overset{iid}{\sim} N(0, \sigma^2)$ where $f(t) = \int_0^t \frac{dY}{ds} ds$, then this this method is equivalent to finding the maximum likelihood estimator (MLE) of $\beta$ and $\gamma$.
+If we assume $Y= f(t) + \epsilon_t$ with $\epsilon_t \overset{iid}{\sim} N(0, \sigma^2)$ where $f(t) = \int_0^t \frac{dY}{dt} dt$, then this this method is equivalent to finding the maximum likelihood estimator (MLE) of $\beta$ and $\gamma$.
 However, we note that the the $\epsilon_t$ are certainly not independent from one another and so valid inference is still an issue in this framework.
@@
 \end{align*}
-We use the \(\rr\) directly from the above estimation process, which again can be done with gradient descent or another optimization process.
+We use the \(\rr\) directly from the above estimation process, which again be done with gradient descent or another optimization process.
 Again, we cannot assume $\rr$ has a normal distribution as the errors are correlated.
@@
 Specifically, we fit two linear polynomials in \(t\) with degree \(K= 10\) to \(X_{obs}\)
-and \(Y_{obs}\) using least squares to find the coefficients $\{(\hat{x}_k,
-\hat{y}_k)\}_{k=1, \dots, K}$,
+and \(Y_{obs}\) using least squares to find the coefficients (\(\hat{x}_k\),
+\(\hat{y}_k\)),
 \begin{align*}
@@
 \end{align*}
+
 Following,  \(\rr\) is derived from the ODEs in Equation \eqref{eq::sir},
 \begin{align}
-  - \frac{X^\prime}{Z^\prime}&= \rr \frac{X}{N} \nonumber\\
-\rr &=       \frac{X}{ X + Y} \cdot \frac{N}{X} \nonumber\\
-  \hat{\rr} &= \frac{\hat{X}^\prime(0)}{ \hat{X}^\prime(0) + \hat{Y}^\prime(0)} \cdot \frac{N}{\hat{X}(0)}. \nonumber
+  - \frac{X^\prime}{Z^\prime}&= \rr \frac{X}{N}\\
+\rr &=       \frac{X}{ X + Y} \cdot \frac{N}{X}\\
+  \hat{\rr} &= \frac{\hat{X}^\prime(0)}{ \hat{X}^\prime(0) + \hat{Y}^\prime(0)} \cdot \frac{N}{X(0)}.
   \end{align}
@@
 Besides optionally deciding on the degree of polynomials to fit, this model is simple to implement and gives very comparable results to using least squares with the SIR model.
-The time $t=0$ to best capture the initial outbreak.
+The time $t=0$ is used as this is best capture the initial outbreak.
 Again, we cannot use this model for inference, as our error assumption is violated.
@@
 \subsubsection{Max of Data (SIR Max)}\label{max-of-data}
-The I ($Y$) compartment of the SIR model has one peak.    We note that \(Y^\prime(t) = \frac{\beta XY}{N} - \gamma Y = 0\) when \(\rr = \frac{\beta}{\gamma} = \frac{N(t)}{X(t)}\). The max
+The I compartment of the SIR model has one peak.    We note that \(Y^\prime(t) = \frac{\beta XY}{N} - \gamma Y = 0\) when \(\rr = \frac{\beta}{\gamma} = \frac{N(t)}{X(t)}\). The max
 of \(Y\) occurs when \(Y^\prime = 0\) and hence
@@
 Here we assume that the time step is small enough to approximate the incidence.
-The advantage of this method is that incidence data is generally readily available as is prevalence data for certain diseases such as HIV.
+The advantage of this method are that incidence data is generally readily available as is prevalence data for certain diseases such as HIV.
 However, as one is required to have prior knowledge about $\gamma$, it may be easier to directly estimate $\rr$ with one of the many other methods described that does not require a prior knowledge about $\gamma$.
@@
 
-In this framework, the number of susceptibles at the next step, $X_{t}$, has a Binomial distribution based on the contacts with the current number of infected, $Y_{t}$ and the current number of susceptibles.
+In this framework, the number of susceptibles at the current time, $X_{t+1}$, has a binomial distribution based on the contacts with the current number of infected, $Y_t$ and the current number of susceptibles.
 That is $X_{t+1} \sim \text{Binom}(X_t, \alpha^{Y_t}$), where $\alpha$ is the probability of avoiding infection from an infective.
@@
 The advantages of this method are that of the ability to obtain an entire posterior distribution, whereas many other methods are difficult to even find an estimate of the variance.
-Disadvantages include strict assumptions about the distributions and computational time required to estimate the relevant parameters.
+Disadvantages include strict assumptions about the distributions and computational time required.
 
@@
 \end{align*}
-For all these above estimates of $\rr$, we observe the duration of serial intervals in a period of exponential growth.
+For all these above estimations of $\rr$, we observe the duration of serial intervals in a period of exponential growth.
 Deciding when exponential growth occurs and which data points to use to estimate $\rr$ is difficult.
@@
 We write the calculation for $\rr$ under contact tracing in Equation \ref{eq:r0_contacttracing}, where $X_0$ is a random variable denoting the number of individuals in the population individual 0 infects.
-Here $X(0)=N$ indicates that the initial number of susceptibles is $N$, the size of the population,
+Here $S_0=N$ indicates that the initial number of susceptibles is $N$, the size of the population,
 \begin{align}\label{eq:r0_contacttracing}
-\rr = E[ X_0 | X(0) = N].
-\end{align}
-\cite{eames2003} describe contact tracing as an ``extreme form of targeted control, where the potential next-generation cases are the primary focus.'' That is, one focuses on treating the contacts of an infected individual rather than applying uniform treatment across a population.
+\rr = E[ X_0 | S_0 = N].
+\end{align}
+\cite{eames2003} describe contact tracing as an ``extreme form of targeted control, where the potential next-generation cases are the primary focus,'' that is, one focuses on treating the contacts of an infected individual rather than applying uniform treatment across a population.
 
@@
 
+\subsubsection{Sensitivity Analysis}\label{sensitivity-analysis}
+
+Sensitivity analysis is used to measure how robust estimates are to change.
+If small changes in our estimated parameters do not change our estimates for the number of infected by much, then we may prefer this method to another.
+For the SIR model, we estimate \(\frac{\partial Y}{\partial \beta}\) (or \(\frac{\partial Y}{\partial \rr}\) when using reparametrized least squares) and \(\frac{\partial Y}{\partial \gamma}\) at each time point \(t\) using the estimated \(\beta\) or \(\gamma\) from the estimation process where $Y(t)$ stands for the number of infected at time $t$.
+More generally,  let $Y(t_i) = f(t_i, \theta)$ where $\theta$ is a vector of parameters.
+Then an approximation of the covariance matrix is \((\chi^T \chi)^{-1}\) where \(\chi_{ij} = \frac{ \partial Y(t_i)}{ \partial \theta_j}\) \cite{capaldi2012}.
+Advantages of this method include that it is a standard way to estimate the variance of $\rr$.
+
 \subsubsection{Posterior Distribution}
@@
 
-\subsection{Comparing Models}
-
-\textit{With the many different ways to estimate $\rr$, we want to compare them all to see which ones give more accurate and reliable estimates.}  As there is no way to know the ``true'' $\rr$ of an actual disease we compare our methods by fitting them to simulated data sets and look at the resulting $\rr$ estimates and standard errors.
-Our main comparisons are done with data generated from the SIR model with random error added so that we can compare model estimates to a known true $\rr$ value.
-We vary the conditions under which the SIR data is generated to see how the results from the estimators change.
-We also look at data from different functional models to see how the estimators perform under different circumstances.
-
-%% What do you mean by describing the standard standard deviations?
-%% SG: delta_t is normally distributed with sd sigma and similarly for other methods.  How and why did we pick the numbers that we did?
-
-\textit{To see the effects of different conditions for the data, we change one of the starting conditions while holding all of the other ones constant.
-We try changing two conditions: the starting number of susceptible and infected people in the population (holding the total population constant) and the variance of the errors added on to the The starting population sizes for the $X$ and $Y$ are set to ($X_0$, $Y_0$) =  (99999, 1), (99990, 10), (99900, 100), and (99000, 1000), holding all other conditions constant.
-The variance of the added errors by scaling them to: ($\sigma_X$, $\sigma_Y$) = (10, 1), (100, 10), (500, 50), (1000, 100), (2000, 200), (5000, 500), (10000, 1000), (15000, 1500); these are done with ($X_0$, $Y_0$) = (99000, 1000).
-The errors start very small but become very large relative to the actual compartment sizes; this allows us to more clearly see the effects of the changes.}
-
-%% Describe the initial S, I, R values
-%% Describe the different error methods
-%% Describe the standard standard deviations involved
-%% High level fitting comments
-
-%% Throw fitting details into the appendix
-
-%% Talk about top 3-5 methods
-%%%%%% Honestly, all the methods seem to have flaws in one way or another - maybe point out why?
-%% Talk about methods that really miss
-%% Talk about the standard error reports
-%% Talk about 4 categories.  Does anyone do worse/better?
-%% What about the one where we cannot estimate std
-%% What happens when SD is large?
-
-%% Throw some graphs in
-
-\subsubsection{SIR Data}
-
-\textit{The datasets from the SIR model are simulated from the ODEs: \begin{align}
-	\systeme{\frac{dX}{dt} = -\frac{\beta XY}{N}, \frac{dY}{dt} = \frac{\beta XY}{N} - \gamma Y, \frac{dZ}{dt} = \gamma Z}.
-	\end{align}
-}
-
-We use arbitrarily selected the parameters to generate the time series data set from the SIR model.
-We first generate "baseline" data from which we can make comparisons by changing the  conditions under which the data were generated.
-The baseline SIR data are simulated with the following conditions:
-
-\begin{center}
-
-	$\beta$ = 0.006, $\gamma$ = 0.003; $\rr$ = 2
-
-	Number of time steps: 365
-
-	Starting X size ($X_0$): 99950
-
-	Starting Y size ($Y_0$): 50
-
-	Starting Z size: 0
-
-	Total population: 100000.
-
-\end{center}
-
-%% Is this enough detail on describing the initial S, I, and R, or are we looking for something more specific?
-
-\textbf{Our baseline initial size for $Y_0$ is 0.05\% of the total population in our simulated environment, which is fairly large compared to the starting point for the data in most epidemics.
-However, I/we felt that this was necessary in order to see the differences between the models.}
-
-We then generate random errors that are added to the X and Y compartments at each time step in order to simulate some form of randomness.
-\textbf{The sum of these was then subtracted from the Z compartment in order to keep the total population in the system constant.} We generated data sets with two different types of errors: Gaussian and autoregressive.
-\textbf{The Gaussian errors are independently distributed with mean 0 and constant variance for the entire length of the time series.
-The autoregressive errors are generated from a first-order process of the form: $$ \epsilon_{t} = \rho * \epsilon_{t-1} + \delta_{t}$$, where $\rho$ is fixed throughout and $\delta_{t} \sim N(0, \sigma)$ with constant $\sigma$.
-The values of the variances are chosen somewhat arbitrarily, although they are intended to be fairly small at first.}
-
-We also add a monotonicity adjustment to the X and Z compartments with both of the error models.
-\textbf{This is done using the Pool Adjacent Violators Algorithm (PAVA); the algorithm checks if the observation at each time step is greater than or equal to the previous one, and changes the value of both observations to be equal to the mean of the two if the condition is not fulfilled.} This gives us a total of four error models: autoregressive (Auto), autoregressive monotonic (AutoM), Gaussian (Norm), and Gaussian monotonic (NormM).
-For our baseline data, we have ($\sigma_X, \sigma_Y$) = (10, 1).
-We simulate 5 different baseline sets to see how consistent each of the models are.
-
-\subsubsection{Other Models}
-We simulated data from 6 other models besides the classic SIR model.
-5 of these were polynomial linear regression models, from order 0 to 4 with respect to time/day.
-The coefficients used to generate data from the linear regression models were originally chosen by taking the approximate coefficient estimates from fitting the linear regression models to classic SIR data.
-This does not necessarily guarantee that the true $\rr$ will be 1.2, especially for the 0th and 1st order linear regression models.
-What we are interested in for these cases is to see if the models still give reasonable estimates for $\rr$.
-The other model is the linear SIR model, which is a three compartment model with the following ODEs:
-
-	\begin{eqnarray*}
-		\frac{dS}{dt} &=& -\beta I \\
-		\frac{dI}{dt} &=& \beta I - \gamma I \\
-		\frac{dR}{dt} &=& \gamma I.
-	\end{eqnarray*}
-
- The data from the linear SIR model were simulated with parameters $\beta$ = 0.06 and $\gamma$ = 0.05, such that $\rr$ = 1.2.
-
- \textbf{Random error is added to the data generated here as well, using the same 4 error models as above.
-The other specifications for these datasets are the same as for the baseline classic SIR data, namely the variance of the added errors and starting compartment sizes.}
-
-\subsubsection{Variance Estimation}
-
-\textbf{The method for calculating the standard errors for $\hat{\rr}$ depend on the model used.
-The following models take the standard error directly from the regression methods used to estimate $\rr$:}
-
-	\begin{itemize}
-		\item SIR LMA
-		\item SIR LMAT
-		\item SIR Log Linear
-		\item SIR MC
-		\item Exponential Growth
-		\item Initial Growth
-	\end{itemize}
-
-\textbf{The ones that are calculated using the delta method:}
-
-	\begin{itemize}
-		\item-SIR LS
-		\item SIR ReLS
-		\item SEIR
-		\item MSEIR
-		%% \item AMP
-		\item NGM
-		\item Final Size
-	\end{itemize}
-
-\textbf{Standard errors for the maximum likelihood and branching process models are calculated using the negative reciprocal of the Fisher information.
-The sequential Bayes model uses the standard deviation from the posterior distribution.
-We use the jackknife method to estimate the standard error of the SIR smoothed max and the SIR smoothed IPR methods.
-The jackknife calculates the standard error using the standard deviation of estimates of $\rr$ after leaving out a single data point.}
-
-\textbf{Some of the models use only one single point in time to estimate $\rr$ and therefore do not seem to have an innate source of variance.
-We currently do not have standard errors for the SIR Max and SIR IPR models.}
-
-\subsection{Results}
-
-\subsubsection{Survival Function}
-
-%% How do we want to estimate this?????????
-%% SG:
-%% 3.1.1:  So I think this one we should mention there is no direct way to estimate this from the simple data we have.
-Alternatively, you could do a parameter sweep with different beta distributions for F(a) and beta for b(a).
-%% 3.1.2:  So assuming the population is homogeneous, k = N, D = gamma^{-1}.
-So we have to estimmate $b$ from the data.
-So I would average over (change in % susceptible ) / old # infected and use that as $b$.
-%% 3.1.3:  No clue.
-
-\subsubsection{Compartment Models}
-
-\paragraph{SIR Models}
-In general, the models derived from the SIR ordinary differential equations are relatively accurate.
-Most estimates are within the range of $\pm$ 0.2 of the true $\rr$ for the baseline SIR datasets.
-The magnitude of the standard errors on the baseline datasets for the different SIR models generally depend on the method used.
-The models that estimate $\rr$ using optimization or gradient descent, namely SIR LS and SIR ReLS, have standard errors on the order of $10^{-4}$ to $10^{-7}$.
-The standard errors are somewhat larger for the methods that use regression, on the order of $10^{-1}$ to $10^{-3}$.
-
-The SIR models as a whole also do fairly well with changing the starting number of infected people.
-However, they tend not to do as well when the variance of the added errors is very large; in these situations, many of the models tend to have very large standard errors, sometimes on the order of $10^{1}$ or $10^{2}$.
-
-The exceptions to the above are the linear, incidence-prevalence ratio, and Markov chain models.
-The estimates tend to be far from the true $\rr$, and they fluctuate greatly between the baseline data sets.
-The standard errors for these models (if they exist) also tend to be very large under some of the changing conditions.
-
-Many of the SIR models do not work as well with the data that come from the polynomial linear models.
-Specifically, for the 0th and 1st order model datasets, the estimates can be unreasonably large or negative.
-The Log linear, Sequential Bayes, Smoothed max, and Max models do not have this problem.
-
-\paragraph{Other Compartment Models}
-
-All of these models use optimization to estimate the parameters that are used to calculate $\rr$.
-They tend to be fairly accurate and consistent on the baseline data, often as good as or better than those from the various SIR models.
-The SEIR models (least squares and next generation) do not handle the large added errors very well, as it tends to give ridiculously large results.
-The standard errors tend to be on the order of $10^{-2}$ to $10^{-3}$.
-However, as with the estimates above, the standard errors from the SEIR models can become unreasonably large when the variance of the added errors becomes large.
-Also, as with the SIR models, the $\rr$ estimates and standard errors from the 0th and 1st order polynomial linear model datasets are not very good, as the former is sometimes negative and the latter can be very large.
-The MSEIR model tends to do fairly well in all situations.
-
-\subsubsection{Exponential Growth}
-
-We have assumed for these models that the number of infections follows an exponential curve throughout the entire time frame of the data sets.
-This assumption certainly does not hold for the data from the 0th and 1st order linear models, and may not hold for some of the other data sets either.
-
-These models are generally not as accurate as the compartment models above.
-Most of them, except for the initial size model, tend to underestimate $\rr$, with estimates between 0.8 and 1.0.
-The standard errors are all fairly small, on the order of magnitude from $10^{-1}$ to $10^{-3}$.
-This is larger than many of the compartment model standard errors.
-The main advantage that all of these models, except for the maximum likelihood model, seem to be more robust to changing conditions and functional forms of the data compared to the compartment models.
-The estimates and standard errors for $\rr$ appear to be reasonable in all cases, including for very large added errors and for the 0th and 1st order linear model data.
-
-\subsubsection{Network Models}
-
-The branching process model (the only one in this category so far) is not very accurate.
-In particular, for the SIR datasets, the $\rr$ estimates are less than 1, while the true $\rr$ is greater than 1.
-The standard errors, though, are relatively small, around the same order of magnitude as the SIR least squares models.
-Model estimates and standard errors are fairly consistent across the baseline datasets and changing starting compartment sizes.
-However, increasing the standard deviation of the added errors tends to have a somewhat noticeable impact on model results.
-The model also does not handle the 0th order polynomial linear data well.
-
 \section{Discussion}
@@
 
-Below, we compare the estimates and standard errors of $\rr$ given by each of the models for a few selected data sets.
-The forest plots below show the point estimates and 95\% confidence intervals of $\rr$ from each model for a particular data set.
-The true $\rr$ value is also shown by the vertical red line in the graphs.
-
-All of the figures below show results from data sets that were simulated with added Gaussian errors.
-The plot on the top left shows the estimates from one of the baseline data sets.
-For the most part, the $\rr$ values range from approximately 0.8 to 1.4.
-We note that the range of values includes $\rr$ values of less than 1, meaning that some models predict that a disease is not spreading when it actually is.
-However, on the whole, the variation of estimates between models is smaller than we originally expected.
-The confidence intervals are very small relative to the estimates for the most part; exceptions to this include the SIR IPR and the Markov chain model as the exception.
-
-The plot on the top right shows the estimates for the data set where $Y_0$ is increased to 100.
-The $\rr$ estimates generally seem to be further away from the true $\rr$ value compared to the results from the baseline data sets, and the overall .
-The confidence intervals are also more often than not larger than those in the baseline.
-In particular, the standard errors for some of the SIR models increase by several orders of magnitude.
-
-The bottom left plot shows the estimates for the data set where the error added to the generated data is increased.
-For most of the models, the estimates of $\rr$ do not change very much compared to the baseline data set.
-The confidence intervals for this data set increase for some of the models, such as the SIR LMA, while they remain approximately the same as the baseline data sets for most of the others.
-
-On the bottom right, we see the $\rr$ estimates for the case when disease prevalence grows linearly, as opposed to exponentially as seen in the data sets generated under SIR conditions, where the disease prevalence grows exponentially.
-Model performance here varies depending on how well the models the aforementioned linear growth.
-Some of the models, such as the SIR Markov chain and the IPR models, do not handle this case very well, giving much larger estimates and standard errors than are reasonable.
-On the other hand, models such as the SIR maximum and final size seem to be more flexible to different trajectories.
-[The error bars do not show up if they go outside of the x-axis limits - how do we fix this?]
-
-\begin{minipage}{0.5\linewidth}
-	\includegraphics[width=3in]{BaselineFP.jpeg}
-\end{minipage}
-\begin{minipage}{0.5\linewidth}
-	\includegraphics[width=3in]{StartPopFP.jpeg}
-\end{minipage}
-
-\begin{minipage}{0.5\linewidth}
-	\includegraphics[width=3in]{VarianceFP.jpeg}
-\end{minipage}
-\begin{minipage}{0.5\linewidth}
-	\includegraphics[width=3in]{OtherFP.jpeg}
-\end{minipage}
-
 \bibliographystyle{apa}%Choose a bibliograhpic style
 \bibliography{Master}
-
-\section{Appendix}
-
-Included here are the technical details for fitting the models in R, as well as tables of the results.
-
-\subsection{Compartment Models}
-
-\subsubsection{SIR Least Squares}
-
-\paragraph{Estimation}
-
-For the least-squares SIR model, we estimate the parameters $\beta$ and $\gamma$ of the ODEs using optim() in R to minimize the sum of squared errors using Nelder-Mead.
-The standard errors for $\beta$ and $\gamma$ were estimated using the Hessian matrix given to us by optim(), and then the standard error for $\rr$ was calculated using the delta method.
-
-\paragraph{Comments}
-
-The $\rr$ estimates for the baseline data are both close to the actual value, and the standard errors were relatively small.
-We would probably expect this given that this is the same model from which the data were generated.
-
-For the changing compartment sizes, we see that the standard errors are much larger for smaller values of $Y_0$.
-This might be because the variance of the I compartment is much larger relative to the size of the compartment, since we hold $\sigma_Y$ at 1 no matter if $Y_0$ is 1 or 1000.
-
-For the changing error variances, we see that the estimates for $\rr$ become less accurate for increasing $\sigma_X$ and $\sigma_Y$.
-However, it is interesting to note that the standard errors are increasing for the monotonic Gaussian error data, but decreasing for the autoregressive and monotonic autoregressive error data and relatively stable for the Gaussian non-monotonic data.
-It is not clear why this would be the case.
-
-\begin{table}[H]
-	\caption{$\rr$ Estimates and Std. Errs, SIR Model, 
-		Baseline Data, $X_0 = 99950, Y_0 = 50$, 
-		$\sigma_X = 10, \sigma_Y = 1$}
-	\begin{footnotesize}
-		\hskip -1cm
-		\begin{tabular}{l|r|r|r|r|r|r|r|r}
-			\hline
-			Data Set & Auto Est & Auto SE & AutoM Est & AutoM SE & Norm Est & Norm SE & NormM Est & NormM SE\\
-			\hline
-			Baseline 1 & 1.965322 & 0.000869 & 1.983308 & 0.000882 & 2.005867 & 0.000893 & 1.994736 & 0.000884\\
-			\hline
-			Baseline 2 & 2.039608 & 0.000913 & 1.999245 & 0.000895 & 2.001217 & 0.000889 & 1.995960 & 0.000887\\
-			\hline
-			Baseline 3 & 2.005609 & 0.000901 & 1.966631 & 0.000873 & 2.016216 & 0.000901 & 1.997175 & 0.000887\\
-			\hline
-			Baseline 4 & 1.958037 & 0.000859 & 1.941109 & 0.000852 & 1.995544 & 0.000886 & 2.004107 & 0.000892\\
-			\hline
-			Baseline 5 & 2.050904 & 0.000926 & 1.964328 & 0.000866 & 2.009912 & 0.000896 & 1.990153 & 0.000881\\
-			\hline
-		\end{tabular}
-	\end{footnotesize}
-\end{table}
-
-\begin{table}[H]
-
-	\caption{$\rr$ Estimates and Std. Errs, SIR Model,
-		Different Starting Populations, 
-		$\sigma_X = 10, \sigma_Y = 1$}
-	\begin{footnotesize}
-		\hskip -1.7cm
-		\begin{tabular}{l|r|r|r|r|r|r|r|r}
-			\hline
-			Data Set & Auto Est & Auto SE & AutoM Est & AutoM SE & Norm Est & Norm SE & NormM Est & NormM SE\\
-			\hline
-			$X_0 = 99999, Y_0 = 1$ & 1.689292 & 0.020794 & -4.167216 & 0.168853 & 1.976254 & 0.045081 & 1.688593 & 0.029680\\
-			\hline
-			$X_0 = 99990, Y_0 = 10$ & 2.038297 & 0.004597 & 2.207003 & 0.005165 & 2.037726 & 0.004596 & 1.883084 & 0.004049\\
-			\hline
-			$X_0 = 99900, Y_0 = 100$ & 2.009581 & 0.000449 & 2.006557 & 0.000448 & 1.999236 & 0.000445 & 1.995467 & 0.000444\\
-			\hline
-			$X_0 = 99000, Y_0 = 1000$ & 2.002521 & 0.000046 & 2.001829 & 0.000046 & 1.999519 & 0.000046 & 2.000098 & 0.000046\\
-			\hline
-		\end{tabular}
-	\end{footnotesize}
-\end{table}
-
-\begin{table}[H]
-
-	\caption{$\rr$ Estimates and Std. Errs, SIR Model, 
-		Different Variances, $X_0 = 99000$, $Y_0 = 1000$}
-	\begin{footnotesize}
-		\hskip -1.7cm
-		\begin{tabular}{l|r|r|r|r|r|r|r|r}
-			\hline
-			Data Set & Auto Est & Auto SE & AutoM Est & AutoM SE & Norm Est & Norm SE & NormM Est & NormM SE\\
-			\hline
-			$\sigma_X = 10, \sigma_Y = 1$ & 1.927444 & 0.000059 & 1.927090 & 5.9e-05 & 1.925768 & 5.9e-05 & 1.926576 & 5.9e-05\\
-			\hline
-			$\sigma_X = 100, \sigma_Y = 10$ & 1.926786 & 0.000058 & 1.916571 & 5.9e-05 & 1.919778 & 5.9e-05 & 1.929422 & 5.9e-05\\
-			\hline
-			$\sigma_X = 300, \sigma_Y = 30$ & 1.994211 & 0.000062 & 1.976893 & 5.8e-05 & 1.930660 & 5.8e-05 & 1.920869 & 5.9e-05\\
-			\hline
-			$\sigma_X = 1000, \sigma_Y = 100$ & 1.749057 & 0.000050 & 1.736588 & 5.9e-05 & 2.003654 & 6.2e-05 & 1.897049 & 5.9e-05\\
-			\hline
-			$\sigma_X = 3000, \sigma_Y = 300$ & 1.992674 & 0.000063 & 1.681585 & 4.9e-05 & 1.959815 & 6.7e-05 & 1.770816 & 5.3e-05\\
-			\hline
-			$\sigma_X = 10000, \sigma_Y = 1000$ & -42.264180 & 0.002694 & 1.773388 & 3.1e-05 & 1.889188 & 6.6e-05 & 2.649010 & 6.3e-05\\
-			\hline
-		\end{tabular}
-	\end{footnotesize}
-\end{table}
-
-\begin{table}[H]
-
-	\caption{$\rr$ Estimates and Std. Errs, SIR Model,
-		Other Models, $X_0 = 99950, Y_0 = 50$, 
-		$\sigma_X = 10, \sigma_Y = 1$}
-	\begin{footnotesize}
-		\hskip -1cm
-		\begin{tabular}{l|r|r|r|r|r|r|r|r}
-			\hline
-			Data Set & Auto Est & Auto SE & AutoM Est & AutoM SE & Norm Est & Norm SE & NormM Est & NormM SE\\
-			\hline
-			0th & -2.032207 & 0.244058 & 1.229963 & 0.069020 & 1.021369 & 0.056558 & 0.795460 & 0.331468\\
-			\hline
-			1st & 1.038117 & 0.000029 & 1.037889 & 0.000029 & 1.037902 & 0.000029 & 1.037887 & 0.000029\\
-			\hline
-			2nd & 1.198856 & 0.000045 & 1.198589 & 0.000045 & 1.198772 & 0.000045 & 1.198501 & 0.000045\\
-			\hline
-			3rd & 1.202405 & 0.000043 & 1.202098 & 0.000043 & 1.202350 & 0.000043 & 1.202240 & 0.000043\\
-			\hline
-			4th & 1.210597 & 0.000047 & 1.210597 & 0.000047 & 1.210194 & 0.000047 & 1.210304 & 0.000047\\
-			\hline
-			Linear SIR & 1.274346 & 0.000034 & 1.273739 & 0.000034 & 1.274229 & 0.000034 & 1.273950 & 0.000034\\
-			\hline
-		\end{tabular}
-	\end{footnotesize}
-\end{table}
-
-\subsubsection{SIR Reparameterized Least Squares}
-
-\paragraph{Estimation}
-
-For the reparameterized SIR model, we estimate the parameters $\rr$ (as well as $\gamma$) of the ODEs using optim() in R to minimize the sum of squared errors using Nelder-Mead.
-The standard errors were taken directly from the term in the Hessian matrix corresponding to $\rr$.
-(I also have the sensitivity analysis code here.)
-
-\paragraph{Comments}
-
-The $\rr$ estimates for the baseline data are both close to the actual value, and the estimated standard errors are extremely small.
-We would probably expect this given that this is the same model from which the data were generated.
-In general, the estimates for the standard errors are also much smaller than normal least squares.
-(Some of the errors are 0 because I rounded to 6 digits - all of them are around the magnitude of $10 ^ {-7}$)
-
-The $\rr$ estimates for the data with non-monotonic errors become more accurate as ($\sigma_X$, $\sigma_Y$) increase, while they become less accurate for the monotonic values.
-The standard errors are also positively correlated with the estimate of $\rr$ as long as all other parts, i.e., $X_0$ and $Y_0$ are held constant.
-
-\begin{table}[H]
-
-	\caption{$\rr$ Estimates and Std. Errs, SIRREPAR Model,
-		Baseline Data, $X_0 = 99950, Y_0 = 50$, 
-		$\sigma_X = 10, \sigma_Y = 1$}
-	\begin{footnotesize}
-		\hskip -1cm
-		\begin{tabular}{l|r|r|r|r|r|r|r|r}
-			\hline
-			Data Set & Auto Est & Auto SE & AutoM Est & AutoM SE & Norm Est & Norm SE & NormM Est & NormM SE\\
-			\hline
-			Baseline 1 & 3.226272 & 0.000000 & 1.984447 & 0.000294 & 2.006310 & 0.000298 & 1.993558 & 0.000294\\
-			\hline
-			Baseline 2 & 3.203012 & 0.000000 & 1.999589 & 0.000299 & 2.000868 & 0.000297 & 1.995630 & 0.000296\\
-			\hline
-			Baseline 3 & 4.180804 & 0.000733 & 4.180810 & 0.000724 & 5.043355 & 0.000908 & 1.997399 & 0.000296\\
-			\hline
-			Baseline 4 & 4.180819 & 0.000709 & 4.180816 & 0.000715 & 1.995237 & 0.000295 & 2.003821 & 0.000298\\
-			\hline
-			Baseline 5 & 4.180799 & 0.000741 & 4.180815 & 0.000717 & 2.008713 & 0.000300 & 4.180812 & 0.000720\\
-			\hline
-		\end{tabular}
-	\end{footnotesize}
-\end{table}
-
-\begin{table}[H]
-
-	\caption{$\rr$ Estimates and Std. Errs, SIRREPAR Model,
-		Different Starting Populations, 
-		$\sigma_X = 10, \sigma_Y = 1$}
-	\begin{footnotesize}
-		\hskip -1.7cm
-		\begin{tabular}{l|r|r|r|r|r|r|r|r}
-			\hline
-			Data Set & Auto Est & Auto SE & AutoM Est & AutoM SE & Norm Est & Norm SE & NormM Est & NormM SE\\
-			\hline
-			$X_0 = 99999, Y_0 = 1$ & 5.899238 & 0.025009 & 5.898924 & 0.051433 & 3.394166 & 0.000021 & 3.217232 & 0.000016\\
-			\hline
-			$X_0 = 99990, Y_0 = 10$ & 3.162364 & 0.000001 & 3.170152 & 0.000002 & 3.131576 & 0.000001 & 3.108194 & 0.000001\\
-			\hline
-			$X_0 = 99900, Y_0 = 100$ & 4.564960 & 0.000404 & 2.005939 & 0.000150 & 1.998880 & 0.000149 & 1.994592 & 0.000148\\
-			\hline
-			$X_0 = 99000, Y_0 = 1000$ & 47.241714 & 0.002493 & 47.241714 & 0.002511 & 47.241714 & 0.002502 & 47.241714 & 0.002502\\
-			\hline
-		\end{tabular}
-	\end{footnotesize}
-\end{table}
-
-\begin{table} [H]
-
-	\caption{$\rr$ Estimates and Std. Errs, SIRREPAR Model,
-		Different Variances, 
-		$X_0 = 99000, Y_0 = 1000$}
-	\begin{footnotesize}
-		\hskip -1.7cm
-		\begin{tabular}{l|r|r|r|r|r|r|r|r}
-			\hline
-			Data Set & Auto Est & Auto SE & AutoM Est & AutoM SE & Norm Est & Norm SE & NormM Est & NormM SE\\
-			\hline
-			$\sigma_X = 10, \sigma_Y = 1$ & 1.926733 & 4e-06 & 1.927468 & 4e-06 & 1.925555 & 4e-06 & 1.926689 & 4e-06\\
-			\hline
-			$\sigma_X = 100, \sigma_Y = 10$ & 1.925550 & 4e-06 & 1.916380 & 4e-06 & 1.920075 & 4e-06 & 1.928987 & 4e-06\\
-			\hline
-			$\sigma_X = 300, \sigma_Y = 30$ & 1.994719 & 4e-06 & 1.976273 & 4e-06 & 1.930978 & 4e-06 & 1.920290 & 4e-06\\
-			\hline
-			$\sigma_X = 1000, \sigma_Y = 100$ & 1.749031 & 3e-06 & 1.736589 & 3e-06 & 2.003044 & 4e-06 & 1.897328 & 4e-06\\
-			\hline
-			$\sigma_X = 3000, \sigma_Y = 300$ & 1.990948 & 4e-06 & 1.681748 & 3e-06 & 1.959229 & 4e-06 & 1.770892 & 3e-06\\
-			\hline
-			$\sigma_X = 10000, \sigma_Y = 1000$ & 5.329640 & 3e-05 & 1.773820 & 2e-06 & 1.887014 & 4e-06 & 2.648300 & 6e-06\\
-			\hline
-		\end{tabular}
-	\end{footnotesize}
-\end{table}
-\begin{table}[H]
-
-	\caption{$\rr$ Estimates and Std. Errs, SIRREPAR Model,
-		Other Models, $X_0 = 99950, Y_0 = 50$, 
-		$\sigma_X = 10, \sigma_Y = 1$}
-	\begin{footnotesize}
-		\hskip -1cm
-		\begin{tabular}{l|r|r|r|r|r|r|r|r}
-			\hline
-			Data Set & Auto Est & Auto SE & AutoM Est & AutoM SE & Norm Est & Norm SE & NormM Est & NormM SE\\
-			\hline
-			0th & 3.272461 & 0e+00 & 3.275535 & 0e+00 & 3.135010 & 0e+00 & 3.235197 & 0e+00\\
-			\hline
-			1st & 1.038125 & 0e+00 & 1.037880 & 0e+00 & 1.037903 & 0e+00 & 1.037881 & 0e+00\\
-			\hline
-			2nd & 1.198760 & 1e-06 & 1.198526 & 1e-06 & 1.198864 & 1e-06 & 1.198616 & 1e-06\\
-			\hline
-			3rd & 1.202489 & 1e-06 & 1.202071 & 1e-06 & 1.202246 & 1e-06 & 1.202355 & 1e-06\\
-			\hline
-			4th & 1.210605 & 1e-06 & 1.210605 & 1e-06 & 1.210220 & 1e-06 & 1.210199 & 1e-06\\
-			\hline
-			Linear SIR & 1.274175 & 1e-06 & 1.273799 & 1e-06 & 1.274314 & 1e-06 & 1.274075 & 1e-06\\
-			\hline
-		\end{tabular}
-	\end{footnotesize}
-\end{table}
-
-\subsubsection{SIR Linear Model}
-
-\paragraph{Estimation}
-
-To calculate the estimates for $\rr$, we use lm() in R to fit the linear polynomial regression functions for X(t) and Y(t) with respect to t.
-We then calculated the derivatives for the linear regression functions at time 0 using our own function.
-For the standard error, the jackknife was used on the data set.
-This calculates the standard error of the $\rr$ by using the standard deviation of the $R-0$ estimates after leaving out a single observation.
-
-\paragraph{Comments}
-
-The estimates for $\rr$ are very inaccurate compared to the other models.
-The inaccuracy is most glaring for the data with large error variances and the simulated from models besides the SIR one.
-The estimates seem to become less accurate for the data generated from the classic SIR model as $\sigma_X$ and $\sigma_Y$ increase and as $Y_0$ decreases relative to the total population.
-There are also many estimates of $\rr$ that are negative, which implies that there is a negative derivative estimate for either the susceptible or infected compartment (is this supposed to be possible?).
-
-The standard errors are also extremely large in some cases.
-This seems to hold consistently when $\sigma_X$ and $\sigma_Y$ are large for the classic SIR data as well as for the 4th order linear data.
-However, there are also other places where large standard error estimates appear for no apparent reason.
-The large standard errors and the above issue with the $\rr$ estimates being negative could be because we are only using a point estimate.
-In general, the only time we see fairly consistent and reasonable estimates from the linear model is for the SIR data with relatively large $Y_0$ as well as small $\sigma_X$ and $\sigma_Y$ (i.e., for the data sets where ($X_0$, $Y_0$) = (99000, 1000)).
-
-\begin{table}[H]
-
-	\caption{$\rr$ Estimates and Std. Errs, SIRLMA Model,
-		Baseline Data, $X_0 = 99950, Y_0 = 50$, 
-		$\sigma_X = 10, \sigma_Y = 1$}
-	\begin{footnotesize}
-		\hskip -1cm
-		\begin{tabular}{l|r|r|r|r|r|r|r|r}
-			\hline
-			Data Set & Auto Est & Auto SE & AutoM Est & AutoM SE & Norm Est & Norm SE & NormM Est & NormM SE\\
-			\hline
-			Baseline 1 & 0.864913 & 0.203402 & 1.739858 & 0.122632 & 1.359771 & 0.919629 & 1.173245 & 0.096355\\
-			\hline
-			Baseline 2 & 1.485664 & 1.654031 & 1.294412 & 0.618155 & 1.461580 & 0.734907 & 1.186621 & 0.091633\\
-			\hline
-			Baseline 3 & 1.082572 & 0.088142 & 1.389817 & 0.117032 & 0.615994 & 1.496674 & 1.088364 & 0.089282\\
-			\hline
-			Baseline 4 & 1.088362 & 0.367110 & 1.370251 & 0.159221 & 2.683109 & 27.981877 & 1.090398 & 0.059515\\
-			\hline
-			Baseline 5 & 0.871650 & 0.100744 & 1.438190 & 0.382650 & 0.829591 & 0.120911 & 0.924238 & 0.188305\\
-			\hline
-		\end{tabular}
-	\end{footnotesize}
-\end{table}
-
-\begin{table}[H]
-
-	\caption{$\rr$ Estimates and Std. Errs, SIRLMA Model,
-		Different Starting Populations, 
-		$\sigma_X = 10, \sigma_Y = 1$}
-	\begin{footnotesize}
-		\hskip -1.7cm
-		\begin{tabular}{l|r|r|r|r|r|r|r|r}
-			\hline
-			Data Set & Auto Est & Auto SE & AutoM Est & AutoM SE & Norm Est & Norm SE & NormM Est & NormM SE\\
-			\hline
-			$X_0 = 99999, Y_0 = 1$ & 1.713333 & 1.757278 & 1.187769 & 0.515629 & 1.451833 & 59.746089 & 1.068633 & 1.809596\\
-			\hline
-			$X_0 = 99990, Y_0 = 10$ & 1.197457 & 0.192831 & 0.674259 & 0.090168 & 1.095143 & 0.143146 & 1.199437 & 0.106393\\
-			\hline
-			$X_0 = 99900, Y_0 = 100$ & 0.796760 & 0.488540 & 1.022290 & 0.070029 & 0.024273 & 4.564245 & 1.265712 & 0.080303\\
-			\hline
-			$X_0 = 99000, Y_0 = 1000$ & 1.431229 & 0.105605 & 2.091993 & 0.340798 & 2.183329 & 0.428090 & 1.812900 & 0.122106\\
-			\hline
-		\end{tabular}
-	\end{footnotesize}
-\end{table}
-
-\begin{table}[H]
-
-	\caption{$\rr$ Estimates and Std. Errs, SIRLMA Model,
-		Different Variances, 
-		$X_0 = 99000, Y_0 = 1000$}
-	\begin{footnotesize}
-		\hskip -2cm
-		\begin{tabular}{l|r|r|r|r|r|r|r|r}
-			\hline
-			Data Set & Auto Est & Auto SE & AutoM Est & AutoM SE & Norm Est & Norm SE & NormM Est & NormM SE\\
-			\hline
-			$\sigma_X = 10, \sigma_Y = 1$ & -121.825007 & 904.236167 & 121.668347 & 308.632857 & 21.981032 & 21.204959 & 18.741699 & 17.132467\\
-			\hline
-			$\sigma_X = 100, \sigma_Y = 10$ & 2.918979 & 2.010263 & -1.223521 & 2.355038 & -5.683542 & 10.750736 & 4.071392 & 2.877332\\
-			\hline
-			$\sigma_X = 300, \sigma_Y = 30$ & 2.769860 & 1.759591 & 2.002654 & 0.949586 & 3.227956 & 3.934407 & 87.874782 & 456.259511\\
-			\hline
-			$\sigma_X = 1000, \sigma_Y = 100$ & 0.915434 & 0.114881 & 1.887797 & 0.672640 & 1.759919 & 1.537275 & -0.235868 & 1.519149\\
-			\hline
-			$\sigma_X = 3000, \sigma_Y = 300$ & -4.547791 & 970.222266 & 1.037930 & 0.112483 & 0.748209 & 0.437849 & 0.235119 & 1.489429\\
-			\hline
-			$\sigma_X = 10000, \sigma_Y = 1000$ & 1.033786 & 0.068722 & 0.958868 & 0.066758 & 0.862043 & 0.168471 & 0.671640 & 0.206497\\
-			\hline
-		\end{tabular}
-	\end{footnotesize}
-\end{table}
-
-\begin{table}[H]
-
-	\caption{$\rr$ Estimates and Std. Errs, SIRLMA Model,
-		Other Models, $X_0 = 99950, Y_0 = 50$, 
-		$\sigma_X = 10, \sigma_Y = 1$}
-	\begin{footnotesize}
-		\hskip -1cm
-		\begin{tabular}{l|r|r|r|r|r|r|r|r}
-			\hline
-			Data Set & Auto Est & Auto SE & AutoM Est & AutoM SE & Norm Est & Norm SE & NormM Est & NormM SE\\
-			\hline
-			0th & 1.049994 & 0.056806 & 1.286339 & 0.479624 & 0.980772 & 0.060085 & 0.474270 & 2.099030\\
-			\hline
-			1st & 1.000767 & 0.125396 & 1.009129 & 0.142918 & 1.018881 & 0.127098 & 1.019178 & 0.159078\\
-			\hline
-			2nd & 1.467835 & 0.558738 & 0.085260 & 1.223959 & 1.215482 & 0.162666 & -6.656798 & 265.908999\\
-			\hline
-			3rd & 4.877978 & 57.939148 & 1.111925 & 0.120143 & 1.238831 & 0.321316 & 1.227691 & 0.221355\\
-			\hline
-			4th & -0.892745 & 7.391321 & -0.892745 & 7.391321 & 0.445461 & 2.016867 & 1.709272 & 2.038759\\
-			\hline
-			Linear SIR & 1.214648 & 0.062718 & 0.950315 & 0.106644 & 1.219144 & 0.109327 & 1.206761 & 0.048099\\
-			\hline
-		\end{tabular}
-	\end{footnotesize}
-\end{table}
-
-\subsubsection{SIR Linear Model Average}
-
-\paragraph{Estimation}
-
-To calculate the estimates for $\rr$, we use lm() in R to fit the linear polynomial regression functions for X(t) and Y(t) with respect to t.
-We then calculated the derivatives of the linear regression functions at each time using our own function in order to find the average linear model $\rr$ estimate across all time steps $t$.
-The standard error was calculated by taking the standard deviation of the $\rr$ estimates across all time steps $t$ in the data set.
-
-\paragraph{Comments}
-
-The estimates for $\rr$ here are somewhat more accurate compared to the linear model estimates that only looked at the derivative first time point.
-In particular, it does a lot better for the baseline data and the data from other models.
-The estimates are still really bad for large values of $\sigma_X$ and $\sigma_Y$ (including some negative estimates).
-Interestingly, the linear model average estimates here are less accurate for large values of $Y_0$, whereas they were more accurate for the linear model.
-
-The standard errors are not as ridiculously large compared to the linear model with only one time point, but are still extremely big compared to what we might expect to see from other models.
-This is especially apparent when $\sigma_X$ and $\sigma_Y$ are large.
-
-\begin{table}[H]
-
-	\caption{$\rr$ Estimates and Std. Errs, SIRLMAT Model,
-		Baseline Data, $X_0 = 99950, Y_0 = 50$, 
-		$\sigma_X = 10, \sigma_Y = 1$}
-	\begin{footnotesize}
-		\hskip -1cm
-		\begin{tabular}{l|r|r|r|r|r|r|r|r}
-			\hline
-			Data Set & Auto Est & Auto SE & AutoM Est & AutoM SE & Norm Est & Norm SE & NormM Est & NormM SE\\
-			\hline
-			Baseline 1 & 2.015772 & 27.518508 & 2.150988 & 4.474655 & 1.988515 & 0.232390 & 2.103188 & 0.636825\\
-			\hline
-			Baseline 2 & 2.571202 & 21.048981 & -0.131021 & 51.293391 & 2.192838 & 1.684886 & 2.189259 & 1.514415\\
-			\hline
-			Baseline 3 & 1.201407 & 21.526024 & 3.023289 & 26.322188 & 2.079618 & 1.027536 & 2.007048 & 0.269511\\
-			\hline
-			Baseline 4 & 1.991370 & 1.164132 & 3.604925 & 35.619793 & 2.028025 & 0.205708 & 1.875085 & 2.080861\\
-			\hline
-			Baseline 5 & 1.873964 & 7.146565 & 1.032642 & 11.871804 & 1.372848 & 10.164198 & 1.926002 & 0.166963\\
-			\hline
-		\end{tabular}
-	\end{footnotesize}
-\end{table}
-
-\begin{table}[H]
-
-	\caption{$\rr$ Estimates and Std. Errs, SIRLMAT Model,
-		Different Starting Populations, 
-		$\sigma_X = 10, \sigma_Y = 1$}
-	\begin{footnotesize}
-		\hskip -1.7cm
-		\begin{tabular}{l|r|r|r|r|r|r|r|r}
-			Data Set & Auto Est & Auto SE & AutoM Est & AutoM SE & Norm Est & Norm SE & NormM Est & NormM SE\\
-			\hline
-			$X_0 = 99999, Y_0 = 1$ & 2.558602 & 32.007328 & 1.296284 & 9.652181 & 1.178839 & 3.901942 & 1.287102 & 0.524929\\
-			\hline
-			$X_0 = 99990, Y_0 = 10$ & 0.745127 & 3.487290 & 1.541772 & 1.793288 & 3.227689 & 26.102558 & 2.489264 & 26.755104\\
-			\hline
-			$X_0 = 99900, Y_0 = 100$ & 2.082023 & 2.181500 & -1.745769 & 81.529981 & 1.946617 & 1.232307 & 2.008181 & 0.191259\\
-			\hline
-			$X_0 = 99000, Y_0 = 1000$ & 1.969496 & 0.088997 & 1.979588 & 0.093847 & 1.972052 & 0.036176 & 1.964677 & 0.028405\\
-			\hline
-		\end{tabular}
-	\end{footnotesize}
-\end{table}
-
-\begin{table}[H]
-
-	\caption{$\rr$ Estimates and Std. Errs, SIRLMAT Model,
-		Different Variances, 
-		$X_0 = 99000, Y_0 = 1000$}
-	\begin{footnotesize}
-		\hskip -1.7cm
-		\begin{tabular}{l|r|r|r|r|r|r|r|r}
-			\hline
-			Data Set & Auto Est & Auto SE & AutoM Est & AutoM SE & Norm Est & Norm SE & NormM Est & NormM SE\\
-			\hline
-			$\sigma_X = 10, \sigma_Y = 1$ & 5.556294 & 63.868004 & 2.843995 & 7.555433 & 2.277135 & 2.022040 & 2.234907 & 1.759312\\
-			\hline
-			$\sigma_X = 100, \sigma_Y = 10$ & 2.020666 & 0.483996 & 1.802373 & 5.248770 & 1.891754 & 5.445302 & 2.034306 & 0.527914\\
-			\hline
-			$\sigma_X = 300, \sigma_Y = 30$ & -1.064869 & 50.543183 & 4.685491 & 53.111296 & 2.076773 & 0.646978 & 2.521401 & 5.276258\\
-			\hline
-			$\sigma_X = 1000, \sigma_Y = 100$ & 0.999874 & 3.992411 & 2.390182 & 3.320493 & 1.456545 & 6.796468 & 1.703567 & 2.395295\\
-			\hline
-			$\sigma_X = 3000, \sigma_Y = 300$ & -2.740783 & 70.416107 & 1.371659 & 1.347889 & 0.030795 & 21.459099 & 2.023080 & 6.341932\\
-			\hline
-			$\sigma_X = 10000, \sigma_Y = 1000$ & 0.877306 & 0.815504 & 1.106179 & 2.530188 & 1.183491 & 3.063114 & 0.949453 & 6.173692\\
-			\hline
-		\end{tabular}
-	\end{footnotesize}
-\end{table}
-\begin{table}[H]
-
-	\caption{$\rr$ Estimates and Std. Errs, SIRLMAT Model,
-		Other Models, $X_0 = 99950, Y_0 = 50$, 
-		$\sigma_X = 10, \sigma_Y = 1$}
-	\begin{footnotesize}
-		\hskip -1cm
-		\begin{tabular}{l|r|r|r|r|r|r|r|r}
-			\hline
-			Data Set & Auto Est & Auto SE & AutoM Est & AutoM SE & Norm Est & Norm SE & NormM Est & NormM SE\\
-			\hline
-			0th & 0.848941 & 3.448691 & 0.925502 & 18.031398 & -0.885541 & 17.053815 & 0.898319 & 0.676024\\
-			\hline
-			1st & 1.152006 & 0.018441 & 1.152364 & 0.017047 & 1.151996 & 0.016274 & 1.152495 & 0.015856\\
-			\hline
-			2nd & 1.164984 & 0.747602 & 1.181705 & 1.419549 & 1.233825 & 1.421868 & 3.337101 & 40.581184\\
-			\hline
-			3rd & 1.184468 & 0.202637 & 1.168692 & 0.036327 & 1.170062 & 0.035413 & 1.170215 & 0.034933\\
-			\hline
-			4th & 1.188048 & 0.446227 & 1.188048 & 0.446227 & 1.172531 & 0.353343 & 1.189808 & 0.052711\\
-			\hline
-			Linear SIR & 1.203159 & 0.014809 & 1.195875 & 0.024693 & 1.200980 & 0.003110 & 1.200130 & 0.004076\\
-			\hline
-		\end{tabular}
-	\end{footnotesize}
-\end{table}
-
-\subsubsection{SIR Maximum}
-
-\paragraph{Estimation}
-
-We use the which.max() function in R to find the point in time where the number people in the infected compartment is at a maximum.
-We need to come up with a way to calculate the variance (sensitivity analysis?).
-
-\paragraph{Comments}
-
-The SIR max model gives decent results.
-Most of the estimates are fairly close to the true $\rr$ value.
-It appears to give more accurate $\rr$ estimates better for higher values of $Y_0$ relative to the total population as well as smaller values of $\sigma_X$ and $\sigma_Y$.
-This model also consistently gives reasonable results, unlike for other models, which sometimes fail miserably when dealing with larger errors or data generated from other models (especially for 0th and 1st order data).
-
-\begin{table}[H]
-
-	\caption{$\rr$ Estimates and Std. Errs, SIRMAX Model,
-		Baseline Data, $X_0 = 99950, Y_0 = 50$, 
-		$\sigma_X = 10, \sigma_Y = 1$}
-	\begin{footnotesize}
-		\hskip -1cm
-		\begin{tabular}{l|r|r|r|r|r|r|r|r}
-			\hline
-			Data Set & Auto Est & Auto SE & AutoM Est & AutoM SE & Norm Est & Norm SE & NormM Est & NormM SE\\
-			\hline
-			Baseline 1 & 1.002232 & NA & 1.002359 & NA & 1.002498 & NA & 1.002531 & NA\\
-			\hline
-			Baseline 2 & 1.002405 & NA & 1.002802 & NA & 1.002364 & NA & 1.002561 & NA\\
-			\hline
-			Baseline 3 & 1.002261 & NA & 1.002393 & NA & 1.002516 & NA & 1.002501 & NA\\
-			\hline
-			Baseline 4 & 1.002469 & NA & 1.002494 & NA & 1.002511 & NA & 1.002481 & NA\\
-			\hline
-			Baseline 5 & 1.002392 & NA & 1.002496 & NA & 1.002509 & NA & 1.002529 & NA\\
-			\hline
-		\end{tabular}
-	\end{footnotesize}
-\end{table}
-
-\begin{table}[H]
-
-	\caption{$\rr$ Estimates and Std. Errs, SIRMAX Model,
-		Different Starting Populations, 
-		$\sigma_X = 10, \sigma_Y = 1$}
-	\begin{footnotesize}
-		\hskip -1.7cm
-		\begin{tabular}{l|r|r|r|r|r|r|r|r}
-			\hline
-			Data Set & Auto Est & Auto SE & AutoM Est & AutoM SE & Norm Est & Norm SE & NormM Est & NormM SE\\
-			\hline
-			$X_0 = 99999, Y_0 = 1$ & 0.999831 & NA & 1.000028 & NA & 0.999962 & NA & 1.000079 & NA\\
-			\hline
-			$X_0 = 99990, Y_0 = 10$ & 1.000442 & NA & 1.000473 & NA & 1.000514 & NA & 1.000527 & NA\\
-			\hline
-			$X_0 = 99900, Y_0 = 100$ & 1.005071 & NA & 1.004938 & NA & 1.005222 & NA & 1.004948 & NA\\
-			\hline
-			$X_0 = 99000, Y_0 = 1000$ & 1.049518 & NA & 1.050139 & NA & 1.050143 & NA & 1.050157 & NA\\
-			\hline
-		\end{tabular}
-	\end{footnotesize}
-\end{table}
-
-\begin{table}[H]
-
-	\caption{$\rr$ Estimates and Std. Errs, SIRMAX Model,
-		Different Variances, 
-		$X_0 = 99000, Y_0 = 1000$}
-	\begin{footnotesize}
-		\hskip -1.7cm
-		\begin{tabular}{l|r|r|r|r|r|r|r|r}
-			\hline
-			Data Set & Auto Est & Auto SE & AutoM Est & AutoM SE & Norm Est & Norm SE & NormM Est & NormM SE\\
-			\hline
-			$\sigma_X = 10, \sigma_Y = 1$ & 1.050097 & NA & 1.049897 & NA & 1.050034 & NA & 1.050098 & NA\\
-			\hline
-			$\sigma_X = 100, \sigma_Y = 10$ & 1.048816 & NA & 1.051191 & NA & 1.051560 & NA & 1.051580 & NA\\
-			\hline
-			$\sigma_X = 300, \sigma_Y = 30$ & 1.045130 & NA & 1.054815 & NA & 1.042187 & NA & 1.049682 & NA\\
-			\hline
-			$\sigma_X = 1000, \sigma_Y = 100$ & 1.038396 & NA & 1.051101 & NA & 1.057586 & NA & 1.049022 & NA\\
-			\hline
-			$\sigma_X = 3000, \sigma_Y = 300$ & 1.005515 & NA & 1.081383 & NA & 1.002521 & NA & 1.064796 & NA\\
-			\hline
-			$\sigma_X = 10000, \sigma_Y = 1000$ & 1.076870 & NA & 1.114340 & NA & 1.077148 & NA & 1.037306 & NA\\
-			\hline
-		\end{tabular}
-	\end{footnotesize}
-\end{table}
-
-\begin{table}[H]
-
-	\caption{$\rr$ Estimates and Std. Errs, SIRMAX Model,
-		Other Models, $X_0 = 99950, Y_0 = 50$, 
-		$\sigma_X = 10, \sigma_Y = 1$}
-	\begin{footnotesize}
-		\hskip -1cm
-		\begin{tabular}{l|r|r|r|r|r|r|r|r}
-			\hline
-			Data Set & Auto Est & Auto SE & AutoM Est & AutoM SE & Norm Est & Norm SE & NormM Est & NormM SE\\
-			\hline
-			0th & 1.000604 & NA & 1.000563 & NA & 1.000366 & NA & 1.000543 & NA\\
-			\hline
-			1st & 1.083647 & NA & 1.083855 & NA & 1.083627 & NA & 1.083315 & NA\\
-			\hline
-			2nd & 1.086218 & NA & 1.085875 & NA & 1.085994 & NA & 1.085891 & NA\\
-			\hline
-			3rd & 1.096363 & NA & 1.095844 & NA & 1.096062 & NA & 1.095990 & NA\\
-			\hline
-			4th & 1.085509 & NA & 1.085509 & NA & 1.085289 & NA & 1.085296 & NA\\
-			\hline
-			Linear SIR & 1.127272 & NA & 1.127403 & NA & 1.127313 & NA & 1.127194 & NA\\
-			\hline
-		\end{tabular}
-	\end{footnotesize}
-\end{table}
-
-\subsubsection{SIR Smoothed Max}
-
-\paragraph{Estimation}
-
-We use the smooth.spline() function in R to get estimates for X(t) and Y(t), then used the which.max() function to find the point in time where the number of people in the infected compartment is at a maximum.
-The standard error for the $\rr$ estimates was done using the jackknife method; we estimate a new spline function after removing one observation from the data set and calculate a new $\rr$ estimate, then use the standard deviation of those estimates as the estimate of the standard error.
-
-\paragraph{Comments}
-
-The results are fairly similar to the non-smoothed SIR maximum model.
-While the model does not give perfectly accurate estimates for $\rr$, the results are decent.
-It does not give us any crazy results.
-The model is more accurate for larger values of $Y_0$ relative to the total size of the population and for smaller $\sigma_X$ and $\sigma_Y$.
-
-The standard errors are reasonable values, and they seem to be fairly consistent under repeated estimation with the same conditions for the generated data.
-They increase with increases in $Y_0$ as well as $\sigma_X$ and $\sigma_Y$.
-
-\begin{table}[H]
-
-	\caption{$\rr$ Estimates and Std. Errs, SIRSMAX Model,
-		Baseline Data, $X_0 = 99950, Y_0 = 50$, 
-		$\sigma_X = 10, \sigma_Y = 1$}
-	\begin{footnotesize}
-		\hskip -1cm
-		\begin{tabular}{l|r|r|r|r|r|r|r|r}
-			\hline
-			Data Set & Auto Est & Auto SE & AutoM Est & AutoM SE & Norm Est & Norm SE & NormM Est & NormM SE\\
-			\hline
-			Baseline 1 & 1.002424 & 2.3e-05 & 1.002460 & 1.1e-05 & 1.002435 & 1.5e-05 & 1.002459 & 8e-06\\
-			\hline
-			Baseline 2 & 1.002507 & 2.2e-05 & 1.002444 & 1.8e-05 & 1.002462 & 1.9e-05 & 1.002436 & 7e-06\\
-			\hline
-			Baseline 3 & 1.002374 & 1.9e-05 & 1.002418 & 5.0e-06 & 1.002439 & 1.7e-05 & 1.002445 & 5e-06\\
-			\hline
-			Baseline 4 & 1.002434 & 1.5e-05 & 1.002479 & 8.0e-06 & 1.002461 & 1.5e-05 & 1.002449 & 6e-06\\
-			\hline
-			Baseline 5 & 1.002403 & 2.0e-05 & 1.002469 & 9.0e-06 & 1.002439 & 1.6e-05 & 1.002468 & 6e-06\\
-			\hline
-		\end{tabular}
-	\end{footnotesize}
-\end{table}
-
-\begin{table}[H]
-
-	\caption{$\rr$ Estimates and Std. Errs, SIRSMAX Model,
-		Different Starting Populations, 
-		$\sigma_X = 10, \sigma_Y = 1$}
-	\begin{footnotesize}
-		\hskip -1.7cm
-		\begin{tabular}{l|r|r|r|r|r|r|r|r}
-			\hline
-			Data Set & Auto Est & Auto SE & AutoM Est & AutoM SE & Norm Est & Norm SE & NormM Est & NormM SE\\
-			\hline
-			$X_0 = 99999, Y_0 = 1$ & 1.000116 & 1.7e-05 & 1.000063 & 2.0e-06 & 1.000040 & 1.5e-05 & 1.000064 & 1.0e-06\\
-			\hline
-			$X_0 = 99990, Y_0 = 10$ & 1.000474 & 2.1e-05 & 1.000477 & 8.0e-06 & 1.000473 & 1.6e-05 & 1.000507 & 4.0e-06\\
-			\hline
-			$X_0 = 99900, Y_0 = 100$ & 1.004902 & 1.8e-05 & 1.004880 & 1.4e-05 & 1.004899 & 1.9e-05 & 1.004895 & 8.0e-06\\
-			\hline
-			$X_0 = 99000, Y_0 = 1000$ & 1.049155 & 7.4e-05 & 1.049349 & 6.6e-05 & 1.049271 & 6.9e-05 & 1.049277 & 6.6e-05\\
-			\hline
-		\end{tabular}
-	\end{footnotesize}
-\end{table}
-
-\begin{table}[H]
-
-	\caption{$\rr$ Estimates and Std. Errs, SIRSMAX Model,
-		Different Variances, 
-		$X_0 = 99000, Y_0 = 1000$}
-	\begin{footnotesize}
-		\hskip -1.7cm
-		\begin{tabular}{l|r|r|r|r|r|r|r|r}
-			\hline
-			Data Set & Auto Est & Auto SE & AutoM Est & AutoM SE & Norm Est & Norm SE & NormM Est & NormM SE\\
-			\hline
-			$\sigma_X = 10, \sigma_Y = 1$ & 1.049351 & 0.000076 & 1.049219 & 0.000065 & 1.049294 & 0.000069 & 1.049283 & 0.000069\\
-			\hline
-			$\sigma_X = 100, \sigma_Y = 10$ & 1.049675 & 0.000154 & 1.049504 & 0.000132 & 1.049443 & 0.000182 & 1.049389 & 0.000131\\
-			\hline
-			$\sigma_X = 300, \sigma_Y = 30$ & 1.046190 & 0.000566 & 1.051219 & 0.000676 & 1.050137 & 0.000623 & 1.049671 & 0.000141\\
-			\hline
-			$\sigma_X = 1000, \sigma_Y = 100$ & 1.061907 & 0.002419 & 1.048793 & 0.000363 & 1.046010 & 0.001880 & 1.048815 & 0.000172\\
-			\hline
-			$\sigma_X = 3000, \sigma_Y = 300$ & 1.045117 & 0.005651 & 1.068145 & 0.004304 & 1.039168 & 0.005444 & 1.057089 & 0.000372\\
-			\hline
-			$\sigma_X = 10000, \sigma_Y = 1000$ & 0.979174 & 0.014889 & 1.111583 & 0.002687 & 1.039000 & 0.015012 & 1.070210 & 0.004720\\
-			\hline
-		\end{tabular}
-	\end{footnotesize}
-\end{table}
-
-\begin{table}[H]
-
-	\caption{$\rr$ Estimates and Std. Errs, SIRSMAX Model,
-		Other Models, $X_0 = 99950, Y_0 = 50$, 
-		$\sigma_X = 10, \sigma_Y = 1$}
-	\begin{footnotesize}
-		\hskip -1cm
-		\begin{tabular}{l|r|r|r|r|r|r|r|r}
-			\hline
-			Data Set & Auto Est & Auto SE & AutoM Est & AutoM SE & Norm Est & Norm SE & NormM Est & NormM SE\\
-			\hline
-			0th & 1.000515 & 0.000022 & 1.000486 & 0.000000 & 1.000466 & 0.000015 & 1.000462 & 0.000003\\
-			\hline
-			1st & 1.083710 & 0.000021 & 1.083642 & 0.000032 & 1.083606 & 0.000022 & 1.083580 & 0.000018\\
-			\hline
-			2nd & 1.082579 & 0.000320 & 1.082543 & 0.000312 & 1.082600 & 0.000314 & 1.082560 & 0.000310\\
-			\hline
-			3rd & 1.089734 & 0.000559 & 1.089601 & 0.000550 & 1.089649 & 0.000554 & 1.089634 & 0.000553\\
-			\hline
-			4th & 1.079535 & 0.000524 & 1.079535 & 0.000524 & 1.079467 & 0.000516 & 1.079486 & 0.000511\\
-			\hline
-			Linear SIR & 1.112544 & 0.001242 & 1.112644 & 0.001239 & 1.112546 & 0.001238 & 1.112544 & 0.001236\\
-			\hline
-		\end{tabular}
-	\end{footnotesize}
-\end{table}
-
-\subsubsection{SIR Incidence-Prevalence Ratio}
-
-\paragraph{Estimation}
-
-We use basic arithmetic functions in R to calculate the incidence to prevalence ratio for each observation in time and to derive the $\rr$ estimate.
-
-I don't currently have a standard error calculation.
-One thing we could do is use the standard deviation of the $\frac{IPR(t)}{D}$ values for across all $t$, since we take the mean of them as the estimate for $\rr$.
-
-\paragraph{Comments}
-
-The estimates for $\rr$ seem to be very imprecise for this model.
-Looking at the baseline data, the $\rr$ estimates vary fairly widely given that the 5 data sets were generated under the same conditions.
-We also see that the estimates change massively as we change $Y_0$ or $\sigma_X$ and $\sigma_Y$.
-As we see below, the estimate for $\rr$ drops from around 45 for $Y_0 = 1$ to 0.9 for $Y_0 = 1000$.
-Similarly, the estimate for $\rr$ grows from around 0.9 for ($\sigma_X$, $\sigma_Y$) = (10, 1) to 200 for ($\sigma_X$, $\sigma_Y$) = (15000, 1500).
-In general, we see that the estimates for $\rr$ decrease as $Y_0$ increases relative to the total population and that they increase as we increase
-$\sigma_X$ and $\sigma_Y$.
-
-\begin{table}[H]
-
-	\caption{$\rr$ Estimates and Std. Errs, SIRIPR Model,
-		Baseline Data, $X_0 = 99950, Y_0 = 50$, 
-		$\sigma_X = 10, \sigma_Y = 1$}
-	\begin{footnotesize}
-		\hskip -1cm
-		\begin{tabular}{l|r|r|r|r|r|r|r|r}
-			\hline
-			Data Set & Auto Est & Auto SE & AutoM Est & AutoM SE & Norm Est & Norm SE & NormM Est & NormM SE\\
-			\hline
-			Data Set & Auto Est & Auto SE & AutoM Est & AutoM SE & Norm Est & Norm SE & NormM Est & NormM SE\\
-			\hline
-			Baseline 1 & 1.170709 & 1.726236 & 0.118299 & 0.501111 & 1.566640 & 2.461139 & 0.129870 & 0.410478\\
-			\hline
-			Baseline 2 & 1.169642 & 1.736295 & 0.134667 & 0.531973 & 1.405544 & 1.947442 & 0.143524 & 0.470087\\
-			\hline
-			Baseline 3 & 1.024440 & 1.566376 & 0.147574 & 0.616955 & 1.327697 & 2.039847 & 0.135604 & 0.415505\\
-			\hline
-			Baseline 4 & 1.078543 & 1.635074 & 0.126040 & 0.518177 & 1.538235 & 2.350329 & 0.136590 & 0.461818\\
-			\hline
-			Baseline 5 & 1.082233 & 1.600320 & 0.130253 & 0.464518 & 1.287763 & 2.121584 & 0.124392 & 0.379142\\
-			\hline
-		\end{tabular}
-	\end{footnotesize}
-\end{table}
-
-\begin{table}[H]
-
-	\caption{$\rr$ Estimates and Std. Errs, SIRIPR Model,
-		Different Starting Populations, 
-		$\sigma_X = 10, \sigma_Y = 1$}
-	\begin{footnotesize}
-		\hskip -1.7cm
-		\begin{tabular}{l|r|r|r|r|r|r|r|r}
-			\hline
-			Data Set & Auto Est & Auto SE & AutoM Est & AutoM SE & Norm Est & Norm SE & NormM Est & NormM SE\\
-			\hline
-			$X_0 = 99999, Y_0 = 1$ & 132.003557 & 979.057271 & 0.637136 & 7.591563 & 571.022230 & 8311.989460 & 0.917525 & 10.386777\\
-			\hline
-			$X_0 = 99990, Y_0 = 10$ & 5.566424 & 8.249684 & 0.169114 & 1.056522 & 6.817065 & 11.495275 & 0.152642 & 0.759693\\
-			\hline
-			$X_0 = 99900, Y_0 = 100$ & 0.595913 & 0.859870 & 0.129120 & 0.359819 & 0.764240 & 1.162178 & 0.121305 & 0.305982\\
-			\hline
-			$X_0 = 99000, Y_0 = 1000$ & 0.132387 & 0.115126 & 0.117280 & 0.105368 & 0.146490 & 0.145274 & 0.117695 & 0.116146\\
-			\hline
-		\end{tabular}
-	\end{footnotesize}
-\end{table}
-
-\begin{table}[H]
-
-	\caption{$\rr$ Estimates and Std. Errs, SIRIPR Model,
-		Different Variances, 
-		$X_0 = 99000, Y_0 = 1000$}
-	\begin{footnotesize}
-		\hskip -1.7cm
-		\begin{tabular}{l|r|r|r|r|r|r|r|r}
-			\hline
-			Data Set & Auto Est & Auto SE & AutoM Est & AutoM SE & Norm Est & Norm SE & NormM Est & NormM SE\\
-			\hline
-			$\sigma_X = 10, \sigma_Y = 1$ & 1.178938 & 20.02418 & 1.185635 & 20.439407 & 1.219909 & 20.39467 & 1.167064 & 20.057929\\
-			\hline
-			$\sigma_X = 100, \sigma_Y = 10$ & 1.356012 & 15.38756 & 1.089050 & 18.415936 & 1.859016 & 20.12416 & 1.150239 & 19.649800\\
-			\hline
-			$\sigma_X = 300, \sigma_Y = 30$ & 3.142258 & 28.52150 & 0.969976 & 15.934410 & 3.201584 & 21.35673 & 1.085100 & 18.379016\\
-			\hline
-			$\sigma_X = 1000, \sigma_Y = 100$ & 7.305927 & 39.63369 & 1.088018 & 18.236155 & 7.978626 & 16.64830 & 0.787073 & 12.417881\\
-			\hline
-			$\sigma_X = 3000, \sigma_Y = 300$ & 16.009523 & 27.38954 & 0.424419 & 3.054379 & 30.351952 & 130.75442 & 1.357313 & 23.125797\\
-			\hline
-			$\sigma_X = 10000, \sigma_Y = 1000$ & 142.030295 & 894.32779 & 0.718062 & 8.362043 & 174.629326 & 706.02904 & 0.390140 & 3.444363\\
-			\hline
-		\end{tabular}
-	\end{footnotesize}
-\end{table}
-
-\begin{table}[H]
-
-	\caption{$\rr$ Estimates and Std. Errs, SIRIPR Model,
-		Other Models, $X_0 = 99950, Y_0 = 50$, 
-		$\sigma_X = 10, \sigma_Y = 1$}
-	\begin{footnotesize}
-		\hskip -1cm
-		\begin{tabular}{l|r|r|r|r|r|r|r|r}
-			\hline
-			Data Set & Auto Est & Auto SE & AutoM Est & AutoM SE & Norm Est & Norm SE & NormM Est & NormM SE\\
-			\hline
-			0th & 1.779121 & 2.482162 & 0.009267 & 0.089052 & 2.352857 & 3.236806 & 0.021922 & 0.200162\\
-			\hline
-			1st & 3.392178 & 21.127161 & 5.199764 & 44.528822 & 4.615877 & 45.803986 & 4.032633 & 33.360591\\
-			\hline
-			2nd & 1.453143 & 5.808155 & 1.073020 & 2.586030 & 1.608484 & 5.985549 & 1.076302 & 2.657303\\
-			\hline
-			3rd & 1.360154 & 1.299049 & 1.192788 & 1.033645 & 1.453000 & 1.464758 & 1.195854 & 1.134779\\
-			\hline
-			4th & 1.300569 & 1.215989 & 1.300569 & 1.215989 & 1.426296 & 1.590604 & 1.153380 & 1.398890\\
-			\hline
-			Linear SIR & 1.379871 & 1.311406 & 1.208521 & 1.003065 & 1.516803 & 1.554241 & 1.208578 & 1.125440\\
-			\hline
-		\end{tabular}
-	\end{footnotesize}
-\end{table}
-
-\subsubsection{SIR Smoothed Incidence-Prevalence Ratio}
-
-\paragraph{Estimation}
-
-This estimation method is similar to the incidence-prevalence ratio estimator above, except that we first smooth the data for the number of susceptible and infected using a smoothing spline.
-We use basic arithmetic functions in R to calculate the incidence to prevalence ratio for each observation in time and to derive the $\rr$ estimate.
-
-We estimate the standard error using the jackknife method.
-For every observation, we fit a smoothed spline to X(t) and Y(t) after taking out that specific observation, then take the standard deviation of $t$ the $\rr$ estimates.
-
-\paragraph{Comments}
-
-Smoothing the data actually makes the $\rr$ estimates much more precise.
-All of them are now within $\pm$ 0.5 of the true $\rr$ for the data generated from the classic SIR model.
-We still see the same pattern as in the non-smoothed incidence-prevalence ratio estimates, that the estimates for $\rr$ decrease as $Y_0$ increases relative to the total population and that they increase as we increase $\sigma_X$ and $\sigma_Y$.
-For the data generated from other models, we see that there are some negative $\rr$ estimates for the 0th and 1st order linear model data.
-(This does not happen to be a problem with the non-smoothed IPR estimates.) This might be the results of having some negative incidences in the data affecting the results of the smoothing spline.
-
-The standard errors for this model are large compared to the SIR models that are soled using optim() (such as least squares).
-We also see that the standard error increases dramatically as $\sigma_X$ and $\sigma_Y$ increase.
-
-\begin{table}[H]
-
-	\caption{$\rr$ Estimates and Std. Errs, SIRSIPR Model,
-		Baseline Data, $X_0 = 99950, Y_0 = 50$, 
-		$\sigma_X = 10, \sigma_Y = 1$}
-	\begin{footnotesize}
-		\hskip -1cm
-		\begin{tabular}{l|r|r|r|r|r|r|r|r}
-			\hline
-			Data Set & Auto Est & Auto SE & AutoM Est & AutoM SE & Norm Est & Norm SE & NormM Est & NormM SE\\
-			\hline
-			Baseline 1 & 0.111589 & 0.013332 & 0.116714 & 0.004233 & 0.121731 & 0.008296 & 0.122992 & 0.002911\\
-			\hline
-			Baseline 2 & 0.113638 & 0.011308 & 0.115175 & 0.003540 & 0.119221 & 0.007058 & 0.120945 & 0.005199\\
-			\hline
-			Baseline 3 & 0.118958 & 0.009720 & 0.126134 & 0.011717 & 0.121108 & 0.007978 & 0.122298 & 0.004135\\
-			\hline
-			Baseline 4 & 0.129073 & 0.010889 & 0.125410 & 0.004537 & 0.120610 & 0.008410 & 0.120497 & 0.003618\\
-			\hline
-			Baseline 5 & 0.122424 & 0.011326 & 0.129906 & 0.004269 & 0.118352 & 0.007518 & 0.124123 & 0.002414\\
-			\hline
-		\end{tabular}
-	\end{footnotesize}
-\end{table}
-
-\begin{table}[H]
-
-	\caption{$\rr$ Estimates and Std. Errs, SIRSIPR Model,
-		Different Starting Populations, 
-		$\sigma_X = 10, \sigma_Y = 1$}
-	\begin{footnotesize}
-		\hskip -1.7cm
-		\begin{tabular}{l|r|r|r|r|r|r|r|r}
-			\hline
-			Data Set & Auto Est & Auto SE & AutoM Est & AutoM SE & Norm Est & Norm SE & NormM Est & NormM SE\\
-			\hline
-			$X_0 = 99999, Y_0 = 1$ & 0.496556 & 0.404930 & 0.249343 & 0.080732 & 0.244347 & 0.534660 & 0.279914 & 0.121538\\
-			\hline
-			$X_0 = 99990, Y_0 = 10$ & 0.082031 & 0.045989 & 0.130633 & 0.007387 & 0.116519 & 0.052715 & 0.119705 & 0.006003\\
-			\hline
-			$X_0 = 99900, Y_0 = 100$ & 0.120003 & 0.006022 & 0.119230 & 0.003320 & 0.121229 & 0.004296 & 0.120600 & 0.001531\\
-			\hline
-			$X_0 = 99000, Y_0 = 1000$ & 0.117786 & 0.001351 & 0.117888 & 0.001247 & 0.117755 & 0.001314 & 0.117804 & 0.001219\\
-			\hline
-		\end{tabular}
-	\end{footnotesize}
-\end{table}
-
-\begin{table}[H]
-
-	\caption{$\rr$ Estimates and Std. Errs, SIRSIPR Model,
-		Different Variances, 
-		$X_0 = 99000, Y_0 = 1000$}
-	\begin{footnotesize}
-		\hskip -1.7cm
-		\begin{tabular}{l|r|r|r|r|r|r|r|r}
-			\hline
-			Data Set & Auto Est & Auto SE & AutoM Est & AutoM SE & Norm Est & Norm SE & NormM Est & NormM SE\\
-			\hline
-			$\sigma_X = 10, \sigma_Y = 1$ & 0.119798 & 0.015080 & 0.119388 & 0.015098 & 0.119629 & 0.015084 & 0.119720 & 0.015063\\
-			\hline
-			$\sigma_X = 100, \sigma_Y = 10$ & 0.123180 & 0.015455 & 0.120830 & 0.015114 & 0.119951 & 0.015663 & 0.118726 & 0.015203\\
-			\hline
-			$\sigma_X = 300, \sigma_Y = 30$ & 0.108430 & 0.022619 & 0.119020 & 0.014395 & 0.118138 & 0.018715 & 0.120683 & 0.015369\\
-			\hline
-			$\sigma_X = 1000, \sigma_Y = 100$ & 0.151905 & 0.067340 & 0.122619 & 0.016990 & 0.120618 & 0.053277 & 0.138444 & 0.013843\\
-			\hline
-			$\sigma_X = 3000, \sigma_Y = 300$ & 0.148289 & 0.141669 & 0.237888 & 0.054042 & 0.121580 & 0.147039 & 0.135574 & 0.015871\\
-			\hline
-			$\sigma_X = 10000, \sigma_Y = 1000$ & -0.253735 & 0.457021 & 0.636736 & 0.092124 & 0.113880 & 0.555507 & 0.208459 & 0.018114\\
-			\hline
-		\end{tabular}
-	\end{footnotesize}
-\end{table}
-
-\begin{table}[H]
-
-	\caption{$\rr$ Estimates and Std. Errs, SIRSIPR Model,
-		Other Models, $X_0 = 99950, Y_0 = 50$, 
-		$\sigma_X = 10, \sigma_Y = 1$}
-	\begin{footnotesize}
-		\hskip -1cm
-		\begin{tabular}{l|r|r|r|r|r|r|r|r}
-			\hline
-			Data Set & Auto Est & Auto SE & AutoM Est & AutoM SE & Norm Est & Norm SE & NormM Est & NormM SE\\
-			\hline
-			0th & 0.014270 & 0.013561 & 0.002865 & 0.000486 & -0.006574 & 0.010632 & 0.005140 & 0.001792\\
-			\hline
-			1st & -9.099024 & 94554.286222 & 2.655271 & 153.459186 & 2.339653 & 190.604213 & 2.285211 & 200.382597\\
-			\hline
-			2nd & 1.045203 & 0.201211 & 1.076390 & 0.156022 & 1.042439 & 0.205421 & 1.074918 & 0.152346\\
-			\hline
-			3rd & 1.279959 & 0.299658 & 1.290916 & 0.315063 & 1.279248 & 0.295345 & 1.280916 & 0.294307\\
-			\hline
-			4th & 1.192312 & 0.187455 & 1.192312 & 0.187455 & 1.203649 & 0.200467 & 1.202999 & 0.198628\\
-			\hline
-			Linear SIR & 1.296731 & 0.221560 & 1.304489 & 0.229988 & 1.294079 & 0.215113 & 1.294364 & 0.215810\\
-			\hline
-		\end{tabular}
-	\end{footnotesize}
-\end{table}
-
-\subsubsection{SIR Log Linear}
-
-\paragraph{Estimation}
-
-We use lm() in R to fit a linear regression model to $log(\frac{X(t)}{X(0)})$ on $Z(t)$. We use the coefficient estimate and standard error from lm() as our estimate of and standard error of $\rr$.
-
-\paragraph{Comments}
-
-The $\rr$ estimates here are very accurate and precise.
-The estimates from the baseline and changing $Y_0$ data sets are all very close to the true $\rr$ value.
-Increasing $\sigma_X$ and $\sigma_Y$ does decrease the model accuracy.
-It is not quite as good for data generated from other models, but it gives reasonable results for the 0th and 1st order linear models unlike some of the other models.
-
-The standard errors for the model are relatively small as well, and there are no unusually large values.
-They tend to decrease as $Y_0$ increases and they tend to increase as $\sigma_X$ and $\sigma_Y$ increase.
-
-\begin{table}[H]
-
-	\caption{$\rr$ Estimates and Std. Errs, HARKO Model, 
-		Baseline Data, $X_0 = 99950, Y_0 = 50$, 
-		$\sigma_X = 10, \sigma_Y = 1$}
-	\begin{footnotesize}
-		\hskip -1cm
-		\begin{tabular}{l|r|r|r|r|r|r|r|r}
-			\hline
-			Data Set & Auto Est & Auto SE & AutoM Est & AutoM SE & Norm Est & Norm SE & NormM Est & NormM SE\\
-			\hline
-			Baseline 1 & 1.780655 & 0.023436 & 1.983971 & 0.006583 & 1.887483 & 0.016523 & 1.977806 & 0.004671\\
-			\hline
-			Baseline 2 & 1.831444 & 0.021164 & 2.002378 & 0.010261 & 1.902733 & 0.017365 & 2.003881 & 0.003998\\
-			\hline
-			Baseline 3 & 1.839113 & 0.022491 & 1.914380 & 0.010098 & 1.906137 & 0.016586 & 1.987737 & 0.003900\\
-			\hline
-			Baseline 4 & 1.794680 & 0.017916 & 1.840147 & 0.007965 & 1.894689 & 0.016695 & 1.990857 & 0.003619\\
-			\hline
-			Baseline 5 & 1.823206 & 0.020518 & 1.892921 & 0.011397 & 1.914908 & 0.016957 & 1.965487 & 0.003611\\
-			\hline
-		\end{tabular}
-	\end{footnotesize}
-\end{table}
-
-\begin{table}[H]
-
-	\caption{$\rr$ Estimates and Std. Errs, HARKO Model,
-		Different Starting Populations, 
-		$\sigma_X = 10, \sigma_Y = 1$}
-	\begin{footnotesize}
-		\hskip -1.7cm
-		\begin{tabular}{l|r|r|r|r|r|r|r|r}
-			\hline
-			Data Set & Auto Est & Auto SE & AutoM Est & AutoM SE & Norm Est & Norm SE & NormM Est & NormM SE\\
-			\hline
-			$X_0 = 99999, Y_0 = 1$ & 0.990097 & 0.006292 & 1.257974 & 0.013850 & 1.007582 & 0.005853 & 1.191026 & 0.003584\\
-			\hline
-			$X_0 = 99990, Y_0 = 10$ & 1.123018 & 0.021783 & 1.735611 & 0.013290 & 1.245655 & 0.024246 & 1.923291 & 0.010741\\
-			\hline
-			$X_0 = 99900, Y_0 = 100$ & 1.951514 & 0.012785 & 2.016594 & 0.005770 & 1.966805 & 0.008724 & 1.983918 & 0.002413\\
-			\hline
-			$X_0 = 99000, Y_0 = 1000$ & 2.006289 & 0.001318 & 1.995221 & 0.001014 & 2.000268 & 0.000951 & 2.000158 & 0.000544\\
-			\hline
-		\end{tabular}
-	\end{footnotesize}
-\end{table}
-
-\begin{table}[H]
-
-	\caption{$\rr$ Estimates and Std. Errs, HARKO Model, 
-		Different Variances, $X_0 = 99000$, $Y_0 = 1000$}
-	\begin{footnotesize}
-		\hskip -1.7cm
-		\begin{tabular}{l|r|r|r|r|r|r|r|r}
-			\hline
-			Data Set & Auto Est & Auto SE & AutoM Est & AutoM SE & Norm Est & Norm SE & NormM Est & NormM SE\\
-			\hline
-			$\sigma_X = 10, \sigma_Y = 1$ & 1.997497 & 0.001357 & 2.002070 & 0.001141 & 2.000316 & 0.001015 & 2.000543 & 0.000521\\
-			\hline
-			$\sigma_X = 100, \sigma_Y = 10$ & 1.918862 & 0.009993 & 1.990296 & 0.004519 & 1.968331 & 0.009515 & 2.002762 & 0.002228\\
-			\hline
-			$\sigma_X = 300, \sigma_Y = 30$ & 1.688924 & 0.027688 & 1.822846 & 0.012972 & 1.786604 & 0.021708 & 1.999747 & 0.003704\\
-			\hline
-			$\sigma_X = 1000, \sigma_Y = 100$ & 1.202346 & 0.017004 & 1.860911 & 0.008679 & 1.244529 & 0.023070 & 1.863444 & 0.009948\\
-			\hline
-			$\sigma_X = 3000, \sigma_Y = 300$ & 1.043406 & 0.009591 & 1.414679 & 0.009365 & 1.027725 & 0.010276 & 1.696223 & 0.003871\\
-			\hline
-			$\sigma_X = 10000, \sigma_Y = 1000$ & 1.044556 & 0.008612 & 1.093328 & 0.004175 & 1.021139 & 0.007466 & 1.203591 & 0.007232\\
-			\hline
-		\end{tabular}
-	\end{footnotesize}
-\end{table}
-
-\begin{table}[H]
-
-	\caption{$\rr$ Estimates and Std. Errs, HARKO Model,
-		Other Models, $X_0 = 99950, Y_0 = 50$, 
-		$\sigma_X = 10, \sigma_Y = 1$}
-	\begin{footnotesize}
-		\hskip -1cm
-		\begin{tabular}{l|r|r|r|r|r|r|r|r}
-			\hline
-			Data Set & Auto Est & Auto SE & AutoM Est & AutoM SE & Norm Est & Norm SE & NormM Est & NormM SE\\
-			\hline
-			0th & 0.992417 & 0.005898 & 0.801563 & 0.009989 & 0.984522 & 0.004772 & 0.960479 & 0.003184\\
-			\hline
-			1st & 1.200537 & 0.000641 & 1.200834 & 0.000664 & 1.201039 & 0.000648 & 1.200963 & 0.000653\\
-			\hline
-			2nd & 1.198169 & 0.000451 & 1.198486 & 0.000407 & 1.198510 & 0.000451 & 1.198652 & 0.000403\\
-			\hline
-			3rd & 1.207309 & 0.000196 & 1.207754 & 0.000188 & 1.207402 & 0.000184 & 1.207447 & 0.000187\\
-			\hline
-			4th & 1.210499 & 0.000136 & 1.210499 & 0.000136 & 1.210671 & 0.000121 & 1.210571 & 0.000114\\
-			\hline
-			Linear SIR & 1.258920 & 0.000994 & 1.258462 & 0.001020 & 1.258587 & 0.001008 & 1.258670 & 0.001007\\
-			\hline
-		\end{tabular}
-	\end{footnotesize}
-\end{table}
-
-\subsubsection{SIR Markov Chain}
-
-\paragraph{Estimation}
-
-I'm not sure what I did was correct in this case. Based on what I understood from the paper, the method I came up with for estimating $\rr$ here was using lm() in R. My logic in algebra:
-
-$$X_{t+1} \sim Binom(X_{t}, \alpha^Y)$$ $$ X_{t+1} = X_{t} * \alpha^Y$$ $$X _{t+1} / X_{t} = \alpha^Y$$ $$log(X _{t+1}) - log(X_{t}) = Y * log(\alpha)$$
-
-We would regress $log(X _{t+1}) - log(X_{t})$ on $Y$.
-However, that regression function returned unreasonable estimates for $\rr$.
-I then attempted scale $Y$ by the number of contacts per day of each member of the susceptible compartment such that $Y^*$ would be the expected number of infectious people contacted for the day (i.e., $log(X _{t+1}) - log(X_{t}) = Y^* log(\alpha)$ where $Y^* = Y/N * contacts$) The results below were on the assumption that each susceptible person would contact one other person per time step (i.e., $Y^* = Y/N$).
-
-\begin{table}[H]
-
-	\caption{$\rr$ Estimates and Std. Errs, SIRMC Model,
-		Baseline Data, $X_0 = 99950, Y_0 = 50$, 
-		$\sigma_X = 10, \sigma_Y = 1$}
-	\begin{footnotesize}
-		\hskip -1cm
-		\begin{tabular}{l|r|r|r|r|r|r|r|r}
-			\hline
-			Data Set & Auto Est & Auto SE & AutoM Est & AutoM SE & Norm Est & Norm SE & NormM Est & NormM SE\\
-			\hline
-			Baseline 1 & 5.127170 & 179.2844 & 5.234313 & 40.61000 & 5.086225 & 216.3083 & 5.100115 & 26.56028\\
-			\hline
-			Baseline 2 & 5.158918 & 189.7917 & 4.921625 & 26.13471 & 5.126360 & 217.2991 & 5.074875 & 29.08225\\
-			\hline
-			Baseline 3 & 5.160983 & 169.1971 & 5.207777 & 47.15165 & 5.083438 & 195.1501 & 5.119470 & 28.63475\\
-			\hline
-			Baseline 4 & 5.161219 & 170.6186 & 5.127197 & 34.24464 & 5.207647 & 269.4176 & 5.135429 & 31.76258\\
-			\hline
-			Baseline 5 & 5.187114 & 187.2816 & 5.131911 & 31.43095 & 5.077370 & 186.0883 & 5.098254 & 23.90021\\
-			\hline
-		\end{tabular}
-	\end{footnotesize}
-\end{table}
-
-\begin{table}[H]
-
-	\caption{$\rr$ Estimates and Std. Errs, SIRMC Model,
-		Different Starting Populations, 
-		$\sigma_X = 10, \sigma_Y = 1$}
-	\begin{footnotesize}
-		\hskip -1.7cm
-		\begin{tabular}{l|r|r|r|r|r|r|r|r}
-			\hline
-			Data Set & Auto Est & Auto SE & AutoM Est & AutoM SE & Norm Est & Norm SE & NormM Est & NormM SE\\
-			\hline
-			$X_0 = 99999, Y_0 = 1$ & 2.502983 & 31.703174 & 4.587050 & 141.618120 & NaN & 1.53910 & 4.584606 & 211.645976\\
-			\hline
-			$X_0 = 99990, Y_0 = 10$ & 5.370030 & 1402.395267 & 4.745881 & 37.893736 & 5.079299 & 1000.80267 & 4.808880 & 30.644371\\
-			\hline
-			$X_0 = 99900, Y_0 = 100$ & 5.083530 & 78.047548 & 5.141349 & 24.913580 & 5.036936 & 94.46507 & 5.137375 & 21.960267\\
-			\hline
-			$X_0 = 99000, Y_0 = 1000$ & 5.134886 & 9.346665 & 5.112739 & 7.046171 & 5.112847 & 11.30402 & 5.113115 & 7.906187\\
-			\hline
-		\end{tabular}
-	\end{footnotesize}
-\end{table}
-
-\begin{table}[H]
-
-	\caption{$\rr$ Estimates and Std. Errs, SIRMC Model,
-		Different Variances, 
-		$X_0 = 99000, Y_0 = 1000$}
-	\begin{footnotesize}
-		\hskip -1.7cm
-		\begin{tabular}{l|r|r|r|r|r|r|r|r}
-			\hline
-			Data Set & Auto Est & Auto SE & AutoM Est & AutoM SE & Norm Est & Norm SE & NormM Est & NormM SE\\
-			\hline
-			$\sigma_X = 10, \sigma_Y = 1$ & 5.107658 & 40.07322 & 5.118340 & 41.49349 & 5.112780 & 42.18333 & 5.109027 & 40.04422\\
-			\hline
-			$\sigma_X = 100, \sigma_Y = 10$ & 5.047351 & 74.96928 & 5.074418 & 40.22261 & 5.055666 & 112.53820 & 5.060957 & 39.62932\\
-			\hline
-			$\sigma_X = 300, \sigma_Y = 30$ & 5.183856 & 304.55113 & 4.929632 & 34.41019 & 5.459944 & 664.73372 & 5.129073 & 49.24831\\
-			\hline
-			$\sigma_X = 1000, \sigma_Y = 100$ & 4.267588 & 158.79219 & 5.122059 & 60.44409 & 5.741570 & 4073.99286 & 5.133013 & 56.70565\\
-			\hline
-			$\sigma_X = 3000, \sigma_Y = 300$ & NaN & 74731.85279 & 3.918439 & 18.49635 & 3.200282 & 80.21230 & 4.771752 & 45.08632\\
-			\hline
-			$\sigma_X = 10000, \sigma_Y = 1000$ & 2.120699 & 17.46286 & 3.767783 & 35.76247 & NaN & 83.73480 & 3.884264 & 23.04767\\
-			\hline
-		\end{tabular}
-	\end{footnotesize}
-\end{table}
-
-\begin{table}[H]
-
-	\caption{$\rr$ Estimates and Std. Errs, SIRMC Model,
-		Other Models, $X_0 = 99950, Y_0 = 50$, 
-		$\sigma_X = 10, \sigma_Y = 1$}
-	\begin{footnotesize}
-		\hskip -1cm
-		\begin{tabular}{l|r|r|r|r|r|r|r|r}
-			\hline
-			Data Set & Auto Est & Auto SE & AutoM Est & AutoM SE & Norm Est & Norm SE & NormM Est & NormM SE\\
-			\hline
-			0th & NaN & 6.511016e+05 & 7.976281 & 2207.762640 & NaN & 3.490885e+05 & 7.207505 & 1149.944668\\
-			\hline
-			1st & 3.436204 & 1.350819e+00 & 3.433055 & 1.341851 & 3.437569 & 1.599346e+00 & 3.444295 & 1.441849\\
-			\hline
-			2nd & 2.934223 & 6.682100e-01 & 2.939971 & 0.443031 & 2.938814 & 7.756630e-01 & 2.939860 & 0.504643\\
-			\hline
-			3rd & 2.859786 & 3.352930e-01 & 2.867682 & 0.292399 & 2.864071 & 4.244820e-01 & 2.865659 & 0.388652\\
-			\hline
-			4th & 2.907566 & 3.910230e-01 & 2.907566 & 0.391023 & 2.912020 & 5.038200e-01 & 2.912098 & 0.438172\\
-			\hline
-			Linear SIR & 2.762911 & 2.063760e-01 & 2.762471 & 0.193776 & 2.762182 & 2.720380e-01 & 2.764241 & 0.229030\\
-			\hline
-		\end{tabular}
-	\end{footnotesize}
-\end{table}
-
-\subsubsection{SIR Sequential Bayes}
-
-\paragraph{Estimation}
-
-For the results below, I assume that the prior distribution is distributed as Gamma(1,1).
-Thus, multiplying by the Poisson likelihood will give us a posterior that is also gamma distributed.
-The distribution is updated for every day in the data.
-The final estimate and standard error for $\rr$ are the mean and standard deviation of the final posterior distribution, respectively.
-In this case, these are the expected value and standard deviation of the final gamma distribution.
-
-\paragraph{Comments}
-
-The $\rr$ estimates and standard errors are dependent on the choice of prior distribution and parameters.
-The Gamma(1,1) was chosen only for mathematical simplicity, and other choices could give us vastly different results.
-
-The $\rr$ estimates here are all fairly reasonable, though they are not as close to the true $\rr$ as some of the other SIR models.
-Also, in general, the model seems to underestimate $\rr$.
-There is a spread in the estimates for the baseline data of around 0.2.
-The estimates appear to increase as $\sigma_X$ and $\sigma_Y$ increase.
-On the other hand, there does not seem to be a consistent effect with changing $Y_0$
-
-The standard errors are larger than some of the other SIR models, especially those that use lm() or optim() for estimation.
-There does not seem to be any noticeable patterns with changing the variance of the errors or the starting compartment populations.
-
-\begin{table}[H]
-
-	\caption{$\rr$ Estimates and Std. Errs, SEQBAYES Model, 
-		Baseline Data, $X_0 = 99950, Y_0 = 50$, 
-		$\sigma_X = 10, \sigma_Y = 1$}
-	\begin{footnotesize}
-		\hskip -1cm
-		\begin{tabular}{l|r|r|r|r|r|r|r|r}
-			Data Set & Auto Est & Auto SE & AutoM Est & AutoM SE & Norm Est & Norm SE & NormM Est & NormM SE\\
-			\hline
-			Baseline 1 & 1.120551 & 0.055408 & 0.834692 & 0.047821 & 1.085113 & 0.054524 & 1.000000 & 0.052342\\
-			\hline
-			Baseline 2 & 1.163677 & 0.056464 & 0.963486 & 0.051378 & 1.086715 & 0.054565 & 1.266142 & 0.058897\\
-			\hline
-			Baseline 3 & 1.006208 & 0.052505 & 1.000000 & 0.052342 & 1.000000 & 0.052342 & 1.200584 & 0.057352\\
-			\hline
-			Baseline 4 & 0.874006 & 0.048934 & 1.200129 & 0.057341 & 0.840416 & 0.047984 & 1.000000 & 0.052342\\
-			\hline
-			Baseline 5 & 1.000000 & 0.052342 & 1.000000 & 0.052342 & 1.200097 & 0.057341 & 1.169565 & 0.056606\\
-			\hline
-		\end{tabular}
-	\end{footnotesize}
-\end{table}
-
-\begin{table}[H]
-
-	\caption{$\rr$ Estimates and Std. Errs, SEQBAYES Model,
-		Different Starting Populations, 
-		$\sigma_X = 10, \sigma_Y = 1$}
-	\begin{footnotesize}
-		\hskip -1.7cm
-		\begin{tabular}{l|r|r|r|r|r|r|r|r}
-			\hline
-			Data Set & Auto Est & Auto SE & AutoM Est & AutoM SE & Norm Est & Norm SE & NormM Est & NormM SE\\
-			\hline
-			$X_0 = 99999, Y_0 = 1$ & 1.284355 & 0.059319 & 1.000000 & 0.052342 & 1.255944 & 0.058660 & 1.129035 & 0.055617\\
-			\hline
-			$X_0 = 99990, Y_0 = 10$ & 0.745920 & 0.045206 & 1.000000 & 0.052342 & 1.000000 & 0.052342 & 1.241239 & 0.058315\\
-			\hline
-			$X_0 = 99900, Y_0 = 100$ & 0.716836 & 0.044316 & 1.236492 & 0.058204 & 1.064823 & 0.054012 & 1.000000 & 0.052342\\
-			\hline
-			$X_0 = 99000, Y_0 = 1000$ & 0.753867 & 0.045447 & 1.158019 & 0.056326 & 1.048705 & 0.053602 & 1.322988 & 0.060205\\
-			\hline
-		\end{tabular}
-	\end{footnotesize}
-\end{table}
-
-\begin{table}[H]
-
-	\caption{$\rr$ Estimates and Std. Errs, SEQBAYES Model, 
-		Different Variances, $X_0 = 99000$, $Y_0 = 1000$}
-	\begin{footnotesize}
-		\hskip -1.7cm
-		\begin{tabular}{l|r|r|r|r|r|r|r|r}
-			\hline
-			Data Set & Auto Est & Auto SE & AutoM Est & AutoM SE & Norm Est & Norm SE & NormM Est & NormM SE\\
-			\hline
-			$\sigma_X = 10, \sigma_Y = 1$ & 0.793499 & 0.046626 & 0.787881 & 0.046461 & 0.762945 & 0.045719 & 0.796512 & 0.046714\\
-			\hline
-			$\sigma_X = 100, \sigma_Y = 10$ & 0.925325 & 0.050350 & 0.891440 & 0.049420 & 0.888104 & 0.049327 & 0.885803 & 0.049263\\
-			\hline
-			$\sigma_X = 300, \sigma_Y = 30$ & 0.478432 & 0.036205 & 0.510298 & 0.037391 & 0.494531 & 0.036809 & 0.502460 & 0.037103\\
-			\hline
-			$\sigma_X = 1000, \sigma_Y = 100$ & 0.897133 & 0.049577 & 0.973510 & 0.051644 & 0.524686 & 0.037914 & 0.524018 & 0.037890\\
-			\hline
-			$\sigma_X = 3000, \sigma_Y = 300$ & 1.000000 & 0.052342 & 1.000000 & 0.052342 & 1.024154 & 0.052971 & 0.984151 & 0.051926\\
-			\hline
-			$\sigma_X = 10000, \sigma_Y = 1000$ & 1.686398 & 0.067973 & 1.000000 & 0.052342 & 0.379716 & 0.032254 & 1.000000 & 0.052342\\
-			\hline
-		\end{tabular}
-	\end{footnotesize}
-\end{table}
-
-\begin{table}[H]
-
-	\caption{$\rr$ Estimates and Std. Errs, SEQBAYES Model,
-		Other Models, $X_0 = 99950, Y_0 = 50$, 
-		$\sigma_X = 10, \sigma_Y = 1$}
-	\begin{footnotesize}
-		\hskip -1cm
-		\begin{tabular}{l|r|r|r|r|r|r|r|r}
-			\hline
-			Data Set & Auto Est & Auto SE & AutoM Est & AutoM SE & Norm Est & Norm SE & NormM Est & NormM SE\\
-			\hline
-			0th & 0.751296 & 0.045369 & 1.177502 & 0.056798 & 1.000000 & 0.052342 & 1.234447 & 0.058155\\
-			\hline
-			1st & 0.997717 & 0.052283 & 0.999939 & 0.052341 & 1.251535 & 0.058556 & 0.895793 & 0.049540\\
-			\hline
-			2nd & 0.919672 & 0.050196 & 0.946849 & 0.050932 & 0.899505 & 0.049643 & 0.935652 & 0.050630\\
-			\hline
-			3rd & 1.368874 & 0.061240 & 1.343871 & 0.060678 & 1.359729 & 0.061035 & 1.365769 & 0.061171\\
-			\hline
-			4th & 1.357390 & 0.060983 & 1.357390 & 0.060983 & 1.346107 & 0.060729 & 1.345567 & 0.060716\\
-			\hline
-			Linear SIR & 1.123600 & 0.055483 & 1.156827 & 0.056297 & 1.121743 & 0.055437 & 1.194507 & 0.057207\\
-			\hline
-		\end{tabular}
-	\end{footnotesize}
-\end{table}
-
-\subsubsection{SEIR Least Squares}
-
-\paragraph{Estimation}
-
-The parameters of the SEIR model ($\beta$, $\gamma$, and $\alpha$) were estimated using optim() in R to minimize the sum of squared errors using Nelder-Mead.
-The number of iterations was limited to 50 so that the estimate for $\alpha$ did not reach infinity and break the function.
-The covariance matrix of the estimates of $\alpha$, $\beta$, an $\gamma$ were calculated using sensitivity analysis.
-The standard error for $\rr$ was then calculated by resampling $\beta$ and $\gamma$ from a multivariate normal distribution with the aforementioned covariance matrix from sensitivity analysis, then taking the standard deviation of the new $\rr$ estimates derived from the set of resampled parameters.
-
-\paragraph{Comments}
-
-In general, we might have issues with fitting the SEIR model to data generated from the classic SIR model, since $\alpha$ will go off to infinity with too many iterations of unconstrained optimization.
-
-The estimates from the SEIR model are relatively consistent for the baseline data, but they are all slightly larger than the true $\rr$ value.
-The estimates tend to become more accurate as $Y_0$ increases relative to the total population.
-However, we can see that increasing the variance of the added errors to the simulated data dramatically decreases the accuracy and precision of the $\rr$ estimates.
-This includes estimates that are negative, estimates that are orders of magnitude larger than they should be, and estimates that are NA.
-
-The standard errors are an order of magnitude larger than for the SIR model.
-They are fairly consistent between the multiple baseline set.
-The standard errors decrease dramatically as $Y_0$ increases.
-Also, as we see from the estimates above, the standard deviations fluctuate by several orders of magnitude as we change the variance of the added errors.
-
-\begin{table}[H]
-
-	\caption{$\rr$ Estimates and Std. Errs, SEIR Model, 
-		Baseline Data, $X_0 = 99950, Y_0 = 50$, 
-		$\sigma_X = 10, \sigma_Y = 1$}
-	\begin{footnotesize}
-		\hskip -1cm
-		\begin{tabular}{l|r|r|r|r|r|r|r|r}
-			\hline
-			Data Set & Auto Est & Auto SE & AutoM Est & AutoM SE & Norm Est & Norm SE & NormM Est & NormM SE\\
-			\hline
-			Baseline 1 & 1.546608 & 0.171494 & 1.326568 & 0.013100 & 3.206236 & 2.451373 & 2.455754 & 0.066867\\
-			\hline
-			Baseline 2 & 1.772431 & 0.354807 & -0.080128 & 0.031367 & -12.937346 & 150.847519 & 3.675647 & 0.123981\\
-			\hline
-			Baseline 3 & 1.781222 & 0.395661 & 1.375838 & 0.024409 & 2.482198 & 1.391814 & 2.149920 & 0.029209\\
-			\hline
-			Baseline 4 & 1.636096 & 0.594579 & 1.737715 & 0.020414 & 0.616769 & 9.237073 & 1.982802 & 0.025452\\
-			\hline
-			Baseline 5 & -10.920596 & 123.661450 & 1.846941 & 0.027824 & 4.892252 & 12.254994 & 2.457392 & 0.039093\\
-			\hline
-		\end{tabular}
-	\end{footnotesize}
-\end{table}
-
-\begin{table}[H]
-
-	\caption{$\rr$ Estimates and Std. Errs, SEIR Model,
-		Different Starting Populations, 
-		$\sigma_X = 10, \sigma_Y = 1$}
-	\begin{footnotesize}
-		\hskip -1.7cm
-		\begin{tabular}{l|r|r|r|r|r|r|r|r}
-			\hline
-			Data Set & Auto Est & Auto SE & AutoM Est & AutoM SE & Norm Est & Norm SE & NormM Est & NormM SE\\
-			\hline
-			$X_0 = 99999, Y_0 = 1$ & NA & NA & -0.182830 & 0.225877 & NA & NA & -0.484025 & 0.614292\\
-			\hline
-			$X_0 = 99990, Y_0 = 10$ & 0.973745 & 0.546827 & NA & NA & 1.052351 & 444.645186 & NA & NA\\
-			\hline
-			$X_0 = 99900, Y_0 = 100$ & -6.416001 & 25.905717 & 2.451224 & 0.118072 & -3.124388 & 6.766299 & -1.475048 & 1.343083\\
-			\hline
-			$X_0 = 99000, Y_0 = 1000$ & -1.665131 & 1.139871 & 2.132087 & 0.009851 & 2.146884 & 0.018722 & 2.114676 & 0.010444\\
-			\hline
-		\end{tabular}
-	\end{footnotesize}
-\end{table}
-
-\begin{table}[H]
-
-	\caption{$\rr$ Estimates and Std. Errs, SEIR Model, 
-		Different Variances, $X_0 = 99000$, $Y_0 = 1000$}
-	\begin{footnotesize}
-		\hskip -1.7cm
-		\begin{tabular}{l|r|r|r|r|r|r|r|r}
-			\hline
-			Data Set & Auto Est & Auto SE & AutoM Est & AutoM SE & Norm Est & Norm SE & NormM Est & NormM SE\\
-			\hline
-			$\sigma_X = 10, \sigma_Y = 1$ & 19.226063 & 0.289480 & 21.685544 & 0.346537 & 18.468777 & 0.256194 & 23.787444 & 0.195589\\
-			\hline
-			$\sigma_X = 100, \sigma_Y = 10$ & 16.976120 & 0.530197 & 19.896647 & 0.226174 & 13.179504 & 0.570207 & 15.331206 & 0.206823\\
-			\hline
-			$\sigma_X = 300, \sigma_Y = 30$ & 9.925354 & 0.770734 & 16.239531 & 0.234976 & -4.473057 & 0.190822 & 22.862570 & 0.525209\\
-			\hline
-			$\sigma_X = 1000, \sigma_Y = 100$ & -0.007144 & 0.000245 & 25.560238 & 0.933150 & -4.082639 & 0.436790 & 11.261214 & 0.216712\\
-			\hline
-			$\sigma_X = 3000, \sigma_Y = 300$ & -0.262553 & 0.003032 & -0.012345 & 0.000093 & -0.000062 & 0.000004 & 12.886121 & 0.277206\\
-			\hline
-			$\sigma_X = 10000, \sigma_Y = 1000$ & -0.000217 & 0.000031 & -5.959291 & 0.225965 & NA & NA & -0.000064 & 0.000000\\
-			\hline
-		\end{tabular}
-	\end{footnotesize}
-\end{table}
-
-\begin{table}[H]
-
-	\caption{$\rr$ Estimates and Std. Errs, SEIR Model,
-		Other Models, $X_0 = 99950, Y_0 = 50$, 
-		$\sigma_X = 10, \sigma_Y = 1$}
-	\begin{footnotesize}
-		\hskip -1cm
-		\begin{tabular}{l|r|r|r|r|r|r|r|r}
-			\hline
-			Data Set & Auto Est & Auto SE & AutoM Est & AutoM SE & Norm Est & Norm SE & NormM Est & NormM SE\\
-			\hline
-			0th & -1.034177 & 0.143642 & -0.055180 & 0.000244 & 0.055060 & 0.029086 & -0.165649 & 0.000683\\
-			\hline
-			1st & 1.040105 & 0.004954 & 1.190938 & 1.173853 & 1.043084 & 0.004767 & 1.040387 & 0.008804\\
-			\hline
-			2nd & 1.134694 & 0.000507 & 1.150572 & 0.000496 & 1.132432 & 0.000509 & 1.143565 & 0.000470\\
-			\hline
-			3rd & 1.303457 & 0.004479 & 1.288333 & 0.004306 & 1.289472 & 0.005720 & 1.287972 & 0.004548\\
-			\hline
-			4th & 1.264807 & 0.006362 & 1.253277 & 0.004502 & 1.271573 & 0.006568 & 1.247322 & 0.005298\\
-			\hline
-			Linear SIR & 1.495457 & 0.006932 & 1.496117 & 0.006626 & 1.493676 & 0.009262 & 1.488063 & 0.008507\\
-			\hline
-		\end{tabular}
-	\end{footnotesize}
-\end{table}
-
-\subsubsection{MSEIR Least Squares}
-
-\paragraph{Estimation}
-
-The parameters of the MSEIR model were estimated by finding the values that minimizes the sum of squared errors using Nelder-Mead through the constrOptim() in R.
-This function is the same as optim(), except that it allows us to constrain the search space for the parameters by specifying a constraint matrix.
-In this case, all of the parameters were constrained to be positive, as they should be by definition.
-The maximum number of iterations for the optimization was chosen to be 50 in order to limit runtime.
-
-The standard error was calculated using the delta method.
-We find the standard deviation of all of the parameters using the Hessian matrix, then use that as the covariance matrix in our delta method calculations.
-
-\paragraph{Comments}
-
-The estimates of $\rr$ are very precise under the same conditions for the classic SIR model.
-The change in $\sigma_X$ and $\sigma_Y$ do not change the estimates much.
-They do differ somewhat largely as $Y_0$ changes.
-The estimates from the data generated from other models are actually consistent and reasonable as well.
-
-The standard errors are large relative to the SEIR or SIR models solved by least squares. There are also NAs for some of the standard error estimates because of singular Hessian matrices.
-
-One thing to note is that the estimates and standard errors may also be highly dependent on the starting parameter estimates given to the optimization function, as there are 5 parameters for the MSEIR model to optimize over.
-
-\begin{table}[H]
-
-	\caption{$\rr$ Estimates and Std. Errs, COMP Model, 
-		Baseline Data, $X_0 = 99950, Y_0 = 50$, 
-		$\sigma_X = 10, \sigma_Y = 1$}
-	\begin{footnotesize}
-		\hskip -1cm
-		\begin{tabular}{l|r|r|r|r|r|r|r|r}
-			\hline
-			Data Set & Auto Est & Auto SE & AutoM Est & AutoM SE & Norm Est & Norm SE & NormM Est & NormM SE\\
-			\hline
-			Baseline 1 & 1.151396 & 0.007840 & 1.156166 & 0.075838 & 1.168638 & 0.061978 & 1.168638 & 0.069916\\
-			\hline
-			Baseline 2 & 1.151396 & 0.015427 & 1.151396 & 0.015825 & 1.168638 & 0.068289 & 1.168638 & 0.056477\\
-			\hline
-			Baseline 3 & 1.155559 & 0.127438 & 1.151396 & 0.012891 & 1.168638 & 0.068609 & 1.168638 & 0.052676\\
-			\hline
-			Baseline 4 & 1.156588 & 0.049150 & 1.154959 & 0.026360 & 1.168638 & 0.052003 & 1.126363 & 0.038690\\
-			\hline
-			Baseline 5 & 1.149516 & 0.024329 & 1.151396 & 0.014140 & 1.168638 & 0.060267 & 1.149516 & 0.029827\\
-			\hline
-		\end{tabular}
-	\end{footnotesize}
-\end{table}
-
-\begin{table}[H]
-
-	\caption{$\rr$ Estimates and Std. Errs, COMP Model,
-		Different Starting Populations, 
-		$\sigma_X = 10, \sigma_Y = 1$}
-	\begin{footnotesize}
-		\hskip -1.7cm
-		\begin{tabular}{l|r|r|r|r|r|r|r|r}
-			\hline
-			Data Set & Auto Est & Auto SE & AutoM Est & AutoM SE & Norm Est & Norm SE & NormM Est & NormM SE\\
-			\hline
-			$X_0 = 99999, Y_0 = 1$ & 0.205516 & NA & 0.220109 & NA & 0.203940 & NA & 0.220109 & NA\\
-			\hline
-			$X_0 = 99990, Y_0 = 10$ & 1.049647 & 0.040841 & 1.043576 & 0.109889 & 1.049647 & 0.042400 & 1.049647 & 0.070514\\
-			\hline
-			$X_0 = 99900, Y_0 = 100$ & 1.215624 & 0.008808 & 1.163593 & 0.005915 & 1.163593 & 0.004472 & 1.163593 & 0.007400\\
-			\hline
-			$X_0 = 99000, Y_0 = 1000$ & 1.153396 & 0.000940 & 1.143722 & 0.000599 & 1.103664 & 0.000472 & 1.153396 & 0.000970\\
-			\hline
-		\end{tabular}
-	\end{footnotesize}
-\end{table}
-
-\begin{table}[H]
-
-	\caption{$\rr$ Estimates and Std. Errs, COMP Model, 
-		Different Variances, $X_0 = 99000$, $Y_0 = 1000$}
-	\begin{footnotesize}
-		\hskip -1.7cm
-		\begin{tabular}{l|r|r|r|r|r|r|r|r}
-			\hline
-			Data Set & Auto Est & Auto SE & AutoM Est & AutoM SE & Norm Est & Norm SE & NormM Est & NormM SE\\
-			\hline
-			$\sigma_X = 10, \sigma_Y = 1$ & 1.143722 & 0.000584 & 1.147266 & 0.002023 & 1.132778 & 0.000684 & 1.132778 & 0.000696\\
-			\hline
-			$\sigma_X = 100, \sigma_Y = 10$ & 1.140700 & 0.000566 & 1.171033 & 0.000785 & 1.155093 & 0.000925 & 1.183014 & 0.001369\\
-			\hline
-			$\sigma_X = 300, \sigma_Y = 30$ & 1.174448 & 0.001166 & 1.137020 & 0.002796 & 1.236429 & 0.003236 & 1.122842 & 0.001041\\
-			\hline
-			$\sigma_X = 1000, \sigma_Y = 100$ & 1.142865 & 0.000803 & 1.154025 & 0.001744 & 1.205823 & 0.003666 & 1.399630 & 0.002187\\
-			\hline
-			$\sigma_X = 3000, \sigma_Y = 300$ & 1.142642 & 0.001985 & 1.140222 & 0.000500 & 1.150901 & 0.000913 & 1.140190 & 0.000416\\
-			\hline
-			$\sigma_X = 10000, \sigma_Y = 1000$ & 1.198172 & NA & 1.147908 & 0.000934 & 1.162161 & 0.001444 & 1.157407 & 0.000959\\
-			\hline
-		\end{tabular}
-	\end{footnotesize}
-\end{table}
-
-\begin{table}[H]
-
-	\caption{$\rr$ Estimates and Std. Errs, COMP Model,
-		Other Models, $X_0 = 99950, Y_0 = 50$, 
-		$\sigma_X = 10, \sigma_Y = 1$}
-	\begin{footnotesize}
-		\hskip -1cm
-		\begin{tabular}{l|r|r|r|r|r|r|r|r}
-			\hline
-			Data Set & Auto Est & Auto SE & AutoM Est & AutoM SE & Norm Est & Norm SE & NormM Est & NormM SE\\
-			\hline
-			0th & 1.040710 & 0.042623 & 1.038403 & 0.142913 & 1.036463 & 0.053749 & 1.036463 & 0.049341\\
-			\hline
-			1st & 1.287607 & 0.010336 & 1.287607 & 0.009681 & 1.287607 & 0.009128 & 1.287607 & 0.010393\\
-			\hline
-			2nd & 1.397864 & 0.000200 & 1.397864 & 0.002296 & 1.397864 & 0.000618 & 1.397864 & 0.001171\\
-			\hline
-			3rd & 1.364520 & 0.004846 & 1.364520 & 0.004999 & 1.364520 & 0.004918 & 1.364520 & 0.004876\\
-			\hline
-			4th & 1.397864 & 0.002116 & 1.397864 & 0.002116 & 1.397864 & 0.002426 & 1.397864 & 0.001674\\
-			\hline
-			Linear SIR & 1.463399 & 0.011059 & 1.463399 & 0.011044 & 1.463399 & 0.010852 & 1.463399 & 0.010955\\
-			\hline
-		\end{tabular}
-	\end{footnotesize}
-\end{table}
-
-\subsubsection{Next Generation (SEIR)}
-
-\paragraph{Estimation}
-
-The next generation model below is implemented for the SEIR model.
-The parameters of the SEIR model ($\beta$, $\gamma$, and $\alpha$) were estimated using constrOptim() in R to minimize the sum of squared errors using Nelder-Mead.
-The number of iterations was limited to 50 to limit the run time.
-The constrOptim() function is the same as optim(), except that it allows us to constrain the search space for the parameters by specifying a constraint matrix.
-In this case, all of the parameters were constrained to be positive, as they should be by definition.
-We then plug in the parameter estimates to the $F$ and $V$ matrices used to calculate the estimates.
-The standard errors are calculated by resampling the parameters of the SEIR model and taking the standard deviation of the $\rr$ estimates from the resampled parameters.
-The standard deviation of the parameters are calculated using the Hessian matrix given by constrOptim() for Nelder-Mead, and resampling is done under the assumption that the parameters are distributed as a multivariate Gaussian.
-
-\paragraph{Comments}
-
-Since this is the SEIR model fit to SIR data, we might have issues with fitting.
-A lot of the $\rr$ estimates are exactly 1.4.
-It is not clear why this is or why this is not the case for some of the other data sets.
-Like we see from the SEIR least squares model, there are some NA estimates, though there are no negative estimates.
-The estimates do not fluctuate as much as those from the SEIR least squares model either.
-The $\rr$ estimates also become less accurate for large values of $\sigma_X$ and $\sigma_Y$.
-Looking at the data generated from other models, the next generation does not seem to work well for the 0th and 1st order data, where the number of people in the S and aI compartment change linearly rather than exponentially.
-
-The standard errors are approximately of the same order of magnitude as those from the SEIR least squares model.
-Also, like for the SEIR least squares model, they are fairly consistent between the multiple baseline set, the standard errors decrease dramatically as $Y_0$ increases, and the standard deviations fluctuate by several orders of magnitude as we change the variance of the added errors.
-The most extreme values of standard errors that we see here are not as unreasonably gigantic as for the SEIR least squares model either.
-
-\begin{table}[H]
-
-	\caption{$\rr$ Estimates and Std. Errs, NEXTGEN Model, 
-		Baseline Data, $X_0 = 99950, Y_0 = 50$, 
-		$\sigma_X = 10, \sigma_Y = 1$}
-	\begin{footnotesize}
-		\hskip -1cm
-		\begin{tabular}{l|r|r|r|r|r|r|r|r}
-			\hline
-			Data Set & Auto Est & Auto SE & AutoM Est & AutoM SE & Norm Est & Norm SE & NormM Est & NormM SE\\
-			\hline
-			Baseline 1 & 1.421896 & 0.110167 & 1.246864 & 0.010846 & 1.715318 & 0.798366 & 1.781744 & 0.024938\\
-			\hline
-			Baseline 2 & 1.758858 & 0.366701 & 5.512315 & 0.320605 & 1.963196 & 0.846297 & 1.625233 & 0.018011\\
-			\hline
-			Baseline 3 & 1.758858 & 0.569505 & 1.421896 & 0.016966 & 2.297508 & 2.211150 & 1.858192 & 0.036113\\
-			\hline
-			Baseline 4 & 1.447204 & 0.091904 & 1.712243 & 0.020892 & 1.421896 & 0.341627 & 2.172764 & 0.038152\\
-			\hline
-			Baseline 5 & 1.421896 & 0.109030 & 2.162264 & 0.034685 & 1.625233 & 0.626916 & 2.011729 & 0.027506\\
-			\hline
-		\end{tabular}
-	\end{footnotesize}
-\end{table}
-
-\begin{table}[H]
-
-	\caption{$\rr$ Estimates and Std. Errs, NEXTGEN Model,
-		Different Starting Populations, 
-		$\sigma_X = 10, \sigma_Y = 1$}
-	\begin{footnotesize}
-		\hskip -1.7cm
-		\begin{tabular}{l|r|r|r|r|r|r|r|r}
-			\hline
-			Data Set & Auto Est & Auto SE & AutoM Est & AutoM SE & Norm Est & Norm SE & NormM Est & NormM SE\\
-			\hline
-			$X_0 = 99999, Y_0 = 1$ & 575.476314 & 5450.536541 & 205.872462 & 2972.609438 & 16.358566 & 9841.298666 & 34.387930 & 5805.567900\\
-			\hline
-			$X_0 = 99990, Y_0 = 10$ & 1.383532 & 25.019898 & 206.028474 & 18394.966469 & 3.214261 & 12206.722105 & 741.368136 & 15921.351448\\
-			\hline
-			$X_0 = 99900, Y_0 = 100$ & 4.757366 & 2.846210 & 3.819417 & 0.635151 & 15.683199 & 137.249782 & 3.874259 & 0.588683\\
-			\hline
-			$X_0 = 99000, Y_0 = 1000$ & 4.072926 & 0.317373 & 4.130517 & 0.226424 & 4.145316 & 0.359158 & 4.012320 & 0.239506\\
-			\hline
-		\end{tabular}
-	\end{footnotesize}
-\end{table}
-
-\begin{table}[H]
-
-	\caption{$\rr$ Estimates and Std. Errs, NEXTGEN Model, 
-		Different Variances, $X_0 = 99000$, $Y_0 = 1000$}
-	\begin{footnotesize}
-		\hskip -1.7cm
-		\begin{tabular}{l|r|r|r|r|r|r|r|r}
-			\hline
-			Data Set & Auto Est & Auto SE & AutoM Est & AutoM SE & Norm Est & Norm SE & NormM Est & NormM SE\\
-			\hline
-			$\sigma_X = 10, \sigma_Y = 1$ & 12.998364 & 8.755960e-01 & 8.061987 & 0.486050 & 9.400660 & 6.204380e-01 & 5.752931 & 0.261811\\
-			\hline
-			$\sigma_X = 100, \sigma_Y = 10$ & 5.404625 & 5.198610e-01 & 7.387983 & 0.436634 & 8.531874 & 1.705313e+00 & 5.978214 & 1.108848\\
-			\hline
-			$\sigma_X = 300, \sigma_Y = 30$ & 23.186330 & 6.441964e+02 & 5.963903 & 1.266200 & 322.144190 & 2.088921e+05 & 7.276865 & 0.663915\\
-			\hline
-			$\sigma_X = 1000, \sigma_Y = 100$ & 23.983516 & 1.278927e+04 & 67.517641 & 52.327269 & 11.529412 & 4.001174e+02 & 7.010464 & 0.567235\\
-			\hline
-			$\sigma_X = 3000, \sigma_Y = 300$ & 33.846154 & 1.163559e+07 & 21.915757 & 1527.656692 & 162.022472 & 4.124435e+03 & 12.140594 & 4.474694\\
-			\hline
-			$\sigma_X = 10000, \sigma_Y = 1000$ & 174.921348 & 3.898656e+03 & 11.524763 & 165.981987 & 1.025641 & 2.676852e+02 & 101.413031 & 1760.215196\\
-			\hline
-		\end{tabular}
-	\end{footnotesize}
-\end{table}
-
-\begin{table}[H]
-
-	\caption{$\rr$ Estimates and Std. Errs, NEXTGEN Model,
-		Other Models, $X_0 = 99950, Y_0 = 50$, 
-		$\sigma_X = 10, \sigma_Y = 1$}
-	\begin{footnotesize}
-		\hskip -1cm
-		\begin{tabular}{l|r|r|r|r|r|r|r|r}
-			\hline
-			Data Set & Auto Est & Auto SE & AutoM Est & AutoM SE & Norm Est & Norm SE & NormM Est & NormM SE\\
-			\hline
-			0th & 0.001360 & 0.497415 & 0.000040 & 0.004153 & 0.037223 & 146.510831 & 0.000040 & 0.010014\\
-			\hline
-			1st & 6.217726 & 78.702007 & 13.734754 & 240.363417 & 111.232168 & 4202.221696 & 8.742980 & 66.867285\\
-			\hline
-			2nd & 2.451548 & 0.016035 & 5.220397 & 0.437529 & 8.020012 & 2.063975 & 5.220397 & 0.475516\\
-			\hline
-			3rd & 5.406024 & 0.554855 & 4.833617 & 0.256668 & 7.333991 & 1.363208 & 5.536633 & 0.660877\\
-			\hline
-			4th & 9.415812 & 2.964507 & 9.415812 & 2.964507 & 5.964387 & 0.856505 & 5.964387 & 0.748410\\
-			\hline
-			Linear SIR & 14.488463 & 6.089165 & 14.488463 & 5.828112 & 14.488463 & 6.513183 & 5.383303 & 0.308536\\
-			\hline
-		\end{tabular}
-	\end{footnotesize}
-\end{table}
-
-\subsection{Exponential Growth}
-
-\subsubsection{Exponential Growth}
-
-\paragraph{Estimation}
-
-We use Poisson regression with glm() in R to fit the incidence on the time step, without an intercept.
-We then use the coefficient estimate from the regression to find the estimate for $\rr$ by plugging in to the moment generating function.
-The standard error is calculated using the standard error of the coefficient from glm() and the delta method.
-
-\paragraph{Comments}
-
-It is assumed here that the exponential growth in incidence is seen throughout the entire time span of the data set.
-This may not be true, but there doesn't seem to be a mathematical way to determine when "exponential growth" starts or stops.
-We also assume that the generation time distribution is Uniform(10,30).
-This is somewhat arbitrarily chosen.
-While we know that the mean time between a primary and secondary case is 20 days for the classic SIR data sets that we generated, we have to choose the distribution around the mean by ourselves.
-
-The estimates here for $\rr$ are consistently too small compared to the true $\rr$.
-In particular, the estimates are almost all under 1 when the true $\rr$ is above 1, meaning we would predict that there is no epidemic when there actually is.
-It seems that $\rr$ increases as $Y_0$ increases as well as when $\sigma_X$ and $\sigma_Y$ increase.
-It is unclear whether this is because the model becomes more accurate or if it just gets larger in general.
-The model does not break when handling linearly shaped incidence data (although one could argue that this isn't exponential growth, so this model does not apply here in these cases.)
-
-The standard errors are all fairly small, among the order of $10^{-3}$. They seem to decrease as $Y_0$ increases as well as when $\sigma_X$ and $\sigma_Y$ increases.
-
-\begin{table}[H]
-
-	\caption{$\rr$ Estimates and Std. Errs, EXPGROWTH Model, 
-		Baseline Data, $X_0 = 99950, Y_0 = 50$, 
-		$\sigma_X = 10, \sigma_Y = 1$}
-	\begin{footnotesize}
-		\hskip -1cm
-		\begin{tabular}{l|r|r|r|r|r|r|r|r}
-			\hline
-			Data Set & Auto Est & Auto SE & AutoM Est & AutoM SE & Norm Est & Norm SE & NormM Est & NormM SE\\
-			\hline
-			Baseline 1 & 1.012007 & 0.010027 & 0.929414 & 0.030175 & 1.018400 & 0.009172 & 0.935858 & 0.027818\\
-			\hline
-			Baseline 2 & 1.010507 & 0.010239 & 0.943961 & 0.025078 & 1.018896 & 0.009109 & 0.936850 & 0.027469\\
-			\hline
-			Baseline 3 & 1.006199 & 0.010870 & 0.930872 & 0.029627 & 1.016426 & 0.009428 & 0.934886 & 0.028163\\
-			\hline
-			Baseline 4 & 1.006496 & 0.010825 & 0.934524 & 0.028292 & 1.018095 & 0.009211 & 0.934295 & 0.028375\\
-			\hline
-			Baseline 5 & 1.009068 & 0.010445 & 0.934444 & 0.028321 & 1.015362 & 0.009569 & 0.935845 & 0.027819\\
-			\hline
-		\end{tabular}
-	\end{footnotesize}
-\end{table}
-
-\begin{table}[H]
-
-	\caption{$\rr$ Estimates and Std. Errs, EXPGROWTH Model,
-		Different Starting Populations, 
-		$\sigma_X = 10, \sigma_Y = 1$}
-	\begin{footnotesize}
-		\hskip -1.7cm
-		\begin{tabular}{l|r|r|r|r|r|r|r|r}
-			Data Set & Auto Est & Auto SE & AutoM Est & AutoM SE & Norm Est & Norm SE & NormM Est & NormM SE\\
-			\hline
-			$X_0 = 99999, Y_0 = 1$ & 1.004407 & 0.011143 & 0.767095 & 0.146339 & 1.016822 & 0.009376 & 0.757875 & 0.155300\\
-			\hline
-			$X_0 = 99990, Y_0 = 10$ & 1.008824 & 0.010481 & 0.885239 & 0.051071 & 1.015486 & 0.009553 & 0.877557 & 0.055606\\
-			\hline
-			$X_0 = 99900, Y_0 = 100$ & 1.012554 & 0.009951 & 0.961057 & 0.020060 & 1.019685 & 0.009009 & 0.961228 & 0.020014\\
-			\hline
-			$X_0 = 99000, Y_0 = 1000$ & 1.044120 & 0.006387 & 1.042065 & 0.006576 & 1.045714 & 0.006245 & 1.042041 & 0.006578\\
-			\hline
-		\end{tabular}
-	\end{footnotesize}
-\end{table}
-
-\begin{table}[H]
-
-	\caption{$\rr$ Estimates and Std. Errs, EXPGROWTH Model, 
-		Different Variances, $X_0 = 99000$, $Y_0 = 1000$}
-	\begin{footnotesize}
-		\hskip -1.7cm
-		\begin{tabular}{l|r|r|r|r|r|r|r|r}
-			\hline
-			Data Set & Auto Est & Auto SE & AutoM Est & AutoM SE & Norm Est & Norm SE & NormM Est & NormM SE\\
-			\hline
-			$\sigma_X = 10, \sigma_Y = 1$ & 1.044244 & 0.006376 & 1.041733 & 0.006606 & 1.047118 & 0.006122 & 1.042010 & 0.006580\\
-			\hline
-			$\sigma_X = 100, \sigma_Y = 10$ & 1.088170 & 0.003410 & 1.043282 & 0.006463 & 1.099191 & 0.002912 & 1.044029 & 0.006395\\
-			\hline
-			$\sigma_X = 300, \sigma_Y = 30$ & 1.123132 & 0.002065 & 1.049199 & 0.005944 & 1.130609 & 0.001855 & 1.041324 & 0.006645\\
-			\hline
-			$\sigma_X = 1000, \sigma_Y = 100$ & 1.161414 & 0.001192 & 1.041609 & 0.006618 & 1.171423 & 0.001033 & 1.040663 & 0.006707\\
-			\hline
-			$\sigma_X = 3000, \sigma_Y = 300$ & 1.194876 & 0.000738 & 1.077426 & 0.003976 & 1.210073 & 0.000594 & 1.056925 & 0.005326\\
-			\hline
-			$\sigma_X = 10000, \sigma_Y = 1000$ & 1.237467 & 0.000402 & 1.092848 & 0.003189 & 1.246201 & 0.000355 & 1.093502 & 0.003159\\
-			\hline
-		\end{tabular}
-	\end{footnotesize}
-\end{table}
-
-\begin{table}[H]
-
-	\caption{$\rr$ Estimates and Std. Errs, EXPGROWTH Model,
-		Other Models, $X_0 = 99950, Y_0 = 50$, 
-		$\sigma_X = 10, \sigma_Y = 1$}
-	\begin{footnotesize}
-		\hskip -1cm
-		\begin{tabular}{l|r|r|r|r|r|r|r|r}
-			\hline
-			Data Set & Auto Est & Auto SE & AutoM Est & AutoM SE & Norm Est & Norm SE & NormM Est & NormM SE\\
-			\hline
-			0th & 1.006482 & 0.010827 & 0.782299 & 0.131800 & 1.016043 & 0.009479 & 0.829597 & 0.089876\\
-			\hline
-			1st & 1.060996 & 0.005026 & 1.061004 & 0.005025 & 1.061431 & 0.004995 & 1.060633 & 0.005052\\
-			\hline
-			2nd & 1.073885 & 0.004183 & 1.073323 & 0.004216 & 1.074063 & 0.004172 & 1.073328 & 0.004216\\
-			\hline
-			3rd & 1.077530 & 0.003970 & 1.077039 & 0.003998 & 1.077724 & 0.003959 & 1.077108 & 0.003994\\
-			\hline
-			4th & 1.073888 & 0.004182 & 1.073888 & 0.004182 & 1.074161 & 0.004166 & 1.073508 & 0.004205\\
-			\hline
-			Linear SIR & 1.087538 & 0.003441 & 1.087399 & 0.003448 & 1.087742 & 0.003431 & 1.087338 & 0.003451\\
-			\hline04 & 0.001114\\
-			\hline
-		\end{tabular}
-	\end{footnotesize}
-\end{table}
-
-\subsubsection{Maximum Likelihood}
-
-\paragraph{Estimation}
-
-We assume that the generation time distribution is Uniform(10,30).
-This is somewhat arbitrarily chosen.
-While we know that the mean time between a primary and secondary case is 20 days for the classic SIR data sets that we generated, we have to choose the distribution around the mean by ourselves.
-We also round to the nearest integer because our data contains non-integer counts of cases.
-
-Note that the mathematical calculations are the same here as for the branching process.
-
-We use basic algebra in R to find the estimates for $\rr$, and we use the Fisher information to calculate the standard error for the MLE.
-
-\paragraph{Comments}
-
-All of the $\rr$ estimates are less than 1, while the true $\rr$ is greater than 1.
-Changing $Y_0$ does not appear to change much, but increasing $\sigma_X$ and $\sigma_Y$ seems to increase the estimate of $\rr$ towards the true value.
-It is not clear whether the estimate becomes more accurate or simply increases.
-
-The standard errors are really small, around the same order of magnitude as the optim() models.
-They decrease as $Y_0$ increases and increase as $\sigma_X$ and $\sigma_Y$ increase.
-We have some negative standard error estimates for the 0th order model data because the sum of the incidences is actually negative due to our added random error.
-This would probably not happen with actual data.
-
-\begin{table}[H]
-
-	\caption{$\rr$ Estimates and Std. Errs, MLE Model, 
-		Baseline Data, $X_0 = 99950, Y_0 = 50$, 
-		$\sigma_X = 10, \sigma_Y = 1$}
-	\begin{footnotesize}
-		\hskip -1cm
-		\begin{tabular}{l|r|r|r|r|r|r|r|r}
-			\hline
-			Data Set & Auto Est & Auto SE & AutoM Est & AutoM SE & Norm Est & Norm SE & NormM Est & NormM SE\\
-			\hline
-			Baseline 1 & 1.045203 & 0.005596 & 0.954158 & 0.004911 & 1.009321 & 0.005045 & 1.010310 & 0.005043\\
-			\hline
-			Baseline 2 & 0.927409 & 0.004482 & 1.132450 & 0.005591 & 0.985546 & 0.004868 & 1.038093 & 0.005246\\
-			\hline
-			Baseline 3 & 1.035850 & 0.005555 & 0.969459 & 0.004980 & 1.001299 & 0.005062 & 1.006797 & 0.005083\\
-			\hline
-			Baseline 4 & 0.974099 & 0.004834 & 0.971615 & 0.004748 & 0.944385 & 0.004705 & 0.994621 & 0.005009\\
-			\hline
-			Baseline 5 & 0.988996 & 0.005185 & 1.020904 & 0.005239 & 1.053116 & 0.005433 & 1.002879 & 0.004973\\
-			\hline
-		\end{tabular}
-	\end{footnotesize}
-\end{table}
-
-\begin{table}[H]
-
-	\caption{$\rr$ Estimates and Std. Errs, MLE Model,
-		Different Starting Populations, 
-		$\sigma_X = 10, \sigma_Y = 1$}
-	\begin{footnotesize}
-		\hskip -1.7cm
-		\begin{tabular}{l|r|r|r|r|r|r|r|r}
-			\hline
-			Data Set & Auto Est & Auto SE & AutoM Est & AutoM SE & Norm Est & Norm SE & NormM Est & NormM SE\\
-			\hline
-			$X_0 = 99999, Y_0 = 1$ & 1.232756 & 0.069833 & 0.976411 & 0.163888 & 1.615820 & 0.467853 & 1.192249 & 0.205440\\
-			\hline
-			$X_0 = 99990, Y_0 = 10$ & 0.949905 & 0.027784 & 1.301035 & 0.031985 & 0.884468 & 0.022576 & 1.208762 & 0.028717\\
-			\hline
-			$X_0 = 99900, Y_0 = 100$ & 1.016057 & 0.002553 & 0.984204 & 0.002475 & 1.055094 & 0.002654 & 0.990380 & 0.002500\\
-			\hline
-			$X_0 = 99000, Y_0 = 1000$ & 0.983594 & 0.000260 & 0.996016 & 0.000263 & 0.997396 & 0.000264 & 0.998214 & 0.000264\\
-			\hline
-		\end{tabular}
-	\end{footnotesize}
-\end{table}
-
-\begin{table}[H]
-
-	\caption{$\rr$ Estimates and Std. Errs, MLE Model, 
-		Different Variances, $X_0 = 99000$, $Y_0 = 1000$}
-	\begin{footnotesize}
-		\hskip -1.7cm
-		\begin{tabular}{l|r|r|r|r|r|r|r|r}
-			\hline
-			Data Set & Auto Est & Auto SE & AutoM Est & AutoM SE & Norm Est & Norm SE & NormM Est & NormM SE\\
-			\hline
-			$\sigma_X = 10, \sigma_Y = 1$ & 0.985737 & 0.000206 & 0.985711 & 0.000207 & 0.985510 & 0.000206 & 0.986968 & 0.000206\\
-			\hline
-			$\sigma_X = 100, \sigma_Y = 10$ & 0.999005 & 0.000208 & 1.001050 & 0.000208 & 1.006205 & 0.000209 & 1.013000 & 0.000211\\
-			\hline
-			$\sigma_X = 300, \sigma_Y = 30$ & 1.051749 & 0.000236 & 0.966436 & 0.000181 & 0.816122 & 0.000167 & 0.966859 & 0.000200\\
-			\hline
-			$\sigma_X = 1000, \sigma_Y = 100$ & 1.060553 & 0.000161 & 1.065634 & 0.000236 & 0.777445 & 0.000177 & 0.975480 & 0.000206\\
-			\hline
-			$\sigma_X = 3000, \sigma_Y = 300$ & 0.511344 & 0.000110 & 1.003435 & 0.000113 & 2.385157 & 0.000672 & 1.095206 & 0.000199\\
-			\hline
-			$\sigma_X = 10000, \sigma_Y = 1000$ & -3.235191 & 0.000840 & 0.952381 & 0.000089 & -2.850294 & -0.001262 & 1.532468 & 0.000209\\
-			\hline
-		\end{tabular}
-	\end{footnotesize}
-\end{table}
-
-\begin{table}[H]
-
-	\caption{$\rr$ Estimates and Std. Errs, MLE Model,
-		Other Models, $X_0 = 99950, Y_0 = 50$, 
-		$\sigma_X = 10, \sigma_Y = 1$}
-	\begin{footnotesize}
-		\hskip -1cm
-		\begin{tabular}{l|r|r|r|r|r|r|r|r}
-			\hline
-			Data Set & Auto Est & Auto SE & AutoM Est & AutoM SE & Norm Est & Norm SE & NormM Est & NormM SE\\
-			\hline
-			0th & -12.609343 & -14.168066 & 16.325348 & 42.366553 & 3.152735 & -0.768473 & 10.431474 & 8.031043\\
-			\hline
-			1st & 0.978683 & 0.000125 & 0.982726 & 0.000126 & 0.979523 & 0.000125 & 0.976522 & 0.000125\\
-			\hline
-			2nd & 1.017042 & 0.000131 & 1.013913 & 0.000131 & 1.014664 & 0.000131 & 1.014455 & 0.000131\\
-			\hline
-			3rd & 1.029675 & 0.000121 & 1.026645 & 0.000121 & 1.027745 & 0.000121 & 1.027322 & 0.000121\\
-			\hline
-			4th & 1.028890 & 0.000135 & 1.028890 & 0.000135 & 1.027876 & 0.000135 & 1.027989 & 0.000135\\
-			\hline
-			Linear SIR & 1.052763 & 0.000099 & 1.053736 & 0.000099 & 1.053703 & 0.000099 & 1.052769 & 0.000099\\
-			\hline
-		\end{tabular}
-	\end{footnotesize}
-\end{table}
-
-\subsubsection{Initial Growth}
-
-\paragraph{Estimation}
-
-We use Poisson regression with glm() in R to fit the incidence on the time step, without an intercept.
-We then use the coefficient estimate from the regression to find the estimate for $\rr$ by plugging in to the given equation.
-The duration time is assumed to be $\frac{1}{\gamma}$ from the true data.
-We use the delta method to solve for the standard errors, using the standard error given from the regression from glm().
-
-\paragraph{Comments}
-
-This model assumes that the incidence is in a period of exponential growth throughout the entire time period. I have fitted the model on the entire data sets.
-
-The estimates here are actually all pretty close to the true $\rr$, and they are extremely consistent for the baseline data sets.
-They also seem to increase as $Y_0$ increases and as $\sigma_X$ and $\sigma_Y$ increase.
-This seems to occur regardless of accuracy.
-The model also handles the 0th and 1st order data sets well, giving us somewhat reasonable estimates (although one could argue that this isn't exponential growth, so this model does not apply here in these cases.)
-
-The standard error estimates are all reasonable. They seem to decrease slightly as $Y_0$ increases and as $\sigma_X$ and $\sigma_Y$ increase.
-
-\begin{table}[H]
-
-	\caption{$\rr$ Estimates and Std. Errs, INITGROWTH Model, 
-		Baseline Data, $X_0 = 99950, Y_0 = 50$, 
-		$\sigma_X = 10, \sigma_Y = 1$}
-	\begin{footnotesize}
-		\hskip -1cm
-		\begin{tabular}{l|r|r|r|r|r|r|r|r}
-			\hline
-			Data Set & Auto Est & Auto SE & AutoM Est & AutoM SE & Norm Est & Norm SE & NormM Est & NormM SE\\
-			\hline
-			Baseline 1 & 1.122458 & 0.044665 & 0.951335 & 0.152044 & 1.135276 & 0.040490 & 0.965079 & 0.138753\\
-			\hline
-			Baseline 2 & 1.119443 & 0.045703 & 0.982261 & 0.123521 & 1.136269 & 0.040183 & 0.967189 & 0.136801\\
-			\hline
-			Baseline 3 & 1.110763 & 0.048817 & 0.954452 & 0.148942 & 1.131325 & 0.041737 & 0.963011 & 0.140689\\
-			\hline
-			Baseline 4 & 1.111362 & 0.048596 & 0.962238 & 0.141417 & 1.134666 & 0.040680 & 0.961751 & 0.141879\\
-			\hline
-			Baseline 5 & 1.116547 & 0.046721 & 0.962069 & 0.141578 & 1.129191 & 0.042424 & 0.965051 & 0.138765\\
-			\hline
-		\end{tabular}
-	\end{footnotesize}
-\end{table}
-
-\begin{table}[H]
-
-	\caption{$\rr$ Estimates and Std. Errs, INITGROWTH Model,
-		Different Starting Populations, 
-		$\sigma_X = 10, \sigma_Y = 1$}
-	\begin{footnotesize}
-		\hskip -1.7cm
-		\begin{tabular}{l|r|r|r|r|r|r|r|r}
-			\hline
-			Data Set & Auto Est & Auto SE & AutoM Est & AutoM SE & Norm Est & Norm SE & NormM Est & NormM SE\\
-			\hline
-			$X_0 = 99999, Y_0 = 1$ & 1.107146 & 0.050171 & 0.578787 & 0.985754 & 1.132118 & 0.041484 & 0.555865 & 1.065990\\
-			\hline
-			$X_0 = 99990, Y_0 = 10$ & 1.116055 & 0.046896 & 0.855171 & 0.276575 & 1.129441 & 0.042344 & 0.838084 & 0.305078\\
-			\hline
-			$X_0 = 99900, Y_0 = 100$ & 1.123557 & 0.044293 & 1.018159 & 0.096252 & 1.137845 & 0.039698 & 1.018516 & 0.096009\\
-			\hline
-			$X_0 = 99000, Y_0 = 1000$ & 1.186268 & 0.027217 & 1.182227 & 0.028098 & 1.189399 & 0.026553 & 1.182180 & 0.028108\\
-			\hline
-		\end{tabular}
-	\end{footnotesize}
-\end{table}
-
-\begin{table}[H]
-
-	\caption{$\rr$ Estimates and Std. Errs, INITGROWTH Model, 
-		Different Variances, $X_0 = 99000$, $Y_0 = 1000$}
-	\begin{footnotesize}
-		\hskip -1.7cm
-		\begin{tabular}{l|r|r|r|r|r|r|r|r}
-			\hline
-			Data Set & Auto Est & Auto SE & AutoM Est & AutoM SE & Norm Est & Norm SE & NormM Est & NormM SE\\
-			\hline
-			$\sigma_X = 10, \sigma_Y = 1$ & 1.186512 & 0.027164 & 1.181573 & 0.028243 & 1.192153 & 0.025981 & 1.182118 & 0.028121\\
-			\hline
-			$\sigma_X = 100, \sigma_Y = 10$ & 1.271572 & 0.013714 & 1.184621 & 0.027573 & 1.292533 & 0.011548 & 1.186089 & 0.027256\\
-			\hline
-			$\sigma_X = 300, \sigma_Y = 30$ & 1.337570 & 0.007951 & 1.196232 & 0.025155 & 1.351500 & 0.007077 & 1.180770 & 0.028422\\
-			\hline
-			$\sigma_X = 1000, \sigma_Y = 100$ & 1.408230 & 0.004385 & 1.181330 & 0.028297 & 1.426440 & 0.003755 & 1.179468 & 0.028713\\
-			\hline
-			$\sigma_X = 3000, \sigma_Y = 300$ & 1.468704 & 0.002614 & 1.250992 & 0.016213 & 1.495791 & 0.002069 & 1.211323 & 0.022310\\
-			\hline
-			$\sigma_X = 10000, \sigma_Y = 1000$ & 1.544051 & 0.001360 & 1.280487 & 0.012748 & 1.559289 & 0.001190 & 1.281732 & 0.012619\\
-			\hline
-
-		\end{tabular}
-	\end{footnotesize}
-\end{table}
-
-\begin{table}[H]
-
-	\caption{$\rr$ Estimates and Std. Errs, INITGROWTH Model,
-		Other Models, $X_0 = 99950, Y_0 = 50$, 
-		$\sigma_X = 10, \sigma_Y = 1$}
-	\begin{footnotesize}
-		\hskip -1cm
-		\begin{tabular}{l|r|r|r|r|r|r|r|r}
-			\hline
-			Data Set & Auto Est & Auto SE & AutoM Est & AutoM SE & Norm Est & Norm SE & NormM Est & NormM SE\\
-			\hline
-			0th & 1.111335 & 0.048605 & 0.616117 & 0.861217 & 1.130557 & 0.041983 & 0.728769 & 0.536725\\
-			\hline
-			1st & 1.219243 & 0.020940 & 1.219259 & 0.020937 & 1.220088 & 0.020800 & 1.218537 & 0.021060\\
-			\hline
-			2nd & 1.244179 & 0.017134 & 1.243097 & 0.017284 & 1.244522 & 0.017086 & 1.243106 & 0.017283\\
-			\hline
-			3rd & 1.251193 & 0.016187 & 1.250249 & 0.016312 & 1.251565 & 0.016139 & 1.250381 & 0.016295\\
-			\hline
-			4th & 1.244185 & 0.017133 & 1.244185 & 0.017133 & 1.244711 & 0.017060 & 1.243452 & 0.017235\\
-			\hline
-			Linear SIR & 1.270365 & 0.013849 & 1.270099 & 0.013879 & 1.270753 & 0.013805 & 1.269983 & 0.013893\\
-			\hline
-		\end{tabular}
-	\end{footnotesize}
-\end{table}
-
-\subsubsection{Final Size}
-
-\paragraph{Estimation}
-
-We use basic arithmetic functions in R to calculate the $\rr$ estimates.
-We use the delta method to calculate the standard error; treating the final size a proportion of the total population, we use the variance of proportions equation($ \sqrt{\frac{p(1-p)}{N}}$) to calculate the variance of the final size.
-
-\paragraph{Comments}
-
-The model depends on our estimate for the mean infectious period.
-In this case, we use the true infectious period from the classic SIR data ($\frac{1}{\gamma}$).
-The estimates and standard errors for the baseline data sets are almost exactly the same between the different ones.
-The estimates tend to become less accurate as $\sigma_X$ and $\sigma_Y$ increase.
-They increase as $Y_0$ increases as well.
-However, this may be due to the fact that the model does not take starting compartment sizes in to account; thus, it does not differentiate between the change in final size caused by changing $Y_0$ when the data are generated from the classic SIR model.
-Also, the model does not give crazy results for data that are not generated from the classic SIR model, as we might expect since all it uses for estimation is the final number of people who were infected.
-
-The standard errors are all relatively small in magnitude given the way I calculated them, and they increase as $Y_0$ and as $\sigma_X$ and $\sigma_Y$ increase.
-
-\begin{table}[H]
-
-	\caption{$\rr$ Estimates and Std. Errs, FINALSIZE Model,
-		Baseline Data$X_0 = 99950, Y_0 = 50$, 
-		$\sigma_X = 10, \sigma_Y = 1$}
-	\centering
-	\begin{footnotesize}
-		\hskip -1cm
-	\begin{tabular}[t]{l|r|r|r|r|r|r|r|r}
-		\hline
-		Data Set & Auto Est & Auto SE & AutoM Est & AutoM SE & Norm Est & Norm SE & NormM Est & NormM SE\\
-		\hline
-		Baseline 1 & 1.001228 & 7.8e-05 & 1.001179 & 7.7e-05 & 1.001262 & 8.0e-05 & 1.001264 & 8.0e-05\\
-		\hline
-		Baseline 2 & 1.001211 & 7.8e-05 & 1.001400 & 8.4e-05 & 1.001250 & 7.9e-05 & 1.001279 & 8.0e-05\\
-		\hline
-		Baseline 3 & 1.001218 & 7.8e-05 & 1.001196 & 7.7e-05 & 1.001242 & 7.9e-05 & 1.001249 & 7.9e-05\\
-		\hline
-		Baseline 4 & 1.001233 & 7.9e-05 & 1.001246 & 7.9e-05 & 1.001200 & 7.8e-05 & 1.001239 & 7.9e-05\\
-		\hline
-		Baseline 5 & 1.001195 & 7.7e-05 & 1.001247 & 7.9e-05 & 1.001273 & 8.0e-05 & 1.001263 & 8.0e-05\\
-		\hline
-	\end{tabular}
-	\end{footnotesize}
-\end{table}
-
-\begin{table}[H]
-
-	\caption{$\rr$ Estimates and Std. Errs, FINALSIZE Model,
-		Different Starting Populations, 
-		$\sigma_X = 10, \sigma_Y = 1$}
-	\begin{footnotesize}
-		\hskip -1.7cm
-		\begin{tabular}{l|r|r|r|r|r|r|r|r}
-			\hline
-			Data Set & Auto Est & Auto SE & AutoM Est & AutoM SE & Norm Est & Norm SE & NormM Est & NormM SE\\
-			\hline
-			$X_0 = 99999, Y_0 = 1$ & 1.000114 & 0.000024 & 1.000034 & 0.000013 & 1.000033 & 0.000013 & 1.000040 & 0.000014\\
-			\hline
-			$X_0 = 99990, Y_0 = 10$ & 1.000212 & 0.000033 & 1.000315 & 0.000040 & 1.000223 & 0.000033 & 1.000305 & 0.000039\\
-			\hline
-			$X_0 = 99900, Y_0 = 100$ & 1.002530 & 0.000113 & 1.002465 & 0.000111 & 1.002606 & 0.000114 & 1.002470 & 0.000111\\
-			\hline
-			$X_0 = 99000, Y_0 = 1000$ & 1.024360 & 0.000358 & 1.024661 & 0.000360 & 1.024662 & 0.000360 & 1.024670 & 0.000360\\
-			\hline
-		\end{tabular}
-	\end{footnotesize}
-\end{table}
-
-\begin{table}[H]
-
-	\caption{$\rr$ Estimates and Std. Errs, FINALSIZE Model, 
-		Different Variances, $X_0 = 99000$, $Y_0 = 1000$}
-	\begin{footnotesize}
-		\hskip -1.7cm
-		\begin{tabular}{l|r|r|r|r|r|r|r|r}
-			\hline
-			Data Set & Auto Est & Auto SE & AutoM Est & AutoM SE & Norm Est & Norm SE & NormM Est & NormM SE\\
-			\hline
-			$\sigma_X = 10, \sigma_Y = 1$ & 1.024640 & 0.000360 & 1.024544 & 0.000359 & 1.024610 & 0.000359 & 1.024641 & 0.000360\\
-			\hline
-			$\sigma_X = 100, \sigma_Y = 10$ & 1.025024 & 0.000363 & 1.025169 & 0.000364 & 1.025348 & 0.000365 & 1.025358 & 0.000365\\
-			\hline
-			$\sigma_X = 300, \sigma_Y = 30$ & 1.024494 & 0.000359 & 1.026920 & 0.000377 & 1.020803 & 0.000329 & 1.024440 & 0.000358\\
-			\hline
-			$\sigma_X = 1000, \sigma_Y = 100$ & 1.036821 & 0.000445 & 1.025126 & 0.000363 & 1.017766 & 0.000303 & 1.024120 & 0.000356\\
-			\hline
-			$\sigma_X = 3000, \sigma_Y = 300$ & 1.012305 & 0.000251 & 1.047894 & 0.000513 & 1.045165 & 0.000497 & 1.031720 & 0.000411\\
-			\hline
-			$\sigma_X = 10000, \sigma_Y = 1000$ & 1.068277 & 0.000624 & 1.055107 & 0.000554 & 0.969364 & NaN & 1.061044 & 0.000586\\
-			\hline
-		\end{tabular}
-	\end{footnotesize}
-\end{table}
-
-\begin{table}[H]
-
-	\caption{$\rr$ Estimates and Std. Errs, FINALSIZE Model,
-		Other Models, $X_0 = 99950, Y_0 = 50$, 
-		$\sigma_X = 10, \sigma_Y = 1$}
-	\begin{footnotesize}
-		\hskip -1cm
-		\begin{tabular}{l|r|r|r|r|r|r|r|r}
-			\hline
-			Data Set & Auto Est & Auto SE & AutoM Est & AutoM SE & Norm Est & Norm SE & NormM Est & NormM SE\\
-			\hline
-			0th & 1.000194 & 0.000031 & 1.000282 & 0.000038 & 1.000185 & 0.000030 & 1.000318 & 0.000040\\
-			\hline
-			1st & 1.040704 & 0.000470 & 1.040802 & 0.000470 & 1.040694 & 0.000469 & 1.040546 & 0.000469\\
-			\hline
-			2nd & 1.041921 & 0.000477 & 1.041759 & 0.000476 & 1.041815 & 0.000476 & 1.041766 & 0.000476\\
-			\hline
-			3rd & 1.046705 & 0.000506 & 1.046460 & 0.000504 & 1.046563 & 0.000505 & 1.046529 & 0.000505\\
-			\hline
-			4th & 1.041585 & 0.000475 & 1.041585 & 0.000475 & 1.041481 & 0.000474 & 1.041485 & 0.000474\\
-			\hline
-			Linear SIR & 1.061096 & 0.000587 & 1.061156 & 0.000587 & 1.061115 & 0.000587 & 1.061060 & 0.000586\\
-			\hline
-		\end{tabular}
-	\end{footnotesize}
-\end{table}
-
-\subsection{Networks}
-
-\subsubsection{Branching Process}
-
-\paragraph{Estimation}
-
-We assume that the generation time distribution is Uniform(10,30).
-This is somewhat arbitrarily chosen.
-While we know that the mean time between a primary and secondary case is 20 days for the classic SIR data sets that we generated, we have to choose the distribution around the mean by ourselves.
-We also round to the nearest integer because our data contains non-integer counts of cases.
-
-Note that the mathematical calculations are the same here as for the MLE.
-
-We use basic algebra in R to find the estimates for $\rr$, and we use the second derivative to calculate the standard error for the branching process.
-
-\paragraph{Comments}
-
-All of the $\rr$ estimates are less than 1, while the true $\rr$ is greater than 1.
-Changing $Y_0$ does not appear to change much, but increasing $\sigma_X$ and $\sigma_Y$ seems to increase the estimate of $\rr$ towards the true value.
-It is not clear whether the estimate becomes more accurate or simply increases.
-
-The standard errors are really small, around the same order of magnitude as the optim() models.
-They decrease as $Y_0$ increases and increase as $\sigma_X$ and $\sigma_Y$ increase.
-We have some negative standard error estimates for the 0th order model data because the sum of the incidences is actually negative due to our added random error.
-This would probably not happen with actual data.
-
-\begin{table}[H]
-
-	\caption{$\rr$ Estimates and Std. Errs, BRANCH Model,
-		Baseline Data$X_0 = 99950, Y_0 = 50$, 
-		$\sigma_X = 10, \sigma_Y = 1$}
-	\centering
-	\begin{footnotesize}
-		\hskip -1cm
-	\begin{tabular}[t]{l|r|r|r|r|r|r|r|r}
-		\hline
-		Data Set & Auto Est & Auto SE & AutoM Est & AutoM SE & Norm Est & Norm SE & NormM Est & NormM SE\\
-		\hline
-		Baseline 1 & 1.045203 & 0.005596 & 0.954158 & 0.004911 & 1.009321 & 0.005045 & 1.010310 & 0.005043\\
-		\hline
-		Baseline 2 & 0.927409 & 0.004482 & 1.132450 & 0.005591 & 0.985546 & 0.004868 & 1.038093 & 0.005246\\
-		\hline
-		Baseline 3 & 1.035850 & 0.005555 & 0.969459 & 0.004980 & 1.001299 & 0.005062 & 1.006797 & 0.005083\\
-		\hline
-		Baseline 4 & 0.974099 & 0.004834 & 0.971615 & 0.004748 & 0.944385 & 0.004705 & 0.994621 & 0.005009\\
-		\hline
-		Baseline 5 & 0.988996 & 0.005185 & 1.020904 & 0.005239 & 1.053116 & 0.005433 & 1.002879 & 0.004973\\
-		\hline
-	\end{tabular}
-\end{footnotesize}
-\end{table}
-
-\begin{table}[H]
-
-	\caption{$\rr$ Estimates and Std. Errs, BRANCH Model,
-		Different Starting Populations, 
-		$\sigma_X = 10, \sigma_Y = 1$}
-	\begin{footnotesize}
-		\hskip -1.7cm
-		\begin{tabular}{l|r|r|r|r|r|r|r|r}
-			\hline
-			Data Set & Auto Est & Auto SE & AutoM Est & AutoM SE & Norm Est & Norm SE & NormM Est & NormM SE\\
-			\hline
-			$X_0 = 99999, Y_0 = 1$ & 1.232756 & 0.069833 & 0.976411 & 0.163888 & 1.615820 & 0.467853 & 1.192249 & 0.205440\\
-			\hline
-			$X_0 = 99990, Y_0 = 10$ & 0.949905 & 0.027784 & 1.301035 & 0.031985 & 0.884468 & 0.022576 & 1.208762 & 0.028717\\
-			\hline
-			$X_0 = 99900, Y_0 = 100$ & 1.016057 & 0.002553 & 0.984204 & 0.002475 & 1.055094 & 0.002654 & 0.990380 & 0.002500\\
-			\hline
-			$X_0 = 99000, Y_0 = 1000$ & 0.983594 & 0.000260 & 0.996016 & 0.000263 & 0.997396 & 0.000264 & 0.998214 & 0.000264\\
-			\hline
-		\end{tabular}
-	\end{footnotesize}
-\end{table}
-
-\begin{table}[H]
-
-	\caption{$\rr$ Estimates and Std. Errs, BRANCH Model, 
-		Different Variances, $X_0 = 99000$, $Y_0 = 1000$}
-	\begin{footnotesize}
-		\hskip -1.7cm
-		\begin{tabular}{l|r|r|r|r|r|r|r|r}
-			\hline
-			Data Set & Auto Est & Auto SE & AutoM Est & AutoM SE & Norm Est & Norm SE & NormM Est & NormM SE\\
-			\hline
-			$\sigma_X = 10, \sigma_Y = 1$ & 0.985737 & 0.000206 & 0.985711 & 0.000207 & 0.985510 & 0.000206 & 0.986968 & 0.000206\\
-			\hline
-			$\sigma_X = 100, \sigma_Y = 10$ & 0.999005 & 0.000208 & 1.001050 & 0.000208 & 1.006205 & 0.000209 & 1.013000 & 0.000211\\
-			\hline
-			$\sigma_X = 300, \sigma_Y = 30$ & 1.051749 & 0.000236 & 0.966436 & 0.000181 & 0.816122 & 0.000167 & 0.966859 & 0.000200\\
-			\hline
-			$\sigma_X = 1000, \sigma_Y = 100$ & 1.060553 & 0.000161 & 1.065634 & 0.000236 & 0.777445 & 0.000177 & 0.975480 & 0.000206\\
-			\hline
-			$\sigma_X = 3000, \sigma_Y = 300$ & 0.511344 & 0.000110 & 1.003435 & 0.000113 & 2.385157 & 0.000672 & 1.095206 & 0.000199\\
-			\hline
-			$\sigma_X = 10000, \sigma_Y = 1000$& -3.235191 & 0.000840 & 0.952381 & 0.000089 & -2.850294 & 0.001262 & 1.532468 & 0.000209\\
-			\hline
-		\end{tabular}
-	\end{footnotesize}
-\end{table}
-
-\begin{table}[H]
-
-	\caption{$\rr$ Estimates and Std. Errs, BRANCH Model,
-		Other Models, $X_0 = 99950, Y_0 = 50$, 
-		$\sigma_X = 10, \sigma_Y = 1$}
-	\begin{footnotesize}
-		\hskip -1cm
-		\begin{tabular}{l|r|r|r|r|r|r|r|r}
-			\hline
-			Data Set & Auto Est & Auto SE & AutoM Est & AutoM SE & Norm Est & Norm SE & NormM Est & NormM SE\\
-			\hline
-			0th & -12.609343 & -14.168066 & 16.325348 & 42.366553 & 3.152735 & -0.768473 & 10.431474 & 8.031043\\
-			\hline
-			1st & 0.978683 & 0.000125 & 0.982726 & 0.000126 & 0.979523 & 0.000125 & 0.976522 & 0.000125\\
-			\hline
-			2nd & 1.017042 & 0.000131 & 1.013913 & 0.000131 & 1.014664 & 0.000131 & 1.014455 & 0.000131\\
-			\hline
-			3rd & 1.029675 & 0.000121 & 1.026645 & 0.000121 & 1.027745 & 0.000121 & 1.027322 & 0.000121\\
-			\hline
-			4th & 1.028890 & 0.000135 & 1.028890 & 0.000135 & 1.027876 & 0.000135 & 1.027989 & 0.000135\\
-			\hline
-			Linear SIR & 1.052763 & 0.000099 & 1.053736 & 0.000099 & 1.053703 & 0.000099 & 1.052769 & 0.000099\\
-			\hline
-		\end{tabular}
-	\end{footnotesize}
-\end{table}
 

### 2018-07-03 paper/draft_v7/draft_v7.tex
finally on git

--- 2018-07-03 paper/Version9/draft_v9.tex
@@
 Additionally, we describe  interval estimates for the above methods for estimating $\rr$ in order to better answer the question of whether $\rr > 1$, which implies an outbreak of the disease.
-These methods are the delta method and posterior distribution.
+These methods are the delta method, and posterior distribution.
 
@@
 We write the calculation for $\rr$ under contact tracing in Equation \ref{eq:r0_contacttracing}, where $X_0$ is a random variable denoting the number of individuals in the population individual 0 infects.
-Here $X(0)=N$ indicates that the initial number of susceptibles is $N$, the size of the population,
+Here $S_0=N$ indicates that the initial number of susceptibles is $N$, the size of the population,
 \begin{align}\label{eq:r0_contacttracing}
-\rr = E[ X_0 | X(0) = N].
+\rr = E[ X_0 | S_0 = N].
 \end{align}
-\cite{eames2003} describe contact tracing as an ``extreme form of targeted control, where the potential next-generation cases are the primary focus.'' That is, one focuses on treating the contacts of an infected individual rather than applying uniform treatment across a population.
+\cite{eames2003} describe contact tracing as an ``extreme form of targeted control, where the potential next-generation cases are the primary focus,'' that is, one focuses on treating the contacts of an infected individual rather than applying uniform treatment across a population.
 
@@
 
-\textit{With the many different ways to estimate $\rr$, we want to compare them all to see which ones give more accurate and reliable estimates.}  As there is no way to know the ``true'' $\rr$ of an actual disease we compare our methods by fitting them to simulated data sets and look at the resulting $\rr$ estimates and standard errors.
+To compare the different models, we fit each of them to simulated data sets and look at the resulting $\rr$ estimates and standard errors.
 Our main comparisons are done with data generated from the SIR model with random error added so that we can compare model estimates to a known true $\rr$ value.
@@
 We also look at data from different functional models to see how the estimators perform under different circumstances.
-
-%% What do you mean by describing the standard standard deviations?
-%% SG: delta_t is normally distributed with sd sigma and similarly for other methods.  How and why did we pick the numbers that we did?
-
-\textit{To see the effects of different conditions for the data, we change one of the starting conditions while holding all of the other ones constant.
-We try changing two conditions: the starting number of susceptible and infected people in the population (holding the total population constant) and the variance of the errors added on to the The starting population sizes for the $X$ and $Y$ are set to ($X_0$, $Y_0$) =  (99999, 1), (99990, 10), (99900, 100), and (99000, 1000), holding all other conditions constant.
-The variance of the added errors by scaling them to: ($\sigma_X$, $\sigma_Y$) = (10, 1), (100, 10), (500, 50), (1000, 100), (2000, 200), (5000, 500), (10000, 1000), (15000, 1500); these are done with ($X_0$, $Y_0$) = (99000, 1000).
-The errors start very small but become very large relative to the actual compartment sizes; this allows us to more clearly see the effects of the changes.}
 
@@
 %% Talk about top 3-5 methods
-%%%%%% Honestly, all the methods seem to have flaws in one way or another - maybe point out why?
 %% Talk about methods that really miss
@@
 \subsubsection{SIR Data}
-
-\textit{The datasets from the SIR model are simulated from the ODEs: \begin{align}
-	\systeme{\frac{dX}{dt} = -\frac{\beta XY}{N}, \frac{dY}{dt} = \frac{\beta XY}{N} - \gamma Y, \frac{dZ}{dt} = \gamma Z}.
-	\end{align}
-}
-
-We use arbitrarily selected the parameters to generate the time series data set from the SIR model.
+We use arbitrarily selected parameters to generate the time series data set from the SIR model.
 We first generate "baseline" data from which we can make comparisons by changing the  conditions under which the data were generated.
@@
 
-	$\beta$ = 0.006, $\gamma$ = 0.003; $\rr$ = 2
+	$\beta$ = 0.06, $\gamma$ = 0.05; $\rr$ = 1.2
 
@@
 
-	Starting X size ($X_0$): 99950
-
-	Starting Y size ($Y_0$): 50
-
-	Starting Z size: 0
+	Starting S size ($S_0$): 99950
+
+	Starting I size ($I_0$): 50
+
+	Starting R size: 0
 
@@
 
-%% Is this enough detail on describing the initial S, I, and R, or are we looking for something more specific?
-
-\textbf{Our baseline initial size for $Y_0$ is 0.05\% of the total population in our simulated environment, which is fairly large compared to the starting point for the data in most epidemics.
-However, I/we felt that this was necessary in order to see the differences between the models.}
-
-We then generate random errors that are added to the X and Y compartments at each time step in order to simulate some form of randomness.
-\textbf{The sum of these was then subtracted from the Z compartment in order to keep the total population in the system constant.} We generated data sets with two different types of errors: Gaussian and autoregressive.
-\textbf{The Gaussian errors are independently distributed with mean 0 and constant variance for the entire length of the time series.
-The autoregressive errors are generated from a first-order process of the form: $$ \epsilon_{t} = \rho * \epsilon_{t-1} + \delta_{t}$$, where $\rho$ is fixed throughout and $\delta_{t} \sim N(0, \sigma)$ with constant $\sigma$.
-The values of the variances are chosen somewhat arbitrarily, although they are intended to be fairly small at first.}
-
-We also add a monotonicity adjustment to the X and Z compartments with both of the error models.
-\textbf{This is done using the Pool Adjacent Violators Algorithm (PAVA); the algorithm checks if the observation at each time step is greater than or equal to the previous one, and changes the value of both observations to be equal to the mean of the two if the condition is not fulfilled.} This gives us a total of four error models: autoregressive (Auto), autoregressive monotonic (AutoM), Gaussian (Norm), and Gaussian monotonic (NormM).
-For our baseline data, we have ($\sigma_X, \sigma_Y$) = (10, 1).
+We also generate random errors that are added to the S and I compartments at each time step.
+We had two types of errors: independent Gaussian and autoregressive.
+We also added a monotonicity adjustment to the S and R compartments with both of the error models.
+This gives us a total of four error models: autoregressive (Auto), autoregressive monotonic (AutoM), Gaussian (Norm), and Gaussian monotonic (NormM).
+For our baseline data, we have ($\sigma_S, \sigma_I$) = (10, 1).
 We simulate 5 different baseline sets to see how consistent each of the models are.
+
+To see the effects of different conditions for the data, we change the starting population sizes for the $S$ and $I$ are to ($S_0$, $I_0$) =  (99999, 1), (99990, 10), (99900, 100), and (99000, 1000), holding all other conditions constant.
+We also change the variance of the added errors by scaling them to: ($\sigma_S$, $\sigma_I$) = (10, 1), (100, 10), (500, 50), (1000, 100), (2000, 200), (5000, 500), (10000, 1000), (15000, 1500); these are done with ($S_0$, $I_0$) = (99000, 1000) and all other conditions the same as the baseline.
 
@@
 We simulated data from 6 other models besides the classic SIR model.
-5 of these were polynomial linear regression models, from order 0 to 4 with respect to time/day.
+5 of these were linear regression models, from order 0 to 4 with respect to time/day.
 The coefficients used to generate data from the linear regression models were originally chosen by taking the approximate coefficient estimates from fitting the linear regression models to classic SIR data.
 This does not necessarily guarantee that the true $\rr$ will be 1.2, especially for the 0th and 1st order linear regression models.
-What we are interested in for these cases is to see if the models still give reasonable estimates for $\rr$.
 The other model is the linear SIR model, which is a three compartment model with the following ODEs:
@@
  The data from the linear SIR model were simulated with parameters $\beta$ = 0.06 and $\gamma$ = 0.05, such that $\rr$ = 1.2.
-
- \textbf{Random error is added to the data generated here as well, using the same 4 error models as above.
-The other specifications for these datasets are the same as for the baseline classic SIR data, namely the variance of the added errors and starting compartment sizes.}
-
-\subsubsection{Variance Estimation}
-
-\textbf{The method for calculating the standard errors for $\hat{\rr}$ depend on the model used.
-The following models take the standard error directly from the regression methods used to estimate $\rr$:}
-
-	\begin{itemize}
-		\item SIR LMA
-		\item SIR LMAT
-		\item SIR Log Linear
-		\item SIR MC
-		\item Exponential Growth
-		\item Initial Growth
-	\end{itemize}
-
-\textbf{The ones that are calculated using the delta method:}
-
-	\begin{itemize}
-		\item-SIR LS
-		\item SIR ReLS
-		\item SEIR
-		\item MSEIR
-		%% \item AMP
-		\item NGM
-		\item Final Size
-	\end{itemize}
-
-\textbf{Standard errors for the maximum likelihood and branching process models are calculated using the negative reciprocal of the Fisher information.
-The sequential Bayes model uses the standard deviation from the posterior distribution.
-We use the jackknife method to estimate the standard error of the SIR smoothed max and the SIR smoothed IPR methods.
-The jackknife calculates the standard error using the standard deviation of estimates of $\rr$ after leaving out a single data point.}
-
-\textbf{Some of the models use only one single point in time to estimate $\rr$ and therefore do not seem to have an innate source of variance.
-We currently do not have standard errors for the SIR Max and SIR IPR models.}
-
-\subsection{Results}
-
-\subsubsection{Survival Function}
-
-%% How do we want to estimate this?????????
-%% SG:
-%% 3.1.1:  So I think this one we should mention there is no direct way to estimate this from the simple data we have.
-Alternatively, you could do a parameter sweep with different beta distributions for F(a) and beta for b(a).
-%% 3.1.2:  So assuming the population is homogeneous, k = N, D = gamma^{-1}.
-So we have to estimmate $b$ from the data.
-So I would average over (change in % susceptible ) / old # infected and use that as $b$.
-%% 3.1.3:  No clue.
-
-\subsubsection{Compartment Models}
-
-\paragraph{SIR Models}
-In general, the models derived from the SIR ordinary differential equations are relatively accurate.
-Most estimates are within the range of $\pm$ 0.2 of the true $\rr$ for the baseline SIR datasets.
-The magnitude of the standard errors on the baseline datasets for the different SIR models generally depend on the method used.
-The models that estimate $\rr$ using optimization or gradient descent, namely SIR LS and SIR ReLS, have standard errors on the order of $10^{-4}$ to $10^{-7}$.
-The standard errors are somewhat larger for the methods that use regression, on the order of $10^{-1}$ to $10^{-3}$.
-
-The SIR models as a whole also do fairly well with changing the starting number of infected people.
-However, they tend not to do as well when the variance of the added errors is very large; in these situations, many of the models tend to have very large standard errors, sometimes on the order of $10^{1}$ or $10^{2}$.
-
-The exceptions to the above are the linear, incidence-prevalence ratio, and Markov chain models.
-The estimates tend to be far from the true $\rr$, and they fluctuate greatly between the baseline data sets.
-The standard errors for these models (if they exist) also tend to be very large under some of the changing conditions.
-
-Many of the SIR models do not work as well with the data that come from the polynomial linear models.
-Specifically, for the 0th and 1st order model datasets, the estimates can be unreasonably large or negative.
-The Log linear, Sequential Bayes, Smoothed max, and Max models do not have this problem.
-
-\paragraph{Other Compartment Models}
-
-All of these models use optimization to estimate the parameters that are used to calculate $\rr$.
-They tend to be fairly accurate and consistent on the baseline data, often as good as or better than those from the various SIR models.
-The SEIR models (least squares and next generation) do not handle the large added errors very well, as it tends to give ridiculously large results.
-The standard errors tend to be on the order of $10^{-2}$ to $10^{-3}$.
-However, as with the estimates above, the standard errors from the SEIR models can become unreasonably large when the variance of the added errors becomes large.
-Also, as with the SIR models, the $\rr$ estimates and standard errors from the 0th and 1st order polynomial linear model datasets are not very good, as the former is sometimes negative and the latter can be very large.
-The MSEIR model tends to do fairly well in all situations.
-
-\subsubsection{Exponential Growth}
-
-We have assumed for these models that the number of infections follows an exponential curve throughout the entire time frame of the data sets.
-This assumption certainly does not hold for the data from the 0th and 1st order linear models, and may not hold for some of the other data sets either.
-
-These models are generally not as accurate as the compartment models above.
-Most of them, except for the initial size model, tend to underestimate $\rr$, with estimates between 0.8 and 1.0.
-The standard errors are all fairly small, on the order of magnitude from $10^{-1}$ to $10^{-3}$.
-This is larger than many of the compartment model standard errors.
-The main advantage that all of these models, except for the maximum likelihood model, seem to be more robust to changing conditions and functional forms of the data compared to the compartment models.
-The estimates and standard errors for $\rr$ appear to be reasonable in all cases, including for very large added errors and for the 0th and 1st order linear model data.
-
-\subsubsection{Network Models}
-
-The branching process model (the only one in this category so far) is not very accurate.
-In particular, for the SIR datasets, the $\rr$ estimates are less than 1, while the true $\rr$ is greater than 1.
-The standard errors, though, are relatively small, around the same order of magnitude as the SIR least squares models.
-Model estimates and standard errors are fairly consistent across the baseline datasets and changing starting compartment sizes.
-However, increasing the standard deviation of the added errors tends to have a somewhat noticeable impact on model results.
-The model also does not handle the 0th order polynomial linear data well.
-
-\section{Discussion}
-\label{sec:dis}
-
-Below, we compare the estimates and standard errors of $\rr$ given by each of the models for a few selected data sets.
-The forest plots below show the point estimates and 95\% confidence intervals of $\rr$ from each model for a particular data set.
-The true $\rr$ value is also shown by the vertical red line in the graphs.
-
-All of the figures below show results from data sets that were simulated with added Gaussian errors.
-The plot on the top left shows the estimates from one of the baseline data sets.
-For the most part, the $\rr$ values range from approximately 0.8 to 1.4.
-We note that the range of values includes $\rr$ values of less than 1, meaning that some models predict that a disease is not spreading when it actually is.
-However, on the whole, the variation of estimates between models is smaller than we originally expected.
-The confidence intervals are very small relative to the estimates for the most part; exceptions to this include the SIR IPR and the Markov chain model as the exception.
-
-The plot on the top right shows the estimates for the data set where $Y_0$ is increased to 100.
-The $\rr$ estimates generally seem to be further away from the true $\rr$ value compared to the results from the baseline data sets, and the overall .
-The confidence intervals are also more often than not larger than those in the baseline.
-In particular, the standard errors for some of the SIR models increase by several orders of magnitude.
-
-The bottom left plot shows the estimates for the data set where the error added to the generated data is increased.
-For most of the models, the estimates of $\rr$ do not change very much compared to the baseline data set.
-The confidence intervals for this data set increase for some of the models, such as the SIR LMA, while they remain approximately the same as the baseline data sets for most of the others.
-
-On the bottom right, we see the $\rr$ estimates for the case when disease prevalence grows linearly, as opposed to exponentially as seen in the data sets generated under SIR conditions, where the disease prevalence grows exponentially.
-Model performance here varies depending on how well the models the aforementioned linear growth.
-Some of the models, such as the SIR Markov chain and the IPR models, do not handle this case very well, giving much larger estimates and standard errors than are reasonable.
-On the other hand, models such as the SIR maximum and final size seem to be more flexible to different trajectories.
-[The error bars do not show up if they go outside of the x-axis limits - how do we fix this?]
-
-\begin{minipage}{0.5\linewidth}
-	\includegraphics[width=3in]{BaselineFP.jpeg}
-\end{minipage}
-\begin{minipage}{0.5\linewidth}
-	\includegraphics[width=3in]{StartPopFP.jpeg}
-\end{minipage}
-
-\begin{minipage}{0.5\linewidth}
-	\includegraphics[width=3in]{VarianceFP.jpeg}
-\end{minipage}
-\begin{minipage}{0.5\linewidth}
-	\includegraphics[width=3in]{OtherFP.jpeg}
-\end{minipage}
-
-\bibliographystyle{apa}%Choose a bibliograhpic style
-\bibliography{Master}
-
-\section{Appendix}
-
-Included here are the technical details for fitting the models in R, as well as tables of the results.
+The other specifications for the data simulated from the other models (the variance of the errors and starting compartment sizes) are the same as for the baseline classic SIR data.
 
@@
 
-For the changing compartment sizes, we see that the standard errors are much larger for smaller values of $Y_0$.
-This might be because the variance of the I compartment is much larger relative to the size of the compartment, since we hold $\sigma_Y$ at 1 no matter if $Y_0$ is 1 or 1000.
-
-For the changing error variances, we see that the estimates for $\rr$ become less accurate for increasing $\sigma_X$ and $\sigma_Y$.
+For the changing compartment sizes, we see that the standard errors are much larger for smaller values of $I_0$.
+This might be because the variance of the I compartment is much larger relative to the size of the compartment, since we hold $\sigma_I$ at 1 no matter if $I_0$ is 1 or 1000.
+
+For the changing error variances, we see that the estimates for $\rr$ become less accurate for increasing $\sigma_S$ and $\sigma_I$.
 However, it is interesting to note that the standard errors are increasing for the monotonic Gaussian error data, but decreasing for the autoregressive and monotonic autoregressive error data and relatively stable for the Gaussian non-monotonic data.
@@
 	\caption{$\rr$ Estimates and Std. Errs, SIR Model, 
-		Baseline Data, $X_0 = 99950, Y_0 = 50$, 
-		$\sigma_X = 10, \sigma_Y = 1$}
-	\begin{footnotesize}
-		\hskip -1cm
-		\begin{tabular}{l|r|r|r|r|r|r|r|r}
-			\hline
-			Data Set & Auto Est & Auto SE & AutoM Est & AutoM SE & Norm Est & Norm SE & NormM Est & NormM SE\\
-			\hline
-			Baseline 1 & 1.965322 & 0.000869 & 1.983308 & 0.000882 & 2.005867 & 0.000893 & 1.994736 & 0.000884\\
-			\hline
-			Baseline 2 & 2.039608 & 0.000913 & 1.999245 & 0.000895 & 2.001217 & 0.000889 & 1.995960 & 0.000887\\
-			\hline
-			Baseline 3 & 2.005609 & 0.000901 & 1.966631 & 0.000873 & 2.016216 & 0.000901 & 1.997175 & 0.000887\\
-			\hline
-			Baseline 4 & 1.958037 & 0.000859 & 1.941109 & 0.000852 & 1.995544 & 0.000886 & 2.004107 & 0.000892\\
-			\hline
-			Baseline 5 & 2.050904 & 0.000926 & 1.964328 & 0.000866 & 2.009912 & 0.000896 & 1.990153 & 0.000881\\
+		Baseline Data, $S_0 = 99950, I_0 = 50$, 
+		$\sigma_S = 10, \sigma_I = 1$}
+	\begin{footnotesize}
+		\hskip -1cm
+	\begin{tabular}{l|r|r|r|r|r|r|r|r}
+		\hline
+		Data Set & Auto Est & Auto SE & AutoM Est & AutoM SE & Norm Est & Norm SE & NormM Est & NormM SE\\
+		\hline
+		Baseline1 & 1.199737 & 0.000476 & 1.199557 & 0.000476 & 1.199737 & 0.000475 & 1.199737 & 0.000475\\
+		\hline
+		Baseline2 & 1.199737 & 0.000475 & 1.200380 & 0.000481 & 1.200376 & 0.000482 & 1.200376 & 0.000482\\
+		\hline
+		Baseline3 & 1.200615 & 0.000485 & 1.199557 & 0.000476 & 1.200380 & 0.000481 & 1.200298 & 0.000479\\
+		\hline
+		Baseline4 & 1.200376 & 0.000481 & 1.200061 & 0.000479 & 1.200380 & 0.000481 & 1.200602 & 0.000484\\
+		\hline
+		Baseline5 & 1.200376 & 0.000481 & 1.200376 & 0.000482 & 1.200298 & 0.000479 & 1.200602 & 0.000484\\
 		\hline
@@
 		Different Starting Populations, 
-		$\sigma_X = 10, \sigma_Y = 1$}
-	\begin{footnotesize}
-		\hskip -1.7cm
-		\begin{tabular}{l|r|r|r|r|r|r|r|r}
-			\hline
-			Data Set & Auto Est & Auto SE & AutoM Est & AutoM SE & Norm Est & Norm SE & NormM Est & NormM SE\\
-			\hline
-			$X_0 = 99999, Y_0 = 1$ & 1.689292 & 0.020794 & -4.167216 & 0.168853 & 1.976254 & 0.045081 & 1.688593 & 0.029680\\
-			\hline
-			$X_0 = 99990, Y_0 = 10$ & 2.038297 & 0.004597 & 2.207003 & 0.005165 & 2.037726 & 0.004596 & 1.883084 & 0.004049\\
-			\hline
-			$X_0 = 99900, Y_0 = 100$ & 2.009581 & 0.000449 & 2.006557 & 0.000448 & 1.999236 & 0.000445 & 1.995467 & 0.000444\\
-			\hline
-			$X_0 = 99000, Y_0 = 1000$ & 2.002521 & 0.000046 & 2.001829 & 0.000046 & 1.999519 & 0.000046 & 2.000098 & 0.000046\\
+		$\sigma_S = 10, \sigma_I = 1$}
+	\begin{footnotesize}
+		\hskip -1.7cm
+	\begin{tabular}{l|r|r|r|r|r|r|r|r}
+		\hline
+		Data Set & Auto Est & Auto SE & AutoM Est & AutoM SE & Norm Est & Norm SE & NormM Est & NormM SE\\
+		\hline
+		$S_0 = 99999, I_0 = 1$ & 1.214968 & 0.095795 & 1.236988 & 0.096579 & 1.218704 & 0.099065 & 1.218704 & 0.096508\\
+		\hline
+		$S_0 = 99990, I_0 = 10$ & 1.219334 & 0.006658 & 1.219334 & 0.006854 & 1.219596 & 0.017843 & 1.219596 & 0.017847\\
+		\hline
+		$S_0 = 99900, I_0 = 100$ & 1.199311 & 0.000166 & 1.200216 & 0.000170 & 1.200336 & 0.000170 & 1.200336 & 0.000170\\
+		\hline
+		$S_0 = 99000, I_0 = 1000$ & 1.199807 & 0.000029 & 1.199807 & 0.000029 & 1.200160 & 0.000029 & 1.200160 & 0.000029\\
 		\hline
@@
 	\caption{$\rr$ Estimates and Std. Errs, SIR Model, 
-		Different Variances, $X_0 = 99000$, $Y_0 = 1000$}
-	\begin{footnotesize}
-		\hskip -1.7cm
-		\begin{tabular}{l|r|r|r|r|r|r|r|r}
-			\hline
-			Data Set & Auto Est & Auto SE & AutoM Est & AutoM SE & Norm Est & Norm SE & NormM Est & NormM SE\\
-			\hline
-			$\sigma_X = 10, \sigma_Y = 1$ & 1.927444 & 0.000059 & 1.927090 & 5.9e-05 & 1.925768 & 5.9e-05 & 1.926576 & 5.9e-05\\
-			\hline
-			$\sigma_X = 100, \sigma_Y = 10$ & 1.926786 & 0.000058 & 1.916571 & 5.9e-05 & 1.919778 & 5.9e-05 & 1.929422 & 5.9e-05\\
-			\hline
-			$\sigma_X = 300, \sigma_Y = 30$ & 1.994211 & 0.000062 & 1.976893 & 5.8e-05 & 1.930660 & 5.8e-05 & 1.920869 & 5.9e-05\\
-			\hline
-			$\sigma_X = 1000, \sigma_Y = 100$ & 1.749057 & 0.000050 & 1.736588 & 5.9e-05 & 2.003654 & 6.2e-05 & 1.897049 & 5.9e-05\\
-			\hline
-			$\sigma_X = 3000, \sigma_Y = 300$ & 1.992674 & 0.000063 & 1.681585 & 4.9e-05 & 1.959815 & 6.7e-05 & 1.770816 & 5.3e-05\\
-			\hline
-			$\sigma_X = 10000, \sigma_Y = 1000$ & -42.264180 & 0.002694 & 1.773388 & 3.1e-05 & 1.889188 & 6.6e-05 & 2.649010 & 6.3e-05\\
+		Different Variances, $S_0 = 99000$, $I_0 = 1000$}
+	\begin{footnotesize}
+		\hskip -1.7cm
+	\begin{tabular}{l|r|r|r|r|r|r|r|r}
+		\hline
+		Error & Auto Est & Auto SE & AutoM Est & AutoM SE & Norm Est & Norm SE & NormM Est & NormM SE\\
+		\hline
+		$\sigma_S = 10, \sigma_I = 1$ & 1.200061 & 2.9e-05 & 1.200005 & 2.9e-05 & 1.199951 & 2.9e-05 & 1.199916 & 2.9e-05\\
+		\hline
+		$\sigma_S = 100, \sigma_I = 10$ & 1.200317 & 2.9e-05 & 1.200427 & 2.9e-05 & 1.199963 & 2.9e-05 & 1.199734 & 2.9e-05\\
+		\hline
+		$\sigma_S = 500, \sigma_I = 50$ & 1.201642 & 2.8e-05 & 1.201871 & 2.8e-05 & 1.199996 & 2.9e-05 & 1.198872 & 2.9e-05\\
+		\hline
+		$\sigma_S = 1000, \sigma_I = 100$ & 1.203259 & 2.8e-05 & 1.203811 & 2.8e-05 & 1.199985 & 2.9e-05 & 1.197928 & 2.9e-05\\
+		\hline
+		$\sigma_S = 2000, \sigma_I = 200$ & 1.206447 & 2.7e-05 & 1.207566 & 2.7e-05 & 1.199849 & 2.9e-05 & 1.195612 & 3.0e-05\\
+		\hline
+		$\sigma_S = 5000, \sigma_I = 500$ & 1.215906 & 2.5e-05 & 1.218773 & 2.5e-05 & 1.199656 & 2.9e-05 & 1.189382 & 3.2e-05\\
+		\hline
+		$\sigma_S = 7500, \sigma_I = 750$ & 1.223554 & 2.3e-05 & 1.228037 & 2.4e-05 & 1.199579 & 2.9e-05 & 1.183914 & 3.4e-05\\
+		\hline
+		$\sigma_S = 10000, \sigma_I = 1000$ & 1.230992 & 2.2e-05 & 1.237142 & 2.2e-05 & 1.199381 & 2.9e-05 & 1.178701 & 3.6e-05\\
+		\hline
+		$\sigma_S = 15000, \sigma_I = 1500$ & 1.245535 & 1.9e-05 & 1.255188 & 2.0e-05 & 1.199137 & 2.9e-05 & 1.168520 & 4.1e-05\\
 		\hline
@@
 	\caption{$\rr$ Estimates and Std. Errs, SIR Model,
-		Other Models, $X_0 = 99950, Y_0 = 50$, 
-		$\sigma_X = 10, \sigma_Y = 1$}
-	\begin{footnotesize}
-		\hskip -1cm
-		\begin{tabular}{l|r|r|r|r|r|r|r|r}
-			\hline
-			Data Set & Auto Est & Auto SE & AutoM Est & AutoM SE & Norm Est & Norm SE & NormM Est & NormM SE\\
-			\hline
-			0th & -2.032207 & 0.244058 & 1.229963 & 0.069020 & 1.021369 & 0.056558 & 0.795460 & 0.331468\\
-			\hline
-			1st & 1.038117 & 0.000029 & 1.037889 & 0.000029 & 1.037902 & 0.000029 & 1.037887 & 0.000029\\
-			\hline
-			2nd & 1.198856 & 0.000045 & 1.198589 & 0.000045 & 1.198772 & 0.000045 & 1.198501 & 0.000045\\
-			\hline
-			3rd & 1.202405 & 0.000043 & 1.202098 & 0.000043 & 1.202350 & 0.000043 & 1.202240 & 0.000043\\
-			\hline
-			4th & 1.210597 & 0.000047 & 1.210597 & 0.000047 & 1.210194 & 0.000047 & 1.210304 & 0.000047\\
-			\hline
-			Linear SIR & 1.274346 & 0.000034 & 1.273739 & 0.000034 & 1.274229 & 0.000034 & 1.273950 & 0.000034\\
+		Other Models, $S_0 = 99950, I_0 = 50$, 
+		$\sigma_S = 10, \sigma_I = 1$}
+	\begin{footnotesize}
+		\hskip -1cm
+	\begin{tabular}{l|r|r|r|r|r|r|r|r}
+		\hline
+		Data Set & Auto Est & Auto SE & AutoM Est & AutoM SE & Norm Est & Norm SE & NormM Est & NormM SE\\
+		\hline
+		0th & 0.005468 & 0.000745 & -0.034869 & 0.000764 & -0.034869 & 7.4e-04 & 0.034869 & 0.000673\\
+		\hline
+		1st & 1.038004 & 0.000029 & 1.037900 & 0.000029 & 1.037826 & 2.9e-05 & 1.037826 & 0.000029\\
+		\hline
+		2nd & 1.199663 & 0.000045 & 1.198627 & 0.000045 & 1.199663 & 4.5e-05 & 1.198949 & 0.000045\\
+		\hline
+		3rd & 1.202909 & 0.000043 & 1.202282 & 0.000043 & 1.202414 & 4.3e-05 & 1.202414 & 0.000043\\
+		\hline
+		4th & 1.211092 & 0.000047 & 1.211092 & 0.000047 & 1.209479 & 4.7e-05 & 1.211092 & 0.000047\\
+		\hline
+		Linear SIR & 1.274529 & 0.000034 & 1.273768 & 0.000034 & 1.275057 & 3.4e-05 & 1.274745 & 0.000034\\
 		\hline
@@
 
-The $\rr$ estimates for the data with non-monotonic errors become more accurate as ($\sigma_X$, $\sigma_Y$) increase, while they become less accurate for the monotonic values.
-The standard errors are also positively correlated with the estimate of $\rr$ as long as all other parts, i.e., $X_0$ and $Y_0$ are held constant.
+The $\rr$ estimates for the data with non-monotonic errors become more accurate as ($\sigma_S$, $\sigma_I$) increase, while they become less accurate for the monotonic values.
+The standard errors are also positively correlated with the estimate of $\rr$ as long as all other parts, i.e., $S_0$ and $I_0$ are held constant.
 
@@
 	\caption{$\rr$ Estimates and Std. Errs, SIRREPAR Model,
-		Baseline Data, $X_0 = 99950, Y_0 = 50$, 
-		$\sigma_X = 10, \sigma_Y = 1$}
-	\begin{footnotesize}
-		\hskip -1cm
-		\begin{tabular}{l|r|r|r|r|r|r|r|r}
-			\hline
-			Data Set & Auto Est & Auto SE & AutoM Est & AutoM SE & Norm Est & Norm SE & NormM Est & NormM SE\\
-			\hline
-			Baseline 1 & 3.226272 & 0.000000 & 1.984447 & 0.000294 & 2.006310 & 0.000298 & 1.993558 & 0.000294\\
-			\hline
-			Baseline 2 & 3.203012 & 0.000000 & 1.999589 & 0.000299 & 2.000868 & 0.000297 & 1.995630 & 0.000296\\
-			\hline
-			Baseline 3 & 4.180804 & 0.000733 & 4.180810 & 0.000724 & 5.043355 & 0.000908 & 1.997399 & 0.000296\\
-			\hline
-			Baseline 4 & 4.180819 & 0.000709 & 4.180816 & 0.000715 & 1.995237 & 0.000295 & 2.003821 & 0.000298\\
-			\hline
-			Baseline 5 & 4.180799 & 0.000741 & 4.180815 & 0.000717 & 2.008713 & 0.000300 & 4.180812 & 0.000720\\
+		Baseline Data, $S_0 = 99950, I_0 = 50$, 
+		$\sigma_S = 10, \sigma_I = 1$}
+	\begin{footnotesize}
+		\hskip -1cm
+	\begin{tabular}{l|r|r|r|r|r|r|r|r}
+		\hline
+		Data Set & Auto Est & Auto SE & AutoM Est & AutoM SE & Norm Est & Norm SE & NormM Est & NormM SE\\
+		\hline
+		Baseline 1 & 1.200196 & 1e-06 & 1.199942 & 1e-06 & 1.200108 & 1e-06 & 1.199965 & 1e-06\\
+		\hline
+		Baseline 2 & 1.199574 & 1e-06 & 1.199733 & 1e-06 & 1.199922 & 1e-06 & 1.200219 & 1e-06\\
+		\hline
+		Baseline 3 & 1.199909 & 1e-06 & 1.200416 & 1e-06 & 1.199851 & 1e-06 & 1.200009 & 1e-06\\
+		\hline
+		Baseline 4 & 1.199735 & 1e-06 & 1.199829 & 1e-06 & 1.199925 & 1e-06 & 1.199856 & 1e-06\\
+		\hline
+		Baseline 5 & 1.199849 & 1e-06 & 1.199683 & 1e-06 & 1.199943 & 1e-06 & 1.199900 & 1e-06\\
 		\hline
@@
 		Different Starting Populations, 
-		$\sigma_X = 10, \sigma_Y = 1$}
-	\begin{footnotesize}
-		\hskip -1.7cm
-		\begin{tabular}{l|r|r|r|r|r|r|r|r}
-			\hline
-			Data Set & Auto Est & Auto SE & AutoM Est & AutoM SE & Norm Est & Norm SE & NormM Est & NormM SE\\
-			\hline
-			$X_0 = 99999, Y_0 = 1$ & 5.899238 & 0.025009 & 5.898924 & 0.051433 & 3.394166 & 0.000021 & 3.217232 & 0.000016\\
-			\hline
-			$X_0 = 99990, Y_0 = 10$ & 3.162364 & 0.000001 & 3.170152 & 0.000002 & 3.131576 & 0.000001 & 3.108194 & 0.000001\\
-			\hline
-			$X_0 = 99900, Y_0 = 100$ & 4.564960 & 0.000404 & 2.005939 & 0.000150 & 1.998880 & 0.000149 & 1.994592 & 0.000148\\
-			\hline
-			$X_0 = 99000, Y_0 = 1000$ & 47.241714 & 0.002493 & 47.241714 & 0.002511 & 47.241714 & 0.002502 & 47.241714 & 0.002502\\
+		$\sigma_S = 10, \sigma_I = 1$}
+	\begin{footnotesize}
+		\hskip -1.7cm
+	\begin{tabular}{l|r|r|r|r|r|r|r|r}
+		\hline
+		Data Set & Auto Est & Auto SE & AutoM Est & AutoM SE & Norm Est & Norm SE & NormM Est & NormM SE\\
+		\hline
+		$S_0 = 99999, I_0 = 1$ & 1.209265 & 3.7e-05 & 1.198242 & 3.5e-05 & 1.193162 & 3.4e-05 & 1.200995 & 3.5e-05\\
+		\hline
+		$S_0 = 99990, I_0 = 10$ & 1.202026 & 4.0e-06 & 1.198515 & 4.0e-06 & 1.199652 & 4.0e-06 & 1.199716 & 4.0e-06\\
+		\hline
+		$S_0 = 99900, I_0 = 100$ & 1.199895 & 1.0e-06 & 1.199895 & 1.0e-06 & 1.199976 & 1.0e-06 & 1.200103 & 1.0e-06\\
+		\hline
+		$S_0 = 99000, I_0 = 1000$ & 1.200089 & 0.0e+00 & 1.200008 & 0.0e+00 & 1.199992 & 0.0e+00 & 1.199985 & 0.0e+00\\
 		\hline
@@
 		Different Variances, 
-		$X_0 = 99000, Y_0 = 1000$}
-	\begin{footnotesize}
-		\hskip -1.7cm
-		\begin{tabular}{l|r|r|r|r|r|r|r|r}
-			\hline
-			Data Set & Auto Est & Auto SE & AutoM Est & AutoM SE & Norm Est & Norm SE & NormM Est & NormM SE\\
-			\hline
-			$\sigma_X = 10, \sigma_Y = 1$ & 1.926733 & 4e-06 & 1.927468 & 4e-06 & 1.925555 & 4e-06 & 1.926689 & 4e-06\\
-			\hline
-			$\sigma_X = 100, \sigma_Y = 10$ & 1.925550 & 4e-06 & 1.916380 & 4e-06 & 1.920075 & 4e-06 & 1.928987 & 4e-06\\
-			\hline
-			$\sigma_X = 300, \sigma_Y = 30$ & 1.994719 & 4e-06 & 1.976273 & 4e-06 & 1.930978 & 4e-06 & 1.920290 & 4e-06\\
-			\hline
-			$\sigma_X = 1000, \sigma_Y = 100$ & 1.749031 & 3e-06 & 1.736589 & 3e-06 & 2.003044 & 4e-06 & 1.897328 & 4e-06\\
-			\hline
-			$\sigma_X = 3000, \sigma_Y = 300$ & 1.990948 & 4e-06 & 1.681748 & 3e-06 & 1.959229 & 4e-06 & 1.770892 & 3e-06\\
-			\hline
-			$\sigma_X = 10000, \sigma_Y = 1000$ & 5.329640 & 3e-05 & 1.773820 & 2e-06 & 1.887014 & 4e-06 & 2.648300 & 6e-06\\
+		$S_0 = 99000, I_0 = 1000$}
+	\begin{footnotesize}
+		\hskip -1.7cm
+	\begin{tabular}{l|r|r|r|r|r|r|r|r}
+		\hline
+		Data Set & Auto Est & Auto SE & AutoM Est & AutoM SE & Norm Est & Norm SE & NormM Est & NormM SE\\
+		\hline
+		$\sigma_S = 10, \sigma_I = 1$ & 1.199998 & 0 & 1.199990 & 0 & 1.200027 & 0 & 1.199972 & 0\\
+		\hline
+		$\sigma_S = 100, \sigma_I = 10$ & 1.199920 & 0 & 1.199837 & 0 & 1.200179 & 0 & 1.199611 & 0\\
+		\hline
+		$\sigma_S = 500, \sigma_I = 50$ & 1.199579 & 0 & 1.199154 & 0 & 1.200844 & 0 & 1.198063 & 0\\
+		\hline
+		$\sigma_S = 1000, \sigma_I = 100$ & 1.199168 & 0 & 1.198307 & 0 & 1.201705 & 0 & 1.196130 & 0\\
+		\hline
+		$\sigma_S = 2000, \sigma_I = 200$ & 1.198425 & 0 & 1.196618 & 0 & 1.203466 & 0 & 1.192321 & 0\\
+		\hline
+		$\sigma_S = 5000, \sigma_I = 500$ & 1.196655 & 0 & 1.191769 & 0 & 1.208871 & 0 & 1.181229 & 0\\
+		\hline
+		$\sigma_S = 7500, \sigma_I = 750$ & 1.195734 & 0 & 1.187972 & 0 & 1.213591 & 0 & 1.172358 & 0\\
+		\hline
+		$\sigma_S = 10000, \sigma_I = 1000$ & 1.195439 & 0 & 1.184353 & 0 & 1.218455 & 0 & 1.163930 & 0\\
+		\hline
+		$\sigma_S = 15000, \sigma_I = 1500$ & 1.196493 & 0 & 1.177762 & 0 & 1.228683 & 0 & 1.148505 & 0\\
 		\hline
@@
 	\caption{$\rr$ Estimates and Std. Errs, SIRREPAR Model,
-		Other Models, $X_0 = 99950, Y_0 = 50$, 
-		$\sigma_X = 10, \sigma_Y = 1$}
-	\begin{footnotesize}
-		\hskip -1cm
-		\begin{tabular}{l|r|r|r|r|r|r|r|r}
-			\hline
-			Data Set & Auto Est & Auto SE & AutoM Est & AutoM SE & Norm Est & Norm SE & NormM Est & NormM SE\\
-			\hline
-			0th & 3.272461 & 0e+00 & 3.275535 & 0e+00 & 3.135010 & 0e+00 & 3.235197 & 0e+00\\
-			\hline
-			1st & 1.038125 & 0e+00 & 1.037880 & 0e+00 & 1.037903 & 0e+00 & 1.037881 & 0e+00\\
-			\hline
-			2nd & 1.198760 & 1e-06 & 1.198526 & 1e-06 & 1.198864 & 1e-06 & 1.198616 & 1e-06\\
-			\hline
-			3rd & 1.202489 & 1e-06 & 1.202071 & 1e-06 & 1.202246 & 1e-06 & 1.202355 & 1e-06\\
-			\hline
-			4th & 1.210605 & 1e-06 & 1.210605 & 1e-06 & 1.210220 & 1e-06 & 1.210199 & 1e-06\\
-			\hline
-			Linear SIR & 1.274175 & 1e-06 & 1.273799 & 1e-06 & 1.274314 & 1e-06 & 1.274075 & 1e-06\\
+		Other Models, $S_0 = 99950, I_0 = 50$, 
+		$\sigma_S = 10, \sigma_I = 1$}
+	\begin{footnotesize}
+		\hskip -1cm
+	\begin{tabular}{l|r|r|r|r|r|r|r|r}
+		\hline
+		Data Set & Auto Est & Auto SE & AutoM Est & AutoM SE & Norm Est & Norm SE & NormM Est & NormM SE\\
+		\hline
+		0th & 2.803084 & 0e+00 & 2.663158 & 0e+00 & 2.688581 & 0e+00 & 2.756037 & 0e+00\\
+		\hline
+		1st & 1.038122 & 0e+00 & 1.037880 & 0e+00 & 1.037900 & 0e+00 & 1.037887 & 0e+00\\
+		\hline
+		2nd & 1.198739 & 1e-06 & 1.198544 & 1e-06 & 1.198847 & 1e-06 & 1.198585 & 1e-06\\
+		\hline
+		3rd & 1.202419 & 1e-06 & 1.202136 & 1e-06 & 1.202294 & 1e-06 & 1.202288 & 1e-06\\
+		\hline
+		4th & 1.210605 & 1e-06 & 1.210605 & 1e-06 & 1.210274 & 1e-06 & 1.210259 & 1e-06\\
+		\hline
+		Linear SIR & 1.274256 & 1e-06 & 1.273828 & 1e-06 & 1.274260 & 1e-06 & 1.274030 & 1e-06\\
 		\hline
@@
 
-To calculate the estimates for $\rr$, we use lm() in R to fit the linear polynomial regression functions for X(t) and Y(t) with respect to t.
+To calculate the estimates for $\rr$, we use lm() in R to fit the linear polynomial regression functions for S(t) and I(t) with respect to t.
 We then calculated the derivatives for the linear regression functions at time 0 using our own function.
@@
 The inaccuracy is most glaring for the data with large error variances and the simulated from models besides the SIR one.
-The estimates seem to become less accurate for the data generated from the classic SIR model as $\sigma_X$ and $\sigma_Y$ increase and as $Y_0$ decreases relative to the total population.
+The estimates seem to become less accurate for the data generated from the classic SIR model as $\sigma_S$ and $\sigma_I$ increase and as $I_0$ decreases relative to the total population.
 There are also many estimates of $\rr$ that are negative, which implies that there is a negative derivative estimate for either the susceptible or infected compartment (is this supposed to be possible?).
@@
 The standard errors are also extremely large in some cases.
-This seems to hold consistently when $\sigma_X$ and $\sigma_Y$ are large for the classic SIR data as well as for the 4th order linear data.
+This seems to hold consistently when $\sigma_S$ and $\sigma_I$ are large for the classic SIR data as well as for the 4th order linear data.
 However, there are also other places where large standard error estimates appear for no apparent reason.
 The large standard errors and the above issue with the $\rr$ estimates being negative could be because we are only using a point estimate.
-In general, the only time we see fairly consistent and reasonable estimates from the linear model is for the SIR data with relatively large $Y_0$ as well as small $\sigma_X$ and $\sigma_Y$ (i.e., for the data sets where ($X_0$, $Y_0$) = (99000, 1000)).
+In general, the only time we see fairly consistent and reasonable estimates from the linear model is for the SIR data with relatively large $I_0$ as well as small $\sigma_S$ and $\sigma_I$ (i.e., for the data sets where ($S_0$, $I_0$) = (99000, 1000)).
 
@@
 	\caption{$\rr$ Estimates and Std. Errs, SIRLMA Model,
-		Baseline Data, $X_0 = 99950, Y_0 = 50$, 
-		$\sigma_X = 10, \sigma_Y = 1$}
-	\begin{footnotesize}
-		\hskip -1cm
-		\begin{tabular}{l|r|r|r|r|r|r|r|r}
-			\hline
-			Data Set & Auto Est & Auto SE & AutoM Est & AutoM SE & Norm Est & Norm SE & NormM Est & NormM SE\\
-			\hline
-			Baseline 1 & 0.864913 & 0.203402 & 1.739858 & 0.122632 & 1.359771 & 0.919629 & 1.173245 & 0.096355\\
-			\hline
-			Baseline 2 & 1.485664 & 1.654031 & 1.294412 & 0.618155 & 1.461580 & 0.734907 & 1.186621 & 0.091633\\
-			\hline
-			Baseline 3 & 1.082572 & 0.088142 & 1.389817 & 0.117032 & 0.615994 & 1.496674 & 1.088364 & 0.089282\\
-			\hline
-			Baseline 4 & 1.088362 & 0.367110 & 1.370251 & 0.159221 & 2.683109 & 27.981877 & 1.090398 & 0.059515\\
-			\hline
-			Baseline 5 & 0.871650 & 0.100744 & 1.438190 & 0.382650 & 0.829591 & 0.120911 & 0.924238 & 0.188305\\
+		Baseline Data, $S_0 = 99950, I_0 = 50$, 
+		$\sigma_S = 10, \sigma_I = 1$}
+	\begin{footnotesize}
+		\hskip -1cm
+	\begin{tabular}{l|r|r|r|r|r|r|r|r}
+		\hline
+		Data Set & Auto Est & Auto SE & AutoM Est & AutoM SE & Norm Est & Norm SE & NormM Est & NormM SE\\
+		\hline
+		Baseline 1 & 1.269690 & 0.120767 & 0.130935 & 10.750359 & 1.154610 & 0.065991 & 1.397649 & 0.233918\\
+		\hline
+		Baseline 2 & 1.187856 & 0.168327 & 1.121309 & 0.040461 & 1.193012 & 0.081712 & 1.357958 & 0.199538\\
+		\hline
+		Baseline 3 & 0.812981 & 0.084419 & 1.092512 & 0.034560 & 1.179720 & 0.054902 & 1.233221 & 0.060551\\
+		\hline
+		Baseline 4 & 1.180116 & 0.078889 & 1.168589 & 0.134484 & 1.293454 & 0.271027 & 1.165880 & 0.036177\\
+		\hline
+		Baseline 5 & 0.853203 & 0.064872 & 1.116949 & 0.041290 & 1.198555 & 0.106425 & 1.161485 & 0.032157\\
 		\hline
@@
 		Different Starting Populations, 
-		$\sigma_X = 10, \sigma_Y = 1$}
-	\begin{footnotesize}
-		\hskip -1.7cm
-		\begin{tabular}{l|r|r|r|r|r|r|r|r}
-			\hline
-			Data Set & Auto Est & Auto SE & AutoM Est & AutoM SE & Norm Est & Norm SE & NormM Est & NormM SE\\
-			\hline
-			$X_0 = 99999, Y_0 = 1$ & 1.713333 & 1.757278 & 1.187769 & 0.515629 & 1.451833 & 59.746089 & 1.068633 & 1.809596\\
-			\hline
-			$X_0 = 99990, Y_0 = 10$ & 1.197457 & 0.192831 & 0.674259 & 0.090168 & 1.095143 & 0.143146 & 1.199437 & 0.106393\\
-			\hline
-			$X_0 = 99900, Y_0 = 100$ & 0.796760 & 0.488540 & 1.022290 & 0.070029 & 0.024273 & 4.564245 & 1.265712 & 0.080303\\
-			\hline
-			$X_0 = 99000, Y_0 = 1000$ & 1.431229 & 0.105605 & 2.091993 & 0.340798 & 2.183329 & 0.428090 & 1.812900 & 0.122106\\
+		$\sigma_S = 10, \sigma_I = 1$}
+	\begin{footnotesize}
+		\hskip -1.7cm
+	\begin{tabular}{l|r|r|r|r|r|r|r|r}
+		\hline
+		Data Set & Auto Est & Auto SE & AutoM Est & AutoM SE & Norm Est & Norm SE & NormM Est & NormM SE\\
+		\hline
+		$S_0 = 99999, I_0 = 1$ & 1.185172 & 1.285564 & 1.158778 & 0.152094 & 0.859907 & 83.111543 & 1.116840 & 0.064265\\
+		\hline
+		$S_0 = 99990, I_0 = 10$ & 0.934239 & 3.360057 & 0.944893 & 0.034982 & 1.101154 & 0.287314 & 6.153800 & 1525.693988\\
+		\hline
+		$S_0 = 99900, I_0 = 100$ & 1.255903 & 0.086428 & 0.462015 & 0.325984 & 1.250009 & 0.051102 & 1.193192 & 0.035157\\
+		\hline
+		$S_0 = 99000, I_0 = 1000$ & 1.205152 & 0.005332 & 1.193656 & 0.005244 & 1.199742 & 0.004094 & 1.208671 & 0.004804\\
 		\hline
@@
 		Different Variances, 
-		$X_0 = 99000, Y_0 = 1000$}
+		$S_0 = 99000, I_0 = 1000$}
 	\begin{footnotesize}
@@
 		\hline
-			$\sigma_X = 10, \sigma_Y = 1$ & -121.825007 & 904.236167 & 121.668347 & 308.632857 & 21.981032 & 21.204959 & 18.741699 & 17.132467\\
-			\hline
-			$\sigma_X = 100, \sigma_Y = 10$ & 2.918979 & 2.010263 & -1.223521 & 2.355038 & -5.683542 & 10.750736 & 4.071392 & 2.877332\\
-			\hline
-			$\sigma_X = 300, \sigma_Y = 30$ & 2.769860 & 1.759591 & 2.002654 & 0.949586 & 3.227956 & 3.934407 & 87.874782 & 456.259511\\
-			\hline
-			$\sigma_X = 1000, \sigma_Y = 100$ & 0.915434 & 0.114881 & 1.887797 & 0.672640 & 1.759919 & 1.537275 & -0.235868 & 1.519149\\
-			\hline
-			$\sigma_X = 3000, \sigma_Y = 300$ & -4.547791 & 970.222266 & 1.037930 & 0.112483 & 0.748209 & 0.437849 & 0.235119 & 1.489429\\
-			\hline
-			$\sigma_X = 10000, \sigma_Y = 1000$ & 1.033786 & 0.068722 & 0.958868 & 0.066758 & 0.862043 & 0.168471 & 0.671640 & 0.206497\\
+		$\sigma_S = 10, \sigma_I = 1$ & 1.185958 & 0.008294 & 1.202635 & 0.006364 & 1.200712 & 0.004175 & 1.194196 & 0.003282\\
+		\hline
+		$\sigma_S = 100, \sigma_I = 10$ & 1.080974 & 0.051432 & 1.242575 & 0.081110 & 1.221708 & 0.047705 & 1.160408 & 0.025033\\
+		\hline
+		$\sigma_S = 500, \sigma_I = 50$ & 0.738334 & 0.237194 & 1.506353 & 7.578001 & 1.343488 & 0.552568 & 1.071472 & 0.060095\\
+		\hline
+		$\sigma_S = 1000, \sigma_I = 100$ & 0.474710 & 0.832353 & 2.427907 & 19.349048 & 1.619427 & 3.245662 & 1.018452 & 0.087734\\
+		\hline
+		$\sigma_S = 2000, \sigma_I = 200$ & 0.184986 & 2.306650 & -1.221597 & 233.262939 & 9.803881 & 71195.968128 & 0.969270 & 0.132646\\
+		\hline
+		$\sigma_S = 5000, \sigma_I = 500$ & -0.130854 & 10.533729 & 0.351591 & 7.021734 & 0.389501 & 23.012536 & 0.916922 & 0.188977\\
+		\hline
+		$\sigma_S = 7500, \sigma_I = 750$ & -0.225912 & 37.943501 & 0.472891 & 14.985760 & 0.546636 & 21.663742 & 0.895990 & 0.203207\\
+		\hline
+		$\sigma_S = 10000, \sigma_I = 1000$ & -0.276013 & 174.714931 & 0.530827 & 128.425370 & 0.603731 & 28.923054 & 0.880411 & 0.207996\\
+		\hline
+		$\sigma_S = 15000, \sigma_I = 1500$ & -0.324210 & 28.295186 & 0.598720 & 14.309696 & 0.650612 & 8.469867 & 0.855473 & 0.206941\\
 		\hline
@@
 	\caption{$\rr$ Estimates and Std. Errs, SIRLMA Model,
-		Other Models, $X_0 = 99950, Y_0 = 50$, 
-		$\sigma_X = 10, \sigma_Y = 1$}
+		Other Models, $S_0 = 99950, I_0 = 50$, 
+		$\sigma_S = 10, \sigma_I = 1$}
 	\begin{footnotesize}
@@
 
-To calculate the estimates for $\rr$, we use lm() in R to fit the linear polynomial regression functions for X(t) and Y(t) with respect to t.
+To calculate the estimates for $\rr$, we use lm() in R to fit the linear polynomial regression functions for S(t) and I(t) with respect to t.
 We then calculated the derivatives of the linear regression functions at each time using our own function in order to find the average linear model $\rr$ estimate across all time steps $t$.
@@
 In particular, it does a lot better for the baseline data and the data from other models.
-The estimates are still really bad for large values of $\sigma_X$ and $\sigma_Y$ (including some negative estimates).
-Interestingly, the linear model average estimates here are less accurate for large values of $Y_0$, whereas they were more accurate for the linear model.
+The estimates are still really bad for large values of $\sigma_S$ and $\sigma_I$ (including some negative estimates).
+Interestingly, the linear model average estimates here are less accurate for large values of $I_0$, whereas they were more accurate for the linear model.
 
 The standard errors are not as ridiculously large compared to the linear model with only one time point, but are still extremely big compared to what we might expect to see from other models.
-This is especially apparent when $\sigma_X$ and $\sigma_Y$ are large.
+This is especially apparent when $\sigma_S$ and $\sigma_I$ are large.
 
@@
 	\caption{$\rr$ Estimates and Std. Errs, SIRLMAT Model,
-		Baseline Data, $X_0 = 99950, Y_0 = 50$, 
-		$\sigma_X = 10, \sigma_Y = 1$}
-	\begin{footnotesize}
-		\hskip -1cm
-		\begin{tabular}{l|r|r|r|r|r|r|r|r}
-			\hline
-			Data Set & Auto Est & Auto SE & AutoM Est & AutoM SE & Norm Est & Norm SE & NormM Est & NormM SE\\
-			\hline
-			Baseline 1 & 2.015772 & 27.518508 & 2.150988 & 4.474655 & 1.988515 & 0.232390 & 2.103188 & 0.636825\\
-			\hline
-			Baseline 2 & 2.571202 & 21.048981 & -0.131021 & 51.293391 & 2.192838 & 1.684886 & 2.189259 & 1.514415\\
-			\hline
-			Baseline 3 & 1.201407 & 21.526024 & 3.023289 & 26.322188 & 2.079618 & 1.027536 & 2.007048 & 0.269511\\
-			\hline
-			Baseline 4 & 1.991370 & 1.164132 & 3.604925 & 35.619793 & 2.028025 & 0.205708 & 1.875085 & 2.080861\\
-			\hline
-			Baseline 5 & 1.873964 & 7.146565 & 1.032642 & 11.871804 & 1.372848 & 10.164198 & 1.926002 & 0.166963\\
+		Baseline Data, $S_0 = 99950, I_0 = 50$, 
+		$\sigma_S = 10, \sigma_I = 1$}
+	\begin{footnotesize}
+		\hskip -1cm
+	\begin{tabular}{l|r|r|r|r|r|r|r|r}
+		\hline
+		Data Set & Auto Est & Auto SE & AutoM Est & AutoM SE & Norm Est & Norm SE & NormM Est & NormM SE\\
+		\hline
+		Baseline 1 & 1.171311 & 0.029553 & 1.169409 & 0.077379 & 1.171164 & 0.028320 & 1.173413 & 0.034820\\
+		\hline
+		Baseline 2 & 1.170936 & 0.027339 & 1.172363 & 0.032062 & 1.171630 & 0.027987 & 1.173249 & 0.032884\\
+		\hline
+		Baseline 3 & 1.023937 & 2.701492 & 1.168652 & 0.029267 & 1.171384 & 0.028419 & 1.171662 & 0.028700\\
+		\hline
+		Baseline 4 & 1.171501 & 0.028128 & 1.174232 & 0.033153 & 1.173115 & 0.031369 & 1.171094 & 0.028392\\
+		\hline
+		Baseline 5 & 1.178276 & 0.303469 & 1.167810 & 0.027766 & 1.170475 & 0.027259 & 1.171001 & 0.027792\\
 		\hline
@@
 		Different Starting Populations, 
-		$\sigma_X = 10, \sigma_Y = 1$}
-	\begin{footnotesize}
-		\hskip -1.7cm
-		\begin{tabular}{l|r|r|r|r|r|r|r|r}
-			Data Set & Auto Est & Auto SE & AutoM Est & AutoM SE & Norm Est & Norm SE & NormM Est & NormM SE\\
-			\hline
-			$X_0 = 99999, Y_0 = 1$ & 2.558602 & 32.007328 & 1.296284 & 9.652181 & 1.178839 & 3.901942 & 1.287102 & 0.524929\\
-			\hline
-			$X_0 = 99990, Y_0 = 10$ & 0.745127 & 3.487290 & 1.541772 & 1.793288 & 3.227689 & 26.102558 & 2.489264 & 26.755104\\
-			\hline
-			$X_0 = 99900, Y_0 = 100$ & 2.082023 & 2.181500 & -1.745769 & 81.529981 & 1.946617 & 1.232307 & 2.008181 & 0.191259\\
-			\hline
-			$X_0 = 99000, Y_0 = 1000$ & 1.969496 & 0.088997 & 1.979588 & 0.093847 & 1.972052 & 0.036176 & 1.964677 & 0.028405\\
+		$\sigma_S = 10, \sigma_I = 1$}
+	\begin{footnotesize}
+		\hskip -1.7cm
+	\begin{tabular}{l|r|r|r|r|r|r|r|r}
+		\hline
+		Data Set & Auto Est & Auto SE & AutoM Est & AutoM SE & Norm Est & Norm SE & NormM Est & NormM SE\\
+		\hline
+		$S_0 = 99999, I_0 = 1$ & 1.422018 & 8.968638 & 1.188044 & 0.101487 & 1.151807 & 0.275409 & 1.195600 & 0.033124\\
+		\hline
+		$S_0 = 99990, I_0 = 10$ & 1.191840 & 0.033509 & 1.178169 & 0.160530 & 1.194324 & 0.017020 & 1.209180 & 0.263013\\
+		\hline
+		$S_0 = 99900, I_0 = 100$ & 1.148868 & 0.047200 & 1.144860 & 0.232458 & 1.149893 & 0.047661 & 1.149490 & 0.046846\\
+		\hline
+		$S_0 = 99000, I_0 = 1000$ & 0.995113 & 0.120333 & 0.994869 & 0.120456 & 0.994867 & 0.120621 & 0.995126 & 0.120503\\
 		\hline
@@
 		Different Variances, 
-		$X_0 = 99000, Y_0 = 1000$}
-	\begin{footnotesize}
-		\hskip -1.7cm
-		\begin{tabular}{l|r|r|r|r|r|r|r|r}
-			\hline
-			Data Set & Auto Est & Auto SE & AutoM Est & AutoM SE & Norm Est & Norm SE & NormM Est & NormM SE\\
-			\hline
-			$\sigma_X = 10, \sigma_Y = 1$ & 5.556294 & 63.868004 & 2.843995 & 7.555433 & 2.277135 & 2.022040 & 2.234907 & 1.759312\\
-			\hline
-			$\sigma_X = 100, \sigma_Y = 10$ & 2.020666 & 0.483996 & 1.802373 & 5.248770 & 1.891754 & 5.445302 & 2.034306 & 0.527914\\
-			\hline
-			$\sigma_X = 300, \sigma_Y = 30$ & -1.064869 & 50.543183 & 4.685491 & 53.111296 & 2.076773 & 0.646978 & 2.521401 & 5.276258\\
-			\hline
-			$\sigma_X = 1000, \sigma_Y = 100$ & 0.999874 & 3.992411 & 2.390182 & 3.320493 & 1.456545 & 6.796468 & 1.703567 & 2.395295\\
-			\hline
-			$\sigma_X = 3000, \sigma_Y = 300$ & -2.740783 & 70.416107 & 1.371659 & 1.347889 & 0.030795 & 21.459099 & 2.023080 & 6.341932\\
-			\hline
-			$\sigma_X = 10000, \sigma_Y = 1000$ & 0.877306 & 0.815504 & 1.106179 & 2.530188 & 1.183491 & 3.063114 & 0.949453 & 6.173692\\
+		$S_0 = 99000, I_0 = 1000$}
+	\begin{footnotesize}
+		\hskip -1.7cm
+	\begin{tabular}{l|r|r|r|r|r|r|r|r}
+		\hline
+		Data Set & Auto Est & Auto SE & AutoM Est & AutoM SE & Norm Est & Norm SE & NormM Est & NormM SE\\
+		\hline
+		$\sigma_S = 10, \sigma_I = 1$ & 0.994680 & 0.119919 & 0.995029 & 0.120570 & 0.994850 & 0.120592 & 0.994886 & 0.120268\\
+		\hline
+		$\sigma_S = 100, \sigma_I = 10$ & 0.993618 & 0.116092 & 0.996562 & 0.121634 & 0.994779 & 0.121739 & 0.994844 & 0.119207\\
+		\hline
+		$\sigma_S = 500, \sigma_I = 50$ & 1.018998 & 0.741279 & 1.003635 & 0.132436 & 0.994571 & 0.130529 & 0.992271 & 0.121888\\
+		\hline
+		$\sigma_S = 1000, \sigma_I = 100$ & 0.976732 & 0.194268 & 1.015891 & 0.176727 & 0.994170 & 0.159640 & 0.986805 & 0.133885\\
+		\hline
+		$\sigma_S = 2000, \sigma_I = 200$ & 0.584095 & 6.159187 & 1.131111 & 2.595232 & 1.118248 & 1.135767 & 0.954122 & 0.243263\\
+		\hline
+		$\sigma_S = 5000, \sigma_I = 500$ & 2.054850 & 17.230026 & 0.731612 & 3.124131 & 1.018859 & 0.714739 & 1.087380 & 0.674374\\
+		\hline
+		$\sigma_S = 7500, \sigma_I = 750$ & 1.326042 & 5.616124 & 0.741682 & 6.293590 & 1.061083 & 1.608632 & 0.914812 & 2.034648\\
+		\hline
+		$\sigma_S = 10000, \sigma_I = 1000$ & -0.424508 & 24.406119 & 0.728716 & 7.478691 & 0.984529 & 0.261810 & 0.677480 & 4.972663\\
+		\hline
+		$\sigma_S = 15000, \sigma_I = 1500$ & -0.352152 & 22.171481 & 0.774077 & 3.857982 & 0.984145 & 0.158089 & 0.980721 & 1.294957\\
 		\hline
@@
 	\caption{$\rr$ Estimates and Std. Errs, SIRLMAT Model,
-		Other Models, $X_0 = 99950, Y_0 = 50$, 
-		$\sigma_X = 10, \sigma_Y = 1$}
+		Other Models, $S_0 = 99950, I_0 = 50$, 
+		$\sigma_S = 10, \sigma_I = 1$}
 	\begin{footnotesize}
@@
 Most of the estimates are fairly close to the true $\rr$ value.
-It appears to give more accurate $\rr$ estimates better for higher values of $Y_0$ relative to the total population as well as smaller values of $\sigma_X$ and $\sigma_Y$.
+It appears to give more accurate $\rr$ estimates better for higher values of $I_0$ relative to the total population as well as smaller values of $\sigma_S$ and $\sigma_I$.
 This model also consistently gives reasonable results, unlike for other models, which sometimes fail miserably when dealing with larger errors or data generated from other models (especially for 0th and 1st order data).
@@
 	\caption{$\rr$ Estimates and Std. Errs, SIRMAX Model,
-		Baseline Data, $X_0 = 99950, Y_0 = 50$, 
-		$\sigma_X = 10, \sigma_Y = 1$}
-	\begin{footnotesize}
-		\hskip -1cm
-		\begin{tabular}{l|r|r|r|r|r|r|r|r}
-			\hline
-			Data Set & Auto Est & Auto SE & AutoM Est & AutoM SE & Norm Est & Norm SE & NormM Est & NormM SE\\
-			\hline
-			Baseline 1 & 1.002232 & NA & 1.002359 & NA & 1.002498 & NA & 1.002531 & NA\\
-			\hline
-			Baseline 2 & 1.002405 & NA & 1.002802 & NA & 1.002364 & NA & 1.002561 & NA\\
-			\hline
-			Baseline 3 & 1.002261 & NA & 1.002393 & NA & 1.002516 & NA & 1.002501 & NA\\
-			\hline
-			Baseline 4 & 1.002469 & NA & 1.002494 & NA & 1.002511 & NA & 1.002481 & NA\\
-			\hline
-			Baseline 5 & 1.002392 & NA & 1.002496 & NA & 1.002509 & NA & 1.002529 & NA\\
+		Baseline Data, $S_0 = 99950, I_0 = 50$, 
+		$\sigma_S = 10, \sigma_I = 1$}
+	\begin{footnotesize}
+		\hskip -1cm
+	\begin{tabular}{l|r|r|r|r|r|r|r|r}
+		\hline
+		Data Set & Auto Est & Auto SE & AutoM Est & AutoM SE & Norm Est & Norm SE & NormM Est & NormM SE\\
+		\hline
+		Baseline 1 & 1.091592 & NA & 1.091985 & NA & 1.091542 & NA & 1.091488 & NA\\
+		\hline
+		Baseline 2 & 1.091390 & NA & 1.091563 & NA & 1.091537 & NA & 1.091717 & NA\\
+		\hline
+		Baseline 3 & 1.091540 & NA & 1.091211 & NA & 1.091819 & NA & 1.091724 & NA\\
+		\hline
+		Baseline 4 & 1.091405 & NA & 1.091566 & NA & 1.091414 & NA & 1.091801 & NA\\
+		\hline
+		Baseline 5 & 1.091637 & NA & 1.091549 & NA & 1.091627 & NA & 1.091556 & NA\\
 		\hline
@@
 		Different Starting Populations, 
-		$\sigma_X = 10, \sigma_Y = 1$}
-	\begin{footnotesize}
-		\hskip -1.7cm
-		\begin{tabular}{l|r|r|r|r|r|r|r|r}
-			\hline
-			Data Set & Auto Est & Auto SE & AutoM Est & AutoM SE & Norm Est & Norm SE & NormM Est & NormM SE\\
-			\hline
-			$X_0 = 99999, Y_0 = 1$ & 0.999831 & NA & 1.000028 & NA & 0.999962 & NA & 1.000079 & NA\\
-			\hline
-			$X_0 = 99990, Y_0 = 10$ & 1.000442 & NA & 1.000473 & NA & 1.000514 & NA & 1.000527 & NA\\
-			\hline
-			$X_0 = 99900, Y_0 = 100$ & 1.005071 & NA & 1.004938 & NA & 1.005222 & NA & 1.004948 & NA\\
-			\hline
-			$X_0 = 99000, Y_0 = 1000$ & 1.049518 & NA & 1.050139 & NA & 1.050143 & NA & 1.050157 & NA\\
+		$\sigma_S = 10, \sigma_I = 1$}
+	\begin{footnotesize}
+		\hskip -1.7cm
+	\begin{tabular}{l|r|r|r|r|r|r|r|r}
+		\hline
+		Data Set & Auto Est & Auto SE & AutoM Est & AutoM SE & Norm Est & Norm SE & NormM Est & NormM SE\\
+		\hline
+		$S_0 = 99999, I_0 = 1$ & 1.002195 & NA & 1.002221 & NA & 1.002225 & NA & 1.002435 & NA\\
+		\hline
+		$S_0 = 99990, I_0 = 10$ & 1.021510 & NA & 1.021948 & NA & 1.021505 & NA & 1.021804 & NA\\
+		\hline
+		$S_0 = 99900, I_0 = 100$ & 1.154799 & NA & 1.155005 & NA & 1.155144 & NA & 1.155213 & NA\\
+		\hline
+		$S_0 = 99000, I_0 = 1000$ & 1.202948 & NA & 1.201580 & NA & 1.199981 & NA & 1.196597 & NA\\
 		\hline
@@
 		Different Variances, 
-		$X_0 = 99000, Y_0 = 1000$}
-	\begin{footnotesize}
-		\hskip -1.7cm
-		\begin{tabular}{l|r|r|r|r|r|r|r|r}
-			\hline
-			Data Set & Auto Est & Auto SE & AutoM Est & AutoM SE & Norm Est & Norm SE & NormM Est & NormM SE\\
-			\hline
-			$\sigma_X = 10, \sigma_Y = 1$ & 1.050097 & NA & 1.049897 & NA & 1.050034 & NA & 1.050098 & NA\\
-			\hline
-			$\sigma_X = 100, \sigma_Y = 10$ & 1.048816 & NA & 1.051191 & NA & 1.051560 & NA & 1.051580 & NA\\
-			\hline
-			$\sigma_X = 300, \sigma_Y = 30$ & 1.045130 & NA & 1.054815 & NA & 1.042187 & NA & 1.049682 & NA\\
-			\hline
-			$\sigma_X = 1000, \sigma_Y = 100$ & 1.038396 & NA & 1.051101 & NA & 1.057586 & NA & 1.049022 & NA\\
-			\hline
-			$\sigma_X = 3000, \sigma_Y = 300$ & 1.005515 & NA & 1.081383 & NA & 1.002521 & NA & 1.064796 & NA\\
-			\hline
-			$\sigma_X = 10000, \sigma_Y = 1000$ & 1.076870 & NA & 1.114340 & NA & 1.077148 & NA & 1.037306 & NA\\
+		$S_0 = 99000, I_0 = 1000$}
+	\begin{footnotesize}
+		\hskip -1.7cm
+	\begin{tabular}{l|r|r|r|r|r|r|r|r}
+		\hline
+		Data Set & Auto Est & Auto SE & AutoM Est & AutoM SE & Norm Est & Norm SE & NormM Est & NormM SE\\
+		\hline
+		$\sigma_S = 10, \sigma_I = 1$ & 1.200091 & NA & 1.199638 & NA & 1.203322 & NA & 1.198214 & NA\\
+		\hline
+		$\sigma_S = 100, \sigma_I = 10$ & 1.192908 & NA & 1.206314 & NA & 1.204461 & NA & 1.189191 & NA\\
+		\hline
+		$\sigma_S = 500, \sigma_I = 50$ & 1.164184 & NA & 1.198797 & NA & 1.209550 & NA & 1.220847 & NA\\
+		\hline
+		$\sigma_S = 1000, \sigma_I = 100$ & 1.177169 & NA & 1.226796 & NA & 1.215972 & NA & 1.223427 & NA\\
+		\hline
+		$\sigma_S = 2000, \sigma_I = 200$ & 1.204027 & NA & 1.212075 & NA & 1.175607 & NA & 1.228619 & NA\\
+		\hline
+		$\sigma_S = 5000, \sigma_I = 500$ & 1.292498 & NA & 1.169958 & NA & 1.208613 & NA & 1.348023 & NA\\
+		\hline
+		$\sigma_S = 7500, \sigma_I = 750$ & 1.376802 & NA & 1.137033 & NA & 1.237567 & NA & 1.359960 & NA\\
+		\hline
+		$\sigma_S = 10000, \sigma_I = 1000$ & 1.288801 & NA & 1.105911 & NA & 1.075749 & NA & 1.372110 & NA\\
+		\hline
+		$\sigma_S = 15000, \sigma_I = 1500$ & 1.228367 & NA & 1.158778 & NA & 0.978885 & NA & 1.397074 & NA\\
 		\hline
@@
 	\caption{$\rr$ Estimates and Std. Errs, SIRMAX Model,
-		Other Models, $X_0 = 99950, Y_0 = 50$, 
-		$\sigma_X = 10, \sigma_Y = 1$}
+		Other Models, $S_0 = 99950, I_0 = 50$, 
+		$\sigma_S = 10, \sigma_I = 1$}
 	\begin{footnotesize}
@@
 
-We use the smooth.spline() function in R to get estimates for X(t) and Y(t), then used the which.max() function to find the point in time where the number of people in the infected compartment is at a maximum.
+We use the smooth.spline() function in R to get estimates for S(t) and I(t), then used the which.max() function to find the point in time where the number of people in the infected compartment is at a maximum.
 The standard error for the $\rr$ estimates was done using the jackknife method; we estimate a new spline function after removing one observation from the data set and calculate a new $\rr$ estimate, then use the standard deviation of those estimates as the estimate of the standard error.
@@
 It does not give us any crazy results.
-The model is more accurate for larger values of $Y_0$ relative to the total size of the population and for smaller $\sigma_X$ and $\sigma_Y$.
+The model is more accurate for larger values of $I_0$ relative to the total size of the population and for smaller $\sigma_S$ and $\sigma_I$.
 
 The standard errors are reasonable values, and they seem to be fairly consistent under repeated estimation with the same conditions for the generated data.
-They increase with increases in $Y_0$ as well as $\sigma_X$ and $\sigma_Y$.
+They increase with increases in $I_0$ as well as $\sigma_S$ and $\sigma_I$.
 
@@
 	\caption{$\rr$ Estimates and Std. Errs, SIRSMAX Model,
-		Baseline Data, $X_0 = 99950, Y_0 = 50$, 
-		$\sigma_X = 10, \sigma_Y = 1$}
-	\begin{footnotesize}
-		\hskip -1cm
-		\begin{tabular}{l|r|r|r|r|r|r|r|r}
-			\hline
-			Data Set & Auto Est & Auto SE & AutoM Est & AutoM SE & Norm Est & Norm SE & NormM Est & NormM SE\\
-			\hline
-			Baseline 1 & 1.002424 & 2.3e-05 & 1.002460 & 1.1e-05 & 1.002435 & 1.5e-05 & 1.002459 & 8e-06\\
-			\hline
-			Baseline 2 & 1.002507 & 2.2e-05 & 1.002444 & 1.8e-05 & 1.002462 & 1.9e-05 & 1.002436 & 7e-06\\
-			\hline
-			Baseline 3 & 1.002374 & 1.9e-05 & 1.002418 & 5.0e-06 & 1.002439 & 1.7e-05 & 1.002445 & 5e-06\\
-			\hline
-			Baseline 4 & 1.002434 & 1.5e-05 & 1.002479 & 8.0e-06 & 1.002461 & 1.5e-05 & 1.002449 & 6e-06\\
-			\hline
-			Baseline 5 & 1.002403 & 2.0e-05 & 1.002469 & 9.0e-06 & 1.002439 & 1.6e-05 & 1.002468 & 6e-06\\
+		Baseline Data, $S_0 = 99950, I_0 = 50$, 
+		$\sigma_S = 10, \sigma_I = 1$}
+	\begin{footnotesize}
+		\hskip -1cm
+	\begin{tabular}{l|r|r|r|r|r|r|r|r}
+		\hline
+		Data Set & Auto Est & Auto SE & AutoM Est & AutoM SE & Norm Est & Norm SE & NormM Est & NormM SE\\
+		\hline
+		Baseline 1 & 1.085556 & 0.000544 & 1.085589 & 0.000546 & 1.085551 & 0.000538 & 1.085568 & 0.000534\\
+		\hline
+		Baseline 2 & 1.085512 & 0.000531 & 1.085507 & 0.000536 & 1.085553 & 0.000534 & 1.085583 & 0.000536\\
+		\hline
+		Baseline 3 & 1.085561 & 0.000542 & 1.085528 & 0.000538 & 1.085557 & 0.000539 & 1.085600 & 0.000538\\
+		\hline
+		Baseline 4 & 1.085540 & 0.000527 & 1.085527 & 0.000537 & 1.085556 & 0.000536 & 1.085547 & 0.000538\\
+		\hline
+		Baseline 5 & 1.085546 & 0.000531 & 1.085561 & 0.000531 & 1.085545 & 0.000535 & 1.085510 & 0.000535\\
 		\hline
@@
 		Different Starting Populations, 
-		$\sigma_X = 10, \sigma_Y = 1$}
-	\begin{footnotesize}
-		\hskip -1.7cm
-		\begin{tabular}{l|r|r|r|r|r|r|r|r}
-			\hline
-			Data Set & Auto Est & Auto SE & AutoM Est & AutoM SE & Norm Est & Norm SE & NormM Est & NormM SE\\
-			\hline
-			$X_0 = 99999, Y_0 = 1$ & 1.000116 & 1.7e-05 & 1.000063 & 2.0e-06 & 1.000040 & 1.5e-05 & 1.000064 & 1.0e-06\\
-			\hline
-			$X_0 = 99990, Y_0 = 10$ & 1.000474 & 2.1e-05 & 1.000477 & 8.0e-06 & 1.000473 & 1.6e-05 & 1.000507 & 4.0e-06\\
-			\hline
-			$X_0 = 99900, Y_0 = 100$ & 1.004902 & 1.8e-05 & 1.004880 & 1.4e-05 & 1.004899 & 1.9e-05 & 1.004895 & 8.0e-06\\
-			\hline
-			$X_0 = 99000, Y_0 = 1000$ & 1.049155 & 7.4e-05 & 1.049349 & 6.6e-05 & 1.049271 & 6.9e-05 & 1.049277 & 6.6e-05\\
+		$\sigma_S = 10, \sigma_I = 1$}
+	\begin{footnotesize}
+		\hskip -1.7cm
+	\begin{tabular}{l|r|r|r|r|r|r|r|r}
+		\hline
+		Data Set & Auto Est & Auto SE & AutoM Est & AutoM SE & Norm Est & Norm SE & NormM Est & NormM SE\\
+		\hline
+		$S_0 = 99999, I_0 = 1$ & 1.001996 & 0.000024 & 1.002034 & 0.000029 & 1.002019 & 0.000026 & 1.002030 & 0.000021\\
+		\hline
+		$S_0 = 99990, I_0 = 10$ & 1.019557 & 0.000186 & 1.019528 & 0.000183 & 1.019498 & 0.000179 & 1.019483 & 0.000175\\
+		\hline
+		$S_0 = 99900, I_0 = 100$ & 1.148891 & 0.000569 & 1.148956 & 0.000563 & 1.148938 & 0.000580 & 1.148941 & 0.000581\\
+		\hline
+		$S_0 = 99000, I_0 = 1000$ & 1.201700 & 0.000907 & 1.201844 & 0.000881 & 1.201772 & 0.000939 & 1.201768 & 0.000898\\
 		\hline
@@
 		Different Variances, 
-		$X_0 = 99000, Y_0 = 1000$}
-	\begin{footnotesize}
-		\hskip -1.7cm
-		\begin{tabular}{l|r|r|r|r|r|r|r|r}
-			\hline
-			Data Set & Auto Est & Auto SE & AutoM Est & AutoM SE & Norm Est & Norm SE & NormM Est & NormM SE\\
-			\hline
-			$\sigma_X = 10, \sigma_Y = 1$ & 1.049351 & 0.000076 & 1.049219 & 0.000065 & 1.049294 & 0.000069 & 1.049283 & 0.000069\\
-			\hline
-			$\sigma_X = 100, \sigma_Y = 10$ & 1.049675 & 0.000154 & 1.049504 & 0.000132 & 1.049443 & 0.000182 & 1.049389 & 0.000131\\
-			\hline
-			$\sigma_X = 300, \sigma_Y = 30$ & 1.046190 & 0.000566 & 1.051219 & 0.000676 & 1.050137 & 0.000623 & 1.049671 & 0.000141\\
-			\hline
-			$\sigma_X = 1000, \sigma_Y = 100$ & 1.061907 & 0.002419 & 1.048793 & 0.000363 & 1.046010 & 0.001880 & 1.048815 & 0.000172\\
-			\hline
-			$\sigma_X = 3000, \sigma_Y = 300$ & 1.045117 & 0.005651 & 1.068145 & 0.004304 & 1.039168 & 0.005444 & 1.057089 & 0.000372\\
-			\hline
-			$\sigma_X = 10000, \sigma_Y = 1000$ & 0.979174 & 0.014889 & 1.111583 & 0.002687 & 1.039000 & 0.015012 & 1.070210 & 0.004720\\
+		$S_0 = 99000, I_0 = 1000$}
+	\begin{footnotesize}
+		\hskip -1.7cm
+	\begin{tabular}{l|r|r|r|r|r|r|r|r}
+		\hline
+		Data Set & Auto Est & Auto SE & AutoM Est & AutoM SE & Norm Est & Norm SE & NormM Est & NormM SE\\
+		\hline
+		$\sigma_S = 10, \sigma_I = 1$ & 1.201804 & 0.000884 & 1.201778 & 0.000936 & 1.201741 & 0.000950 & 1.201767 & 0.000849\\
+		\hline
+		$\sigma_S = 100, \sigma_I = 10$ & 1.202440 & 0.001007 & 1.202091 & 0.000898 & 1.201573 & 0.000962 & 1.202069 & 0.000931\\
+		\hline
+		$\sigma_S = 500, \sigma_I = 50$ & 1.205047 & 0.001527 & 1.203529 & 0.001342 & 1.200705 & 0.001254 & 1.203260 & 0.001214\\
+		\hline
+		$\sigma_S = 1000, \sigma_I = 100$ & 1.208288 & 0.002605 & 1.205228 & 0.002166 & 1.199591 & 0.001887 & 1.204841 & 0.001929\\
+		\hline
+		$\sigma_S = 2000, \sigma_I = 200$ & 1.214838 & 0.004880 & 1.208625 & 0.004127 & 1.197394 & 0.003492 & 1.207986 & 0.003637\\
+		\hline
+		$\sigma_S = 5000, \sigma_I = 500$ & 1.234608 & 0.011933 & 1.218729 & 0.009843 & 1.190987 & 0.008460 & 1.218687 & 0.009868\\
+		\hline
+		$\sigma_S = 7500, \sigma_I = 750$ & 1.251420 & 0.017892 & 1.226979 & 0.014483 & 1.185867 & 0.012567 & 1.229087 & 0.016029\\
+		\hline
+		$\sigma_S = 10000, \sigma_I = 1000$ & 1.268381 & 0.023922 & 1.235083 & 0.019214 & 1.180831 & 0.016460 & 1.241403 & 0.023960\\
+		\hline
+		$\sigma_S = 15000, \sigma_I = 1500$ & 1.303020 & 0.035872 & 1.251167 & 0.028997 & 1.171122 & 0.024414 & 1.273194 & 0.037158\\
 		\hline
@@
 	\caption{$\rr$ Estimates and Std. Errs, SIRSMAX Model,
-		Other Models, $X_0 = 99950, Y_0 = 50$, 
-		$\sigma_X = 10, \sigma_Y = 1$}
+		Other Models, $S_0 = 99950, I_0 = 50$, 
+		$\sigma_S = 10, \sigma_I = 1$}
 	\begin{footnotesize}
@@
 Looking at the baseline data, the $\rr$ estimates vary fairly widely given that the 5 data sets were generated under the same conditions.
-We also see that the estimates change massively as we change $Y_0$ or $\sigma_X$ and $\sigma_Y$.
-As we see below, the estimate for $\rr$ drops from around 45 for $Y_0 = 1$ to 0.9 for $Y_0 = 1000$.
-Similarly, the estimate for $\rr$ grows from around 0.9 for ($\sigma_X$, $\sigma_Y$) = (10, 1) to 200 for ($\sigma_X$, $\sigma_Y$) = (15000, 1500).
-In general, we see that the estimates for $\rr$ decrease as $Y_0$ increases relative to the total population and that they increase as we increase
-$\sigma_X$ and $\sigma_Y$.
+We also see that the estimates change massively as we change $I_0$ or $\sigma_S$ and $\sigma_I$.
+As we see below, the estimate for $\rr$ drops from around 45 for $I_0 = 1$ to 0.9 for $I_0 = 1000$.
+Similarly, the estimate for $\rr$ grows from around 0.9 for ($\sigma_S$, $\sigma_I$) = (10, 1) to 200 for ($\sigma_S$, $\sigma_I$) = (15000, 1500).
+In general, we see that the estimates for $\rr$ decrease as $I_0$ increases relative to the total population and that they increase as we increase
+$\sigma_S$ and $\sigma_I$.
 
@@
 	\caption{$\rr$ Estimates and Std. Errs, SIRIPR Model,
-		Baseline Data, $X_0 = 99950, Y_0 = 50$, 
-		$\sigma_X = 10, \sigma_Y = 1$}
-	\begin{footnotesize}
-		\hskip -1cm
-		\begin{tabular}{l|r|r|r|r|r|r|r|r}
-			\hline
-			Data Set & Auto Est & Auto SE & AutoM Est & AutoM SE & Norm Est & Norm SE & NormM Est & NormM SE\\
-			\hline
-			Data Set & Auto Est & Auto SE & AutoM Est & AutoM SE & Norm Est & Norm SE & NormM Est & NormM SE\\
-			\hline
-			Baseline 1 & 1.170709 & 1.726236 & 0.118299 & 0.501111 & 1.566640 & 2.461139 & 0.129870 & 0.410478\\
-			\hline
-			Baseline 2 & 1.169642 & 1.736295 & 0.134667 & 0.531973 & 1.405544 & 1.947442 & 0.143524 & 0.470087\\
-			\hline
-			Baseline 3 & 1.024440 & 1.566376 & 0.147574 & 0.616955 & 1.327697 & 2.039847 & 0.135604 & 0.415505\\
-			\hline
-			Baseline 4 & 1.078543 & 1.635074 & 0.126040 & 0.518177 & 1.538235 & 2.350329 & 0.136590 & 0.461818\\
-			\hline
-			Baseline 5 & 1.082233 & 1.600320 & 0.130253 & 0.464518 & 1.287763 & 2.121584 & 0.124392 & 0.379142\\
+		Baseline Data, $S_0 = 99950, I_0 = 50$, 
+		$\sigma_S = 10, \sigma_I = 1$}
+	\begin{footnotesize}
+		\hskip -1cm
+	\begin{tabular}{l|r|r|r|r|r|r|r|r}
+		\hline
+		Data Set & Auto Est & Auto SE & AutoM Est & AutoM SE & Norm Est & Norm SE & NormM Est & NormM SE\\
+		\hline
+		Baseline 1 & 1.568114 & NA & 1.171336 & NA & 1.332469 & NA & 1.062616 & NA\\
+		\hline
+		Baseline 2 & 1.273631 & NA & 1.287347 & NA & 1.287487 & NA & 1.324252 & NA\\
+		\hline
+		Baseline 3 & 1.302008 & NA & 1.352183 & NA & 1.528039 & NA & 1.059422 & NA\\
+		\hline
+		Baseline 4 & 1.181609 & NA & 1.001200 & NA & 1.624614 & NA & 1.128520 & NA\\
+		\hline
+		Baseline 5 & 1.168529 & NA & 1.278128 & NA & 1.289007 & NA & 1.207690 & NA\\
 		\hline
@@
 		Different Starting Populations, 
-		$\sigma_X = 10, \sigma_Y = 1$}
-	\begin{footnotesize}
-		\hskip -1.7cm
-		\begin{tabular}{l|r|r|r|r|r|r|r|r}
-			\hline
-			Data Set & Auto Est & Auto SE & AutoM Est & AutoM SE & Norm Est & Norm SE & NormM Est & NormM SE\\
-			\hline
-			$X_0 = 99999, Y_0 = 1$ & 132.003557 & 979.057271 & 0.637136 & 7.591563 & 571.022230 & 8311.989460 & 0.917525 & 10.386777\\
-			\hline
-			$X_0 = 99990, Y_0 = 10$ & 5.566424 & 8.249684 & 0.169114 & 1.056522 & 6.817065 & 11.495275 & 0.152642 & 0.759693\\
-			\hline
-			$X_0 = 99900, Y_0 = 100$ & 0.595913 & 0.859870 & 0.129120 & 0.359819 & 0.764240 & 1.162178 & 0.121305 & 0.305982\\
-			\hline
-			$X_0 = 99000, Y_0 = 1000$ & 0.132387 & 0.115126 & 0.117280 & 0.105368 & 0.146490 & 0.145274 & 0.117695 & 0.116146\\
+		$\sigma_S = 10, \sigma_I = 1$}
+	\begin{footnotesize}
+		\hskip -1.7cm
+	\begin{tabular}{l|r|r|r|r|r|r|r|r}
+		\hline
+		Data Set & Auto Est & Auto SE & AutoM Est & AutoM SE & Norm Est & Norm SE & NormM Est & NormM SE\\
+		\hline
+		$S_0 = 99999, I_0 = 1$ & 45.345556 & NA & 2.018195 & NA & 62.416323 & NA & 1.301719 & NA\\
+		\hline
+		$S_0 = 99990, I_0 = 10$ & 2.756187 & NA & 1.161055 & NA & 3.991245 & NA & 1.288459 & NA\\
+		\hline
+		$S_0 = 99900, I_0 = 100$ & 1.250277 & NA & 1.199468 & NA & 1.237907 & NA & 1.291419 & NA\\
+		\hline
+		$S_0 = 99000, I_0 = 1000$ & 0.834190 & NA & 0.969512 & NA & 0.900419 & NA & 0.936967 & NA\\
 		\hline
@@
 		Different Variances, 
-		$X_0 = 99000, Y_0 = 1000$}
-	\begin{footnotesize}
-		\hskip -1.7cm
-		\begin{tabular}{l|r|r|r|r|r|r|r|r}
-			\hline
-			Data Set & Auto Est & Auto SE & AutoM Est & AutoM SE & Norm Est & Norm SE & NormM Est & NormM SE\\
-			\hline
-			$\sigma_X = 10, \sigma_Y = 1$ & 1.178938 & 20.02418 & 1.185635 & 20.439407 & 1.219909 & 20.39467 & 1.167064 & 20.057929\\
-			\hline
-			$\sigma_X = 100, \sigma_Y = 10$ & 1.356012 & 15.38756 & 1.089050 & 18.415936 & 1.859016 & 20.12416 & 1.150239 & 19.649800\\
-			\hline
-			$\sigma_X = 300, \sigma_Y = 30$ & 3.142258 & 28.52150 & 0.969976 & 15.934410 & 3.201584 & 21.35673 & 1.085100 & 18.379016\\
-			\hline
-			$\sigma_X = 1000, \sigma_Y = 100$ & 7.305927 & 39.63369 & 1.088018 & 18.236155 & 7.978626 & 16.64830 & 0.787073 & 12.417881\\
-			\hline
-			$\sigma_X = 3000, \sigma_Y = 300$ & 16.009523 & 27.38954 & 0.424419 & 3.054379 & 30.351952 & 130.75442 & 1.357313 & 23.125797\\
-			\hline
-			$\sigma_X = 10000, \sigma_Y = 1000$ & 142.030295 & 894.32779 & 0.718062 & 8.362043 & 174.629326 & 706.02904 & 0.390140 & 3.444363\\
+		$S_0 = 99000, I_0 = 1000$}
+	\begin{footnotesize}
+		\hskip -1.7cm
+	\begin{tabular}{l|r|r|r|r|r|r|r|r}
+		\hline
+		Data Set & Auto Est & Auto SE & AutoM Est & AutoM SE & Norm Est & Norm SE & NormM Est & NormM SE\\
+		\hline
+		$\sigma_S = 10, \sigma_I = 1$ & 0.985846 & NA & 0.986994 & NA & 0.983782 & NA & 0.985415 & NA \\
+		\hline
+		$\sigma_S = 100, \sigma_I = 10$ & 1.237596 & NA & 1.209899 & NA & 1.338217 & NA & 1.323149 & NA\\
+		\hline
+		$\sigma_S = 500, \sigma_I = 50$ & 3.544004 & NA & 3.412129 & NA & 4.189569 & NA & 4.080100 & NA\\
+		\hline
+		$\sigma_S = 1000, \sigma_I = 100$ & 6.610460 & NA & 6.295422 & NA & 7.893167 & NA & 7.650749 & NA\\
+		\hline
+		$\sigma_S = 2000, \sigma_I = 200$ & 13.301341 & NA & 12.235214 & NA & 15.500513 & NA & 15.026026 & NA\\
+		\hline
+		$\sigma_S = 5000, \sigma_I = 500$ & 62.640210 & NA & 57.764547 & NA & 49.217485 & NA & 104.547880 & NA\\
+		\hline
+		$\sigma_S = 7500, \sigma_I = 750$ & 71.225178 & NA & 105.181361 & NA & 320.428071 & NA & 107.391030 & NA\\
+		\hline
+		$\sigma_S = 10000, \sigma_I = 1000$ & 116.249187 & NA & 161.948194 & NA & 126.671661 & NA & 121.517490 & NA \\
+		\hline
+		$\sigma_S = 15000, \sigma_I = 1500$ & 235.468174 & NA & 148.873668 & NA & 220.585394 & NA & 308.061708 & NA\\
 		\hline
@@
 	\caption{$\rr$ Estimates and Std. Errs, SIRIPR Model,
-		Other Models, $X_0 = 99950, Y_0 = 50$, 
-		$\sigma_X = 10, \sigma_Y = 1$}
-	\begin{footnotesize}
-		\hskip -1cm
-		\begin{tabular}{l|r|r|r|r|r|r|r|r}
-			\hline
-			Data Set & Auto Est & Auto SE & AutoM Est & AutoM SE & Norm Est & Norm SE & NormM Est & NormM SE\\
-			\hline
-			0th & 1.779121 & 2.482162 & 0.009267 & 0.089052 & 2.352857 & 3.236806 & 0.021922 & 0.200162\\
-			\hline
-			1st & 3.392178 & 21.127161 & 5.199764 & 44.528822 & 4.615877 & 45.803986 & 4.032633 & 33.360591\\
-			\hline
-			2nd & 1.453143 & 5.808155 & 1.073020 & 2.586030 & 1.608484 & 5.985549 & 1.076302 & 2.657303\\
-			\hline
-			3rd & 1.360154 & 1.299049 & 1.192788 & 1.033645 & 1.453000 & 1.464758 & 1.195854 & 1.134779\\
-			\hline
-			4th & 1.300569 & 1.215989 & 1.300569 & 1.215989 & 1.426296 & 1.590604 & 1.153380 & 1.398890\\
-			\hline
-			Linear SIR & 1.379871 & 1.311406 & 1.208521 & 1.003065 & 1.516803 & 1.554241 & 1.208578 & 1.125440\\
+		Other Models, $S_0 = 99950, I_0 = 50$, 
+		$\sigma_S = 10, \sigma_I = 1$}
+	\begin{footnotesize}
+		\hskip -1cm
+	\begin{tabular}{l|r|r|r|r|r|r|r|r}
+		\hline
+		Data Set & Auto Est & Auto SE & AutoM Est & AutoM SE & Norm Est & Norm SE & NormM Est & NormM SE\\
+		\hline
+		0th & 1.706806 & NA & 0.009605 & NA & 2.134967 & NA & 0.025424 & NA\\
+		\hline
+		1st & 3.051573 & NA & 4.744311 & NA & 4.477775 & NA & 4.457355 & NA\\
+		\hline
+		2nd & 1.425110 & NA & 1.181736 & NA & 1.688859 & NA & 1.088296 & NA\\
+		\hline
+		3rd & 1.280001 & NA & 1.497495 & NA & 1.569176 & NA & 1.232182 & NA\\
+		\hline
+		4th & 1.065715 & NA & 1.284711 & NA & 1.373762 & NA & 1.148073 & NA\\
+		\hline
+		Linear SIR & 1.537546 & NA & 1.212776 & NA & 1.611864 & NA & 1.161722 & NA\\
 		\hline
@@
 We estimate the standard error using the jackknife method.
-For every observation, we fit a smoothed spline to X(t) and Y(t) after taking out that specific observation, then take the standard deviation of $t$ the $\rr$ estimates.
+For every observation, we fit a smoothed spline to S(t) and I(t) after taking out that specific observation, then take the standard deviation of $t$ the $\rr$ estimates.
 
@@
 All of them are now within $\pm$ 0.5 of the true $\rr$ for the data generated from the classic SIR model.
-We still see the same pattern as in the non-smoothed incidence-prevalence ratio estimates, that the estimates for $\rr$ decrease as $Y_0$ increases relative to the total population and that they increase as we increase $\sigma_X$ and $\sigma_Y$.
+We still see the same pattern as in the non-smoothed incidence-prevalence ratio estimates, that the estimates for $\rr$ decrease as $I_0$ increases relative to the total population and that they increase as we increase $\sigma_S$ and $\sigma_I$.
 For the data generated from other models, we see that there are some negative $\rr$ estimates for the 0th and 1st order linear model data.
@@
 The standard errors for this model are large compared to the SIR models that are soled using optim() (such as least squares).
-We also see that the standard error increases dramatically as $\sigma_X$ and $\sigma_Y$ increase.
+We also see that the standard error increases dramatically as $\sigma_S$ and $\sigma_I$ increase.
 
@@
 	\caption{$\rr$ Estimates and Std. Errs, SIRSIPR Model,
-		Baseline Data, $X_0 = 99950, Y_0 = 50$, 
-		$\sigma_X = 10, \sigma_Y = 1$}
-	\begin{footnotesize}
-		\hskip -1cm
-		\begin{tabular}{l|r|r|r|r|r|r|r|r}
-			\hline
-			Data Set & Auto Est & Auto SE & AutoM Est & AutoM SE & Norm Est & Norm SE & NormM Est & NormM SE\\
-			\hline
-			Baseline 1 & 0.111589 & 0.013332 & 0.116714 & 0.004233 & 0.121731 & 0.008296 & 0.122992 & 0.002911\\
-			\hline
-			Baseline 2 & 0.113638 & 0.011308 & 0.115175 & 0.003540 & 0.119221 & 0.007058 & 0.120945 & 0.005199\\
-			\hline
-			Baseline 3 & 0.118958 & 0.009720 & 0.126134 & 0.011717 & 0.121108 & 0.007978 & 0.122298 & 0.004135\\
-			\hline
-			Baseline 4 & 0.129073 & 0.010889 & 0.125410 & 0.004537 & 0.120610 & 0.008410 & 0.120497 & 0.003618\\
-			\hline
-			Baseline 5 & 0.122424 & 0.011326 & 0.129906 & 0.004269 & 0.118352 & 0.007518 & 0.124123 & 0.002414\\
+		Baseline Data, $S_0 = 99950, I_0 = 50$, 
+		$\sigma_S = 10, \sigma_I = 1$}
+	\begin{footnotesize}
+		\hskip -1cm
+	\begin{tabular}{l|r|r|r|r|r|r|r|r}
+		\hline
+		Data Set & Auto Est & Auto SE & AutoM Est & AutoM SE & Norm Est & Norm SE & NormM Est & NormM SE\\
+		\hline
+		Baseline 1 & 1.168689 & 0.211967 & 1.433300 & 0.275675 & 1.247849 & 0.225538 & 1.403578 & 0.253343\\
+		\hline
+		Baseline 2 & 1.243889 & 0.223453 & 1.169045 & 0.207609 & 1.165859 & 0.213966 & 1.539909 & 0.280209\\
+		\hline
+		Baseline 3 & 1.351741 & 0.236626 & 1.236297 & 0.210356 & 1.188403 & 0.218132 & 1.214513 & 0.219824\\
+		\hline
+		Baseline 4 & 1.222771 & 0.209546 & 1.507230 & 0.279421 & 1.260300 & 0.230086 & 1.290742 & 0.231896\\
+		\hline
+		Baseline 5 & 1.106582 & 0.190483 & 1.261039 & 0.229758 & 1.248270 & 0.227508 & 1.327386 & 0.239792\\
 		\hline
@@
 		Different Starting Populations, 
-		$\sigma_X = 10, \sigma_Y = 1$}
-	\begin{footnotesize}
-		\hskip -1.7cm
-		\begin{tabular}{l|r|r|r|r|r|r|r|r}
-			\hline
-			Data Set & Auto Est & Auto SE & AutoM Est & AutoM SE & Norm Est & Norm SE & NormM Est & NormM SE\\
-			\hline
-			$X_0 = 99999, Y_0 = 1$ & 0.496556 & 0.404930 & 0.249343 & 0.080732 & 0.244347 & 0.534660 & 0.279914 & 0.121538\\
-			\hline
-			$X_0 = 99990, Y_0 = 10$ & 0.082031 & 0.045989 & 0.130633 & 0.007387 & 0.116519 & 0.052715 & 0.119705 & 0.006003\\
-			\hline
-			$X_0 = 99900, Y_0 = 100$ & 0.120003 & 0.006022 & 0.119230 & 0.003320 & 0.121229 & 0.004296 & 0.120600 & 0.001531\\
-			\hline
-			$X_0 = 99000, Y_0 = 1000$ & 0.117786 & 0.001351 & 0.117888 & 0.001247 & 0.117755 & 0.001314 & 0.117804 & 0.001219\\
+		$\sigma_S = 10, \sigma_I = 1$}
+	\begin{footnotesize}
+		\hskip -1.7cm
+	\begin{tabular}{l|r|r|r|r|r|r|r|r}
+		\hline
+		Data Set & Auto Est & Auto SE & AutoM Est & AutoM SE & Norm Est & Norm SE & NormM Est & NormM SE\\
+		\hline
+		$S_0 = 99999, I_0 = 1$ & 1.125861 & 2.218581 & 1.396896 & 0.546108 & 1.414820 & 1.323435 & 1.265761 & 0.231673\\
+		\hline
+		$S_0 = 99990, I_0 = 10$ & 1.316003 & 0.274009 & 1.446887 & 0.251195 & 1.224239 & 0.233581 & 1.430876 & 0.234044\\
+		\hline
+		$S_0 = 99900, I_0 = 100$ & 1.456975 & 0.246964 & 1.203267 & 0.208514 & 1.169186 & 0.197852 & 1.388754 & 0.233590\\
+		\hline
+		$S_0 = 99000, I_0 = 1000$ & 1.061364 & 0.014436 & 0.913393 & 0.012300 & 1.027593 & 0.013776 & 1.258616 & 0.016998\\
 		\hline
@@
 		Different Variances, 
-		$X_0 = 99000, Y_0 = 1000$}
-	\begin{footnotesize}
-		\hskip -1.7cm
-		\begin{tabular}{l|r|r|r|r|r|r|r|r}
-			\hline
-			Data Set & Auto Est & Auto SE & AutoM Est & AutoM SE & Norm Est & Norm SE & NormM Est & NormM SE\\
-			\hline
-			$\sigma_X = 10, \sigma_Y = 1$ & 0.119798 & 0.015080 & 0.119388 & 0.015098 & 0.119629 & 0.015084 & 0.119720 & 0.015063\\
-			\hline
-			$\sigma_X = 100, \sigma_Y = 10$ & 0.123180 & 0.015455 & 0.120830 & 0.015114 & 0.119951 & 0.015663 & 0.118726 & 0.015203\\
-			\hline
-			$\sigma_X = 300, \sigma_Y = 30$ & 0.108430 & 0.022619 & 0.119020 & 0.014395 & 0.118138 & 0.018715 & 0.120683 & 0.015369\\
-			\hline
-			$\sigma_X = 1000, \sigma_Y = 100$ & 0.151905 & 0.067340 & 0.122619 & 0.016990 & 0.120618 & 0.053277 & 0.138444 & 0.013843\\
-			\hline
-			$\sigma_X = 3000, \sigma_Y = 300$ & 0.148289 & 0.141669 & 0.237888 & 0.054042 & 0.121580 & 0.147039 & 0.135574 & 0.015871\\
-			\hline
-			$\sigma_X = 10000, \sigma_Y = 1000$ & -0.253735 & 0.457021 & 0.636736 & 0.092124 & 0.113880 & 0.555507 & 0.208459 & 0.018114\\
+		$S_0 = 99000, I_0 = 1000$}
+	\begin{footnotesize}
+		\hskip -1.7cm
+	\begin{tabular}{l|r|r|r|r|r|r|r|r}
+		\hline
+		Data Set & Auto Est & Auto SE & AutoM Est & AutoM SE & Norm Est & Norm SE & NormM Est & NormM SE\\
+		\hline
+		$\sigma_S = 10, \sigma_I = 1$ & 1.077390 & 0.014486 & 1.076946 & 0.014585 & 1.077445 & 0.014524 & 1.077259 & 0.014413\\
+		\hline
+		$\sigma_S = 100, \sigma_I = 10$ & 1.078355 & 0.015334 & 1.073914 & 0.016113 & 1.078901 & 0.015210 & 1.077047 & 0.014195\\
+		\hline
+		$\sigma_S = 500, \sigma_I = 50$ & 1.082754 & 0.030646 & 1.060547 & 0.031371 & 1.085358 & 0.024514 & 1.076160 & 0.022713\\
+		\hline
+		$\sigma_S = 1000, \sigma_I = 100$ & 1.088511 & 0.055954 & 1.044073 & 0.055612 & 1.093402 & 0.041119 & 1.075184 & 0.041831\\
+		\hline
+		$\sigma_S = 2000, \sigma_I = 200$ & 1.100914 & 0.107709 & 1.011872 & 0.106957 & 1.109397 & 0.076725 & 1.073676 & 0.086113\\
+		\hline
+		$\sigma_S = 5000, \sigma_I = 500$ & 1.146051 & 0.254897 & 0.920643 & 0.268296 & 1.156686 & 0.182392 & 1.072789 & 0.244303\\
+		\hline
+		$\sigma_S = 7500, \sigma_I = 750$ & 1.194432 & 0.367353 & 0.849913 & 0.409537 & 1.195360 & 0.266558 & 1.076371 & 0.408601\\
+		\hline
+		$\sigma_S = 10000, \sigma_I = 1000$ & 1.255238 & 0.471487 & 0.783195 & 0.557464 & 1.233433 & 0.347153 & 1.084101 & 0.611590\\
+		\hline
+		$\sigma_S = 15000, \sigma_I = 1500$ & 1.431325 & 0.658287 & 0.659451 & 0.877489 & 1.307994 & 0.498443 & 1.113121 & 1.186676\\
 		\hline
@@
 	\caption{$\rr$ Estimates and Std. Errs, SIRSIPR Model,
-		Other Models, $X_0 = 99950, Y_0 = 50$, 
-		$\sigma_X = 10, \sigma_Y = 1$}
-	\begin{footnotesize}
-		\hskip -1cm
-		\begin{tabular}{l|r|r|r|r|r|r|r|r}
-			\hline
-			Data Set & Auto Est & Auto SE & AutoM Est & AutoM SE & Norm Est & Norm SE & NormM Est & NormM SE\\
-			\hline
-			0th & 0.014270 & 0.013561 & 0.002865 & 0.000486 & -0.006574 & 0.010632 & 0.005140 & 0.001792\\
-			\hline
-			1st & -9.099024 & 94554.286222 & 2.655271 & 153.459186 & 2.339653 & 190.604213 & 2.285211 & 200.382597\\
-			\hline
-			2nd & 1.045203 & 0.201211 & 1.076390 & 0.156022 & 1.042439 & 0.205421 & 1.074918 & 0.152346\\
-			\hline
-			3rd & 1.279959 & 0.299658 & 1.290916 & 0.315063 & 1.279248 & 0.295345 & 1.280916 & 0.294307\\
-			\hline
-			4th & 1.192312 & 0.187455 & 1.192312 & 0.187455 & 1.203649 & 0.200467 & 1.202999 & 0.198628\\
-			\hline
-			Linear SIR & 1.296731 & 0.221560 & 1.304489 & 0.229988 & 1.294079 & 0.215113 & 1.294364 & 0.215810\\
+		Other Models, $S_0 = 99950, I_0 = 50$, 
+		$\sigma_S = 10, \sigma_I = 1$}
+	\begin{footnotesize}
+		\hskip -1cm
+	\begin{tabular}{l|r|r|r|r|r|r|r|r}
+		\hline
+		Data Set & Auto Est & Auto SE & AutoM Est & AutoM SE & Norm Est & Norm SE & NormM Est & NormM SE\\
+		\hline
+		0th & 0.016596 & 1.57720e-02 & 0.003117 & 0.000529 & -0.006329 & 0.010236 & 0.005741 & 0.002001\\
+		\hline
+		1st & -7.847035 & 8.15440e+04 & 2.832331 & 163.692234 & 2.274969 & 185.334642 & 2.106098 & 184.676745\\
+		\hline
+		2nd & 1.094937 & 2.10785e-01 & 0.956905 & 0.138703 & 0.869594 & 0.171361 & 1.179674 & 0.167192\\
+		\hline
+		3rd & 1.238637 & 2.89984e-01 & 1.446063 & 0.352928 & 1.237537 & 0.285715 & 1.230995 & 0.282838\\
+		\hline
+		4th & 1.103846 & 1.73547e-01 & 1.131708 & 0.177927 & 1.133062 & 0.188710 & 1.349997 & 0.222899\\
+		\hline
+		Linear SIR & 1.396246 & 2.38564e-01 & 1.269573 & 0.223832 & 1.305677 & 0.217041 & 1.333689 & 0.222366\\
 		\hline
@@
 The $\rr$ estimates here are very accurate and precise.
-The estimates from the baseline and changing $Y_0$ data sets are all very close to the true $\rr$ value.
-Increasing $\sigma_X$ and $\sigma_Y$ does decrease the model accuracy.
+The estimates from the baseline and changing $I_0$ data sets are all very close to the true $\rr$ value.
+Increasing $\sigma_S$ and $\sigma_I$ does decrease the model accuracy.
 It is not quite as good for data generated from other models, but it gives reasonable results for the 0th and 1st order linear models unlike some of the other models.
@@
 The standard errors for the model are relatively small as well, and there are no unusually large values.
-They tend to decrease as $Y_0$ increases and they tend to increase as $\sigma_X$ and $\sigma_Y$ increase.
+They tend to decrease as $I_0$ increases and they tend to increase as $\sigma_S$ and $\sigma_I$ increase.
 
@@
 	\caption{$\rr$ Estimates and Std. Errs, HARKO Model, 
-		Baseline Data, $X_0 = 99950, Y_0 = 50$, 
-		$\sigma_X = 10, \sigma_Y = 1$}
-	\begin{footnotesize}
-		\hskip -1cm
-		\begin{tabular}{l|r|r|r|r|r|r|r|r}
-			\hline
-			Data Set & Auto Est & Auto SE & AutoM Est & AutoM SE & Norm Est & Norm SE & NormM Est & NormM SE\\
-			\hline
-			Baseline 1 & 1.780655 & 0.023436 & 1.983971 & 0.006583 & 1.887483 & 0.016523 & 1.977806 & 0.004671\\
-			\hline
-			Baseline 2 & 1.831444 & 0.021164 & 2.002378 & 0.010261 & 1.902733 & 0.017365 & 2.003881 & 0.003998\\
-			\hline
-			Baseline 3 & 1.839113 & 0.022491 & 1.914380 & 0.010098 & 1.906137 & 0.016586 & 1.987737 & 0.003900\\
-			\hline
-			Baseline 4 & 1.794680 & 0.017916 & 1.840147 & 0.007965 & 1.894689 & 0.016695 & 1.990857 & 0.003619\\
-			\hline
-			Baseline 5 & 1.823206 & 0.020518 & 1.892921 & 0.011397 & 1.914908 & 0.016957 & 1.965487 & 0.003611\\
+		Baseline Data, $S_0 = 99950, I_0 = 50$, 
+		$\sigma_S = 10, \sigma_I = 1$}
+	\begin{footnotesize}
+		\hskip -1cm
+	\begin{tabular}{l|r|r|r|r|r|r|r|r}
+		\hline
+		Data Set & Auto Est & Auto SE & AutoM Est & AutoM SE & Norm Est & Norm SE & NormM Est & NormM SE\\
+		\hline
+		Baseline1 & 1.199905 & 7.6e-05 & 1.199973 & 7.2e-05 & 1.200018 & 5.4e-05 & 1.199920 & 4.1e-05\\
+		\hline
+		Baseline2 & 1.200098 & 7.9e-05 & 1.200053 & 7.3e-05 & 1.200032 & 5.5e-05 & 1.200006 & 4.4e-05\\
+		\hline
+		Baseline3 & 1.200107 & 7.8e-05 & 1.199705 & 6.7e-05 & 1.200028 & 5.3e-05 & 1.199869 & 4.1e-05\\
+		\hline
+		Baseline4 & 1.200077 & 7.0e-05 & 1.200122 & 6.0e-05 & 1.200015 & 5.7e-05 & 1.199999 & 4.4e-05\\
+		\hline
+		Baseline5 & 1.199987 & 8.5e-05 & 1.200006 & 6.5e-05 & 1.199914 & 5.7e-05 & 1.200139 & 4.0e-05\\
 		\hline
@@
 		Different Starting Populations, 
-		$\sigma_X = 10, \sigma_Y = 1$}
-	\begin{footnotesize}
-		\hskip -1.7cm
-		\begin{tabular}{l|r|r|r|r|r|r|r|r}
-			\hline
-			Data Set & Auto Est & Auto SE & AutoM Est & AutoM SE & Norm Est & Norm SE & NormM Est & NormM SE\\
-			\hline
-			$X_0 = 99999, Y_0 = 1$ & 0.990097 & 0.006292 & 1.257974 & 0.013850 & 1.007582 & 0.005853 & 1.191026 & 0.003584\\
-			\hline
-			$X_0 = 99990, Y_0 = 10$ & 1.123018 & 0.021783 & 1.735611 & 0.013290 & 1.245655 & 0.024246 & 1.923291 & 0.010741\\
-			\hline
-			$X_0 = 99900, Y_0 = 100$ & 1.951514 & 0.012785 & 2.016594 & 0.005770 & 1.966805 & 0.008724 & 1.983918 & 0.002413\\
-			\hline
-			$X_0 = 99000, Y_0 = 1000$ & 2.006289 & 0.001318 & 1.995221 & 0.001014 & 2.000268 & 0.000951 & 2.000158 & 0.000544\\
+		$\sigma_S = 10, \sigma_I = 1$}
+	\begin{footnotesize}
+		\hskip -1.7cm
+	\begin{tabular}{l|r|r|r|r|r|r|r|r}
+		\hline
+		Data Set & Auto Est & Auto SE & AutoM Est & AutoM SE & Norm Est & Norm SE & NormM Est & NormM SE\\
+		\hline
+		$S_0 = 99999, I_0 = 1$ & 1.183163 & 0.002609 & 1.196363 & 0.001372 & 1.188140 & 0.002488 & 1.192792 & 0.000554\\
+		\hline
+		$S_0 = 99990, I_0 = 10$ & 1.200267 & 0.000295 & 1.200320 & 0.000269 & 1.200069 & 0.000259 & 1.200435 & 0.000146\\
+		\hline
+		$S_0 = 99900, I_0 = 100$ & 1.200009 & 0.000044 & 1.200014 & 0.000043 & 1.199928 & 0.000032 & 1.200010 & 0.000024\\
+		\hline
+		$S_0 = 99000, I_0 = 1000$ & 1.199986 & 0.000013 & 1.199981 & 0.000013 & 1.200009 & 0.000010 & 1.200006 & 0.000010\\
 		\hline
@@
 	\caption{$\rr$ Estimates and Std. Errs, HARKO Model, 
-		Different Variances, $X_0 = 99000$, $Y_0 = 1000$}
-	\begin{footnotesize}
-		\hskip -1.7cm
-		\begin{tabular}{l|r|r|r|r|r|r|r|r}
-			\hline
-			Data Set & Auto Est & Auto SE & AutoM Est & AutoM SE & Norm Est & Norm SE & NormM Est & NormM SE\\
-			\hline
-			$\sigma_X = 10, \sigma_Y = 1$ & 1.997497 & 0.001357 & 2.002070 & 0.001141 & 2.000316 & 0.001015 & 2.000543 & 0.000521\\
-			\hline
-			$\sigma_X = 100, \sigma_Y = 10$ & 1.918862 & 0.009993 & 1.990296 & 0.004519 & 1.968331 & 0.009515 & 2.002762 & 0.002228\\
-			\hline
-			$\sigma_X = 300, \sigma_Y = 30$ & 1.688924 & 0.027688 & 1.822846 & 0.012972 & 1.786604 & 0.021708 & 1.999747 & 0.003704\\
-			\hline
-			$\sigma_X = 1000, \sigma_Y = 100$ & 1.202346 & 0.017004 & 1.860911 & 0.008679 & 1.244529 & 0.023070 & 1.863444 & 0.009948\\
-			\hline
-			$\sigma_X = 3000, \sigma_Y = 300$ & 1.043406 & 0.009591 & 1.414679 & 0.009365 & 1.027725 & 0.010276 & 1.696223 & 0.003871\\
-			\hline
-			$\sigma_X = 10000, \sigma_Y = 1000$ & 1.044556 & 0.008612 & 1.093328 & 0.004175 & 1.021139 & 0.007466 & 1.203591 & 0.007232\\
+		Different Variances, $S_0 = 99000$, $I_0 = 1000$}
+	\begin{footnotesize}
+		\hskip -1.7cm
+	\begin{tabular}{l|r|r|r|r|r|r|r|r}
+		\hline
+		Error & Auto Est & Auto SE & AutoM Est & AutoM SE & Norm Est & Norm SE & NormM Est & NormM SE\\
+		\hline
+		$\sigma_S = 10, \sigma_I = 1$ & 1.200000 & 0.000014 & 1.200036 & 0.000013 & 1.199988 & 0.000010 & 1.199994 & 0.000010\\
+		\hline
+		$\sigma_S = 100, \sigma_I = 10$ & 1.200006 & 0.000142 & 1.200377 & 0.000133 & 1.199883 & 0.000102 & 1.199945 & 0.000097\\
+		\hline
+		$\sigma_S = 500, \sigma_I = 50$ & 1.200213 & 0.000712 & 1.202158 & 0.000666 & 1.199521 & 0.000506 & 1.199785 & 0.000482\\
+		\hline
+		$\sigma_S = 1000, \sigma_I = 100$ & 1.200917 & 0.001424 & 1.204990 & 0.001330 & 1.199297 & 0.001008 & 1.199719 & 0.000959\\
+		\hline
+		$\sigma_S = 2000, \sigma_I = 200$ & 1.203936 & 0.002827 & 1.212506 & 0.002627 & 1.199576 & 0.001988 & 1.200005 & 0.001892\\
+		\hline
+		$\sigma_S = 5000, \sigma_I = 500$ & 1.225212 & 0.006669 & 1.244598 & 0.006057 & 1.205086 & 0.004659 & 1.203631 & 0.004414\\
+		\hline
+		$\sigma_S = 7500, \sigma_I = 750$ & 1.254099 & 0.009665 & 1.274883 & 0.008428 & 1.213266 & 0.006584 & 1.208945 & 0.006158\\
+		\hline
+		$\sigma_S = 10000, \sigma_I = 1000$ & 1.292933 & 0.013084 & 1.306145 & 0.010690 & 1.223726 & 0.008360 & 1.215920 & 0.007672\\
+		\hline
+		$\sigma_S = 15000, \sigma_I = 1500$ & 1.439976 & 0.028186 & 1.381843 & 0.016227 & 1.252207 & 0.012016 & 1.235788 & 0.010554\\
 		\hline
@@
 	\caption{$\rr$ Estimates and Std. Errs, HARKO Model,
-		Other Models, $X_0 = 99950, Y_0 = 50$, 
-		$\sigma_X = 10, \sigma_Y = 1$}
-	\begin{footnotesize}
-		\hskip -1cm
-		\begin{tabular}{l|r|r|r|r|r|r|r|r}
-			\hline
-			Data Set & Auto Est & Auto SE & AutoM Est & AutoM SE & Norm Est & Norm SE & NormM Est & NormM SE\\
-			\hline
-			0th & 0.992417 & 0.005898 & 0.801563 & 0.009989 & 0.984522 & 0.004772 & 0.960479 & 0.003184\\
-			\hline
-			1st & 1.200537 & 0.000641 & 1.200834 & 0.000664 & 1.201039 & 0.000648 & 1.200963 & 0.000653\\
-			\hline
-			2nd & 1.198169 & 0.000451 & 1.198486 & 0.000407 & 1.198510 & 0.000451 & 1.198652 & 0.000403\\
-			\hline
-			3rd & 1.207309 & 0.000196 & 1.207754 & 0.000188 & 1.207402 & 0.000184 & 1.207447 & 0.000187\\
-			\hline
-			4th & 1.210499 & 0.000136 & 1.210499 & 0.000136 & 1.210671 & 0.000121 & 1.210571 & 0.000114\\
-			\hline
-			Linear SIR & 1.258920 & 0.000994 & 1.258462 & 0.001020 & 1.258587 & 0.001008 & 1.258670 & 0.001007\\
+		Other Models, $S_0 = 99950, I_0 = 50$, 
+		$\sigma_S = 10, \sigma_I = 1$}
+	\begin{footnotesize}
+		\hskip -1cm
+	\begin{tabular}{l|r|r|r|r|r|r|r|r}
+		\hline
+		Data Set & Auto Est & Auto SE & AutoM Est & AutoM SE & Norm Est & Norm SE & NormM Est & NormM SE\\
+		\hline
+		0th & 1.000194 & 0.000021  & 1.000282 & 0.000013 & 1.000185 & 0.000015 & 1.000318 & 0.000018\\
+		\hline
+		1st & 1.040689 & 0.000468 & 1.040788 & 0.000469 & 1.040680 & 0.000468 & 1.040532 & 0.000467\\
+		\hline
+		2nd & 1.041906 & 0.000476 & 1.041744 & 0.000475 & 1.041800 & 0.000475 & 1.041751 & 0.000475\\
+		\hline
+		3rd & 1.046688 & 0.000504 & 1.046444 & 0.000503 & 1.046546 & 0.000504 & 1.046513 & 0.000503\\
+		\hline
+		4th & 1.041571 & 0.000473 & 1.041571 & 0.000473 & 1.041466 & 0.000473 & 1.041470 & 0.000473\\
+		\hline
+		Linear SIR & 1.061074 & 0.000585 & 1.061134 & 0.000586 & 1.061092 & 0.000585 & 1.061038 & 0.000585\\
 		\hline
@@
 	\caption{$\rr$ Estimates and Std. Errs, SIRMC Model,
-		Baseline Data, $X_0 = 99950, Y_0 = 50$, 
-		$\sigma_X = 10, \sigma_Y = 1$}
-	\begin{footnotesize}
-		\hskip -1cm
-		\begin{tabular}{l|r|r|r|r|r|r|r|r}
-			\hline
-			Data Set & Auto Est & Auto SE & AutoM Est & AutoM SE & Norm Est & Norm SE & NormM Est & NormM SE\\
-			\hline
-			Baseline 1 & 5.127170 & 179.2844 & 5.234313 & 40.61000 & 5.086225 & 216.3083 & 5.100115 & 26.56028\\
-			\hline
-			Baseline 2 & 5.158918 & 189.7917 & 4.921625 & 26.13471 & 5.126360 & 217.2991 & 5.074875 & 29.08225\\
-			\hline
-			Baseline 3 & 5.160983 & 169.1971 & 5.207777 & 47.15165 & 5.083438 & 195.1501 & 5.119470 & 28.63475\\
-			\hline
-			Baseline 4 & 5.161219 & 170.6186 & 5.127197 & 34.24464 & 5.207647 & 269.4176 & 5.135429 & 31.76258\\
-			\hline
-			Baseline 5 & 5.187114 & 187.2816 & 5.131911 & 31.43095 & 5.077370 & 186.0883 & 5.098254 & 23.90021\\
+		Baseline Data, $S_0 = 99950, I_0 = 50$, 
+		$\sigma_S = 10, \sigma_I = 1$}
+	\begin{footnotesize}
+		\hskip -1cm
+	\begin{tabular}{l|r|r|r|r|r|r|r|r}
+		\hline
+		Data Set & Auto Est & Auto SE & AutoM Est & AutoM SE & Norm Est & Norm SE & NormM Est & NormM SE\\
+		\hline
+		Baseline 1 & 2.840123 & 0.346977 & 2.832794 & 0.298082 & 2.841003 & 0.427146 & 2.842369 & 0.361052\\
+		\hline
+		Baseline 2 & 2.844091 & 0.344388 & 2.840638 & 0.305295 & 2.841250 & 0.443461 & 2.838177 & 0.336637\\
+		\hline
+		Baseline 3 & 2.841266 & 0.324645 & 2.845812 & 0.316198 & 2.836290 & 0.398227 & 2.837974 & 0.365954\\
+		\hline
+		Baseline 4 & 2.843637 & 0.362766 & 2.840456 & 0.300201 & 2.843353 & 0.462407 & 2.836546 & 0.364677\\
+		\hline
+		Baseline 5 & 2.839390 & 0.335435 & 2.841471 & 0.309445 & 2.839357 & 0.445301 & 2.840635 & 0.376539\\
 		\hline
@@
 		Different Starting Populations, 
-		$\sigma_X = 10, \sigma_Y = 1$}
-	\begin{footnotesize}
-		\hskip -1.7cm
-		\begin{tabular}{l|r|r|r|r|r|r|r|r}
-			\hline
-			Data Set & Auto Est & Auto SE & AutoM Est & AutoM SE & Norm Est & Norm SE & NormM Est & NormM SE\\
-			\hline
-			$X_0 = 99999, Y_0 = 1$ & 2.502983 & 31.703174 & 4.587050 & 141.618120 & NaN & 1.53910 & 4.584606 & 211.645976\\
-			\hline
-			$X_0 = 99990, Y_0 = 10$ & 5.370030 & 1402.395267 & 4.745881 & 37.893736 & 5.079299 & 1000.80267 & 4.808880 & 30.644371\\
-			\hline
-			$X_0 = 99900, Y_0 = 100$ & 5.083530 & 78.047548 & 5.141349 & 24.913580 & 5.036936 & 94.46507 & 5.137375 & 21.960267\\
-			\hline
-			$X_0 = 99000, Y_0 = 1000$ & 5.134886 & 9.346665 & 5.112739 & 7.046171 & 5.112847 & 11.30402 & 5.113115 & 7.906187\\
+		$\sigma_S = 10, \sigma_I = 1$}
+	\begin{footnotesize}
+		\hskip -1.7cm
+	\begin{tabular}{l|r|r|r|r|r|r|r|r}
+		\hline
+		Data Set & Auto Est & Auto SE & AutoM Est & AutoM SE & Norm Est & Norm SE & NormM Est & NormM SE\\
+		\hline
+		$S_0 = 99999, I_0 = 1$ & 2.837554 & 11.144048 & 2.923039 & 2.862449 & 2.735179 & 12.672799 & 2.706761 & 1.871524\\
+		\hline
+		$S_0 = 99990, I_0 = 10$ & 2.846620 & 1.220266 & 2.811991 & 0.715527 & 2.844415 & 1.710048 & 2.819770 & 0.863066\\
+		\hline
+		$S_0 = 99900, I_0 = 100$ & 2.843606 & 0.232628 & 2.841555 & 0.218771 & 2.839663 & 0.279605 & 2.839069 & 0.256531\\
+		\hline
+		$S_0 = 99000, I_0 = 1000$ & 2.843031 & 0.115210 & 2.843140 & 0.112716 & 2.843451 & 0.141831 & 2.843141 & 0.156539\\
 		\hline
@@
 		Different Variances, 
-		$X_0 = 99000, Y_0 = 1000$}
-	\begin{footnotesize}
-		\hskip -1.7cm
-		\begin{tabular}{l|r|r|r|r|r|r|r|r}
-			\hline
-			Data Set & Auto Est & Auto SE & AutoM Est & AutoM SE & Norm Est & Norm SE & NormM Est & NormM SE\\
-			\hline
-			$\sigma_X = 10, \sigma_Y = 1$ & 5.107658 & 40.07322 & 5.118340 & 41.49349 & 5.112780 & 42.18333 & 5.109027 & 40.04422\\
-			\hline
-			$\sigma_X = 100, \sigma_Y = 10$ & 5.047351 & 74.96928 & 5.074418 & 40.22261 & 5.055666 & 112.53820 & 5.060957 & 39.62932\\
-			\hline
-			$\sigma_X = 300, \sigma_Y = 30$ & 5.183856 & 304.55113 & 4.929632 & 34.41019 & 5.459944 & 664.73372 & 5.129073 & 49.24831\\
-			\hline
-			$\sigma_X = 1000, \sigma_Y = 100$ & 4.267588 & 158.79219 & 5.122059 & 60.44409 & 5.741570 & 4073.99286 & 5.133013 & 56.70565\\
-			\hline
-			$\sigma_X = 3000, \sigma_Y = 300$ & NaN & 74731.85279 & 3.918439 & 18.49635 & 3.200282 & 80.21230 & 4.771752 & 45.08632\\
-			\hline
-			$\sigma_X = 10000, \sigma_Y = 1000$ & 2.120699 & 17.46286 & 3.767783 & 35.76247 & NaN & 83.73480 & 3.884264 & 23.04767\\
+		$S_0 = 99000, I_0 = 1000$}
+	\begin{footnotesize}
+		\hskip -1.7cm
+	\begin{tabular}{l|r|r|r|r|r|r|r|r}
+		\hline
+		Data Set & Auto Est & Auto SE & AutoM Est & AutoM SE & Norm Est & Norm SE & NormM Est & NormM SE\\
+		\hline
+		$\sigma_S = 10, \sigma_I = 1$ & 2.843163 & 0.121074 & 2.843082 & 0.117326 & 2.843563 & 0.163321 & 2.843234 & 0.150979\\
+		\hline
+		$\sigma_S = 100, \sigma_I = 10$ & 2.841006 & 1.204736 & 2.840820 & 1.166121 & 2.845950 & 1.640665 & 2.841955 & 1.506458\\
+		\hline
+		$\sigma_S = 500, \sigma_I = 50$ & 2.822320 & 5.786070 & 2.834664 & 5.718502 & 2.867771 & 8.562812 & 2.831940 & 7.395682\\
+		\hline
+		$\sigma_S = 1000, \sigma_I = 100$ & 2.780071 & 10.584335 & 2.835372 & 11.341172 & 2.922494 & 19.076712 & 2.809761 & 14.170863\\
+		\hline
+		$\sigma_S = 2000, \sigma_I = 200$ & 2.649423 & 16.091219 & 2.861883 & 23.385320 & 3.148266 & 59.622533 & 2.736588 & 24.502271\\
+		\hline
+		$\sigma_S = 5000, \sigma_I = 500$ & 2.167158 & 14.446035 & 3.104554 & 87.360354 & NaN & 385.053384 & 2.384396 & 29.821372\\
+		\hline
+		$\sigma_S = 7500, \sigma_I = 750$ & 1.860269 & 11.045568 & 3.451396 & 242.494249 & NaN & 26.530063 & 2.075721 & 23.380306\\
+		\hline
+		$\sigma_S = 10000, \sigma_I = 1000$ & 1.661519 & 9.373613 & 3.871861 & 694.256270 & NaN & 8.570973 & 1.821502 & 17.932802\\
+		\hline
+		$\sigma_S = 15000, \sigma_I = 1500$ & 1.494903 & 10.226678 & 4.113933 & 1491.975630 & NaN & 2.509313 & 1.485749 & 12.308392\\
 		\hline
@@
 	\caption{$\rr$ Estimates and Std. Errs, SIRMC Model,
-		Other Models, $X_0 = 99950, Y_0 = 50$, 
-		$\sigma_X = 10, \sigma_Y = 1$}
+		Other Models, $S_0 = 99950, I_0 = 50$, 
+		$\sigma_S = 10, \sigma_I = 1$}
 	\begin{footnotesize}
@@
 There is a spread in the estimates for the baseline data of around 0.2.
-The estimates appear to increase as $\sigma_X$ and $\sigma_Y$ increase.
-On the other hand, there does not seem to be a consistent effect with changing $Y_0$
+The estimates appear to increase as $\sigma_S$ and $\sigma_I$ increase.
+On the other hand, there does not seem to be a consistent effect with changing $I_0$
 
@@
 	\caption{$\rr$ Estimates and Std. Errs, SEQBAYES Model, 
-		Baseline Data, $X_0 = 99950, Y_0 = 50$, 
-		$\sigma_X = 10, \sigma_Y = 1$}
-	\begin{footnotesize}
-		\hskip -1cm
-		\begin{tabular}{l|r|r|r|r|r|r|r|r}
-			Data Set & Auto Est & Auto SE & AutoM Est & AutoM SE & Norm Est & Norm SE & NormM Est & NormM SE\\
-			\hline
-			Baseline 1 & 1.120551 & 0.055408 & 0.834692 & 0.047821 & 1.085113 & 0.054524 & 1.000000 & 0.052342\\
-			\hline
-			Baseline 2 & 1.163677 & 0.056464 & 0.963486 & 0.051378 & 1.086715 & 0.054565 & 1.266142 & 0.058897\\
-			\hline
-			Baseline 3 & 1.006208 & 0.052505 & 1.000000 & 0.052342 & 1.000000 & 0.052342 & 1.200584 & 0.057352\\
-			\hline
-			Baseline 4 & 0.874006 & 0.048934 & 1.200129 & 0.057341 & 0.840416 & 0.047984 & 1.000000 & 0.052342\\
-			\hline
-			Baseline 5 & 1.000000 & 0.052342 & 1.000000 & 0.052342 & 1.200097 & 0.057341 & 1.169565 & 0.056606\\
+		Baseline Data, $S_0 = 99950, I_0 = 50$, 
+		$\sigma_S = 10, \sigma_I = 1$}
+	\begin{footnotesize}
+		\hskip -1cm
+	\begin{tabular}{l|r|r|r|r|r|r|r|r}
+		\hline
+		Data Set & Auto Est & Auto SE & AutoM Est & AutoM SE & Norm Est & Norm SE & NormM Est & NormM SE\\
+		\hline
+		Baseline1 & 1.201518 & 0.057374 & 1.058315 & 0.053847 & 1.239136 & 0.058266 & 1.237283 & 0.058222\\
+		\hline
+		Baseline2 & 1.246696 & 0.058443 & 1.079252 & 0.054377 & 1.124592 & 0.055507 & 1.097017 & 0.054823\\
+		\hline
+		Baseline3 & 1.039525 & 0.053367 & 1.229901 & 0.058048 & 1.253978 & 0.058614 & 1.082712 & 0.054464\\
+		\hline
+		Baseline4 & 1.118929 & 0.055368 & 1.137825 & 0.055833 & 1.215893 & 0.057717 & 1.063300 & 0.053974\\
+		\hline
+		Baseline5 & 1.086093 & 0.054549 & 1.243553 & 0.058369 & 1.050861 & 0.053657 & 1.068459 & 0.054104\\
 		\hline
@@
 		Different Starting Populations, 
-		$\sigma_X = 10, \sigma_Y = 1$}
-	\begin{footnotesize}
-		\hskip -1.7cm
-		\begin{tabular}{l|r|r|r|r|r|r|r|r}
-			\hline
-			Data Set & Auto Est & Auto SE & AutoM Est & AutoM SE & Norm Est & Norm SE & NormM Est & NormM SE\\
-			\hline
-			$X_0 = 99999, Y_0 = 1$ & 1.284355 & 0.059319 & 1.000000 & 0.052342 & 1.255944 & 0.058660 & 1.129035 & 0.055617\\
-			\hline
-			$X_0 = 99990, Y_0 = 10$ & 0.745920 & 0.045206 & 1.000000 & 0.052342 & 1.000000 & 0.052342 & 1.241239 & 0.058315\\
-			\hline
-			$X_0 = 99900, Y_0 = 100$ & 0.716836 & 0.044316 & 1.236492 & 0.058204 & 1.064823 & 0.054012 & 1.000000 & 0.052342\\
-			\hline
-			$X_0 = 99000, Y_0 = 1000$ & 0.753867 & 0.045447 & 1.158019 & 0.056326 & 1.048705 & 0.053602 & 1.322988 & 0.060205\\
+		$\sigma_S = 10, \sigma_I = 1$}
+	\begin{footnotesize}
+		\hskip -1.7cm
+	\begin{tabular}{l|r|r|r|r|r|r|r|r}
+		\hline
+		Data Set & Auto Est & Auto SE & AutoM Est & AutoM SE & Norm Est & Norm SE & NormM Est & NormM SE\\
+		\hline
+		$S_0 = 99999, I_0 = 1$ & 0.929118 & 0.050453 & 1.000000 & 0.052342 & 0.842065 & 0.048032 & 1.216860 & 0.057740\\
+		\hline
+		$S_0 = 99990, I_0 = 10$ & 1.038604 & 0.053343 & 1.167800 & 0.056564 & 0.877679 & 0.049037 & 1.250160 & 0.058524\\
+		\hline
+		$S_0 = 99900, I_0 = 100$ & 1.253176 & 0.058595 & 1.069024 & 0.054119 & 1.165760 & 0.056514 & 1.077389 & 0.054330\\
+		\hline
+		$S_0 = 99000, I_0 = 1000$ & 0.975516 & 0.051698 & 0.970667 & 0.051569 & 0.947781 & 0.050957 & 0.969026 & 0.051525\\
 		\hline
@@
 	\caption{$\rr$ Estimates and Std. Errs, SEQBAYES Model, 
-		Different Variances, $X_0 = 99000$, $Y_0 = 1000$}
-	\begin{footnotesize}
-		\hskip -1.7cm
-		\begin{tabular}{l|r|r|r|r|r|r|r|r}
-			\hline
-			Data Set & Auto Est & Auto SE & AutoM Est & AutoM SE & Norm Est & Norm SE & NormM Est & NormM SE\\
-			\hline
-			$\sigma_X = 10, \sigma_Y = 1$ & 0.793499 & 0.046626 & 0.787881 & 0.046461 & 0.762945 & 0.045719 & 0.796512 & 0.046714\\
-			\hline
-			$\sigma_X = 100, \sigma_Y = 10$ & 0.925325 & 0.050350 & 0.891440 & 0.049420 & 0.888104 & 0.049327 & 0.885803 & 0.049263\\
-			\hline
-			$\sigma_X = 300, \sigma_Y = 30$ & 0.478432 & 0.036205 & 0.510298 & 0.037391 & 0.494531 & 0.036809 & 0.502460 & 0.037103\\
-			\hline
-			$\sigma_X = 1000, \sigma_Y = 100$ & 0.897133 & 0.049577 & 0.973510 & 0.051644 & 0.524686 & 0.037914 & 0.524018 & 0.037890\\
-			\hline
-			$\sigma_X = 3000, \sigma_Y = 300$ & 1.000000 & 0.052342 & 1.000000 & 0.052342 & 1.024154 & 0.052971 & 0.984151 & 0.051926\\
-			\hline
-			$\sigma_X = 10000, \sigma_Y = 1000$ & 1.686398 & 0.067973 & 1.000000 & 0.052342 & 0.379716 & 0.032254 & 1.000000 & 0.052342\\
+		Different Variances, $S_0 = 99000$, $I_0 = 1000$}
+	\begin{footnotesize}
+		\hskip -1.7cm
+	\begin{tabular}{l|r|r|r|r|r|r|r|r}
+		\hline
+		Error & Auto Est & Auto SE & AutoM Est & AutoM SE & Norm Est & Norm SE & NormM Est & NormM SE\\
+		\hline
+		$\sigma_S = 10, \sigma_I = 1$ & 0.984301 & 0.051930 & 0.990291 & 0.052088 & 0.967635 & 0.051488 & 0.968098 & 0.051501\\
+		\hline
+		$\sigma_S = 100, \sigma_I = 10$ & 1.014390 & 0.052718 & 1.270396 & 0.058996 & 1.000000 & 0.052342 & 1.000000 & 0.052342\\
+		\hline
+		$\sigma_S = 500, \sigma_I = 50$ & 1.033386 & 0.053209 & 1.322754 & 0.060200 & 1.000000 & 0.052342 & 1.000000 & 0.052342\\
+		\hline
+		$\sigma_S = 1000, \sigma_I = 100$ & 1.037446 & 0.053313 & 1.347989 & 0.060771 & 1.000000 & 0.052342 & 1.000000 & 0.052342\\
+		\hline
+		$\sigma_S = 2000, \sigma_I = 200$ & 1.039747 & 0.053372 & 1.373882 & 0.061352 & 1.000000 & 0.052342 & 1.000000 & 0.052342\\
+		\hline
+		$\sigma_S = 5000, \sigma_I = 500$ & 1.041231 & 0.053411 & 1.408605 & 0.062122 & 1.000000 & 0.052342 & 1.000000 & 0.052342\\
+		\hline
+		$\sigma_S = 7500, \sigma_I = 750$ & 1.041573 & 0.053419 & 1.424062 & 0.062462 & 1.000000 & 0.052342 & 1.000000 & 0.052342\\
+		\hline
+		$\sigma_S = 10000, \sigma_I = 1000$ & 1.041745 & 0.053424 & 1.435049 & 0.062703 & 1.000000 & 0.052342 & 1.000000 & 0.052342\\
+		\hline
+		$\sigma_S = 15000, \sigma_I = 1500$ & 1.041919 & 0.053428 & 1.450554 & 0.063041 & 1.000000 & 0.052342 & 1.000000 & 0.052342\\
 		\hline
@@
 	\caption{$\rr$ Estimates and Std. Errs, SEQBAYES Model,
-		Other Models, $X_0 = 99950, Y_0 = 50$, 
-		$\sigma_X = 10, \sigma_Y = 1$}
+		Other Models, $S_0 = 99950, I_0 = 50$, 
+		$\sigma_S = 10, \sigma_I = 1$}
 	\begin{footnotesize}
@@
 
-\subsubsection{SEIR Least Squares}
+\subsubsection{SEIR Least Sqares}
 
@@
 The estimates from the SEIR model are relatively consistent for the baseline data, but they are all slightly larger than the true $\rr$ value.
-The estimates tend to become more accurate as $Y_0$ increases relative to the total population.
+The estimates tend to become more accurate as $I_0$ increases relative to the total population.
 However, we can see that increasing the variance of the added errors to the simulated data dramatically decreases the accuracy and precision of the $\rr$ estimates.
@@
 They are fairly consistent between the multiple baseline set.
-The standard errors decrease dramatically as $Y_0$ increases.
+The standard errors decrease dramatically as $I_0$ increases.
 Also, as we see from the estimates above, the standard deviations fluctuate by several orders of magnitude as we change the variance of the added errors.
@@
 	\caption{$\rr$ Estimates and Std. Errs, SEIR Model, 
-		Baseline Data, $X_0 = 99950, Y_0 = 50$, 
-		$\sigma_X = 10, \sigma_Y = 1$}
-	\begin{footnotesize}
-		\hskip -1cm
-		\begin{tabular}{l|r|r|r|r|r|r|r|r}
-			\hline
-			Data Set & Auto Est & Auto SE & AutoM Est & AutoM SE & Norm Est & Norm SE & NormM Est & NormM SE\\
-			\hline
-			Baseline 1 & 1.546608 & 0.171494 & 1.326568 & 0.013100 & 3.206236 & 2.451373 & 2.455754 & 0.066867\\
-			\hline
-			Baseline 2 & 1.772431 & 0.354807 & -0.080128 & 0.031367 & -12.937346 & 150.847519 & 3.675647 & 0.123981\\
-			\hline
-			Baseline 3 & 1.781222 & 0.395661 & 1.375838 & 0.024409 & 2.482198 & 1.391814 & 2.149920 & 0.029209\\
-			\hline
-			Baseline 4 & 1.636096 & 0.594579 & 1.737715 & 0.020414 & 0.616769 & 9.237073 & 1.982802 & 0.025452\\
-			\hline
-			Baseline 5 & -10.920596 & 123.661450 & 1.846941 & 0.027824 & 4.892252 & 12.254994 & 2.457392 & 0.039093\\
+		Baseline Data, $S_0 = 99950, I_0 = 50$, 
+		$\sigma_S = 10, \sigma_I = 1$}
+	\begin{footnotesize}
+		\hskip -1cm
+	\begin{tabular}{l|r|r|r|r|r|r|r|r}
+		\hline
+		Data Set & Auto Est & Auto SE & AutoM Est & AutoM SE & Norm Est & Norm SE & NormM Est & NormM SE\\
+		\hline
+		Baseline1 & 1.302870 & 0.005985 & 1.304978 & 0.004832 & 1.295983 & 0.007049 & 1.289180 & 0.006367\\
+		\hline
+		Baseline2 & 1.291782 & 0.005861 & 1.303073 & 0.004532 & 1.293540 & 0.008068 & 1.285063 & 0.004522\\
+		\hline
+		Baseline3 & 1.291790 & 0.005322 & 1.305184 & 0.004968 & 1.313588 & 0.006870 & 1.295970 & 0.007001\\
+		\hline
+		Baseline4 & 1.305709 & 0.005793 & 1.303979 & 0.005127 & 1.307232 & 0.006962 & 1.313454 & 0.006151\\
+		\hline
+		Baseline5 & 1.302877 & 0.005851 & 1.297431 & 0.004872 & 1.302505 & 0.007885 & 1.297394 & 0.006744\\
 		\hline
@@
 		Different Starting Populations, 
-		$\sigma_X = 10, \sigma_Y = 1$}
-	\begin{footnotesize}
-		\hskip -1.7cm
-		\begin{tabular}{l|r|r|r|r|r|r|r|r}
-			\hline
-			Data Set & Auto Est & Auto SE & AutoM Est & AutoM SE & Norm Est & Norm SE & NormM Est & NormM SE\\
-			\hline
-			$X_0 = 99999, Y_0 = 1$ & NA & NA & -0.182830 & 0.225877 & NA & NA & -0.484025 & 0.614292\\
-			\hline
-			$X_0 = 99990, Y_0 = 10$ & 0.973745 & 0.546827 & NA & NA & 1.052351 & 444.645186 & NA & NA\\
-			\hline
-			$X_0 = 99900, Y_0 = 100$ & -6.416001 & 25.905717 & 2.451224 & 0.118072 & -3.124388 & 6.766299 & -1.475048 & 1.343083\\
-			\hline
-			$X_0 = 99000, Y_0 = 1000$ & -1.665131 & 1.139871 & 2.132087 & 0.009851 & 2.146884 & 0.018722 & 2.114676 & 0.010444\\
+		$\sigma_S = 10, \sigma_I = 1$}
+	\begin{footnotesize}
+		\hskip -1.7cm
+	\begin{tabular}{l|r|r|r|r|r|r|r|r}
+		\hline
+		Data Set & Auto Est & Auto SE & AutoM Est & AutoM SE & Norm Est & Norm SE & NormM Est & NormM SE\\
+		\hline
+		$S_0 = 99999, I_0 = 1$ & 1.327236 & 0.062503 & 1.305344 & 0.029386 & 1.319333 & 0.104531 & 1.486194 & 0.042930\\
+		\hline
+		$S_0 = 99990, I_0 = 10$ & 1.320088 & 0.014535 & 1.327951 & 0.010927 & 1.323049 & 0.019898 & 1.330877 & 0.012021\\
+		\hline
+		$S_0 = 99900, I_0 = 100$ & 1.257056 & 0.001195 & 1.252233 & 0.001181 & 1.269212 & 0.001844 & 1.196636 & 0.000635\\
+		\hline
+		$S_0 = 99000, I_0 = 1000$ & 1.198724 & 0.000024 & 1.198618 & 0.000026 & 1.198275 & 0.000036 & 1.198736 & 0.000036\\
 		\hline
@@
 	\caption{$\rr$ Estimates and Std. Errs, SEIR Model, 
-		Different Variances, $X_0 = 99000$, $Y_0 = 1000$}
-	\begin{footnotesize}
-		\hskip -1.7cm
-		\begin{tabular}{l|r|r|r|r|r|r|r|r}
-			\hline
-			Data Set & Auto Est & Auto SE & AutoM Est & AutoM SE & Norm Est & Norm SE & NormM Est & NormM SE\\
-			\hline
-			$\sigma_X = 10, \sigma_Y = 1$ & 19.226063 & 0.289480 & 21.685544 & 0.346537 & 18.468777 & 0.256194 & 23.787444 & 0.195589\\
-			\hline
-			$\sigma_X = 100, \sigma_Y = 10$ & 16.976120 & 0.530197 & 19.896647 & 0.226174 & 13.179504 & 0.570207 & 15.331206 & 0.206823\\
-			\hline
-			$\sigma_X = 300, \sigma_Y = 30$ & 9.925354 & 0.770734 & 16.239531 & 0.234976 & -4.473057 & 0.190822 & 22.862570 & 0.525209\\
-			\hline
-			$\sigma_X = 1000, \sigma_Y = 100$ & -0.007144 & 0.000245 & 25.560238 & 0.933150 & -4.082639 & 0.436790 & 11.261214 & 0.216712\\
-			\hline
-			$\sigma_X = 3000, \sigma_Y = 300$ & -0.262553 & 0.003032 & -0.012345 & 0.000093 & -0.000062 & 0.000004 & 12.886121 & 0.277206\\
-			\hline
-			$\sigma_X = 10000, \sigma_Y = 1000$ & -0.000217 & 0.000031 & -5.959291 & 0.225965 & NA & NA & -0.000064 & 0.000000\\
+		Different Variances, $S_0 = 99000$, $I_0 = 1000$}
+	\begin{footnotesize}
+		\hskip -1.7cm
+	\begin{tabular}{l|r|r|r|r|r|r|r|r}
+		\hline
+		Error & Auto Est & Auto SE & AutoM Est & AutoM SE & Norm Est & Norm SE & NormM Est & NormM SE\\
+		\hline
+		$\sigma_S = 10, \sigma_I = 1$ & 1.198667 & 0.022446 & 1.199697 & 0.000054 & 1.184111 & 0.000033 & 1.215409 & 0.003919\\
+		\hline
+		$\sigma_S = 100, \sigma_I = 10$ & 0.738349 & 0.027172 & 0.326262 & 0.148088 & 0.876058 & 0.009180 & 1.261285 & 0.000093\\
+		\hline
+		$\sigma_S = 500, \sigma_I = 50$ & -0.003532 & 0.056577 & -0.070816 & 0.504898 & 0.094900 & 0.691400 & 1.326638 & 0.030156\\
+		\hline
+		$\sigma_S = 1000, \sigma_I = 100$ & 1.652869 & 0.287473 & -0.261653 & 2.358679 & NA & NA & 0.951134 & 0.007148\\
+		\hline
+		$\sigma_S = 2000, \sigma_I = 200$ & -0.000235 & 0.079270 & -0.043704 & 0.553296 & 9.411580 & 0.032052 & 1.142685 & 0.017086\\
+		\hline
+		$\sigma_S = 5000, \sigma_I = 500$ & 0.014666 & 0.124467 & 0.790375 & 0.015295 & -0.014291 & 0.124025 & 9.473186 & 33.693741\\
+		\hline
+		$\sigma_S = 7500, \sigma_I = 750$ & 2.856424 & 11.133371 & 0.808988 & 0.038110 & -0.007135 & 0.109857 & NA & NA\\
+		\hline
+		$\sigma_S = 10000, \sigma_I = 1000$ & 2.755506 & 5.698171 & 2250.629540 & 62419.537578 & 0.016792 & 0.294779 & NA & NA\\
+		\hline
+		$\sigma_S = 15000, \sigma_I = 1500$ & NA & NA & 0.845936 & 0.056621 & 0.020971 & 0.117313 & 0.703029 & 0.145427\\
 		\hline
@@
 	\caption{$\rr$ Estimates and Std. Errs, SEIR Model,
-		Other Models, $X_0 = 99950, Y_0 = 50$, 
-		$\sigma_X = 10, \sigma_Y = 1$}
+		Other Models, $S_0 = 99950, I_0 = 50$, 
+		$\sigma_S = 10, \sigma_I = 1$}
 	\begin{footnotesize}
@@
 The estimates of $\rr$ are very precise under the same conditions for the classic SIR model.
-The change in $\sigma_X$ and $\sigma_Y$ do not change the estimates much.
-They do differ somewhat largely as $Y_0$ changes.
+The change in $\sigma_S$ and $\sigma_I$ do not change the estimates much.
+They do differ somewhat largely as $I_0$ changes.
 The estimates from the data generated from other models are actually consistent and reasonable as well.
@@
 	\caption{$\rr$ Estimates and Std. Errs, COMP Model, 
-		Baseline Data, $X_0 = 99950, Y_0 = 50$, 
-		$\sigma_X = 10, \sigma_Y = 1$}
-	\begin{footnotesize}
-		\hskip -1cm
-		\begin{tabular}{l|r|r|r|r|r|r|r|r}
-			\hline
-			Data Set & Auto Est & Auto SE & AutoM Est & AutoM SE & Norm Est & Norm SE & NormM Est & NormM SE\\
-			\hline
-			Baseline 1 & 1.151396 & 0.007840 & 1.156166 & 0.075838 & 1.168638 & 0.061978 & 1.168638 & 0.069916\\
-			\hline
-			Baseline 2 & 1.151396 & 0.015427 & 1.151396 & 0.015825 & 1.168638 & 0.068289 & 1.168638 & 0.056477\\
-			\hline
-			Baseline 3 & 1.155559 & 0.127438 & 1.151396 & 0.012891 & 1.168638 & 0.068609 & 1.168638 & 0.052676\\
-			\hline
-			Baseline 4 & 1.156588 & 0.049150 & 1.154959 & 0.026360 & 1.168638 & 0.052003 & 1.126363 & 0.038690\\
-			\hline
-			Baseline 5 & 1.149516 & 0.024329 & 1.151396 & 0.014140 & 1.168638 & 0.060267 & 1.149516 & 0.029827\\
+		Baseline Data, $S_0 = 99950, I_0 = 50$, 
+		$\sigma_S = 10, \sigma_I = 1$}
+	\begin{footnotesize}
+		\hskip -1cm
+	\begin{tabular}{l|r|r|r|r|r|r|r|r}
+		\hline
+		Data Set & Auto Est & Auto SE & AutoM Est & AutoM SE & Norm Est & Norm SE & NormM Est & NormM SE\\
+		\hline
+		Baseline1 & 1.412982 & 0.010767 & 1.412982 & 0.009669 & 1.412982 & 0.010947 & 1.412982 & 0.010842\\
+		\hline
+		Baseline2 & 1.412982 & 0.010613 & 1.412982 & 0.010721 & 1.412982 & 0.011393 & 1.412982 & 0.010027\\
+		\hline
+		Baseline3 & 1.412982 & 0.010823 & 1.412982 & 0.009758 & 1.412982 & 0.010201 & 1.412982 & 0.009806\\
+		\hline
+		Baseline4 & 1.412982 & 0.011042 & 1.412982 & 0.009580 & 1.412982 & 0.010452 & 1.412982 & 0.011354\\
+		\hline
+		Baseline5 & 1.412982 & 0.011579 & 1.412982 & 0.009644 & 1.412982 & 0.010383 & 1.412982 & 0.009949\\
 		\hline
@@
 		Different Starting Populations, 
-		$\sigma_X = 10, \sigma_Y = 1$}
-	\begin{footnotesize}
-		\hskip -1.7cm
-		\begin{tabular}{l|r|r|r|r|r|r|r|r}
-			\hline
-			Data Set & Auto Est & Auto SE & AutoM Est & AutoM SE & Norm Est & Norm SE & NormM Est & NormM SE\\
-			\hline
-			$X_0 = 99999, Y_0 = 1$ & 0.205516 & NA & 0.220109 & NA & 0.203940 & NA & 0.220109 & NA\\
-			\hline
-			$X_0 = 99990, Y_0 = 10$ & 1.049647 & 0.040841 & 1.043576 & 0.109889 & 1.049647 & 0.042400 & 1.049647 & 0.070514\\
-			\hline
-			$X_0 = 99900, Y_0 = 100$ & 1.215624 & 0.008808 & 1.163593 & 0.005915 & 1.163593 & 0.004472 & 1.163593 & 0.007400\\
-			\hline
-			$X_0 = 99000, Y_0 = 1000$ & 1.153396 & 0.000940 & 1.143722 & 0.000599 & 1.103664 & 0.000472 & 1.153396 & 0.000970\\
+		$\sigma_S = 10, \sigma_I = 1$}
+	\begin{footnotesize}
+		\hskip -1.7cm
+	\begin{tabular}{l|r|r|r|r|r|r|r|r}
+		\hline
+		Data Set & Auto Est & Auto SE & AutoM Est & AutoM SE & Norm Est & Norm SE & NormM Est & NormM SE\\
+		\hline
+		$S_0 = 99999, I_0 = 1$ & 0.600000 & NA & 0.600000 & NA & 0.600000 & NA & 0.600000 & NA\\
+		\hline
+		$S_0 = 99990, I_0 = 10$ & 1.417811 & 0.009877 & 1.428718 & 0.030051 & 1.428718 & 0.025660 & 1.428718 & 0.027867\\
+		\hline
+		$S_0 = 99900, I_0 = 100$ & 1.360667 & 0.003552 & 1.360667 & 0.004453 & 1.382555 & 0.003572 & 1.382555 & 0.003212\\
+		\hline
+		$S_0 = 99000, I_0 = 1000$ & 1.265725 & 0.002888 & 1.265725 & 0.002807 & 1.265725 & 0.002835 & 1.265725 & 0.003064\\
 		\hline
@@
 	\caption{$\rr$ Estimates and Std. Errs, COMP Model, 
-		Different Variances, $X_0 = 99000$, $Y_0 = 1000$}
-	\begin{footnotesize}
-		\hskip -1.7cm
-		\begin{tabular}{l|r|r|r|r|r|r|r|r}
-			\hline
-			Data Set & Auto Est & Auto SE & AutoM Est & AutoM SE & Norm Est & Norm SE & NormM Est & NormM SE\\
-			\hline
-			$\sigma_X = 10, \sigma_Y = 1$ & 1.143722 & 0.000584 & 1.147266 & 0.002023 & 1.132778 & 0.000684 & 1.132778 & 0.000696\\
-			\hline
-			$\sigma_X = 100, \sigma_Y = 10$ & 1.140700 & 0.000566 & 1.171033 & 0.000785 & 1.155093 & 0.000925 & 1.183014 & 0.001369\\
-			\hline
-			$\sigma_X = 300, \sigma_Y = 30$ & 1.174448 & 0.001166 & 1.137020 & 0.002796 & 1.236429 & 0.003236 & 1.122842 & 0.001041\\
-			\hline
-			$\sigma_X = 1000, \sigma_Y = 100$ & 1.142865 & 0.000803 & 1.154025 & 0.001744 & 1.205823 & 0.003666 & 1.399630 & 0.002187\\
-			\hline
-			$\sigma_X = 3000, \sigma_Y = 300$ & 1.142642 & 0.001985 & 1.140222 & 0.000500 & 1.150901 & 0.000913 & 1.140190 & 0.000416\\
-			\hline
-			$\sigma_X = 10000, \sigma_Y = 1000$ & 1.198172 & NA & 1.147908 & 0.000934 & 1.162161 & 0.001444 & 1.157407 & 0.000959\\
+		Different Variances, $S_0 = 99000$, $I_0 = 1000$}
+	\begin{footnotesize}
+		\hskip -1.7cm
+	\begin{tabular}{l|r|r|r|r|r|r|r|r}
+		\hline
+		Error & Auto Est & Auto SE & AutoM Est & AutoM SE & Norm Est & Norm SE & NormM Est & NormM SE\\
+		\hline
+		$\sigma_S = 10, \sigma_I = 1$ & 1.265725 & 0.002680 & 1.265725 & 0.003044 & 1.265725 & 0.003271 & 1.265725 & 0.002967\\
+		\hline
+		$\sigma_S = 100, \sigma_I = 10$ & 1.265725 & 0.003122 & 1.232704 & 0.025432 & 1.265725 & 0.002878 & 1.264345 & 0.003288\\
+		\hline
+		$\sigma_S = 500, \sigma_I = 50$ & 1.263026 & 0.006370 & 1.269345 & 0.004386 & 1.258118 & 0.007414 & 1.255344 & 0.006144\\
+		\hline
+		$\sigma_S = 1000, \sigma_I = 100$ & 1.234715 & 0.006079 & 1.273213 & 0.000652 & 1.258118 & 0.010512 & 1.260004 & 0.006172\\
+		\hline
+		$\sigma_S = 2000, \sigma_I = 200$ & 1.236510 & 0.010850 & 1.246699 & 0.024112 & 1.233704 & 0.011063 & 1.219023 & 0.018803\\
+		\hline
+		$\sigma_S = 5000, \sigma_I = 500$ & 1.285391 & 0.002340 & 1.288921 & 0.002671 & 1.222575 & 0.030681 & 1.237872 & 0.017728\\
+		\hline
+		$\sigma_S = 7500, \sigma_I = 750$ & 1.267481 & 0.001248 & 1.265853 & 0.001334 & 1.229195 & 0.004376 & 1.208398 & 0.003815\\
+		\hline
+		$\sigma_S = 10000, \sigma_I = 1000$ & 1.304435 & 0.001567 & 1.298856 & 0.003194 & 1.246699 & 0.006051 & 1.224308 & 0.001915\\
+		\hline
+		$\sigma_S = 15000, \sigma_I = 1500$ & 1.334921 & 0.000388 & 1.319515 & 0.001724 & 1.228979 & 0.002101 & 1.188725 & 0.007633\\
 		\hline
@@
 	\caption{$\rr$ Estimates and Std. Errs, COMP Model,
-		Other Models, $X_0 = 99950, Y_0 = 50$, 
-		$\sigma_X = 10, \sigma_Y = 1$}
+		Other Models, $S_0 = 99950, I_0 = 50$, 
+		$\sigma_S = 10, \sigma_I = 1$}
 	\begin{footnotesize}
@@
 The estimates do not fluctuate as much as those from the SEIR least squares model either.
-The $\rr$ estimates also become less accurate for large values of $\sigma_X$ and $\sigma_Y$.
+The $\rr$ estimates also become less accurate for large values of $\sigma_S$ and $\sigma_I$.
 Looking at the data generated from other models, the next generation does not seem to work well for the 0th and 1st order data, where the number of people in the S and aI compartment change linearly rather than exponentially.
@@
 The standard errors are approximately of the same order of magnitude as those from the SEIR least squares model.
-Also, like for the SEIR least squares model, they are fairly consistent between the multiple baseline set, the standard errors decrease dramatically as $Y_0$ increases, and the standard deviations fluctuate by several orders of magnitude as we change the variance of the added errors.
+Also, like for the SEIR least squares model, they are fairly consistent between the multiple baseline set, the standard errors decrease dramatically as $I_0$ increases, and the standard deviations fluctuate by several orders of magnitude as we change the variance of the added errors.
 The most extreme values of standard errors that we see here are not as unreasonably gigantic as for the SEIR least squares model either.
@@
 	\caption{$\rr$ Estimates and Std. Errs, NEXTGEN Model, 
-		Baseline Data, $X_0 = 99950, Y_0 = 50$, 
-		$\sigma_X = 10, \sigma_Y = 1$}
-	\begin{footnotesize}
-		\hskip -1cm
-		\begin{tabular}{l|r|r|r|r|r|r|r|r}
-			\hline
-			Data Set & Auto Est & Auto SE & AutoM Est & AutoM SE & Norm Est & Norm SE & NormM Est & NormM SE\\
-			\hline
-			Baseline 1 & 1.421896 & 0.110167 & 1.246864 & 0.010846 & 1.715318 & 0.798366 & 1.781744 & 0.024938\\
-			\hline
-			Baseline 2 & 1.758858 & 0.366701 & 5.512315 & 0.320605 & 1.963196 & 0.846297 & 1.625233 & 0.018011\\
-			\hline
-			Baseline 3 & 1.758858 & 0.569505 & 1.421896 & 0.016966 & 2.297508 & 2.211150 & 1.858192 & 0.036113\\
-			\hline
-			Baseline 4 & 1.447204 & 0.091904 & 1.712243 & 0.020892 & 1.421896 & 0.341627 & 2.172764 & 0.038152\\
-			\hline
-			Baseline 5 & 1.421896 & 0.109030 & 2.162264 & 0.034685 & 1.625233 & 0.626916 & 2.011729 & 0.027506\\
+		Baseline Data, $S_0 = 99950, I_0 = 50$, 
+		$\sigma_S = 10, \sigma_I = 1$}
+	\begin{footnotesize}
+		\hskip -1cm
+	\begin{tabular}{l|r|r|r|r|r|r|r|r}
+		\hline
+		Data Set & Auto Est & Auto SE & AutoM Est & AutoM SE & Norm Est & Norm SE & NormM Est & NormM SE\\
+		\hline
+		Baseline1 & 1.4 & 0.008096 & 1.4 & 0.005942 & 1.4 & 0.010666 & 1.4 & 0.007859\\
+		\hline
+		Baseline2 & 1.4 & 0.007525 & 1.4 & 0.006043 & 1.4 & 0.009713 & 1.4 & 0.006573\\
+		\hline
+		Baseline3 & 1.4 & 0.006891 & 1.4 & 0.006580 & 1.4 & 0.008550 & 1.4 & 0.007550\\
+		\hline
+		Baseline4 & 1.4 & 0.006989 & 1.4 & 0.005960 & 1.4 & 0.009015 & 1.4 & 0.006935\\
+		\hline
+		Baseline5 & 1.4 & 0.006809 & 1.4 & 0.006459 & 1.4 & 0.009730 & 1.4 & 0.007778\\
 		\hline
@@
 		Different Starting Populations, 
-		$\sigma_X = 10, \sigma_Y = 1$}
-	\begin{footnotesize}
-		\hskip -1.7cm
-		\begin{tabular}{l|r|r|r|r|r|r|r|r}
-			\hline
-			Data Set & Auto Est & Auto SE & AutoM Est & AutoM SE & Norm Est & Norm SE & NormM Est & NormM SE\\
-			\hline
-			$X_0 = 99999, Y_0 = 1$ & 575.476314 & 5450.536541 & 205.872462 & 2972.609438 & 16.358566 & 9841.298666 & 34.387930 & 5805.567900\\
-			\hline
-			$X_0 = 99990, Y_0 = 10$ & 1.383532 & 25.019898 & 206.028474 & 18394.966469 & 3.214261 & 12206.722105 & 741.368136 & 15921.351448\\
-			\hline
-			$X_0 = 99900, Y_0 = 100$ & 4.757366 & 2.846210 & 3.819417 & 0.635151 & 15.683199 & 137.249782 & 3.874259 & 0.588683\\
-			\hline
-			$X_0 = 99000, Y_0 = 1000$ & 4.072926 & 0.317373 & 4.130517 & 0.226424 & 4.145316 & 0.359158 & 4.012320 & 0.239506\\
+		$\sigma_S = 10, \sigma_I = 1$}
+	\begin{footnotesize}
+		\hskip -1.7cm
+	\begin{tabular}{l|r|r|r|r|r|r|r|r}
+		\hline
+		Data Set & Auto Est & Auto SE & AutoM Est & AutoM SE & Norm Est & Norm SE & NormM Est & NormM SE\\
+		\hline
+		$S_0 = 99999, I_0 = 1$ & 1.400000 & 0.085421 & 1.400000 & 0.041869 & 1.400000 & 0.159127 & 1.437675 & 0.030005\\
+		\hline
+		$S_0 = 99990, I_0 = 10$ & 1.400000 & 0.020310 & 1.400000 & 0.012602 & 1.400000 & 0.023171 & 1.400000 & 0.015834\\
+		\hline
+		$S_0 = 99900, I_0 = 100$ & 1.400000 & 0.004333 & 1.400000 & 0.004476 & 1.400000 & 0.005505 & 1.400000 & 0.004161\\
+		\hline
+		$S_0 = 99000, I_0 = 1000$ & 1.200159 & 0.000029 & 1.198359 & 0.000025 & 1.198359 & 0.000028 & 1.200159 & 0.000031\\
 		\hline
@@
 	\caption{$\rr$ Estimates and Std. Errs, NEXTGEN Model, 
-		Different Variances, $X_0 = 99000$, $Y_0 = 1000$}
-	\begin{footnotesize}
-		\hskip -1.7cm
-		\begin{tabular}{l|r|r|r|r|r|r|r|r}
-			\hline
-			Data Set & Auto Est & Auto SE & AutoM Est & AutoM SE & Norm Est & Norm SE & NormM Est & NormM SE\\
-			\hline
-			$\sigma_X = 10, \sigma_Y = 1$ & 12.998364 & 8.755960e-01 & 8.061987 & 0.486050 & 9.400660 & 6.204380e-01 & 5.752931 & 0.261811\\
-			\hline
-			$\sigma_X = 100, \sigma_Y = 10$ & 5.404625 & 5.198610e-01 & 7.387983 & 0.436634 & 8.531874 & 1.705313e+00 & 5.978214 & 1.108848\\
-			\hline
-			$\sigma_X = 300, \sigma_Y = 30$ & 23.186330 & 6.441964e+02 & 5.963903 & 1.266200 & 322.144190 & 2.088921e+05 & 7.276865 & 0.663915\\
-			\hline
-			$\sigma_X = 1000, \sigma_Y = 100$ & 23.983516 & 1.278927e+04 & 67.517641 & 52.327269 & 11.529412 & 4.001174e+02 & 7.010464 & 0.567235\\
-			\hline
-			$\sigma_X = 3000, \sigma_Y = 300$ & 33.846154 & 1.163559e+07 & 21.915757 & 1527.656692 & 162.022472 & 4.124435e+03 & 12.140594 & 4.474694\\
-			\hline
-			$\sigma_X = 10000, \sigma_Y = 1000$ & 174.921348 & 3.898656e+03 & 11.524763 & 165.981987 & 1.025641 & 2.676852e+02 & 101.413031 & 1760.215196\\
+		Different Variances, $S_0 = 99000$, $I_0 = 1000$}
+	\begin{footnotesize}
+		\hskip -1.7cm
+	\begin{tabular}{l|r|r|r|r|r|r|r|r}
+		\hline
+		Error & Auto Est & Auto SE & AutoM Est & AutoM SE & Norm Est & Norm SE & NormM Est & NormM SE\\
+		\hline
+		$\sigma_S = 10, \sigma_I = 1$ & 1.213314 & 0.000028 & 1.212138 & 0.000030 & 1.213314 & 0.000034 & 1.213314 & 0.000031\\
+		\hline
+		$\sigma_S = 100, \sigma_I = 10$ & 1.213314 & 0.000231 & 1.204838 & 0.000240 & 1.213314 & 0.000333 & 1.213314 & 0.000345\\
+		\hline
+		$\sigma_S = 500, \sigma_I = 50$ & 1.213314 & 0.001189 & 1.204838 & 0.001234 & 1.179108 & 0.002048 & 1.213314 & 0.001537\\
+		\hline
+		$\sigma_S = 1000, \sigma_I = 100$ & 1.198952 & 0.005372 & 1.207589 & 0.005635 & 1.198952 & 0.007195 & 1.198952 & 0.006684\\
+		\hline
+		$\sigma_S = 2000, \sigma_I = 200$ & 1.237684 & 0.006507 & 1.249285 & 0.025702 & 1.215487 & 0.007941 & 1.198952 & 0.024341\\
+		\hline
+		$\sigma_S = 5000, \sigma_I = 500$ & 1.210888 & 0.039465 & 0.772582 & 0.018292 & 0.002250 & 4.601591 & 1.167461 & 0.382569\\
+		\hline
+		$\sigma_S = 7500, \sigma_I = 750$ & 1.191708 & 0.326848 & 0.820021 & 0.011286 & 0.002250 & 0.100353 & NA & NA\\
+		\hline
+		$\sigma_S = 10000, \sigma_I = 1000$ & 3.356645 & 0.725374 & 0.836387 & 0.011665 & 0.002250 & 1.020673 & NA & NA\\
+		\hline
+		$\sigma_S = 15000, \sigma_I = 1500$ & 2.072075 & 21.625905 & 0.655863 & 0.052597 & 0.002250 & 0.688567 & NA & NA\\
 		\hline
@@
 	\caption{$\rr$ Estimates and Std. Errs, NEXTGEN Model,
-		Other Models, $X_0 = 99950, Y_0 = 50$, 
-		$\sigma_X = 10, \sigma_Y = 1$}
-	\begin{footnotesize}
-		\hskip -1cm
-		\begin{tabular}{l|r|r|r|r|r|r|r|r}
-			\hline
-			Data Set & Auto Est & Auto SE & AutoM Est & AutoM SE & Norm Est & Norm SE & NormM Est & NormM SE\\
-			\hline
-			0th & 0.001360 & 0.497415 & 0.000040 & 0.004153 & 0.037223 & 146.510831 & 0.000040 & 0.010014\\
-			\hline
-			1st & 6.217726 & 78.702007 & 13.734754 & 240.363417 & 111.232168 & 4202.221696 & 8.742980 & 66.867285\\
-			\hline
-			2nd & 2.451548 & 0.016035 & 5.220397 & 0.437529 & 8.020012 & 2.063975 & 5.220397 & 0.475516\\
-			\hline
-			3rd & 5.406024 & 0.554855 & 4.833617 & 0.256668 & 7.333991 & 1.363208 & 5.536633 & 0.660877\\
-			\hline
-			4th & 9.415812 & 2.964507 & 9.415812 & 2.964507 & 5.964387 & 0.856505 & 5.964387 & 0.748410\\
-			\hline
-			Linear SIR & 14.488463 & 6.089165 & 14.488463 & 5.828112 & 14.488463 & 6.513183 & 5.383303 & 0.308536\\
+		Other Models, $S_0 = 99950, I_0 = 50$, 
+		$\sigma_S = 10, \sigma_I = 1$}
+	\begin{footnotesize}
+		\hskip -1cm
+	\begin{tabular}{l|r|r|r|r|r|r|r|r}
+		\hline
+		Data Set & Auto Est & Auto SE & AutoM Est & AutoM SE & Norm Est & Norm SE & NormM Est & NormM SE\\
+		\hline
+		0th & 0.000095 & 0.009102 & 0.000082 & 0.000173 & 0.055180 & 0.017657 & 0.000023 & 0.000389\\
+		\hline
+		1st & 1.318348 & 258.153117 & 1.318348 & 5.918045 & 1.318348 & 0.628775 & 1.318348 & 8.692243\\
+		\hline
+		2nd & 1.400000 & 0.009256 & 1.400000 & 0.005901 & 1.400000 & 0.009429 & 1.400000 & 0.006330\\
+		\hline
+		3rd & 1.400000 & 0.006097 & 1.400000 & 0.006041 & 1.400000 & 0.007921 & 1.400000 & 0.007604\\
+		\hline
+		4th & 1.400000 & 0.007004 & 1.400000 & 0.006878 & 1.400000 & 0.009211 & 1.400000 & 0.008413\\
+		\hline
+		Linear SIR & 1.483791 & 0.007043 & 1.483791 & 0.006561 & 1.483791 & 0.008190 & 1.483791 & 0.007156\\
 		\hline
@@
 In particular, the estimates are almost all under 1 when the true $\rr$ is above 1, meaning we would predict that there is no epidemic when there actually is.
-It seems that $\rr$ increases as $Y_0$ increases as well as when $\sigma_X$ and $\sigma_Y$ increase.
+It seems that $\rr$ increases as $I_0$ increases as well as when $\sigma_S$ and $\sigma_I$ increase.
 It is unclear whether this is because the model becomes more accurate or if it just gets larger in general.
@@
 
-The standard errors are all fairly small, among the order of $10^{-3}$. They seem to decrease as $Y_0$ increases as well as when $\sigma_X$ and $\sigma_Y$ increases.
+The standard errors are all fairly small, among the order of $10^{-3}$. They seem to decrease as $I_0$ increases as well as when $\sigma_S$ and $\sigma_I$ increases.
 
@@
 	\caption{$\rr$ Estimates and Std. Errs, EXPGROWTH Model, 
-		Baseline Data, $X_0 = 99950, Y_0 = 50$, 
-		$\sigma_X = 10, \sigma_Y = 1$}
-	\begin{footnotesize}
-		\hskip -1cm
-		\begin{tabular}{l|r|r|r|r|r|r|r|r}
-			\hline
-			Data Set & Auto Est & Auto SE & AutoM Est & AutoM SE & Norm Est & Norm SE & NormM Est & NormM SE\\
-			\hline
-			Baseline 1 & 1.012007 & 0.010027 & 0.929414 & 0.030175 & 1.018400 & 0.009172 & 0.935858 & 0.027818\\
-			\hline
-			Baseline 2 & 1.010507 & 0.010239 & 0.943961 & 0.025078 & 1.018896 & 0.009109 & 0.936850 & 0.027469\\
-			\hline
-			Baseline 3 & 1.006199 & 0.010870 & 0.930872 & 0.029627 & 1.016426 & 0.009428 & 0.934886 & 0.028163\\
-			\hline
-			Baseline 4 & 1.006496 & 0.010825 & 0.934524 & 0.028292 & 1.018095 & 0.009211 & 0.934295 & 0.028375\\
-			\hline
-			Baseline 5 & 1.009068 & 0.010445 & 0.934444 & 0.028321 & 1.015362 & 0.009569 & 0.935845 & 0.027819\\
+		Baseline Data, $S_0 = 99950, I_0 = 50$, 
+		$\sigma_S = 10, \sigma_I = 1$}
+	\begin{footnotesize}
+		\hskip -1cm
+	\begin{tabular}{l|r|r|r|r|r|r|r|r}
+		\hline
+		Data Set & Auto Est & Auto SE & AutoM Est & AutoM SE & Norm Est & Norm SE & NormM Est & NormM SE\\
+		\hline
+		Baseline 1 & 0.928417 & 0.001317 & 0.928356 & 0.001319 & 0.928626 & 0.001310 & 0.928228 & 0.001323\\
+		\hline
+		Baseline 2 & 0.928419 & 0.001317 & 0.928247 & 0.001323 & 0.928594 & 0.001311 & 0.928289 & 0.001321\\
+		\hline
+		Baseline 3 & 0.928401 & 0.001318 & 0.928163 & 0.001326 & 0.928614 & 0.001310 & 0.928290 & 0.001321\\
+		\hline
+		Baseline 4 & 0.928395 & 0.001318 & 0.928248 & 0.001323 & 0.928611 & 0.001310 & 0.928310 & 0.001321\\
+		\hline
+		Baseline 5 & 0.928412 & 0.001317 & 0.928241 & 0.001323 & 0.928651 & 0.001309 & 0.928247 & 0.001323\\
 		\hline
@@
 		Different Starting Populations, 
-		$\sigma_X = 10, \sigma_Y = 1$}
-	\begin{footnotesize}
-		\hskip -1.7cm
-		\begin{tabular}{l|r|r|r|r|r|r|r|r}
-			Data Set & Auto Est & Auto SE & AutoM Est & AutoM SE & Norm Est & Norm SE & NormM Est & NormM SE\\
-			\hline
-			$X_0 = 99999, Y_0 = 1$ & 1.004407 & 0.011143 & 0.767095 & 0.146339 & 1.016822 & 0.009376 & 0.757875 & 0.155300\\
-			\hline
-			$X_0 = 99990, Y_0 = 10$ & 1.008824 & 0.010481 & 0.885239 & 0.051071 & 1.015486 & 0.009553 & 0.877557 & 0.055606\\
-			\hline
-			$X_0 = 99900, Y_0 = 100$ & 1.012554 & 0.009951 & 0.961057 & 0.020060 & 1.019685 & 0.009009 & 0.961228 & 0.020014\\
-			\hline
-			$X_0 = 99000, Y_0 = 1000$ & 1.044120 & 0.006387 & 1.042065 & 0.006576 & 1.045714 & 0.006245 & 1.042041 & 0.006578\\
+		$\sigma_S = 10, \sigma_I = 1$}
+	\begin{footnotesize}
+		\hskip -1.7cm
+	\begin{tabular}{l|r|r|r|r|r|r|r|r}
+		\hline
+		Data Set & Auto Est & Auto SE & AutoM Est & AutoM SE & Norm Est & Norm SE & NormM Est & NormM SE\\
+		\hline
+		$S_0 = 99999, I_0 = 1$ & 0.889241 & 0.003555 & 0.854729 & 0.008135 & 0.895513 & 0.003041 & 0.857664 & 0.007601\\
+		\hline
+		$S_0 = 99990, I_0 = 10$ & 0.905684 & 0.002355 & 0.902737 & 0.002537 & 0.908169 & 0.002211 & 0.902592 & 0.002546\\
+		\hline
+		$S_0 = 99900, I_0 = 100$ & 0.936306 & 0.001074 & 0.936309 & 0.001074 & 0.936398 & 0.001072 & 0.936345 & 0.001073\\
+		\hline
+		$S_0 = 99000, I_0 = 1000$ & 0.944202 & 0.000875 & 0.944200 & 0.000875 & 0.944184 & 0.000876 & 0.944210 & 0.000875\\
 		\hline
@@
 	\caption{$\rr$ Estimates and Std. Errs, EXPGROWTH Model, 
-		Different Variances, $X_0 = 99000$, $Y_0 = 1000$}
-	\begin{footnotesize}
-		\hskip -1.7cm
-		\begin{tabular}{l|r|r|r|r|r|r|r|r}
-			\hline
-			Data Set & Auto Est & Auto SE & AutoM Est & AutoM SE & Norm Est & Norm SE & NormM Est & NormM SE\\
-			\hline
-			$\sigma_X = 10, \sigma_Y = 1$ & 1.044244 & 0.006376 & 1.041733 & 0.006606 & 1.047118 & 0.006122 & 1.042010 & 0.006580\\
-			\hline
-			$\sigma_X = 100, \sigma_Y = 10$ & 1.088170 & 0.003410 & 1.043282 & 0.006463 & 1.099191 & 0.002912 & 1.044029 & 0.006395\\
-			\hline
-			$\sigma_X = 300, \sigma_Y = 30$ & 1.123132 & 0.002065 & 1.049199 & 0.005944 & 1.130609 & 0.001855 & 1.041324 & 0.006645\\
-			\hline
-			$\sigma_X = 1000, \sigma_Y = 100$ & 1.161414 & 0.001192 & 1.041609 & 0.006618 & 1.171423 & 0.001033 & 1.040663 & 0.006707\\
-			\hline
-			$\sigma_X = 3000, \sigma_Y = 300$ & 1.194876 & 0.000738 & 1.077426 & 0.003976 & 1.210073 & 0.000594 & 1.056925 & 0.005326\\
-			\hline
-			$\sigma_X = 10000, \sigma_Y = 1000$ & 1.237467 & 0.000402 & 1.092848 & 0.003189 & 1.246201 & 0.000355 & 1.093502 & 0.003159\\
+		Different Variances, $S_0 = 99000$, $I_0 = 1000$}
+	\begin{footnotesize}
+		\hskip -1.7cm
+	\begin{tabular}{l|r|r|r|r|r|r|r|r}
+		\hline
+		Data Set & Auto Est & Auto SE & AutoM Est & AutoM SE & Norm Est & Norm SE & NormM Est & NormM SE\\
+		\hline
+		$\sigma_S = 10, \sigma_I = 1$ & 0.944214 & 0.000875 & 0.944234 & 0.000875 & 0.944177 & 0.000876 & 0.944203 & 0.000875\\
+		\hline
+		$\sigma_S = 100, \sigma_I = 10$ & 0.948437 & 0.000784 & 0.948047 & 0.000792 & 0.950444 & 0.000744 & 0.949851 & 0.000756\\
+		\hline
+		$\sigma_S = 500, \sigma_I = 50$ & 0.968003 & 0.000471 & 0.967327 & 0.000479 & 0.972145 & 0.000422 & 0.970875 & 0.000437\\
+		\hline
+		$\sigma_S = 1000, \sigma_I = 100$ & 0.979365 & 0.000350 & 0.978657 & 0.000356 & 0.983895 & 0.000311 & 0.982500 & 0.000322\\
+		\hline
+		$\sigma_S = 2000, \sigma_I = 200$ & 0.991428 & 0.000255 & 0.990718 & 0.000260 & 0.996167 & 0.000225 & 0.994717 & 0.000234\\
+		\hline
+		$\sigma_S = 5000, \sigma_I = 500$ & 1.007910 & 0.000166 & 1.007211 & 0.000169 & 1.012791 & 0.000146 & 1.011292 & 0.000152\\
+		\hline
+		$\sigma_S = 7500, \sigma_I = 750$ & 1.015308 & 0.000136 & 1.014610 & 0.000139 & 1.020223 & 0.000120 & 1.018711 & 0.000125\\
+		\hline
+		$\sigma_S = 10000, \sigma_I = 1000$ & 1.020580 & 0.000119 & 1.019883 & 0.000121 & 1.025513 & 0.000104 & 1.023994 & 0.000109\\
+		\hline
+		$\sigma_S = 15000, \sigma_I = 1500$ & 1.028035 & 0.000098 & 1.027339 & 0.000100 & 1.032988 & 0.000086 & 1.031463 & 0.000089\\
 		\hline
@@
 	\caption{$\rr$ Estimates and Std. Errs, EXPGROWTH Model,
-		Other Models, $X_0 = 99950, Y_0 = 50$, 
-		$\sigma_X = 10, \sigma_Y = 1$}
-	\begin{footnotesize}
-		\hskip -1cm
-		\begin{tabular}{l|r|r|r|r|r|r|r|r}
-			\hline
-			Data Set & Auto Est & Auto SE & AutoM Est & AutoM SE & Norm Est & Norm SE & NormM Est & NormM SE\\
-			\hline
-			0th & 1.006482 & 0.010827 & 0.782299 & 0.131800 & 1.016043 & 0.009479 & 0.829597 & 0.089876\\
-			\hline
-			1st & 1.060996 & 0.005026 & 1.061004 & 0.005025 & 1.061431 & 0.004995 & 1.060633 & 0.005052\\
-			\hline
-			2nd & 1.073885 & 0.004183 & 1.073323 & 0.004216 & 1.074063 & 0.004172 & 1.073328 & 0.004216\\
-			\hline
-			3rd & 1.077530 & 0.003970 & 1.077039 & 0.003998 & 1.077724 & 0.003959 & 1.077108 & 0.003994\\
-			\hline
-			4th & 1.073888 & 0.004182 & 1.073888 & 0.004182 & 1.074161 & 0.004166 & 1.073508 & 0.004205\\
-			\hline
-			Linear SIR & 1.087538 & 0.003441 & 1.087399 & 0.003448 & 1.087742 & 0.003431 & 1.087338 & 0.003451\\
-			\hline04 & 0.001114\\
+		Other Models, $S_0 = 99950, I_0 = 50$, 
+		$\sigma_S = 10, \sigma_I = 1$}
+	\begin{footnotesize}
+		\hskip -1cm
+	\begin{tabular}{l|r|r|r|r|r|r|r|r}
+		\hline
+		Data Set & Auto Est & Auto SE & AutoM Est & AutoM SE & Norm Est & Norm SE & NormM Est & NormM SE\\
+		\hline
+		0th & 0.888590 & 0.003613 & 0.755287 & 0.049788 & 0.894107 & 0.003150 & 0.784131 & 0.032919\\
+		\hline
+		1st & 0.919899 & 0.001639 & 0.919904 & 0.001639 & 0.920147 & 0.001629 & 0.919691 & 0.001648\\
+		\hline
+		2nd & 0.927250 & 0.001357 & 0.926930 & 0.001368 & 0.927352 & 0.001354 & 0.926933 & 0.001368\\
+		\hline
+		3rd & 0.929326 & 0.001287 & 0.929046 & 0.001296 & 0.929436 & 0.001283 & 0.929085 & 0.001295\\
+		\hline
+		4th & 0.927252 & 0.001357 & 0.927252 & 0.001357 & 0.927408 & 0.001352 & 0.927035 & 0.001365\\
+		\hline
+		Linear SIR & 0.935018 & 0.001111 & 0.934939 & 0.001113 & 0.935133 & 0.001107 & 0.934904 & 0.001114\\
 		\hline
@@
 All of the $\rr$ estimates are less than 1, while the true $\rr$ is greater than 1.
-Changing $Y_0$ does not appear to change much, but increasing $\sigma_X$ and $\sigma_Y$ seems to increase the estimate of $\rr$ towards the true value.
+Changing $I_0$ does not appear to change much, but increasing $\sigma_S$ and $\sigma_I$ seems to increase the estimate of $\rr$ towards the true value.
 It is not clear whether the estimate becomes more accurate or simply increases.
@@
 The standard errors are really small, around the same order of magnitude as the optim() models.
-They decrease as $Y_0$ increases and increase as $\sigma_X$ and $\sigma_Y$ increase.
+They decrease as $I_0$ increases and increase as $\sigma_S$ and $\sigma_I$ increase.
 We have some negative standard error estimates for the 0th order model data because the sum of the incidences is actually negative due to our added random error.
@@
 	\caption{$\rr$ Estimates and Std. Errs, MLE Model, 
-		Baseline Data, $X_0 = 99950, Y_0 = 50$, 
-		$\sigma_X = 10, \sigma_Y = 1$}
-	\begin{footnotesize}
-		\hskip -1cm
-		\begin{tabular}{l|r|r|r|r|r|r|r|r}
-			\hline
-			Data Set & Auto Est & Auto SE & AutoM Est & AutoM SE & Norm Est & Norm SE & NormM Est & NormM SE\\
-			\hline
-			Baseline 1 & 1.045203 & 0.005596 & 0.954158 & 0.004911 & 1.009321 & 0.005045 & 1.010310 & 0.005043\\
-			\hline
-			Baseline 2 & 0.927409 & 0.004482 & 1.132450 & 0.005591 & 0.985546 & 0.004868 & 1.038093 & 0.005246\\
-			\hline
-			Baseline 3 & 1.035850 & 0.005555 & 0.969459 & 0.004980 & 1.001299 & 0.005062 & 1.006797 & 0.005083\\
-			\hline
-			Baseline 4 & 0.974099 & 0.004834 & 0.971615 & 0.004748 & 0.944385 & 0.004705 & 0.994621 & 0.005009\\
-			\hline
-			Baseline 5 & 0.988996 & 0.005185 & 1.020904 & 0.005239 & 1.053116 & 0.005433 & 1.002879 & 0.004973\\
+		Baseline Data, $S_0 = 99950, I_0 = 50$, 
+		$\sigma_S = 10, \sigma_I = 1$}
+	\begin{footnotesize}
+		\hskip -1cm
+	\begin{tabular}{l|r|r|r|r|r|r|r|r}
+		\hline
+		Data Set & Auto Est & Auto SE & AutoM Est & AutoM SE & Norm Est & Norm SE & NormM Est & NormM SE\\
+		\hline
+		Baseline1 & 0.899074 & 9.7e-05 & 0.902153 & 9.7e-05 & 0.899199 & 9.7e-05 & 0.898925 & 9.7e-05\\
+		\hline
+		Baseline2 & 0.898625 & 9.7e-05 & 0.899906 & 9.7e-05 & 0.899591 & 9.7e-05 & 0.900620 & 9.7e-05\\
+		\hline
+		Baseline3 & 0.898641 & 9.7e-05 & 0.896045 & 9.7e-05 & 0.901772 & 9.7e-05 & 0.900604 & 9.7e-05\\
+		\hline
+		Baseline4 & 0.898617 & 9.7e-05 & 0.899951 & 9.7e-05 & 0.898398 & 9.7e-05 & 0.901681 & 9.7e-05\\
+		\hline
+		Baseline5 & 0.901165 & 9.7e-05 & 0.899822 & 9.7e-05 & 0.900534 & 9.7e-05 & 0.900186 & 9.7e-05\\
 		\hline
@@
 		Different Starting Populations, 
-		$\sigma_X = 10, \sigma_Y = 1$}
-	\begin{footnotesize}
-		\hskip -1.7cm
-		\begin{tabular}{l|r|r|r|r|r|r|r|r}
-			\hline
-			Data Set & Auto Est & Auto SE & AutoM Est & AutoM SE & Norm Est & Norm SE & NormM Est & NormM SE\\
-			\hline
-			$X_0 = 99999, Y_0 = 1$ & 1.232756 & 0.069833 & 0.976411 & 0.163888 & 1.615820 & 0.467853 & 1.192249 & 0.205440\\
-			\hline
-			$X_0 = 99990, Y_0 = 10$ & 0.949905 & 0.027784 & 1.301035 & 0.031985 & 0.884468 & 0.022576 & 1.208762 & 0.028717\\
-			\hline
-			$X_0 = 99900, Y_0 = 100$ & 1.016057 & 0.002553 & 0.984204 & 0.002475 & 1.055094 & 0.002654 & 0.990380 & 0.002500\\
-			\hline
-			$X_0 = 99000, Y_0 = 1000$ & 0.983594 & 0.000260 & 0.996016 & 0.000263 & 0.997396 & 0.000264 & 0.998214 & 0.000264\\
+		$\sigma_S = 10, \sigma_I = 1$}
+	\begin{footnotesize}
+		\hskip -1.7cm
+	\begin{tabular}{l|r|r|r|r|r|r|r|r}
+		\hline
+		Data Set & Auto Est & Auto SE & AutoM Est & AutoM SE & Norm Est & Norm SE & NormM Est & NormM SE\\
+		\hline
+		$S_0 = 99999, I_0 = 1$ & 0.912953 & 0.003775 & 0.861994 & 0.003368 & 0.877184 & 0.003582 & 0.995336 & 0.004095\\
+		\hline
+		$S_0 = 99990, I_0 = 10$ & 0.903707 & 0.000390 & 0.926002 & 0.000401 & 0.909539 & 0.000395 & 0.923559 & 0.000402\\
+		\hline
+		$S_0 = 99900, I_0 = 100$ & 0.890136 & 0.000060 & 0.891312 & 0.000060 & 0.891251 & 0.000060 & 0.891413 & 0.000060\\
+		\hline
+		$S_0 = 99000, I_0 = 1000$ & 0.863266 & 0.000024 & 0.863334 & 0.000024 & 0.862902 & 0.000024 & 0.863491 & 0.000024\\
 		\hline
@@
 	\caption{$\rr$ Estimates and Std. Errs, MLE Model, 
-		Different Variances, $X_0 = 99000$, $Y_0 = 1000$}
-	\begin{footnotesize}
-		\hskip -1.7cm
-		\begin{tabular}{l|r|r|r|r|r|r|r|r}
-			\hline
-			Data Set & Auto Est & Auto SE & AutoM Est & AutoM SE & Norm Est & Norm SE & NormM Est & NormM SE\\
-			\hline
-			$\sigma_X = 10, \sigma_Y = 1$ & 0.985737 & 0.000206 & 0.985711 & 0.000207 & 0.985510 & 0.000206 & 0.986968 & 0.000206\\
-			\hline
-			$\sigma_X = 100, \sigma_Y = 10$ & 0.999005 & 0.000208 & 1.001050 & 0.000208 & 1.006205 & 0.000209 & 1.013000 & 0.000211\\
-			\hline
-			$\sigma_X = 300, \sigma_Y = 30$ & 1.051749 & 0.000236 & 0.966436 & 0.000181 & 0.816122 & 0.000167 & 0.966859 & 0.000200\\
-			\hline
-			$\sigma_X = 1000, \sigma_Y = 100$ & 1.060553 & 0.000161 & 1.065634 & 0.000236 & 0.777445 & 0.000177 & 0.975480 & 0.000206\\
-			\hline
-			$\sigma_X = 3000, \sigma_Y = 300$ & 0.511344 & 0.000110 & 1.003435 & 0.000113 & 2.385157 & 0.000672 & 1.095206 & 0.000199\\
-			\hline
-			$\sigma_X = 10000, \sigma_Y = 1000$ & -3.235191 & 0.000840 & 0.952381 & 0.000089 & -2.850294 & -0.001262 & 1.532468 & 0.000209\\
+		Different Variances, $S_0 = 99000$, $I_0 = 1000$}
+	\begin{footnotesize}
+		\hskip -1.7cm
+	\begin{tabular}{l|r|r|r|r|r|r|r|r}
+		\hline
+		Error & Auto Est & Auto SE & AutoM Est & AutoM SE & Norm Est & Norm SE & NormM Est & NormM SE\\
+		\hline
+		$\sigma_S = 10, \sigma_I = 1$ & 0.863286 & 2.4e-05 & 0.864054 & 0.000024 & 0.862686 & 2.4e-05 & 0.863301 & 2.4e-05\\
+		\hline
+		$\sigma_S = 100, \sigma_I = 10$ & 0.865256 & 2.4e-05 & 0.872950 & 0.000025 & 0.859255 & 2.4e-05 & 0.865412 & 2.4e-05\\
+		\hline
+		$\sigma_S = 500, \sigma_I = 50$ & 0.873904 & 2.4e-05 & 0.912804 & 0.000026 & 0.844000 & 2.4e-05 & 0.874831 & 2.5e-05\\
+		\hline
+		$\sigma_S = 1000, \sigma_I = 100$ & 0.884476 & 2.4e-05 & 0.963357 & 0.000027 & 0.824913 & 2.3e-05 & 0.886693 & 2.5e-05\\
+		\hline
+		$\sigma_S = 2000, \sigma_I = 200$ & 0.904863 & 2.4e-05 & 1.066995 & 0.000031 & 0.786679 & 2.2e-05 & 0.910719 & 2.6e-05\\
+		\hline
+		$\sigma_S = 5000, \sigma_I = 500$ & 0.960573 & 2.4e-05 & 1.399764 & 0.000043 & 0.671494 & 1.9e-05 & 0.985289 & 2.9e-05\\
+		\hline
+		$\sigma_S = 7500, \sigma_I = 750$ & 1.001610 & 2.4e-05 & 1.705105 & 0.000054 & 0.574948 & 1.6e-05 & 1.050455 & 3.1e-05\\
+		\hline
+		$\sigma_S = 10000, \sigma_I = 1000$ & 1.038534 & 2.3e-05 & 2.039858 & 0.000068 & 0.477891 & 1.4e-05 & 1.118568 & 3.4e-05\\
+		\hline
+		$\sigma_S = 15000, \sigma_I = 1500$ & 1.102292 & 2.2e-05 & 2.816412 & 0.000104 & 0.282224 & 8.0e-06 & 1.264474 & 4.1e-05\\
 		\hline
@@
 	\caption{$\rr$ Estimates and Std. Errs, MLE Model,
-		Other Models, $X_0 = 99950, Y_0 = 50$, 
-		$\sigma_X = 10, \sigma_Y = 1$}
-	\begin{footnotesize}
-		\hskip -1cm
-		\begin{tabular}{l|r|r|r|r|r|r|r|r}
-			\hline
-			Data Set & Auto Est & Auto SE & AutoM Est & AutoM SE & Norm Est & Norm SE & NormM Est & NormM SE\\
-			\hline
-			0th & -12.609343 & -14.168066 & 16.325348 & 42.366553 & 3.152735 & -0.768473 & 10.431474 & 8.031043\\
-			\hline
-			1st & 0.978683 & 0.000125 & 0.982726 & 0.000126 & 0.979523 & 0.000125 & 0.976522 & 0.000125\\
-			\hline
-			2nd & 1.017042 & 0.000131 & 1.013913 & 0.000131 & 1.014664 & 0.000131 & 1.014455 & 0.000131\\
-			\hline
-			3rd & 1.029675 & 0.000121 & 1.026645 & 0.000121 & 1.027745 & 0.000121 & 1.027322 & 0.000121\\
-			\hline
-			4th & 1.028890 & 0.000135 & 1.028890 & 0.000135 & 1.027876 & 0.000135 & 1.027989 & 0.000135\\
-			\hline
-			Linear SIR & 1.052763 & 0.000099 & 1.053736 & 0.000099 & 1.053703 & 0.000099 & 1.052769 & 0.000099\\
+		Other Models, $S_0 = 99950, I_0 = 50$, 
+		$\sigma_S = 10, \sigma_I = 1$}
+	\begin{footnotesize}
+		\hskip -1cm
+	\begin{tabular}{l|r|r|r|r|r|r|r|r}
+		\hline
+		Data Set & Auto Est & Auto SE & AutoM Est & AutoM SE & Norm Est & Norm SE & NormM Est & NormM SE\\
+		\hline
+		0th & 3.167663 & -0.894136 & 8.459310 & 11.375438 & 3.145217 & -0.764813 & 6.433011 & 3.054282\\
+		\hline
+		1st & 0.872597 & 0.000099 & 0.875151 & 0.000100 & 0.873004 & 0.000099 & 0.870359 & 0.000099\\
+		\hline
+		2nd & 0.894306 & 0.000101 & 0.892614 & 0.000101 & 0.892698 & 0.000101 & 0.892652 & 0.000101\\
+		\hline
+		3rd & 0.902721 & 0.000093 & 0.900171 & 0.000093 & 0.901044 & 0.000093 & 0.900592 & 0.000093\\
+		\hline
+		4th & 0.901112 & 0.000104 & 0.901112 & 0.000104 & 0.900495 & 0.000104 & 0.900734 & 0.000104\\
+		\hline
+		Linear SIR & 0.915222 & 0.000075 & 0.915111 & 0.000074 & 0.915750 & 0.000075 & 0.914991 & 0.000075\\
 		\hline
@@
 The estimates here are actually all pretty close to the true $\rr$, and they are extremely consistent for the baseline data sets.
-They also seem to increase as $Y_0$ increases and as $\sigma_X$ and $\sigma_Y$ increase.
+They also seem to increase as $I_0$ increases and as $\sigma_S$ and $\sigma_I$ increase.
 This seems to occur regardless of accuracy.
@@
 
-The standard error estimates are all reasonable. They seem to decrease slightly as $Y_0$ increases and as $\sigma_X$ and $\sigma_Y$ increase.
+The standard error estimates are all reasonable. They seem to decrease slightly as $I_0$ increases and as $\sigma_S$ and $\sigma_I$ increase.
 
@@
 	\caption{$\rr$ Estimates and Std. Errs, INITGROWTH Model, 
-		Baseline Data, $X_0 = 99950, Y_0 = 50$, 
-		$\sigma_X = 10, \sigma_Y = 1$}
-	\begin{footnotesize}
-		\hskip -1cm
-		\begin{tabular}{l|r|r|r|r|r|r|r|r}
-			\hline
-			Data Set & Auto Est & Auto SE & AutoM Est & AutoM SE & Norm Est & Norm SE & NormM Est & NormM SE\\
-			\hline
-			Baseline 1 & 1.122458 & 0.044665 & 0.951335 & 0.152044 & 1.135276 & 0.040490 & 0.965079 & 0.138753\\
-			\hline
-			Baseline 2 & 1.119443 & 0.045703 & 0.982261 & 0.123521 & 1.136269 & 0.040183 & 0.967189 & 0.136801\\
-			\hline
-			Baseline 3 & 1.110763 & 0.048817 & 0.954452 & 0.148942 & 1.131325 & 0.041737 & 0.963011 & 0.140689\\
-			\hline
-			Baseline 4 & 1.111362 & 0.048596 & 0.962238 & 0.141417 & 1.134666 & 0.040680 & 0.961751 & 0.141879\\
-			\hline
-			Baseline 5 & 1.116547 & 0.046721 & 0.962069 & 0.141578 & 1.129191 & 0.042424 & 0.965051 & 0.138765\\
+		Baseline Data, $S_0 = 99950, I_0 = 50$, 
+		$\sigma_S = 10, \sigma_I = 1$}
+	\begin{footnotesize}
+		\hskip -1cm
+	\begin{tabular}{l|r|r|r|r|r|r|r|r}
+		\hline
+		Data Set & Auto Est & Auto SE & AutoM Est & AutoM SE & Norm Est & Norm SE & NormM Est & NormM SE\\
+		\hline
+		Baseline 1 & 1.248124 & 0.016595 & 1.247917 & 0.016623 & 1.248829 & 0.016501 & 1.247484 & 0.016682\\
+		\hline
+		Baseline 2 & 1.248130 & 0.016594 & 1.247549 & 0.016673 & 1.248722 & 0.016515 & 1.247691 & 0.016653\\
+		\hline
+		Baseline 3 & 1.248067 & 0.016603 & 1.247266 & 0.016711 & 1.248788 & 0.016506 & 1.247692 & 0.016654\\
+		\hline
+		Baseline 4 & 1.248048 & 0.016605 & 1.247550 & 0.016672 & 1.248777 & 0.016508 & 1.247761 & 0.016644\\
+		\hline
+		Baseline 5 & 1.248106 & 0.016597 & 1.247529 & 0.016675 & 1.248912 & 0.016490 & 1.247547 & 0.016673\\
 		\hline
@@
 		Different Starting Populations, 
-		$\sigma_X = 10, \sigma_Y = 1$}
-	\begin{footnotesize}
-		\hskip -1.7cm
-		\begin{tabular}{l|r|r|r|r|r|r|r|r}
-			\hline
-			Data Set & Auto Est & Auto SE & AutoM Est & AutoM SE & Norm Est & Norm SE & NormM Est & NormM SE\\
-			\hline
-			$X_0 = 99999, Y_0 = 1$ & 1.107146 & 0.050171 & 0.578787 & 0.985754 & 1.132118 & 0.041484 & 0.555865 & 1.065990\\
-			\hline
-			$X_0 = 99990, Y_0 = 10$ & 1.116055 & 0.046896 & 0.855171 & 0.276575 & 1.129441 & 0.042344 & 0.838084 & 0.305078\\
-			\hline
-			$X_0 = 99900, Y_0 = 100$ & 1.123557 & 0.044293 & 1.018159 & 0.096252 & 1.137845 & 0.039698 & 1.018516 & 0.096009\\
-			\hline
-			$X_0 = 99000, Y_0 = 1000$ & 1.186268 & 0.027217 & 1.182227 & 0.028098 & 1.189399 & 0.026553 & 1.182180 & 0.028108\\
+		$\sigma_S = 10, \sigma_I = 1$}
+	\begin{footnotesize}
+		\hskip -1.7cm
+	\begin{tabular}{l|r|r|r|r|r|r|r|r}
+		\hline
+		Data Set & Auto Est & Auto SE & AutoM Est & AutoM SE & Norm Est & Norm SE & NormM Est & NormM SE\\
+		\hline
+		$S_0 = 99999, I_0 = 1$ & 1.113607 & 0.047776 & 0.991326 & 0.116074 & 1.135438 & 0.040440 & 1.001869 & 0.107891\\
+		\hline
+		$S_0 = 99990, I_0 = 10$ & 1.170601 & 0.030784 & 1.160443 & 0.033326 & 1.179144 & 0.028787 & 1.159943 & 0.033456\\
+		\hline
+		$S_0 = 99900, I_0 = 100$ & 1.274694 & 0.013368 & 1.274705 & 0.013367 & 1.275003 & 0.013334 & 1.274823 & 0.013354\\
+		\hline
+		$S_0 = 99000, I_0 = 1000$ & 1.301122 & 0.010759 & 1.301114 & 0.010760 & 1.301059 & 0.010765 & 1.301148 & 0.010757\\
 		\hline
@@
 	\caption{$\rr$ Estimates and Std. Errs, INITGROWTH Model, 
-		Different Variances, $X_0 = 99000$, $Y_0 = 1000$}
-	\begin{footnotesize}
-		\hskip -1.7cm
-		\begin{tabular}{l|r|r|r|r|r|r|r|r}
-			\hline
-			Data Set & Auto Est & Auto SE & AutoM Est & AutoM SE & Norm Est & Norm SE & NormM Est & NormM SE\\
-			\hline
-			$\sigma_X = 10, \sigma_Y = 1$ & 1.186512 & 0.027164 & 1.181573 & 0.028243 & 1.192153 & 0.025981 & 1.182118 & 0.028121\\
-			\hline
-			$\sigma_X = 100, \sigma_Y = 10$ & 1.271572 & 0.013714 & 1.184621 & 0.027573 & 1.292533 & 0.011548 & 1.186089 & 0.027256\\
-			\hline
-			$\sigma_X = 300, \sigma_Y = 30$ & 1.337570 & 0.007951 & 1.196232 & 0.025155 & 1.351500 & 0.007077 & 1.180770 & 0.028422\\
-			\hline
-			$\sigma_X = 1000, \sigma_Y = 100$ & 1.408230 & 0.004385 & 1.181330 & 0.028297 & 1.426440 & 0.003755 & 1.179468 & 0.028713\\
-			\hline
-			$\sigma_X = 3000, \sigma_Y = 300$ & 1.468704 & 0.002614 & 1.250992 & 0.016213 & 1.495791 & 0.002069 & 1.211323 & 0.022310\\
-			\hline
-			$\sigma_X = 10000, \sigma_Y = 1000$ & 1.544051 & 0.001360 & 1.280487 & 0.012748 & 1.559289 & 0.001190 & 1.281732 & 0.012619\\
-			\hline
-
+		Different Variances, $S_0 = 99000$, $I_0 = 1000$}
+	\begin{footnotesize}
+		\hskip -1.7cm
+	\begin{tabular}{l|r|r|r|r|r|r|r|r}
+		\hline
+		Data Set & Auto Est & Auto SE & AutoM Est & AutoM SE & Norm Est & Norm SE & NormM Est & NormM SE\\
+		\hline
+		$\sigma_S = 10, \sigma_I = 1$ & 1.301161 & 0.010756 & 1.301227 & 0.010750 & 1.301037 & 0.010767 & 1.301125 & 0.010759\\
+		\hline
+		$\sigma_S = 100, \sigma_I = 10$ & 1.315228 & 0.009575 & 1.313931 & 0.009678 & 1.321895 & 0.009059 & 1.319928 & 0.009209\\
+		\hline
+		$\sigma_S = 500, \sigma_I = 50$ & 1.379802 & 0.005578 & 1.377586 & 0.005683 & 1.393349 & 0.004975 & 1.389199 & 0.005152\\
+		\hline
+		$\sigma_S = 1000, \sigma_I = 100$ & 1.416861 & 0.004075 & 1.414559 & 0.004155 & 1.431547 & 0.003595 & 1.427030 & 0.003736\\
+		\hline
+		$\sigma_S = 2000, \sigma_I = 200$ & 1.455861 & 0.002919 & 1.453577 & 0.002977 & 1.471091 & 0.002561 & 1.466436 & 0.002665\\
+		\hline
+		$\sigma_S = 5000, \sigma_I = 500$ & 1.508597 & 0.001851 & 1.506373 & 0.001888 & 1.524094 & 0.001618 & 1.519341 & 0.001686\\
+		\hline
+		$\sigma_S = 7500, \sigma_I = 750$ & 1.532063 & 0.001510 & 1.529856 & 0.001539 & 1.547585 & 0.001318 & 1.542817 & 0.001374\\
+		\hline
+		$\sigma_S = 10000, \sigma_I = 1000$ & 1.548711 & 0.001305 & 1.546515 & 0.001331 & 1.564232 & 0.001140 & 1.559461 & 0.001188\\
+		\hline
+		$\sigma_S = 15000, \sigma_I = 1500$ & 1.572148 & 0.001063 & 1.569966 & 0.001084 & 1.587649 & 0.000928 & 1.582882 & 0.000968\\
+		\hline
 	\end{tabular}
@@
 	\caption{$\rr$ Estimates and Std. Errs, INITGROWTH Model,
-		Other Models, $X_0 = 99950, Y_0 = 50$, 
-		$\sigma_X = 10, \sigma_Y = 1$}
+		Other Models, $S_0 = 99950, I_0 = 50$, 
+		$\sigma_S = 10, \sigma_I = 1$}
 	\begin{footnotesize}
@@
 The estimates and standard errors for the baseline data sets are almost exactly the same between the different ones.
-The estimates tend to become less accurate as $\sigma_X$ and $\sigma_Y$ increase.
-They increase as $Y_0$ increases as well.
-However, this may be due to the fact that the model does not take starting compartment sizes in to account; thus, it does not differentiate between the change in final size caused by changing $Y_0$ when the data are generated from the classic SIR model.
+The estimates tend to become less accurate as $\sigma_S$ and $\sigma_I$ increase.
+They increase as $I_0$ increases as well.
+However, this may be due to the fact that the model does not take starting compartment sizes in to account; thus, it does not differentiate between the change in final size caused by changing $I_0$ when the data are generated from the classic SIR model.
 Also, the model does not give crazy results for data that are not generated from the classic SIR model, as we might expect since all it uses for estimation is the final number of people who were infected.
 
-The standard errors are all relatively small in magnitude given the way I calculated them, and they increase as $Y_0$ and as $\sigma_X$ and $\sigma_Y$ increase.
+The standard errors are all relatively small in magnitude given the way I calculated them, and they increase as $I_0$ and as $\sigma_S$ and $\sigma_I$ increase.
 
@@
 	\caption{$\rr$ Estimates and Std. Errs, FINALSIZE Model, 
-		Baseline Data$X_0 = 99950, Y_0 = 50$, 
-		$\sigma_X = 10, \sigma_Y = 1$}
-	\centering
-	\begin{footnotesize}
-		\hskip -1cm
-	\begin{tabular}[t]{l|r|r|r|r|r|r|r|r}
-		\hline
-		Data Set & Auto Est & Auto SE & AutoM Est & AutoM SE & Norm Est & Norm SE & NormM Est & NormM SE\\
-		\hline
-		Baseline 1 & 1.001228 & 7.8e-05 & 1.001179 & 7.7e-05 & 1.001262 & 8.0e-05 & 1.001264 & 8.0e-05\\
-		\hline
-		Baseline 2 & 1.001211 & 7.8e-05 & 1.001400 & 8.4e-05 & 1.001250 & 7.9e-05 & 1.001279 & 8.0e-05\\
-		\hline
-		Baseline 3 & 1.001218 & 7.8e-05 & 1.001196 & 7.7e-05 & 1.001242 & 7.9e-05 & 1.001249 & 7.9e-05\\
-		\hline
-		Baseline 4 & 1.001233 & 7.9e-05 & 1.001246 & 7.9e-05 & 1.001200 & 7.8e-05 & 1.001239 & 7.9e-05\\
-		\hline
-		Baseline 5 & 1.001195 & 7.7e-05 & 1.001247 & 7.9e-05 & 1.001273 & 8.0e-05 & 1.001263 & 8.0e-05\\
+		Baseline Data, $S_0 = 99950, I_0 = 50$, 
+		$\sigma_S = 10, \sigma_I = 1$}
+	\begin{footnotesize}
+		\hskip -1cm
+	\begin{tabular}{l|r|r|r|r|r|r|r|r}
+		\hline
+		Data Set & Auto Est & Auto SE & AutoM Est & AutoM SE & Norm Est & Norm SE & NormM Est & NormM SE\\
+		\hline
+		Baseline1 & 1.044459 & 0.000492 & 1.044644 & 0.000494 & 1.044435 & 0.000492 & 1.044410 & 0.000492\\
+		\hline
+		Baseline2 & 1.044363 & 0.000492 & 1.044445 & 0.000492 & 1.044433 & 0.000492 & 1.044518 & 0.000493\\
+		\hline
+		Baseline3 & 1.044434 & 0.000492 & 1.044279 & 0.000491 & 1.044565 & 0.000493 & 1.044521 & 0.000493\\
+		\hline
+		Baseline4 & 1.044370 & 0.000492 & 1.044446 & 0.000492 & 1.044374 & 0.000492 & 1.044557 & 0.000493\\
+		\hline
+		Baseline5 & 1.044480 & 0.000493 & 1.044438 & 0.000492 & 1.044475 & 0.000493 & 1.044442 & 0.000492\\
 		\hline
@@
 		Different Starting Populations, 
-		$\sigma_X = 10, \sigma_Y = 1$}
-	\begin{footnotesize}
-		\hskip -1.7cm
-		\begin{tabular}{l|r|r|r|r|r|r|r|r}
-			\hline
-			Data Set & Auto Est & Auto SE & AutoM Est & AutoM SE & Norm Est & Norm SE & NormM Est & NormM SE\\
-			\hline
-			$X_0 = 99999, Y_0 = 1$ & 1.000114 & 0.000024 & 1.000034 & 0.000013 & 1.000033 & 0.000013 & 1.000040 & 0.000014\\
-			\hline
-			$X_0 = 99990, Y_0 = 10$ & 1.000212 & 0.000033 & 1.000315 & 0.000040 & 1.000223 & 0.000033 & 1.000305 & 0.000039\\
-			\hline
-			$X_0 = 99900, Y_0 = 100$ & 1.002530 & 0.000113 & 1.002465 & 0.000111 & 1.002606 & 0.000114 & 1.002470 & 0.000111\\
-			\hline
-			$X_0 = 99000, Y_0 = 1000$ & 1.024360 & 0.000358 & 1.024661 & 0.000360 & 1.024662 & 0.000360 & 1.024670 & 0.000360\\
+		$\sigma_S = 10, \sigma_I = 1$}
+	\begin{footnotesize}
+		\hskip -1.7cm
+	\begin{tabular}{l|r|r|r|r|r|r|r|r}
+		\hline
+		Data Set & Auto Est & Auto SE & AutoM Est & AutoM SE & Norm Est & Norm SE & NormM Est & NormM SE\\
+		\hline
+		$S_0 = 99999, I_0 = 1$ & 1.001111 & 0.000075 & 1.001110 & 0.000075 & 1.001081 & 0.000074 & 1.001217 & 0.000078\\
+		\hline
+		$S_0 = 99990, I_0 = 10$ & 1.010679 & 0.000234 & 1.010895 & 0.000236 & 1.010676 & 0.000234 & 1.010824 & 0.000235\\
+		\hline
+		$S_0 = 99900, I_0 = 100$ & 1.073689 & 0.000652 & 1.073782 & 0.000652 & 1.073845 & 0.000653 & 1.073877 & 0.000653\\
+		\hline
+		$S_0 = 99000, I_0 = 1000$ & 1.203036 & 0.001215 & 1.203028 & 0.001215 & 1.202929 & 0.001214 & 1.203088 & 0.001215\\
 		\hline
@@
 	\caption{$\rr$ Estimates and Std. Errs, FINALSIZE Model, 
-		Different Variances, $X_0 = 99000$, $Y_0 = 1000$}
-	\begin{footnotesize}
-		\hskip -1.7cm
-		\begin{tabular}{l|r|r|r|r|r|r|r|r}
-			\hline
-			Data Set & Auto Est & Auto SE & AutoM Est & AutoM SE & Norm Est & Norm SE & NormM Est & NormM SE\\
-			\hline
-			$\sigma_X = 10, \sigma_Y = 1$ & 1.024640 & 0.000360 & 1.024544 & 0.000359 & 1.024610 & 0.000359 & 1.024641 & 0.000360\\
-			\hline
-			$\sigma_X = 100, \sigma_Y = 10$ & 1.025024 & 0.000363 & 1.025169 & 0.000364 & 1.025348 & 0.000365 & 1.025358 & 0.000365\\
-			\hline
-			$\sigma_X = 300, \sigma_Y = 30$ & 1.024494 & 0.000359 & 1.026920 & 0.000377 & 1.020803 & 0.000329 & 1.024440 & 0.000358\\
-			\hline
-			$\sigma_X = 1000, \sigma_Y = 100$ & 1.036821 & 0.000445 & 1.025126 & 0.000363 & 1.017766 & 0.000303 & 1.024120 & 0.000356\\
-			\hline
-			$\sigma_X = 3000, \sigma_Y = 300$ & 1.012305 & 0.000251 & 1.047894 & 0.000513 & 1.045165 & 0.000497 & 1.031720 & 0.000411\\
-			\hline
-			$\sigma_X = 10000, \sigma_Y = 1000$ & 1.068277 & 0.000624 & 1.055107 & 0.000554 & 0.969364 & NaN & 1.061044 & 0.000586\\
+		Different Variances, $S_0 = 99000$, $I_0 = 1000$}
+	\begin{footnotesize}
+		\hskip -1.7cm
+	\begin{tabular}{l|r|r|r|r|r|r|r|r}
+		\hline
+		Error & Auto Est & Auto SE & AutoM Est & AutoM SE & Norm Est & Norm SE & NormM Est & NormM SE\\
+		\hline
+		$\sigma_S = 10, \sigma_I = 1$ & 1.203100 & 0.001215 & 1.203221 & 0.001215 & 1.202857 & 0.001214 & 1.203019 & 0.001215\\
+		\hline
+		$\sigma_S = 100, \sigma_I = 10$ & 1.204254 & 0.001220 & 1.205468 & 0.001224 & 1.201827 & 0.001210 & 1.203450 & 0.001216\\
+		\hline
+		$\sigma_S = 500, \sigma_I = 50$ & 1.209421 & 0.001240 & 1.215599 & 0.001265 & 1.197277 & 0.001191 & 1.205368 & 0.001224\\
+		\hline
+		$\sigma_S = 1000, \sigma_I = 100$ & 1.215968 & 0.001267 & 1.228600 & 0.001317 & 1.191656 & 0.001169 & 1.207778 & 0.001234\\
+		\hline
+		$\sigma_S = 2000, \sigma_I = 200$ & 1.229359 & 0.001320 & 1.255800 & 0.001425 & 1.180629 & 0.001124 & 1.212639 & 0.001253\\
+		\hline
+		$\sigma_S = 5000, \sigma_I = 500$ & 1.272128 & 0.001489 & 1.348709 & 0.001792 & 1.149157 & 0.000994 & 1.227550 & 0.001313\\
+		\hline
+		$\sigma_S = 7500, \sigma_I = 750$ & 1.311117 & 0.001643 & 1.442921 & 0.002168 & 1.124616 & 0.000888 & 1.240372 & 0.001364\\
+		\hline
+		$\sigma_S = 10000, \sigma_I = 1000$ & 1.353661 & 0.001811 & 1.558903 & 0.002644 & 1.101452 & 0.000785 & 1.253572 & 0.001416\\
+		\hline
+		$\sigma_S = 15000, \sigma_I = 1500$ & 1.451957 & 0.002204 & 1.908199 & 0.004213 & 1.058763 & 0.000574 & 1.281188 & 0.001525\\
 		\hline
@@
 	\caption{$\rr$ Estimates and Std. Errs, FINALSIZE Model,
-		Other Models, $X_0 = 99950, Y_0 = 50$, 
-		$\sigma_X = 10, \sigma_Y = 1$}
+		Other Models, $S_0 = 99950, I_0 = 50$, 
+		$\sigma_S = 10, \sigma_I = 1$}
 	\begin{footnotesize}
@@
 	\caption{$\rr$ Estimates and Std. Errs, BRANCH Model, 
-		Baseline Data$X_0 = 99950, Y_0 = 50$, 
-		$\sigma_X = 10, \sigma_Y = 1$}
-	\centering
-	\begin{footnotesize}
-		\hskip -1cm
-	\begin{tabular}[t]{l|r|r|r|r|r|r|r|r}
-		\hline
-		Data Set & Auto Est & Auto SE & AutoM Est & AutoM SE & Norm Est & Norm SE & NormM Est & NormM SE\\
-		\hline
-		Baseline 1 & 1.045203 & 0.005596 & 0.954158 & 0.004911 & 1.009321 & 0.005045 & 1.010310 & 0.005043\\
-		\hline
-		Baseline 2 & 0.927409 & 0.004482 & 1.132450 & 0.005591 & 0.985546 & 0.004868 & 1.038093 & 0.005246\\
-		\hline
-		Baseline 3 & 1.035850 & 0.005555 & 0.969459 & 0.004980 & 1.001299 & 0.005062 & 1.006797 & 0.005083\\
-		\hline
-		Baseline 4 & 0.974099 & 0.004834 & 0.971615 & 0.004748 & 0.944385 & 0.004705 & 0.994621 & 0.005009\\
-		\hline
-		Baseline 5 & 0.988996 & 0.005185 & 1.020904 & 0.005239 & 1.053116 & 0.005433 & 1.002879 & 0.004973\\
+		Baseline Data, $S_0 = 99950, I_0 = 50$, 
+		$\sigma_S = 10, \sigma_I = 1$}
+	\begin{footnotesize}
+		\hskip -1cm
+	\begin{tabular}{l|r|r|r|r|r|r|r|r}
+		\hline
+		Data Set & Auto Est & Auto SE & AutoM Est & AutoM SE & Norm Est & Norm SE & NormM Est & NormM SE\\
+		\hline
+		Baseline1 & 0.899074 & 9.7e-05 & 0.902153 & 9.7e-05 & 0.899199 & 9.7e-05 & 0.898925 & 9.7e-05\\
+		\hline
+		Baseline2 & 0.898625 & 9.7e-05 & 0.899906 & 9.7e-05 & 0.899591 & 9.7e-05 & 0.900620 & 9.7e-05\\
+		\hline
+		Baseline3 & 0.898641 & 9.7e-05 & 0.896045 & 9.7e-05 & 0.901772 & 9.7e-05 & 0.900604 & 9.7e-05\\
+		\hline
+		Baseline4 & 0.898617 & 9.7e-05 & 0.899951 & 9.7e-05 & 0.898398 & 9.7e-05 & 0.901681 & 9.7e-05\\
+		\hline
+		Baseline5 & 0.901165 & 9.7e-05 & 0.899822 & 9.7e-05 & 0.900534 & 9.7e-05 & 0.900186 & 9.7e-05\\
 		\hline
@@
 	\caption{$\rr$ Estimates and Std. Errs, BRANCH Model, 
-		Different Variances, $X_0 = 99000$, $Y_0 = 1000$}
-	\begin{footnotesize}
-		\hskip -1.7cm
-		\begin{tabular}{l|r|r|r|r|r|r|r|r}
-			\hline
-			Data Set & Auto Est & Auto SE & AutoM Est & AutoM SE & Norm Est & Norm SE & NormM Est & NormM SE\\
-			\hline
-			$\sigma_X = 10, \sigma_Y = 1$ & 0.985737 & 0.000206 & 0.985711 & 0.000207 & 0.985510 & 0.000206 & 0.986968 & 0.000206\\
-			\hline
-			$\sigma_X = 100, \sigma_Y = 10$ & 0.999005 & 0.000208 & 1.001050 & 0.000208 & 1.006205 & 0.000209 & 1.013000 & 0.000211\\
-			\hline
-			$\sigma_X = 300, \sigma_Y = 30$ & 1.051749 & 0.000236 & 0.966436 & 0.000181 & 0.816122 & 0.000167 & 0.966859 & 0.000200\\
-			\hline
-			$\sigma_X = 1000, \sigma_Y = 100$ & 1.060553 & 0.000161 & 1.065634 & 0.000236 & 0.777445 & 0.000177 & 0.975480 & 0.000206\\
-			\hline
-			$\sigma_X = 3000, \sigma_Y = 300$ & 0.511344 & 0.000110 & 1.003435 & 0.000113 & 2.385157 & 0.000672 & 1.095206 & 0.000199\\
-			\hline
-			$\sigma_X = 10000, \sigma_Y = 1000$& -3.235191 & 0.000840 & 0.952381 & 0.000089 & -2.850294 & 0.001262 & 1.532468 & 0.000209\\
+		Different Variances, $S_0 = 99000$, $I_0 = 1000$}
+	\begin{footnotesize}
+		\hskip -1.7cm
+	\begin{tabular}{l|r|r|r|r|r|r|r|r}
+		\hline
+		Error & Auto Est & Auto SE & AutoM Est & AutoM SE & Norm Est & Norm SE & NormM Est & NormM SE\\
+		\hline
+		$\sigma_S = 10, \sigma_I = 1$ & 0.863286 & 2.4e-05 & 0.864054 & 0.000024 & 0.862686 & 2.4e-05 & 0.863301 & 2.4e-05\\
+		\hline
+		$\sigma_S = 100, \sigma_I = 10$ & 0.865256 & 2.4e-05 & 0.872950 & 0.000025 & 0.859255 & 2.4e-05 & 0.865412 & 2.4e-05\\
+		\hline
+		$\sigma_S = 500, \sigma_I = 50$ & 0.873904 & 2.4e-05 & 0.912804 & 0.000026 & 0.844000 & 2.4e-05 & 0.874831 & 2.5e-05\\
+		\hline
+		$\sigma_S = 1000, \sigma_I = 100$ & 0.884476 & 2.4e-05 & 0.963357 & 0.000027 & 0.824913 & 2.3e-05 & 0.886693 & 2.5e-05\\
+		\hline
+		$\sigma_S = 2000, \sigma_I = 200$ & 0.904863 & 2.4e-05 & 1.066995 & 0.000031 & 0.786679 & 2.2e-05 & 0.910719 & 2.6e-05\\
+		\hline
+		$\sigma_S = 5000, \sigma_I = 500$ & 0.960573 & 2.4e-05 & 1.399764 & 0.000043 & 0.671494 & 1.9e-05 & 0.985289 & 2.9e-05\\
+		\hline
+		$\sigma_S = 7500, \sigma_I = 750$ & 1.001610 & 2.4e-05 & 1.705105 & 0.000054 & 0.574948 & 1.6e-05 & 1.050455 & 3.1e-05\\
+		\hline
+		$\sigma_S = 10000, \sigma_I = 1000$ & 1.038534 & 2.3e-05 & 2.039858 & 0.000068 & 0.477891 & 1.4e-05 & 1.118568 & 3.4e-05\\
+		\hline
+		$\sigma_S = 15000, \sigma_I = 1500$ & 1.102292 & 2.2e-05 & 2.816412 & 0.000104 & 0.282224 & 8.0e-06 & 1.264474 & 4.1e-05\\
 		\hline
@@
 	\caption{$\rr$ Estimates and Std. Errs, BRANCH Model,
-		Other Models, $X_0 = 99950, Y_0 = 50$, 
-		$\sigma_X = 10, \sigma_Y = 1$}
-	\begin{footnotesize}
-		\hskip -1cm
-		\begin{tabular}{l|r|r|r|r|r|r|r|r}
-			\hline
-			Data Set & Auto Est & Auto SE & AutoM Est & AutoM SE & Norm Est & Norm SE & NormM Est & NormM SE\\
-			\hline
-			0th & -12.609343 & -14.168066 & 16.325348 & 42.366553 & 3.152735 & -0.768473 & 10.431474 & 8.031043\\
-			\hline
-			1st & 0.978683 & 0.000125 & 0.982726 & 0.000126 & 0.979523 & 0.000125 & 0.976522 & 0.000125\\
-			\hline
-			2nd & 1.017042 & 0.000131 & 1.013913 & 0.000131 & 1.014664 & 0.000131 & 1.014455 & 0.000131\\
-			\hline
-			3rd & 1.029675 & 0.000121 & 1.026645 & 0.000121 & 1.027745 & 0.000121 & 1.027322 & 0.000121\\
-			\hline
-			4th & 1.028890 & 0.000135 & 1.028890 & 0.000135 & 1.027876 & 0.000135 & 1.027989 & 0.000135\\
-			\hline
-			Linear SIR & 1.052763 & 0.000099 & 1.053736 & 0.000099 & 1.053703 & 0.000099 & 1.052769 & 0.000099\\
-			\hline
-		\end{tabular}
-	\end{footnotesize}
-\end{table}
+		Other Models, $S_0 = 99950, I_0 = 50$, 
+		$\sigma_S = 10, \sigma_I = 1$}
+	\begin{footnotesize}
+		\hskip -1cm
+	\begin{tabular}{l|r|r|r|r|r|r|r|r}
+		\hline
+		Data Set & Auto Est & Auto SE & AutoM Est & AutoM SE & Norm Est & Norm SE & NormM Est & NormM SE\\
+		\hline
+		0th & 3.167663 & -0.894136 & 8.459310 & 11.375438 & 3.145217 & -0.764813 & 6.433011 & 3.054282\\
+		\hline
+		1st & 0.872597 & 0.000099 & 0.875151 & 0.000100 & 0.873004 & 0.000099 & 0.870359 & 0.000099\\
+		\hline
+		2nd & 0.894306 & 0.000101 & 0.892614 & 0.000101 & 0.892698 & 0.000101 & 0.892652 & 0.000101\\
+		\hline
+		3rd & 0.902721 & 0.000093 & 0.900171 & 0.000093 & 0.901044 & 0.000093 & 0.900592 & 0.000093\\
+		\hline
+		4th & 0.901112 & 0.000104 & 0.901112 & 0.000104 & 0.900495 & 0.000104 & 0.900734 & 0.000104\\
+		\hline
+		Linear SIR & 0.915222 & 0.000075 & 0.915111 & 0.000074 & 0.915750 & 0.000075 & 0.914991 & 0.000075\\
+		\hline
+	\end{tabular}
+\end{footnotesize}
+\end{table}
+
+\section{Discussion}
+\label{sec:dis}
+
+\bibliographystyle{apa}%Choose a bibliograhpic style
+\bibliography{Master}
 

### 2018-07-03 paper/draft_v8/draft_v8.tex
finally on git

--- 2018-07-03 paper/Version9/draft_v9.tex
@@
 \textbf{The Gaussian errors are independently distributed with mean 0 and constant variance for the entire length of the time series.
-The autoregressive errors are generated from a first-order process of the form: $$ \epsilon_{t} = \rho * \epsilon_{t-1} + \delta_{t}$$, where $\rho$ is fixed throughout and $\delta_{t} \sim N(0, \sigma)$ with constant $\sigma$.
-The values of the variances are chosen somewhat arbitrarily, although they are intended to be fairly small at first.}
+The autoregressive errors are generated from a first-order process of the form: $$ \epsilon_{t} = \rho * \epsilon_{t-1} + \delta_{t}$$, where $\rho$ is fixed throughout and $\delta_{t} \sim N(0, \sigma)$ with constant $\sigma$.}
 
@@
 
-Below, we compare the estimates and standard errors of $\rr$ given by each of the models for a few selected data sets.
-The forest plots below show the point estimates and 95\% confidence intervals of $\rr$ from each model for a particular data set.
-The true $\rr$ value is also shown by the vertical red line in the graphs.
-
-All of the figures below show results from data sets that were simulated with added Gaussian errors.
-The plot on the top left shows the estimates from one of the baseline data sets.
-For the most part, the $\rr$ values range from approximately 0.8 to 1.4.
-We note that the range of values includes $\rr$ values of less than 1, meaning that some models predict that a disease is not spreading when it actually is.
-However, on the whole, the variation of estimates between models is smaller than we originally expected.
-The confidence intervals are very small relative to the estimates for the most part; exceptions to this include the SIR IPR and the Markov chain model as the exception.
-
-The plot on the top right shows the estimates for the data set where $Y_0$ is increased to 100.
-The $\rr$ estimates generally seem to be further away from the true $\rr$ value compared to the results from the baseline data sets, and the overall .
-The confidence intervals are also more often than not larger than those in the baseline.
-In particular, the standard errors for some of the SIR models increase by several orders of magnitude.
-
-The bottom left plot shows the estimates for the data set where the error added to the generated data is increased.
-For most of the models, the estimates of $\rr$ do not change very much compared to the baseline data set.
-The confidence intervals for this data set increase for some of the models, such as the SIR LMA, while they remain approximately the same as the baseline data sets for most of the others.
-
-On the bottom right, we see the $\rr$ estimates for the case when disease prevalence grows linearly, as opposed to exponentially as seen in the data sets generated under SIR conditions, where the disease prevalence grows exponentially.
-Model performance here varies depending on how well the models the aforementioned linear growth.
-Some of the models, such as the SIR Markov chain and the IPR models, do not handle this case very well, giving much larger estimates and standard errors than are reasonable.
-On the other hand, models such as the SIR maximum and final size seem to be more flexible to different trajectories.
-[The error bars do not show up if they go outside of the x-axis limits - how do we fix this?]
-
-\begin{minipage}{0.5\linewidth}
-	\includegraphics[width=3in]{BaselineFP.jpeg}
-\end{minipage}
-\begin{minipage}{0.5\linewidth}
-	\includegraphics[width=3in]{StartPopFP.jpeg}
-\end{minipage}
-
-\begin{minipage}{0.5\linewidth}
-	\includegraphics[width=3in]{VarianceFP.jpeg}
-\end{minipage}
-\begin{minipage}{0.5\linewidth}
-	\includegraphics[width=3in]{OtherFP.jpeg}
-\end{minipage}
+%% What kind of graphs should we throw in?
+
+%% SG: Forest plots!
 
@@
 			\hline
-			Baseline 1 & 1.965322 & 0.000869 & 1.983308 & 0.000882 & 2.005867 & 0.000893 & 1.994736 & 0.000884\\
-			\hline
-			Baseline 2 & 2.039608 & 0.000913 & 1.999245 & 0.000895 & 2.001217 & 0.000889 & 1.995960 & 0.000887\\
-			\hline
-			Baseline 3 & 2.005609 & 0.000901 & 1.966631 & 0.000873 & 2.016216 & 0.000901 & 1.997175 & 0.000887\\
-			\hline
-			Baseline 4 & 1.958037 & 0.000859 & 1.941109 & 0.000852 & 1.995544 & 0.000886 & 2.004107 & 0.000892\\
-			\hline
-			Baseline 5 & 2.050904 & 0.000926 & 1.964328 & 0.000866 & 2.009912 & 0.000896 & 1.990153 & 0.000881\\
+			Baseline1 & 1.199737 & 0.000476 & 1.199557 & 0.000476 & 1.199737 & 0.000475 & 1.199737 & 0.000475\\
+			\hline
+			Baseline2 & 1.199737 & 0.000475 & 1.200380 & 0.000481 & 1.200376 & 0.000482 & 1.200376 & 0.000482\\
+			\hline
+			Baseline3 & 1.200615 & 0.000485 & 1.199557 & 0.000476 & 1.200380 & 0.000481 & 1.200298 & 0.000479\\
+			\hline
+			Baseline4 & 1.200376 & 0.000481 & 1.200061 & 0.000479 & 1.200380 & 0.000481 & 1.200602 & 0.000484\\
+			\hline
+			Baseline5 & 1.200376 & 0.000481 & 1.200376 & 0.000482 & 1.200298 & 0.000479 & 1.200602 & 0.000484\\
 			\hline
@@
 			\hline
-			$X_0 = 99999, Y_0 = 1$ & 1.689292 & 0.020794 & -4.167216 & 0.168853 & 1.976254 & 0.045081 & 1.688593 & 0.029680\\
-			\hline
-			$X_0 = 99990, Y_0 = 10$ & 2.038297 & 0.004597 & 2.207003 & 0.005165 & 2.037726 & 0.004596 & 1.883084 & 0.004049\\
-			\hline
-			$X_0 = 99900, Y_0 = 100$ & 2.009581 & 0.000449 & 2.006557 & 0.000448 & 1.999236 & 0.000445 & 1.995467 & 0.000444\\
-			\hline
-			$X_0 = 99000, Y_0 = 1000$ & 2.002521 & 0.000046 & 2.001829 & 0.000046 & 1.999519 & 0.000046 & 2.000098 & 0.000046\\
+			$X_0 = 99999, Y_0 = 1$ & 1.214968 & 0.095795 & 1.236988 & 0.096579 & 1.218704 & 0.099065 & 1.218704 & 0.096508\\
+			\hline
+			$X_0 = 99990, Y_0 = 10$ & 1.219334 & 0.006658 & 1.219334 & 0.006854 & 1.219596 & 0.017843 & 1.219596 & 0.017847\\
+			\hline
+			$X_0 = 99900, Y_0 = 100$ & 1.199311 & 0.000166 & 1.200216 & 0.000170 & 1.200336 & 0.000170 & 1.200336 & 0.000170\\
+			\hline
+			$X_0 = 99000, Y_0 = 1000$ & 1.199807 & 0.000029 & 1.199807 & 0.000029 & 1.200160 & 0.000029 & 1.200160 & 0.000029\\
 			\hline
@@
 			\hline
-			Data Set & Auto Est & Auto SE & AutoM Est & AutoM SE & Norm Est & Norm SE & NormM Est & NormM SE\\
-			\hline
-			$\sigma_X = 10, \sigma_Y = 1$ & 1.927444 & 0.000059 & 1.927090 & 5.9e-05 & 1.925768 & 5.9e-05 & 1.926576 & 5.9e-05\\
-			\hline
-			$\sigma_X = 100, \sigma_Y = 10$ & 1.926786 & 0.000058 & 1.916571 & 5.9e-05 & 1.919778 & 5.9e-05 & 1.929422 & 5.9e-05\\
-			\hline
-			$\sigma_X = 300, \sigma_Y = 30$ & 1.994211 & 0.000062 & 1.976893 & 5.8e-05 & 1.930660 & 5.8e-05 & 1.920869 & 5.9e-05\\
-			\hline
-			$\sigma_X = 1000, \sigma_Y = 100$ & 1.749057 & 0.000050 & 1.736588 & 5.9e-05 & 2.003654 & 6.2e-05 & 1.897049 & 5.9e-05\\
-			\hline
-			$\sigma_X = 3000, \sigma_Y = 300$ & 1.992674 & 0.000063 & 1.681585 & 4.9e-05 & 1.959815 & 6.7e-05 & 1.770816 & 5.3e-05\\
-			\hline
-			$\sigma_X = 10000, \sigma_Y = 1000$ & -42.264180 & 0.002694 & 1.773388 & 3.1e-05 & 1.889188 & 6.6e-05 & 2.649010 & 6.3e-05\\
+			Error & Auto Est & Auto SE & AutoM Est & AutoM SE & Norm Est & Norm SE & NormM Est & NormM SE\\
+			\hline
+			$\sigma_X = 10, \sigma_Y = 1$ & 1.200061 & 2.9e-05 & 1.200005 & 2.9e-05 & 1.199951 & 2.9e-05 & 1.199916 & 2.9e-05\\
+			\hline
+			$\sigma_X = 100, \sigma_Y = 10$ & 1.200317 & 2.9e-05 & 1.200427 & 2.9e-05 & 1.199963 & 2.9e-05 & 1.199734 & 2.9e-05\\
+			\hline
+			$\sigma_X = 500, \sigma_Y = 50$ & 1.201642 & 2.8e-05 & 1.201871 & 2.8e-05 & 1.199996 & 2.9e-05 & 1.198872 & 2.9e-05\\
+			\hline
+			$\sigma_X = 1000, \sigma_Y = 100$ & 1.203259 & 2.8e-05 & 1.203811 & 2.8e-05 & 1.199985 & 2.9e-05 & 1.197928 & 2.9e-05\\
+			\hline
+			$\sigma_X = 2000, \sigma_Y = 200$ & 1.206447 & 2.7e-05 & 1.207566 & 2.7e-05 & 1.199849 & 2.9e-05 & 1.195612 & 3.0e-05\\
+			\hline
+			$\sigma_X = 5000, \sigma_Y = 500$ & 1.215906 & 2.5e-05 & 1.218773 & 2.5e-05 & 1.199656 & 2.9e-05 & 1.189382 & 3.2e-05\\
+			\hline
+			$\sigma_X = 7500, \sigma_Y = 750$ & 1.223554 & 2.3e-05 & 1.228037 & 2.4e-05 & 1.199579 & 2.9e-05 & 1.183914 & 3.4e-05\\
+			\hline
+			$\sigma_X = 10000, \sigma_Y = 1000$ & 1.230992 & 2.2e-05 & 1.237142 & 2.2e-05 & 1.199381 & 2.9e-05 & 1.178701 & 3.6e-05\\
+			\hline
+			$\sigma_X = 15000, \sigma_Y = 1500$ & 1.245535 & 1.9e-05 & 1.255188 & 2.0e-05 & 1.199137 & 2.9e-05 & 1.168520 & 4.1e-05\\
 			\hline
@@
 			\hline
-			0th & -2.032207 & 0.244058 & 1.229963 & 0.069020 & 1.021369 & 0.056558 & 0.795460 & 0.331468\\
-			\hline
-			1st & 1.038117 & 0.000029 & 1.037889 & 0.000029 & 1.037902 & 0.000029 & 1.037887 & 0.000029\\
-			\hline
-			2nd & 1.198856 & 0.000045 & 1.198589 & 0.000045 & 1.198772 & 0.000045 & 1.198501 & 0.000045\\
-			\hline
-			3rd & 1.202405 & 0.000043 & 1.202098 & 0.000043 & 1.202350 & 0.000043 & 1.202240 & 0.000043\\
-			\hline
-			4th & 1.210597 & 0.000047 & 1.210597 & 0.000047 & 1.210194 & 0.000047 & 1.210304 & 0.000047\\
-			\hline
-			Linear SIR & 1.274346 & 0.000034 & 1.273739 & 0.000034 & 1.274229 & 0.000034 & 1.273950 & 0.000034\\
+			0th & 0.005468 & 0.000745 & -0.034869 & 0.000764 & -0.034869 & 7.4e-04 & 0.034869 & 0.000673\\
+			\hline
+			1st & 1.038004 & 0.000029 & 1.037900 & 0.000029 & 1.037826 & 2.9e-05 & 1.037826 & 0.000029\\
+			\hline
+			2nd & 1.199663 & 0.000045 & 1.198627 & 0.000045 & 1.199663 & 4.5e-05 & 1.198949 & 0.000045\\
+			\hline
+			3rd & 1.202909 & 0.000043 & 1.202282 & 0.000043 & 1.202414 & 4.3e-05 & 1.202414 & 0.000043\\
+			\hline
+			4th & 1.211092 & 0.000047 & 1.211092 & 0.000047 & 1.209479 & 4.7e-05 & 1.211092 & 0.000047\\
+			\hline
+			Linear SIR & 1.274529 & 0.000034 & 1.273768 & 0.000034 & 1.275057 & 3.4e-05 & 1.274745 & 0.000034\\
 			\hline
@@
 			\hline
-			Baseline 1 & 3.226272 & 0.000000 & 1.984447 & 0.000294 & 2.006310 & 0.000298 & 1.993558 & 0.000294\\
-			\hline
-			Baseline 2 & 3.203012 & 0.000000 & 1.999589 & 0.000299 & 2.000868 & 0.000297 & 1.995630 & 0.000296\\
-			\hline
-			Baseline 3 & 4.180804 & 0.000733 & 4.180810 & 0.000724 & 5.043355 & 0.000908 & 1.997399 & 0.000296\\
-			\hline
-			Baseline 4 & 4.180819 & 0.000709 & 4.180816 & 0.000715 & 1.995237 & 0.000295 & 2.003821 & 0.000298\\
-			\hline
-			Baseline 5 & 4.180799 & 0.000741 & 4.180815 & 0.000717 & 2.008713 & 0.000300 & 4.180812 & 0.000720\\
+			Baseline 1 & 1.200196 & 1e-06 & 1.199942 & 1e-06 & 1.200108 & 1e-06 & 1.199965 & 1e-06\\
+			\hline
+			Baseline 2 & 1.199574 & 1e-06 & 1.199733 & 1e-06 & 1.199922 & 1e-06 & 1.200219 & 1e-06\\
+			\hline
+			Baseline 3 & 1.199909 & 1e-06 & 1.200416 & 1e-06 & 1.199851 & 1e-06 & 1.200009 & 1e-06\\
+			\hline
+			Baseline 4 & 1.199735 & 1e-06 & 1.199829 & 1e-06 & 1.199925 & 1e-06 & 1.199856 & 1e-06\\
+			\hline
+			Baseline 5 & 1.199849 & 1e-06 & 1.199683 & 1e-06 & 1.199943 & 1e-06 & 1.199900 & 1e-06\\
 			\hline
@@
 			\hline
-			$X_0 = 99999, Y_0 = 1$ & 5.899238 & 0.025009 & 5.898924 & 0.051433 & 3.394166 & 0.000021 & 3.217232 & 0.000016\\
-			\hline
-			$X_0 = 99990, Y_0 = 10$ & 3.162364 & 0.000001 & 3.170152 & 0.000002 & 3.131576 & 0.000001 & 3.108194 & 0.000001\\
-			\hline
-			$X_0 = 99900, Y_0 = 100$ & 4.564960 & 0.000404 & 2.005939 & 0.000150 & 1.998880 & 0.000149 & 1.994592 & 0.000148\\
-			\hline
-			$X_0 = 99000, Y_0 = 1000$ & 47.241714 & 0.002493 & 47.241714 & 0.002511 & 47.241714 & 0.002502 & 47.241714 & 0.002502\\
+			$X_0 = 99999, Y_0 = 1$ & 1.209265 & 3.7e-05 & 1.198242 & 3.5e-05 & 1.193162 & 3.4e-05 & 1.200995 & 3.5e-05\\
+			\hline
+			$X_0 = 99990, Y_0 = 10$ & 1.202026 & 4.0e-06 & 1.198515 & 4.0e-06 & 1.199652 & 4.0e-06 & 1.199716 & 4.0e-06\\
+			\hline
+			$X_0 = 99900, Y_0 = 100$ & 1.199895 & 1.0e-06 & 1.199895 & 1.0e-06 & 1.199976 & 1.0e-06 & 1.200103 & 1.0e-06\\
+			\hline
+			$X_0 = 99000, Y_0 = 1000$ & 1.200089 & 0.0e+00 & 1.200008 & 0.0e+00 & 1.199992 & 0.0e+00 & 1.199985 & 0.0e+00\\
 			\hline
@@
 			\hline
-			$\sigma_X = 10, \sigma_Y = 1$ & 1.926733 & 4e-06 & 1.927468 & 4e-06 & 1.925555 & 4e-06 & 1.926689 & 4e-06\\
-			\hline
-			$\sigma_X = 100, \sigma_Y = 10$ & 1.925550 & 4e-06 & 1.916380 & 4e-06 & 1.920075 & 4e-06 & 1.928987 & 4e-06\\
-			\hline
-			$\sigma_X = 300, \sigma_Y = 30$ & 1.994719 & 4e-06 & 1.976273 & 4e-06 & 1.930978 & 4e-06 & 1.920290 & 4e-06\\
-			\hline
-			$\sigma_X = 1000, \sigma_Y = 100$ & 1.749031 & 3e-06 & 1.736589 & 3e-06 & 2.003044 & 4e-06 & 1.897328 & 4e-06\\
-			\hline
-			$\sigma_X = 3000, \sigma_Y = 300$ & 1.990948 & 4e-06 & 1.681748 & 3e-06 & 1.959229 & 4e-06 & 1.770892 & 3e-06\\
-			\hline
-			$\sigma_X = 10000, \sigma_Y = 1000$ & 5.329640 & 3e-05 & 1.773820 & 2e-06 & 1.887014 & 4e-06 & 2.648300 & 6e-06\\
+			$\sigma_X = 10, \sigma_Y = 1$ & 1.199998 & 0 & 1.199990 & 0 & 1.200027 & 0 & 1.199972 & 0\\
+			\hline
+			$\sigma_X = 100, \sigma_Y = 10$ & 1.199920 & 0 & 1.199837 & 0 & 1.200179 & 0 & 1.199611 & 0\\
+			\hline
+			$\sigma_X = 500, \sigma_Y = 50$ & 1.199579 & 0 & 1.199154 & 0 & 1.200844 & 0 & 1.198063 & 0\\
+			\hline
+			$\sigma_X = 1000, \sigma_Y = 100$ & 1.199168 & 0 & 1.198307 & 0 & 1.201705 & 0 & 1.196130 & 0\\
+			\hline
+			$\sigma_X = 2000, \sigma_Y = 200$ & 1.198425 & 0 & 1.196618 & 0 & 1.203466 & 0 & 1.192321 & 0\\
+			\hline
+			$\sigma_X = 5000, \sigma_Y = 500$ & 1.196655 & 0 & 1.191769 & 0 & 1.208871 & 0 & 1.181229 & 0\\
+			\hline
+			$\sigma_X = 7500, \sigma_Y = 750$ & 1.195734 & 0 & 1.187972 & 0 & 1.213591 & 0 & 1.172358 & 0\\
+			\hline
+			$\sigma_X = 10000, \sigma_Y = 1000$ & 1.195439 & 0 & 1.184353 & 0 & 1.218455 & 0 & 1.163930 & 0\\
+			\hline
+			$\sigma_X = 15000, \sigma_Y = 1500$ & 1.196493 & 0 & 1.177762 & 0 & 1.228683 & 0 & 1.148505 & 0\\
 			\hline
@@
 			\hline
-			0th & 3.272461 & 0e+00 & 3.275535 & 0e+00 & 3.135010 & 0e+00 & 3.235197 & 0e+00\\
-			\hline
-			1st & 1.038125 & 0e+00 & 1.037880 & 0e+00 & 1.037903 & 0e+00 & 1.037881 & 0e+00\\
-			\hline
-			2nd & 1.198760 & 1e-06 & 1.198526 & 1e-06 & 1.198864 & 1e-06 & 1.198616 & 1e-06\\
-			\hline
-			3rd & 1.202489 & 1e-06 & 1.202071 & 1e-06 & 1.202246 & 1e-06 & 1.202355 & 1e-06\\
-			\hline
-			4th & 1.210605 & 1e-06 & 1.210605 & 1e-06 & 1.210220 & 1e-06 & 1.210199 & 1e-06\\
-			\hline
-			Linear SIR & 1.274175 & 1e-06 & 1.273799 & 1e-06 & 1.274314 & 1e-06 & 1.274075 & 1e-06\\
+			0th & 2.803084 & 0e+00 & 2.663158 & 0e+00 & 2.688581 & 0e+00 & 2.756037 & 0e+00\\
+			\hline
+			1st & 1.038122 & 0e+00 & 1.037880 & 0e+00 & 1.037900 & 0e+00 & 1.037887 & 0e+00\\
+			\hline
+			2nd & 1.198739 & 1e-06 & 1.198544 & 1e-06 & 1.198847 & 1e-06 & 1.198585 & 1e-06\\
+			\hline
+			3rd & 1.202419 & 1e-06 & 1.202136 & 1e-06 & 1.202294 & 1e-06 & 1.202288 & 1e-06\\
+			\hline
+			4th & 1.210605 & 1e-06 & 1.210605 & 1e-06 & 1.210274 & 1e-06 & 1.210259 & 1e-06\\
+			\hline
+			Linear SIR & 1.274256 & 1e-06 & 1.273828 & 1e-06 & 1.274260 & 1e-06 & 1.274030 & 1e-06\\
 			\hline
@@
 			\hline
-			Baseline 1 & 0.864913 & 0.203402 & 1.739858 & 0.122632 & 1.359771 & 0.919629 & 1.173245 & 0.096355\\
-			\hline
-			Baseline 2 & 1.485664 & 1.654031 & 1.294412 & 0.618155 & 1.461580 & 0.734907 & 1.186621 & 0.091633\\
-			\hline
-			Baseline 3 & 1.082572 & 0.088142 & 1.389817 & 0.117032 & 0.615994 & 1.496674 & 1.088364 & 0.089282\\
-			\hline
-			Baseline 4 & 1.088362 & 0.367110 & 1.370251 & 0.159221 & 2.683109 & 27.981877 & 1.090398 & 0.059515\\
-			\hline
-			Baseline 5 & 0.871650 & 0.100744 & 1.438190 & 0.382650 & 0.829591 & 0.120911 & 0.924238 & 0.188305\\
+			Baseline 1 & 1.269690 & 0.120767 & 0.130935 & 10.750359 & 1.154610 & 0.065991 & 1.397649 & 0.233918\\
+			\hline
+			Baseline 2 & 1.187856 & 0.168327 & 1.121309 & 0.040461 & 1.193012 & 0.081712 & 1.357958 & 0.199538\\
+			\hline
+			Baseline 3 & 0.812981 & 0.084419 & 1.092512 & 0.034560 & 1.179720 & 0.054902 & 1.233221 & 0.060551\\
+			\hline
+			Baseline 4 & 1.180116 & 0.078889 & 1.168589 & 0.134484 & 1.293454 & 0.271027 & 1.165880 & 0.036177\\
+			\hline
+			Baseline 5 & 0.853203 & 0.064872 & 1.116949 & 0.041290 & 1.198555 & 0.106425 & 1.161485 & 0.032157\\
 			\hline
@@
 			\hline
-			$X_0 = 99999, Y_0 = 1$ & 1.713333 & 1.757278 & 1.187769 & 0.515629 & 1.451833 & 59.746089 & 1.068633 & 1.809596\\
-			\hline
-			$X_0 = 99990, Y_0 = 10$ & 1.197457 & 0.192831 & 0.674259 & 0.090168 & 1.095143 & 0.143146 & 1.199437 & 0.106393\\
-			\hline
-			$X_0 = 99900, Y_0 = 100$ & 0.796760 & 0.488540 & 1.022290 & 0.070029 & 0.024273 & 4.564245 & 1.265712 & 0.080303\\
-			\hline
-			$X_0 = 99000, Y_0 = 1000$ & 1.431229 & 0.105605 & 2.091993 & 0.340798 & 2.183329 & 0.428090 & 1.812900 & 0.122106\\
+			$X_0 = 99999, Y_0 = 1$ & 1.185172 & 1.285564 & 1.158778 & 0.152094 & 0.859907 & 83.111543 & 1.116840 & 0.064265\\
+			\hline
+			$X_0 = 99990, Y_0 = 10$ & 0.934239 & 3.360057 & 0.944893 & 0.034982 & 1.101154 & 0.287314 & 6.153800 & 1525.693988\\
+			\hline
+			$X_0 = 99900, Y_0 = 100$ & 1.255903 & 0.086428 & 0.462015 & 0.325984 & 1.250009 & 0.051102 & 1.193192 & 0.035157\\
+			\hline
+			$X_0 = 99000, Y_0 = 1000$ & 1.205152 & 0.005332 & 1.193656 & 0.005244 & 1.199742 & 0.004094 & 1.208671 & 0.004804\\
 			\hline
@@
 			\hline
-			$\sigma_X = 10, \sigma_Y = 1$ & -121.825007 & 904.236167 & 121.668347 & 308.632857 & 21.981032 & 21.204959 & 18.741699 & 17.132467\\
-			\hline
-			$\sigma_X = 100, \sigma_Y = 10$ & 2.918979 & 2.010263 & -1.223521 & 2.355038 & -5.683542 & 10.750736 & 4.071392 & 2.877332\\
-			\hline
-			$\sigma_X = 300, \sigma_Y = 30$ & 2.769860 & 1.759591 & 2.002654 & 0.949586 & 3.227956 & 3.934407 & 87.874782 & 456.259511\\
-			\hline
-			$\sigma_X = 1000, \sigma_Y = 100$ & 0.915434 & 0.114881 & 1.887797 & 0.672640 & 1.759919 & 1.537275 & -0.235868 & 1.519149\\
-			\hline
-			$\sigma_X = 3000, \sigma_Y = 300$ & -4.547791 & 970.222266 & 1.037930 & 0.112483 & 0.748209 & 0.437849 & 0.235119 & 1.489429\\
-			\hline
-			$\sigma_X = 10000, \sigma_Y = 1000$ & 1.033786 & 0.068722 & 0.958868 & 0.066758 & 0.862043 & 0.168471 & 0.671640 & 0.206497\\
+			$\sigma_X = 10, \sigma_Y = 1$ & 1.185958 & 0.008294 & 1.202635 & 0.006364 & 1.200712 & 0.004175 & 1.194196 & 0.003282\\
+			\hline
+			$\sigma_X = 100, \sigma_Y = 10$ & 1.080974 & 0.051432 & 1.242575 & 0.081110 & 1.221708 & 0.047705 & 1.160408 & 0.025033\\
+			\hline
+			$\sigma_X = 500, \sigma_Y = 50$ & 0.738334 & 0.237194 & 1.506353 & 7.578001 & 1.343488 & 0.552568 & 1.071472 & 0.060095\\
+			\hline
+			$\sigma_X = 1000, \sigma_Y = 100$ & 0.474710 & 0.832353 & 2.427907 & 19.349048 & 1.619427 & 3.245662 & 1.018452 & 0.087734\\
+			\hline
+			$\sigma_X = 2000, \sigma_Y = 200$ & 0.184986 & 2.306650 & -1.221597 & 233.262939 & 9.803881 & 71195.968128 & 0.969270 & 0.132646\\
+			\hline
+			$\sigma_X = 5000, \sigma_Y = 500$ & -0.130854 & 10.533729 & 0.351591 & 7.021734 & 0.389501 & 23.012536 & 0.916922 & 0.188977\\
+			\hline
+			$\sigma_X = 7500, \sigma_Y = 750$ & -0.225912 & 37.943501 & 0.472891 & 14.985760 & 0.546636 & 21.663742 & 0.895990 & 0.203207\\
+			\hline
+			$\sigma_X = 10000, \sigma_Y = 1000$ & -0.276013 & 174.714931 & 0.530827 & 128.425370 & 0.603731 & 28.923054 & 0.880411 & 0.207996\\
+			\hline
+			$\sigma_X = 15000, \sigma_Y = 1500$ & -0.324210 & 28.295186 & 0.598720 & 14.309696 & 0.650612 & 8.469867 & 0.855473 & 0.206941\\
 			\hline
@@
 			\hline
-			Baseline 1 & 2.015772 & 27.518508 & 2.150988 & 4.474655 & 1.988515 & 0.232390 & 2.103188 & 0.636825\\
-			\hline
-			Baseline 2 & 2.571202 & 21.048981 & -0.131021 & 51.293391 & 2.192838 & 1.684886 & 2.189259 & 1.514415\\
-			\hline
-			Baseline 3 & 1.201407 & 21.526024 & 3.023289 & 26.322188 & 2.079618 & 1.027536 & 2.007048 & 0.269511\\
-			\hline
-			Baseline 4 & 1.991370 & 1.164132 & 3.604925 & 35.619793 & 2.028025 & 0.205708 & 1.875085 & 2.080861\\
-			\hline
-			Baseline 5 & 1.873964 & 7.146565 & 1.032642 & 11.871804 & 1.372848 & 10.164198 & 1.926002 & 0.166963\\
+			Baseline 1 & 1.171311 & 0.029553 & 1.169409 & 0.077379 & 1.171164 & 0.028320 & 1.173413 & 0.034820\\
+			\hline
+			Baseline 2 & 1.170936 & 0.027339 & 1.172363 & 0.032062 & 1.171630 & 0.027987 & 1.173249 & 0.032884\\
+			\hline
+			Baseline 3 & 1.023937 & 2.701492 & 1.168652 & 0.029267 & 1.171384 & 0.028419 & 1.171662 & 0.028700\\
+			\hline
+			Baseline 4 & 1.171501 & 0.028128 & 1.174232 & 0.033153 & 1.173115 & 0.031369 & 1.171094 & 0.028392\\
+			\hline
+			Baseline 5 & 1.178276 & 0.303469 & 1.167810 & 0.027766 & 1.170475 & 0.027259 & 1.171001 & 0.027792\\
 			\hline
@@
 		\begin{tabular}{l|r|r|r|r|r|r|r|r}
-			Data Set & Auto Est & Auto SE & AutoM Est & AutoM SE & Norm Est & Norm SE & NormM Est & NormM SE\\
-			\hline
-			$X_0 = 99999, Y_0 = 1$ & 2.558602 & 32.007328 & 1.296284 & 9.652181 & 1.178839 & 3.901942 & 1.287102 & 0.524929\\
-			\hline
-			$X_0 = 99990, Y_0 = 10$ & 0.745127 & 3.487290 & 1.541772 & 1.793288 & 3.227689 & 26.102558 & 2.489264 & 26.755104\\
-			\hline
-			$X_0 = 99900, Y_0 = 100$ & 2.082023 & 2.181500 & -1.745769 & 81.529981 & 1.946617 & 1.232307 & 2.008181 & 0.191259\\
-			\hline
-			$X_0 = 99000, Y_0 = 1000$ & 1.969496 & 0.088997 & 1.979588 & 0.093847 & 1.972052 & 0.036176 & 1.964677 & 0.028405\\
+			\hline
+			Data Set & Auto Est & Auto SE & AutoM Est & AutoM SE & Norm Est & Norm SE & NormM Est & NormM SE\\
+			\hline
+			$X_0 = 99999, Y_0 = 1$ & 1.422018 & 8.968638 & 1.188044 & 0.101487 & 1.151807 & 0.275409 & 1.195600 & 0.033124\\
+			\hline
+			$X_0 = 99990, Y_0 = 10$ & 1.191840 & 0.033509 & 1.178169 & 0.160530 & 1.194324 & 0.017020 & 1.209180 & 0.263013\\
+			\hline
+			$X_0 = 99900, Y_0 = 100$ & 1.148868 & 0.047200 & 1.144860 & 0.232458 & 1.149893 & 0.047661 & 1.149490 & 0.046846\\
+			\hline
+			$X_0 = 99000, Y_0 = 1000$ & 0.995113 & 0.120333 & 0.994869 & 0.120456 & 0.994867 & 0.120621 & 0.995126 & 0.120503\\
 			\hline
@@
 			\hline
-			$\sigma_X = 10, \sigma_Y = 1$ & 5.556294 & 63.868004 & 2.843995 & 7.555433 & 2.277135 & 2.022040 & 2.234907 & 1.759312\\
-			\hline
-			$\sigma_X = 100, \sigma_Y = 10$ & 2.020666 & 0.483996 & 1.802373 & 5.248770 & 1.891754 & 5.445302 & 2.034306 & 0.527914\\
-			\hline
-			$\sigma_X = 300, \sigma_Y = 30$ & -1.064869 & 50.543183 & 4.685491 & 53.111296 & 2.076773 & 0.646978 & 2.521401 & 5.276258\\
-			\hline
-			$\sigma_X = 1000, \sigma_Y = 100$ & 0.999874 & 3.992411 & 2.390182 & 3.320493 & 1.456545 & 6.796468 & 1.703567 & 2.395295\\
-			\hline
-			$\sigma_X = 3000, \sigma_Y = 300$ & -2.740783 & 70.416107 & 1.371659 & 1.347889 & 0.030795 & 21.459099 & 2.023080 & 6.341932\\
-			\hline
-			$\sigma_X = 10000, \sigma_Y = 1000$ & 0.877306 & 0.815504 & 1.106179 & 2.530188 & 1.183491 & 3.063114 & 0.949453 & 6.173692\\
+			$\sigma_X = 10, \sigma_Y = 1$ & 0.994680 & 0.119919 & 0.995029 & 0.120570 & 0.994850 & 0.120592 & 0.994886 & 0.120268\\
+			\hline
+			$\sigma_X = 100, \sigma_Y = 10$ & 0.993618 & 0.116092 & 0.996562 & 0.121634 & 0.994779 & 0.121739 & 0.994844 & 0.119207\\
+			\hline
+			$\sigma_X = 500, \sigma_Y = 50$ & 1.018998 & 0.741279 & 1.003635 & 0.132436 & 0.994571 & 0.130529 & 0.992271 & 0.121888\\
+			\hline
+			$\sigma_X = 1000, \sigma_Y = 100$ & 0.976732 & 0.194268 & 1.015891 & 0.176727 & 0.994170 & 0.159640 & 0.986805 & 0.133885\\
+			\hline
+			$\sigma_X = 2000, \sigma_Y = 200$ & 0.584095 & 6.159187 & 1.131111 & 2.595232 & 1.118248 & 1.135767 & 0.954122 & 0.243263\\
+			\hline
+			$\sigma_X = 5000, \sigma_Y = 500$ & 2.054850 & 17.230026 & 0.731612 & 3.124131 & 1.018859 & 0.714739 & 1.087380 & 0.674374\\
+			\hline
+			$\sigma_X = 7500, \sigma_Y = 750$ & 1.326042 & 5.616124 & 0.741682 & 6.293590 & 1.061083 & 1.608632 & 0.914812 & 2.034648\\
+			\hline
+			$\sigma_X = 10000, \sigma_Y = 1000$ & -0.424508 & 24.406119 & 0.728716 & 7.478691 & 0.984529 & 0.261810 & 0.677480 & 4.972663\\
+			\hline
+			$\sigma_X = 15000, \sigma_Y = 1500$ & -0.352152 & 22.171481 & 0.774077 & 3.857982 & 0.984145 & 0.158089 & 0.980721 & 1.294957\\
 			\hline
@@
 			\hline
-			Baseline 1 & 1.002232 & NA & 1.002359 & NA & 1.002498 & NA & 1.002531 & NA\\
-			\hline
-			Baseline 2 & 1.002405 & NA & 1.002802 & NA & 1.002364 & NA & 1.002561 & NA\\
-			\hline
-			Baseline 3 & 1.002261 & NA & 1.002393 & NA & 1.002516 & NA & 1.002501 & NA\\
-			\hline
-			Baseline 4 & 1.002469 & NA & 1.002494 & NA & 1.002511 & NA & 1.002481 & NA\\
-			\hline
-			Baseline 5 & 1.002392 & NA & 1.002496 & NA & 1.002509 & NA & 1.002529 & NA\\
+			Baseline 1 & 1.091592 & NA & 1.091985 & NA & 1.091542 & NA & 1.091488 & NA\\
+			\hline
+			Baseline 2 & 1.091390 & NA & 1.091563 & NA & 1.091537 & NA & 1.091717 & NA\\
+			\hline
+			Baseline 3 & 1.091540 & NA & 1.091211 & NA & 1.091819 & NA & 1.091724 & NA\\
+			\hline
+			Baseline 4 & 1.091405 & NA & 1.091566 & NA & 1.091414 & NA & 1.091801 & NA\\
+			\hline
+			Baseline 5 & 1.091637 & NA & 1.091549 & NA & 1.091627 & NA & 1.091556 & NA\\
 			\hline
@@
 			\hline
-			$X_0 = 99999, Y_0 = 1$ & 0.999831 & NA & 1.000028 & NA & 0.999962 & NA & 1.000079 & NA\\
-			\hline
-			$X_0 = 99990, Y_0 = 10$ & 1.000442 & NA & 1.000473 & NA & 1.000514 & NA & 1.000527 & NA\\
-			\hline
-			$X_0 = 99900, Y_0 = 100$ & 1.005071 & NA & 1.004938 & NA & 1.005222 & NA & 1.004948 & NA\\
-			\hline
-			$X_0 = 99000, Y_0 = 1000$ & 1.049518 & NA & 1.050139 & NA & 1.050143 & NA & 1.050157 & NA\\
+			$X_0 = 99999, Y_0 = 1$ & 1.002195 & NA & 1.002221 & NA & 1.002225 & NA & 1.002435 & NA\\
+			\hline
+			$X_0 = 99990, Y_0 = 10$ & 1.021510 & NA & 1.021948 & NA & 1.021505 & NA & 1.021804 & NA\\
+			\hline
+			$X_0 = 99900, Y_0 = 100$ & 1.154799 & NA & 1.155005 & NA & 1.155144 & NA & 1.155213 & NA\\
+			\hline
+			$X_0 = 99000, Y_0 = 1000$ & 1.202948 & NA & 1.201580 & NA & 1.199981 & NA & 1.196597 & NA\\
 			\hline
@@
 			\hline
-			$\sigma_X = 10, \sigma_Y = 1$ & 1.050097 & NA & 1.049897 & NA & 1.050034 & NA & 1.050098 & NA\\
-			\hline
-			$\sigma_X = 100, \sigma_Y = 10$ & 1.048816 & NA & 1.051191 & NA & 1.051560 & NA & 1.051580 & NA\\
-			\hline
-			$\sigma_X = 300, \sigma_Y = 30$ & 1.045130 & NA & 1.054815 & NA & 1.042187 & NA & 1.049682 & NA\\
-			\hline
-			$\sigma_X = 1000, \sigma_Y = 100$ & 1.038396 & NA & 1.051101 & NA & 1.057586 & NA & 1.049022 & NA\\
-			\hline
-			$\sigma_X = 3000, \sigma_Y = 300$ & 1.005515 & NA & 1.081383 & NA & 1.002521 & NA & 1.064796 & NA\\
-			\hline
-			$\sigma_X = 10000, \sigma_Y = 1000$ & 1.076870 & NA & 1.114340 & NA & 1.077148 & NA & 1.037306 & NA\\
+			$\sigma_X = 10, \sigma_Y = 1$ & 1.200091 & NA & 1.199638 & NA & 1.203322 & NA & 1.198214 & NA\\
+			\hline
+			$\sigma_X = 100, \sigma_Y = 10$ & 1.192908 & NA & 1.206314 & NA & 1.204461 & NA & 1.189191 & NA\\
+			\hline
+			$\sigma_X = 500, \sigma_Y = 50$ & 1.164184 & NA & 1.198797 & NA & 1.209550 & NA & 1.220847 & NA\\
+			\hline
+			$\sigma_X = 1000, \sigma_Y = 100$ & 1.177169 & NA & 1.226796 & NA & 1.215972 & NA & 1.223427 & NA\\
+			\hline
+			$\sigma_X = 2000, \sigma_Y = 200$ & 1.204027 & NA & 1.212075 & NA & 1.175607 & NA & 1.228619 & NA\\
+			\hline
+			$\sigma_X = 5000, \sigma_Y = 500$ & 1.292498 & NA & 1.169958 & NA & 1.208613 & NA & 1.348023 & NA\\
+			\hline
+			$\sigma_X = 7500, \sigma_Y = 750$ & 1.376802 & NA & 1.137033 & NA & 1.237567 & NA & 1.359960 & NA\\
+			\hline
+			$\sigma_X = 10000, \sigma_Y = 1000$ & 1.288801 & NA & 1.105911 & NA & 1.075749 & NA & 1.372110 & NA\\
+			\hline
+			$\sigma_X = 15000, \sigma_Y = 1500$ & 1.228367 & NA & 1.158778 & NA & 0.978885 & NA & 1.397074 & NA\\
 			\hline
@@
 			\hline
-			Baseline 1 & 1.002424 & 2.3e-05 & 1.002460 & 1.1e-05 & 1.002435 & 1.5e-05 & 1.002459 & 8e-06\\
-			\hline
-			Baseline 2 & 1.002507 & 2.2e-05 & 1.002444 & 1.8e-05 & 1.002462 & 1.9e-05 & 1.002436 & 7e-06\\
-			\hline
-			Baseline 3 & 1.002374 & 1.9e-05 & 1.002418 & 5.0e-06 & 1.002439 & 1.7e-05 & 1.002445 & 5e-06\\
-			\hline
-			Baseline 4 & 1.002434 & 1.5e-05 & 1.002479 & 8.0e-06 & 1.002461 & 1.5e-05 & 1.002449 & 6e-06\\
-			\hline
-			Baseline 5 & 1.002403 & 2.0e-05 & 1.002469 & 9.0e-06 & 1.002439 & 1.6e-05 & 1.002468 & 6e-06\\
+			Baseline 1 & 1.085556 & 0.000544 & 1.085589 & 0.000546 & 1.085551 & 0.000538 & 1.085568 & 0.000534\\
+			\hline
+			Baseline 2 & 1.085512 & 0.000531 & 1.085507 & 0.000536 & 1.085553 & 0.000534 & 1.085583 & 0.000536\\
+			\hline
+			Baseline 3 & 1.085561 & 0.000542 & 1.085528 & 0.000538 & 1.085557 & 0.000539 & 1.085600 & 0.000538\\
+			\hline
+			Baseline 4 & 1.085540 & 0.000527 & 1.085527 & 0.000537 & 1.085556 & 0.000536 & 1.085547 & 0.000538\\
+			\hline
+			Baseline 5 & 1.085546 & 0.000531 & 1.085561 & 0.000531 & 1.085545 & 0.000535 & 1.085510 & 0.000535\\
 			\hline
@@
 			\hline
-			$X_0 = 99999, Y_0 = 1$ & 1.000116 & 1.7e-05 & 1.000063 & 2.0e-06 & 1.000040 & 1.5e-05 & 1.000064 & 1.0e-06\\
-			\hline
-			$X_0 = 99990, Y_0 = 10$ & 1.000474 & 2.1e-05 & 1.000477 & 8.0e-06 & 1.000473 & 1.6e-05 & 1.000507 & 4.0e-06\\
-			\hline
-			$X_0 = 99900, Y_0 = 100$ & 1.004902 & 1.8e-05 & 1.004880 & 1.4e-05 & 1.004899 & 1.9e-05 & 1.004895 & 8.0e-06\\
-			\hline
-			$X_0 = 99000, Y_0 = 1000$ & 1.049155 & 7.4e-05 & 1.049349 & 6.6e-05 & 1.049271 & 6.9e-05 & 1.049277 & 6.6e-05\\
+			$X_0 = 99999, Y_0 = 1$ & 1.001996 & 0.000024 & 1.002034 & 0.000029 & 1.002019 & 0.000026 & 1.002030 & 0.000021\\
+			\hline
+			$X_0 = 99990, Y_0 = 10$ & 1.019557 & 0.000186 & 1.019528 & 0.000183 & 1.019498 & 0.000179 & 1.019483 & 0.000175\\
+			\hline
+			$X_0 = 99900, Y_0 = 100$ & 1.148891 & 0.000569 & 1.148956 & 0.000563 & 1.148938 & 0.000580 & 1.148941 & 0.000581\\
+			\hline
+			$X_0 = 99000, Y_0 = 1000$ & 1.201700 & 0.000907 & 1.201844 & 0.000881 & 1.201772 & 0.000939 & 1.201768 & 0.000898\\
 			\hline
@@
 			\hline
-			$\sigma_X = 10, \sigma_Y = 1$ & 1.049351 & 0.000076 & 1.049219 & 0.000065 & 1.049294 & 0.000069 & 1.049283 & 0.000069\\
-			\hline
-			$\sigma_X = 100, \sigma_Y = 10$ & 1.049675 & 0.000154 & 1.049504 & 0.000132 & 1.049443 & 0.000182 & 1.049389 & 0.000131\\
-			\hline
-			$\sigma_X = 300, \sigma_Y = 30$ & 1.046190 & 0.000566 & 1.051219 & 0.000676 & 1.050137 & 0.000623 & 1.049671 & 0.000141\\
-			\hline
-			$\sigma_X = 1000, \sigma_Y = 100$ & 1.061907 & 0.002419 & 1.048793 & 0.000363 & 1.046010 & 0.001880 & 1.048815 & 0.000172\\
-			\hline
-			$\sigma_X = 3000, \sigma_Y = 300$ & 1.045117 & 0.005651 & 1.068145 & 0.004304 & 1.039168 & 0.005444 & 1.057089 & 0.000372\\
-			\hline
-			$\sigma_X = 10000, \sigma_Y = 1000$ & 0.979174 & 0.014889 & 1.111583 & 0.002687 & 1.039000 & 0.015012 & 1.070210 & 0.004720\\
+			$\sigma_X = 10, \sigma_Y = 1$ & 1.201804 & 0.000884 & 1.201778 & 0.000936 & 1.201741 & 0.000950 & 1.201767 & 0.000849\\
+			\hline
+			$\sigma_X = 100, \sigma_Y = 10$ & 1.202440 & 0.001007 & 1.202091 & 0.000898 & 1.201573 & 0.000962 & 1.202069 & 0.000931\\
+			\hline
+			$\sigma_X = 500, \sigma_Y = 50$ & 1.205047 & 0.001527 & 1.203529 & 0.001342 & 1.200705 & 0.001254 & 1.203260 & 0.001214\\
+			\hline
+			$\sigma_X = 1000, \sigma_Y = 100$ & 1.208288 & 0.002605 & 1.205228 & 0.002166 & 1.199591 & 0.001887 & 1.204841 & 0.001929\\
+			\hline
+			$\sigma_X = 2000, \sigma_Y = 200$ & 1.214838 & 0.004880 & 1.208625 & 0.004127 & 1.197394 & 0.003492 & 1.207986 & 0.003637\\
+			\hline
+			$\sigma_X = 5000, \sigma_Y = 500$ & 1.234608 & 0.011933 & 1.218729 & 0.009843 & 1.190987 & 0.008460 & 1.218687 & 0.009868\\
+			\hline
+			$\sigma_X = 7500, \sigma_Y = 750$ & 1.251420 & 0.017892 & 1.226979 & 0.014483 & 1.185867 & 0.012567 & 1.229087 & 0.016029\\
+			\hline
+			$\sigma_X = 10000, \sigma_Y = 1000$ & 1.268381 & 0.023922 & 1.235083 & 0.019214 & 1.180831 & 0.016460 & 1.241403 & 0.023960\\
+			\hline
+			$\sigma_X = 15000, \sigma_Y = 1500$ & 1.303020 & 0.035872 & 1.251167 & 0.028997 & 1.171122 & 0.024414 & 1.273194 & 0.037158\\
 			\hline
@@
 			\hline
-			Data Set & Auto Est & Auto SE & AutoM Est & AutoM SE & Norm Est & Norm SE & NormM Est & NormM SE\\
-			\hline
-			Baseline 1 & 1.170709 & 1.726236 & 0.118299 & 0.501111 & 1.566640 & 2.461139 & 0.129870 & 0.410478\\
-			\hline
-			Baseline 2 & 1.169642 & 1.736295 & 0.134667 & 0.531973 & 1.405544 & 1.947442 & 0.143524 & 0.470087\\
-			\hline
-			Baseline 3 & 1.024440 & 1.566376 & 0.147574 & 0.616955 & 1.327697 & 2.039847 & 0.135604 & 0.415505\\
-			\hline
-			Baseline 4 & 1.078543 & 1.635074 & 0.126040 & 0.518177 & 1.538235 & 2.350329 & 0.136590 & 0.461818\\
-			\hline
-			Baseline 5 & 1.082233 & 1.600320 & 0.130253 & 0.464518 & 1.287763 & 2.121584 & 0.124392 & 0.379142\\
+			Baseline 1 & 1.568114 & NA & 1.171336 & NA & 1.332469 & NA & 1.062616 & NA\\
+			\hline
+			Baseline 2 & 1.273631 & NA & 1.287347 & NA & 1.287487 & NA & 1.324252 & NA\\
+			\hline
+			Baseline 3 & 1.302008 & NA & 1.352183 & NA & 1.528039 & NA & 1.059422 & NA\\
+			\hline
+			Baseline 4 & 1.181609 & NA & 1.001200 & NA & 1.624614 & NA & 1.128520 & NA\\
+			\hline
+			Baseline 5 & 1.168529 & NA & 1.278128 & NA & 1.289007 & NA & 1.207690 & NA\\
 			\hline
@@
 			\hline
-			$X_0 = 99999, Y_0 = 1$ & 132.003557 & 979.057271 & 0.637136 & 7.591563 & 571.022230 & 8311.989460 & 0.917525 & 10.386777\\
-			\hline
-			$X_0 = 99990, Y_0 = 10$ & 5.566424 & 8.249684 & 0.169114 & 1.056522 & 6.817065 & 11.495275 & 0.152642 & 0.759693\\
-			\hline
-			$X_0 = 99900, Y_0 = 100$ & 0.595913 & 0.859870 & 0.129120 & 0.359819 & 0.764240 & 1.162178 & 0.121305 & 0.305982\\
-			\hline
-			$X_0 = 99000, Y_0 = 1000$ & 0.132387 & 0.115126 & 0.117280 & 0.105368 & 0.146490 & 0.145274 & 0.117695 & 0.116146\\
+			$X_0 = 99999, Y_0 = 1$ & 45.345556 & NA & 2.018195 & NA & 62.416323 & NA & 1.301719 & NA\\
+			\hline
+			$X_0 = 99990, Y_0 = 10$ & 2.756187 & NA & 1.161055 & NA & 3.991245 & NA & 1.288459 & NA\\
+			\hline
+			$X_0 = 99900, Y_0 = 100$ & 1.250277 & NA & 1.199468 & NA & 1.237907 & NA & 1.291419 & NA\\
+			\hline
+			$X_0 = 99000, Y_0 = 1000$ & 0.834190 & NA & 0.969512 & NA & 0.900419 & NA & 0.936967 & NA\\
 			\hline
@@
 			\hline
-			$\sigma_X = 10, \sigma_Y = 1$ & 1.178938 & 20.02418 & 1.185635 & 20.439407 & 1.219909 & 20.39467 & 1.167064 & 20.057929\\
-			\hline
-			$\sigma_X = 100, \sigma_Y = 10$ & 1.356012 & 15.38756 & 1.089050 & 18.415936 & 1.859016 & 20.12416 & 1.150239 & 19.649800\\
-			\hline
-			$\sigma_X = 300, \sigma_Y = 30$ & 3.142258 & 28.52150 & 0.969976 & 15.934410 & 3.201584 & 21.35673 & 1.085100 & 18.379016\\
-			\hline
-			$\sigma_X = 1000, \sigma_Y = 100$ & 7.305927 & 39.63369 & 1.088018 & 18.236155 & 7.978626 & 16.64830 & 0.787073 & 12.417881\\
-			\hline
-			$\sigma_X = 3000, \sigma_Y = 300$ & 16.009523 & 27.38954 & 0.424419 & 3.054379 & 30.351952 & 130.75442 & 1.357313 & 23.125797\\
-			\hline
-			$\sigma_X = 10000, \sigma_Y = 1000$ & 142.030295 & 894.32779 & 0.718062 & 8.362043 & 174.629326 & 706.02904 & 0.390140 & 3.444363\\
+			$\sigma_X = 10, \sigma_Y = 1$ & 0.985846 & NA & 0.986994 & NA & 0.983782 & NA & 0.985415 & NA \\
+			\hline
+			$\sigma_X = 100, \sigma_Y = 10$ & 1.237596 & NA & 1.209899 & NA & 1.338217 & NA & 1.323149 & NA\\
+			\hline
+			$\sigma_X = 500, \sigma_Y = 50$ & 3.544004 & NA & 3.412129 & NA & 4.189569 & NA & 4.080100 & NA\\
+			\hline
+			$\sigma_X = 1000, \sigma_Y = 100$ & 6.610460 & NA & 6.295422 & NA & 7.893167 & NA & 7.650749 & NA\\
+			\hline
+			$\sigma_X = 2000, \sigma_Y = 200$ & 13.301341 & NA & 12.235214 & NA & 15.500513 & NA & 15.026026 & NA\\
+			\hline
+			$\sigma_X = 5000, \sigma_Y = 500$ & 62.640210 & NA & 57.764547 & NA & 49.217485 & NA & 104.547880 & NA\\
+			\hline
+			$\sigma_X = 7500, \sigma_Y = 750$ & 71.225178 & NA & 105.181361 & NA & 320.428071 & NA & 107.391030 & NA\\
+			\hline
+			$\sigma_X = 10000, \sigma_Y = 1000$ & 116.249187 & NA & 161.948194 & NA & 126.671661 & NA & 121.517490 & NA \\
+			\hline
+			$\sigma_X = 15000, \sigma_Y = 1500$ & 235.468174 & NA & 148.873668 & NA & 220.585394 & NA & 308.061708 & NA\\
 			\hline
@@
 			\hline
-			0th & 1.779121 & 2.482162 & 0.009267 & 0.089052 & 2.352857 & 3.236806 & 0.021922 & 0.200162\\
-			\hline
-			1st & 3.392178 & 21.127161 & 5.199764 & 44.528822 & 4.615877 & 45.803986 & 4.032633 & 33.360591\\
-			\hline
-			2nd & 1.453143 & 5.808155 & 1.073020 & 2.586030 & 1.608484 & 5.985549 & 1.076302 & 2.657303\\
-			\hline
-			3rd & 1.360154 & 1.299049 & 1.192788 & 1.033645 & 1.453000 & 1.464758 & 1.195854 & 1.134779\\
-			\hline
-			4th & 1.300569 & 1.215989 & 1.300569 & 1.215989 & 1.426296 & 1.590604 & 1.153380 & 1.398890\\
-			\hline
-			Linear SIR & 1.379871 & 1.311406 & 1.208521 & 1.003065 & 1.516803 & 1.554241 & 1.208578 & 1.125440\\
+			0th & 1.706806 & NA & 0.009605 & NA & 2.134967 & NA & 0.025424 & NA\\
+			\hline
+			1st & 3.051573 & NA & 4.744311 & NA & 4.477775 & NA & 4.457355 & NA\\
+			\hline
+			2nd & 1.425110 & NA & 1.181736 & NA & 1.688859 & NA & 1.088296 & NA\\
+			\hline
+			3rd & 1.280001 & NA & 1.497495 & NA & 1.569176 & NA & 1.232182 & NA\\
+			\hline
+			4th & 1.065715 & NA & 1.284711 & NA & 1.373762 & NA & 1.148073 & NA\\
+			\hline
+			Linear SIR & 1.537546 & NA & 1.212776 & NA & 1.611864 & NA & 1.161722 & NA\\
 			\hline
@@
 			\hline
-			Baseline 1 & 0.111589 & 0.013332 & 0.116714 & 0.004233 & 0.121731 & 0.008296 & 0.122992 & 0.002911\\
-			\hline
-			Baseline 2 & 0.113638 & 0.011308 & 0.115175 & 0.003540 & 0.119221 & 0.007058 & 0.120945 & 0.005199\\
-			\hline
-			Baseline 3 & 0.118958 & 0.009720 & 0.126134 & 0.011717 & 0.121108 & 0.007978 & 0.122298 & 0.004135\\
-			\hline
-			Baseline 4 & 0.129073 & 0.010889 & 0.125410 & 0.004537 & 0.120610 & 0.008410 & 0.120497 & 0.003618\\
-			\hline
-			Baseline 5 & 0.122424 & 0.011326 & 0.129906 & 0.004269 & 0.118352 & 0.007518 & 0.124123 & 0.002414\\
+			Baseline 1 & 1.168689 & 0.211967 & 1.433300 & 0.275675 & 1.247849 & 0.225538 & 1.403578 & 0.253343\\
+			\hline
+			Baseline 2 & 1.243889 & 0.223453 & 1.169045 & 0.207609 & 1.165859 & 0.213966 & 1.539909 & 0.280209\\
+			\hline
+			Baseline 3 & 1.351741 & 0.236626 & 1.236297 & 0.210356 & 1.188403 & 0.218132 & 1.214513 & 0.219824\\
+			\hline
+			Baseline 4 & 1.222771 & 0.209546 & 1.507230 & 0.279421 & 1.260300 & 0.230086 & 1.290742 & 0.231896\\
+			\hline
+			Baseline 5 & 1.106582 & 0.190483 & 1.261039 & 0.229758 & 1.248270 & 0.227508 & 1.327386 & 0.239792\\
 			\hline
@@
 			\hline
-			$X_0 = 99999, Y_0 = 1$ & 0.496556 & 0.404930 & 0.249343 & 0.080732 & 0.244347 & 0.534660 & 0.279914 & 0.121538\\
-			\hline
-			$X_0 = 99990, Y_0 = 10$ & 0.082031 & 0.045989 & 0.130633 & 0.007387 & 0.116519 & 0.052715 & 0.119705 & 0.006003\\
-			\hline
-			$X_0 = 99900, Y_0 = 100$ & 0.120003 & 0.006022 & 0.119230 & 0.003320 & 0.121229 & 0.004296 & 0.120600 & 0.001531\\
-			\hline
-			$X_0 = 99000, Y_0 = 1000$ & 0.117786 & 0.001351 & 0.117888 & 0.001247 & 0.117755 & 0.001314 & 0.117804 & 0.001219\\
+			$X_0 = 99999, Y_0 = 1$ & 1.125861 & 2.218581 & 1.396896 & 0.546108 & 1.414820 & 1.323435 & 1.265761 & 0.231673\\
+			\hline
+			$X_0 = 99990, Y_0 = 10$ & 1.316003 & 0.274009 & 1.446887 & 0.251195 & 1.224239 & 0.233581 & 1.430876 & 0.234044\\
+			\hline
+			$X_0 = 99900, Y_0 = 100$ & 1.456975 & 0.246964 & 1.203267 & 0.208514 & 1.169186 & 0.197852 & 1.388754 & 0.233590\\
+			\hline
+			$X_0 = 99000, Y_0 = 1000$ & 1.061364 & 0.014436 & 0.913393 & 0.012300 & 1.027593 & 0.013776 & 1.258616 & 0.016998\\
 			\hline
@@
 			\hline
-			$\sigma_X = 10, \sigma_Y = 1$ & 0.119798 & 0.015080 & 0.119388 & 0.015098 & 0.119629 & 0.015084 & 0.119720 & 0.015063\\
-			\hline
-			$\sigma_X = 100, \sigma_Y = 10$ & 0.123180 & 0.015455 & 0.120830 & 0.015114 & 0.119951 & 0.015663 & 0.118726 & 0.015203\\
-			\hline
-			$\sigma_X = 300, \sigma_Y = 30$ & 0.108430 & 0.022619 & 0.119020 & 0.014395 & 0.118138 & 0.018715 & 0.120683 & 0.015369\\
-			\hline
-			$\sigma_X = 1000, \sigma_Y = 100$ & 0.151905 & 0.067340 & 0.122619 & 0.016990 & 0.120618 & 0.053277 & 0.138444 & 0.013843\\
-			\hline
-			$\sigma_X = 3000, \sigma_Y = 300$ & 0.148289 & 0.141669 & 0.237888 & 0.054042 & 0.121580 & 0.147039 & 0.135574 & 0.015871\\
-			\hline
-			$\sigma_X = 10000, \sigma_Y = 1000$ & -0.253735 & 0.457021 & 0.636736 & 0.092124 & 0.113880 & 0.555507 & 0.208459 & 0.018114\\
+			$\sigma_X = 10, \sigma_Y = 1$ & 1.077390 & 0.014486 & 1.076946 & 0.014585 & 1.077445 & 0.014524 & 1.077259 & 0.014413\\
+			\hline
+			$\sigma_X = 100, \sigma_Y = 10$ & 1.078355 & 0.015334 & 1.073914 & 0.016113 & 1.078901 & 0.015210 & 1.077047 & 0.014195\\
+			\hline
+			$\sigma_X = 500, \sigma_Y = 50$ & 1.082754 & 0.030646 & 1.060547 & 0.031371 & 1.085358 & 0.024514 & 1.076160 & 0.022713\\
+			\hline
+			$\sigma_X = 1000, \sigma_Y = 100$ & 1.088511 & 0.055954 & 1.044073 & 0.055612 & 1.093402 & 0.041119 & 1.075184 & 0.041831\\
+			\hline
+			$\sigma_X = 2000, \sigma_Y = 200$ & 1.100914 & 0.107709 & 1.011872 & 0.106957 & 1.109397 & 0.076725 & 1.073676 & 0.086113\\
+			\hline
+			$\sigma_X = 5000, \sigma_Y = 500$ & 1.146051 & 0.254897 & 0.920643 & 0.268296 & 1.156686 & 0.182392 & 1.072789 & 0.244303\\
+			\hline
+			$\sigma_X = 7500, \sigma_Y = 750$ & 1.194432 & 0.367353 & 0.849913 & 0.409537 & 1.195360 & 0.266558 & 1.076371 & 0.408601\\
+			\hline
+			$\sigma_X = 10000, \sigma_Y = 1000$ & 1.255238 & 0.471487 & 0.783195 & 0.557464 & 1.233433 & 0.347153 & 1.084101 & 0.611590\\
+			\hline
+			$\sigma_X = 15000, \sigma_Y = 1500$ & 1.431325 & 0.658287 & 0.659451 & 0.877489 & 1.307994 & 0.498443 & 1.113121 & 1.186676\\
 			\hline
@@
 			\hline
-			0th & 0.014270 & 0.013561 & 0.002865 & 0.000486 & -0.006574 & 0.010632 & 0.005140 & 0.001792\\
-			\hline
-			1st & -9.099024 & 94554.286222 & 2.655271 & 153.459186 & 2.339653 & 190.604213 & 2.285211 & 200.382597\\
-			\hline
-			2nd & 1.045203 & 0.201211 & 1.076390 & 0.156022 & 1.042439 & 0.205421 & 1.074918 & 0.152346\\
-			\hline
-			3rd & 1.279959 & 0.299658 & 1.290916 & 0.315063 & 1.279248 & 0.295345 & 1.280916 & 0.294307\\
-			\hline
-			4th & 1.192312 & 0.187455 & 1.192312 & 0.187455 & 1.203649 & 0.200467 & 1.202999 & 0.198628\\
-			\hline
-			Linear SIR & 1.296731 & 0.221560 & 1.304489 & 0.229988 & 1.294079 & 0.215113 & 1.294364 & 0.215810\\
+			0th & 0.016596 & 1.57720e-02 & 0.003117 & 0.000529 & -0.006329 & 0.010236 & 0.005741 & 0.002001\\
+			\hline
+			1st & -7.847035 & 8.15440e+04 & 2.832331 & 163.692234 & 2.274969 & 185.334642 & 2.106098 & 184.676745\\
+			\hline
+			2nd & 1.094937 & 2.10785e-01 & 0.956905 & 0.138703 & 0.869594 & 0.171361 & 1.179674 & 0.167192\\
+			\hline
+			3rd & 1.238637 & 2.89984e-01 & 1.446063 & 0.352928 & 1.237537 & 0.285715 & 1.230995 & 0.282838\\
+			\hline
+			4th & 1.103846 & 1.73547e-01 & 1.131708 & 0.177927 & 1.133062 & 0.188710 & 1.349997 & 0.222899\\
+			\hline
+			Linear SIR & 1.396246 & 2.38564e-01 & 1.269573 & 0.223832 & 1.305677 & 0.217041 & 1.333689 & 0.222366\\
 			\hline
@@
 			\hline
-			Baseline 1 & 1.780655 & 0.023436 & 1.983971 & 0.006583 & 1.887483 & 0.016523 & 1.977806 & 0.004671\\
-			\hline
-			Baseline 2 & 1.831444 & 0.021164 & 2.002378 & 0.010261 & 1.902733 & 0.017365 & 2.003881 & 0.003998\\
-			\hline
-			Baseline 3 & 1.839113 & 0.022491 & 1.914380 & 0.010098 & 1.906137 & 0.016586 & 1.987737 & 0.003900\\
-			\hline
-			Baseline 4 & 1.794680 & 0.017916 & 1.840147 & 0.007965 & 1.894689 & 0.016695 & 1.990857 & 0.003619\\
-			\hline
-			Baseline 5 & 1.823206 & 0.020518 & 1.892921 & 0.011397 & 1.914908 & 0.016957 & 1.965487 & 0.003611\\
+			Baseline1 & 1.199905 & 7.6e-05 & 1.199973 & 7.2e-05 & 1.200018 & 5.4e-05 & 1.199920 & 4.1e-05\\
+			\hline
+			Baseline2 & 1.200098 & 7.9e-05 & 1.200053 & 7.3e-05 & 1.200032 & 5.5e-05 & 1.200006 & 4.4e-05\\
+			\hline
+			Baseline3 & 1.200107 & 7.8e-05 & 1.199705 & 6.7e-05 & 1.200028 & 5.3e-05 & 1.199869 & 4.1e-05\\
+			\hline
+			Baseline4 & 1.200077 & 7.0e-05 & 1.200122 & 6.0e-05 & 1.200015 & 5.7e-05 & 1.199999 & 4.4e-05\\
+			\hline
+			Baseline5 & 1.199987 & 8.5e-05 & 1.200006 & 6.5e-05 & 1.199914 & 5.7e-05 & 1.200139 & 4.0e-05\\
 			\hline
@@
 			\hline
-			$X_0 = 99999, Y_0 = 1$ & 0.990097 & 0.006292 & 1.257974 & 0.013850 & 1.007582 & 0.005853 & 1.191026 & 0.003584\\
-			\hline
-			$X_0 = 99990, Y_0 = 10$ & 1.123018 & 0.021783 & 1.735611 & 0.013290 & 1.245655 & 0.024246 & 1.923291 & 0.010741\\
-			\hline
-			$X_0 = 99900, Y_0 = 100$ & 1.951514 & 0.012785 & 2.016594 & 0.005770 & 1.966805 & 0.008724 & 1.983918 & 0.002413\\
-			\hline
-			$X_0 = 99000, Y_0 = 1000$ & 2.006289 & 0.001318 & 1.995221 & 0.001014 & 2.000268 & 0.000951 & 2.000158 & 0.000544\\
+			$X_0 = 99999, Y_0 = 1$ & 1.183163 & 0.002609 & 1.196363 & 0.001372 & 1.188140 & 0.002488 & 1.192792 & 0.000554\\
+			\hline
+			$X_0 = 99990, Y_0 = 10$ & 1.200267 & 0.000295 & 1.200320 & 0.000269 & 1.200069 & 0.000259 & 1.200435 & 0.000146\\
+			\hline
+			$X_0 = 99900, Y_0 = 100$ & 1.200009 & 0.000044 & 1.200014 & 0.000043 & 1.199928 & 0.000032 & 1.200010 & 0.000024\\
+			\hline
+			$X_0 = 99000, Y_0 = 1000$ & 1.199986 & 0.000013 & 1.199981 & 0.000013 & 1.200009 & 0.000010 & 1.200006 & 0.000010\\
 			\hline
@@
 			\hline
-			Data Set & Auto Est & Auto SE & AutoM Est & AutoM SE & Norm Est & Norm SE & NormM Est & NormM SE\\
-			\hline
-			$\sigma_X = 10, \sigma_Y = 1$ & 1.997497 & 0.001357 & 2.002070 & 0.001141 & 2.000316 & 0.001015 & 2.000543 & 0.000521\\
-			\hline
-			$\sigma_X = 100, \sigma_Y = 10$ & 1.918862 & 0.009993 & 1.990296 & 0.004519 & 1.968331 & 0.009515 & 2.002762 & 0.002228\\
-			\hline
-			$\sigma_X = 300, \sigma_Y = 30$ & 1.688924 & 0.027688 & 1.822846 & 0.012972 & 1.786604 & 0.021708 & 1.999747 & 0.003704\\
-			\hline
-			$\sigma_X = 1000, \sigma_Y = 100$ & 1.202346 & 0.017004 & 1.860911 & 0.008679 & 1.244529 & 0.023070 & 1.863444 & 0.009948\\
-			\hline
-			$\sigma_X = 3000, \sigma_Y = 300$ & 1.043406 & 0.009591 & 1.414679 & 0.009365 & 1.027725 & 0.010276 & 1.696223 & 0.003871\\
-			\hline
-			$\sigma_X = 10000, \sigma_Y = 1000$ & 1.044556 & 0.008612 & 1.093328 & 0.004175 & 1.021139 & 0.007466 & 1.203591 & 0.007232\\
+			Error & Auto Est & Auto SE & AutoM Est & AutoM SE & Norm Est & Norm SE & NormM Est & NormM SE\\
+			\hline
+			$\sigma_X = 10, \sigma_Y = 1$ & 1.200000 & 0.000014 & 1.200036 & 0.000013 & 1.199988 & 0.000010 & 1.199994 & 0.000010\\
+			\hline
+			$\sigma_X = 100, \sigma_Y = 10$ & 1.200006 & 0.000142 & 1.200377 & 0.000133 & 1.199883 & 0.000102 & 1.199945 & 0.000097\\
+			\hline
+			$\sigma_X = 500, \sigma_Y = 50$ & 1.200213 & 0.000712 & 1.202158 & 0.000666 & 1.199521 & 0.000506 & 1.199785 & 0.000482\\
+			\hline
+			$\sigma_X = 1000, \sigma_Y = 100$ & 1.200917 & 0.001424 & 1.204990 & 0.001330 & 1.199297 & 0.001008 & 1.199719 & 0.000959\\
+			\hline
+			$\sigma_X = 2000, \sigma_Y = 200$ & 1.203936 & 0.002827 & 1.212506 & 0.002627 & 1.199576 & 0.001988 & 1.200005 & 0.001892\\
+			\hline
+			$\sigma_X = 5000, \sigma_Y = 500$ & 1.225212 & 0.006669 & 1.244598 & 0.006057 & 1.205086 & 0.004659 & 1.203631 & 0.004414\\
+			\hline
+			$\sigma_X = 7500, \sigma_Y = 750$ & 1.254099 & 0.009665 & 1.274883 & 0.008428 & 1.213266 & 0.006584 & 1.208945 & 0.006158\\
+			\hline
+			$\sigma_X = 10000, \sigma_Y = 1000$ & 1.292933 & 0.013084 & 1.306145 & 0.010690 & 1.223726 & 0.008360 & 1.215920 & 0.007672\\
+			\hline
+			$\sigma_X = 15000, \sigma_Y = 1500$ & 1.439976 & 0.028186 & 1.381843 & 0.016227 & 1.252207 & 0.012016 & 1.235788 & 0.010554\\
 			\hline
@@
 			\hline
-			0th & 0.992417 & 0.005898 & 0.801563 & 0.009989 & 0.984522 & 0.004772 & 0.960479 & 0.003184\\
-			\hline
-			1st & 1.200537 & 0.000641 & 1.200834 & 0.000664 & 1.201039 & 0.000648 & 1.200963 & 0.000653\\
-			\hline
-			2nd & 1.198169 & 0.000451 & 1.198486 & 0.000407 & 1.198510 & 0.000451 & 1.198652 & 0.000403\\
-			\hline
-			3rd & 1.207309 & 0.000196 & 1.207754 & 0.000188 & 1.207402 & 0.000184 & 1.207447 & 0.000187\\
-			\hline
-			4th & 1.210499 & 0.000136 & 1.210499 & 0.000136 & 1.210671 & 0.000121 & 1.210571 & 0.000114\\
-			\hline
-			Linear SIR & 1.258920 & 0.000994 & 1.258462 & 0.001020 & 1.258587 & 0.001008 & 1.258670 & 0.001007\\
+			0th & 1.000194 & 0.000021  & 1.000282 & 0.000013 & 1.000185 & 0.000015 & 1.000318 & 0.000018\\
+			\hline
+			1st & 1.040689 & 0.000468 & 1.040788 & 0.000469 & 1.040680 & 0.000468 & 1.040532 & 0.000467\\
+			\hline
+			2nd & 1.041906 & 0.000476 & 1.041744 & 0.000475 & 1.041800 & 0.000475 & 1.041751 & 0.000475\\
+			\hline
+			3rd & 1.046688 & 0.000504 & 1.046444 & 0.000503 & 1.046546 & 0.000504 & 1.046513 & 0.000503\\
+			\hline
+			4th & 1.041571 & 0.000473 & 1.041571 & 0.000473 & 1.041466 & 0.000473 & 1.041470 & 0.000473\\
+			\hline
+			Linear SIR & 1.061074 & 0.000585 & 1.061134 & 0.000586 & 1.061092 & 0.000585 & 1.061038 & 0.000585\\
 			\hline
@@
 			\hline
-			Baseline 1 & 5.127170 & 179.2844 & 5.234313 & 40.61000 & 5.086225 & 216.3083 & 5.100115 & 26.56028\\
-			\hline
-			Baseline 2 & 5.158918 & 189.7917 & 4.921625 & 26.13471 & 5.126360 & 217.2991 & 5.074875 & 29.08225\\
-			\hline
-			Baseline 3 & 5.160983 & 169.1971 & 5.207777 & 47.15165 & 5.083438 & 195.1501 & 5.119470 & 28.63475\\
-			\hline
-			Baseline 4 & 5.161219 & 170.6186 & 5.127197 & 34.24464 & 5.207647 & 269.4176 & 5.135429 & 31.76258\\
-			\hline
-			Baseline 5 & 5.187114 & 187.2816 & 5.131911 & 31.43095 & 5.077370 & 186.0883 & 5.098254 & 23.90021\\
+			Baseline 1 & 2.840123 & 0.346977 & 2.832794 & 0.298082 & 2.841003 & 0.427146 & 2.842369 & 0.361052\\
+			\hline
+			Baseline 2 & 2.844091 & 0.344388 & 2.840638 & 0.305295 & 2.841250 & 0.443461 & 2.838177 & 0.336637\\
+			\hline
+			Baseline 3 & 2.841266 & 0.324645 & 2.845812 & 0.316198 & 2.836290 & 0.398227 & 2.837974 & 0.365954\\
+			\hline
+			Baseline 4 & 2.843637 & 0.362766 & 2.840456 & 0.300201 & 2.843353 & 0.462407 & 2.836546 & 0.364677\\
+			\hline
+			Baseline 5 & 2.839390 & 0.335435 & 2.841471 & 0.309445 & 2.839357 & 0.445301 & 2.840635 & 0.376539\\
 			\hline
@@
 			\hline
-			$X_0 = 99999, Y_0 = 1$ & 2.502983 & 31.703174 & 4.587050 & 141.618120 & NaN & 1.53910 & 4.584606 & 211.645976\\
-			\hline
-			$X_0 = 99990, Y_0 = 10$ & 5.370030 & 1402.395267 & 4.745881 & 37.893736 & 5.079299 & 1000.80267 & 4.808880 & 30.644371\\
-			\hline
-			$X_0 = 99900, Y_0 = 100$ & 5.083530 & 78.047548 & 5.141349 & 24.913580 & 5.036936 & 94.46507 & 5.137375 & 21.960267\\
-			\hline
-			$X_0 = 99000, Y_0 = 1000$ & 5.134886 & 9.346665 & 5.112739 & 7.046171 & 5.112847 & 11.30402 & 5.113115 & 7.906187\\
+			$X_0 = 99999, Y_0 = 1$ & 2.837554 & 11.144048 & 2.923039 & 2.862449 & 2.735179 & 12.672799 & 2.706761 & 1.871524\\
+			\hline
+			$X_0 = 99990, Y_0 = 10$ & 2.846620 & 1.220266 & 2.811991 & 0.715527 & 2.844415 & 1.710048 & 2.819770 & 0.863066\\
+			\hline
+			$X_0 = 99900, Y_0 = 100$ & 2.843606 & 0.232628 & 2.841555 & 0.218771 & 2.839663 & 0.279605 & 2.839069 & 0.256531\\
+			\hline
+			$X_0 = 99000, Y_0 = 1000$ & 2.843031 & 0.115210 & 2.843140 & 0.112716 & 2.843451 & 0.141831 & 2.843141 & 0.156539\\
 			\hline
@@
 			\hline
-			$\sigma_X = 10, \sigma_Y = 1$ & 5.107658 & 40.07322 & 5.118340 & 41.49349 & 5.112780 & 42.18333 & 5.109027 & 40.04422\\
-			\hline
-			$\sigma_X = 100, \sigma_Y = 10$ & 5.047351 & 74.96928 & 5.074418 & 40.22261 & 5.055666 & 112.53820 & 5.060957 & 39.62932\\
-			\hline
-			$\sigma_X = 300, \sigma_Y = 30$ & 5.183856 & 304.55113 & 4.929632 & 34.41019 & 5.459944 & 664.73372 & 5.129073 & 49.24831\\
-			\hline
-			$\sigma_X = 1000, \sigma_Y = 100$ & 4.267588 & 158.79219 & 5.122059 & 60.44409 & 5.741570 & 4073.99286 & 5.133013 & 56.70565\\
-			\hline
-			$\sigma_X = 3000, \sigma_Y = 300$ & NaN & 74731.85279 & 3.918439 & 18.49635 & 3.200282 & 80.21230 & 4.771752 & 45.08632\\
-			\hline
-			$\sigma_X = 10000, \sigma_Y = 1000$ & 2.120699 & 17.46286 & 3.767783 & 35.76247 & NaN & 83.73480 & 3.884264 & 23.04767\\
+			$\sigma_X = 10, \sigma_Y = 1$ & 2.843163 & 0.121074 & 2.843082 & 0.117326 & 2.843563 & 0.163321 & 2.843234 & 0.150979\\
+			\hline
+			$\sigma_X = 100, \sigma_Y = 10$ & 2.841006 & 1.204736 & 2.840820 & 1.166121 & 2.845950 & 1.640665 & 2.841955 & 1.506458\\
+			\hline
+			$\sigma_X = 500, \sigma_Y = 50$ & 2.822320 & 5.786070 & 2.834664 & 5.718502 & 2.867771 & 8.562812 & 2.831940 & 7.395682\\
+			\hline
+			$\sigma_X = 1000, \sigma_Y = 100$ & 2.780071 & 10.584335 & 2.835372 & 11.341172 & 2.922494 & 19.076712 & 2.809761 & 14.170863\\
+			\hline
+			$\sigma_X = 2000, \sigma_Y = 200$ & 2.649423 & 16.091219 & 2.861883 & 23.385320 & 3.148266 & 59.622533 & 2.736588 & 24.502271\\
+			\hline
+			$\sigma_X = 5000, \sigma_Y = 500$ & 2.167158 & 14.446035 & 3.104554 & 87.360354 & NaN & 385.053384 & 2.384396 & 29.821372\\
+			\hline
+			$\sigma_X = 7500, \sigma_Y = 750$ & 1.860269 & 11.045568 & 3.451396 & 242.494249 & NaN & 26.530063 & 2.075721 & 23.380306\\
+			\hline
+			$\sigma_X = 10000, \sigma_Y = 1000$ & 1.661519 & 9.373613 & 3.871861 & 694.256270 & NaN & 8.570973 & 1.821502 & 17.932802\\
+			\hline
+			$\sigma_X = 15000, \sigma_Y = 1500$ & 1.494903 & 10.226678 & 4.113933 & 1491.975630 & NaN & 2.509313 & 1.485749 & 12.308392\\
 			\hline
@@
 		\begin{tabular}{l|r|r|r|r|r|r|r|r}
-			Data Set & Auto Est & Auto SE & AutoM Est & AutoM SE & Norm Est & Norm SE & NormM Est & NormM SE\\
-			\hline
-			Baseline 1 & 1.120551 & 0.055408 & 0.834692 & 0.047821 & 1.085113 & 0.054524 & 1.000000 & 0.052342\\
-			\hline
-			Baseline 2 & 1.163677 & 0.056464 & 0.963486 & 0.051378 & 1.086715 & 0.054565 & 1.266142 & 0.058897\\
-			\hline
-			Baseline 3 & 1.006208 & 0.052505 & 1.000000 & 0.052342 & 1.000000 & 0.052342 & 1.200584 & 0.057352\\
-			\hline
-			Baseline 4 & 0.874006 & 0.048934 & 1.200129 & 0.057341 & 0.840416 & 0.047984 & 1.000000 & 0.052342\\
-			\hline
-			Baseline 5 & 1.000000 & 0.052342 & 1.000000 & 0.052342 & 1.200097 & 0.057341 & 1.169565 & 0.056606\\
+			\hline
+			Data Set & Auto Est & Auto SE & AutoM Est & AutoM SE & Norm Est & Norm SE & NormM Est & NormM SE\\
+			\hline
+			Baseline1 & 1.201518 & 0.057374 & 1.058315 & 0.053847 & 1.239136 & 0.058266 & 1.237283 & 0.058222\\
+			\hline
+			Baseline2 & 1.246696 & 0.058443 & 1.079252 & 0.054377 & 1.124592 & 0.055507 & 1.097017 & 0.054823\\
+			\hline
+			Baseline3 & 1.039525 & 0.053367 & 1.229901 & 0.058048 & 1.253978 & 0.058614 & 1.082712 & 0.054464\\
+			\hline
+			Baseline4 & 1.118929 & 0.055368 & 1.137825 & 0.055833 & 1.215893 & 0.057717 & 1.063300 & 0.053974\\
+			\hline
+			Baseline5 & 1.086093 & 0.054549 & 1.243553 & 0.058369 & 1.050861 & 0.053657 & 1.068459 & 0.054104\\
 			\hline
@@
 			\hline
-			$X_0 = 99999, Y_0 = 1$ & 1.284355 & 0.059319 & 1.000000 & 0.052342 & 1.255944 & 0.058660 & 1.129035 & 0.055617\\
-			\hline
-			$X_0 = 99990, Y_0 = 10$ & 0.745920 & 0.045206 & 1.000000 & 0.052342 & 1.000000 & 0.052342 & 1.241239 & 0.058315\\
-			\hline
-			$X_0 = 99900, Y_0 = 100$ & 0.716836 & 0.044316 & 1.236492 & 0.058204 & 1.064823 & 0.054012 & 1.000000 & 0.052342\\
-			\hline
-			$X_0 = 99000, Y_0 = 1000$ & 0.753867 & 0.045447 & 1.158019 & 0.056326 & 1.048705 & 0.053602 & 1.322988 & 0.060205\\
+			$X_0 = 99999, Y_0 = 1$ & 0.929118 & 0.050453 & 1.000000 & 0.052342 & 0.842065 & 0.048032 & 1.216860 & 0.057740\\
+			\hline
+			$X_0 = 99990, Y_0 = 10$ & 1.038604 & 0.053343 & 1.167800 & 0.056564 & 0.877679 & 0.049037 & 1.250160 & 0.058524\\
+			\hline
+			$X_0 = 99900, Y_0 = 100$ & 1.253176 & 0.058595 & 1.069024 & 0.054119 & 1.165760 & 0.056514 & 1.077389 & 0.054330\\
+			\hline
+			$X_0 = 99000, Y_0 = 1000$ & 0.975516 & 0.051698 & 0.970667 & 0.051569 & 0.947781 & 0.050957 & 0.969026 & 0.051525\\
 			\hline
@@
 			\hline
-			Data Set & Auto Est & Auto SE & AutoM Est & AutoM SE & Norm Est & Norm SE & NormM Est & NormM SE\\
-			\hline
-			$\sigma_X = 10, \sigma_Y = 1$ & 0.793499 & 0.046626 & 0.787881 & 0.046461 & 0.762945 & 0.045719 & 0.796512 & 0.046714\\
-			\hline
-			$\sigma_X = 100, \sigma_Y = 10$ & 0.925325 & 0.050350 & 0.891440 & 0.049420 & 0.888104 & 0.049327 & 0.885803 & 0.049263\\
-			\hline
-			$\sigma_X = 300, \sigma_Y = 30$ & 0.478432 & 0.036205 & 0.510298 & 0.037391 & 0.494531 & 0.036809 & 0.502460 & 0.037103\\
-			\hline
-			$\sigma_X = 1000, \sigma_Y = 100$ & 0.897133 & 0.049577 & 0.973510 & 0.051644 & 0.524686 & 0.037914 & 0.524018 & 0.037890\\
-			\hline
-			$\sigma_X = 3000, \sigma_Y = 300$ & 1.000000 & 0.052342 & 1.000000 & 0.052342 & 1.024154 & 0.052971 & 0.984151 & 0.051926\\
-			\hline
-			$\sigma_X = 10000, \sigma_Y = 1000$ & 1.686398 & 0.067973 & 1.000000 & 0.052342 & 0.379716 & 0.032254 & 1.000000 & 0.052342\\
+			Error & Auto Est & Auto SE & AutoM Est & AutoM SE & Norm Est & Norm SE & NormM Est & NormM SE\\
+			\hline
+			$\sigma_X = 10, \sigma_Y = 1$ & 0.984301 & 0.051930 & 0.990291 & 0.052088 & 0.967635 & 0.051488 & 0.968098 & 0.051501\\
+			\hline
+			$\sigma_X = 100, \sigma_Y = 10$ & 1.014390 & 0.052718 & 1.270396 & 0.058996 & 1.000000 & 0.052342 & 1.000000 & 0.052342\\
+			\hline
+			$\sigma_X = 500, \sigma_Y = 50$ & 1.033386 & 0.053209 & 1.322754 & 0.060200 & 1.000000 & 0.052342 & 1.000000 & 0.052342\\
+			\hline
+			$\sigma_X = 1000, \sigma_Y = 100$ & 1.037446 & 0.053313 & 1.347989 & 0.060771 & 1.000000 & 0.052342 & 1.000000 & 0.052342\\
+			\hline
+			$\sigma_X = 2000, \sigma_Y = 200$ & 1.039747 & 0.053372 & 1.373882 & 0.061352 & 1.000000 & 0.052342 & 1.000000 & 0.052342\\
+			\hline
+			$\sigma_X = 5000, \sigma_Y = 500$ & 1.041231 & 0.053411 & 1.408605 & 0.062122 & 1.000000 & 0.052342 & 1.000000 & 0.052342\\
+			\hline
+			$\sigma_X = 7500, \sigma_Y = 750$ & 1.041573 & 0.053419 & 1.424062 & 0.062462 & 1.000000 & 0.052342 & 1.000000 & 0.052342\\
+			\hline
+			$\sigma_X = 10000, \sigma_Y = 1000$ & 1.041745 & 0.053424 & 1.435049 & 0.062703 & 1.000000 & 0.052342 & 1.000000 & 0.052342\\
+			\hline
+			$\sigma_X = 15000, \sigma_Y = 1500$ & 1.041919 & 0.053428 & 1.450554 & 0.063041 & 1.000000 & 0.052342 & 1.000000 & 0.052342\\
 			\hline
@@
 			\hline
-			Baseline 1 & 1.546608 & 0.171494 & 1.326568 & 0.013100 & 3.206236 & 2.451373 & 2.455754 & 0.066867\\
-			\hline
-			Baseline 2 & 1.772431 & 0.354807 & -0.080128 & 0.031367 & -12.937346 & 150.847519 & 3.675647 & 0.123981\\
-			\hline
-			Baseline 3 & 1.781222 & 0.395661 & 1.375838 & 0.024409 & 2.482198 & 1.391814 & 2.149920 & 0.029209\\
-			\hline
-			Baseline 4 & 1.636096 & 0.594579 & 1.737715 & 0.020414 & 0.616769 & 9.237073 & 1.982802 & 0.025452\\
-			\hline
-			Baseline 5 & -10.920596 & 123.661450 & 1.846941 & 0.027824 & 4.892252 & 12.254994 & 2.457392 & 0.039093\\
+			Baseline1 & 1.302870 & 0.005985 & 1.304978 & 0.004832 & 1.295983 & 0.007049 & 1.289180 & 0.006367\\
+			\hline
+			Baseline2 & 1.291782 & 0.005861 & 1.303073 & 0.004532 & 1.293540 & 0.008068 & 1.285063 & 0.004522\\
+			\hline
+			Baseline3 & 1.291790 & 0.005322 & 1.305184 & 0.004968 & 1.313588 & 0.006870 & 1.295970 & 0.007001\\
+			\hline
+			Baseline4 & 1.305709 & 0.005793 & 1.303979 & 0.005127 & 1.307232 & 0.006962 & 1.313454 & 0.006151\\
+			\hline
+			Baseline5 & 1.302877 & 0.005851 & 1.297431 & 0.004872 & 1.302505 & 0.007885 & 1.297394 & 0.006744\\
 			\hline
@@
 			\hline
-			$X_0 = 99999, Y_0 = 1$ & NA & NA & -0.182830 & 0.225877 & NA & NA & -0.484025 & 0.614292\\
-			\hline
-			$X_0 = 99990, Y_0 = 10$ & 0.973745 & 0.546827 & NA & NA & 1.052351 & 444.645186 & NA & NA\\
-			\hline
-			$X_0 = 99900, Y_0 = 100$ & -6.416001 & 25.905717 & 2.451224 & 0.118072 & -3.124388 & 6.766299 & -1.475048 & 1.343083\\
-			\hline
-			$X_0 = 99000, Y_0 = 1000$ & -1.665131 & 1.139871 & 2.132087 & 0.009851 & 2.146884 & 0.018722 & 2.114676 & 0.010444\\
+			$X_0 = 99999, Y_0 = 1$ & 1.327236 & 0.062503 & 1.305344 & 0.029386 & 1.319333 & 0.104531 & 1.486194 & 0.042930\\
+			\hline
+			$X_0 = 99990, Y_0 = 10$ & 1.320088 & 0.014535 & 1.327951 & 0.010927 & 1.323049 & 0.019898 & 1.330877 & 0.012021\\
+			\hline
+			$X_0 = 99900, Y_0 = 100$ & 1.257056 & 0.001195 & 1.252233 & 0.001181 & 1.269212 & 0.001844 & 1.196636 & 0.000635\\
+			\hline
+			$X_0 = 99000, Y_0 = 1000$ & 1.198724 & 0.000024 & 1.198618 & 0.000026 & 1.198275 & 0.000036 & 1.198736 & 0.000036\\
 			\hline
@@
 			\hline
-			Data Set & Auto Est & Auto SE & AutoM Est & AutoM SE & Norm Est & Norm SE & NormM Est & NormM SE\\
-			\hline
-			$\sigma_X = 10, \sigma_Y = 1$ & 19.226063 & 0.289480 & 21.685544 & 0.346537 & 18.468777 & 0.256194 & 23.787444 & 0.195589\\
-			\hline
-			$\sigma_X = 100, \sigma_Y = 10$ & 16.976120 & 0.530197 & 19.896647 & 0.226174 & 13.179504 & 0.570207 & 15.331206 & 0.206823\\
-			\hline
-			$\sigma_X = 300, \sigma_Y = 30$ & 9.925354 & 0.770734 & 16.239531 & 0.234976 & -4.473057 & 0.190822 & 22.862570 & 0.525209\\
-			\hline
-			$\sigma_X = 1000, \sigma_Y = 100$ & -0.007144 & 0.000245 & 25.560238 & 0.933150 & -4.082639 & 0.436790 & 11.261214 & 0.216712\\
-			\hline
-			$\sigma_X = 3000, \sigma_Y = 300$ & -0.262553 & 0.003032 & -0.012345 & 0.000093 & -0.000062 & 0.000004 & 12.886121 & 0.277206\\
-			\hline
-			$\sigma_X = 10000, \sigma_Y = 1000$ & -0.000217 & 0.000031 & -5.959291 & 0.225965 & NA & NA & -0.000064 & 0.000000\\
+			Error & Auto Est & Auto SE & AutoM Est & AutoM SE & Norm Est & Norm SE & NormM Est & NormM SE\\
+			\hline
+			$\sigma_X = 10, \sigma_Y = 1$ & 1.198667 & 0.022446 & 1.199697 & 0.000054 & 1.184111 & 0.000033 & 1.215409 & 0.003919\\
+			\hline
+			$\sigma_X = 100, \sigma_Y = 10$ & 0.738349 & 0.027172 & 0.326262 & 0.148088 & 0.876058 & 0.009180 & 1.261285 & 0.000093\\
+			\hline
+			$\sigma_X = 500, \sigma_Y = 50$ & -0.003532 & 0.056577 & -0.070816 & 0.504898 & 0.094900 & 0.691400 & 1.326638 & 0.030156\\
+			\hline
+			$\sigma_X = 1000, \sigma_Y = 100$ & 1.652869 & 0.287473 & -0.261653 & 2.358679 & NA & NA & 0.951134 & 0.007148\\
+			\hline
+			$\sigma_X = 2000, \sigma_Y = 200$ & -0.000235 & 0.079270 & -0.043704 & 0.553296 & 9.411580 & 0.032052 & 1.142685 & 0.017086\\
+			\hline
+			$\sigma_X = 5000, \sigma_Y = 500$ & 0.014666 & 0.124467 & 0.790375 & 0.015295 & -0.014291 & 0.124025 & 9.473186 & 33.693741\\
+			\hline
+			$\sigma_X = 7500, \sigma_Y = 750$ & 2.856424 & 11.133371 & 0.808988 & 0.038110 & -0.007135 & 0.109857 & NA & NA\\
+			\hline
+			$\sigma_X = 10000, \sigma_Y = 1000$ & 2.755506 & 5.698171 & 2250.629540 & 62419.537578 & 0.016792 & 0.294779 & NA & NA\\
+			\hline
+			$\sigma_X = 15000, \sigma_Y = 1500$ & NA & NA & 0.845936 & 0.056621 & 0.020971 & 0.117313 & 0.703029 & 0.145427\\
 			\hline
@@
 			\hline
-			Baseline 1 & 1.151396 & 0.007840 & 1.156166 & 0.075838 & 1.168638 & 0.061978 & 1.168638 & 0.069916\\
-			\hline
-			Baseline 2 & 1.151396 & 0.015427 & 1.151396 & 0.015825 & 1.168638 & 0.068289 & 1.168638 & 0.056477\\
-			\hline
-			Baseline 3 & 1.155559 & 0.127438 & 1.151396 & 0.012891 & 1.168638 & 0.068609 & 1.168638 & 0.052676\\
-			\hline
-			Baseline 4 & 1.156588 & 0.049150 & 1.154959 & 0.026360 & 1.168638 & 0.052003 & 1.126363 & 0.038690\\
-			\hline
-			Baseline 5 & 1.149516 & 0.024329 & 1.151396 & 0.014140 & 1.168638 & 0.060267 & 1.149516 & 0.029827\\
+			Baseline1 & 1.412982 & 0.010767 & 1.412982 & 0.009669 & 1.412982 & 0.010947 & 1.412982 & 0.010842\\
+			\hline
+			Baseline2 & 1.412982 & 0.010613 & 1.412982 & 0.010721 & 1.412982 & 0.011393 & 1.412982 & 0.010027\\
+			\hline
+			Baseline3 & 1.412982 & 0.010823 & 1.412982 & 0.009758 & 1.412982 & 0.010201 & 1.412982 & 0.009806\\
+			\hline
+			Baseline4 & 1.412982 & 0.011042 & 1.412982 & 0.009580 & 1.412982 & 0.010452 & 1.412982 & 0.011354\\
+			\hline
+			Baseline5 & 1.412982 & 0.011579 & 1.412982 & 0.009644 & 1.412982 & 0.010383 & 1.412982 & 0.009949\\
 			\hline
@@
 			\hline
-			$X_0 = 99999, Y_0 = 1$ & 0.205516 & NA & 0.220109 & NA & 0.203940 & NA & 0.220109 & NA\\
-			\hline
-			$X_0 = 99990, Y_0 = 10$ & 1.049647 & 0.040841 & 1.043576 & 0.109889 & 1.049647 & 0.042400 & 1.049647 & 0.070514\\
-			\hline
-			$X_0 = 99900, Y_0 = 100$ & 1.215624 & 0.008808 & 1.163593 & 0.005915 & 1.163593 & 0.004472 & 1.163593 & 0.007400\\
-			\hline
-			$X_0 = 99000, Y_0 = 1000$ & 1.153396 & 0.000940 & 1.143722 & 0.000599 & 1.103664 & 0.000472 & 1.153396 & 0.000970\\
+			$X_0 = 99999, Y_0 = 1$ & 0.600000 & NA & 0.600000 & NA & 0.600000 & NA & 0.600000 & NA\\
+			\hline
+			$X_0 = 99990, Y_0 = 10$ & 1.417811 & 0.009877 & 1.428718 & 0.030051 & 1.428718 & 0.025660 & 1.428718 & 0.027867\\
+			\hline
+			$X_0 = 99900, Y_0 = 100$ & 1.360667 & 0.003552 & 1.360667 & 0.004453 & 1.382555 & 0.003572 & 1.382555 & 0.003212\\
+			\hline
+			$X_0 = 99000, Y_0 = 1000$ & 1.265725 & 0.002888 & 1.265725 & 0.002807 & 1.265725 & 0.002835 & 1.265725 & 0.003064\\
 			\hline
@@
 			\hline
-			Data Set & Auto Est & Auto SE & AutoM Est & AutoM SE & Norm Est & Norm SE & NormM Est & NormM SE\\
-			\hline
-			$\sigma_X = 10, \sigma_Y = 1$ & 1.143722 & 0.000584 & 1.147266 & 0.002023 & 1.132778 & 0.000684 & 1.132778 & 0.000696\\
-			\hline
-			$\sigma_X = 100, \sigma_Y = 10$ & 1.140700 & 0.000566 & 1.171033 & 0.000785 & 1.155093 & 0.000925 & 1.183014 & 0.001369\\
-			\hline
-			$\sigma_X = 300, \sigma_Y = 30$ & 1.174448 & 0.001166 & 1.137020 & 0.002796 & 1.236429 & 0.003236 & 1.122842 & 0.001041\\
-			\hline
-			$\sigma_X = 1000, \sigma_Y = 100$ & 1.142865 & 0.000803 & 1.154025 & 0.001744 & 1.205823 & 0.003666 & 1.399630 & 0.002187\\
-			\hline
-			$\sigma_X = 3000, \sigma_Y = 300$ & 1.142642 & 0.001985 & 1.140222 & 0.000500 & 1.150901 & 0.000913 & 1.140190 & 0.000416\\
-			\hline
-			$\sigma_X = 10000, \sigma_Y = 1000$ & 1.198172 & NA & 1.147908 & 0.000934 & 1.162161 & 0.001444 & 1.157407 & 0.000959\\
+			Error & Auto Est & Auto SE & AutoM Est & AutoM SE & Norm Est & Norm SE & NormM Est & NormM SE\\
+			\hline
+			$\sigma_X = 10, \sigma_Y = 1$ & 1.265725 & 0.002680 & 1.265725 & 0.003044 & 1.265725 & 0.003271 & 1.265725 & 0.002967\\
+			\hline
+			$\sigma_X = 100, \sigma_Y = 10$ & 1.265725 & 0.003122 & 1.232704 & 0.025432 & 1.265725 & 0.002878 & 1.264345 & 0.003288\\
+			\hline
+			$\sigma_X = 500, \sigma_Y = 50$ & 1.263026 & 0.006370 & 1.269345 & 0.004386 & 1.258118 & 0.007414 & 1.255344 & 0.006144\\
+			\hline
+			$\sigma_X = 1000, \sigma_Y = 100$ & 1.234715 & 0.006079 & 1.273213 & 0.000652 & 1.258118 & 0.010512 & 1.260004 & 0.006172\\
+			\hline
+			$\sigma_X = 2000, \sigma_Y = 200$ & 1.236510 & 0.010850 & 1.246699 & 0.024112 & 1.233704 & 0.011063 & 1.219023 & 0.018803\\
+			\hline
+			$\sigma_X = 5000, \sigma_Y = 500$ & 1.285391 & 0.002340 & 1.288921 & 0.002671 & 1.222575 & 0.030681 & 1.237872 & 0.017728\\
+			\hline
+			$\sigma_X = 7500, \sigma_Y = 750$ & 1.267481 & 0.001248 & 1.265853 & 0.001334 & 1.229195 & 0.004376 & 1.208398 & 0.003815\\
+			\hline
+			$\sigma_X = 10000, \sigma_Y = 1000$ & 1.304435 & 0.001567 & 1.298856 & 0.003194 & 1.246699 & 0.006051 & 1.224308 & 0.001915\\
+			\hline
+			$\sigma_X = 15000, \sigma_Y = 1500$ & 1.334921 & 0.000388 & 1.319515 & 0.001724 & 1.228979 & 0.002101 & 1.188725 & 0.007633\\
 			\hline
@@
 			\hline
-			Baseline 1 & 1.421896 & 0.110167 & 1.246864 & 0.010846 & 1.715318 & 0.798366 & 1.781744 & 0.024938\\
-			\hline
-			Baseline 2 & 1.758858 & 0.366701 & 5.512315 & 0.320605 & 1.963196 & 0.846297 & 1.625233 & 0.018011\\
-			\hline
-			Baseline 3 & 1.758858 & 0.569505 & 1.421896 & 0.016966 & 2.297508 & 2.211150 & 1.858192 & 0.036113\\
-			\hline
-			Baseline 4 & 1.447204 & 0.091904 & 1.712243 & 0.020892 & 1.421896 & 0.341627 & 2.172764 & 0.038152\\
-			\hline
-			Baseline 5 & 1.421896 & 0.109030 & 2.162264 & 0.034685 & 1.625233 & 0.626916 & 2.011729 & 0.027506\\
+			Baseline1 & 1.4 & 0.008096 & 1.4 & 0.005942 & 1.4 & 0.010666 & 1.4 & 0.007859\\
+			\hline
+			Baseline2 & 1.4 & 0.007525 & 1.4 & 0.006043 & 1.4 & 0.009713 & 1.4 & 0.006573\\
+			\hline
+			Baseline3 & 1.4 & 0.006891 & 1.4 & 0.006580 & 1.4 & 0.008550 & 1.4 & 0.007550\\
+			\hline
+			Baseline4 & 1.4 & 0.006989 & 1.4 & 0.005960 & 1.4 & 0.009015 & 1.4 & 0.006935\\
+			\hline
+			Baseline5 & 1.4 & 0.006809 & 1.4 & 0.006459 & 1.4 & 0.009730 & 1.4 & 0.007778\\
 			\hline
@@
 			\hline
-			$X_0 = 99999, Y_0 = 1$ & 575.476314 & 5450.536541 & 205.872462 & 2972.609438 & 16.358566 & 9841.298666 & 34.387930 & 5805.567900\\
-			\hline
-			$X_0 = 99990, Y_0 = 10$ & 1.383532 & 25.019898 & 206.028474 & 18394.966469 & 3.214261 & 12206.722105 & 741.368136 & 15921.351448\\
-			\hline
-			$X_0 = 99900, Y_0 = 100$ & 4.757366 & 2.846210 & 3.819417 & 0.635151 & 15.683199 & 137.249782 & 3.874259 & 0.588683\\
-			\hline
-			$X_0 = 99000, Y_0 = 1000$ & 4.072926 & 0.317373 & 4.130517 & 0.226424 & 4.145316 & 0.359158 & 4.012320 & 0.239506\\
+			$X_0 = 99999, Y_0 = 1$ & 1.400000 & 0.085421 & 1.400000 & 0.041869 & 1.400000 & 0.159127 & 1.437675 & 0.030005\\
+			\hline
+			$X_0 = 99990, Y_0 = 10$ & 1.400000 & 0.020310 & 1.400000 & 0.012602 & 1.400000 & 0.023171 & 1.400000 & 0.015834\\
+			\hline
+			$X_0 = 99900, Y_0 = 100$ & 1.400000 & 0.004333 & 1.400000 & 0.004476 & 1.400000 & 0.005505 & 1.400000 & 0.004161\\
+			\hline
+			$X_0 = 99000, Y_0 = 1000$ & 1.200159 & 0.000029 & 1.198359 & 0.000025 & 1.198359 & 0.000028 & 1.200159 & 0.000031\\
 			\hline
@@
 			\hline
-			Data Set & Auto Est & Auto SE & AutoM Est & AutoM SE & Norm Est & Norm SE & NormM Est & NormM SE\\
-			\hline
-			$\sigma_X = 10, \sigma_Y = 1$ & 12.998364 & 8.755960e-01 & 8.061987 & 0.486050 & 9.400660 & 6.204380e-01 & 5.752931 & 0.261811\\
-			\hline
-			$\sigma_X = 100, \sigma_Y = 10$ & 5.404625 & 5.198610e-01 & 7.387983 & 0.436634 & 8.531874 & 1.705313e+00 & 5.978214 & 1.108848\\
-			\hline
-			$\sigma_X = 300, \sigma_Y = 30$ & 23.186330 & 6.441964e+02 & 5.963903 & 1.266200 & 322.144190 & 2.088921e+05 & 7.276865 & 0.663915\\
-			\hline
-			$\sigma_X = 1000, \sigma_Y = 100$ & 23.983516 & 1.278927e+04 & 67.517641 & 52.327269 & 11.529412 & 4.001174e+02 & 7.010464 & 0.567235\\
-			\hline
-			$\sigma_X = 3000, \sigma_Y = 300$ & 33.846154 & 1.163559e+07 & 21.915757 & 1527.656692 & 162.022472 & 4.124435e+03 & 12.140594 & 4.474694\\
-			\hline
-			$\sigma_X = 10000, \sigma_Y = 1000$ & 174.921348 & 3.898656e+03 & 11.524763 & 165.981987 & 1.025641 & 2.676852e+02 & 101.413031 & 1760.215196\\
+			Error & Auto Est & Auto SE & AutoM Est & AutoM SE & Norm Est & Norm SE & NormM Est & NormM SE\\
+			\hline
+			$\sigma_X = 10, \sigma_Y = 1$ & 1.213314 & 0.000028 & 1.212138 & 0.000030 & 1.213314 & 0.000034 & 1.213314 & 0.000031\\
+			\hline
+			$\sigma_X = 100, \sigma_Y = 10$ & 1.213314 & 0.000231 & 1.204838 & 0.000240 & 1.213314 & 0.000333 & 1.213314 & 0.000345\\
+			\hline
+			$\sigma_X = 500, \sigma_Y = 50$ & 1.213314 & 0.001189 & 1.204838 & 0.001234 & 1.179108 & 0.002048 & 1.213314 & 0.001537\\
+			\hline
+			$\sigma_X = 1000, \sigma_Y = 100$ & 1.198952 & 0.005372 & 1.207589 & 0.005635 & 1.198952 & 0.007195 & 1.198952 & 0.006684\\
+			\hline
+			$\sigma_X = 2000, \sigma_Y = 200$ & 1.237684 & 0.006507 & 1.249285 & 0.025702 & 1.215487 & 0.007941 & 1.198952 & 0.024341\\
+			\hline
+			$\sigma_X = 5000, \sigma_Y = 500$ & 1.210888 & 0.039465 & 0.772582 & 0.018292 & 0.002250 & 4.601591 & 1.167461 & 0.382569\\
+			\hline
+			$\sigma_X = 7500, \sigma_Y = 750$ & 1.191708 & 0.326848 & 0.820021 & 0.011286 & 0.002250 & 0.100353 & NA & NA\\
+			\hline
+			$\sigma_X = 10000, \sigma_Y = 1000$ & 3.356645 & 0.725374 & 0.836387 & 0.011665 & 0.002250 & 1.020673 & NA & NA\\
+			\hline
+			$\sigma_X = 15000, \sigma_Y = 1500$ & 2.072075 & 21.625905 & 0.655863 & 0.052597 & 0.002250 & 0.688567 & NA & NA\\
 			\hline
@@
 			\hline
-			0th & 0.001360 & 0.497415 & 0.000040 & 0.004153 & 0.037223 & 146.510831 & 0.000040 & 0.010014\\
-			\hline
-			1st & 6.217726 & 78.702007 & 13.734754 & 240.363417 & 111.232168 & 4202.221696 & 8.742980 & 66.867285\\
-			\hline
-			2nd & 2.451548 & 0.016035 & 5.220397 & 0.437529 & 8.020012 & 2.063975 & 5.220397 & 0.475516\\
-			\hline
-			3rd & 5.406024 & 0.554855 & 4.833617 & 0.256668 & 7.333991 & 1.363208 & 5.536633 & 0.660877\\
-			\hline
-			4th & 9.415812 & 2.964507 & 9.415812 & 2.964507 & 5.964387 & 0.856505 & 5.964387 & 0.748410\\
-			\hline
-			Linear SIR & 14.488463 & 6.089165 & 14.488463 & 5.828112 & 14.488463 & 6.513183 & 5.383303 & 0.308536\\
+			0th & 0.000095 & 0.009102 & 0.000082 & 0.000173 & 0.055180 & 0.017657 & 0.000023 & 0.000389\\
+			\hline
+			1st & 1.318348 & 258.153117 & 1.318348 & 5.918045 & 1.318348 & 0.628775 & 1.318348 & 8.692243\\
+			\hline
+			2nd & 1.400000 & 0.009256 & 1.400000 & 0.005901 & 1.400000 & 0.009429 & 1.400000 & 0.006330\\
+			\hline
+			3rd & 1.400000 & 0.006097 & 1.400000 & 0.006041 & 1.400000 & 0.007921 & 1.400000 & 0.007604\\
+			\hline
+			4th & 1.400000 & 0.007004 & 1.400000 & 0.006878 & 1.400000 & 0.009211 & 1.400000 & 0.008413\\
+			\hline
+			Linear SIR & 1.483791 & 0.007043 & 1.483791 & 0.006561 & 1.483791 & 0.008190 & 1.483791 & 0.007156\\
 			\hline
@@
 			\hline
-			Baseline 1 & 1.012007 & 0.010027 & 0.929414 & 0.030175 & 1.018400 & 0.009172 & 0.935858 & 0.027818\\
-			\hline
-			Baseline 2 & 1.010507 & 0.010239 & 0.943961 & 0.025078 & 1.018896 & 0.009109 & 0.936850 & 0.027469\\
-			\hline
-			Baseline 3 & 1.006199 & 0.010870 & 0.930872 & 0.029627 & 1.016426 & 0.009428 & 0.934886 & 0.028163\\
-			\hline
-			Baseline 4 & 1.006496 & 0.010825 & 0.934524 & 0.028292 & 1.018095 & 0.009211 & 0.934295 & 0.028375\\
-			\hline
-			Baseline 5 & 1.009068 & 0.010445 & 0.934444 & 0.028321 & 1.015362 & 0.009569 & 0.935845 & 0.027819\\
+			Baseline 1 & 0.928417 & 0.001317 & 0.928356 & 0.001319 & 0.928626 & 0.001310 & 0.928228 & 0.001323\\
+			\hline
+			Baseline 2 & 0.928419 & 0.001317 & 0.928247 & 0.001323 & 0.928594 & 0.001311 & 0.928289 & 0.001321\\
+			\hline
+			Baseline 3 & 0.928401 & 0.001318 & 0.928163 & 0.001326 & 0.928614 & 0.001310 & 0.928290 & 0.001321\\
+			\hline
+			Baseline 4 & 0.928395 & 0.001318 & 0.928248 & 0.001323 & 0.928611 & 0.001310 & 0.928310 & 0.001321\\
+			\hline
+			Baseline 5 & 0.928412 & 0.001317 & 0.928241 & 0.001323 & 0.928651 & 0.001309 & 0.928247 & 0.001323\\
 			\hline
@@
 		\begin{tabular}{l|r|r|r|r|r|r|r|r}
-			Data Set & Auto Est & Auto SE & AutoM Est & AutoM SE & Norm Est & Norm SE & NormM Est & NormM SE\\
-			\hline
-			$X_0 = 99999, Y_0 = 1$ & 1.004407 & 0.011143 & 0.767095 & 0.146339 & 1.016822 & 0.009376 & 0.757875 & 0.155300\\
-			\hline
-			$X_0 = 99990, Y_0 = 10$ & 1.008824 & 0.010481 & 0.885239 & 0.051071 & 1.015486 & 0.009553 & 0.877557 & 0.055606\\
-			\hline
-			$X_0 = 99900, Y_0 = 100$ & 1.012554 & 0.009951 & 0.961057 & 0.020060 & 1.019685 & 0.009009 & 0.961228 & 0.020014\\
-			\hline
-			$X_0 = 99000, Y_0 = 1000$ & 1.044120 & 0.006387 & 1.042065 & 0.006576 & 1.045714 & 0.006245 & 1.042041 & 0.006578\\
+			\hline
+			Data Set & Auto Est & Auto SE & AutoM Est & AutoM SE & Norm Est & Norm SE & NormM Est & NormM SE\\
+			\hline
+			$X_0 = 99999, Y_0 = 1$ & 0.889241 & 0.003555 & 0.854729 & 0.008135 & 0.895513 & 0.003041 & 0.857664 & 0.007601\\
+			\hline
+			$X_0 = 99990, Y_0 = 10$ & 0.905684 & 0.002355 & 0.902737 & 0.002537 & 0.908169 & 0.002211 & 0.902592 & 0.002546\\
+			\hline
+			$X_0 = 99900, Y_0 = 100$ & 0.936306 & 0.001074 & 0.936309 & 0.001074 & 0.936398 & 0.001072 & 0.936345 & 0.001073\\
+			\hline
+			$X_0 = 99000, Y_0 = 1000$ & 0.944202 & 0.000875 & 0.944200 & 0.000875 & 0.944184 & 0.000876 & 0.944210 & 0.000875\\
 			\hline
@@
 			\hline
-			$\sigma_X = 10, \sigma_Y = 1$ & 1.044244 & 0.006376 & 1.041733 & 0.006606 & 1.047118 & 0.006122 & 1.042010 & 0.006580\\
-			\hline
-			$\sigma_X = 100, \sigma_Y = 10$ & 1.088170 & 0.003410 & 1.043282 & 0.006463 & 1.099191 & 0.002912 & 1.044029 & 0.006395\\
-			\hline
-			$\sigma_X = 300, \sigma_Y = 30$ & 1.123132 & 0.002065 & 1.049199 & 0.005944 & 1.130609 & 0.001855 & 1.041324 & 0.006645\\
-			\hline
-			$\sigma_X = 1000, \sigma_Y = 100$ & 1.161414 & 0.001192 & 1.041609 & 0.006618 & 1.171423 & 0.001033 & 1.040663 & 0.006707\\
-			\hline
-			$\sigma_X = 3000, \sigma_Y = 300$ & 1.194876 & 0.000738 & 1.077426 & 0.003976 & 1.210073 & 0.000594 & 1.056925 & 0.005326\\
-			\hline
-			$\sigma_X = 10000, \sigma_Y = 1000$ & 1.237467 & 0.000402 & 1.092848 & 0.003189 & 1.246201 & 0.000355 & 1.093502 & 0.003159\\
+			$\sigma_X = 10, \sigma_Y = 1$ & 0.944214 & 0.000875 & 0.944234 & 0.000875 & 0.944177 & 0.000876 & 0.944203 & 0.000875\\
+			\hline
+			$\sigma_X = 100, \sigma_Y = 10$ & 0.948437 & 0.000784 & 0.948047 & 0.000792 & 0.950444 & 0.000744 & 0.949851 & 0.000756\\
+			\hline
+			$\sigma_X = 500, \sigma_Y = 50$ & 0.968003 & 0.000471 & 0.967327 & 0.000479 & 0.972145 & 0.000422 & 0.970875 & 0.000437\\
+			\hline
+			$\sigma_X = 1000, \sigma_Y = 100$ & 0.979365 & 0.000350 & 0.978657 & 0.000356 & 0.983895 & 0.000311 & 0.982500 & 0.000322\\
+			\hline
+			$\sigma_X = 2000, \sigma_Y = 200$ & 0.991428 & 0.000255 & 0.990718 & 0.000260 & 0.996167 & 0.000225 & 0.994717 & 0.000234\\
+			\hline
+			$\sigma_X = 5000, \sigma_Y = 500$ & 1.007910 & 0.000166 & 1.007211 & 0.000169 & 1.012791 & 0.000146 & 1.011292 & 0.000152\\
+			\hline
+			$\sigma_X = 7500, \sigma_Y = 750$ & 1.015308 & 0.000136 & 1.014610 & 0.000139 & 1.020223 & 0.000120 & 1.018711 & 0.000125\\
+			\hline
+			$\sigma_X = 10000, \sigma_Y = 1000$ & 1.020580 & 0.000119 & 1.019883 & 0.000121 & 1.025513 & 0.000104 & 1.023994 & 0.000109\\
+			\hline
+			$\sigma_X = 15000, \sigma_Y = 1500$ & 1.028035 & 0.000098 & 1.027339 & 0.000100 & 1.032988 & 0.000086 & 1.031463 & 0.000089\\
 			\hline
@@
 			\hline
-			0th & 1.006482 & 0.010827 & 0.782299 & 0.131800 & 1.016043 & 0.009479 & 0.829597 & 0.089876\\
-			\hline
-			1st & 1.060996 & 0.005026 & 1.061004 & 0.005025 & 1.061431 & 0.004995 & 1.060633 & 0.005052\\
-			\hline
-			2nd & 1.073885 & 0.004183 & 1.073323 & 0.004216 & 1.074063 & 0.004172 & 1.073328 & 0.004216\\
-			\hline
-			3rd & 1.077530 & 0.003970 & 1.077039 & 0.003998 & 1.077724 & 0.003959 & 1.077108 & 0.003994\\
-			\hline
-			4th & 1.073888 & 0.004182 & 1.073888 & 0.004182 & 1.074161 & 0.004166 & 1.073508 & 0.004205\\
-			\hline
-			Linear SIR & 1.087538 & 0.003441 & 1.087399 & 0.003448 & 1.087742 & 0.003431 & 1.087338 & 0.003451\\
-			\hline04 & 0.001114\\
+			0th & 0.888590 & 0.003613 & 0.755287 & 0.049788 & 0.894107 & 0.003150 & 0.784131 & 0.032919\\
+			\hline
+			1st & 0.919899 & 0.001639 & 0.919904 & 0.001639 & 0.920147 & 0.001629 & 0.919691 & 0.001648\\
+			\hline
+			2nd & 0.927250 & 0.001357 & 0.926930 & 0.001368 & 0.927352 & 0.001354 & 0.926933 & 0.001368\\
+			\hline
+			3rd & 0.929326 & 0.001287 & 0.929046 & 0.001296 & 0.929436 & 0.001283 & 0.929085 & 0.001295\\
+			\hline
+			4th & 0.927252 & 0.001357 & 0.927252 & 0.001357 & 0.927408 & 0.001352 & 0.927035 & 0.001365\\
+			\hline
+			Linear SIR & 0.935018 & 0.001111 & 0.934939 & 0.001113 & 0.935133 & 0.001107 & 0.934904 & 0.001114\\
 			\hline
@@
 			\hline
-			Baseline 1 & 1.045203 & 0.005596 & 0.954158 & 0.004911 & 1.009321 & 0.005045 & 1.010310 & 0.005043\\
-			\hline
-			Baseline 2 & 0.927409 & 0.004482 & 1.132450 & 0.005591 & 0.985546 & 0.004868 & 1.038093 & 0.005246\\
-			\hline
-			Baseline 3 & 1.035850 & 0.005555 & 0.969459 & 0.004980 & 1.001299 & 0.005062 & 1.006797 & 0.005083\\
-			\hline
-			Baseline 4 & 0.974099 & 0.004834 & 0.971615 & 0.004748 & 0.944385 & 0.004705 & 0.994621 & 0.005009\\
-			\hline
-			Baseline 5 & 0.988996 & 0.005185 & 1.020904 & 0.005239 & 1.053116 & 0.005433 & 1.002879 & 0.004973\\
+			Baseline1 & 0.899074 & 9.7e-05 & 0.902153 & 9.7e-05 & 0.899199 & 9.7e-05 & 0.898925 & 9.7e-05\\
+			\hline
+			Baseline2 & 0.898625 & 9.7e-05 & 0.899906 & 9.7e-05 & 0.899591 & 9.7e-05 & 0.900620 & 9.7e-05\\
+			\hline
+			Baseline3 & 0.898641 & 9.7e-05 & 0.896045 & 9.7e-05 & 0.901772 & 9.7e-05 & 0.900604 & 9.7e-05\\
+			\hline
+			Baseline4 & 0.898617 & 9.7e-05 & 0.899951 & 9.7e-05 & 0.898398 & 9.7e-05 & 0.901681 & 9.7e-05\\
+			\hline
+			Baseline5 & 0.901165 & 9.7e-05 & 0.899822 & 9.7e-05 & 0.900534 & 9.7e-05 & 0.900186 & 9.7e-05\\
 			\hline
@@
 			\hline
-			$X_0 = 99999, Y_0 = 1$ & 1.232756 & 0.069833 & 0.976411 & 0.163888 & 1.615820 & 0.467853 & 1.192249 & 0.205440\\
-			\hline
-			$X_0 = 99990, Y_0 = 10$ & 0.949905 & 0.027784 & 1.301035 & 0.031985 & 0.884468 & 0.022576 & 1.208762 & 0.028717\\
-			\hline
-			$X_0 = 99900, Y_0 = 100$ & 1.016057 & 0.002553 & 0.984204 & 0.002475 & 1.055094 & 0.002654 & 0.990380 & 0.002500\\
-			\hline
-			$X_0 = 99000, Y_0 = 1000$ & 0.983594 & 0.000260 & 0.996016 & 0.000263 & 0.997396 & 0.000264 & 0.998214 & 0.000264\\
+			$X_0 = 99999, Y_0 = 1$ & 0.912953 & 0.003775 & 0.861994 & 0.003368 & 0.877184 & 0.003582 & 0.995336 & 0.004095\\
+			\hline
+			$X_0 = 99990, Y_0 = 10$ & 0.903707 & 0.000390 & 0.926002 & 0.000401 & 0.909539 & 0.000395 & 0.923559 & 0.000402\\
+			\hline
+			$X_0 = 99900, Y_0 = 100$ & 0.890136 & 0.000060 & 0.891312 & 0.000060 & 0.891251 & 0.000060 & 0.891413 & 0.000060\\
+			\hline
+			$X_0 = 99000, Y_0 = 1000$ & 0.863266 & 0.000024 & 0.863334 & 0.000024 & 0.862902 & 0.000024 & 0.863491 & 0.000024\\
 			\hline
@@
 			\hline
-			Data Set & Auto Est & Auto SE & AutoM Est & AutoM SE & Norm Est & Norm SE & NormM Est & NormM SE\\
-			\hline
-			$\sigma_X = 10, \sigma_Y = 1$ & 0.985737 & 0.000206 & 0.985711 & 0.000207 & 0.985510 & 0.000206 & 0.986968 & 0.000206\\
-			\hline
-			$\sigma_X = 100, \sigma_Y = 10$ & 0.999005 & 0.000208 & 1.001050 & 0.000208 & 1.006205 & 0.000209 & 1.013000 & 0.000211\\
-			\hline
-			$\sigma_X = 300, \sigma_Y = 30$ & 1.051749 & 0.000236 & 0.966436 & 0.000181 & 0.816122 & 0.000167 & 0.966859 & 0.000200\\
-			\hline
-			$\sigma_X = 1000, \sigma_Y = 100$ & 1.060553 & 0.000161 & 1.065634 & 0.000236 & 0.777445 & 0.000177 & 0.975480 & 0.000206\\
-			\hline
-			$\sigma_X = 3000, \sigma_Y = 300$ & 0.511344 & 0.000110 & 1.003435 & 0.000113 & 2.385157 & 0.000672 & 1.095206 & 0.000199\\
-			\hline
-			$\sigma_X = 10000, \sigma_Y = 1000$ & -3.235191 & 0.000840 & 0.952381 & 0.000089 & -2.850294 & -0.001262 & 1.532468 & 0.000209\\
+			Error & Auto Est & Auto SE & AutoM Est & AutoM SE & Norm Est & Norm SE & NormM Est & NormM SE\\
+			\hline
+			$\sigma_X = 10, \sigma_Y = 1$ & 0.863286 & 2.4e-05 & 0.864054 & 0.000024 & 0.862686 & 2.4e-05 & 0.863301 & 2.4e-05\\
+			\hline
+			$\sigma_X = 100, \sigma_Y = 10$ & 0.865256 & 2.4e-05 & 0.872950 & 0.000025 & 0.859255 & 2.4e-05 & 0.865412 & 2.4e-05\\
+			\hline
+			$\sigma_X = 500, \sigma_Y = 50$ & 0.873904 & 2.4e-05 & 0.912804 & 0.000026 & 0.844000 & 2.4e-05 & 0.874831 & 2.5e-05\\
+			\hline
+			$\sigma_X = 1000, \sigma_Y = 100$ & 0.884476 & 2.4e-05 & 0.963357 & 0.000027 & 0.824913 & 2.3e-05 & 0.886693 & 2.5e-05\\
+			\hline
+			$\sigma_X = 2000, \sigma_Y = 200$ & 0.904863 & 2.4e-05 & 1.066995 & 0.000031 & 0.786679 & 2.2e-05 & 0.910719 & 2.6e-05\\
+			\hline
+			$\sigma_X = 5000, \sigma_Y = 500$ & 0.960573 & 2.4e-05 & 1.399764 & 0.000043 & 0.671494 & 1.9e-05 & 0.985289 & 2.9e-05\\
+			\hline
+			$\sigma_X = 7500, \sigma_Y = 750$ & 1.001610 & 2.4e-05 & 1.705105 & 0.000054 & 0.574948 & 1.6e-05 & 1.050455 & 3.1e-05\\
+			\hline
+			$\sigma_X = 10000, \sigma_Y = 1000$ & 1.038534 & 2.3e-05 & 2.039858 & 0.000068 & 0.477891 & 1.4e-05 & 1.118568 & 3.4e-05\\
+			\hline
+			$\sigma_X = 15000, \sigma_Y = 1500$ & 1.102292 & 2.2e-05 & 2.816412 & 0.000104 & 0.282224 & 8.0e-06 & 1.264474 & 4.1e-05\\
 			\hline
@@
 			\hline
-			0th & -12.609343 & -14.168066 & 16.325348 & 42.366553 & 3.152735 & -0.768473 & 10.431474 & 8.031043\\
-			\hline
-			1st & 0.978683 & 0.000125 & 0.982726 & 0.000126 & 0.979523 & 0.000125 & 0.976522 & 0.000125\\
-			\hline
-			2nd & 1.017042 & 0.000131 & 1.013913 & 0.000131 & 1.014664 & 0.000131 & 1.014455 & 0.000131\\
-			\hline
-			3rd & 1.029675 & 0.000121 & 1.026645 & 0.000121 & 1.027745 & 0.000121 & 1.027322 & 0.000121\\
-			\hline
-			4th & 1.028890 & 0.000135 & 1.028890 & 0.000135 & 1.027876 & 0.000135 & 1.027989 & 0.000135\\
-			\hline
-			Linear SIR & 1.052763 & 0.000099 & 1.053736 & 0.000099 & 1.053703 & 0.000099 & 1.052769 & 0.000099\\
+			0th & 3.167663 & -0.894136 & 8.459310 & 11.375438 & 3.145217 & -0.764813 & 6.433011 & 3.054282\\
+			\hline
+			1st & 0.872597 & 0.000099 & 0.875151 & 0.000100 & 0.873004 & 0.000099 & 0.870359 & 0.000099\\
+			\hline
+			2nd & 0.894306 & 0.000101 & 0.892614 & 0.000101 & 0.892698 & 0.000101 & 0.892652 & 0.000101\\
+			\hline
+			3rd & 0.902721 & 0.000093 & 0.900171 & 0.000093 & 0.901044 & 0.000093 & 0.900592 & 0.000093\\
+			\hline
+			4th & 0.901112 & 0.000104 & 0.901112 & 0.000104 & 0.900495 & 0.000104 & 0.900734 & 0.000104\\
+			\hline
+			Linear SIR & 0.915222 & 0.000075 & 0.915111 & 0.000074 & 0.915750 & 0.000075 & 0.914991 & 0.000075\\
 			\hline
@@
 			\hline
-			Baseline 1 & 1.122458 & 0.044665 & 0.951335 & 0.152044 & 1.135276 & 0.040490 & 0.965079 & 0.138753\\
-			\hline
-			Baseline 2 & 1.119443 & 0.045703 & 0.982261 & 0.123521 & 1.136269 & 0.040183 & 0.967189 & 0.136801\\
-			\hline
-			Baseline 3 & 1.110763 & 0.048817 & 0.954452 & 0.148942 & 1.131325 & 0.041737 & 0.963011 & 0.140689\\
-			\hline
-			Baseline 4 & 1.111362 & 0.048596 & 0.962238 & 0.141417 & 1.134666 & 0.040680 & 0.961751 & 0.141879\\
-			\hline
-			Baseline 5 & 1.116547 & 0.046721 & 0.962069 & 0.141578 & 1.129191 & 0.042424 & 0.965051 & 0.138765\\
+			Baseline 1 & 1.248124 & 0.016595 & 1.247917 & 0.016623 & 1.248829 & 0.016501 & 1.247484 & 0.016682\\
+			\hline
+			Baseline 2 & 1.248130 & 0.016594 & 1.247549 & 0.016673 & 1.248722 & 0.016515 & 1.247691 & 0.016653\\
+			\hline
+			Baseline 3 & 1.248067 & 0.016603 & 1.247266 & 0.016711 & 1.248788 & 0.016506 & 1.247692 & 0.016654\\
+			\hline
+			Baseline 4 & 1.248048 & 0.016605 & 1.247550 & 0.016672 & 1.248777 & 0.016508 & 1.247761 & 0.016644\\
+			\hline
+			Baseline 5 & 1.248106 & 0.016597 & 1.247529 & 0.016675 & 1.248912 & 0.016490 & 1.247547 & 0.016673\\
 			\hline
@@
 			\hline
-			$X_0 = 99999, Y_0 = 1$ & 1.107146 & 0.050171 & 0.578787 & 0.985754 & 1.132118 & 0.041484 & 0.555865 & 1.065990\\
-			\hline
-			$X_0 = 99990, Y_0 = 10$ & 1.116055 & 0.046896 & 0.855171 & 0.276575 & 1.129441 & 0.042344 & 0.838084 & 0.305078\\
-			\hline
-			$X_0 = 99900, Y_0 = 100$ & 1.123557 & 0.044293 & 1.018159 & 0.096252 & 1.137845 & 0.039698 & 1.018516 & 0.096009\\
-			\hline
-			$X_0 = 99000, Y_0 = 1000$ & 1.186268 & 0.027217 & 1.182227 & 0.028098 & 1.189399 & 0.026553 & 1.182180 & 0.028108\\
+			$X_0 = 99999, Y_0 = 1$ & 1.113607 & 0.047776 & 0.991326 & 0.116074 & 1.135438 & 0.040440 & 1.001869 & 0.107891\\
+			\hline
+			$X_0 = 99990, Y_0 = 10$ & 1.170601 & 0.030784 & 1.160443 & 0.033326 & 1.179144 & 0.028787 & 1.159943 & 0.033456\\
+			\hline
+			$X_0 = 99900, Y_0 = 100$ & 1.274694 & 0.013368 & 1.274705 & 0.013367 & 1.275003 & 0.013334 & 1.274823 & 0.013354\\
+			\hline
+			$X_0 = 99000, Y_0 = 1000$ & 1.301122 & 0.010759 & 1.301114 & 0.010760 & 1.301059 & 0.010765 & 1.301148 & 0.010757\\
 			\hline
@@
 			\hline
-			$\sigma_X = 10, \sigma_Y = 1$ & 1.186512 & 0.027164 & 1.181573 & 0.028243 & 1.192153 & 0.025981 & 1.182118 & 0.028121\\
-			\hline
-			$\sigma_X = 100, \sigma_Y = 10$ & 1.271572 & 0.013714 & 1.184621 & 0.027573 & 1.292533 & 0.011548 & 1.186089 & 0.027256\\
-			\hline
-			$\sigma_X = 300, \sigma_Y = 30$ & 1.337570 & 0.007951 & 1.196232 & 0.025155 & 1.351500 & 0.007077 & 1.180770 & 0.028422\\
-			\hline
-			$\sigma_X = 1000, \sigma_Y = 100$ & 1.408230 & 0.004385 & 1.181330 & 0.028297 & 1.426440 & 0.003755 & 1.179468 & 0.028713\\
-			\hline
-			$\sigma_X = 3000, \sigma_Y = 300$ & 1.468704 & 0.002614 & 1.250992 & 0.016213 & 1.495791 & 0.002069 & 1.211323 & 0.022310\\
-			\hline
-			$\sigma_X = 10000, \sigma_Y = 1000$ & 1.544051 & 0.001360 & 1.280487 & 0.012748 & 1.559289 & 0.001190 & 1.281732 & 0.012619\\
-			\hline
-
+			$\sigma_X = 10, \sigma_Y = 1$ & 1.301161 & 0.010756 & 1.301227 & 0.010750 & 1.301037 & 0.010767 & 1.301125 & 0.010759\\
+			\hline
+			$\sigma_X = 100, \sigma_Y = 10$ & 1.315228 & 0.009575 & 1.313931 & 0.009678 & 1.321895 & 0.009059 & 1.319928 & 0.009209\\
+			\hline
+			$\sigma_X = 500, \sigma_Y = 50$ & 1.379802 & 0.005578 & 1.377586 & 0.005683 & 1.393349 & 0.004975 & 1.389199 & 0.005152\\
+			\hline
+			$\sigma_X = 1000, \sigma_Y = 100$ & 1.416861 & 0.004075 & 1.414559 & 0.004155 & 1.431547 & 0.003595 & 1.427030 & 0.003736\\
+			\hline
+			$\sigma_X = 2000, \sigma_Y = 200$ & 1.455861 & 0.002919 & 1.453577 & 0.002977 & 1.471091 & 0.002561 & 1.466436 & 0.002665\\
+			\hline
+			$\sigma_X = 5000, \sigma_Y = 500$ & 1.508597 & 0.001851 & 1.506373 & 0.001888 & 1.524094 & 0.001618 & 1.519341 & 0.001686\\
+			\hline
+			$\sigma_X = 7500, \sigma_Y = 750$ & 1.532063 & 0.001510 & 1.529856 & 0.001539 & 1.547585 & 0.001318 & 1.542817 & 0.001374\\
+			\hline
+			$\sigma_X = 10000, \sigma_Y = 1000$ & 1.548711 & 0.001305 & 1.546515 & 0.001331 & 1.564232 & 0.001140 & 1.559461 & 0.001188\\
+			\hline
+			$\sigma_X = 15000, \sigma_Y = 1500$ & 1.572148 & 0.001063 & 1.569966 & 0.001084 & 1.587649 & 0.000928 & 1.582882 & 0.000968\\
+			\hline
 		\end{tabular}
@@
 	\caption{$\rr$ Estimates and Std. Errs, FINALSIZE Model, 
-		Baseline Data$X_0 = 99950, Y_0 = 50$, 
-		$\sigma_X = 10, \sigma_Y = 1$}
-	\centering
-	\begin{footnotesize}
-		\hskip -1cm
-	\begin{tabular}[t]{l|r|r|r|r|r|r|r|r}
-		\hline
-		Data Set & Auto Est & Auto SE & AutoM Est & AutoM SE & Norm Est & Norm SE & NormM Est & NormM SE\\
-		\hline
-		Baseline 1 & 1.001228 & 7.8e-05 & 1.001179 & 7.7e-05 & 1.001262 & 8.0e-05 & 1.001264 & 8.0e-05\\
-		\hline
-		Baseline 2 & 1.001211 & 7.8e-05 & 1.001400 & 8.4e-05 & 1.001250 & 7.9e-05 & 1.001279 & 8.0e-05\\
-		\hline
-		Baseline 3 & 1.001218 & 7.8e-05 & 1.001196 & 7.7e-05 & 1.001242 & 7.9e-05 & 1.001249 & 7.9e-05\\
-		\hline
-		Baseline 4 & 1.001233 & 7.9e-05 & 1.001246 & 7.9e-05 & 1.001200 & 7.8e-05 & 1.001239 & 7.9e-05\\
-		\hline
-		Baseline 5 & 1.001195 & 7.7e-05 & 1.001247 & 7.9e-05 & 1.001273 & 8.0e-05 & 1.001263 & 8.0e-05\\
+		Baseline Data, $X_0 = 99950, Y_0 = 50$, 
+		$\sigma_X = 10, \sigma_Y = 1$}
+	\begin{footnotesize}
+		\hskip -1cm
+		\begin{tabular}{l|r|r|r|r|r|r|r|r}
+			\hline
+			Data Set & Auto Est & Auto SE & AutoM Est & AutoM SE & Norm Est & Norm SE & NormM Est & NormM SE\\
+			\hline
+			Baseline1 & 1.044459 & 0.000492 & 1.044644 & 0.000494 & 1.044435 & 0.000492 & 1.044410 & 0.000492\\
+			\hline
+			Baseline2 & 1.044363 & 0.000492 & 1.044445 & 0.000492 & 1.044433 & 0.000492 & 1.044518 & 0.000493\\
+			\hline
+			Baseline3 & 1.044434 & 0.000492 & 1.044279 & 0.000491 & 1.044565 & 0.000493 & 1.044521 & 0.000493\\
+			\hline
+			Baseline4 & 1.044370 & 0.000492 & 1.044446 & 0.000492 & 1.044374 & 0.000492 & 1.044557 & 0.000493\\
+			\hline
+			Baseline5 & 1.044480 & 0.000493 & 1.044438 & 0.000492 & 1.044475 & 0.000493 & 1.044442 & 0.000492\\
 			\hline
@@
 			\hline
-			$X_0 = 99999, Y_0 = 1$ & 1.000114 & 0.000024 & 1.000034 & 0.000013 & 1.000033 & 0.000013 & 1.000040 & 0.000014\\
-			\hline
-			$X_0 = 99990, Y_0 = 10$ & 1.000212 & 0.000033 & 1.000315 & 0.000040 & 1.000223 & 0.000033 & 1.000305 & 0.000039\\
-			\hline
-			$X_0 = 99900, Y_0 = 100$ & 1.002530 & 0.000113 & 1.002465 & 0.000111 & 1.002606 & 0.000114 & 1.002470 & 0.000111\\
-			\hline
-			$X_0 = 99000, Y_0 = 1000$ & 1.024360 & 0.000358 & 1.024661 & 0.000360 & 1.024662 & 0.000360 & 1.024670 & 0.000360\\
+			$X_0 = 99999, Y_0 = 1$ & 1.001111 & 0.000075 & 1.001110 & 0.000075 & 1.001081 & 0.000074 & 1.001217 & 0.000078\\
+			\hline
+			$X_0 = 99990, Y_0 = 10$ & 1.010679 & 0.000234 & 1.010895 & 0.000236 & 1.010676 & 0.000234 & 1.010824 & 0.000235\\
+			\hline
+			$X_0 = 99900, Y_0 = 100$ & 1.073689 & 0.000652 & 1.073782 & 0.000652 & 1.073845 & 0.000653 & 1.073877 & 0.000653\\
+			\hline
+			$X_0 = 99000, Y_0 = 1000$ & 1.203036 & 0.001215 & 1.203028 & 0.001215 & 1.202929 & 0.001214 & 1.203088 & 0.001215\\
 			\hline
@@
 			\hline
-			Data Set & Auto Est & Auto SE & AutoM Est & AutoM SE & Norm Est & Norm SE & NormM Est & NormM SE\\
-			\hline
-			$\sigma_X = 10, \sigma_Y = 1$ & 1.024640 & 0.000360 & 1.024544 & 0.000359 & 1.024610 & 0.000359 & 1.024641 & 0.000360\\
-			\hline
-			$\sigma_X = 100, \sigma_Y = 10$ & 1.025024 & 0.000363 & 1.025169 & 0.000364 & 1.025348 & 0.000365 & 1.025358 & 0.000365\\
-			\hline
-			$\sigma_X = 300, \sigma_Y = 30$ & 1.024494 & 0.000359 & 1.026920 & 0.000377 & 1.020803 & 0.000329 & 1.024440 & 0.000358\\
-			\hline
-			$\sigma_X = 1000, \sigma_Y = 100$ & 1.036821 & 0.000445 & 1.025126 & 0.000363 & 1.017766 & 0.000303 & 1.024120 & 0.000356\\
-			\hline
-			$\sigma_X = 3000, \sigma_Y = 300$ & 1.012305 & 0.000251 & 1.047894 & 0.000513 & 1.045165 & 0.000497 & 1.031720 & 0.000411\\
-			\hline
-			$\sigma_X = 10000, \sigma_Y = 1000$ & 1.068277 & 0.000624 & 1.055107 & 0.000554 & 0.969364 & NaN & 1.061044 & 0.000586\\
+			Error & Auto Est & Auto SE & AutoM Est & AutoM SE & Norm Est & Norm SE & NormM Est & NormM SE\\
+			\hline
+			$\sigma_X = 10, \sigma_Y = 1$ & 1.203100 & 0.001215 & 1.203221 & 0.001215 & 1.202857 & 0.001214 & 1.203019 & 0.001215\\
+			\hline
+			$\sigma_X = 100, \sigma_Y = 10$ & 1.204254 & 0.001220 & 1.205468 & 0.001224 & 1.201827 & 0.001210 & 1.203450 & 0.001216\\
+			\hline
+			$\sigma_X = 500, \sigma_Y = 50$ & 1.209421 & 0.001240 & 1.215599 & 0.001265 & 1.197277 & 0.001191 & 1.205368 & 0.001224\\
+			\hline
+			$\sigma_X = 1000, \sigma_Y = 100$ & 1.215968 & 0.001267 & 1.228600 & 0.001317 & 1.191656 & 0.001169 & 1.207778 & 0.001234\\
+			\hline
+			$\sigma_X = 2000, \sigma_Y = 200$ & 1.229359 & 0.001320 & 1.255800 & 0.001425 & 1.180629 & 0.001124 & 1.212639 & 0.001253\\
+			\hline
+			$\sigma_X = 5000, \sigma_Y = 500$ & 1.272128 & 0.001489 & 1.348709 & 0.001792 & 1.149157 & 0.000994 & 1.227550 & 0.001313\\
+			\hline
+			$\sigma_X = 7500, \sigma_Y = 750$ & 1.311117 & 0.001643 & 1.442921 & 0.002168 & 1.124616 & 0.000888 & 1.240372 & 0.001364\\
+			\hline
+			$\sigma_X = 10000, \sigma_Y = 1000$ & 1.353661 & 0.001811 & 1.558903 & 0.002644 & 1.101452 & 0.000785 & 1.253572 & 0.001416\\
+			\hline
+			$\sigma_X = 15000, \sigma_Y = 1500$ & 1.451957 & 0.002204 & 1.908199 & 0.004213 & 1.058763 & 0.000574 & 1.281188 & 0.001525\\
 			\hline
@@
 	\caption{$\rr$ Estimates and Std. Errs, BRANCH Model, 
-		Baseline Data$X_0 = 99950, Y_0 = 50$, 
-		$\sigma_X = 10, \sigma_Y = 1$}
-	\centering
-	\begin{footnotesize}
-		\hskip -1cm
-	\begin{tabular}[t]{l|r|r|r|r|r|r|r|r}
-		\hline
-		Data Set & Auto Est & Auto SE & AutoM Est & AutoM SE & Norm Est & Norm SE & NormM Est & NormM SE\\
-		\hline
-		Baseline 1 & 1.045203 & 0.005596 & 0.954158 & 0.004911 & 1.009321 & 0.005045 & 1.010310 & 0.005043\\
-		\hline
-		Baseline 2 & 0.927409 & 0.004482 & 1.132450 & 0.005591 & 0.985546 & 0.004868 & 1.038093 & 0.005246\\
-		\hline
-		Baseline 3 & 1.035850 & 0.005555 & 0.969459 & 0.004980 & 1.001299 & 0.005062 & 1.006797 & 0.005083\\
-		\hline
-		Baseline 4 & 0.974099 & 0.004834 & 0.971615 & 0.004748 & 0.944385 & 0.004705 & 0.994621 & 0.005009\\
-		\hline
-		Baseline 5 & 0.988996 & 0.005185 & 1.020904 & 0.005239 & 1.053116 & 0.005433 & 1.002879 & 0.004973\\
+		Baseline Data, $X_0 = 99950, Y_0 = 50$, 
+		$\sigma_X = 10, \sigma_Y = 1$}
+	\begin{footnotesize}
+		\hskip -1cm
+		\begin{tabular}{l|r|r|r|r|r|r|r|r}
+			\hline
+			Data Set & Auto Est & Auto SE & AutoM Est & AutoM SE & Norm Est & Norm SE & NormM Est & NormM SE\\
+			\hline
+			Baseline1 & 0.899074 & 9.7e-05 & 0.902153 & 9.7e-05 & 0.899199 & 9.7e-05 & 0.898925 & 9.7e-05\\
+			\hline
+			Baseline2 & 0.898625 & 9.7e-05 & 0.899906 & 9.7e-05 & 0.899591 & 9.7e-05 & 0.900620 & 9.7e-05\\
+			\hline
+			Baseline3 & 0.898641 & 9.7e-05 & 0.896045 & 9.7e-05 & 0.901772 & 9.7e-05 & 0.900604 & 9.7e-05\\
+			\hline
+			Baseline4 & 0.898617 & 9.7e-05 & 0.899951 & 9.7e-05 & 0.898398 & 9.7e-05 & 0.901681 & 9.7e-05\\
+			\hline
+			Baseline5 & 0.901165 & 9.7e-05 & 0.899822 & 9.7e-05 & 0.900534 & 9.7e-05 & 0.900186 & 9.7e-05\\
 			\hline
@@
 			\hline
-			Data Set & Auto Est & Auto SE & AutoM Est & AutoM SE & Norm Est & Norm SE & NormM Est & NormM SE\\
-			\hline
-			$\sigma_X = 10, \sigma_Y = 1$ & 0.985737 & 0.000206 & 0.985711 & 0.000207 & 0.985510 & 0.000206 & 0.986968 & 0.000206\\
-			\hline
-			$\sigma_X = 100, \sigma_Y = 10$ & 0.999005 & 0.000208 & 1.001050 & 0.000208 & 1.006205 & 0.000209 & 1.013000 & 0.000211\\
-			\hline
-			$\sigma_X = 300, \sigma_Y = 30$ & 1.051749 & 0.000236 & 0.966436 & 0.000181 & 0.816122 & 0.000167 & 0.966859 & 0.000200\\
-			\hline
-			$\sigma_X = 1000, \sigma_Y = 100$ & 1.060553 & 0.000161 & 1.065634 & 0.000236 & 0.777445 & 0.000177 & 0.975480 & 0.000206\\
-			\hline
-			$\sigma_X = 3000, \sigma_Y = 300$ & 0.511344 & 0.000110 & 1.003435 & 0.000113 & 2.385157 & 0.000672 & 1.095206 & 0.000199\\
-			\hline
-			$\sigma_X = 10000, \sigma_Y = 1000$& -3.235191 & 0.000840 & 0.952381 & 0.000089 & -2.850294 & 0.001262 & 1.532468 & 0.000209\\
+			Error & Auto Est & Auto SE & AutoM Est & AutoM SE & Norm Est & Norm SE & NormM Est & NormM SE\\
+			\hline
+			$\sigma_X = 10, \sigma_Y = 1$ & 0.863286 & 2.4e-05 & 0.864054 & 0.000024 & 0.862686 & 2.4e-05 & 0.863301 & 2.4e-05\\
+			\hline
+			$\sigma_X = 100, \sigma_Y = 10$ & 0.865256 & 2.4e-05 & 0.872950 & 0.000025 & 0.859255 & 2.4e-05 & 0.865412 & 2.4e-05\\
+			\hline
+			$\sigma_X = 500, \sigma_Y = 50$ & 0.873904 & 2.4e-05 & 0.912804 & 0.000026 & 0.844000 & 2.4e-05 & 0.874831 & 2.5e-05\\
+			\hline
+			$\sigma_X = 1000, \sigma_Y = 100$ & 0.884476 & 2.4e-05 & 0.963357 & 0.000027 & 0.824913 & 2.3e-05 & 0.886693 & 2.5e-05\\
+			\hline
+			$\sigma_X = 2000, \sigma_Y = 200$ & 0.904863 & 2.4e-05 & 1.066995 & 0.000031 & 0.786679 & 2.2e-05 & 0.910719 & 2.6e-05\\
+			\hline
+			$\sigma_X = 5000, \sigma_Y = 500$ & 0.960573 & 2.4e-05 & 1.399764 & 0.000043 & 0.671494 & 1.9e-05 & 0.985289 & 2.9e-05\\
+			\hline
+			$\sigma_X = 7500, \sigma_Y = 750$ & 1.001610 & 2.4e-05 & 1.705105 & 0.000054 & 0.574948 & 1.6e-05 & 1.050455 & 3.1e-05\\
+			\hline
+			$\sigma_X = 10000, \sigma_Y = 1000$ & 1.038534 & 2.3e-05 & 2.039858 & 0.000068 & 0.477891 & 1.4e-05 & 1.118568 & 3.4e-05\\
+			\hline
+			$\sigma_X = 15000, \sigma_Y = 1500$ & 1.102292 & 2.2e-05 & 2.816412 & 0.000104 & 0.282224 & 8.0e-06 & 1.264474 & 4.1e-05\\
 			\hline

### 2018-07-03 paper/draft_v9/draft_v8.tex
finally on git

--- 2018-07-03 paper/draft_v8/draft_v8.tex
@@
 \textbf{The Gaussian errors are independently distributed with mean 0 and constant variance for the entire length of the time series.
-The autoregressive errors are generated from a first-order process of the form: $$ \epsilon_{t} = \rho * \epsilon_{t-1} + \delta_{t}$$, where $\rho$ is fixed throughout and $\delta_{t} \sim N(0, \sigma)$ with constant $\sigma$.}
+The autoregressive errors are generated from a first-order process of the form: $$ \epsilon_{t} = \rho * \epsilon_{t-1} + \delta_{t}$$, where $\rho$ is fixed throughout and $\delta_{t} \sim N(0, \sigma)$ with constant $\sigma$.
+The values of the variances are chosen somewhat arbitrarily, although they are intended to be fairly small at first.}
 
@@
 
-%% What kind of graphs should we throw in?
-
-%% SG: Forest plots!
+Below, we compare the estimates and standard errors of $\rr$ given by each of the models for a few selected data sets.
+The forest plots below show the point estimates and 95\% confidence intervals of $\rr$ from each model for a particular data set.
+The true $\rr$ value is also shown by the vertical red line in the graphs.
+
+All of the figures below show results from data sets that were simulated with added Gaussian errors.
+The plot on the top left shows the estimates from one of the baseline data sets.
+For the most part, the $\rr$ values range from approximately 0.8 to 1.4.
+We note that the range of values includes $\rr$ values of less than 1, meaning that some models predict that a disease is not spreading when it actually is.
+However, on the whole, the variation of estimates between models is smaller than we originally expected.
+The confidence intervals are very small relative to the estimates for the most part; exceptions to this include the SIR IPR and the Markov chain model as the exception.
+
+The plot on the top right shows the estimates for the data set where $Y_0$ is increased to 100.
+The $\rr$ estimates generally seem to be further away from the true $\rr$ value compared to the results from the baseline data sets, and the overall .
+The confidence intervals are also more often than not larger than those in the baseline.
+In particular, the standard errors for some of the SIR models increase by several orders of magnitude.
+
+The bottom left plot shows the estimates for the data set where the error added to the generated data is increased.
+For most of the models, the estimates of $\rr$ do not change very much compared to the baseline data set.
+The confidence intervals for this data set increase for some of the models, such as the SIR LMA, while they remain approximately the same as the baseline data sets for most of the others.
+
+On the bottom right, we see the $\rr$ estimates for the case when disease prevalence grows linearly, as opposed to exponentially as seen in the data sets generated under SIR conditions, where the disease prevalence grows exponentially.
+Model performance here varies depending on how well the models the aforementioned linear growth.
+Some of the models, such as the SIR Markov chain and the IPR models, do not handle this case very well, giving much larger estimates and standard errors than are reasonable.
+On the other hand, models such as the SIR maximum and final size seem to be more flexible to different trajectories.
+[The error bars do not show up if they go outside of the x-axis limits - how do we fix this?]
+
+\begin{minipage}{0.5\linewidth}
+	\includegraphics[width=3in]{BaselineFP.jpeg}
+\end{minipage}
+\begin{minipage}{0.5\linewidth}
+	\includegraphics[width=3in]{StartPopFP.jpeg}
+\end{minipage}
+
+\begin{minipage}{0.5\linewidth}
+	\includegraphics[width=3in]{VarianceFP.jpeg}
+\end{minipage}
+\begin{minipage}{0.5\linewidth}
+	\includegraphics[width=3in]{OtherFP.jpeg}
+\end{minipage}
 
@@
 			\hline
-			0th & 0.005468 & 0.000745 & -0.034869 & 0.000764 & -0.034869 & 7.4e-04 & 0.034869 & 0.000673\\
-			\hline
-			1st & 1.038004 & 0.000029 & 1.037900 & 0.000029 & 1.037826 & 2.9e-05 & 1.037826 & 0.000029\\
-			\hline
-			2nd & 1.199663 & 0.000045 & 1.198627 & 0.000045 & 1.199663 & 4.5e-05 & 1.198949 & 0.000045\\
-			\hline
-			3rd & 1.202909 & 0.000043 & 1.202282 & 0.000043 & 1.202414 & 4.3e-05 & 1.202414 & 0.000043\\
-			\hline
-			4th & 1.211092 & 0.000047 & 1.211092 & 0.000047 & 1.209479 & 4.7e-05 & 1.211092 & 0.000047\\
+			0th & 0.005468 & 0.000745 & -0.034869 & 0.000764 & -0.034869 & 0.000743 & 0.034869 & 0.000673\\
+			\hline
+			1st & 1.149386 & 0.000099 & 1.149386 & 0.000099 & 1.149386 & 0.000099 & 1.149386 & 0.000099\\
+			\hline
+			2nd & 1.199663 & 0.000045 & 1.198627 & 0.000045 & 1.199663 & 0.000045 & 1.198949 & 0.000045\\
+			\hline
+			3rd & 1.202909 & 0.000043 & 1.202282 & 0.000043 & 1.202414 & 0.000043 & 1.202414 & 0.000043\\
+			\hline
+			4th & 1.211092 & 0.000047 & 1.211092 & 0.000047 & 1.209479 & 0.000047 & 1.211092 & 0.000047\\
 			\hline

### 2018-07-03 paper/old_drafts/draft_v1.tex
finally on git

\documentclass[12pt]{article}
\usepackage[margin=1in]{geometry}
%\usepackage{hyperref}
\usepackage{amsmath}
\usepackage{graphicx}% http://ctan.org/pkg/graphicx
\usepackage{xcolor}
\usepackage{systeme}

%% Make red text
\newcommand{\com}[1]{\textcolor{red}{ #1}}

\usepackage[authoryear]{natbib}

\usepackage{tikz}
\tikzset{
  int/.style={circle, draw, fill=blue!20, minimum size=3em},
  init/.style={pin distance=1.2cm,pin edge={loop,thin,black}}
}
\usetikzlibrary{arrows}

\pagenumbering{arabic}

\newcommand{\XX}{15 } % number of methods
\newcommand{\rr}{\ensuremath{\mathcal{R}_0}}

\setlength{\itemsep}{0pt}

\begin{document}

\title{\XX dubious ways of estimating $\rr$}
\author{ Department of Statistics, Carnegie Mellon University}
\date{\today}
\maketitle

%\tableofcontents

\section{Introduction}\label{sec:intro}
What has been called ``arguably the most important quantity in the study of epidemics'' \cite{Heesterbeek2002}, $\mathcal{R}_0$, the reproduction number, has been found to be an elusive quantity for epidemic modelers.  \citet{anderson1992} define $\rr$ as the ``the average number of secondary infections produced when one infected individual is introduced into a host population where everyone is susceptible.''  $\rr$, in some ways, summarizes an entire epidemic.  Despite the clear definition of $\rr$, epidemiologists have struggled to create a \textit{de facto} estimator for $\rr$  \citep{hethcote2000}.  The primary issue in estimating $\rr$ is that the quantity is a \textit{property of the model}, meaning that $\rr$ is dependent not only on the usual noise that comes with statistical modeling but also on a variety of assumptions on how we assume disease is transmitted through a population.

Typically, disease models are a part of the ``SI-framework'' pioneered by Kermack and McKendrick in the 1920s \citep{getz2006}.  Here, S and I are known as compartments where S stands for ``susceptible'' and I for ``infected.''  The issue arises in how individuals are assumed, either implicitly or explicitly, to travel from the S compartment to the I compartment and any other compartments that have been introduced such as immune, recovered, dead, or exposed, just to name a few.  One of the simplest models in the SI framework is the SIR model where R stands for a ``recovered'' state where individuals are no longer susceptible to the disease \citep{Kermack700}.  Even in this simplified framework, there is much dispute on how $\rr$ is to be estimated arising from assumptions on how populations intermix, if any intervention has been put in place, or whether disease parameters are constant over time.

Other issues of estimating $\rr$ include mathematical versus statistical approaches (i.e. solving for $\rr$ versus estimating $\rr$), how to estimate the quantity in the presence of different available data, and when to estimate $\rr$.  For instance, if only weekly incidence counts of a disease are available, then we have different methods of estimating $\rr$ than when we have more data available such as contact networks, the serial interval of a disease, and more.  Another major issue is that since $\rr$ is so difficult to estimate in the first place, the variance of the quantity has largely been glossed over or completely ignored.  This is especially troubling as we are concerned whether $\rr > 1$, which implies an outbreak of a disease.  A point estimate is simply not enough.

In short, $\rr$ is difficult to estimate because it is a property of the model.  This, in turn, makes it difficult to compare $\rr$ across different diseases or even instances of the same disease.  Many of these issues with estimating $\rr$ are summarized by \cite{li2011}

We review \XX ways to estimate $\rr$ over varying assumptions and in the presence of different sources of data.  The title of this manuscript is in reference to a seminal paper by Moler and Van Loan in 1978 (and its subsequent update in 2003) in which they calculate ``Nineteen dubious ways to compute the exponential of a matrix'' \citep{moler2003}.  We not only review the methods of estimating of $\rr$ but also include recommendations on when to use the methods as well as introducing techniques in which to estimate the variance of estimators.  We compare these estimates, when applicable, and find that $\rr$ is more robust than what was initially presumed.

The rest of this manuscript is organized as follows.  In Section \ref{sec:overview}, we introduce the methods in 3 broad groups:.  In Section \ref{sec:term} we introduce terminology used throughout.  In Section \ref{sec:details} we give an overview of each of the methods, what assumptions are used, what data is required, and where the method has been used in the past along with ways in which to calculate the variance for methods, when applicable.  In Section \ref{sec:results}, we give estimates of $\rr$ for the different methods and their variances over a variety of data sets.  Finally, in Section \ref{sec:dis} we discuss our conclusions and recommendations.

\section{Overview of \XX Methods to estimate $\rr$}
\label{sec:overview}

In order to organize the methods for estimating \rr, we have chosen four broad classes of estimation methods.  These partitions are meant to be guidelines of distinguishing methods from one another rather than immutable groups.  Many of these methods could have fit in another category or even all four of the classes chosen.  We have chosen the four classes as application of the survival function and its close relatives, classic compartment models, exponential growth assumptions, and estimation from networks.

\textbf{Point Estimation}
\begin{enumerate}
\item Survival function
  \begin{enumerate}
  \item Survival function
  \item Direct estimation of the survival function
  \item Direct parameter estimation
  \item Contact tracing
  \end{enumerate}
\item Compartment models
  \begin{enumerate}
  \item SIR model
  \item Harko's SIR model
  \item SEIR model
  \item Attack rate
  \item Amplifier model
  \item Sequential Bayes
  \item Next generation model
  \end{enumerate}
\item Exponential growth
  \begin{enumerate}
  \item Exponential growth
  \item Maximum likelihood of secondary infections
  \item Time dependent reproduction number
  \item Initial growth rate and final size
  \end{enumerate}
\item Estimation from networks
  \begin{enumerate}
  \item Branching process
  \item Agent-based models
  \end{enumerate}
  \end{enumerate}

  Additionally, we describe a number of variance estimates for the above methods for estimating \rr.  These are

\textbf{Variance Estimation}
  \begin{enumerate}
  \item Inverse Hessian
  \item Sensitivity analysis
  \item Jackknife
  \item Delta method
  \end{enumerate}

\section{Epidemic Modeling Terminology}
\label{sec:term}

Here we introduce some of the terminology and notation used throughout the methods for estimating \rr.

{$\rr$} -- reproduction number --

{$\mathcal{R}_t$} -- effective reproduction number
{$\beta$} -- the infection rate

{$\gamma$} -- inverse of the expected recovery time

{$N_t$} -- Number of individuals in the population at time $t$

{$S(t)$} -- the number of susceptible individuals in a population at time $t$

{$I(t)$} -- the number of infected individuals in a population at time $t$

{$R(t)$} -- the number of recovered individuals in a population at time $t$

{$\omega$} -- serial or generation interval -- the time from onset of symptoms in an index case to the onset of symptoms in a subsequent case infected by the index patient

AR -- attack rate -- percentage of the population eventually infected

\section{Details for the \XX Methods}
\label{sec:details}
\subsection{Survival function}
\label{sec:direct}
\subsubsection{Survival Function}
\label{sec:survival_fxn}
We describe the approach to the survival function as presented by \cite{Heffernan2005}.  They note that this is the most storied of the methods of calculating $\rr$.

Let $F(a)$ be the probability that a newly infected individual remains infectious for at least time $a$ (the survival probability).  Moreover, let $b(a)$ the average number of newly infected individuals that an infectious individual will produce per unit time when infected for total time $a$.  Then $\rr$ is derived in Equation (\ref{eq:r0_survivalfxn}),

\begin{align}\label{eq:r0_survivalfxn}
  \rr = \int_0^\infty b(a)F(a)da.
\end{align}

\com{this should be method \#1}

Estimating this quantity is hard, and it is due to this difficulty that the rest of these methods for estimating $\rr$ even exist.  This derivation of $\rr$ is not restricted to compartment models.

%%%%%%%

\subsubsection{Direct Estimation of the Survival Function }
\label{sec:direct-estim-surv}

\cite{fraser2004factors} have a derivation of $\rr$ related to the survival function, where $\beta (\tau)$ represents the infectiousness at time $\tau$ since the infection,

\begin{align*}
\rr = \int_0^\infty \beta(\tau) d\tau
\end{align*}

\com{move in with survival function.  These are \textit{mathematical quantities}}

%%%%%%%%%%%%%%%

\subsubsection{Direct Parameter Estimation }
\label{sec:dpe}

This `plug and chug` approach as described by \cite{lipsitch2003} and \cite{dietz1993estimation} is the idea of estimating parameters $k$, the number of contacts per unit time, $b$, the probability of transmission per contact, and $D$ the mean duration of infectiousness to estimate $\rr$ directly from available data,
\begin{align}
\rr = kbD.
\end{align}

\cite{lipsitch2003} note that directly estimating $\rr$ in such a way may be difficult due to the available data.  They instead estimate $\rr$ using the mean serial interval, the time from onset of symptoms in an index case to the onset of symptoms in a subsequent case infected by the index patient in addition to the assumption of exponential growth of infected cases.

In the event, that th is estimate may be calculated, uncertainty estimates must be compounded, and to estimate $\rr$ alone, we only use data from the initial outbreak of the disease, so often we have a small sample size, resulting in large error bars.

%%%%%%%% 5
\subsubsection{Contact Tracing}
\label{sec:contact_tracing}
As $\rr$ is defined by \cite{anderson1992} ``the average number of secondary infections produced when one infected individual is introduced into a host population where everyone is susceptible,'' it seems reasonable to actually observe the number of secondary infections produced by an initial infector.  This is generally impractical, with seriological information, we could conceivably recreate the transmission of the disease throughout a population.  Even in this case, however, $\rr$ it is unlikely, for one, to have a completely susceptible population.  Moreover, this would implicitly be conditioning on Patient 0's contacts which may not reflect those of the rest of the population.  We write the calculation for $\rr$ under contact tracing in Equation \ref{eq:r0_contacttracing}, where $X_0$ is a random variable denoting the number of individuals in the population person 0 infects.  Here $S_0=N$ indicates that the initial number of susceptibles is $N$, the size of the population.

\begin{align}\label{eq:r0_contacttracing}
\rr = E[ X_0 | S_0 = N]
\end{align}

\cite{eames2003} describe contact tracing as an ``extreme form of targeted control, where the potential next-generation cases are the primary focus,'' i.e. one focuses on treating the contacts of an infected individual rather than applying random treatment.  They note that this method has ``proved to be a highly successful strategy when the number of infectious cases is low'' and ``is the standard tool for eliminating minor outbreaks in the latter stages of disease eradication.''

They describe an estimate for $\rr$ for a SIR model on a network, with $r$ as the rate of transmission across a contact multiplied by the infectious period and $n$ as the number of contacts as

\begin{align*}
\rr = r(n-2)
\end{align*}

\com{should this be split into two methods?}

\subsection{Compartment Models}
\label{sec:cms}

\subsubsection{SIR Model}
\label{sec:sir-model}

In practice, $\rr$ is often calculated through fitted compartment models.  An example of this is the SIR model.  First introduced to epidemiology by  \cite{Kermack700}, the SIR model governs how individuals transition from \textbf{S}usceptible, \textbf{I}nfected, and \textbf{R}ecovered compartments.  The model is represented graphically in Figure \ref{fig::sir}.  In this model, once a person is recovered she cannot have the disease again.

\begin{figure}[h]
\centering
\begin{tikzpicture}[node distance=3cm,auto,>=latex',every node/.append style={align=center}]
    \node [int,  fill = white!50!green] (a)              {$S$};
    \node [int,  fill = white!50!green]           (c) [right of=a] {$I$};
    \node [int,  fill = white!50!green] (e) [right of=c] {$R$};
    \path[->, auto=false] (a) edge node {} (c)
                          (c) edge node {} (e) ;

\end{tikzpicture}
\caption{Depiction of a SIR model.  One can only get the disease once in this model.}\label{fig::sir}
\end{figure}

Here, $N$ is the number of individuals in the system, and  $\beta$ and $\gamma$ are epidemiological parameters.  We have that $\beta$ is the infection rate and $\gamma$ is the inverse of the recovery rate.  The model is represented through the ODEs in \ref{eq::sir}.

A key point in the SIR model and other CMs is the assumption of homogeneity.  This means that ``any infected individual has a probability of infecting any susceptible individual that is reasonably well approximated by the average'' \cite{jonesjh2007}

\begin{align}
\systeme{\frac{dS}{dt} = -\frac{\beta I S}{N}, \frac{dI}{dt} = \frac{\beta I S}{N} - \gamma I, \frac{dR}{dt} = \gamma I}. \label{eq::sir}
\end{align}

In words, susceptible individuals become infected at a rate that is proportional to the proportion of infected people multiplied by $\beta$, the infection rate, and the number of susceptible individuals.  Infected individuals recover at a rate of $\gamma$ multiplied by the number of infected individuals.

An epidemic occurs if 
\begin{align*}
\frac{\beta S I}{N}  - \gamma I > 0 ,
\end{align*}

e.g. the rate of infection is greater than the rate of recovery.  So as long as $\frac{S}{N} \approx 1$ then an epidemic will occur if $\rr := \frac{\beta}{\gamma} > 1$.

In the ideal set up, we have data of the number of Susceptibles, Infectious, and Recovered individuals at different points over time.  From this we would like to estimate the $\beta$ and $\gamma$ parameters and ultimately $\rr$.

There are many ways to estimate the parameters in this model.  We detail a few of them here.

\textbf{Fitting Methods}\label{fitting-methods}

\textbf{E{[}\(\rr\){]}}:

\begin{enumerate}
\item
  Least Squares (\(\beta\), \(\gamma\)) (LS)
\item
  Reparametrized Least Squares (\(\rr\), \(\gamma\)) (ReLS)
\item
  Linear Model Approximation (LMA)
\item
  Linear Model Approximation, all time points (LMAT)
\item
  Max of data (Max)
\item
  Smooth max (spline) (SMax)
\item
  Incidence-Prevalence Ratio (IPR)
\end{enumerate}

\textbf{Least Squares ($\beta$, $\gamma$)}\label{least-squares-beta-gamma}

We find

\begin{align*}
(\hat{\beta}, \hat{\gamma} )&=\text{argmin}_{\beta, \gamma} \sum_{t} (Y(t)_{obs} - Y(t))^2 
\end{align*}

Then \(\hat{R}_0= \frac{\hat{\beta}}{\hat{\gamma}}\).

\textbf{Reparametrized Least Squares ($\rr$, $\gamma$)}\label{reparametrized-least-squares-rux5f0-gamma}

We Reparametrized the the ODES directly with \(\rr\) and \(\gamma\),

\begin{align*}
  \left \{
  \begin{array}{cl}
    \dot{X} &= - \rr \gamma Y \frac{X}{N}\\
    \dot{Y} &=  \rr \gamma Y \frac{X}{N}  - \gamma Y \\
    \dot{Z} &=  - \gamma Y 
  \end{array}
  \right .
  \end{align*}

We find

\begin{align*}
(\hat{R}_0, \hat{\gamma} )&=\text{argmin}_{\rr, \gamma} \sum_{t} (Y(t)_{obs} - Y(t))^2 .
\end{align*}

We use the \(\rr\) directly from the above calculation.

\textbf{Linear Model Approximation (degree
10)}\label{linear-model-approximation-degree-10}

We fit a linear polynomial of \(t\) with degree \(K= 10\) model to \(X\)
and \(Y\) using least squares to find the coefficients (\(\hat{x}_k\),
\(\hat{y}_k\)),

\begin{align*}
X(t) &= \sum_{k=0}^K x_k t^k\\
{Y}(t) &= \sum_{k=0}^K y_k t^k
\end{align*}

Then, we estimate the derivatives as

\begin{align*}
\hat{X}^\prime(t) &= \sum_{k=1}^K k \hat{x}_k t^{k-1}\\
\hat{Y}^\prime(t) &= \sum_{k=0}^K k \hat{y}_k t^{k-1}
\end{align*}

Then \(\rr\) is calculated as according to the ODES,
\[\rr = \frac{\hat{X}^\prime(0)}{ \hat{X}^\prime(0) + \hat{Y}^\prime(0)} \cdot \frac{N}{X(0)} \]

\textbf{Linear Model Approximation, All Time Points (degree
10)}\label{linear-model-approximation-all-time-points-degree-10}

We fit a linear polynomial of \(t\) with degree \(K= 10\) model to \(X\)
and \(Y\) as above, with a slight modification in how we calculate
\(\rr\),
\[\rr = \frac{1}{\# \text{ Obs}}\sum_t \frac{\hat{X}^\prime(t)}{ \hat{X}^\prime(t) + \hat{Y}^\prime(t)} \cdot \frac{N}{X(0)} \]

\textbf{Max of Data}\label{max-of-data}

We note that \(Y^\prime(t) > 0\) when \(\rr < \frac{S(t)}{N}\). The max
of \(Y\) occurs when \(Y^\prime = 0\) and hence
\[\rr = \frac{N}{X_{obs}(t^*)},\] where
\(t^* = \text{arg max}_{t} Y_{obs}(t)\).

\textbf{Smooth Maximum of Data}\label{smooth-maximum-of-data}

We use a spline with 4 degrees of freedom to fit both \(X\), and \(Y\).
We then apply the principle as above, except now,
\[\rr = \frac{N}{\hat{X}(t^*)},\] where
\(t^* = \text{arg max}_{t} \hat{Y}(t)\).

\textbf{Incidence to Prevalence
Ratio}\label{incidence-to-prevalence-ratio}
In terms of data from the SIR model, incidence $J(t) \approx -(X(t+1) - X(t))$, and the IPR$(t) = \frac{J(t)}{Y(t)}$. The actual reproduction number, $R_a(t) = \text{IPR}(t)\cdot D$ where $D = 1 /\gamma$.  This method assumes that we have some prior knowledge about $\gamma$.  Thus we use as our estimate,

\textbf{Smoothed Incidence to Prevalence Ratio}
We use the same method as above, but first fit splines with 4 degrees of freedom, $\hat{X}(t)$ and $\hat{Y}(t)$, , to fit to $X_{obs}$ and $Y_{obs}$, respectively.  Then  $J(t) \approx -(\hat{X}(t+1) - \hat{X}(t))$, and the IPR$(t) = \frac{J(t)}{\hat{Y}(t)}$.  Then
\begin{align*}
\rr &= R_a(0).
\end{align*}

\subsubsection{Harko}
Recently, \citep{harko2014exact} were able to reduce the SIR (XYZ) equation, with the following set of ODEs

\begin{align*}
  \left \{ \begin{array}{ll}
             \frac{dX}{dt} &= - b X(t) Y(t) \\
             \frac{dY}{dt} &=  b X(t) Y(t) - \gamma Y\\
             \frac{dZ}{dt} &= \gamma Y,\\
           \end{array}
  \right .
\end{align*}

where $b = \frac{\beta}{N}$, to one differential equation.  Here,
\begin{align*}
  u(t) &= e^{\frac{\beta}{\gamma}Z} \\
  t - t_0 &= \int_{u_0}^u \frac{d \xi}{\xi (C_1 - \gamma \log \xi + X(0) b \xi)},
\end{align*}
where $C_1 = -\frac{b}{N}$.  Then

\begin{align}
  X &= X(0) u  = X(0) e^{\frac{\beta}{\gamma}Z} \nonumber\\
  \log \frac{X(t)}{X(0)} &=  \frac{b}{\gamma}Z =  \frac{\beta}{\gamma N} Z \nonumber\\
  \log \frac{X(t)}{X(0)} &=  \frac{\rr}{N} Z \label{eq:harko_lin}
\end{align}

If we add error into Equation (\ref{eq:harko_lin}), then we have

\begin{align}
  \log \frac{X(t)}{X(0)} &=  \frac{\rr}{N} Z  + \epsilon_t\label{eq:r0_harko}
\end{align}

If we assume $\epsilon_t \overset{iid}{\sim}N(0, \sigma^2)$, then we can directly estimate $\rr/N$ through linear regression and also an expression for the variance.

\subsubsection{SEIR Model}
\label{sec:seir-model}

The SEIR model is a common adaptation of the SIR compartment model \com{add references that use this method}.  The ``E'' compartment stands for ``exposed'' and represents the stage where individuals are infected but not yet infectious.  The model is described by the following set of equations, as described by \cite{cintronarias2009}.  This particular model here does not include birth and death rates but the model can be adapted to do so,
\begin{align*}
  \frac{dS}{dt} &= - \beta SI \\
  \frac{dE}{dt} &= \beta SI  - \alpha E\\
  \frac{dI}{dt} &= \alpha E - \gamma I \\
  \frac{dR}{dt} &= \gamma I.
\end{align*}

Then $\rr = \beta / \gamma$, just as in the SIR model.  As in the SIR model, $\beta$ and $\gamma$ have the same interpretation and $\alpha$  The methods used to estimate the parameters $\beta$ and $\gamma$ are of the same class used to estimate those parameters in the SIR model.  The important difference between this method and the SIR model is the assumption of a 4th class, the exposed compartment.  Typically, this method will result in ...\com{figure out}... than the SIR method.  Homogeneous mixing of the compartments is assumed.

\subsubsection{Attack Rate}

The Attack Rate (AR) as described by \cite{obadia2012r0} as the ``percentage of the population eventually infected.''  Under the SIR (XYZ) model with constant population,

\begin{align}\label{r0_attackrate}
\rr =  \frac{\log \left (\frac{1  - AR}{X(0)/N}  \right ) }{AR - (1 - X(0)/N)}
\end{align}

This method can only be used to assess $\rr$ after a disease has passed through a population.  Additionally, if either $\beta$ or $\gamma$ were affected during the disease's lifetime then $\rr$ cannot be properly assessed.

\subsubsection{Amplifier Model}
\cite{blower2004} discuss the case where a disease may have different strains and consequently different reproduction numbers.  They introduce a multistrain mathematical model called the amplifier model to deal with this case.

This model is based on the the SLIR compartment model, where $L$ stands for latently infected individuals and $T$ for diseased individuals.  There is a set of ODEs for each strain of the disease with different parameters for recovery, transmission rate, and drug resistance of the disease.  Additionally, there is mixing of the different strains as ``the amplifier model also allows for immigrating and emigration of individuals of all types, as well as reinfection of latently infected individuals'' \citep{blower2004}.  The details may be found in their manuscript.  Ultimately, their derivation for $\rr^{(i)}$ where $i$ stands for the $i$th strain is
\begin{align*}
\rr^{(i)} = S^* \frac{ ( \beta_i^T + \beta_i^L)\nu_i + \beta_i^T \mu_i^L}{(\nu_i + \mu_i^L)(c_i + k_{i,i+1} + \mu_i^T)}.
\end{align*}
Here, $S^*$ is the number of susceptibles individuals in the disease-free equilibrium in the absence of immigration or emigration; $\beta_i^L$ and $\beta_i^T$ are strain-specific transmission rates for the latent, and infected classes, respectively; $\nu_i$ denotes the progression rate from latently infected to disease for strain $i$; $\mu_i^L$ and $\mu_i^{T}$ are death rates for the latent and infected classes, respectively; $c_i$ is the cure rate; and $k_{i, i+1}$ is a parameter associated with the degree of amplification of drug resistance.  They note that ``each reproduction number is the product of average number of secondary infections caused per unit time, the aver time a case remains infectious and the probability that an infected individual develops disease.''

In the amplifier model, $\rr$ was derived from using the stability point (disease free equilibrium state in the absence of immigration or emigration) of the system by solving the characteristic polynomial for the eigenvalues

\subsubsection{Sequential Bayes}\label{sec:seqbayes}

Posited by \cite{bettencourt2008} and summarized in \cite{obadia2012r0}, this method uses a Bayesian approach to an approximation to the classic SIR model.  The approximated SIR model is that the incidence at $t+1$, $N(t+1)$ is, with $\gamma$ as the  average inverse of the infectious period and $R$ as the effective reproduction number, which we take here to be $\rr$.

In order to estimate $\rr$, we must have some idea about the serial interval.
\begin{align*}
N(t+1)  \sim Poisson( N(t) \exp \left \{  \gamma (R-1)\right \})
\end{align*}
Then, the posterior distribution of $\rr$ given the previous days' incidences is
\begin{align*}
  P(\rr | N_0, \dots, N_{t+1}) = \frac{P(N_{t+1} | \rr, N_0, \dots, N_t)P(\rr| N_0, \dots, N_t)}{P(N_0, \dots, N_{t+1})}.
\end{align*}
This method is sequential in that the prior distribution for $\rr$ comes from the previous day.  The initial prior for $\rr$ is assumed to be flat.  This method results in a posterior distribution from which credible intervals may be obtained.  This method assumes, initial growth in incidence to be exponential, and homogeneous mixing of populations as with any compartment model.

\subsubsection{Next Generation Model}
\label{sec:ngm}

The Next Generation Model (NGM) is a generalization of any compartment model at the infection-free steady state. This model solves the problem of having an expression of $\rr$ in terms of the epidemiological parameters for a general compartment model.  Originally, introduced by \citep{diekmann1990}, \cite{diekmann2009} posit a recipe to find $\rr$ for a wide class of compartment models, including commonly used ones such as MSEIR, MSEIRS, SEIR, SEIRS, SIR, SIRS, SEI, SEIS, SI, and SIS \citep{hethcote2000}. We summarize that recipe here, using their notation.  First, define the \textit{infected subsystem} as ``the equations of the ODE system that describe the production of new infections and changes in state among infected individuals.''  Let $x = (C_1, C_2, \dots, C_m)$ where $C_i$ are the different compartments of the infected subsystem.  The steps to find $\rr$ are as follows:

\begin{enumerate}
\item ``Linearize the infected subsystem of nonlinear ODEs about the infection-free steady state''
\item Decompose the linearized infected subsystem into $(T + \Sigma )x$ where $T$ is a $m\times m$ matrix of transmissions and $\Sigma$ is a $m \times m$ matrix of transitions.
\item $\rr$ is the spectral radius (i.e. the dominant eigenvalue) of $K_L=-T \Sigma^{-1}$.  Here $L$ stands for the ``large domain.''
\end{enumerate}

Here, $T_{ij}$ is the rate of transmission of \textit{newly} infected individuals in state $i$ created by individuals in state $j$.  $\Sigma_{ij}$ is the transition rate of individuals into compartment $i$ from compartment $j$.

This method has the assumptions of a basic compartment model: homogeneous mixing among populations and the law of mass action, i.e., that the number of infected individuals are proportional to the number of infected individuals at the previous step.  The estimations for $\rr$ from the previously mentioned compartment models \com{list those models} may also be derived using this method.  Additionally, this method is advantageous in that the matrix $\Sigma$ has an intuitive meaning.

%%%%%%%%%%%%%%%%%%%%%%%%

\subsection{Exponential Growth}\label{sec:exp-growth}

\subsubsection{Exponential Growth}
\label{sec:expgrowth}
\cite{wallinga2007generation} report that $R$ and hence $\rr$ may calculated by using the fact that infection ``counts increase exponentially in the initial phase of an epidemic.''  We need to know $r$, the \textit{per capita} change in the number of new cases per unit of time and $T_c$ the serial interval.

Then, we have

\begin{align}\label{eq:lotka}
\rr = \exp{(r T_c)}
\end{align}
or its first order approximation

\begin{align}\label{eq:anderson}
\rr = 1 + rT_c.
\end{align}

Equation \eqref{eq:lotka} is derived from a demographic view using the Lotka Euler survival equations, whereas Equation \eqref{eq:anderson}is derived through an epidemiologist view using \com{something}.  These estimates for $\rr$ can be strikingly different.

\cite{wallinga2007generation}, instead, use a moment generating function expression for $\rr$.  With $g(a)$ as the distribution of age at infection, then

\begin{align*}
\rr^{-1} &= \int_{a=0}^\infty e^{-ra}g(a)da.
\end{align*}

For all these above estimations of $\rr$, we observe the duration of generation intervals in a period of exponential growth.  Deciding when this occurs and which data to use to estimate $\rr$ is difficult.

\subsubsection{Maximum Likelihood Estimator of Secondary Infections}\label{sec:mle-si}
This method, described by \cite{forsberg2008}, finds the estimate of $\rr$ by maximizing the likelihood function under certain assumptions.  We assume that the ``the number of secondary cases produced by an infected individual follows a Poisson distribution, and that the serial interval is described by a multinomial distribution.''  Recall, the serial interval is the distribution of time between a primary case developing symptoms and a case infected by the primary case developing symptoms.  An approximate version of this can be simplified to a thinned Poisson where
\begin{align*}
  L(\rr, \mathbf{p}) = \prod_{t=1}^T \frac{e^{- \rr \sum_{j=1}^{\min(k,t)}N_{t-j}p_j}\left (\rr \sum_{j=1}^{\min(k,t)}N_{t-j}p_j \right )^{N_t}}{N_t!}.
\end{align*}
Here, $k$ is the maximal amount of the serial number, $T$ is the total time and $p_j$ is the probability of displaying symptoms on day $j$ after being infected.  The number of cases on day $t$, $N_t$ are incidence counts that are known. \com{more on this}

This method assumes a Poisson distribution on $\rr$ in addition to assuming a multinomial on the serial interval.  Additionally, this method also assumes a period of exponential growth,

%%%%%%%%%%%%% 5

\subsubsection{Time Dependent Reproduction Number}\label{sec:timedep}
This is another method used to estimate the effective reproduction number, which again, we take to be $\rr$ here.  This is a likelihood based approach where we maximize the likelihood that case $i$ has been infected by case $j$ at certain time using the generation interval, $w(t)$ where $t_i$ is when case $i$ was infected and $N$ is the total number of cases,
\begin{align*}
  p_{ij} &= w(t_i- t_j) / \sum_{i \neq k} w(t_i - t_k),\\
  \rr &= \frac{1}{I(0)}\sum_{i=1}^N p_{i0}
  \end{align*}

  It is assumed we have some idea about the generation interval distribution, $w(t)$.

\subsubsection{Initial Growth Rate and Final Size}
\label{sec:igr-fs}
\com{this should be moved up in the order}

Initial growth rate, as described by \cite{dietz1993estimation} is derived by \cite{anderson1986}, initially to study the spread of HIV estimates $\rr$ through
\begin{align*}
\rr = \frac{D \ln 2} {t_d} + 1,
  \end{align*}
  with $D$ as the average incubation period and $t_d$ as the doubling time during its early stages.  This method assumes you know when the ``early stage'' ends along with knowledge of the average incubation period.

  Final size, on the other hand, only looks at $\rr$ once a disease has run its course.  Here $\mu_0$ is the proportion of initially infected individuals and $\mu_\infty$ is the final, cumulative proportion of infected cases at the end of an epidemic,
  \begin{align*}
    \rr = (1- \mu_\infty)^{-1} \ln \mu_\infty^{-1}.
  \end{align*}

  For final size, it is assumed the initial population is completely susceptible

  These two methods allow one to look at $\rr$ at both the beginning and the end of an epidemic.  However, both approaches have advantages and disadvantages.  For the initial growth rate, while one can see $\rr$ at early stages, estimating the doubling time $t_d$ is difficult.  For final size, one has to wait until the end of an epidemic to calculate $\rr$, and preventions of the disease may skew the final result.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\subsection{Estimation from Networks}
\label{sec:network}

\subsubsection{Branching Process}
\label{sec:branching-process}

Branching processes are models for population growth and track generations of new offspring.  We describe branching processes as adapted for epidemic modelling from the notation of \cite{grimmett1991}.  An infector $i$ may produce new infections over the lifetime of the original infector's disease.  These new infections are called offspring.  The offspring the orginal infector directly produces are called the first generation.  The offspring of the first generation are called the second generation and so on.  An important assumption in branching processes is that the offspring produced from different infectors are indepedent and identically distributed (iid), and so for a branching process to be a valid model in epidemic modeling, we must assume a large population with few infectors.

\com{draw a family tree}

\cite{getz2006} give a way to calculate \rr from a branching process.  ```We define the offspring distribution $\{q_i \}_{i=0}^\infty$, where $q_i$  is the probability that an infectious individual infects $i$ other individuals.  Thus we require $\sum_{i=0}^\infty q_i =1$ and note that $\rr$, the mean number of cases contracting disease from each infective, is simply given by
  \begin{align*}
    \rr = \sum_{i=0}^\infty iq_i."
  \end{align*}

\subsubsection{Agent-Based Models}
\label{sec:agent-based-models}
Agent-based models (ABMs) or individual level models (ILMs) are ``bottom-up'' models, meaning micro patterns are simulated and macro patterns are then derived from these simulations.  ABMs consist of agents which represent individuals (e.g., people, mosquitoes, poultry) and a series of activities in which the agents can affect one another and evolve over time.

In epidemiology, ABMs can be used to simulate the spread of disease.  ABMs allow for every detail of an infection to be known, in the context of $\rr$, meaning we know exactly who is infected by whom and when.  Thus $\rr$ can be calculated simply through
\begin{align*}
  \rr = \frac{1}{\# \text{ simulations}}\frac{1}{\#\text{ initial infections}} \sum_{s \in \text{ simulations}}\sum_{i \text{ infected at } t=0} \# \text{ infected by $i$ in simulation }s.
\end{align*}

The difficulty in the ABM lies in calibrating the evolving of agents over time to accurately reflect reality, which is a problem that cannot be understated.  Thus, it is typically necessary to have knowledge of parameters such as the serial interval, transmission rate, recovery rate, etc.  Additionally, we need to have accurate agents in number, characteristics, and their implied network structure.

Examples of calculations of $\rr$ using ABMs/ILMs are in \cite{breban2007,ahmed2013variance}.

\subsection{Variance Methods}
\label{sec:methods}

\subsubsection{Inverse Hessian}\label{inverse-hessian}

When we calculate the parameters of the SIR model using least squares,
we use gradient descent using the Hessian along the way. We use the
relationship with the Fisher Information Matrix to approximate the
covariance matrix of the parameters.

\subsubsection{Sensitivity Analysis}\label{sensitivity-analysis}

We estimate \(\frac{\partial Y}{\partial \beta}\) (or
\(\frac{\partial Y}{\partial \rr}\) when using reparametrized least
squares) and \(\frac{\partial Y}{\partial \gamma}\) at each time point
\(t\) using the estimated \(\beta\) or \(\gamma\) from the estimation
process. Then an approximation of the covariance matrix is
\((\chi^T \chi)^{-1}\) where
\(\chi_{ij} = \frac{ \partial Y(t_i)}{ \partial \theta_j}\) where
\(\theta\) is a vector of parameters. (add citation)

\subsubsection{Jackknife}\label{jackknife}

We estimate \(\rr^{(-i)}\) using a specified method leaving out the
\(i\)th time point out from the data each iteration. Assume there are
\(n\) time points of observations. Then
\[V[\rr] = \frac{n-1}{n} \sum_{i=1}^n \left (\rr^{(-i)} - \frac{1}{n} \sum_{i=1}^n\rr^{(-i)} \right)^2\]

\subsubsection{Delta Method}\label{delta-method}

When the method estimates \(\beta\) and \(\gamma\) instead of \(\rr\)
directly, we use the delta method approximation to calculate the
variance of \(\rr\). Here, we know
\(\rr = h(\beta, \gamma) = \frac{\beta}{\gamma}\). Then
\(\bigtriangleup h = (\frac{1}{\gamma}, - \frac{\beta}{\gamma^2})^T\).
Then
\(V[\rr] = \bigtriangleup h^T \Sigma_{\beta, \gamma} \bigtriangleup h\),
where \(\Sigma_{\beta, \gamma}\) is covariance matrix of \(\beta\) and
\(\gamma\).
%%%%%%%%%%%%%%%

\section{Analysis and Results}
\label{sec:results}

\section{Discussion}
\label{sec:dis}

\label{sec:details}

\bibliographystyle{apa}%Choose a bibliograhpic style
\bibliography{Master}

\end{document}

%%%%%%%%%%%%%%%%%%%%%%
\Textbf{Title:}

\textbf{Author:}

\textbf{Citation:}

\textbf{Major themes:}

\textbf{Notes:}
\\
%%%%%%%%%%%%%%%%%%%%%%

\begin{figure}[h]
\begin{center}
\includegraphics[width=4in]{mvhw7_3c.pdf}
\end{center}
\caption{
Biplots of the different continuous variables and their correlations.}\label{fig3c}
\end{figure}

%% two pictures whoa

\begin{figure}[h]
\centering

\begin{figure}[h]
\centering

\begin{subfigure}{.5\textwidth}
  \centering
  \includegraphics[width=1\linewidth]{mt_eda_cont_hists.pdf}
  \caption{Histograms of Arrival Delay and continuous covariates.  Arrival delay seems to have a right skewed distribution.  This may indicate that we will be transforming this variable later on.  After transforming Air Time and Distance by a log transformation, we don't really seem to have many outliers in our covariates.  We seem to have outliers in the CRS Dep. Time and Arrival Time; however, time is cyclical and so these are not, in fact outliers.}
  \label{hists}
\end{subfigure}%
\begin{subfigure}{.5\textwidth}
  \centering
  \includegraphics[width=1\linewidth]{mt_eda_cont_hists.pdf}
  \caption{\textcolor{red}{Placeholder}}
  \label{tabs1}
\end{subfigure}
\caption{}
\end{figure}

%%
\begin{figure}
\centering
\begin{subfigure}{.5\textwidth}
  \centering
  \includegraphics[width=1\linewidth]{resids_full.pdf}
  \caption{}
  \label{residsf}
\end{subfigure}%
\begin{subfigure}{.5\textwidth}
  \centering
  \includegraphics[width=1\linewidth]{diags_full.pdf}
  \caption{ }
  \label{diagsf}
\end{subfigure}
\caption{}
\end{figure}

### 2018-07-03 paper/old_drafts/draft_v2.tex
finally on git

--- 2018-07-03 paper/old_drafts/draft_v1.tex
@@
 
-\newcommand{\XX}{15 } % number of methods
+\newcommand{\XX}{21 } % number of methods
 \newcommand{\rr}{\ensuremath{\mathcal{R}_0}}
@@
 
-%\tableofcontents
+\tableofcontents
 
@@
   \begin{enumerate}
-  \item Survival function
-  \item Direct estimation of the survival function
-  \item Direct parameter estimation
-  \item Contact tracing
+  \item Survival function (SF)
+  \item Direct estimation of the survival function (DESF)
+  \item Direct parameter estimation (DPE)
+
   \end{enumerate}
@@
   \item SIR model
-  \item Harko's SIR model
-  \item SEIR model
-  \item Attack rate
-  \item Amplifier model
-  \item Sequential Bayes
-  \item Next generation model
+    \begin{itemize}
+    \item Least Squares (SIR LS)
+    \item Reparametrized Least Squares (SIR ReLS)
+    \item Linear Model Approximation (SIR LMA)
+    \item Linear Model Approximation, all time points (SIR LMAT)
+    \item Max of Data (SIR Max)
+    \item Smooth max (SIR SMax)
+    \item Incidence-Prevalence Ratio (SIR IPR)
+    \item Smooth Incidence-Prevalence Ratio (SIR SIPR)
+    \item Harko's SIR model (HSIR)
+    \item Attack rate (SIR AR)
+    \item Sequential Bayes (SIR SB)
+    \end{itemize}
+  \item SEIR model (SEIR)
+  \item Amplifier model (Amp)
+  \item Next generation model (NGM)
   \end{enumerate}
@@
   \begin{enumerate}
-  \item Exponential growth
-  \item Maximum likelihood of secondary infections
-  \item Time dependent reproduction number
-  \item Initial growth rate and final size
+  \item Exponential growth (EG)
+  \item Maximum likelihood of secondary infections (MLSI)
+  \item Time dependent reproduction number (TDR)
+  \item Initial growth rate and final size (IGRFS)
   \end{enumerate}
@@
   \begin{enumerate}
-  \item Branching process
-  \item Agent-based models
+  \item Contact tracing (CT)
+  \item Branching process (BP)
+  \item Agent-based models (ABM)
   \end{enumerate}
@@
 \begin{enumerate}
-  \item Inverse Hessian
+\item Delta method
   \item Sensitivity analysis
-  \item Jackknife
-  \item Delta method
   \end{enumerate}
@@
 {$\mathcal{R}_t$} -- effective reproduction number
+
 {$\beta$} -- the infection rate
@@
 
-{$S(t)$} -- the number of susceptible individuals in a population at time $t$
+{$X(t)$} -- the number of susceptible individuals in a population at time $t$
 
-{$I(t)$} -- the number of infected individuals in a population at time $t$
+{$Y(t)$} -- the number of infected individuals in a population at time $t$
 
-{$R(t)$} -- the number of recovered individuals in a population at time $t$
+{$Z(t)$} -- the number of recovered individuals in a population at time $t$
 
@@
 \label{sec:direct}
+
+The survival function is intimately with the theory of \rr, which originates from the field of demography in the late 1800s \citep{dietz1993estimation}.  Here, we have
+\begin{align*}
+\rr = \int_0^\infty p(a) \beta(a) da
+\end{align*}
+where $p(a)$ denotes the probability of a woman surviving age $a$ and $\beta(a)$ the rate of giving birth to a  girl for an individual of age $a$.
+This interpretation also explains the name of \rr, which was imported to the field of epidemiology by MacDonald and Smith as reported by \cite{dietz1993estimation}.
+
+This is the chronologically first and in many ways, the most direct method to calculate \rr.
+However, this model also seems to be the most difficult to estimate.
+We describe four methods attempting to estimate \rr from the survival function:  the survival function alone, direct estimation, and direct parameter estimation.
+
 \subsubsection{Survival Function}
 \label{sec:survival_fxn}
-We describe the approach to the survival function as presented by \cite{Heffernan2005}.  They note that this is the most storied of the methods of calculating $\rr$.
+The first method, is not surprisingly, the survival function itself as presented by \cite{Heffernan2005}.
+We change the notation from the above survival function to conform to the standards of epidemiology.
 
-Let $F(a)$ be the probability that a newly infected individual remains infectious for at least time $a$ (the survival probability).
-Moreover, let $b(a)$ the average number of newly infected individuals that an infectious individual will produce per unit time when infected for total time $a$.
+Let $F(a)$ be the probability that a newly infected individual remains infectious for at least time $a$ (the survival probability,$p(a)$ in the previous definition).
+Moreover, let $b(a)$ $(\beta (a)$ in the previous definition), the average number of newly infected individuals that an infectious individual will produce per unit time when infected for total time $a$.
 Then $\rr$ is derived in Equation (\ref{eq:r0_survivalfxn}),
@@
 
-\com{this should be method \#1}
+This method presents one of the unique difficulties of estimating \rr: that of reconciling generations and time.
+On the other hand, this method is flexible in that this derivation of $\rr$ is not restricted to compartment models.
 
-Estimating this quantity is hard, and it is due to this difficulty that the rest of these methods for estimating $\rr$ even exist.  This derivation of $\rr$ is not restricted to compartment models.
+Estimating this quantity in Equation \ref{eq:r0_survivalfxn} is hard and rarely attempted as $b(a)$ is particularly difficult to estimate, and it is due to this difficulty that the rest of these methods for estimating $\rr$ even exist.
 
@@
 
-\cite{fraser2004factors} have a derivation of $\rr$ related to the survival function, where $\beta (\tau)$ represents the infectiousness at time $\tau$ since the infection,
-
+\cite{fraser2004factors} have a derivation of $\rr$ related to the survival function based on the Van Forester equations,
 \begin{align*}
-\rr = \int_0^\infty \beta(\tau) d\tau
+  \frac{\partial Y(t, \tau)}{\partial t} +   \frac{\partial Y(t, \tau)}{\partial \tau} = 0,
 \end{align*}
-
-\com{move in with survival function.  These are \textit{mathematical quantities}}
+where $Y(t,\tau)$ is the number of people at time $t$ who were infected time $\tau$ ago.  $\rr$  is then,
+\begin{align*}
+\rr = \int_0^\infty \beta(\tau) d\tau,
+\end{align*}
+ where $\beta (\tau)$ represents the infectiousness at time $\tau$ since the infection.  Again, this model is more mathematical in nature than practical as $\beta (\tau)$ is difficult to estimate.
 
@@
 \cite{lipsitch2003} note that directly estimating $\rr$ in such a way may be difficult due to the available data.
-They instead estimate $\rr$ using the mean serial interval, the time from onset of symptoms in an index case to the onset of symptoms in a subsequent case infected by the index patient in addition to the assumption of exponential growth of infected cases.
+They instead estimate $\rr$ using the mean serial interval, $\omega$, the time from onset of symptoms in an index case to the onset of symptoms in a subsequent case infected by the index patient in addition to the assumption of exponential growth of infected cases.
 
-In the event, that th is estimate may be calculated, uncertainty estimates must be compounded, and to estimate $\rr$ alone, we only use data from the initial outbreak of the disease, so often we have a small sample size, resulting in large error bars.
-
-%%%%%%%% 5
-\subsubsection{Contact Tracing}
-\label{sec:contact_tracing}
-As $\rr$ is defined by \cite{anderson1992} ``the average number of secondary infections produced when one infected individual is introduced into a host population where everyone is susceptible,'' it seems reasonable to actually observe the number of secondary infections produced by an initial infector.
-This is generally impractical, with seriological information, we could conceivably recreate the transmission of the disease throughout a population.
-Even in this case, however, $\rr$ it is unlikely, for one, to have a completely susceptible population.
-Moreover, this would implicitly be conditioning on Patient 0's contacts which may not reflect those of the rest of the population.
-We write the calculation for $\rr$ under contact tracing in Equation \ref{eq:r0_contacttracing}, where $X_0$ is a random variable denoting the number of individuals in the population person 0 infects.
-Here $S_0=N$ indicates that the initial number of susceptibles is $N$, the size of the population.
-
-\begin{align}\label{eq:r0_contacttracing}
-\rr = E[ X_0 | S_0 = N]
-\end{align}
-
-\cite{eames2003} describe contact tracing as an ``extreme form of targeted control, where the potential next-generation cases are the primary focus,'' i.e.
-one focuses on treating the contacts of an infected individual rather than applying random treatment.
-They note that this method has ``proved to be a highly successful strategy when the number of infectious cases is low'' and ``is the standard tool for eliminating minor outbreaks in the latter stages of disease eradication.''
-
-They describe an estimate for $\rr$ for a SIR model on a network, with $r$ as the rate of transmission across a contact multiplied by the infectious period and $n$ as the number of contacts as
-
-\begin{align*}
-\rr = r(n-2)
-\end{align*}
-
-\com{should this be split into two methods?}
+In the event, that the is estimate may be calculated, uncertainty estimates must be compounded, and to estimate $\rr$ alone, we only use data from the initial outbreak of the disease, so often we have a small sample size, resulting in large confidence sets.
 
@@
 
+A large and the most common set of models used estimate $\mathcal{R}$ is that of compartment models (CMs).
+As adapted from \cite{daley2001epidemic}, compartment models consist of the following.
+CMs model how objects move from one discrete state to another over time and provide a high-level model of how diseases, for instance, may be spread throughout a population.
+Within the CM framework, we make four essential assumptions,
+\begin{enumerate}
+\item The compartments are discrete and have no overlap.
+\item The transition of objects among compartments is described by a set of equations.
+\item The populations mix homogeneously
+  \item The number of objects in each compartment at time $t=0$ is known
+  \end{enumerate}
+
+Another common assumption for compartment models, is the law of mass action, a property borrowed from chemistry which says that the mass of the product is proportional to the mass of the reactants.
+\cite{anderson1992} adapt the law of mass action to epidemiology saying, ``the net rate of spread of infection is assumed to be proportional to the product of the susceptible people times the density of the infectious individuals.''  Therefore, the flow of objects from one compartment to another may be dependent on the percentage of objects within one or many of the compartments.
+The assumption of mass action holds particularly true for disease modelling as typically we assume infections at the next state are proportional to both the susceptible and infected populations at the current state.
+
+For disease modelling, all CMs fall into the SI-framework, meaning that there is a susceptible population that may become infected.
+Thus, we require there be susceptible and infected compartments but also allow additional compartments such as the recovered state, an exposed latent state, and more.
+CMs may either work with continuous or discrete time and \cite{getz2006} remark that there is no preference between the two.
+
+In this section, we discuss 14 different CMs used to estimate $\mathcal{R}$.
+Different CMs yield different mathematical derivations of \rr.
+That is, for example, an SIR model with non-constant population will yield a different mathematical formula for \rr than the SEIR model with non-constant population.
+The only difference in these two models is that the SEIR model.
+That said, even within the same compartment model framework, we can estimate \rr in a variety of ways.
+In fact, we describe 11 methods to estimate \rr, within the SIR framework!
+Most of these methods are descriptions to estimate parameters within the SIR model and may be extended to other CM frameworks such as the SEIR model.
+The final method described in this section, the next generation operator (NGM) is a general method to derive \rr from almost any CM.
+
 \subsubsection{SIR Model}
@@
 
-In practice, $\rr$ is often calculated through fitted compartment models.
-An example of this is the SIR model.
+We begin with the most popular CM, the SIR model.
 First introduced to epidemiology by  \cite{Kermack700}, the SIR model governs how individuals transition from \textbf{S}usceptible, \textbf{I}nfected, and \textbf{R}ecovered compartments.
@@
 \begin{tikzpicture}[node distance=3cm,auto,>=latex',every node/.append style={align=center}]
-    \node [int,  fill = white!50!green] (a)              {$S$};
-    \node [int,  fill = white!50!green]           (c) [right of=a] {$I$};
-    \node [int,  fill = white!50!green] (e) [right of=c] {$R$};
+    \node [int,  fill = white!70!blue] (a)              {$S$};
+    \node [int,  fill = white!70!blue]           (c) [right of=a] {$I$};
+    \node [int,  fill = white!70!blue] (e) [right of=c] {$R$};
     \path[->, auto=false] (a) edge node {} (c)
@@
 \end{figure}
-
-Here, $N$ is the number of individuals in the system, and  $\beta$ and $\gamma$ are epidemiological parameters.
-We have that $\beta$ is the infection rate and $\gamma$ is the inverse of the recovery rate.
-The model is represented through the ODEs in \ref{eq::sir}.
-
-A key point in the SIR model and other CMs is the assumption of homogeneity.
-This means that ``any infected individual has a probability of infecting any susceptible individual that is reasonably well approximated by the average'' \cite{jonesjh2007}
-
+Here, $N$ is the number of individuals in the system.
+As we see no arrows pointing out of the nodes in Figure \ref{fig::sir}, then $N$ is constant throughout time.
+Simple adaptations of the SIR model can include birth and death rates, which will correspondingly change the derivation of $\rr$ (for more information, see Section \ref{sec:ngm}).
+The parameters  $\beta$ and $\gamma$ have epidemiological meaning.
+Here, $\beta$ is the average infection rate and $\gamma$ is the inverse of the average recovery rate.
+The model is represented through the ODEs in Equation \ref{eq::sir}.
+Throughout, we use $X$, $Y$, and $Z$ for stand-ins for the ODEs for $S$, $I$, and $R$, respectively to avoid confusion.
 \begin{align}
-\systeme{\frac{dS}{dt} = -\frac{\beta I S}{N}, \frac{dI}{dt} = \frac{\beta I S}{N} - \gamma I, \frac{dR}{dt} = \gamma I}. \label{eq::sir}
+\systeme{\frac{dX}{dt} = -\frac{\beta XY}{N}, \frac{dY}{dt} = \frac{\beta XY}{N} - \gamma Y, \frac{dZ}{dt} = \gamma Z}. \label{eq::sir}
 \end{align}
@@
 \begin{align*}
-\frac{\beta S I}{N}  - \gamma I > 0 ,
+  \frac{dY}{dt} &> 0 \\
+\frac{\beta S I}{N}  - \gamma I &> 0 ,
 \end{align*}
 
-e.g. the rate of infection is greater than the rate of recovery.  So as long as $\frac{S}{N} \approx 1$ then an epidemic will occur if $\rr := \frac{\beta}{\gamma} > 1$.
+that is,  the rate of new infections is greater than the rate of recovery.  So as long as $\frac{S}{N} \approx 1$ then an epidemic will occur if
+\begin{align}\label{eq:sir_r0}
+  \rr = \frac{\beta}{\gamma} > 1.
+  \end{align}
 
@@
 
-There are many ways to estimate the parameters in this model.  We detail a few of them here.
+There are numerous ways to estimate the parameters in this model.  We detail 11 of them here.
 
-\textbf{Fitting Methods}\label{fitting-methods}
-
-\textbf{E{[}\(\rr\){]}}:
-
-\begin{enumerate}
-\item
-  Least Squares (\(\beta\), \(\gamma\)) (LS)
-\item
-  Reparametrized Least Squares (\(\rr\), \(\gamma\)) (ReLS)
-\item
-  Linear Model Approximation (LMA)
-\item
-  Linear Model Approximation, all time points (LMAT)
-\item
-  Max of data (Max)
-\item
-  Smooth max (spline) (SMax)
-\item
-  Incidence-Prevalence Ratio (IPR)
-\end{enumerate}
-
-\textbf{Least Squares ($\beta$, $\gamma$)}\label{least-squares-beta-gamma}
+\subsubsection{Least Squares ($\beta$, $\gamma$) (SIR LS)}\label{least-squares-beta-gamma}
 
@@
 
-\textbf{Reparametrized Least Squares ($\rr$, $\gamma$)}\label{reparametrized-least-squares-rux5f0-gamma}
+We can use gradient descent or another solver to find estimates for these parameters.
+If we assume $Y= f(t) + \epsilon_t$ with $\epsilon_t \overset{iid}{\sim} N(0, \sigma^2)$ where $f(t) = \int_0^t \frac{dY}{dt} dt$, then this this method is equivalent to finding the MLE of $\beta$ and $\gamma$.
+However, we note that the assumption of homoskedasticity seems likely to be violated in this model.
 
-We Reparametrized the the ODES directly with \(\rr\) and \(\gamma\),
+\subsubsection{Reparametrized Least Squares ($\rr$, $\gamma$) ( SIR ReLS)}\label{reparametrized-least-squares-rux5f0-gamma}
+
+We reparametrized the the ODES in Equation \eqref{eq::sir} directly with \(\rr\) and \(\gamma\), using the relation $\rr = \frac{\beta}{\gamma}$.
 
@@
 
-\textbf{Linear Model Approximation (degree
-10)}\label{linear-model-approximation-degree-10}
+\subsubsection{Linear Model Approximation (SIR LMA)}\label{linear-model-approximation-degree-10}
 
-We fit a linear polynomial of \(t\) with degree \(K= 10\) model to \(X\)
+The idea that an SIR model may be well approximated by a linear model is described by \cite{chang2017}.
+We borrow this idea here and use the ODES in Equation \eqref{eq::sir} to relate our estimated linear model to that of $\rr$.
+
+Specifically, we fit a linear polynomial of \(t\) with degree \(K= 10\) model to \(X\)
 and \(Y\) using least squares to find the coefficients (\(\hat{x}_k\),
@@
 
-\textbf{Linear Model Approximation, All Time Points (degree
-10)}\label{linear-model-approximation-all-time-points-degree-10}
+Like in the previous cases, we are assuming the true models are
+
+\begin{align*}
+X(t) &= \sum_{k=0}^K x_k t^k + \epsilon_t\\
+  {Y}(t) &= \sum_{k=0}^K y_k t^k + \epsilon_t,
+\end{align*}
+with $\epsilon_t \overset{iid}{\sim} N(0, \sigma^2)$.
+Here, $K=10$ is rather arbitrary and should be selected using some criterion such as AIC in addition with the knowledge of how many data points are available to the user.
+
+\subsubsection{Linear Model Approximation, All Time Points (SIR LMAT)}\label{linear-model-approximation-all-time-points-degree-10}
 
@@
 \[\rr = \frac{1}{\# \text{ Obs}}\sum_t \frac{\hat{X}^\prime(t)}{ \hat{X}^\prime(t) + \hat{Y}^\prime(t)} \cdot \frac{N}{X(0)} \]
+The idea is that $\frac{\mathcal{X}^\prime}{X^\prime + Y^\prime}$ is constant in $t$, but due to our approximations with the linear model, this is no longer the case.
+Here, we average over the different possible values of $\rr(t)$.
 
-\textbf{Max of Data}\label{max-of-data}
+\subsubsection{Max of Data (SIR Max)}\label{max-of-data}
 
-We note that \(Y^\prime(t) > 0\) when \(\rr < \frac{S(t)}{N}\). The max
+Under the SIR model there is an absolute maximum of infected individuals $Y$.
+We note that \(Y^\prime(t) = \frac{\beta XY}{N} - \gamma Y > 0\) when \(\rr = \frac{\beta}{\gamma} < \frac{X(t)}{N}\).
+The max
 of \(Y\) occurs when \(Y^\prime = 0\) and hence
@@
 
-\textbf{Smooth Maximum of Data}\label{smooth-maximum-of-data}
+\subsubsection{Smooth Maximum of Data (SIR SMax)}\label{smooth-maximum-of-data}
 
@@
 
-\textbf{Incidence to Prevalence
-Ratio}\label{incidence-to-prevalence-ratio}
+\subsubsection{Incidence to Prevalence
+Ratio (SIR IPR)}\label{incidence-to-prevalence-ratio}
 In terms of data from the SIR model, incidence $J(t) \approx -(X(t+1) - X(t))$, and the IPR$(t) = \frac{J(t)}{Y(t)}$.
@@
 Thus we use as our estimate,
+\begin{align*}
+\rr &= R_a(0).
+\end{align*}
 
-\textbf{Smoothed Incidence to Prevalence Ratio}
+Here we assume that the time step is small enough to approximate the incidence.
+
+\subsubsection{Smoothed Incidence to Prevalence Ratio (SIR SIPR)}
 We use the same method as above, but first fit splines with 4 degrees of freedom, $\hat{X}(t)$ and $\hat{Y}(t)$, , to fit to $X_{obs}$ and $Y_{obs}$, respectively.
@@
 
-\subsubsection{Harko}
-Recently, \citep{harko2014exact} were able to reduce the SIR (XYZ) equation, with the following set of ODEs
+%%%%%%%%%%%%%%%%%%%%%
+\subsubsection{Harko (HSIR)}
+Recently, \citep{harko2014exact} were able to reduce the SIR equation, with the following set of ODEs (reparametrized from Equation \eqref{eq::sir})
 
@@
 If we assume $\epsilon_t \overset{iid}{\sim}N(0, \sigma^2)$, then we can directly estimate $\rr/N$ through linear regression and also an expression for the variance.
+This method greatly simplifies the computational effort needed to estimate $\rr$ as it does not require any numerical integration.
 
-\subsubsection{SEIR Model}
-\label{sec:seir-model}
-
-The SEIR model is a common adaptation of the SIR compartment model \com{add references that use this method}.
-The ``E'' compartment stands for ``exposed'' and represents the stage where individuals are infected but not yet infectious.
-The model is described by the following set of equations, as described by \cite{cintronarias2009}.
-This particular model here does not include birth and death rates but the model can be adapted to do so,
-\begin{align*}
-  \frac{dS}{dt} &= - \beta SI \\
-  \frac{dE}{dt} &= \beta SI  - \alpha E\\
-  \frac{dI}{dt} &= \alpha E - \gamma I \\
-  \frac{dR}{dt} &= \gamma I.
-\end{align*}
-
-Then $\rr = \beta / \gamma$, just as in the SIR model.
-As in the SIR model, $\beta$ and $\gamma$ have the same interpretation and $\alpha$  The methods used to estimate the parameters $\beta$ and $\gamma$ are of the same class used to estimate those parameters in the SIR model.
-The important difference between this method and the SIR model is the assumption of a 4th class, the exposed compartment.
-Typically, this method will result in ...\com{figure out}...
-than the SIR method.
-Homogeneous mixing of the compartments is assumed.
-
-\subsubsection{Attack Rate}
+\subsubsection{Attack Rate (SIR AR)}
 
@@
 
-\subsubsection{Amplifier Model}
-\cite{blower2004} discuss the case where a disease may have different strains and consequently different reproduction numbers.
-They introduce a multistrain mathematical model called the amplifier model to deal with this case.
-
-This model is based on the the SLIR compartment model, where $L$ stands for latently infected individuals and $T$ for diseased individuals.
-There is a set of ODEs for each strain of the disease with different parameters for recovery, transmission rate, and drug resistance of the disease.
-Additionally, there is mixing of the different strains as ``the amplifier model also allows for immigrating and emigration of individuals of all types, as well as reinfection of latently infected individuals'' \citep{blower2004}.
-The details may be found in their manuscript.
-Ultimately, their derivation for $\rr^{(i)}$ where $i$ stands for the $i$th strain is
-\begin{align*}
-\rr^{(i)} = S^* \frac{ ( \beta_i^T + \beta_i^L)\nu_i + \beta_i^T \mu_i^L}{(\nu_i + \mu_i^L)(c_i + k_{i,i+1} + \mu_i^T)}.
-\end{align*}
-Here, $S^*$ is the number of susceptibles individuals in the disease-free equilibrium in the absence of immigration or emigration; $\beta_i^L$ and $\beta_i^T$ are strain-specific transmission rates for the latent, and infected classes, respectively; $\nu_i$ denotes the progression rate from latently infected to disease for strain $i$; $\mu_i^L$ and $\mu_i^{T}$ are death rates for the latent and infected classes, respectively; $c_i$ is the cure rate; and $k_{i, i+1}$ is a parameter associated with the degree of amplification of drug resistance.
-They note that ``each reproduction number is the product of average number of secondary infections caused per unit time, the aver time a case remains infectious and the probability that an infected individual develops disease.''
-
-In the amplifier model, $\rr$ was derived from using the stability point (disease free equilibrium state in the absence of immigration or emigration) of the system by solving the characteristic polynomial for the eigenvalues
-
-\subsubsection{Sequential Bayes}\label{sec:seqbayes}
+\subsubsection{Sequential Bayes (SIR SB)}\label{sec:seqbayes}
 
 Posited by \cite{bettencourt2008} and summarized in \cite{obadia2012r0}, this method uses a Bayesian approach to an approximation to the classic SIR model.
-The approximated SIR model is that the incidence at $t+1$, $N(t+1)$ is, with $\gamma$ as the  average inverse of the infectious period and $R$ as the effective reproduction number, which we take here to be $\rr$.
+The approximated SIR model is that the incidence at $t+1$, $N(t+1)$ is from a Poisson distribution, with $\gamma$ as the  average inverse of the infectious period and $R$ as the effective reproduction number, which we take here to be $\rr$.
 
-In order to estimate $\rr$, we must have some idea about the serial interval.
+In order to estimate $\rr$, we must have some idea about the serial interval, $\omega$.
 \begin{align*}
@@
 
-\subsubsection{Next Generation Model}
+\subsubsection{SEIR Model (SEIR)}
+\label{sec:seir-model}
+
+The SEIR model is a common adaptation of the SIR compartment model \com{add references that use this method}.
+The ``E'' compartment stands for ``exposed'' and represents the stage where individuals are infected but not yet infectious.
+The model is described by the following set of equations, as described by \cite{cintronarias2009}.
+This particular model here does not include birth and death rates but the model can be adapted to do so,
+\begin{align*}
+  \frac{dX}{dt} &= - \frac{\beta XY}{N} \\
+  \frac{dE}{dt} &= \frac{\beta XY}{N}  - \alpha E\\
+  \frac{dY}{dt} &= \alpha E - \gamma Y \\
+  \frac{dZ}{dt} &= \gamma Y.
+\end{align*}
+
+Then $\rr = \beta / \gamma$, just as in the SIR model.
+As in the SIR model, $\beta$ and $\gamma$ have the same interpretation and $\alpha$  The methods used to estimate the parameters $\beta$ and $\gamma$ are of the same class used to estimate those parameters in the SIR model.
+The important difference between this method and the SIR model is the assumption of a 4th class, the exposed compartment.
+Even though the derivation for $\rr$ is the same for both the SIR and SEIR model, this will still result in different results for estimations of \rr, even if the same method for estimation of parameters is used.
+Typically, some form of least squares is used to estimate parameters in this model.
+We demonstrate this in Section \com{results}.
+
+\subsubsection{Amplifier Model (Amp)}
+\cite{blower2004} discuss the case where a disease may have different strains and consequently different reproduction numbers.
+They introduce a multistrain mathematical model called the amplifier model to deal with this case.
+
+This model is based on the the SLIR compartment model, where $L$ stands for latently infected individuals and $T$ for diseased individuals.
+There is a set of ODEs for each strain of the disease with different parameters for recovery, transmission rate, and drug resistance of the disease.
+Additionally, there is mixing of the different strains as ``the amplifier model also allows for immigrating and emigration of individuals of all types, as well as reinfection of latently infected individuals'' \citep{blower2004}.
+The details may be found in their manuscript.
+Ultimately, their derivation for $\rr^{(i)}$ where $i$ stands for the $i$th strain is
+\begin{align*}
+\rr^{(i)} = X^* \frac{ ( \beta_i^T + \beta_i^L)\nu_i + \beta_i^T \mu_i^L}{(\nu_i + \mu_i^L)(c_i + k_{i,i+1} + \mu_i^T)}.
+\end{align*}
+Here, $X^*$ is the number of susceptibles individuals in the disease-free equilibrium in the absence of immigration or emigration; $\beta_i^L$ and $\beta_i^T$ are strain-specific transmission rates for the latent, and infected classes, respectively; $\nu_i$ denotes the progression rate from latently infected to disease for strain $i$; $\mu_i^L$ and $\mu_i^{T}$ are death rates for the latent and infected classes, respectively; $c_i$ is the cure rate; and $k_{i, i+1}$ is a parameter associated with the degree of amplification of drug resistance.
+They note that ``each reproduction number is the product of average number of secondary infections caused per unit time, the aver time a case remains infectious and the probability that an infected individual develops disease.''
+
+In the amplifier model, $\rr$ was derived from using the stability point (disease free equilibrium state in the absence of immigration or emigration) of the system by solving the characteristic polynomial for the eigenvalues.
+This derivation is explained in more depth in Section \ref{sec:ngm}.
+
+We show this model as opposed to other available CMs because it shows not only how complicated CMs can become, but researchers approach the assumption of homogeneity, which is often suspect.
+In this case, we have strain-specific transmission rates, so we are essentially stratifying compartments into different sets.
+Very complicated models have arisen from stratification of compartments, which leads to the approach of ABMs.
+
+\subsubsection{Next Generation Model (NGM)}
 \label{sec:ngm}
@@
 This method has the assumptions of a basic compartment model: homogeneous mixing among populations and the law of mass action, i.e., that the number of infected individuals are proportional to the number of infected individuals at the previous step.
-The estimations for $\rr$ from the previously mentioned compartment models \com{list those models} may also be derived using this method.
+The estimations for $\rr$ from the previously mentioned compartment models may also be derived using this method.
 Additionally, this method is advantageous in that the matrix $\Sigma$ has an intuitive meaning.
+
+Although, this method is useful for deriving an expression of $\rr$, it is unclear how to actually estimate it from these expressions.
+Again, least squares is typically used to estimate such parameters.
 
@@
 \subsection{Exponential Growth}\label{sec:exp-growth}
+Typically, CMs expressed through differential equations have no known analytical solutions, which makes deriving an expression for $\rr$ difficult.
+One simplification that is often introduced to the SI framework is that of exponential growth, or that the number of infections in the beginning of an epidemic mimics that of an exponential curve.
+This simplified expression has allowed for many derivations of $\rr$.
 
-\subsubsection{Exponential Growth}
+\subsubsection{Exponential Growth (EG)}
 \label{sec:expgrowth}
-\cite{wallinga2007generation} report that $R$ and hence $\rr$ may calculated by using the fact that infection ``counts increase exponentially in the initial phase of an epidemic.''  We need to know $r$, the \textit{per capita} change in the number of new cases per unit of time and $T_c$ the serial interval.
+\cite{wallinga2007generation} report that $\mathcal{R}_t$ and hence $\rr$ may calculated by using the fact that infection ``counts increase exponentially in the initial phase of an epidemic.''  We need to know $r$, the \textit{per capita} change in the number of new cases per unit of time and $\omega$ the serial interval.
 
@@
 \begin{align}\label{eq:lotka}
-\rr = \exp{(r T_c)}
+\rr = \exp{(r \omega)}
 \end{align}
@@
 \begin{align}\label{eq:anderson}
-\rr = 1 + rT_c.
+\rr = 1 + r \omega.
 \end{align}
 
-Equation \eqref{eq:lotka} is derived from a demographic view using the Lotka Euler survival equations, whereas Equation \eqref{eq:anderson}is derived through an epidemiologist view using \com{something}.
-These estimates for $\rr$ can be strikingly different.
+Equation \eqref{eq:lotka} is derived from a demographic view using the Lotka-Euler survival equations which come from the fields of demography, ecology, and evolutionary biology, whereas Equation \eqref{eq:anderson}is derived through an epidemiologist.
+These estimates for $\rr$ between Equation \eqref{eq:lotka} and Equation \eqref{eq:anderson} can be strikingly different.
 
@@
 
-For all these above estimations of $\rr$, we observe the duration of generation intervals in a period of exponential growth.
+For all these above estimations of $\rr$, we observe the duration of serial intervals in a period of exponential growth.
 Deciding when this occurs and which data to use to estimate $\rr$ is difficult.
+\citeauthor{wallinga2007generation} specify $\omega$ in multiple ways.
+Parametrically, they use exponential, normal, and delta distributions.
+They also observe the empirical distribution of the serial interval.
 
-\subsubsection{Maximum Likelihood Estimator of Secondary Infections}\label{sec:mle-si}
+\citeauthor{wallinga2007generation} note that specifying the epidemic model ``implicitly specifies a generation interval distribution.''  The above equations are under the SIR framework, but this method may be adapted to acommodate frameworks such as the SEIR model and more.
+
+\subsubsection{Maximum Likelihood Estimator of Secondary Infections (MLSI)}\label{sec:mle-si}
 This method, described by \cite{forsberg2008}, finds the estimate of $\rr$ by maximizing the likelihood function under certain assumptions.
 We assume that the ``the number of secondary cases produced by an infected individual follows a Poisson distribution, and that the serial interval is described by a multinomial distribution.''  Recall, the serial interval is the distribution of time between a primary case developing symptoms and a case infected by the primary case developing symptoms.
-An approximate version of this can be simplified to a thinned Poisson where
-\begin{align*}
+An approximate version of this can be simplified to a thinned Poisson in Equation \ref{eq:mlesi} where
+\begin{align}\label{eq:mlesi}
   L(\rr, \mathbf{p}) = \prod_{t=1}^T \frac{e^{- \rr \sum_{j=1}^{\min(k,t)}N_{t-j}p_j}\left (\rr \sum_{j=1}^{\min(k,t)}N_{t-j}p_j \right )^{N_t}}{N_t!}.
-\end{align*}
+\end{align}
 Here, $k$ is the maximal amount of the serial number, $T$ is the total time and $p_j$ is the probability of displaying symptoms on day $j$ after being infected.
 The number of cases on day $t$, $N_t$ are incidence counts that are known.
-\com{more on this}
 
@@
 
-\subsubsection{Time Dependent Reproduction Number}\label{sec:timedep}
-This is another method used to estimate the effective reproduction number, which again, we take to be $\rr$ here.
-This is a likelihood based approach where we maximize the likelihood that case $i$ has been infected by case $j$ at certain time using the generation interval, $w(t)$ where $t_i$ is when case $i$ was infected and $N$ is the total number of cases,
+\subsubsection{Time Dependent Reproduction Number (TDR)}\label{sec:timedep}
+This is another method used to estimate the effective reproduction number, which again, we take to be $\rr$ here \citep{forsberg2008}.
+This is a likelihood based approach where we maximize the likelihood that case $i$ has been infected by case $j$ at certain time using the serial interval, $w(t)$ where $t_i$ is when case $i$ was infected and $N$ is the total number of cases,
 \begin{align*}
@@
 
-\subsubsection{Initial Growth Rate and Final Size}
+\subsubsection{Initial Growth Rate and Final Size (IGRFS)}
 \label{sec:igr-fs}
-\com{this should be moved up in the order}
 
@@
   \subsection{Estimation from Networks}
+As estimating $\rr$ is concerned with the number of new infections based on the contacts of old infections, a network is a seemingly intuitive way to model such contacts and hence infections.
+Networks that have been used to estimate $\rr$ include contact tracing of actual individuals, branching processes to model the spread of disease, and agent-based models (also known as individual level models), which all leverage networks along with different sources of data to generate estimates of $\rr$.
+
 \label{sec:network}
 
-\subsubsection{Branching Process}
+%%%%%%%% 5
+\subsubsection{Contact Tracing (CT)}
+\label{sec:contact_tracing}
+A reasonable idea to estimate $\rr$ is to directly observe the number of secondary infections produced by an initial infector.
+However, this approach is generally impractical as it is difficult to estimate every single contact a person may have.
+With seriological information, we could conceivably recreate the transmission of the disease throughout a population.
+Even in this case, it is unlikely, for one, to have a completely susceptible population, which is essential to an unbiased estimate of $\rr$.
+Moreover, this would implicitly be conditioning on the initial infector's contacts which may not reflect those of the rest of the population.
+We write the calculation for $\rr$ under contact tracing in Equation \ref{eq:r0_contacttracing}, where $X_0$ is a random variable denoting the number of individuals in the population person 0 infects.
+Here $S_0=N$ indicates that the initial number of susceptibles is $N$, the size of the population.
+
+\begin{align}\label{eq:r0_contacttracing}
+\rr = E[ X_0 | S_0 = N]
+\end{align}
+
+\cite{eames2003} describe contact tracing as an ``extreme form of targeted control, where the potential next-generation cases are the primary focus,'' that is, one focuses on treating the contacts of an infected individual rather than applying uniform treatment across a population.
+They note that this method has ``proved to be a highly successful strategy when the number of infectious cases is low'' and ``is the standard tool for eliminating minor outbreaks in the latter stages of disease eradication.''
+
+They describe an estimate for $\rr$ for a SIR model on a network, with $r$ as the rate of transmission across a contact multiplied by the infectious period and $n$ as the number of contacts as
+
+\begin{align*}
+\rr = r(n-2)
+\end{align*}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\subsubsection{Branching Process (BP)}
 \label{sec:branching-process}
 
-Branching processes are models for population growth and track generations of new offspring.
+Branching processes are models for population growth which track generations of new offspring.
 We describe branching processes as adapted for epidemic modelling from the notation of \cite{grimmett1991}.
-An infector $i$ may produce new infections over the lifetime of the original infector's disease.
-These new infections are called offspring.
-The offspring the orginal infector directly produces are called the first generation.
+An infector $i$ may produce new infections, called offspring, over the lifetime of the  infector $i$'s disease.
+The offspring of the original infector directly produces are called the first generation.
 The offspring of the first generation are called the second generation and so on.
-An important assumption in branching processes is that the offspring produced from different infectors are indepedent and identically distributed (iid), and so for a branching process to be a valid model in epidemic modeling, we must assume a large population with few infectors.
+An important assumption in branching processes is that the offspring produced from different infectors are indepedent and identically distributed, and so for a branching process to be a valid model in epidemic modeling, we must assume a large population with few infectors.
 
@@
 
-\cite{getz2006} give a way to calculate \rr from a branching process.
+\cite{getz2006} give a way to calculate $\rr$ from a branching process.
 ```We define the offspring distribution $\{q_i \}_{i=0}^\infty$, where $q_i$  is the probability that an infectious individual infects $i$ other individuals.
-Thus we require $\sum_{i=0}^\infty q_i =1$ and note that $\rr$, the mean number of cases contracting disease from each infective, is simply given by
+Thus we require $\sum_{i=0}^\infty q_i =1$ and note that $\rr$, the mean number of cases contracting disease from each infective'', is simply given by'
   \begin{align*}
-    \rr = \sum_{i=0}^\infty iq_i."
+    \rr = \sum_{i=0}^\infty iq_i.
   \end{align*}
+  It is assumed we know how to estimate the $q_i$, and so in this case, it is not too different of an approach than contact tracing.
 
-\subsubsection{Agent-Based Models}
+\subsubsection{Agent-Based Models (ABM)}
 \label{sec:agent-based-models}
@@
 In epidemiology, ABMs can be used to simulate the spread of disease.
-ABMs allow for every detail of an infection to be known, in the context of $\rr$, meaning we know exactly who is infected by whom and when.
+ABMs allow for every detail of an infection to be known, in the context of estimating $\rr$, meaning we know exactly who is infected by whom and when.
 Thus $\rr$ can be calculated simply through
 \begin{align*}
-  \rr = \frac{1}{\# \text{ simulations}}\frac{1}{\#\text{ initial infections}} \sum_{s \in \text{ simulations}}\sum_{i \text{ infected at } t=0} \# \text{ infected by $i$ in simulation }s.
+  \rr = \frac{1}{\# \text{ simulations}}\frac{1}{\#\text{ initial infections}} \sum_{(s \in \text{ simulations})}\sum_{(i \text{ infected at } t=0)} \# \text{ infected by $i$ in simulation }s.
 \end{align*}
@@
 Additionally, we need to have accurate agents in number, characteristics, and their implied network structure.
+ABMs are flexible in that they may be used with any model in the SI framework.
+The only requirement is that we track who infects whom and can adapt for any additional states such as exposure.
 
@@
 
-\subsubsection{Inverse Hessian}\label{inverse-hessian}
+Estimating $\rr$ is difficult, and it is no different for estimating $V[\rr]$, the variance.
+We describe four general methods which may be applicable to estimate the variance.
+We describe the assumptions that are used along with each method and note that they may not be applicable to all of our methods.
+Still, reporting confidence intervals, or at the very least, standard deviations is very important and we hope to encourage this practice by describing the below methods used to estimate models in the SIR framework in a simple fashion.
 
-When we calculate the parameters of the SIR model using least squares,
-we use gradient descent using the Hessian along the way. We use the
-relationship with the Fisher Information Matrix to approximate the
-covariance matrix of the parameters.
+\subsubsection{Delta Method}\label{delta-method}
+
+When the method estimates \(\beta\) and \(\gamma\) instead of \(\rr\) directly, we use the delta method approximation to calculate the
+variance of \(\rr\).
+Here, we know \(\rr = h(\beta, \gamma) = \frac{\beta}{\gamma}\).
+Then \(\bigtriangleup h = (\frac{1}{\gamma},  -\frac{\beta}{\gamma^2})^T\).
+Then \(V[\rr] = \bigtriangleup h^T \Sigma_{\beta, \gamma} \bigtriangleup h\), where \(\Sigma_{\beta, \gamma}\) is covariance matrix of \(\beta\) and \(\gamma\).
+
+This estimate assumes that the distribution of $\beta$ and $\gamma$ are asymptotically normal.
+Here, we use the relationship of $\rr = \frac{\beta}{\gamma}$ in order to estimate the variance, which is specific to the SIR framework.
+This method may be extended to other frameworks and we provide some examples here \com{add examples of SEIR, serial interval, etc.}
 
@@
 We estimate \(\frac{\partial Y}{\partial \beta}\) (or
-\(\frac{\partial Y}{\partial \rr}\) when using reparametrized least
-squares) and \(\frac{\partial Y}{\partial \gamma}\) at each time point
-\(t\) using the estimated \(\beta\) or \(\gamma\) from the estimation
-process. Then an approximation of the covariance matrix is
-\((\chi^T \chi)^{-1}\) where
-\(\chi_{ij} = \frac{ \partial Y(t_i)}{ \partial \theta_j}\) where
-\(\theta\) is a vector of parameters. (add citation)
+\(\frac{\partial Y}{\partial \rr}\) when using reparametrized least squares) and \(\frac{\partial Y}{\partial \gamma}\) at each time point \(t\) using the estimated \(\beta\) or \(\gamma\) from the estimation process where $Y$ stands for the number of infected at time $t$.
+Then an approximation of the covariance matrix is \((\chi^T \chi)^{-1}\) where \(\chi_{ij} = \frac{ \partial Y(t_i)}{ \partial \theta_j}\) where \(\theta\) is a vector of parameters.
+\com{add citation}
 
-\subsubsection{Jackknife}\label{jackknife}
-
-We estimate \(\rr^{(-i)}\) using a specified method leaving out the
-\(i\)th time point out from the data each iteration. Assume there are
-\(n\) time points of observations. Then
-\[V[\rr] = \frac{n-1}{n} \sum_{i=1}^n \left (\rr^{(-i)} - \frac{1}{n} \sum_{i=1}^n\rr^{(-i)} \right)^2\]
-
-\subsubsection{Delta Method}\label{delta-method}
-
-When the method estimates \(\beta\) and \(\gamma\) instead of \(\rr\)
-directly, we use the delta method approximation to calculate the
-variance of \(\rr\). Here, we know
-\(\rr = h(\beta, \gamma) = \frac{\beta}{\gamma}\). Then
-\(\bigtriangleup h = (\frac{1}{\gamma}, - \frac{\beta}{\gamma^2})^T\).
-Then
-\(V[\rr] = \bigtriangleup h^T \Sigma_{\beta, \gamma} \bigtriangleup h\),
-where \(\Sigma_{\beta, \gamma}\) is covariance matrix of \(\beta\) and
-\(\gamma\).
 %%%%%%%%%%%%%%%

### 2018-07-03 paper/old_drafts/draft_v3.tex
finally on git

--- 2018-07-03 paper/old_drafts/draft_v2.tex
@@
 
-\newcommand{\XX}{21 } % number of methods
+\newcommand{\XX}{\ensuremath{23}} % number of methods
 \newcommand{\rr}{\ensuremath{\mathcal{R}_0}}
@@
 
-\title{\XX dubious ways of estimating $\rr$}
+\title{$\XX$ dubious ways of estimating $\rr$}
 \author{ Department of Statistics, Carnegie Mellon University}
@@
 
-\tableofcontents
+%\tableofcontents
 
 \section{Introduction}\label{sec:intro}
-What has been called ``arguably the most important quantity in the study of epidemics'' \cite{Heesterbeek2002}, $\mathcal{R}_0$, the reproduction number, has been found to be an elusive quantity for epidemic modelers.
+What has been called ``arguably the most important quantity in the study of epidemics'', $\mathcal{R}_0$, the reproduction number, is an elusive quantity for epidemic modelers \cite{Heesterbeek2002}.
 \citet{anderson1992} define $\rr$ as the ``the average number of secondary infections produced when one infected individual is introduced into a host population where everyone is susceptible.''  $\rr$, in some ways, summarizes an entire epidemic.
+This quantity is used whether to assess if a disease will break out and what percentage of the population needs to be vaccinated to avoid such an outbreak.
 Despite the clear definition of $\rr$, epidemiologists have struggled to create a \textit{de facto} estimator for $\rr$  \citep{hethcote2000}.
-The primary issue in estimating $\rr$ is that the quantity is a \textit{property of the model}, meaning that $\rr$ is dependent not only on the usual noise that comes with statistical modeling but also on a variety of assumptions on how we assume disease is transmitted through a population.
+The primary issue in estimating $\rr$ is that the quantity is a \textit{property of the model}, meaning that $\rr$ is dependent not only on the usual noise that comes with statistical modeling but also on a variety of assumptions on how researchers assume disease is transmitted through a population.
 
@@
 Other issues of estimating $\rr$ include mathematical versus statistical approaches (i.e.
-solving for $\rr$ versus estimating $\rr$), how to estimate the quantity in the presence of different available data, and when to estimate $\rr$.
-For instance, if only weekly incidence counts of a disease are available, then we have different methods of estimating $\rr$ than when we have more data available such as contact networks, the serial interval of a disease, and more.
+solving for $\rr$ versus estimating $\rr$), how to estimate the quantity in the presence of different available data, and when to estimate $\rr$ in a disease's lifetime.
+For instance, if only weekly incidence counts of a disease are available, then we have different methods of estimating $\rr$ available than when we have more data available such as contact networks or the serial interval of a disease.
 Another major issue is that since $\rr$ is so difficult to estimate in the first place, the variance of the quantity has largely been glossed over or completely ignored.
-This is especially troubling as we are concerned whether $\rr > 1$, which implies an outbreak of a disease.
-A point estimate is simply not enough.
+This is especially troubling as we are concerned whether $\rr > 1$, which indicates an outbreak of a disease.
+A point estimate is simply not enough to determine the truth of this inequality.
 
@@
 This, in turn, makes it difficult to compare $\rr$ across different diseases or even instances of the same disease.
-Many of these issues with estimating $\rr$ are summarized by \cite{li2011}
+Many of these issues with estimating $\rr$ are summarized and expanded upon by \cite{li2011}.
 
-We review \XX ways to estimate $\rr$ over varying assumptions and in the presence of different sources of data.
+We review $\XX$ ways to estimate $\rr$ over varying assumptions and in the presence of different sources of data.
 The title of this manuscript is in reference to a seminal paper by Moler and Van Loan in 1978 (and its subsequent update in 2003) in which they calculate ``Nineteen dubious ways to compute the exponential of a matrix'' \citep{moler2003}.
@@
 The rest of this manuscript is organized as follows.
-In Section \ref{sec:overview}, we introduce the methods in 3 broad groups:.
-In Section \ref{sec:term} we introduce terminology used throughout.
+In Section \ref{sec:overview}, we introduce the methods in four broad groups: the survival function, compartment models, exponential growth models, and agent-based models.
 In Section \ref{sec:details} we give an overview of each of the methods, what assumptions are used, what data is required, and where the method has been used in the past along with ways in which to calculate the variance for methods, when applicable.
@@
 Finally, in Section \ref{sec:dis} we discuss our conclusions and recommendations.
+In Appendix \ref{sec:term} we include a glossary of terminology used throughout our manuscript.
 
-\section{Overview of \XX Methods to estimate $\rr$}
+\section{Overview of $\XX$ Methods to estimate $\rr$}
 \label{sec:overview}
+
+We chose $\XX$ methods to review in this paper, which does not cover all possible methods to estimate $\rr$.
+These methods were chosen as  they represent the type of models typically used to model $\rr$ in addition to their high impact in the field of epidemiology.
 
@@
 These partitions are meant to be guidelines of distinguishing methods from one another rather than immutable groups.
-Many of these methods could have fit in another category or even all four of the classes chosen.
+Many of these methods could have fit in another category or even multiple of the groups chosen.
 We have chosen the four classes as application of the survival function and its close relatives, classic compartment models, exponential growth assumptions, and estimation from networks.
+We list these point estimate methods below.
 
@@
   \item Survival function (SF)
-  \item Direct estimation of the survival function (DESF)
   \item Direct parameter estimation (DPE)
+  \item Von Foerster equations (VFE)
 
@@
     \begin{itemize}
-    \item Least Squares (SIR LS)
-    \item Reparametrized Least Squares (SIR ReLS)
-    \item Linear Model Approximation (SIR LMA)
-    \item Linear Model Approximation, all time points (SIR LMAT)
-    \item Max of Data (SIR Max)
+    \item Least squares (SIR LS)
+    \item Reparametrized least squares (SIR ReLS)
+    \item Linear model approximation (SIR LMA)
+    \item Linear model approximation, all time points (SIR LMAT)
+    \item Max of data (SIR Max)
     \item Smooth max (SIR SMax)
-    \item Incidence-Prevalence Ratio (SIR IPR)
-    \item Smooth Incidence-Prevalence Ratio (SIR SIPR)
+    \item Incidence-prevalence ratio (SIR IPR)
+    \item Smooth incidence-prevalence Ratio (SIR SIPR)
     \item Harko's SIR model (HSIR)
-    \item Attack rate (SIR AR)
     \item Sequential Bayes (SIR SB)
@@
   \item SEIR model (SEIR)
-  \item Amplifier model (Amp)
+  \item Amplifier model (AMP)
   \item Next generation model (NGM)
@@
 
-  Additionally, we describe a number of variance estimates for the above methods for estimating \rr.  These are
+  Additionally, we describe  interval estimates for the above methods for estimating \rr.
 
-\textbf{Variance Estimation}
+\textbf{Interval Estimation}
 \begin{enumerate}
@@
 \item Sensitivity analysis
+  \item Posterior distribution
   \end{enumerate}
 
-\section{Epidemic Modeling Terminology}
-\label{sec:term}
-
-Here we introduce some of the terminology and notation used throughout the methods for estimating \rr.
-
-{$\rr$} -- reproduction number --
-
-{$\mathcal{R}_t$} -- effective reproduction number
-
-{$\beta$} -- the infection rate
-
-{$\gamma$} -- inverse of the expected recovery time
-
-{$N_t$} -- Number of individuals in the population at time $t$
-
-{$X(t)$} -- the number of susceptible individuals in a population at time $t$
-
-{$Y(t)$} -- the number of infected individuals in a population at time $t$
-
-{$Z(t)$} -- the number of recovered individuals in a population at time $t$
-
-{$\omega$} -- serial or generation interval -- the time from onset of symptoms in an index case to the onset of symptoms in a subsequent case infected by the index patient
-
-AR -- attack rate -- percentage of the population eventually infected
-
-\section{Details for the \XX Methods}
+\section{Details for the $\XX$ methods}
 \label{sec:details}
@@
 
-The survival function is intimately with the theory of \rr, which originates from the field of demography in the late 1800s \citep{dietz1993estimation}.  Here, we have
+The survival function is intimately tied to the history of \rr, which originates from the field of demography in the late 1800s \citep{dietz1993estimation}.  In demography, we have
 \begin{align*}
@@
 This interpretation also explains the name of \rr, which was imported to the field of epidemiology by MacDonald and Smith as reported by \cite{dietz1993estimation}.
+In epidemiology, the analogy is that $p(a)$ is like the duration of a disease, and $\beta$ is the transmission rate.
 
-This is the chronologically first and in many ways, the most direct method to calculate \rr.
-However, this model also seems to be the most difficult to estimate.
-We describe four methods attempting to estimate \rr from the survival function:  the survival function alone, direct estimation, and direct parameter estimation.
+This is the chronologically first and in many ways, the most direct method to calculate $\rr$.
+However, this model also seems to be the most difficult to estimate as both $p(a)$ and $\beta (a)$ as $p(a)$ is a distribution and $\beta$ is a function.
+Typically, it is also difficult to estimate these parameters before or even during an ongoing disease outbreak.
+In turn, $\rr$ can only be estimated after an outbreak, which means it cannot be used to determine ongoing prevention strategies.
+However, even having a post-disease estimation of $\rr$ may be useful for future outbreaks of the disease or being able to compare the severity of many diseases.
+Another advantage is that this method is very intuitive and is independent of assumptions of how the disease is transmitted from person to person.
+We describe four methods attempting to estimate $\rr$ from the survival function:  the survival function alone, direct estimation, and direct parameter estimation.
 
-\subsubsection{Survival Function}
+\subsubsection{Survival Function (SF)}
 \label{sec:survival_fxn}
-The first method, is not surprisingly, the survival function itself as presented by \cite{Heffernan2005}.
+The first method of estimating $\rr$, is not surprisingly, the survival function itself as presented by \cite{Heffernan2005}.
 We change the notation from the above survival function to conform to the standards of epidemiology.
 
-Let $F(a)$ be the probability that a newly infected individual remains infectious for at least time $a$ (the survival probability,$p(a)$ in the previous definition).
+Let $F(a)$ be the probability that a newly infected individual remains infectious for at least time $a$ (the survival probability, $p(a)$ in the previous definition).
 Moreover, let $b(a)$ $(\beta (a)$ in the previous definition), the average number of newly infected individuals that an infectious individual will produce per unit time when infected for total time $a$.
@@
 
-This method presents one of the unique difficulties of estimating \rr: that of reconciling generations and time.
-On the other hand, this method is flexible in that this derivation of $\rr$ is not restricted to compartment models.
+This method presents one of the unique difficulties of estimating $\rr$: that of reconciling generations and time.
+$\rr$ tells us about the generations of disease, meaning a parent and their offspring but not necessarily when they occur.
+This is made difficult as the data we have about a disease is often of a temporal nature.
+On the other hand, this method is flexible in that this derivation of $\rr$ is does not place assumptions on how the disease transmits through a population.
+A disadvantage is that data about $F(a)$ and especially $b(a)$ is difficult to collect directly.
 
-Estimating this quantity in Equation \ref{eq:r0_survivalfxn} is hard and rarely attempted as $b(a)$ is particularly difficult to estimate, and it is due to this difficulty that the rest of these methods for estimating $\rr$ even exist.
+Estimating this quantity in Equation \ref{eq:r0_survivalfxn} is hard and, in fact, rarely attempted. It is due to this difficulty that many of rest of the methods for estimating $\rr$ even exist.
+
+\subsubsection{Direct Parameter Estimation (DPE)}
+\label{sec:dpe}
+
+Direct parameter estimation is a `plug and chug' approach as described by \cite{lipsitch2003} and \cite{dietz1993estimation}, which has an intuitive derivation, similar to that of the survival function.
+The intuition is that we can estimate $\rr$ if we know how many others an individual comes into contact, the probability of transmission given contact, and how long the disease lasts within an individual  In this method, we estimate the parameters $k$, the number of contacts per unit time, $b$, the probability of transmission per contact, and $D$ the mean duration of infectiousness.
+Then $\rr$ is the product of these parameters, (Equation \ref{eq:dpe}),
+\begin{align}\label{eq:dpe}
+\rr = kbD.
+\end{align}
+
+\cite{lipsitch2003} note that directly estimating $\rr$ in such a way may be difficult due to the available data, as in the previous two methods.
+They instead estimate $\rr$ using the mean serial interval, $\omega$, the time from onset of symptoms in an index case to the onset of symptoms in a subsequent case infected by the index patient in addition to the assumption of exponential growth of infected cases.
+
+In the event, that the is estimate may be calculated, uncertainty estimates must be compounded, and to estimate $\rr$ alone, we only use data from the initial outbreak of the disease, so often we have a small sample size, resulting in large confidence sets or credible intervals.
 
@@
 
-\subsubsection{Direct Estimation of the Survival Function }
+\subsubsection{Von Foerster Equations (VFE)}
 \label{sec:direct-estim-surv}
 
-\cite{fraser2004factors} have a derivation of $\rr$ related to the survival function based on the Van Forester equations,
+A derivation of $\rr$ similar to the survival function comes from the idea of the von Foerster equations, which were used to calculate the number of cells in a ``deme,'' a constrained area.
+The function $n(t, a)$ calculated the number of cells at age $a$ at time $t$ \citep{trucco1965}.
+In epidemiology, $n$ represents the number of infections instead of cells.
+Letting $Y(t, \tau)$ be the number of infections at time $t$ of age $\tau$ and assuming no deaths in the population, the von Foerster equation is
 \begin{align*}
@@
 \end{align*}
-where $Y(t,\tau)$ is the number of people at time $t$ who were infected time $\tau$ ago.  $\rr$  is then,
+where $Y(t,\tau)$ is the number of people at time $t$ who were infected time $\tau$ ago.    \cite{fraser2004factors} derived an expression for $\rr$ from the von Foerster equation and the equation
 \begin{align*}
-\rr = \int_0^\infty \beta(\tau) d\tau,
+  Y(t,0) &= \int_0^t \beta( \tau) Y(t, \tau) d \tau
 \end{align*}
- where $\beta (\tau)$ represents the infectiousness at time $\tau$ since the infection.  Again, this model is more mathematical in nature than practical as $\beta (\tau)$ is difficult to estimate.
+and initial conditions $Y(0,0) = Y_i$ and $Y(t, \tau) = 0$ when $t < \tau$.  Then the reproduction number is Equation \ref{eq:vfr0},
+\begin{align}\label{eq:vfr0}
+\rr &= \int_0^\infty \beta(\tau) d\tau,
+\end{align}
+ where $\beta (\tau)$ represents the infectiousness at time $\tau$ since the infection.
+Again, this model is more mathematical in nature than practical as $\beta (\tau)$ is difficult to estimate.
+To estimate $\rr$, \cite{fraser2004factors} make an approximation that $Y(t, \tau) = K(\tau) \exp (rt)$ and further adapt the method to include prevention strategies, but the main parameter to be estimated is $\beta (\tau)$.
+Estimating $\beta (\tau)$ has many of the same issues as trying to estimate $b(a)$ in the survival function in that it is difficult to estimate while the disease is ongoing but is also advantageous in that it does not place assumptions on how the disease is transmitted throughout the population.
 
 %%%%%%%%%%%%%%%
-
-\subsubsection{Direct Parameter Estimation }
-\label{sec:dpe}
-
-This `plug and chug` approach as described by \cite{lipsitch2003} and \cite{dietz1993estimation} is the idea of estimating parameters $k$, the number of contacts per unit time, $b$, the probability of transmission per contact, and $D$ the mean duration of infectiousness to estimate $\rr$ directly from available data,
-\begin{align}
-\rr = kbD.
-\end{align}
-
-\cite{lipsitch2003} note that directly estimating $\rr$ in such a way may be difficult due to the available data.
-They instead estimate $\rr$ using the mean serial interval, $\omega$, the time from onset of symptoms in an index case to the onset of symptoms in a subsequent case infected by the index patient in addition to the assumption of exponential growth of infected cases.
-
-In the event, that the is estimate may be calculated, uncertainty estimates must be compounded, and to estimate $\rr$ alone, we only use data from the initial outbreak of the disease, so often we have a small sample size, resulting in large confidence sets.
 
@@
 
-A large and the most common set of models used estimate $\mathcal{R}$ is that of compartment models (CMs).
+A large and likely the most common set of models used estimate $\mathcal{R}$ is that of compartment models (CMs).
 As adapted from \cite{daley2001epidemic}, compartment models consist of the following.
@@
 \item The compartments are discrete and have no overlap.
-\item The transition of objects among compartments is described by a set of equations.
+\item The transition of objects into and out of compartments is described by a set of equations.
 \item The populations mix homogeneously
@@
 
-In this section, we discuss 14 different CMs used to estimate $\mathcal{R}$.
+In this section, we discuss 14 different CMs used to estimate $\rr$.
 Different CMs yield different mathematical derivations of \rr.
-That is, for example, an SIR model with non-constant population will yield a different mathematical formula for \rr than the SEIR model with non-constant population.
-The only difference in these two models is that the SEIR model.
-That said, even within the same compartment model framework, we can estimate \rr in a variety of ways.
-In fact, we describe 11 methods to estimate \rr, within the SIR framework!
-Most of these methods are descriptions to estimate parameters within the SIR model and may be extended to other CM frameworks such as the SEIR model.
+That is, for example, an SIR model with non-constant population will yield a different mathematical formula for $\rr$ than the SEIR model (here ``E'' stands for exposed) with non-constant population.
+The difference in these two models is that the SEIR model has an extra compartment which in turn effects the estimates of transitions from one compartment to another.
+That said, even within the same compartment model framework, we can estimate $\rr$ in a variety of ways.
+In fact, we describe 10 methods to estimate $\rr$, within the SIR framework!
+Most of these methods are recipes to estimate parameters within the SIR model and may be extended to other CM frameworks such as the SEIR model.
 The final method described in this section, the next generation operator (NGM) is a general method to derive \rr from almost any CM.
+A particular advantage to CMs compared to survival function methods is that it better fits data that is typically collected about diseases: namely, incidence counts.
+CMs are very popular and still in wide-use today.
 
@@
 
-We begin with the most popular CM, the SIR model.
+We begin with the most well known CM, the SIR model.
 First introduced to epidemiology by  \cite{Kermack700}, the SIR model governs how individuals transition from \textbf{S}usceptible, \textbf{I}nfected, and \textbf{R}ecovered compartments.
 The model is represented graphically in Figure \ref{fig::sir}.
-In this model, once a person is recovered she cannot have the disease again.
+In this model, once an individual is recovered it cannot be infected again.
 
@@
 \end{figure}
-Here, $N$ is the number of individuals in the system.
-As we see no arrows pointing out of the nodes in Figure \ref{fig::sir}, then $N$ is constant throughout time.
+In this model particular  SIR model, $N$, the number of individuals in constant, as we see no arrows pointing out of the nodes in Figure \ref{fig::sir}.
 Simple adaptations of the SIR model can include birth and death rates, which will correspondingly change the derivation of $\rr$ (for more information, see Section \ref{sec:ngm}).
@@
 Here, $\beta$ is the average infection rate and $\gamma$ is the inverse of the average recovery rate.
-The model is represented through the ODEs in Equation \ref{eq::sir}.
+The model is represented through the ODEs below.
 Throughout, we use $X$, $Y$, and $Z$ for stand-ins for the ODEs for $S$, $I$, and $R$, respectively to avoid confusion.
@@
 
-In words, susceptible individuals become infected at a rate that is proportional to the proportion of infected people multiplied by $\beta$, the infection rate, and the number of susceptible individuals.
+In words, susceptible individuals become infected at a rate that is proportional to the proportion of infected individuals multiplied by $\beta$, the infection rate, and the number of susceptible individuals.
 Infected individuals recover at a rate of $\gamma$ multiplied by the number of infected individuals.
 
-An epidemic occurs if 
+An epidemic occurs if the rate of change of susceptibles is positive,
 \begin{align*}
   \frac{dY}{dt} &> 0 \\
-\frac{\beta S I}{N}  - \gamma I &> 0 ,
+\rightarrow \frac{\beta S I}{N}  - \gamma I &> 0 ,
 \end{align*}
 
-that is,  the rate of new infections is greater than the rate of recovery.  So as long as $\frac{S}{N} \approx 1$ then an epidemic will occur if
-\begin{align}\label{eq:sir_r0}
+that is,  the rate of new infections is greater than the rate of recovery.
+So as long as the number of susceptibles is very large, $\frac{S}{N} \approx 1$, then an epidemic will occur if $\rr >1$, as seen in Equation \ref{eq:deriv_sir},
+\begin{align}\label{eq:deriv_sir}
   \rr = \frac{\beta}{\gamma} > 1.
   \end{align}
-
 In the ideal set up, we have data of the number of Susceptibles, Infectious, and Recovered individuals at different points over time.
@@
 
-There are numerous ways to estimate the parameters in this model.  We detail 11 of them here.
+There are numerous ways to estimate the parameters in this model.  We detail 10 of them here.  In every one of these methods, it is assumed that the SIR model is the correct specification.
 
 \subsubsection{Least Squares ($\beta$, $\gamma$) (SIR LS)}\label{least-squares-beta-gamma}
-
-We find
+The first approach to estimate $\rr$ in the SIR model is simply to minimize the mean square error for the data collected at each time point.  In particular, we find
 
@@
 
-Then \(\hat{R}_0= \frac{\hat{\beta}}{\hat{\gamma}}\).
+Then  the estimate for $\rr$ is in Equation \ref{eq:sirls},
+\begin{align}\label{eq:sirls}
+  \hat{R}_0= \frac{\hat{\beta}}{\hat{\gamma}}.
+  \end{align}
 
-We can use gradient descent or another solver to find estimates for these parameters.
-If we assume $Y= f(t) + \epsilon_t$ with $\epsilon_t \overset{iid}{\sim} N(0, \sigma^2)$ where $f(t) = \int_0^t \frac{dY}{dt} dt$, then this this method is equivalent to finding the MLE of $\beta$ and $\gamma$.
-However, we note that the assumption of homoskedasticity seems likely to be violated in this model.
+We can use gradient descent or another optimizer to find estimates for these parameters.
+If we assume $Y= f(t) + \epsilon_t$ with $\epsilon_t \overset{iid}{\sim} N(0, \sigma^2)$ where $f(t) = \int_0^t \frac{dY}{dt} dt$, then this this method is equivalent to finding the maximum likelihood estimator (MLE) of $\beta$ and $\gamma$.
+However, we note that the the $\epsilon_t$ are very likely not independent from one another and so valid inference is still an issue in this framework.
+The advantage is that this is a very straightforward way to estimate $\rr$ from the SIR model.
 
@@
 
-We reparametrized the the ODES in Equation \eqref{eq::sir} directly with \(\rr\) and \(\gamma\), using the relation $\rr = \frac{\beta}{\gamma}$.
+In the previous method, we estimated $\beta$ and $\gamma$ and then estimated $\rr$.
+However, it is possible to directly estimate $\rr$ if we reparametrize the ODEs in Equation \eqref{eq::sir} directly with \(\rr\) and \(\gamma\), using the relation $\rr = \frac{\beta}{\gamma}$.
 
@@
   \begin{array}{cl}
-    \dot{X} &= - \rr \gamma Y \frac{X}{N}\\
-    \dot{Y} &=  \rr \gamma Y \frac{X}{N}  - \gamma Y \\
-    \dot{Z} &=  - \gamma Y 
+    \frac{dX}{dt} &= - \rr \gamma Y \frac{X}{N}\\
+    \frac{dY}{dt} &=  \rr \gamma Y \frac{X}{N}  - \gamma Y \\
+    \frac{dZ}{dt} &=  - \gamma Y 
   \end{array}
@@
 
-We use the \(\rr\) directly from the above calculation.
+We use the \(\rr\) directly from the above estimation process, which again be done with gradient descent or another optimization process.
+Again, we cannot assume $\rr$ has a normal distribution as the errors are likely correlated.
+This method is more advantageous in least squares in finding the distribution of $\rr$ as ratios typically give large confidence intervals.
 
@@
 
-The idea that an SIR model may be well approximated by a linear model is described by \cite{chang2017}.
-We borrow this idea here and use the ODES in Equation \eqref{eq::sir} to relate our estimated linear model to that of $\rr$.
+The SIR model has no known closed form solution and so we are already using approximations using numerical integration, albeit small approximations.
+In addition to this, data collected from real diseases are typically very noisy in the first place.
+\cite{chang2017} posit the idea that an SIR model may be well approximated by a linear model.
+We use this idea here and use the ODEs in Equation \eqref{eq::sir} to relate our estimated linear model to that of $\rr$.
 
@@
 
-Then \(\rr\) is calculated as according to the ODES,
+Then \(\rr\) is calculated as according to the ODEs,
 \[\rr = \frac{\hat{X}^\prime(0)}{ \hat{X}^\prime(0) + \hat{Y}^\prime(0)} \cdot \frac{N}{X(0)} \]
@@
 Here, $K=10$ is rather arbitrary and should be selected using some criterion such as AIC in addition with the knowledge of how many data points are available to the user.
+Besides this select of the degree of polynomials to fit, this model is simplest to implement and gives very comparable results to using least squares with the SIR model.
+Again, we cannot use this model for inference, as our error assumption is violated.
+Another disadvantage is that we are using the first time point $t=0$ to estimate $\rr$.
+However, as this is the edge of the support of time values, it is likely be a biased estimate.
 
@@
 
+The above formulation of a linear model approximation only uses the estimate at time $t=0$ to estimate $\rr$.
+We can instead, use all time points available to use.
 We fit a linear polynomial of \(t\) with degree \(K= 10\) model to \(X\)
@@
 \(\rr\),
-\[\rr = \frac{1}{\# \text{ Obs}}\sum_t \frac{\hat{X}^\prime(t)}{ \hat{X}^\prime(t) + \hat{Y}^\prime(t)} \cdot \frac{N}{X(0)} \]
-The idea is that $\frac{\mathcal{X}^\prime}{X^\prime + Y^\prime}$ is constant in $t$, but due to our approximations with the linear model, this is no longer the case.
+\[\rr = \frac{1}{\# \text{ Obs }t}\sum_t \frac{\hat{X}^\prime(t)}{ \hat{X}^\prime(t) + \hat{Y}^\prime(t)} \cdot \frac{N}{X(0)} \]
+The intuition is that $\frac{X^\prime}{X^\prime + Y^\prime}$ is constant in $t$, but due to our approximations with the linear model, this is no longer the case.
 Here, we average over the different possible values of $\rr(t)$.
+An advantage to this approach is that we have a more robust estimate of $\rr$ than just using one time point.
 
 \subsubsection{Max of Data (SIR Max)}\label{max-of-data}
-
-Under the SIR model there is an absolute maximum of infected individuals $Y$.
-We note that \(Y^\prime(t) = \frac{\beta XY}{N} - \gamma Y > 0\) when \(\rr = \frac{\beta}{\gamma} < \frac{X(t)}{N}\).
-The max
+The $I$ curve in the SIR model has one peak.    We note that \(Y^\prime(t) = \frac{\beta XY}{N} - \gamma Y > 0\) when \(\rr = \frac{\beta}{\gamma} < \frac{X(t)}{N}\). The max
 of \(Y\) occurs when \(Y^\prime = 0\) and hence
@@
 \(t^* = \text{arg max}_{t} Y_{obs}(t)\).
+A major disadvantage to this method is that it only uses one sample, making it a terrible a statistic and more of a mathematical quantity.
+An advantage is that if for instance, one is running simulations, one can make this quick calculation for $\rr$ instead of having to fit a model each time.
 
@@
 \(t^* = \text{arg max}_{t} \hat{Y}(t)\).
+This approach is a statistical approximation of the previous method where we fit a spline to the data.
+In this way, we can utilize all our data.
+Another advantage is the flexibility in model choice in that we can use a spline, linear regression, or any other model which gives us predictions for the number infected and number of susceptibles.
+However, despite this flexibility in model choice, one must remember that we assumed an underlying SIR model for the derivation of $\rr$.
 
@@
   Ratio (SIR IPR)}\label{incidence-to-prevalence-ratio}
+The incidence to prevalence ratio (IPR), described by \cite{Nishiura2009} is another intuitive method to calculate $\rr$ as it uses some of most basic epidemiological quantities, incidence and prevalence.
+
 In terms of data from the SIR model, incidence $J(t) \approx -(X(t+1) - X(t))$, and the IPR$(t) = \frac{J(t)}{Y(t)}$.
 The actual reproduction number, $R_a(t) = \text{IPR}(t)\cdot D$ where $D = 1 /\gamma$.
-This method assumes that we have some prior knowledge about $\gamma$.
+This method assumes that we have some prior knowledge about $\gamma$, the transmission rate.
 Thus we use as our estimate,
@@
 Here we assume that the time step is small enough to approximate the incidence.
+The advantages of this method are that incidence data is generally readily available as is prevalence data for certain diseases such as HIV.
+However, as one is required to have prior knowledge about $\gamma$, it may be easier to directly estimate $\rr$ with one of the many other methods described that does not require a prior knowledge about $\gamma$.
 
@@
 \end{align*}
+The advantage of this method is that it creates a less noisy estimate than estimating IPR using only one point.
+It has the same disadvantages as the regular IPR ratio in that it requires knowledge about $\gamma$.
 
@@
 \subsubsection{Harko (HSIR)}
-Recently, \citep{harko2014exact} were able to reduce the SIR equation, with the following set of ODEs (reparametrized from Equation \eqref{eq::sir})
+Recently, \cite{harko2014exact} were able to reduce the SIR model, with the following set of ODEs (reparametrized from Equation \eqref{eq::sir})
 
@@
 
-where $b = \frac{\beta}{N}$, to one differential equation.  Here,
+where $b = \frac{\beta}{N}$, to one differential equation.  Letting
 \begin{align*}
-  u(t) &= e^{\frac{\beta}{\gamma}Z} \\
+  u(t) &= e^{\frac{b}{\gamma}Z} \\
   t - t_0 &= \int_{u_0}^u \frac{d \xi}{\xi (C_1 - \gamma \log \xi + X(0) b \xi)},
 \end{align*}
-where $C_1 = -\frac{b}{N}$.  Then
+where $C_1 = -\frac{b}{N}$, then
 
 \begin{align}
-  X &= X(0) u  = X(0) e^{\frac{\beta}{\gamma}Z} \nonumber\\
-  \log \frac{X(t)}{X(0)} &=  \frac{b}{\gamma}Z =  \frac{\beta}{\gamma N} Z \nonumber\\
-  \log \frac{X(t)}{X(0)} &=  \frac{\rr}{N} Z \label{eq:harko_lin}
+  X(t) &= X(0) u  = X(0) e^{\frac{b}{\gamma}Z(t)} \nonumber\\
+  \log \frac{X(t)}{X(0)} &=  \frac{b}{\gamma}Z(t) =  \frac{\beta}{\gamma N} Z \nonumber\\
+  \log \frac{X(t)}{X(0)} &=  \frac{\rr}{N} Z(t) \label{eq:harko_lin}
 \end{align}
@@
 This method greatly simplifies the computational effort needed to estimate $\rr$ as it does not require any numerical integration.
-
-\subsubsection{Attack Rate (SIR AR)}
-
-The Attack Rate (AR) as described by \cite{obadia2012r0} as the ``percentage of the population eventually infected.''  Under the SIR (XYZ) model with constant population,
-
-\begin{align}\label{r0_attackrate}
-\rr =  \frac{\log \left (\frac{1  - AR}{X(0)/N}  \right ) }{AR - (1 - X(0)/N)}
-\end{align}
-
-This method can only be used to assess $\rr$ after a disease has passed through a population.
-Additionally, if either $\beta$ or $\gamma$ were affected during the disease's lifetime then $\rr$ cannot be properly assessed.
+Another advantage is that unlike using the linear approximation method, the coefficients from this model are directly interpretable.
+A disadvantage is that again, $\epsilon_t$ are  not likely iid.
 
@@
 
-Posited by \cite{bettencourt2008} and summarized in \cite{obadia2012r0}, this method uses a Bayesian approach to an approximation to the classic SIR model.
-The approximated SIR model is that the incidence at $t+1$, $N(t+1)$ is from a Poisson distribution, with $\gamma$ as the  average inverse of the infectious period and $R$ as the effective reproduction number, which we take here to be $\rr$.
+Posited by \cite{bettencourt2008} and summarized in \cite{obadia2012r0}, this sequential Bayes is a Bayesian approach to an approximation of the classic SIR model.
+The approximated SIR model assumes that the incidence at $t+1$, $J(t+1)$ has a Poisson distribution, with $\gamma$ as the  average inverse of the infectious period and $\mathcal{R}$ as the effective reproduction number, which we take here to be $\rr$.
 
@@
 \begin{align*}
-N(t+1)  \sim Poisson( N(t) \exp \left \{  \gamma (R-1)\right \})
+J(t+1)  \sim Poisson( J(t) \exp \left \{  \gamma (R-1)\right \})
 \end{align*}
@@
 \begin{align*}
-  P(\rr | N_0, \dots, N_{t+1}) = \frac{P(N_{t+1} | \rr, N_0, \dots, N_t)P(\rr| N_0, \dots, N_t)}{P(N_0, \dots, N_{t+1})}.
+  P(\rr | J_0, \dots, J_{t+1}) = \frac{P(J_{t+1} | \rr, J_0, \dots, J_t)P(\rr| J_0, \dots, J_t)}{P(J_0, \dots, J_{t+1})}.
 \end{align*}
@@
 This method assumes, initial growth in incidence to be exponential, and homogeneous mixing of populations as with any compartment model.
+The advantages of this method are that of the ability to obtain an entire posterior distribution, whereas many other methods are difficult to even find an estimate of the variance.
+Disadvantages include strict assumptions about the distributions and computational time.
 
@@
 
-The SEIR model is a common adaptation of the SIR compartment model \com{add references that use this method}.
+The SEIR model is a common adaptation, especially in recent years, of the SIR compartment model used to model infectious diseases such as influenza and Ebola\citep{mills2004,althaus2014}.
 The ``E'' compartment stands for ``exposed'' and represents the stage where individuals are infected but not yet infectious.
@@
 Then $\rr = \beta / \gamma$, just as in the SIR model.
-As in the SIR model, $\beta$ and $\gamma$ have the same interpretation and $\alpha$  The methods used to estimate the parameters $\beta$ and $\gamma$ are of the same class used to estimate those parameters in the SIR model.
-The important difference between this method and the SIR model is the assumption of a 4th class, the exposed compartment.
-Even though the derivation for $\rr$ is the same for both the SIR and SEIR model, this will still result in different results for estimations of \rr, even if the same method for estimation of parameters is used.
-Typically, some form of least squares is used to estimate parameters in this model.
-We demonstrate this in Section \com{results}.
+As in the SIR model, $\beta$ and $\gamma$ have the same interpretation and $\alpha$ is the transmission rate from the exposed compartment to infected.
+The methods used to estimate the parameters $\beta$ and $\gamma$ are of the same types used to estimate those parameters in the SIR model.
+An important difference between this method and the SIR model is the assumption of a 4th class, the exposed compartment.
+Even though the derivation for $\rr$ is the same for both the SIR and SEIR model, this will still result in different results for estimations of $\rr$, even if the same method for estimation of parameters is used.
+Here, we see a very strong example of how $\rr$ is a property of the model and why it is important to argue why this model specification is correct.
+An advantage of this model is that the addition of an exposed class is thought to better mimic reality of the transmission of many diseases.
+A major disadvantage is that exposure counts are typically back-calculated from incidence counts, and noise is rarely considered in this estimation of the exposure counts.
+Typically, some form of maximum likelihood estimation is used to estimate parameters in this model, which places strict assumptions on each of the compartments.
+We demonstrate this in Section \ref{sec:results}.
 
-\subsubsection{Amplifier Model (Amp)}
-\cite{blower2004} discuss the case where a disease may have different strains and consequently different reproduction numbers.
+\subsubsection{Amplifier Model (AMP)}
+As mentioned previously, compartment models may become arbitrarily complex.
+We demonstrate this idea with a set of compartments for each strain of the disease.
+\cite{blower2004} discuss the case where a disease may have different strains and consequently a different reproduction number for each strain.
 They introduce a multistrain mathematical model called the amplifier model to deal with this case.
 
-This model is based on the the SLIR compartment model, where $L$ stands for latently infected individuals and $T$ for diseased individuals.
+This model is based on the the SEIR compartment model, accounting for a number of strains of a disease at once.
 There is a set of ODEs for each strain of the disease with different parameters for recovery, transmission rate, and drug resistance of the disease.
@@
 \begin{align*}
-\rr^{(i)} = X^* \frac{ ( \beta_i^T + \beta_i^L)\nu_i + \beta_i^T \mu_i^L}{(\nu_i + \mu_i^L)(c_i + k_{i,i+1} + \mu_i^T)}.
+\rr^{(i)} = X^* \frac{ ( \beta_i^Y + \beta_i^E)\nu_i + \beta_i^Y \mu_i^E}{(\nu_i + \mu_i^E)(c_i + k_{i,i+1} + \mu_i^Y)}.
 \end{align*}
-Here, $X^*$ is the number of susceptibles individuals in the disease-free equilibrium in the absence of immigration or emigration; $\beta_i^L$ and $\beta_i^T$ are strain-specific transmission rates for the latent, and infected classes, respectively; $\nu_i$ denotes the progression rate from latently infected to disease for strain $i$; $\mu_i^L$ and $\mu_i^{T}$ are death rates for the latent and infected classes, respectively; $c_i$ is the cure rate; and $k_{i, i+1}$ is a parameter associated with the degree of amplification of drug resistance.
+Here, $X^*$ is the number of susceptibles individuals in the disease-free equilibrium in the absence of immigration or emigration; $\beta_i^E$ and $\beta_i^I$ are strain-specific transmission rates for the latent, and infected classes, respectively; $\nu_i$ denotes the progression rate from latently infected to disease for strain $i$; $\mu_i^E$ and $\mu_i^{Y}$ are death rates for the latent and infected classes, respectively; $c_i$ is the cure rate; and $k_{i, i+1}$ is a parameter associated with the degree of amplification of drug resistance.
 They note that ``each reproduction number is the product of average number of secondary infections caused per unit time, the aver time a case remains infectious and the probability that an infected individual develops disease.''
@@
 
-We show this model as opposed to other available CMs because it shows not only how complicated CMs can become, but researchers approach the assumption of homogeneity, which is often suspect.
-In this case, we have strain-specific transmission rates, so we are essentially stratifying compartments into different sets.
-Very complicated models have arisen from stratification of compartments, which leads to the approach of ABMs.
+We show this model as opposed to other available CMs because it shows not only how complicated CMs can become, but also to demonstrate how researchers approach the often controversial assumption of homogeneity.
+In this case, we have strain-specific transmission rates, so we are essentially stratifying compartments into different groups, allowing for some heterogeneity.
+A disadvantage is the amount of typed data required to estimate all parameters.
 
@@
 
+Very complicated models have arisen from stratification of compartments such as one with 26 compartments \citep{pandey2014}.
+As such, there is a need to be able to derive an expression for $\rr$ regardless of the structure of the compartment model.
 The Next Generation Model (NGM) is a generalization of any compartment model at the infection-free steady state.
@@
 Although, this method is useful for deriving an expression of $\rr$, it is unclear how to actually estimate it from these expressions.
-Again, least squares is typically used to estimate such parameters.
+One could use maximum likelihood if one specifies distributions for how compartments change over time or least squares may be applied.
+In general, adding numerous compartments is often tedious and difficult and needlessly complicated.
+Instead, researchers have started to lean towards ABMs when introducing heterogeneity in their models, which we describe in Section \ref{sec:network}.
+Another disadvantage, is that $\rr$ is only derived using the disease-free steady state, which is an approximation.
+Other approximations such as exponential growth are also used to derive expressions for $\rr$.
 
@@
 Typically, CMs expressed through differential equations have no known analytical solutions, which makes deriving an expression for $\rr$ difficult.
-One simplification that is often introduced to the SI framework is that of exponential growth, or that the number of infections in the beginning of an epidemic mimics that of an exponential curve.
+One simplification that is often introduced to the SI-framework is that of exponential growth, or that the number of infections in the beginning of an epidemic mimics that of an exponential curve.
 This simplified expression has allowed for many derivations of $\rr$.
@@
 \label{sec:expgrowth}
-\cite{wallinga2007generation} report that $\mathcal{R}_t$ and hence $\rr$ may calculated by using the fact that infection ``counts increase exponentially in the initial phase of an epidemic.''  We need to know $r$, the \textit{per capita} change in the number of new cases per unit of time and $\omega$ the serial interval.
+\cite{wallinga2007generation} report that the effective reproduction number $\mathcal{R}_t$ and hence the initial reproduction number $\rr$ may derived using the fact that infection ``counts increase exponentially in the initial phase of an epidemic.''  We then have to estimate $r$, the \textit{per capita} change in the number of new cases per unit of time and $\omega$ the serial interval.
 
@@
 
-Equation \eqref{eq:lotka} is derived from a demographic view using the Lotka-Euler survival equations which come from the fields of demography, ecology, and evolutionary biology, whereas Equation \eqref{eq:anderson}is derived through an epidemiologist.
+Equation \eqref{eq:lotka} is derived from a demographic view using the Lotka-Euler survival equations which come from the fields of demography, ecology, and evolutionary biology, whereas Equation \eqref{eq:anderson} is derived through an epidemiologist's perspective.
 These estimates for $\rr$ between Equation \eqref{eq:lotka} and Equation \eqref{eq:anderson} can be strikingly different.
@@
 For all these above estimations of $\rr$, we observe the duration of serial intervals in a period of exponential growth.
-Deciding when this occurs and which data to use to estimate $\rr$ is difficult.
+Deciding when exponential growth occurs and which data points to use to estimate $\rr$ is difficult.
 \citeauthor{wallinga2007generation} specify $\omega$ in multiple ways.
@@
 \subsubsection{Maximum Likelihood Estimator of Secondary Infections (MLSI)}\label{sec:mle-si}
-This method, described by \cite{forsberg2008}, finds the estimate of $\rr$ by maximizing the likelihood function under certain assumptions.
+A common method used to estimate parameters in the SI-framework is maximum likelihood estimation.
+This method requires knowing the likelihood of a parameter given the data observed, and hence many distributions may be specified.
+The MLE of secondary infections, described by \cite{forsberg2008}, finds the estimate of $\rr$ by maximizing the likelihood function under a certain assumptions.
 We assume that the ``the number of secondary cases produced by an infected individual follows a Poisson distribution, and that the serial interval is described by a multinomial distribution.''  Recall, the serial interval is the distribution of time between a primary case developing symptoms and a case infected by the primary case developing symptoms.
@@
 
-This method assumes a Poisson distribution on $\rr$ in addition to assuming a multinomial on the serial interval.  Additionally, this method also assumes a period of exponential growth,
+This method assumes a Poisson distribution on $\rr$ in addition to assuming a multinomial on the serial interval.
+Additionally, this method also assumes a period of exponential growth.
+An advantage is that assuming a Poisson distribution has nice properties which make it possible derive some of the statistical properties of this method.
+Two disadvantages are that we only have an approximation for $\rr$ and have to decide which data points to use in a certain time frame.
 
@@
 \subsubsection{Time Dependent Reproduction Number (TDR)}\label{sec:timedep}
-This is another method used to estimate the effective reproduction number, which again, we take to be $\rr$ here \citep{forsberg2008}.
-This is a likelihood based approach where we maximize the likelihood that case $i$ has been infected by case $j$ at certain time using the serial interval, $w(t)$ where $t_i$ is when case $i$ was infected and $N$ is the total number of cases,
+Another likelihood based approach to estimate $\rr$ is that of the time dependent reproduction number, as described by \cite{forsberg2008}.
+Here, we maximize the likelihood that case $i$ has been infected by case $j$ at certain time using the serial interval, $w(t)$ where $t_i$ is when case $i$ was infected and $N$ is the total number of cases,
 \begin{align*}
   p_{ij} &= w(t_i- t_j) / \sum_{i \neq k} w(t_i - t_k),\\
-  \rr &= \frac{1}{I(0)}\sum_{i=1}^N p_{i0}
+  \rr &= \frac{1}{Y(0)}\sum_{i=1}^N p_{i0}
   \end{align*}
 
-  It is assumed we have some idea about the generation interval distribution, $w(t)$.
+  A disadvantage is that it assumed we have some idea about the generation interval distribution, $w(t)$, but an advantage is that there is no explicit assumption on the distribution of $\omega$.
+This method also requires that the initial number of susceptibles $X(0)$ is roughly the size of the total population.
+Another advantage is the explicit use of distributions to indicate that $\rr$ is indeed a random variable.
 
@@
 Here $\mu_0$ is the proportion of initially infected individuals and $\mu_\infty$ is the final, cumulative proportion of infected cases at the end of an epidemic,
-  \begin{align*}
-    \rr = (1- \mu_\infty)^{-1} \ln \mu_\infty^{-1}.
-  \end{align*}
 
-  For final size, it is assumed the initial population is completely susceptible
+\begin{align}\label{r0_attackrate}
+\rr =  \frac{\log \left (\frac{1  - \mu_0}{X(0)/N}  \right ) }{\mu_0 - (1 - X(0)/N)}
+\end{align}
+
+This method can only be used to assess $\rr$ after a disease has passed through a population.
+Additionally, if either $\beta$ or $\gamma$ were affected during the disease's lifetime then $\rr$ cannot be properly assessed.
 
@@
 Networks that have been used to estimate $\rr$ include contact tracing of actual individuals, branching processes to model the spread of disease, and agent-based models (also known as individual level models), which all leverage networks along with different sources of data to generate estimates of $\rr$.
+Additionally, networks provide an intuitive framework in which individuals can act with varying levels of heterogeneity.
 
@@
 A reasonable idea to estimate $\rr$ is to directly observe the number of secondary infections produced by an initial infector.
-However, this approach is generally impractical as it is difficult to estimate every single contact a person may have.
+However, this approach is generally impractical as it is difficult to estimate every single contact an individual may have.
 With seriological information, we could conceivably recreate the transmission of the disease throughout a population.
-Even in this case, it is unlikely, for one, to have a completely susceptible population, which is essential to an unbiased estimate of $\rr$.
+Even in this case it is unlikely, for one, to have a completely susceptible population, which is essential to an unbiased estimate of $\rr$.
 Moreover, this would implicitly be conditioning on the initial infector's contacts which may not reflect those of the rest of the population.
 We write the calculation for $\rr$ under contact tracing in Equation \ref{eq:r0_contacttracing}, where $X_0$ is a random variable denoting the number of individuals in the population person 0 infects.
-Here $S_0=N$ indicates that the initial number of susceptibles is $N$, the size of the population.
-
+Here $S_0=N$ indicates that the initial number of susceptibles is $N$, the size of the population,
 \begin{align}\label{eq:r0_contacttracing}
-\rr = E[ X_0 | S_0 = N]
+\rr = E[ X_0 | S_0 = N].
 \end{align}
-
 \cite{eames2003} describe contact tracing as an ``extreme form of targeted control, where the potential next-generation cases are the primary focus,'' that is, one focuses on treating the contacts of an infected individual rather than applying uniform treatment across a population.
@@
 They describe an estimate for $\rr$ for a SIR model on a network, with $r$ as the rate of transmission across a contact multiplied by the infectious period and $n$ as the number of contacts as
-
 \begin{align*}
@@
 \end{align*}
+Again, the difficulties of this method lies in one, estimating the contacts of an initial infected individual and two, extrapolating the contacts of the initial individual to the population as a whole.
+As such, although very useful in helping eradicate a disease, is not the best choice to estimate $\rr$ due to the high level of uncertainty inherit in the method.
 
@@
 \cite{getz2006} give a way to calculate $\rr$ from a branching process.
-```We define the offspring distribution $\{q_i \}_{i=0}^\infty$, where $q_i$  is the probability that an infectious individual infects $i$ other individuals.
+``We define the offspring distribution $\{q_i \}_{i=0}^\infty$, where $q_i$  is the probability that an infectious individual infects $i$ other individuals.
 Thus we require $\sum_{i=0}^\infty q_i =1$ and note that $\rr$, the mean number of cases contracting disease from each infective'', is simply given by'
@@
   It is assumed we know how to estimate the $q_i$, and so in this case, it is not too different of an approach than contact tracing.
+The disadvantages of this method are similar to that of contact tracing as data about which individuals  infected another is very difficult to obtain.
+An advantage is that this method directly gets to the ``heart'' of the definition of $\rr$ as it is concerned with generations of infected individuals rather than incidence at a certain time.
 
@@
 ABMs allow for every detail of an infection to be known, in the context of estimating $\rr$, meaning we know exactly who is infected by whom and when.
-Thus $\rr$ can be calculated simply through
+Thus $\rr$ can be estimated simply through
 \begin{align*}
@@
 \end{align*}
+Examples of estimations of $\rr$ using ABMs/ILMs are in \cite{breban2007,ahmed2013variance}.
 
-The difficulty in the ABM lies in calibrating the evolving of agents over time to accurately reflect reality, which is a problem that cannot be understated.
+There are many advantages to using ABMs, a main one being able to run the model multiple times under the same initial conditions.
+In this way, one can easily obtain variance estimates of $\rr$ in addition to empirical distributions of it.
+Moreover, this estimation can accommodate any model specification of how the disease is transmitted through the population.
+However, the major disadvantage of the ABM lies in calibrating the evolving of agents over time to accurately reflect reality, which is a problem that cannot be understated.
 Thus, it is typically necessary to have knowledge of parameters such as the serial interval, transmission rate, recovery rate, etc.
@@
 ABMs are flexible in that they may be used with any model in the SI framework.
-The only requirement is that we track who infects whom and can adapt for any additional states such as exposure.
-
-Examples of calculations of $\rr$ using ABMs/ILMs are in \cite{breban2007,ahmed2013variance}.
+The only requirement is that we track who infects whom and can adapt for any additional states such as exposure or prevention strategies such as closing schools \citep{fred}.
 
@@
 Estimating $\rr$ is difficult, and it is no different for estimating $V[\rr]$, the variance.
-We describe four general methods which may be applicable to estimate the variance.
+We describe general methods which may be applicable to estimate the variance.
 We describe the assumptions that are used along with each method and note that they may not be applicable to all of our methods.
 Still, reporting confidence intervals, or at the very least, standard deviations is very important and we hope to encourage this practice by describing the below methods used to estimate models in the SIR framework in a simple fashion.
+Regardless of the method used to estimate variance, confidence intervals, or credible intervals, we encourage researchers to carefully examine all underlying assumptions of 1) deriving an expression for $\rr$, 2) estimating $\rr$, and three the method used to estimate the variance or CIs.
 
@@
 Here, we use the relationship of $\rr = \frac{\beta}{\gamma}$ in order to estimate the variance, which is specific to the SIR framework.
-This method may be extended to other frameworks and we provide some examples here \com{add examples of SEIR, serial interval, etc.}
+This method may be extended to other frameworks and we provide some examples here \com{add examples of SEIR, serial interval, etc.}.
+The advantages of the delta method are that as the NGM posits a recipe to derive $\rr$, then it is theoretically possible to use the delta method to derive an expression for an expression of variance for any compartment model.
 
@@
 
-We estimate \(\frac{\partial Y}{\partial \beta}\) (or
-\(\frac{\partial Y}{\partial \rr}\) when using reparametrized least squares) and \(\frac{\partial Y}{\partial \gamma}\) at each time point \(t\) using the estimated \(\beta\) or \(\gamma\) from the estimation process where $Y$ stands for the number of infected at time $t$.
-Then an approximation of the covariance matrix is \((\chi^T \chi)^{-1}\) where \(\chi_{ij} = \frac{ \partial Y(t_i)}{ \partial \theta_j}\) where \(\theta\) is a vector of parameters.
-\com{add citation}
+Sensitivity analysis is used to measure how robust estimates are to change.
+If small changes in our estimated parameters do not change our estimates for the number of infected by much, then we may prefer this method to another.
+For the SIR model, we estimate \(\frac{\partial Y}{\partial \beta}\) (or \(\frac{\partial Y}{\partial \rr}\) when using reparametrized least squares) and \(\frac{\partial Y}{\partial \gamma}\) at each time point \(t\) using the estimated \(\beta\) or \(\gamma\) from the estimation process where $Y(t)$ stands for the number of infected at time $t$.
+More generally,  let $Y(t_i) = f(t_i, \theta)$ where $\theta$ is a vector of parameters.
+Then an approximation of the covariance matrix is \((\chi^T \chi)^{-1}\) where \(\chi_{ij} = \frac{ \partial Y(t_i)}{ \partial \theta_j}\).
+\com{add citation}.
+Advantages of this method include that it is a standard way to estimate the variance of $\rr$.
+
+\subsubsection{Posterior Distribution}
+When one uses a Bayesian approach to estimate $\rr$, the result should be a distribution rather than an estimate.
+One can simply look at the quantities of this distribution to form a credible interval for this estimate.
+Advantages of this method include having an entire distribution compared to a point estimate.
+One disadvantage is that the Bayesian approach will not guarantee frequentist properties that epidemiologists may be looking for such as being able to compare diseases to one another.
 
@@
 \bibliography{Master}
+
+\appendix
+
+\section{Epidemic Modeling Terminology}
+\label{sec:term}
+
+Here we summarize some of the terminology and notation used throughout the methods for estimating \rr.
+
+{$\rr$} -- reproduction number --
+
+{$\mathcal{R}_t$} -- effective reproduction number
+
+{$\beta$} -- the infection rate
+
+{$\gamma$} -- inverse of the expected recovery time
+
+{$N_t$} -- Number of individuals in the population at time $t$
+
+{$X(t)$} -- the number of susceptible individuals in a population at time $t$
+
+{$Y(t)$} -- the number of infected individuals in a population at time $t$
+
+{$Z(t)$} -- the number of recovered individuals in a population at time $t$
+
+{$\omega$} -- serial or generation interval -- the time from onset of symptoms in an index case to the onset of symptoms in a subsequent case infected by the index patient
+
+AR -- attack rate -- percentage of the population eventually infected
 

### 2018-07-03 paper/old_drafts/draft_v4.tex
finally on git

--- 2018-07-03 paper/old_drafts/draft_v3.tex
@@
 \newcommand{\com}[1]{\textcolor{red}{ #1}}
+
+\linespread{2}
 
@@
 
-\newcommand{\XX}{\ensuremath{23}} % number of methods
+\newcommand{\XX}{\ensuremath{25}} % number of methods
+
+\newcommand{\xxsir}{\ensuremath{11}} % number of SIR methods
+
 \newcommand{\rr}{\ensuremath{\mathcal{R}_0}}
@@
 \section{Introduction}\label{sec:intro}
-What has been called ``arguably the most important quantity in the study of epidemics'', $\mathcal{R}_0$, the reproduction number, is an elusive quantity for epidemic modelers \cite{Heesterbeek2002}.
-\citet{anderson1992} define $\rr$ as the ``the average number of secondary infections produced when one infected individual is introduced into a host population where everyone is susceptible.''  $\rr$, in some ways, summarizes an entire epidemic.
-This quantity is used whether to assess if a disease will break out and what percentage of the population needs to be vaccinated to avoid such an outbreak.
-Despite the clear definition of $\rr$, epidemiologists have struggled to create a \textit{de facto} estimator for $\rr$  \citep{hethcote2000}.
-The primary issue in estimating $\rr$ is that the quantity is a \textit{property of the model}, meaning that $\rr$ is dependent not only on the usual noise that comes with statistical modeling but also on a variety of assumptions on how researchers assume disease is transmitted through a population.
+What has been called ``arguably the most important quantity in the study of epidemics'', $\mathcal{R}_0$, the reproduction number, has been an elusive quantity to estimate for epidemic modelers for nearly a century \citep{Heesterbeek2002}.
+Defined by \citet{anderson1992}, $\rr$ (pronounced $R$-naught) is the ``the average number of secondary infections produced when one infected individual is introduced into a host population where everyone is susceptible.''  $\rr$, in some ways, summarizes an entire outbreak of a disease as this quantity is used whether to assess if a disease will break out and what percentage of the population needs to be vaccinated to avoid such an outbreak.
+Despite a clear definition of $\rr$, epidemiologists have struggled to create a \textit{de facto} estimator for $\rr$  \citep{hethcote2000}.
+The primary issue in estimating $\rr$ is that the quantity is a \textit{property of the model}, meaning that $\rr$ is dependent not only on the usual noise that comes with statistical modeling but also on a variety of assumptions on how researchers assume a disease is transmitted through a population.
 
 Typically, disease models are a part of the ``SI-framework'' pioneered by Kermack and McKendrick in the 1920s \citep{getz2006}.
-Here, S and I are known as compartments where S stands for ``susceptible'' and I for ``infected.''  The issue arises in how individuals are assumed, either implicitly or explicitly, to travel from the S compartment to the I compartment and any other compartments that have been introduced such as immune, recovered, dead, or exposed, just to name a few.
+Here, S and I are known as compartments where S stands for ``susceptible'' and I for ``infected.''  The issue of estimating $\rr$ stems from how individuals are assumed, either implicitly or explicitly, to move from the S compartment to the I compartment and any other compartments that have been introduced such as immune, recovered, dead, or exposed, for example.
 One of the simplest models in the SI framework is the SIR model where R stands for a ``recovered'' state where individuals are no longer susceptible to the disease \citep{Kermack700}.
@@
 
-Other issues of estimating $\rr$ include mathematical versus statistical approaches (i.e.
-solving for $\rr$ versus estimating $\rr$), how to estimate the quantity in the presence of different available data, and when to estimate $\rr$ in a disease's lifetime.
+Other issues about estimating $\rr$ include mathematical versus statistical approaches (i.e.
+solving for $\rr$ versus estimating $\rr$), how to estimate the quantity in the presence of types of available data, and when to estimate $\rr$ in a disease's lifetime.
+This depends on the ``time constant'' of the disease.
+Some diseases like measles progress rapdily, HIV more slowly, prion diseases glacially so.
 For instance, if only weekly incidence counts of a disease are available, then we have different methods of estimating $\rr$ available than when we have more data available such as contact networks or the serial interval of a disease.
-Another major issue is that since $\rr$ is so difficult to estimate in the first place, the variance of the quantity has largely been glossed over or completely ignored.
-This is especially troubling as we are concerned whether $\rr > 1$, which indicates an outbreak of a disease.
+Another major issue is that since $\rr$ is so difficult to estimate in the first place, the estimated variance of the estimated $\rr$ has largely been glossed over or  ignored.
+This is especially troubling because we are concerned whether $\rr > 1$, which indicates an outbreak of a disease.
 A point estimate is simply not enough to determine the truth of this inequality.
 
-In short, $\rr$ is difficult to estimate because it is a property of the model.
-This, in turn, makes it difficult to compare $\rr$ across different diseases or even instances of the same disease.
-Many of these issues with estimating $\rr$ are summarized and expanded upon by \cite{li2011}.
+In short, $\rr$ is an important parameter in epidemiology that is very difficult to estimate.
+This, in turn, makes it difficult to compare $\rr$ across different diseases or even different instances of the same disease.
+Many of these challenges with estimating $\rr$ and more are summarized and expanded upon by \cite{li2011}.
 
-We review $\XX$ ways to estimate $\rr$ over varying assumptions and in the presence of different sources of data.
-The title of this manuscript is in reference to a seminal paper by Moler and Van Loan in 1978 (and its subsequent update in 2003) in which they calculate ``Nineteen dubious ways to compute the exponential of a matrix'' \citep{moler2003}.
-We not only review the methods of estimating of $\rr$ but also include recommendations on when to use the methods as well as introducing techniques in which to estimate the variance of estimators.
+We review $\XX$ ways to estimate $\rr$ across a number of model assumptions and in the presence of different sources of data.
+The title of this manuscript is in reference to a seminal paper by Moler and Van Loan (1978) (and its subsequent update in 2003) in which they calculate ``Nineteen dubious ways to compute the exponential of a matrix'' \citep{moler2003}.
+We not only review the methods of estimating of $\rr$ but also include recommendations on when to use the methods as well as introducing techniques which allow estimation the variance.
 We compare these estimates, when applicable, and find that $\rr$ is more robust than what was initially presumed.
@@
 In Section \ref{sec:overview}, we introduce the methods in four broad groups: the survival function, compartment models, exponential growth models, and agent-based models.
-In Section \ref{sec:details} we give an overview of each of the methods, what assumptions are used, what data is required, and where the method has been used in the past along with ways in which to calculate the variance for methods, when applicable.
-In Section \ref{sec:results}, we give estimates of $\rr$ for the different methods and their variances over a variety of data sets.
+In Section \ref{sec:details}, we give an overview of each of the methods, what assumptions are used, what data is required, and where the method has been used in the past along with ways in which to estimate the variance of the methods, when applicable.
+In Section \ref{sec:results}, for a number of different data sets, we give estimates of $\rr$ for the different methods and their variances over a variety of data sets.
 Finally, in Section \ref{sec:dis} we discuss our conclusions and recommendations.
@@
 
-We chose $\XX$ methods to review in this paper, which does not cover all possible methods to estimate $\rr$.
-These methods were chosen as  they represent the type of models typically used to model $\rr$ in addition to their high impact in the field of epidemiology.
+In this manuscript, $\XX$ methods were chosen to give a survey of methods for estimating $\rr$.
+This is not exhaustive, but these methods were chosen due to historical importance and impact in the field of epidemiology.
 
-In order to organize the methods for estimating \rr, we have chosen four broad classes of estimation methods.
-These partitions are meant to be guidelines of distinguishing methods from one another rather than immutable groups.
-Many of these methods could have fit in another category or even multiple of the groups chosen.
-We have chosen the four classes as application of the survival function and its close relatives, classic compartment models, exponential growth assumptions, and estimation from networks.
+In order to organize the methods for estimating $\rr$, we have chosen four broad classes of estimation methods.
+These partitions are meant to be guidelines for distinguishing methods from one another rather than immutable groups.
+Many of these methods  fit in  mulltiple of the groups shown.
+We have chosen the four groups as application of the survival function and its close relatives, classic compartment models, exponential growth assumptions, and estimation from networks.
 We list these point estimate methods below.
@@
     \item Incidence-prevalence ratio (SIR IPR)
-    \item Smooth incidence-prevalence Ratio (SIR SIPR)
-    \item Harko's SIR model (HSIR)
+    \item Smooth incidence-prevalence ratio (SIR SIPR)
+    \item Log-linear (SIR LL)
+    \item Markov chain estimation (SIR MC)
     \item Sequential Bayes (SIR SB)
@@
   \item SEIR model (SEIR)
+  \item MSEIR model (MSEIR)
   \item Amplifier model (AMP)
@@
 
-  Additionally, we describe  interval estimates for the above methods for estimating \rr.
+  Additionally, we describe  interval estimates for the above methods for estimating $\rr$ in order to better answer the question of whether $\rr > 1$, which implies an outbreak of the disease..
 
@@
 
-The survival function is intimately tied to the history of \rr, which originates from the field of demography in the late 1800s \citep{dietz1993estimation}.  In demography, we have
+The first set of methods are those of hte survival function, which is closely  tied to the history of $\rr$.
+The survival function, which originates from the field of demography in the late 1800s, described how many female offspring a woman would produce \citep{dietz1993estimation}.
+In demography, we have
 \begin{align*}
@@
 \end{align*}
-where $p(a)$ denotes the probability of a woman surviving age $a$ and $\beta(a)$ the rate of giving birth to a  girl for an individual of age $a$.
-This interpretation also explains the name of \rr, which was imported to the field of epidemiology by MacDonald and Smith as reported by \cite{dietz1993estimation}.
-In epidemiology, the analogy is that $p(a)$ is like the duration of a disease, and $\beta$ is the transmission rate.
+where $p(a)$ denotes the probability of a woman surviving to age $a$ and $\beta(a)$ the rate of an individual of age $a$ giving birth to a  girl.
+This interpretation also explains the origins of the name of $\rr$, which was imported to the field of epidemiology by MacDonald and Smith \citep{dietz1993estimation}.
+Analogously in epidemiology, $p(a)$ is the age of a disease, and $\beta$ is the transmission rate.
 
-This is the chronologically first and in many ways, the most direct method to calculate $\rr$.
-However, this model also seems to be the most difficult to estimate as both $p(a)$ and $\beta (a)$ as $p(a)$ is a distribution and $\beta$ is a function.
-Typically, it is also difficult to estimate these parameters before or even during an ongoing disease outbreak.
-In turn, $\rr$ can only be estimated after an outbreak, which means it cannot be used to determine ongoing prevention strategies.
-However, even having a post-disease estimation of $\rr$ may be useful for future outbreaks of the disease or being able to compare the severity of many diseases.
-Another advantage is that this method is very intuitive and is independent of assumptions of how the disease is transmitted from person to person.
+This is the chronologically first, and in many ways, the most direct method to estimate $\rr$.
+However, this model also seems to be the most difficult to estimate as both $p(a)$ and $\beta (a)$ are random variables, a distribution and funtion, respectively.
+In general, it is also difficult to estimate these parameters before or even during an ongoing disease outbreak.
+As such, $\rr$ can only be estimated after an outbreak, which means it cannot be used to determine ongoing prevention strategies.
+However, having a post-disease estimation of $\rr$ may be useful for preventing future outbreaks of the disease and aids in comparing the severity of many diseases.
+Another advantage of this method is it intuitive and independent of assumptions of how the disease is transmitted from person to person.
 We describe four methods attempting to estimate $\rr$ from the survival function:  the survival function alone, direct estimation, and direct parameter estimation.
@@
 
-Let $F(a)$ be the probability that a newly infected individual remains infectious for at least time $a$ (the survival probability, $p(a)$ in the previous definition).
-Moreover, let $b(a)$ $(\beta (a)$ in the previous definition), the average number of newly infected individuals that an infectious individual will produce per unit time when infected for total time $a$.
+Let $F(a)$ be the probability that a newly infected individual remains infectious for at least time $a$, and let $b(a)$ be the average number of newly infected individuals that an infectious individual will produce per unit time when infected for total time $a$.
 Then $\rr$ is derived in Equation (\ref{eq:r0_survivalfxn}),
@@
 This method presents one of the unique difficulties of estimating $\rr$: that of reconciling generations and time.
-$\rr$ tells us about the generations of disease, meaning a parent and their offspring but not necessarily when they occur.
-This is made difficult as the data we have about a disease is often of a temporal nature.
-On the other hand, this method is flexible in that this derivation of $\rr$ is does not place assumptions on how the disease transmits through a population.
+$\rr$ tells us about the generations of disease, meaning an infector and tis infected, but not necessarily when the infections occur.
+This reconciliation between generations and time is made difficult as the data we have about a disease is often of a temporal nature.
+On the other hand, this method is flexible in that this derivation of $\rr$  does not place assumptions on how the disease transmits through a population.
 A disadvantage is that data about $F(a)$ and especially $b(a)$ is difficult to collect directly.
@@
 
-Direct parameter estimation is a `plug and chug' approach as described by \cite{lipsitch2003} and \cite{dietz1993estimation}, which has an intuitive derivation, similar to that of the survival function.
-The intuition is that we can estimate $\rr$ if we know how many others an individual comes into contact, the probability of transmission given contact, and how long the disease lasts within an individual  In this method, we estimate the parameters $k$, the number of contacts per unit time, $b$, the probability of transmission per contact, and $D$ the mean duration of infectiousness.
-Then $\rr$ is the product of these parameters, (Equation \ref{eq:dpe}),
+This approach for parameters that are readily available to be estimated, direct parameter estimation, is a `plug and chug' approach as described by \cite{lipsitch2003} and \cite{dietz1993estimation}.
+This method has an intuitive derivation, similar to that of the survival function.
+We assume that we can estimate $\rr$ if we know how many others an individual comes into contact $(k)$, the probability of transmission given contact $(b)$, and how long the disease lasts within an individual $(D)$.
+Then $\rr$ is the product of these parameters,
 \begin{align}\label{eq:dpe}
@@
 \end{align}
-
 \cite{lipsitch2003} note that directly estimating $\rr$ in such a way may be difficult due to the available data, as in the previous two methods.
-They instead estimate $\rr$ using the mean serial interval, $\omega$, the time from onset of symptoms in an index case to the onset of symptoms in a subsequent case infected by the index patient in addition to the assumption of exponential growth of infected cases.
-
-In the event, that the is estimate may be calculated, uncertainty estimates must be compounded, and to estimate $\rr$ alone, we only use data from the initial outbreak of the disease, so often we have a small sample size, resulting in large confidence sets or credible intervals.
+They instead estimate $\rr$ using the mean serial interval, $\omega$, the time from onset of symptoms in an index case to the onset of symptoms in a subsequent case infected by the index patient in addition to the assumption of exponential growth of infected cases, as a proxy for the duration $D$.
+Advantages of this method is that it reduces the problem to three parameters that we can conceivably collect data on.
+Disadvantages include in the event, that the is estimate may be calculated, uncertainty estimates must be compounded, and to estimate $\rr$ alone, we only use data from the initial outbreak of the disease, so often we have a small sample size, resulting in large confidence sets or credible intervals.
 
@@
 A derivation of $\rr$ similar to the survival function comes from the idea of the von Foerster equations, which were used to calculate the number of cells in a ``deme,'' a constrained area.
-The function $n(t, a)$ calculated the number of cells at age $a$ at time $t$ \citep{trucco1965}.
-In epidemiology, $n$ represents the number of infections instead of cells.
+The function $Y(t, \tau)$ calculated the number of cells at age $\tau$ at time $t$ \citep{trucco1965}.
+In epidemiology, $Y$ represents the number of infections instead of cells.
 Letting $Y(t, \tau)$ be the number of infections at time $t$ of age $\tau$ and assuming no deaths in the population, the von Foerster equation is
@@
 \end{align*}
-where $Y(t,\tau)$ is the number of people at time $t$ who were infected time $\tau$ ago.    \cite{fraser2004factors} derived an expression for $\rr$ from the von Foerster equation and the equation
+where $Y(t,\tau)$ is the number of people at time $t$ who were infected time $\tau$ ago.
+\cite{fraser2004factors} derived an expression for $\rr$ from the von Foerster equation, the additional equation
 \begin{align*}
-  Y(t,0) &= \int_0^t \beta( \tau) Y(t, \tau) d \tau
+  Y(t,0) &= \int_0^t \beta( \tau) Y(t, \tau) d \tau,
 \end{align*}
-and initial conditions $Y(0,0) = Y_i$ and $Y(t, \tau) = 0$ when $t < \tau$.  Then the reproduction number is Equation \ref{eq:vfr0},
+and initial conditions $Y(0,0) = Y_0$ and $Y(t, \tau) = 0$ when $t < \tau$.  Then the reproduction number is Equation \ref{eq:vfr0},
 \begin{align}\label{eq:vfr0}
@@
 
-A large and likely the most common set of models used estimate $\mathcal{R}$ is that of compartment models (CMs).
-As adapted from \cite{daley2001epidemic}, compartment models consist of the following.
-CMs model how objects move from one discrete state to another over time and provide a high-level model of how diseases, for instance, may be spread throughout a population.
+A large and the most common set of methods used estimate $\rr$ is that of fitting compartment models (CMs) to disease data.
+As described in \cite{daley2001epidemic},  CMs model how objects move from one discrete state to another over time.
+These models provide a high-level model of how diseases may be spread throughout a population.
 Within the CM framework, we make four essential assumptions,
@@
 \item The compartments are discrete and have no overlap.
-\item The transition of objects into and out of compartments is described by a set of equations.
+\item The transition of objects into and out of compartments is described by a set of known equations, possibly dependent on unknown parameters.
 \item The populations mix homogeneously
@@
 Another common assumption for compartment models, is the law of mass action, a property borrowed from chemistry which says that the mass of the product is proportional to the mass of the reactants.
-\cite{anderson1992} adapt the law of mass action to epidemiology saying, ``the net rate of spread of infection is assumed to be proportional to the product of the susceptible people times the density of the infectious individuals.''  Therefore, the flow of objects from one compartment to another may be dependent on the percentage of objects within one or many of the compartments.
-The assumption of mass action holds particularly true for disease modelling as typically we assume infections at the next state are proportional to both the susceptible and infected populations at the current state.
+In epidemiology, this means that the proportion of new infections is proportional to the current  number of susceptible and infected \citep{anderson1992}.
+Therefore, the flow of objects from one compartment to another may be dependent on the percentage of objects within one or many of the compartments.
 
@@
 
-In this section, we discuss 14 different CMs used to estimate $\rr$.
-Different CMs yield different mathematical derivations of \rr.
+In this section, we discuss $\xxsir$ different CM-based methods used to estimate $\rr$.
+Different model specifications of CMs yield different mathematical derivations of $\rr$.
 That is, for example, an SIR model with non-constant population will yield a different mathematical formula for $\rr$ than the SEIR model (here ``E'' stands for exposed) with non-constant population.
 The difference in these two models is that the SEIR model has an extra compartment which in turn effects the estimates of transitions from one compartment to another.
-That said, even within the same compartment model framework, we can estimate $\rr$ in a variety of ways.
-In fact, we describe 10 methods to estimate $\rr$, within the SIR framework!
+That said, even within the same compartment model framework, we can estimate $\rr$ in a variety of ways using statistical techniques.
+In fact, we describe 10 methods to estimate $\rr$, within the SIR framework alone!
 Most of these methods are recipes to estimate parameters within the SIR model and may be extended to other CM frameworks such as the SEIR model.
-The final method described in this section, the next generation operator (NGM) is a general method to derive \rr from almost any CM.
-A particular advantage to CMs compared to survival function methods is that it better fits data that is typically collected about diseases: namely, incidence counts.
-CMs are very popular and still in wide-use today.
+The final method described in this section, the next generation operator (NGM) is a general method to derive $\rr$ from any valid CM.
+A particular advantage to CMs compared to survival function methods is that it easier to estimate as data that is typically collected about diseases: namely, incidence counts.
 
@@
 We begin with the most well known CM, the SIR model.
-First introduced to epidemiology by  \cite{Kermack700}, the SIR model governs how individuals transition from \textbf{S}usceptible, \textbf{I}nfected, and \textbf{R}ecovered compartments.
+First introduced to epidemiology by  \cite{Kermack700}, the SIR model governs how individuals transition from susceptible, infected, and recovered compartments.
 The model is represented graphically in Figure \ref{fig::sir}.
-In this model, once an individual is recovered it cannot be infected again.
 
@@
 In this model particular  SIR model, $N$, the number of individuals in constant, as we see no arrows pointing out of the nodes in Figure \ref{fig::sir}.
-Simple adaptations of the SIR model can include birth and death rates, which will correspondingly change the derivation of $\rr$ (for more information, see Section \ref{sec:ngm}).
+Simple adaptations of the SIR model can include birth and death rates, which may correspondingly change the derivation of $\rr$ (for more information, see Section \ref{sec:ngm}).
 The parameters  $\beta$ and $\gamma$ have epidemiological meaning.
 Here, $\beta$ is the average infection rate and $\gamma$ is the inverse of the average recovery rate.
-The model is represented through the ODEs below.
-Throughout, we use $X$, $Y$, and $Z$ for stand-ins for the ODEs for $S$, $I$, and $R$, respectively to avoid confusion.
+The transition of individuals from one compartment to another is represented through the ODEs below.
+Throughout, we use $X$, $Y$, and $Z$ for stand-ins for the ODEs for $S$, $I$, and $R$, respectively to avoid any confusion with $\rr$.
 \begin{align}
@@
   \frac{dY}{dt} &> 0 \\
-\rightarrow \frac{\beta S I}{N}  - \gamma I &> 0 ,
+ \frac{\beta S I}{N}  - \gamma I &> 0 ,
 \end{align*}
 
-that is,  the rate of new infections is greater than the rate of recovery.
-So as long as the number of susceptibles is very large, $\frac{S}{N} \approx 1$, then an epidemic will occur if $\rr >1$, as seen in Equation \ref{eq:deriv_sir},
+that is,  the rate of new infections is greater than the rate of recovery.  So as long as the number of susceptibles is large, $\frac{S}{N} \approx 1$, then an outbreak will occur if $\rr >1$,
 \begin{align}\label{eq:deriv_sir}
@@
   \end{align}
-In the ideal set up, we have data of the number of Susceptibles, Infectious, and Recovered individuals at different points over time.
+Ideally, we have access data of the number of susceptibles, infecteds, and recovered individuals at different points over time.
 From this we would like to estimate the $\beta$ and $\gamma$ parameters and ultimately $\rr$.
@@
 However, we note that the the $\epsilon_t$ are very likely not independent from one another and so valid inference is still an issue in this framework.
+Even if it were correct, ratio estimations typically result in large, uninformative confidence intervals.
 The advantage is that this is a very straightforward way to estimate $\rr$ from the SIR model.
@@
 Again, we cannot assume $\rr$ has a normal distribution as the errors are likely correlated.
-This method is more advantageous in least squares in finding the distribution of $\rr$ as ratios typically give large confidence intervals.
+This method is more advantageous in least squares in finding the distribution of $\rr$ as there is no ratio estimation.
 
@@
 In addition to this, data collected from real diseases are typically very noisy in the first place.
-\cite{chang2017} posit the idea that an SIR model may be well approximated by a linear model.
+\cite{chang2017} discover that an SIR model may be well approximated by a linear model.
 We use this idea here and use the ODEs in Equation \eqref{eq::sir} to relate our estimated linear model to that of $\rr$.
@@
 \(\hat{y}_k\)),
-
 \begin{align*}
@@
 \end{align*}
-
 Then, we estimate the derivatives as
-
 \begin{align*}
@@
 
-Then \(\rr\) is calculated as according to the ODEs,
-\[\rr = \frac{\hat{X}^\prime(0)}{ \hat{X}^\prime(0) + \hat{Y}^\prime(0)} \cdot \frac{N}{X(0)} \]
-
+Following,  \(\rr\) is derived the ODEs,
+\[\rr = \frac{\hat{X}^\prime(0)}{ \hat{X}^\prime(0) + \hat{Y}^\prime(0)} \cdot \frac{N}{X(0)}. \]
 Like in the previous cases, we are assuming the true models are
@@
 Here, $K=10$ is rather arbitrary and should be selected using some criterion such as AIC in addition with the knowledge of how many data points are available to the user.
-Besides this select of the degree of polynomials to fit, this model is simplest to implement and gives very comparable results to using least squares with the SIR model.
+Besides this selection of the degree of polynomials to fit, this model is simple to implement and gives very comparable results to using least squares with the SIR model.
 Again, we cannot use this model for inference, as our error assumption is violated.
 Another disadvantage is that we are using the first time point $t=0$ to estimate $\rr$.
-However, as this is the edge of the support of time values, it is likely be a biased estimate.
+However, as $t=0$ is the boundary of the support of time values, it is likely be a biased estimate as typically seen in linear regression.
 
@@
 The above formulation of a linear model approximation only uses the estimate at time $t=0$ to estimate $\rr$.
-We can instead, use all time points available to use.
+We can instead, use all time points available to estimate $\rr$.
 We fit a linear polynomial of \(t\) with degree \(K= 10\) model to \(X\)
@@
 An advantage to this approach is that we have a more robust estimate of $\rr$ than just using one time point.
+The disadvantages are the same as before.
 
 \subsubsection{Max of Data (SIR Max)}\label{max-of-data}
-The $I$ curve in the SIR model has one peak.    We note that \(Y^\prime(t) = \frac{\beta XY}{N} - \gamma Y > 0\) when \(\rr = \frac{\beta}{\gamma} < \frac{X(t)}{N}\). The max
+The number of infections curve in the SIR model has one peak.    We note that \(Y^\prime(t) = \frac{\beta XY}{N} - \gamma Y = 0\) when \(\rr = \frac{\beta}{\gamma} = \frac{N(t)}{X(t)}\). The max
 of \(Y\) occurs when \(Y^\prime = 0\) and hence
@@
 \(t^* = \text{arg max}_{t} \hat{Y}(t)\).
-This approach is a statistical approximation of the previous method where we fit a spline to the data.
+This approach is a statistical approximation of the previous method of using the smoothed max to estimate $\rr$ instead of a single point.
 In this way, we can utilize all our data.
@@
   Ratio (SIR IPR)}\label{incidence-to-prevalence-ratio}
-The incidence to prevalence ratio (IPR), described by \cite{Nishiura2009} is another intuitive method to calculate $\rr$ as it uses some of most basic epidemiological quantities, incidence and prevalence.
+The incidence to prevalence ratio (IPR), described by \cite{Nishiura2009} is another intuitive method to calculate $\rr$ as it incorporates some of most basic epidemiological quantities, incidence and prevalence.
 
 In terms of data from the SIR model, incidence $J(t) \approx -(X(t+1) - X(t))$, and the IPR$(t) = \frac{J(t)}{Y(t)}$.
-The actual reproduction number, $R_a(t) = \text{IPR}(t)\cdot D$ where $D = 1 /\gamma$.
+The \textit{actual} reproduction number, $R_a(t) = \text{IPR}(t)\cdot D$ where $D = 1 /\gamma$ the duration of the infection.
 This method assumes that we have some prior knowledge about $\gamma$, the transmission rate.
@@
 Here we assume that the time step is small enough to approximate the incidence.
-The advantages of this method are that incidence data is generally readily available as is prevalence data for certain diseases such as HIV.
+The advantage of this method are that incidence data is generally readily available as is prevalence data for certain diseases such as HIV.
 However, as one is required to have prior knowledge about $\gamma$, it may be easier to directly estimate $\rr$ with one of the many other methods described that does not require a prior knowledge about $\gamma$.
+Again, we are using only one point to estimate $\rr$.
 
 \subsubsection{Smoothed Incidence to Prevalence Ratio (SIR SIPR)}
-We use the same method as above, but first fit splines with 4 degrees of freedom, $\hat{X}(t)$ and $\hat{Y}(t)$, , to fit to $X_{obs}$ and $Y_{obs}$, respectively.
+We use the same method as above, IPR, but first fit splines with 4 degrees of freedom, $\hat{X}(t)$ and $\hat{Y}(t)$, , to fit to $X_{obs}$ and $Y_{obs}$, respectively.
 Then  $J(t) \approx -(\hat{X}(t+1) - \hat{X}(t))$, and the IPR$(t) = \frac{J(t)}{\hat{Y}(t)}$.
@@
 %%%%%%%%%%%%%%%%%%%%%
-\subsubsection{Harko (HSIR)}
+\subsubsection{Log-Linear (SIR LL)}
 Recently, \cite{harko2014exact} were able to reduce the SIR model, with the following set of ODEs (reparametrized from Equation \eqref{eq::sir})
@@
   \log \frac{X(t)}{X(0)} &=  \frac{b}{\gamma}Z(t) =  \frac{\beta}{\gamma N} Z \nonumber\\
-  \log \frac{X(t)}{X(0)} &=  \frac{\rr}{N} Z(t) \label{eq:harko_lin}
+  N\log \frac{X(t)}{X(0)} &=  \rr Z(t) \label{eq:harko_lin}
 \end{align}
@@
 \begin{align}
-  \log \frac{X(t)}{X(0)} &=  \frac{\rr}{N} Z  + \epsilon_t\label{eq:r0_harko}
+N  \log \frac{X(t)}{X(0)} &=  \rr Z  + \epsilon_t\label{eq:r0_harko}
 \end{align}
 
-If we assume $\epsilon_t \overset{iid}{\sim}N(0, \sigma^2)$, then we can directly estimate $\rr/N$ through linear regression and also an expression for the variance.
+If we assume $\epsilon_t \overset{iid}{\sim}N(0, \sigma^2)$, then we can directly estimate $\rr$ via linear regression and least squares.
 This method greatly simplifies the computational effort needed to estimate $\rr$ as it does not require any numerical integration.
-Another advantage is that unlike using the linear approximation method, the coefficients from this model are directly interpretable.
+Another advantage is that unlike using the linear approximation method, the coefficient from this model are directly interpretable.
+In words, $\rr$ means that for one more recovery, we expect $N \log \frac{X(t)}{X(0)}$ to increase by $\rr$.
 A disadvantage is that again, $\epsilon_t$ are  not likely iid.
+
+\subsubsection{Markov chain estimation (SIR MC)}
+A natural approach to epidemic modeling is that of Markov chains (MC), since it is assumed an individual's next state is only dependent on its current state.
+Much work has been done over the years in this specific field including asymptotic behavior, continuous time MC, confidence intervals, and more \citep{jacquez1991,gani1995,daley2001epidemic}.
+We present one the simpler instantiation of the model, the discrete time case, which traces its origin back to the Reed-Frost model \citep{abbey1952}.
+
+In this framework, the number of susceptibles at the current time, $X_{t+1}$ has a binomial distribution based on the contacts with the current number of infected, $Y_t$ and the current number of susceptibles.
+That is $X_{t+1} \sim \text{Binom}(X_t, \alpha^{Y_t}$), where $\alpha$ is the probability of avoiding infection from an infective.
+\cite{barbour2004} report that the reproduction number is thus,
+\begin{align}\label{eq:r0-mc}
+\rr &= \log \left ( \frac{1}{1-\alpha}\right ).
+\end{align}
+
+Thus, using regression on susceptible/infection counts will lead to an estimate of $\alpha$ and hence $\rr$.
+However, this framework typically allows more than just the reproduction number to be estimated.
+Through recursion, one can calculate the probability of having a given number of susceptibles and infected at each time step, and hence the entire distribution may be known.
 
@@
 
-Posited by \cite{bettencourt2008} and summarized in \cite{obadia2012r0}, this sequential Bayes is a Bayesian approach to an approximation of the classic SIR model.
+Described by \cite{bettencourt2008} and summarized in \cite{obadia2012r0}, the sequential Bayes method is a Bayesian approach to an approximation of the classic SIR model.
 The approximated SIR model assumes that the incidence at $t+1$, $J(t+1)$ has a Poisson distribution, with $\gamma$ as the  average inverse of the infectious period and $\mathcal{R}$ as the effective reproduction number, which we take here to be $\rr$.
-
-In order to estimate $\rr$, we must have some idea about the serial interval, $\omega$.
+In order to estimate $\rr$, we must have some idea about $\gamma$,
 \begin{align*}
-J(t+1)  \sim Poisson( J(t) \exp \left \{  \gamma (R-1)\right \})
+J(t+1)  \sim Poisson( J(t) \exp \left \{  \gamma (\rr-1)\right \})
 \end{align*}
@@
 The advantages of this method are that of the ability to obtain an entire posterior distribution, whereas many other methods are difficult to even find an estimate of the variance.
-Disadvantages include strict assumptions about the distributions and computational time.
+Disadvantages include strict assumptions about the distributions and computational time required.
+Additionally, this method will produce credible intervals as opposed to confidence intervals which, depending on one's perspective, may either be an advantage or disadvantage.
 
@@
 
-The SEIR model is a common adaptation, especially in recent years, of the SIR compartment model used to model infectious diseases such as influenza and Ebola\citep{mills2004,althaus2014}.
+The SEIR model is a common adaptation, especially in recent years, of the SIR compartment model used to model infectious diseases such as influenza and Ebola \citep{mills2004,althaus2014}.
 The ``E'' compartment stands for ``exposed'' and represents the stage where individuals are infected but not yet infectious.
-The model is described by the following set of equations, as described by \cite{cintronarias2009}.
+The model is described by the following set of equations, as shown by \cite{cintronarias2009}.
 This particular model here does not include birth and death rates but the model can be adapted to do so,
@@
 Then $\rr = \beta / \gamma$, just as in the SIR model.
-As in the SIR model, $\beta$ and $\gamma$ have the same interpretation and $\alpha$ is the transmission rate from the exposed compartment to infected.
-The methods used to estimate the parameters $\beta$ and $\gamma$ are of the same types used to estimate those parameters in the SIR model.
+As in the SIR model, $\beta$ and $\gamma$ have the same interpretation and additionally, $\alpha$ is the transmission rate from the exposed compartment to infected.
+The methods used to estimate the parameters $\beta$ and $\gamma$ are of the same types used to estimate those parameters in the SIR model such as least squares and maximum likelihood estimation.
 An important difference between this method and the SIR model is the assumption of a 4th class, the exposed compartment.
@@
 An advantage of this model is that the addition of an exposed class is thought to better mimic reality of the transmission of many diseases.
-A major disadvantage is that exposure counts are typically back-calculated from incidence counts, and noise is rarely considered in this estimation of the exposure counts.
+A major disadvantage is that exposure counts are typically back-estimated from incidence counts, and noise is rarely considered in this estimation of the exposure counts.
 Typically, some form of maximum likelihood estimation is used to estimate parameters in this model, which places strict assumptions on each of the compartments.
-We demonstrate this in Section \ref{sec:results}.
+
+\subsubsection{MSEIR model (MSEIR)}
+\label{sec:mseir-model-mseir}
 
 \subsubsection{Amplifier Model (AMP)}
-As mentioned previously, compartment models may become arbitrarily complex.
-We demonstrate this idea with a set of compartments for each strain of the disease.
+As mentioned previously, compartment models may be arbitrarily complex.
+We demonstrate this idea with a set of compartments for each strain of a disease like influenza or HIV.
 \cite{blower2004} discuss the case where a disease may have different strains and consequently a different reproduction number for each strain.
@@
 
-This model is based on the the SEIR compartment model, accounting for a number of strains of a disease at once.
+This model is based on the the SEIR compartment model, accounting for a number of strains simultaneously.
 There is a set of ODEs for each strain of the disease with different parameters for recovery, transmission rate, and drug resistance of the disease.
@@
 
-We show this model as opposed to other available CMs because it shows not only how complicated CMs can become, but also to demonstrate how researchers approach the often controversial assumption of homogeneity.
+We describe this model as opposed to other available CMs because it shows not only how complicated CMs can become, but also to demonstrate how researchers approach the often controversial assumption of homogeneity.
 In this case, we have strain-specific transmission rates, so we are essentially stratifying compartments into different groups, allowing for some heterogeneity.
-A disadvantage is the amount of typed data required to estimate all parameters.
+A disadvantage is the amount of typed data required to estimate all parameters, which may be estimated through the usual methods such as least squares and maximum likelihood estimation.
 
@@
 The estimations for $\rr$ from the previously mentioned compartment models may also be derived using this method.
-Additionally, this method is advantageous in that the matrix $\Sigma$ has an intuitive meaning.
+Additionally, this method is advantageous in that the matrix $\Sigma$ and $T$ have intuitive.
 
@@
 Another disadvantage, is that $\rr$ is only derived using the disease-free steady state, which is an approximation.
-Other approximations such as exponential growth are also used to derive expressions for $\rr$.
 
@@
 \subsection{Exponential Growth}\label{sec:exp-growth}
-Typically, CMs expressed through differential equations have no known analytical solutions, which makes deriving an expression for $\rr$ difficult.
+Typically, CMs expressed through differential equations have no known analytical solutions, which makes deriving a closed-form expression for $\rr$ difficult or impossible.
 One simplification that is often introduced to the SI-framework is that of exponential growth, or that the number of infections in the beginning of an epidemic mimics that of an exponential curve.
@@
 \cite{wallinga2007generation} report that the effective reproduction number $\mathcal{R}_t$ and hence the initial reproduction number $\rr$ may derived using the fact that infection ``counts increase exponentially in the initial phase of an epidemic.''  We then have to estimate $r$, the \textit{per capita} change in the number of new cases per unit of time and $\omega$ the serial interval.
-
 Then, we have
-
 \begin{align}\label{eq:lotka}
@@
 or its first order approximation
-
 \begin{align}\label{eq:anderson}
@@
 \end{align}
+Equation \eqref{eq:lotka} is derived from a demographic view using the Lotka-Euler survival equations which come from the fields of demography, ecology, and evolutionary biology, whereas Equation \eqref{eq:anderson} is derived through an epidemiologist's perspective.
+These estimates for $\rr$ obtained from Equation \eqref{eq:lotka} and Equation \eqref{eq:anderson} can be strikingly different in practice.
 
-Equation \eqref{eq:lotka} is derived from a demographic view using the Lotka-Euler survival equations which come from the fields of demography, ecology, and evolutionary biology, whereas Equation \eqref{eq:anderson} is derived through an epidemiologist's perspective.
-These estimates for $\rr$ between Equation \eqref{eq:lotka} and Equation \eqref{eq:anderson} can be strikingly different.
-
-\cite{wallinga2007generation}, instead, use a moment generating function expression for $\rr$.  With $g(a)$ as the distribution of age at infection, then
-
+\cite{wallinga2007generation}, instead, use a moment generating function expression for $\rr$.  With $\omega(t)$ as the serial interval, then
 \begin{align*}
-\rr^{-1} &= \int_{a=0}^\infty e^{-ra}g(a)da.
+\rr^{-1} &= \int_{a=0}^\infty e^{-rt}\omega(t)dt.
 \end{align*}
-
 For all these above estimations of $\rr$, we observe the duration of serial intervals in a period of exponential growth.
@@
 \begin{align}\label{eq:mlesi}
-  L(\rr, \mathbf{p}) = \prod_{t=1}^T \frac{e^{- \rr \sum_{j=1}^{\min(k,t)}N_{t-j}p_j}\left (\rr \sum_{j=1}^{\min(k,t)}N_{t-j}p_j \right )^{N_t}}{N_t!}.
+  L(\rr, \mathbf{p}) = \prod_{t=1}^T \frac{e^{- \rr \sum_{j=1}^{\min(k,t)}J_{t-j}p_j}\left (\rr \sum_{j=1}^{\min(k,t)}J_{t-j}p_j \right )^{J_t}}{J_t!}.
 \end{align}
 Here, $k$ is the maximal amount of the serial number, $T$ is the total time and $p_j$ is the probability of displaying symptoms on day $j$ after being infected.
-The number of cases on day $t$, $N_t$ are incidence counts that are known.
+The number of cases on day $t$, $J_t$ are incidence counts that are known.
 
@@
 Two disadvantages are that we only have an approximation for $\rr$ and have to decide which data points to use in a certain time frame.
+Additionally, we hae to validate our assumptions about the specified distributions.
 
@@
   \end{align*}
-
-  A disadvantage is that it assumed we have some idea about the generation interval distribution, $w(t)$, but an advantage is that there is no explicit assumption on the distribution of $\omega$.
+  A disadvantage of this method is that it assumed we have some idea about the generation interval distribution, $w(t)$, but an advantage is that there is no explicit assumption on the distribution of $\omega$.
 This method also requires that the initial number of susceptibles $X(0)$ is roughly the size of the total population.
@@
   with $D$ as the average incubation period and $t_d$ as the doubling time during its early stages.
-This method assumes you know when the ``early stage'' ends along with knowledge of the average incubation period.
+This method assumes one knows when the ``early stage'' ends along with knowledge of the average incubation period.
 
@@
 Here $\mu_0$ is the proportion of initially infected individuals and $\mu_\infty$ is the final, cumulative proportion of infected cases at the end of an epidemic,
-
 \begin{align}\label{r0_attackrate}
@@
 \end{align}
-
 This method can only be used to assess $\rr$ after a disease has passed through a population.
-Additionally, if either $\beta$ or $\gamma$ were affected during the disease's lifetime then $\rr$ cannot be properly assessed.
+Additionally, if either $\beta$ or $\gamma$ were affected during the disease's lifetime, such as through interventions and prevention strategies, then $\rr$ cannot be properly assessed.
 
@@
 Networks that have been used to estimate $\rr$ include contact tracing of actual individuals, branching processes to model the spread of disease, and agent-based models (also known as individual level models), which all leverage networks along with different sources of data to generate estimates of $\rr$.
-Additionally, networks provide an intuitive framework in which individuals can act with varying levels of heterogeneity.
+Additionally, networks provide an intuitive framework in which individuals can interact with varying levels of heterogeneity.
 
@@
 Moreover, this would implicitly be conditioning on the initial infector's contacts which may not reflect those of the rest of the population.
-We write the calculation for $\rr$ under contact tracing in Equation \ref{eq:r0_contacttracing}, where $X_0$ is a random variable denoting the number of individuals in the population person 0 infects.
+We write the calculation for $\rr$ under contact tracing in Equation \ref{eq:r0_contacttracing}, where $X_0$ is a random variable denoting the number of individuals in the population individual 0 infects.
 Here $S_0=N$ indicates that the initial number of susceptibles is $N$, the size of the population,
@@
 Again, the difficulties of this method lies in one, estimating the contacts of an initial infected individual and two, extrapolating the contacts of the initial individual to the population as a whole.
-As such, although very useful in helping eradicate a disease, is not the best choice to estimate $\rr$ due to the high level of uncertainty inherit in the method.
+As such, although vital in treating a disease in its early stages, is not the best choice to estimate $\rr$ due to the high level of uncertainty inherit in the method.
 
@@
 Branching processes are models for population growth which track generations of new offspring.
-We describe branching processes as adapted for epidemic modelling from the notation of \cite{grimmett1991}.
-An infector $i$ may produce new infections, called offspring, over the lifetime of the  infector $i$'s disease.
+We describe branching processes as adapted for epidemic modelling from the notation of \cite{grimmett1992}.
+An infector $i$ may produce new infections, called offspring, over the duration of the  infector $i$'s disease.
 The offspring of the original infector directly produces are called the first generation.
@@
 \label{sec:agent-based-models}
-Agent-based models (ABMs) or individual level models (ILMs) are ``bottom-up'' models, meaning micro patterns are simulated and macro patterns are then derived from these simulations.
+Agent-based models (ABMs) or individual level models (ILMs) are ``bottom-up'' models, meaning micro patterns are simulated and macro patterns are then inferred from these micro patterns.
 ABMs consist of agents which represent individuals (e.g., people, mosquitoes, poultry) and a series of activities in which the agents can affect one another and evolve over time.
 
-In epidemiology, ABMs can be used to simulate the spread of disease.
+In epidemiology, ABMs may be used to simulate the spread of disease.
 ABMs allow for every detail of an infection to be known, in the context of estimating $\rr$, meaning we know exactly who is infected by whom and when.
@@
 Moreover, this estimation can accommodate any model specification of how the disease is transmitted through the population.
-However, the major disadvantage of the ABM lies in calibrating the evolving of agents over time to accurately reflect reality, which is a problem that cannot be understated.
+However, the major disadvantage of the ABM lies in calibrating the evolving of agents over time to accurately reflect reality, a problem that cannot be understated.
 Thus, it is typically necessary to have knowledge of parameters such as the serial interval, transmission rate, recovery rate, etc.
@@
 
-Estimating $\rr$ is difficult, and it is no different for estimating $V[\rr]$, the variance.
+Estimating $\rr$ is difficult and estimating $V[\rr]$, the variance, is even moreso.
 We describe general methods which may be applicable to estimate the variance.
@@
 Still, reporting confidence intervals, or at the very least, standard deviations is very important and we hope to encourage this practice by describing the below methods used to estimate models in the SIR framework in a simple fashion.
-Regardless of the method used to estimate variance, confidence intervals, or credible intervals, we encourage researchers to carefully examine all underlying assumptions of 1) deriving an expression for $\rr$, 2) estimating $\rr$, and three the method used to estimate the variance or CIs.
+Regardless of the method used to estimate variance, confidence intervals, or credible intervals, we encourage researchers to carefully examine all underlying assumptions of 1) specifying a model for disease transmission, 2) estimating $\rr$ from the model, and 3) the method used to estimate the variance or CIs.
 

### 2018-07-03 paper/old_drafts/draft_v5.tex
finally on git

--- 2018-07-03 paper/old_drafts/draft_v4.tex
@@
 }
-\usetikzlibrary{arrows}
+\usetikzlibrary{arrows,automata}
 
@@
 
+\begin{figure}[h]
+\centering
+\begin{tikzpicture}[node distance=3cm,auto,>=latex',every node/.append style={align=center}]
+  \node [int,  fill = white!70!blue] (s)              {$S$};
+  \node[int, fill = white!70!blue] (m) [left of=s] {$M$};
+  \node [int,  fill = white!70!blue] (e) [right of=s]             {$E$};
+  \node [int,  fill = white!70!blue]           (i) [right of=e] {$I$};
+  \node [int,  fill = white!70!blue] (r) [right of=i] {$R$};
+  \path[->, auto=false] (s) edge node {} (e)
+  (e) edge node {} (i)
+  (i) edge node {} (r)
+  (m) edge node {} (s)
+  ++ (0,2) edge node {} (m)
+  (s) ++ (0,2) edge node {} (s);
+  \path[->]   (m) edge node {}  ++ (0,-2);
+  \path[->]   (s) edge node {}  ++ (0,-2);
+  \path[->]   (e) edge node {}  ++ (0,-2);
+  \path[->]   (i) edge node {}  ++ (0,-2);
+  \path[->]   (r) edge node {}  ++ (0,-2);
+\end{tikzpicture}
+\caption{Depiction of a MSEIR model with birth and death rates.
+$M$ - passive immunity from mother; $S$ - susceptible; $E$- exposed, latent; $I$- infected; $R$ - recovered.
+An arrow out of the system indicates death and an arrow into the system indicates birth.}\label{fig::mseir}
+\end{figure}
+
+This compartment model approach includes not only the exposed, latent compartment $E$ but additionally the compartment $M$ which stands for passive immunity inherited from the mother.
+\cite{hethcote2000} describe the model with the following set of ODES that include birth rate $b$ and death rate $d$, respectively (reparametrized to match our notation),
+\begin{align*}
+  \frac{dM}{dt} &= b(N-X) - (\delta + d) M \\
+  \frac{dX}{dt} &= bX + \delta M - \frac{\beta XY}{N} - dX\\
+  \frac{dE}{dt} &= \frac{\beta XY}{N} - (\epsilon + d) E\\ 
+  \frac{dY}{dt} &= \epsilon E - (\gamma + d) Y \\
+  \frac{dZ}{dt} &= \gamma Y - dZ \\
+  \frac{dN}{dt} &= (b-d)N,
+\end{align*}
+where $\delta$ is the rate of passive, immune infants transitioning to susceptible indnividuals, $\beta$ is the rate of infection, $\epsilon$ is the rate of of infectious individuals becoming infected, and $\gamma$ is the rate of recovery.
+The reproduction number is then,
+\begin{align}\label{eq:r0-mseir}
+\rr &= \frac{ \beta \epsilon}{ (\gamma + b)(\epsilon + b)}.
+\end{align}
+\cite{hethcote2000} note that this calculation of $\rr$ may be strictly derived through interpretation of the definition, as ``$\rr$ is the product of the contact rate $\beta$ per unit time, the average infectious period adjusted for population growth of $\frac{1}{\gamma + b}$ and the fraction $\frac{\epsilon}{\epsilon + b}$ of exposed people surviving the latent class $E$.''  They note that this model is well suited for diseases such as measles, rubella, or mumps where the mother may pass some antibodies to her newborn.
+This expression may also be derived using the next generation operator described in Section \ref{sec:ngm}.
+
+Again, this is simply a model for transmission of a disease and does not specify how $\rr$ is to be estimated once the model is written down.
+Method such as least squares and MLE are common ways to estimate the parameters in this model and hence $\rr$.
+When the birth rate $b=0$, note that $\rr$ reduces to the derived expression for both the SIR and SEIR models without birth and death rates, $\frac{\beta}{\gamma}$.
+
 \subsubsection{Amplifier Model (AMP)}

### 2018-07-03 references/_region_.tex
finally on git

\message{ !name(refs.tex)}\documentclass{article}
\usepackage[margin=1in]{geometry}
\usepackage{hyperref}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{scalerel}
\usepackage{graphicx}% http://ctan.org/pkg/graphicx
\usepackage{enumerate}
\usepackage{xcolor}
\usepackage{caption}
\usepackage{subcaption}
\usepackage{listings}

\pagenumbering{arabic}
\pagestyle{plain}
\setlength{\parindent}{0in}

\setlength{\itemsep}{0pt}
\setlength{\parsep}{0pt}

\makeatletter
\newcommand{\distas}[1]{\mathbin{\overset{#1}{\kern\z@\sim}}}%
\newsavebox{\mybox}\newsavebox{\mysim}
\newcommand{\distras}[1]{%
  \savebox{\mybox}{\hbox{\kern3pt$\scriptstyle#1$\kern3pt}}%
  \savebox{\mysim}{\hbox{$\sim$}}%
  \mathbin{\overset{#1}{\kern\z@\resizebox{\wd\mybox}{\ht\mysim}{$\sim$}}}%
}

\newcommand\rz{$R_0$}

\begin{document}

\message{ !name(refs.tex) !offset(-3) }

\bibliographystyle{plain}%Choose a bibliograhpic style

\title{ABM and CM Reference Notes}
\author{Shannon Gallagher}
\date{January 11, 2016}
\maketitle

\section{Papers}

%%%%%%%%%%%%%%%%%%%%%%
\textbf{Title:} Variance in System Dynamics and Agent Based Modelling Using the SIR Model of Infectious Disease

\textbf{Author:}  Ahmed A, Ahmed G, Aickelin, Uwe

\textbf{Citation:} \cite{ahmed2013variance}

\textbf{Link:} \url{ahmed2013variance.pdf}

\textbf{Major themes:}  SIR model for variance in ABMs

\textbf{Notes:} See r0.tex.  Read May 19, 2016.\\
\\
%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%
\textbf{Title:} Dynamically Modeling SARS and Other Newly Emerging Respiratory Illnesses: Past, Present, and Future

\textbf{Author:}  Bauch et al

\textbf{Citation:} \cite{bauch2005}

\textbf{Link:} \url{bauch2005.pdf}

\textbf{Major themes:}  r0, SARS, lit review

\textbf{Notes:} See r0.tex.  Read 6/16/2016
%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\textbf{Title:}  Comparing methods for estimating R0 from the size distribution of subcritical transimission chains

\textbf{Author:}  Blumberg and Lloyd-Smith

\textbf{Citation:}\cite{blumberg2013comparing}

\textbf{Link:} \url{blumberg2013comparing.pdf}

\textbf{Major themes:}  R0, SIR, outbreak, R0 does not represent epidemic threshold parameter.  (Individual-level model)

\textbf{Notes:} $R_0$ is the ``average number of secondary infections caused by one infected individual during his/her entire infectious period at the start of an outbreak.''

\textbf{Intro:}  Anderson and May have $R_0$ def.

$R_0$ can be found through individual-level contact tracing data or the more general Mass Action principle.  These two yield very different results.

Are ILMs like ABMs?  Why does $R_0$ change over time?
\\
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\textbf{Title:}  Theory versus Data:  How to Calculate $R_0$?

\textbf{Author:}  Breban, Vardvas, and Blower

\textbf{Citation:}\cite{breban2007}

\textbf{Link:} \url{breban2007.pdf}

\textbf{Major themes:}  R0, SIR, outbreak

\textbf{Notes:} Read 6/17/16.  See r0.tex\\
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\textbf{Title:}  On the definition and the computation of the basic reproduction ratio R0 in models for infectious diseases in heterogeneous populations

\textbf{Author:} Diekmann 1990

\textbf{Citation:}\cite{diekmann1990}

\textbf{Link:} \url{diekmann1990.pdf}

\textbf{Major themes:}  R0, heterogeneous population

\textbf{Notes:}  
\\
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\textbf{Title:}  The estimation of the basic reproduction number for infectious diseases

\textbf{Author:} Dietz 1993

\textbf{Citation:}\cite{dietz1993estimation}

\textbf{Link:} \url{dietz1993estimation.pdf}

\textbf{Major themes:}  R0, population dynamics, epidemiology, estimation

\textbf{Notes:}  See r0.tex
\\
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\textbf{Title:}  On the True Rate of Natural Increase

\textbf{Author:} Dublin, Lotka 1925

\textbf{Citation:}\cite{dublin1925}

\textbf{Link:} \url{dublin1925.pdf}

\textbf{Major themes:}  R0 origination of subscript, net fertility

\textbf{Notes:}  See r0.tex
\\
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%
\textbf{Title:} The Asymptotic Distributions of Kernel Estimators of the Mode

\textbf{Author:}  William F. Eddy

\textbf{Citation:} \cite{eddy1982asymptotic}

\textbf{Link:} \url{eddy1982asymptotic.pdf}

\textbf{Major themes:}  Kernel estimate converges to a non-stationary Gaussian random process.  The covariance is related to the smoothness of the kernel.

\textbf{Notes:}  
\textbf{Intro}  Parzen shows that the empirical kernel estimate of the mode converges weakly in distribution to a Normal distribution centered at zero.  Eddy shows that a normalized version of the kernel estimate converges weakly to a random parabola through the origin.  From this, he shows that the maximum of the normalized kernel estimator converges weakly to a Gaussian.  Eddy gives a more general result here.
\textbf{Theorem}  The main theorem says that $Z_n \overset{d}{\to} Z$ where $Z_n$ and $Z$ are random processes.  Specifically, $E[Z_n]$ converges to a parabola.
Eddy analyzes the tail behavior under different conditions.
\\
%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%
\textbf{Title:} Agent-Based Computational Models and Generative Social Science

\textbf{Author:} Joshua Epstein

\textbf{Citation:} \cite{epstein2007agent}

\textbf{Link:} \url{epstein2007agent.pdf}

\textbf{Major themes:} over view of computation based abm; first order logic; emergentism

\textbf{Notes:}

\begin{itemize}
\item no top down control, rather bottom up
\item happens in an explicit space - well posed def'n of local
\item assume bounded information and bdd computing power
\item micro agents $\rightarrow$ macro effects
\item underlying networks are key
\item agents are dynamic, interacting differently over time
\item Young 1998 for Markov process properties
\item equations vs. abm
\item Diff Eq $\iff$ abm and hybrids
\item standard for model comparison and replication of results
\item Sweeping parameter space for sensitivity analysis
\item Metrizing the rule space
\item Emergentism - large scale patterns arising from local interactions
\end{itemize}

%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\textbf{Title:}  Detail in network models of epidemiology: are we there yet?

\textbf{Author:}  Stephen Eubank et al

\textbf{Citation:}\cite{eubank2010detail}

\textbf{Link:} \url{eubank2010detail.pdf}

\textbf{Major themes:}  ABMs limit and networks

\textbf{Notes:}
\begin{itemize}
\item \textbf{Intro}  ``Network-based models for infectious disease epidemiology are natural generalizations of stochastic mass action models'' (p446)
An ABM is a network model
\item ``In appropriate limits, specifically as the number of vertices tends to infinity and time becomes continuous, the mean behavior of this model reduces to a standard Kermack and McKendrick SIR model if the graph $G$ is a complete graph and all edge labels are the same.''  (p446-447)
\item Look at the underlyinig structure of a graph to look at variability
\item references for network estimating
\item When is enough enough? p448
\end{itemize}
\textbf{Approach:}
\begin{itemize}
\item three outcomes variables:  \% infected, time to peak, peak height
\item dynamically inducdd vs model induced (p448)
\end{itemize}

\textbf{Methods:}
\begin{itemize}
\item generation a la Beckman (p449)
\item graph generation
\item smaller graph use and sigmoid assumption (p451)
\end{itemize}

ANOVA - (p454)

disease is stochastic enough (p454)
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\textbf{Title:}  Factors That Make an Infectious Disease Outbreak Controllable

\textbf{Author:}  Anderson, Roy M, Fraser, Christophe, Riley, Steven, May, Robert, Ferguson, Neil M

\textbf{Citation:}\cite{fraser2004factors}

\textbf{Link:} \url{fraser2004factors.pdf}

\textbf{Major themes:}  r0 estimation, influenza, control factors, contact tracing

\textbf{Notes:}
\\
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\textbf{Title:}  A likelihood‐based method for real‐time estimation of the serial interval and reproductive number of an epidemic

\textbf{Author:}  Forsberg and Pagano

\textbf{Citation:}\cite{forsberg2008}

\textbf{Link:} \url{forsberg2008.pdf}

\textbf{Major themes:}  r0 estimation, MLE, serial interval, generation interval

\textbf{Notes:} See r0.tex.  Read 5/11/16
\\
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\textbf{Title:}  Introduction to ``The First Draft Report on the EDVAC'' by John von Neumann

\textbf{Author:}  Michael D. Godfrey

\textbf{Citation:}\cite{von1993first}

\textbf{Link:} \url{von1993first.pdf}

\textbf{Major themes:}  Computing history

\textbf{Notes:}
\\
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%
\textbf{Title:} Pattern-Oriented Modeling of Agent Based Complex Systems

\textbf{Author:} Grimm, Volker, et al.

\textbf{Citation:} \cite{Grimm11112005}

\textbf{Link:} \url{Grimm11112005.pdf}

\textbf{Major themes:} Ecology and ABM

\textbf{Notes:}
\\
%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%
\textbf{Title:} Exact analytical solutions of the SIR Model

\textbf{Author:}  Tiberiu Harko, Francisco Lobo, MK Mak

\textbf{Citation:} \cite{harko2014exact}

\textbf{Link:} \url{harko2014exact.pdf}

\textbf{Major themes:}  Exact solutions to SIR model, death and birth rates

\textbf{Notes:}  See write up in ABM
\\
%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\textbf{Title:}  A Brief History of R0 and a Recipe for its Calculation

\textbf{Author:}  Heesterbeek 2002

\textbf{Citation:}\cite{Heesterbeek2002}

\textbf{Link:} \url{Heesterbeek2002.pdf}

\textbf{Major themes:}  r0, estimation, history of r0

\textbf{Notes:} See r0.tex
\\
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\textbf{Title:}  The Concept of R0 in Epidemic Theory

\textbf{Author:}  Heesterbeek and Dietz

\textbf{Citation:}\cite{heesterbeek1996concept}

\textbf{Link:} \url{heesterbeek1996concept.pdf}

\textbf{Major themes:}  epidemic theory, threshold behvaior, demography, structured populations

\textbf{Notes:}
\\
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\textbf{Title:}  mathemetical epidemiology of infectious diseases: model building, analysis and interpretation

\textbf{Author:}  Heesterbeek 2000

\textbf{Citation:}\cite{heesterbeek2000mathematical}

\textbf{Link:} \url{heesterbeek2000mathematical.pdf}

\textbf{Major themes:}  epidemic theory, next generation operator, r0,  structured populations

\textbf{Notes:}
\\
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\textbf{Title:}  Perspectives on the basic reproduction ratio

\textbf{Author:} Heffernan et al

\textbf{Citation:}\cite{Heffernan2005}

\textbf{Link:} \url{Heffernan2007.pdf}

\textbf{Major themes:}  R0, population dynamics, epidemiology

\textbf{Notes:}
\begin{itemize}
\item  $R_0$ can be used to indicate the magnitude of an outbreak (P1).
\item stochastic and finite systems Nasell 1995
\item great references
\end{itemize}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\textbf{Title:} The Mathematics of Infectious Diseases

\textbf{Author:}  Hethcote 2000

\textbf{Citation:}\cite{hethcote2000}

\textbf{Link:} \url{hethcote2000.pdf}

\textbf{Major themes:}  R0, SIR, epidemiology

\textbf{Notes:} See r0.tex
\\
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%
\textbf{Title:}  A Contribution to the Mathematical Theory of Epidemics

\textbf{Author:} Kermack and McKendrick

\textbf{Citation:} \cite{Kermack700}

\textbf{Link:} \url{Kermack700.pdf}

\textbf{Major themes:} Compartment Models, foundation

\textbf{Notes:}
\begin{itemize}
\item set up of the SIR model
\item Begins with discrete time steps and a derivation of the differential equations
\item Says they can't find explicit integral and spend pretty much the rest of the paper solving a similar integral
\end{itemize}
%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%
\textbf{Title:} Solutions of Ordinary Differential Equations as Limits of Pure Jump Markov Processes

\textbf{Author:}  Thomas G. Kurtz

\textbf{Citation:} \cite{kurtz_diff_eq_1970}

\textbf{Link:} \url{kurtz_diff_eq_1970.pdf}

\textbf{Major themes:}

\textbf{Notes:}
\\
%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%
\textbf{Title:} Sex ratio features of two-group SIR model for asymmetrie transmission of heterosexual disease

\textbf{Author:}  Koide 1996

\textbf{Citation:} \cite{koide1996}

\textbf{Link:} \url{koide1996.pdf}

\textbf{Major themes:}  Parallel CMs (male vs female)

\textbf{Notes:}
\\
%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%
\textbf{Title:} The analysis of equilibirum in malaria

\textbf{Author:}  Macdonald

\textbf{Citation:} \cite{macdonald1952analysis}

\textbf{Link:} Having trouble finding this paper

\textbf{Major themes:}  r0, threshold

\textbf{Notes:}
\\
%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%
\textbf{Title:} The analysis of equilibrium in malaria

\textbf{Author:}  Macdonald 1952

\textbf{Citation:} \cite{macdonald1952analysis}

\textbf{Link:} Having trouble finding this paper

\textbf{Major themes:}  r0, threshold

\textbf{Notes:}
\\
%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\textbf{Title:}  19 Dubious ways to compute the Exponential of a Matrix, 25 Years Later

\textbf{Author:} Moler and Van Loan 2003

\textbf{Citation:}\cite{moler2003}

\textbf{Link:} \url{moler2003.pdf}

\textbf{Major themes:}  19 ways to of computing $e^A$

\textbf{Notes:}  lots of ways to calculate $e^{tA}$ but none are completely satisfactory (p3)

\begin{itemize}
\item focused on generality, reliability, efficiency, storage requirements, ease of use, and simplicity
\end{itemize}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%
\textbf{Title:} Methods and Techniques of Complex Systems Science:  An Overview

\textbf{Author:} Cosma Shalizi

\textbf{Citation:} \cite{shalizi2006methods}

\textbf{Link:} \url{shalizi2006methods.pdf}

\textbf{Major themes:} overview; definitions; complex systems science; relationship with SIR models; object oriented programming; what are NOT ABMs

\textbf{Notes:}
\begin{itemize}
\item ``An agent is a thing which does things to things'' - Stuart Kauffman.  
\item An ABM consists of collection of agents and their states, rules governing interactions, and their environment
\item Relationship with SIR models/aggregate models
\item Only become interesting with more rules but balance with run time
\item ABMs \textbf{are} equation-based models, but very difficult to write down
\item ABMS should \textit{model} some phenomena 
\end{itemize}

%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\textbf{Title:}  Revisiting the Basic Reproductive Number for Malaria and Its Implications for Malaria Control

\textbf{Author:}  Smith

\textbf{Citation:}\cite{Smith2007}

\textbf{Link:} \url{Smith2007.pdf}

\textbf{Major themes:}  r0, problems with it, malaria

\textbf{Notes:}
\\
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\textbf{Title:}  A note on generation times in epidemic models

\textbf{Author:}  Svensson, A

\textbf{Citation:}\cite{Svensson2007}

\section{Books}

%%%%%%%%%%%%%%%%%%%%%%
\textbf{Title:} Epidemic Modelling:  An Introduction

\textbf{Author:}  Daley and Gani

\textbf{Citation:} \cite{daley2001epidemic}

\textbf{Major themes:} Compartment models, SIR models, discrete and continuous time, stochastic models

\textbf{Summary:}  Daley and Gani take the reader through the basics of epidemic modelling, beginning with deterministic modelling and working up to stochastic modelling.  They work out many mathematical properties of the models they introduce along the way, one of which is Markov processes.

One of the models they use in depth is the S-R model also known as the death process.  The most discussed model is the classic SIR model.  Little attention is given to larger compartment models, but many variants of the SIR model are studied.  These variants include discrete or continuous time and strata.  The first is resolved with a method similar to Euler's method, while the second is resolved with matrix notation.

Very few of these equations have explicit solutions and generally seem to be toy examples when they do.

With regards to ABM, little is said although the book does provide guidance, especially with regards to strata.  One interesting direction would be look at multiple strata models together.  This would be very close to an ABM.

\textbf{Notes:}
\textbf{Ch. 1:}  Some history of disease modeling

``Hamer (1906) first foreshadowed the simple 'mass action' principle for a deterministic epidemic model in discrete time'' P. 8

The reproductive rate is $\rho= \gamma/\beta$ and if the intial number of susceptible persons must be greater than this for the epidemic to grow.  P. 9-10.

\textbf{1.4 Stochastic Modelling}

Most apparent in small populations that deterministic models do not work.  We can think that deterministic models are predicting the mean.

First stochastic model by McKendrick in 1926 but most famous by Frost -- chain binomial model.

Markov chains come up a lot.  P. 13.

\textbf{Ch. 2 Deterministic Models}

Again iterating that we are estimating the mean when using a deterministic model.

Derivation of SIR model on P. 28 and discussion of properties in the following pages.

Theorem 2.2 (P. 37) Stratified population.  Have a transition matrix $\textbf{B}$.

Ball and Clancy (1994) - paper on moving from strata to strata

2.7 Vector borne diseases

P. 51-52 discretizing time

\textbf{Ch. 3 Stochastic Models in Continuous Time}
Self described most important part of the book.

nonnegative integer valued Markov processes P.56

A \textit{simple stochastic (Markovian) epidemic model} in continuous time is one of which the population consists only of susceptibles and infectives (P. 57).  eg AIDS.

Changing notation,

\begin{align*}
S(t) + I(t) = S(0) + I(0) \text{ (for all } t \ge 0)
\end{align*}
Time $t$ is continuous and $S$ and $I$ are nonnegative integers.  We thus have

\begin{align*}
P\{(S, I) (t + \delta t) = (i -1, j+1) | (S, I)(t)=(i,j) \} &= \beta i j \delta t + o(\delta t),\\
P\{ (S,I)(t + \delta t) = (i,j)|(S,I)(t)=(i,j) \} &= 1 - \beta ij \delta t - o(\delta t)
\end{align*}

We can thus compute the expected value at a certain time.  See P. 58.  Some difficult math follows.  Lots of Kolmogorov, and probability generating functions.

\textbf{3.5 The general stochastic epidemic in a stratified population}
You have $N$ people in $m$ strata.  There is no analytic solution known.  Can find general outbreak number.  Ex 3.5.2 (P. 92) m households.  Ex. 3.6 (carrier borne epidemic)

\textbf{4 Stochastic Models in Discrete Time}  Use of chain binomials.  4.5 graph theory

\textbf{5 Rumor Modelling}

\textbf{6 Fitting Epidemic Data}

\textbf{7 The Control of Epidemics}
The idea that if you're preventing an epidemic, then you should get better results that what you initially expect.  Three possible ways of curbing epidemics are :
\begin{itemize}
\item keeping \# of susceptibles low, eg immunization
\item accelerating rate of removal of infectives, eg medicine
\item lowering pairwise rate of infectius contact, eg isolation
\end{itemize}

%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%
\textbf{Title:} Infectious Diseases of Humans

\textbf{Author:}  Anderson R and May R

\textbf{Citation:} \cite{anderson1992}

\textbf{Major themes:}  Infectious diseases, $R_0$

\textbf{Notes:}

\textbf{Ch.1}
\begin{itemize}
\item Principle of mass action - 1906 Hamer (p7)
\item ``the net rate of spread of infection is assumed to be proportional to the product of the density of susceptible people times the density of infectious individuals''  (p7)
\item Ross did first continuous time framework
\item epidemic thresholds - ``this theory...is, in conjunction with the mass action principle, a cornerstone of modern theoretical epidemiology'' (p7)
\item due to the abstract math and little impact of the math epi advances, ``a much greater emphasis must be placed on data-oriented studies'' (p8)
\item ``In the absence of such a unified framework each infectin tends to develop its own, often, arcane literature'' (p9) need unifying framework
\item ``The relation between infection in \textit{individuals} and infection in p opulations often has aspects that the most experienced medical practicioner would find hard to intuit'' (p9)
\item Need to look at part I about microparasites, up to Ch. 13 (p11)
\end{itemize}

\textbf{Ch. 2} A framework for discussing the population biology of infectious diseases
\begin{itemize}
\item ``\textit{Microparasites} may be thought of as those parasites which have direct reproduction--usually at very high rates--within the host (Anderson and May 1976b).  THey tend to be characterized by small size and a short generation time.  Hosts that recover from infection usually acquire immunity against reinfection for some time, and often for life.'' (p13)
\item ``The essential feature of these compartmental models, however, is that little or no account is taken of the degree of severity of the infection'' (p14)
\item ``The period from the point of infection to the appearance of the symptoms is termed the \textit{incubation period}'' (p14)
\item ``The period from the point of infection to to teh beginning of the state of infectiousness is termed the \textit{latent period}'' (p14)
\item ``The sum of the average latent and anverage infectious periods is referred to as the \textit{average generation time of the infection}'' (p14)
\item ``The transmission rate, however, combines many biological, social, and environmental factors and...is thus rarely amenable to direct measurement'' (p14-15)
\item macroparasites like helminths (worms) and arthropods (bugs, spiders) (p15)
\item ``It is uncommon to find 80 per cent or more of the macroparasites contained in 20 per cent or fewer of their human hosts'' (p16).  macroparaistes are not iid!
\item Basic Reproduction rate of a parasite $R_0$ (p17)
  \begin{itemize}
  \item ``The \textbf{reproductive rate, $R_0$}, is essentially the average number of successful offspring that a parasite is intrinsically capable of producing'' (p17)
  \item ``For a microparasite (represented by a compartment model), $R_0$ is more precisely defined as the average number of secondary infections producted when one infected individual is introduced into a host population where everyone is susceptible'' (p17)
  \item ``At equilibrium, each infection will on average produce exactly one secondary infection; that is, at equilibrium the effective reproductive rate of the parasite is $R=1$'' (p17)
  \item ``If we assume the host population is homogeneously mixed...then the number of secondary infections produced by an infected individual will be linearly proportional to the probability that any one random contact is with a susceptible individual.  In this event, the effective reproductive rate, $R$, is equal to the basic rate, $R_0$, discounted by $x$, the fraction of the host population that is susceptible: $R=R_0x$'' (p17)
  \item When the population is homogeneously mixed, then at the equilibrium condition $R_0x^* = 1$ where $x^*$ is the fraction of the host population that is susceptible at equilibrium (p17)
  \item ``It is notoriously difficult to assess the intrinsic reproductive capacity, $R_0$, of any species of organism (even humans)'' (p17)  Thus $R_0$ can be computed at equilibrium
  \item  $x^*$ is a way to calculate $R_0$ (p17)
  \item  (p19) Way to estimate $R_0$:  The initial infection can be exponential (special case)
    \begin{align*}
      P(t) = P(0) \exp ( \Lambda t)
    \end{align*}
with $P(t)$ is the proportion of hosts infected, $\Lambda$ is the rate at which new infectives are being produced.  We have with $D$ as the duration of infectiousness
\begin{align*}
  \Lambda = (R_0 - 1)/D
\end{align*}
  \end{itemize}
\item Threshold host densities
  \begin{itemize}
  \item (p19) way to estimate \rz.  \rz $= N/N_T$ with $N$ the host population size and $N_T$ is a threshold.  When $N > N_T $ then $R_0  > 1$.
  \item ``the number of latent plus infected hosts at equilibirum, $H^* + Y^*$ is roughly
\begin{align*}
H^* + Y^* = (1 - 1/R_0)[ (D + D^\prime)/L] N,
\end{align*}
$D$ is the average duration of infectiousness, $D^\prime$ is the average duration of latency, $L$ is the average life expectancy of a host, and $N$ is the total number of hosts'' (p20) endemic fadeout
  \end{itemize}
\end{itemize}

\textbf{Ch. 3} Biology of host-microparasite associations
\begin{itemize}
\item ``The degree of specific population immunity, often referred to as the \textit{herd immunity}, is a major determinant of the transmission dynamics of most microparasitic organisms'' (p31)
\item ``\textit{Incidence} is defined as the \textit{rate} of appearance of new cases (of infection or disease) per upnit of time.  The second measure is the \textit{prevalence}...it records the proportion of people infected (or with symptoms of the disease) at one point in time, or over a short sampling interval'' (p39)
\item ``For most common viral and bacterial infections, incidence changes substantially with host age owing to a variety of factors'' (p44)
\item ``In practice these [latent, incubation, and infectious] periods vary for any particular infection among individuals...However, for many common viral and bacterial infections the degree of variability appears to be small relative to the average length of the period'' (p55) variability

\item Type I mortality:  ``Everyone survives to exactly age $L$ and then promptly dies'' (p62)
\item Type II mortality: ``a newly born individual's probability of surviving to age $a$ declines exponentially with $a$'' (p62)
\end{itemize}

\textbf{Ch.4} The basic model: statistics
\begin{itemize}
\item ``Thus for Type I mortality, which is a good approximation to reality in developed countries, $R_0$ is given by
\begin{align*}
R_0 = \frac{ \lambda L}{1 - e^{- \lambda L}}
\end{align*}
where $\lambda$ is the 'force of infection (p 58)' at time $t$, $L$ is the life expectancy''.  For Type II mortality, this is $R_0 = 1 + \lambda L$, assuming births = deaths (p69-71)
\item If there is population growh, ``a good approximation can be obtained by observing that the life expectancy parameter $L$, enters the above expression as the ratio $\frac{\bar{N}}{N(0)} = L$'' (p71)
\item $R_0 \approx L/A$ where $A$ is the average age of infection (p72-74)
\item After substiutions and approximations, for Type I mortality, $R_0 \approx \frac{\beta \bar{N}}{\nu}$ where $\bar{N}$ is the total host population density and $\nu$ is the per capita recovery rate (p75)
\item initial duration $M$ and latent periods estimation of $R_0$ (p80-81)
\end{itemize}

\textbf{Ch. 5} Static aspects of eradication and control
\begin{itemize}
\item ``Implementation of a vaccination or other immunization programme has two main effects.  The first and most obvious is the \textit{direct} effect whereby a fraction of the host population is removed directly into the immune class by successful immunization...The second and less obvious effect is indirect:  a smaller number of infections implies a weaker force of infection, $\lambda$.  Unimmunized individuals are indirectly protected, to a degree, by the diminution in the number of their fellows who are candidates for spreading infection'' (p87) herd immunity
\item Calculating the overall fraction of pop to be immunized.  So critical immunity proportion is $p_c = 1 - \frac{1}{R_0} \approx 1 - A/L$, where $A = \frac{1}{\lambda} \int_0^\infty a \lambda X(a) da \left ( \int_0^\infty \lambda X(a)da\right )^{-1}$ and $X(a)$ is the number of hosts of age $a$ who are susceptible (p73, 87)
\item During Type I mortality, $A = \frac{1}{\lambda} \left (\frac{1 - (1 + \lambda L)\exp(- \lambda L)}{1 - \exp ( - \lambda L)} \right ) $ (p72)
\item During Type II mortality, $A = 1/ (\lambda + \mu)$ (p73)
\item $p$ is the fraction of each age-cohort of hosts that is successfully immunized.  $\lambda^\prime$ is the force of infection. $Z$ is the immune class.  $X^\prime(a)$ is
\begin{align*}
X^\prime(a) = (1-p)N(0)\ell (a) \exp ( - \lambda^\prime a),
\end{align*}
where $\ell (a)$ is the age specific survivorship, $\ell (a) = \exp \left ( - \int_0^a \mu (s) ds \right )$. (p91)
\item Then For Type I mortality: $R_0 = \frac{ \lambda^\prime L}{(1-p)[1 - \exp ( - \lambda^\prime L)]}$ and for Type II is:  $R_0 = \frac{\lambda^\prime + \mu}{(1-p) \mu}$ (p91)
\item Age effects (p101)

\end{itemize}

\textbf{Ch. 6} The basic model: dynamics

\begin{itemize}
\item SIR model with birth and death rates (p123)
\item epidemics - ``essentially concerned with the introduction of a `seed' of infection into a largely susceptible population'' (p123)
\item endemic - ``deal with the properties of the eqns (6.5) and (6.6) at or near equilibrium'' (p123)
\item Check out Bailey 1975 (p126) for mathematical aspects
\item Table with epidemic values on P. 129
\end{itemize}

\textbf{Ch. 7} Dynamic aspects of eradication and control

k \textbf{}

%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\textbf{Title:}  Agent-based modeling: Methods and techniques for simulating human systems

\textbf{Author:}  Eric Bonabeau

\textbf{Citation:}\cite{bonabeau2002agent}

\textbf{Link:} \url{bonabeau2002agent.pdf}

\textbf{Major themes:}  ABMs

\textbf{Notes:}  Specifically, I'm looking at Ch. 2 by D. Helbing.  This is about ABMs in the social sciences but is a good literature review for many things as well as guidelines for making good ABMs.  Check out write up in 
\\
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%
\textbf{Title:} Growing Artificial Socities:  Social Science from the Bottom Up (Complex Adaptive Systems) 1st Edition

\textbf{Author:}  Epstein and Axtell (1996)

\textbf{Citation:}

\textbf{Major themes:}

\textbf{Notes:}
\\
%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%5
\textbf{Title:}  Complex Adaptive Systems:  An Introduction to Computational Models of Social Life

\textbf{Author:}  John H. Miller \& Scott E. Page (2007)

\textbf{Citation:} \cite{Grimm11112005}

\textbf{Major themes:}

\textbf{Notes:}
\\
%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%
\textbf{Title:} Advanced Data Analysis from an Elementary Point of View

\textbf{Author:}  Cosma Shalizi

\textbf{Citation:} \cite{shalizi2013advanced}

\textbf{Link:} \url{shalizi2013advanced.pdf}

\textbf{Major themes:}  data analysis

\textbf{Notes:}
\\
%%%%%%%%%%%%%%%%%%%%%%

\section{Blogs \& Notes}

%%%%%%%%%%%%%%%%%%%%%%
\textbf{Title:} Notes on $R_0$

\textbf{Author:} James Holland Jones

\textbf{Citation:} \cite{jonesjh2007}

\textbf{Link:} \url{http://web.stanford.edu/~jhj1/teachingdocs/Jones-on-R0.pdf}

\textbf{Major themes:} $R_0$

\textbf{Notes:}
\\
%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%
\textbf{Title:} Agent-Based Modeling

\textbf{Author:} Shalizi

\textbf{Citation:} \url{http://bactra.org/notebooks/agent-based-modeling.html}

\textbf{Major themes:}

\textbf{Notes:}
\\
%%%%%%%%%%%%%%%%%%%%%%

\section{Questions to Pursue}

\begin{itemize}
\item How does underlying topology effect everything?
\item What can we do with the US population?
\item Variance?
\item What sort of ABMs can we build from strata compartment models?
\item How do we fit parameters in an ABM?
\end{itemize}

\bibliography{Master}
\end{document}

%%%%%%%%%%%%%%%%%%%%%%
\textbf{Title:}

\textbf{Author:}

\textbf{Citation:}

\textbf{Major themes:}

\textbf{Notes:}
\\
%%%%%%%%%%%%%%%%%%%%%%

\begin{figure}[h]
\begin{center}
\includegraphics[width=4in]{mvhw7_3c.pdf}
\end{center}
\caption{
Biplots of the different continuous variables and their correlations.}\label{fig3c}
\end{figure}

%% two pictures whoa

\begin{figure}[h]
\centering

\begin{figure}[h]
\centering

\begin{subfigure}{.5\textwidth}
  \centering
  \includegraphics[width=1\linewidth]{mt_eda_cont_hists.pdf}
  \caption{Histograms of Arrival Delay and continuous covariates.  Arrival delay seems to have a right skewed distribution.  This may indicate that we will be transforming this variable later on.  After transforming Air Time and Distance by a log transformation, we don't really seem to have many outliers in our covariates.  We seem to have outliers in the CRS Dep. Time and Arrival Time; however, time is cyclical and so these are not, in fact outliers.}
  \label{hists}
\end{subfigure}%
\begin{subfigure}{.5\textwidth}
  \centering
  \includegraphics[width=1\linewidth]{mt_eda_cont_hists.pdf}
  \caption{\textcolor{red}{Placeholder}}
  \label{tabs1}
\end{subfigure}
\caption{}
\end{figure}

%%
\begin{figure}
\centering
\begin{subfigure}{.5\textwidth}
  \centering
  \includegraphics[width=1\linewidth]{resids_full.pdf}
  \caption{}
  \label{residsf}
\end{subfigure}%
\begin{subfigure}{.5\textwidth}
  \centering
  \includegraphics[width=1\linewidth]{diags_full.pdf}
  \caption{ }
  \label{diagsf}
\end{subfigure}
\caption{}
\end{figure}

\message{ !name(refs.tex) !offset(-936) }

### 2018-07-03 references/refs.tex
finally on git

--- 2018-07-03 references/_region_.tex
@@
-\message{ !name(refs.tex)}\documentclass{article}
+\documentclass{article}
 \usepackage[margin=1in]{geometry}
@@
 \begin{document}
-
-\message{ !name(refs.tex) !offset(-3) }
-
 \bibliographystyle{plain}%Choose a bibliograhpic style
@@
 %%%%%%%%%%%%%%%%%%%%%%
+\vskip .2cm
+\hrule
+\vskip .2cm
+\textbf{Title:} An Examination of the Reed-Frost Theory of Epidemics
+
+\textbf{Author:}  Helen Abbey
+
+\textbf{Citation:} \cite{abbey1952}
+
+\textbf{Link:} \url{abbey1952.pdf}
+
+\textbf{Major themes:}  Reed Frost theory of epidemics, r0
+
+\textbf{Notes:}
+
+\textbf{Intro:}
+\begin{itemize}
+\item ``If certain combinations of assumptions give a better fit [to the model] than others, the simplest assumptions which give a good fit provide working leads for further study of the actual relationships which have been approximated by the model'' (p201)
+\item ``The estimates of the parameters of the model may be useful in comparing different diseases or the same disease under different environmental conditions'' (p201)
+\item Reed and Frost adapted a model proposed by Soper (1927), the Reed-Frost model
+\item ``Soper had postulated a community in which all individuals had equal susceptibility to a disease, equal capacity to transmit it, and the power of passing out of observation when the transmitting period was over'' (p202)
+\item law of mass action is ``assumed to apply to the transmission of the disease'' (p202)
+\item model adapated so susceptible person can only get disease once from infected people
+\item assumptions of Reed-Frost (p202)
+  \begin{itemize}
+  \item spread through a particular kind of contact and ``in no other way''
+  \item person infected will become immune after time period
+  \item ``Each individual has a fixed probability of coming into adequate contact with any other specified individual in the group...andthis probability is the same for every member of the group'' (homogeneity (p202)
+  \item this is a closed system, during which the model is constant
+  \item ``infective period is short relative to the incubation period'' (p203)
+  \end{itemize}
+\item ``$p$ is the probability of contact between any two specified individuals'' (p202)
+\item contact == infecting a new person.
+$p$ is dependent on the susceptiblity/resistance of person, infectivity of disease, ``length of exposure and size of dose necessary to produce the disease, as well as environmental conditions necessary'' (p202-203)
+\item  The model of Soper (verbatim)  (p203)
+  \begin{itemize}
+  \item $p$ contact probability
+  \item  $q = 1-p$
+  \item $C_t$ is the number of cases produced at time $t$
+  \item $q^{C_t}$ is the probability that the specified individual will not have contact with any of the $C_t$ cases
+  \item $1-q^{C_t}$ is the probablity he will have contact with at least one
+  \item$S_t$ number of susceptibles at time $t$
+  \item $C_{t+1} = S_t (1 - q^{C_t})$ (expected number of cases)
+  \end{itemize}
+\item Reed-Frost have a probabilistic model where infection is a binomial distribution (p204-205)
+  \item assumptions of diseases to fit this model (206) 
+  \item measles, German measles and chicken pox fit these failry well
+  \item Scarlet fever, polio, diphtheria, mumps, and whooping cough do not fit this model well
+  \item ``The method which is used for estimating a single contact rate for the entire epidemic if the ``method of maximum likelihood" (p210)
+  \end{itemize}
+
+%%%%%%%%%%%%%%%%%%%%%%
+
+%%%%%%%%%%%%%%%%%%%%%%
+\vskip .2cm
+\hrule
+\vskip .2cm
 \textbf{Title:} Variance in System Dynamics and Agent Based Modelling Using the SIR Model of Infectious Disease
@@
 %%%%%%%%%%%%%%%%%%%%%%
+\vskip .2cm
+\hrule
+\vskip .2cm
 \textbf{Title:} Dynamically Modeling SARS and Other Newly Emerging Respiratory Illnesses: Past, Present, and Future
@@
 
-\textbf{Notes:} See r0.tex.  Read 6/16/2016
-%%%%%%%%%%%%%%%%%%%%%%
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\textbf{Notes:} See r0.tex.  Read 6/16/2016\\
+%%%%%%%%%%%%%%%%%%%%%%
+
+%%%%%%%%%%%%%%%%%%%%%%
+\vskip .2cm
+\hrule
+\vskip .2cm
+\textbf{Title:} Modeling the emergence of the'hot zones': tuberculosis and the amplification dynamics of drug resistance
+
+\textbf{Author:}  blower and chou
+
+\textbf{Citation:} \cite{blower2004}
+
+\textbf{Link:} \url{blower2004.pdf}
+
+\textbf{Major themes:} tuberculosis, r0, amplifier model
+
+\textbf{Notes:} 
+%%%%%%%%%%%%%%%%%%%%%%
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\vskip .2cm
+\hrule
+\vskip .2cm
 \textbf{Title:}  Comparing methods for estimating R0 from the size distribution of subcritical transimission chains
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\vskip .2cm
+\hrule
+\vskip .2cm
 \textbf{Title:}  Theory versus Data:  How to Calculate $R_0$?
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%
+\vskip .2cm
+\hrule
+\vskip .2cm
+\textbf{Title:} Dynamically Modeling SARS and Other Newly Emerging Respiratory Illnesses: Past, Present, and Future
+
+\textbf{Author:}  Cintron-Arias et al
+
+\textbf{Citation:} \cite{cintronarias2009}
+
+\textbf{Link:} \url{cintronarias2009.pdf}
+
+\textbf{Major themes:}  r0, least squares, asymptotic theories for mean and variance.
+
+\textbf{Notes:} 
+%%%%%%%%%%%%%%%%%%%%%%
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\vskip .2cm
+\hrule
+\vskip .2cm
+\textbf{Title:} Pharamacokinetic-Pharmacodynamic Modelling:  History and Perspectives
+
+\textbf{Author:} Csajka and Verotta
+
+\textbf{Citation:}\cite{csajka2006}
+
+\textbf{Link:} \url{csajka2006.pdf}
+
+\textbf{Major themes:} Compartment Models
+
+\textbf{Notes:}  
+\\
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\vskip .2cm
+\hrule
+\vskip .2cm
 \textbf{Title:}  On the definition and the computation of the basic reproduction ratio R0 in models for infectious diseases in heterogeneous populations
@@
 
-\textbf{Notes:}  
-\\
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\textbf{Notes:}  See r0.tex.  Read 6/20/16
+\\
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\vskip .2cm
+\hrule
+\vskip .2cm
 \textbf{Title:}  The estimation of the basic reproduction number for infectious diseases
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\vskip .2cm
+\hrule
+\vskip .2cm
 \textbf{Title:}  On the True Rate of Natural Increase
@@
 %%%%%%%%%%%%%%%%%%%%%%
+\vskip .2cm
+\hrule
+\vskip .2cm
 \textbf{Title:} The Asymptotic Distributions of Kernel Estimators of the Mode
@@
 %%%%%%%%%%%%%%%%%%%%%%
+\vskip .2cm
+\hrule
+\vskip .2cm
+\textbf{Title:} The Asymptotic Distributions of Kernel Estimators of the Mode
+
+\textbf{Author:}  William F. Eddy
+
+\textbf{Citation:} \cite{eddy1980}
+
+\textbf{Link:} \url{eddy1980.pdf}
+
+\textbf{Major themes:}  modes, convergence, normal distribution
+
+\textbf{Notes:}
+
+\textbf{Read:}
+\\
+%%%%%%%%%%%%%%%%%%%%%%
+
+%%%%%%%%%%%%%%%%%%%%%%
+\vskip .2cm
+\hrule
+\vskip .2cm
 \textbf{Title:} Agent-Based Computational Models and Generative Social Science
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\vskip .2cm
+\hrule
+\vskip .2cm
 \textbf{Title:}  Detail in network models of epidemiology: are we there yet?
@@
 
-disease is stochastic enough (p454)
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+disease is stochastic enough (p454)\\
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\vskip .2cm
+\hrule
+\vskip .2cm
 \textbf{Title:}  Factors That Make an Infectious Disease Outbreak Controllable
@@
 
-\textbf{Notes:}
-\\
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\textbf{Notes:}  See r0.tex.  Read 6/21/16
+\\
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\vskip .2cm
+\hrule
+\vskip .2cm
 \textbf{Title:}  A likelihood‐based method for real‐time estimation of the serial interval and reproductive number of an epidemic
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\vskip .2cm
+\hrule
+\vskip .2cm
 \textbf{Title:}  Introduction to ``The First Draft Report on the EDVAC'' by John von Neumann
@@
 %%%%%%%%%%%%%%%%%%%%%%
+\vskip .2cm
+\hrule
+\vskip .2cm
 \textbf{Title:} Pattern-Oriented Modeling of Agent Based Complex Systems
@@
 
-%%%%%%%%%%%%%%%%%%%%%%
-\textbf{Title:} Exact analytical solutions of the SIR Model
-
-\textbf{Author:}  Tiberiu Harko, Francisco Lobo, MK Mak
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\vskip .2cm
+\hrule
+\vskip .2cm
+\textbf{Title:}  Modeling of epidemic spreading with white Gaussian noise
+
+\textbf{Author:} Gu, Gao, and Li
+
+\textbf{Citation:}\cite{gu2011}
+
+\textbf{Link:} \url{gu2011.pdf}
+
+\textbf{Major themes:}  Adding noise to SIR
+
+\textbf{Notes:}
+
+\textbf{Read: }
+\\
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+%%%%%%%%%%%%%%%%%%%%%%
+\vskip .2cm
+\hrule
+\vskip .2cm
+\textbf{Title:} Exact analytical solutions of the Susceptible-Infected-Recovered (SIR) epidemic model and of the SIR model with equal death and birth rates
+
+\textbf{Author:}  Harko
 
@@
 \textbf{Notes:}  See write up in ABM
-\\
-%%%%%%%%%%%%%%%%%%%%%%
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\textbf{Read:}
+\\
+%%%%%%%%%%%%%%%%%%%%%%
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\vskip .2cm
+\hrule
+\vskip .2cm
 \textbf{Title:}  A Brief History of R0 and a Recipe for its Calculation
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\vskip .2cm
+\hrule
+\vskip .2cm
 \textbf{Title:}  The Concept of R0 in Epidemic Theory
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\vskip .2cm
+\hrule
+\vskip .2cm
 \textbf{Title:}  Perspectives on the basic reproduction ratio
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\vskip .2cm
+\hrule
+\vskip .2cm
 \textbf{Title:} The Mathematics of Infectious Diseases
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\vskip .2cm
+\hrule
+\vskip .2cm
+\textbf{Title:} Effects of quarantine in six endemic models for infectious diseases
+
+\textbf{Author:}  Hethcote 2002
+
+\textbf{Citation:}\cite{hethcote2000}
+
+\textbf{Link:} \url{hethcote2002.pdf}
+
+\textbf{Major themes:}  R0, SIR, epidemiology
+
+\textbf{Notes:}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\vskip .2cm
+\hrule
+\vskip .2cm
+\textbf{Title:} Transmission Dynamics and Control of Severe Acute Respiratory Syndrome
+
+\textbf{Author:} Lipsitch et al 2003
+
+\textbf{Citation:}\cite{lipsitch2003}
+
+\textbf{Link:} \url{lipsitch2003.pdf}
+
+\textbf{Major themes:}  R0, SIR, epidemiology
+
+\textbf{Notes:}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+%%%%%%%%%%%%%%%%%%%%%%
+\vskip .2cm
+\hrule
+\vskip .2cm
 \textbf{Title:}  A Contribution to the Mathematical Theory of Epidemics
@@
 %%%%%%%%%%%%%%%%%%%%%%
+\vskip .2cm
+\hrule
+\vskip .2cm
 \textbf{Title:} Solutions of Ordinary Differential Equations as Limits of Pure Jump Markov Processes
@@
 %%%%%%%%%%%%%%%%%%%%%%
+\vskip .2cm
+\hrule
+\vskip .2cm
 \textbf{Title:} Sex ratio features of two-group SIR model for asymmetrie transmission of heterosexual disease
@@
 %%%%%%%%%%%%%%%%%%%%%%
+\vskip .2cm
+\hrule
+\vskip .2cm
 \textbf{Title:} The analysis of equilibirum in malaria
@@
 %%%%%%%%%%%%%%%%%%%%%%
+\vskip .2cm
+\hrule
+\vskip .2cm
 \textbf{Title:} The analysis of equilibrium in malaria
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\vskip .2cm
+\hrule
+\vskip .2cm
 \textbf{Title:}  19 Dubious ways to compute the Exponential of a Matrix, 25 Years Later
@@
 %%%%%%%%%%%%%%%%%%%%%%
+\vskip .2cm
+\hrule
+\vskip .2cm
 \textbf{Title:} Methods and Techniques of Complex Systems Science:  An Overview
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\vskip .2cm
+\hrule
+\vskip .2cm
 \textbf{Title:}  Revisiting the Basic Reproductive Number for Malaria and Its Implications for Malaria Control
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\vskip .2cm
+\hrule
+\vskip .2cm
 \textbf{Title:}  A note on generation times in epidemic models
@@
 
-\textbf{Citation:}\cite{Svensson2007}
+\textbf{Citation:}\cite{Svensson2007}\\
+
+%%%%%%%%%%%%%%%%%%%%%%
+\vskip .2cm
+\hrule
+\vskip .2cm
+\textbf{Title:} A methodology to match distributions of both household and person attributes in the generation of synthetic populations
+
+\textbf{Author:}  Ye et al
+
+\textbf{Citation:} \cite{ye2009methodology}
+
+\textbf{Link:} \url{ye2009methodology.pdf}
+
+\textbf{Major themes:}  IPU
+
+\textbf{Notes:} 
+\\
+%%%%%%%%%%%%%%%%%%%%%%
 
@@
 %%%%%%%%%%%%%%%%%%%%%%
+
+%%%%%%%%%%%%%%%%%%%%%%
+\vskip .2cm
+\hrule
+\vskip .2cm
+\textbf{Title:} Random Fields and Geometry
+
+\textbf{Author:}  Adler and Taylor
+
+\textbf{Citation:} \cite{adler2007}
+
+\textbf{Link:} \url{adler2007.pdf}
+
+\textbf{Major themes:}  Gaussian processes, smoothness conditions for covariance matrix
+
+\textbf{Notes:}
+\textbf{Ch. 1}
+
+\vskip .2cm
+\hrule
+\vskip .2cm
+%%%%%%%%%%%%%%%%%%%%%%
+
 \textbf{Title:} Epidemic Modelling:  An Introduction
@@
 %%%%%%%%%%%%%%%%%%%%%%
+
+\vskip .2cm
+\hrule
+\vskip .2cm
 \textbf{Title:} Infectious Diseases of Humans
@@
 \textbf{Ch. 7} Dynamic aspects of eradication and control
-
-k \textbf{}
-
-%%%%%%%%%%%%%%%%%%%%%%
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\begin{itemize}
+\item ``the time steps used in approximating partial differential equations (7.6)-(7.9) must be chosen carefully'' (p153)
+\end{itemize}
+
+\textbf{Beyond the basic model: empirical evidence for inhomogeneous mixing}
+\begin{itemize}
+\item d
+\end{itemize}
+
+%%%%%%%%%%%%%%%%%%%%%%
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\vskip .2cm
+\hrule
+\vskip .2cm
 \textbf{Title:}  Agent-based modeling: Methods and techniques for simulating human systems
@@
 %%%%%%%%%%%%%%%%%%%%%%
+\vskip .2cm
+\hrule
+\vskip .2cm
 \textbf{Title:} Growing Artificial Socities:  Social Science from the Bottom Up (Complex Adaptive Systems) 1st Edition
@@
 %%%%%%%%%%%%%%%%%%%%%%
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%5
+\vskip .2cm
+\hrule
+\vskip .2cm
+\textbf{Title:} Wiley series in mathematical and computational biology
+
+\textbf{Author:}  Diekmann and Heesterbeek 2000
+
+\textbf{Citation:} \cite{diekmann2000}
+
+\textbf{Major themes:} r0
+
+%%%%%%%%%%%%%%%%%%%%%%
+\vskip .2cm
+\hrule
+\vskip .2cm
+\textbf{Title:} Statistical Analysis of Stationary Time Series
+
+\textbf{Author:}  Grenander and Rosenblatt
+
+\textbf{Citation:} \cite{grenander1957}
+
+\textbf{Major themes:}  time series, gaussian processes
+
+\textbf{Notes:} See scans.  Due Jan 17, 2017
+
+\textbf{Preface}
+\begin{itemize}
+\item k
+\end{itemize}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% 5
+\vskip .2cm
+\hrule
+\vskip .2cm
 \textbf{Title:}  Complex Adaptive Systems:  An Introduction to Computational Models of Social Life
@@
 \textbf{Notes:}
-\\
 %%%%%%%%%%%%%%%%%
@@
 %%%%%%%%%%%%%%%%%%%%%%
+\vskip .2cm
+\hrule
+\vskip .2cm
+\textbf{Title:} Gaussian Processes for Machine Learning
+
+\textbf{Author:}  Rasmussen and Williams 2006
+
+\textbf{Citation:} \cite{rasmussen2006}
+
+\textbf{Link:} \url{rasmussen2006.pdf}
+
+\textbf{Major themes:}  gaussian processes, covariance
+
+\textbf{Notes:}
+
+\textbf{Ch. 4 Covariance Functions}
+
+\begin{itemize}
+  \item ``The notion of \textit{similarity} between data points is crucial; it is a basic assumption that points with inputs \textbf{x} which are close are likely to have similar target values $y$'' (p79)
+  \item ``Under the Gaussian process view it is the covariance function that defines nearness or similarity'' (p79)
+  \item ``A \textit{stationary} covariance function is a function of $\mathbf{x} - \mathbf{x}^\prime$...If further the covariance is a function of $|\mathbf{x}-\mathbf{x}^\prime|$ then it is called \textit{isotropic}...These are also known as \textit{radial basis functions} (RBFs)'' (p79-80)
+  \item ``If a covariance function depends only on $\mathbf{x}$ and $\mathbf{x}^\prime$ through $\mathbf{x} \cdot \mathbf{x}^\prime$ we call it a \textit{dot product} covariance function...Dot product covariance functions are invariant to a rotation of the coordinates about the origin, but not translations'' (p80)
+  \item ``A general name for a function $k$ of two arguments mapping a pair of inputs $\mathbf{x}, \mathbf{x}^\prime \in \mathcal{X}$ into $\mathbb{R}$ is a \textit{kernel}...clearly covariance functions must be symmetric'' (p80)
+  \item ``A \textit{kernel} is said to be positive semidefinite if $\int k(\mathbf{x}, \mathbf{x}^\prime)f(\mathbf{x})f(\mathbf{x}^\prime) d\mu(\mathbf{x}) d\mu(\mathbf{x}^\prime) \ge 0,$ for all $f \in L_2(\mathcal{X}, \mu)$.'' (p80)
+  \item ``Let $\mathbf{x}_1, \mathbf{x}_2, \dots$ be a sequence of points and $\mathbf{x}_*$ be a fixed point in $\mathbb{R}^D$ such that $|\mathbf{x}_k - \mathbf{x}*| \to 0$ as $k \to \infty$.
+Then a process $f(\mathbf{x})$ is continuous in mean square at $\mathbf{x}_*$ if $E[|f(\mathbf{x}_k) - f(\mathbf{x}_*)|^2] \to 0$ as $k \to \infty$.
+If this holds for all $\mathbf{x}_*\in A$ where $A$ is a subset of $\mathbb{R}^D$ then $f(\mathbf{x})$ is said to be continuous in mean square (MS) over $A$'' (p81)
+\end{itemize}
+%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%
+\vskip .2cm
+\hrule
+\vskip .2cm
+\textbf{Title:} Epidemic MOdels: Their Structure and RElation to Data
+
+\textbf{Author:}  Mollison 1995
+
+\textbf{Citation:} \cite{mollison1995}
+
+\textbf{Major themes:} r0, epidemic models
+
+%%%%%%%%%%%%%%%%%%%%%%
+
+%%%%%%%%%%%%%%%%%%%%%%
+\vskip .2cm
+\hrule
+\vskip .2cm
 \textbf{Title:} Advanced Data Analysis from an Elementary Point of View
@@
 %%%%%%%%%%%%%%%%%%%%%%
+\vskip .2cm
+\hrule
+\vskip .2cm
 
@@
 %%%%%%%%%%%%%%%%%%%%%%
+\vskip .2cm
+\hrule
+\vskip .2cm
+
 \textbf{Title:} Notes on $R_0$
@@
 %%%%%%%%%%%%%%%%%%%%%%
+\vskip .2cm
+\hrule
+\vskip .2cm
 \textbf{Title:} Agent-Based Modeling
@@
 \end{figure}
-
-\message{ !name(refs.tex) !offset(-936) }

### 2018-07-03 ModFolder/11_NextGen.R

@@ -1,0 +1,165 @@
+# Done for the SEIR model.
+
+# Infected data should be a vector of counts that start on day 1. 
+# Assume we know starting comp sizes, have data for every evenly spaced time step.
+# start_parm should be a named vector
+# k is the number of samples to do when calculating standard deviation.
+# Returns R0 estimate and standard deviation
+seir_ng <- function(start_parm = c("beta" = 0.06, "gamma" = 0.05, "alpha" = 0.1),
+                    start_pop = c("S" = 99950, "E" = 0, "I" = 50, "R" = 0), sus_data, inf_data, 
+                    k = 100) {
+  library(EpiModel)
+  # SEIR Model
+  SEIR <- function(t, t0, parms) {
+    with(as.list(c(t0, parms)), {    
+      
+      num <- S + E + I + R
+      
+      #uses B, sigma, gamma
+      ## Differential Equations
+      dS <- -beta*S*(I)/num
+      dE <- beta*I*S/num - (alpha)*E
+      dI <- alpha*E - (gamma)*I
+      dR <- gamma*I
+      
+      ## Output
+      list(c(dS, dE, dI, dR))
+    })
+  }
+  
+  # Parameters for EpiModel
+  init <- init.dcm(S = start_pop["S"], E = start_pop["E"], 
+                   I = start_pop["I"], R = start_pop["R"])
+  cont <- control.dcm(nsteps = length(inf_data) + 1, dt = 1, new.mod = SEIR)
+  # Take prevalence data, convert to incidence for the time step. 
+  incidence_dat <- -diff(c(start_pop["S"], sus_data))
+  
+  
+  # Weighted SSE bjective function to minimize
+  obj_func <- function(data, estimate) {
+    return(sum((data - estimate)^2))
+  }
+  # Compare model to data, return weighted SSE
+  comp <- function(param, initdcm, contdcm, data) {
+    paramdcm <- param.dcm(beta = param["beta"], gamma = param["gamma"], alpha = param["alpha"])
+    model1 <- dcm(paramdcm, initdcm, contdcm)
+    inc_est <- -diff(model1$epi$S[,1])
+    return(obj_func(data, inc_est))
+  }
+  
+  
+  # Parameter estimation with OLS.
[…]

### 2018-07-03 ModFolder/12_Branching.R

@@ -1,0 +1,37 @@
+### Estimation of R0 through modeling it as a branching process.
+# Variance?
+
+# Assume data are taken periodically
+# inf_data are the prevalence counts at each equidistant time step.
+# init_pop is the starting compartment sizes
+# p is the vector of probabilities that someone at time t infects someone at
+## time t + j. Maximum length of k. [Generation time distribution]
+# times can be defined if for some reason the prob dist is not defined from 0 
+## or if there are skips in time steps
+likelihood_branching <- function(start_pop = c("S" = 99950, "I" = 50, "R" = 0), 
+                                 sus_data, inf_data, p,
+                                 times = c(0:(length(p) - 1))) {
+  prob <- matrix(c(times, p), nrow = 2, byrow = TRUE)
+  incidence_dat <- -diff(c(start_pop["S"], sus_data))
+  num <- sum(incidence_dat, na.rm = TRUE)
+  denom_vec <- rep(NA, length(incidence_dat))
+  for (ii in 1:length(denom_vec)) {
+    pnij_vec <- rep(NA, ncol(prob))
+    for (jj in 1:ncol(prob)) {
+      if (ii - prob[1, jj] > 0) {
+        pij <- prob[2, jj]
+        pnij <- pij * incidence_dat[ii - prob[1, jj]]
+        pnij_vec[jj] <- pnij
+      }
+    }
+    denom_vec[ii] <- sum(pnij_vec, na.rm = TRUE)
+  }
+  denom <- sum(denom_vec, na.rm = TRUE)
+  r0 <- num/denom
+  r0_sd <- r0 ^ 2 * 1 / sum(incidence_dat) 
+  return(list(est = r0, sd = r0_sd))
+}
+
+# dat <- read.csv("https://raw.githubusercontent.com/atzechang/datasets/master/GitData/Baseline1Norm.csv?token=AUVaoJY0HB6ltuwnAYGcjJUe15u8rzgOks5Yz1SWwA%3D%3D")[,-1]
+# p_test <- pgamma(1:14, 3, 1) / sum(pnorm(1:14, 3, 1))
+# likelihood_branching(dat[, 3], p_test, c("S" = 99950, "I" = 50, "R" = 0))

### 2018-07-03 ModFolder/13_Blower.R

@@ -1,0 +1,98 @@
+### Constant term of characteristic equation
+
+# inf_data is the infeacted prevalence
+# Starting parameters for the model: 
+## beta are transmission rates for E and I compartment
+## theta is rate of cured people who become susceptible again
+## alpha is inverse latent period
+## mu are the death rates. 0 is for S and R, e for E, i for I
+## c is cure rate of infected
+# start_pop are starting compartment sizes
+seir_blower <- function(start_parm = c("beta_e" = 0.000000006, "beta_i" = 0.000000006,
+                                                 "theta" = 0, "alpha" = 0.1,
+                                                 "mu_0" = 0.000000005, "mu_e" = 0.000000005, 
+                                                 "mu_i" = 0.000000005, "c" = 0),
+                     start_pop = c("S" = 99950, "E" = 0, "I" = 50, "R" = 0), inf_data, days) {
+  library("EpiModel")
+  
+  SEIR <- function(t, t0, parms) {
+    with(as.list(c(t0, parms)), {    
+      
+      #uses B, sigma, gamma
+      ## Differential Equations
+      dS <- -((beta_e + beta_i)* I + mu_0)*S + theta * c * I
+      dE <- - (alpha + mu_e) * E + beta_e * S * I
+      dI <- beta_i * S * I + alpha * E - (mu_i + c) * I
+      dR <- c * (1 - theta) * I - mu_0 * R
+      
+      ## Output
+      list(c(dS, dE, dI, dR))
+    })
+    
+  }
+  
+  init <- init.dcm(S = start_pop["S"], E = start_pop["E"], 
+                   I = start_pop["I"], R = start_pop["R"])
+  cont <- control.dcm(nsteps = max(days) + 1, dt = 1, new.mod = SEIR)
+  
+  obj_func <- function(data, estimate, days) {
+    return(sum((data - estimate[days])^2))
+  }
+  
+  comp <- function(param, initdcm, contdcm, data) {
+    paramdcm <- param.dcm(beta_e = param["beta_e"], beta_i = param["beta_i"],
+                          theta = param["theta"], alpha = param["alpha"],
[…]

### 2018-07-03 ModFolder/14_ExpGrowth.R

@@ -1,0 +1,41 @@
+# Find the exponential growth rate, then use MGF on the GTD
+# Variance by resampling from sampling dist of hat(r)
+
+# Assume we know starting comp sizes, have data for every evenly spaced time step.
+# w is the GTD probabilities; should have one entry for each interval
+# discrete_steps if for some reason the intervals are not defined from 0 
+## or if there are skips
+# k is number of times to resample for std. dev. estimates
+# Returns R0 estimate and standard deviation
+exp_growth <- function(start_pop = c("S" = 99950, "E" = 0, "I" = 50, "R" = 0), 
+                       sus_data, inf_data, w, discrete_steps = c(0:(length(w) - 1)),
+                       k = 100) {
+  
+  # Find incidence data
+  incidence_dat <- -diff(c(start_pop["S"], sus_data))
+  incidence_dat <- ifelse(incidence_dat <= 0, 0, incidence_dat)
+  days <- c(1:length(incidence_dat))
+  
+  # Fit Poisson regression
+  model1 <- glm(incidence_dat ~ days - 1, family = "poisson")
+    
+  # Find expected value
+  r <- coef(summary(model1))[1, 1]
+  r_se <- coef(summary(model1))[1, 2]
+
+  # MGF
+  Mt <- sum(w * exp(-r * (discrete_steps)))
+  r0 <- 1 / Mt
+  
+  dh <- -1 / (Mt ^ 2) * sum(-discrete_steps * w * exp(-discrete_steps * r))
+  
+  
+  r0_sd <- dh ^ 2 * r_se
+  
+  return(list(est = r0, sd = r0_sd))
+}
+
+# dat <- read.csv("https://raw.githubusercontent.com/atzechang/datasets/master/GitData/Baseline1Norm.csv?token=AUVaoJY0HB6ltuwnAYGcjJUe15u8rzgOks5Yz1SWwA%3D%3D")[,-1]
+# p_test <- pgamma(1:14, 3, 1) / sum(pnorm(1:14, 3, 1))
+# exp_growth(dat[, 3], c("S" = 99950, "E" = 0, "I" = 50, "R" = 0),
+#            p_test)

### 2018-07-03 ModFolder/15_MLE.R

@@ -1,0 +1,39 @@
+### MLE estimation of secondary infections
+# Variance through Fisher information 
+
+# Assumes Poisson likelihood for # of secondary infections 
+## caused by one primary infection
+# Must be during an exponential growth period
+# Assumes periodic incidence with equal interval time steps
+# w is the GTD probabilities; should have one entry for each interval
+# times if for some reason the intervals are not defined from 0 
+## or if there are skips
+# Returns R0 estimate and standard deviation
+
+mle_secondary <- function(start_pop = c("S" = 99950, "I" = 50, "R" = 0), 
+                          sus_data, inf_data, w,
+                          times = c(0:(length(w) - 1))) {
+  prob <- matrix(c(times, w), nrow = 2, byrow = TRUE)
+  incidence_dat <- -diff(c(start_pop["S"], sus_data))
+  num <- sum(incidence_dat, na.rm = TRUE)
+  denom_vec <- rep(NA, length(incidence_dat))
+  for (ii in 1:length(denom_vec)) {
+    pnij_vec <- rep(NA, ncol(prob))
+    for (jj in 1:ncol(prob)) {
+      if (ii - prob[1, jj] > 0) {
+        pij <- prob[2, jj]
+        pnij <- pij * incidence_dat[ii - prob[1, jj]]
+        pnij_vec[jj] <- pnij
+      }
+    }
+    denom_vec[ii] <- sum(pnij_vec, na.rm = TRUE)
+  }
+  denom <- sum(denom_vec)
+  r0 <- num/denom
+  r0_sd <- r0 ^ 2 * 1 / sum(incidence_dat) 
+  return(list(est = r0, sd = r0_sd))
+}
+
+# dat <- read.csv("https://raw.githubusercontent.com/atzechang/datasets/master/GitData/Baseline1Norm.csv?token=AUVaoJY0HB6ltuwnAYGcjJUe15u8rzgOks5Yz1SWwA%3D%3D")[,-1]
+# w_test <- pgamma(1:14, 3, 1) / sum(pnorm(1:14, 3, 1))
+# mle_secondary(dat[1:100, 3], w_test, c("S" = 99950, "I" = 50, "R" = 0))

### 2018-07-03 ModFolder/16_SeqBayes.R

@@ -1,0 +1,29 @@
+# Sequential Bayes Method
+
+# For now, we assume that the prior is Gamma
+# Likelihood is known Poisson
+# Posterior is then gamma (conjugate priors)
+# Need an assumption for gam parameters (avg duration of infectious period)
+# Assume data at every time step of equal interval
+# Returns R0 estimate and standard deviation
+# inv_inf is the assumed inverse of mean infectivity period
+seq_bayes_gamma <- function(start_pop = c("S" = 99950, "E" = 0, "I" = 50, "R" = 0), 
+                            sus_data, inf_data, inv_inf,
+                            gamma_prior = c("alpha" = 1, "beta" = 1)) {
+  
+  # Find incidence data
+  incidence_dat <- -diff(c(start_pop["S"], sus_data))
+  incidence_dat <- ifelse(incidence_dat <= 0.1, 0.1, incidence_dat)
+  alpha_par <- gamma_prior["alpha"]
+  beta_par <- gamma_prior["beta"]
+  for(ii in 2:length(incidence_dat)) {
+    alpha_par <- log(incidence_dat[ii] / incidence_dat[ii - 1]) / inv_inf + 1 + alpha_par
+    beta_par <- beta_par + 1
+  }
+  r0 <- alpha_par / beta_par
+  r0_sd <- sqrt(alpha_par / beta_par ^ 2)
+  return(list(est = r0, sd = r0_sd))
+}
+
+# dat <- read.csv("https://raw.githubusercontent.com/atzechang/datasets/master/GitData/Baseline1Norm.csv?token=AUVaoJY0HB6ltuwnAYGcjJUe15u8rzgOks5Yz1SWwA%3D%3D")[,-1]
+# seq_bayes_gamma(dat[, 3], c("S" = 99950, "E" = 0, "I" = 50, "R" = 0), inv_inf = 1/7)

### 2018-07-03 ModFolder/17_TimeDep.R

@@ -1,0 +1,50 @@
+# Time dependent reproduction number
+
+# w is the GTD probabilities; should have one entry for each interval
+# Assume we know starting comp sizes, have data for every evenly spaced time step.
+# Assumes everyone has equal prob of getting infected at time i from case at time j.
+# Assumes transmission takes at least one time step.
+td_rep <- function(start_pop = c("S" = 99950, "E" = 0, "I" = 50, "R" = 0),
+                   sus_data, inf_data, w) {
+  
+  # Find incidence data
+  incidence_dat <- -diff(c(start_pop["S"], sus_data))
+  incidence_dat <- ifelse(incidence_dat <= 0, 0, incidence_dat)
+  max_days <- length(w)
+  days <- c(1:length(w))
+  
+  # Find R0
+  # Sum over all pi0.
+  # Assume GTD is lower bounded by 0.
+  pij <- rep(NA, length(days))
+  
+  # Every time of onset
+  for(ii in (2:(max_days + 1))) {
+    if (round(incidence_dat[ii]) == 0) {
+      next
+    }
+    plj <- rep(NA, round(incidence_dat[ii]))
+    
+    # Every person at that time
+    for(ll in (1:round(incidence_dat[ii]))){
+      nom <- round(incidence_dat[ii]) * w[ii] 
+      gtd <- rep(0, max_days)
+      gtd[1:ii] <- w[ii:1]
+      denom <- 0
+      for (kk in 1:(ii-1)) {
+        denom <- denom + incidence_dat[ii - kk] * gtd[kk]
+      }
+      plj[ll] <- nom / denom
+    }
+    plj <- sum(plj, na.rm = TRUE)
+    pij[ii] <- plj
+  }
+  r0 <- sum(pij, na.rm = TRUE) / max(1, incidence_dat[1])
+  
+  return(list(est = r0))
+}
+
+# dat <- read.csv("https://raw.githubusercontent.com/atzechang/datasets/master/GitData/Baseline1Norm.csv?token=AUVaoJY0HB6ltuwnAYGcjJUe15u8rzgOks5Yz1SWwA%3D%3D")[,-1]
+# w_test <- c(rep(0, 5), pgamma(1:14, 3, 1)) / sum(pnorm(1:14, 3, 1))
+# td_rep(dat[, 3], c("S" = 99950, "E" = 0, "I" = 50, "R" = 0),
+#        w_test)

### 2018-07-03 ModFolder/18_InitFinal.R

@@ -1,0 +1,44 @@
+### Two estimation methods
+
+# Initial growth rate and duration of infection
+## Variance through ???
+
+init_growth <- function(start_pop = c("S" = 99950, "E" = 0, "I" = 50, "R" = 0), dur, 
+                        sus_data, inf_data, k = 100){
+  # Find incidence data
+  incidence_dat <- -diff(c(start_pop["S"], sus_data))
+  incidence_dat <- ifelse(incidence_dat <= 0, 0, incidence_dat)
+  
+
+  # Divide incidence at each time point by starting incidence
+  days <- c(1:length(incidence_dat))
+  
+  # Fit Poisson regression to find exp growth rate
+  model1 <- glm(incidence_dat ~ days - 1, family = "poisson")
+  r <- coef(summary(model1))[1, 1]
+  r_se <- coef(summary(model1))[1, 2]
+
+  # Find doubling time
+  doub_time <- log(2) / log(1 + r)
+  # Find r0
+  r0 <- 1 + dur * log(2) / doub_time
+  
+  dh <- 1 + dur / (1 + r)
+  # Calculate variance
+  r0_sd <- dh ^ 2 * r_se
+  return(list(est = r0, sd = r0_sd))
+}
+
+# Proportion of susceptibles after epidemic ends
+final_size <- function(mu_inf){
+  r0 <- (1 - mu_inf) ^ (-1) * log(mu_inf ^ (-1))
+  dh <- (mu_inf + mu_inf * log(1 / mu_inf) - 1) / 
+    ((1 - mu_inf) ^ 2 * mu_inf)
+  r0_sd <- sqrt(dh ^ 2 * (mu_inf * (1 - mu_inf) / sum(start_pop)))
+  return(list(est = r0, sd = r0_sd))
+}
+
+#Test
+# dat <- read.csv("https://raw.githubusercontent.com/atzechang/datasets/master/GitData/Baseline1Norm.csv?token=AUVaoJY0HB6ltuwnAYGcjJUe15u8rzgOks5Yz1SWwA%3D%3D")[,-1]
+# init_growth(14, dat[, 3], 100, c("S" = 99950, "E" = 0, "I" = 50, "R" = 0))
+# final_size(0.95)

### 2018-07-03 ModFolder/1_ConTrace.R

@@ -1,0 +1,1 @@
+#????????????????

### 2018-07-03 ModFolder/2_Survival.R

@@ -1,0 +1,37 @@
+### Integration of survival function
+
+# bxf is a function that returns the density estimate of the 
+## survival function at each time.
+# Known distributions for b,f need to be multiplied together beforehand
+# b is average number of secondaries per primary at time t after infection
+# f is function for probability of survival until time t after infection
+# max is the maximum length to measure
+cont_surv_func <- function(bxf, max) {
+  r0 <- integrate(bxf, 0, max, subdivisions = 10000)$value
+  return(list(est = r0))
+}
+
+# For discrete distributions.
+# b is average number of secondaries per primary at time t
+# f is function for probability of survival until time t
+disc_surv_func <- function(b, f) {
+  r0 <- sum(b * f)
+  return(list(est = r0))
+}
+
+# Variance: resampling?
+# Figure out sampling distributions?
+# Would probably need to be created for each separate instance
+# 
+# 
+# # Average number of infections is bell-shaped with peak at 5.
+# # Infection time survival is exponential distribution.
+# bxf_test <- function(t) {return(dexp(t / 2) * dnorm(t, 5, 3) * 10)}
+# cont_surv_func(bxf_test, 100)
+# 
+# # Discretized test
+# # Number of people infected at every time is randomly normally distributed
+# # Infection time survival is discretized exponential distribution.
+# b_dtest <- rnorm(11, 0.6, 0.1)
+# f_dtest <- dexp(0:10)
+# disc_surv_func(b_dtest, f_dtest)

### 2018-07-03 ModFolder/3_Direct.R

@@ -1,0 +1,32 @@
+### Direct parameter estimation 
+
+# Find R from an equation involving the serial interval and mean latent period
+## Serial interval = latent + infection period
+## Assumes exponential rate of growth and known serial interval/mean latent period
+## v is the mean latent period + mean infectious period
+## l is the mean latent period
+## Only works if data starts from single first case
+exp_dpe <- function(cases, prop_suscept, t, v, l) {
+  lambda = log(cases - 1) / t
+  f = l/v
+  R = 1 + lambda * v + f * (1 - f) * (lambda * v) ^ 2
+  r0 = R / prop_suscept
+  
+  # The variance estimation makes me want to cry
+  
+  return(list(est = r0))
+}
+
+# This other direct parameter estimation thing
+## k is number of contacts per unit of time
+## b is the probability of transmission between a contact of I and S
+## D = mean duration of infectiousness
+plug_chug <- function(k, b, D) {return(list(est = k * b * D))}
+
+# l_test <- 7
+# v_test <- 14
+# dat <- read.csv("https://raw.githubusercontent.com/atzechang/datasets/master/GitData/Baseline1Norm.csv?token=AUVaoJY0HB6ltuwnAYGcjJUe15u8rzgOks5Yz1SWwA%3D%3D")[,-1]
+# dat[,2:4] <- dat[,2:4] / 50
+# exp_dpe(dat[100,"I"], dat[100,"S"] / 2000, 100, v_test, l_test)
+
+# plug_chug(2, 0.1, 7)

### 2018-07-03 ModFolder/4_SurvFunc.R

@@ -1,0 +1,29 @@
+### Directly estimated survival function
+# Variance?
+
+# start_pop is the starting compartment sizes
+# inf_dens should be the infectivity of an individual at the range of 
+## possible infectivity times. 
+discrete_dsf <- function(start_pop = c("S" = 99950, "E" = 0, "I" = 50, "R" = 0),
+                             inf_dens) {
+  sus_prop <- start_pop["S"] / sum(start_pop)
+  r0 <- sus_prop * sum(inf_dens)
+  return(list(est = r0))
+}
+
+cont_dsf <- function(start_pop = c("S" = 99950, "E" = 0, "I" = 50, "R" = 0),
+                     inf_dens) {
+  sus_prop <- start_pop["S"] / sum(start_pop)
+  r0 <- sus_prop * integrate(inf_dens, 0, Inf)$value
+  return(list(est = r0))
+}
+
+# # Continuous test
+# # Infectiousness is exponentially distributed.
+# id_test <- function(t) {return(dexp(t / 1.5))}
+# cont_dsf(c("S" = 9900, "I" = 100), id_test)
+# 
+# # Discretized test
+# # Infectiousness is beta distributed
+# id_test <- rbeta(11, 0.2, 1)
+# discrete_dsf(c("S" = 9900, "I" = 100), id_test)

### 2018-07-03 ModFolder/5-2_SIRrepar.R

@@ -1,0 +1,145 @@
+#Load package
+library("EpiModel")
+library(deSolve)
+
+#Remember to change desired directory 
+
+#####Arguments:
+#init.param is a vector of the initial guess of the beta and gamma parameters. It needs to be a vector that has elements
+###named "r0" and "gamma". All other elements will be ignored. 
+#init.pop is a vector of the initial size of each compartment. It needs to be a vector that has elements named "S", "I", and 
+###"R". All other elements will be ignored. 
+#days is a vector of the day numbers of the data points. This allows for data without regular intervals between points.
+###Must be a vector of numbers. 
+#inf.data is the number of people in I for each day in the data. Must be a vector of numbers, same length as days.
+#sus.data is the number of people in S for each day in the data. Must be a vector of numbers, same length as days.
+#ndeps is for optim calculating the hessian
+
+SIRrepar <- function(init.param, init.pop, days, inf.data, sus.data, ndeps = 0.00001) {
+  #Get initial populations of each compartment S, I, R.
+  init.sus <- init.pop["S"]
+  init.inf <- init.pop["I"]
+  init.rem <- init.pop["R"]
+  
+  #Get total number of people in population
+  #Assume population stays constant except for deaths by disease
+  N <-sum(init.pop)
+  
+  #Function for EpiModel for solving for first iteration with the linearization adjustment.
+  
+  #Regular SIR
+  SIR <- function(t, t0, parms) {
+    with(as.list(c(t0, parms)), {    
+      num <- S + I + R
+      ## Differential Equations
+      dS <- -R0*gamma*S*I/num
+      dI <- R0*gamma*S*I/num - gamma*I 
+      dR <- gamma*I
+      ## Output
+      list(c(dS, dI, dR), 
+           num = num,
+           si.flow = R0*gamma*S*I/num,
+           pr = I+R)
+    })
+  }
+  
+  
+  #Find the SSE between the real data and the simulation from the model. 
+  opt <- function(inf.model,inf.data, sus.model, sus.data, days) {
+    err<- sum((inf.model[days] - inf.data)^2 + (sus.model[days] - sus.data)^2)
+    return(err)
+  }
+  
+  
[…]

### 2018-07-03 ModFolder/5-3_SIRlma.R

@@ -1,0 +1,94 @@
+#Load package
+library("EpiModel")
+library(deSolve)
+
+#Remember to change desired directory 
+
+#####Arguments:
+#days is a vector of the day numbers of the data points. This allows for data without regular intervals between points.
+###Must be a vector of numbers. 
+#init.pop is a vector of the initial size of each compartment. It needs to be a vector that has elements named "S", "I", and 
+#inf.data is the number of people in I for each day in the data. Must be a vector of numbers, same length as days.
+#sus.data is the number of people in S for each day in the data. Must be a vector of numbers, same length as days.
+#k is the number of degress for the polynomial regression.
+
+poly_deriv <- function(t, mod){
+  deriv_coefs <- as.numeric(mod$coef[-1])
+  deriv_coefs <- ifelse(is.na(deriv_coefs), 0, deriv_coefs)
+  k <- length(deriv_coefs)
+  if( k == 1 ) {return(deriv_coefs)}
+  else if ( k < 1){ return(0)}
+  poly_mat <- cbind(rep(1, length(t)), poly(t, k-1, raw = TRUE))
+  deriv_val <- poly_mat %*% (deriv_coefs * 1:k)
+  return(deriv_val)
+}
+
+SIRlma <- function(init.pop, days, inf.data, sus.data, k = 10) {
+  #Get initial populations of each compartment S, I, R.
+  init.sus <- init.pop["S"]
+  init.inf <- init.pop["I"]
+  init.rem <- init.pop["R"]
+  inf.data <- c(init.inf, inf.data)
+  sus.data <- c(init.sus, sus.data)
+  days <- c(0, days)
+  df <- data.frame(days = days, inf.data = inf.data, sus.data = sus.data)
+  
+  #Get total number of people in population
+  #Assume population stays constant except for deaths by disease
+  N <-sum(init.pop)
+  n <- N
+
+  y_mod <- lm(inf.data ~ poly(days, k, raw = TRUE), data = df)
+  y_hat <- predict(y_mod, df)
+  x_mod <- lm(sus.data ~ poly(days, k, raw = TRUE), data = df)
+  x_hat <- predict(x_mod, df)
[…]

### 2018-07-03 ModFolder/5-4_SIRmax.R

@@ -1,0 +1,38 @@
+#init.pop is a vector of the initial size of each compartment. It needs to be a vector that has elements named "S", "I", and "R"
+#days is a vector of the day numbers of the data points. This allows for data without regular intervals between points.
+###Must be a vector of numbers. 
+#inf.data is the number of people in I for each day in the data. Must be a vector of numbers, same length as days.
+#sus.data is the number of people in S for each day in the data. Must be a vector of numbers, same length as days.
+
+SIRmax <- function(init.pop, days, inf.data, sus.data) {
+  N <- sum(init.pop)
+  tstar <- which.max(inf.data)
+  r0_est <- N / sus.data[tstar]
+  return(list(est = r0_est))
+}
+
+SIRsmax <- function(init.pop, days, inf.data, sus.data) {
+  sus.data <- c(init.pop["S"], sus.data)
+  inf.data <- c(init.pop["I"], inf.data)
+  days <- c(0, days)
+  N <- sum(init.pop)
+  modx <- smooth.spline(days, sus.data, df = 4)
+  xhat <- predict(modx, seq(0, max(days), by = .05))
+  mody <- smooth.spline(days, inf.data, df = 4)
+  yhat <- predict(mody, seq(0, max(days), by = .05))
+  tstar <- which.max(yhat$y)
+  r0_est <- N / xhat$y[tstar]
+  
+  # Jackknife
+  r0 <- numeric(length(days))
+  for(ii in 1:length(r0)){
+    modx <- smooth.spline(days[-ii], sus.data[-ii], df = 4)
+    xhat <- predict(modx, seq(0, max(days), by = .05))
+    mody <- smooth.spline(days[-ii], inf.data[-ii], df = 4)
+    yhat <- predict(mody, seq(0, max(days), by = .05))
+    tstar <- which.max(yhat$y)
+    r0[ii] <- N / xhat$y[tstar]
+  }
+  r0_sd <- sqrt((N-1) / N * sum((r0 - mean(r0))^2))
+  return(list(est = r0_est, sd = r0_sd))
+}

### 2018-07-03 ModFolder/5-5_SIRipr.R

@@ -1,0 +1,48 @@
+#init.pop is a vector of the initial size of each compartment. It needs to be a vector that has elements named "S", "I", and "R"
+#days is a vector of the day numbers of the data points. This allows for data without regular intervals between points.
+###Must be a vector of numbers. 
+#inf.data is the number of people in I for each day in the data. Must be a vector of numbers, same length as days.
+#sus.data is the number of people in S for each day in the data. Must be a vector of numbers, same length as days.
+#gamma is 1/infectivity period.
+
+set.seed(1)
+SIRipr <- function(init.pop, days, inf.data, sus.data, gam) {
+  sus.data <- c(init.pop["S"], sus.data)
+  inf.data <- c(init.pop["I"], inf.data)
+  days <- c(0, days)
+  n <- sum(init.pop)
+  J <- -diff(sus.data, lag = 1) #incidence
+  IPR <- J / inf.data[- length(days)]
+  IPR <- ifelse(IPR < 0, 0, IPR)
+  r0_est <- mean(IPR / gam)
+  r0_sd <- sd(IPR / gam)
+  return(list(est = r0_est, sd = r0_sd))
+}
+
+SIRsipr <- function(init.pop, days, inf.data, sus.data, gam) {
+  sus.data <- c(init.pop["S"], sus.data)
+  inf.data <- c(init.pop["I"], inf.data)
+  days <- c(0, days)
+  n <- sum(init.pop)
+  modx <- smooth.spline(days, sus.data, df = 4)
+  xhat <- predict(modx, seq(0, max(days), by = 1))
+  mody <- smooth.spline(days, inf.data, df = 4)
+  yhat <- predict(mody, seq(0, max(days), by = 1))
+  J <- - diff(xhat$y, lag=1) #incidence
+  IPR <- J / yhat$y[- length(days)]
+  r0_est <- mean(IPR / gam)
+  
+  # Jackknife
+  r0 <- numeric(length(days))
+  for(ii in 1:length(r0)){
+    modx <- smooth.spline(days[-ii], sus.data[-ii], df = 4)
+    xhat <- predict(modx, seq(0, max(days), by = 1))
+    mody <- smooth.spline(days[-ii], inf.data[-ii], df = 4)
+    yhat <- predict(mody, seq(0, max(days), by = 1))
+    J <- -diff(xhat$y, lag = 1) #incidence
[…]

### 2018-07-03 ModFolder/5-6_SIRmc.R

@@ -1,0 +1,16 @@
+
+
+SIRmc <- function(init.pop, inf.data, sus.data, k = 100) {
+  N <- sum(init.pop)
+  sus.data <- c(init.pop["S"], sus.data)
+  inf.data <- c(init.pop["I"], inf.data)
+  resp <- (log(sus.data[2:(length(sus.data))]) - log(sus.data[1:(length(sus.data) - 1)]))
+  pred <- inf.data[1:(length(sus.data) - 1)] / N
+  reg <- lm(resp ~ pred - 1)
+  alpha <- exp(coef(summary(reg))[1, 1])
+  r0_est <- log(1 / (1 - alpha))
+  alpha_se <- (coef(summary(reg))[1, 2])
+  
+  r0_sd <- (1 / (1 - alpha)) ^ 2 * alpha_se
+  return(list(est = r0_est, sd = r0_sd))
+}

### 2018-07-03 ModFolder/5_SIR.R

@@ -1,0 +1,160 @@
+#Load package
+library("EpiModel")
+library(deSolve)
+
+#Remember to change desired directory 
+
+#####Arguments:
+#init.param is a vector of the initial guess of the beta and gamma parameters. It needs to be a vector that has elements
+###named "beta" and "gamma". All other elements will be ignored. 
+#init.pop is a vector of the initial size of each compartment. It needs to be a vector that has elements named "S", "I", and 
+###"R". All other elements will be ignored. 
+#days is a vector of the day numbers of the data points. This allows for data without regular intervals between points.
+###Must be a vector of numbers. 
+#inf.data is the number of people in I for each day in the data. Must be a vector of numbers, same length as days.
+#sus.data is the number of people in S for each day in the data. Must be a vector of numbers, same length as days.
+
+
+SIRreg <- function(init.param, init.pop, days, inf.data, sus.data, ndeps = 0.00001) {
+  #Get initial populations of each compartment S, I, R.
+  init.sus <- init.pop["S"]
+  init.inf <- init.pop["I"]
+  init.rem <- init.pop["R"]
+  
+  #Get total number of people in population
+  #Assume population stays constant except for deaths by disease
+  N <-sum(init.pop)
+  
+  #Function for EpiModel for solving for first iteration with the linearization adjustment.
+  
+  #Regular SIR
+  SIR <- function(t, t0, parms) {
+    with(as.list(c(t0, parms)), {    
+      num <- S + I + R
+      ## Differential Equations
+      dS <- -B*S*I/num
+      dI <- B*S*I/num - gamma*I 
+      dR <- gamma*I
+      ## Output
+      list(c(dS, dI, dR), 
+           num = num,
+           si.flow = B*S*I/num,
+           pr = I+R)
+    })
+  }
+  
+  
+  #Find the SSE between the real data and the simulation from the model. 
+  opt <- function(inf.model,inf.data, sus.model, sus.data, days) {
+    err<- sum((inf.model[days] - inf.data)^2 + (sus.model[days] - sus.data)^2)
+    return(err)
+  }
+  
+  
[…]

### 2018-07-03 ModFolder/6_Harko.R

@@ -1,0 +1,61 @@
+### SIR through Harko's 
+
+# start_param is the intial parameter guess for SIR model
+# start_pop is the initial compartment populations
+# inf.data is the infected prevalence data
+# rem.data is the removed prevalence data
+harkoSIR <- function(start_param = c("beta" = 0.06, "gamma" = 0.05),
+                             start_pop = c("S" = 99950, "I" = 50, "R" = 0), 
+                             sus_data, inf_data, rem_data) {
+  tot_pop <- sum(start_pop)
+  init_pop <- start_pop / tot_pop
+  rem_data <- rem_data / tot_pop
+  inf_data <- inf_data / tot_pop
+  sus_data <- sus_data / tot_pop
+#   #Find the SSE between the real data and the simulation from the model. 
+#   opt <- function(inf.model,inf.data) {
+#     err<- sum((inf.model - inf.data)^2)
+#     return(err)
+#   }
+#   
+#   harko_sir <- function(params, inf.data, rem.data, init) {
+#     beta <- params["beta"]
+#     gamma <- params["gamma"]
+#     sus <- init["S"] * exp(beta / gamma * init["R"]) * exp(- beta / gamma * rem.data)
+#     inf <- sum(init) - rem.data - sus
+#     return(inf)
+#   }
+#   
+#   #Function to be put in the function which finds the minimum SSE.
+#   model.sir <- function(params,initialpop, inf.data, rem.data) {
+#     inf.model <- harko_sir(params, inf.data, rem.data, initialpop)
+#     sse <- opt(inf.model,inf.data)
+#     return(sse)
+#   }
+#   
+#   
+#   #Function to minimize the SSE between the data and the model. 
+#   sir.optim <- function(parm, initpop, inf.data, rem.data) {
+#     par.sir <- optim(parm, model.sir, initialpop = initpop,
+#                      inf.data = inf.data,
+#                      rem.data = rem.data, hessian = T)
+#     return(c(par.sir))
+#   }
+#   
+#   #Solving as regular SIR.
+#   regularSIR <- matrix(ncol = length(init.param))
+#   colnames(regularSIR) <- c("beta","gamma")
+#   model_est <- sir.optim(init.param, init.pop, inf.data, rem.data)
[…]

### 2018-07-03 ModFolder/7_SEIR.R

@@ -1,0 +1,174 @@
+### SEIR Model with GLS
+
+# Need to source files
+#setwd("~/R0/Models")
+
+# Infected data should be a vector of counts that start on day 1. 
+# rho is a parameter for the weights for the GLS algorithm
+# Assume we know starting comp sizes, have data for every evenly spaced time step.
+# start_parm should be a named vector
+# k is the number of samples to do when calculating standard deviation.
+# stop_par is the stopping threshhold for the algorithm
+# Returns R0 estimate and standard deviation
+seir_gls <- function(start_parm = c("beta" = 0.06, "gamma" = 0.05, "alpha" = 0.1),
+                     start_pop = c("S" = 99950, "E" = 0, "I" = 50, "R" = 0), sus_data, inf_data,
+                     rho = 1, stop_par = 0.0001, k = 100, ndeps = 0.0000001) {
+  library(EpiModel)
+  # SEIR Model
+  SEIR <- function(t, t0, parms) {
+    with(as.list(c(t0, parms)), {    
+      
+      num <- S + E + I + R
+      
+      #uses B, sigma, gamma
+      ## Differential Equations
+      dS <- -beta*S*(I)/num
+      dE <- beta*I*S/num - (alpha)*E
+      dI <- alpha*E - (gamma)*I
+      dR <- gamma*I
+      
+      ## Output
+      list(c(dS, dE, dI, dR))
+    })
+  }
+  
+  # Parameters for EpiModel
+  init <- init.dcm(S = start_pop["S"], E = start_pop["E"], 
+                   I = start_pop["I"], R = start_pop["R"])
+  cont <- control.dcm(nsteps = length(inf_data) + 1, dt = 1, new.mod = SEIR)
+  # Take prevalence data, convert to incidence for the time step. 
+  incidence_dat <- -diff(c(start_pop["S"], sus_data))
+  
+  
+  # Weighted SSE bjective function to minimize
+  obj_func <- function(weights, data, estimate) {
+    return(sum(weights * (data - estimate)^2))
+  }
+  
+  # Compare model to data, return weighted SSE
+  comp <- function(param, initdcm, contdcm, weights, data) {
+    paramdcm <- param.dcm(beta = param["beta"], gamma = param["gamma"], alpha = param["alpha"])
+    model1 <- dcm(paramdcm, initdcm, contdcm)
[…]

### 2018-07-03 ModFolder/8_AR.R

@@ -1,0 +1,20 @@
+### Attack rate
+
+# Attack rate is the proportion of people eventually infected
+# start_pop is the starting compartment sizes. 
+# sus_prop is the initial proportion of susceptible
+R0_attack_rate <- function(AR, start_pop) {
+  sus_prop <- start_pop["S"] / sum(start_pop)
+  r0 <- -log((1 - AR) / sus_prop) / (AR - (1 - sus_prop))
+  dh <- (log((1 - AR) / sus_prop) - (AR + sus_prop - 1)/(AR - 1)) /
+    (AR + sus_prop - 1) ^ 2
+  r0_sd <- sqrt(dh ^ 2 * (AR * (1 - AR) / sum(start_pop)))
+  return(list(est = r0, sd = r0_sd))
+}
+
+# Does this even have a model variance? 
+
+# dat <- read.csv("https://raw.githubusercontent.com/atzechang/datasets/master/GitData/Baseline1Norm.csv?token=AUVaoJY0HB6ltuwnAYGcjJUe15u8rzgOks5Yz1SWwA%3D%3D")[,-1]
+# sus_prop_test <- 99950 / 100000
+# ar_test <- (99950 - 0 - dat[nrow(dat), 2]) / sum(dat[nrow(dat), -1])
+# R0_attack_rate(ar_test, sus_prop_test)

### 2018-07-03 ModFolder/9_Comp.R

@@ -1,0 +1,175 @@
+### Other compartment models with Gauss Newton OLS?
+
+# inf_data = infected counts
+# start_param = initialization parameters
+# start_pop = starting compartment sizes
+# t = time steps associated with the infected data
+# b is the birth rate per 1 person in each time step
+# Returns R0 estimate and standard deviation
+comp_gls <- function(inf_data, N, times, start_parm = c("beta" = 0.06, "epsilon" = 0.05, "delta" = 0.1, "b" = 0.00000001, "gamma" = 0.1), 
+                     start_pop = c("M" = 0, "E" = 0, "I" = 50, "R" = 0), k = 100, ndeps = 0.0000000000003) {
+  # Solve for the model parameters
+  # Compartment sizes should be scaled to be proportion of population
+  library(EpiModel)
+  library(numDeriv)
+  # MSEIR Model
+  meir <- function(t, t0, parms) {
+    with(as.list(c(t0, parms)), {    
+      
+      #uses B, sigma, gamma
+      ## Differential Equations
+      dm <- birth*(e + ii + rr) - delta * m
+      de <- beta * ii * (1 - m - e - ii - rr) - (epsilon + birth) * e
+      dii <- epsilon * e - (gamma + birth) * ii
+      drr <- gamma * ii - birth * rr
+      
+      ## Output
+      list(c(dm, de, dii, drr))
+    })
+  }
+  
+  init <- init.dcm(m = start_pop["M"] / N, e = start_pop["E"] / N, 
+                   ii = start_pop["I"] / N, rr = start_pop["R"] / N)
+  cont <- control.dcm(nsteps = max(times) + 1, dt = 1, new.mod = meir)
+  
+  # Standardize
+  inf_data <- inf_data / N
+  
+  # SSE objective function to minimize
+  obj_func <- function(data, estimate, times) {
+    return(sum((data - estimate[times])^2))
+  }
+  
+  # Compare model to data, return weighted SSE
+  comp <- function(param, initdcm, contdcm, weights, data, times) {
+    paramdcm <- param.dcm(beta = param["beta"], epsilon = param["epsilon"], delta = param["delta"],
+                          birth = param["b"], gamma = param["gamma"])
+    model1 <- dcm(paramdcm, initdcm, contdcm)
[…]

### 2018-07-03 ModFolder/r0_model.R

@@ -1,0 +1,197 @@
+setwd("C:/Users/Andersen/Documents/R0/Models")
+library(knitr)
+library(ggplot2)
+
+### FIle path to folder where the data are located
+data_path <- "C:/Users/Andersen/Documents/R0/DataSim/NewPar2/Data/"
+
+
+### Character vector of all files within data_path
+files_list <- paste0(data_path, "/", 
+                     list.files(data_path, recursive = TRUE))
+results_matlist <- list()
+
+xx <- files_list
+### Test dataset
+for (jj in 1:length(xx)) {
+# for (jj in 21:24) {
+  dat <- read.csv(xx[jj])[,-1]
+  start_pop <- c("S" = 99950, "E" = 0, "I" = 50, "R" = 0)
+  start_par_sir <- c("beta" = 0.002, "gamma" = 0.007)
+  start_par_rep <- c("r0" = 2, "gamma" = 0.007)
+  start_par_seir <- c("beta" = 0.008, "gamma" = 0.007, "alpha" = 0.0014)
+  
+  # Surv func
+  bxf <- function(t) {return(dexp(t / 2) * dnorm(t, 5, 3) * 10)}
+  b_dis <- rnorm(11, 0.6, 0.1)
+  f_dis <- dexp(0:10)
+  max_surv <- 100
+  
+  # Direct
+  latent_per <- 7
+  inf_per <- 20
+  latent_inf_per <- latent_per + inf_per
+  k <- 2
+  b <- 0.1
+  d <- 14
+  
+  # Direct survival
+  cont_inf <- function(t) {return(dexp(t / 1.5))}
+  dis_inf <- rbeta(11, 0.2, 1)
+  
+  # Attack Rate
+  ar <- (start_pop["S"] - dat[nrow(dat), 2]) / sum(dat[nrow(dat), -1])
+  
+  # GTD
+  pw <- c(dunif(10:30, 10, 30))
+  
+  
+  # Final size
+  fs <- dat[nrow(dat), 2] / sum(dat[nrow(dat), -1])
+  
+  
+  results_r0 <- list()
+  print(results_r0)
+  # ##### Survival function #####
+  # source("2_Survival.R")
+  # # Continuous test
+  # results_r0$survival_cont <- try(cont_surv_func(bxf, max_surv))
+  # # Discretized test
+  # results_r0$survival_dis <- try(disc_surv_func(b_dis, f_dis))
+  # 
+  # 
[…]

### 2018-07-03 plots/plot-seir.R

@@ -1,0 +1,155 @@
+## SKG
+## Plotting the baseline SEIR
+
+## Package from simCAM (https://github.com/shannong19/simCAM)
+
+devtools::load_all("~/simCAM")
+library(ggplot2)
+library(reshape2)
+library(latex2exp)
+
+r0_var <- function(hess, beta, gamma){
+    hess <- hess[1:2, 1:2]
+    grad <- c(1/gamma, -beta/gamma^2)
+    var <- t(grad) %*% solve(hess) %*% grad
+    return(var)
+}
+
+T <- 365
+init_vals <- c(9500, 0, 500, 0) #SEIR 
+beta <- .06
+gamma <- .03
+alpha <- .01
+N <- sum(init_vals)
+step <- 7
+## Plot the SIR
+do_plot = FALSE
+seir_data <- SEIR(T, init_vals, beta, gamma,
+                  alpha, step, SEIR_inner, do_plot)
+## Optimize function
+inits <- c(beta, gamma, alpha)
+params <- optim(par=inits, fn=SSE_SEIR, data=seir_data,
+                init_vals = init_vals, hessian = TRUE)
+colnames(seir_data)[2:5] <- c("X", "E", "Y", "Z")
+## Change SEIR data to SIR
+sir_data <- data.frame(t = seir_data$t,
+                       X = seir_data$X + seir_data$E,
+                       Y = seir_data$Y,
+                       Z = seir_data$Z)
+init_sir <- inits[-3]
+init_vals_sir <- init_vals[-2]
+params_sir <- optim(par=init_sir, fn=SSE_SIR, data=sir_data,
+                    init_vals = init_vals_sir, hessian = TRUE)
+r0_seir <- params$par[1] / params$par[2]
+r0_sir <- params_sir$par[1] / params_sir$par[2]
+c(r0_seir, r0_sir)
+
+sir_data$type <- "XYZ"
+sir_melt <- melt(sir_data, id.vars=c("t", "type"))
+seir_data$type <- "XEYZ"
+seir_melt <- melt(seir_data, id.vars = c("t", "type"))
+df_melt <- rbind(seir_melt, sir_melt)
+colnames(df_melt)[3] <- "Compartment"
+
+g <- ggplot(data = df_melt, aes(x=t, y=value, linetype=Compartment,
+                                col=Compartment)) + facet_wrap(~type, dir="v") + 
+    geom_line(size=1) +
+    scale_colour_grey() + 
[…]

### 2018-07-03 plots/plot_sir.R

@@ -1,0 +1,88 @@
+## SKG
+## Plotting the baseline SIR
+
+## Package from simCAM (https://github.com/shannong19/simCAM)
+
+devtools::load_all("~/simCAM")
+library(ggplot2)
+library(reshape2)
+library(latex2exp)
+
+
+T <- 364
+init_vals <- c(99950, 50, 0)
+beta <- .06
+gamma <- .03
+
+sir_out <- SIR(T, init_vals, beta, gamma)
+head(sir_out)
+colnames(sir_out)[2:4] <- c("X", "Y", "Z")
+
+df_melt <- melt(sir_out, id.vars = "t")
+colnames(df_melt)[2] <- "Compartment"
+
+g <- ggplot(data = df_melt, aes(x=t, y=value, linetype=Compartment,
+                                col=Compartment)) +
+    geom_line(size=2) +
+    scale_colour_grey() + 
+    theme_bw() +
+    labs(x = "Time", y = "Number of Individuals",
+         title = "SIR (XYZ) Curves",
+         subtitle = TeX(sprintf("$N$ = %.2e; $\\beta$ = %.2f; $\\gamma=$%.2f; $(X(0), Y(0))$= (%.3e, %.1e)",
+                                sum(init_vals), beta, gamma, init_vals[1], init_vals[2]))) +
+    ggplot2::theme(
+                     axis.text.x = ggplot2::element_text(size = 16),
+                     axis.text.y= ggplot2::element_text(size = 16),
+                     axis.title.x= ggplot2::element_text(size = 18),
+                     axis.title.y= ggplot2::element_text(size = 18),
+                     plot.title = ggplot2::element_text(size = 24),
+                     legend.title = ggplot2::element_text(size = 20),
+                     legend.text = ggplot2::element_text(size=16),
+                     legend.key.size = ggplot2::unit(3, "line"),
+                     plot.subtitle = ggplot2::element_text(size=16)
+                     )
+g
+
+
+## H1N1 Data plot
+
+library(lubridate)
+library(ggplot2)
+library(reshape2)
+library(scales)
+sir_out <- read.csv("../Data/H1N12009/data.csv")
+sir_out$t <- 15:34
+sir_out$date <- ymd(sir_out$date)
+
+colnames(sir_out)[1:3] <- c("X", "Y", "Z")
+
+df_melt <- melt(sir_out, id.vars = c("t", "date"))
[…]

### 2018-07-05 ModFolder/14_ExpGrowth.R
New files

@@ -30,7 +30,7 @@ exp_growth <- function(start_pop = c("S" = 99950, "E" = 0, "I" = 50, "R" = 0),
   dh <- -1 / (Mt ^ 2) * sum(-discrete_steps * w * exp(-discrete_steps * r))
   
   
-  r0_sd <- dh ^ 2 * r_se
+  r0_sd <- dh * r_se
   
   return(list(est = r0, sd = r0_sd))
 }

### 2018-07-05 ModFolder/19_wallinga.R

@@ -1,0 +1,13 @@
+
+wallinga <- function(start_pop = c("S" = 99950, "E" = 0, "I" = 50, "R" = 0), 
+                     inf_data, gam){
+  log_it <- log(pmax(inf_data, 0.1))
+  log_istart <- log(start_pop["I"])
+  y <- log_it - log_istart
+  tt <- 1:length(inf_data)
+  lm_r <- lm(y ~ tt - 1)
+  r0 <- 1 + coef(lm_r)[1] / gam
+  r0_sd <- summary(lm_r)$coef[1, 2] / gam
+  
+  return(list(est = r0, sd = r0_sd))
+}

### 2018-07-05 ModFolder/5-2_SIRrepar.R

@@ -15,7 +15,8 @@ library(deSolve)
 #sus.data is the number of people in S for each day in the data. Must be a vector of numbers, same length as days.
 #ndeps is for optim calculating the hessian
 
-SIRrepar <- function(init.param, init.pop, days, inf.data, sus.data, ndeps = 0.00001) {
+SIRrepar <- function(init.param, init.pop, days, inf.data, sus.data, 
+                     ndeps = 0.00001, parscale_o = c(0.01, 0.01)) {
   #Get initial populations of each compartment S, I, R.
   init.sus <- init.pop["S"]
   init.inf <- init.pop["I"]
@@ -68,7 +69,8 @@ SIRrepar <- function(init.param, init.pop, days, inf.data, sus.data, ndeps = 0.0
                      sus.data = sus.data,
                      inf.data = inf.data,
                      hessian = T,
-                     control = list(ndeps = rep(ndeps, length(init.param))))
+                     control = list(ndeps = rep(ndeps, length(init.param)),
+                                    parscale = parscale_o))
     return(c(par.sir))
   }
   
@@ -133,13 +135,13 @@ SIRrepar <- function(init.param, init.pop, days, inf.data, sus.data, ndeps = 0.0
   model_est <- sir.optim(init.param, init.sir, control.sir, inf.data, sus.data, days)
   
   r0_est <- model_est$par["r0"]
-  r0_hessian <- model_est$hessian
-  r0_sd <- sqrt(1 / r0_hessian[1,1])
+  r0_hessian <- solve(model_est$hessian / model_est$value)
+  r0_sd <- sqrt(abs(r0_hessian[1,1]))
   
   r0var <- sensitivity_analysis(model_est$par["r0"], model_est$par["gamma"], 
                                  T = max(days), t = days, X0 = init.sus, 
                                  Y0 = init.inf, Z0 = init.rem, inner_fxn = SIR_re_inner)
   r0sd2 <- sqrt(1 / r0var[1,1])
   
-  return(list(est = r0_est, sd = r0_sd, sd2 = r0sd2))
+  return(list(est = r0_est, sd = r0_sd, sd2 = r0sd2, output = model_est))
 }

### 2018-07-05 ModFolder/5-3_SIRlma.R

@@ -19,7 +19,7 @@ poly_deriv <- function(t, mod){
   if( k == 1 ) {return(deriv_coefs)}
   else if ( k < 1){ return(0)}
   poly_mat <- cbind(rep(1, length(t)), poly(t, k-1, raw = TRUE))
-  deriv_val <- poly_mat %*% (deriv_coefs * 1:k)
+  deriv_val <- poly_mat %*% (deriv_coefs * (1:k))
   return(deriv_val)
 }
 
@@ -45,7 +45,7 @@ SIRlma <- function(init.pop, days, inf.data, sus.data, k = 10) {
   dxt <- poly_deriv(days, x_mod)
   dyt <- poly_deriv(days, y_mod)
   xt <- x_hat
-  r0_est <- (dxt[1] * N) / ((dyt[1] + dxt[1]) * xt[1])
+  r0_est <- (dxt[1] * N) / ((dyt[1] + dxt[1]) * init.sus )
   
   r01 <- numeric(length(days))
   for(ii in 1:length(r01)){
@@ -56,12 +56,14 @@ SIRlma <- function(init.pop, days, inf.data, sus.data, k = 10) {
     dxt <- poly_deriv(days, x_mod)
     dyt <- poly_deriv(days, y_mod)
     xt <- x_hat
-    r01[ii] <- (dxt[1] * N) / ((dyt[1] + dxt[1]) * xt[1])
+    r01[ii] <- (dxt[1] * N) / ((dyt[1] + dxt[1]) * init.sus)
   }
   r01var <- (n-1) / n * sum((r01 - mean(r01))^2)
   r0_sd <- sqrt(r01var)
   
-  return(list(est = r0_est, sd = r0_sd))
+  return(list(est = r0_est, sd = r0_sd, 
+              output = list(y_mod = y_mod, x_mod = x_mod, x_hat = x_hat,
+                            y_hat = y_hat, dxt = dxt, dyt = dyt)))
 }
 
 SIRlmat <- function(init.pop, days, inf.data, sus.data, k = 10) {
@@ -86,9 +88,11 @@ SIRlmat <- function(init.pop, days, inf.data, sus.data, k = 10) {
   dxt <- poly_deriv(days, x_mod)
   dyt <- poly_deriv(days, y_mod)
   xt <- x_hat
-  r0_est <- mean((dxt  / (dyt + dxt)) )* N / xt[1]
-  r0_sd <- sd((dxt  / (dyt + dxt)) )* N / xt[1]
[…]

### 2018-07-05 ModFolder/5-5_SIRipr.R

@@ -16,7 +16,7 @@ SIRipr <- function(init.pop, days, inf.data, sus.data, gam) {
   IPR <- ifelse(IPR < 0, 0, IPR)
   r0_est <- mean(IPR / gam)
   r0_sd <- sd(IPR / gam)
-  return(list(est = r0_est, sd = r0_sd))
+  return(list(est = r0_est, sd = r0_sd, output = list(IPR = IPR)))
 }
 
 SIRsipr <- function(init.pop, days, inf.data, sus.data, gam) {
@@ -41,8 +41,9 @@ SIRsipr <- function(init.pop, days, inf.data, sus.data, gam) {
     yhat <- predict(mody, seq(0, max(days), by = 1))
     J <- -diff(xhat$y, lag = 1) #incidence
     IPR <- J[1] / yhat$y[1]
+    IPR <- ifelse(IPR < 0, 0, IPR)
     r0[ii] <- IPR / gam
   }
   r0_sd <- sqrt((n-1) / n * sum((r0 - mean(r0))^2))
-  return(list(est = r0_est, sd = r0_sd))
+  return(list(est = r0_est, sd = r0_sd, output = list(IPR = IPR, xhat = xhat, yhat = yhat)))
 }

### 2018-07-05 ModFolder/5-6_SIRmc.R

@@ -4,13 +4,14 @@ SIRmc <- function(init.pop, inf.data, sus.data, k = 100) {
   N <- sum(init.pop)
   sus.data <- c(init.pop["S"], sus.data)
   inf.data <- c(init.pop["I"], inf.data)
-  resp <- (log(sus.data[2:(length(sus.data))]) - log(sus.data[1:(length(sus.data) - 1)]))
-  pred <- inf.data[1:(length(sus.data) - 1)] / N
-  reg <- lm(resp ~ pred - 1)
-  alpha <- exp(coef(summary(reg))[1, 1])
-  r0_est <- log(1 / (1 - alpha))
-  alpha_se <- (coef(summary(reg))[1, 2])
-  
-  r0_sd <- (1 / (1 - alpha)) ^ 2 * alpha_se
-  return(list(est = r0_est, sd = r0_sd))
+  resp <- exp((log(sus.data[2:(length(sus.data))]) - log(sus.data[1:(length(sus.data) - 1)])) /
+    inf.data[1:(length(sus.data) - 1)])
+  reg <- glm(resp ~ 1)
+  g <- coef(summary(reg))[1, 1]
+  alpha <- exp(g) / (1-exp(g))
+  r0_est <- -log((1 - exp(g)) / (1 - 2 * exp(g)))
+  g_se <- (coef(summary(reg))[1, 2])
+  alpha_se <- (exp(g) / (1 - exp(g)) ^ 2) ^ 2 * g_se
+  r0_sd <- (exp(g) / (-3*exp(g) + 2 * exp(2 * g) + 1)) ^ 2 * g_se
+  return(list(est = r0_est, sd = r0_sd, output = summary(reg)))
 }

### 2018-07-05 ModFolder/5_SIR.R

@@ -15,7 +15,8 @@ library(deSolve)
 #sus.data is the number of people in S for each day in the data. Must be a vector of numbers, same length as days.
 
 
-SIRreg <- function(init.param, init.pop, days, inf.data, sus.data, ndeps = 0.00001) {
+SIRreg <- function(init.param, init.pop, days, inf.data, sus.data, 
+                   ndeps = 0.00001, parscale_o = c(0.01, 0.01)) {
   #Get initial populations of each compartment S, I, R.
   init.sus <- init.pop["S"]
   init.inf <- init.pop["I"]
@@ -68,7 +69,8 @@ SIRreg <- function(init.param, init.pop, days, inf.data, sus.data, ndeps = 0.000
                      sus.data = sus.data,
                      inf.data = inf.data,
                      hessian = T,
-                     control = list(ndeps = rep(ndeps, length(init.param))))
+                     control = list(ndeps = rep(ndeps, length(init.param)),
+                                    parscale = parscale_o))
     return(c(par.sir))
   }
   
@@ -80,7 +82,7 @@ SIRreg <- function(init.param, init.pop, days, inf.data, sus.data, ndeps = 0.000
   model_est <- sir.optim(init.param, init.sir, control.sir, inf.data, sus.data, days)
   
   r0_est <- model_est$par["beta"] / model_est$par["gamma"]
-  r0_hessian <- model_est$hessian
+  r0_hessian <- model_est$hessian / model_est$value
   dh <- c(1/model_est$par["gamma"], - model_est$par["beta"] / (1/model_est$par["gamma"]^2))
   r0_sd <- sqrt(t(dh) %*% solve(r0_hessian) %*% dh)
 
@@ -153,7 +155,7 @@ SIRreg <- function(init.param, init.pop, days, inf.data, sus.data, ndeps = 0.000
   
   r0_sd2 <- sqrt(t(dh) %*% solve(r0_sd2) %*% dh)
   
-  return(list(est = r0_est, sd = r0_sd, sd2 = r0_sd2))
+  return(list(est = r0_est, sd = r0_sd, sd2 = r0_sd2, output = model_est))
 }
 
 #dat <- read.csv("https://raw.githubusercontent.com/atzechang/data/master/Baseline/Baseline1_99950_10_1/Baseline1Norm.csv")[,-1]

### 2018-07-05 ModFolder/6_Harko.R

@@ -52,7 +52,7 @@ harkoSIR <- function(start_param = c("beta" = 0.06, "gamma" = 0.05),
   r0_est <- abs(summary(lin_reg)$coef[2, 1])
   r0_sd <- summary(lin_reg)$coef[2, 2]
 #   return(list(est = r0_est))
-  return(list(est = r0_est, sd = r0_sd))
+  return(list(est = r0_est, sd = r0_sd, output = summary(lin_reg)))
 }
 
 

### 2018-07-05 ModFolder/r0_methods.R

@@ -1,0 +1,396 @@
+setwd("C:/Users/Andersen/Documents/R0/Models")
+library(knitr)
+library(ggplot2)
+
+### File path to folder where the data are located
+data_path <- "C:/Users/Andersen/Documents/R0/DataSim/PaperData/Baseline"
+# data_path <- "C:/Users/Andersen/Dropbox/r0_refs/Data/H1N12009"
+
+### Character vector of all files within data_path
+files_list <- paste0(data_path, "/", 
+                     list.files(data_path, recursive = TRUE))
+
+results_matlist<- list()
+output_lists <- list()
+
+# xx <- files_list
+
+xx <- files_list
+
+### Test dataset
+for (jj in 1:length(xx)) {
+# for (jj in 5:20) {
+  print(xx[jj])
+  # r0_true <- as.numeric(strsplit(xx[[jj]], split = "_")[[1]][12]) / 
+  #   as.numeric(strsplit(xx[[jj]], split = "_")[[1]][13])
+  # beta_true <- as.numeric(strsplit(xx[[jj]], split = "_")[[1]][12])
+  # gamma_true <- as.numeric(strsplit(xx[[jj]], split = "_")[[1]][11])
+  gamma_true <- 0.03
+  # output_list <- list()
+  # sink(zz, append = TRUE)
+  # sink(zz, type = "message", append = TRUE)
+  # print("------------------------------------------")
+  # print("------------------------------------------")
+  # print("------------------------------------------")
+  # print("------------------------------------------")
+  # print("------------------------------------------")
+  # print((paste0("Beta = ", beta_true, ", Gamma = ", gamma_true, ", R0 = ", r0_true)))
+  # print("")
+  # print("")
+  dat <- read.csv(xx[jj])
+  # start_pop <- c("S" = as.numeric(strsplit(strsplit(xx[jj], "/")[[1]][10], "_")[[1]][1]), "E" = 0,
+  #                "I" = as.numeric(strsplit(strsplit(xx[jj], "/")[[1]][10], "_")[[1]][2]), "R" = 0)
+  
+  start_pop <- c("S" = 99950, "E" = 0,
[…]

### 2018-07-05 plots/functions/altplot.R

@@ -1,0 +1,207 @@
+res_path <- "C:/Users/Andersen/Documents/R0/DataSim/PaperResults/3"
+
+res_list <- paste0(res_path, "/", 
+                     list.files(res_path))[-3]
+
+alt_list <- vector("list", length = 3)
+
+alt_list2 <- vector("list", length = 3)
+
+for(kk in 1:length(alt_list)){
+  alt_list[[kk]] <- matrix(NA, nrow = 1, ncol = 2)
+  colnames(alt_list[[kk]]) <- c("Estimate", "Std. Dev")
+}
+
+blah <- c(1, 2, 7)
+for(ii in 1:length(res_list)){
+  load(res_list[ii])
+  for(jj in 1:length(results_matlist)){
+    for(kk in 1:3) {
+      alt_list[[kk]] <- rbind(alt_list[[kk]], results_matlist[[jj]][blah[kk], 2:3])
+    }
+  }
+}
+
+dat_names_anorder <- c("Base", "Base3", "Base4", "Base5",
+                       "SD = (10000, 500)", "SD = (2500, 100)", "SD = (50, 2)", "SD = (500, 20)", 
+                       "R0 = 60", "R0 = 1.5", "R0 = 1", "R0 = 0.25",
+                       "N = 100", "N = 1000", "N = 10000", "N = 1000000",
+                       "Y(0) = 1000", "Y(0) = 100", "Y(0) = 10", "Y(0) = 1",
+                       "T = 100", "T = 20", "T = 200", "T = 50")
+
+dat_names_paperorder <- c("Base",
+                          "R0 = 60", "R0 = 1.5", "R0 = 1", "R0 = 0.25",
+                          "T = 200", "T = 100", "T = 50", "T = 20",
+                          "Y(0) = 1000", "Y(0) = 100", "Y(0) = 10", "Y(0) = 1",
+                          "N = 1000000", "N = 10000", "N = 1000", "N = 100",
+                          "SD = (50, 2)", "SD = (500, 20)", "SD = (2500, 100)", "SD = (10000, 500)")
+  
+mod_names <- c("RE", "rRE", "LL")
+
+for(kk in 1:length(alt_list2)){
+  alt_list2[[kk]] <- data.frame(Estimate = alt_list[[kk]][-1, 1],
+                                SE = alt_list[[kk]][-1, 2])
+  alt_list2[[kk]]$Error <- rep(c("AR", "AR-M", "Norm", "Norm-M"), times = 96/4)
[…]

### 2018-07-05 plots/functions/plotmaker.R

@@ -1,0 +1,226 @@
+# setwd("~/R0/Models")
+# load("sir_all.RData")
+# load("refit.RData")
+
+
+library(ggplot2)
+library(latex2exp)
+library(knitr)
+
+results_matlist_comb <- list()
+results_matlistbase_comb <- list()
+results_matlistvar_comb <- list()
+results_matlistpop_comb <- list()
+results_matlistoth_comb <- list()
+
+# for(ii in 1:length(results_matlistbase)){
+for(ii in 3:3){
+  rez <- results_matlistbase[[ii]]
+  results_matlistbase_comb[[ii]] <- data.frame(rbind(rez[c(1:10, 18, 11, 13, 14, 16, 17, 19, 20, 15),]))
+  results_matlistbase_comb[[ii]]$Method <- factor(results_matlistbase_comb[[ii]]$Model,
+                                                  levels = rev(results_matlistbase_comb[[ii]]$Model),
+                                                  labels = rev(c("SIR", "SIR Repar.",
+                                                             "SIR LMA", "SIR LMAT",
+                                                             "SIR Max", "SIR SMax",
+                                                             "SIR IPR", "SIR SIPR",
+                                                             "SIR MC", "SIR LL",
+                                                             "Seq. Bayes",
+                                                             "SEIR",
+                                                             "Comp", "Next Gen.",
+                                                             "Exp. Growth", "MLE",
+                                                             "Init. Growth", "Final Size",
+                                                             "Branch")))
+  results_matlistbase_comb[[ii]]$Lower <- results_matlistbase_comb[[ii]]$Estimate - 2 * results_matlistbase_comb[[ii]]$Std..Dev
+  results_matlistbase_comb[[ii]]$Upper <- results_matlistbase_comb[[ii]]$Estimate + 2 * results_matlistbase_comb[[ii]]$Std..Dev
+}
+
+for(ii in 1:length(results_matlistvar)){
+  rez <- results_matlistvar[[ii]]
+  results_matlistvar_comb[[ii]] <- data.frame(rbind(rez[c(1:10, 18, 11, 13, 14, 16, 17, 19, 20, 15),]))
+  results_matlistvar_comb[[ii]]$Method <- factor(results_matlistvar_comb[[ii]]$Model,
+                                                 levels = rev(results_matlistvar_comb[[ii]]$Model),
[…]

### 2018-07-05 plots/functions/plots2.R

@@ -1,0 +1,301 @@
+xx <- results_matlist[[4]]
+
+xx$Model <- toupper(xx$Model)
+
+xxx <- cbind(xx, results_matlist[[2]][, 2:3])
+
+colnames(xxx)[-1] <- c("Norm-M Est.", "Norm-M SE", "AR-M Est.", "AR-E SE")
+
+kable(xxx, format = "latex", caption = "")
+
+library(tidyverse)
+library(gridExtra)
+library(latex2exp)
+
+for(ii in c(2, 4)){
+  results_matlist[[ii]]$Model[11] <- "IDEA"
+  results_matlist[[ii]]$Model <- toupper(results_matlist[[ii]]$Model)
+  results_matlist[[ii]]$Lower <- pmax(results_matlist[[ii]]$Estimate - 2 * results_matlist[[ii]][["Std. Dev"]], -4)
+  results_matlist[[ii]]$Upper <- pmin(results_matlist[[ii]]$Estimate + 2 * results_matlist[[ii]][["Std. Dev"]], 5.5)
+  results_matlist[[ii]]$Model <- factor(results_matlist[[ii]]$Model, levels = rev(results_matlist[[ii]]$Model))
+}
+
+g1 <- ggplot(data = results_matlist[[2]][1:9,], aes(y = Model, x = Estimate, xmin = Lower, xmax = Upper)) +
+  geom_vline(xintercept = 2, col = "black", size = 2) +
+  geom_errorbarh(height = .3, size = 2, col = "black") +
+  geom_point(size = 4) +
+  labs(x = TeX('$\\textbf{R}_0$'), y = "Method", title = "AR-M Errors") + 
+  xlim(-4, 5.5) + theme_bw() + theme(title = element_text(size = 18),
+                                       axis.text = element_text(size = 11))
+
+g2 <- ggplot(data = results_matlist[[4]][1:9,], aes(y = Model, x = Estimate, xmin = Lower, xmax = Upper)) +
+  geom_vline(xintercept = 2, col = "black", size = 2) +
+  geom_errorbarh(height = .3, size = 2, col = "black") +
+  geom_point(size = 4) +
+  labs(x = TeX('$\\textbf{R}_0$'), y = "Method", title = "Norm-M Errors") + 
+  xlim(-4, 5.5) + theme_bw() + theme(title = element_text(size = 18),
+                                       axis.text = element_text(size = 11))
+
+grid.arrange(g2, g1, ncol = 2)
+
[…]

### 2018-07-05 plots/functions/plots3.R

@@ -1,0 +1,301 @@
+library(tidyverse)
+library(latex2exp)
+library(data.table)
+
+errcomp <- list()
+# yy <- matrix(c(1:16),
+#              ncol = 4, byrow = TRUE)
+yy <- matrix(c(9:12, 1:4, 13:16, 5:8),
+                     ncol = 4, byrow = TRUE)
+for (ii in 1:length(results_matlist[[1]]$Model)) {
+  errcomp_mat <- matrix(NA, ncol = 8, nrow = 0)
+  for (jj in 1:nrow(yy)) {
+    newrow <- c()
+    for(kk in yy[jj,]) {
+      newrow <- c(newrow, results_matlist[[kk]][ii, 2], results_matlist[[kk]][ii, 3])
+    }
+    # errcomp_mat <- rbind(errcomp_mat, round(newrow, 4))
+    errcomp_mat <- rbind(errcomp_mat, signif(newrow, 4))
+  }
+  r0_true <- unlist(lapply(strsplit(files_list, "_"),
+                           function(x){as.numeric(x[11]) / as.numeric(x[12])}))[1]
+  err_names <- rep("", 4)
+  # err_names <- unlist(lapply(strsplit(files_list, "_"),
+  #                     function(x){paste0("$beta = ",
+  #                                        x[12],
+  #                                        ", gamma = ",
+  #                                        x[13],
+  #                                        ", rrr = ",
+  #                                        round(as.numeric(x[12]) / as.numeric(x[13]), 3),
+  #                                        "$")}))[(1:(length(results_matlist[[1]]$Model) / 4))*4]
+  errcomp_mat <- data.frame(err_names, errcomp_mat)[, c(1, 6:9, 2:5)]
+  errcomp_mat2 <- errcomp_mat
+  #  apply(errcomp_mat[, -1], 2, sd)
+  errcomp_mat2[, (1:4)*2] <-  round((errcomp_mat[, (1:4)*2] - r0_true) / errcomp_mat[, (1:4)*2 + 1], 4)
+  colnames(errcomp_mat) <- c("Data Set", "AR Est", "AR SE",
+                             "AR-M Est", "AR-M SE",
+                             "Norm Est", "Norm SE",
+                             "Norm-M Est", "Norm-M SE")[c(1, 6:9, 2:5)]
[…]

### 2018-07-06 paper/Version15/draft_v15.tex
Added abstract.  Started cleaning up draft

@@
 
-\tableofcontents
+\begin{abstract}
+For nearly a century, the initial reproduction number ($\rr$) has been used to summarize infectious diseases and to compare them to one another.
+It is difficult to estimate $\rr$ due to assumptions on both how a disease transmits through a population and also due to differences in statistical estimation methods.
+We describe nine methods used to estimate $\rr$ and provide a thorough simulation study of how these methods differ in the presence of disease parameters.
+As real world motivation, we analyze the outbreak of the H1N1 pandemic influenza in the USA and compare the results from our nine methods to a previous study.
+We discuss the most important aspects from our results which effect the estimation of $\rr$ and provide guidelines for estimating point estimates and confidence/credible intervals to create reliable, comparable estimates of $\rr$.
+  \end{abstract}
 
@@
 
+\pagebreak
+
 \bibliographystyle{apa}%Choose a bibliograhpic style
@@
 \appendix
-
-\section{Beyond the SIR Model}
-
-\subsection{Exponential Growth (EG)}\label{sec:expgrowth}
-%% nishiura, majumder, IDEA model -- sum of squares (santillana2016)
-%% An IDEA for short term outbreak projection: nearcasting using the basic reproduction number.  (fisman2013, fisman2014)
-%% chowell 2008 - parametric bootstrap for a model they made up, but sort of empirical.
-Many researchers employ an assumption of exponential growth of the number infectious through times $t=0, 1, \dots, T^* <T$ \citep{wallinga2007generation,fisman2014,nishiura2016,majumder2016,towers2016}.
-\cite{wallinga2007generation} report that the effective reproduction number $\mathcal{R}_t$ and hence the initial reproduction number $\rr$ may derived using the fact that infection ``counts increase exponentially in the initial phase of an epidemic.''  We then have to estimate $r$, the \textit{per capita} change in the number of new cases per unit of time and $\omega$ the serial interval, the distribution of time between a primary and secondary infection.
-Then, we have
-\begin{align}\label{eq:lotka}
-\rr = \exp{(r \omega)}.
-\end{align}
-Equation \eqref{eq:lotka} is derived from a demographic view using the Lotka-Euler survival equations which come from the fields of demography, ecology, and evolutionary biology.
-
-\cite{wallinga2007generation}, instead, use a moment generating function expression for $\rr$.  With $\omega(t)$ as the serial interval, then
-\begin{align*}
-\rr^{-1} &= \frac{1}{M(-r)} = \int_{a=0}^\infty e^{-rt}\omega(t)dt.
-\end{align*}
-\cite{wallinga2007generation} specifically derive an estimate for $\rr$ for the SIR model.
-They assume that the rate of leaving the infectious stsate, $\gamma$ is constant as is the rate of making contacts during the infectious stage.
-Then, the duration of a serial interval is an exponential distribution with mean $\gamma^{-1}$, the moment generating function is then,
-\begin{align*}
-  M(-r) = \frac{\gamma}{\gamma + r},
-\end{align*}
-and the resulting estimate of $\rr$ is
-\begin{align}\label{eq:r0-eg}
-  \hat{\rr} = 1 + \frac{\hat{r}}{\hat{\gamma}},
-\end{align}
-where $\hat{r}$ and $\hat{\gamma}$ are estimates of the $r$ and $\gamma$, respectively.
-The parameter $\hat{\gamma}$ is often assumed to be fixed or tried with a sequence of serial intervals \citep{majumder2016}.
-We may estimate $\hat{r}$ using ordinary linear regression, namely,
-\begin{align*}
-  \log \left  (\frac{Y(t)}{Y(0)} \right )&= rt + \epsilon_t\\
-  \epsilon_t &\sim N(0, \sigma^2),
-\end{align*}
-which results in
-\begin{align}\label{eq:exp-growth}
-  \hat{r} &= \frac{\sum_{t=0}^{T^*}\left  (\frac{Y(t)}{Y(0)} \right )}{\sum_{t=0}^{T^*}t}.
-\end{align}
-
-For the above estimate of $\rr$, we observe the duration of serial intervals in a period of exponential growth of the number of infectious individuals.
-Deciding when exponential growth occurs and which data points to use to estimate $\rr$ may be difficult.
-We can use sensitivity analysis to see how the estimate $\rr$ changes as we increase or decrease $T^{*}$, the number of time points used.
-The reader may notice some similarities to the estimate for $\rr$ in Equation \eqref{eq:r0-eg} and from the log-linear model in Equation \eqref{eq:harko_lin}.
-This method is a simplification of the SIR model as the $R$ compartment is not used to estimate $r$, only the recovery rate $\gamma$.
-
-An advantage to this method are its (and similar adaptations) prevalence in the literature, which allows one to to compare estimates $\rr$ directly to one another.
-Another advantage is its flexibility.
-Here, we assumed the serial interval has an exponential distribution, but this is not a requirement and other distributions may be used (see \cite{wallinga2007generation}).
-However, a disadvantage to this method is that is a simplification of the SIR model and we have to decide how many time points to use to estimate the exponential growth rate of infectious.
-Another disadvantage to this method is that one also has to decide how to treat the additional paramter $\gamma$, whether to treat it as fixed, a range of fixed values, or even having a probability distribution.
-
-\subsection{Incidence Decay and Exponential Adjustment (IDEA)}\label{sec:idea}
-The Incidence Decay and Exponential Adjustment (IDEA) model is a way to account for a declining effective reproduction number of time due to both depletion of susceptible individuals and spontaneous and planned control activities and behaviors'' \citep{fisman2013}.
-They model the incidence counts during the $t$th interval, $J(t),$ as
-\begin{align*}
-  J(t) &= \left [ \frac{\rr}{(1 + d)^t}\right ]^t,
-\end{align*}
-where $d$ is a dampening or discount factor.  As $d$ increases, the incidence at time $t$  decreases. The estimate for $\rr$ is obtained by minimizing the sum of squares over $\rr$ and $d$,
-\begin{align*}
-  \hat{\rr} &=  \arg \min_{\rr,d} \sum_{t=0}^{T^*}\left (J_{obs}(t) - J(t) \right)^2.
-\end{align*}
-The method presented in \cite{fisman2013} is a mathematical, rather than statistical model.
-Sensitivity analysis is conducted with respect to $T^*$, $d$, and $\rr$, and lower and upper bounds are reported as a result of this analysis.
-
-Advantages of this method include its usefulness in estimating $\rr$ \citep{fisman2014,majumder2016}.
-Additionally, this model may be employed to address the effect of control measures and better estimate the final size of an epidemic.
-A disadvantage to this model is that its upper and lower bounds do not correspond to CIs.
-Also, in this method we must select the interval size of the incidence counts.
 

### 2018-07-09 paper/Version15/draft_v15.tex
Added in Bill edits up to Section 3

@@
 \section{Introduction}\label{sec:intro}
-What has been called ``arguably the most important quantity in the study of epidemics'' \citep{Heesterbeek2002},  $\mathcal{R}_0$, the reproduction number (by convention pronounced ``R-naught'') is an elusive quantity for epidemic modelers to estimate.
-As defined by \citet{anderson1992}, $\rr$ is the ``the average number of secondary infections produced when one infected individual is introduced into a host population where everyone is susceptible.''  In some ways,$\rr$ summarizes an entire outbreak of a disease; it  is used to assess whether a disease epidemic will occur.
+What has been called ``arguably the most important quantity in the study of epidemics'' \citep{Heesterbeek2002},  $\mathcal{R}_0$, the initial reproduction number (by convention pronounced ``R-naught'') is an elusive quantity for epidemic modelers to estimate.
+As defined by \citet{anderson1992}, $\rr$ is the ``the average number of secondary infections produced when one infected individual is introduced into a host population where everyone is susceptible.''  In some ways, $\rr$ summarizes an entire outbreak of a disease; it  is used to assess whether a disease epidemic will occur.
 Additionally, it describes what percentage of the population needs to be vaccinated to avoid such an epidemic, roughly $1-\rr^{-1}$ \citep{anderson1992}).
@@
 time), boundary cases of infection and recovery rates, population size, initial SI ratio, and restrictive assumptions placed upon the noise.
-These are all issues for producing point estimates, to say nothing of confidence or credible intervals (CI).
+These are all issues for producing point estimates, to say nothing of issues related to confidence or credible intervals (CI).
 
@@
 For example, Gaussian error, a common assumption in statistics and machine learning, becomes highly suspect due to the non-symmetrical nature of the noise imposed by the monotonic $S$ and $R$ compartments.
-Without assuming Gaussian noise, it becomes much more difficult to guarantee properties such as unbiased and consistent estimates of $\rr$.
-
-These difficulties in estimating $\rr$ have led a number of researchers to analyze the sensitivity of $\rr$ with respect to small perturbations in parameters like $\beta$ and $\gamma$ \citep{lash2003,epstein2007agent,capaldi2012}.
+
+These difficulties in estimating $\rr$ have led a number of researchers to analyze the sensitivity of $\rr$ with respect to small perturbations in parameters like $\beta$ and $\gamma$, respectively the infection and recovery rate \citep{lash2003,epstein2007agent,capaldi2012}.
 Even this is a difficult task.
@@
 Some of the accompanying intervals (not necessarily CIs) are extremely large such as the case of HIV in Uganda (Interval: [0.12, 14.17]) whereas some are extremely small such as for Ebola in Guinea (Interval: [1.50, 1.52]).
-We see that estimates for H1N1 influenza are generally higher than that of seasonal influenza, which is evidence that H1N1 should have been treated as a more critical concern than the typical seasonal influenza cycle.
+We see that estimates for H1N1 influenza are generally higher than that of seasonal influenza, although we find that in our appliaction the opposite is true.
 The lowest possible estimate for $\rr$ reported is 0.12 for HIV in Uganda, and the largest is 17.00 for measles in England and Wales.
@@
   HIV                 & Uganda                           & 2002-2008                     & 3.31        & 0.12           & 14.17          & \cite{nsubuga2014}    \\ \hline
+ZIK-V                & Colombia - Barrranquilla           & 2015                          & 2.40         & 3.80            & 5.60            & \cite{towers2016}    \\  
+  ZIK-V                & Colombia                         & 2015-2016                     & 2.50         & 3.00              & 3.60            & \cite{nishiura2016}    \\
+ZIK-V                & Colombia                         & 2015                          & 1.42        & 2.56           & 3.83           & \cite{majumder2016}  \\
 ZIK-V                & French Polynesia                 & 2013                          & 2.75        & 2.53           & 2.98           & \cite{zhang2017}        \\
 ZIK-V                & French Polynesia                 & 2013-2014                     & 3.70         & 2.60            & 4.80            & \cite{kucharski2016}  \\
-ZIK-V                & Florida                          & 2016                          & 0.16        & 0.13           & 0.19           & \cite{dinh2016}       \\
-ZIK-V                & Colombia                         & 2015-2016                     & 2.50         & 3.00              & 3.60            & \cite{nishiura2016}    \\
-ZIK-V                & Barranquilla, Colombia           & 2015                          & 2.40         & 3.80            & 5.60            & \cite{towers2016}    \\
-ZIK-V                & Colombia                         & 2015                          & 1.42        & 2.56           & 3.83           & \cite{majumder2016}  \\ \hline
+  ZIK-V                & Florida                          & 2016                          & 0.16        & 0.13           & 0.19           & \cite{dinh2016}       \\ \hline
+EVD                 & Democratic Republic of the Congo & 1995                          & 2.70         & 1.90            & 2.80            & \cite{legrand2007}  \\  
+  EVD                 & Guinea                           & 2014                          & 1.51        & 1.50            & 1.52           & \cite{althaus2014}   \\
+  EVD                 & Liberia                          & 2014                          & 1.59        & 1.57           & 1.60            & \cite{althaus2014}    \\
+  EVD                 & Sierra Leone                     & 2014                          & 2.53        & 2.41           & 2.67           & \cite{althaus2014}      \\
+EVD                 & Uganda                           & 2000                          & 2.70         & 2.50            & 4.10            & \cite{legrand2007}   \\
 EVD               & Western Africa                   & 2014                          & 1.78        & 1.60            & 2.00              & \cite{fisman2014}   \\
 EVD & Western Africa                   & 2014                          & 1.60         & 1.40            & 1.80            & \cite{towers2014}  \\
-EVD                 & Democratic Republic of the Congo & 1995                          & 2.70         & 1.90            & 2.80            & \cite{legrand2007}  \\
-EVD                 & Uganda                           & 2000                          & 2.70         & 2.50            & 4.10            & \cite{legrand2007}  \\
-EVD                 & Guinea                           & 2014                          & 1.51        & 1.50            & 1.52           & \cite{althaus2014}   \\
-EVD                 & Sierra Leone                     & 2014                          & 2.53        & 2.41           & 2.67           & \cite{althaus2014}      \\
-EVD                 & Liberia                          & 2014                          & 1.59        & 1.57           & 1.60            & \cite{althaus2014}    \\ \hline
-Seasonal Influenza  & USA, France, Australia           & 1972-2002 & 1.30         & 1.20            & 1.40            & \cite{chowell2008}  \\
+\hline
+Seasonal Influenza  & Australia, France, USA          & 1972-2002 & 1.30         & 1.20            & 1.40            & \cite{chowell2008}  \\
   Seasonal Influenza  & French Territory                 & 1987-1995                     & 1.50         & 1.09           & 1.73           & \cite{bonabeau1998}  \\ \hline
+  H1N1 Influenza     & Peru                             & 2009                          & 1.37        & 1.20            & 1.70            & \cite{desilva2009}        \\   
 H1N1 Influenza      & Thailand                         & 2009                          & 2.07        & 1.92           & 2.22           & \cite{desilva2009}     \\
-  H1N1 Influenza     & Peru                             & 2009                          & 1.37        & 1.20            & 1.70            & \cite{desilva2009}        \\ 
   H1N1 Influenza & USA & 2009 & 1.50 & 1.30 & 1.70 &\cite{towers2009} \\ \hline
-Measles      & Cirencester, England                 & 1947-1950                          & 13.50   & 13.00           & 14.00            & \cite{fraser2009}     \\ 
+  Measles     & Canada - Ontario                  & 1912-1913                         & 11.50        & 11.00           & 12.00            & \cite{anderson1992}     \\
+    Measles     & Eastern Nigeria                  & 1960-1968                          & 16.50        & 16.00          & 17.00           & \cite{anderson1992}     \\
+  Measles      & England - Cirencester                 & 1947-1950                          & 13.50   & 13.00           & 14.00            & \cite{fraser2009}     \\
+  Measles     & England - Willesden                  & 1912-1913                          & 11.00        & 12.00       & 1.60            & \cite{anderson1992}     \\  
   Measles     & England and Wales                  & 1950-1968                          & 17.00      & 16.00          & 18.00         & \cite{anderson1992}     \\
-  Measles     & Kansas, USA                  & 1918-1921                          & 5.50    & 5.00          & 6.00          & \cite{anderson1992}     \\
-  Measles     & Ontario, Canada                  & 1912-1913                         & 11.50        & 11.00           & 12.00            & \cite{anderson1992}     \\
-  Measles     & Willesden, England                  & 1912-1913                          & 11.00        & 12.00       & 1.60            & \cite{anderson1992}     \\
     Measles     & Ghana                 & 1960-1968                          & 14.50       & 14.00          & 15.00            & \cite{anderson1992}     \\
-  Measles     & Eastern Nigeria                  & 1960-1968                          & 16.50        & 16.00          & 17.00           & \cite{anderson1992}     \\
+  Measles     & USA - Kansas                  & 1918-1921                          & 5.50    & 5.00          & 6.00          & \cite{anderson1992}     \\
 \bottomrule
@@
                           (c) edge node {} (e) ;
-
 \end{tikzpicture}
-\caption{Depiction of a SIR model where $X=S, Y=I,$ and $Z=R$.  One can only get the disease once in this model.}\label{fig::sir}
-\end{figure}
+\caption{Depiction of a SIR model where $X=S, Y=I,$ and $Z=R$.  One can only be infected once by the disease once in this model.}\label{fig::sir}
+\end{figure}
+Since the number of indivudals ($N$) is constant, then
+\begin{align*}
+\frac{dX}{dt} + \frac{dY}{dt} + \frac{dZ}{dt} \equiv 0.
+\end{align*}
 An outbreak occurs if the rate of change of infectious individuals is positive, $\frac{dY}{dt} > 0$, or equivalently,
@@
   \begin{align}\label{eq:sir-noise}
-    X_{obs}(t) &= X(t) + \epsilon_{X,t}\\
-    Y_{obs}(t) &=  N - X_{obs}(t) -Y_{obs}(t)  \nonumber\\
-    Z_{obs}(t) &= Z(t) - \epsilon_{Z,t}. \nonumber
+    \hat{X}(t) &= X(t) + \epsilon_{X,t}\\
+    \hat{Y}(t) &=  N - \hat{X}(t) -\hat{Y}(t)  \nonumber\\
+    \hat{Z}(t) &= Z(t) - \epsilon_{Z,t}. \nonumber
   \end{align}
-We are assuming the observations are generated based on the discrete-time,  deterministic ODEs presented in Equation \ref{eq:sir} with time and compartment dependent noise $\epsilon_{X,t}$ and $\epsilon_{Z,t}$.
-Since, $N$, the total population, is constant, then $Y_{obs}$ is adjusted accordingly.
+The ``hats'' in Equation \ref{eq:sir-noise} are used to distinguish real, observed data from the ``true'' model without hats.
+We are assuming the observations are generated based on the  deterministic ODEs (possibly converted to discrete time) presented in Equation \ref{eq:sir} with time and compartment dependent noise $\epsilon_{X,t}$ and $\epsilon_{Z,t}$.
+Since, $N$, the total population, is constant, then $\hat{Y}$ is adjusted accordingly.
 
@@
 \begin{align*}
-  \textnormal{Data} &= \left \{\left (X_{obs}(t), Y_{obs}(t), Z_{obs}(t) \right ) : t=0, 1, \dots, T\right \}, \\
+  \textnormal{Data} &= \left \{\left (\hat{X}(t), \hat{Y}(t), \hat{Z}(t) \right ) : t=0, 1, \dots, T\right \}, \\
   \hat{\rr} &= m(\textnormal{Data}),
@@
 At any rate, we would like to be able to estimate $\hat{a}$ and $\hat{b}$ for a given $\alpha$-level.
-We could also make similar hypothesis tests such as that the $\rr$ for H1N1 influenza is greater than the reported $\rr$ for seasonal influenza.
+We could also make similar hypothesis tests such as  $\rr$ for H1N1 influenza is greater than the reported $\rr$ for seasonal influenza.
 
@@
 \begin{align*}
-  \frac{dX}{dt} &= - \frac{\beta XY}{N} \\
-  \frac{dE}{dt} &= \frac{\beta XY}{N}  - \mu E\\
-  \frac{dY}{dt} &= \mu E - \gamma Y \\
-  \frac{dZ}{dt} &= \gamma Y,
+  \frac{dX}{dt} &= - \frac{\beta_{XEYZ} XY}{N} \\
+  \frac{dE}{dt} &= \frac{\beta_{XEYZ} XY}{N}  - \mu_{XEYZ} E\\
+  \frac{dY}{dt} &= \mu_{XEYZ} E - \gamma_{XEYZ} Y \\
+  \frac{dZ}{dt} &= \gamma_{XEYZ} Y,
 \end{align*}
 again, with a constant population of $N$ and with known initial values $(X(0), E(0), Y(0), Z(t))$.
-The parameter $\mu$ is the rate at which latently exposed individuals become infectious.
-As $\frac{1}{\mu} \to 0$, the XEYZ model converges to the XYZ model.
-The reproduction number for the XEYZ model, is the same as in the XYZ model, that is $\rr = \frac{\beta}{\gamma}$.
+The parameter $\mu_{XEYZ}$ is the rate at which latently exposed individuals become infectious.
+The subscript ``XEYZ'' is used to emphasize that the disease parameters (especially $\beta$ and $\gamma$) are different than of those in the XYZ model.
+The reproduction number for the XEYZ model, is $\rr = \frac{\beta_{XEYZ}}{\gamma_{XEYZ}}$, compared to the reproduction number for the XYZ model, $\rr = \frac{\beta_{XYZ}}{\gamma_{XYZ}}$.
 
@@
 This generates data set
-$$D_{\mu, \textnormal{XEYZ}} = \left \{(X_{XEYZ}(t), E_{XEYZ}(t), Y_{XEYZ}(t), Z_{XEYZ}(t): t=1, \dots, 365\right \},$$
+$$D_{\mu, \textnormal{XEYZ}} = \left \{(\hat{X}_{XEYZ}(t), \hat{E}_{XEYZ}(t), \hat{Y}_{XEYZ}(t), \hat{Z}_{XEYZ}(t): t=1, \dots, 365\right \},$$
 which is observed by researcher 1.
 Now assume that researcher 2 misspecifies the model for this data as a SIR (XYZ) model rather than a SEIR (XEYZ) model.
-That is, she observes all non-infectious individuals as susceptibles and generates data set
-$$D_{\mu, XYZ} = \left \{X_{XYZ}(t) = X_{XEYZ} + E_{XEYZ}, Y_{XYZ} = Y_{XEYZ}, Z_{XEYZ} = Z_{XYZ} : t = 1, \dots, 365 \right \}.$$
+That is, she observes all non-infectious individuals as susceptibles and generates data sets
+$$D_{\mu, XYZ} = \left \{\hat{X}_{XYZ}(t) = X_{XEYZ} + E_{XEYZ}, \hat{Y}_{XYZ} = Y_{XEYZ}, \hat{Z}_{XEYZ} = Z_{XYZ} : t = 1, \dots, 365 \right \}.$$
 This situation is illustrated in Figure \ref{fig:sir-vs-seir} with $\mu = 0.01$.
@@
 Both researchers want to estimate $\rr$ and so fit models to their data sets.
-The details of how the models are fit are not so important at the moment (but will be in the later sections), but to fairly compare one model to another assume both researchers use ratio estimator to fit the models.
+The details of how the models are fit are not so important at the moment (but will be in the later sections), but to fairly compare one model to another assume both researchers use a least squares estimates to estimate the disease parameters and then use the ratio estimator (RE) as an estimate of $\rr$.
 That is researcher 1 estimates $\rr$ with
@@
 \end{align*}
-Then the ratio estimator (RE) for $\rr^{XYZ}$ is given by Equation \ref{eq:sirls},
+Then the RE for $\rr^{XYZ}$ is given by Equation \ref{eq:sirls},
 \begin{align*}
@@
 We plot the two estimates of $\rr$ as a function of $\mu$ in Figure \ref{fig:r0-alpha}.
-We should note we also estimated the standard errors of the estimates of $\rr$, but they do not show up on the figure because they are of an order of magnitude of $10^{-4}$.
+We should note we also estimated the standard errors of the estimates of $\rr$, but they do not show up on the figure because they are of the order of $10^{-4}$.
 Unsurprisingly, under the SEIR model speficiation, $\hat{\rr}=2$ because there is no noise in the data, and the data is generated from an SEIR model.
-However, as we see in Figure \ref{fig:r0-alpha} the RE of $\rr$ from the SIR model for small values of $\mu$ can result in double the size of $\rr$!
+However, as we see in Figure \ref{fig:r0-alpha} the RE of $\hat{\rr}$ from the SIR model for small values of $\mu$ (a long latency period) can result in double the size of $\hat{\rr}$!
 
@@
 
-For those readers not convinced by our small example, we offer this mathematical intuition of the next generation model (NGM) described by \cite{diekmann1990} and \cite{diekmann2009}.
-
- The NGM is a generalization of any compartment model at the infection-free steady state ($Y(t)=0$ and all derivatives with respect to time sum equal zero).
+For those readers not convinced by our small example, we offer this mathematical intuition of the next generation model (NGM).
+
+ The NGM is a generalization of any compartment model at the infection-free steady state ($Y(t)=0$ and all derivatives with respect to time sum to zero).
 Originally, introduced by \citep{diekmann1990}, \cite{diekmann2009} posit a recipe to find $\rr$ for a wide class of compartment models, including SIR and SEIR.
@@
 \begin{enumerate}
-\item ``Linearize the infected subsystem of nonlinear ODEs about the infection-free steady state''
-\item Decompose the linearized infected subsystem into $(T + \Sigma )x$ where $T$ is a $m\times m$ matrix of transmissions and $\Sigma$ is a $m \times m$ matrix of transitions.
+\item Take the subsystem of infected individuals (including those who have been exposed) and linearize the subsystem about the infection-free steady state
+\item Decompose the linearized infected subsystem into $(T + \Sigma )x$ where $T$ is a $m\times m$ matrix of \textit{transmissions} and $\Sigma$ is a $m \times m$ matrix of \textit{transitions}.
 \item $\rr$ is the spectral radius (i.e. the dominant eigenvalue) of $-T \Sigma^{-1}$.  
@@
 
-The NGM shows that $\rr$ is the spectral radius of a matrix composed from the transmission of disease from diseased individuals to susceptible individuals and transition of diseased individuals from among the compartments and is indeed a property of the model.
+The NGM shows that $\rr$ is the spectral radius of a matrix which is composed of the transmission of disease to susceptible individuals and transition of individuals among the other compartments.
+$\rr$ is thus a property of the model.
 
@@
 \begin{align*}
-\frac{dY}{dt} &= - \beta Y + \gamma .
-  \end{align*}
-  Then $T= \beta$ (the rate of new infections) and $\Sigma = -\gamma$.  Then $K = \frac{\beta}{\gamma}$ which is also, trivially, the dominant eigenvalue, and hence, $R_0 = \frac{\beta}{\gamma}$.
+\frac{dY}{dt} &= - \beta_{XYZ} Y + \gamma_{XYZ} .
+  \end{align*}
+  Then $T= \beta_{XYZ}$ (the rate of new infections) and $\Sigma = -\gamma_{XYZ}$.
+Then $K = \frac{\beta_{XYZ}}{\gamma_{XYZ}}$ which is also, trivially, the dominant eigenvalue, and hence, $R_0 = \frac{\beta_{XYZ}}{\gamma_{XYZ}}$.
 
@@
   \begin{array}{cl}
-    \frac{dE}{dt} &= \beta Y  - \mu E\\
-    \frac{dY}{dt} &=  \mu E  - \gamma Y \\
+    \frac{dE}{dt} &= \beta_{XEYZ} Y  - \mu_{XEYZ} E\vspace{.5em}\\
+    \frac{dY}{dt} &=  \mu_{XEYZ} E  - \gamma_{XEYZ} Y \\
   \end{array}
@@
   \end{align*}
-  Then $T = \left ( \begin{array}{cc} 0 & \beta \\ 0 & 0  \end{array} \right )$ and $\Sigma = \left ( \begin{array}{cc} -\mu & 0 \\ \mu & - \gamma \end{array} \right )$, and $K =\left ( \begin{array}{cc} 0& \beta / \gamma \\ 0 & 0 \end{array} \right  )$.
-
-  Then the spectral radius of $K$ is $R_0  = \frac{\beta}{\gamma}$, which makes sense as exposed individuals are not contributing to the number of new infections.
-
-  Although the $\rr$ ends up being the same in both the SIR (XYZ) and SEIR (XEYZ) case, the derivations were different because of the addition of the exposed compartment.
+  Then $T = \left ( \begin{array}{cc} 0 & \beta_{XEYZ} \\ 0 & 0  \end{array} \right )$ and $\Sigma = \left ( \begin{array}{cc} -\mu_{XEYZ} & 0 \\ \mu_{XEYZ} & - \gamma_{XEYZ} \end{array} \right )$, and $K =\left ( \begin{array}{cc} 0& \beta_{XEYZ} / \gamma_{XEYZ} \\ 0 & 0 \end{array} \right  )$.
+
+  Then the spectral radius of $K$ is $R_0  = \frac{\beta_{XEYZ}}{\gamma_{XEYZ}}$, which makes sense as exposed individuals are not contributing to the number of new infections.
+
+  Although the $\rr$ ends up being similar (infection rate over the recovery rate) for both the SIR (XYZ) and SEIR (XEYZ) models, the derivations are different because of the addition of the exposed compartment.
 In general, more complicated models will result in different derivations of $T$ and $\Sigma$ and hence different derivations of $\rr$.
@@
 \begin{align*}
-(\hat{\beta}, \hat{\gamma} )&=\arg \min_{\beta, \gamma} \sum_{t} \left [ \left (X_{obs}(t) - X(t)\right )^2 + \left ( Z_{obs}(t) - Z(t) \right )^2 \right ]
+(\hat{\beta}, \hat{\gamma} )&=\arg \min_{\beta, \gamma} \sum_{t} \left [ \left (\hat{X}(t) - X(t)\right )^2 + \left ( \hat{Z}(t) - Z(t) \right )^2 \right ]
 \end{align*}
@@
 
-The estimates from this method are often found via a grid search of $\beta$ and $\gamma$ values to minimize the squared error.
+The estimate resulting from this method are often found via a grid search of $\beta$ and $\gamma$ values to minimize the squared error.
 The estimates can also be found using more sophisticated optimization algorithms.
@@
 Another answer is that $L_2$ is a continuously differentiable function and hence is easier to compute, which allows for sensitivity analysis to  be conducted more easily.
-Finally, RE can be done without writing down explicit assumptions about the noise within the model.
+Finally, the RE estimate can be found without writing down explicit assumptions about the noise within the model.
 However, we cannot guarantee properties of consistency or convergence to a known distribution without more explicit assumptions.
 
-On the other hand, ratio estimator is easy to implement using software such as \texttt{R}'s \texttt{optim()} function, and as we see in Section \ref{sec:results} produces results comparable methods to other methods.
-ratio estimator and similar variants have been used to estimate $\rr$ in \cite{majumder2016}.
+On the other hand, the RE is easy to implement using software such as \texttt{R}'s \texttt{optim()} function, and as we see in Section \ref{sec:results} produces results comparable to other methods.
+RE and similar variants have been used to estimate $\rr$ in \cite{majumder2016}.
 
@@
 It may be surprising that this method leads to different results than simply using the ratio estimator.
-One reason we obtain different results may be attributed to better numerical precision as there is one fewer step to obtain an estimate for $\rr$.
-However, we also see that this method further results in different standard errors than in RE, which we discuss further in Section \ref{sec:results}.
-
-Like RE, while we have no explicit parametric assumptions on the noise, we also are uncertain of any theoretical properties of our resulting estimator.
-Likewise, the same difficulties in sensitivity analysis arise as in RE.
+We also see that this method results in different standard errors than in RE, which we discuss further in Section \ref{sec:results}.
+
+Like RE, while we have no explicit assumptions on the noise, we also are uncertain of any theoretical properties of our resulting estimator.
+Likewise, the same difficulties in sensitivity analysis arise as for the RE.
 
@@
 In addition to this, data collected from real diseases are typically very noisy to begin with.
-\cite{chang2017} showed that the SIR model may be well approximated by a linear model.
+The SIR model may be approximated by a linear model.
 We use this method to estimate $\rr$.
@@
   An advantage of using this estimation method is that it is simple to implement using any linear modelling software.
-On the other hand, we assume $X(t) = \hat{X}(t)$ and so if the data truly follows the SIR model, then we will likely have a biased result for $\rr$.
+On the other hand, we assume $X(t) = \hat{X}(t)$ and so if the data truly follows the SIR model, then we will likely have a result for $\rr$.
 Using this method, we are able to check how sensitive $\rr$ is to the degree of the fitted polynomial.
@@
 
-In terms of data from the SIR model, the incidence $J(t) \approx -(X(t+1) - X(t))$, and the IPR$(t) = \frac{J(t)}{Y(t)}$.
+In terms of data from the SIR model, the incidence is approximately the number of new infectious for a given time step, $J(t) \approx -(X(t+1) - X(t))$, and the IPR is the ratio of incidence to prevalence, IPR$(t) = \frac{J(t)}{Y(t)}$.
 This method assumes that we have some prior knowledge about $\gamma$, the recovery rate.
@@
 Again, we are using only one time point to estimate $\rr$.
-This model is agnostic to assumptions on the noise, but again, it is unclear if this is an unbiased or consistent estimator of $\rr$.
+This model is agnostic to assumptions on the noise, but the estimate is biased, as we show in Section \ref{sec:results}.
 
 \subsection{Smoothed Incidence to Prevalence Ratio (SIPR)}
-We use the same method as above, IPR, but first estimate splines with $K$ degrees of freedom, where $K$ is small, $\hat{X}(t)$ and $\hat{Y}(t)$, to fit to $X_{obs}$ and $Y_{obs}$, respectively.
+We use the same method as above, IPR, but first smooth the epidemic curves $\left(\hat{X}(t), \hat{Y}(t) \right)$ (we use B-splines with $K=4$ degrees of freedom).
 Then  $J(t) \approx -(\hat{X}(t+1) - \hat{X}(t))$, and the $\hat{\textnormal{IPR}}(t) = \frac{J(t)}{\hat{Y}(t)}$.
@@
 \end{align*}
-Splines are used because they typically can be used to make good approximations of epidemic curves (CITE).
-However, the researcher could use their method of choice to estimate $\hat{X}(t)$ and $\hat{Y}(t)$.
+We use splines to smooth the curves because they  have proven to be good approximations of epidemic curves \citep{brooks2015}.
+However, the researcher can use their method of choice to estimate $\hat{X}(t)$ and $\hat{Y}(t)$.
 An advantage of this method is that it creates a less variable estimate than estimating IPR using only one point.
@@
 We can regress the left hand side in Eq.
-\ref{eq:harko_lin} on $Z(t)$ (using least squares to estimate the coefficients), the number of recovered individuals, with $\rr$ as the coefficient of $Z(t)/N$ and with an intercept term of $\beta_0=0$ to obtain an estimate of $\rr$,
+\ref{eq:harko_lin} on $Z(t)$ (using least squares to estimate the coefficients), the number of recovered individuals, with $\rr$ as the coefficient of $Z(t)/N$ to obtain an estimate of $\rr$,
 \begin{align*}
@@
 
-In this method, the number of susceptibles at the next step, $X_{obs}(t+1)$, has a Binomial distribution based on the contacts with the current number of infectious, $Y(t)$ and the current number of susceptibles.
-That is $X_{obs}(t+1) \sim \text{Binomial}\left(X_{obs}(t), \alpha^{Y_{obs}(t)}\right)$, where $\alpha$ is the probability of avoiding infection from an infective.
+In this model, the number of susceptibles at the next step, $\hat{X}(t+1)$, has a Binomial distribution based on the contacts with the current number of infectious, $\hat{Y}(t)$ and the current number of susceptibles.
+That is $\hat{X}(t+1) \sim \text{Binomial}\left(\hat{X}(t), \alpha^{\hat{Y}(t)}\right)$, where $\alpha$ is the probability of avoiding infection from an infective.
 One can write out the likelihood for $\alpha$ in this model, namely,
 \begin{align*}
-\mathcal{L}\left ( \alpha ; \textnormal{data}\right ) \propto \prod_{t=0}^{T-1}\left ( \alpha^{Y_{obs}(t)} \right )^{X_{obs}(t+1)} \left (1- \alpha^{Y_{obs}(t)} \right )^{X_{obs}(t) - X_{obs}(t+1)}.
+\mathcal{L}\left ( \alpha ; \textnormal{Data}\right ) \propto \prod_{t=0}^{T-1}\left ( \alpha^{\hat{Y}(t)} \right )^{\hat{X}(t+1)} \left (1- \alpha^{\hat{Y}(t)} \right )^{\hat{X}(t) - \hat{X}(t+1)}.
 \end{align*}
@@
   \end{align*}
- \cite{barbour2004} report that the estimated reproduction number is thus,
+ \cite{barbour2004} report that the estimated reproduction number is then,
 \begin{align}\label{eq:r0-mc}

### 2018-07-14 paper/Version15/draft_v15.tex
Added Bill edits to confidence intervals

@@
 When the method estimates \(\beta\) and \(\gamma\) instead of \(\rr\) directly, we use the delta method approximation to estimate the
-variance of \(\rr\).   From \cite{wasserman2004}, if
-\begin{align*}
-  \sqrt{O} \left ( \hat{\mathbf{\theta}}_o - \mathbf{\theta} \right ) \overset{d}{\to} N\left ( \mathbf{0}, \Sigma \right  )
+variance of \(\rr\).   From \cite{wasserman2004}, if we have $\hat{\theta}_T$, an estimate of some parameter(s) $\theta$, dependent on the data and number of observations $T$ and
+\begin{align*}
+  \sqrt{T} \left ( \hat{\mathbf{\theta}}_n - \mathbf{\theta} \right ) \overset{d}{\to} N\left ( \mathbf{0}, \Sigma \right  )
 \end{align*}
@@
 \begin{align*}
-  \sqrt{O} \left (h(\hat{\theta}_n) - h(\theta) \right ) \overset{d}{\to} N \left (0,  \nabla^T h(\theta)\Sigma \nabla h(\theta)\right ),
-\end{align*}
-where $\nabla h$ is the gradient of $h$.
+  \sqrt{T} \left (h(\hat{\theta}_T) - h(\theta) \right ) \overset{d}{\to} N \left (0,  \nabla^T h(\theta)\Sigma \nabla h(\theta)\right ),
+\end{align*}
+where $\nabla h$ is the gradient of $h$ and $\Sigma$ is some variance-covariance matrix..
 
@@
 \begin{align*}
-  V[\rr] &\to \nabla h^T(\beta, \gamma) \Sigma_{\beta, \gamma} \nabla h(\beta, \gamma) \\
-  &= \frac{\Sigma_{\beta, \beta}}{\gamma^2} - \frac{\Sigma_{\beta, \gamma}(1 + \beta)}{\gamma^3} + \frac{\beta \Sigma_{\gamma, \gamma}}{\gamma^4}
-\end{align*}
-where $\Sigma_{\beta, \gamma}$ is covariance matrix of \(\beta\) and \(\gamma\).
-.
+  V[\rr] &\approx \frac{1}{T} \left ( \nabla h^T(\beta, \gamma) \Sigma_{\beta, \gamma} \nabla h(\beta, \gamma) \right ) \\
+  &= \frac{1}{T} \left ( \frac{\Sigma_{\beta, \beta}}{\gamma^2} - \frac{\Sigma_{\beta, \gamma}(1 + \beta)}{\gamma^3} + \frac{\beta \Sigma_{\gamma, \gamma}}{\gamma^4} \right )
+\end{align*}
+where $\Sigma_{\beta, \gamma}$ is covariance matrix of $\hat{\beta}$ and $\hat{\gamma}$.
 We can use plug-in estimators for $\hat{\beta}$, $\hat{\gamma}$, and $\hat{\Sigma}$ where $\hat{\Sigma}$ is taken to the inverse of the Hessian from the least squares optimzation of $\beta$ and $\gamma$.
@@
 
-We can also use the Delta method to derive variance estimates of LMA, LMAT, LL, and MC as for all these estimates first estimate parameters $\theta$ since $\rr= h(\theta)$
+We can also use the Delta method to derive variance estimates of LMA, LMAT, LL, (which in this case is equivalent to the standard error derived from the typical linear regression framework).
+We can also use the Delta method to derive an estimate for MC, where $\hat{\sigma}$ is the estimate of the variance of the $\hat{\alpha}$,
+\begin{align*}
+h(\alpha) &= \log \left ( \frac{1}{1-\alpha}\right ) \\
+  h^\prime(\alpha) &= \alpha - 1 \\
+                     V[ \hat{\rr} ] & \approx \frac{1}{T}  \left (\frac{\hat{\sigma}}{(\hat{\alpha} - 1)^2} \right ).
+\end{align*}
 
@@
 These are non-overlapping partitions or blocks, although there are variants for which blocks can overlap.
-For iteration $\ell$, one samples $k$ blocks uniformly with replacement and estimates $\hat{\rr}_b$, repeating the sampling and estimation for a total of $L$ estimates of $\hat{\rr}$.
+For iteration $\ell$, one samples $k$ blocks uniformly with replacement and estimates $\hat{\rr}_{,\ell}$, repeating the sampling and estimation for a total of $L$ estimates of $\hat{\rr}_{,\ell}$.
 Then
 \begin{align*}
-  V\left [ \hat{\rr} \right ] &\approx V\left [\hat{\rr}_\ell \right ].
+  V\left [ \hat{\rr} \right ] &\approx V\left [\hat{\rr}_{,\ell} \right ].
 \end{align*}
@@
 The asympotic properties hold when the stochastic process is $m$-dependent, that is when observations $t, t+1, \dots$ and $s, s+1, \dots$ are independent whenever $s+m < t$ and under some smoothness conditions on the statistic used to estimate $\rr$.
-
-A choice one must make in the block bootstrap is the block size $b$.  One simple recommendation is to take $b= n^{1/3}$, which is derived from the asymptotic mean square error of the method.
+Intuitively, this means if two observations are at least length $b$ apart (in terms of observation sequence), then one observation has no effect on the other.
+The conventional bootstrap is recovered by setting $b=1$.
+
+A choice one must make in the block bootstrap is the block size $b$.  The asymptotic mean squared error of the estimate is
+\begin{align*}
+\frac{1}{n^2} \left (\frac{C_1}{b^2} + \frac{C_2b}{n} \right ),
+\end{align*}
+where $C_1$ and $C_2$ are unknown constants which are dependent on the estimation problem.  \cite{cao1999} reports then that the optimal block size is order $n^{1/3}$.
 

### 2018-07-14 paper/Version15/draft_v15.tex
Finished Bill edits.  Need to look at H1N1 data again

@@
 We compare the performance of the SIR models on simulated data from both the SIR as well as other models.
-In all of these simulations, we first generate data from the models under known conditions, then we add noise drawn from known distributions to each time point,
+In all of these simulations, we first generate data from the models under known functions ($f$, $g$), then we add noise drawn from known distributions to each time point $(\epsilon_{X,t}, \epsilon_{Y,t})$,
 \begin{align}\label{eq:sim-models}
-  X_{model}(t) &= f(t) + \epsilon_{X,t} \\
-  Y_{model}(t) &= g(t) + \epsilon_{Y,t} \nonumber\\
-  Z_{model}(t) &= N - X_{model}(t) - Y_{model}(t)\nonumber 
+  \hat{X}(t) &= f(t) + \epsilon_{X,t} \\
+  \hat{Y}(t) &= g(t) + \epsilon_{Y,t} \nonumber\\
+  \hat{Z}(t) &= N - \hat{X}(t) - \hat{Y}(t)\nonumber 
 \end{align}
@@
 
-\noindent \textbf{Gaussian (Norm) ($X_{G}$, $Y_{G}$, $Z_{G}$)}
-\begin{align*}
-  f(t) &= X(t) \\
-  g(t) &= Y(t) \\
+\noindent \textbf{Gaussian (Norm) ($\hat{X}_{G}$, $\hat{Y}_{G}$, $\hat{Z}_{G}$)}
+\begin{align*}
+  f(t) &= X_{XYZ}(t) \\
+  g(t) &= Y_{XYZ}(t) \\
   \epsilon_{X,t} &\sim N(0, \sigma_X^2) \\
@@
 \end{align*}
-\textbf{Autoregressive (AR) ($X_{AR}$, $Y_{AR}$, $Z_{AR}$)}
-\begin{align*}
-  f(t) &= \rho X_{obs}(t-1) \\
-  g(t) &= \rho Y_{obs}(t-1) \\
+\textbf{Autoregressive(1) (AR) ($\hat{X}_{AR}$, $\hat{Y}_{AR}$, $\hat{Z}_{AR}$)}
+\begin{align*}
+  f(t) &= \rho \hat{X}(t-1) \\
+  g(t) &= \rho \hat{Y}(t-1) \\
   \epsilon_{X,t} &\sim N(0, \sigma_X^2) \\
@@
 
-\noindent \textbf{Gaussian-PAVA (Norm-M) ($X_{GP}$, $Y_{GP}$, $Z_{GP}$)}
+\noindent \textbf{Gaussian-PAVA (Norm-M) ($\hat{X}_{GP}$, $\hat{Y}_{GP}$, $\hat{Z}_{GP}$)}
 \begin{align*}
@@
 \end{align*}
-\textbf{Autoregressive-PAVA (AR-M) ($X_{ARP}$, $Y_{ARP}$, $Z_{ARP}$)}
+\textbf{Autoregressive-PAVA (AR-M) ($\hat{X}_{ARP}$, $\hat{Y}_{ARP}$, $\hat{Z}_{ARP}$)}
 \begin{align*}
@@
 \end{align*}
-This gives us a total of four error structures for each set of conditions: (1) Gaussian (Norm), (2) Gaussian monotone (Norm-M),  (3) autoregressive (AR), and (4) autoregressive monotone (AR-M) for a given set of parameters $\left (\beta, \gamma, O, X(0), Y(0), Z(0), \sigma_X, \sigma_Y, N \right )$, the infection rate; recovery rate; total number of time points; initial conditions for number of susceptibles, infectious, and recovered individuals, the standard deviation for both the susceptible and infectious compartments, and the total population, respectively.
+This gives us a total of four error structures for each set of conditions  (1) Gaussian (Norm), (2) Gaussian monotone (Norm-M),  (3) autoregressive (AR), and (4) autoregressive monotone (AR-M) for a given set of parameters $\left (\beta, \gamma, T, X(0), Y(0), Z(0), \sigma_X, \sigma_Y, N \right )$, the infection rate; recovery rate; total number of time points; initial conditions for number of susceptibles, infectious, and recovered individuals; the standard deviations for both the susceptible and infectious compartments; and the total population, respectively.
 
@@
 
-\textbf{Baseline Simulated Data}
+\noindent \textbf{Baseline Simulated Data}
 \begin{center}
@@
       \end{center}
-      The baseline data set is generated from the deterministic SIR curves in Figure \ref{fig:baseline-data} and then noise is added to two of the compartments, and the third is adjusted to maintain a constant population value.
-In this scenario, we see that the infection begins slowly but over the course of a year infects over 75\% of the population.
+      The baseline data set is generated by adding noise to the susceptible and infectious curves from  deterministic SIR, shown in Figure \ref{fig:baseline-data}.
+The recovered compartment is then adjusted to maintain a constant population value.
+In this scenario, we see that the infection begins slowly but over the course of a year infects nearly 75\% of the population.
       \begin{figure}
@@
 The infection rate and recovery rate ($\beta, \gamma$) are relatively small.
-The duration of the disease, under the true model, spans nearly a whole year (see Figure \ref{fig:baseline-data}) and so we use a year's worth of daily observations of (X,Y, Z) values.
-Initially, 50 (0.05\%) diseased individuals are introduced to the population of 100,000 individuals (approximately the average population size of a US county in 2015).
-The standard deviation of the X values is 0.1\% of the magnitude of the initial number of susceptible individuals, and the standard deviation of the Y values is 10\% of the magnitude of the initial number of infectious individuals, although in this setting the number of infectious individuals increases rapidly.
-We use Gaussian error as our baseline error because of its known statistical properties so we can more easily compare methods to one another.
+The duration of the disease, under the true model, spans nearly a whole year (see Figure \ref{fig:baseline-data}) and so we use a year's worth of daily observations of ($\hat{X},\hat{Y}, \hat{Z}$) values.
+Initially, 50 (0.05\%) infectious individuals are introduced to the population of 100,000 individuals (approximately the average population size of a US county in 2015).
+The standard deviation of the $\hat{X}$ values is 0.1\% of the magnitude of the initial number of susceptible individuals, and the standard deviation of the $\hat{Y}$ values is 10\% of the magnitude of the initial number of infectious individuals, although in this setting the number of infectious individuals increases rapidly.
+We use Gaussian error as our baseline error because of its well known statistical properties so we can more easily compare methods to one another.
 Later on, we explore the autoregressive and PAVA simulated data, which may better simulate data from  a real epidemic.
@@
 
-We then generate different SIR data sets, changing one of the conditions for simulation each time and using the baseline SIR data as our basis for comparison.
+After generating our baseline data set, we  generate different SIR data sets, changing one of the conditions for simulation each time and using the baseline SIR data as our basis for comparison.
 The different data sets for a given error structure are shown in Table \ref{tab:simulated-data}.
-We repeat  Table \ref{tab:simulated-data}, replacing the ``Error'' with  one of the other three error structures for a total of 96 simulated data sets.
+We repeat  Table \ref{tab:simulated-data}, replacing the ``Error'' feature with  one of the other three error structures for a total of 96 simulated data sets.
 
@@
 
-To test how well the SIR models perform under misspecification, we simulate data from three other models.
-While we will not have a ``known'' $\rr$ value to compare our results to, this will allow us to see if the models give us reasonable (and comparable) results when the data do not come from the SIR model.
-We simulate data from polynomials of degree 1 and 4 with respect to the time step, using a monotonically increasing function for the $X$ compartment and a monotonically decreasing function for the $Z$ compartment.
+To test how well the $\rr$ estimation methods perform when the underlying model is different from the SIR model, we simulate data from three other models.
+This will allow us to see if the models give us reasonable (and comparable) results when the data do not come from the SIR model.
+We simulate data from polynomials in $t$, the time step, of degree 1 and 4, using a monotonically increasing function for the $X$ compartment and a monotonically decreasing function for the $Z$ compartment.
 The sum of these is subtracted from the constant total population to obtain the $Y$ compartment.
@@
 
-\subsection{Model Fitting Details}\label{sec:model-fit}
+%\subsection{Model Fitting Details}\label{sec:model-fit}
 
 Table \ref{tab:se-methods} shows the ways that standard errors are acquired for the different methods.
-The IPR, SIPR, and IDEA models require knowledge of the length of the average infectious period.
+The IPR, and SIPRknowledge of the length of the average infectious period $\gamma^{-1}$.
 For the results below, we assume the best case scenario that the true average infectious period is known for the data set.
-The exponential growth model requires knowledge of the probability distribution of the infectious period.
-In this case, we assume a uniform probability distribution with a mean equivalent to the true average infectious period.
 
@@
 Among these include values of both the infection and recovery rate $(\beta, \gamma)$, the number of data points available ($T$), the total size of the population ($N$), the initial percent of susceptible individuals $\left (\frac{X(0)}{N}\right)$, the initial percent of infectious individuals $\left (\frac{Y(0)}{N}\right )$, the magnitude of the variance of both the number of susceptibles and the number of infectious ($\sigma_X^2, \sigma_Y^2$), the error modeling assumptions, data generated from a non-SIR model, and finally methods of estimating the variance.
-For the sake of length, we do not provide the full results here, and instead examine the results at a high-level.
+For the sake of brevity, we do not provide the full results here, and instead examine the results at a high-level.
 However, for those interested readers, we provide results in the form of forest plots and tables of estimates of $\rr$ for each data set in the Supplementary Material.
@@
 Err.
-The vertical line is the true $\rr$ value.}\label{fig:baseline-res}
+The vertical, dashed line is the true $\rr$ value.}\label{fig:baseline-res}
   \end{figure}
@@
 The methods of RE, rRE, LMA, LMAT, IPR, and LL all result in a CI that covers the true value of $\rr=2$.
-Of these, LMAT and IPR result in a CI which contains 1, so we would not be able to say whether or not we would have an outbreak of a disease or not.
+Of these, LMAT and IPR result in a CI which contains 1, so we would not be able to say whether or the outbreak of a disease was due to random chance or not.
 On the other hand, many methods appear to underestimate the true $\rr$, and many of them have larger standard errors as well.
@@
 
-The results of the baseline data set with normal error are a good summary of the methods as a whole.
+From examining the full results in the Supplementary Material, the results of the baseline data set with normal error seem to be a good summary of the methods as a whole.
 In Figure \ref{fig:baseline-res}, we see that the clear best methods to estimate $\rr$ are RE, rRE, and LL.
-These three methods consistently outperform the other methods of accurately estimating $\rr$ with a reasonably small CI when varying a single parameter of our generated data sets and holding the others as fixed.
-
-In particular, MC, SB almost never capture the true estimate and typically underestimate the true value of $\rr$.
-This leads us to believe that these models are misspecified.
-Both models rely on assumptions of simple underlying statistical distributions for the course of infection (binomial and Poisson, respectively).
-These are good examples of why we need to consider other reasons for using a model other than the math is nice.
-That is not to say that variants of these models are unviable for modelling $\rr$, but they do not make good defaults.
+These three methods consistently outperform the other methods of accurately estimating $\rr$ with a reasonably small CI when varying a single parameter and holding the others as fixed from the generated data sets.
+
+In particular, MC and SB almost never capture the true estimate and typically underestimate the true value of $\rr$.
+This leads us to believe that these models are inadequate in capturing the effects of a disease following the SIR model.
+Both models rely on parametric assumptions of simple underlying statistical distributions for the course of infection (binomial and Poisson, respectively) with time-invariant transmission parameters ($\alpha$ and $\lambda$).
+These are good examples of why we need to consider other reasons for using a model other than the math resulting in a simple estimate.
+That is not to say that variants of these models are unviable for modelling $\rr$, but they do not make for good defaults.
 It is also an issue that these two methods consistently underestimate $\rr$, which could possibly lead to an unchecked epidemic.
@@
 
-However, we do see instances when LL completely outperforms RE and rRE such as in the case where the initial percent of infectious is $\frac{I(0)}{N} \cdot 100\%$ $< .01$\%.
-In Figure \ref{fig:ar-small-I}, we examine the data set generated with autoregressive monotonic error, our most stringent data, with varying levels of initial percent of infectious individuals.
+However, we do see instances where LL completely outperforms RE and rRE such as in the case where the initial percent of infectious $\frac{I(0)}{N} \cdot 100\%$ is $< .01$\%.
+In Figure \ref{fig:ar-small-I}, we examine the data set generated with autoregressive monotonic error, our most stringent data structure, with varying levels of initial percent of infectious individuals.
 When $\frac{I(0)}{N} \cdot 100\%$ $< .01$\%, RE and rRE tend to overestimate $\rr$ and have large CIs.
@@
 
-One interesting feature we see in the results is that RE and rRE can result in very different estimations and CIs, which may seem unusual since rRE is a simple reparametrization of RE.
+\begin{figure}[H]
+	\centering
+	\includegraphics[scale=0.5]{images/start_arm.jpeg}
+	\caption{Forest plots of estimates of $\rr$ for the \xxsir methods for the Inits1-4 data sets $(\beta=.06, \gamma=.03, T=365,  Z(0)=0, \sigma_X=100, \sigma_Y=5, N=10^5)$.
+The points are the estimates of $\rr$ and the lines denote $\pm 2\cdot $Std.
+Err.
+The vertical line is the true $\rr$ value.}\label{fig:ar-small-I}
+\end{figure}
+
+One interesting feature we see in the results is that RE and rRE can result in very different estimations and CIs, which may seem unintuitive since rRE is a simple reparametrization of RE.
 For example, in Figure \ref{fig:ar-r0}, we see that when the true value of $\rr$ is 1, rRE has a much smaller CI than RE, which may be one reason to prefer rRE to RE.
@@
 %% robust to different error assumptions
-Reassuringly, we find that $\hat{\rr}$ seems to be generally robust to the error structure.
+ In Figure \ref{fig:err-rob} we plot the standardized residuals of $\rr$ focusing on the three best methods: RE, rRE, and LL.
+Reassuringly, we find that $\hat{\rr}$ estimates from RE and rRE seem to be generally robust to the error structure.
+((What's up with varying $\rr$ values??))
 
@@
 	\includegraphics[scale=0.5]{images/err_robust.jpeg}
-	\caption{}\label{fig:err-rob}
+	\caption{Standardized residuals $\frac{\left ( \hat{\rr} - \rr\right ) }{SE \left [\hat{\rr}\right ]}$ for the estimation methods of RE, rRE, and LL faceted by the four error structures.
+The dashed line is the true value of $\rr$.}\label{fig:err-rob}
 \end{figure}
@@
 %% model misspecification is fairly robust
-We also examine circumstances where the SIR is misspecified.  The Poly1 and LSIR data sets result is good estimates of $\hat{\rr}$ but Poly4 results in
+We also examine circumstances where the data is generated from underlying models different from the SIR model.
+The results are plotted in Figure \ref{fig:mod-rob}  All three sets of data sets generated by the non-SIR models result in poor estimates of $\rr$, meaning that even our best estimation techniques fail under model misspecification.
 
@@
 	\includegraphics[scale=0.5]{images/model_robust.jpeg}
-	\caption{}\label{fig:mod-rob}
+	\caption{Standardized residuals $\frac{\left ( \hat{\rr} - \rr\right ) }{SE \left [\hat{\rr}\right ]}$ for the estimation methods of RE, rRE, and LL faceted by the four model structures (SIR, Poly1, Poly4, and LSIR).
+The dashed line is the true value of $\rr$.}\label{fig:mod-rob}
 \end{figure}
@@
 
-We find that when $N$, the number of individuals decreases, then the CIs of our estimates generally increase.
-This is likely due to the central limit theorem (CLT).
-When the noise parameters $\sigma_X$ or $\sigma_Y$ increase then the CIs of our estimates also generally increase.
-For both small $N$ and large $\sigma_X$ or $\sigma_Y$, we find that the point estimates of $\rr$ seem to be unbiased, it is just the CIs that are large.
-
-We find the parameter $T$ to be particularly important.
+Generally, we find that when $N$, the number of individuals decreases, then the CIs of our estimates generally increase.
+This is evidence of the  central limit theorem (CLT) taking effect.
+When the noise parameters $\sigma_X$ or $\sigma_Y$ increase then the CI lengths of our estimates also generally increase.
+For both small $N$ and large $\sigma_X$ or $\sigma_Y$, we find that the point estimates of $\rr$ seem to be unbiased but may have large CIs depending on the other parameters.
+
+We find the parameter $T$ to be particularly important, as shown in Figure \ref{fig:timediff}.
 For one, it can play a role similar to that of $N$, directly giving the researcher more data for a larger $T$.
 Thus, larger $T$ may result in smaller CIs due to the CLT.
-However, $T$ also plays an even greater role in estimating $\rr$.
+However, $T$ also plays an even more important role than $N$ in estimating $\rr$.
+We see in Figure \ref{fig:timediff}, for RE and rRE, about 50 time points are required to accurately capture the true estimate of $\rr$.
+For LL, we see that at least 100 time points are required.
 
@@
 	\includegraphics[scale=0.5]{images/timediff.jpeg}
-	\caption{}\label{fig:timediff}
-\end{figure}
-
-\begin{figure}[H]
-	\centering
-	\includegraphics[scale=0.5]{images/obs.jpeg}
-	\caption{}\label{fig:obs}
-\end{figure}
-
-\begin{figure}[H]
-	\centering
-	\includegraphics[scale=0.5]{images/start_arm.jpeg}
-	\caption{Forest plots of estimates of $\rr$ for the \xxsir methods for the Inits1-4 data sets $(\beta=.06, \gamma=.03, T=365,  Z(0)=0, \sigma_X=100, \sigma_Y=5, N=10^5)$.
+	\caption{Forest plots of estimates of $\rr$ for the \xxsir methods for the baseline data sets with normall error structure $(\beta=.06, \gamma=.03, X(0) = 99950, Y(0)=50, \sigma_X=100, \sigma_Y=5, N=10^5)$.
 The points are the estimates of $\rr$ and the lines denote $\pm 2\cdot $Std.
 Err.
-The vertical line is the true $\rr$ value.}\label{fig:ar-small-I}
+The vertical line is the true $\rr$ value.}\label{fig:timediff}
+      \end{figure}
+
+      In the previous plots, we assume that we are collecting incidence and recovery date every day.
+However, it is also possible to observe weekly, monthly, or even quarterly incidence and recovery data, for example.
+In Figure \ref{fig:obs}, we explore estimates of $\rr$ using the data collected at different time periods for our baseline data structure with a full year's worth of data.
+We find that we can reliably estimate $\hat{\rr}$ regardless of the collection period.
+This is likely due to the previously discussed relationship between $\rr$ and the final size of the epidemic.
+(ADD IN)
+
+However, if we are given only 200 days worth of data, we find that the collection period size does result in different estimates.  PLOT
+
+\begin{figure}[H]
+	\centering
+	\includegraphics[scale=0.5]{images/obs.jpeg}
+	\caption{Forest plots of estimates of $\rr$ for the \xxsir methods for the baseline data sets with normall error structure $(\beta=.06, \gamma=.03, X(0) = 99950, Y(0)=50, \sigma_X=100, \sigma_Y=5, N=10^5)$.
+The points are the estimates of $\rr$ and the lines denote $\pm 2\cdot $Std.
+Err.
+The vertical line is the true $\rr$ value.}\label{fig:obs}
 \end{figure}
@@
 The other methods CI width seems to depend heavily on the initial parameters used.
-Unsurprisingly, an $\sigma_X$, $\sigma_Y$, $T$, or $N$ seems to be positively associated with a larger CI.
+Unsurprisingly, larger $\sigma_X$, $\sigma_Y$ values seem to correspond to larger CI sizes.
+On the other hand, large $N$ and $T$ values seem to correspond to smaller CI sizes.
 Having a larger percentage of initially infected individuals also seems to be associated with a larger CI.
@@
 
-We estimate the variance through the delta method, block bootstrap, and posterior and compare it to the empirical variance of $\rr$ for each of the methods, using 1000 data sets generated from the baseline data structure.
+We estimate the variance through the delta method, posterior, and  block bootstrap and compare it to the empirical variance of $\rr$ for each of the methods, using 1000 data sets generated from the baseline data structure.
 
@@
 
-      \subsubsection{Posterior}
-
       \subsection{Application to USA H1N1 data}\label{sec:real-data}
@@
 
-      \subsubsection{USA H1N1 Influenza 2009 data}
+      \subsubsection{USA H1N1 Influenza A 2009 data}
       The data is from the Center of Disease Control and Prevention's (CDC) FluView \citep{cdc-fluview}.
@@
 Feature       & Notes                                                                   \\ \midrule
-Disease       & H1N1 Pandemic Influenza                                                 \\
+Disease       & H1N1 Pandemic Influenza A                                                 \\ 
 Source        & CDC FluView (\url{https://www.cdc.gov/flu/weekly/})      \\

### 2018-07-20 paper/Version15/draft_v15.tex
Updated references

@@
 We describe nine methods used to estimate $\rr$ and provide a thorough simulation study of how these methods differ in the presence of disease parameters.
-As real world motivation, we analyze the outbreak of the H1N1 pandemic influenza in the USA and compare the results from our nine methods to a previous study.
+As real world motivation, we analyze the 2009 outbreak of the H1N1 pandemic influenza in the USA and compare the results from our nine methods to a previous study.
 We discuss the most important aspects from our results which effect the estimation of $\rr$ and provide guidelines for estimating point estimates and confidence/credible intervals to create reliable, comparable estimates of $\rr$.
@@
 What has been called ``arguably the most important quantity in the study of epidemics'' \citep{Heesterbeek2002},  $\mathcal{R}_0$, the initial reproduction number (by convention pronounced ``R-naught'') is an elusive quantity for epidemic modelers to estimate.
-As defined by \citet{anderson1992}, $\rr$ is the ``the average number of secondary infections produced when one infected individual is introduced into a host population where everyone is susceptible.''  In some ways, $\rr$ summarizes an entire outbreak of a disease; it  is used to assess whether a disease epidemic will occur.
-Additionally, it describes what percentage of the population needs to be vaccinated to avoid such an epidemic, roughly $1-\rr^{-1}$ \citep{anderson1992}).
+As defined by \citet{anderson1992}, $\rr$ is the ``the average number of secondary infections produced when one infected individual is introduced into a host population where everyone is susceptible.''  In some ways, $\rr$ summarizes an entire outbreak of a disease; it is used to assess whether a disease outbreak will occur and its severity.
+Additionally, it describes what percentage of the population needs to be vaccinated to avoid such an epidemic, roughly $1-\rr^{-1}$, tells us the final size of the total number of infected individuals, and describes the probability of which we will observe an outbreak under the same conditions \citep{anderson1992,britton2010}).
 Despite a clear definition of $\rr$, epidemiologists have struggled to create a standard  estimator for $\rr$  \citep{hethcote2000}.
@@
 
-In addition to making a point estimate of $\rr$, we are particularly interested in the hypothesis  $H_0: \rr < 1$, as a value less than 1 indicates that a disease will not outbreak.
-We also discuss techniques for estimating CIs for each of the different methods.
+In addition to making a point estimate of $\rr$, we are particularly interested in the hypothesis  $H_0: \rr < \rr^*$, where $\rr^*$ may be the estimated reproduction number of a past outbreak of the same disease, an outbreak of a different disease in the same region, or a value of 1.
+An estimate of $\hat{\rr}<1$ indicates that the disease will die out.
+We do note that the last situation is situational as we generally only observe diseases that do outbreak and inherently have a value greater than 1.
+However, \cite{britton2010} provides guidance on the probability of observing an outbreak based on $\rr$ (that is estimated final size is approximately the probability of us observing an outbreak)   We also discuss techniques for estimating CIs for each of the different methods.
 These CI estimates are created using the delta method, the block bootstrap, and the posterior distribution \citep{cao1999,wasserman2004}.
@@
 Below in Table \ref{tab:r0-real-ex}, we provide examples of estimates of $\rr$ for specific diseases from publications over the past three decades, spanning various locations over the world.
-We display estimates for HIV, Zika (ZIK-V), Ebola (EVD), seasonal influenza, H1N1 influenza, and measles.
-These estimates of $\rr$ are not limited to those of SIR model, although most results come from this framework.
-Of these, Zika seems to have the largest estimated values of $\rr$, followed by Ebola, H1N1 influenza, seasonal influenza, and HIV.
+We display estimates for HIV, Zika (ZIK-V), Ebola (EVD), seasonal influenza, pandemic H1N1 influenza A, and measles.
+These estimates of $\rr$ are not limited to those of the SIR model, although many of the displayed results originate from this framework.
+Of these, measles seems to have the largest estimated values of $\rr$, followed by Zika, Ebola, H1N1 influenza, seasonal influenza, and HIV.
 We see that African locations seem to have larger estimates of $\rr$ compared to the rest of the world, regardless of the disease.
@@
 \begin{align*}
-  H_0:\;& \rr < 1 \\
-  H_A:\;& \rr \ge 1
+  H_0:\;& \rr < \rr^* \\
+  H_A:\;& \rr \ge \rr^*
 \end{align*}

### 2018-08-23 paper/Version16/draft_v16.tex
Put in dependency for lag function

--- 2018-07-20 paper/Version15/draft_v15.tex
@@
 
-\title{\Wxxsir methods to estimate $\rr$ in the SIR model}
+\title{\Wxxsir methods to estimate $\rr$ in the SIR model with applications to 2009 pandemic influenza}
 \author{ Shannon Gallagher$^{\dag}$, Andersen Chang$^{\ddag}$, and William F.
@@
 For nearly a century, the initial reproduction number ($\rr$) has been used to summarize infectious diseases and to compare them to one another.
-It is difficult to estimate $\rr$ due to assumptions on both how a disease transmits through a population and also due to differences in statistical estimation methods.
-We describe nine methods used to estimate $\rr$ and provide a thorough simulation study of how these methods differ in the presence of disease parameters.
+Difficulties in estimating $\rr$ arise from on both how a disease transmits through a population as well as from differences in statistical estimation methods.
+We describe \wxxsir methods used to estimate $\rr$ and provide a thorough simulation study of how these methods differ in the presence of disease parameters.
 As real world motivation, we analyze the 2009 outbreak of the H1N1 pandemic influenza in the USA and compare the results from our nine methods to a previous study.
@@
 \section{Introduction}\label{sec:intro}
-What has been called ``arguably the most important quantity in the study of epidemics'' \citep{Heesterbeek2002},  $\mathcal{R}_0$, the initial reproduction number (by convention pronounced ``R-naught'') is an elusive quantity for epidemic modelers to estimate.
+What has been called ``arguably the most important quantity in the study of epidemics,''  $\mathcal{R}_0$, the initial reproduction number (by convention pronounced ``R-naught'') is an elusive quantity for epidemic modelers to estimate \citep{Heesterbeek2002}.
 As defined by \citet{anderson1992}, $\rr$ is the ``the average number of secondary infections produced when one infected individual is introduced into a host population where everyone is susceptible.''  In some ways, $\rr$ summarizes an entire outbreak of a disease; it is used to assess whether a disease outbreak will occur and its severity.
@@
 Despite a clear definition of $\rr$, epidemiologists have struggled to create a standard  estimator for $\rr$  \citep{hethcote2000}.
-A major issue in estimating $\rr$ is that the quantity is a \textit{property of the model}, meaning that $\rr$ is dependent not only on the usual noise that comes with statistical modeling but also on a variety of assumptions on how researchers assume a disease is transmitted through a population \citep{diekmann2009}.
+A major issue in estimating $\rr$ is that the quantity is a \textit{property of the model}, meaning that $\rr$ is dependent not only on the usual noise that comes with statistical modeling but also on a variety of assumptions on how researchers assume a disease is transmitted through a population \citep{diekmann2009,brown2016}.
 
@@
 
-\subsection{Smoothed Incidence to Prevalence Ratio (SIPR)}
-We use the same method as above, IPR, but first smooth the epidemic curves $\left(\hat{X}(t), \hat{Y}(t) \right)$ (we use B-splines with $K=4$ degrees of freedom).
-Then  $J(t) \approx -(\hat{X}(t+1) - \hat{X}(t))$, and the $\hat{\textnormal{IPR}}(t) = \frac{J(t)}{\hat{Y}(t)}$.
-Then
-\begin{align*}
-\rr &= \hat{\textnormal{IPR}}(t) \cdot \frac{1}{\gamma}
-\end{align*}
-We use splines to smooth the curves because they  have proven to be good approximations of epidemic curves \citep{brooks2015}.
-However, the researcher can use their method of choice to estimate $\hat{X}(t)$ and $\hat{Y}(t)$.
-An advantage of this method is that it creates a less variable estimate than estimating IPR using only one point.
-It has the same disadvantages as the usual IPR ratio in that it requires knowledge about $\gamma$.
-
 %%%%%%%%%%%%%%%%%%%%%
@@
 
+%%%%%%%%%%%%55
+\subsection{Likelihood Based Estimation (LBE)}\label{sec:likelihood-based}
+A common way used to estimate epidemiological quantities is through likelihood based estimation.
+We saw an example of this in the previous section with MC estimation.
+However, as we will see in the results, this particular model does not seem to fit the data well.
+As a result, many researchers specify different models for the generation of SIR data; estimate the likelihood through direct calculation, a method such as iterated filtering used in ``plug-and-play'' and derive point estimates and CIs from this likelihood estimate  \citep{abbey1952,forsberg2008,king2009,bhadra2012}.
+%https://kingaa.github.io/sbied/stochsim/stochsim.html
+
+Here, we examine one model that has been used to model the (discrete) SIR: A binomial chain with probabilities based on the deterministic SIR model.
+
+The discrete XYZ model is as follows, where we have replaced $d$ with $\Delta$,
+\begin{align*}
+\systeme{\frac{\Delta X}{\Delta t} = -\frac{\beta Y X}{N}, \frac{\Delta Y}{\Delta t} = \frac{\beta Y X}{N} - \gamma Y, \frac{\Delta Z}{\Delta t} = \gamma Y}. 
+\end{align*}
+
+Denote $\hat{X}(t), \hat{Y}(t)$, and $\hat{Z}(t)$ to be the observed number in each compartment and $X(t), Y(t)$, and $Z(t)$ to be the true, underlying model at time $t$.
+The stochastic model is given by
+\begin{align*}
+  \hat{X}(t+1) &= \hat{X}(t) - x_t\\
+  \hat{Y}(t+1) & = N - \hat{X}(t+1) - \hat{Z}(t+1), \nonumber \\
+  \hat{Z}(t+1) &= \hat{Z}(t) + z_t \nonumber
+\end{align*}
+with $\hat{X}(0)$, $\hat{Y}(0)$, and $\hat{Z}(0)$ known. The random terms in Equation \eqref{eq:sir-noise} are random variables.  Here we choose them to have the following distributions,
+\begin{align*}
+  x_{t} &\sim \text{Binomial}\left (\hat{X}(t), \frac{\beta Y(t)}{N} \right ) \\
+  z_{t} &\sim \text{Binomial}\left (\hat{Y}(t), \gamma \right ). \nonumber
+\end{align*}
+Then the likelihood is
+\begin{align*}
+  \mathcal{L}\left (\beta, \gamma ; \textnormal{Data} \right ) &=
+                                                  \prod_{t=0}^{T-1} \left [\left ( \begin{array}{c}
+                                                            \hat{X}(t) \\
+                                                            x_t
+                                                            \end{array}
+  \right ) \left (\beta - \frac{\beta\left (X(t) +Z(t) \right )}{N} \right )^{x_t} \left (1 - \beta +  \frac{\beta\left (X(t) + Z(t) \right )}{N} \right )^{\hat{X}(t) - x_t} \cdot \right .\\
+  & \left . \left ( \begin{array}{c}
+              N -\left ( \hat{X}(t) + \hat{Z}(t)\right ) \\
+              z_t
+            \end{array}
+  \right ) \left (\gamma \right )^{z_t} \left ( 1 - \gamma\right )^{N -\left ( \hat{X}(t) + \hat{Z}(t)\right ) - z_t} \right ].
+\end{align*}
+The estimate for $\rr$ is then
+\begin{align}\label{eq:lbe-r0}
+  (\hat{\beta}, \hat{\gamma}) &= \arg \max_{\beta, \gamma} \mathcal{L}\left ( \beta, \gamma; \textnormal{Data} \right )\nonumber\\
+  \rr &= \frac{\hat{\beta}}{\hat{\gamma}}
+  \end{align}
+  The two main differences between this model and MC are  that 1) LBE looks at the joint likelihood of the XYZ model, whereas MC only explicitly models the number of susceptibles and 2) the difference in the structure of the transition probability.
+We use the delta method in conjunction with $\hat{\beta}$ and $\hat{\gamma}$ estimations from the Nelder-Mead optimization of the likelihood to estimate the stantard error of $\hat{\rr}$.
+  The advantages of using this model are that our modelling assumptions are explicit and that the modelling process seeems to be reasonable.
+Disadvantages include maximizing the liklihood may be more time consuming than other methods due to the need to approximate $X(t)$ and $Z(t)$ for many different values of $\beta$ and $\gamma$ and that if we are wrong about the generating process of the XYZ curves, then this model may be a very poor fit.
+
 \subsection{Sequential Bayes (SB)}\label{sec:seqbayes}
 
-Described by \cite{bettencourt2008} and summarized in \cite{obadia2012r0}, the sequential Bayes method is a Bayesian approach to an approximation of the classic SIR model.
+Described by \cite{bettencourt2008} and summarized in \cite{obadia2012r0}, the sequential Bayes method is a Bayesian approach to an approximation of the classic SYR model.
 The approximated SIR model assumes that the incidence at $t+1$, $J(t+1)$ has a Poisson distribution, with $\gamma$ as the  average inverse of the infectious period.
@@
 We adjust the recovered compartment $Z$ so the total population is constant. 
-We simulate with two different types of error distributions, Gaussian (Norm) and autoregressive (AR),
-
-\noindent \textbf{Gaussian (Norm) ($\hat{X}_{G}$, $\hat{Y}_{G}$, $\hat{Z}_{G}$)}
+We simulate with two different types of error distributions, Gaussian (G) and autoregressive (AR),
+
+\noindent \textbf{Gaussian (G) ($\hat{X}_{G}$, $\hat{Y}_{G}$, $\hat{Z}_{G}$)}
 \begin{align*}
@@
 For each of the Gaussian and AR datasets, we form an additional dataset which enforces the monotonicity of X and Z via the Pool Adjacent Violators Algorithm (PAVA), described by \cite{friedman1984}).
-Here, we adjust $Y$ so the constant population is maintained.
-
-\noindent \textbf{Gaussian-PAVA (Norm-M) ($\hat{X}_{GP}$, $\hat{Y}_{GP}$, $\hat{Z}_{GP}$)}
+Here, we adjust $Y$ so the constant population is maintained, since X and Z are both monotone functions.
+
+\noindent \textbf{Gaussian-Monotone-PAVA (G-M) ($\hat{X}_{GP}$, $\hat{Y}_{GP}$, $\hat{Z}_{GP}$)}
 \begin{align*}
@@
 \end{align*}
-\textbf{Autoregressive-PAVA (AR-M) ($\hat{X}_{ARP}$, $\hat{Y}_{ARP}$, $\hat{Z}_{ARP}$)}
+\textbf{Autoregressive-Monotone-PAVA (AR-M) ($\hat{X}_{ARP}$, $\hat{Y}_{ARP}$, $\hat{Z}_{ARP}$)}
 \begin{align*}
@@
 \end{align*}
-This gives us a total of four error structures for each set of conditions  (1) Gaussian (Norm), (2) Gaussian monotone (Norm-M),  (3) autoregressive (AR), and (4) autoregressive monotone (AR-M) for a given set of parameters $\left (\beta, \gamma, T, X(0), Y(0), Z(0), \sigma_X, \sigma_Y, N \right )$, the infection rate; recovery rate; total number of time points; initial conditions for number of susceptibles, infectious, and recovered individuals; the standard deviations for both the susceptible and infectious compartments; and the total population, respectively.
+This gives us a total of four error structures for each set of conditions  (1) Gaussian (G), (2) Gaussian monotone - PAVA (G-M),  (3) autoregressive (AR), and (4) autoregressive monotone - PAVA (AR-M) for a given set of parameters $\left (\beta, \gamma, T, X(0), Y(0), Z(0), \sigma_X, \sigma_Y, N \right )$, the infection rate; recovery rate; total number of time points; initial conditions for number of susceptibles, infectious, and recovered individuals; the standard deviations for both the susceptible and infectious compartments; and the total population, respectively.
 
@@
       \end{center}
-      The baseline data set is generated by adding noise to the susceptible and infectious curves from  deterministic SIR, shown in Figure \ref{fig:baseline-data}.
+      The baseline data set is generated by adding noise to the susceptible and infectious curves from deterministic SIR paths, which are shown in Figure \ref{fig:baseline-data}.
 The recovered compartment is then adjusted to maintain a constant population value.
@@
         \includegraphics[width=.8\textwidth]{images/baseline_sir.pdf}
-        \caption{SIR (XYZ) curves for the deterministic model with $N=100,000$, $(\beta, \gamma) = (0.06, 0.03)$, $(X(0), Y(0), Z(0))= (99,950, 50,0)$  The baseline data is generated by using this model and adding noise to the $X$ and $Y$ compartments and adjusting $Z$ at each time step to maintain a constant population.}\label{fig:baseline-data}
+        \caption{SIR (XYZ) curves for the deterministic model with $N=100,000$, $(\beta, \gamma) = (0.06, 0.03)$, $(X(0), Y(0), Z(0))= (99950, 50,0)$  The baseline data is generated by using this model and adding noise to the $X$ and $Y$ compartments and adjusting $Z$ at each time step to maintain a constant population.}\label{fig:baseline-data}
         \end{figure}
@@
 Initially, 50 (0.05\%) infectious individuals are introduced to the population of 100,000 individuals (approximately the average population size of a US county in 2015).
-The standard deviation of the $\hat{X}$ values is 0.1\% of the magnitude of the initial number of susceptible individuals, and the standard deviation of the $\hat{Y}$ values is 10\% of the magnitude of the initial number of infectious individuals, although in this setting the number of infectious individuals increases rapidly.
+The standard deviation of the $\hat{X}$ values is 0.1\% of the magnitude of the initial number of susceptible individuals, and the standard deviation of the $\hat{Y}$ values is 10\% of the magnitude of the initial number of infectious individuals, although in this setting, the number of infectious individuals increases rapidly.
 We use Gaussian error as our baseline error because of its well known statistical properties so we can more easily compare methods to one another.
@@
 \textbf{Name} & \textbf{Model} & \textbf{Error} & \textbf{($\beta, \gamma$)} & \textbf{T} & \textbf{(X(0),Y(0))} & \textbf{N} & \textbf{($\sigma_X, \sigma_Y$)} \\ \midrule
-  Baseline        & SIR      &   Norm             & (.06, .03)                          &     365       &  (99950, 50)                    &  $10^5$          &(100, 5)                                \\ \midrule
+  Baseline        & SIR      &   G             & (.06, .03)                          &     365       &  (99950, 50)                    &  $10^5$          &(100, 5)                                \\ \midrule
 %  AR       & SIR      &   AR ($\rho=.7)$             & (.06, .03)                          &     365       &  (99950, 50)                    &  $10^5$          &(100, 5)                            \\ \midrule
-%  \textcolor{red}{Norm-M}      & SIR      &   Norm-M ($\rho=.7)$             & (.06, .03)                          &     365       &  (99950, 50)                    &  $10^5$          &(100, 5)                            \\ \midrule
+%  \textcolor{red}{G-M}      & SIR      &   G-M ($\rho=.7)$             & (.06, .03)                          &     365       &  (99950, 50)                    &  $10^5$          &(100, 5)                            \\ \midrule
 %   \textcolor{red}{AR-M}       & SIR      &   AR-M ($\rho=.7)$             & (.06, .03)                          &     365       &  (99950, 50)                    &  $10^5$          &(100, 5)                            \\ \midrule
-  InfRec1       & SIR      &   Norm             & (60000, .001)                          &     365       &  (99950, 50)                    &  $10^5$          &(100, 5)\\
-  InfRec2      & SIR      &   Norm             & (1500, .001)                          &     365       &  (99950, 50)                    &  $10^5$          &(100, 5)                                \\
-  InfRec3      & SIR      &   Norm             & (1000, .001)                          &     365       &  (99950, 50)                    &  $10^5$          &(100, 5)                                \\
-InfRec4      & SIR      &   Norm             & (06, .24)                          &     365       &  (99950, 50)                    &  $10^5$          &(100, 5)                                \\ \midrule
-  Time1 & SIR &  Norm   & (.06, .03)                          &     200       &  (99950, 50)                    &  $10^5$          &(100, 5)                                \\
-  Time2 & SIR &  Norm   & (.06, .03)                          &     100       &  (99950, 50)                    &  $10^5$          &(100, 5)                                \\
-  Time3 & SIR &  Norm   & (.06, .03)                          &     50       &  (99950, 50)                    &  $10^5$          &(100, 5)                                \\
-  Time4 & SIR &  Norm   & (.06, .03)                          &     20       &  (99950, 50)                    &  $10^5$          &(100, 5)                                \\ \midrule
-  Initial1    & SIR      &   Norm             & (.06, .03)                          &     365       &  (99999, 1)                    &  $10^5$          &(100, 5)                                \\
-  Initial2    & SIR      &   Norm             & (.06, .03)                          &     365       &  (99990, 10)                    &  $10^5$          &(100, 5)                                \\
-  Initial3    & SIR      &   Norm             & (.06, .03)                          &     365       &  (99900, 100)                    &  $10^5$          &(100, 5)                                \\
-  Initial4    & SIR      &   Norm             & (.06, .03)                          &     365       &  (99000, 1000)                    &  $10^5$          &(100, 5)                                \\ \midrule
-PopSize1      & SIR      &   Norm             & (.06, .03)                          &     365       &  (999500, 500)                    &  $10^6$          &(100, 5)                                \\
-PopSize2      & SIR      &   Norm             & (.06, .03)                          &     365       &  (9995, 5)                    &  $10^4$          &(100, 5)                                \\
-PopSize3        & SIR      &   Norm             & (.06, .03)                          &     365       &  (990, 10)                    &  $10^3$          &(100, 5)                                \\
-PopSize4       & SIR      &   Norm             & (.06, .03)                          &     365       &  (99, 1)                    &  $10^2$          &(100, 5)                                \\
+  InfRec1       & SIR      &   G             & (60000, .001)                          &     365       &  (99950, 50)                    &  $10^5$          &(100, 5)\\
+  InfRec2      & SIR      &   G             & (1500, .001)                          &     365       &  (99950, 50)                    &  $10^5$          &(100, 5)                                \\
+  InfRec3      & SIR      &   G             & (1000, .001)                          &     365       &  (99950, 50)                    &  $10^5$          &(100, 5)                                \\
+InfRec4      & SIR      &   G             & (06, .24)                          &     365       &  (99950, 50)                    &  $10^5$          &(100, 5)                                \\ \midrule
+  Time1 & SIR &  G   & (.06, .03)                          &     200       &  (99950, 50)                    &  $10^5$          &(100, 5)                                \\
+  Time2 & SIR &  G   & (.06, .03)                          &     100       &  (99950, 50)                    &  $10^5$          &(100, 5)                                \\
+  Time3 & SIR &  G   & (.06, .03)                          &     50       &  (99950, 50)                    &  $10^5$          &(100, 5)                                \\
+  Time4 & SIR &  G   & (.06, .03)                          &     20       &  (99950, 50)                    &  $10^5$          &(100, 5)                                \\ \midrule
+  Initial1    & SIR      &   G             & (.06, .03)                          &     365       &  (99999, 1)                    &  $10^5$          &(100, 5)                                \\
+  Initial2    & SIR      &   G             & (.06, .03)                          &     365       &  (99990, 10)                    &  $10^5$          &(100, 5)                                \\
+  Initial3    & SIR      &   G             & (.06, .03)                          &     365       &  (99900, 100)                    &  $10^5$          &(100, 5)                                \\
+  Initial4    & SIR      &   G             & (.06, .03)                          &     365       &  (99000, 1000)                    &  $10^5$          &(100, 5)                                \\ \midrule
+PopSize1      & SIR      &   G             & (.06, .03)                          &     365       &  (999500, 500)                    &  $10^6$          &(100, 5)                                \\
+PopSize2      & SIR      &   G             & (.06, .03)                          &     365       &  (9995, 5)                    &  $10^4$          &(100, 5)                                \\
+PopSize3        & SIR      &   G             & (.06, .03)                          &     365       &  (990, 10)                    &  $10^3$          &(100, 5)                                \\
+PopSize4       & SIR      &   G             & (.06, .03)                          &     365       &  (99, 1)                    &  $10^2$          &(100, 5)                                \\
  \midrule
-  Var1      & SIR      &   Norm             & (.06, .03)                          &     365       &  (99950, 50)                    &  $10^5$          &(50, 2)                                \\
-  Var2      & SIR      &   Norm             & (.06, .03)                          &     365       &  (99950, 50)                    &  $10^5$          &(500, 20)                                \\
-  Var3      & SIR      &   Norm             & (.06, .03)                          &     365       &  (99950, 50)                    &  $10^5$          &(2500, 1000)                                \\
-  Var4      & SIR      &   Norm             & (.06, .03)                          &     365       &  (99950, 50)                    &  $10^5$          &(10000, 500)                                \\ \midrule
- Poly1     & Linear      &   Norm             & (.06, .03)                          &     365       &  (99950, 50)                    &  $10^5$          &(100, 5)                            \\
-Poly2    & Quartic      &   Norm             & (.06, .03)                          &     365       &  (99950, 50)                    &  $10^5$          &(100, 5)                            \\ 
-   LSIR     & LSIR      &   Norm             & (.06, .03)                          &     365       &  (99950, 50)                    &  $10^5$          &(100, 5)                            \\ 
+  Var1      & SIR      &   G             & (.06, .03)                          &     365       &  (99950, 50)                    &  $10^5$          &(50, 2)                                \\
+  Var2      & SIR      &   G             & (.06, .03)                          &     365       &  (99950, 50)                    &  $10^5$          &(500, 20)                                \\
+  Var3      & SIR      &   G             & (.06, .03)                          &     365       &  (99950, 50)                    &  $10^5$          &(2500, 1000)                                \\
+  Var4      & SIR      &   G             & (.06, .03)                          &     365       &  (99950, 50)                    &  $10^5$          &(10000, 500)                                \\ \midrule
+ Poly1     & Linear      &   G             & (.06, .03)                          &     365       &  (99950, 50)                    &  $10^5$          &(100, 5)                            \\
+Poly2    & Quartic      &   G             & (.06, .03)                          &     365       &  (99950, 50)                    &  $10^5$          &(100, 5)                            \\ 
+   LSIR     & LSIR      &   G             & (.06, .03)                          &     365       &  (99950, 50)                    &  $10^5$          &(100, 5)                            \\ 
   \bottomrule
 \end{tabular}
-\caption{Table of simulated data sets for the ``Norm'' error.
-We also repeat the tables with the other three error structures: AR, Norm-M, and AR-M.
+\caption{Table of simulated data sets for the ``G'' error.
+We also repeat the tables with the other three error structures: AR, G-M, and AR-M.
 We simulate $4 \times 24=96$ total data sets with the given parameters.}
@@
 Table \ref{tab:se-methods} shows the ways that standard errors are acquired for the different methods.
-The IPR, and SIPRknowledge of the length of the average infectious period $\gamma^{-1}$.
+The method IPR requires knowledge of the length of the average infectious period $\gamma^{-1}$ in order to fit the model.
 For the results below, we assume the best case scenario that the true average infectious period is known for the data set.
@@
 		Incidence-Prevalence Ratio (IPR) & Block bootstrap (Jackknife resampling) \\
-		Smoothed IPR (SIPR) & Block bootstrap (Jackknife resampling) \\
 		Log-Linear (LL) & Delta (linear regression SE) \\
           Markov Chain (MC) & Delta (GLM regression SE) \\
+          Likelihood Based Estimation (LBE) & Delta (with inverse Hessian)\\
 		Sequential Bayes (SB) & Posterior\\
@@
 This leads us to believe that these models are inadequate in capturing the effects of a disease following the SIR model.
-Both models rely on parametric assumptions of simple underlying statistical distributions for the course of infection (binomial and Poisson, respectively) with time-invariant transmission parameters ($\alpha$ and $\lambda$).
+Both models rely on parametric assumptions of simple underlying statistical distributions for the course of infection (Binomial and Poisson, respectively) with time-invariant transmission parameters ($\alpha$ and $\lambda$).
 These are good examples of why we need to consider other reasons for using a model other than the math resulting in a simple estimate.
@@
 For example, in Figure \ref{fig:ar-r0}, we see that when the true value of $\rr$ is 1, rRE has a much smaller CI than RE, which may be one reason to prefer rRE to RE.
+In this figure, we see when $\gamma << \beta$ (e.g.
+$\rr=60)$, then all estimates for $\hat{\rr}$ fail.
+However, when $\beta >.
+\gamma$, we the estimates (besides IPR) can accurately obtain the true estimate.
 This figure also helps to show that RE and rRE are robust to extreme values of $\beta$ and $\gamma$ and hence extreme values of $\rr$.
+\com{(For R0=1.5, why do RE, rRE, and LL look like 2 and not 1.5?))}
 
@@
 Reassuringly, we find that $\hat{\rr}$ estimates from RE and rRE seem to be generally robust to the error structure.
-((What's up with varying $\rr$ values??))
+\com{((What's up with varying $\rr$ values??))}
 
@@
 
-%% model misspecification is fairly robust
+%% model misspecification is terribly unrobust
 We also examine circumstances where the data is generated from underlying models different from the SIR model.
 The results are plotted in Figure \ref{fig:mod-rob}  All three sets of data sets generated by the non-SIR models result in poor estimates of $\rr$, meaning that even our best estimation techniques fail under model misspecification.
+Moreover, these methods seem to underestimate $\rr$ \com{Illegible plot, Does RR=0?}
 
@@
 We find the parameter $T$ to be particularly important, as shown in Figure \ref{fig:timediff}.
-For one, it can play a role similar to that of $N$, directly giving the researcher more data for a larger $T$.
+For one, it can play a role similar to that of $N$, as the number of time points increases, the amount of data may also increase.
 Thus, larger $T$ may result in smaller CIs due to the CLT.
@@
 This is likely due to the previously discussed relationship between $\rr$ and the final size of the epidemic.
-(ADD IN)
-
-However, if we are given only 200 days worth of data, we find that the collection period size does result in different estimates.  PLOT
+
+However, if we are given only 200 days worth of data, we find that the collection period size does result in different estimates.
 
@@
 The data source is ILINet and is the national level data.
-We use the same date range as in \cite{towers2009}, namely Epiweeks 21-33 (May 23, 2009-August 22, 2009).
+We use the same date range as in \cite{towers2009}, namely \com{Epiweeks 21-33} (May 23, 2009-August 22, 2009).
 From the United States Census Bureau, we use the population estimate of the USA as of April 1, 2010 of $N=308,740,000$ individuals \citep{census-2010}.
@@
 
+\com{Table and graph do not agree}
 \begin{table}[H]
@@
 	\centering
-	\begin{tabular}[t]{l|r|r}
+	\begin{tabular}{l|r|r}
 		\hline
@@
 		\hline
-		RE & 0.0625 & 0.0134 \\
-		\hline
-		rRE & 0.0278 & 0.0220 \\
-		\hline
-		LMA & 0.9857 & 0.3465 \\
-		\hline
-		LMAT & 0.9701 & 0.0549 \\
-		\hline
-		IPR & 5.5463 & 1.7744 \\
-		\hline
-		SIPR & -3.1438 & 4.1129 \\
-		\hline
-		LL & 0.0040 & $<$ 1e-04\\
-		\hline
-		MC & 0.9462 & 0.0008 \\
-		\hline
-		SB & 29.9850 & 1.2907 \\
+		RE & 1.1851 & 0.3540\\
+		\hline
+		rRE & 1.1851 & 0.0212\\
+		\hline
+		LMA & 4.3811 & 7.4772\\
+		\hline
+		LMAT & 1.2169 & 0.7584\\
+		\hline
+		IPR & 82.8540 & 24.5098\\
+		\hline
+		SIPR & 75.1963 & 16.2124\\
+		\hline
+		LL & 1.2128 & 0.0170\\
+		\hline
+		MC & 0.9486 & $<$ 1e-04 \\
+		\hline
+		SB & 2.0596 & 0.3292\\
 		\hline
@@
 \end{table}
+
+\com{Results TBD}
 
@@
   \item The noisier the epidemic data is, the larger our CIs for $\rr$ are
-  \item Estimations of $\rr$ are fairly robust to the misspecified data sets we presented, at least for the methods of RE, rRE, and LL
+  \item Estimations of $\rr$ are not robust to the misspecified data sets we presented.  Therefore, model specification is of grave importance
   \item Block bootstrap matches the empirical variance through repeated sampling.

### 2018-08-23 ModFolder/likelihood_shell.R

@@ -17,6 +17,7 @@ sir_likelihood <- function(start_pop, sus_data, inf_data,
     data <- data.frame(tt = times, X1 = sus_data,
                        X2 = inf_data,
                        X3 = N - sus_data - inf_data)
+    data$ll <- 1
     do_plug_in <- FALSE # Run the SIR CM each time as opposed to using observed values
     disease_list <- list(params = starting_params,
                          params_names = c("beta", "gamma"),
@@ -200,9 +201,9 @@ get_SIR_diffs <- function(data){
     ## Return X1, X2, X3 diffs and previous values
     sub_df <- plyr::ddply(.data = data, .variables = c("ll"),
                           .fun =  function(x){
-                              obs_X1 <- c(lag(x$X1, 1))
-                              obs_X2 <- c(lag(x$X2, 1))
-                              obs_X3 <- c(lag(x$X3, 1))
+                              obs_X1 <- c(dplyr::lag(x$X1, 1))
+                              obs_X2 <- c(dplyr::lag(x$X2, 1))
+                              obs_X3 <- c(dplyr::lag(x$X3, 1))
                               return(data.frame(obs_X1= obs_X1,
                                                 obs_X2 = obs_X2,
                                                 obs_X3 = obs_X3))

### 2018-08-31 ModFolder/LBE_run.R
Updated supplementary materials

@@ -1,0 +1,98 @@
+this.dir <- dirname(parent.frame(2)$ofile)
+setwd(this.dir)
+
+source("./likelihood_shell.R")
+
+data_path <- "C:/Users/Andersen/Documents/R0/DataSim/PaperData/Other"
+
+files_list <- paste0(data_path, "/", 
+                     list.files(data_path, recursive = TRUE))
+
+results_matlist_exp <- list()
+output_lists <- list()
+
+xx <- files_list
+
+### Test dataset
+for (jj in 1:length(xx)) {
+  # for (jj in 5:20) {
+  print(xx[jj])
+  # r0_true <- as.numeric(strsplit(xx[[jj]], split = "_")[[1]][12]) / 
+  #   as.numeric(strsplit(xx[[jj]], split = "_")[[1]][13])
+  # beta_true <- as.numeric(strsplit(xx[[jj]], split = "_")[[1]][12])
+  # gamma_true <- as.numeric(strsplit(xx[[jj]], split = "_")[[1]][11])
+  gamma_true <- 0.03
+  # output_list <- list()
+  # sink(zz, append = TRUE)
+  # sink(zz, type = "message", append = TRUE)
+  # print("------------------------------------------")
+  # print("------------------------------------------")
+  # print("------------------------------------------")
+  # print("------------------------------------------")
+  # print("------------------------------------------")
+  # print((paste0("Beta = ", beta_true, ", Gamma = ", gamma_true, ", R0 = ", r0_true)))
+  # print("")
+  # print("")
+  dat <- read.csv(xx[jj])
+  # start_pop <- c("S" = as.numeric(strsplit(strsplit(xx[jj], "/")[[1]][10], "_")[[1]][1]),
+  #                "I" = as.numeric(strsplit(strsplit(xx[jj], "/")[[1]][10], "_")[[1]][2]), "R" = 0)
+  # print(start_pop)
+
+  start_pop <- c("S" = 99950,
+                 "I" = 50, "R" = 0)
+  start_par_sir <- c("beta" = 0.02, "gamma" = 0.08)
+  start_par_rep <- c("r0" = 2, "gamma" = 0.007)
+  start_par_seir <- c("beta" = 0.02, "gamma" = 0.07, "alpha" = 0.014)
+  
+  # Surv func
[…]

### 2018-08-31 ModFolder/likelihood_shell.R

@@ -1,3 +1,5 @@
+require(plyr)
+require(dplyr)
 
 # Inputs:
 ### start_pop is the starting estimate for the number of people in each compartment
@@ -27,16 +29,17 @@ sir_likelihood <- function(start_pop, sus_data, inf_data,
     optim_pars <- optim(par = starting_params,
                            fn = loglike_sir, data = data,
                            disease_list = disease_list,
-                           do_plug_in = do_plug_in)
-    r0_est <- optim_pars$par[2] / optim_pars$par[2]
-
+                           do_plug_in = do_plug_in,
+                        hessian = TRUE)
+    
+    r0_est <- optim_pars$par[1] / optim_pars$par[2]
+    
     ## Using delta method and hessian
-    r0_hessian <- model_est$hessian / model_est$value
-    dh <- c(1/model_est$par["gamma"], - model_est$par["beta"] / (1/model_est$par["gamma"]^2))
+    r0_hessian <- optim_pars$hessian / optim_pars$value
+    dh <- c(1/optim_pars$par[2], - optim_pars$par[1] / (1/optim_pars$par[2]^2))
     r0_sd <- sqrt(t(dh) %*% solve(r0_hessian) %*% dh)
                          
-    
-  return(list(est = r0, sd = r0_sd))
+  return(list(est = r0_est, sd = r0_sd, output = optim_pars$par))
 }
 
 
@@ -64,7 +67,7 @@ loglike_sir <- function(params,
 
     N <- sum(disease_list$init_vals)
     beta <- params[1]
-    gamma <- params[2]
+    gamma <-params[2]
 
     new_data <- get_SIR_diffs(data)
     new_data <- get_SIR_lags(params, new_data, disease_list, do_plug_in)
@@ -76,11 +79,11 @@ loglike_sir <- function(params,
     s_loglike <- sum(apply(new_data, 1,
                            function(row){
                                prob <- beta * row["lag_X2"] / N
-                               if(prob == 0){
+                               if(prob < 0 | prob > 1){
                                   return(-10e3)
                                }
-                               like <- dbinom(row["diff_X1"],
-                                          size = row["obs_X1"],
+                               like <- dbinom(round(row["diff_X1"]),
+                                          size = pmax(round(row["obs_X1"]), 0),
                                           prob = prob)
                                if(like > 0) return(log(like))
                                return(-10e3)
@@ -88,9 +91,12 @@ loglike_sir <- function(params,
[…]

### 2018-09-04 paper/Version16/draft_v16.tex
Added changes to version 16

@@
   \centering
-  \includegraphics[scale=0.5]{images/BaseBase.jpeg}
+  \includegraphics[scale=0.5]{images/BaseBase.tiff}
   \caption{Forest plots of estimates of $\rr$ for the \xxsir methods for the baseline data set $(\beta=.06, \gamma=.03, O=365, X(0)=99950, Y(0)=50, Z(0)=0, \sigma_X=100, \sigma_Y=5, N=10^5)$.
@@
 		\hline
-		RE & 1.9999711 & 0.0056\\
-		\hline
-		rRE & 1.9997340 & 0.0050\\
-		\hline
-		LMA & 1.6863640 & 0.1886\\
-		\hline
-		LMAT & 1.4063227 & 0.6309\\
-		\hline
-		IPR & 4.4203632 & 12.3593\\
-		\hline
-		SIPR & 1.8870824 & $<$ 1e-04 \\
-		\hline
-		LL & 1.9998072 & 0.0002\\
-		\hline
-		MC & 0.9485555 &  $<$ 1e-04 \\
-		\hline
-		SB & 1.6495023 & 0.0672\\
+		RE & 2.0000 & 0.0056\\
+		\hline
+		rRE & 1.9997 & 0.0050\\
+		\hline
+		LMA & 1.6864 & 0.1886\\
+		\hline
+		LMAT & 1.4063 & 0.6309\\
+		\hline
+		IPR & 4.4204 & 12.3593\\
+		\hline
+		SIPR & 1.8871 & $<$ 1e-04 \\
+		\hline
+		LL & 1.9998 & 0.0002\\
+		\hline
+		MC & 0.9486 &  $<$ 1e-04 \\
+		\hline
+		SB & 1.6495 & 0.0672\\
+		\hline
+		LBE & 14.6784 & 1.2330\\
 		\hline
@@
 	\centering
-	\includegraphics[scale=0.5]{images/start_arm.jpeg}
+	\includegraphics[scale=0.5]{images/start_arm.tiff}
 	\caption{Forest plots of estimates of $\rr$ for the \xxsir methods for the Inits1-4 data sets $(\beta=.06, \gamma=.03, T=365,  Z(0)=0, \sigma_X=100, \sigma_Y=5, N=10^5)$.
@@
 	\centering
-	\includegraphics[scale=0.5]{images/parchange_ar.jpeg}
+	\includegraphics[scale=0.5]{images/parchange_ar.tiff}
 	\caption{Forest plots of estimates of $\rr$ for the \xxsir methods for the infrec1-4 data sets $(T=365, X(0)=99950, Y(0)=50, Z(0)=0, \sigma_X=100, \sigma_Y=5, N=10^5)$.
@@
 	\centering
-	\includegraphics[scale=0.5]{images/timediff.jpeg}
-	\caption{Forest plots of estimates of $\rr$ for the \xxsir methods for the baseline data sets with normall error structure $(\beta=.06, \gamma=.03, X(0) = 99950, Y(0)=50, \sigma_X=100, \sigma_Y=5, N=10^5)$.
+	\includegraphics[scale=0.5]{images/timediff.tiff}
+	\caption{Forest plots of estimates of $\rr$ for the \xxsir methods for the baseline data sets with normal error structure $(\beta=.06, \gamma=.03, X(0) = 99950, Y(0)=50, \sigma_X=100, \sigma_Y=5, N=10^5)$.
 The points are the estimates of $\rr$ and the lines denote $\pm 2\cdot $Std.
@@
 	\centering
-	\includegraphics[scale=0.5]{images/obs.jpeg}
-	\caption{Forest plots of estimates of $\rr$ for the \xxsir methods for the baseline data sets with normall error structure $(\beta=.06, \gamma=.03, X(0) = 99950, Y(0)=50, \sigma_X=100, \sigma_Y=5, N=10^5)$.
+	\includegraphics[scale=0.5]{images/obs.tiff}
+	\caption{Forest plots of estimates of $\rr$ for the \xxsir methods for the baseline data sets with normal error structure $(\beta=.06, \gamma=.03, X(0) = 99950, Y(0)=50, \sigma_X=100, \sigma_Y=5, N=10^5)$.
 The points are the estimates of $\rr$ and the lines denote $\pm 2\cdot $Std.
@@
 		\hline
+		LBE & 8.5926 & 6.9421 & 14.6784 & 1.2330\\
+		\hline
 	\end{tabular}
@@
 		\hline
+		LBE & 19.9935 & 10.640 & 14.6784 & 1.2330 \\
+		\hline
 	\end{tabular}
@@
 	\begin{center}
-		\includegraphics[scale=0.5]{images/coverage2.jpeg}
+		\includegraphics[scale=0.5]{images/coverage2.tiff}
 		\caption{Proportion of coverage of true $\rr$ of confidence intervals produced by repeated sampling for the applicable method (delta method or posterior).}
@@
 	\begin{center}
-		\includegraphics[scale=0.5]{images/coverage.jpeg}
+		\includegraphics[scale=0.5]{images/coverage.tiff}
 		\caption{Proportion of coverage of true $\rr$ of confidence intervals produced by repeated sampling for block bootstrapping.}
@@
 	\centering
-	\includegraphics[width=\textwidth]{images/h1n1.jpeg}
+	\includegraphics[width=\textwidth]{images/h1n1.tiff}
 	\caption{Results from SIR models on H1N1 data.}

### 2018-09-04 ModFolder/LBE_run.R

@@ -3,10 +3,10 @@ setwd(this.dir)
 
 source("./likelihood_shell.R")
 
-data_path <- "C:/Users/Andersen/Documents/R0/DataSim/PaperData/Other"
+data_path <- "C:/Users/Andersen/Documents/R0/DataSim/PaperData/Baseline"
 
 files_list <- paste0(data_path, "/", 
-                     list.files(data_path, recursive = TRUE))
+                     list.files(data_path, recursive = TRUE))[7]
 
 results_matlist_exp <- list()
 output_lists <- list()

### 2018-09-04 ModFolder/likelihood_shell.R

@@ -36,6 +36,7 @@ sir_likelihood <- function(start_pop, sus_data, inf_data,
     
     ## Using delta method and hessian
     r0_hessian <- optim_pars$hessian / optim_pars$value
+    print(r0_hessian)
     dh <- c(1/optim_pars$par[2], - optim_pars$par[1] / (1/optim_pars$par[2]^2))
     r0_sd <- sqrt(t(dh) %*% solve(r0_hessian) %*% dh)
                          
@@ -68,13 +69,11 @@ loglike_sir <- function(params,
     N <- sum(disease_list$init_vals)
     beta <- params[1]
     gamma <-params[2]
-
     new_data <- get_SIR_diffs(data)
     new_data <- get_SIR_lags(params, new_data, disease_list, do_plug_in)
    
 
     new_data <- na.omit(new_data)
-
     
     s_loglike <- sum(apply(new_data, 1,
                            function(row){

### 2018-09-06 paper/Version16/draft_v16.tex
Changed images to jpg
pulling
Added new explanations and comments

@@
 \usepackage{url}
-
 %% Make red text
@@
 \setlength{\itemsep}{0pt}
+
+%% tif -> png
+\usepackage{epstopdf}
+\epstopdfDeclareGraphicsRule{.tiff}{png}{.png}{convert #1 \OutputFile}
+\AppendGraphicsExtensions{.tiff}
 
@@
 \end{align*}
-with $\hat{X}(0)$, $\hat{Y}(0)$, and $\hat{Z}(0)$ known. The random terms in Equation \eqref{eq:sir-noise} are random variables.  Here we choose them to have the following distributions,
+with $\hat{X}(0)$, $\hat{Y}(0)$, and $\hat{Z}(0)$ known.
+The terms $x_t$ and $z_t$ in Equation \eqref{eq:sir-noise} incorporate random behavior into the model.
+Here we choose them to have the following distributions,
 \begin{align*}
@@
 \end{align*}
-Then the likelihood is
+Then the likelihood is, when we assuming that the next time step is dependent only on the current time step,
 \begin{align*}
@@
 
-We can also use the Delta method to derive variance estimates of LMA, LMAT, LL, (which in this case is equivalent to the standard error derived from the typical linear regression framework).
+We can also use the Delta method to derive variance estimates of LMA, LMAT, LL, (which in this case is equivalent to the standard error derived from the typical linear regression framework), and LBE.
 We can also use the Delta method to derive an estimate for MC, where $\hat{\sigma}$ is the estimate of the variance of the $\hat{\alpha}$,
@@
         \end{figure}
-      This baseline data set is chosen due to its moderate $\rr$ size of 2, which indicates an outbreak of an epidemic similar to HIV, ZIK-V, EVD, and H1N1 pandemic influenza (see Table \ref{tab:r0-real-ex}).
+      This baseline data set is chosen due to its moderate $\rr$ size of 2, which indicates an outbreak of an epidemic similar to ZIK-V, EVD, and H1N1 pandemic influenza (see Table \ref{tab:r0-real-ex}).
 The infection rate and recovery rate ($\beta, \gamma$) are relatively small.
@@
 There are numerous aspects to consider when using a method to estimate $\rr$.
-Among these include values of both the infection and recovery rate $(\beta, \gamma)$, the number of data points available ($T$), the total size of the population ($N$), the initial percent of susceptible individuals $\left (\frac{X(0)}{N}\right)$, the initial percent of infectious individuals $\left (\frac{Y(0)}{N}\right )$, the magnitude of the variance of both the number of susceptibles and the number of infectious ($\sigma_X^2, \sigma_Y^2$), the error modeling assumptions, data generated from a non-SIR model, and finally methods of estimating the variance.
+Among these include values of both the infection and recovery rate $(\beta, \gamma)$, the number of data points available ($T$), the total size of the population ($N$), the initial percent of susceptible individuals $\left (\frac{X(0)}{N}\right)$, the initial percent of infectious individuals $\left (\frac{Y(0)}{N}\right )$, the magnitude of the variance of both the number of susceptibles and the number of infectious ($\sigma_X^2, \sigma_Y^2$), the error modeling assumptions, data generated non-SIR models, and finally methods of estimating the variance.
 For the sake of brevity, we do not provide the full results here, and instead examine the results at a high-level.
@@
   \centering
-  \includegraphics[scale=0.5]{images/BaseBase.tiff}
+  \includegraphics[scale=0.5]{images/BaseBase.jpg}
   \caption{Forest plots of estimates of $\rr$ for the \xxsir methods for the baseline data set $(\beta=.06, \gamma=.03, O=365, X(0)=99950, Y(0)=50, Z(0)=0, \sigma_X=100, \sigma_Y=5, N=10^5)$.
@@
 	\end{tabular}
-        \caption{Baseline Data $(\beta=.06, \gamma=.03, T=365, X(0)=99950, Y(0)=50, Z(0)=0, \sigma_X=100, \sigma_Y=5, N=10^5)$, $\rr$ Estimates and Standard Errors}\label{tab:baseline-res}
+        \caption{Baseline Data $(\beta=.06, \gamma=.03, T=365, X(0)=99950, Y(0)=50, Z(0)=0, \sigma_X=100, \sigma_Y=5, N=10^5)$, $\rr$ Estimates and Standard Errors.
+\com{Table does not match graph --LBE}}\label{tab:baseline-res}
 \end{table}
@@
 We would say that IPR's CI is uninformative.
-Additionally, MC and SIPR models seem to have unreasonably small standard errors.
+Additionally, MC has a standard error of less than $.0001$, which is very small.
+LBE, which assumes that the X and Z compartments are monotonic, overestimates $\rr$
 
@@
 
+On the other hand, LBE also tends to have larger CIs and does not often capture the true value of $\rr$ when the the X and Z compartments are not monotonic functions.
+This method tends to overestimate the value of $\rr$.
+
 IPR often produces reasonable point estimates for $\rr$ but simply has too large of a CI to be considered a useful estimate.
-On the other hand, SIPR has a small CI that typically underestimates the true value of $\rr$.
-Since SIPR smoothes the SIR curves, the peak of the infectious curve in particular is shrunken.
-Since the peak value of number of infectious is crucial in estimating $\rr$ for SIPR, it makes sense that SIPR underestimates this result.
 Additionally, LMA and LMAT are generally prone to underestimating $\rr$ and have large and generally uninformative CIs.
@@
 	\centering
-	\includegraphics[scale=0.5]{images/start_arm.tiff}
+	\includegraphics[scale=0.5]{images/start_arm.jpg}
 	\caption{Forest plots of estimates of $\rr$ for the \xxsir methods for the Inits1-4 data sets $(\beta=.06, \gamma=.03, T=365,  Z(0)=0, \sigma_X=100, \sigma_Y=5, N=10^5)$.
@@
 $\rr=60)$, then all estimates for $\hat{\rr}$ fail.
-However, when $\beta >.
-\gamma$, we the estimates (besides IPR) can accurately obtain the true estimate.
+However, when $\beta >> \gamma$, we the estimates (besides IPR) can accurately obtain the true estimate.
 This figure also helps to show that RE and rRE are robust to extreme values of $\beta$ and $\gamma$ and hence extreme values of $\rr$.
@@
 	\centering
-	\includegraphics[scale=0.5]{images/parchange_ar.tiff}
+	\includegraphics[scale=0.5]{images/parchange_ar.jpg}
 	\caption{Forest plots of estimates of $\rr$ for the \xxsir methods for the infrec1-4 data sets $(T=365, X(0)=99950, Y(0)=50, Z(0)=0, \sigma_X=100, \sigma_Y=5, N=10^5)$.
@@
 The results are plotted in Figure \ref{fig:mod-rob}  All three sets of data sets generated by the non-SIR models result in poor estimates of $\rr$, meaning that even our best estimation techniques fail under model misspecification.
-Moreover, these methods seem to underestimate $\rr$ \com{Illegible plot, Does RR=0?}
+Moreover, these methods seem to underestimate $\rr$ \com{Illegible axes}
 
@@
 	\centering
-	\includegraphics[scale=0.5]{images/timediff.tiff}
+	\includegraphics[scale=0.5]{images/timediff.jpg}
 	\caption{Forest plots of estimates of $\rr$ for the \xxsir methods for the baseline data sets with normal error structure $(\beta=.06, \gamma=.03, X(0) = 99950, Y(0)=50, \sigma_X=100, \sigma_Y=5, N=10^5)$.
@@
 	\centering
-	\includegraphics[scale=0.5]{images/obs.tiff}
+	\includegraphics[scale=0.5]{images/obs.jpg}
 	\caption{Forest plots of estimates of $\rr$ for the \xxsir methods for the baseline data sets with normal error structure $(\beta=.06, \gamma=.03, X(0) = 99950, Y(0)=50, \sigma_X=100, \sigma_Y=5, N=10^5)$.
@@
 In some cases, such as LL, LMA, and LMAT, the standard deviation and standard error estimates are fairly close.
-The least squares methods, RE and rRE, on the other hand, tend to overestimate the error by an order of magnitude.
-Finally, SIPR and SB, have individual standard error estimates that are smaller than the standard deviation of repeated estimates, meaning that they understate how much error is in the estimates.
+The least squares methods, RE and rRE, on the other hand, tend to overestimate the sampling error by an order of magnitude.
+The methods SB and LBE have individual standard error estimates that are smaller than the standard deviation of repeated estimates, meaning that they understate how much error is in the estimates.
 
@@
 Only 5 of the models provide predictions of the size of the compartments and thus allow for detrending of the data.
-The ratio estimator and reparameterized model have a fairly accurate estimates of $\rr$ as well as very small standard errors.
-On the other hand, the linear approximation and smoothed IPR models both have very inaccurate estimates as well as very large standard errors of $\rr$.
+The ratio estimator and reparameterized model have fairly accurate estimates of $\rr$ as well as very small standard errors.
+On the other hand, LMA, LMAT, and LBE  have very inaccurate estimates as well as very large standard errors of $\rr$.
 
@@
 Figure \ref{fig:coverage} shows the proportion of repetitions for both repeated sampling and block bootstrapping in which the 95\% confidence intervals of the estimates produced by the models covers the true $\rr$.
-The ratio estimator, reparameterized, and linear model approximation with all time points contain the true $\rr$ value in all of the confidence intervals.
-On the other hand, the smoothed IPR, Markov chain, and sequential Bayes models never cover the true $\rr$ in any of the confidence intervals.
+The methods RE, rRE, LMAT, and IPR with all time points contain the true $\rr$ value in all of the confidence intervals.
+On the other hand, MC and SB models never cover the true $\rr$ in any of the confidence intervals.
+Both LL and LBE cover the true value about 80\% of the time.
 
@@
 	\begin{center}
-		\includegraphics[scale=0.5]{images/coverage2.tiff}
+		\includegraphics[scale=0.5]{images/coverage2.jpg}
 		\caption{Proportion of coverage of true $\rr$ of confidence intervals produced by repeated sampling for the applicable method (delta method or posterior).}
@@
 
-\begin{figure}[H]
-	\begin{center}
-		\includegraphics[scale=0.5]{images/coverage.tiff}
-		\caption{Proportion of coverage of true $\rr$ of confidence intervals produced by repeated sampling for block bootstrapping.}
-		\label{fig:coverage}	
-	\end{center}
-\end{figure}
+% \begin{figure}[H]
+% 	\begin{center}
+% 		\includegraphics[scale=0.5]{images/coverage.jpg}
+% 		\caption{Proportion of coverage of true $\rr$ of confidence intervals produced by repeated sampling for block bootstrapping.}
+% 		\label{fig:coverage}	
+% 	\end{center}
+% \end{figure}
 
@@
 	\centering
-	\includegraphics[width=\textwidth]{images/h1n1.tiff}
+	\includegraphics[width=\textwidth]{images/h1n1.jpg}
 	\caption{Results from SIR models on H1N1 data.}
@@
 
-\com{Results TBD}
+The results of this analysis are that most of our methods yield point estimates of $\hat{\rr}$ of about 1.2, except for SB and IPR.
+The methods RE, rRE, LMA, LMAT, LL, and MC all have small CIs as well.
+These estimates show that one, H1N1 was indeed an epidemic in the time period but are also smaller compared to the Towers and Feng estimate.
+As Towers and Feng used a different method of estimating $\rr$ than the above, this shows the importance of testing modelling assumptions when using on epidemic data.
+
+\com{Graph to show how R0 changes with T.  X axis is time start, y axis is time end of data used.  size is magnitude of r0}
 
@@
 A value of $\rr > 1$ indicates an outbreak of a disease whereas a value of $\rr < 1$ indicates that no outbreak will occur.
-Thus, it is critical to not only estimate the value of $\rr$ but also a corresponding CI.
+It is critical to not only estimate the value of $\rr$ but also a corresponding CI.
 Overestimating $\rr > 1$ during an ongoing epidemic may result in a waste of resources in prevention and control measures; underestimating $\rr < 1$ may result in a worldwide crisis.
@@
 In Table \ref{tab:r0-real-ex}, we provide estimates for $\rr$ that have appeared over the past three decades for HIV, Zika, Ebola, influenza, H1N1 influenza, and measles in different locations throughout the world.
-Each of these estimates is greater than $1$, which makes sense as we tend to care more when a disease has the potential outbreak.
+Each of these estimates is greater than $1$, which makes sense as we tend observe prolonged outbreaks of a disease as opposed to isolated occurences.
 The largest $\rr$ point estimate we have seen is for measles at $\rr=17$ in England in the 20th century, and the smallest is $\rr=1.08$ for HIV in Portugal in the 21st century.
@@
 We provide three statistical methods to estimate the standard error of $\hat{\rr}$ and use the interval $\left [\hat{\rr} \pm 2\hat{SE}\left (\hat{\rr} \right ) \right ]$ as our 95\% CI.
-We hope that these intervals will provide guidance on how to create more meaningful intervals when estimating $\rr$ and discerning which diseases are more severe than others, and whether the true $\rr >1$.
+We hope that these estimates will provide guidance on how to create more meaningful intervals when estimating $\rr$ and discerning which diseases are more severe than others.
 
@@
 \begin{itemize}
-  \item RE and rRE provide the most accurate estimates of $\rr$
+  \item RE and rRE provide the most accurate and reliable estimates of $\rr$
   \item LL results in close point estimates but occasionally has too small standard errors
   \item The specific MC and SB methods presented here do not seem to fit our generated data well
-  \item There is not much difference in estimations of $\rr$ due to different error generating processes
+  \item There is not much difference in estimations of $\rr$ for the four different error generating processes we use
   \item Very small values of $\gamma$ can result in very poor estimates of $\rr$
-  \item 100 data points seems to be a reasonable number of data points to have enough power for testing that $\rr > 1$ and 50 seems to be too few
   \item The larger the number of initially infectious individuals is, the smaller the CIs for $\rr$ are
   \item The noisier the epidemic data is, the larger our CIs for $\rr$ are
-  \item Estimations of $\rr$ are not robust to the misspecified data sets we presented.  Therefore, model specification is of grave importance
+  \item Estimations of $\rr$ are not robust to the misspecified data sets we presented.  Therefore, model specification is extremely important
+    \item $\hat{rr}$ is most sensitive to changes in $T$, the amount of time points used as well as the beginning and end of this time period
   \item Block bootstrap matches the empirical variance through repeated sampling.
@@
 We also see a large difference between the $\rr$ estimates and standard errors on the exact same data sets as well, even though all derived from the same model.
-Thus, when estimating $\rr$ based on an SIR or any other kind of compartment model, we need to take careful consideration.
+Thus, when estimating $\rr$ based on an SIR or any other kind of compartment model, we need to take careful consideration of all known factors.
 

### 2020-02-28 paper/SiM-v1/draft_v16.tex
added h1n1 data

--- 2018-09-06 paper/Version16/draft_v16.tex
@@
 
-\noindent \textbf{Gaussian-Monotone-PAVA (G-M) ($\hat{X}_{GP}$, $\hat{Y}_{GP}$, $\hat{Z}_{GP}$)}
-\begin{align*}
- X_{GP}(t) &= \textnormal{PAVA}(X_G(t)) \\
-  Y_{GP}(t) &= N - X_{GP}(t) - Z_{GP}(t) \\
-  Z_{GP}(t) &= \textnormal{PAVA}(Z_G(t)) \\
-\end{align*}
-\textbf{Autoregressive-Monotone-PAVA (AR-M) ($\hat{X}_{ARP}$, $\hat{Y}_{ARP}$, $\hat{Z}_{ARP}$)}
-\begin{align*}
-  X_{ARP}(t) &= \textnormal{PAVA}(X_{AR}(t)) \\
-  Y_{ARP}(t) &= N - X_{ARP}(t) - Z_{ARP}(t) \\
-  Z_{ARP}(t) &= \textnormal{PAVA}(Z_{AR}(t)) 
-\end{align*}
-This gives us a total of four error structures for each set of conditions  (1) Gaussian (G), (2) Gaussian monotone - PAVA (G-M),  (3) autoregressive (AR), and (4) autoregressive monotone - PAVA (AR-M) for a given set of parameters $\left (\beta, \gamma, T, X(0), Y(0), Z(0), \sigma_X, \sigma_Y, N \right )$, the infection rate; recovery rate; total number of time points; initial conditions for number of susceptibles, infectious, and recovered individuals; the standard deviations for both the susceptible and infectious compartments; and the total population, respectively.
-
-\subsection{SIR Data}
-
-We first simulate a ``baseline'' data set from the SIR model under the following conditions:
-
-\noindent \textbf{Baseline Simulated Data}
-\begin{center}
-
-	$\beta$ = 0.06, $\gamma$ = 0.03 ($\rr$ = 2)
-
-	Number of time steps $T$: 365
-
-	Starting X size ($X(0)$): 99950
-
-	Starting Y size ($Y(0)$): 50
-
-	Starting Z size ($Z(0)$): 0
-
-	$\sigma_X$: 100
-
-	$\sigma_Y$: 5
-
-	Total population $N$: 100,000
-
-        Error: Gaussian.
-
-      \end{center}
+\noindent \textbf{Gaussian-Monotone-PAVA (GP) ($\hat{X}_{GP}$, $\hat{Y}_{GP}$, $\hat{Z}_{GP}$)}
+\begin{align*}
+ \hat{X}_{GP}(t) &= \textnormal{PAVA}(\hat{X}_G(t)) \\
+  \hat{Y}_{GP}(t) &= N - \hat{X}_{GP}(t) - \hat{Z}_{GP}(t) \\
+  \hat{Z}_{GP}(t) &= \textnormal{PAVA}(\hat{Z}_G(t)) \\
+\end{align*}
+\textbf{Autoregressive-Monotone-PAVA (ARP) ($\hat{X}_{ARP}$, $\hat{Y}_{ARP}$, $\hat{Z}_{ARP}$)}
+\begin{align*}
+  \hat{X}_{ARP}(t) &= \textnormal{PAVA}(\hat{X}_{AR}(t)) \\
+  \hat{Y}_{ARP}(t) &= N - \hat{X}_{ARP}(t) - \hat{Z}_{ARP}(t) \\
+  \hat{Z}_{ARP}(t) &= \textnormal{PAVA}(\hat{Z}_{AR}(t)) 
+\end{align*}
+This gives us a total of four error structures for each set of conditions  (1) Gaussian (G), (2) Gaussian monotone - PAVA (GP),  (3) autoregressive (AR), and (4) autoregressive monotone - PAVA (ARP) for a given set of parameters $\left (\beta, \gamma, T, X(0), Y(0), Z(0), \sigma_X, \sigma_Y, N \right )$, the infection rate; recovery rate; total number of time points; initial conditions for number of susceptibles, infectious, and recovered individuals; the standard deviations for both the susceptible and infectious compartments; and the total population, respectively.
+
+\subsection{Simulated SIR Data}
+
+We first generate ``baseline'' simulated data from the SIR model under the following conditions shown in Table \ref{tab:baseline-params}:
+
+  % Please add the following required packages to your document preamble:
+  % \usepackage{booktabs}
+  % \usepackage{graphicx}
+  \begin{table}[H]
+    \resizebox{\textwidth}{!}{%
+      \begin{tabular}{@{}lllllllllll@{}}
+        \toprule
+        \textbf{Parameter} & $\beta$ & $\gamma$ & $\rr$ & $T$ & $X(0)$ & $Y(0)$ & $Z(0)$ & $\sigma_X$ & $\sigma_Y$ & Error Type \\ \midrule
+        \textbf{Value}     & 0.06    & 0.03     & 2.00  & 365 & 99950  & 50     & 0      & 100        & 5          & Gaussian   \\ \bottomrule
+      \end{tabular}%
+    }
+    \caption{Parameter values for baseline simulated data.}
+    \label{tab:baseline-params}
+  \end{table}
       The baseline data set is generated by adding noise to the susceptible and infectious curves from deterministic SIR paths, which are shown in Figure \ref{fig:baseline-data}.
@@
         \includegraphics[width=.8\textwidth]{images/baseline_sir.pdf}
-        \caption{SIR (XYZ) curves for the deterministic model with $N=100,000$, $(\beta, \gamma) = (0.06, 0.03)$, $(X(0), Y(0), Z(0))= (99950, 50,0)$  The baseline data is generated by using this model and adding noise to the $X$ and $Y$ compartments and adjusting $Z$ at each time step to maintain a constant population.}\label{fig:baseline-data}
-        \end{figure}
-      This baseline data set is chosen due to its moderate $\rr$ size of 2, which indicates an outbreak of an epidemic similar to ZIK-V, EVD, and H1N1 pandemic influenza (see Table \ref{tab:r0-real-ex}).
+        \caption{SIR (XYZ) curves for the deterministic model with $N=100,000$, $(\beta, \gamma) = (0.06, 0.03)$, $(X(0), Y(0), Z(0))= (99950, 50,0)$  The baseline data is generated by using this model and adding noise to the $X$ and $Y$ compartments and adjusting $Z$ at each time step to maintain a constant population $N$.}\label{fig:baseline-data}
+        \end{figure}
+      The parameters for the baseline simulated data is chosen due to its moderate $\rr$ size of 2, which indicates an outbreak of an epidemic similar to ZIK-V, EVD, and H1N1 pandemic influenza (see Table \ref{tab:r0-real-ex}).
 The infection rate and recovery rate ($\beta, \gamma$) are relatively small.
@@
 Initially, 50 (0.05\%) infectious individuals are introduced to the population of 100,000 individuals (approximately the average population size of a US county in 2015).
-The standard deviation of the $\hat{X}$ values is 0.1\% of the magnitude of the initial number of susceptible individuals, and the standard deviation of the $\hat{Y}$ values is 10\% of the magnitude of the initial number of infectious individuals, although in this setting, the number of infectious individuals increases rapidly.
+The standard deviation of the $\hat{X}$ values is 0.1\% of the magnitude of the initial number of susceptible individuals, and the standard deviation of the $\hat{Y}$ values is 10\% of the magnitude of the initial number of infectious individuals.
+In this setting, the number of infectious individuals increases rapidly.
 We use Gaussian error as our baseline error because of its well known statistical properties so we can more easily compare methods to one another.
@@
 
-After generating our baseline data set, we  generate different SIR data sets, changing one of the conditions for simulation each time and using the baseline SIR data as our basis for comparison.
+After generating our baseline simulated data, we generate different SIR data by changing one of the parameters for simulation each time and using the baseline simulated data as our basis for comparison.
 The different data sets for a given error structure are shown in Table \ref{tab:simulated-data}.
@@
 There are numerous aspects to consider when using a method to estimate $\rr$.
-Among these include values of both the infection and recovery rate $(\beta, \gamma)$, the number of data points available ($T$), the total size of the population ($N$), the initial percent of susceptible individuals $\left (\frac{X(0)}{N}\right)$, the initial percent of infectious individuals $\left (\frac{Y(0)}{N}\right )$, the magnitude of the variance of both the number of susceptibles and the number of infectious ($\sigma_X^2, \sigma_Y^2$), the error modeling assumptions, data generated non-SIR models, and finally methods of estimating the variance.
+Among these include values of both the infection and recovery rate $(\beta, \gamma)$, the number of data points available ($T$), the total size of the population ($N$), the initial percent of susceptible individuals $\left (\frac{X(0)}{N}\right)$, the initial percent of infectious individuals $\left (\frac{Y(0)}{N}\right )$, the magnitude of the variance of both the number of susceptibles and the number of infectious ($\sigma_X^2, \sigma_Y^2$), the error modeling assumptions, the structure of the ``true'' model, and finally methods of estimating the variance.
 For the sake of brevity, we do not provide the full results here, and instead examine the results at a high-level.
 However, for those interested readers, we provide results in the form of forest plots and tables of estimates of $\rr$ for each data set in the Supplementary Material.
-
-We implement the \wxxsir methods to estimate $\rr$ using the baseline data set with normal error.  The results are displayed in Figure \ref{fig:baseline-res} and Table \ref{tab:baseline-res}.
+We estimate $\rr$ on the baseline simulated data for each of the \wxxsir methods.
+The results are displayed in Figure \ref{fig:baseline-res} and Table \ref{tab:baseline-res}.
 \begin{figure}[H]
@@
 
-Overall, in Figure \ref{fig:baseline-res}, we see a wide variation in the estimates of and the standard errors for $\rr$ depending on the method used.
+Overall, in Figure \ref{fig:baseline-res}, we see a wide variation in the estimates of and the standard errors for the estimates of $\rr$ depending on the method used.
 The methods of RE, rRE, LMA, LMAT, IPR, and LL all result in a CI that covers the true value of $\rr=2$.
@@
 We would say that IPR's CI is uninformative.
-Additionally, MC has a standard error of less than $.0001$, which is very small.
+Additionally, MC and SIPR have standard errors of less than $.0001$, which is unusually small.
 LBE, which assumes that the X and Z compartments are monotonic, overestimates $\rr$
 
-From examining the full results in the Supplementary Material, the results of the baseline data set with normal error seem to be a good summary of the methods as a whole.
-In Figure \ref{fig:baseline-res}, we see that the clear best methods to estimate $\rr$ are RE, rRE, and LL.
-These three methods consistently outperform the other methods of accurately estimating $\rr$ with a reasonably small CI when varying a single parameter and holding the others as fixed from the generated data sets.
+From examining the full results in the Supplementary Material, the results of the baseline simulated data with normal error seem to be representative of the methods as a whole.
+In Figure \ref{fig:baseline-res}, we see that the best methods to estimate $\rr$ are RE, rRE, and LL.
+These three methods consistently outperform the other methods in the task of accurately estimating $\rr$ with a reasonably small CI when varying a single parameter and holding the others as fixed from the simulated data sets.
 
 In particular, MC and SB almost never capture the true estimate and typically underestimate the true value of $\rr$.
-This leads us to believe that these models are inadequate in capturing the effects of a disease following the SIR model.
+This leads us to believe that these models are for estimating $\rr$ for the SIR model.
 Both models rely on parametric assumptions of simple underlying statistical distributions for the course of infection (Binomial and Poisson, respectively) with time-invariant transmission parameters ($\alpha$ and $\lambda$).
-These are good examples of why we need to consider other reasons for using a model other than the math resulting in a simple estimate.
-That is not to say that variants of these models are unviable for modelling $\rr$, but they do not make for good defaults.
+That is not to say that variants of these models are unviable for modelling $\rr$, but they do not make for good defaults, as they seem to be too simple to accurately model data from a deterministic SIR model.
 It is also an issue that these two methods consistently underestimate $\rr$, which could possibly lead to an unchecked epidemic.
 
-On the other hand, LBE also tends to have larger CIs and does not often capture the true value of $\rr$ when the the X and Z compartments are not monotonic functions.
+On the other hand, LBE tends to have larger CIs and does not often capture the true value of $\rr$ when the the X and Z compartments are not monotonic functions.
 This method tends to overestimate the value of $\rr$.
@@
 
-Of the three well-performing methods, RE and rRE seem to be slightly more robust to parameter variation than LL.
+Of the three best-performing methods, RE and rRE seem to be  more robust to parameter variation than LL.
 This is mostly due to the fact that LL produces such small CIs that even a small discrepancy in estimation of $\rr$ can cause the CI to not capture the true value of $\rr$.
 
-However, we do see instances where LL completely outperforms RE and rRE such as in the case where the initial percent of infectious $\frac{I(0)}{N} \cdot 100\%$ is $< .01$\%.
+However, we do see instances where LL completely outperforms RE and rRE such as in the case where the initial percent of infectious $\frac{I(0)}{N} \cdot 100\%$ is less than $0.01$\%.
 In Figure \ref{fig:ar-small-I}, we examine the data set generated with autoregressive monotonic error, our most stringent data structure, with varying levels of initial percent of infectious individuals.
-When $\frac{I(0)}{N} \cdot 100\%$ $< .01$\%, RE and rRE tend to overestimate $\rr$ and have large CIs.
+When $\frac{I(0)}{N} \cdot 100\%$ is less than $ 0.01$\%, RE and rRE tend to overestimate $\rr$ and have large CIs.
 However, LL accurately estimates $\rr$ even when the initial percentage of infectious is very small.
@@
 
-One interesting feature we see in the results is that RE and rRE can result in very different estimations and CIs, which may seem unintuitive since rRE is a simple reparametrization of RE.
+One interesting feature we see in the results is that RE and rRE can result in very different estimates and CIs, which may seem unintuitive since rRE is a reparametrization of RE.
 For example, in Figure \ref{fig:ar-r0}, we see that when the true value of $\rr$ is 1, rRE has a much smaller CI than RE, which may be one reason to prefer rRE to RE.
-In this figure, we see when $\gamma << \beta$ (e.g.
+In this figure, we see that when $\gamma << \beta$ (e.g.
 $\rr=60)$, then all estimates for $\hat{\rr}$ fail.
@@
 	\includegraphics[scale=0.5]{images/err_robust.jpeg}
-	\caption{Standardized residuals $\frac{\left ( \hat{\rr} - \rr\right ) }{SE \left [\hat{\rr}\right ]}$ for the estimation methods of RE, rRE, and LL faceted by the four error structures.
-The dashed line is the true value of $\rr$.}\label{fig:err-rob}
+	\caption{Standardized residuals $\frac{\left ( \hat{\rr} - \rr\right ) }{SE \left [\hat{\rr}\right ]}$ for the estimation methods of RE, rRE, and LL grouped by the four error structures.
+The dashed line is the true value of $\rr$.
+\com{X axis should be standardized R0 residuals.
+Same for Figure 9.
+Try increasing element\_text size}}\label{fig:err-rob}
 \end{figure}
@@
 	\includegraphics[scale=0.5]{images/model_robust.jpeg}
-	\caption{Standardized residuals $\frac{\left ( \hat{\rr} - \rr\right ) }{SE \left [\hat{\rr}\right ]}$ for the estimation methods of RE, rRE, and LL faceted by the four model structures (SIR, Poly1, Poly4, and LSIR).
+	\caption{Standardized residuals $\frac{\left ( \hat{\rr} - \rr\right ) }{SE \left [\hat{\rr}\right ]}$ for the estimation methods of RE, rRE, and LL grouped by the four model structures (SIR, Poly1, Poly4, and LSIR).
 The dashed line is the true value of $\rr$.}\label{fig:mod-rob}
@@
 
-      In the previous plots, we assume that we are collecting incidence and recovery date every day.
+      In the previous plots, we assume that we are collecting incidence and recovery data every day.
 However, it is also possible to observe weekly, monthly, or even quarterly incidence and recovery data, for example.
@@
 On the other hand, methods such as IPR have consistently large CIs, to the point of being uninformative.
-The other methods CI width seems to depend heavily on the initial parameters used.
+For the other methods, the CI width seems to depend heavily on the parameters used to simulate the data.
 Unsurprisingly, larger $\sigma_X$, $\sigma_Y$ values seem to correspond to larger CI sizes.
@@
 		\hline
-		Model & Mean Estimate & Std. Dev. & Baseline Estimate & Baseline SE\\
+		Model & Empirical Mean Est. & Empirical Std. Dev. & Method Mean & Method SE\\
 		\hline
@@
           \hline
-		Model & Mean Estimate & Std. Dev. & Baseline Estimate & Baseline SE\\
+          	Model & Empirical Mean Est. & Empirical Std. Dev. & Method Mean & Method SE\\
 		\hline
@@
 	\end{tabular}
-	\caption{Mean $\rr$ Estimates and Std. Devs, Block Bootstrap}
+	\caption{Mean $\rr$ Estimates and Std. Devs, Block Bootstrap.  \com{Shouldn't we get predictions from LL?}}
 	\label{tab:bb-samp}
@@
   \includegraphics[width=\textwidth]{images/h1n1-sir.pdf}
-  \caption{Data from H1N1 influenza 2009 in the USA.
+  \caption{\com{Change title to ``Observed SIR...''}Observed data from H1N1 influenza 2009 in the USA.
 Weekly wILI is reported, which we transform into S, I, and R counts shown for weeks (start of weeks: April 16-Aug.
@@
 	\includegraphics[width=\textwidth]{images/h1n1.jpg}
-	\caption{Results from SIR models on H1N1 data.}
+	\caption{\com{Just show 0-12 R0.  Need LBE in table}Results from SIR models on H1N1 data.}
 \end{figure}
@@
 The methods RE, rRE, LMA, LMAT, LL, and MC all have small CIs as well.
+Using RE, we are 95\% confident that $\hat{\rr} \in [0.48, 1.89]$, which is in agreement with the Towers and Feng estimate.
+However, when using rRE we are 95\% confident that $\hat{\rr} \in [1.14, 1.23]$ and for LL $\hat{\rr} \in [1.15, 1.22]$, which is much smaller than their estimate of $\rr$.
 These estimates show that one, H1N1 was indeed an epidemic in the time period but are also smaller compared to the Towers and Feng estimate.
@@
 
-In Table \ref{tab:r0-real-ex}, we provide estimates for $\rr$ that have appeared over the past three decades for HIV, Zika, Ebola, influenza, H1N1 influenza, and measles in different locations throughout the world.
-Each of these estimates is greater than $1$, which makes sense as we tend observe prolonged outbreaks of a disease as opposed to isolated occurences.
+In Table \ref{tab:r0-real-ex}, we provide estimates for $\rr$ that have appeared in the literature over the past three decades for HIV, Zika, Ebola, influenza, H1N1 influenza, and measles in different locations throughout the world.
+Each of these estimates is greater than $1$, which makes sense as we tend to observe prolonged outbreaks of a disease as opposed to isolated occurences.
 The largest $\rr$ point estimate we have seen is for measles at $\rr=17$ in England in the 20th century, and the smallest is $\rr=1.08$ for HIV in Portugal in the 21st century.
@@
 %% the methods we chose and why we chose them
-Additionally, the SIR model assumes monotoncity of S and R compartments which can make estimating $\rr$ more difficult, especially in terms of the CIs since the usual assumption of Gaussian noise may be violated due to the non-symmetric nature of the noise.
+Additionally, the SIR model assumes monotoncity of S and R compartments which can make estimating $\rr$ more difficult, especially in terms of the CIs since the usual assumption of Gaussian noise is violated due to the non-symmetric nature of the noise.
 We examine \wxxsir methods to estimate $\rr$ which include least squares optimization, polynomial approximations, estimates from incidence and variance, a log-linear model, a Markov chain method, and a simple Bayesian method.
 Although, this is not an exhaustive list of methods used to estimate $\rr$ from the SIR model, we believe it covers common and popular methods used throughout the decades.
-We overview the details of methods in Section \ref{sec:methods}.
+We reviewed the details of methods in Section \ref{sec:methods}.
 
@@
 \begin{itemize}
-  \item RE and rRE provide the most accurate and reliable estimates of $\rr$
+  \item RE and rRE provide the most accurate and precie estimates of $\rr$
   \item LL results in close point estimates but occasionally has too small standard errors
   \item The specific MC and SB methods presented here do not seem to fit our generated data well
-  \item There is not much difference in estimations of $\rr$ for the four different error generating processes we use
+  \item There is not much difference in estimates of $\rr$ for the four different error generating processes we use
   \item Very small values of $\gamma$ can result in very poor estimates of $\rr$
@@
   \item The noisier the epidemic data is, the larger our CIs for $\rr$ are
-  \item Estimations of $\rr$ are not robust to the misspecified data sets we presented.  Therefore, model specification is extremely important
-    \item $\hat{rr}$ is most sensitive to changes in $T$, the amount of time points used as well as the beginning and end of this time period
+  \item Estimates of $\rr$ are not robust to the misspecified data sets we presented.  Therefore, model specification is extremely important
+    \item $\hat{\rr}$ is most sensitive to changes in $T$, both through the number of time points used as well as the beginning and end of this time period.
   \item Block bootstrap matches the empirical variance through repeated sampling.
@@
 
-$\rr$ is a difficult quantity to estimate, but we hope that our methods of estimation and simulations provide some guidance for researchers on how to generate statistically reliable estimates for $\rr$ and its CIs.
+  We also analyzed estimates of $\rr$ using data from the 2009 H1N1 pandemic in the USA.
+Our most reliable methods (RE, rRE, and LL) estimate $\hat{\rr}\approx 1.2$ where RE has a standard error estimate of $0.35$ while rRE and LL have a standard error of approximately 0.02.
+This estimate is lower than the reported estimate from Towers and Feng of $\hat{\rr} \in [1.7, 1.9]$.
+
+Overall, we find that $\rr$ is a difficult quantity to estimate, but we hope that our methods of estimation and simulations provide some guidance for researchers on how to generate statistically reliable estimates for $\rr$ and its CIs.
 Our analysis is limited in scope to the SIR model, and in the future we would like to see how methods of estimating $\rr$ compare for models such as the SEIR and beyond.
@@
 
-The SIR estimation methods generally give reliable estimates when the data are from an underlying SIR model with little or no noise.
+The $\rr$ estimation methods generally give reliable estimates when the data are from an underlying SIR model with little or no noise.
 However, the estimates from the various methods are generally much worse when the magnitude of the noise is larger, when the parameters of the model are unusual, and when the model is misspecified.
-We also see a large difference between the $\rr$ estimates and standard errors on the exact same data sets as well, even though all derived from the same model.
-Thus, when estimating $\rr$ based on an SIR or any other kind of compartment model, we need to take careful consideration of all known factors.
+Thus, when estimating $\rr$ based on an SIR or any other kind of compartment model, we need to take careful consideration of all known factors and carefully state and verify the assumptions of the chosen model(s).
 
@@
 \end{figure}
+
+%%% Local Variables:
+%%% mode: latex
+%%% TeX-master: t
+%%% End:

### 2020-02-28 paper/SiM-v1/sim_v1.tex
added h1n1 data

--- 2020-02-28 paper/SiM-v1/draft_v16.tex
@@
-\documentclass[12pt]{article}
-\usepackage[margin=1in]{geometry}
-%\usepackage{hyperref}
-\usepackage{amsmath}
-\usepackage{amsfonts}
-\usepackage{graphicx}% http://ctan.org/pkg/graphicx
-\usepackage{systeme}
-%% tables
+\documentclass[AMA,demo,STIX1COL]{WileyNJD-v2}
+\usepackage{moreverb}
+
+%% big red comments
+\newcommand{\com}[1]{\textcolor{red}{ #1}}
+
+%% nice tables and sideways table
 \usepackage{booktabs}
 \usepackage{lscape}
-\usepackage[table,xcdraw]{xcolor}
-\usepackage{float}
-\usepackage{url}
-%% Make red text
-\newcommand{\com}[1]{\textcolor{red}{ #1}}
-
-%\linespread{2}
-
-\usepackage[authoryear]{natbib}
-
+
+%% Our custom macros
+\newcommand{\XX}{\ensuremath{25}} % number of methods
+\newcommand{\xxsir}{\ensuremath{9} } % number of SIR methods
+s\newcommand{\wxxsir}{nine } % lower case number as word
+\newcommand{\Wxxsir}{Nine } % capitalized word
+\newcommand{\rr}{\ensuremath{\mathcal{R}_0}} %% r0 macro
+
+%% diff eqs
+\usepackage{systeme}
+
+%% tikz graph
 \usepackage{tikz}
@@
 
-\pagenumbering{arabic}
-
-\newcommand{\XX}{\ensuremath{25}} % number of methods
-
-\newcommand{\xxsir}{\ensuremath{9} } % number of SIR methods
-\newcommand{\wxxsir}{nine } % lower case number as word
-\newcommand{\Wxxsir}{Nine } % capitalized word
-
-\newcommand{\rr}{\ensuremath{\mathcal{R}_0}}
-
-\setlength{\itemsep}{0pt}
-
-%% tif -> png
-\usepackage{epstopdf}
-\epstopdfDeclareGraphicsRule{.tiff}{png}{.png}{convert #1 \OutputFile}
-\AppendGraphicsExtensions{.tiff}
+\newcommand\BibTeX{{\rmfamily B\kern-.05em \textsc{i\kern-.025em b}\kern-.08em
+T\kern-.1667em\lower.7ex\hbox{E}\kern-.125emX}}
+
+\articletype{Statistics in Medicine}%
+
+\received{<day> <Month>, <year>}
+\revised{<day> <Month>, <year>}
+\accepted{<day> <Month>, <year>}
+
+% \raggedbottom
+
+\usepackage{tikz}
+\usetikzlibrary{positioning}
+
+\tikzset{
+  int/.style={circle, draw, fill=blue!20, minimum size=3em},
+  init/.style={pin distance=1.2cm,pin edge={loop,thin,black}}
+}
+\usetikzlibrary{arrows}
 
@@
 \title{\Wxxsir methods to estimate $\rr$ in the SIR model with applications to 2009 pandemic influenza}
-\author{ Shannon Gallagher$^{\dag}$, Andersen Chang$^{\ddag}$, and William F.
-Eddy$^{\dag}$ \\$\dag$ Department of Statistics and Data Science, Carnegie Mellon University\\ $\ddag$ Department of Statistics, Rice University}
-\date{\today}
-\maketitle
-
-\begin{abstract}
-For nearly a century, the initial reproduction number ($\rr$) has been used to summarize infectious diseases and to compare them to one another.
+
+\author[1]{Shannon Gallagher*}
+\author[2]{Andersen Chang}
+\author[1]{William F. Eddy}
+
+\authormark{Gallagher, Chang, and Eddy}
+
+\address[1]{\orgdiv{Department of Statistics}, \orgname{Carnegie Mellon University}, \orgaddress{\state{PA}, \country{USA}}}
+\address[1]{\orgdiv{Department of Statistics}, \orgname{Rice University}, \orgaddress{\state{TX}, \country{USA}}}
+
+\corres{*Shannon Gallagher, Department of Statistics; Carnegie Mellon University; Pittsburgh, PA 15213. \email{sgallagh@stat.cmu.edu}}
+
+\presentaddress{Department of Statistics; Carnegie Mellon University}
+
+\abstract[Abstract]{For nearly a century, the initial reproduction number ($\rr$) has been used to summarize infectious diseases and to compare them to one another.
 Difficulties in estimating $\rr$ arise from on both how a disease transmits through a population as well as from differences in statistical estimation methods.
@@
 As real world motivation, we analyze the 2009 outbreak of the H1N1 pandemic influenza in the USA and compare the results from our nine methods to a previous study.
-We discuss the most important aspects from our results which effect the estimation of $\rr$ and provide guidelines for estimating point estimates and confidence/credible intervals to create reliable, comparable estimates of $\rr$.
-  \end{abstract}
+We discuss the most important aspects from our results which effect the estimation of $\rr$ and provide guidelines for estimating point estimates and confidence/credible intervals to create reliable, comparable estimates of $\rr$}
+
+\keywords{Compartment models; Agent-based models; Stochastic infectious disease modelling.}
+
+\jnlcitation{\cname{%
+\author{Gallagher, S.} and
+\author{Eddy, W. F}.
+\ctitle{Stochastic Equivalence of Compartment and Agent-based Models}.  September 2017.}
+}
+
+\maketitle
 
@@
 \begin{align*}
-  (\hat{\beta}_{XEYZ}, \hat{\gamma}_{XEYZ}, \hat{\mu}_{XEYZ} )&=\arg \min_{\beta, \gamma, \mu} \sum_{t} \left [ \left (X_{XEYZ}(t) - X(t)\right )^2  \right . \\
-                                                                 &+\left ( E_{XEYZ}(t) - E(t) \right )^2  \\
-  &\left . + \left ( Z_{XEYZ}(t) - Z(t) \right )^2\right ]
+  (\hat{\beta}_{XEYZ}, \hat{\gamma}_{XEYZ}, \hat{\mu}_{XEYZ} )&=\textnormal{arg} \textnormal{min}_{\beta, \gamma, \mu} \sum_{t} \left [ \left (X_{XEYZ}(t) - X(t)\right )^2  \right . \\
+                                                                 & + \left ( E_{XEYZ}(t) - E(t) \right )^2  \\
+  & \left . + \left ( Z_{XEYZ}(t) - Z(t) \right )^2\right ]
 \end{align*}
@@
 \begin{align*}
-(\hat{\beta}_{XYZ}, \hat{\gamma}_{XYZ} )&=\arg \min_{\beta, \gamma} \sum_{t} \left [ \left (X_{XYZ}(t) - X(t)\right )^2 + \left ( Z_{XYZ}(t) - Z(t) \right )^2 \right ].
+(\hat{\beta}_{XYZ}, \hat{\gamma}_{XYZ} )&=\textnormal{arg} \textnormal{min}_{\beta, \gamma} \sum_{t} \left [ \left (X_{XYZ}(t) - X(t)\right )^2 + \left ( Z_{XYZ}(t) - Z(t) \right )^2 \right ].
 \end{align*}
@@
 \begin{align*}
-(\hat{\beta}, \hat{\gamma} )&=\arg \min_{\beta, \gamma} \sum_{t} \left [ \left (\hat{X}(t) - X(t)\right )^2 + \left ( \hat{Z}(t) - Z(t) \right )^2 \right ]
+(\hat{\beta}, \hat{\gamma} )&=\textnormal{arg} \textnormal{min}_{\beta, \gamma} \sum_{t} \left [ \left (\hat{X}(t) - X(t)\right )^2 + \left ( \hat{Z}(t) - Z(t) \right )^2 \right ]
 \end{align*}
@@
   X(t) &=  X(0) e^{\frac{-\beta}{\gamma}\frac{Z(t)}{N}} \nonumber\\
-  \log \frac{X(t)}{X(0)} &=  \frac{-\beta }{\gamma}\frac{Z(t)}{N} \nonumber\\
-  \log \frac{X(t)}{X(0)} &=  -\rr \frac{Z(t)}{N}. \label{eq:harko_lin}
+  \textnormal{log} \frac{X(t)}{X(0)} &=  \frac{-\beta }{\gamma}\frac{Z(t)}{N} \nonumber\\
+  \textnormal{log} \frac{X(t)}{X(0)} &=  -\rr \frac{Z(t)}{N}. \label{eq:harko_lin}
 \end{align}
@@
 \begin{align*}
-  \hat{\rr} = -\frac{\sum_{t=0}^T \log \frac{ X(t)}{X(0)}}{\sum_{t=0}^T\frac{Z(t)}{N}}.
+  \hat{\rr} = -\frac{\sum_{t=0}^T \textnormal{log} \frac{ X(t)}{X(0)}}{\sum_{t=0}^T\frac{Z(t)}{N}}.
 \end{align*}
@@
 \begin{align*}
-  \log \left ( X(t+1)/ X(0) \right ) - \log \left ( X(t)/X(0) \right ) &= - .01\rr\\
-  \log \left ( X(t+1) \right ) - \log \left ( X(t) \right )  &=- .01\rr\\
-  \log \left ( X(t) / X(t+1) \right ) &= .01\rr\\
+  \textnormal{log} \left ( X(t+1)/ X(0) \right ) - \textnormal{log} \left ( X(t)/X(0) \right ) &= - .01\rr\\
+  \textnormal{log} \left ( X(t+1) \right ) - \textnormal{log} \left ( X(t) \right )  &=- .01\rr\\
+  \textnormal{log} \left ( X(t) / X(t+1) \right ) &= .01\rr\\
   \frac{X(t)}{X(t+1)}  &\approx e^{\rr}.
@@
 \begin{align*}
-\hat{\alpha} = \arg \max_{\alpha} \mathcal{L}\left(\alpha; \textnormal{data} \right ).
+\hat{\alpha} = \textnormal{arg} \textnormal{max}_{\alpha} \mathcal{L}\left(\alpha; \textnormal{data} \right ).
   \end{align*}
@@
 \begin{align}\label{eq:r0-mc}
-\hat{\rr} &= \log \left ( \frac{1}{1-\hat{\alpha}}\right ).
+\hat{\rr} &= \textnormal{log} \left ( \frac{1}{1-\hat{\alpha}}\right ).
 \end{align}
@@
 \begin{align}\label{eq:lbe-r0}
-  (\hat{\beta}, \hat{\gamma}) &= \arg \max_{\beta, \gamma} \mathcal{L}\left ( \beta, \gamma; \textnormal{Data} \right )\nonumber\\
+  (\hat{\beta}, \hat{\gamma}) &= \textnormal{arg} \textnormal{max}_{\beta, \gamma} \mathcal{L}\left ( \beta, \gamma; \textnormal{Data} \right )\nonumber\\
   \rr &= \frac{\hat{\beta}}{\hat{\gamma}}
@@
 \begin{align*}
-J(t+1) | \rr  \sim \textnormal{Poisson}( J(t) \exp \left \{  \gamma (\rr-1)\right \})
+J(t+1) | \rr  \sim \textnormal{Poisson}( J(t) \textnormal{exp} \left \{  \gamma (\rr-1)\right \})
 \end{align*}
@@
 \begin{align*}
-h(\alpha) &= \log \left ( \frac{1}{1-\alpha}\right ) \\
+h(\alpha) &= \textnormal{log} \left ( \frac{1}{1-\alpha}\right ) \\
   h^\prime(\alpha) &= \alpha - 1 \\
@@
 
-\pagebreak
-
-\bibliographystyle{apa}%Choose a bibliograhpic style
-\bibliography{Master}
-
-\appendix
-
 \end{document}
 
-%%%%%%%%%%%%%%%%%%%%%%
-\textbf{Title:}
-
-\textbf{Author:}
-
-\textbf{Citation:}
-
-\textbf{Major themes:}
-
-\textbf{Notes:}
-\\
-%%%%%%%%%%%%%%%%%%%%%%
-
-\begin{figure}[h]
-\begin{center}
-\includegraphics[width=4in]{images/mvhw7_3c.pdf}
-\end{center}
-\caption{
-Biplots of the different continuous variables and their correlations.}\label{fig3c}
-\end{figure}
-
-%% two pictures whoa
-
-\begin{figure}[h]
-\centering
-
-\begin{figure}[h]
-\centering
-
-\begin{subfigure}{.5\textwidth}
-  \centering
-  \includegraphics[width=1\linewidth]{images/mt_eda_cont_hists.pdf}
-  \caption{Histograms of Arrival Delay and continuous covariates.
-Arrival delay seems to have a right skewed distribution.
-This may indicate that we will be transforming this variable later on.
-After transforming Air Time and Distance by a log transformation, we don't really seem to have many outliers in our covariates.
-We seem to have outliers in the CRS Dep.
-Time and Arrival Time; however, time is cyclical and so these are not, in fact outliers.}
-  \label{hists}
-\end{subfigure}%
-\begin{subfigure}{.5\textwidth}
-  \centering
-  \includegraphics[width=1\linewidth]{images/mt_eda_cont_hists.pdf}
-  \caption{\textcolor{red}{Placeholder}}
-  \label{tabs1}
-\end{subfigure}
-\caption{}
-\end{figure}
-
-\begin{table}
-\begin{tabular}{l | a | b | a | b}
-\hline
-\rowcolor{LightCyan}
-\mc{1}{}  & \mc{1}{x} & \mc{1}{y} & \mc{1}{w} & \mc{1}{z} \    \hline
-variable 1 & a & b & c & d \    variable 2 & a & b & c & d \\ \hline
-\end{tabular}
-\end{table}
-
-%%
-\begin{figure}
-\centering
-\begin{subfigure}{.5\textwidth}
-  \centering
-  \includegraphics[width=1\linewidth]{images/resids_full.pdf}
-  \caption{}
-  \label{residsf}
-\end{subfigure}%
-\begin{subfigure}{.5\textwidth}
-  \centering
-  \includegraphics[width=1\linewidth]{images/diags_full.pdf}
-  \caption{ }
-  \label{diagsf}
-\end{subfigure}
-\caption{}
-\end{figure}
+%  To get the journal style of heading for an appendix, mimic the following.
 

### 2020-02-28 h1n1/get_data.R

@@ -1,0 +1,174 @@
+## SKG
+## June 15, 2018
+## Getting data for Pandemic influenza between April and August 2009 for USA
+
+
+## DOWNLOAD DATA:
+## SEASON: 2008-2009
+## REGION: NATIONAL
+## DATA SOURCES: WHO/NREVSS and ILINET
+## https://gis.cdc.gov/grasp/fluview/fluportaldashboard.html
+
+## READ IN DATA
+flu <- read.csv("ILINet.csv", skip = 1)
+head(flu)
+
+plot(flu$X..WEIGHTED.ILI, type = "l", lwd=2)
+
+## Get the date
+## https://www.cmmcp.org/sites/cmmcp/files/uploads/2009epi.pdf
+## Week 21 start 5/23, Week 33 ends 8/22
+## Week 15 (April 13) - Week 34 (ends Aug. 29)
+
+start_week <- 15
+end_week <- 34
+start_ind <- which(flu$WEEK == start_week)
+end_ind <- which(flu$WEEK == end_week)
+flu_sub <- flu[start_ind:end_ind,]
+plot(flu_sub$X.UNWEIGHTED.ILI, type = "l", lwd = 2)
+
+## Population of US on April 1, 2010:
+## https://www.census.gov/quickfacts/fact/table/US/POP010210#viewtop
+## 308,740,000 individuals
+
+## So assumingggg
+
+us_pop <- 30874000
+incidence <- flu_sub$X.UNWEIGHTED.ILI / 100 * us_pop
+S <- us_pop - cumsum(incidence)
+
+## gamma = 1/3 days{-1} (inverse of avg recovery = 3days)
+## gamma = (7/3) weeks^{-1} (i.e. average duration is 3/7 week)
+## SOURCE: VESPIGNANI ET AL 2006 http://journals.plos.org/plosmedicine/article/file?id=10.1371/journal.pmed.0040013&type=printable
+
+## Incidence to prevalence
+## http://sphweb.bumc.bu.edu/otlt/MPH-Modules/EP/EP713_DiseaseFrequency/EP713_DiseaseFrequency7.html
+I <- 3/7 * incidence
+R <- us_pop - S - I
+
+
+library(lubridate)
+df <- data.frame(S = S, I =I, R=R)
+df$date <- ymd("2009-01-01") + weeks(flu_sub$WEEK )
+write.csv(df, "../../r0_refs/Data/H1N12009/data.csv", row.names = FALSE)
+
+
+##############################################################
+## Trying again with confirmed cases
[…]
